### 2017-06-10 main.tex
Preliminary text for intro and part of Buzzard

%\documentclass[a4paper]{article}
\documentclass[usenatbib]{mn2e}
%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}

\usepackage{natbib}
%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

%% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}

\newcommand       \apj  {ApJ}
\newcommand       \apjl  {ApJL}
\newcommand       \apjs  {ApJS}
\newcommand       \aj  {AJ}
\newcommand       \aap {A \& Ap.}
\newcommand       \mnras {MNRAS}
\newcommand       \nat {Nature}
\newcommand       \araa {ARA\&A}
\newcommand       \physrep {Phys.~Reports}
\newcommand       \prd {PhysRevD}

\title{Photo-z for LSST: comparison of PDFs}
\author{LSST PZ WG}

\begin{document}
\maketitle

\begin{abstract}
The rationale: - to test different PZ codes on DC1 simulations

- to focus on the pdf p(z) and to compare results via various metrics

- This is a good way to compare quantitatively results by various groups, and also to credit collaborators,

The timeline for the paper:

Which Journal? AJ, ApJ, MNRAS? (leaning toward MNRAS, no page charge…)

\end{abstract}

\section{Introduction}

(Sam Schmidt, Ofer Lahav, Jeff Newman, Alex Malz)

%* The challenges of PZ in LSST 
%* Many papers compared point estimates; here we focus on the pdf $p(z)$
We are entering a new regime with Stage III and Stage IV dark energy experiments[add text about DES, LSST, WFIRST, Euclid, along with references].  The move to imaging based surveys, rather than spectroscopic based, for cosmological measurements makes proper understanding of photometric redshifts ("photo-z's") of paramount importance, as all cosmological distance measures for statistical samples will be dependent on photo-z measurements.  As we push to fainter magnitudes and deeper imaging, we expand our samples to include lower luminosity and higher redshift galaxies.  These populations introduce major degeneracies, for example the Lyman break/Balmer break degeneracy, that were not present in previous large area surveys (SDSS, 2MASS, etc...[spell out and put references]).  The increased sample size that comes with the added depth means that stringent constraints on photo-z accuracy are necessary if systematic errors are not to dominate the statistical errors.  The LSST Science Requirements Document (SRD)\footnote{available at:https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17} lists the photometric redshift goals for a magnitude limited sample with $i<25$ as: root mean square error of $\sigma<0.02(1+z)$; $3\sigma$ "catastrophic outlier'' rate below 10\%; bias below 0.003.  Note that at the time the SRD was written, these goals were stated in terms of one signle number estimate of photometric redshift for each galaxy.  Many of the previous studies of photo-z performance focused on single point estimates of the redshift.  Such a simple redshift measure fails to capture the full information in the presence redshift degeneracies, where multiple redshift solutions have considerable likelihood.  In this paper we focus on the full probability density function (PDF), or p(z) for each galaxy.  [What else to say about p(z)?]

In order to meet these ambitious goals for photo-z accuracy, every aspect of photo-z estimation will have to be optimized: the algorithms employed, both template and machine-learning based (both in design and implementation),  the spectroscopic data used as a training set for machine learning algorithms or to estimate template sets and train Bayesian priors in template based methods
storage formats for p(z) that are accurate enough for all science cases while simultaneously compact enough to fit in a reasonably spec'd database.
In this paper, we focus on current generation photo-z algorithms, comparing the performance of several of the most widely employed codes.  This is similar in spirit to the PHoto-z Accuracy And Testing \citep[PHAT;][]{Hildebrandt:10} effort, though that comparison used many more photometric bands, and compared only point-estimates.  The Dark Energy Survey (DES, REFERENCE!) compared several codes, both for point estimates \citep[]{Sanchez:14}, and using sum of p(z) for collections of galaxies in tomographic bins \citep[]{Bonnett:16}. (mention that this is not the correct way to combine and cite Malz in prep).

In order to isolate the effects of the codes from issues with the training set and template library, we will use two sets of simulated galaxies and construct a training sample that is {\it representative and spanning}.  We will address the very important issue of incomplete spectroscopic coverage in a future paper.

The outline of the paper is as follows: in \S\,\ref{pzcodes} we describe the current generation codes employed in the paper; in \S\,\ref{sims} we present the two simulated data sets; in \S\,\ref{metrics} we discuss the "meaning'' of redshift PDFs, and lay out the metrics that we use to evaluate them; in \S\,\ref{results} we show our results and compare the performance of the codes; in \S\,\ref{discussion} we offer our conclusions and discuss future extensions of this work.

\section{PZ Codes}\label{pzcodes}

(Sam Schmidt, Rongpu Zhou, Ibrahim Almosallam, Eric Nuss, Johann Cohen Tanugi, John Soo, Alex Malz)

* Existing training and testing codes (briefly, possibly a Table)

* What is the meaning of $p(z)$ in different codes? (Ofer)

* ACTION ITEM: who has what code in place? Are any important codes missing?

* Action item: make a table of codes (Sam)

\section{The DC1 simulations}\label{sims}

(Sam Schmidt)

* Describe the simulations (Sam, Eve, etc…)

* Describe the template sets (Sam)

* Discuss limitations wrt to the galaxies in nature (Eve, Tina Peters)

In order to test the current generation codes, we created two simulated datasets.  The simulations are completely catalog-based, with no image construction or mock measurements made [phrase this better].  We describe these in detail below.

\subsection{Buzzard}\label{buzzard}

   The Buzzard-highres-v1.0 catalog construction began with an dark matter only simulation.  This N-body simulation contained 2048$^3$ particles in a 400 Mpc/h box.  [HOW MANY SNAPSHOTS?] time snapshots were saved in order to construct a lightcone [Any smoothing/interpolation done between snapshots?].  Dark matter halos were identified using the {\it Rockstar} software package \citep[]{Behroozi:13}.  These dark matter halos were populated with galaxies with a stellar mass and  absolute r-band magnitude (in the [SDSS or DES?] system) determined using a sub-halo abundance matching model constrained to match both projected two-point galaxy clustering statistics and an observed conditional stellar mass function \citep[]{Reddick:13}.

To assign an spectral energy distribution (SED) to each galaxy, the Adding Density Dependent Spectral Energy Distributions (ADDSEDS) [Is there a reference paper for ADDSEDS?] procedure was used.  This consisted of training an empirical relation between absolute r-band magnitude, local galaxy density, and SED using a sample of $\sim 5e^{5}$ galaxies from the Sloan Digital Sky Survey [PUT PROPER REF FOR specz sample]. [should we mention that we actually only have 5 NNMF components? and they are same as k-correct?].  The distance to the projected fifth-nearest neighbor was used as a proxy for local density in the SDSS training sample.  For each simulated galaxy, a ``random'' [How exactly is this done? random in a certain volume of parameter space, maybe?] galaxy with ``similar'' [again, this needs detail] absolute r-band magnitude and local galaxy density was chosen from the training set, and that training galaxy's SED was assigned to the simulated galaxy.  Given the SED, absolute r-band magnitude, and redshift, we computed apparent magnitudes in the six LSST filter passbands, $ugrizy$.  We assigned magnitude errors in the six bands using the simple model described in Section 3.2 of \citet[]{Ivezic:08}, assuming full 10-year depth observations had been completed [should we list the number of visits and such?].

The total catalog covered 400 square degrees and contained 238 million galaxies to an apparent magnitude limit of $r\!=\!29$ and spanning the redshift range $0\!<\!z\!\leq\!8.7$.  This catalog contained two orders of magnitude more galaxies than were needed for this study, so only $\sim\!8$ square degrees were used.  Systematic problems with galaxy colours above $z\!>\!2$ were observed, so the catalog was trimmed to include only galaxies in the redshift range$0\!<\!z\!\leq\!2.0$.

break up into test/train, list numbers in each

\section{Metrics for quantifying pdf comparisons}\label{metrics}

(Alex Malz)

What is the meaning of “pdf”?

- QQ (Rongpu, Jeff)

- KS (Ofer)

- $D_{KL}$ (Alex)

- etc

(JAN: note the distinction between these: the QQ test allows us to test the accuracy of individual  photo-z PDFs taken one at a time, KS and $D_{KL}$ do not.  I'd also suggest expanding our view from KS to include Cramer-von Mises (which is more sensitive to errors on the tails, which matter a lot for cosmology) and Anderson-Darling (which is more sensitive to the tails still).

\section{Results}\label{results}

(Sam Schmidt, Johann Cohen Tanugi, Rongpu Zhou)

- plots

- comparison tables per metric

\section{Summary and Discussion}\label{discussion}

(Sam Schmidt, Ofer Lahav, Jeff Newman, Alex Malz, Eve Kovacs, Tony Tyson)

- challenges

Why some codes work better?

(possibly outside scope after this point?)
Combination of codes?

-  Future inclusion of AGN effects (Tina Peters)

- future work: from $p(z)$ to $N(z)$; end-to-end analysis towards cosmological parameter, DC2

- future work: $p(z,\alpha)$
% \section{Some examples to get started}

% \subsection{How to add Comments}

% Comments can be added to your project by clicking on the comment icon in the toolbar above. % * <john.hammersley@gmail.com> 2014-09-03T09:54:16.211Z:
% %
% % Here's an example comment!
% %
% To reply to a comment, simply click the reply button in the lower right corner of the comment, and you can close them when you're done.

% \subsection{How to include Figures}

% First you have to upload the image file from your computer using the upload link the project menu. Then use the includegraphics command to include it in your document. Use the figure environment and the caption command to add a number and a caption to your figure. See the code for Figure \ref{fig:frog} in this section for an example.

% \begin{figure}
% \centering
% \includegraphics[width=0.3\textwidth]{frog.jpg}
% \caption{\label{fig:frog}This frog was uploaded via the project menu.}
% \end{figure}

% \subsection{How to add Tables}

% Use the table and tabular commands for basic tables --- see Table~\ref{tab:widgets}, for example.

% \begin{table}
% \centering
% \begin{tabular}{l|r}
% Item & Quantity \\\hline
% Widgets & 42 \\
% Gadgets & 13
% \end{tabular}
% \caption{\label{tab:widgets}An example table.}
% \end{table}

% \subsection{How to write Mathematics}

% \LaTeX{} is great at typesetting mathematics. Let $X_1, X_2, \ldots, X_n$ be a sequence of independent and identically distributed random variables with $\text{E}[X_i] = \mu$ and $\text{Var}[X_i] = \sigma^2 < \infty$, and let
% \[S_n = \frac{X_1 + X_2 + \cdots + X_n}{n}
%       = \frac{1}{n}\sum_{i}^{n} X_i\]
% denote their mean. Then as $n$ approaches infinity, the random variables $\sqrt{n}(S_n - \mu)$ converge in distribution to a normal $\mathcal{N}(0, \sigma^2)$.

% \subsection{How to create Sections and Subsections}

% Use section and subsections to organize your document. Simply use the section and subsection buttons in the toolbar to create them, and we'll handle all the formatting and numbering automatically.

% \subsection{How to add Lists}

% You can make lists with automatic numbering \dots

% \begin{enumerate}
% \item Like this,
% \item and like this.
% \end{enumerate}
% \dots or bullet points \dots
% \begin{itemize}
% \item Like this,
% \item and like this.
% \end{itemize}

% \subsection{How to add Citations and a References List}

% You can upload a \verb|.bib| file containing your BibTeX entries, created with JabRef; or import your \href{https://www.overleaf.com/blog/184}{Mendeley}, CiteULike or Zotero library as a \verb|.bib| file. You can then cite entries from it, like this: \cite{greenwade93}. Just remember to specify a bibliography style, as well as the filename of the \verb|.bib|.

% You can find a \href{https://www.overleaf.com/help/97-how-to-include-a-bibliography-using-bibtex}{video tutorial here} to learn more about BibTeX.

% We hope you find Overleaf useful, and please let us know if you have any feedback using the help menu above --- or use the contact form at \url{https://www.overleaf.com/contact}!

\bibliographystyle{mn2e}
\bibliography{references}

\end{document}

### 2017-06-15 main.tex
Update on Overleaf.

@@
 - This is a good way to compare quantitatively results by various groups, and also to credit collaborators, 
+\end{abstract}
 
@@
 Which Journal? AJ, ApJ, MNRAS? (leaning toward MNRAS, no page charge…)
-
-\end{abstract}
 
@@
 In this paper, we focus on current generation photo-z algorithms, comparing the performance of several of the most widely employed codes.
-This is similar in spirit to the PHoto-z Accuracy And Testing \citep[PHAT;][]{Hildebrandt:10} effort, though that comparison used many more photometric bands, and compared only point-estimates.
+This is similar in spirit to the PHoto-z Accuracy And Testing \citep[PHAT,][]{Hildebrandt:10} effort, though that comparison used many more photometric bands, and compared only point-estimates.
 The Dark Energy Survey (DES, REFERENCE!) compared several codes, both for point estimates \citep[]{Sanchez:14}, and using sum of p(z) for collections of galaxies in tomographic bins \citep[]{Bonnett:16}.
 (mention that this is not the correct way to combine and cite Malz in prep).
+
+ToDo: Mention SRM, multiple data challenges (DCs), and that this is the first of multiple papers, which will grow in sophistication.
 
@@
 
+\subsection{ANNz2}\label{sec:annz2}
+ANNz2\footnote{\url{https://github.com/IftachSadeh/ANNZ}} \citep{sadeh_annz2:_2016} is a machine learning photo-z estimation method which uses the Toolkit for Multivariate Data Analysis (TMVA) with ROOT\footnote{\url{http://tmva.sourceforge.net/}}.
+It is a powerful package that incorporates several machine learning algorithms including ANN, boosted decision tree (BDT) and k-nearest neighbour (KNN).
+ANNz2 can run multiple machine learning algorithms for a single training and outputs photo-z's based on a weighted average of their performances.
+
+Other than the usual regression method to obtain point estimates for photo-z's, ANNz2 is also able to produce redshift posterior probability distributions $P(z)$, conduct classification and support reweighting between samples.
+The pdfs are produced by propagating the intrinsic uncertainty on the input parameters and the uncertainty in the machine learning method to the expected photo-z solution.
+ANNz2 also differs from many other codes by its presentation of photo-z uncertainty, it derives uncertainties using the KNN method: first it estimates the photo-z bias between each object and a fixed number of nearest neighbours in parameter space, it then takes the 68th percentile width of the distribution of the bias.
+This asserts that objects with similar photometric properties should have similar uncertainties, and this error presentation has been shown to perform better than the former \citep{sadeh_annz2:_2016}.
+
+ANNz2 produces four point estimates for each run, but in this study we use the pdf average photo-z, which is the mean photo-z calculated over the pdf.
+The pdf average photo-z will act as the best point estimate photo-z output for ANNz2.
+In this study, a set of 5 ANNs with different random seeds are used during each training.
+
 \section{The DC1 simulations}\label{sims}
 
-(Sam Schmidt)
+(Sam Schmidt, Eve Kovacs, Tina Peters)
 
@@
 
-\subsection{Buzzard}\label{buzzard}
+\subsection{Buzzard (what should we call this officially?)}\label{buzzard}
 
@@
 Systematic problems with galaxy colours above $z\!>\!2$ were observed, so the catalog was trimmed to include only galaxies in the redshift range$0\!<\!z\!\leq\!2.0$.
-
-break up into test/train, list numbers in each
+A random subset of the the remaining galaxies was chose, and placed, again at random, into either a "training'' set (10\% of the sample), for which the galaxies true redshifts will be supplied, or a "test'' set (the remaining 90\% of the sample), for which each code will need to predict a redshift PDF for each galaxy.
+The resulting catalogues contain 111,171 training galaxies and 1,000,883 test galaxies.
 
@@
 
-(Alex Malz)
+(Alex Malz, Rongpu Zhou, Jeff Newman, Ofer Lahav)
 
 What is the meaning of “pdf”?
+
+The overloaded "$p(z)$" is a widespread abuse of notation; we would like the outputs of our codes to be interpretable as probabilities.
+Each PDF is a posterior probability distribution conditioned on a number of things that will differ depending on the method used to derive it; these things will certainly include the photometry associated with the galaxy in question but also the assumptions necessary for the method to be valid and any inputs it takes, such as a template library or training set.
+Furthermore, each PDF must not take negative values and must integrate to unity over the range of possible redshifts.
+
+There are a number of metrics that can be used to test the accuracy of an estimate of a probability distribution if the true probability distribution is known.
+Even for simulated data, the true photo-$z$ PDF is in general not accessible, and because the PDFs produced by each method are conditioned on different information, it is unclear what would be gleaned from direct comparisons between individual PDFs.
+However, we do know the overall redshift distribution $n(z)$ for the \texttt{Buzzard} simulation, which, notably, is itself a probability distribution over redshift and so can be used as the basis for a metric of the photo-$z$ PDFs used to calculate it.
+
+Though alternatives exist (Malz, et al., in prep), stacking is the most widely accepted method for estimating $n(z)$ from photo-$z$ PDFs.
+The exact implementation of the stacked estimator $\hat{n}(z)$ will depend on the parametrization of the photo-$z$ PDFs, which may differ for different codes.
+Once an estimator is derived, it may be compared to the true redshift distribution using a number of metrics, described below.
 
@@
 I'd also suggest expanding our view from KS to include Cramer-von Mises (which is more sensitive to errors on the tails, which matter a lot for cosmology) and Anderson-Darling (which is more sensitive to the tails still).
+(AIM: I'm not sure I agree with that statement.
+In what way can any of them not be used for individual photo-z PDFs?
+I think the bigger issue is that we  can't test accuracy of any of the PDFs because we don't have a concept of a "true" photo-z PDF for each galaxy.)
+
+\subsection{Qualitative Metrics}
+\label{sec:qualmet}
+
+\subsubsection{Q-Q Plot}
+
+\subsubsection{Moments}
+
+\subsection{Quantitative Metrics}
+\label{sec:quantmet}
+
+\subsubsection{RMSE}
+
+\subsubsection{Kolmogorov-Smirnov}
+
+\subsubsection{Anderson-Darling}
+
+\subsubsection{Kullback-Leibler}
 
@@
 
-(Sam Schmidt, Ofer Lahav, Jeff Newman, Alex Malz, Eve Kovacs, Tony Tyson)
+(Sam Schmidt, Ofer Lahav, Jeff Newman, Alex Malz, Eve Kovacs, Tony Tyson, Tina Peters)
 

### 2017-06-16 main.tex
Update on Overleaf.

@@
 
+\subsection{BPZ}\label{sec:BPZ}
+\texttt{BPZ}\footnote{We use \texttt{BPZ v 1.99.3}, available at: \url{http://www.stsci.edu/~dcoe/BPZ/}} \citep[Bayesian Photometric-$z$, see][for details]{Benitez:00} is a template-based photo-$z$ code which compares the expected colors (C) calculated for a set of SED types/templates (T) to the actual observed colors to calculate the likelihood of observing colors $p(C|z,T)$ at each redshift for each type.
+The code employs an empirically determined Bayesian prior in apparent magnitude ($m_0$) and SED-type.
+Assuming that the SED-types are spanning and exclusive, we can determine the redshift posterior by marginalizing over all SED-types with a simple sum (Eq.~3 from \citet[]{Benitez:00}):
+\begin{equation}
+p(z|C,m_0)\,\propto\, \sum_{T}p(z,T|m_0)\,p(C|z,T)
+\end{equation}
+\noindent where the first term is the Bayesian prior and the second term is the traditional likelihood.
+The prior is assumed to have the form: $p(z,T|m_0) = p(T|m_0)\,p(z|T,m_0)$that is, it parameterizes the prior as an evolving type fraction with apparent magnitude, combined with a prior on the expected redshift probability distribution as a function of both apparent magnitude and SED-type.\\
+
+The template set used for the Buzzard dataset is the set of 100 SEDs described in Section~\ref{sec:buzztemplates}, and use the 129 Brown SEDs for the Galacticus dataset.
+For both datasets, to keep the number of free parameters to a manageable level, the SEDs in the training set are sorted by rest-frame $u-g$ colour and split into three ``broad'' SED classes, equivalent to the \texttt{E}, \texttt{Sp} and \texttt{Im/SB} types in \citet[]{Benitez:00}.
+We assume the same functional form for the Bayesian prios as used by \citet[]{Benitez:00}, and utilize the training-set galaxies with known SED-type, redshift, and apparent magnitude to determine the type fractions and the best fit for the 11 free parameters of the prior.
+For photo-$z$ point estimate we use the \texttt{Z\_B} output parameter.
+For the type-marginalized p(z) is generated by setting the parameter \texttt{PROBS\_LITE=TRUE}.
+
 \subsection{ANNz2}\label{sec:annz2}
@@
 The resulting catalogues contain 111,171 training galaxies and 1,000,883 test galaxies.
+
+\subsubsection{Templates}\label{sec:buzztemplates}
+[Describe the templates here? Or describe them earlier?]
+
+\subsection{Galacticus [ABC?]}\label{Galacticus}
 

### 2018-08-10 paper/acknowledgments.tex
paper from overleaf

% This is here for compliance with the DESC Publication Policy, as the paper will ultimately have to be compiled with the start_paper format.
%
%This is the text imported from \code{acknowledgments.tex}, and will be replaced by some standard LSST DESC boilerplate at some point.
%

The DESC acknowledges ongoing support from the Institut National de Physique Nucl\'eaire et de Physique des Particules in France; the Science \& Technology Facilities Council in the United Kingdom; and the Department of Energy, the National Science Foundation, and the LSST Corporation in the United States.
DESC uses resources of the IN2P3 Computing Center (CC-IN2P3--Lyon/Villeurbanne - France) funded by the Centre National de la Recherche Scientifique; the National Energy Research Scientific Computing Center, a DOE Office of Science User Facility supported by the Office of Science of the U.S.\ Department of Energy under Contract No.\ DE-AC02-05CH11231; STFC DiRAC HPC Facilities, funded by UK BIS National E-infrastructure capital grants; and the UK particle physics grid, supported by the GridPP Collaboration.
This work was performed in part under DOE Contract DE-AC02-76SF00515.

% hack event acknowledgments are important
We thank Stony Brook University for hosting the Summer 2017 LSST-DESC Hack Day at which this work was partially completed.

% individual acknowledgments for everyone's grants, probably should be in alphabetical order
AIM is advised by David Hogg and was supported by National Science Foundation grant AST-1517237 and the U.S. Department of Energy, Office of Science, Office of Workforce Development for Teachers and Scientists, Office of Science Graduate Student Research (SCGSR) program, administered by the Oak Ridge Institute for Science and Education for the DOE under contract number DE‐SC0014664.

JYHS is advised by Ofer Lahav and acknowledges support from the MyBrainSc Scholarship, provided by the Ministry of Education, Malaysia. JS would also like to thank Antonella Palmese and Boris Leistedt for guidance on the use of the algorithms \textsc{ANNz2} and \textsc{Delight} respectively.

### 2018-08-12 paper/main.tex

%\documentclass[a4paper]{article}
\documentclass[usenatbib]{mn2e}
%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lineno}
\linenumbers
\modulolinenumbers[2]

\usepackage{natbib}
%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

%% Useful packages
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{multirow}

\newcommand       \apj  {ApJ}
\newcommand       \apjl  {ApJL}
\newcommand       \apjs  {ApJS}
\newcommand       \aj  {AJ}
\newcommand       \aap {A\&A}
\newcommand       \mnras {MNRAS}
\newcommand       \nat {Nature}
\newcommand       \araa {ARA\&A}
\newcommand       \physrep {Phys.~Reports}
\newcommand       \prd {Phys. Rev. D}

\newcommand{\trainz}{\textsc{trainZ}}

\newcommand{\textul}{\underline}
\newcommand*\mathinhead[2]{\texorpdfstring{$\boldsymbol{#1}$}{#2}}

\newcommand{\red}[1]{\textcolor{red}{#1}}
% Claim a color for your comments
\newcommand{\aim}[1]{\textcolor{green}{#1}}%Alex Malz comments in green
\newcommand{\blue}[1]{\textcolor{blue}{#1}}%unknown person comments in blue
\definecolor{scc}{rgb}{0.0, 0.26, 0.15}
\newcommand{\scc}[1]{\textcolor{scc}{#1}}%Stefano Cavuoti Comments in dark green
\definecolor{newpink}{rgb}{0.858, 0.188, 0.478}
\newcommand{\erfan}[1]{\textcolor{newpink}{#1}} % Erfan Nourbakhsh comments in pink
\newcommand{\jan}[1]{\textcolor{orange}{#1}}%Alex Malz comments in green

\def\X{{\mathbf{X}}}
\def\x{{\mathbf{x}}}
\def\E{{\mathbb E}}

\title{An assessment of photometric redshift PDF performance in the context of LSST}
% Alternate title proposal: Assessing the characterization of photometric redshift uncertainties in the regime of LSST

\author{LSST-DESC Photometric Redshift Working Group}

\begin{document}
\maketitle

\begin{abstract}
Photometric redshift (photo-$z$) probability distribution functions (PDFs) are a key data product of nearly all upcoming galaxy imaging surveys.
However, the photo-$z$ PDFs resulting from different techniques are not in general consistent with one another, and an optimal method for obtaining an accurate PDF remains unclear.  
We present the results of an initial study of the Large Synoptic Survey Telescope Dark Energy Science Collaboration (LSST-DESC), the first in a planned series of papers testing multiple photo-$z$ codes on simulations of upcoming LSST galaxy photometry catalogues.  This initial test evaluates photo-$z$ algorithms in the presence of representative training data and in the absence of several common sources of systematic errors that affect the procedures by which photo-$z$ PDFs are derived.  
The photo-$z$ PDFs are evaluated using multiple metrics including the Kolmogorov-Smirnoff statistic, Cramer-von Mises statistic, Anderson-Darling statistic, Kullback-Leibler divergence, $N(z)$ moments, quantile-quantile plots and probability integral transform.  We observe several trends, including an overall over/under-prediction in the broadness of the PDFs for several of the codes.  A careful accounting of all systematics discovered will be necessary for the codes employed in upcoming analyses in order to achieve unbiased cosmological measurements.
\end{abstract}
\begin{keywords}
galaxies: distances and redshifts -- galaxies: statistics -- methods: statistical
\end{keywords}
%\noindent comments color-code:\\
%\red{SJS},
%\aim{AIM},
%\scc{SC},
%\textcolor{brown}{MB},\jan{JAN}
%\blue{unknown}\\
%\red{note: not all red text is mine, at some point various people were using red --SJS.}

% Questions left unanswered:
% 1. Which journal will this paper be published? 
% Answer: AJ / ApJ / MNRAS (no charges, but not sure if it will accept a paper based on a DC1 effort and the associated simulations)
% 2. What is the meaning of $p(z)$ in different codes? (Ofer)
% Answer: ?
%

\section{Introduction}
\label{sec:intro}

%(Sam Schmidt, Ofer Lahav, Jeff Newman, Alex Malz, John Soo, Tony Tyson)

%\jan{Is the intention to keep these parenthetical author sets?  I find it a bit offputting in the text.  It might be better to put this for sections in the acknowledgements section, and for individual codes/statistics specify who ran/developed it with a sentence.}\red{I was just about to comment these out before finalizing the internal review draft of the paper. They are in place to remind me to contact a few people who still have not filled out their authors.csv info.}

Large-scale photometric galaxy surveys are entering a new era with currently or soon-to-be running Stage III and Stage IV dark energy experiments like the Dark Energy Survey \citep[DES,][]{Abbott:05}, the Kilo-Degree Survey \citep[KiDS,][]{de_Jong:13}, Large Synoptic Survey Telescope \citep[LSST,][]{Abell:09}, Euclid \citep{Laureijs:11}, and Wide-Field Infrared Survey Telescope \citep[WFIRST,][]{Green:12}. 
The move to imaging based surveys, rather than spectroscopic based, for cosmological measurements makes proper understanding of photometric redshifts (``photo-$z$'s'') of paramount importance, as cosmological distance measures for statistical samples are directly dependent on photo-$z$ measurements.

The unprecedented sample size of LSST galaxies, expected to number several billion for the main cosmological sample, necessitates stringent constraints on photo-$z$ accuracy if systematic errors are not to dominate the statistical errors. 
The LSST Science Requirements Document (SRD)\footnote{available at \url{https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17}} lists the photometric redshift goals for a magnitude limited sample with $i<25$ as: root-mean-square error with a goal of $\sigma_z<0.02(1+z)$; $3\sigma$ ``catastrophic outlier'' rate below $10\%$; bias below $0.003$\footnote{ 
Note that at the time the SRD was written, these goals were stated in terms of a photo-$z$ point estimate for each galaxy, as was standard in many previous studies, while in this paper we emphasize the importance of using a full photo-$z$ PDF.}.

The tremendous size of LSST's galaxy catalog will be enabled by its exceptional depth, pushing to fainter magnitudes and deeper imaging and including galaxies of lower luminosity and higher redshift than ever before. 
In addition to the contribution of low signal-to-noise photometry to the systematic error of photo-$z$s, these populations introduce major physical degeneracies, for example the Lyman break/Balmer break degeneracy, that were not present in the populations covered in previous large area surveys like the Sloan Digital Sky Survey \citep[SDSS,][]{York:00} and the Two Micron All Sky Survey \citep[2MASS,][]{Skrutskie:06}.  In order to meet LSST's demanding error budget, it will be necessary to fully characterize those degeneracies wherein multiple redshift solutions have comparable likelihood.

There is often a desire to have a single valued ``point-estimate'' redshift for an individual galaxy.  However, the complex, non-linear (and often non-unique) nature of the mapping between broad band fluxes and redshift means that a single value is unable to capture the full redshift information encoded in a galaxy's magnitudes.  For example, a common point-estimate for a template-based method is taking the highest likelihood solution as the point photo-$z$.  A single valued redshift ignores degenerate redshift solutions of lower probability, potentially biasing photometric redshift estimates both for individual galaxies and ensemble distributions.  Storing more information is necessary, most often photo-$z$ codes output the redshift probability density function (PDF), also often referred to as $p(z)$, describing the relative likelihood as a function of redshift. %\red{should we move some of AIM's p(z) vs p(z|d,I) to here?}
Early template methods such as \citet{Fernandezsoto:99} converted relative $\chi^2$ values of template spectra to likelihoods to estimate $p(z)$.  Soon after, codes such as \citet{Benitez:00} added a Bayesian prior and output a posterior probability distribution.  While many early machine learning based algorithms focused on a point-estimate, \citet{Firth:03} used a neural net with 1000 realizations scattered within the photometric errors to estimate a $p(z)$. 
As more groups began to employ photometric redshifts in their cosmological analyses, realization that point-estimate photo-$z$'s were inadequate for precision cosmology measurements \citep{Mandelbaum:2008}.  From around this point onward, most photo-$z$ algorithms have attempted to implement some estimate of the overall redshift probability in their outputs, and some surveys began supplying a full $p(z)$ rather than a simple redshift point-estimate and error \citep[e.~g.~][]{de_Jong:17}.

%\aim{We should spend a whole paragraph establishing that photo-$z$ PDFs are a thing, i.e. review the motivation for PDFs over point estimates, basic definition, early methods papers, surveys that have published them, etc.}
%\scc{I agree with Alex, we should introduce here what a pdf is and which survey have already published some pdfs (KiDS dr3 for example) or will produce them (LSST, Euclid etc.)}
%\red{See above paragraph, does this adequately address this?--SJS}\scc{fine for me}

There are numerous techniques for deriving photo-$z$ PDFs from photometry, yet no one method has yet been established as clearly superior.
Quantitative comparisons of photo-$z$ methods have been made before.  
The Photo-$z$ Accuracy And Testing \citep[PHAT,][]{Hildebrandt:10} effort focused on point estimates derived from many photometric bands. 
DES compared several codes for point estimates \citep{Sanchez:14} and a summary statistic of photo-$z$ interim posteriors for tomographically binned galaxy subsamples \citep{Bonnett:16}. 
This paper is distinguished by its inclusion of metrics of photo-$z$ interim posteriors themselves and consideration of both classic and state-of-the-art photo-$z$ algorithms, comparing the performance of several of the most widely employed codes as well as some that have been developed only recently on the basis of metrics appropriate for a probabilistic data product. 
The results presented in this work are 
%part of the Data Challenge 1 (DC1) 
 a major focus of the Photometric Redshift working group of the LSST Dark Energy Science Collaboration (LSST-DESC).  This work is laid out in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}} as one of the critical activities to be completed in preparation for dark energy science analysis on the first year LSST data. 
This is the first of multiple papers by the working group, which will grow in sophistication.
In this initial paper we focus on evaluating the performance of photometric redshift codes and their ability to produce accurate PDFs in the presence of representative training sets.  This can be thought of as an initial test under near perfect conditions, before further complexities are added in future papers.  Comparing the relative performance of the codes enables us to evaluate whether each code is using information in an optimal way, and may reveal enhancements in some codes and deficiencies in others, either in the fundamental algorithm, or in specific implementation.

Certain science cases need redshift information on individual objects, e.~g.~identification of host galaxy redshift for supernova classification, or identifying potential cluster membership.  
Other science cases need only ensemble redshift information; for instance current weak lensing techniques require the overall redshift distribution $N(z)$ for tomographic redshift samples, but do not need single galaxy estimates. 
In the case of the multiple types of probes of cosmology enabled by the
LSST cosmology sample of several billion galaxies, the number of redshift bins and their photo-$z$ requirements vary with the specific probe;  2-point angular correlations benefit from many bins, while weak lensing probes do not (due to the wide lensing kernel).
Large photometric surveys such as LSST must develop algorithms that meet the needs of all such science cases.  
In order to meet these ambitious goals for photo-$z$ accuracy, every aspect of photo-$z$ estimation will have to be optimized: the algorithms employed, both template and machine-learning based (both in design and implementation); the spectroscopic data used as a training set for machine learning algorithms or to estimate template sets and train Bayesian priors; and probabilistic catalog compression schemes that balance information retention against limited storage resources.

Before moving forward, we must address how the best methods may be unique to the performance metrics and science cases considered and what distinguishes photo-$z$ PDFs of different methods from one another.
Though photo-$z$ PDFs are often written simply as $p(z)$, the PDF itself must be an interim posterior distribution $p(z | d, I)$, the probability of redshift conditioned on photometric data $d$ that has actually been observed and the prior information $I$ that guides how a redshift is extracted from the photometry.
If we run multiple photo-$z$ codes on a single dataset, the photo-$z$ interim posteriors will not be identical because each code is based on assumptions in the form of an interim prior --- these assumptions form the premise for photo-$z$ estimation as a whole and are the only way to introduce differences in estimates of what would otherwise be a shared photo-$z$ posterior $p(z | d)$ regardless of the code used to obtain it.
Though explicit knowledge of the interim prior is necessary to use photo-$z$ interim posteriors self-consistently in physical inference, the interim prior of a particular methodology is often implicit and not necessarily shared among all galaxies in the catalog.

This paper therefore aims (1) to constrain the impact of the interim prior $I$ by separating it into a component $I_{H}$ representing the method itself and a component $I_{D}$ representing physical information, such as a training set or SED template library and (2) to present a procedure for evaluating the performance of photo-$z$ codes in generic tests that may include many more systematics in the interim prior $I$.
In order to isolate the effects encapsulated by $I_{H}$ of variation between codes from issues with the training set or template library encapsulated by $I_{D}$, we use an identical set of simulated galaxies for every code and construct a template library and training sample that are {\it complete and representative} and shared among all codes; that is, our training sample for machine learning codes is drawn from the same underlying galaxy population as our test set, with no additional selections, and the SED library used for template-based codes is the same as the one used to generate the photometric data.  
We explore a number of performance metrics in this paper, not to make a conclusion regarding the superiority or even relative favorability of each code but to establish a method for comparing photo-$z$ PDFs derived by different methods.
These test conditions set the stage for addressing in a future paper the crucial issue of incomplete and non-representative prior information.

The outline of the paper is as follows: in \S\,\ref{sec:sims} we present the simulated data set; in \S\,\ref{sec:pzcodes} we describe the current generation codes employed in the paper;  in \S\,\ref{sec:metrics} we discuss the interpretation of photo-$z$ PDFs in terms of metrics of accuracy; in \S\,\ref{sec:results} we show our results and compare the performance of the codes; in \S\,\ref{sec:discussion} we offer our conclusions and discuss future extensions of this work.

\section{The simulation and mock galaxy catalog}
\label{sec:sims}
%(Sam Schmidt, Eve Kovacs, Tina Peters)

In order to test the current generation codes, we employ a simulated galaxy catalogue. The simulation is completely catalogue-based, with no image construction or mock measurements made. We describe these in detail below.

\subsection{Buzzard-v1.0 simulation}
\label{sec:buzzard}
The \textsc{Buzzard-highres-v1.0} \red{put in cites to in prep Buzzard papers} catalogue construction started with a dark matter only simulation. This N-body simulation contained $2048^3$ particles in a $400$ Mpc h$^{-1}$ box. \red{[N]} snapshots (with smoothing and interpolation between snapshots) were saved in order to construct a lightcone. Dark matter halos were identified using the \textsc{Rockstar} software package \citep{Behroozi:13}. These dark matter halos were populated with galaxies with a stellar mass and absolute $r$-band magnitude in the SDSS system determined using a sub-halo abundance matching model constrained to match both projected two-point galaxy clustering statistics and an observed conditional stellar mass function \citep{Reddick:13}.

To assign an SED to each galaxy, the {\it Adding Density Dependent Spectral Energy Distributions} (\textsc{ADDSEDS}, deRose in prep.)\footnote{\url{https://github.com/vipasu/addseds}} procedure was used. This consisted of training an empirical relation between absolute $r$-band magnitude, local galaxy density, and SED using a sample of $\sim 5e^{5}$ galaxies from the magnitude-limited Sloan Digital Sky Survey Data Release 6 Value Added Galaxy Catalog \citep{Blanton:05}\red{[Note: is this the proper reference to SDSS-NYU VAGC? File is called combined\_dr6\_cooper.fits, but I don't see which Cooper et al 2006 this is supposed to refer to?]}. Each SDSS spectrum is fit with a sum of five SED components using the \textsc{k-correct \red{v?}} software package\footnote{\url{http://kcorrect.org}} \citep{Blanton:07}, thus each galaxy SED is parameterized as five weights for the basis SEDs. The distance to the spatial projected fifth-nearest neighbour was used as a proxy for local density in the SDSS training sample. For each simulated galaxy, a ``random'' \red{[details]} galaxy with ``similar'' \red{[details]} absolute $r$-band magnitude and local galaxy density was chosen from the training set, and that training galaxy's SED was assigned to the simulated galaxy. Given the SED, absolute $r$-band magnitude and redshift, we computed apparent magnitudes in the six LSST filter passbands, $ugrizy$. We assigned magnitude errors in the six bands using the simple model described in \citet{Ivezic:08}, assuming full 10-year depth observations had been completed.  The number of total 30-second visits assumed when generating the photometric errors differs slightly from the fiducial numbers assumed for LSST: we assume 60 visits in u-band, 80 visits in g-band, 180 visits in r-band, 180 visits in i-band, 160 visits in z-band, and 160 visits in y-band.
%numbers taken from Alex Abate's photErrorModel.py script in PhotoZDC1 repository, which was used with nYrObs=10.

\subsubsection{Selection of training and test sets}
\label{sec:buzztraining}
The total catalogue covered $400$ square degrees and contained $238$ million galaxies to an apparent magnitude limit of $r\!=\!29$ and spanning the redshift range $0\!<\!z\!\leq\!8.7$. This catalogue contained two orders of magnitude more galaxies than were needed for this study, so only $\sim\!8$ square degrees were used. Systematic problems with galaxy colors above $z\!>\!2$ were observed, so the catalogue was trimmed to include only galaxies in the redshift range $0\!<\!z\!\leq\!2.0$. A random subset of the the remaining galaxies was chosen, and placed at random into either a ``training'' set ($10$ per cent of the sample), for which the galaxies true redshifts will be supplied, or a ``test'' set (the remaining $90$ per cent of the sample), for which each code will need to predict a redshift PDF for each galaxy. The resulting catalogues contain $111\,171$ training galaxies and $1\,000\,883$ test galaxies.  We restrict our analysis to a sample with $i<25.3$, which give a signal-to-noise $\sim\,30$ for most galaxies, a cut often referred to as the expected ``LSST Gold Sample''.  This magnitude cut results in a training set with $44\,404$ galaxies and a test set containing $399\,356$ galaxies.  All subsequent results will evaluate this ``gold sample'' test set.

\subsubsection{Templates}
\label{sec:buzztemplates}
As mentioned in Section~\ref{sec:buzzard}, the SEDs in the Buzzard simulation are drawn from an empirical set of SEDs taken from the SDSS DR6 NYU-VAGC, a sample of roughly $\sim5e^{5}$ galaxies with spectra in SDSS. To determine a finite set of templates to use with template fitting codes we take the five SED weight coefficients for each of the $\sim 500\,000$ galaxies in the SDSS sample and run a simple K-means clustering algorithm on this five dimensional space. The K-means cluster centres span the space of coefficients and properly reflect the underlying density in the coefficient space, thus providing a reasonable approximation for a spanning SED set. An ad-hoc number of $K=100$ was chosen and the $100$ K-means centre positions are taken as the weights for the \textsc{k-correct} SED components to construct one hundred template SEDs. These $100$ templates were provided, 
%\red{(JS: if I remembered correctly, weren't there 150 templates given? Answer: no, initially I supplied a list of 150 that were drawn with a different algorithm, but once we switched to the final k-means set I only supplied 100 (the initial algorithm over-weighted outlier SEDs, which is why we needed the extra 50)--SJS)}
however not every template code uses this set of one hundred templates: because EAZY was designed and written to use the same five basis templates employed by \textsc{k-correct} when constructing our mock galaxies, EAZY was run using linear combinations of these five templates rather than using the 100 discrete templates.

\subsubsection{Limitations}
\label{sec:buzzlimitations}
For our initial investigation of photometric redshift codes, we begin with a data set that is somewhat idealized, and does not contain all of the complicating factors present in real data.  In several cases, the simplification is done with a purpose, with potentially confounding effects excluded in order to better isolate the differences between current-generation photo-$z$ codes, and their causes.  We list several of the simulations limitations in this section.  
As the simulation is catalogue-based, no image level effects, such as photometric measurement effects, object blending, contamination from sky background (Zodiacal light, scattered light, etc...), lensing magnification, or Galactic reddening are included.  No stars are included in the catalogue, nor are the effects of AGN.
As all SEDs are constructed from only five basis templates, properties of the galaxy population will be restricted to follow linear combinations of the characteristics of the five basis templates, so certain non-linear features, for example the full range of emission line fluxes relative to the continuum, will not be included in the model galaxy population.  No additional dust reddening intrinsic to the host galaxy is included, the only approximation of dust extinction comes in the form of dust encoded in the five basis SEDs via the training set used to create the basis templates.  Simple linear combinations of these basis templats will, once again, not explore the full range of realistic dust extinction observed in galaxy populations.

%\subsection{ABC-Galacticus simulation}
%\label{sec:galacticus}
%\red{The Argonne-Berkeley-Carnegie (ABC) Galacticus cosmological simulation uses a semi-analytic model to simulate the galaxy properties and is not directly tuned to observations. Each simulated SED is replaced with a SED drawn from the distribution of empirical SEDs given by the SED Atlas of \citet{Brown:14} (diffusion maps method in development).}

\section{Methods}
\label{sec:pzcodes}

%(Sam Schmidt, Rongpu Zhou, Ibrahim Almosallam, Eric Nuss, Johann Cohen Tanugi, John Soo)

Here we outline the photo-$z$ PDF codes tested in this study. In total, eleven distinct codes are tested.  This sample is not comprehensive, but does cover a broad range of current-generation codes.  Both template-based and machine learning approaches are included and each are described separately in Secs.~\ref{sec:templatecodes} and \ref{sec:trainingcodes} respectively. The list of codes are summarized in Table.~\ref{tab:list_of_codes}.

\begin{table*}  %%% DATA TABLE %%%
\caption{List of photo-$z$ codes featured in this study. ML here means machine learning.} \label{tab:list_of_codes}\resizebox{\textwidth}{!}{
\begin{tabular}{llll}
\hline
\bf Code 			& \bf Type & \bf Paper & \bf Website \\
\hline
\textsc{BPz} 		& template 			& \citet{Benitez:00}		& \url{http://www.stsci.edu/~dcoe/BPZ/} \\
\textsc{EAZY} 		& template 			& \citet{Brammer:08}		& \url{https://github.com/gbrammer/eazy-photoz} \\
\textsc{LePhare} 	& template 			& \citet{Arnouts:99}		& \url{http://www.cfht.hawaii.edu/~arnouts/lephare.html} \\
\textsc{ANNz2} 		& ML 	& \citet{Sadeh:16}			& \url{https://github.com/IftachSadeh/ANNZ} \\
\textsc{Delight} 	& ML/template 	& \citet{Leistedt:17}		& \url{https://github.com/ixkael/Delight} \\
\multirow{ 2}{*}{\textsc{FlexZBoost}} & \multirow{ 2}{*}{ML} 	& \multirow{ 2}{*}{\citet{Izbicki:17}}		& \url{https://github.com/tpospisi/flexcode};\\ 
					& 		&							& \url{https://github.com/rizbicki/FlexCoDE}\\
%\textsc{Frankenz} 	& ML 	& \citet{Speagle:frankenz}	& \url{https://github.com/joshspeagle/frankenz} \\
\textsc{GPz} 		& ML	& \citet{Almosallam:15b}	& \url{https://github.com/OxfordML/GPz} \\
\textsc{METAPhoR} 	& ML 	& \citet{Cavuoti:17}		& \url{http://dame.dsf.unina.it}\\
\textsc{CMNN} 		& ML 	& \citet{Graham:17}			& - \\
\textsc{SkyNet} 	& ML 	& \citet{Graff:14}			& \url{http://ccpforge.cse.rl.ac.uk/gf/project/skynet/} \\
\textsc{TPZ} 		& ML 	& \citet{Carrasco_Kind:13}	& \url{https://github.com/mgckind/MLZ} \\
\hline
\trainz 	& N/A	& See Section~\ref{sec:method:trainz} & \\
\end{tabular}}
\end{table*}

%\red{The questions that must be answered for each code are\dots}
The questions that must be answered for each code are: what unique features are included in the specific implementation that influence the output $p(z)$.  What form of validation was performed with the training data, how were photometric uncertainties employed in the analysis, how were negative fluxes treated, what specific prior form was employed (for template based codes), or what specific machine learning architcture was used (for ML codes)? 
%\scc{which parameter (in a general meaning) have been used, which kind of validation has been performed (if any)}
\subsection{Template-based Approaches}
\label{sec:templatecodes}

\subsubsection{BPZ}
\label{sec:BPZ}
%(Sam Schmidt)

\textsc{BPZ}\footnote{\url{http://www.stsci.edu/~dcoe/BPZ/}} \citep[Bayesian Photometric Redshift,][]{Benitez:00} is a template-based photo-$z$ code that compares the expected colors ($C$) calculated for a set of spectral energy distribution (SED) types/templates ($T$) to the observed colors to calculate the likelihood of observing colors at each redshift for each type, $p(C|z,T)$. The code employs an empirically determined Bayesian prior in apparent magnitude ($m_0$) and SED-type. Assuming that the SED-types are spanning and exclusive, we can determine the redshift posterior $p(z|C,m_0)$ by marginalizing over all SED-types with a simple sum \citep[Eq.~3 from][]{Benitez:00}:

\begin{equation} \label{eq:redshift_posterior}
p(z|C,m_0)\,\propto\, \sum_{T}p(z,T|m_0)\,p(C|z,T)
\end{equation}
\noindent where the first term on the right-hand side is the Bayesian prior and the second term is the traditional likelihood. The prior is assumed to have the form: $p(z,T|m_0)=p(T|m_0)\,p(z|T,m_0)$, i.e. it parameterizes the prior as an evolving type fraction with apparent magnitude, combined with a prior on the expected redshift probability distribution as a function of both apparent magnitude and SED-type.

In this paper we use \textsc{BPZ v 1.99.3}. The template set employed here is the set of $100$ discrete SEDs described in Section~\ref{sec:buzztemplates}
%, and the $129$ Brown SEDs for the Galacticus dataset. 
To keep the number of free parameters to a manageable level the SEDs in the training set are sorted by the rest-frame $u$-$g$ colour and split into three ``broad'' SED classes, equivalent to the \texttt{E}, \texttt{Sp} and \texttt{Im/SB} types in \citet{Benitez:00}. We assume the same functional form for the Bayesian priors as used by \citet{Benitez:00}, and utilize the training-set galaxies with known SED-type, redshift, and apparent magnitude to determine the type fractions and the best fit for the eleven free parameters of the prior. 
%For photo-$z$ point estimates we use the \texttt{Z\_B} output parameter. 
For galaxies with negative flux in a measured band, the placeholder value is replaced with an estimate one $\sigma$ detection limit in that particular band, i.~e.~a value close to the estimated sky noise threshold. 
The type-marginalized $p(z)$ is generated by setting the parameter \texttt{PROBS\_LITE=TRUE} in the \textsc{BPZ} parameter file.

\subsubsection{EAZY} \label{sec:eazy}
%(Rongpu Zhou)

\textsc{EAZY}\footnote{\url{https://github.com/gbrammer/eazy-photoz}} \citep[Easy and Accurate Photometric Redshifts from Yale,][]{Brammer:08} is a template-based photo-$z$ code that includes several features that improve on the basic $\chi^2$ fit used in many template codes. It can fit the observed photometry with SEDs created from a linear combination of a set of templates at each redshift, and the best-fit SED is found by simultaneously fitting one, two or all of the templates by minimizing $\chi^2$. The minimized $\chi^2(z)$ is then combined with an apparent magnitude prior to obtain the posterior redshift probability distribution, although some argue that this is not the mathematically correct way of calculating the posteriors. \textsc{EAZY} can also account for the uncertainties in the templates by adding an empirically derived template error in quadrature as a function of redshift to the flux errors.

In this paper we use the all-templates mode, which fits the photometric data with a linear combination of the five basis templates. We employed the $5$ basis templates described in Section~\ref{sec:buzzard}, and set the template error to zero since these same templates were used to produce the simulated catalog photometry. %\red{The quality parameter \texttt{Qz} is...}
The likelihoods are calculated on a 200-point redshift grid spanning $0\leq z \leq 2$, and include the application of a type-independent apparent magnitude prior estimated from the training data.

\subsubsection{LePhare}\label{sec:lephare}
%(C\'ecile Roucelle, Eric Nuss, Johann Cohen-Tanugi)

\textsc{LePhare}\footnote{\url{http://www.cfht.hawaii.edu/~arnouts/lephare.html}}\citep[Photometric Analysis for Redshift Estimate,][]{Arnouts:99,Ilbert:06} is a photo-$z$ reconstruction code based on a $\chi^2$ template-fitting procedure. The observed colors are matched with the colours predicted from a set of spectral energy distribution (SED) which can be either synthetic or based on a semi-empirical approach. \textsc{LePhare} has been used to produce the COSMOS2015 photo-$z$ catalogue \citep{Laigle:16}.

Each SED is redshifted in steps of $\Delta z = 0.01$ and convolved with the simulated LSST filter transmission curves (accounting for instrument efficiency). The opacity of the inter-galactic medium has been set to zero as no additional reddening has been included in the Buzzard simulations. The computed photo-$z$ is then the value that minimizes the merit function $\chi^2 (z,T,A)$ from \citet{Arnouts:99}:

\begin{equation}
\chi^2(z,T,A) = \sum_f^{N_f}\left(\frac{F^f_{\rm{obs}}−A \times F^f_{\rm{pred}}(T,z)}{\sigma^f_{\rm{obs}}}\right)^2 
\end{equation}
\noindent where $A$ is a normalization factor, $F^f_{\rm{pred}}(T,z)$ is the flux predicted for a template $T$ at redshift $z$. $F^f_{\rm{obs}}$ is the observed flux in a given band $f$ and $\sigma^f_{\rm{obs}}$ the associated observational error. The index $f$ refers to the considered band and $N_f$ is the total number of filters.

In this paper we use \textsc{LePhare v 2.2}. 
The set of templates used for fitting the photo-$z$'s are the 100 discrete Buzzard SED templates as described in section~\ref{sec:buzztemplates}.
%No correction for the photometric zero-points was applied for this work. 
The full $p(z)$ corresponds to the likelihoods calculated at each point on our $z$-grid.

\subsection{Training-based Codes}
\label{sec:trainingcodes}

\subsubsection{ANNz2}
\label{sec:annz2}
%(John Soo)

\textsc{ANNz2}\footnote{\url{https://github.com/IftachSadeh/ANNZ}} \citep{Sadeh:16} is a powerful package that has the ability to employ several machine learning algorithms, including artificial neural networks (ANN), boosted decision tree (BDT) and k-nearest neighbour (KNN). Using the Toolkit for Multivariate Data Analysis (TMVA) with ROOT\footnote{\url{http://tmva.sourceforge.net/}}, it can run multiple machine learning algorithms for a single training and outputs photo-$z$'s based on a weighted average of their performances.

\textsc{ANNz2} is capable of producing both photo-$z$ point estimates and redshift posterior probability distributions $p(z)$, it could also conduct classifications and supports reweighting between samples. The PDFs are produced by propagating the intrinsic uncertainty on the input parameters and the uncertainty in the machine learning method to the expected photo-$z$ solution, averaged over multiple runs weighted based on the performance of each run. \textsc{ANNz2} presents its photo-$z$ uncertainty different from many codes by using the KNN method: it estimates the photo-$z$ bias between each object and a fixed number of nearest neighbours in parameter space, it then takes the $68$th percentile width of the distribution of the bias. This is based on the implication that objects with similar photometric properties would have similar uncertainties, and therefore the photometric errors of the inputs are not propagated into the code.

In this study, \textsc{ANNz2 v. 2.0.4} was used. The full PDF for each galaxy is also produced with a linear stepsize of $z=0.01$ for $0 \leq z \leq 2$. A set of $5$ ANNs with architecture $6:12:12:1$ ($6$ $ugrizy$ inputs, $2$ hidden layers with $12$ nodes each, and $1$ output) with different random seeds are used during each training. %Training was conducted on a subset of the given training sample which satisfies $i\le25.3$ as it would have improved the results compared to training on the entire sample. 
Half of the training set is used as a validation set to prevent overtraining. All training objects are set to have detected magnitudes, however the non-detections (mag$=-99$) in the testing set are replaced with the mean of that particular band.

\subsubsection{Color-Matched Nearest-Neighbours}
\label{sec:cmnn}
%(Melissa Graham)

The nearest-neighbours color-matching photometric redshift estimator (\textsc{CMNN}) is presented in \cite[][herafter G18]{Graham:17}. This method uses a training set of galaxies with known redshifts that has equivalent or better photometry as the test set in terms of quality and filter coverage. For each galaxy in the test set we identify a color-matched subset of training galaxies, choose one (e.g. the nearest-neighbour or a random selection), and use its known redshift as the photo-$z$. This color-matched subset is identified by first calculating the Mahalanobis distance $D_M$ in color-space between the test galaxy and all training-set galaxies: the difference between the test and a training set galaxy's color divided by the photometric error, summed over all colors (i.e., $u$-$g$, $g$-$r$, $r$-$i$, $i$-$z$, and $z$-$y$). Then, the threshold value for $D_M$ that define a good color match is set by the percent point function (PPF): for example, for $N_{\rm dof}=5$, PPF$=95$ per cent of all training galaxies consistent with the test galaxy will have $D_M < 11.07$ (where $N_{\rm dof}$, the number of degrees of freedom, is the number of colors). For a given test galaxy, the $p(z)$ is the normalized distribution of the true catalogue redshifts of this color-matched subset of training galaxies, and the standard deviation of the color-matched subset is used as the photo-$z$ uncertainty.

We have applied the nearest-neighbours color-matching photometric redshift estimator described in G18 to the simulated data. Compared to its application in G18, there are some minor differences in the application of this estimator to the Buzzard catalogue. First, we do not impose non-detections on galaxies with a magnitude fainter than the expected LSST 10-year limiting magnitude or bright enough to saturate with LSST: {\it all} of the photometry for all the galaxies in the test and training sets are used for this experiment. Second, as in G18 we do apply an initial cut in color to the training set before calculating the Mahalanobis distance in order to accelerate processing, and also use a magnitude pseudo-prior to improve photo-$z$ estimates, but for both we have used different cut-off values that are appropriate for the Buzzard galaxies' colors and magnitudes. Third, we set different parameters for the identification of the color-matched subset of training galaxies and the selection of a photometric redshift estimate. In G18 we used a percent point function (PPF) value of 0.68 to identify the color-matched subset of training galaxies and used the redshift of nearest neighbour in color-space as the photo-$z$ estimate. These choices work well when the desire is to obtain accurate photo-$z$ estimates for most test-set galaxies, but does not return a robust $p(z)$ in all cases -- especially for galaxies that are bright and/or have few matches in color-space. Since a robust estimate of $p(z)$ is desired for this work we make several changes to our implementation of the \textsc{CMNN} photo-$z$ estimator. We continue to use a percent point function of ${\rm PPF} = 0.95$ to generate the subset of color-matched training galaxies, but weight them by the inverse of their Mahalanobis distance. This weighting maintains some of the accuracy that was previously achieved by simply using the nearest neighbour in color-space. We then use the weights to create the $p(z)$ instead of having the redshift of each color-matched training-set galaxy count equally. To obtain a robust estimate of the $p(z)$ for galaxies with a small number of color-matched training set galaxies, when this number is less than $20$ the nearest $20$ neighbours in color-space are used instead, and we convolve the $p(z)$ with a Gaussian with a standard deviation of:

\begin{equation}
\sigma = \sigma_{\rm train} \sqrt{({\rm PPF}_{20}/0.95)^2 -1}
\end{equation}

\noindent to appropriately broaden it so that the $p(z)$ for these test galaxies represents the enlarged PPF value associated with it. Overall, these three changes will yield poorer accuracy photo-$z$ compared to those presented in G18, but they will all have significantly more robust estimates of the $p(z)$, particularly for the brightest test galaxies. This is sufficient for this work because, as described in G18, the goal of the \textsc{CMNN} photo-$z$ estimator was never to provide the ``best'' (or even competitive) estimates in the first place, given its reliance on a deep training set, but rather to provide a means for direct comparisons between LSST photometric quality and photo-$z$ estimates. With this work we show how the input parameters should be set in order to return robust $p(z)$ estimates in addition to point value estimates.

\subsubsection{Delight}
\label{sec:delight}
%(John Soo)

\textsc{Delight}\footnote{\url{https://github.com/ixkael/Delight}} \citep{Leistedt:17} infers photo-$z$'s by using a data-driven model of latent SEDs and a physical model of photometric fluxes as a function of redshift. Generally, machine learning methods rely on representative training data with similar band passes, while template based methods rely on a complete library of templates based on physical models constructed. \textsc{Delight} is constructed in attempt to combine the advantages and eliminate the disadvantages of both template-based and machine learning algorithms: it constructs a large collection of latent SED templates (or physical flux-redshift models) from training data, with a template SED library as a guide to the learning of the model. The advantage of \textsc{Delight} is that it neither needs representative training data in the same photometric bands, nor does it need detailed galaxy SED models to work.

This conceptually novel approach is done by using Gaussian processes operating in flux-redshift space. The posterior distribution on the redshift of a target galaxy is obtained via a pairwise comparison with training galaxies,

\begin{equation}
p(z|\mathbf{\hat{F}}) \approx \sum_i p(\mathbf{\hat{F}}|z,t_i)\, p(z|t_i)p(t_i),
\end{equation}
\noindent where $p(z|t_i)p(t_i)$ captures prior information about the redshift distributions and abundances of the galaxies, with $t_i$ denoting the galaxy template; while $p(\mathbf{\hat{F}}|z,t_i)$ is the posterior of noisy flux $\mathbf{\hat{F}}$ at redshift $z$. For each training-target pair, $p(\mathbf{\hat{F}}|z,t_i)$ is evaluated as follows:

\begin{equation} \label{eq:delight_noisy}
p(\mathbf{\hat{F}}|z,t_i) = \int p(\mathbf{\hat{F}}|\mathbf{F})\, p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)\, d\mathbf{F},
\end{equation}
where $p(\mathbf{\hat{F}}|\mathbf{F})$ is the likelihood function, it compares the noisy real flux $\mathbf{\hat{F}}$ with the noiseless flux $\mathbf{F}$ obtained from the linear combination of template models, carefully constructed to account for model uncertainties and different normalization of the same SED; while $p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)$ is the prediction of flux at a different redshift $z$ with respect to the training object with redshift $z_i$ and flux $\mathbf{\hat{F}}_i$. Eq.~\ref{eq:delight_noisy} is essentially the probability that the training and the target galaxies having the same SED but at a different redshift. The flux prediction $p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)$ of the training galaxy at redshift $z$ is modeled via a Gaussian process,

\begin{equation} \label{eq:delight_gp}
F_b \sim \mathcal{GP}\left( \mu^F,k^F \right),
\end{equation}
\noindent with mean function $\mu^F$ and kernel $k^F$, both imposed to capture expected correlations resulting from the known underlying physics (i.e., fluxes resulting from observing SEDs through filter response, and the SEDs being redshifted). The reader should refer to \citet{Leistedt:17} for further details.

In this study, all $100$ ordered Buzzard templates, as described in Section~\ref{sec:buzztemplates}, were used in \textsc{Delight}, and the Gaussian process was trained with a subset of $50\,000$ galaxies. Photometric uncertainties from the inputs are propagated into the code, while non-detections for each band are set to the mean of the respective bands. Default settings of \textsc{Delight} were use, with the exception that the PDF bins were set to be linear instead of logarithmic, with $200$ equally-spaced bins between $0.0<z<2.0$. In this study a flat prior is assumed.

\subsubsection{FlexZBoost}
\label{sec:flexzboost}
%(Ann Lee, Rafael Izbicki, Taylor Pospisil, Peter Freeman)

\textsc{FlexZBoost}\footnote{\url{https://github.com/tpospisi/flexcode};\\ \url{https://github.com/rizbicki/FlexCoDE}} \citep{Izbicki:17} is a particular realization of \texttt{FlexCode}, which is a general-purpose methodology for converting any conditional mean point estimator of $z$ to a conditional {\em density} estimator $f(z \vert \x)$, where $\x$ here represents our photometric covariates and errors.\footnote{Instead of $p(z)$, we use the notation $f(z \vert \x)$ to explicitly show the dependence on $\x$.} The key idea is to expand the unknown function $f(z \vert \x)$ in an orthonormal basis $\{\phi_i(z)\}_{i}$:

\begin{equation}
f(z|\x)=\sum_{i}\beta_{i }(\x)\phi_i(z). \label{eq::series_expansion}
\end{equation}
By the orthogonality property, the expansion coefficients are just conditional means

\begin{equation}
\beta_{i }(\x) =  \mathbb{E}\left[\phi_i(z)|\x\right] \equiv \int f(z|\x)   \phi_i(z) dz. \label{eq:FlexZBoost_coeffs}
\end{equation}
These coefficients can easily be estimated from data by regression.

In this paper, we use \textsc{xgboost}~\citep{Chen:16} for the regression part as these techniques scale well for massive data; it should however be noted that \textsc{FlexCode-RF} (also on GitHub), based on Random Forests, generally performs better for smaller data sets. As our basis, we choose a standard Fourier basis. There are two tuning parameters in our $p(z)$ estimate: (i) the number of terms, $I$, in the series expansion in Eq.~\ref{eq::series_expansion}, and (ii) an exponent $\alpha$ that we use to sharpen the computed density estimates $\widehat{f}(z|\x)$, according to $\widetilde{f}(z|\x) \propto \widehat{f}(z|\x)^\alpha$. We split the ``train data'' into a training set (85\%) and a validation set (15\%), and choose both $I$ and $\alpha$ in an automated way by minimizing the weighted $L_2$-loss function \citep[Eq. 5 in][]{Izbicki:17} on the  validation set.

Although {\tt FlexCode} offers a {\em lossless compression} of the photo-$z$ estimates (in this study, one can reconstruct $\widetilde{f}(z|\x)$ exactly at any resolution from estimates of the first 35 coefficients, Eq.~\ref{eq:FlexZBoost_coeffs}, for a Fourier basis $\{\phi_i(z)\}_{i}$), we discretize our final estimates into 200 bins linearly spaced in $0 < z < 2$ for easy comparison with other algorithms. Using a higher resolution may yield better results (with no added cost in storage).

%\subsubsection{Frankenz}
%\label{sec:frankenz}
%(seeking volunteers)
%
%\red{\textsc{Frankenz}\footnote{\url{https://github.com/joshspeagle/frankenz}} \cite{Speagle:frankenz} is...}
%

\subsubsection{GPz}
\label{sec:gpz}
%(Ibrahim Almosallam)

\textsc{GPz}\footnote{\url{https://github.com/OxfordML/GPz}} \citep{Almosallam:16a,Almosallam:15b} is a sparse Gaussian process based code, a fast and a scalable approximation of full Gaussian Processes \citep{Rasmussen:06}, with the added feature of being able to produce input-dependent variance estimations (heteroscedastic noise). The model assumes that the probability of the output $y$, the redshift, given the input $x$, the photometry, is $p(y|x)=\mathcal{N}\left(y|\mu(x),\sigma(x)^{2}\right)$. The mean function, $\mu(x)$, and the variance function $\sigma(x)^{2}$ are both linear combinations of basis functions that take the following form:

\begin{align}
f(x)=\sum_{i=1}^{m}\phi_{i}(x)w_{i},
\end{align}
where $\left\{\phi_{i}(x)\right\}_{i=1}^{m}$ and $\left\{w_{i}\right\}_{i=1}^{m}$ are sets of $m$ basis functions and their associated weights respectively. Basis function models (BFM), for specific classes of basis functions such as the sigmoid or the squared exponential, have the advantage of being universal approximators, i.e. there exist a function of that form that can approximate any function, with mild assumptions, to any desired degree of accuracy. The details on how to learn the parameters of the model and the hyper-parameters of the basis functions are described in \citet{Almosallam:15b}.

A unique feature in \textsc{GPz}, is that the variance estimate is composed of two terms each quantifying a different source of uncertainty. One term (the model uncertainty) reflects how much of the uncertainty is due to lack of training samples at the location of interest, whereas the second term (the noise uncertainty) reflects how much of the uncertainty is caused from observing many noisy samples at that location. Thus, the predictive variance can determine whether we need more representative samples or more precise samples for any particular location in the input space. \textsc{GPz} can also emphasize the importance of some samples as weights. This weight can be for example $|z_{\rm{spec}}-z_{\rm{phot}}|/(1+z_{\rm{spec}})$ to target the desired objective of minimizing the normalized redshift error or as a function of their probability in the test set relative to the training set in order to pressure the model to better fit samples that are rare in the training set but are expected to be abundant during testing.

The data is prepared for \textsc{GPz} by taking the log of the magnitude errors, decorrelating the data set using PCA and imputing the missing values using a simple linear model that estimates the missing variables given the observed ones. The log transformation helps to smooth the long tail distribution of the magnitude errors, which is more stable numerically and makes the optimization process unconstrained. The missing values are imputed by computing the mean of the training set $\mu$ and its covariance $\Sigma$, then we use the following equation to estimate the missing values from the observed ones %\[

\begin{equation}
x_{u} = \mu_{u}+\Sigma_{uo}\Sigma_{oo}^{-1}(x_{o}-\mu_{o}),
\end{equation}  %\]
where the subscript $o$ in $x_{o}$ indexes the \emph{observed} part of the input $x$, whereas the subscript $u$ indexes the \emph{unobserved} set (similarly for $\mu$ and $\Sigma$). This is the optimal expected value of the unobserved variables given the observed ones if the distribution is jointly Gaussian, note that if the variables are independent, i.e. $\Sigma_{uo}=0$, this will reduce to a simple average predictor.

We use the Variable Covariance (VC) option in \textsc{GPz} with 200 basis functions after we note that there is no significant increase in the performance on the validation set (using 80\%-20\% training-validation split) and with no cost-sensitive learning applied.  We do note that \textsc{GPz} is trained using a data set that includes galaxies fainter than those in the test set, which was truncated at $i<25.3$.  This mismatch in training vs.~test results in a slightly different distribution in galaxy photometry and redshift distribution between the training and test set, which most likely degraded results compared to a run where the training set more closely matched the test set.

\subsubsection{METAPhoR}
\label{sec:metaphor}
%(Stefano Cavuoti, Massimo Brescia, Giuseppe Longo)
%\textcolor{brown}{general review done on this section}
\textsc{METAPhoR} \citep[Machine-learning Estimation Tool for Accurate Photometric Redshifts,][]{Cavuoti:17} is a pipeline designed to provide photo-$z$'s point estimates and a reliable PDF for machine learning (ML) based techniques. It includes pre- and post-processing phases, hosting a photo-$z$ prediction engine based on the Multi Layer Perceptron with Quasi Newton Algorithm (MLPQNA), already validated on photo-$z$'s in several cases \citep{de_Jong:17,Cavuoti:17b,Cavuoti:15,Brescia:14,Brescia:13,Biviano:13}. Due to its plug-in based modular nature, \textsc{METAPhoR} can be easily replaced by any other photo-$z$ prediction kernel, regardless its implementation, by taking the I/O interface compliance as unique constrain.

At a higher level, the pipeline mainly consists of three modules: (i) \textit{data pre-processing}, including a catalogue cross-matching sub-module \citep[based on the tool C3, ][]{Riccio:17}, a sub-module for photometric evaluation and error estimation of the multi-band catalogue used as Knowledge Base (KB), and a sub-module dedicated to the perturbation of the photometric KB, propaedeutic to the PDF estimation; (ii) \textit{photo-$z$ prediction}, which is the training/validation/test phase, producing the photo-$z$'s point estimates, based on a pre-selected ML method; (iii) \textit{PDF estimation}, specifically designed to calculate the PDF of the photo-$z$ estimation errors. The last module includes also a post-processing tool, providing some statistics on the produced point estimates and PDFs.

The photometry perturbation law is based on the formula $m_{ij} = m_{ij} + \alpha_{i}F_{ij}*u_{\mu=0,\sigma=1}$, where $\alpha_{i}$ is a user selected multiplicative constant (useful in case of multi-survey photometry), $u_{\mu=0,\sigma=1}$ is a random value from the standard normal distribution and $F_{ij}$ is a bimodal function (a constant function + polynomial fitting of the mean magnitude errors on the binned bands), heuristically tuned in such a way that the constant component is the threshold under which the polynomial function is considered too low to provide a significant noise contribution to the photometry perturbation.

As introduced, the photo-$z$ point estimate prediction engine of \textsc{METAPhoR} is based on the MLPQNA model, whose photo-$z$ regression training error, used by the quasi Newton learning rule, is based on the least square error and Tikhonov $L_{2}$-norm regularization \citep{Hofmann:18}.

As main prerogative, \textsc{METAPhoR} is able to provide a PDF for ML methods by taking into account the photometric errors provided with data, by running $N$ trainings on the same training set, or $M$ trainings on $M$ different random extractions from the KB. The different test sets, used to produce the PDF, are thus obtained by introducing a proper perturbation, parametrized from the photometric error distribution in each band, on the photometric data populating the original test set \citep{Brescia:18}.

For the present work since it was required to produce a redshift (and a PDF) for each object of the test set we decided to apply a hierarchical kNN to fill the missing detection, it goes without saying that for such points the reliability of PDFs and point estimation is lower. No cross validation has been used.

\subsubsection{SkyNet}
\label{sec:skynet}
%(J. Cohen-Tanugi and Hugo Tranin)

\textsc{SkyNet}\footnote{\url{http://ccpforge.cse.rl.ac.uk/gf/project/skynet/}} \citep{Graff:14} is a publicly available neural network software, based on a 2nd order conjugate gradient optimization scheme \citep[see][for further details]{Graff:14}. It has been used efficiently for redshift PDF estimates \citep{Sanchez:14,Bonnett:15,Bonnett:16}.

%In the present work, we use \textsc{SkyNet} both as a regressor (for redshift point estimates) and a classifier (for PDF estimates). 
The neural network is configured as a standard multilayer perceptron with three hidden layers and one input layer with $12$ nodes (the $6$ magnitudes and their errors). 
%For the regressor, the output layer is a single linear node, while each hidden layer has $10$ nodes with a tanh activation function. 
The classifier is laid out such that the hidden layers have 20:40:40 nodes each, all rectified linear units, and the output layer has $200$ nodes (corresponding to 200 bins for the PDF) activated with a ``softmax'' function so that they automatically sum to $1$.

To avoid over-fitting, a $30$ per cent fraction of the training set is used as validation, and the training is stopped as soon as the error rate begins to increase in the validation set. The weights are randomly initialized based on normal sampling. The error function is a standard chi-square function for the regressor, and a cross-entropy function for the classifier. Finally, the data are all whitened before processing, with magnitudes pegged to (45,45,40,35,42,42) and their errors pegged to (20,20,10,5,15,15) for $ugrizy$ filters, respectively.

\subsubsection{TPZ}
\label{sec:tpz}
%(Erfan Nourbakhsh)

\textsc{TPZ}\footnote{\url{https://github.com/mgckind/MLZ}} \citep[Trees for Photo-$z$,][]{Carrasco_Kind:13,Carrascokind:14} is a parallel machine learning algorithm that generates photometric redshift PDFs using prediction trees and random forest techniques.
The code recursively splits the input data (i.~e.~the training sample), into two branches, one after another, until a terminal leaf is created that meets a termination criterion (e.~g.~a minimum leaf size or a variance threshold). 
Bootstrap samples from the training data and associated errors are used to build a set of prediction trees.  In order to minimize correlation between the trees, the data is divided in such a way that the highest information gain among the random subsample of features is obtained at every point. The regions in each terminal leaf node corresponds to a specific subsample of the entire data that possesses similar properties.

The training data is examined before running TPZ. Since TPZ does not handle non-detections (magnitudes flagged as 99.0), we replace these values with an approximation of the $1\sigma$ detection threshold, i.~e.~a signal to noise ratio of 1 in terms of magnitude uncertainty using the equation $dm = 2.5 ~ \log ( 1 + N/S )$ where $dm \sim 0.7526 ~ mag$ for $N/S=1$.  That is, for each band, we replace the non-detection with the magnitude corresponding to the error of 0.7526 from the error model forecasted for 10-year LSST data. The Out-of-Bag \citep{Breiman:84,Carrasco_Kind:13} cross-validation technique is used within TPZ to evaluate its predictive validity and determine the relative importance of the different input attributes. We employed this information to calibrate our algorithm.

In the present work, the LSST magnitudes $u,~g,~r,~i$ and colors $u-g,~g-r,~r-i,~i-z,~z-y$ and their associated errors are used in the process of growing 100 trees with a minimum leaf size of 5 (the $z$ and $y$ magnitudes did not show significant correlation with the redshift in our cross-validation, so we did not use them when constructing our trees). We partitioned our redshift space into 100 bins from $z=0.005$ to $z=2.0$ and smoothed each individual PDF with a smoothing scale of twice the bin size.

\subsection{Simple Ensemble Estimator}\label{sec:method:trainz}
In addition to the main photo-$z$ algorithms described above we also include a very simple method.  For \trainz, as we will we call this simple estimator, we well define $p(z)$ as simply:

\begin{equation}
p(z) = \frac{1}{N_{train}}\sum_{i=1}^{N_{train}}z_{train}
\end{equation}
That is, we simply set the redshift PDF of every galaxy equal to the normalized $N(z)$ of the training sample.  As the training sample is drawn from the same underlying distribution as the test sample, modulo small deviations due to sample size, the quantiles of the training and test distributions should be identical.  This is a wildly unrealistic estimator, as it assigns all galaxies, no matter their apparent magnitude, colour, or true redshift, the same redshift PDF, and is thus uninformative at the level of individual object redshifts, but is designed to perform very well for the ensemble of all objects.  We will discuss this method and cautions relative to metrics in Section~\ref{sec:caution}.

\section{Metrics for quantifying PDF comparisons}
\label{sec:metrics}
%(Alex Malz, Rongpu Zhou, Jeff Newman, Ofer Lahav)

%\red{Maybe it would be best to introduce the summary statistics (KS, CvM, AD) up front and then the ensemble functions and stacked n(z).}

The overloaded ``$p(z)$'' is a widespread abuse of notation; we would like the outputs of photo-$z$ PDF codes to be interpretable as probabilities. 
Obviously photo-$z$ PDFs must not take negative values and must integrate to unity over the range of possible redshifts. 
Additionally, an estimator derived by method $H$ for the photo-$z$ PDF of galaxy $i$ must be understood as a posterior probability distribution

\begin{equation}
\hat{p}_{j}(z_{i}) = p(z | d_{i}, I_{D}, I_{H}),
\end{equation}
conditioned not only on the photometric data $d_{i}$ for that galaxy but also on parameters encompassing a number of things that will differ depending on the method $H$ used to produce it, namely the assumptions $I_{H}$ necessary for the method to be valid and any inputs $I_{D}$ it takes as prior information, such as a template library or training set. 
Because of this, direct comparison of photo-$z$ PDFs produced by different methods is in some sense impossible; even if they share the same prior information $I_{D}$, by definition they cannot be conditioned on the same assumptions $I_{H}$, otherwise they would not be distinct methods at all.

In this study, we isolate the differences in prior information specific to each method by using a single training set $I_{D}^{ML}$ for all machine learning-based codes and a single template library $I_{D}^{T}$ for all template-based codes, and these sets of prior information are carefully constructed to be representative and complete, we have $I_{D}^{ML}\equiv I_{D}^{T}$ for every method $H$. 
Thus, we are saying

\begin{equation}
\frac{\hat{p}_{i, H}(z)}{\hat{p}_{i, H'}(z)} \approx \frac{p(z | d_{i}, I_{H})}{p(z | d_{i}, I_{H'})},
\end{equation}
meaning that we assume comparisons of $\hat{p}_{i, H}(z)$ isolate the effect of the method used to obtain the estimator, which should make examination of differences caused by specifics of the method implementations easier to isolate.

As mentioned previously, there are cosmology probes that require knowledge of individual galaxy $p(z)$ and others that require only knowledge of the ensemble redshift distribution, $N(z)$.  
Due to the paucity of principled techniques for using and validating photo-$z$ PDFs, there are few alternatives to the common practice of reducing photo-$z$ PDFs to point estimates. 
Though this practice should not be encouraged, we also calculate traditional metrics based on the most common point estimators derived from photo-$z$ PDFs.  
Those seeking to establish a connection to traditional ways of thinking about redshift estimation may consult the Appendix for these results.

There are a number of metrics that can be used to test the accuracy of a photo-$z$ interim posterior as an estimator of a true photo-$z$ posterior if it is known.
Even for simulated data, the true photo-$z$ PDF is in general not accessible unless the redshifts are in fact drawn from the true photo-$z$ PDFs, a mock catalogue generation procedure that has not yet appeared in the literature. 
Furthermore, only limited applications of photo-$z$ PDFs that could be used as the basis for a metric have been presented in the literature. 
The most popular application by far is the calculation of the overall redshift distribution $N(z)$, the true value of which is known for the \textsc{Buzzard} simulation and will be denoted as $N'(z)$. 
Though alternatives exist \citep{Malz:chippr}, stacking according to

\begin{equation} \label{eq:stacked}
\hat{N}^{H}(z) \approx \frac{1}{N_{tot}}\ \sum_{i}^{N_{tot}}\ \hat{p}^{H}_{i}(z)
\end{equation}
is the most widely accepted method for estimating the redshift distribution from photo-$z$ PDFs. 
If we assume that the response of estimators of $N(z)$ is uniform across all approaches $H$, then we may interpret metrics on the accuracy of $\hat{N}(z)$ obtained in this way. 
We must note, however, that this is a poor assumption in general. 
Under the setup of this paper, the true redshift distribution $N'(z) = p(z | I_{D})$ (i.e. because our training data is representative, the interim prior is the truth).  In this ideal case, the method that would give the best approximation to $N'(z)$ would be one that neglects all the information contained in the photometry $\{d_{i}\}_{N_{tot}}$ and gives every galaxy the same photo-$z$ PDF $\hat{p}_{i}(z) = N'(z)$ for all $i$.  This is the exact estimator, \trainz, that we have described in Section~\ref{sec:method:trainz}, and which will serve as a point of reference for the other codes.

The exact implementation of the stacked estimator $\hat{N}^{H}(z)$ will depend on the parametrization of the photo-$z$ PDFs, which may differ across codes and can affect the precision of the estimator \citep{Malz:qp}; even considering a single method under the same parametrization, say a piecewise constant function over bins or a set of samples from the posterior, an estimator using $2N$ bins or samples will trivially be more precise than an estimator using $N$ bins or samples.
%\red{Note: Let's address how the gridding (endpoints and spacing) affects the metrics, from the email correspondence Re: [LSST-DESC-MEMBERS] Summary of Nov 29 PZ Telecon. (AIM)}
In order to minimize the effects of such choices, we asked those running all eleven codes to output $p(z)$ parameterized with a generous $\approx 200$ piecewise constant bins spanning $0<z<2$.  
The piecewise constant format is chosen because of its established presence in the literature, and the choice of 200 bins was motivated by the approximate number of columns expected to be available for storage of $p(z)$ for the final LSST Project tables.\footnote{See, e.~g.~the LSST Data Products Definition Document, available at: \url{https://ls.st/dpdd}}  
All $p(z)$ catalogues are processed using the \textsc{qp} software package \citep{Malz:qp}\footnote{available at: \url{http://github.com/aimalz/qp/}} for manipulating and calculating metrics of 1-dimensional PDFs.  
We will discuss the choice of $p(z)$ parameterization further in Section~\ref{sec:results}.

%\red{Divide metrics by application to ensemble of PDFs vs. estimators of the redshift distribution}

\subsection{Metrics of an ensemble of photo-\mathinhead{$z$}{z} interim posteriors}
\label{sec:qualmet}

\subsubsection{Probability integral transform (PIT)}
\label{sec:pit}
%(Alex Malz, Peter Freeman, Sam Schmidt)

The probability integral transform (PIT) \citep{Polsterer:16} is defined for each individual galaxy as:%essentially the derivative of the QQ plot, a histogram of

\begin{align}
\label{eq:pit}
\mathrm{PIT} &= \int_{-\infty}^{z_{\rm true}}p(z)\, dz \,.
\end{align}
The distribution of PIT values quantifies the behavior of the {\it ensemble} of photo-$z$ PDFs, enabling us to evaluate whether the $p(z)$ is, on average, accurate:  The PIT value is the Cumulative Distribution Function (CDF) of the $p(z)$ evaluated at the true redshift.  A catalogue of photo-$z$ PDFs that are accurate should have a flat PIT histogram (i.e., the individual PIT values as samples from each CDF should match a Uniform(0,1) distribution if the CDFs are accurate).  Specific deviations from flatness indicate inaccuracy: overly broad photo-$z$ PDFs would manifest as underrepresentation of the lowest and highest PIT values, whereas overly narrow photo-$z$ PDFs would manifest as over-representation of the lowest and highest PIT values. 
High frequency at only $\mathrm{PIT}\approx0$ and $\mathrm{PIT}\approx1$ indicates the presence of catastrophic outliers with highly inaccurate photo-$z$ PDFs where the true redshift is outside of the support of $p(z)$.
\citet{Tanaka:17} use the histogram of PIT values as a diagnostic indicator of overall code performance, while \citet{Freeman:17} independently define the PIT and demonstrate how its individual values may be used both to perform hypothesis testing (via, e.g., the KS, CvM, and AD tests; see below) and to construct quantile-quantile plots. 
%\red{We might want to divide this by redshift to get trends, as the overall PIT could be impacted by true n(z).  
%For example, a PIT with nontrivial shape could be due to that behavior being uniform as a function of redshift or significant inaccuracy only in a small range of redshifts where $n(z)$ is large and a flat PIT elsewhere.}

\subsubsection{Quantile-quantile (QQ) plot}
\label{sec:qq}
%(Rongpu Zhou, Jeffrey Newman, Alex Malz)

The quantile-quantile (QQ) plot is a graphical method for comparing two distributions, where the quantiles of one distribution are plotted against the quantiles of the other distribution (A quantile being defined by partitioning a distribution into consecutive intervals containing equal amounts of probability, or equal numbers of objects in each interval). 
%\blue{Include an equation to ensure quantile definition is crystal clear before building upon it for PIT.}\red{Is maybe the above text clear enough?  Text may be more clear than any equation I can come up with--SJS.}
In this paper we show the quantiles of the PIT values compared to the quantiles of the Uniform distribution that we expect the PIT values to match if $p(z)$ is an accurate probability distribution for all objects.
The QQ plot provides an easy way to qualitatively assess the differences in various properties such as the moments of an estimating distribution relative to a true distribution.  
%location, scale and skewness.  
In this paper, QQ plots are used for two purposes: (1) for comparing $N(z)$ from photo-$z$ PDFs (estimated using Eq.~\ref{eq:stacked}) with the true $N(z)$, i.e. comparing the estimated distribution of redshifts with the true redshift distribution, and (2) for assessing the overall consistency of an ensemble of photo-$z$ PDFs with their true redshifts on a population level, where the distribution of the PIT values (see previous section) is compared to a uniform distribution between $0$ and $1$.  The QQ plot contains very similar information to that shown in the PIT histogram plot, we include both forms, as visually they each convey the information in a somewhat distinct manner. %\red{is this enough reason to include both?--SJS}
%\aim{Maybe we should explain why we expect the distribution of CDFs to be uniform.}\red{Note: this is just describing the PIT.  I'm going to swap the order of PIT to come before this, and then say we are evaluating the distribution of PIT values.  I will also add why we expect this to be uniform--SJS}

\subsubsection{Conditional density estimation loss}
\label{sec:CDE_loss}
%(Alex Malz, Ann Lee)

With the conditional density estimation loss (CDE loss) we can compare how well different methods estimate individual PDFs for photometric covariates $\x$ rather than looking only at the ensemble distribution.  As in Section~\ref{sec:flexzboost}, we use the notation $f(z \vert \x)$ instead of $p(z)$ to explicitly show the dependence on $\x$.

%\aim{Perhaps this should be rewritten in the same notation used elsewhere in the paper.} \red{I agree, does someone who speaks both stats and astro want to take a shot at this?}
%\textcolor{cyan}{Ann: We are looking at individual PDFs so the dependence on $\x$ needs to be explicit; this is probably the most accurate notation as it shows both the conditioning clearly as well as whether $\x$ is fixed or random.}

The CDE loss is defined as:
\begin{equation} \label{eq:cde-loss}
L(f, \widehat{f}) = \int \int (f(z \mid \x) - \widehat{f}(z \mid \x))^{2} dz dP(\x)
\end{equation}
This loss is the CDE equivalent of the RMSE in regression. To estimate this loss we rewrite the loss as
\begin{equation} \label{eq:estimated-cde-loss}
\mathbb{E}_\X \left[\int \widehat{f}(z \mid \X)^{2} dz\right] - 2 \mathbb{E}_{\X,Z}\left[\widehat{f}(Z \mid \X)\right] + K_{f},
\end{equation}
where the first expectation is with respect to the marginal distribution of the covariates $\X$, the second expectation is with respect to the joint distribution of $\X$ and $Z$, and $K_{f}$ is a constant depending only upon the true conditional densities $f(z\mid \x)$. 
For each method we can estimate these expectations as empirical expectations on the test or validation data \citep[Eq.~7 in][]{Izbicki:17b} without knowledge of the true densities.

%\subsubsection{Continuous Ranked Probability Score (CRPS)}
%\label{sec:crps}
%\red{The CRPS is used to assess the accuracy of two probabilistic models respectively.}

\subsection{Metrics over estimated probability distributions}
\label{sec:quantmet}

In tandem with the QQ and PIT metrics introduced above, we additionally compute the following metrics comparing the empirical CDF of a distribution to the true or expected distribution.  These metrics give a more quantitative measure of the departure from ideal than the more visual PIT histogram and QQ plot.  We compute metrics comparing the CDF of PIT values to a the CDF of a Uniform distribution, and also compute the CDF of the true redshift distribution $N'(z)$ compared the $\hat{N}(z)$ distribution derived from summing the $p(z)$ as described in Eq.~\ref{eq:stacked}.

\subsubsection{Root-mean-square error (RMSE)}
\label{sec:rmse}
%(Alex Malz)

We employ the familiar root-mean-square error:

\begin{equation}
\mathrm{RMSE} = \sqrt{\int_{-\infty}^{\infty}\left(\hat{f}(z)-f'(z)\right)^{2}\,dz},
\end{equation}
%may be used for two purposes in this analysis. 
%We calculate it between the true redshift distribution $N'(z)$ and each estimator $\hat{N}(z)$ (derived by Eq.~\ref{eq:stacked}) as well as between the ideal QQ and PIT curves from each method studied and the ideal curves.  
Though this metric does not account for the fact that the redshift distribution function is, in fact, a probability distribution, it can still be interpreted as a measure of the integrated difference between the estimated distribution and the true distribution, and it can be used to quantify the otherwise qualitative metrics.

\subsubsection{Kolmogorov-Smirnov (KS) and related statistics}
\label{sec:ks}
%(John Soo, Rongpu Zhou)

The \textit{Kolmogorov-Smirnov statistic} $N_{\rm KS}$ is the maximum difference between $F_{\rm phot}(z)$ and $F_{\rm spec}(z)$, the CDFs of the photo-$z$ and spectroscopic redshift respectively:

\begin{equation}
N_{\rm KS} = \max_{z}\left( \left| F_{\rm phot}(z) - F_{\rm spec}(z) \right| \right).
\end{equation}
%The KS test quantifies whether the two redshift distributions are drawn from a similar distribution, independent of binning. 
The KS test quantifies the similarity between two distributions, independent of binning. 
A lower $N_{\rm KS}$ value corresponds to more similar distributions.
%We calculate two KS test statistics, one to test the distribution of $p(z)$ values, and one to test the stacked estimator $\hat{n}(z)$ of the redshift distribution.  
%The $p(z)$ test compares the CDF of the PIT values described in the previous section to that of a uniform distribution.  
%The $n(z)$ test compares the CDF of the distribution formed by stacking the $p(z)$ for every object in our test set against the CDF of the corresponding true redshifts.
%In this case a lower $N_{\rm KS}$ value means that the photo-$z$ distribution fits the spectroscopic redshift distribution well. $F_{\rm phot}(z)$ is calculated by stacking the PDFs of every single object in the sample, producing a smooth cumulative $n(z)$ to be compared with $F_{\rm spec}(z)$.

We also consider two variants of the KS statistic: the Cramer-von Mises (CvM) and Anderson-Darling (AD) statistics.
The CvM statistic is similar to the KS statistic as it is also computed from the distance between the measured CDF and the ideal CDF, but instead of the maximum distance, the CvM statistic calculates the average of the distance squared:

\begin{equation}
\label{eq:cvm}
\omega^2 = \int_{-\infty}^{+\infty}\big(F_{\mathrm{meas.}}(x) - F_{\mathrm{ideal}}(x)\big)^2\mathrm{d}F_{\mathrm{ideal}}
\end{equation}
% In this paper we use the CvM statistic to evaluate the performance of both $n(z)$ and $p(z)$.
The AD statistic is a weighted version of the CvM statistic, making it more sensitive to the tails of the distribution:

\begin{equation} \label{eq:ad}
A^2 = n\int_{-\infty}^{+\infty} \frac{\big(F_{\mathrm{meas.}}(x) - F_{\mathrm{ideal}}(x)\big)^2} {F_{\mathrm{ideal}}(x) (1-F_{\mathrm{ideal}}(x))}\mathrm{d}F_{\mathrm{ideal}}
\end{equation}
where $n$ is the sample size. 
%The KS statistic and the CvM and AD statistics related to it are all used to evaluate the performance of both $\hat{n}(z)$ and the QQ and PIT of the CDFs of the ensemble of photo-$z$ interim posteriors.

%\subsubsection{Kullback-Leibler Divergence (KLD)}
%\label{sec:kl}
%(Alex Malz)
%
%The KLD is a measure of the loss in information when using a probability distribution that is an estimate of a known true probability distribution, so we use 
%\begin{equation}
%\mathrm{KLD} = \int_{-\infty}^{\infty}n'(z)\log\left[\frac{n'(z)}{\hat{n}(z)}\right]dz
%\end{equation}
%as a metric on the stacked redshift distribution estimator of Eq.~\ref{eq:stacked}. Noting the assumptions made in using Eq.~\ref{eq:stacked}, we say that a photo-$z$ PDFs method whose stacked redshift distribution estimator $\hat{n}(z)$ more closely resembles the true redshift distribution $n'(z)$ is desired, i.e. we seek to minimize the KLD.

\subsubsection{Moments}
\label{sec:moments}
%(Alex Malz)
For the $\hat{N}(z)$ distributions we additionaly calculate the first three moments of the estimated redshift distribution for each code and compare them to the moments of the true redshift distribution $N'(z)$.
The $m$th moment of a distribution is defined as

\begin{equation}
\langle z^{m}\rangle = \int_{-\infty}^{\infty}z^{m}N(z)dz.
\end{equation}
Here, we use the moments of the stacked estimator of the redshift distribution function as the basis for a metric. The closer the moments of $\hat{N}(z)$ for a photo-$z$ PDF method are to the moments of the true redshift distribution function $N'(z)$, the better the photo-$z$ PDF method.

\section{Results}
\label{sec:results}
%(Sam Schmidt, Bryce Kalmbach, Johann Cohen Tanugi, Rongpu Zhou)

\subsection{Ensembles of photo-\mathinhead{$z$}{z} interim posteriors}

\begin{figure*}
%\centering
\begin{subfigure}[b]{0.49\textwidth}
\includegraphics[width=\textwidth]{figures/pz_12codes_261931_crop.jpg}
\end{subfigure}
\begin{subfigure}[b]{0.49\textwidth}
\includegraphics[width=\textwidth]{figures/pz_12codes_471167_crop.jpg}
\end{subfigure}
\begin{subfigure}[b]{0.49\textwidth}
\includegraphics[width=\textwidth]
{figures/pz_12codes_713178_crop.jpg}
\end{subfigure}
\begin{subfigure}[b]{0.49\textwidth}
\includegraphics[width=\textwidth]{figures/pz_12codes_982747_crop.jpg}
\end{subfigure}
\caption{Four illustrative examples of individual p(z) distributions produced by the codes.  The red vertical line represents the true redshift.  Examples are chosen with common features seen in PDFs: tight unimodal $p(z)$ (upper left), broad unimodal $p(z)$ (upper right), bimodal $p(z)$ (lower right), and complex/multimodal $p(z)$ (lower left).  Codes show varying amounts of small-scale structure in their reconstruction of the posterior distribution.  We see varying responses from the codes in the presence of color degeneracies and photometric errors, resulting in narrow and broad unimodal, bimodal, and multi-modal $p(z)$ curves.} \label{fig:pz_examples}
\end{figure*}

%\begin{figure*}
%\centering
%\includegraphics[width=\textwidth]{figures/QQplot_11codes.jpg}
%\caption{The QQ plots produced for every photo-$z$ code.  \aim{Consider making this a plot of QQ minus that diagonal for better contrast between methods.}\red{The PIT histogram is essentially this, which is one of the main reasons to include both.  QQ is a fairly standard form, so plotting in this fashion rather than the difference seems better to me, given that we also ahve the PIT histograms. --SJS}} \label{fig:qq}
%\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{figures/PITANDQQplot_12codes_crop.jpg}
\caption{Summary plots for all eleven photo-$z$ codes illustrating performance for the interim posterior statistics. The top panel of each pair shows both the Quantile-Quantile (QQ) plot (red) and the histogram of PIT values (blue).  The desired behavior is a QQ plot that matches the diagonal dashed line, and a PIT histogram that matches a uniform distribution matching the thin horizontal black line.  The bottom panel of each pair shows the difference between the QQ quantile and the diagonal, illustrating departure from the desired performance.  Histograms with an overabundance of PIT values at the centre of the distribution indicate $p(z)$ distributions that are overly broad, while an excess of values at the extrema indicate $p(z)$ distributions that are overly narrow.  Values of PIT=0 and PIT=1 indicate ``catastrophic failures'' where the true redshift is completely outside the support of $p(z)$.  Asymmetric features are indicative of systematic bias in the redshift predictions.  A variety of behaviors are evident, and specific details are discussed in the text.}
\label{fig:pitqq}
\end{figure*}

Fig.~\ref{fig:pz_examples} Shows the $p(z)$ produced by each of our eleven photo-$z$ codes for four example galaxies which exemplify some prominent cases that arise when estimating photo-$z$ PDFs: a narrow, unimodal redshift solution, a broader unimodal solution, a bimodal distribution, and a complex, multimodal distribution.   
The red vertical line represents the true redshift of the individual galaxy, and the blue curve represents the redshift probability.  
Several features are obvious even in these illustrative examples.  
\textsc{ANNz2, METAPhoR, NN,} and \textsc{SkyNet} all show an excess of small-scale features, which appear to be print-through of the underlying training set galaxies.  
\textsc{GPZ} (in its current implementation), on the other hand, always produces a single Gaussian, which broadens to cover the multi-modal redshift solutions seen in other codes.

As stated in Section~\ref{sec:metrics}, $p(z)$ is parameterized as $\approx 200$ piecewise constant bins covering $0<z<2$ for all eleven codes, giving a grid size of roughly $\delta z = 0.01$ for each code.  
A piecewise constant grid was a natural choice for some photo-$z$ codes, for instance most template-based codes compute likelihoods on a fixed grid.  
In contrast, FlexZBoost, for example, can return estimates on any grid without compression errors as it’s a basis expansion method where only the expansion coefficients need to be stored.
Codes with a native output format other than the shared piecewise constant binning scheme (or one that can be losslessly converted to it) may suffer from loss of information when converting to it, which could artificially favor some codes over others.

Furthermore, the fidelity of photo-$z$ interim posteriors in this format varies with the quality of the photometry.  
For faint galaxies, this redshift resolution is sufficient to capture the shape of $p(z)$ for the majority of the test sample, where photometric errors on the faint galaxies lead to somewhat broad peaks in the redshift posterior.  
However, as can be seen in e.~g.~the top left panel of Fig.~\ref{fig:pz_examples}, for bright galaxies with narrow $p(z)$ the grid spacing of $\delta z = 0.01$ is not sufficient to resolve the peak.  
This is consistent with the results described in \citet[]{Malz:qp}, who find that quantiles (and, to a lesser degree, samples) often outperform gridded $p(z)$, particularly for bright objects and in the presence of harsher storage constraints.  
With a full 200 numbers to capture the information of each photo-$z$ PDF, any parametrization will perform adequately, but other storage parametrizations and limits on storage resources may be considered in future work.  
%Switching to a quantile based parameterization may be more costly computationally, for example template-based codes would need to test more grid point in order to resolve the quantiles for bright galaxies.  However, the computational time for template based codes scales roughly linearly with the number of grid points, so this may be a reasonable thing to implement.  
We will discuss this further in Section~\ref{sec:discussion}.
% \red{someone review this statement to make sure that I'm saying this correctly!}

Fig.~\ref{fig:pitqq} shows both the quantile-quantile plots (red) and the histogram of PIT values (blue) summarizing the results from each photo-$z$ code.  The red line shows the measured quantiles, while the black diagonal represents the ideal QQ values if the distribution were perfectly reproduced.  A second panel below the main panel for each code shows the difference between $Q_{data}$ and $Q_{theory}$, i.~e.~the departure from the diagonal, for clarity.
Biases and trends in whether the average width of the $p(z)$ values being over/under-predicted are evident.  An overall bias where the predicted redshift is systematically low manifests as the measured QQ value falling above the diagonal, as is the case for \textsc{BPZ} and \textsc{EAZY}, while a systematic overprediction shows up as the measured QQ value falling below the diagonal, as seen in \textsc{TPZ}.  In terms of PIT histograms, a systematic underprediction of redshift corresponds to fewer PIT values at $PIT<0.5$ and more at $PIT>0.5$, while a systematic overprediction will show the opposite.

Examination of the PIT histograms and QQ plots shows that there are fairly generic issues with the width of $p(z)$ uncertainties: \textsc{Delight, NN, SkyNet} and \textsc{TPZ} all show a PIT histogram with an dearth of low values and and an excess of high values, signs that, on average, their $p(z)$ are more broad than the true distribution of redshifts.  \textsc{METAPhoR} shows the opposite trend, indicating the the $p(z)$  are more narrow than the distributions given by the true redshifts.  In all of these code cases there is a free parameter or bandwidth that can be used to tune uncertainties.  The sensitivity of multiple codes to this bandwidth choice emphasizes the fact that great care must be taken in setting user-defined parameters in photo-$z$ codes, even in the presence of representative training/validation data.  for \textsc{FlexZBoost} the ``sharpening'' parameter (described in Section~\ref{sec:flexzboost}) plays a key role in improving the results, resulting in a QQ plot that is very nearly diagonal.  A similar sharpening procedure could be beneficial for several codes.
Interestingly, the three purely template-based codes, \textsc{BPZ, EAZY}, and \textsc{LePhare}, show relatively well behaved $p(z)$ statistics (albeit with some bias), which may indicate that the likelihood estimation with representative templates is accurately capturing the uncertainties on individual redshifts.

%Fig.~\ref{fig:pitqq} shows the PIT histogram plots produced by each photo-$z$ code. 
%These plots show information that is very similar to that shown in the QQ plot in many ways. 
The ideal PIT histogram would follow the black dashed line, representing a uniform distribution of PIT values, equivalent to the diagonal line in the QQ plot.
Overly broad $p(z)$ values show up as an excess of PIT values near $0.5$ and a dearth of values at the edges, while overly narrow $p(z)$ will have an excess at the edges and will be missing values at the centre.
Another feature evident in the PIT histograms is the number of ``catastrophic outlier'' values where the true redshift falls outside of the non-zero support of $p(z)$, corresponding to PIT$=0.0$ or $1.0$ is more apparent than in the QQ plots.  Following Kodra \& Newman (in prep.) we define $f_{\rm O}$ as the fraction of objects with $PIT<0.0001$ or $PIT>0.9999$.  Table~\ref{tab:pitoutlier} lists these fractions for each of the codes. For a proper Uniform distribution we expect a value of $0.0002$.  Several codes show a marked excess, with \textsc{ANNz2, FlexZBoost, LePhare, and METAPhoR} with $f_{\rm O} > 0.02$, indicating a sizeable number of catastrophic redshift solutions where the true redshift is not covered by the extent of $p(z)$.  For \textsc{METAPhoR} this may be partially due to an overall underprediction of the $p(z)$ width, however this is not the case for the other codes.  \textsc{LePhare} is a particular outlier with nearly 5 per cent of objects outside of $p(z)$ support.  Further study will be necessary to determine what is causing these misclassifications for \textsc{LePhare}.  As expected, and by design, \textsc{TrainZ} has the proper fraction of outliers for the $f_{\rm O}$ statistic.

\begin{table}
\setlength{\tabcolsep}{2pt}
\centering
\caption{The fraction of ``catastrophic outlier'' PIT values.  We expect a value of 0.0002 for a proper Uniform distribution.  An excess over this small value indicates true redshifts that fall outside the non-zero support of the $p(z)$.  }\label{tab:pitoutlier}
\begin{tabular}{lc}
\hline
\hline
 ``catastrophic outlier''\\ PIT fraction\\
\hline
Photo-z Code & fraction PIT$<10^{-4}$ or $>$0.9999\\
\hline
\textsc{ANNz2} & 0.0265\\
\textsc{BPZ} & 0.0192\\
\textsc{Delight} & 0.0006\\
\textsc{EAZY} & 0.0154\\ 
\textsc{FlexZBoost} & 0.0202\\
\textsc{GPz} & 0.0068\\ 
\textsc{LePhare} & 0.0486\\
\textsc{METAPhoR}& 0.0229\\
\textsc{CMNN} & 0.0034\\ 
\textsc{Skynet} & 0.0001\\  
\textsc{TPZ} & 0.0130\\
\hline
\textsc{TrainZ} & 0.0002\\
\end{tabular}
\end{table}

%\begin{figure*}
%\centering
%\includegraphics[width=\textwidth]{figures/PIT_histogram_plot_11codes.jpg}
%\caption{The PIT histogram plots produced for every photo-$z$ code.  Ideal behaviour would match a uniform distribution, indicated by the horizontal dashed line.  Histograms with an overabundance of PIT values at the centre of the distribution indicate $p(z)$ distributions that are overly broad, while an excess of values at the extrema indicate $p(z)$ distributions that are overly narrow.  Values of PIT=0 and PIT=1 indicate ``catastrophic failures'' where the true redshift is completely outside the support of $p(z)$.  Asymmetric features are indicative of systematic bias in the redshift predictions.
%\aim{Could we combine this with the previous figure by plotting both as differences from ideal on one plot with left and right axis labels?  The information seems degenerate (which makes sense!)}\red{as you say, the information is degenerat, this form emphasizes the differences a bit more, and shows off the catastrophic outliers with PIT=0 and 1 a bit better in my mind.} } \label{fig:pit}
%\end{figure*}
\begin{figure*}
\centering
\includegraphics[width=\textwidth]{figures/KSvsCvMvsAD_PIT_withnull_jpg.jpg}
\caption{A visual representation of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the PIT distributions.  The statistics are often highly correlated, though the AD statistic truncates the extrema of the distribution and can have disparate values compared to KS and CvM.} \label{fig:pit_stats}
\end{figure*}

Fig.~\ref{fig:pit_stats} shows comparative metric values for the quantitative Kolmogorov-Smirnoff (KS), Cramer-Von Mises (CvM), and Anderson Darling (AD) test statistics for each of the codes based on comparing the distribution of their PIT values to the expected uniform distribution over the interval [0,1].  The individual values of the statistic are not as important as the comparative score between the different codes.   
%\red{Can p-values be supplied for each statistic? The statistics themselves are difficult to interpret, other than ``lower is better'' (p-value in skgof was broken, having trouble finding 1-sample KS calculation for uniform distribution)}
The AD test statistic diverges for values that include the extrema, and thus is calculated by excluding the edges of the distribution.  
We calculate the AD statistic over the range of PIT values $v=[0.01,0.99]$.  \textsc{ANNz2} and \textsc{FlexZBoost} score very well for the PIT metrics.  
\textsc{METAPhoR} and \textsc{LePhare} score very well in the PIT AD statistic, but both have a large number of catastrophic outliers, resulting in higher KS and CvM scores.

Given the near-perfect training data, examining the individual codes for explanations for departures from the expected behaviour will be instructive in avoiding similar problems in future tests. 
\textsc{ANNz2} performs quite well in $p(z)$ based metrics.  In the specific implementation employed in this paper, the final $p(z)$ is a weighted average of five neural-nets.  During the training process \textsc{ANNz2} compares the percentiles of the redshift training sample against the CDFs of the $p(z)$ sample.  Distributions that more closely match are given extra weight, and the final weights are designed to produce accurate percentiles.  Given that our metrics are focused on the percentile distributions, it is unsurprising that \textsc{ANNz2} performs well in the given metrics.  The discreteness in the individual $p(z)$ estimated by \textsc{ANNz2} can be attributed to the fact that the code was run as a classifier, assigning weights to discrete bins of redshift.  While multiple bins may receive weight, the bins themselves will still be discretized, and no additional smoothing was performed.
Overall, \textsc{FlexZBoost} and \textsc{ANNz2} show the best ensemble agreement in their distribution of PIT values.

\subsection{Metrics of the stacked estimator of the redshift distribution}\label{sec:stackedmetrics}

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{figures/NZsumplot_12codes.jpg}
\caption{The stacked $p(z)$ produced by each photo-$z$ code ($\hat{N}(z)$, red) compared to the spectroscopic redshift distribution ($N'(z)$, blue).  Varying levels of small-scale structure are seen in the codes.  $N'(z)$ is smoothed using a single bandwidth chosen via Scott's rule for all codes.} \label{fig:nz}
%\aim{Arguably this could also be clarified by showing the true $n'(z)$ once and then differences from it for each cod's stacked estimator\dots}\scc{what if we plot the difference between true and stacked rather than soverposing the two lines? Sam please do not hate me.}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{figures/KSvsCvMvsAD_NZ_withnull_jpg.jpg}
\caption{A visual representation of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the $\hat{N}(z)$ distributions. The statistics are correlated, the codes with the lowest KS statistics tend to have the lowest CvM and AD statistics.  \textsc{CMNN} performs markedly better than the others in reconstructing the overall $N(z)$ distribution, while \textsc{SkyNet} scores poorly due to an overall bias in its redshift predictions.} \label{fig:nz_stats}
\end{figure*}

Fig.~\ref{fig:nz} shows the stacked $\hat{N}(z)$ distribution compared to the true redshift distribution $N'(z)$ for all tested codes.  The red line indicates the summed $p(z)$ for each code, while the blue line shows the true redshift distribution smoothed via kernel density estimation (KDE), with a bandwidth chosen via Scott's rule \citep{Scott:1992}.  While Scott's rule is used to display $N'(z)$ in the figure, all quantitative statistics are computed via the empirical CDF, and are thus unaffected by bandwidth/smoothing choice.  
Several of the codes show an excess at $z \sim 1.4$, particularly the template-based codes \textsc{BPZ}, \textsc{EAZY}, and \textsc{LePhare}.  
This is likely due to the $4000$ angstrom break passing through the gap between the $z$ and $y$ filters.  
This feature is one of the most prominent in individual galaxy p(z), and is readily seen in the point-estimate plots shown in Fig.~\ref{fig:pz_pointestimates} and described in the Appendix.  
Several of the machine learning based codes appear to be over-trained, adding excess galaxy probability to the redshift peaks and missing probability in the troughs.  
Given that our training data is drawn from the same galaxy population as the test set, and our data has prominent peaks in $N'(z)$, perhaps it is not unexpected that such overtraining occurs. 
A more extensive training/validation set might allow for a better choice of smoothing parameters in individual codes that would avoid such overtraining.

%\aim{Move the text on Scott's Rule to Discussion?}
%Scott's rule \citep{Scott:1992} was used to determine the smoothing bandwidth of the true redshift sample: that is, a single smoothing scale was chosen to represent the true redshift sample for all eleven codes.

As with the $p(z)$ values in Figure~\ref{fig:pitqq}, different levels of substructure are obvious for the different codes.  
While Scott's rule provides a relatively good general smoothing scale to represent the true $N'(z)$, there are smaller scale fluctuations: while \textsc{FlexZBoost} and \textsc{CMNN} appear somewhat discrepant in Fig.~\ref{fig:nz}, they are actually the two most accurate in terms of their quantitative measurements.   
%\textsc{SkyNet} also shows excess small-scale structure, but also shows significant overestimate of the high redshift galaxy population.  
Interestingly, while \textsc{ANNz2} shows an abundance of small scale structure in individual $p(z)$ measurements (see Fig.~\ref{fig:pz_examples}), the summed $\hat{N}(z)$ is rather smooth, where the small scale features average out.  This is not the case for the two other codes that show an abundance of substructure in their individual $p(z)$: both \textsc{CMNN} and \textsc{SkyNet} show small scale features both in $p(z)$ and $\hat{N}(z)$.  For \textsc{CMNN} the $p(z)$ are a simply a weighted histogram of all spectroscopic training galaxies in nearby colour space with no smoothing applied, so the substructure is not unexpected.  The PIT histogram and shape of the QQ plot in Figure~\ref{fig:pitqq} show that \textsc{CMNN} is producing $p(z)$ that are overly broad, additional smoothing of the $p(z)$ would exacerbate this problem.  While the $\hat{N}(z)$ plot shows more small scale features than other codes, these features are actually representative of real structure in the true $N'(z)$, as evidenced by the very good metric scores for \textsc{CMNN}.  \textsc{SkyNet} $p(z)$ were also not smoothed: while previous implementations of the code such as \citet{Sanchez:14} and \citet{Bonnett:15} (see Appendix C.3) implement a ``sliding bin'' smoothing, no such procedure was used in this study.  In addition to excess substructure, \textsc{SkyNet} shows an obvious redshift bias, evident both visually in Figure~\ref{fig:nz} and in the first moment of N(z) listed in Table~\ref{tab:moments}, where it is clearly an outlier.  \textsc{SkyNet} employed a method where a random sample of training galaxies was chosen, but there was no test that the subset was completely representative of the overall redshift distribution.  Also unlike \citet{Bonnett:15}, no effort was made to add extra weight to more rare low and high redshift galaxies.  Either of these decisions could be the cause of the bias seen in our results.  Future runs of \textsc{SkyNet} will explore these implementation choices and their effects.

Figure~\ref{fig:nz_stats} shows the quantitative Kolmogorov-Smirnoff (KS), Cramer-Von Mises (CvM), and Anderson Darling (AD) test statistics for each of the codes for the $\hat{N}(z)$ based measures.
%\red{Can p-values be supplied for each statistic? The statistics themselves are difficult to interpret, other than ``lower is better'' (no, p-values are very difficult to compute for non-uniform distributions)}
\textsc{FlexZBoost}, \textsc{CMNN}, and \textsc{TPZ} outperform the other codes in the $\hat{N}(z)$ metrics.  
It is unsurprising that \textsc{CMNN} scores well, as with a near perfectly representative training set means that choosing neighbouring points in color/magnitude space should lead to excellent agreement in the final $\hat{N}(z)$ estimate.  \textsc{TPZ} performed quite poorly in $p(z)$ statistics, but results in a good fit to the overall $N(z)$.  This is somewhat surprising, as performance was optimized for accurate $p(z)$, not $\hat{N}(z)$.  During the validation stage for \textsc{TPZ}, there was a trade off between the width of the $p(z)$ when adjusting a smoothing parameter and overall redshift bias.  The optimal result in the PIT metrics, as illustrated in the shape of the QQ plot, does contain some level of bias as well as a slight underprediction of mean $p(z)$ width, which translates to poor metric scores.  This is something that will be looked into for \textsc{TPZ} in the future.

It is also of note that all three template-based codes show an excess in their stacked $p(z)$ at $z\,\sim\,1.3-1.4$.  This redshift range corresponds to the wavelengths where the $4000$ Angstrom break is passing between the borders of the $z$ and $y$ filters.  This strong break entering the gap between the two reddest filters can cause problems with redshift estimation of individual galaxies, as can be seen in the point-estimate photo-$z$'s shown in Figure~\ref{fig:pz_pointestimates}.  This is not unique to this dataset, it is a common occurrence in photo-$z$ estimation.  The fact that similar excesses appear in Figure~\ref{fig:nz} for \textsc{ANNz2} and \textsc{METAPhoR} shows that the effect is not limited to template-based codes.  However, the lack of such a feature in the other codes shows that it is possible to eliminate the degeneracies.  Further study on this issue may provide a solution for codes that suffer from this shortcoming.

%\red{discussion of why NN and FZBoost are doing well, how we can try to improve the other codes} 
%\textcolor{cyan}{Ann: }

Table~\ref{tab:cdeloss} shows the CDE loss statistic for each photo-$z$ code.  Once again \textsc{FlexZBoost} and \textsc{CMNN} score very well for the stacked $\hat{N}(z)$ metrics, as do \textsc{GPz} and \textsc{TPZ}.  The CDE loss measures how well individual PDFs are estimated, and codes with a low CDE loss tend to have good $\hat{N}(z)$ estimates (though the reverse is not necessarily true). \textsc{FlexZBoost} is optimized to minimize CDE loss which may explain why the method has good ensemble metrics as well. Note from Table~\ref{tab:cdeloss} that both \textsc{FlexZBoost} and \textsc{CMNN} have low CDE losses.  Empirically, we have found that PIT RMSE is not as closely correlated to CDE loss as it is to the $N(z)$ statistics.  As CDE loss is a better measure of individual redshift performance, rather than ensemble distribution performance, this statistic is a better indicator of which codes will be most likely to perform well for science cases where single objects are employed.

Table~\ref{tab:rmse} gives the root-mean-square-error (RMSE) statistics for both the PIT and N(z) estimators.  The PIT value calculates the RMSE between the quantiles shown in the QQ plot in Figure~\ref{fig:pitqq} and the diagonal, while the N(z) calculates the RMSE between the cumulative distribution of the stacked $\hat{N}(z)$ and the true redshift distribution $N'(z)$.  %\red{more about RMSE?  Or remove it? It's barely described in Section~\ref{sec:metrics} and wasn't even referred to in the text until I added this on June 29th--SJS.}

Table~\ref{tab:moments} lists the first three moments of the stacked $\hat{N}(z)$ distribution, including the moments of the ``truth'' distribution for comparison.  Several codes are able to reproduce the mean and variance of the distribution to less than a per cent, while several codes do not, which may be a cause for concern, given that mean and variance of the redshift distribution are key properties in cosmological analyses.  We note that this stated goal of the study as defined for participants was to accurately reproduce $p(z)$, the ``stacking'' of the probability distributions to estimate $\hat{N}(z)$ was not the focus as stated to the participants.  This explains why some of the best-performing empirical codes in terms of $p(z)$ measures (e.~g.~\textsc{FlexZBoost}) do not do as well at reproducing $\hat{N}(z)$ moments.  Had we defined a different parameter to optimize, in this case overall accuracy of $\hat{N}(z)$ rather than individual $p(z)$, would result in improved performance in a particular metric.  That is, optimizing photo-$z$ performance for one metric does not automatically give optimal performance for other metrics.  As previously stated, there are a variety of scientific use cases for photo-$z$'s in large upcoming surveys, and care must be taken in how the metrics used to optimize catalog photometric redshifts are defined as well as in how they are used.  In addition, very few scientific use cases will employ the overall $\hat{N}(z)$ with no cuts, as we explore in this paper.  We discuss more realistic tomographic bin selections that will be explored in a follow-up paper in Section~\ref{sec:futurework}.
%\red{[mention that we are calculating moments for the entire sample, will look at fiducial tomographic bins in a follow-up paper (unless group and reviewers realy think we should include in this paper.]}
%\red{FlexZBoost has some of the best $n(z)$ statistics, but the moments are not good.  Why is this?  Is it the over/underprediction of the high redshift part of the distribution?  Discuss this after talking with Rafael.}

%\subsection{Response of Individual Codes}
%\label{sec:res:pz_indiv_codes}
%\red{this may be incorporated in to 5.1 and 5.2 rather than live in its own section!}

%\red{HOW CODES deal with negative fluxes, and magnitude uncertainties}

%\red{overall conclusions of the Results section}

\subsection{Interpretation of metrics}\label{sec:caution}
%Caution on metrics and photo-$z$ codes}\label{sec:caution}
%(Alex Malz, Sam Schmidt)

Samples from accurate photo-$z$ posteriors should reproduce the space of $p(z, data)$. However, it is difficult to test this reconstruction given our data set, as the galaxy distributions arise from mock objects pasted on to an underlying dark matter halo catalogue with properties designed to match empirical relations, rather than being drawn from statistical distributions in redshift.  In previous sections we have mentioned that optimizing for a specific metric does not guarantee good performance on other metrics, nor is there any guarantee that good performance by our metrics corresponds to \textit{accurate} photo-$z$ posteriors.  
%\aim{Explain that samples from accurate photo-$z$ posteriors should reconstruct the space of $p(z, data)$, a test we can't check with our datasets.}  
In other words, we can construct photo-$z$ estimators that provide good coverage in many of our tests, but which have very little predictive power.

The \trainz\ estimator, which assigns every galaxy a $p(z)$ equal to $N(z)$ of the training set as described in Section~\ref{sec:method:trainz}, is introduced as a ``null test'' to demonstrate this point via \textit{reductio ad absurdum}.  
\trainz\ outperforms all codes on the PIT-based metrics, and all but one code on the $N(z)$ based statistics.  
Because our training set is perfectly representative of the test set, $N(z)$ should be identical for both sets down to statistical noise.
%\textit{Explain the connection between \trainz\ and these metrics.}

The CDE loss and point estimate metrics, however, successfully identify problems with \trainz.  
As shown in Appendix~\ref{sec:pointmetrics}, \trainz has identical $ZPEAK$ and $ZWEIGHT$ values for every galaxy, and thus the photo-$z$s are constant as a function of spec-$z$s, i.e. a horizontal line at the mode and mean of the training set distribution respectively.  The explicit dependence on the \it{individual} posteriors in the calculation of the CDE loss, described in Section~\ref{sec:CDE_loss}, distinguishes this metric from the other $p(z)$ metrics that test the overall ensemble of $p(z)$ distributions.  With a representative training set, \trainz\ will score well on the ensemble metrics, but fails miserably for metrics tied to individual redshifts.  We note that many of the ensemble-based metrics are prominent in the photo-$z$ literature despite their inability to identify problems such as those exemplified by \trainz.

%\aim{Explain why CDE loss is different from other metrics (isn't fooled by \trainz) and note that the others have more of a presence in the literature despite their flaw of failing to identify this severe problem.}

% While looking at one, or even most of our metrics would have given the impression that this estimator was nearly optimal, red flags in CDE loss and point estimates reveal the problems.  
%[more caution on over-reliance on metrics, and making sure that metrics test all of what you want] \red{this definitely still needs work}
In summary, context is crucial to interpreting metrics and defending against the likes of \trainz.
The best photo-$z$ method is the one that most effectively achieves our science goals, not the one that performs best on a metric that does not accurately reflect those goals.
In the absence of clear goals or the information necessary for a principled metric definition, we must think carefully before choosing a single metric

%\begin{table*}  %%% DATA TABLE %%%
%\caption{KS, CvM and AD statistics for each photo-$z$ code.} \label{tab:results}
%PIT statistic
%
%\begin{tabular}{lrrrr}
%\hline
%\bf Photo-$z$ Code & \bf KS & \bf CvM & \bf AD & \bf AD \\
%                   &        &         & $v=[0.05,0.95]$ & $v=[0.01,0.99]$        \\
%\hline
%\textsc{ANNz2} 		& $0.0200$ &   $52.25$ &   $583.0$ &   $759.2$ \\
%\textsc{BPZ} 		& $0.0388$ &  $280.79$ &  $1361.4$ &  $1557.5$ \\
%%\textsc{Delight} 	& $0.1380$ & $3040.59$ & $13880.2$ & $17761.4$ \\
%\textsc{Delight}    & $0.0876$ & $1075.17$ & $4407.8$  & $6167.5$\\
%%\textsc{EAZY} 		& $0.0690$ &  $994.19$ &  $3479.5$ &  $4158.8$ \\
%\textsc{EAZY}       & $0.0723$ & $1105.58$ & $3475.0$  & $4418.6$\\
%%\textsc{FlexZBoost} & $0.0260$ &   $70.58$ &   $595.0$ &   $533.0$ \\
%\textsc{FlexZBoost} & $0.0240$ &   $68.83$ &   $567.9$ &   $478.8$\\
%\textsc{GPz} 		& $0.0449$ &  $258.56$ &   $904.8$ &  $1414.8$ \\
%\textsc{LePhare} 	& $0.0663$ &  $473.05$ &    $85.9$ &   $383.8$ \\
%\textsc{METAPhoR} 	& $0.0438$ &  $298.56$ &   $248.6$ &   $715.5$ \\
%%\textsc{CMNN} 		& $-$ & $-$ & $-$ & $-$ \\
%\textsc{CMNN}         & $0.0795$  &$1011.11$ &  $3908.2$ &  $6307.5$ \\
%\textsc{SkyNet} 	& $0.0747$ &  $763.00$ &  $2740.2$ &  $4216.4$ \\
%\textsc{TPZ} 		& $0.1138$ & $1801.74$ &  $9046.2$ & $10565.7$ \\
%%\textsc{Frankenz}	& $-$ & $-$ & $-$ & $-$ \\
%\hline
%\end{tabular}

%Stacked $n(z)$ statistic
%
%\begin{tabular}{lrrrr}
%\hline
%\bf Photo-$z$ Code & \bf KS & \bf CvM & \bf AD & \bf AD \\
%                   &        &         & $z=[0.002,2.0]$ & $z=[0.0,2.0]$        \\
%\hline
%\textsc{ANNz2} 		& $0.0237$ &  $25.676$ &   $276.4$ &   $276.0$ \\
%\textsc{BPZ} 		& $0.0131$ &  $17.587$ &   $169.0$ &   $170.9$ \\
%%\textsc{Delight} 	& $0.0184$ &  $18.975$ &   $151.5$ &   $151.5$ \\
%\textsc{Delight}    & $0.0256$ &  $42.341$ &   $253.3$ &   $253.2$ \\
%%\textsc{EAZY} 		& $0.0434$ & $167.329$ &   $811.3$ &   $817.0$ \\
%\textsc{EAZY}       & $0.0441$ & $169.977$ &   $827.7$ &   $833.3$ \\
%%\textsc{FlexZBoost} & $0.0065$ &   $2.764$ &    $24.7$ &    $24.6$ \\
%\textsc{FlexZBoost} & $0.0128$ &   $7.007$ &  $127.02$ &   $126.9$\\
%\textsc{GPz} 		& $0.0168$ &  $25.605$ &   $307.1$ &   $307.1$ \\
%\textsc{LePhare} 	& $0.0254$ &  $58.900$ &   $477.4$ &   $472.0$ \\
%\textsc{METAPhoR} 	& $0.0350$ &  $60.790$ &   $671.9$ &   $672.3$ \\
%%\textsc{CMNN} 		& $-$ & $-$ & $-$ & $-$ \\
%\textsc{CMNN}         & $0.0043$ & $0.538$   &   $5.6$   &      $5.8$ \\
%\textsc{SkyNet} 	& $0.0483$ & $385.830$ &  $2119.4$ &  $2110.3$ \\
%\textsc{TPZ} 		& $0.0126$ &   $9.483$ &   $110.7$ &   $110.6$ \\
%%\textsc{Frankenz}	& $-$ & $-$ & $-$ & $-$ \\
%\hline
%\end{tabular}
%\end{table*}

\begin{table}  %%% DATA TABLE %%%
\centering
\caption{CDE loss statistic for each photo-$z$ code.} \label{tab:cdeloss}
\begin{tabular}{lr}
\hline
\bf Photo-$z$ Code & \bf CDE Loss \\
\hline
\textsc{ANNz2} 		& $-6.88$ \\
\textsc{BPZ} 		& $-7.82$ \\
%\textsc{Delight} 	& $-4.06$ \\
\textsc{Delight}    & $-8.33$\\
%\textsc{EAZY} 		& $-7.97$ \\
\textsc{EAZY}       & $-7.07$ \\
%\textsc{FlexZBoost} & $-11.51$ \\
\textsc{FlexZBoost} & $-10.60$\\
\textsc{GPz} 		& $-9.56$ \\
\textsc{LePhare} 	& $-1.66$ \\
\textsc{METAPhoR} 	& $-6.28$ \\
%\textsc{CMNN} 		& $-$ \\
\textsc{CMNN}         & $-10.43$ \\
\textsc{SkyNet} 	& $-7.89$ \\
\textsc{TPZ} 		& $-9.55$ \\
\hline
\trainz		& $-0.83$ \\
%\textsc{Frankenz}	& $-$  \\
%\hline
\end{tabular}
\end{table}

\begin{table}
\setlength{\tabcolsep}{2pt}
\begin{center}
\caption{Root-Mean-Square-Error (RMSE) statistics for the eleven photo-z codes for both PIT and $\hat{N}(z)$ distributions.}\label{tab:rmse}
\begin{tabular}{lcc}
\hline
\hline
Root-Mean-Square-Error \\
(RMSE) statistics \\
\hline
Photo-z Code & PIT RMSE & N(z) RMSE\\
\hline
\textsc{ANNz2}      & 0.019 & 0.0054\\
\textsc{BPZ}        & 0.032 & 0.0050\\
\textsc{Delight}    & 0.111 & 0.0056\\
\textsc{EAZY}       & 0.054 & 0.0102\\
\textsc{FlexZBoost} & 0.021 & 0.0022\\
\textsc{GPz}        & 0.048 & 0.0042\\
\textsc{LePhare}    & 0.028 & 0.0062\\
\textsc{METAPhoR}   & 0.064 & 0.0081\\
\textsc{CMNN}         & 0.108 & 0.0009\\
\textsc{Skynet}     & 0.054 & 0.0144\\ 
\textsc{TPZ}        & 0.082 & 0.0031\\
\hline
\trainz		& 0.0025 & 0.0013\\
\end{tabular}
\end{center} 
\end{table}

%\begin{table}
%\begin{center}
%caption{RMSE Statistics for the 11 Photo-z codes for DC1 $i<25.3$ ``gold'' sample}\label{tab:pitrmse}
%%\begin{tabular}{|l|c|c|c|c|}
%\begin{tabular}{lc}
%\hline
%\hline
% PIT RMSE statistics \\
%\hline
%Photo-z Code & PIT RMSE\\
%\hline
%\texttt{ANNz2}      & 0.019\\
%\texttt{BPZ}        & 0.032\\
%\texttt{Delight}    & 0.111\\
%\texttt{EAZY}       & 0.054\\
%\texttt{FlexZBoost} & 0.021\\
%\texttt{GPz}        & 0.048\\
%texttt{LePhare}    & 0.028\\
%\texttt{METAPhoR}   & 0.064\\
%\texttt{NN}         & 0.108 \\
%\texttt{Skynet}     & 0.054\\ 
%texttt{TPZ}        & 0.082\\
%\hline
%\end{tabular}
%\end{center} 
%\end{table}
%
%\begin{table}
%\begin{center}
%\caption{RMSE Statistics for the 11 Photo-z codes for DC1 $i<25.3$ ``gold'' sample}\label{tab:nzrmse}
%%\begin{tabular}{|l|c|c|c|c|}
%\begin{tabular}{lc}
%\hline
%\hline
% N(z) RMSE statistics \\
%\hline
%hoto-z Code & N(z) RMSE\\
%\hline
%\texttt{ANNz2}      & 0.0054\\
%\texttt{BPZ}        & 0.0050\\
%\texttt{Delight}    & 0.0056\\
%\texttt{EAZY}       & 0.0102\\
%\texttt{FlexZBoost} & 0.0022\\
%\texttt{GPz}        & 0.0042\\
%\texttt{LePhare}    & 0.0062\\
%\texttt{METAPhoR}   & 0.0081\\
%\texttt{NN}         & 0.0009\\
%\texttt{Skynet}     & 0.0144\\ 
%\texttt{TPZ}        & 0.0031\\
%\end{tabular}
%\end{center} 
%\end{table}

\begin{table}
\setlength{\tabcolsep}{2pt}
\caption{Moments of the stacked $\hat{N}(z)$ distribution}\label{tab:moments}
\begin{tabular}{lccc}
\hline
\hline
 \multicolumn{4}{l}{Stacked $n(z)$ Moments} \\
\hline
              & 1st Moment & 2nd Moment & 3rd Moment \\
\textsc{Truth}     & 0.701      &   0.630    & 0.671  \\
\hline
Photo-z Code       & 1st Moment & 2nd Moment & 3rd Moment\\
\hline
\textsc{ANNz2}     & 0.702      & 0.625      & 0.653    \\ 
\textsc{BPZ}       & 0.699      & 0.629      & 0.671    \\ 
\textsc{Delight}   & 0.692      & 0.609      & 0.638    \\ 
\textsc{EAZY}      & 0.681      & 0.595      & 0.619    \\ 
\textsc{FlexZBoost}& 0.694      & 0.610      & 0.631    \\ 
\textsc{GPz}       & 0.692      & 0.605      & 0.619    \\ 
\textsc{LePhare}   & 0.718      & 0.668      & 0.741    \\ 
\textsc{METAPhoR}  & 0.705      & 0.628      & 0.657    \\ 
\textsc{CMNN}        & 0.701      & 0.628      & 0.667    \\
\textsc{Skynet}    & 0.743      & 0.708      & 0.797    \\
\textsc{TPZ}       & 0.700      & 0.619      & 0.643    \\ 
\hline
\trainz	   & 0.699 		& 0.627 	& 0.666 \\
\end{tabular}
\end{table}

\section{Summary and Discussion}
\label{sec:discussion}
%(Sam Schmidt, Ofer Lahav, Jeff Newman, Alex Malz, Eve Kovacs, Tony Tyson, Tina Peters)
In this paper we presented results evaluating the photometric redshift PDF computation for eleven photo-$z$ codes.  As discussed in Section~\ref{sec:metrics} the $p(z)$ should accurately reflect the relative likelihood as a function of redshift for each galaxy.  All codes were provided a set of representative training data and tested on an idealized set of model galaxies with high signal-to-noise and photometry with no confounding effects due to blending, instrumental effects, the night sky, etc... included.  The goal was not to determine a ``best'' photo-$z$ code: in many ways, this was a baseline test of a ``best case scenario'' to predict the expected photo-$z$ performance if a stage IV dark energy survey was to obtain complete training samples and perfectly calibrated their multi-band photometry.  Given these idealized conditions, any deficiencies observed in a photo-$z$ code's performance should be a cause for concern, and may be evidence in a problem with either/both of the specific code implementation or the underlying algorithm.  In order to meet the stringent LSST requirements on photo-$z$ performance, identifying and correcting such problems is an important first step before tackling more realistic data in future challenges.  Most of the codes tested performed well, however, several did not meet the stringent goals that have been laid out for LSST photometric redshift performance.  This is a cause for concern, given the idealized conditions, and the individual code responses will be studied in detail moving forward.  One obvious trend in several of the codes tested was an overall over or underprediction of the widths of $p(z)$, as evidenced by the QQ plots and PIT histograms shown in Fig.~\ref{fig:pitqq}.  A more careful tuning of bandwidth or smoothing during the validation process appears to be necessary for many of the machine learning based codes in order to improve the accuracy of $p(z)$.  For narrow peaked $p(z)$ the parameterization of the PDF as evaluated on a fixed redshift grid could also have contributed to some overestimates of $p(z)$ width simply due to the finite resolution.  After evaluating resuts such as those presented in \citet[]{Malz:qp}, in future analyses we plan to switch from a fixed grid to quantile-based storage of $p(z)$ in order to more efficiently and accurately store redshift PDF results.

Another important factor to keep in mind when examining the results presented in this paper is the fact that they are at some level dependent on the metrics that we aim to optimize: in this case code participants were asked to submit their optimal measures of an accurate $p(z)$, so participants used the training/validation data to optimize their codes accordingly.  Had we, instead, asked for an optimal $\hat{N(z)}$ the resulting metrics would be different for most, if not all, of the codes, as they would optimize toward a different goal.  Specific metric choice can affect which codes are among the ``best'' codes.  As stated earlier, there are cosmological science cases that require either individual galaxy photo-$z$ measures, or ensemble $\hat{N}(z)$ measures.  We must be aware of that the optimal method for one is not necessarily optimal for the other, and in fact several photo-$z$ algorithms may be necessary in the final cosmological analysis in order to satisfy the requirements of all science use cases.  The example of the simple \trainz\ estimator described in Section~\ref{sec:caution} shows a simple model with a $p(z)$ that is unrealistic for individual objects can still score very well on many of our metrics.  It is important to look at {\it all} metrics, and keep in mind what information each metric conveys.
%\red{mention 'null' and how 'dumb' algorithms can score well on metrics, need to keep all metrics in mind}
%\aim{[reiterate the meaning of $p(z)$ and the goal -- emphasize finding ``best'' code in terms of impact of assumptions underlying the method \textit{or} establishing methodology for the realistic case of biased prior information and other science goals]}
%\red{[the challenges faced in this project]}
%\red{mention z$<$2, missing some important degeneracies, new sim will go to higher z}
%\red{no quality cuts included, could identify outliers, but could also induce biases}
%\red{[how optimization defined has impact, here we optimized p(z), not n(z), could be science-case specific if stringent requirements are needed, though that may be a computational/storage challenge if too many use cases needed.]}
%\red{some discussion of output parameterization, limitations of 200 point fixed grid, in future switch to something like quantiles (cite qp paper)}
We re-emphasize that the dataset tested was quite idealized, and discuss enhancements that will be added in future simulations to test photo-$z$ codes on increasingly realistic conditions in the following section.

\subsection{Future work}
\label{sec:futurework}
The work presented in this paper is only the first step in characterizing current photo-$z$ codes and moving toward an improved photometric redshift estimator.  This initial paper explored code performance in idealized conditions with perfect catalog-based photometry and representative training data.  As mentioned in Section~\ref{sec:stackedmetrics} for the stacked $N(z)$ metrics we examined only the entire galaxy population with no selections in either photo-$z$ ``quality" or redshift.  The cosmological analyses for weak lensing and large scale structure based measures plan to break galaxy samples into tomographic redshift bins, using photo-$z$ $p(z)$ to infer the redshift distribution for each bin.  The specific selection used to determine these bins, both algorithmically and the specific bin boundaries, could induce biases due to indirect selections inherent in the photo-$z$ or other bin selection parameters.  The effects of tomographic bin selection will be explored in a dedicated future paper. \red{[are there any references for this?  I remember Gary Bernstein talking about this at a photo-z workshop in Japan, but I don't know that it was published.  I believe Michael Troxel has discussed this as well.]}
We also plan to propagate the uncertainties measured in a set of fiducial tomographic redshift bins in order to estimate impact on cosmological parameter estimation.
%\item \red{cosmological parameters in DC2}
%\item \red{mention z$<$2, missing some important degeneracies, new sim will go to higher z}
%\red{no quality cuts included, could identify outliers, but could also induce biases}

In future papers we will add more and more complexity to our simulated data in order to test photo-$z$ algorithms in increasingly realistic conditions.  The most pressing concern is the impact of incomplete spectroscopic training samples.  As discussed extensively in \citet{Newman:2015} a representative set of spectroscopically confirmed galaxies spanning the full range of both redshift and apparent magnitude is necessary as a training set to characterize the mapping from broad-band fluxes to photometric redshifts.  However, due to a combination of factors due to both the galaxy SEDs and limitations of spectrographic instruments, redshift samples are known to be systematically incomplete, where certain galaxy types and redshift intervals fail to yield a redshift even at the longest integration times on current and near-future instruments.  The more representative the training data, the better the performance of photo-$z$ algorithms will be.  Current and upcoming surveys are putting in significant effort into obtaining these training samples \citep[e.~g.\,][]{Masters:2017}, however we still expect significant incompleteness for LSST-like samples, particularly at faint magnitudes.  One major focus of an upcoming {\it LSST Dark Energy Science Collaboration Photo-z Working Group} data challenge is to produce a realistically incomplete training set of spectroscopic galaxies, modeling the performance of spectrographs, emission-line properties, and expected signal-to-noise to determine which galaxies will fail to yield a secure redshift.  In addition to outright redshift failures we will model the inclusion of a small number of falsely identified secure redshifts where misidentified emission lines or noise spikes cause an incorrect redshift solution to be marked as a high quality identification.  Even sub-per cent level contamination by false redshifts can impact photo-$z$ solutions at levels comparable to the stringent  requirements of some LSST science cases.
We expect different systematics to occur in different photo-$z$ codes in response to training on incomplete data, particularly some of the machine learning methods.  The response of the codes will inform future directions of code development.

This initial paper explored a data set that was constructed at the catalog level, with no inclusion of the complications that come from measuring photometry from images.  Future data challenges will move to catalogs constructed from mock images, including effects that will have great impact on photo-$z$ measurements.  Object blending will be a major area of investigation, as the mixing of flux from multiple objects and the resultant change in measured colours is predicted to affect a large fraction of LSST galaxies \citep{Dawson:2016}, and will be one of the major contributing systematics for photo-$z$'s.  Inclusion of differing observing conditions (seeing, clouds, variations in filter curves, Galactic dust, ...), as well as models for instrumental and system effects, sky masks, will all impact object photometry, and will be explored in the upcoming data challenge and their impacts described in upcoming papers.
All underlying SEDs were parameterized as a weighted combination of five basis SEDs, with no additional accounting for host galaxy dust obscuration beyond what was encoded in the basis templates.  This, in effect, limited the simulation to a very simple model of internal obscuration.  Future simulations will include a more complicated and realistic treatment of host galaxy dust.

The underlying simulation used in this work was based on a light-cone constructed to a maximum redshift of $z=2$.  LSST imaging after 10 years of observations will include a significant number of $z>2$ galaxies in expected cosmology samples, and their inclusion does have potential significant implications for photo-$z$ measures: the high redshift galaxies lie at fainter apparent magnitudes and can have anomalous colours due to evolution of stellar populations and the shift to rest-frame magnitudes probing UV features of the underlying SED.  More importantly, one of the most common ``catastrophic outlier'' degeneracies observed in deep photometric samples occurs when the Lyman break is mistaken for the Balmer break, leading to multiple redshift solutions at $z\sim0.2-0.3$ and $z\sim2-3$ \citep{Massarotti:2001}.  This degeneracy, along with other potential degeneracies, are currently not covered by the limited redshift range of this initial paper, which could mean that we are not probing the full range of potential extreme outlier populations and how our photo-$z$ estimators respond to them. Extending simulations to include the high-redshift galaxy population will be a priority in future data challenges.

%\item \red{inclusion of incompleteness effects on training sample, ``wrong redshifts'', filter curves}
%\item \red{inclusion of photometric errors, image based effects, blending, lensing, sky, mask, etc...}

In this study we have not accounted for the presence of Active Galactic Nuclei (AGN) contributions to galaxy fluxes. In some cases, AGN will be easily identified from the colors and morphologies, i.e. the case of the brightest quasars where the nuclear activity outshines the host galaxy, and numerous studies have utilized color selection to create large samples of quasars \citep[e.g.][]{Richards:06,Maddox:08,Richards:15}.  In current deep fields, similar in depth to what we expect from LSST, variability information and multi-wavelength data have been critical to not only identify AGN dominated galaxies, but also obtain more accurate photometric redshifts \citep[e.g][]{Salvato:11}.

In addition to AGN dominated galaxies, those with lower levels of nuclear activity present a more insidious problem, where AGN features may not be apparent, but the colors and other host galaxy properties are perturbed relative to galaxies with an inactive nucleus.  In such cases, the presence of the AGN may induce a bias if the template SEDs or empirical datasets do not include low-level AGN counterparts. 
For LSST, we will need to identify and obtain accurate photometric redshifts of all types of AGN for a range of science goals, whether it is to eliminate such objects from cosmology experiments, or to use them with confidence, all the way through to understanding galaxy evolution and the role that AGN may play in influencing galaxy properties over cosmic time.

A promising route to classifying and obtaining accurate photometric redshifts for the AGN population is by combining machine learning with template-fitting techniques, as has recently been demonstrated by \citet{Duncan:18} for radio-selected AGN. This is because AGN are relatively easy to obtain spectroscopic redshifts for over all redshifts due to the strong emission lines that they exhibit, allowing very good training sets for machine learning algorithms to use. Whereas for those galaxies where the AGN is sub-dominant the galaxy templates are still adequate for obtaining reasonable photometric redshifts.

In addition to these improvements, the DESC Photo-$z$ group plans to look at all potential methods to combine the results from multiple photo-$z$ codes to improve $p(z)$ accuracy, similar to the work presented in \citet{Dahlen:13,Carrascokind:14,Duncan:18}.  Taking advantage of multiple algorithms that use observables in slightly different ways has shown promise, however we must be very conscious of whether a potential combination properly treats the covariance between the methods, given that they are estimating quantities based on the same underlying observables.  
Several science cases wish to estimate physical quantities along with redshift, for example galaxy stellar mass and star formation rate.  Proper joint estimation of redshift and physical quantities requires an in depth understanding of galaxy evolution, and progress on accurate bivariate redshift probability distributions will go hand in had with progress on understanding galaxies themselves.  Parameterization and storage of a complex 2-dimensional probability surface for potentially billions of galaxies (or even subsets of hundreds of thousand of particular interest) pose a potential challenge.  These issues will be examined in another future paper.

Finally, while this paper and future papers discussed above focus on photometric redshift codes and estimating accurate $p(z)$ from training data, we plan a separate, but complementary, project to examine calibration of the resultant redshifts via spatial cross-correlations \citep{Newman:2008}, which will be explored in a separate series of future papers.  The overarching plan describing everything laid out in this section is described in more detail in the LSST DESC Science Roadmap (see Footnote in Section~\ref{sec:intro}).  These plans will require significant effort, but they are necessary if we are to make optimal use of the LSST data for astrophysical and cosmological analyses.   
%\begin{enumerate}
%\item \red{the combination of codes}

%\item \red{$p(z,\alpha)$}
%\item \red{tomographic analysis later on}
%\red{this paper deals with training, future papers will also explore calibration via cross-correlation methods}
%\end{enumerate}
%\red{mention the SRM once again, say that the plans are extensive, but we do have a plan and a rough timeline.}

\section*{acknowledgements}
\red{[summary of effort breakdown], will need to be in authors.csv for auto-generation by start\_paper}

\red{[personal funding sources]}

\red{[NERSC computation acknowledgement}

The LSST Dark Energy Science Collaboration acknowledges generous ongoing support from the agencies and institutes supporting the collaboration. These include the Institut National de Physique Nucl\'eaire et de Physique des Particules in France, the Science \& Technology Facilities Council in the United Kingdom, and the Department of Energy, the National Science Foundation, and the LSST Corporation in the United States. This research used resources of the National Energy Research Scientific Computing Center, a DOE Office of Science User Facility supported by the Office of Science of the U.S. Department of Energy under Contract No. DE-AC02-05CH11231. Part of this work was undertaken on STFC DiRAC HPC Facilities, funded by UK BIS National E-infrastructure capital grants, and on the UK particle physics grid, supported by the GridPP Collaboration.

\appendix

 \section{Point Estimate Photometric Redshifts}
 \label{sec:pointmetrics}
 While we do not recommend the use of single point estimates of redshift for most science applications, plots of the point estimates can be a useful qualitative diagnostic of photo-z code performance, i.~e.~examining point photo-$z$ vs.~spec-$z$ plots visually can give a quick impression of some common trends in different codes.  Computing point estimate statistics may also be useful for more direct comparisons with previous photo-z evaluations.  If a point-estimate is preferred for a specific science case, it is fairly simple to compute the mean, mode, or some other simple estimator from each $p(z)$, so these point estimates can be easily derived from the stored $p(z)$.

There are several common point estimators of photo-$z$ posteriors employed by different codes, e.~g. the mode, mean, median of the $p(z)$ distribution.  In addition, many of the machine learning based estimators can be set up to return a single redshift solution.  For example, SkyNet can be configured to run as a regressor that returns a single float rather than a classifier that returns a 200-bin $p(z)$ estimate.  The single value returned by a machine learning based code may not correspond to a particular measure such as the mode or mean, and so to avoid interpretation of results that might be introduced by variations in choice of specific point-estimate implementation per code, we discard the code-specific point estimates. We instead calculate point estimates more uniformly across the codes directly from the $p(z)$ using two measures, $z_{PEAK}$ and $z_{WEIGHT}$.  $z_{PEAK}$  is simply the maximum value attained for each galaxy p(z), the mode of the probability distribution.  $z_{WEIGHT}$ is defined similarly to how it is defined in \citet{Dahlen:13}, as the weighted mean of the redshift over the {\it main peak} of $p(z)$ containing the $z_{PEAK}$ value.  The main peak is defined by subtracting 0.05$\,\times\,z_{PEAK}$ from $p(z)$ and identifying the roots to isolate the peak containing $z_{PEAK}$, $z_{WEIGHT}$ is defined as the weighted mean redshift within this peak.  We restrict to a single peak in order to avoid confusion from bimodal and multimodal $p(z)$ such as those shown in bottom panels of Figure~\ref{fig:pz_examples}.  For example, for a bimodal probability distribution a weighted mean calculated over both peaks would fall between the peaks, at a redshift where the probability is minimal. Restricting the weighting to a single peak ensures that the point estimate will fall in the region of maximum redshift probability. %\red{Have Dritan check zweight description to make sure this is what was done.}
% The most common point estimators of photo-$z$ PDFs are the mode, median, and mean of the distribution.  Following \cite{Tanaka:17}, we report the values of the following metrics using whichever of these point estimators performs best for each method, noting which point reduction is used in each case.

 \subsection{Point Estimate Metrics}
 \label{sec:point_metrics}
We calculate the commonly used point estimate metrics of the overall photo-$z$ scatter ($\sigma_{z}$, the standard deviation of the photo-$z$ residuals), bias, and ``catastrophic outlier rate''.  Specifically, we calculate the metrics as follows:
we define $e_{z}$ as

\begin{equation}
e_{z} = \frac{z_{P} -z_{S}}{1+z_{S}} 
\end{equation}
where $z_{P}$ is the point estimate and $z_{S}$ is the true redshift.
In practice, because the standard deviation calculation is quite sensitive to the outliers, we define the photo-$z$ scatter, $\sigma$ in terms of the Interquartile Range (IQR), the difference between the 75th and 25th percentiles of the $e_{z}$ distribution.  In order to match the usual meaning of a 1$\sigma$ interval, we scale the IQR and define $\sigma_{IQR} = IQR/1.349$, as there is a factor of 1.349 difference between the IQR and the standard deviation of a Normal distribution.
%$\sigma_{z}$ is, in practice, defined in terms the Interquartile Range (IQR) of $e_{z}$ values as $IQR/1.349$, where $IQR$ is the difference between the 75th and 25th percentiles of the $e_{z}$ distribution.  The factor of 1.349 relates the IQR of a Normal distribution to the standard 68\% range of a one $\sigma$ uncertainty.
%\scc{usually the statistical indicator defined as $\sigma$ is the standard deviation I know that is not a rule but if someone skips the test and go directly to the table (and we all know that a lot of people skips the definition of statistical indicator) could be confusing, could we report it as $\frac{IQR(e_z)}{1.349}$ or $\sigma_{IQR}$ that should be clear without any risk of confusion?}\red{Using IQR is done because IQR is less sensitive to outliers than calculating a sigma.  I've rewritten the text explaining this above.--SJS}
While many other studies define the bias based on the {\it mean} offset between true and estimated redshift, in this study we define the bias as the median value of $e_{z}$ for the sample.  We use median as it is, once again, less sensitive to outliers than the mean.  The catastrophic outlier fraction is defined as the fraction of galaxies with $e_{z}$ greater than the {\it larger} of $3\sigma_{IQR}$ or 0.06, i.e. 3$\sigma$ outliers with a floor of $\sigma_{IQR}$=0.02.
%\scc{again, is clearly not a rule but a huge number of paper refers to the mean as bias, could we use the word median, which is clear without any doubt rather than bias?}
%\red{I also have code to calculate $\sigma_{MAD}$, should we include this as well?  It's almost always within a few thousandths of $\sigma_{IQR}$, so I left it out for now}.
%\scc{we may just say that the two indicators reports almost the same value therefore there is no need for both of them?}
For reference, the goals stated in Section 3.8 of the LSST Science Book \citep{Abell:09} for photo-$z$ performance in these metrics, assuming perfect training knowledge (as we are testing in this paper) are:
\begin{itemize}
\item RMS scatter$ < 0.02(1+z)$ 
%\scc{actually the SB reports in page 75 that the goal of 0.02 is for the RMS of $\frac{\sigma_{z}}{(1+z)}$ while we are using IQR}
\item bias $<$ 0.003 %\scc{in page 518 of the SB is clearly reported that the mean is expected to be less than 0.003 in our case we have defined as bias the median and not the mean }
\item catastrophic outlier rate $<$ 10\% %\scc{clearly the number of outliers since are defined above $3\sigma$ strictly depends on the sigma definition, in the science book it seems to refer to the RMS of the $e_z$ distribution even ef it could be the standard deviation, but again I don't think it refers to IQR, therefore we could not consider exactly this numbers as reference.}
%\red{I am going off of conversations with Zeljko and how we {\it actually} computed the metrics in the Science Book (I have the scripts).  You are correct that several of these are slightly different than actually stated in the Science Book, or where the Science Book does not use very precise language as to what was done to compute the RMS or define the bias.  We can also cut out those specifics, or maybe say Ivezic private communication, maybe?  Also $\sigma_{z}$ defined in terms of IQR is exactly equivalent to the standard deviation when scaled by 1.349 for a Gaussian distribution, so the numbers are appropriate to cite in this context.  IQR is just a more robust way of calculating sigma for a distribution with outliers.--SJS }
%\scc{I am very sorry I am not saying that IQR is not fine, I want to keep the IQR my concerns are related to the notation and the \textit{reader understanding}, if I wrote $\sigma$ somewhere in a table the reader will understand for sure that is the traditional standard deviation (a lot of people skips the indicators definition for the \textit{obvious} ones), I am suggesting to replace the notation $\sigma$ with the notation $\sigma_{IQR}$ (or something like this) to avoid any confusion, and also to change from $bias$ to $median$ for the same reason. I agree that say Ivezic private communication could be better.}
\end{itemize}
These definitions are similar, but not exactly the same, as the $\sigma_{IQR}$ and median bias calculated here, but are similar enough for qualitative comparisons to the LSST goals.

%Scatter: SRD $\sigma<0.02(1+z)$
%Catastrophic Outler Rate: SRD $3\sigma$ ``catastrophic outlier'' rate below 10 per cent
%Bias: SRD bias$<0.003$

Fig.~\ref{fig:pz_pointestimates}  shows the point estimates for both $z_{PEAK}$ and $z_{WEIGHT}$.  Point density is shown with mixed contours to emphasize that most of the galaxies do fall close to the $z_{phot}=z_{spec}$ line, while blue points show differing characteristics of the outlier populations.  The red dashed lines indicated the cutoff for catastrophic outliers, defined as: $max(0.06,3\sigma_{IQR})$.  As with the full $p(z)$ results, a variety of behaviours are evident in the different codes.  Table~\ref{tab:pointestimates} lists the scatter, bias, and catastrophic outlier fractions for the codes.  The performance of the codes for point metrics is highly correlated with performance on $p(z)$ based tests, which is to be expected, given that the point-estimates were derived from the $p(z)$.  Some discretization is evident in $z_{PEAK}$, particularly for \textsc{SkyNet}, due to the finite grid spacing of the reported $p(z)$.  These discreteness effects are mitigated by the weighting of $z_{WEIGHT}$, resulting in a smoother distribution of redshift estimates.  Several features perpendicular to the main $z_{phot}=z_{spec}$ line are evident.  These features are due to the 4000 angstrom break passing through the gaps between adjacent LSST filters.  These features are most prominent in template-based codes, but appear to some degree in all codes tested.

In even the best performing codes, there are visible occupied regions away from the $z_{phot}=z_{spec}$ line, corresponding to degenerate redshift solutions for certain LSST magnitudes and colors.  While use of the full information available via $p(z)$ mitigates their impact, a full understanding of the outlier population is critical for LSST science, particularly in tomographic applications %\red{I forget what exactly I was trying to say here, if someone else wants to take a crack at this it might be helpful --SJS}
%\scc{may be worth to say that the usage of further bands could mitigate this effect by removing the degeneration  usually induced from emission lines moving across different filters?}

Finally, we note that all eleven codes are at or near the goals for point-estimates as outlined in the LSST Science Requirements Document\footnote{available at: \url{http://ls.st/srd}} and \citet{Graham:17}.  This is to be expected, given that the requirements were designed such that a point estimate photo-z would meet these requirements for perfect training data to a depth of $i<25$.  But, it is still an encouraging sign, given an updated mock galaxy simulation and the expanded set of photo-$z$ codes tested.
%\red{maybe say consistent with expectations, and a reference to Melissa's paper in here?  What else to say about nearly meeting point goals?}

\begin{figure*}
\centering
\includegraphics[width=0.9\textwidth]{figures/ZPEAK_szpz_threecolumn_12codes_navy.jpg}
\includegraphics[width=0.9\textwidth]{figures/ZWEIGHT_szpz_threecolumn_12codes_navy.jpg}
\caption{Point estimate photo-z's derived from the posteriors. Top panel shows $z_{PEAK}$, while bottom panel shows $z_{WEIGHT}$.  Point estimate density is represented with fixed density contours, while outliers at lower density are represented by blue points.  While use of point-estimate photo-$z$'s is not recommended, they do make for useful comparative and visual diagnostics.  In the lower-right panel of each plot, the \trainz\ estimator results in identical photo-$z$ estimates at the mode and mean of the training set $N'(z)$ distribution for all galaxies.} \label{fig:pz_pointestimates}
\end{figure*}

\begin{table*}
\begin{center}
\caption{Point estimate statistics}\label{tab:pointestimates}
\begin{tabular}{lcccccc}
\hline
\hline
                  &            & $Z_{PEAK}$  &          &  & $Z_{WEIGHT}$          &\\
\hline
Photo-z Code       & $\frac{\sigma_{IQR}}{(1+z)}$ & median  & \multicolumn{1}{|p{0.75cm}|}{\centering outlier \\fraction} & $\frac{\sigma_{IQR}}{(1+z)}$ & median & \multicolumn{1}{|p{0.75cm}|}{\centering outlier \\fraction}\\
\hline
\textsc{ANNz2}     & 0.0270  &  0.00063  & 0.044      & 0.0244  &  0.000307  & 0.047  \\ 
\textsc{BPZ}       & 0.0215  & -0.00175  & 0.035      & 0.0215  & -0.002005  & 0.032 \\ 
\textsc{Delight}   & 0.0212  & -0.00185  & 0.038      & 0.0216  & -0.002158  & 0.038 \\ 
\textsc{EAZY}      & 0.0225  & -0.00218  & 0.034      & 0.0226  & -0.003765  & 0.029 \\ 
\textsc{FlexZBoost}& 0.0154  & -0.00027  & 0.020      & 0.0148  & -0.000211  & 0.017 \\ 
\textsc{GPz}       & 0.0202  & -0.00091  & 0.036      & 0.0201  & -0.000950  & 0.037 \\ 
\textsc{LePhare}   & 0.0236  & -0.00161  & 0.058      & 0.0239  & -0.002007  & 0.056 \\ 
\textsc{METAPhoR}  & 0.0264  &  0.00000  & 0.037      & 0.0262  &  0.001333  & 0.048 \\ 
\textsc{CMNN}        & 0.0184  & -0.00132  & 0.035      & 0.0170  & -0.001049  & 0.034 \\
\textsc{Skynet}    & 0.0219  & -0.00167  & 0.036      & 0.0218  &  0.000174  & 0.037 \\
\textsc{TPZ}       & 0.0161  &  0.00309  & 0.033      & 0.0166  &  0.003048  & 0.031 \\ 
\hline
\trainz	   & 0.1808  &  -0.2086  & 0.000	  & 0.2335  & 0.022135  & 0.000\\
\end{tabular}
\end{center} 
\end{table*}

\bibliographystyle{mn2e}
\bibliography{references}
\end{document}

### 2018-08-19 paper/tex/appendix.tex
should now compile!

--- 2018-08-12 paper/main.tex
@@
-%\documentclass[a4paper]{article}
-\documentclass[usenatbib]{mn2e}
-%% Language and font encodings
-\usepackage[english]{babel}
-\usepackage[utf8x]{inputenc}
-\usepackage[T1]{fontenc}
-\usepackage{lineno}
-\linenumbers
-\modulolinenumbers[2]
-
-\usepackage{natbib}
-%% Sets page size and margins
-\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
-
-%% Useful packages
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{graphicx}
-\usepackage{subcaption}
-\usepackage[colorinlistoftodos]{todonotes}
-\usepackage[colorlinks=true, allcolors=blue]{hyperref}
-\usepackage{multirow}
-
-\newcommand       \apj  {ApJ}
-\newcommand       \apjl  {ApJL}
-\newcommand       \apjs  {ApJS}
-\newcommand       \aj  {AJ}
-\newcommand       \aap {A\&A}
-\newcommand       \mnras {MNRAS}
-\newcommand       \nat {Nature}
-\newcommand       \araa {ARA\&A}
-\newcommand       \physrep {Phys.~Reports}
-\newcommand       \prd {Phys. Rev. D}
-
-\newcommand{\trainz}{\textsc{trainZ}}
-
-\newcommand{\textul}{\underline}
-\newcommand*\mathinhead[2]{\texorpdfstring{$\boldsymbol{#1}$}{#2}}
-
-\newcommand{\red}[1]{\textcolor{red}{#1}}
-% Claim a color for your comments
-\newcommand{\aim}[1]{\textcolor{green}{#1}}%Alex Malz comments in green
-\newcommand{\blue}[1]{\textcolor{blue}{#1}}%unknown person comments in blue
-\definecolor{scc}{rgb}{0.0, 0.26, 0.15}
-\newcommand{\scc}[1]{\textcolor{scc}{#1}}%Stefano Cavuoti Comments in dark green
-\definecolor{newpink}{rgb}{0.858, 0.188, 0.478}
-\newcommand{\erfan}[1]{\textcolor{newpink}{#1}} % Erfan Nourbakhsh comments in pink
-\newcommand{\jan}[1]{\textcolor{orange}{#1}}%Alex Malz comments in green
-
-\def\X{{\mathbf{X}}}
-\def\x{{\mathbf{x}}}
-\def\E{{\mathbb E}}
-
-\title{An assessment of photometric redshift PDF performance in the context of LSST}
-% Alternate title proposal: Assessing the characterization of photometric redshift uncertainties in the regime of LSST
-
-\author{LSST-DESC Photometric Redshift Working Group}
-
-\begin{document}
-\maketitle
-
-\begin{abstract}
-Photometric redshift (photo-$z$) probability distribution functions (PDFs) are a key data product of nearly all upcoming galaxy imaging surveys.
-However, the photo-$z$ PDFs resulting from different techniques are not in general consistent with one another, and an optimal method for obtaining an accurate PDF remains unclear.  
-We present the results of an initial study of the Large Synoptic Survey Telescope Dark Energy Science Collaboration (LSST-DESC), the first in a planned series of papers testing multiple photo-$z$ codes on simulations of upcoming LSST galaxy photometry catalogues.
-This initial test evaluates photo-$z$ algorithms in the presence of representative training data and in the absence of several common sources of systematic errors that affect the procedures by which photo-$z$ PDFs are derived.
-The photo-$z$ PDFs are evaluated using multiple metrics including the Kolmogorov-Smirnoff statistic, Cramer-von Mises statistic, Anderson-Darling statistic, Kullback-Leibler divergence, $N(z)$ moments, quantile-quantile plots and probability integral transform.
-We observe several trends, including an overall over/under-prediction in the broadness of the PDFs for several of the codes.
-A careful accounting of all systematics discovered will be necessary for the codes employed in upcoming analyses in order to achieve unbiased cosmological measurements.
-\end{abstract}
-\begin{keywords}
-galaxies: distances and redshifts -- galaxies: statistics -- methods: statistical
-\end{keywords}
-%\noindent comments color-code:\\
-%\red{SJS},
-%\aim{AIM},
-%\scc{SC},
-%\textcolor{brown}{MB},\jan{JAN}
-%\blue{unknown}\\
-%\red{note: not all red text is mine, at some point various people were using red --SJS.}
-
-% Questions left unanswered:
-% 1. Which journal will this paper be published? 
-% Answer: AJ / ApJ / MNRAS (no charges, but not sure if it will accept a paper based on a DC1 effort and the associated simulations)
-% 2. What is the meaning of $p(z)$ in different codes? (Ofer)
-% Answer: ?
-%
-
-\section{Introduction}
-\label{sec:intro}
-
-%(Sam Schmidt, Ofer Lahav, Jeff Newman, Alex Malz, John Soo, Tony Tyson)
-
-%\jan{Is the intention to keep these parenthetical author sets?
-I find it a bit offputting in the text.
-It might be better to put this for sections in the acknowledgements section, and for individual codes/statistics specify who ran/developed it with a sentence.}\red{I was just about to comment these out before finalizing the internal review draft of the paper.
-They are in place to remind me to contact a few people who still have not filled out their authors.csv info.}
-
-Large-scale photometric galaxy surveys are entering a new era with currently or soon-to-be running Stage III and Stage IV dark energy experiments like the Dark Energy Survey \citep[DES,][]{Abbott:05}, the Kilo-Degree Survey \citep[KiDS,][]{de_Jong:13}, Large Synoptic Survey Telescope \citep[LSST,][]{Abell:09}, Euclid \citep{Laureijs:11}, and Wide-Field Infrared Survey Telescope \citep[WFIRST,][]{Green:12}.
-The move to imaging based surveys, rather than spectroscopic based, for cosmological measurements makes proper understanding of photometric redshifts (``photo-$z$'s'') of paramount importance, as cosmological distance measures for statistical samples are directly dependent on photo-$z$ measurements.
-
-The unprecedented sample size of LSST galaxies, expected to number several billion for the main cosmological sample, necessitates stringent constraints on photo-$z$ accuracy if systematic errors are not to dominate the statistical errors.
-The LSST Science Requirements Document (SRD)\footnote{available at \url{https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17}} lists the photometric redshift goals for a magnitude limited sample with $i<25$ as: root-mean-square error with a goal of $\sigma_z<0.02(1+z)$; $3\sigma$ ``catastrophic outlier'' rate below $10\%$; bias below $0.003$\footnote{
-Note that at the time the SRD was written, these goals were stated in terms of a photo-$z$ point estimate for each galaxy, as was standard in many previous studies, while in this paper we emphasize the importance of using a full photo-$z$ PDF.}.
-
-The tremendous size of LSST's galaxy catalog will be enabled by its exceptional depth, pushing to fainter magnitudes and deeper imaging and including galaxies of lower luminosity and higher redshift than ever before.
-In addition to the contribution of low signal-to-noise photometry to the systematic error of photo-$z$s, these populations introduce major physical degeneracies, for example the Lyman break/Balmer break degeneracy, that were not present in the populations covered in previous large area surveys like the Sloan Digital Sky Survey \citep[SDSS,][]{York:00} and the Two Micron All Sky Survey \citep[2MASS,][]{Skrutskie:06}.
-In order to meet LSST's demanding error budget, it will be necessary to fully characterize those degeneracies wherein multiple redshift solutions have comparable likelihood.
-
-There is often a desire to have a single valued ``point-estimate'' redshift for an individual galaxy.
-However, the complex, non-linear (and often non-unique) nature of the mapping between broad band fluxes and redshift means that a single value is unable to capture the full redshift information encoded in a galaxy's magnitudes.
-For example, a common point-estimate for a template-based method is taking the highest likelihood solution as the point photo-$z$.
-A single valued redshift ignores degenerate redshift solutions of lower probability, potentially biasing photometric redshift estimates both for individual galaxies and ensemble distributions.
-Storing more information is necessary, most often photo-$z$ codes output the redshift probability density function (PDF), also often referred to as $p(z)$, describing the relative likelihood as a function of redshift.
-%\red{should we move some of AIM's p(z) vs p(z|d,I) to here?}
-Early template methods such as \citet{Fernandezsoto:99} converted relative $\chi^2$ values of template spectra to likelihoods to estimate $p(z)$.
-Soon after, codes such as \citet{Benitez:00} added a Bayesian prior and output a posterior probability distribution.
-While many early machine learning based algorithms focused on a point-estimate, \citet{Firth:03} used a neural net with 1000 realizations scattered within the photometric errors to estimate a $p(z)$.
-As more groups began to employ photometric redshifts in their cosmological analyses, realization that point-estimate photo-$z$'s were inadequate for precision cosmology measurements \citep{Mandelbaum:2008}.
-From around this point onward, most photo-$z$ algorithms have attempted to implement some estimate of the overall redshift probability in their outputs, and some surveys began supplying a full $p(z)$ rather than a simple redshift point-estimate and error \citep[e.~g.~][]{de_Jong:17}.
-
-%\aim{We should spend a whole paragraph establishing that photo-$z$ PDFs are a thing, i.e.
-review the motivation for PDFs over point estimates, basic definition, early methods papers, surveys that have published them, etc.}
-%\scc{I agree with Alex, we should introduce here what a pdf is and which survey have already published some pdfs (KiDS dr3 for example) or will produce them (LSST, Euclid etc.)}
-%\red{See above paragraph, does this adequately address this?--SJS}\scc{fine for me}
-
-There are numerous techniques for deriving photo-$z$ PDFs from photometry, yet no one method has yet been established as clearly superior.
-Quantitative comparisons of photo-$z$ methods have been made before.  
-The Photo-$z$ Accuracy And Testing \citep[PHAT,][]{Hildebrandt:10} effort focused on point estimates derived from many photometric bands. 
-DES compared several codes for point estimates \citep{Sanchez:14} and a summary statistic of photo-$z$ interim posteriors for tomographically binned galaxy subsamples \citep{Bonnett:16}. 
-This paper is distinguished by its inclusion of metrics of photo-$z$ interim posteriors themselves and consideration of both classic and state-of-the-art photo-$z$ algorithms, comparing the performance of several of the most widely employed codes as well as some that have been developed only recently on the basis of metrics appropriate for a probabilistic data product.
-The results presented in this work are 
-%part of the Data Challenge 1 (DC1) 
- a major focus of the Photometric Redshift working group of the LSST Dark Energy Science Collaboration (LSST-DESC).
-This work is laid out in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}} as one of the critical activities to be completed in preparation for dark energy science analysis on the first year LSST data.
-This is the first of multiple papers by the working group, which will grow in sophistication.
-In this initial paper we focus on evaluating the performance of photometric redshift codes and their ability to produce accurate PDFs in the presence of representative training sets.
-This can be thought of as an initial test under near perfect conditions, before further complexities are added in future papers.
-Comparing the relative performance of the codes enables us to evaluate whether each code is using information in an optimal way, and may reveal enhancements in some codes and deficiencies in others, either in the fundamental algorithm, or in specific implementation.
-
-Certain science cases need redshift information on individual objects, e.~g.~identification of host galaxy redshift for supernova classification, or identifying potential cluster membership.  
-Other science cases need only ensemble redshift information; for instance current weak lensing techniques require the overall redshift distribution $N(z)$ for tomographic redshift samples, but do not need single galaxy estimates.
-In the case of the multiple types of probes of cosmology enabled by the
-LSST cosmology sample of several billion galaxies, the number of redshift bins and their photo-$z$ requirements vary with the specific probe;  2-point angular correlations benefit from many bins, while weak lensing probes do not (due to the wide lensing kernel).
-Large photometric surveys such as LSST must develop algorithms that meet the needs of all such science cases.  
-In order to meet these ambitious goals for photo-$z$ accuracy, every aspect of photo-$z$ estimation will have to be optimized: the algorithms employed, both template and machine-learning based (both in design and implementation); the spectroscopic data used as a training set for machine learning algorithms or to estimate template sets and train Bayesian priors; and probabilistic catalog compression schemes that balance information retention against limited storage resources.
-
-Before moving forward, we must address how the best methods may be unique to the performance metrics and science cases considered and what distinguishes photo-$z$ PDFs of different methods from one another.
-Though photo-$z$ PDFs are often written simply as $p(z)$, the PDF itself must be an interim posterior distribution $p(z | d, I)$, the probability of redshift conditioned on photometric data $d$ that has actually been observed and the prior information $I$ that guides how a redshift is extracted from the photometry.
-If we run multiple photo-$z$ codes on a single dataset, the photo-$z$ interim posteriors will not be identical because each code is based on assumptions in the form of an interim prior --- these assumptions form the premise for photo-$z$ estimation as a whole and are the only way to introduce differences in estimates of what would otherwise be a shared photo-$z$ posterior $p(z | d)$ regardless of the code used to obtain it.
-Though explicit knowledge of the interim prior is necessary to use photo-$z$ interim posteriors self-consistently in physical inference, the interim prior of a particular methodology is often implicit and not necessarily shared among all galaxies in the catalog.
-
-This paper therefore aims (1) to constrain the impact of the interim prior $I$ by separating it into a component $I_{H}$ representing the method itself and a component $I_{D}$ representing physical information, such as a training set or SED template library and (2) to present a procedure for evaluating the performance of photo-$z$ codes in generic tests that may include many more systematics in the interim prior $I$.
-In order to isolate the effects encapsulated by $I_{H}$ of variation between codes from issues with the training set or template library encapsulated by $I_{D}$, we use an identical set of simulated galaxies for every code and construct a template library and training sample that are {\it complete and representative} and shared among all codes; that is, our training sample for machine learning codes is drawn from the same underlying galaxy population as our test set, with no additional selections, and the SED library used for template-based codes is the same as the one used to generate the photometric data.
-We explore a number of performance metrics in this paper, not to make a conclusion regarding the superiority or even relative favorability of each code but to establish a method for comparing photo-$z$ PDFs derived by different methods.
-These test conditions set the stage for addressing in a future paper the crucial issue of incomplete and non-representative prior information.
-
-The outline of the paper is as follows: in \S\,\ref{sec:sims} we present the simulated data set; in \S\,\ref{sec:pzcodes} we describe the current generation codes employed in the paper;  in \S\,\ref{sec:metrics} we discuss the interpretation of photo-$z$ PDFs in terms of metrics of accuracy; in \S\,\ref{sec:results} we show our results and compare the performance of the codes; in \S\,\ref{sec:discussion} we offer our conclusions and discuss future extensions of this work.
-
-\section{The simulation and mock galaxy catalog}
-\label{sec:sims}
-%(Sam Schmidt, Eve Kovacs, Tina Peters)
-
-In order to test the current generation codes, we employ a simulated galaxy catalogue.
-The simulation is completely catalogue-based, with no image construction or mock measurements made.
-We describe these in detail below.
-
-\subsection{Buzzard-v1.0 simulation}
-\label{sec:buzzard}
-The \textsc{Buzzard-highres-v1.0} \red{put in cites to in prep Buzzard papers} catalogue construction started with a dark matter only simulation.
-This N-body simulation contained $2048^3$ particles in a $400$ Mpc h$^{-1}$ box.
-\red{[N]} snapshots (with smoothing and interpolation between snapshots) were saved in order to construct a lightcone.
-Dark matter halos were identified using the \textsc{Rockstar} software package \citep{Behroozi:13}.
-These dark matter halos were populated with galaxies with a stellar mass and absolute $r$-band magnitude in the SDSS system determined using a sub-halo abundance matching model constrained to match both projected two-point galaxy clustering statistics and an observed conditional stellar mass function \citep{Reddick:13}.
-
-To assign an SED to each galaxy, the {\it Adding Density Dependent Spectral Energy Distributions} (\textsc{ADDSEDS}, deRose in prep.)\footnote{\url{https://github.com/vipasu/addseds}} procedure was used.
-This consisted of training an empirical relation between absolute $r$-band magnitude, local galaxy density, and SED using a sample of $\sim 5e^{5}$ galaxies from the magnitude-limited Sloan Digital Sky Survey Data Release 6 Value Added Galaxy Catalog \citep{Blanton:05}\red{[Note: is this the proper reference to SDSS-NYU VAGC?
-File is called combined\_dr6\_cooper.fits, but I don't see which Cooper et al 2006 this is supposed to refer to?]}.
-Each SDSS spectrum is fit with a sum of five SED components using the \textsc{k-correct \red{v?}} software package\footnote{\url{http://kcorrect.org}} \citep{Blanton:07}, thus each galaxy SED is parameterized as five weights for the basis SEDs.
-The distance to the spatial projected fifth-nearest neighbour was used as a proxy for local density in the SDSS training sample.
-For each simulated galaxy, a ``random'' \red{[details]} galaxy with ``similar'' \red{[details]} absolute $r$-band magnitude and local galaxy density was chosen from the training set, and that training galaxy's SED was assigned to the simulated galaxy.
-Given the SED, absolute $r$-band magnitude and redshift, we computed apparent magnitudes in the six LSST filter passbands, $ugrizy$.
-We assigned magnitude errors in the six bands using the simple model described in \citet{Ivezic:08}, assuming full 10-year depth observations had been completed.
-The number of total 30-second visits assumed when generating the photometric errors differs slightly from the fiducial numbers assumed for LSST: we assume 60 visits in u-band, 80 visits in g-band, 180 visits in r-band, 180 visits in i-band, 160 visits in z-band, and 160 visits in y-band.
-%numbers taken from Alex Abate's photErrorModel.py script in PhotoZDC1 repository, which was used with nYrObs=10.
-
-\subsubsection{Selection of training and test sets}
-\label{sec:buzztraining}
-The total catalogue covered $400$ square degrees and contained $238$ million galaxies to an apparent magnitude limit of $r\!=\!29$ and spanning the redshift range $0\!<\!z\!\leq\!8.7$.
-This catalogue contained two orders of magnitude more galaxies than were needed for this study, so only $\sim\!8$ square degrees were used.
-Systematic problems with galaxy colors above $z\!>\!2$ were observed, so the catalogue was trimmed to include only galaxies in the redshift range $0\!<\!z\!\leq\!2.0$.
-A random subset of the the remaining galaxies was chosen, and placed at random into either a ``training'' set ($10$ per cent of the sample), for which the galaxies true redshifts will be supplied, or a ``test'' set (the remaining $90$ per cent of the sample), for which each code will need to predict a redshift PDF for each galaxy.
-The resulting catalogues contain $111\,171$ training galaxies and $1\,000\,883$ test galaxies.
-We restrict our analysis to a sample with $i<25.3$, which give a signal-to-noise $\sim\,30$ for most galaxies, a cut often referred to as the expected ``LSST Gold Sample''.
-This magnitude cut results in a training set with $44\,404$ galaxies and a test set containing $399\,356$ galaxies.
-All subsequent results will evaluate this ``gold sample'' test set.
-
-\subsubsection{Templates}
-\label{sec:buzztemplates}
-As mentioned in Section~\ref{sec:buzzard}, the SEDs in the Buzzard simulation are drawn from an empirical set of SEDs taken from the SDSS DR6 NYU-VAGC, a sample of roughly $\sim5e^{5}$ galaxies with spectra in SDSS.
-To determine a finite set of templates to use with template fitting codes we take the five SED weight coefficients for each of the $\sim 500\,000$ galaxies in the SDSS sample and run a simple K-means clustering algorithm on this five dimensional space.
-The K-means cluster centres span the space of coefficients and properly reflect the underlying density in the coefficient space, thus providing a reasonable approximation for a spanning SED set.
-An ad-hoc number of $K=100$ was chosen and the $100$ K-means centre positions are taken as the weights for the \textsc{k-correct} SED components to construct one hundred template SEDs.
-These $100$ templates were provided,
-%\red{(JS: if I remembered correctly, weren't there 150 templates given?
-Answer: no, initially I supplied a list of 150 that were drawn with a different algorithm, but once we switched to the final k-means set I only supplied 100 (the initial algorithm over-weighted outlier SEDs, which is why we needed the extra 50)--SJS)}
-however not every template code uses this set of one hundred templates: because EAZY was designed and written to use the same five basis templates employed by \textsc{k-correct} when constructing our mock galaxies, EAZY was run using linear combinations of these five templates rather than using the 100 discrete templates.
-
-\subsubsection{Limitations}
-\label{sec:buzzlimitations}
-For our initial investigation of photometric redshift codes, we begin with a data set that is somewhat idealized, and does not contain all of the complicating factors present in real data.
-In several cases, the simplification is done with a purpose, with potentially confounding effects excluded in order to better isolate the differences between current-generation photo-$z$ codes, and their causes.
-We list several of the simulations limitations in this section.
-As the simulation is catalogue-based, no image level effects, such as photometric measurement effects, object blending, contamination from sky background (Zodiacal light, scattered light, etc...), lensing magnification, or Galactic reddening are included.
-No stars are included in the catalogue, nor are the effects of AGN.
-As all SEDs are constructed from only five basis templates, properties of the galaxy population will be restricted to follow linear combinations of the characteristics of the five basis templates, so certain non-linear features, for example the full range of emission line fluxes relative to the continuum, will not be included in the model galaxy population.
-No additional dust reddening intrinsic to the host galaxy is included, the only approximation of dust extinction comes in the form of dust encoded in the five basis SEDs via the training set used to create the basis templates.
-Simple linear combinations of these basis templats will, once again, not explore the full range of realistic dust extinction observed in galaxy populations.
-
-%\subsection{ABC-Galacticus simulation}
-%\label{sec:galacticus}
-%\red{The Argonne-Berkeley-Carnegie (ABC) Galacticus cosmological simulation uses a semi-analytic model to simulate the galaxy properties and is not directly tuned to observations.
-Each simulated SED is replaced with a SED drawn from the distribution of empirical SEDs given by the SED Atlas of \citet{Brown:14} (diffusion maps method in development).}
-
-\section{Methods}
-\label{sec:pzcodes}
-
-%(Sam Schmidt, Rongpu Zhou, Ibrahim Almosallam, Eric Nuss, Johann Cohen Tanugi, John Soo)
-
-Here we outline the photo-$z$ PDF codes tested in this study.
-In total, eleven distinct codes are tested.
-This sample is not comprehensive, but does cover a broad range of current-generation codes.
-Both template-based and machine learning approaches are included and each are described separately in Secs.~\ref{sec:templatecodes} and \ref{sec:trainingcodes} respectively.
-The list of codes are summarized in Table.~\ref{tab:list_of_codes}.
-
-\begin{table*}  %%% DATA TABLE %%%
-\caption{List of photo-$z$ codes featured in this study. ML here means machine learning.} \label{tab:list_of_codes}\resizebox{\textwidth}{!}{
-\begin{tabular}{llll}
-\hline
-\bf Code 			& \bf Type & \bf Paper & \bf Website \\
-\hline
-\textsc{BPz} 		& template 			& \citet{Benitez:00}		& \url{http://www.stsci.edu/~dcoe/BPZ/} \\
-\textsc{EAZY} 		& template 			& \citet{Brammer:08}		& \url{https://github.com/gbrammer/eazy-photoz} \\
-\textsc{LePhare} 	& template 			& \citet{Arnouts:99}		& \url{http://www.cfht.hawaii.edu/~arnouts/lephare.html} \\
-\textsc{ANNz2} 		& ML 	& \citet{Sadeh:16}			& \url{https://github.com/IftachSadeh/ANNZ} \\
-\textsc{Delight} 	& ML/template 	& \citet{Leistedt:17}		& \url{https://github.com/ixkael/Delight} \\
-\multirow{ 2}{*}{\textsc{FlexZBoost}} & \multirow{ 2}{*}{ML} 	& \multirow{ 2}{*}{\citet{Izbicki:17}}		& \url{https://github.com/tpospisi/flexcode};\\ 
-					& 		&							& \url{https://github.com/rizbicki/FlexCoDE}\\
-%\textsc{Frankenz} 	& ML 	& \citet{Speagle:frankenz}	& \url{https://github.com/joshspeagle/frankenz} \\
-\textsc{GPz} 		& ML	& \citet{Almosallam:15b}	& \url{https://github.com/OxfordML/GPz} \\
-\textsc{METAPhoR} 	& ML 	& \citet{Cavuoti:17}		& \url{http://dame.dsf.unina.it}\\
-\textsc{CMNN} 		& ML 	& \citet{Graham:17}			& - \\
-\textsc{SkyNet} 	& ML 	& \citet{Graff:14}			& \url{http://ccpforge.cse.rl.ac.uk/gf/project/skynet/} \\
-\textsc{TPZ} 		& ML 	& \citet{Carrasco_Kind:13}	& \url{https://github.com/mgckind/MLZ} \\
-\hline
-\trainz 	& N/A	& See Section~\ref{sec:method:trainz} & \\
-\end{tabular}}
-\end{table*}
-
-%\red{The questions that must be answered for each code are\dots}
-The questions that must be answered for each code are: what unique features are included in the specific implementation that influence the output $p(z)$.
-What form of validation was performed with the training data, how were photometric uncertainties employed in the analysis, how were negative fluxes treated, what specific prior form was employed (for template based codes), or what specific machine learning architcture was used (for ML codes)?
-%\scc{which parameter (in a general meaning) have been used, which kind of validation has been performed (if any)}
-\subsection{Template-based Approaches}
-\label{sec:templatecodes}
-
-\subsubsection{BPZ}
-\label{sec:BPZ}
-%(Sam Schmidt)
-
-\textsc{BPZ}\footnote{\url{http://www.stsci.edu/~dcoe/BPZ/}} \citep[Bayesian Photometric Redshift,][]{Benitez:00} is a template-based photo-$z$ code that compares the expected colors ($C$) calculated for a set of spectral energy distribution (SED) types/templates ($T$) to the observed colors to calculate the likelihood of observing colors at each redshift for each type, $p(C|z,T)$.
-The code employs an empirically determined Bayesian prior in apparent magnitude ($m_0$) and SED-type.
-Assuming that the SED-types are spanning and exclusive, we can determine the redshift posterior $p(z|C,m_0)$ by marginalizing over all SED-types with a simple sum \citep[Eq.~3 from][]{Benitez:00}:
-
-\begin{equation} \label{eq:redshift_posterior}
-p(z|C,m_0)\,\propto\, \sum_{T}p(z,T|m_0)\,p(C|z,T)
-\end{equation}
-\noindent where the first term on the right-hand side is the Bayesian prior and the second term is the traditional likelihood.
-The prior is assumed to have the form: $p(z,T|m_0)=p(T|m_0)\,p(z|T,m_0)$, i.e.
-it parameterizes the prior as an evolving type fraction with apparent magnitude, combined with a prior on the expected redshift probability distribution as a function of both apparent magnitude and SED-type.
-
-In this paper we use \textsc{BPZ v 1.99.3}. The template set employed here is the set of $100$ discrete SEDs described in Section~\ref{sec:buzztemplates}
-%, and the $129$ Brown SEDs for the Galacticus dataset. 
-To keep the number of free parameters to a manageable level the SEDs in the training set are sorted by the rest-frame $u$-$g$ colour and split into three ``broad'' SED classes, equivalent to the \texttt{E}, \texttt{Sp} and \texttt{Im/SB} types in \citet{Benitez:00}.
-We assume the same functional form for the Bayesian priors as used by \citet{Benitez:00}, and utilize the training-set galaxies with known SED-type, redshift, and apparent magnitude to determine the type fractions and the best fit for the eleven free parameters of the prior.
-%For photo-$z$ point estimates we use the \texttt{Z\_B} output parameter. 
-For galaxies with negative flux in a measured band, the placeholder value is replaced with an estimate one $\sigma$ detection limit in that particular band, i.~e.~a value close to the estimated sky noise threshold.
-The type-marginalized $p(z)$ is generated by setting the parameter \texttt{PROBS\_LITE=TRUE} in the \textsc{BPZ} parameter file.
-
-\subsubsection{EAZY} \label{sec:eazy}
-%(Rongpu Zhou)
-
-\textsc{EAZY}\footnote{\url{https://github.com/gbrammer/eazy-photoz}} \citep[Easy and Accurate Photometric Redshifts from Yale,][]{Brammer:08} is a template-based photo-$z$ code that includes several features that improve on the basic $\chi^2$ fit used in many template codes.
-It can fit the observed photometry with SEDs created from a linear combination of a set of templates at each redshift, and the best-fit SED is found by simultaneously fitting one, two or all of the templates by minimizing $\chi^2$.
-The minimized $\chi^2(z)$ is then combined with an apparent magnitude prior to obtain the posterior redshift probability distribution, although some argue that this is not the mathematically correct way of calculating the posteriors.
-\textsc{EAZY} can also account for the uncertainties in the templates by adding an empirically derived template error in quadrature as a function of redshift to the flux errors.
-
-In this paper we use the all-templates mode, which fits the photometric data with a linear combination of the five basis templates.
-We employed the $5$ basis templates described in Section~\ref{sec:buzzard}, and set the template error to zero since these same templates were used to produce the simulated catalog photometry.
-%\red{The quality parameter \texttt{Qz} is...}
-The likelihoods are calculated on a 200-point redshift grid spanning $0\leq z \leq 2$, and include the application of a type-independent apparent magnitude prior estimated from the training data.
-
-\subsubsection{LePhare}\label{sec:lephare}
-%(C\'ecile Roucelle, Eric Nuss, Johann Cohen-Tanugi)
-
-\textsc{LePhare}\footnote{\url{http://www.cfht.hawaii.edu/~arnouts/lephare.html}}\citep[Photometric Analysis for Redshift Estimate,][]{Arnouts:99,Ilbert:06} is a photo-$z$ reconstruction code based on a $\chi^2$ template-fitting procedure.
-The observed colors are matched with the colours predicted from a set of spectral energy distribution (SED) which can be either synthetic or based on a semi-empirical approach.
-\textsc{LePhare} has been used to produce the COSMOS2015 photo-$z$ catalogue \citep{Laigle:16}.
-
-Each SED is redshifted in steps of $\Delta z = 0.01$ and convolved with the simulated LSST filter transmission curves (accounting for instrument efficiency).
-The opacity of the inter-galactic medium has been set to zero as no additional reddening has been included in the Buzzard simulations.
-The computed photo-$z$ is then the value that minimizes the merit function $\chi^2 (z,T,A)$ from \citet{Arnouts:99}:
-
-\begin{equation}
-\chi^2(z,T,A) = \sum_f^{N_f}\left(\frac{F^f_{\rm{obs}}−A \times F^f_{\rm{pred}}(T,z)}{\sigma^f_{\rm{obs}}}\right)^2 
-\end{equation}
-\noindent where $A$ is a normalization factor, $F^f_{\rm{pred}}(T,z)$ is the flux predicted for a template $T$ at redshift $z$.
-$F^f_{\rm{obs}}$ is the observed flux in a given band $f$ and $\sigma^f_{\rm{obs}}$ the associated observational error.
-The index $f$ refers to the considered band and $N_f$ is the total number of filters.
-
-In this paper we use \textsc{LePhare v 2.2}. 
-The set of templates used for fitting the photo-$z$'s are the 100 discrete Buzzard SED templates as described in section~\ref{sec:buzztemplates}.
-%No correction for the photometric zero-points was applied for this work. 
-The full $p(z)$ corresponds to the likelihoods calculated at each point on our $z$-grid.
-
-\subsection{Training-based Codes}
-\label{sec:trainingcodes}
-
-\subsubsection{ANNz2}
-\label{sec:annz2}
-%(John Soo)
-
-\textsc{ANNz2}\footnote{\url{https://github.com/IftachSadeh/ANNZ}} \citep{Sadeh:16} is a powerful package that has the ability to employ several machine learning algorithms, including artificial neural networks (ANN), boosted decision tree (BDT) and k-nearest neighbour (KNN).
-Using the Toolkit for Multivariate Data Analysis (TMVA) with ROOT\footnote{\url{http://tmva.sourceforge.net/}}, it can run multiple machine learning algorithms for a single training and outputs photo-$z$'s based on a weighted average of their performances.
-
-\textsc{ANNz2} is capable of producing both photo-$z$ point estimates and redshift posterior probability distributions $p(z)$, it could also conduct classifications and supports reweighting between samples.
-The PDFs are produced by propagating the intrinsic uncertainty on the input parameters and the uncertainty in the machine learning method to the expected photo-$z$ solution, averaged over multiple runs weighted based on the performance of each run.
-\textsc{ANNz2} presents its photo-$z$ uncertainty different from many codes by using the KNN method: it estimates the photo-$z$ bias between each object and a fixed number of nearest neighbours in parameter space, it then takes the $68$th percentile width of the distribution of the bias.
-This is based on the implication that objects with similar photometric properties would have similar uncertainties, and therefore the photometric errors of the inputs are not propagated into the code.
-
-In this study, \textsc{ANNz2 v.
-2.0.4} was used.
-The full PDF for each galaxy is also produced with a linear stepsize of $z=0.01$ for $0 \leq z \leq 2$.
-A set of $5$ ANNs with architecture $6:12:12:1$ ($6$ $ugrizy$ inputs, $2$ hidden layers with $12$ nodes each, and $1$ output) with different random seeds are used during each training.
-%Training was conducted on a subset of the given training sample which satisfies $i\le25.3$ as it would have improved the results compared to training on the entire sample.
-Half of the training set is used as a validation set to prevent overtraining.
-All training objects are set to have detected magnitudes, however the non-detections (mag$=-99$) in the testing set are replaced with the mean of that particular band.
-
-\subsubsection{Color-Matched Nearest-Neighbours}
-\label{sec:cmnn}
-%(Melissa Graham)
-
-The nearest-neighbours color-matching photometric redshift estimator (\textsc{CMNN}) is presented in \cite[][herafter G18]{Graham:17}.
-This method uses a training set of galaxies with known redshifts that has equivalent or better photometry as the test set in terms of quality and filter coverage.
-For each galaxy in the test set we identify a color-matched subset of training galaxies, choose one (e.g.
-the nearest-neighbour or a random selection), and use its known redshift as the photo-$z$.
-This color-matched subset is identified by first calculating the Mahalanobis distance $D_M$ in color-space between the test galaxy and all training-set galaxies: the difference between the test and a training set galaxy's color divided by the photometric error, summed over all colors (i.e., $u$-$g$, $g$-$r$, $r$-$i$, $i$-$z$, and $z$-$y$).
-Then, the threshold value for $D_M$ that define a good color match is set by the percent point function (PPF): for example, for $N_{\rm dof}=5$, PPF$=95$ per cent of all training galaxies consistent with the test galaxy will have $D_M < 11.07$ (where $N_{\rm dof}$, the number of degrees of freedom, is the number of colors).
-For a given test galaxy, the $p(z)$ is the normalized distribution of the true catalogue redshifts of this color-matched subset of training galaxies, and the standard deviation of the color-matched subset is used as the photo-$z$ uncertainty.
-
-We have applied the nearest-neighbours color-matching photometric redshift estimator described in G18 to the simulated data.
-Compared to its application in G18, there are some minor differences in the application of this estimator to the Buzzard catalogue.
-First, we do not impose non-detections on galaxies with a magnitude fainter than the expected LSST 10-year limiting magnitude or bright enough to saturate with LSST: {\it all} of the photometry for all the galaxies in the test and training sets are used for this experiment.
-Second, as in G18 we do apply an initial cut in color to the training set before calculating the Mahalanobis distance in order to accelerate processing, and also use a magnitude pseudo-prior to improve photo-$z$ estimates, but for both we have used different cut-off values that are appropriate for the Buzzard galaxies' colors and magnitudes.
-Third, we set different parameters for the identification of the color-matched subset of training galaxies and the selection of a photometric redshift estimate.
-In G18 we used a percent point function (PPF) value of 0.68 to identify the color-matched subset of training galaxies and used the redshift of nearest neighbour in color-space as the photo-$z$ estimate.
-These choices work well when the desire is to obtain accurate photo-$z$ estimates for most test-set galaxies, but does not return a robust $p(z)$ in all cases -- especially for galaxies that are bright and/or have few matches in color-space.
-Since a robust estimate of $p(z)$ is desired for this work we make several changes to our implementation of the \textsc{CMNN} photo-$z$ estimator.
-We continue to use a percent point function of ${\rm PPF} = 0.95$ to generate the subset of color-matched training galaxies, but weight them by the inverse of their Mahalanobis distance.
-This weighting maintains some of the accuracy that was previously achieved by simply using the nearest neighbour in color-space.
-We then use the weights to create the $p(z)$ instead of having the redshift of each color-matched training-set galaxy count equally.
-To obtain a robust estimate of the $p(z)$ for galaxies with a small number of color-matched training set galaxies, when this number is less than $20$ the nearest $20$ neighbours in color-space are used instead, and we convolve the $p(z)$ with a Gaussian with a standard deviation of:
-
-\begin{equation}
-\sigma = \sigma_{\rm train} \sqrt{({\rm PPF}_{20}/0.95)^2 -1}
-\end{equation}
-
-\noindent to appropriately broaden it so that the $p(z)$ for these test galaxies represents the enlarged PPF value associated with it.
-Overall, these three changes will yield poorer accuracy photo-$z$ compared to those presented in G18, but they will all have significantly more robust estimates of the $p(z)$, particularly for the brightest test galaxies.
-This is sufficient for this work because, as described in G18, the goal of the \textsc{CMNN} photo-$z$ estimator was never to provide the ``best'' (or even competitive) estimates in the first place, given its reliance on a deep training set, but rather to provide a means for direct comparisons between LSST photometric quality and photo-$z$ estimates.
-With this work we show how the input parameters should be set in order to return robust $p(z)$ estimates in addition to point value estimates.
-
-\subsubsection{Delight}
-\label{sec:delight}
-%(John Soo)
-
-\textsc{Delight}\footnote{\url{https://github.com/ixkael/Delight}} \citep{Leistedt:17} infers photo-$z$'s by using a data-driven model of latent SEDs and a physical model of photometric fluxes as a function of redshift.
-Generally, machine learning methods rely on representative training data with similar band passes, while template based methods rely on a complete library of templates based on physical models constructed.
-\textsc{Delight} is constructed in attempt to combine the advantages and eliminate the disadvantages of both template-based and machine learning algorithms: it constructs a large collection of latent SED templates (or physical flux-redshift models) from training data, with a template SED library as a guide to the learning of the model.
-The advantage of \textsc{Delight} is that it neither needs representative training data in the same photometric bands, nor does it need detailed galaxy SED models to work.
-
-This conceptually novel approach is done by using Gaussian processes operating in flux-redshift space.
-The posterior distribution on the redshift of a target galaxy is obtained via a pairwise comparison with training galaxies,
-
-\begin{equation}
-p(z|\mathbf{\hat{F}}) \approx \sum_i p(\mathbf{\hat{F}}|z,t_i)\, p(z|t_i)p(t_i),
-\end{equation}
-\noindent where $p(z|t_i)p(t_i)$ captures prior information about the redshift distributions and abundances of the galaxies, with $t_i$ denoting the galaxy template; while $p(\mathbf{\hat{F}}|z,t_i)$ is the posterior of noisy flux $\mathbf{\hat{F}}$ at redshift $z$.
-For each training-target pair, $p(\mathbf{\hat{F}}|z,t_i)$ is evaluated as follows:
-
-\begin{equation} \label{eq:delight_noisy}
-p(\mathbf{\hat{F}}|z,t_i) = \int p(\mathbf{\hat{F}}|\mathbf{F})\, p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)\, d\mathbf{F},
-\end{equation}
-where $p(\mathbf{\hat{F}}|\mathbf{F})$ is the likelihood function, it compares the noisy real flux $\mathbf{\hat{F}}$ with the noiseless flux $\mathbf{F}$ obtained from the linear combination of template models, carefully constructed to account for model uncertainties and different normalization of the same SED; while $p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)$ is the prediction of flux at a different redshift $z$ with respect to the training object with redshift $z_i$ and flux $\mathbf{\hat{F}}_i$.
-Eq.~\ref{eq:delight_noisy} is essentially the probability that the training and the target galaxies having the same SED but at a different redshift.
-The flux prediction $p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)$ of the training galaxy at redshift $z$ is modeled via a Gaussian process,
-
-\begin{equation} \label{eq:delight_gp}
-F_b \sim \mathcal{GP}\left( \mu^F,k^F \right),
-\end{equation}
-\noindent with mean function $\mu^F$ and kernel $k^F$, both imposed to capture expected correlations resulting from the known underlying physics (i.e., fluxes resulting from observing SEDs through filter response, and the SEDs being redshifted).
-The reader should refer to \citet{Leistedt:17} for further details.
-
-In this study, all $100$ ordered Buzzard templates, as described in Section~\ref{sec:buzztemplates}, were used in \textsc{Delight}, and the Gaussian process was trained with a subset of $50\,000$ galaxies.
-Photometric uncertainties from the inputs are propagated into the code, while non-detections for each band are set to the mean of the respective bands.
-Default settings of \textsc{Delight} were use, with the exception that the PDF bins were set to be linear instead of logarithmic, with $200$ equally-spaced bins between $0.0<z<2.0$.
-In this study a flat prior is assumed.
-
-\subsubsection{FlexZBoost}
-\label{sec:flexzboost}
-%(Ann Lee, Rafael Izbicki, Taylor Pospisil, Peter Freeman)
-
-\textsc{FlexZBoost}\footnote{\url{https://github.com/tpospisi/flexcode};\\ \url{https://github.com/rizbicki/FlexCoDE}} \citep{Izbicki:17} is a particular realization of \texttt{FlexCode}, which is a general-purpose methodology for converting any conditional mean point estimator of $z$ to a conditional {\em density} estimator $f(z \vert \x)$, where $\x$ here represents our photometric covariates and errors.\footnote{Instead of $p(z)$, we use the notation $f(z \vert \x)$ to explicitly show the dependence on $\x$.} The key idea is to expand the unknown function $f(z \vert \x)$ in an orthonormal basis $\{\phi_i(z)\}_{i}$:
-
-\begin{equation}
-f(z|\x)=\sum_{i}\beta_{i }(\x)\phi_i(z). \label{eq::series_expansion}
-\end{equation}
-By the orthogonality property, the expansion coefficients are just conditional means
-
-\begin{equation}
-\beta_{i }(\x) =  \mathbb{E}\left[\phi_i(z)|\x\right] \equiv \int f(z|\x)   \phi_i(z) dz. \label{eq:FlexZBoost_coeffs}
-\end{equation}
-These coefficients can easily be estimated from data by regression.
-
-In this paper, we use \textsc{xgboost}~\citep{Chen:16} for the regression part as these techniques scale well for massive data; it should however be noted that \textsc{FlexCode-RF} (also on GitHub), based on Random Forests, generally performs better for smaller data sets.
-As our basis, we choose a standard Fourier basis.
-There are two tuning parameters in our $p(z)$ estimate: (i) the number of terms, $I$, in the series expansion in Eq.~\ref{eq::series_expansion}, and (ii) an exponent $\alpha$ that we use to sharpen the computed density estimates $\widehat{f}(z|\x)$, according to $\widetilde{f}(z|\x) \propto \widehat{f}(z|\x)^\alpha$.
-We split the ``train data'' into a training set (85\%) and a validation set (15\%), and choose both $I$ and $\alpha$ in an automated way by minimizing the weighted $L_2$-loss function \citep[Eq.
-5 in][]{Izbicki:17} on the  validation set.
-
-Although {\tt FlexCode} offers a {\em lossless compression} of the photo-$z$ estimates (in this study, one can reconstruct $\widetilde{f}(z|\x)$ exactly at any resolution from estimates of the first 35 coefficients, Eq.~\ref{eq:FlexZBoost_coeffs}, for a Fourier basis $\{\phi_i(z)\}_{i}$), we discretize our final estimates into 200 bins linearly spaced in $0 < z < 2$ for easy comparison with other algorithms.
-Using a higher resolution may yield better results (with no added cost in storage).
-
-%\subsubsection{Frankenz}
-%\label{sec:frankenz}
-%(seeking volunteers)
-%
-%\red{\textsc{Frankenz}\footnote{\url{https://github.com/joshspeagle/frankenz}} \cite{Speagle:frankenz} is...}
-%
-
-\subsubsection{GPz}
-\label{sec:gpz}
-%(Ibrahim Almosallam)
-
-\textsc{GPz}\footnote{\url{https://github.com/OxfordML/GPz}} \citep{Almosallam:16a,Almosallam:15b} is a sparse Gaussian process based code, a fast and a scalable approximation of full Gaussian Processes \citep{Rasmussen:06}, with the added feature of being able to produce input-dependent variance estimations (heteroscedastic noise).
-The model assumes that the probability of the output $y$, the redshift, given the input $x$, the photometry, is $p(y|x)=\mathcal{N}\left(y|\mu(x),\sigma(x)^{2}\right)$.
-The mean function, $\mu(x)$, and the variance function $\sigma(x)^{2}$ are both linear combinations of basis functions that take the following form:
-
-\begin{align}
-f(x)=\sum_{i=1}^{m}\phi_{i}(x)w_{i},
-\end{align}
-where $\left\{\phi_{i}(x)\right\}_{i=1}^{m}$ and $\left\{w_{i}\right\}_{i=1}^{m}$ are sets of $m$ basis functions and their associated weights respectively.
-Basis function models (BFM), for specific classes of basis functions such as the sigmoid or the squared exponential, have the advantage of being universal approximators, i.e.
-there exist a function of that form that can approximate any function, with mild assumptions, to any desired degree of accuracy.
-The details on how to learn the parameters of the model and the hyper-parameters of the basis functions are described in \citet{Almosallam:15b}.
-
-A unique feature in \textsc{GPz}, is that the variance estimate is composed of two terms each quantifying a different source of uncertainty.
-One term (the model uncertainty) reflects how much of the uncertainty is due to lack of training samples at the location of interest, whereas the second term (the noise uncertainty) reflects how much of the uncertainty is caused from observing many noisy samples at that location.
-Thus, the predictive variance can determine whether we need more representative samples or more precise samples for any particular location in the input space.
-\textsc{GPz} can also emphasize the importance of some samples as weights.
-This weight can be for example $|z_{\rm{spec}}-z_{\rm{phot}}|/(1+z_{\rm{spec}})$ to target the desired objective of minimizing the normalized redshift error or as a function of their probability in the test set relative to the training set in order to pressure the model to better fit samples that are rare in the training set but are expected to be abundant during testing.
-
-The data is prepared for \textsc{GPz} by taking the log of the magnitude errors, decorrelating the data set using PCA and imputing the missing values using a simple linear model that estimates the missing variables given the observed ones.
-The log transformation helps to smooth the long tail distribution of the magnitude errors, which is more stable numerically and makes the optimization process unconstrained.
-The missing values are imputed by computing the mean of the training set $\mu$ and its covariance $\Sigma$, then we use the following equation to estimate the missing values from the observed ones %\[
-
-\begin{equation}
-x_{u} = \mu_{u}+\Sigma_{uo}\Sigma_{oo}^{-1}(x_{o}-\mu_{o}),
-\end{equation}  %\]
-where the subscript $o$ in $x_{o}$ indexes the \emph{observed} part of the input $x$, whereas the subscript $u$ indexes the \emph{unobserved} set (similarly for $\mu$ and $\Sigma$).
-This is the optimal expected value of the unobserved variables given the observed ones if the distribution is jointly Gaussian, note that if the variables are independent, i.e.
-$\Sigma_{uo}=0$, this will reduce to a simple average predictor.
-
-We use the Variable Covariance (VC) option in \textsc{GPz} with 200 basis functions after we note that there is no significant increase in the performance on the validation set (using 80\%-20\% training-validation split) and with no cost-sensitive learning applied.
-We do note that \textsc{GPz} is trained using a data set that includes galaxies fainter than those in the test set, which was truncated at $i<25.3$.
-This mismatch in training vs.~test results in a slightly different distribution in galaxy photometry and redshift distribution between the training and test set, which most likely degraded results compared to a run where the training set more closely matched the test set.
-
-\subsubsection{METAPhoR}
-\label{sec:metaphor}
-%(Stefano Cavuoti, Massimo Brescia, Giuseppe Longo)
-%\textcolor{brown}{general review done on this section}
-\textsc{METAPhoR} \citep[Machine-learning Estimation Tool for Accurate Photometric Redshifts,][]{Cavuoti:17} is a pipeline designed to provide photo-$z$'s point estimates and a reliable PDF for machine learning (ML) based techniques.
-It includes pre- and post-processing phases, hosting a photo-$z$ prediction engine based on the Multi Layer Perceptron with Quasi Newton Algorithm (MLPQNA), already validated on photo-$z$'s in several cases \citep{de_Jong:17,Cavuoti:17b,Cavuoti:15,Brescia:14,Brescia:13,Biviano:13}.
-Due to its plug-in based modular nature, \textsc{METAPhoR} can be easily replaced by any other photo-$z$ prediction kernel, regardless its implementation, by taking the I/O interface compliance as unique constrain.
-
-At a higher level, the pipeline mainly consists of three modules: (i) \textit{data pre-processing}, including a catalogue cross-matching sub-module \citep[based on the tool C3, ][]{Riccio:17}, a sub-module for photometric evaluation and error estimation of the multi-band catalogue used as Knowledge Base (KB), and a sub-module dedicated to the perturbation of the photometric KB, propaedeutic to the PDF estimation; (ii) \textit{photo-$z$ prediction}, which is the training/validation/test phase, producing the photo-$z$'s point estimates, based on a pre-selected ML method; (iii) \textit{PDF estimation}, specifically designed to calculate the PDF of the photo-$z$ estimation errors.
-The last module includes also a post-processing tool, providing some statistics on the produced point estimates and PDFs.
-
-The photometry perturbation law is based on the formula $m_{ij} = m_{ij} + \alpha_{i}F_{ij}*u_{\mu=0,\sigma=1}$, where $\alpha_{i}$ is a user selected multiplicative constant (useful in case of multi-survey photometry), $u_{\mu=0,\sigma=1}$ is a random value from the standard normal distribution and $F_{ij}$ is a bimodal function (a constant function + polynomial fitting of the mean magnitude errors on the binned bands), heuristically tuned in such a way that the constant component is the threshold under which the polynomial function is considered too low to provide a significant noise contribution to the photometry perturbation.
-
-As introduced, the photo-$z$ point estimate prediction engine of \textsc{METAPhoR} is based on the MLPQNA model, whose photo-$z$ regression training error, used by the quasi Newton learning rule, is based on the least square error and Tikhonov $L_{2}$-norm regularization \citep{Hofmann:18}.
-
-As main prerogative, \textsc{METAPhoR} is able to provide a PDF for ML methods by taking into account the photometric errors provided with data, by running $N$ trainings on the same training set, or $M$ trainings on $M$ different random extractions from the KB.
-The different test sets, used to produce the PDF, are thus obtained by introducing a proper perturbation, parametrized from the photometric error distribution in each band, on the photometric data populating the original test set \citep{Brescia:18}.
-
-For the present work since it was required to produce a redshift (and a PDF) for each object of the test set we decided to apply a hierarchical kNN to fill the missing detection, it goes without saying that for such points the reliability of PDFs and point estimation is lower.
-No cross validation has been used.
-
-\subsubsection{SkyNet}
-\label{sec:skynet}
-%(J. Cohen-Tanugi and Hugo Tranin)
-
-\textsc{SkyNet}\footnote{\url{http://ccpforge.cse.rl.ac.uk/gf/project/skynet/}} \citep{Graff:14} is a publicly available neural network software, based on a 2nd order conjugate gradient optimization scheme \citep[see][for further details]{Graff:14}.
-It has been used efficiently for redshift PDF estimates \citep{Sanchez:14,Bonnett:15,Bonnett:16}.
-
-%In the present work, we use \textsc{SkyNet} both as a regressor (for redshift point estimates) and a classifier (for PDF estimates). 
-The neural network is configured as a standard multilayer perceptron with three hidden layers and one input layer with $12$ nodes (the $6$ magnitudes and their errors). 
-%For the regressor, the output layer is a single linear node, while each hidden layer has $10$ nodes with a tanh activation function. 
-The classifier is laid out such that the hidden layers have 20:40:40 nodes each, all rectified linear units, and the output layer has $200$ nodes (corresponding to 200 bins for the PDF) activated with a ``softmax'' function so that they automatically sum to $1$.
-
-To avoid over-fitting, a $30$ per cent fraction of the training set is used as validation, and the training is stopped as soon as the error rate begins to increase in the validation set.
-The weights are randomly initialized based on normal sampling.
-The error function is a standard chi-square function for the regressor, and a cross-entropy function for the classifier.
-Finally, the data are all whitened before processing, with magnitudes pegged to (45,45,40,35,42,42) and their errors pegged to (20,20,10,5,15,15) for $ugrizy$ filters, respectively.
-
-\subsubsection{TPZ}
-\label{sec:tpz}
-%(Erfan Nourbakhsh)
-
-\textsc{TPZ}\footnote{\url{https://github.com/mgckind/MLZ}} \citep[Trees for Photo-$z$,][]{Carrasco_Kind:13,Carrascokind:14} is a parallel machine learning algorithm that generates photometric redshift PDFs using prediction trees and random forest techniques.
-The code recursively splits the input data (i.~e.~the training sample), into two branches, one after another, until a terminal leaf is created that meets a termination criterion (e.~g.~a minimum leaf size or a variance threshold).
-Bootstrap samples from the training data and associated errors are used to build a set of prediction trees.
-In order to minimize correlation between the trees, the data is divided in such a way that the highest information gain among the random subsample of features is obtained at every point.
-The regions in each terminal leaf node corresponds to a specific subsample of the entire data that possesses similar properties.
-
-The training data is examined before running TPZ.
-Since TPZ does not handle non-detections (magnitudes flagged as 99.0), we replace these values with an approximation of the $1\sigma$ detection threshold, i.~e.~a signal to noise ratio of 1 in terms of magnitude uncertainty using the equation $dm = 2.5 ~ \log ( 1 + N/S )$ where $dm \sim 0.7526 ~ mag$ for $N/S=1$.
-That is, for each band, we replace the non-detection with the magnitude corresponding to the error of 0.7526 from the error model forecasted for 10-year LSST data.
-The Out-of-Bag \citep{Breiman:84,Carrasco_Kind:13} cross-validation technique is used within TPZ to evaluate its predictive validity and determine the relative importance of the different input attributes.
-We employed this information to calibrate our algorithm.
-
-In the present work, the LSST magnitudes $u,~g,~r,~i$ and colors $u-g,~g-r,~r-i,~i-z,~z-y$ and their associated errors are used in the process of growing 100 trees with a minimum leaf size of 5 (the $z$ and $y$ magnitudes did not show significant correlation with the redshift in our cross-validation, so we did not use them when constructing our trees).
-We partitioned our redshift space into 100 bins from $z=0.005$ to $z=2.0$ and smoothed each individual PDF with a smoothing scale of twice the bin size.
-
-\subsection{Simple Ensemble Estimator}\label{sec:method:trainz}
-In addition to the main photo-$z$ algorithms described above we also include a very simple method.  For \trainz, as we will we call this simple estimator, we well define $p(z)$ as simply:
-
-\begin{equation}
-p(z) = \frac{1}{N_{train}}\sum_{i=1}^{N_{train}}z_{train}
-\end{equation}
-That is, we simply set the redshift PDF of every galaxy equal to the normalized $N(z)$ of the training sample.
-As the training sample is drawn from the same underlying distribution as the test sample, modulo small deviations due to sample size, the quantiles of the training and test distributions should be identical.
-This is a wildly unrealistic estimator, as it assigns all galaxies, no matter their apparent magnitude, colour, or true redshift, the same redshift PDF, and is thus uninformative at the level of individual object redshifts, but is designed to perform very well for the ensemble of all objects.
-We will discuss this method and cautions relative to metrics in Section~\ref{sec:caution}.
-
-\section{Metrics for quantifying PDF comparisons}
-\label{sec:metrics}
-%(Alex Malz, Rongpu Zhou, Jeff Newman, Ofer Lahav)
-
-%\red{Maybe it would be best to introduce the summary statistics (KS, CvM, AD) up front and then the ensemble functions and stacked n(z).}
-
-The overloaded ``$p(z)$'' is a widespread abuse of notation; we would like the outputs of photo-$z$ PDF codes to be interpretable as probabilities. 
-Obviously photo-$z$ PDFs must not take negative values and must integrate to unity over the range of possible redshifts. 
-Additionally, an estimator derived by method $H$ for the photo-$z$ PDF of galaxy $i$ must be understood as a posterior probability distribution
-
-\begin{equation}
-\hat{p}_{j}(z_{i}) = p(z | d_{i}, I_{D}, I_{H}),
-\end{equation}
-conditioned not only on the photometric data $d_{i}$ for that galaxy but also on parameters encompassing a number of things that will differ depending on the method $H$ used to produce it, namely the assumptions $I_{H}$ necessary for the method to be valid and any inputs $I_{D}$ it takes as prior information, such as a template library or training set.
-Because of this, direct comparison of photo-$z$ PDFs produced by different methods is in some sense impossible; even if they share the same prior information $I_{D}$, by definition they cannot be conditioned on the same assumptions $I_{H}$, otherwise they would not be distinct methods at all.
-
-In this study, we isolate the differences in prior information specific to each method by using a single training set $I_{D}^{ML}$ for all machine learning-based codes and a single template library $I_{D}^{T}$ for all template-based codes, and these sets of prior information are carefully constructed to be representative and complete, we have $I_{D}^{ML}\equiv I_{D}^{T}$ for every method $H$.
-Thus, we are saying
-
-\begin{equation}
-\frac{\hat{p}_{i, H}(z)}{\hat{p}_{i, H'}(z)} \approx \frac{p(z | d_{i}, I_{H})}{p(z | d_{i}, I_{H'})},
-\end{equation}
-meaning that we assume comparisons of $\hat{p}_{i, H}(z)$ isolate the effect of the method used to obtain the estimator, which should make examination of differences caused by specifics of the method implementations easier to isolate.
-
-As mentioned previously, there are cosmology probes that require knowledge of individual galaxy $p(z)$ and others that require only knowledge of the ensemble redshift distribution, $N(z)$.  
-Due to the paucity of principled techniques for using and validating photo-$z$ PDFs, there are few alternatives to the common practice of reducing photo-$z$ PDFs to point estimates. 
-Though this practice should not be encouraged, we also calculate traditional metrics based on the most common point estimators derived from photo-$z$ PDFs.  
-Those seeking to establish a connection to traditional ways of thinking about redshift estimation may consult the Appendix for these results.
-
-There are a number of metrics that can be used to test the accuracy of a photo-$z$ interim posterior as an estimator of a true photo-$z$ posterior if it is known.
-Even for simulated data, the true photo-$z$ PDF is in general not accessible unless the redshifts are in fact drawn from the true photo-$z$ PDFs, a mock catalogue generation procedure that has not yet appeared in the literature.
-Furthermore, only limited applications of photo-$z$ PDFs that could be used as the basis for a metric have been presented in the literature. 
-The most popular application by far is the calculation of the overall redshift distribution $N(z)$, the true value of which is known for the \textsc{Buzzard} simulation and will be denoted as $N'(z)$.
-Though alternatives exist \citep{Malz:chippr}, stacking according to
-
-\begin{equation} \label{eq:stacked}
-\hat{N}^{H}(z) \approx \frac{1}{N_{tot}}\ \sum_{i}^{N_{tot}}\ \hat{p}^{H}_{i}(z)
-\end{equation}
-is the most widely accepted method for estimating the redshift distribution from photo-$z$ PDFs. 
-If we assume that the response of estimators of $N(z)$ is uniform across all approaches $H$, then we may interpret metrics on the accuracy of $\hat{N}(z)$ obtained in this way. 
-We must note, however, that this is a poor assumption in general. 
-Under the setup of this paper, the true redshift distribution $N'(z) = p(z | I_{D})$ (i.e.
-because our training data is representative, the interim prior is the truth).
-In this ideal case, the method that would give the best approximation to $N'(z)$ would be one that neglects all the information contained in the photometry $\{d_{i}\}_{N_{tot}}$ and gives every galaxy the same photo-$z$ PDF $\hat{p}_{i}(z) = N'(z)$ for all $i$.
-This is the exact estimator, \trainz, that we have described in Section~\ref{sec:method:trainz}, and which will serve as a point of reference for the other codes.
-
-The exact implementation of the stacked estimator $\hat{N}^{H}(z)$ will depend on the parametrization of the photo-$z$ PDFs, which may differ across codes and can affect the precision of the estimator \citep{Malz:qp}; even considering a single method under the same parametrization, say a piecewise constant function over bins or a set of samples from the posterior, an estimator using $2N$ bins or samples will trivially be more precise than an estimator using $N$ bins or samples.
-%\red{Note: Let's address how the gridding (endpoints and spacing) affects the metrics, from the email correspondence Re: [LSST-DESC-MEMBERS] Summary of Nov 29 PZ Telecon. (AIM)}
-In order to minimize the effects of such choices, we asked those running all eleven codes to output $p(z)$ parameterized with a generous $\approx 200$ piecewise constant bins spanning $0<z<2$.  
-The piecewise constant format is chosen because of its established presence in the literature, and the choice of 200 bins was motivated by the approximate number of columns expected to be available for storage of $p(z)$ for the final LSST Project tables.\footnote{See, e.~g.~the LSST Data Products Definition Document, available at: \url{https://ls.st/dpdd}}
-All $p(z)$ catalogues are processed using the \textsc{qp} software package \citep{Malz:qp}\footnote{available at: \url{http://github.com/aimalz/qp/}} for manipulating and calculating metrics of 1-dimensional PDFs.
-We will discuss the choice of $p(z)$ parameterization further in Section~\ref{sec:results}.
-
-%\red{Divide metrics by application to ensemble of PDFs vs. estimators of the redshift distribution}
-
-\subsection{Metrics of an ensemble of photo-\mathinhead{$z$}{z} interim posteriors}
-\label{sec:qualmet}
-
-\subsubsection{Probability integral transform (PIT)}
-\label{sec:pit}
-%(Alex Malz, Peter Freeman, Sam Schmidt)
-
-The probability integral transform (PIT) \citep{Polsterer:16} is defined for each individual galaxy as:%essentially the derivative of the QQ plot, a histogram of
-
-\begin{align}
-\label{eq:pit}
-\mathrm{PIT} &= \int_{-\infty}^{z_{\rm true}}p(z)\, dz \,.
-\end{align}
-The distribution of PIT values quantifies the behavior of the {\it ensemble} of photo-$z$ PDFs, enabling us to evaluate whether the $p(z)$ is, on average, accurate:  The PIT value is the Cumulative Distribution Function (CDF) of the $p(z)$ evaluated at the true redshift.
-A catalogue of photo-$z$ PDFs that are accurate should have a flat PIT histogram (i.e., the individual PIT values as samples from each CDF should match a Uniform(0,1) distribution if the CDFs are accurate).
-Specific deviations from flatness indicate inaccuracy: overly broad photo-$z$ PDFs would manifest as underrepresentation of the lowest and highest PIT values, whereas overly narrow photo-$z$ PDFs would manifest as over-representation of the lowest and highest PIT values.
-High frequency at only $\mathrm{PIT}\approx0$ and $\mathrm{PIT}\approx1$ indicates the presence of catastrophic outliers with highly inaccurate photo-$z$ PDFs where the true redshift is outside of the support of $p(z)$.
-\citet{Tanaka:17} use the histogram of PIT values as a diagnostic indicator of overall code performance, while \citet{Freeman:17} independently define the PIT and demonstrate how its individual values may be used both to perform hypothesis testing (via, e.g., the KS, CvM, and AD tests; see below) and to construct quantile-quantile plots.
-%\red{We might want to divide this by redshift to get trends, as the overall PIT could be impacted by true n(z).  
-%For example, a PIT with nontrivial shape could be due to that behavior being uniform as a function of redshift or significant inaccuracy only in a small range of redshifts where $n(z)$ is large and a flat PIT elsewhere.}
-
-\subsubsection{Quantile-quantile (QQ) plot}
-\label{sec:qq}
-%(Rongpu Zhou, Jeffrey Newman, Alex Malz)
-
-The quantile-quantile (QQ) plot is a graphical method for comparing two distributions, where the quantiles of one distribution are plotted against the quantiles of the other distribution (A quantile being defined by partitioning a distribution into consecutive intervals containing equal amounts of probability, or equal numbers of objects in each interval).
-%\blue{Include an equation to ensure quantile definition is crystal clear before building upon it for PIT.}\red{Is maybe the above text clear enough?
-Text may be more clear than any equation I can come up with--SJS.}
-In this paper we show the quantiles of the PIT values compared to the quantiles of the Uniform distribution that we expect the PIT values to match if $p(z)$ is an accurate probability distribution for all objects.
-The QQ plot provides an easy way to qualitatively assess the differences in various properties such as the moments of an estimating distribution relative to a true distribution.  
-%location, scale and skewness.  
-In this paper, QQ plots are used for two purposes: (1) for comparing $N(z)$ from photo-$z$ PDFs (estimated using Eq.~\ref{eq:stacked}) with the true $N(z)$, i.e.
-comparing the estimated distribution of redshifts with the true redshift distribution, and (2) for assessing the overall consistency of an ensemble of photo-$z$ PDFs with their true redshifts on a population level, where the distribution of the PIT values (see previous section) is compared to a uniform distribution between $0$ and $1$.
-The QQ plot contains very similar information to that shown in the PIT histogram plot, we include both forms, as visually they each convey the information in a somewhat distinct manner.
-%\red{is this enough reason to include both?--SJS}
-%\aim{Maybe we should explain why we expect the distribution of CDFs to be uniform.}\red{Note: this is just describing the PIT.
-I'm going to swap the order of PIT to come before this, and then say we are evaluating the distribution of PIT values.
-I will also add why we expect this to be uniform--SJS}
-
-\subsubsection{Conditional density estimation loss}
-\label{sec:CDE_loss}
-%(Alex Malz, Ann Lee)
-
-With the conditional density estimation loss (CDE loss) we can compare how well different methods estimate individual PDFs for photometric covariates $\x$ rather than looking only at the ensemble distribution.
-As in Section~\ref{sec:flexzboost}, we use the notation $f(z \vert \x)$ instead of $p(z)$ to explicitly show the dependence on $\x$.
-
-%\aim{Perhaps this should be rewritten in the same notation used elsewhere in the paper.} \red{I agree, does someone who speaks both stats and astro want to take a shot at this?}
-%\textcolor{cyan}{Ann: We are looking at individual PDFs so the dependence on $\x$ needs to be explicit; this is probably the most accurate notation as it shows both the conditioning clearly as well as whether $\x$ is fixed or random.}
-
-The CDE loss is defined as:
-\begin{equation} \label{eq:cde-loss}
-L(f, \widehat{f}) = \int \int (f(z \mid \x) - \widehat{f}(z \mid \x))^{2} dz dP(\x)
-\end{equation}
-This loss is the CDE equivalent of the RMSE in regression. To estimate this loss we rewrite the loss as
-\begin{equation} \label{eq:estimated-cde-loss}
-\mathbb{E}_\X \left[\int \widehat{f}(z \mid \X)^{2} dz\right] - 2 \mathbb{E}_{\X,Z}\left[\widehat{f}(Z \mid \X)\right] + K_{f},
-\end{equation}
-where the first expectation is with respect to the marginal distribution of the covariates $\X$, the second expectation is with respect to the joint distribution of $\X$ and $Z$, and $K_{f}$ is a constant depending only upon the true conditional densities $f(z\mid \x)$.
-For each method we can estimate these expectations as empirical expectations on the test or validation data \citep[Eq.~7 in][]{Izbicki:17b} without knowledge of the true densities.
-
-%\subsubsection{Continuous Ranked Probability Score (CRPS)}
-%\label{sec:crps}
-%\red{The CRPS is used to assess the accuracy of two probabilistic models respectively.}
-
-\subsection{Metrics over estimated probability distributions}
-\label{sec:quantmet}
-
-In tandem with the QQ and PIT metrics introduced above, we additionally compute the following metrics comparing the empirical CDF of a distribution to the true or expected distribution.
-These metrics give a more quantitative measure of the departure from ideal than the more visual PIT histogram and QQ plot.
-We compute metrics comparing the CDF of PIT values to a the CDF of a Uniform distribution, and also compute the CDF of the true redshift distribution $N'(z)$ compared the $\hat{N}(z)$ distribution derived from summing the $p(z)$ as described in Eq.~\ref{eq:stacked}.
-
-\subsubsection{Root-mean-square error (RMSE)}
-\label{sec:rmse}
-%(Alex Malz)
-
-We employ the familiar root-mean-square error:
-
-\begin{equation}
-\mathrm{RMSE} = \sqrt{\int_{-\infty}^{\infty}\left(\hat{f}(z)-f'(z)\right)^{2}\,dz},
-\end{equation}
-%may be used for two purposes in this analysis. 
-%We calculate it between the true redshift distribution $N'(z)$ and each estimator $\hat{N}(z)$ (derived by Eq.~\ref{eq:stacked}) as well as between the ideal QQ and PIT curves from each method studied and the ideal curves.
-Though this metric does not account for the fact that the redshift distribution function is, in fact, a probability distribution, it can still be interpreted as a measure of the integrated difference between the estimated distribution and the true distribution, and it can be used to quantify the otherwise qualitative metrics.
-
-\subsubsection{Kolmogorov-Smirnov (KS) and related statistics}
-\label{sec:ks}
-%(John Soo, Rongpu Zhou)
-
-The \textit{Kolmogorov-Smirnov statistic} $N_{\rm KS}$ is the maximum difference between $F_{\rm phot}(z)$ and $F_{\rm spec}(z)$, the CDFs of the photo-$z$ and spectroscopic redshift respectively:
-
-\begin{equation}
-N_{\rm KS} = \max_{z}\left( \left| F_{\rm phot}(z) - F_{\rm spec}(z) \right| \right).
-\end{equation}
-%The KS test quantifies whether the two redshift distributions are drawn from a similar distribution, independent of binning. 
-The KS test quantifies the similarity between two distributions, independent of binning. 
-A lower $N_{\rm KS}$ value corresponds to more similar distributions.
-%We calculate two KS test statistics, one to test the distribution of $p(z)$ values, and one to test the stacked estimator $\hat{n}(z)$ of the redshift distribution.  
-%The $p(z)$ test compares the CDF of the PIT values described in the previous section to that of a uniform distribution.  
-%The $n(z)$ test compares the CDF of the distribution formed by stacking the $p(z)$ for every object in our test set against the CDF of the corresponding true redshifts.
-%In this case a lower $N_{\rm KS}$ value means that the photo-$z$ distribution fits the spectroscopic redshift distribution well.
-$F_{\rm phot}(z)$ is calculated by stacking the PDFs of every single object in the sample, producing a smooth cumulative $n(z)$ to be compared with $F_{\rm spec}(z)$.
-
-We also consider two variants of the KS statistic: the Cramer-von Mises (CvM) and Anderson-Darling (AD) statistics.
-The CvM statistic is similar to the KS statistic as it is also computed from the distance between the measured CDF and the ideal CDF, but instead of the maximum distance, the CvM statistic calculates the average of the distance squared:
-
-\begin{equation}
-\label{eq:cvm}
-\omega^2 = \int_{-\infty}^{+\infty}\big(F_{\mathrm{meas.}}(x) - F_{\mathrm{ideal}}(x)\big)^2\mathrm{d}F_{\mathrm{ideal}}
-\end{equation}
-% In this paper we use the CvM statistic to evaluate the performance of both $n(z)$ and $p(z)$.
-The AD statistic is a weighted version of the CvM statistic, making it more sensitive to the tails of the distribution:
-
-\begin{equation} \label{eq:ad}
-A^2 = n\int_{-\infty}^{+\infty} \frac{\big(F_{\mathrm{meas.}}(x) - F_{\mathrm{ideal}}(x)\big)^2} {F_{\mathrm{ideal}}(x) (1-F_{\mathrm{ideal}}(x))}\mathrm{d}F_{\mathrm{ideal}}
-\end{equation}
-where $n$ is the sample size. 
-%The KS statistic and the CvM and AD statistics related to it are all used to evaluate the performance of both $\hat{n}(z)$ and the QQ and PIT of the CDFs of the ensemble of photo-$z$ interim posteriors.
-
-%\subsubsection{Kullback-Leibler Divergence (KLD)}
-%\label{sec:kl}
-%(Alex Malz)
-%
-%The KLD is a measure of the loss in information when using a probability distribution that is an estimate of a known true probability distribution, so we use 
-%\begin{equation}
-%\mathrm{KLD} = \int_{-\infty}^{\infty}n'(z)\log\left[\frac{n'(z)}{\hat{n}(z)}\right]dz
-%\end{equation}
-%as a metric on the stacked redshift distribution estimator of Eq.~\ref{eq:stacked}.
-Noting the assumptions made in using Eq.~\ref{eq:stacked}, we say that a photo-$z$ PDFs method whose stacked redshift distribution estimator $\hat{n}(z)$ more closely resembles the true redshift distribution $n'(z)$ is desired, i.e.
-we seek to minimize the KLD.
-
-\subsubsection{Moments}
-\label{sec:moments}
-%(Alex Malz)
-For the $\hat{N}(z)$ distributions we additionaly calculate the first three moments of the estimated redshift distribution for each code and compare them to the moments of the true redshift distribution $N'(z)$.
-The $m$th moment of a distribution is defined as
-
-\begin{equation}
-\langle z^{m}\rangle = \int_{-\infty}^{\infty}z^{m}N(z)dz.
-\end{equation}
-Here, we use the moments of the stacked estimator of the redshift distribution function as the basis for a metric.
-The closer the moments of $\hat{N}(z)$ for a photo-$z$ PDF method are to the moments of the true redshift distribution function $N'(z)$, the better the photo-$z$ PDF method.
-
-\section{Results}
-\label{sec:results}
-%(Sam Schmidt, Bryce Kalmbach, Johann Cohen Tanugi, Rongpu Zhou)
-
-\subsection{Ensembles of photo-\mathinhead{$z$}{z} interim posteriors}
-
-\begin{figure*}
-%\centering
-\begin{subfigure}[b]{0.49\textwidth}
-\includegraphics[width=\textwidth]{figures/pz_12codes_261931_crop.jpg}
-\end{subfigure}
-\begin{subfigure}[b]{0.49\textwidth}
-\includegraphics[width=\textwidth]{figures/pz_12codes_471167_crop.jpg}
-\end{subfigure}
-\begin{subfigure}[b]{0.49\textwidth}
-\includegraphics[width=\textwidth]
-{figures/pz_12codes_713178_crop.jpg}
-\end{subfigure}
-\begin{subfigure}[b]{0.49\textwidth}
-\includegraphics[width=\textwidth]{figures/pz_12codes_982747_crop.jpg}
-\end{subfigure}
-\caption{Four illustrative examples of individual p(z) distributions produced by the codes.
-The red vertical line represents the true redshift.
-Examples are chosen with common features seen in PDFs: tight unimodal $p(z)$ (upper left), broad unimodal $p(z)$ (upper right), bimodal $p(z)$ (lower right), and complex/multimodal $p(z)$ (lower left).
-Codes show varying amounts of small-scale structure in their reconstruction of the posterior distribution.
-We see varying responses from the codes in the presence of color degeneracies and photometric errors, resulting in narrow and broad unimodal, bimodal, and multi-modal $p(z)$ curves.} \label{fig:pz_examples}
-\end{figure*}
-
-%\begin{figure*}
-%\centering
-%\includegraphics[width=\textwidth]{figures/QQplot_11codes.jpg}
-%\caption{The QQ plots produced for every photo-$z$ code.
-\aim{Consider making this a plot of QQ minus that diagonal for better contrast between methods.}\red{The PIT histogram is essentially this, which is one of the main reasons to include both.
-QQ is a fairly standard form, so plotting in this fashion rather than the difference seems better to me, given that we also ahve the PIT histograms.
---SJS}} \label{fig:qq}
-%\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{figures/PITANDQQplot_12codes_crop.jpg}
-\caption{Summary plots for all eleven photo-$z$ codes illustrating performance for the interim posterior statistics.
-The top panel of each pair shows both the Quantile-Quantile (QQ) plot (red) and the histogram of PIT values (blue).
-The desired behavior is a QQ plot that matches the diagonal dashed line, and a PIT histogram that matches a uniform distribution matching the thin horizontal black line.
-The bottom panel of each pair shows the difference between the QQ quantile and the diagonal, illustrating departure from the desired performance.
-Histograms with an overabundance of PIT values at the centre of the distribution indicate $p(z)$ distributions that are overly broad, while an excess of values at the extrema indicate $p(z)$ distributions that are overly narrow.
-Values of PIT=0 and PIT=1 indicate ``catastrophic failures'' where the true redshift is completely outside the support of $p(z)$.
-Asymmetric features are indicative of systematic bias in the redshift predictions.
-A variety of behaviors are evident, and specific details are discussed in the text.}
-\label{fig:pitqq}
-\end{figure*}
-
-Fig.~\ref{fig:pz_examples} Shows the $p(z)$ produced by each of our eleven photo-$z$ codes for four example galaxies which exemplify some prominent cases that arise when estimating photo-$z$ PDFs: a narrow, unimodal redshift solution, a broader unimodal solution, a bimodal distribution, and a complex, multimodal distribution.
-The red vertical line represents the true redshift of the individual galaxy, and the blue curve represents the redshift probability.  
-Several features are obvious even in these illustrative examples.  
-\textsc{ANNz2, METAPhoR, NN,} and \textsc{SkyNet} all show an excess of small-scale features, which appear to be print-through of the underlying training set galaxies.  
-\textsc{GPZ} (in its current implementation), on the other hand, always produces a single Gaussian, which broadens to cover the multi-modal redshift solutions seen in other codes.
-
-As stated in Section~\ref{sec:metrics}, $p(z)$ is parameterized as $\approx 200$ piecewise constant bins covering $0<z<2$ for all eleven codes, giving a grid size of roughly $\delta z = 0.01$ for each code.
-A piecewise constant grid was a natural choice for some photo-$z$ codes, for instance most template-based codes compute likelihoods on a fixed grid.  
-In contrast, FlexZBoost, for example, can return estimates on any grid without compression errors as it’s a basis expansion method where only the expansion coefficients need to be stored.
-Codes with a native output format other than the shared piecewise constant binning scheme (or one that can be losslessly converted to it) may suffer from loss of information when converting to it, which could artificially favor some codes over others.
-
-Furthermore, the fidelity of photo-$z$ interim posteriors in this format varies with the quality of the photometry.  
-For faint galaxies, this redshift resolution is sufficient to capture the shape of $p(z)$ for the majority of the test sample, where photometric errors on the faint galaxies lead to somewhat broad peaks in the redshift posterior.
-However, as can be seen in e.~g.~the top left panel of Fig.~\ref{fig:pz_examples}, for bright galaxies with narrow $p(z)$ the grid spacing of $\delta z = 0.01$ is not sufficient to resolve the peak.  
-This is consistent with the results described in \citet[]{Malz:qp}, who find that quantiles (and, to a lesser degree, samples) often outperform gridded $p(z)$, particularly for bright objects and in the presence of harsher storage constraints.
-With a full 200 numbers to capture the information of each photo-$z$ PDF, any parametrization will perform adequately, but other storage parametrizations and limits on storage resources may be considered in future work.
-%Switching to a quantile based parameterization may be more costly computationally, for example template-based codes would need to test more grid point in order to resolve the quantiles for bright galaxies.
-However, the computational time for template based codes scales roughly linearly with the number of grid points, so this may be a reasonable thing to implement.
-We will discuss this further in Section~\ref{sec:discussion}.
-% \red{someone review this statement to make sure that I'm saying this correctly!}
-
-Fig.~\ref{fig:pitqq} shows both the quantile-quantile plots (red) and the histogram of PIT values (blue) summarizing the results from each photo-$z$ code.
-The red line shows the measured quantiles, while the black diagonal represents the ideal QQ values if the distribution were perfectly reproduced.
-A second panel below the main panel for each code shows the difference between $Q_{data}$ and $Q_{theory}$, i.~e.~the departure from the diagonal, for clarity.
-Biases and trends in whether the average width of the $p(z)$ values being over/under-predicted are evident.
-An overall bias where the predicted redshift is systematically low manifests as the measured QQ value falling above the diagonal, as is the case for \textsc{BPZ} and \textsc{EAZY}, while a systematic overprediction shows up as the measured QQ value falling below the diagonal, as seen in \textsc{TPZ}.
-In terms of PIT histograms, a systematic underprediction of redshift corresponds to fewer PIT values at $PIT<0.5$ and more at $PIT>0.5$, while a systematic overprediction will show the opposite.
-
-Examination of the PIT histograms and QQ plots shows that there are fairly generic issues with the width of $p(z)$ uncertainties: \textsc{Delight, NN, SkyNet} and \textsc{TPZ} all show a PIT histogram with an dearth of low values and and an excess of high values, signs that, on average, their $p(z)$ are more broad than the true distribution of redshifts.
-\textsc{METAPhoR} shows the opposite trend, indicating the the $p(z)$  are more narrow than the distributions given by the true redshifts.
-In all of these code cases there is a free parameter or bandwidth that can be used to tune uncertainties.
-The sensitivity of multiple codes to this bandwidth choice emphasizes the fact that great care must be taken in setting user-defined parameters in photo-$z$ codes, even in the presence of representative training/validation data.
-for \textsc{FlexZBoost} the ``sharpening'' parameter (described in Section~\ref{sec:flexzboost}) plays a key role in improving the results, resulting in a QQ plot that is very nearly diagonal.
-A similar sharpening procedure could be beneficial for several codes.
-Interestingly, the three purely template-based codes, \textsc{BPZ, EAZY}, and \textsc{LePhare}, show relatively well behaved $p(z)$ statistics (albeit with some bias), which may indicate that the likelihood estimation with representative templates is accurately capturing the uncertainties on individual redshifts.
-
-%Fig.~\ref{fig:pitqq} shows the PIT histogram plots produced by each photo-$z$ code. 
-%These plots show information that is very similar to that shown in the QQ plot in many ways. 
-The ideal PIT histogram would follow the black dashed line, representing a uniform distribution of PIT values, equivalent to the diagonal line in the QQ plot.
-Overly broad $p(z)$ values show up as an excess of PIT values near $0.5$ and a dearth of values at the edges, while overly narrow $p(z)$ will have an excess at the edges and will be missing values at the centre.
-Another feature evident in the PIT histograms is the number of ``catastrophic outlier'' values where the true redshift falls outside of the non-zero support of $p(z)$, corresponding to PIT$=0.0$ or $1.0$ is more apparent than in the QQ plots.
-Following Kodra \& Newman (in prep.) we define $f_{\rm O}$ as the fraction of objects with $PIT<0.0001$ or $PIT>0.9999$.
-Table~\ref{tab:pitoutlier} lists these fractions for each of the codes.
-For a proper Uniform distribution we expect a value of $0.0002$.
-Several codes show a marked excess, with \textsc{ANNz2, FlexZBoost, LePhare, and METAPhoR} with $f_{\rm O} > 0.02$, indicating a sizeable number of catastrophic redshift solutions where the true redshift is not covered by the extent of $p(z)$.
-For \textsc{METAPhoR} this may be partially due to an overall underprediction of the $p(z)$ width, however this is not the case for the other codes.
-\textsc{LePhare} is a particular outlier with nearly 5 per cent of objects outside of $p(z)$ support.
-Further study will be necessary to determine what is causing these misclassifications for \textsc{LePhare}.
-As expected, and by design, \textsc{TrainZ} has the proper fraction of outliers for the $f_{\rm O}$ statistic.
-
-\begin{table}
-\setlength{\tabcolsep}{2pt}
-\centering
-\caption{The fraction of ``catastrophic outlier'' PIT values.
-We expect a value of 0.0002 for a proper Uniform distribution.
-An excess over this small value indicates true redshifts that fall outside the non-zero support of the $p(z)$.
-}\label{tab:pitoutlier}
-\begin{tabular}{lc}
-\hline
-\hline
- ``catastrophic outlier''\\ PIT fraction\\
-\hline
-Photo-z Code & fraction PIT$<10^{-4}$ or $>$0.9999\\
-\hline
-\textsc{ANNz2} & 0.0265\\
-\textsc{BPZ} & 0.0192\\
-\textsc{Delight} & 0.0006\\
-\textsc{EAZY} & 0.0154\\ 
-\textsc{FlexZBoost} & 0.0202\\
-\textsc{GPz} & 0.0068\\ 
-\textsc{LePhare} & 0.0486\\
-\textsc{METAPhoR}& 0.0229\\
-\textsc{CMNN} & 0.0034\\ 
-\textsc{Skynet} & 0.0001\\  
-\textsc{TPZ} & 0.0130\\
-\hline
-\textsc{TrainZ} & 0.0002\\
-\end{tabular}
-\end{table}
-
-%\begin{figure*}
-%\centering
-%\includegraphics[width=\textwidth]{figures/PIT_histogram_plot_11codes.jpg}
-%\caption{The PIT histogram plots produced for every photo-$z$ code.
-Ideal behaviour would match a uniform distribution, indicated by the horizontal dashed line.
-Histograms with an overabundance of PIT values at the centre of the distribution indicate $p(z)$ distributions that are overly broad, while an excess of values at the extrema indicate $p(z)$ distributions that are overly narrow.
-Values of PIT=0 and PIT=1 indicate ``catastrophic failures'' where the true redshift is completely outside the support of $p(z)$.
-Asymmetric features are indicative of systematic bias in the redshift predictions.
-%\aim{Could we combine this with the previous figure by plotting both as differences from ideal on one plot with left and right axis labels?
-The information seems degenerate (which makes sense!)}\red{as you say, the information is degenerat, this form emphasizes the differences a bit more, and shows off the catastrophic outliers with PIT=0 and 1 a bit better in my mind.} } \label{fig:pit}
-%\end{figure*}
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{figures/KSvsCvMvsAD_PIT_withnull_jpg.jpg}
-\caption{A visual representation of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the PIT distributions.
-The statistics are often highly correlated, though the AD statistic truncates the extrema of the distribution and can have disparate values compared to KS and CvM.} \label{fig:pit_stats}
-\end{figure*}
-
-Fig.~\ref{fig:pit_stats} shows comparative metric values for the quantitative Kolmogorov-Smirnoff (KS), Cramer-Von Mises (CvM), and Anderson Darling (AD) test statistics for each of the codes based on comparing the distribution of their PIT values to the expected uniform distribution over the interval [0,1].
-The individual values of the statistic are not as important as the comparative score between the different codes.
-%\red{Can p-values be supplied for each statistic?
-The statistics themselves are difficult to interpret, other than ``lower is better'' (p-value in skgof was broken, having trouble finding 1-sample KS calculation for uniform distribution)}
-The AD test statistic diverges for values that include the extrema, and thus is calculated by excluding the edges of the distribution.  
-We calculate the AD statistic over the range of PIT values $v=[0.01,0.99]$.  \textsc{ANNz2} and \textsc{FlexZBoost} score very well for the PIT metrics.  
-\textsc{METAPhoR} and \textsc{LePhare} score very well in the PIT AD statistic, but both have a large number of catastrophic outliers, resulting in higher KS and CvM scores.
-
-Given the near-perfect training data, examining the individual codes for explanations for departures from the expected behaviour will be instructive in avoiding similar problems in future tests. 
-\textsc{ANNz2} performs quite well in $p(z)$ based metrics.
-In the specific implementation employed in this paper, the final $p(z)$ is a weighted average of five neural-nets.
-During the training process \textsc{ANNz2} compares the percentiles of the redshift training sample against the CDFs of the $p(z)$ sample.
-Distributions that more closely match are given extra weight, and the final weights are designed to produce accurate percentiles.
-Given that our metrics are focused on the percentile distributions, it is unsurprising that \textsc{ANNz2} performs well in the given metrics.
-The discreteness in the individual $p(z)$ estimated by \textsc{ANNz2} can be attributed to the fact that the code was run as a classifier, assigning weights to discrete bins of redshift.
-While multiple bins may receive weight, the bins themselves will still be discretized, and no additional smoothing was performed.
-Overall, \textsc{FlexZBoost} and \textsc{ANNz2} show the best ensemble agreement in their distribution of PIT values.
-
-\subsection{Metrics of the stacked estimator of the redshift distribution}\label{sec:stackedmetrics}
-
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{figures/NZsumplot_12codes.jpg}
-\caption{The stacked $p(z)$ produced by each photo-$z$ code ($\hat{N}(z)$, red) compared to the spectroscopic redshift distribution ($N'(z)$, blue).
-Varying levels of small-scale structure are seen in the codes.
-$N'(z)$ is smoothed using a single bandwidth chosen via Scott's rule for all codes.} \label{fig:nz}
-%\aim{Arguably this could also be clarified by showing the true $n'(z)$ once and then differences from it for each cod's stacked estimator\dots}\scc{what if we plot the difference between true and stacked rather than soverposing the two lines?
-Sam please do not hate me.}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{figures/KSvsCvMvsAD_NZ_withnull_jpg.jpg}
-\caption{A visual representation of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the $\hat{N}(z)$ distributions.
-The statistics are correlated, the codes with the lowest KS statistics tend to have the lowest CvM and AD statistics.
-\textsc{CMNN} performs markedly better than the others in reconstructing the overall $N(z)$ distribution, while \textsc{SkyNet} scores poorly due to an overall bias in its redshift predictions.} \label{fig:nz_stats}
-\end{figure*}
-
-Fig.~\ref{fig:nz} shows the stacked $\hat{N}(z)$ distribution compared to the true redshift distribution $N'(z)$ for all tested codes.
-The red line indicates the summed $p(z)$ for each code, while the blue line shows the true redshift distribution smoothed via kernel density estimation (KDE), with a bandwidth chosen via Scott's rule \citep{Scott:1992}.
-While Scott's rule is used to display $N'(z)$ in the figure, all quantitative statistics are computed via the empirical CDF, and are thus unaffected by bandwidth/smoothing choice.
-Several of the codes show an excess at $z \sim 1.4$, particularly the template-based codes \textsc{BPZ}, \textsc{EAZY}, and \textsc{LePhare}.  
-This is likely due to the $4000$ angstrom break passing through the gap between the $z$ and $y$ filters.  
-This feature is one of the most prominent in individual galaxy p(z), and is readily seen in the point-estimate plots shown in Fig.~\ref{fig:pz_pointestimates} and described in the Appendix.  
-Several of the machine learning based codes appear to be over-trained, adding excess galaxy probability to the redshift peaks and missing probability in the troughs.  
-Given that our training data is drawn from the same galaxy population as the test set, and our data has prominent peaks in $N'(z)$, perhaps it is not unexpected that such overtraining occurs. 
-A more extensive training/validation set might allow for a better choice of smoothing parameters in individual codes that would avoid such overtraining.
-
-%\aim{Move the text on Scott's Rule to Discussion?}
-%Scott's rule \citep{Scott:1992} was used to determine the smoothing bandwidth of the true redshift sample: that is, a single smoothing scale was chosen to represent the true redshift sample for all eleven codes.
-
-As with the $p(z)$ values in Figure~\ref{fig:pitqq}, different levels of substructure are obvious for the different codes.  
-While Scott's rule provides a relatively good general smoothing scale to represent the true $N'(z)$, there are smaller scale fluctuations: while \textsc{FlexZBoost} and \textsc{CMNN} appear somewhat discrepant in Fig.~\ref{fig:nz}, they are actually the two most accurate in terms of their quantitative measurements.
-%\textsc{SkyNet} also shows excess small-scale structure, but also shows significant overestimate of the high redshift galaxy population.  
-Interestingly, while \textsc{ANNz2} shows an abundance of small scale structure in individual $p(z)$ measurements (see Fig.~\ref{fig:pz_examples}), the summed $\hat{N}(z)$ is rather smooth, where the small scale features average out.
-This is not the case for the two other codes that show an abundance of substructure in their individual $p(z)$: both \textsc{CMNN} and \textsc{SkyNet} show small scale features both in $p(z)$ and $\hat{N}(z)$.
-For \textsc{CMNN} the $p(z)$ are a simply a weighted histogram of all spectroscopic training galaxies in nearby colour space with no smoothing applied, so the substructure is not unexpected.
-The PIT histogram and shape of the QQ plot in Figure~\ref{fig:pitqq} show that \textsc{CMNN} is producing $p(z)$ that are overly broad, additional smoothing of the $p(z)$ would exacerbate this problem.
-While the $\hat{N}(z)$ plot shows more small scale features than other codes, these features are actually representative of real structure in the true $N'(z)$, as evidenced by the very good metric scores for \textsc{CMNN}.
-\textsc{SkyNet} $p(z)$ were also not smoothed: while previous implementations of the code such as \citet{Sanchez:14} and \citet{Bonnett:15} (see Appendix C.3) implement a ``sliding bin'' smoothing, no such procedure was used in this study.
-In addition to excess substructure, \textsc{SkyNet} shows an obvious redshift bias, evident both visually in Figure~\ref{fig:nz} and in the first moment of N(z) listed in Table~\ref{tab:moments}, where it is clearly an outlier.
-\textsc{SkyNet} employed a method where a random sample of training galaxies was chosen, but there was no test that the subset was completely representative of the overall redshift distribution.
-Also unlike \citet{Bonnett:15}, no effort was made to add extra weight to more rare low and high redshift galaxies.
-Either of these decisions could be the cause of the bias seen in our results.
-Future runs of \textsc{SkyNet} will explore these implementation choices and their effects.
-
-Figure~\ref{fig:nz_stats} shows the quantitative Kolmogorov-Smirnoff (KS), Cramer-Von Mises (CvM), and Anderson Darling (AD) test statistics for each of the codes for the $\hat{N}(z)$ based measures.
-%\red{Can p-values be supplied for each statistic?
-The statistics themselves are difficult to interpret, other than ``lower is better'' (no, p-values are very difficult to compute for non-uniform distributions)}
-\textsc{FlexZBoost}, \textsc{CMNN}, and \textsc{TPZ} outperform the other codes in the $\hat{N}(z)$ metrics.  
-It is unsurprising that \textsc{CMNN} scores well, as with a near perfectly representative training set means that choosing neighbouring points in color/magnitude space should lead to excellent agreement in the final $\hat{N}(z)$ estimate.
-\textsc{TPZ} performed quite poorly in $p(z)$ statistics, but results in a good fit to the overall $N(z)$.
-This is somewhat surprising, as performance was optimized for accurate $p(z)$, not $\hat{N}(z)$.
-During the validation stage for \textsc{TPZ}, there was a trade off between the width of the $p(z)$ when adjusting a smoothing parameter and overall redshift bias.
-The optimal result in the PIT metrics, as illustrated in the shape of the QQ plot, does contain some level of bias as well as a slight underprediction of mean $p(z)$ width, which translates to poor metric scores.
-This is something that will be looked into for \textsc{TPZ} in the future.
-
-It is also of note that all three template-based codes show an excess in their stacked $p(z)$ at $z\,\sim\,1.3-1.4$.
-This redshift range corresponds to the wavelengths where the $4000$ Angstrom break is passing between the borders of the $z$ and $y$ filters.
-This strong break entering the gap between the two reddest filters can cause problems with redshift estimation of individual galaxies, as can be seen in the point-estimate photo-$z$'s shown in Figure~\ref{fig:pz_pointestimates}.
-This is not unique to this dataset, it is a common occurrence in photo-$z$ estimation.
-The fact that similar excesses appear in Figure~\ref{fig:nz} for \textsc{ANNz2} and \textsc{METAPhoR} shows that the effect is not limited to template-based codes.
-However, the lack of such a feature in the other codes shows that it is possible to eliminate the degeneracies.
-Further study on this issue may provide a solution for codes that suffer from this shortcoming.
-
-%\red{discussion of why NN and FZBoost are doing well, how we can try to improve the other codes} 
-%\textcolor{cyan}{Ann: }
-
-Table~\ref{tab:cdeloss} shows the CDE loss statistic for each photo-$z$ code.
-Once again \textsc{FlexZBoost} and \textsc{CMNN} score very well for the stacked $\hat{N}(z)$ metrics, as do \textsc{GPz} and \textsc{TPZ}.
-The CDE loss measures how well individual PDFs are estimated, and codes with a low CDE loss tend to have good $\hat{N}(z)$ estimates (though the reverse is not necessarily true).
-\textsc{FlexZBoost} is optimized to minimize CDE loss which may explain why the method has good ensemble metrics as well.
-Note from Table~\ref{tab:cdeloss} that both \textsc{FlexZBoost} and \textsc{CMNN} have low CDE losses.
-Empirically, we have found that PIT RMSE is not as closely correlated to CDE loss as it is to the $N(z)$ statistics.
-As CDE loss is a better measure of individual redshift performance, rather than ensemble distribution performance, this statistic is a better indicator of which codes will be most likely to perform well for science cases where single objects are employed.
-
-Table~\ref{tab:rmse} gives the root-mean-square-error (RMSE) statistics for both the PIT and N(z) estimators.
-The PIT value calculates the RMSE between the quantiles shown in the QQ plot in Figure~\ref{fig:pitqq} and the diagonal, while the N(z) calculates the RMSE between the cumulative distribution of the stacked $\hat{N}(z)$ and the true redshift distribution $N'(z)$.
-%\red{more about RMSE?
-Or remove it?
-It's barely described in Section~\ref{sec:metrics} and wasn't even referred to in the text until I added this on June 29th--SJS.}
-
-Table~\ref{tab:moments} lists the first three moments of the stacked $\hat{N}(z)$ distribution, including the moments of the ``truth'' distribution for comparison.
-Several codes are able to reproduce the mean and variance of the distribution to less than a per cent, while several codes do not, which may be a cause for concern, given that mean and variance of the redshift distribution are key properties in cosmological analyses.
-We note that this stated goal of the study as defined for participants was to accurately reproduce $p(z)$, the ``stacking'' of the probability distributions to estimate $\hat{N}(z)$ was not the focus as stated to the participants.
-This explains why some of the best-performing empirical codes in terms of $p(z)$ measures (e.~g.~\textsc{FlexZBoost}) do not do as well at reproducing $\hat{N}(z)$ moments.
-Had we defined a different parameter to optimize, in this case overall accuracy of $\hat{N}(z)$ rather than individual $p(z)$, would result in improved performance in a particular metric.
-That is, optimizing photo-$z$ performance for one metric does not automatically give optimal performance for other metrics.
-As previously stated, there are a variety of scientific use cases for photo-$z$'s in large upcoming surveys, and care must be taken in how the metrics used to optimize catalog photometric redshifts are defined as well as in how they are used.
-In addition, very few scientific use cases will employ the overall $\hat{N}(z)$ with no cuts, as we explore in this paper.
-We discuss more realistic tomographic bin selections that will be explored in a follow-up paper in Section~\ref{sec:futurework}.
-%\red{[mention that we are calculating moments for the entire sample, will look at fiducial tomographic bins in a follow-up paper (unless group and reviewers realy think we should include in this paper.]}
-%\red{FlexZBoost has some of the best $n(z)$ statistics, but the moments are not good.
-Why is this?
-Is it the over/underprediction of the high redshift part of the distribution?
-Discuss this after talking with Rafael.}
-
-%\subsection{Response of Individual Codes}
-%\label{sec:res:pz_indiv_codes}
-%\red{this may be incorporated in to 5.1 and 5.2 rather than live in its own section!}
-
-%\red{HOW CODES deal with negative fluxes, and magnitude uncertainties}
-
-%\red{overall conclusions of the Results section}
-
-\subsection{Interpretation of metrics}\label{sec:caution}
-%Caution on metrics and photo-$z$ codes}\label{sec:caution}
-%(Alex Malz, Sam Schmidt)
-
-Samples from accurate photo-$z$ posteriors should reproduce the space of $p(z, data)$.
-However, it is difficult to test this reconstruction given our data set, as the galaxy distributions arise from mock objects pasted on to an underlying dark matter halo catalogue with properties designed to match empirical relations, rather than being drawn from statistical distributions in redshift.
-In previous sections we have mentioned that optimizing for a specific metric does not guarantee good performance on other metrics, nor is there any guarantee that good performance by our metrics corresponds to \textit{accurate} photo-$z$ posteriors.
-%\aim{Explain that samples from accurate photo-$z$ posteriors should reconstruct the space of $p(z, data)$, a test we can't check with our datasets.}  
-In other words, we can construct photo-$z$ estimators that provide good coverage in many of our tests, but which have very little predictive power.
-
-The \trainz\ estimator, which assigns every galaxy a $p(z)$ equal to $N(z)$ of the training set as described in Section~\ref{sec:method:trainz}, is introduced as a ``null test'' to demonstrate this point via \textit{reductio ad absurdum}.
-\trainz\ outperforms all codes on the PIT-based metrics, and all but one code on the $N(z)$ based statistics.  
-Because our training set is perfectly representative of the test set, $N(z)$ should be identical for both sets down to statistical noise.
-%\textit{Explain the connection between \trainz\ and these metrics.}
-
-The CDE loss and point estimate metrics, however, successfully identify problems with \trainz.  
-As shown in Appendix~\ref{sec:pointmetrics}, \trainz has identical $ZPEAK$ and $ZWEIGHT$ values for every galaxy, and thus the photo-$z$s are constant as a function of spec-$z$s, i.e.
-a horizontal line at the mode and mean of the training set distribution respectively.
-The explicit dependence on the \it{individual} posteriors in the calculation of the CDE loss, described in Section~\ref{sec:CDE_loss}, distinguishes this metric from the other $p(z)$ metrics that test the overall ensemble of $p(z)$ distributions.
-With a representative training set, \trainz\ will score well on the ensemble metrics, but fails miserably for metrics tied to individual redshifts.
-We note that many of the ensemble-based metrics are prominent in the photo-$z$ literature despite their inability to identify problems such as those exemplified by \trainz.
-
-%\aim{Explain why CDE loss is different from other metrics (isn't fooled by \trainz) and note that the others have more of a presence in the literature despite their flaw of failing to identify this severe problem.}
-
-% While looking at one, or even most of our metrics would have given the impression that this estimator was nearly optimal, red flags in CDE loss and point estimates reveal the problems.  
-%[more caution on over-reliance on metrics, and making sure that metrics test all of what you want] \red{this definitely still needs work}
-In summary, context is crucial to interpreting metrics and defending against the likes of \trainz.
-The best photo-$z$ method is the one that most effectively achieves our science goals, not the one that performs best on a metric that does not accurately reflect those goals.
-In the absence of clear goals or the information necessary for a principled metric definition, we must think carefully before choosing a single metric
-
-%\begin{table*}  %%% DATA TABLE %%%
-%\caption{KS, CvM and AD statistics for each photo-$z$ code.} \label{tab:results}
-%PIT statistic
-%
-%\begin{tabular}{lrrrr}
-%\hline
-%\bf Photo-$z$ Code & \bf KS & \bf CvM & \bf AD & \bf AD \\
-%                   &        &         & $v=[0.05,0.95]$ & $v=[0.01,0.99]$        \\
-%\hline
-%\textsc{ANNz2} 		& $0.0200$ &   $52.25$ &   $583.0$ &   $759.2$ \\
-%\textsc{BPZ} 		& $0.0388$ &  $280.79$ &  $1361.4$ &  $1557.5$ \\
-%%\textsc{Delight} 	& $0.1380$ & $3040.59$ & $13880.2$ & $17761.4$ \\
-%\textsc{Delight}    & $0.0876$ & $1075.17$ & $4407.8$  & $6167.5$\\
-%%\textsc{EAZY} 		& $0.0690$ &  $994.19$ &  $3479.5$ &  $4158.8$ \\
-%\textsc{EAZY}       & $0.0723$ & $1105.58$ & $3475.0$  & $4418.6$\\
-%%\textsc{FlexZBoost} & $0.0260$ &   $70.58$ &   $595.0$ &   $533.0$ \\
-%\textsc{FlexZBoost} & $0.0240$ &   $68.83$ &   $567.9$ &   $478.8$\\
-%\textsc{GPz} 		& $0.0449$ &  $258.56$ &   $904.8$ &  $1414.8$ \\
-%\textsc{LePhare} 	& $0.0663$ &  $473.05$ &    $85.9$ &   $383.8$ \\
-%\textsc{METAPhoR} 	& $0.0438$ &  $298.56$ &   $248.6$ &   $715.5$ \\
-%%\textsc{CMNN} 		& $-$ & $-$ & $-$ & $-$ \\
-%\textsc{CMNN}         & $0.0795$  &$1011.11$ &  $3908.2$ &  $6307.5$ \\
-%\textsc{SkyNet} 	& $0.0747$ &  $763.00$ &  $2740.2$ &  $4216.4$ \\
-%\textsc{TPZ} 		& $0.1138$ & $1801.74$ &  $9046.2$ & $10565.7$ \\
-%%\textsc{Frankenz}	& $-$ & $-$ & $-$ & $-$ \\
-%\hline
-%\end{tabular}
-
-%Stacked $n(z)$ statistic
-%
-%\begin{tabular}{lrrrr}
-%\hline
-%\bf Photo-$z$ Code & \bf KS & \bf CvM & \bf AD & \bf AD \\
-%                   &        &         & $z=[0.002,2.0]$ & $z=[0.0,2.0]$        \\
-%\hline
-%\textsc{ANNz2} 		& $0.0237$ &  $25.676$ &   $276.4$ &   $276.0$ \\
-%\textsc{BPZ} 		& $0.0131$ &  $17.587$ &   $169.0$ &   $170.9$ \\
-%%\textsc{Delight} 	& $0.0184$ &  $18.975$ &   $151.5$ &   $151.5$ \\
-%\textsc{Delight}    & $0.0256$ &  $42.341$ &   $253.3$ &   $253.2$ \\
-%%\textsc{EAZY} 		& $0.0434$ & $167.329$ &   $811.3$ &   $817.0$ \\
-%\textsc{EAZY}       & $0.0441$ & $169.977$ &   $827.7$ &   $833.3$ \\
-%%\textsc{FlexZBoost} & $0.0065$ &   $2.764$ &    $24.7$ &    $24.6$ \\
-%\textsc{FlexZBoost} & $0.0128$ &   $7.007$ &  $127.02$ &   $126.9$\\
-%\textsc{GPz} 		& $0.0168$ &  $25.605$ &   $307.1$ &   $307.1$ \\
-%\textsc{LePhare} 	& $0.0254$ &  $58.900$ &   $477.4$ &   $472.0$ \\
-%\textsc{METAPhoR} 	& $0.0350$ &  $60.790$ &   $671.9$ &   $672.3$ \\
-%%\textsc{CMNN} 		& $-$ & $-$ & $-$ & $-$ \\
-%\textsc{CMNN}         & $0.0043$ & $0.538$   &   $5.6$   &      $5.8$ \\
-%\textsc{SkyNet} 	& $0.0483$ & $385.830$ &  $2119.4$ &  $2110.3$ \\
-%\textsc{TPZ} 		& $0.0126$ &   $9.483$ &   $110.7$ &   $110.6$ \\
-%%\textsc{Frankenz}	& $-$ & $-$ & $-$ & $-$ \\
-%\hline
-%\end{tabular}
-%\end{table*}
-
-\begin{table}  %%% DATA TABLE %%%
-\centering
-\caption{CDE loss statistic for each photo-$z$ code.} \label{tab:cdeloss}
-\begin{tabular}{lr}
-\hline
-\bf Photo-$z$ Code & \bf CDE Loss \\
-\hline
-\textsc{ANNz2} 		& $-6.88$ \\
-\textsc{BPZ} 		& $-7.82$ \\
-%\textsc{Delight} 	& $-4.06$ \\
-\textsc{Delight}    & $-8.33$\\
-%\textsc{EAZY} 		& $-7.97$ \\
-\textsc{EAZY}       & $-7.07$ \\
-%\textsc{FlexZBoost} & $-11.51$ \\
-\textsc{FlexZBoost} & $-10.60$\\
-\textsc{GPz} 		& $-9.56$ \\
-\textsc{LePhare} 	& $-1.66$ \\
-\textsc{METAPhoR} 	& $-6.28$ \\
-%\textsc{CMNN} 		& $-$ \\
-\textsc{CMNN}         & $-10.43$ \\
-\textsc{SkyNet} 	& $-7.89$ \\
-\textsc{TPZ} 		& $-9.55$ \\
-\hline
-\trainz		& $-0.83$ \\
-%\textsc{Frankenz}	& $-$  \\
-%\hline
-\end{tabular}
-\end{table}
-
-\begin{table}
-\setlength{\tabcolsep}{2pt}
-\begin{center}
-\caption{Root-Mean-Square-Error (RMSE) statistics for the eleven photo-z codes for both PIT and $\hat{N}(z)$ distributions.}\label{tab:rmse}
-\begin{tabular}{lcc}
-\hline
-\hline
-Root-Mean-Square-Error \\
-(RMSE) statistics \\
-\hline
-Photo-z Code & PIT RMSE & N(z) RMSE\\
-\hline
-\textsc{ANNz2}      & 0.019 & 0.0054\\
-\textsc{BPZ}        & 0.032 & 0.0050\\
-\textsc{Delight}    & 0.111 & 0.0056\\
-\textsc{EAZY}       & 0.054 & 0.0102\\
-\textsc{FlexZBoost} & 0.021 & 0.0022\\
-\textsc{GPz}        & 0.048 & 0.0042\\
-\textsc{LePhare}    & 0.028 & 0.0062\\
-\textsc{METAPhoR}   & 0.064 & 0.0081\\
-\textsc{CMNN}         & 0.108 & 0.0009\\
-\textsc{Skynet}     & 0.054 & 0.0144\\ 
-\textsc{TPZ}        & 0.082 & 0.0031\\
-\hline
-\trainz		& 0.0025 & 0.0013\\
-\end{tabular}
-\end{center} 
-\end{table}
-
-%\begin{table}
-%\begin{center}
-%caption{RMSE Statistics for the 11 Photo-z codes for DC1 $i<25.3$ ``gold'' sample}\label{tab:pitrmse}
-%%\begin{tabular}{|l|c|c|c|c|}
-%\begin{tabular}{lc}
-%\hline
-%\hline
-% PIT RMSE statistics \\
-%\hline
-%Photo-z Code & PIT RMSE\\
-%\hline
-%\texttt{ANNz2}      & 0.019\\
-%\texttt{BPZ}        & 0.032\\
-%\texttt{Delight}    & 0.111\\
-%\texttt{EAZY}       & 0.054\\
-%\texttt{FlexZBoost} & 0.021\\
-%\texttt{GPz}        & 0.048\\
-%texttt{LePhare}    & 0.028\\
-%\texttt{METAPhoR}   & 0.064\\
-%\texttt{NN}         & 0.108 \\
-%\texttt{Skynet}     & 0.054\\ 
-%texttt{TPZ}        & 0.082\\
-%\hline
-%\end{tabular}
-%\end{center} 
-%\end{table}
-%
-%\begin{table}
-%\begin{center}
-%\caption{RMSE Statistics for the 11 Photo-z codes for DC1 $i<25.3$ ``gold'' sample}\label{tab:nzrmse}
-%%\begin{tabular}{|l|c|c|c|c|}
-%\begin{tabular}{lc}
-%\hline
-%\hline
-% N(z) RMSE statistics \\
-%\hline
-%hoto-z Code & N(z) RMSE\\
-%\hline
-%\texttt{ANNz2}      & 0.0054\\
-%\texttt{BPZ}        & 0.0050\\
-%\texttt{Delight}    & 0.0056\\
-%\texttt{EAZY}       & 0.0102\\
-%\texttt{FlexZBoost} & 0.0022\\
-%\texttt{GPz}        & 0.0042\\
-%\texttt{LePhare}    & 0.0062\\
-%\texttt{METAPhoR}   & 0.0081\\
-%\texttt{NN}         & 0.0009\\
-%\texttt{Skynet}     & 0.0144\\ 
-%\texttt{TPZ}        & 0.0031\\
-%\end{tabular}
-%\end{center} 
-%\end{table}
-
-\begin{table}
-\setlength{\tabcolsep}{2pt}
-\caption{Moments of the stacked $\hat{N}(z)$ distribution}\label{tab:moments}
-\begin{tabular}{lccc}
-\hline
-\hline
- \multicolumn{4}{l}{Stacked $n(z)$ Moments} \\
-\hline
-              & 1st Moment & 2nd Moment & 3rd Moment \\
-\textsc{Truth}     & 0.701      &   0.630    & 0.671  \\
-\hline
-Photo-z Code       & 1st Moment & 2nd Moment & 3rd Moment\\
-\hline
-\textsc{ANNz2}     & 0.702      & 0.625      & 0.653    \\ 
-\textsc{BPZ}       & 0.699      & 0.629      & 0.671    \\ 
-\textsc{Delight}   & 0.692      & 0.609      & 0.638    \\ 
-\textsc{EAZY}      & 0.681      & 0.595      & 0.619    \\ 
-\textsc{FlexZBoost}& 0.694      & 0.610      & 0.631    \\ 
-\textsc{GPz}       & 0.692      & 0.605      & 0.619    \\ 
-\textsc{LePhare}   & 0.718      & 0.668      & 0.741    \\ 
-\textsc{METAPhoR}  & 0.705      & 0.628      & 0.657    \\ 
-\textsc{CMNN}        & 0.701      & 0.628      & 0.667    \\
-\textsc{Skynet}    & 0.743      & 0.708      & 0.797    \\
-\textsc{TPZ}       & 0.700      & 0.619      & 0.643    \\ 
-\hline
-\trainz	   & 0.699 		& 0.627 	& 0.666 \\
-\end{tabular}
-\end{table}
-
-\section{Summary and Discussion}
-\label{sec:discussion}
-%(Sam Schmidt, Ofer Lahav, Jeff Newman, Alex Malz, Eve Kovacs, Tony Tyson, Tina Peters)
-In this paper we presented results evaluating the photometric redshift PDF computation for eleven photo-$z$ codes.
-As discussed in Section~\ref{sec:metrics} the $p(z)$ should accurately reflect the relative likelihood as a function of redshift for each galaxy.
-All codes were provided a set of representative training data and tested on an idealized set of model galaxies with high signal-to-noise and photometry with no confounding effects due to blending, instrumental effects, the night sky, etc...
-included.
-The goal was not to determine a ``best'' photo-$z$ code: in many ways, this was a baseline test of a ``best case scenario'' to predict the expected photo-$z$ performance if a stage IV dark energy survey was to obtain complete training samples and perfectly calibrated their multi-band photometry.
-Given these idealized conditions, any deficiencies observed in a photo-$z$ code's performance should be a cause for concern, and may be evidence in a problem with either/both of the specific code implementation or the underlying algorithm.
-In order to meet the stringent LSST requirements on photo-$z$ performance, identifying and correcting such problems is an important first step before tackling more realistic data in future challenges.
-Most of the codes tested performed well, however, several did not meet the stringent goals that have been laid out for LSST photometric redshift performance.
-This is a cause for concern, given the idealized conditions, and the individual code responses will be studied in detail moving forward.
-One obvious trend in several of the codes tested was an overall over or underprediction of the widths of $p(z)$, as evidenced by the QQ plots and PIT histograms shown in Fig.~\ref{fig:pitqq}.
-A more careful tuning of bandwidth or smoothing during the validation process appears to be necessary for many of the machine learning based codes in order to improve the accuracy of $p(z)$.
-For narrow peaked $p(z)$ the parameterization of the PDF as evaluated on a fixed redshift grid could also have contributed to some overestimates of $p(z)$ width simply due to the finite resolution.
-After evaluating resuts such as those presented in \citet[]{Malz:qp}, in future analyses we plan to switch from a fixed grid to quantile-based storage of $p(z)$ in order to more efficiently and accurately store redshift PDF results.
-
-Another important factor to keep in mind when examining the results presented in this paper is the fact that they are at some level dependent on the metrics that we aim to optimize: in this case code participants were asked to submit their optimal measures of an accurate $p(z)$, so participants used the training/validation data to optimize their codes accordingly.
-Had we, instead, asked for an optimal $\hat{N(z)}$ the resulting metrics would be different for most, if not all, of the codes, as they would optimize toward a different goal.
-Specific metric choice can affect which codes are among the ``best'' codes.
-As stated earlier, there are cosmological science cases that require either individual galaxy photo-$z$ measures, or ensemble $\hat{N}(z)$ measures.
-We must be aware of that the optimal method for one is not necessarily optimal for the other, and in fact several photo-$z$ algorithms may be necessary in the final cosmological analysis in order to satisfy the requirements of all science use cases.
-The example of the simple \trainz\ estimator described in Section~\ref{sec:caution} shows a simple model with a $p(z)$ that is unrealistic for individual objects can still score very well on many of our metrics.
-It is important to look at {\it all} metrics, and keep in mind what information each metric conveys.
-%\red{mention 'null' and how 'dumb' algorithms can score well on metrics, need to keep all metrics in mind}
-%\aim{[reiterate the meaning of $p(z)$ and the goal -- emphasize finding ``best'' code in terms of impact of assumptions underlying the method \textit{or} establishing methodology for the realistic case of biased prior information and other science goals]}
-%\red{[the challenges faced in this project]}
-%\red{mention z$<$2, missing some important degeneracies, new sim will go to higher z}
-%\red{no quality cuts included, could identify outliers, but could also induce biases}
-%\red{[how optimization defined has impact, here we optimized p(z), not n(z), could be science-case specific if stringent requirements are needed, though that may be a computational/storage challenge if too many use cases needed.]}
-%\red{some discussion of output parameterization, limitations of 200 point fixed grid, in future switch to something like quantiles (cite qp paper)}
-We re-emphasize that the dataset tested was quite idealized, and discuss enhancements that will be added in future simulations to test photo-$z$ codes on increasingly realistic conditions in the following section.
-
-\subsection{Future work}
-\label{sec:futurework}
-The work presented in this paper is only the first step in characterizing current photo-$z$ codes and moving toward an improved photometric redshift estimator.
-This initial paper explored code performance in idealized conditions with perfect catalog-based photometry and representative training data.
-As mentioned in Section~\ref{sec:stackedmetrics} for the stacked $N(z)$ metrics we examined only the entire galaxy population with no selections in either photo-$z$ ``quality" or redshift.
-The cosmological analyses for weak lensing and large scale structure based measures plan to break galaxy samples into tomographic redshift bins, using photo-$z$ $p(z)$ to infer the redshift distribution for each bin.
-The specific selection used to determine these bins, both algorithmically and the specific bin boundaries, could induce biases due to indirect selections inherent in the photo-$z$ or other bin selection parameters.
-The effects of tomographic bin selection will be explored in a dedicated future paper.
-\red{[are there any references for this?
-I remember Gary Bernstein talking about this at a photo-z workshop in Japan, but I don't know that it was published.
-I believe Michael Troxel has discussed this as well.]}
-We also plan to propagate the uncertainties measured in a set of fiducial tomographic redshift bins in order to estimate impact on cosmological parameter estimation.
-%\item \red{cosmological parameters in DC2}
-%\item \red{mention z$<$2, missing some important degeneracies, new sim will go to higher z}
-%\red{no quality cuts included, could identify outliers, but could also induce biases}
-
-In future papers we will add more and more complexity to our simulated data in order to test photo-$z$ algorithms in increasingly realistic conditions.
-The most pressing concern is the impact of incomplete spectroscopic training samples.
-As discussed extensively in \citet{Newman:2015} a representative set of spectroscopically confirmed galaxies spanning the full range of both redshift and apparent magnitude is necessary as a training set to characterize the mapping from broad-band fluxes to photometric redshifts.
-However, due to a combination of factors due to both the galaxy SEDs and limitations of spectrographic instruments, redshift samples are known to be systematically incomplete, where certain galaxy types and redshift intervals fail to yield a redshift even at the longest integration times on current and near-future instruments.
-The more representative the training data, the better the performance of photo-$z$ algorithms will be.
-Current and upcoming surveys are putting in significant effort into obtaining these training samples \citep[e.~g.\,][]{Masters:2017}, however we still expect significant incompleteness for LSST-like samples, particularly at faint magnitudes.
-One major focus of an upcoming {\it LSST Dark Energy Science Collaboration Photo-z Working Group} data challenge is to produce a realistically incomplete training set of spectroscopic galaxies, modeling the performance of spectrographs, emission-line properties, and expected signal-to-noise to determine which galaxies will fail to yield a secure redshift.
-In addition to outright redshift failures we will model the inclusion of a small number of falsely identified secure redshifts where misidentified emission lines or noise spikes cause an incorrect redshift solution to be marked as a high quality identification.
-Even sub-per cent level contamination by false redshifts can impact photo-$z$ solutions at levels comparable to the stringent  requirements of some LSST science cases.
-We expect different systematics to occur in different photo-$z$ codes in response to training on incomplete data, particularly some of the machine learning methods.
-The response of the codes will inform future directions of code development.
-
-This initial paper explored a data set that was constructed at the catalog level, with no inclusion of the complications that come from measuring photometry from images.
-Future data challenges will move to catalogs constructed from mock images, including effects that will have great impact on photo-$z$ measurements.
-Object blending will be a major area of investigation, as the mixing of flux from multiple objects and the resultant change in measured colours is predicted to affect a large fraction of LSST galaxies \citep{Dawson:2016}, and will be one of the major contributing systematics for photo-$z$'s.
-Inclusion of differing observing conditions (seeing, clouds, variations in filter curves, Galactic dust, ...), as well as models for instrumental and system effects, sky masks, will all impact object photometry, and will be explored in the upcoming data challenge and their impacts described in upcoming papers.
-All underlying SEDs were parameterized as a weighted combination of five basis SEDs, with no additional accounting for host galaxy dust obscuration beyond what was encoded in the basis templates.
-This, in effect, limited the simulation to a very simple model of internal obscuration.
-Future simulations will include a more complicated and realistic treatment of host galaxy dust.
-
-The underlying simulation used in this work was based on a light-cone constructed to a maximum redshift of $z=2$.
-LSST imaging after 10 years of observations will include a significant number of $z>2$ galaxies in expected cosmology samples, and their inclusion does have potential significant implications for photo-$z$ measures: the high redshift galaxies lie at fainter apparent magnitudes and can have anomalous colours due to evolution of stellar populations and the shift to rest-frame magnitudes probing UV features of the underlying SED.
-More importantly, one of the most common ``catastrophic outlier'' degeneracies observed in deep photometric samples occurs when the Lyman break is mistaken for the Balmer break, leading to multiple redshift solutions at $z\sim0.2-0.3$ and $z\sim2-3$ \citep{Massarotti:2001}.
-This degeneracy, along with other potential degeneracies, are currently not covered by the limited redshift range of this initial paper, which could mean that we are not probing the full range of potential extreme outlier populations and how our photo-$z$ estimators respond to them.
-Extending simulations to include the high-redshift galaxy population will be a priority in future data challenges.
-
-%\item \red{inclusion of incompleteness effects on training sample, ``wrong redshifts'', filter curves}
-%\item \red{inclusion of photometric errors, image based effects, blending, lensing, sky, mask, etc...}
-
-In this study we have not accounted for the presence of Active Galactic Nuclei (AGN) contributions to galaxy fluxes.
-In some cases, AGN will be easily identified from the colors and morphologies, i.e.
-the case of the brightest quasars where the nuclear activity outshines the host galaxy, and numerous studies have utilized color selection to create large samples of quasars \citep[e.g.][]{Richards:06,Maddox:08,Richards:15}.
-In current deep fields, similar in depth to what we expect from LSST, variability information and multi-wavelength data have been critical to not only identify AGN dominated galaxies, but also obtain more accurate photometric redshifts \citep[e.g][]{Salvato:11}.
-
-In addition to AGN dominated galaxies, those with lower levels of nuclear activity present a more insidious problem, where AGN features may not be apparent, but the colors and other host galaxy properties are perturbed relative to galaxies with an inactive nucleus.
-In such cases, the presence of the AGN may induce a bias if the template SEDs or empirical datasets do not include low-level AGN counterparts.
-For LSST, we will need to identify and obtain accurate photometric redshifts of all types of AGN for a range of science goals, whether it is to eliminate such objects from cosmology experiments, or to use them with confidence, all the way through to understanding galaxy evolution and the role that AGN may play in influencing galaxy properties over cosmic time.
-
-A promising route to classifying and obtaining accurate photometric redshifts for the AGN population is by combining machine learning with template-fitting techniques, as has recently been demonstrated by \citet{Duncan:18} for radio-selected AGN.
-This is because AGN are relatively easy to obtain spectroscopic redshifts for over all redshifts due to the strong emission lines that they exhibit, allowing very good training sets for machine learning algorithms to use.
-Whereas for those galaxies where the AGN is sub-dominant the galaxy templates are still adequate for obtaining reasonable photometric redshifts.
-
-In addition to these improvements, the DESC Photo-$z$ group plans to look at all potential methods to combine the results from multiple photo-$z$ codes to improve $p(z)$ accuracy, similar to the work presented in \citet{Dahlen:13,Carrascokind:14,Duncan:18}.
-Taking advantage of multiple algorithms that use observables in slightly different ways has shown promise, however we must be very conscious of whether a potential combination properly treats the covariance between the methods, given that they are estimating quantities based on the same underlying observables.
-Several science cases wish to estimate physical quantities along with redshift, for example galaxy stellar mass and star formation rate.
-Proper joint estimation of redshift and physical quantities requires an in depth understanding of galaxy evolution, and progress on accurate bivariate redshift probability distributions will go hand in had with progress on understanding galaxies themselves.
-Parameterization and storage of a complex 2-dimensional probability surface for potentially billions of galaxies (or even subsets of hundreds of thousand of particular interest) pose a potential challenge.
-These issues will be examined in another future paper.
-
-Finally, while this paper and future papers discussed above focus on photometric redshift codes and estimating accurate $p(z)$ from training data, we plan a separate, but complementary, project to examine calibration of the resultant redshifts via spatial cross-correlations \citep{Newman:2008}, which will be explored in a separate series of future papers.
-The overarching plan describing everything laid out in this section is described in more detail in the LSST DESC Science Roadmap (see Footnote in Section~\ref{sec:intro}).
-These plans will require significant effort, but they are necessary if we are to make optimal use of the LSST data for astrophysical and cosmological analyses.
-%\begin{enumerate}
-%\item \red{the combination of codes}
-
-%\item \red{$p(z,\alpha)$}
-%\item \red{tomographic analysis later on}
-%\red{this paper deals with training, future papers will also explore calibration via cross-correlation methods}
-%\end{enumerate}
-%\red{mention the SRM once again, say that the plans are extensive, but we do have a plan and a rough timeline.}
-
-\section*{acknowledgements}
-\red{[summary of effort breakdown], will need to be in authors.csv for auto-generation by start\_paper}
-
-\red{[personal funding sources]}
-
-\red{[NERSC computation acknowledgement}
-
-The LSST Dark Energy Science Collaboration acknowledges generous ongoing support from the agencies and institutes supporting the collaboration.
-These include the Institut National de Physique Nucl\'eaire et de Physique des Particules in France, the Science \& Technology Facilities Council in the United Kingdom, and the Department of Energy, the National Science Foundation, and the LSST Corporation in the United States.
-This research used resources of the National Energy Research Scientific Computing Center, a DOE Office of Science User Facility supported by the Office of Science of the U.S.
-Department of Energy under Contract No.
-DE-AC02-05CH11231.
-Part of this work was undertaken on STFC DiRAC HPC Facilities, funded by UK BIS National E-infrastructure capital grants, and on the UK particle physics grid, supported by the GridPP Collaboration.
-
 \appendix
@@
 \centering
-\includegraphics[width=0.9\textwidth]{figures/ZPEAK_szpz_threecolumn_12codes_navy.jpg}
-\includegraphics[width=0.9\textwidth]{figures/ZWEIGHT_szpz_threecolumn_12codes_navy.jpg}
+\includegraphics[width=0.9\textwidth]{fig/ZPEAK_szpz_threecolumn_12codes_navy.jpg}
+\includegraphics[width=0.9\textwidth]{fig/ZWEIGHT_szpz_threecolumn_12codes_navy.jpg}
 \caption{Point estimate photo-z's derived from the posteriors.
@@
 \end{table*}
-
-\bibliographystyle{mn2e}
-\bibliography{references}
-\end{document}

### 2018-08-19 paper/tex/data.tex
should now compile!

--- 2018-08-12 paper/main.tex
@@
-%\documentclass[a4paper]{article}
-\documentclass[usenatbib]{mn2e}
-%% Language and font encodings
-\usepackage[english]{babel}
-\usepackage[utf8x]{inputenc}
-\usepackage[T1]{fontenc}
-\usepackage{lineno}
-\linenumbers
-\modulolinenumbers[2]
-
-\usepackage{natbib}
-%% Sets page size and margins
-\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
-
-%% Useful packages
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{graphicx}
-\usepackage{subcaption}
-\usepackage[colorinlistoftodos]{todonotes}
-\usepackage[colorlinks=true, allcolors=blue]{hyperref}
-\usepackage{multirow}
-
-\newcommand       \apj  {ApJ}
-\newcommand       \apjl  {ApJL}
-\newcommand       \apjs  {ApJS}
-\newcommand       \aj  {AJ}
-\newcommand       \aap {A\&A}
-\newcommand       \mnras {MNRAS}
-\newcommand       \nat {Nature}
-\newcommand       \araa {ARA\&A}
-\newcommand       \physrep {Phys.~Reports}
-\newcommand       \prd {Phys. Rev. D}
-
-\newcommand{\trainz}{\textsc{trainZ}}
-
-\newcommand{\textul}{\underline}
-\newcommand*\mathinhead[2]{\texorpdfstring{$\boldsymbol{#1}$}{#2}}
-
-\newcommand{\red}[1]{\textcolor{red}{#1}}
-% Claim a color for your comments
-\newcommand{\aim}[1]{\textcolor{green}{#1}}%Alex Malz comments in green
-\newcommand{\blue}[1]{\textcolor{blue}{#1}}%unknown person comments in blue
-\definecolor{scc}{rgb}{0.0, 0.26, 0.15}
-\newcommand{\scc}[1]{\textcolor{scc}{#1}}%Stefano Cavuoti Comments in dark green
-\definecolor{newpink}{rgb}{0.858, 0.188, 0.478}
-\newcommand{\erfan}[1]{\textcolor{newpink}{#1}} % Erfan Nourbakhsh comments in pink
-\newcommand{\jan}[1]{\textcolor{orange}{#1}}%Alex Malz comments in green
-
-\def\X{{\mathbf{X}}}
-\def\x{{\mathbf{x}}}
-\def\E{{\mathbb E}}
-
-\title{An assessment of photometric redshift PDF performance in the context of LSST}
-% Alternate title proposal: Assessing the characterization of photometric redshift uncertainties in the regime of LSST
-
-\author{LSST-DESC Photometric Redshift Working Group}
-
-\begin{document}
-\maketitle
-
-\begin{abstract}
-Photometric redshift (photo-$z$) probability distribution functions (PDFs) are a key data product of nearly all upcoming galaxy imaging surveys.
-However, the photo-$z$ PDFs resulting from different techniques are not in general consistent with one another, and an optimal method for obtaining an accurate PDF remains unclear.  
-We present the results of an initial study of the Large Synoptic Survey Telescope Dark Energy Science Collaboration (LSST-DESC), the first in a planned series of papers testing multiple photo-$z$ codes on simulations of upcoming LSST galaxy photometry catalogues.
-This initial test evaluates photo-$z$ algorithms in the presence of representative training data and in the absence of several common sources of systematic errors that affect the procedures by which photo-$z$ PDFs are derived.
-The photo-$z$ PDFs are evaluated using multiple metrics including the Kolmogorov-Smirnoff statistic, Cramer-von Mises statistic, Anderson-Darling statistic, Kullback-Leibler divergence, $N(z)$ moments, quantile-quantile plots and probability integral transform.
-We observe several trends, including an overall over/under-prediction in the broadness of the PDFs for several of the codes.
-A careful accounting of all systematics discovered will be necessary for the codes employed in upcoming analyses in order to achieve unbiased cosmological measurements.
-\end{abstract}
-\begin{keywords}
-galaxies: distances and redshifts -- galaxies: statistics -- methods: statistical
-\end{keywords}
-%\noindent comments color-code:\\
-%\red{SJS},
-%\aim{AIM},
-%\scc{SC},
-%\textcolor{brown}{MB},\jan{JAN}
-%\blue{unknown}\\
-%\red{note: not all red text is mine, at some point various people were using red --SJS.}
-
-% Questions left unanswered:
-% 1. Which journal will this paper be published? 
-% Answer: AJ / ApJ / MNRAS (no charges, but not sure if it will accept a paper based on a DC1 effort and the associated simulations)
-% 2. What is the meaning of $p(z)$ in different codes? (Ofer)
-% Answer: ?
-%
-
-\section{Introduction}
-\label{sec:intro}
-
-%(Sam Schmidt, Ofer Lahav, Jeff Newman, Alex Malz, John Soo, Tony Tyson)
-
-%\jan{Is the intention to keep these parenthetical author sets?
-I find it a bit offputting in the text.
-It might be better to put this for sections in the acknowledgements section, and for individual codes/statistics specify who ran/developed it with a sentence.}\red{I was just about to comment these out before finalizing the internal review draft of the paper.
-They are in place to remind me to contact a few people who still have not filled out their authors.csv info.}
-
-Large-scale photometric galaxy surveys are entering a new era with currently or soon-to-be running Stage III and Stage IV dark energy experiments like the Dark Energy Survey \citep[DES,][]{Abbott:05}, the Kilo-Degree Survey \citep[KiDS,][]{de_Jong:13}, Large Synoptic Survey Telescope \citep[LSST,][]{Abell:09}, Euclid \citep{Laureijs:11}, and Wide-Field Infrared Survey Telescope \citep[WFIRST,][]{Green:12}.
-The move to imaging based surveys, rather than spectroscopic based, for cosmological measurements makes proper understanding of photometric redshifts (``photo-$z$'s'') of paramount importance, as cosmological distance measures for statistical samples are directly dependent on photo-$z$ measurements.
-
-The unprecedented sample size of LSST galaxies, expected to number several billion for the main cosmological sample, necessitates stringent constraints on photo-$z$ accuracy if systematic errors are not to dominate the statistical errors.
-The LSST Science Requirements Document (SRD)\footnote{available at \url{https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17}} lists the photometric redshift goals for a magnitude limited sample with $i<25$ as: root-mean-square error with a goal of $\sigma_z<0.02(1+z)$; $3\sigma$ ``catastrophic outlier'' rate below $10\%$; bias below $0.003$\footnote{
-Note that at the time the SRD was written, these goals were stated in terms of a photo-$z$ point estimate for each galaxy, as was standard in many previous studies, while in this paper we emphasize the importance of using a full photo-$z$ PDF.}.
-
-The tremendous size of LSST's galaxy catalog will be enabled by its exceptional depth, pushing to fainter magnitudes and deeper imaging and including galaxies of lower luminosity and higher redshift than ever before.
-In addition to the contribution of low signal-to-noise photometry to the systematic error of photo-$z$s, these populations introduce major physical degeneracies, for example the Lyman break/Balmer break degeneracy, that were not present in the populations covered in previous large area surveys like the Sloan Digital Sky Survey \citep[SDSS,][]{York:00} and the Two Micron All Sky Survey \citep[2MASS,][]{Skrutskie:06}.
-In order to meet LSST's demanding error budget, it will be necessary to fully characterize those degeneracies wherein multiple redshift solutions have comparable likelihood.
-
-There is often a desire to have a single valued ``point-estimate'' redshift for an individual galaxy.
-However, the complex, non-linear (and often non-unique) nature of the mapping between broad band fluxes and redshift means that a single value is unable to capture the full redshift information encoded in a galaxy's magnitudes.
-For example, a common point-estimate for a template-based method is taking the highest likelihood solution as the point photo-$z$.
-A single valued redshift ignores degenerate redshift solutions of lower probability, potentially biasing photometric redshift estimates both for individual galaxies and ensemble distributions.
-Storing more information is necessary, most often photo-$z$ codes output the redshift probability density function (PDF), also often referred to as $p(z)$, describing the relative likelihood as a function of redshift.
-%\red{should we move some of AIM's p(z) vs p(z|d,I) to here?}
-Early template methods such as \citet{Fernandezsoto:99} converted relative $\chi^2$ values of template spectra to likelihoods to estimate $p(z)$.
-Soon after, codes such as \citet{Benitez:00} added a Bayesian prior and output a posterior probability distribution.
-While many early machine learning based algorithms focused on a point-estimate, \citet{Firth:03} used a neural net with 1000 realizations scattered within the photometric errors to estimate a $p(z)$.
-As more groups began to employ photometric redshifts in their cosmological analyses, realization that point-estimate photo-$z$'s were inadequate for precision cosmology measurements \citep{Mandelbaum:2008}.
-From around this point onward, most photo-$z$ algorithms have attempted to implement some estimate of the overall redshift probability in their outputs, and some surveys began supplying a full $p(z)$ rather than a simple redshift point-estimate and error \citep[e.~g.~][]{de_Jong:17}.
-
-%\aim{We should spend a whole paragraph establishing that photo-$z$ PDFs are a thing, i.e.
-review the motivation for PDFs over point estimates, basic definition, early methods papers, surveys that have published them, etc.}
-%\scc{I agree with Alex, we should introduce here what a pdf is and which survey have already published some pdfs (KiDS dr3 for example) or will produce them (LSST, Euclid etc.)}
-%\red{See above paragraph, does this adequately address this?--SJS}\scc{fine for me}
-
-There are numerous techniques for deriving photo-$z$ PDFs from photometry, yet no one method has yet been established as clearly superior.
-Quantitative comparisons of photo-$z$ methods have been made before.  
-The Photo-$z$ Accuracy And Testing \citep[PHAT,][]{Hildebrandt:10} effort focused on point estimates derived from many photometric bands. 
-DES compared several codes for point estimates \citep{Sanchez:14} and a summary statistic of photo-$z$ interim posteriors for tomographically binned galaxy subsamples \citep{Bonnett:16}. 
-This paper is distinguished by its inclusion of metrics of photo-$z$ interim posteriors themselves and consideration of both classic and state-of-the-art photo-$z$ algorithms, comparing the performance of several of the most widely employed codes as well as some that have been developed only recently on the basis of metrics appropriate for a probabilistic data product.
-The results presented in this work are 
-%part of the Data Challenge 1 (DC1) 
- a major focus of the Photometric Redshift working group of the LSST Dark Energy Science Collaboration (LSST-DESC).
-This work is laid out in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}} as one of the critical activities to be completed in preparation for dark energy science analysis on the first year LSST data.
-This is the first of multiple papers by the working group, which will grow in sophistication.
-In this initial paper we focus on evaluating the performance of photometric redshift codes and their ability to produce accurate PDFs in the presence of representative training sets.
-This can be thought of as an initial test under near perfect conditions, before further complexities are added in future papers.
-Comparing the relative performance of the codes enables us to evaluate whether each code is using information in an optimal way, and may reveal enhancements in some codes and deficiencies in others, either in the fundamental algorithm, or in specific implementation.
-
-Certain science cases need redshift information on individual objects, e.~g.~identification of host galaxy redshift for supernova classification, or identifying potential cluster membership.  
-Other science cases need only ensemble redshift information; for instance current weak lensing techniques require the overall redshift distribution $N(z)$ for tomographic redshift samples, but do not need single galaxy estimates.
-In the case of the multiple types of probes of cosmology enabled by the
-LSST cosmology sample of several billion galaxies, the number of redshift bins and their photo-$z$ requirements vary with the specific probe;  2-point angular correlations benefit from many bins, while weak lensing probes do not (due to the wide lensing kernel).
-Large photometric surveys such as LSST must develop algorithms that meet the needs of all such science cases.  
-In order to meet these ambitious goals for photo-$z$ accuracy, every aspect of photo-$z$ estimation will have to be optimized: the algorithms employed, both template and machine-learning based (both in design and implementation); the spectroscopic data used as a training set for machine learning algorithms or to estimate template sets and train Bayesian priors; and probabilistic catalog compression schemes that balance information retention against limited storage resources.
-
-Before moving forward, we must address how the best methods may be unique to the performance metrics and science cases considered and what distinguishes photo-$z$ PDFs of different methods from one another.
-Though photo-$z$ PDFs are often written simply as $p(z)$, the PDF itself must be an interim posterior distribution $p(z | d, I)$, the probability of redshift conditioned on photometric data $d$ that has actually been observed and the prior information $I$ that guides how a redshift is extracted from the photometry.
-If we run multiple photo-$z$ codes on a single dataset, the photo-$z$ interim posteriors will not be identical because each code is based on assumptions in the form of an interim prior --- these assumptions form the premise for photo-$z$ estimation as a whole and are the only way to introduce differences in estimates of what would otherwise be a shared photo-$z$ posterior $p(z | d)$ regardless of the code used to obtain it.
-Though explicit knowledge of the interim prior is necessary to use photo-$z$ interim posteriors self-consistently in physical inference, the interim prior of a particular methodology is often implicit and not necessarily shared among all galaxies in the catalog.
-
-This paper therefore aims (1) to constrain the impact of the interim prior $I$ by separating it into a component $I_{H}$ representing the method itself and a component $I_{D}$ representing physical information, such as a training set or SED template library and (2) to present a procedure for evaluating the performance of photo-$z$ codes in generic tests that may include many more systematics in the interim prior $I$.
-In order to isolate the effects encapsulated by $I_{H}$ of variation between codes from issues with the training set or template library encapsulated by $I_{D}$, we use an identical set of simulated galaxies for every code and construct a template library and training sample that are {\it complete and representative} and shared among all codes; that is, our training sample for machine learning codes is drawn from the same underlying galaxy population as our test set, with no additional selections, and the SED library used for template-based codes is the same as the one used to generate the photometric data.
-We explore a number of performance metrics in this paper, not to make a conclusion regarding the superiority or even relative favorability of each code but to establish a method for comparing photo-$z$ PDFs derived by different methods.
-These test conditions set the stage for addressing in a future paper the crucial issue of incomplete and non-representative prior information.
-
-The outline of the paper is as follows: in \S\,\ref{sec:sims} we present the simulated data set; in \S\,\ref{sec:pzcodes} we describe the current generation codes employed in the paper;  in \S\,\ref{sec:metrics} we discuss the interpretation of photo-$z$ PDFs in terms of metrics of accuracy; in \S\,\ref{sec:results} we show our results and compare the performance of the codes; in \S\,\ref{sec:discussion} we offer our conclusions and discuss future extensions of this work.
-
 \section{The simulation and mock galaxy catalog}
@@
 Each simulated SED is replaced with a SED drawn from the distribution of empirical SEDs given by the SED Atlas of \citet{Brown:14} (diffusion maps method in development).}
-
-\section{Methods}
-\label{sec:pzcodes}
-
-%(Sam Schmidt, Rongpu Zhou, Ibrahim Almosallam, Eric Nuss, Johann Cohen Tanugi, John Soo)
-
-Here we outline the photo-$z$ PDF codes tested in this study.
-In total, eleven distinct codes are tested.
-This sample is not comprehensive, but does cover a broad range of current-generation codes.
-Both template-based and machine learning approaches are included and each are described separately in Secs.~\ref{sec:templatecodes} and \ref{sec:trainingcodes} respectively.
-The list of codes are summarized in Table.~\ref{tab:list_of_codes}.
-
-\begin{table*}  %%% DATA TABLE %%%
-\caption{List of photo-$z$ codes featured in this study. ML here means machine learning.} \label{tab:list_of_codes}\resizebox{\textwidth}{!}{
-\begin{tabular}{llll}
-\hline
-\bf Code 			& \bf Type & \bf Paper & \bf Website \\
-\hline
-\textsc{BPz} 		& template 			& \citet{Benitez:00}		& \url{http://www.stsci.edu/~dcoe/BPZ/} \\
-\textsc{EAZY} 		& template 			& \citet{Brammer:08}		& \url{https://github.com/gbrammer/eazy-photoz} \\
-\textsc{LePhare} 	& template 			& \citet{Arnouts:99}		& \url{http://www.cfht.hawaii.edu/~arnouts/lephare.html} \\
-\textsc{ANNz2} 		& ML 	& \citet{Sadeh:16}			& \url{https://github.com/IftachSadeh/ANNZ} \\
-\textsc{Delight} 	& ML/template 	& \citet{Leistedt:17}		& \url{https://github.com/ixkael/Delight} \\
-\multirow{ 2}{*}{\textsc{FlexZBoost}} & \multirow{ 2}{*}{ML} 	& \multirow{ 2}{*}{\citet{Izbicki:17}}		& \url{https://github.com/tpospisi/flexcode};\\ 
-					& 		&							& \url{https://github.com/rizbicki/FlexCoDE}\\
-%\textsc{Frankenz} 	& ML 	& \citet{Speagle:frankenz}	& \url{https://github.com/joshspeagle/frankenz} \\
-\textsc{GPz} 		& ML	& \citet{Almosallam:15b}	& \url{https://github.com/OxfordML/GPz} \\
-\textsc{METAPhoR} 	& ML 	& \citet{Cavuoti:17}		& \url{http://dame.dsf.unina.it}\\
-\textsc{CMNN} 		& ML 	& \citet{Graham:17}			& - \\
-\textsc{SkyNet} 	& ML 	& \citet{Graff:14}			& \url{http://ccpforge.cse.rl.ac.uk/gf/project/skynet/} \\
-\textsc{TPZ} 		& ML 	& \citet{Carrasco_Kind:13}	& \url{https://github.com/mgckind/MLZ} \\
-\hline
-\trainz 	& N/A	& See Section~\ref{sec:method:trainz} & \\
-\end{tabular}}
-\end{table*}
-
-%\red{The questions that must be answered for each code are\dots}
-The questions that must be answered for each code are: what unique features are included in the specific implementation that influence the output $p(z)$.
-What form of validation was performed with the training data, how were photometric uncertainties employed in the analysis, how were negative fluxes treated, what specific prior form was employed (for template based codes), or what specific machine learning architcture was used (for ML codes)?
-%\scc{which parameter (in a general meaning) have been used, which kind of validation has been performed (if any)}
-\subsection{Template-based Approaches}
-\label{sec:templatecodes}
-
-\subsubsection{BPZ}
-\label{sec:BPZ}
-%(Sam Schmidt)
-
-\textsc{BPZ}\footnote{\url{http://www.stsci.edu/~dcoe/BPZ/}} \citep[Bayesian Photometric Redshift,][]{Benitez:00} is a template-based photo-$z$ code that compares the expected colors ($C$) calculated for a set of spectral energy distribution (SED) types/templates ($T$) to the observed colors to calculate the likelihood of observing colors at each redshift for each type, $p(C|z,T)$.
-The code employs an empirically determined Bayesian prior in apparent magnitude ($m_0$) and SED-type.
-Assuming that the SED-types are spanning and exclusive, we can determine the redshift posterior $p(z|C,m_0)$ by marginalizing over all SED-types with a simple sum \citep[Eq.~3 from][]{Benitez:00}:
-
-\begin{equation} \label{eq:redshift_posterior}
-p(z|C,m_0)\,\propto\, \sum_{T}p(z,T|m_0)\,p(C|z,T)
-\end{equation}
-\noindent where the first term on the right-hand side is the Bayesian prior and the second term is the traditional likelihood.
-The prior is assumed to have the form: $p(z,T|m_0)=p(T|m_0)\,p(z|T,m_0)$, i.e.
-it parameterizes the prior as an evolving type fraction with apparent magnitude, combined with a prior on the expected redshift probability distribution as a function of both apparent magnitude and SED-type.
-
-In this paper we use \textsc{BPZ v 1.99.3}. The template set employed here is the set of $100$ discrete SEDs described in Section~\ref{sec:buzztemplates}
-%, and the $129$ Brown SEDs for the Galacticus dataset. 
-To keep the number of free parameters to a manageable level the SEDs in the training set are sorted by the rest-frame $u$-$g$ colour and split into three ``broad'' SED classes, equivalent to the \texttt{E}, \texttt{Sp} and \texttt{Im/SB} types in \citet{Benitez:00}.
-We assume the same functional form for the Bayesian priors as used by \citet{Benitez:00}, and utilize the training-set galaxies with known SED-type, redshift, and apparent magnitude to determine the type fractions and the best fit for the eleven free parameters of the prior.
-%For photo-$z$ point estimates we use the \texttt{Z\_B} output parameter. 
-For galaxies with negative flux in a measured band, the placeholder value is replaced with an estimate one $\sigma$ detection limit in that particular band, i.~e.~a value close to the estimated sky noise threshold.
-The type-marginalized $p(z)$ is generated by setting the parameter \texttt{PROBS\_LITE=TRUE} in the \textsc{BPZ} parameter file.
-
-\subsubsection{EAZY} \label{sec:eazy}
-%(Rongpu Zhou)
-
-\textsc{EAZY}\footnote{\url{https://github.com/gbrammer/eazy-photoz}} \citep[Easy and Accurate Photometric Redshifts from Yale,][]{Brammer:08} is a template-based photo-$z$ code that includes several features that improve on the basic $\chi^2$ fit used in many template codes.
-It can fit the observed photometry with SEDs created from a linear combination of a set of templates at each redshift, and the best-fit SED is found by simultaneously fitting one, two or all of the templates by minimizing $\chi^2$.
-The minimized $\chi^2(z)$ is then combined with an apparent magnitude prior to obtain the posterior redshift probability distribution, although some argue that this is not the mathematically correct way of calculating the posteriors.
-\textsc{EAZY} can also account for the uncertainties in the templates by adding an empirically derived template error in quadrature as a function of redshift to the flux errors.
-
-In this paper we use the all-templates mode, which fits the photometric data with a linear combination of the five basis templates.
-We employed the $5$ basis templates described in Section~\ref{sec:buzzard}, and set the template error to zero since these same templates were used to produce the simulated catalog photometry.
-%\red{The quality parameter \texttt{Qz} is...}
-The likelihoods are calculated on a 200-point redshift grid spanning $0\leq z \leq 2$, and include the application of a type-independent apparent magnitude prior estimated from the training data.
-
-\subsubsection{LePhare}\label{sec:lephare}
-%(C\'ecile Roucelle, Eric Nuss, Johann Cohen-Tanugi)
-
-\textsc{LePhare}\footnote{\url{http://www.cfht.hawaii.edu/~arnouts/lephare.html}}\citep[Photometric Analysis for Redshift Estimate,][]{Arnouts:99,Ilbert:06} is a photo-$z$ reconstruction code based on a $\chi^2$ template-fitting procedure.
-The observed colors are matched with the colours predicted from a set of spectral energy distribution (SED) which can be either synthetic or based on a semi-empirical approach.
-\textsc{LePhare} has been used to produce the COSMOS2015 photo-$z$ catalogue \citep{Laigle:16}.
-
-Each SED is redshifted in steps of $\Delta z = 0.01$ and convolved with the simulated LSST filter transmission curves (accounting for instrument efficiency).
-The opacity of the inter-galactic medium has been set to zero as no additional reddening has been included in the Buzzard simulations.
-The computed photo-$z$ is then the value that minimizes the merit function $\chi^2 (z,T,A)$ from \citet{Arnouts:99}:
-
-\begin{equation}
-\chi^2(z,T,A) = \sum_f^{N_f}\left(\frac{F^f_{\rm{obs}}−A \times F^f_{\rm{pred}}(T,z)}{\sigma^f_{\rm{obs}}}\right)^2 
-\end{equation}
-\noindent where $A$ is a normalization factor, $F^f_{\rm{pred}}(T,z)$ is the flux predicted for a template $T$ at redshift $z$.
-$F^f_{\rm{obs}}$ is the observed flux in a given band $f$ and $\sigma^f_{\rm{obs}}$ the associated observational error.
-The index $f$ refers to the considered band and $N_f$ is the total number of filters.
-
-In this paper we use \textsc{LePhare v 2.2}. 
-The set of templates used for fitting the photo-$z$'s are the 100 discrete Buzzard SED templates as described in section~\ref{sec:buzztemplates}.
-%No correction for the photometric zero-points was applied for this work. 
-The full $p(z)$ corresponds to the likelihoods calculated at each point on our $z$-grid.
-
-\subsection{Training-based Codes}
-\label{sec:trainingcodes}
-
-\subsubsection{ANNz2}
-\label{sec:annz2}
-%(John Soo)
-
-\textsc{ANNz2}\footnote{\url{https://github.com/IftachSadeh/ANNZ}} \citep{Sadeh:16} is a powerful package that has the ability to employ several machine learning algorithms, including artificial neural networks (ANN), boosted decision tree (BDT) and k-nearest neighbour (KNN).
-Using the Toolkit for Multivariate Data Analysis (TMVA) with ROOT\footnote{\url{http://tmva.sourceforge.net/}}, it can run multiple machine learning algorithms for a single training and outputs photo-$z$'s based on a weighted average of their performances.
-
-\textsc{ANNz2} is capable of producing both photo-$z$ point estimates and redshift posterior probability distributions $p(z)$, it could also conduct classifications and supports reweighting between samples.
-The PDFs are produced by propagating the intrinsic uncertainty on the input parameters and the uncertainty in the machine learning method to the expected photo-$z$ solution, averaged over multiple runs weighted based on the performance of each run.
-\textsc{ANNz2} presents its photo-$z$ uncertainty different from many codes by using the KNN method: it estimates the photo-$z$ bias between each object and a fixed number of nearest neighbours in parameter space, it then takes the $68$th percentile width of the distribution of the bias.
-This is based on the implication that objects with similar photometric properties would have similar uncertainties, and therefore the photometric errors of the inputs are not propagated into the code.
-
-In this study, \textsc{ANNz2 v.
-2.0.4} was used.
-The full PDF for each galaxy is also produced with a linear stepsize of $z=0.01$ for $0 \leq z \leq 2$.
-A set of $5$ ANNs with architecture $6:12:12:1$ ($6$ $ugrizy$ inputs, $2$ hidden layers with $12$ nodes each, and $1$ output) with different random seeds are used during each training.
-%Training was conducted on a subset of the given training sample which satisfies $i\le25.3$ as it would have improved the results compared to training on the entire sample.
-Half of the training set is used as a validation set to prevent overtraining.
-All training objects are set to have detected magnitudes, however the non-detections (mag$=-99$) in the testing set are replaced with the mean of that particular band.
-
-\subsubsection{Color-Matched Nearest-Neighbours}
-\label{sec:cmnn}
-%(Melissa Graham)
-
-The nearest-neighbours color-matching photometric redshift estimator (\textsc{CMNN}) is presented in \cite[][herafter G18]{Graham:17}.
-This method uses a training set of galaxies with known redshifts that has equivalent or better photometry as the test set in terms of quality and filter coverage.
-For each galaxy in the test set we identify a color-matched subset of training galaxies, choose one (e.g.
-the nearest-neighbour or a random selection), and use its known redshift as the photo-$z$.
-This color-matched subset is identified by first calculating the Mahalanobis distance $D_M$ in color-space between the test galaxy and all training-set galaxies: the difference between the test and a training set galaxy's color divided by the photometric error, summed over all colors (i.e., $u$-$g$, $g$-$r$, $r$-$i$, $i$-$z$, and $z$-$y$).
-Then, the threshold value for $D_M$ that define a good color match is set by the percent point function (PPF): for example, for $N_{\rm dof}=5$, PPF$=95$ per cent of all training galaxies consistent with the test galaxy will have $D_M < 11.07$ (where $N_{\rm dof}$, the number of degrees of freedom, is the number of colors).
-For a given test galaxy, the $p(z)$ is the normalized distribution of the true catalogue redshifts of this color-matched subset of training galaxies, and the standard deviation of the color-matched subset is used as the photo-$z$ uncertainty.
-
-We have applied the nearest-neighbours color-matching photometric redshift estimator described in G18 to the simulated data.
-Compared to its application in G18, there are some minor differences in the application of this estimator to the Buzzard catalogue.
-First, we do not impose non-detections on galaxies with a magnitude fainter than the expected LSST 10-year limiting magnitude or bright enough to saturate with LSST: {\it all} of the photometry for all the galaxies in the test and training sets are used for this experiment.
-Second, as in G18 we do apply an initial cut in color to the training set before calculating the Mahalanobis distance in order to accelerate processing, and also use a magnitude pseudo-prior to improve photo-$z$ estimates, but for both we have used different cut-off values that are appropriate for the Buzzard galaxies' colors and magnitudes.
-Third, we set different parameters for the identification of the color-matched subset of training galaxies and the selection of a photometric redshift estimate.
-In G18 we used a percent point function (PPF) value of 0.68 to identify the color-matched subset of training galaxies and used the redshift of nearest neighbour in color-space as the photo-$z$ estimate.
-These choices work well when the desire is to obtain accurate photo-$z$ estimates for most test-set galaxies, but does not return a robust $p(z)$ in all cases -- especially for galaxies that are bright and/or have few matches in color-space.
-Since a robust estimate of $p(z)$ is desired for this work we make several changes to our implementation of the \textsc{CMNN} photo-$z$ estimator.
-We continue to use a percent point function of ${\rm PPF} = 0.95$ to generate the subset of color-matched training galaxies, but weight them by the inverse of their Mahalanobis distance.
-This weighting maintains some of the accuracy that was previously achieved by simply using the nearest neighbour in color-space.
-We then use the weights to create the $p(z)$ instead of having the redshift of each color-matched training-set galaxy count equally.
-To obtain a robust estimate of the $p(z)$ for galaxies with a small number of color-matched training set galaxies, when this number is less than $20$ the nearest $20$ neighbours in color-space are used instead, and we convolve the $p(z)$ with a Gaussian with a standard deviation of:
-
-\begin{equation}
-\sigma = \sigma_{\rm train} \sqrt{({\rm PPF}_{20}/0.95)^2 -1}
-\end{equation}
-
-\noindent to appropriately broaden it so that the $p(z)$ for these test galaxies represents the enlarged PPF value associated with it.
-Overall, these three changes will yield poorer accuracy photo-$z$ compared to those presented in G18, but they will all have significantly more robust estimates of the $p(z)$, particularly for the brightest test galaxies.
-This is sufficient for this work because, as described in G18, the goal of the \textsc{CMNN} photo-$z$ estimator was never to provide the ``best'' (or even competitive) estimates in the first place, given its reliance on a deep training set, but rather to provide a means for direct comparisons between LSST photometric quality and photo-$z$ estimates.
-With this work we show how the input parameters should be set in order to return robust $p(z)$ estimates in addition to point value estimates.
-
-\subsubsection{Delight}
-\label{sec:delight}
-%(John Soo)
-
-\textsc{Delight}\footnote{\url{https://github.com/ixkael/Delight}} \citep{Leistedt:17} infers photo-$z$'s by using a data-driven model of latent SEDs and a physical model of photometric fluxes as a function of redshift.
-Generally, machine learning methods rely on representative training data with similar band passes, while template based methods rely on a complete library of templates based on physical models constructed.
-\textsc{Delight} is constructed in attempt to combine the advantages and eliminate the disadvantages of both template-based and machine learning algorithms: it constructs a large collection of latent SED templates (or physical flux-redshift models) from training data, with a template SED library as a guide to the learning of the model.
-The advantage of \textsc{Delight} is that it neither needs representative training data in the same photometric bands, nor does it need detailed galaxy SED models to work.
-
-This conceptually novel approach is done by using Gaussian processes operating in flux-redshift space.
-The posterior distribution on the redshift of a target galaxy is obtained via a pairwise comparison with training galaxies,
-
-\begin{equation}
-p(z|\mathbf{\hat{F}}) \approx \sum_i p(\mathbf{\hat{F}}|z,t_i)\, p(z|t_i)p(t_i),
-\end{equation}
-\noindent where $p(z|t_i)p(t_i)$ captures prior information about the redshift distributions and abundances of the galaxies, with $t_i$ denoting the galaxy template; while $p(\mathbf{\hat{F}}|z,t_i)$ is the posterior of noisy flux $\mathbf{\hat{F}}$ at redshift $z$.
-For each training-target pair, $p(\mathbf{\hat{F}}|z,t_i)$ is evaluated as follows:
-
-\begin{equation} \label{eq:delight_noisy}
-p(\mathbf{\hat{F}}|z,t_i) = \int p(\mathbf{\hat{F}}|\mathbf{F})\, p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)\, d\mathbf{F},
-\end{equation}
-where $p(\mathbf{\hat{F}}|\mathbf{F})$ is the likelihood function, it compares the noisy real flux $\mathbf{\hat{F}}$ with the noiseless flux $\mathbf{F}$ obtained from the linear combination of template models, carefully constructed to account for model uncertainties and different normalization of the same SED; while $p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)$ is the prediction of flux at a different redshift $z$ with respect to the training object with redshift $z_i$ and flux $\mathbf{\hat{F}}_i$.
-Eq.~\ref{eq:delight_noisy} is essentially the probability that the training and the target galaxies having the same SED but at a different redshift.
-The flux prediction $p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)$ of the training galaxy at redshift $z$ is modeled via a Gaussian process,
-
-\begin{equation} \label{eq:delight_gp}
-F_b \sim \mathcal{GP}\left( \mu^F,k^F \right),
-\end{equation}
-\noindent with mean function $\mu^F$ and kernel $k^F$, both imposed to capture expected correlations resulting from the known underlying physics (i.e., fluxes resulting from observing SEDs through filter response, and the SEDs being redshifted).
-The reader should refer to \citet{Leistedt:17} for further details.
-
-In this study, all $100$ ordered Buzzard templates, as described in Section~\ref{sec:buzztemplates}, were used in \textsc{Delight}, and the Gaussian process was trained with a subset of $50\,000$ galaxies.
-Photometric uncertainties from the inputs are propagated into the code, while non-detections for each band are set to the mean of the respective bands.
-Default settings of \textsc{Delight} were use, with the exception that the PDF bins were set to be linear instead of logarithmic, with $200$ equally-spaced bins between $0.0<z<2.0$.
-In this study a flat prior is assumed.
-
-\subsubsection{FlexZBoost}
-\label{sec:flexzboost}
-%(Ann Lee, Rafael Izbicki, Taylor Pospisil, Peter Freeman)
-
-\textsc{FlexZBoost}\footnote{\url{https://github.com/tpospisi/flexcode};\\ \url{https://github.com/rizbicki/FlexCoDE}} \citep{Izbicki:17} is a particular realization of \texttt{FlexCode}, which is a general-purpose methodology for converting any conditional mean point estimator of $z$ to a conditional {\em density} estimator $f(z \vert \x)$, where $\x$ here represents our photometric covariates and errors.\footnote{Instead of $p(z)$, we use the notation $f(z \vert \x)$ to explicitly show the dependence on $\x$.} The key idea is to expand the unknown function $f(z \vert \x)$ in an orthonormal basis $\{\phi_i(z)\}_{i}$:
-
-\begin{equation}
-f(z|\x)=\sum_{i}\beta_{i }(\x)\phi_i(z). \label{eq::series_expansion}
-\end{equation}
-By the orthogonality property, the expansion coefficients are just conditional means
-
-\begin{equation}
-\beta_{i }(\x) =  \mathbb{E}\left[\phi_i(z)|\x\right] \equiv \int f(z|\x)   \phi_i(z) dz. \label{eq:FlexZBoost_coeffs}
-\end{equation}
-These coefficients can easily be estimated from data by regression.
-
-In this paper, we use \textsc{xgboost}~\citep{Chen:16} for the regression part as these techniques scale well for massive data; it should however be noted that \textsc{FlexCode-RF} (also on GitHub), based on Random Forests, generally performs better for smaller data sets.
-As our basis, we choose a standard Fourier basis.
-There are two tuning parameters in our $p(z)$ estimate: (i) the number of terms, $I$, in the series expansion in Eq.~\ref{eq::series_expansion}, and (ii) an exponent $\alpha$ that we use to sharpen the computed density estimates $\widehat{f}(z|\x)$, according to $\widetilde{f}(z|\x) \propto \widehat{f}(z|\x)^\alpha$.
-We split the ``train data'' into a training set (85\%) and a validation set (15\%), and choose both $I$ and $\alpha$ in an automated way by minimizing the weighted $L_2$-loss function \citep[Eq.
-5 in][]{Izbicki:17} on the  validation set.
-
-Although {\tt FlexCode} offers a {\em lossless compression} of the photo-$z$ estimates (in this study, one can reconstruct $\widetilde{f}(z|\x)$ exactly at any resolution from estimates of the first 35 coefficients, Eq.~\ref{eq:FlexZBoost_coeffs}, for a Fourier basis $\{\phi_i(z)\}_{i}$), we discretize our final estimates into 200 bins linearly spaced in $0 < z < 2$ for easy comparison with other algorithms.
-Using a higher resolution may yield better results (with no added cost in storage).
-
-%\subsubsection{Frankenz}
-%\label{sec:frankenz}
-%(seeking volunteers)
-%
-%\red{\textsc{Frankenz}\footnote{\url{https://github.com/joshspeagle/frankenz}} \cite{Speagle:frankenz} is...}
-%
-
-\subsubsection{GPz}
-\label{sec:gpz}
-%(Ibrahim Almosallam)
-
-\textsc{GPz}\footnote{\url{https://github.com/OxfordML/GPz}} \citep{Almosallam:16a,Almosallam:15b} is a sparse Gaussian process based code, a fast and a scalable approximation of full Gaussian Processes \citep{Rasmussen:06}, with the added feature of being able to produce input-dependent variance estimations (heteroscedastic noise).
-The model assumes that the probability of the output $y$, the redshift, given the input $x$, the photometry, is $p(y|x)=\mathcal{N}\left(y|\mu(x),\sigma(x)^{2}\right)$.
-The mean function, $\mu(x)$, and the variance function $\sigma(x)^{2}$ are both linear combinations of basis functions that take the following form:
-
-\begin{align}
-f(x)=\sum_{i=1}^{m}\phi_{i}(x)w_{i},
-\end{align}
-where $\left\{\phi_{i}(x)\right\}_{i=1}^{m}$ and $\left\{w_{i}\right\}_{i=1}^{m}$ are sets of $m$ basis functions and their associated weights respectively.
-Basis function models (BFM), for specific classes of basis functions such as the sigmoid or the squared exponential, have the advantage of being universal approximators, i.e.
-there exist a function of that form that can approximate any function, with mild assumptions, to any desired degree of accuracy.
-The details on how to learn the parameters of the model and the hyper-parameters of the basis functions are described in \citet{Almosallam:15b}.
-
-A unique feature in \textsc{GPz}, is that the variance estimate is composed of two terms each quantifying a different source of uncertainty.
-One term (the model uncertainty) reflects how much of the uncertainty is due to lack of training samples at the location of interest, whereas the second term (the noise uncertainty) reflects how much of the uncertainty is caused from observing many noisy samples at that location.
-Thus, the predictive variance can determine whether we need more representative samples or more precise samples for any particular location in the input space.
-\textsc{GPz} can also emphasize the importance of some samples as weights.
-This weight can be for example $|z_{\rm{spec}}-z_{\rm{phot}}|/(1+z_{\rm{spec}})$ to target the desired objective of minimizing the normalized redshift error or as a function of their probability in the test set relative to the training set in order to pressure the model to better fit samples that are rare in the training set but are expected to be abundant during testing.
-
-The data is prepared for \textsc{GPz} by taking the log of the magnitude errors, decorrelating the data set using PCA and imputing the missing values using a simple linear model that estimates the missing variables given the observed ones.
-The log transformation helps to smooth the long tail distribution of the magnitude errors, which is more stable numerically and makes the optimization process unconstrained.
-The missing values are imputed by computing the mean of the training set $\mu$ and its covariance $\Sigma$, then we use the following equation to estimate the missing values from the observed ones %\[
-
-\begin{equation}
-x_{u} = \mu_{u}+\Sigma_{uo}\Sigma_{oo}^{-1}(x_{o}-\mu_{o}),
-\end{equation}  %\]
-where the subscript $o$ in $x_{o}$ indexes the \emph{observed} part of the input $x$, whereas the subscript $u$ indexes the \emph{unobserved} set (similarly for $\mu$ and $\Sigma$).
-This is the optimal expected value of the unobserved variables given the observed ones if the distribution is jointly Gaussian, note that if the variables are independent, i.e.
-$\Sigma_{uo}=0$, this will reduce to a simple average predictor.
-
-We use the Variable Covariance (VC) option in \textsc{GPz} with 200 basis functions after we note that there is no significant increase in the performance on the validation set (using 80\%-20\% training-validation split) and with no cost-sensitive learning applied.
-We do note that \textsc{GPz} is trained using a data set that includes galaxies fainter than those in the test set, which was truncated at $i<25.3$.
-This mismatch in training vs.~test results in a slightly different distribution in galaxy photometry and redshift distribution between the training and test set, which most likely degraded results compared to a run where the training set more closely matched the test set.
-
-\subsubsection{METAPhoR}
-\label{sec:metaphor}
-%(Stefano Cavuoti, Massimo Brescia, Giuseppe Longo)
-%\textcolor{brown}{general review done on this section}
-\textsc{METAPhoR} \citep[Machine-learning Estimation Tool for Accurate Photometric Redshifts,][]{Cavuoti:17} is a pipeline designed to provide photo-$z$'s point estimates and a reliable PDF for machine learning (ML) based techniques.
-It includes pre- and post-processing phases, hosting a photo-$z$ prediction engine based on the Multi Layer Perceptron with Quasi Newton Algorithm (MLPQNA), already validated on photo-$z$'s in several cases \citep{de_Jong:17,Cavuoti:17b,Cavuoti:15,Brescia:14,Brescia:13,Biviano:13}.
-Due to its plug-in based modular nature, \textsc{METAPhoR} can be easily replaced by any other photo-$z$ prediction kernel, regardless its implementation, by taking the I/O interface compliance as unique constrain.
-
-At a higher level, the pipeline mainly consists of three modules: (i) \textit{data pre-processing}, including a catalogue cross-matching sub-module \citep[based on the tool C3, ][]{Riccio:17}, a sub-module for photometric evaluation and error estimation of the multi-band catalogue used as Knowledge Base (KB), and a sub-module dedicated to the perturbation of the photometric KB, propaedeutic to the PDF estimation; (ii) \textit{photo-$z$ prediction}, which is the training/validation/test phase, producing the photo-$z$'s point estimates, based on a pre-selected ML method; (iii) \textit{PDF estimation}, specifically designed to calculate the PDF of the photo-$z$ estimation errors.
-The last module includes also a post-processing tool, providing some statistics on the produced point estimates and PDFs.
-
-The photometry perturbation law is based on the formula $m_{ij} = m_{ij} + \alpha_{i}F_{ij}*u_{\mu=0,\sigma=1}$, where $\alpha_{i}$ is a user selected multiplicative constant (useful in case of multi-survey photometry), $u_{\mu=0,\sigma=1}$ is a random value from the standard normal distribution and $F_{ij}$ is a bimodal function (a constant function + polynomial fitting of the mean magnitude errors on the binned bands), heuristically tuned in such a way that the constant component is the threshold under which the polynomial function is considered too low to provide a significant noise contribution to the photometry perturbation.
-
-As introduced, the photo-$z$ point estimate prediction engine of \textsc{METAPhoR} is based on the MLPQNA model, whose photo-$z$ regression training error, used by the quasi Newton learning rule, is based on the least square error and Tikhonov $L_{2}$-norm regularization \citep{Hofmann:18}.
-
-As main prerogative, \textsc{METAPhoR} is able to provide a PDF for ML methods by taking into account the photometric errors provided with data, by running $N$ trainings on the same training set, or $M$ trainings on $M$ different random extractions from the KB.
-The different test sets, used to produce the PDF, are thus obtained by introducing a proper perturbation, parametrized from the photometric error distribution in each band, on the photometric data populating the original test set \citep{Brescia:18}.
-
-For the present work since it was required to produce a redshift (and a PDF) for each object of the test set we decided to apply a hierarchical kNN to fill the missing detection, it goes without saying that for such points the reliability of PDFs and point estimation is lower.
-No cross validation has been used.
-
-\subsubsection{SkyNet}
-\label{sec:skynet}
-%(J. Cohen-Tanugi and Hugo Tranin)
-
-\textsc{SkyNet}\footnote{\url{http://ccpforge.cse.rl.ac.uk/gf/project/skynet/}} \citep{Graff:14} is a publicly available neural network software, based on a 2nd order conjugate gradient optimization scheme \citep[see][for further details]{Graff:14}.
-It has been used efficiently for redshift PDF estimates \citep{Sanchez:14,Bonnett:15,Bonnett:16}.
-
-%In the present work, we use \textsc{SkyNet} both as a regressor (for redshift point estimates) and a classifier (for PDF estimates). 
-The neural network is configured as a standard multilayer perceptron with three hidden layers and one input layer with $12$ nodes (the $6$ magnitudes and their errors). 
-%For the regressor, the output layer is a single linear node, while each hidden layer has $10$ nodes with a tanh activation function. 
-The classifier is laid out such that the hidden layers have 20:40:40 nodes each, all rectified linear units, and the output layer has $200$ nodes (corresponding to 200 bins for the PDF) activated with a ``softmax'' function so that they automatically sum to $1$.
-
-To avoid over-fitting, a $30$ per cent fraction of the training set is used as validation, and the training is stopped as soon as the error rate begins to increase in the validation set.
-The weights are randomly initialized based on normal sampling.
-The error function is a standard chi-square function for the regressor, and a cross-entropy function for the classifier.
-Finally, the data are all whitened before processing, with magnitudes pegged to (45,45,40,35,42,42) and their errors pegged to (20,20,10,5,15,15) for $ugrizy$ filters, respectively.
-
-\subsubsection{TPZ}
-\label{sec:tpz}
-%(Erfan Nourbakhsh)
-
-\textsc{TPZ}\footnote{\url{https://github.com/mgckind/MLZ}} \citep[Trees for Photo-$z$,][]{Carrasco_Kind:13,Carrascokind:14} is a parallel machine learning algorithm that generates photometric redshift PDFs using prediction trees and random forest techniques.
-The code recursively splits the input data (i.~e.~the training sample), into two branches, one after another, until a terminal leaf is created that meets a termination criterion (e.~g.~a minimum leaf size or a variance threshold).
-Bootstrap samples from the training data and associated errors are used to build a set of prediction trees.
-In order to minimize correlation between the trees, the data is divided in such a way that the highest information gain among the random subsample of features is obtained at every point.
-The regions in each terminal leaf node corresponds to a specific subsample of the entire data that possesses similar properties.
-
-The training data is examined before running TPZ.
-Since TPZ does not handle non-detections (magnitudes flagged as 99.0), we replace these values with an approximation of the $1\sigma$ detection threshold, i.~e.~a signal to noise ratio of 1 in terms of magnitude uncertainty using the equation $dm = 2.5 ~ \log ( 1 + N/S )$ where $dm \sim 0.7526 ~ mag$ for $N/S=1$.
-That is, for each band, we replace the non-detection with the magnitude corresponding to the error of 0.7526 from the error model forecasted for 10-year LSST data.
-The Out-of-Bag \citep{Breiman:84,Carrasco_Kind:13} cross-validation technique is used within TPZ to evaluate its predictive validity and determine the relative importance of the different input attributes.
-We employed this information to calibrate our algorithm.
-
-In the present work, the LSST magnitudes $u,~g,~r,~i$ and colors $u-g,~g-r,~r-i,~i-z,~z-y$ and their associated errors are used in the process of growing 100 trees with a minimum leaf size of 5 (the $z$ and $y$ magnitudes did not show significant correlation with the redshift in our cross-validation, so we did not use them when constructing our trees).
-We partitioned our redshift space into 100 bins from $z=0.005$ to $z=2.0$ and smoothed each individual PDF with a smoothing scale of twice the bin size.
-
-\subsection{Simple Ensemble Estimator}\label{sec:method:trainz}
-In addition to the main photo-$z$ algorithms described above we also include a very simple method.  For \trainz, as we will we call this simple estimator, we well define $p(z)$ as simply:
-
-\begin{equation}
-p(z) = \frac{1}{N_{train}}\sum_{i=1}^{N_{train}}z_{train}
-\end{equation}
-That is, we simply set the redshift PDF of every galaxy equal to the normalized $N(z)$ of the training sample.
-As the training sample is drawn from the same underlying distribution as the test sample, modulo small deviations due to sample size, the quantiles of the training and test distributions should be identical.
-This is a wildly unrealistic estimator, as it assigns all galaxies, no matter their apparent magnitude, colour, or true redshift, the same redshift PDF, and is thus uninformative at the level of individual object redshifts, but is designed to perform very well for the ensemble of all objects.
-We will discuss this method and cautions relative to metrics in Section~\ref{sec:caution}.
-
-\section{Metrics for quantifying PDF comparisons}
-\label{sec:metrics}
-%(Alex Malz, Rongpu Zhou, Jeff Newman, Ofer Lahav)
-
-%\red{Maybe it would be best to introduce the summary statistics (KS, CvM, AD) up front and then the ensemble functions and stacked n(z).}
-
-The overloaded ``$p(z)$'' is a widespread abuse of notation; we would like the outputs of photo-$z$ PDF codes to be interpretable as probabilities. 
-Obviously photo-$z$ PDFs must not take negative values and must integrate to unity over the range of possible redshifts. 
-Additionally, an estimator derived by method $H$ for the photo-$z$ PDF of galaxy $i$ must be understood as a posterior probability distribution
-
-\begin{equation}
-\hat{p}_{j}(z_{i}) = p(z | d_{i}, I_{D}, I_{H}),
-\end{equation}
-conditioned not only on the photometric data $d_{i}$ for that galaxy but also on parameters encompassing a number of things that will differ depending on the method $H$ used to produce it, namely the assumptions $I_{H}$ necessary for the method to be valid and any inputs $I_{D}$ it takes as prior information, such as a template library or training set.
-Because of this, direct comparison of photo-$z$ PDFs produced by different methods is in some sense impossible; even if they share the same prior information $I_{D}$, by definition they cannot be conditioned on the same assumptions $I_{H}$, otherwise they would not be distinct methods at all.
-
-In this study, we isolate the differences in prior information specific to each method by using a single training set $I_{D}^{ML}$ for all machine learning-based codes and a single template library $I_{D}^{T}$ for all template-based codes, and these sets of prior information are carefully constructed to be representative and complete, we have $I_{D}^{ML}\equiv I_{D}^{T}$ for every method $H$.
-Thus, we are saying
-
-\begin{equation}
-\frac{\hat{p}_{i, H}(z)}{\hat{p}_{i, H'}(z)} \approx \frac{p(z | d_{i}, I_{H})}{p(z | d_{i}, I_{H'})},
-\end{equation}
-meaning that we assume comparisons of $\hat{p}_{i, H}(z)$ isolate the effect of the method used to obtain the estimator, which should make examination of differences caused by specifics of the method implementations easier to isolate.
-
-As mentioned previously, there are cosmology probes that require knowledge of individual galaxy $p(z)$ and others that require only knowledge of the ensemble redshift distribution, $N(z)$.  
-Due to the paucity of principled techniques for using and validating photo-$z$ PDFs, there are few alternatives to the common practice of reducing photo-$z$ PDFs to point estimates. 
-Though this practice should not be encouraged, we also calculate traditional metrics based on the most common point estimators derived from photo-$z$ PDFs.  
-Those seeking to establish a connection to traditional ways of thinking about redshift estimation may consult the Appendix for these results.
-
-There are a number of metrics that can be used to test the accuracy of a photo-$z$ interim posterior as an estimator of a true photo-$z$ posterior if it is known.
-Even for simulated data, the true photo-$z$ PDF is in general not accessible unless the redshifts are in fact drawn from the true photo-$z$ PDFs, a mock catalogue generation procedure that has not yet appeared in the literature.
-Furthermore, only limited applications of photo-$z$ PDFs that could be used as the basis for a metric have been presented in the literature. 
-The most popular application by far is the calculation of the overall redshift distribution $N(z)$, the true value of which is known for the \textsc{Buzzard} simulation and will be denoted as $N'(z)$.
-Though alternatives exist \citep{Malz:chippr}, stacking according to
-
-\begin{equation} \label{eq:stacked}
-\hat{N}^{H}(z) \approx \frac{1}{N_{tot}}\ \sum_{i}^{N_{tot}}\ \hat{p}^{H}_{i}(z)
-\end{equation}
-is the most widely accepted method for estimating the redshift distribution from photo-$z$ PDFs. 
-If we assume that the response of estimators of $N(z)$ is uniform across all approaches $H$, then we may interpret metrics on the accuracy of $\hat{N}(z)$ obtained in this way. 
-We must note, however, that this is a poor assumption in general. 
-Under the setup of this paper, the true redshift distribution $N'(z) = p(z | I_{D})$ (i.e.
-because our training data is representative, the interim prior is the truth).
-In this ideal case, the method that would give the best approximation to $N'(z)$ would be one that neglects all the information contained in the photometry $\{d_{i}\}_{N_{tot}}$ and gives every galaxy the same photo-$z$ PDF $\hat{p}_{i}(z) = N'(z)$ for all $i$.
-This is the exact estimator, \trainz, that we have described in Section~\ref{sec:method:trainz}, and which will serve as a point of reference for the other codes.
-
-The exact implementation of the stacked estimator $\hat{N}^{H}(z)$ will depend on the parametrization of the photo-$z$ PDFs, which may differ across codes and can affect the precision of the estimator \citep{Malz:qp}; even considering a single method under the same parametrization, say a piecewise constant function over bins or a set of samples from the posterior, an estimator using $2N$ bins or samples will trivially be more precise than an estimator using $N$ bins or samples.
-%\red{Note: Let's address how the gridding (endpoints and spacing) affects the metrics, from the email correspondence Re: [LSST-DESC-MEMBERS] Summary of Nov 29 PZ Telecon. (AIM)}
-In order to minimize the effects of such choices, we asked those running all eleven codes to output $p(z)$ parameterized with a generous $\approx 200$ piecewise constant bins spanning $0<z<2$.  
-The piecewise constant format is chosen because of its established presence in the literature, and the choice of 200 bins was motivated by the approximate number of columns expected to be available for storage of $p(z)$ for the final LSST Project tables.\footnote{See, e.~g.~the LSST Data Products Definition Document, available at: \url{https://ls.st/dpdd}}
-All $p(z)$ catalogues are processed using the \textsc{qp} software package \citep{Malz:qp}\footnote{available at: \url{http://github.com/aimalz/qp/}} for manipulating and calculating metrics of 1-dimensional PDFs.
-We will discuss the choice of $p(z)$ parameterization further in Section~\ref{sec:results}.
-
-%\red{Divide metrics by application to ensemble of PDFs vs. estimators of the redshift distribution}
-
-\subsection{Metrics of an ensemble of photo-\mathinhead{$z$}{z} interim posteriors}
-\label{sec:qualmet}
-
-\subsubsection{Probability integral transform (PIT)}
-\label{sec:pit}
-%(Alex Malz, Peter Freeman, Sam Schmidt)
-
-The probability integral transform (PIT) \citep{Polsterer:16} is defined for each individual galaxy as:%essentially the derivative of the QQ plot, a histogram of
-
-\begin{align}
-\label{eq:pit}
-\mathrm{PIT} &= \int_{-\infty}^{z_{\rm true}}p(z)\, dz \,.
-\end{align}
-The distribution of PIT values quantifies the behavior of the {\it ensemble} of photo-$z$ PDFs, enabling us to evaluate whether the $p(z)$ is, on average, accurate:  The PIT value is the Cumulative Distribution Function (CDF) of the $p(z)$ evaluated at the true redshift.
-A catalogue of photo-$z$ PDFs that are accurate should have a flat PIT histogram (i.e., the individual PIT values as samples from each CDF should match a Uniform(0,1) distribution if the CDFs are accurate).
-Specific deviations from flatness indicate inaccuracy: overly broad photo-$z$ PDFs would manifest as underrepresentation of the lowest and highest PIT values, whereas overly narrow photo-$z$ PDFs would manifest as over-representation of the lowest and highest PIT values.
-High frequency at only $\mathrm{PIT}\approx0$ and $\mathrm{PIT}\approx1$ indicates the presence of catastrophic outliers with highly inaccurate photo-$z$ PDFs where the true redshift is outside of the support of $p(z)$.
-\citet{Tanaka:17} use the histogram of PIT values as a diagnostic indicator of overall code performance, while \citet{Freeman:17} independently define the PIT and demonstrate how its individual values may be used both to perform hypothesis testing (via, e.g., the KS, CvM, and AD tests; see below) and to construct quantile-quantile plots.
-%\red{We might want to divide this by redshift to get trends, as the overall PIT could be impacted by true n(z).  
-%For example, a PIT with nontrivial shape could be due to that behavior being uniform as a function of redshift or significant inaccuracy only in a small range of redshifts where $n(z)$ is large and a flat PIT elsewhere.}
-
-\subsubsection{Quantile-quantile (QQ) plot}
-\label{sec:qq}
-%(Rongpu Zhou, Jeffrey Newman, Alex Malz)
-
-The quantile-quantile (QQ) plot is a graphical method for comparing two distributions, where the quantiles of one distribution are plotted against the quantiles of the other distribution (A quantile being defined by partitioning a distribution into consecutive intervals containing equal amounts of probability, or equal numbers of objects in each interval).
-%\blue{Include an equation to ensure quantile definition is crystal clear before building upon it for PIT.}\red{Is maybe the above text clear enough?
-Text may be more clear than any equation I can come up with--SJS.}
-In this paper we show the quantiles of the PIT values compared to the quantiles of the Uniform distribution that we expect the PIT values to match if $p(z)$ is an accurate probability distribution for all objects.
-The QQ plot provides an easy way to qualitatively assess the differences in various properties such as the moments of an estimating distribution relative to a true distribution.  
-%location, scale and skewness.  
-In this paper, QQ plots are used for two purposes: (1) for comparing $N(z)$ from photo-$z$ PDFs (estimated using Eq.~\ref{eq:stacked}) with the true $N(z)$, i.e.
-comparing the estimated distribution of redshifts with the true redshift distribution, and (2) for assessing the overall consistency of an ensemble of photo-$z$ PDFs with their true redshifts on a population level, where the distribution of the PIT values (see previous section) is compared to a uniform distribution between $0$ and $1$.
-The QQ plot contains very similar information to that shown in the PIT histogram plot, we include both forms, as visually they each convey the information in a somewhat distinct manner.
-%\red{is this enough reason to include both?--SJS}
-%\aim{Maybe we should explain why we expect the distribution of CDFs to be uniform.}\red{Note: this is just describing the PIT.
-I'm going to swap the order of PIT to come before this, and then say we are evaluating the distribution of PIT values.
-I will also add why we expect this to be uniform--SJS}
-
-\subsubsection{Conditional density estimation loss}
-\label{sec:CDE_loss}
-%(Alex Malz, Ann Lee)
-
-With the conditional density estimation loss (CDE loss) we can compare how well different methods estimate individual PDFs for photometric covariates $\x$ rather than looking only at the ensemble distribution.
-As in Section~\ref{sec:flexzboost}, we use the notation $f(z \vert \x)$ instead of $p(z)$ to explicitly show the dependence on $\x$.
-
-%\aim{Perhaps this should be rewritten in the same notation used elsewhere in the paper.} \red{I agree, does someone who speaks both stats and astro want to take a shot at this?}
-%\textcolor{cyan}{Ann: We are looking at individual PDFs so the dependence on $\x$ needs to be explicit; this is probably the most accurate notation as it shows both the conditioning clearly as well as whether $\x$ is fixed or random.}
-
-The CDE loss is defined as:
-\begin{equation} \label{eq:cde-loss}
-L(f, \widehat{f}) = \int \int (f(z \mid \x) - \widehat{f}(z \mid \x))^{2} dz dP(\x)
-\end{equation}
-This loss is the CDE equivalent of the RMSE in regression. To estimate this loss we rewrite the loss as
-\begin{equation} \label{eq:estimated-cde-loss}
-\mathbb{E}_\X \left[\int \widehat{f}(z \mid \X)^{2} dz\right] - 2 \mathbb{E}_{\X,Z}\left[\widehat{f}(Z \mid \X)\right] + K_{f},
-\end{equation}
-where the first expectation is with respect to the marginal distribution of the covariates $\X$, the second expectation is with respect to the joint distribution of $\X$ and $Z$, and $K_{f}$ is a constant depending only upon the true conditional densities $f(z\mid \x)$.
-For each method we can estimate these expectations as empirical expectations on the test or validation data \citep[Eq.~7 in][]{Izbicki:17b} without knowledge of the true densities.
-
-%\subsubsection{Continuous Ranked Probability Score (CRPS)}
-%\label{sec:crps}
-%\red{The CRPS is used to assess the accuracy of two probabilistic models respectively.}
-
-\subsection{Metrics over estimated probability distributions}
-\label{sec:quantmet}
-
-In tandem with the QQ and PIT metrics introduced above, we additionally compute the following metrics comparing the empirical CDF of a distribution to the true or expected distribution.
-These metrics give a more quantitative measure of the departure from ideal than the more visual PIT histogram and QQ plot.
-We compute metrics comparing the CDF of PIT values to a the CDF of a Uniform distribution, and also compute the CDF of the true redshift distribution $N'(z)$ compared the $\hat{N}(z)$ distribution derived from summing the $p(z)$ as described in Eq.~\ref{eq:stacked}.
-
-\subsubsection{Root-mean-square error (RMSE)}
-\label{sec:rmse}
-%(Alex Malz)
-
-We employ the familiar root-mean-square error:
-
-\begin{equation}
-\mathrm{RMSE} = \sqrt{\int_{-\infty}^{\infty}\left(\hat{f}(z)-f'(z)\right)^{2}\,dz},
-\end{equation}
-%may be used for two purposes in this analysis. 
-%We calculate it between the true redshift distribution $N'(z)$ and each estimator $\hat{N}(z)$ (derived by Eq.~\ref{eq:stacked}) as well as between the ideal QQ and PIT curves from each method studied and the ideal curves.
-Though this metric does not account for the fact that the redshift distribution function is, in fact, a probability distribution, it can still be interpreted as a measure of the integrated difference between the estimated distribution and the true distribution, and it can be used to quantify the otherwise qualitative metrics.
-
-\subsubsection{Kolmogorov-Smirnov (KS) and related statistics}
-\label{sec:ks}
-%(John Soo, Rongpu Zhou)
-
-The \textit{Kolmogorov-Smirnov statistic} $N_{\rm KS}$ is the maximum difference between $F_{\rm phot}(z)$ and $F_{\rm spec}(z)$, the CDFs of the photo-$z$ and spectroscopic redshift respectively:
-
-\begin{equation}
-N_{\rm KS} = \max_{z}\left( \left| F_{\rm phot}(z) - F_{\rm spec}(z) \right| \right).
-\end{equation}
-%The KS test quantifies whether the two redshift distributions are drawn from a similar distribution, independent of binning. 
-The KS test quantifies the similarity between two distributions, independent of binning. 
-A lower $N_{\rm KS}$ value corresponds to more similar distributions.
-%We calculate two KS test statistics, one to test the distribution of $p(z)$ values, and one to test the stacked estimator $\hat{n}(z)$ of the redshift distribution.  
-%The $p(z)$ test compares the CDF of the PIT values described in the previous section to that of a uniform distribution.  
-%The $n(z)$ test compares the CDF of the distribution formed by stacking the $p(z)$ for every object in our test set against the CDF of the corresponding true redshifts.
-%In this case a lower $N_{\rm KS}$ value means that the photo-$z$ distribution fits the spectroscopic redshift distribution well.
-$F_{\rm phot}(z)$ is calculated by stacking the PDFs of every single object in the sample, producing a smooth cumulative $n(z)$ to be compared with $F_{\rm spec}(z)$.
-
-We also consider two variants of the KS statistic: the Cramer-von Mises (CvM) and Anderson-Darling (AD) statistics.
-The CvM statistic is similar to the KS statistic as it is also computed from the distance between the measured CDF and the ideal CDF, but instead of the maximum distance, the CvM statistic calculates the average of the distance squared:
-
-\begin{equation}
-\label{eq:cvm}
-\omega^2 = \int_{-\infty}^{+\infty}\big(F_{\mathrm{meas.}}(x) - F_{\mathrm{ideal}}(x)\big)^2\mathrm{d}F_{\mathrm{ideal}}
-\end{equation}
-% In this paper we use the CvM statistic to evaluate the performance of both $n(z)$ and $p(z)$.
-The AD statistic is a weighted version of the CvM statistic, making it more sensitive to the tails of the distribution:
-
-\begin{equation} \label{eq:ad}
-A^2 = n\int_{-\infty}^{+\infty} \frac{\big(F_{\mathrm{meas.}}(x) - F_{\mathrm{ideal}}(x)\big)^2} {F_{\mathrm{ideal}}(x) (1-F_{\mathrm{ideal}}(x))}\mathrm{d}F_{\mathrm{ideal}}
-\end{equation}
-where $n$ is the sample size. 
-%The KS statistic and the CvM and AD statistics related to it are all used to evaluate the performance of both $\hat{n}(z)$ and the QQ and PIT of the CDFs of the ensemble of photo-$z$ interim posteriors.
-
-%\subsubsection{Kullback-Leibler Divergence (KLD)}
-%\label{sec:kl}
-%(Alex Malz)
-%
-%The KLD is a measure of the loss in information when using a probability distribution that is an estimate of a known true probability distribution, so we use 
-%\begin{equation}
-%\mathrm{KLD} = \int_{-\infty}^{\infty}n'(z)\log\left[\frac{n'(z)}{\hat{n}(z)}\right]dz
-%\end{equation}
-%as a metric on the stacked redshift distribution estimator of Eq.~\ref{eq:stacked}.
-Noting the assumptions made in using Eq.~\ref{eq:stacked}, we say that a photo-$z$ PDFs method whose stacked redshift distribution estimator $\hat{n}(z)$ more closely resembles the true redshift distribution $n'(z)$ is desired, i.e.
-we seek to minimize the KLD.
-
-\subsubsection{Moments}
-\label{sec:moments}
-%(Alex Malz)
-For the $\hat{N}(z)$ distributions we additionaly calculate the first three moments of the estimated redshift distribution for each code and compare them to the moments of the true redshift distribution $N'(z)$.
-The $m$th moment of a distribution is defined as
-
-\begin{equation}
-\langle z^{m}\rangle = \int_{-\infty}^{\infty}z^{m}N(z)dz.
-\end{equation}
-Here, we use the moments of the stacked estimator of the redshift distribution function as the basis for a metric.
-The closer the moments of $\hat{N}(z)$ for a photo-$z$ PDF method are to the moments of the true redshift distribution function $N'(z)$, the better the photo-$z$ PDF method.
-
-\section{Results}
-\label{sec:results}
-%(Sam Schmidt, Bryce Kalmbach, Johann Cohen Tanugi, Rongpu Zhou)
-
-\subsection{Ensembles of photo-\mathinhead{$z$}{z} interim posteriors}
-
-\begin{figure*}
-%\centering
-\begin{subfigure}[b]{0.49\textwidth}
-\includegraphics[width=\textwidth]{figures/pz_12codes_261931_crop.jpg}
-\end{subfigure}
-\begin{subfigure}[b]{0.49\textwidth}
-\includegraphics[width=\textwidth]{figures/pz_12codes_471167_crop.jpg}
-\end{subfigure}
-\begin{subfigure}[b]{0.49\textwidth}
-\includegraphics[width=\textwidth]
-{figures/pz_12codes_713178_crop.jpg}
-\end{subfigure}
-\begin{subfigure}[b]{0.49\textwidth}
-\includegraphics[width=\textwidth]{figures/pz_12codes_982747_crop.jpg}
-\end{subfigure}
-\caption{Four illustrative examples of individual p(z) distributions produced by the codes.
-The red vertical line represents the true redshift.
-Examples are chosen with common features seen in PDFs: tight unimodal $p(z)$ (upper left), broad unimodal $p(z)$ (upper right), bimodal $p(z)$ (lower right), and complex/multimodal $p(z)$ (lower left).
-Codes show varying amounts of small-scale structure in their reconstruction of the posterior distribution.
-We see varying responses from the codes in the presence of color degeneracies and photometric errors, resulting in narrow and broad unimodal, bimodal, and multi-modal $p(z)$ curves.} \label{fig:pz_examples}
-\end{figure*}
-
-%\begin{figure*}
-%\centering
-%\includegraphics[width=\textwidth]{figures/QQplot_11codes.jpg}
-%\caption{The QQ plots produced for every photo-$z$ code.
-\aim{Consider making this a plot of QQ minus that diagonal for better contrast between methods.}\red{The PIT histogram is essentially this, which is one of the main reasons to include both.
-QQ is a fairly standard form, so plotting in this fashion rather than the difference seems better to me, given that we also ahve the PIT histograms.
---SJS}} \label{fig:qq}
-%\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{figures/PITANDQQplot_12codes_crop.jpg}
-\caption{Summary plots for all eleven photo-$z$ codes illustrating performance for the interim posterior statistics.
-The top panel of each pair shows both the Quantile-Quantile (QQ) plot (red) and the histogram of PIT values (blue).
-The desired behavior is a QQ plot that matches the diagonal dashed line, and a PIT histogram that matches a uniform distribution matching the thin horizontal black line.
-The bottom panel of each pair shows the difference between the QQ quantile and the diagonal, illustrating departure from the desired performance.
-Histograms with an overabundance of PIT values at the centre of the distribution indicate $p(z)$ distributions that are overly broad, while an excess of values at the extrema indicate $p(z)$ distributions that are overly narrow.
-Values of PIT=0 and PIT=1 indicate ``catastrophic failures'' where the true redshift is completely outside the support of $p(z)$.
-Asymmetric features are indicative of systematic bias in the redshift predictions.
-A variety of behaviors are evident, and specific details are discussed in the text.}
-\label{fig:pitqq}
-\end{figure*}
-
-Fig.~\ref{fig:pz_examples} Shows the $p(z)$ produced by each of our eleven photo-$z$ codes for four example galaxies which exemplify some prominent cases that arise when estimating photo-$z$ PDFs: a narrow, unimodal redshift solution, a broader unimodal solution, a bimodal distribution, and a complex, multimodal distribution.
-The red vertical line represents the true redshift of the individual galaxy, and the blue curve represents the redshift probability.  
-Several features are obvious even in these illustrative examples.  
-\textsc{ANNz2, METAPhoR, NN,} and \textsc{SkyNet} all show an excess of small-scale features, which appear to be print-through of the underlying training set galaxies.  
-\textsc{GPZ} (in its current implementation), on the other hand, always produces a single Gaussian, which broadens to cover the multi-modal redshift solutions seen in other codes.
-
-As stated in Section~\ref{sec:metrics}, $p(z)$ is parameterized as $\approx 200$ piecewise constant bins covering $0<z<2$ for all eleven codes, giving a grid size of roughly $\delta z = 0.01$ for each code.
-A piecewise constant grid was a natural choice for some photo-$z$ codes, for instance most template-based codes compute likelihoods on a fixed grid.  
-In contrast, FlexZBoost, for example, can return estimates on any grid without compression errors as it’s a basis expansion method where only the expansion coefficients need to be stored.
-Codes with a native output format other than the shared piecewise constant binning scheme (or one that can be losslessly converted to it) may suffer from loss of information when converting to it, which could artificially favor some codes over others.
-
-Furthermore, the fidelity of photo-$z$ interim posteriors in this format varies with the quality of the photometry.  
-For faint galaxies, this redshift resolution is sufficient to capture the shape of $p(z)$ for the majority of the test sample, where photometric errors on the faint galaxies lead to somewhat broad peaks in the redshift posterior.
-However, as can be seen in e.~g.~the top left panel of Fig.~\ref{fig:pz_examples}, for bright galaxies with narrow $p(z)$ the grid spacing of $\delta z = 0.01$ is not sufficient to resolve the peak.  
-This is consistent with the results described in \citet[]{Malz:qp}, who find that quantiles (and, to a lesser degree, samples) often outperform gridded $p(z)$, particularly for bright objects and in the presence of harsher storage constraints.
-With a full 200 numbers to capture the information of each photo-$z$ PDF, any parametrization will perform adequately, but other storage parametrizations and limits on storage resources may be considered in future work.
-%Switching to a quantile based parameterization may be more costly computationally, for example template-based codes would need to test more grid point in order to resolve the quantiles for bright galaxies.
-However, the computational time for template based codes scales roughly linearly with the number of grid points, so this may be a reasonable thing to implement.
-We will discuss this further in Section~\ref{sec:discussion}.
-% \red{someone review this statement to make sure that I'm saying this correctly!}
-
-Fig.~\ref{fig:pitqq} shows both the quantile-quantile plots (red) and the histogram of PIT values (blue) summarizing the results from each photo-$z$ code.
-The red line shows the measured quantiles, while the black diagonal represents the ideal QQ values if the distribution were perfectly reproduced.
-A second panel below the main panel for each code shows the difference between $Q_{data}$ and $Q_{theory}$, i.~e.~the departure from the diagonal, for clarity.
-Biases and trends in whether the average width of the $p(z)$ values being over/under-predicted are evident.
-An overall bias where the predicted redshift is systematically low manifests as the measured QQ value falling above the diagonal, as is the case for \textsc{BPZ} and \textsc{EAZY}, while a systematic overprediction shows up as the measured QQ value falling below the diagonal, as seen in \textsc{TPZ}.
-In terms of PIT histograms, a systematic underprediction of redshift corresponds to fewer PIT values at $PIT<0.5$ and more at $PIT>0.5$, while a systematic overprediction will show the opposite.
-
-Examination of the PIT histograms and QQ plots shows that there are fairly generic issues with the width of $p(z)$ uncertainties: \textsc{Delight, NN, SkyNet} and \textsc{TPZ} all show a PIT histogram with an dearth of low values and and an excess of high values, signs that, on average, their $p(z)$ are more broad than the true distribution of redshifts.
-\textsc{METAPhoR} shows the opposite trend, indicating the the $p(z)$  are more narrow than the distributions given by the true redshifts.
-In all of these code cases there is a free parameter or bandwidth that can be used to tune uncertainties.
-The sensitivity of multiple codes to this bandwidth choice emphasizes the fact that great care must be taken in setting user-defined parameters in photo-$z$ codes, even in the presence of representative training/validation data.
-for \textsc{FlexZBoost} the ``sharpening'' parameter (described in Section~\ref{sec:flexzboost}) plays a key role in improving the results, resulting in a QQ plot that is very nearly diagonal.
-A similar sharpening procedure could be beneficial for several codes.
-Interestingly, the three purely template-based codes, \textsc{BPZ, EAZY}, and \textsc{LePhare}, show relatively well behaved $p(z)$ statistics (albeit with some bias), which may indicate that the likelihood estimation with representative templates is accurately capturing the uncertainties on individual redshifts.
-
-%Fig.~\ref{fig:pitqq} shows the PIT histogram plots produced by each photo-$z$ code. 
-%These plots show information that is very similar to that shown in the QQ plot in many ways. 
-The ideal PIT histogram would follow the black dashed line, representing a uniform distribution of PIT values, equivalent to the diagonal line in the QQ plot.
-Overly broad $p(z)$ values show up as an excess of PIT values near $0.5$ and a dearth of values at the edges, while overly narrow $p(z)$ will have an excess at the edges and will be missing values at the centre.
-Another feature evident in the PIT histograms is the number of ``catastrophic outlier'' values where the true redshift falls outside of the non-zero support of $p(z)$, corresponding to PIT$=0.0$ or $1.0$ is more apparent than in the QQ plots.
-Following Kodra \& Newman (in prep.) we define $f_{\rm O}$ as the fraction of objects with $PIT<0.0001$ or $PIT>0.9999$.
-Table~\ref{tab:pitoutlier} lists these fractions for each of the codes.
-For a proper Uniform distribution we expect a value of $0.0002$.
-Several codes show a marked excess, with \textsc{ANNz2, FlexZBoost, LePhare, and METAPhoR} with $f_{\rm O} > 0.02$, indicating a sizeable number of catastrophic redshift solutions where the true redshift is not covered by the extent of $p(z)$.
-For \textsc{METAPhoR} this may be partially due to an overall underprediction of the $p(z)$ width, however this is not the case for the other codes.
-\textsc{LePhare} is a particular outlier with nearly 5 per cent of objects outside of $p(z)$ support.
-Further study will be necessary to determine what is causing these misclassifications for \textsc{LePhare}.
-As expected, and by design, \textsc{TrainZ} has the proper fraction of outliers for the $f_{\rm O}$ statistic.
-
-\begin{table}
-\setlength{\tabcolsep}{2pt}
-\centering
-\caption{The fraction of ``catastrophic outlier'' PIT values.
-We expect a value of 0.0002 for a proper Uniform distribution.
-An excess over this small value indicates true redshifts that fall outside the non-zero support of the $p(z)$.
-}\label{tab:pitoutlier}
-\begin{tabular}{lc}
-\hline
-\hline
- ``catastrophic outlier''\\ PIT fraction\\
-\hline
-Photo-z Code & fraction PIT$<10^{-4}$ or $>$0.9999\\
-\hline
-\textsc{ANNz2} & 0.0265\\
-\textsc{BPZ} & 0.0192\\
-\textsc{Delight} & 0.0006\\
-\textsc{EAZY} & 0.0154\\ 
-\textsc{FlexZBoost} & 0.0202\\
-\textsc{GPz} & 0.0068\\ 
-\textsc{LePhare} & 0.0486\\
-\textsc{METAPhoR}& 0.0229\\
-\textsc{CMNN} & 0.0034\\ 
-\textsc{Skynet} & 0.0001\\  
-\textsc{TPZ} & 0.0130\\
-\hline
-\textsc{TrainZ} & 0.0002\\
-\end{tabular}
-\end{table}
-
-%\begin{figure*}
-%\centering
-%\includegraphics[width=\textwidth]{figures/PIT_histogram_plot_11codes.jpg}
-%\caption{The PIT histogram plots produced for every photo-$z$ code.
-Ideal behaviour would match a uniform distribution, indicated by the horizontal dashed line.
-Histograms with an overabundance of PIT values at the centre of the distribution indicate $p(z)$ distributions that are overly broad, while an excess of values at the extrema indicate $p(z)$ distributions that are overly narrow.
-Values of PIT=0 and PIT=1 indicate ``catastrophic failures'' where the true redshift is completely outside the support of $p(z)$.
-Asymmetric features are indicative of systematic bias in the redshift predictions.
-%\aim{Could we combine this with the previous figure by plotting both as differences from ideal on one plot with left and right axis labels?
-The information seems degenerate (which makes sense!)}\red{as you say, the information is degenerat, this form emphasizes the differences a bit more, and shows off the catastrophic outliers with PIT=0 and 1 a bit better in my mind.} } \label{fig:pit}
-%\end{figure*}
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{figures/KSvsCvMvsAD_PIT_withnull_jpg.jpg}
-\caption{A visual representation of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the PIT distributions.
-The statistics are often highly correlated, though the AD statistic truncates the extrema of the distribution and can have disparate values compared to KS and CvM.} \label{fig:pit_stats}
-\end{figure*}
-
-Fig.~\ref{fig:pit_stats} shows comparative metric values for the quantitative Kolmogorov-Smirnoff (KS), Cramer-Von Mises (CvM), and Anderson Darling (AD) test statistics for each of the codes based on comparing the distribution of their PIT values to the expected uniform distribution over the interval [0,1].
-The individual values of the statistic are not as important as the comparative score between the different codes.
-%\red{Can p-values be supplied for each statistic?
-The statistics themselves are difficult to interpret, other than ``lower is better'' (p-value in skgof was broken, having trouble finding 1-sample KS calculation for uniform distribution)}
-The AD test statistic diverges for values that include the extrema, and thus is calculated by excluding the edges of the distribution.  
-We calculate the AD statistic over the range of PIT values $v=[0.01,0.99]$.  \textsc{ANNz2} and \textsc{FlexZBoost} score very well for the PIT metrics.  
-\textsc{METAPhoR} and \textsc{LePhare} score very well in the PIT AD statistic, but both have a large number of catastrophic outliers, resulting in higher KS and CvM scores.
-
-Given the near-perfect training data, examining the individual codes for explanations for departures from the expected behaviour will be instructive in avoiding similar problems in future tests. 
-\textsc{ANNz2} performs quite well in $p(z)$ based metrics.
-In the specific implementation employed in this paper, the final $p(z)$ is a weighted average of five neural-nets.
-During the training process \textsc{ANNz2} compares the percentiles of the redshift training sample against the CDFs of the $p(z)$ sample.
-Distributions that more closely match are given extra weight, and the final weights are designed to produce accurate percentiles.
-Given that our metrics are focused on the percentile distributions, it is unsurprising that \textsc{ANNz2} performs well in the given metrics.
-The discreteness in the individual $p(z)$ estimated by \textsc{ANNz2} can be attributed to the fact that the code was run as a classifier, assigning weights to discrete bins of redshift.
-While multiple bins may receive weight, the bins themselves will still be discretized, and no additional smoothing was performed.
-Overall, \textsc{FlexZBoost} and \textsc{ANNz2} show the best ensemble agreement in their distribution of PIT values.
-
-\subsection{Metrics of the stacked estimator of the redshift distribution}\label{sec:stackedmetrics}
-
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{figures/NZsumplot_12codes.jpg}
-\caption{The stacked $p(z)$ produced by each photo-$z$ code ($\hat{N}(z)$, red) compared to the spectroscopic redshift distribution ($N'(z)$, blue).
-Varying levels of small-scale structure are seen in the codes.
-$N'(z)$ is smoothed using a single bandwidth chosen via Scott's rule for all codes.} \label{fig:nz}
-%\aim{Arguably this could also be clarified by showing the true $n'(z)$ once and then differences from it for each cod's stacked estimator\dots}\scc{what if we plot the difference between true and stacked rather than soverposing the two lines?
-Sam please do not hate me.}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{figures/KSvsCvMvsAD_NZ_withnull_jpg.jpg}
-\caption{A visual representation of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the $\hat{N}(z)$ distributions.
-The statistics are correlated, the codes with the lowest KS statistics tend to have the lowest CvM and AD statistics.
-\textsc{CMNN} performs markedly better than the others in reconstructing the overall $N(z)$ distribution, while \textsc{SkyNet} scores poorly due to an overall bias in its redshift predictions.} \label{fig:nz_stats}
-\end{figure*}
-
-Fig.~\ref{fig:nz} shows the stacked $\hat{N}(z)$ distribution compared to the true redshift distribution $N'(z)$ for all tested codes.
-The red line indicates the summed $p(z)$ for each code, while the blue line shows the true redshift distribution smoothed via kernel density estimation (KDE), with a bandwidth chosen via Scott's rule \citep{Scott:1992}.
-While Scott's rule is used to display $N'(z)$ in the figure, all quantitative statistics are computed via the empirical CDF, and are thus unaffected by bandwidth/smoothing choice.
-Several of the codes show an excess at $z \sim 1.4$, particularly the template-based codes \textsc{BPZ}, \textsc{EAZY}, and \textsc{LePhare}.  
-This is likely due to the $4000$ angstrom break passing through the gap between the $z$ and $y$ filters.  
-This feature is one of the most prominent in individual galaxy p(z), and is readily seen in the point-estimate plots shown in Fig.~\ref{fig:pz_pointestimates} and described in the Appendix.  
-Several of the machine learning based codes appear to be over-trained, adding excess galaxy probability to the redshift peaks and missing probability in the troughs.  
-Given that our training data is drawn from the same galaxy population as the test set, and our data has prominent peaks in $N'(z)$, perhaps it is not unexpected that such overtraining occurs. 
-A more extensive training/validation set might allow for a better choice of smoothing parameters in individual codes that would avoid such overtraining.
-
-%\aim{Move the text on Scott's Rule to Discussion?}
-%Scott's rule \citep{Scott:1992} was used to determine the smoothing bandwidth of the true redshift sample: that is, a single smoothing scale was chosen to represent the true redshift sample for all eleven codes.
-
-As with the $p(z)$ values in Figure~\ref{fig:pitqq}, different levels of substructure are obvious for the different codes.  
-While Scott's rule provides a relatively good general smoothing scale to represent the true $N'(z)$, there are smaller scale fluctuations: while \textsc{FlexZBoost} and \textsc{CMNN} appear somewhat discrepant in Fig.~\ref{fig:nz}, they are actually the two most accurate in terms of their quantitative measurements.
-%\textsc{SkyNet} also shows excess small-scale structure, but also shows significant overestimate of the high redshift galaxy population.  
-Interestingly, while \textsc{ANNz2} shows an abundance of small scale structure in individual $p(z)$ measurements (see Fig.~\ref{fig:pz_examples}), the summed $\hat{N}(z)$ is rather smooth, where the small scale features average out.
-This is not the case for the two other codes that show an abundance of substructure in their individual $p(z)$: both \textsc{CMNN} and \textsc{SkyNet} show small scale features both in $p(z)$ and $\hat{N}(z)$.
-For \textsc{CMNN} the $p(z)$ are a simply a weighted histogram of all spectroscopic training galaxies in nearby colour space with no smoothing applied, so the substructure is not unexpected.
-The PIT histogram and shape of the QQ plot in Figure~\ref{fig:pitqq} show that \textsc{CMNN} is producing $p(z)$ that are overly broad, additional smoothing of the $p(z)$ would exacerbate this problem.
-While the $\hat{N}(z)$ plot shows more small scale features than other codes, these features are actually representative of real structure in the true $N'(z)$, as evidenced by the very good metric scores for \textsc{CMNN}.
-\textsc{SkyNet} $p(z)$ were also not smoothed: while previous implementations of the code such as \citet{Sanchez:14} and \citet{Bonnett:15} (see Appendix C.3) implement a ``sliding bin'' smoothing, no such procedure was used in this study.
-In addition to excess substructure, \textsc{SkyNet} shows an obvious redshift bias, evident both visually in Figure~\ref{fig:nz} and in the first moment of N(z) listed in Table~\ref{tab:moments}, where it is clearly an outlier.
-\textsc{SkyNet} employed a method where a random sample of training galaxies was chosen, but there was no test that the subset was completely representative of the overall redshift distribution.
-Also unlike \citet{Bonnett:15}, no effort was made to add extra weight to more rare low and high redshift galaxies.
-Either of these decisions could be the cause of the bias seen in our results.
-Future runs of \textsc{SkyNet} will explore these implementation choices and their effects.
-
-Figure~\ref{fig:nz_stats} shows the quantitative Kolmogorov-Smirnoff (KS), Cramer-Von Mises (CvM), and Anderson Darling (AD) test statistics for each of the codes for the $\hat{N}(z)$ based measures.
-%\red{Can p-values be supplied for each statistic?
-The statistics themselves are difficult to interpret, other than ``lower is better'' (no, p-values are very difficult to compute for non-uniform distributions)}
-\textsc{FlexZBoost}, \textsc{CMNN}, and \textsc{TPZ} outperform the other codes in the $\hat{N}(z)$ metrics.  
-It is unsurprising that \textsc{CMNN} scores well, as with a near perfectly representative training set means that choosing neighbouring points in color/magnitude space should lead to excellent agreement in the final $\hat{N}(z)$ estimate.
-\textsc{TPZ} performed quite poorly in $p(z)$ statistics, but results in a good fit to the overall $N(z)$.
-This is somewhat surprising, as performance was optimized for accurate $p(z)$, not $\hat{N}(z)$.
-During the validation stage for \textsc{TPZ}, there was a trade off between the width of the $p(z)$ when adjusting a smoothing parameter and overall redshift bias.
-The optimal result in the PIT metrics, as illustrated in the shape of the QQ plot, does contain some level of bias as well as a slight underprediction of mean $p(z)$ width, which translates to poor metric scores.
-This is something that will be looked into for \textsc{TPZ} in the future.
-
-It is also of note that all three template-based codes show an excess in their stacked $p(z)$ at $z\,\sim\,1.3-1.4$.
-This redshift range corresponds to the wavelengths where the $4000$ Angstrom break is passing between the borders of the $z$ and $y$ filters.
-This strong break entering the gap between the two reddest filters can cause problems with redshift estimation of individual galaxies, as can be seen in the point-estimate photo-$z$'s shown in Figure~\ref{fig:pz_pointestimates}.
-This is not unique to this dataset, it is a common occurrence in photo-$z$ estimation.
-The fact that similar excesses appear in Figure~\ref{fig:nz} for \textsc{ANNz2} and \textsc{METAPhoR} shows that the effect is not limited to template-based codes.
-However, the lack of such a feature in the other codes shows that it is possible to eliminate the degeneracies.
-Further study on this issue may provide a solution for codes that suffer from this shortcoming.
-
-%\red{discussion of why NN and FZBoost are doing well, how we can try to improve the other codes} 
-%\textcolor{cyan}{Ann: }
-
-Table~\ref{tab:cdeloss} shows the CDE loss statistic for each photo-$z$ code.
-Once again \textsc{FlexZBoost} and \textsc{CMNN} score very well for the stacked $\hat{N}(z)$ metrics, as do \textsc{GPz} and \textsc{TPZ}.
-The CDE loss measures how well individual PDFs are estimated, and codes with a low CDE loss tend to have good $\hat{N}(z)$ estimates (though the reverse is not necessarily true).
-\textsc{FlexZBoost} is optimized to minimize CDE loss which may explain why the method has good ensemble metrics as well.
-Note from Table~\ref{tab:cdeloss} that both \textsc{FlexZBoost} and \textsc{CMNN} have low CDE losses.
-Empirically, we have found that PIT RMSE is not as closely correlated to CDE loss as it is to the $N(z)$ statistics.
-As CDE loss is a better measure of individual redshift performance, rather than ensemble distribution performance, this statistic is a better indicator of which codes will be most likely to perform well for science cases where single objects are employed.
-
-Table~\ref{tab:rmse} gives the root-mean-square-error (RMSE) statistics for both the PIT and N(z) estimators.
-The PIT value calculates the RMSE between the quantiles shown in the QQ plot in Figure~\ref{fig:pitqq} and the diagonal, while the N(z) calculates the RMSE between the cumulative distribution of the stacked $\hat{N}(z)$ and the true redshift distribution $N'(z)$.
-%\red{more about RMSE?
-Or remove it?
-It's barely described in Section~\ref{sec:metrics} and wasn't even referred to in the text until I added this on June 29th--SJS.}
-
-Table~\ref{tab:moments} lists the first three moments of the stacked $\hat{N}(z)$ distribution, including the moments of the ``truth'' distribution for comparison.
-Several codes are able to reproduce the mean and variance of the distribution to less than a per cent, while several codes do not, which may be a cause for concern, given that mean and variance of the redshift distribution are key properties in cosmological analyses.
-We note that this stated goal of the study as defined for participants was to accurately reproduce $p(z)$, the ``stacking'' of the probability distributions to estimate $\hat{N}(z)$ was not the focus as stated to the participants.
-This explains why some of the best-performing empirical codes in terms of $p(z)$ measures (e.~g.~\textsc{FlexZBoost}) do not do as well at reproducing $\hat{N}(z)$ moments.
-Had we defined a different parameter to optimize, in this case overall accuracy of $\hat{N}(z)$ rather than individual $p(z)$, would result in improved performance in a particular metric.
-That is, optimizing photo-$z$ performance for one metric does not automatically give optimal performance for other metrics.
-As previously stated, there are a variety of scientific use cases for photo-$z$'s in large upcoming surveys, and care must be taken in how the metrics used to optimize catalog photometric redshifts are defined as well as in how they are used.
-In addition, very few scientific use cases will employ the overall $\hat{N}(z)$ with no cuts, as we explore in this paper.
-We discuss more realistic tomographic bin selections that will be explored in a follow-up paper in Section~\ref{sec:futurework}.
-%\red{[mention that we are calculating moments for the entire sample, will look at fiducial tomographic bins in a follow-up paper (unless group and reviewers realy think we should include in this paper.]}
-%\red{FlexZBoost has some of the best $n(z)$ statistics, but the moments are not good.
-Why is this?
-Is it the over/underprediction of the high redshift part of the distribution?
-Discuss this after talking with Rafael.}
-
-%\subsection{Response of Individual Codes}
-%\label{sec:res:pz_indiv_codes}
-%\red{this may be incorporated in to 5.1 and 5.2 rather than live in its own section!}
-
-%\red{HOW CODES deal with negative fluxes, and magnitude uncertainties}
-
-%\red{overall conclusions of the Results section}
-
-\subsection{Interpretation of metrics}\label{sec:caution}
-%Caution on metrics and photo-$z$ codes}\label{sec:caution}
-%(Alex Malz, Sam Schmidt)
-
-Samples from accurate photo-$z$ posteriors should reproduce the space of $p(z, data)$.
-However, it is difficult to test this reconstruction given our data set, as the galaxy distributions arise from mock objects pasted on to an underlying dark matter halo catalogue with properties designed to match empirical relations, rather than being drawn from statistical distributions in redshift.
-In previous sections we have mentioned that optimizing for a specific metric does not guarantee good performance on other metrics, nor is there any guarantee that good performance by our metrics corresponds to \textit{accurate} photo-$z$ posteriors.
-%\aim{Explain that samples from accurate photo-$z$ posteriors should reconstruct the space of $p(z, data)$, a test we can't check with our datasets.}  
-In other words, we can construct photo-$z$ estimators that provide good coverage in many of our tests, but which have very little predictive power.
-
-The \trainz\ estimator, which assigns every galaxy a $p(z)$ equal to $N(z)$ of the training set as described in Section~\ref{sec:method:trainz}, is introduced as a ``null test'' to demonstrate this point via \textit{reductio ad absurdum}.
-\trainz\ outperforms all codes on the PIT-based metrics, and all but one code on the $N(z)$ based statistics.  
-Because our training set is perfectly representative of the test set, $N(z)$ should be identical for both sets down to statistical noise.
-%\textit{Explain the connection between \trainz\ and these metrics.}
-
-The CDE loss and point estimate metrics, however, successfully identify problems with \trainz.  
-As shown in Appendix~\ref{sec:pointmetrics}, \trainz has identical $ZPEAK$ and $ZWEIGHT$ values for every galaxy, and thus the photo-$z$s are constant as a function of spec-$z$s, i.e.
-a horizontal line at the mode and mean of the training set distribution respectively.
-The explicit dependence on the \it{individual} posteriors in the calculation of the CDE loss, described in Section~\ref{sec:CDE_loss}, distinguishes this metric from the other $p(z)$ metrics that test the overall ensemble of $p(z)$ distributions.
-With a representative training set, \trainz\ will score well on the ensemble metrics, but fails miserably for metrics tied to individual redshifts.
-We note that many of the ensemble-based metrics are prominent in the photo-$z$ literature despite their inability to identify problems such as those exemplified by \trainz.
-
-%\aim{Explain why CDE loss is different from other metrics (isn't fooled by \trainz) and note that the others have more of a presence in the literature despite their flaw of failing to identify this severe problem.}
-
-% While looking at one, or even most of our metrics would have given the impression that this estimator was nearly optimal, red flags in CDE loss and point estimates reveal the problems.  
-%[more caution on over-reliance on metrics, and making sure that metrics test all of what you want] \red{this definitely still needs work}
-In summary, context is crucial to interpreting metrics and defending against the likes of \trainz.
-The best photo-$z$ method is the one that most effectively achieves our science goals, not the one that performs best on a metric that does not accurately reflect those goals.
-In the absence of clear goals or the information necessary for a principled metric definition, we must think carefully before choosing a single metric
-
-%\begin{table*}  %%% DATA TABLE %%%
-%\caption{KS, CvM and AD statistics for each photo-$z$ code.} \label{tab:results}
-%PIT statistic
-%
-%\begin{tabular}{lrrrr}
-%\hline
-%\bf Photo-$z$ Code & \bf KS & \bf CvM & \bf AD & \bf AD \\
-%                   &        &         & $v=[0.05,0.95]$ & $v=[0.01,0.99]$        \\
-%\hline
-%\textsc{ANNz2} 		& $0.0200$ &   $52.25$ &   $583.0$ &   $759.2$ \\
-%\textsc{BPZ} 		& $0.0388$ &  $280.79$ &  $1361.4$ &  $1557.5$ \\
-%%\textsc{Delight} 	& $0.1380$ & $3040.59$ & $13880.2$ & $17761.4$ \\
-%\textsc{Delight}    & $0.0876$ & $1075.17$ & $4407.8$  & $6167.5$\\
-%%\textsc{EAZY} 		& $0.0690$ &  $994.19$ &  $3479.5$ &  $4158.8$ \\
-%\textsc{EAZY}       & $0.0723$ & $1105.58$ & $3475.0$  & $4418.6$\\
-%%\textsc{FlexZBoost} & $0.0260$ &   $70.58$ &   $595.0$ &   $533.0$ \\
-%\textsc{FlexZBoost} & $0.0240$ &   $68.83$ &   $567.9$ &   $478.8$\\
-%\textsc{GPz} 		& $0.0449$ &  $258.56$ &   $904.8$ &  $1414.8$ \\
-%\textsc{LePhare} 	& $0.0663$ &  $473.05$ &    $85.9$ &   $383.8$ \\
-%\textsc{METAPhoR} 	& $0.0438$ &  $298.56$ &   $248.6$ &   $715.5$ \\
-%%\textsc{CMNN} 		& $-$ & $-$ & $-$ & $-$ \\
-%\textsc{CMNN}         & $0.0795$  &$1011.11$ &  $3908.2$ &  $6307.5$ \\
-%\textsc{SkyNet} 	& $0.0747$ &  $763.00$ &  $2740.2$ &  $4216.4$ \\
-%\textsc{TPZ} 		& $0.1138$ & $1801.74$ &  $9046.2$ & $10565.7$ \\
-%%\textsc{Frankenz}	& $-$ & $-$ & $-$ & $-$ \\
-%\hline
-%\end{tabular}
-
-%Stacked $n(z)$ statistic
-%
-%\begin{tabular}{lrrrr}
-%\hline
-%\bf Photo-$z$ Code & \bf KS & \bf CvM & \bf AD & \bf AD \\
-%                   &        &         & $z=[0.002,2.0]$ & $z=[0.0,2.0]$        \\
-%\hline
-%\textsc{ANNz2} 		& $0.0237$ &  $25.676$ &   $276.4$ &   $276.0$ \\
-%\textsc{BPZ} 		& $0.0131$ &  $17.587$ &   $169.0$ &   $170.9$ \\
-%%\textsc{Delight} 	& $0.0184$ &  $18.975$ &   $151.5$ &   $151.5$ \\
-%\textsc{Delight}    & $0.0256$ &  $42.341$ &   $253.3$ &   $253.2$ \\
-%%\textsc{EAZY} 		& $0.0434$ & $167.329$ &   $811.3$ &   $817.0$ \\
-%\textsc{EAZY}       & $0.0441$ & $169.977$ &   $827.7$ &   $833.3$ \\
-%%\textsc{FlexZBoost} & $0.0065$ &   $2.764$ &    $24.7$ &    $24.6$ \\
-%\textsc{FlexZBoost} & $0.0128$ &   $7.007$ &  $127.02$ &   $126.9$\\
-%\textsc{GPz} 		& $0.0168$ &  $25.605$ &   $307.1$ &   $307.1$ \\
-%\textsc{LePhare} 	& $0.0254$ &  $58.900$ &   $477.4$ &   $472.0$ \\
-%\textsc{METAPhoR} 	& $0.0350$ &  $60.790$ &   $671.9$ &   $672.3$ \\
-%%\textsc{CMNN} 		& $-$ & $-$ & $-$ & $-$ \\
-%\textsc{CMNN}         & $0.0043$ & $0.538$   &   $5.6$   &      $5.8$ \\
-%\textsc{SkyNet} 	& $0.0483$ & $385.830$ &  $2119.4$ &  $2110.3$ \\
-%\textsc{TPZ} 		& $0.0126$ &   $9.483$ &   $110.7$ &   $110.6$ \\
-%%\textsc{Frankenz}	& $-$ & $-$ & $-$ & $-$ \\
-%\hline
-%\end{tabular}
-%\end{table*}
-
-\begin{table}  %%% DATA TABLE %%%
-\centering
-\caption{CDE loss statistic for each photo-$z$ code.} \label{tab:cdeloss}
-\begin{tabular}{lr}
-\hline
-\bf Photo-$z$ Code & \bf CDE Loss \\
-\hline
-\textsc{ANNz2} 		& $-6.88$ \\
-\textsc{BPZ} 		& $-7.82$ \\
-%\textsc{Delight} 	& $-4.06$ \\
-\textsc{Delight}    & $-8.33$\\
-%\textsc{EAZY} 		& $-7.97$ \\
-\textsc{EAZY}       & $-7.07$ \\
-%\textsc{FlexZBoost} & $-11.51$ \\
-\textsc{FlexZBoost} & $-10.60$\\
-\textsc{GPz} 		& $-9.56$ \\
-\textsc{LePhare} 	& $-1.66$ \\
-\textsc{METAPhoR} 	& $-6.28$ \\
-%\textsc{CMNN} 		& $-$ \\
-\textsc{CMNN}         & $-10.43$ \\
-\textsc{SkyNet} 	& $-7.89$ \\
-\textsc{TPZ} 		& $-9.55$ \\
-\hline
-\trainz		& $-0.83$ \\
-%\textsc{Frankenz}	& $-$  \\
-%\hline
-\end{tabular}
-\end{table}
-
-\begin{table}
-\setlength{\tabcolsep}{2pt}
-\begin{center}
-\caption{Root-Mean-Square-Error (RMSE) statistics for the eleven photo-z codes for both PIT and $\hat{N}(z)$ distributions.}\label{tab:rmse}
-\begin{tabular}{lcc}
-\hline
-\hline
-Root-Mean-Square-Error \\
-(RMSE) statistics \\
-\hline
-Photo-z Code & PIT RMSE & N(z) RMSE\\
-\hline
-\textsc{ANNz2}      & 0.019 & 0.0054\\
-\textsc{BPZ}        & 0.032 & 0.0050\\
-\textsc{Delight}    & 0.111 & 0.0056\\
-\textsc{EAZY}       & 0.054 & 0.0102\\
-\textsc{FlexZBoost} & 0.021 & 0.0022\\
-\textsc{GPz}        & 0.048 & 0.0042\\
-\textsc{LePhare}    & 0.028 & 0.0062\\
-\textsc{METAPhoR}   & 0.064 & 0.0081\\
-\textsc{CMNN}         & 0.108 & 0.0009\\
-\textsc{Skynet}     & 0.054 & 0.0144\\ 
-\textsc{TPZ}        & 0.082 & 0.0031\\
-\hline
-\trainz		& 0.0025 & 0.0013\\
-\end{tabular}
-\end{center} 
-\end{table}
-
-%\begin{table}
-%\begin{center}
-%caption{RMSE Statistics for the 11 Photo-z codes for DC1 $i<25.3$ ``gold'' sample}\label{tab:pitrmse}
-%%\begin{tabular}{|l|c|c|c|c|}
-%\begin{tabular}{lc}
-%\hline
-%\hline
-% PIT RMSE statistics \\
-%\hline
-%Photo-z Code & PIT RMSE\\
-%\hline
-%\texttt{ANNz2}      & 0.019\\
-%\texttt{BPZ}        & 0.032\\
-%\texttt{Delight}    & 0.111\\
-%\texttt{EAZY}       & 0.054\\
-%\texttt{FlexZBoost} & 0.021\\
-%\texttt{GPz}        & 0.048\\
-%texttt{LePhare}    & 0.028\\
-%\texttt{METAPhoR}   & 0.064\\
-%\texttt{NN}         & 0.108 \\
-%\texttt{Skynet}     & 0.054\\ 
-%texttt{TPZ}        & 0.082\\
-%\hline
-%\end{tabular}
-%\end{center} 
-%\end{table}
-%
-%\begin{table}
-%\begin{center}
-%\caption{RMSE Statistics for the 11 Photo-z codes for DC1 $i<25.3$ ``gold'' sample}\label{tab:nzrmse}
-%%\begin{tabular}{|l|c|c|c|c|}
-%\begin{tabular}{lc}
-%\hline
-%\hline
-% N(z) RMSE statistics \\
-%\hline
-%hoto-z Code & N(z) RMSE\\
-%\hline
-%\texttt{ANNz2}      & 0.0054\\
-%\texttt{BPZ}        & 0.0050\\
-%\texttt{Delight}    & 0.0056\\
-%\texttt{EAZY}       & 0.0102\\
-%\texttt{FlexZBoost} & 0.0022\\
-%\texttt{GPz}        & 0.0042\\
-%\texttt{LePhare}    & 0.0062\\
-%\texttt{METAPhoR}   & 0.0081\\
-%\texttt{NN}         & 0.0009\\
-%\texttt{Skynet}     & 0.0144\\ 
-%\texttt{TPZ}        & 0.0031\\
-%\end{tabular}
-%\end{center} 
-%\end{table}
-
-\begin{table}
-\setlength{\tabcolsep}{2pt}
-\caption{Moments of the stacked $\hat{N}(z)$ distribution}\label{tab:moments}
-\begin{tabular}{lccc}
-\hline
-\hline
- \multicolumn{4}{l}{Stacked $n(z)$ Moments} \\
-\hline
-              & 1st Moment & 2nd Moment & 3rd Moment \\
-\textsc{Truth}     & 0.701      &   0.630    & 0.671  \\
-\hline
-Photo-z Code       & 1st Moment & 2nd Moment & 3rd Moment\\
-\hline
-\textsc{ANNz2}     & 0.702      & 0.625      & 0.653    \\ 
-\textsc{BPZ}       & 0.699      & 0.629      & 0.671    \\ 
-\textsc{Delight}   & 0.692      & 0.609      & 0.638    \\ 
-\textsc{EAZY}      & 0.681      & 0.595      & 0.619    \\ 
-\textsc{FlexZBoost}& 0.694      & 0.610      & 0.631    \\ 
-\textsc{GPz}       & 0.692      & 0.605      & 0.619    \\ 
-\textsc{LePhare}   & 0.718      & 0.668      & 0.741    \\ 
-\textsc{METAPhoR}  & 0.705      & 0.628      & 0.657    \\ 
-\textsc{CMNN}        & 0.701      & 0.628      & 0.667    \\
-\textsc{Skynet}    & 0.743      & 0.708      & 0.797    \\
-\textsc{TPZ}       & 0.700      & 0.619      & 0.643    \\ 
-\hline
-\trainz	   & 0.699 		& 0.627 	& 0.666 \\
-\end{tabular}
-\end{table}
-
-\section{Summary and Discussion}
-\label{sec:discussion}
-%(Sam Schmidt, Ofer Lahav, Jeff Newman, Alex Malz, Eve Kovacs, Tony Tyson, Tina Peters)
-In this paper we presented results evaluating the photometric redshift PDF computation for eleven photo-$z$ codes.
-As discussed in Section~\ref{sec:metrics} the $p(z)$ should accurately reflect the relative likelihood as a function of redshift for each galaxy.
-All codes were provided a set of representative training data and tested on an idealized set of model galaxies with high signal-to-noise and photometry with no confounding effects due to blending, instrumental effects, the night sky, etc...
-included.
-The goal was not to determine a ``best'' photo-$z$ code: in many ways, this was a baseline test of a ``best case scenario'' to predict the expected photo-$z$ performance if a stage IV dark energy survey was to obtain complete training samples and perfectly calibrated their multi-band photometry.
-Given these idealized conditions, any deficiencies observed in a photo-$z$ code's performance should be a cause for concern, and may be evidence in a problem with either/both of the specific code implementation or the underlying algorithm.
-In order to meet the stringent LSST requirements on photo-$z$ performance, identifying and correcting such problems is an important first step before tackling more realistic data in future challenges.
-Most of the codes tested performed well, however, several did not meet the stringent goals that have been laid out for LSST photometric redshift performance.
-This is a cause for concern, given the idealized conditions, and the individual code responses will be studied in detail moving forward.
-One obvious trend in several of the codes tested was an overall over or underprediction of the widths of $p(z)$, as evidenced by the QQ plots and PIT histograms shown in Fig.~\ref{fig:pitqq}.
-A more careful tuning of bandwidth or smoothing during the validation process appears to be necessary for many of the machine learning based codes in order to improve the accuracy of $p(z)$.
-For narrow peaked $p(z)$ the parameterization of the PDF as evaluated on a fixed redshift grid could also have contributed to some overestimates of $p(z)$ width simply due to the finite resolution.
-After evaluating resuts such as those presented in \citet[]{Malz:qp}, in future analyses we plan to switch from a fixed grid to quantile-based storage of $p(z)$ in order to more efficiently and accurately store redshift PDF results.
-
-Another important factor to keep in mind when examining the results presented in this paper is the fact that they are at some level dependent on the metrics that we aim to optimize: in this case code participants were asked to submit their optimal measures of an accurate $p(z)$, so participants used the training/validation data to optimize their codes accordingly.
-Had we, instead, asked for an optimal $\hat{N(z)}$ the resulting metrics would be different for most, if not all, of the codes, as they would optimize toward a different goal.
-Specific metric choice can affect which codes are among the ``best'' codes.
-As stated earlier, there are cosmological science cases that require either individual galaxy photo-$z$ measures, or ensemble $\hat{N}(z)$ measures.
-We must be aware of that the optimal method for one is not necessarily optimal for the other, and in fact several photo-$z$ algorithms may be necessary in the final cosmological analysis in order to satisfy the requirements of all science use cases.
-The example of the simple \trainz\ estimator described in Section~\ref{sec:caution} shows a simple model with a $p(z)$ that is unrealistic for individual objects can still score very well on many of our metrics.
-It is important to look at {\it all} metrics, and keep in mind what information each metric conveys.
-%\red{mention 'null' and how 'dumb' algorithms can score well on metrics, need to keep all metrics in mind}
-%\aim{[reiterate the meaning of $p(z)$ and the goal -- emphasize finding ``best'' code in terms of impact of assumptions underlying the method \textit{or} establishing methodology for the realistic case of biased prior information and other science goals]}
-%\red{[the challenges faced in this project]}
-%\red{mention z$<$2, missing some important degeneracies, new sim will go to higher z}
-%\red{no quality cuts included, could identify outliers, but could also induce biases}
-%\red{[how optimization defined has impact, here we optimized p(z), not n(z), could be science-case specific if stringent requirements are needed, though that may be a computational/storage challenge if too many use cases needed.]}
-%\red{some discussion of output parameterization, limitations of 200 point fixed grid, in future switch to something like quantiles (cite qp paper)}
-We re-emphasize that the dataset tested was quite idealized, and discuss enhancements that will be added in future simulations to test photo-$z$ codes on increasingly realistic conditions in the following section.
-
-\subsection{Future work}
-\label{sec:futurework}
-The work presented in this paper is only the first step in characterizing current photo-$z$ codes and moving toward an improved photometric redshift estimator.
-This initial paper explored code performance in idealized conditions with perfect catalog-based photometry and representative training data.
-As mentioned in Section~\ref{sec:stackedmetrics} for the stacked $N(z)$ metrics we examined only the entire galaxy population with no selections in either photo-$z$ ``quality" or redshift.
-The cosmological analyses for weak lensing and large scale structure based measures plan to break galaxy samples into tomographic redshift bins, using photo-$z$ $p(z)$ to infer the redshift distribution for each bin.
-The specific selection used to determine these bins, both algorithmically and the specific bin boundaries, could induce biases due to indirect selections inherent in the photo-$z$ or other bin selection parameters.
-The effects of tomographic bin selection will be explored in a dedicated future paper.
-\red{[are there any references for this?
-I remember Gary Bernstein talking about this at a photo-z workshop in Japan, but I don't know that it was published.
-I believe Michael Troxel has discussed this as well.]}
-We also plan to propagate the uncertainties measured in a set of fiducial tomographic redshift bins in order to estimate impact on cosmological parameter estimation.
-%\item \red{cosmological parameters in DC2}
-%\item \red{mention z$<$2, missing some important degeneracies, new sim will go to higher z}
-%\red{no quality cuts included, could identify outliers, but could also induce biases}
-
-In future papers we will add more and more complexity to our simulated data in order to test photo-$z$ algorithms in increasingly realistic conditions.
-The most pressing concern is the impact of incomplete spectroscopic training samples.
-As discussed extensively in \citet{Newman:2015} a representative set of spectroscopically confirmed galaxies spanning the full range of both redshift and apparent magnitude is necessary as a training set to characterize the mapping from broad-band fluxes to photometric redshifts.
-However, due to a combination of factors due to both the galaxy SEDs and limitations of spectrographic instruments, redshift samples are known to be systematically incomplete, where certain galaxy types and redshift intervals fail to yield a redshift even at the longest integration times on current and near-future instruments.
-The more representative the training data, the better the performance of photo-$z$ algorithms will be.
-Current and upcoming surveys are putting in significant effort into obtaining these training samples \citep[e.~g.\,][]{Masters:2017}, however we still expect significant incompleteness for LSST-like samples, particularly at faint magnitudes.
-One major focus of an upcoming {\it LSST Dark Energy Science Collaboration Photo-z Working Group} data challenge is to produce a realistically incomplete training set of spectroscopic galaxies, modeling the performance of spectrographs, emission-line properties, and expected signal-to-noise to determine which galaxies will fail to yield a secure redshift.
-In addition to outright redshift failures we will model the inclusion of a small number of falsely identified secure redshifts where misidentified emission lines or noise spikes cause an incorrect redshift solution to be marked as a high quality identification.
-Even sub-per cent level contamination by false redshifts can impact photo-$z$ solutions at levels comparable to the stringent  requirements of some LSST science cases.
-We expect different systematics to occur in different photo-$z$ codes in response to training on incomplete data, particularly some of the machine learning methods.
-The response of the codes will inform future directions of code development.
-
-This initial paper explored a data set that was constructed at the catalog level, with no inclusion of the complications that come from measuring photometry from images.
-Future data challenges will move to catalogs constructed from mock images, including effects that will have great impact on photo-$z$ measurements.
-Object blending will be a major area of investigation, as the mixing of flux from multiple objects and the resultant change in measured colours is predicted to affect a large fraction of LSST galaxies \citep{Dawson:2016}, and will be one of the major contributing systematics for photo-$z$'s.
-Inclusion of differing observing conditions (seeing, clouds, variations in filter curves, Galactic dust, ...), as well as models for instrumental and system effects, sky masks, will all impact object photometry, and will be explored in the upcoming data challenge and their impacts described in upcoming papers.
-All underlying SEDs were parameterized as a weighted combination of five basis SEDs, with no additional accounting for host galaxy dust obscuration beyond what was encoded in the basis templates.
-This, in effect, limited the simulation to a very simple model of internal obscuration.
-Future simulations will include a more complicated and realistic treatment of host galaxy dust.
-
-The underlying simulation used in this work was based on a light-cone constructed to a maximum redshift of $z=2$.
-LSST imaging after 10 years of observations will include a significant number of $z>2$ galaxies in expected cosmology samples, and their inclusion does have potential significant implications for photo-$z$ measures: the high redshift galaxies lie at fainter apparent magnitudes and can have anomalous colours due to evolution of stellar populations and the shift to rest-frame magnitudes probing UV features of the underlying SED.
-More importantly, one of the most common ``catastrophic outlier'' degeneracies observed in deep photometric samples occurs when the Lyman break is mistaken for the Balmer break, leading to multiple redshift solutions at $z\sim0.2-0.3$ and $z\sim2-3$ \citep{Massarotti:2001}.
-This degeneracy, along with other potential degeneracies, are currently not covered by the limited redshift range of this initial paper, which could mean that we are not probing the full range of potential extreme outlier populations and how our photo-$z$ estimators respond to them.
-Extending simulations to include the high-redshift galaxy population will be a priority in future data challenges.
-
-%\item \red{inclusion of incompleteness effects on training sample, ``wrong redshifts'', filter curves}
-%\item \red{inclusion of photometric errors, image based effects, blending, lensing, sky, mask, etc...}
-
-In this study we have not accounted for the presence of Active Galactic Nuclei (AGN) contributions to galaxy fluxes.
-In some cases, AGN will be easily identified from the colors and morphologies, i.e.
-the case of the brightest quasars where the nuclear activity outshines the host galaxy, and numerous studies have utilized color selection to create large samples of quasars \citep[e.g.][]{Richards:06,Maddox:08,Richards:15}.
-In current deep fields, similar in depth to what we expect from LSST, variability information and multi-wavelength data have been critical to not only identify AGN dominated galaxies, but also obtain more accurate photometric redshifts \citep[e.g][]{Salvato:11}.
-
-In addition to AGN dominated galaxies, those with lower levels of nuclear activity present a more insidious problem, where AGN features may not be apparent, but the colors and other host galaxy properties are perturbed relative to galaxies with an inactive nucleus.
-In such cases, the presence of the AGN may induce a bias if the template SEDs or empirical datasets do not include low-level AGN counterparts.
-For LSST, we will need to identify and obtain accurate photometric redshifts of all types of AGN for a range of science goals, whether it is to eliminate such objects from cosmology experiments, or to use them with confidence, all the way through to understanding galaxy evolution and the role that AGN may play in influencing galaxy properties over cosmic time.
-
-A promising route to classifying and obtaining accurate photometric redshifts for the AGN population is by combining machine learning with template-fitting techniques, as has recently been demonstrated by \citet{Duncan:18} for radio-selected AGN.
-This is because AGN are relatively easy to obtain spectroscopic redshifts for over all redshifts due to the strong emission lines that they exhibit, allowing very good training sets for machine learning algorithms to use.
-Whereas for those galaxies where the AGN is sub-dominant the galaxy templates are still adequate for obtaining reasonable photometric redshifts.
-
-In addition to these improvements, the DESC Photo-$z$ group plans to look at all potential methods to combine the results from multiple photo-$z$ codes to improve $p(z)$ accuracy, similar to the work presented in \citet{Dahlen:13,Carrascokind:14,Duncan:18}.
-Taking advantage of multiple algorithms that use observables in slightly different ways has shown promise, however we must be very conscious of whether a potential combination properly treats the covariance between the methods, given that they are estimating quantities based on the same underlying observables.
-Several science cases wish to estimate physical quantities along with redshift, for example galaxy stellar mass and star formation rate.
-Proper joint estimation of redshift and physical quantities requires an in depth understanding of galaxy evolution, and progress on accurate bivariate redshift probability distributions will go hand in had with progress on understanding galaxies themselves.
-Parameterization and storage of a complex 2-dimensional probability surface for potentially billions of galaxies (or even subsets of hundreds of thousand of particular interest) pose a potential challenge.
-These issues will be examined in another future paper.
-
-Finally, while this paper and future papers discussed above focus on photometric redshift codes and estimating accurate $p(z)$ from training data, we plan a separate, but complementary, project to examine calibration of the resultant redshifts via spatial cross-correlations \citep{Newman:2008}, which will be explored in a separate series of future papers.
-The overarching plan describing everything laid out in this section is described in more detail in the LSST DESC Science Roadmap (see Footnote in Section~\ref{sec:intro}).
-These plans will require significant effort, but they are necessary if we are to make optimal use of the LSST data for astrophysical and cosmological analyses.
-%\begin{enumerate}
-%\item \red{the combination of codes}
-
-%\item \red{$p(z,\alpha)$}
-%\item \red{tomographic analysis later on}
-%\red{this paper deals with training, future papers will also explore calibration via cross-correlation methods}
-%\end{enumerate}
-%\red{mention the SRM once again, say that the plans are extensive, but we do have a plan and a rough timeline.}
-
-\section*{acknowledgements}
-\red{[summary of effort breakdown], will need to be in authors.csv for auto-generation by start\_paper}
-
-\red{[personal funding sources]}
-
-\red{[NERSC computation acknowledgement}
-
-The LSST Dark Energy Science Collaboration acknowledges generous ongoing support from the agencies and institutes supporting the collaboration.
-These include the Institut National de Physique Nucl\'eaire et de Physique des Particules in France, the Science \& Technology Facilities Council in the United Kingdom, and the Department of Energy, the National Science Foundation, and the LSST Corporation in the United States.
-This research used resources of the National Energy Research Scientific Computing Center, a DOE Office of Science User Facility supported by the Office of Science of the U.S.
-Department of Energy under Contract No.
-DE-AC02-05CH11231.
-Part of this work was undertaken on STFC DiRAC HPC Facilities, funded by UK BIS National E-infrastructure capital grants, and on the UK particle physics grid, supported by the GridPP Collaboration.
-
-\appendix
-
- \section{Point Estimate Photometric Redshifts}
- \label{sec:pointmetrics}
- While we do not recommend the use of single point estimates of redshift for most science applications, plots of the point estimates can be a useful qualitative diagnostic of photo-z code performance, i.~e.~examining point photo-$z$ vs.~spec-$z$ plots visually can give a quick impression of some common trends in different codes.
-Computing point estimate statistics may also be useful for more direct comparisons with previous photo-z evaluations.
-If a point-estimate is preferred for a specific science case, it is fairly simple to compute the mean, mode, or some other simple estimator from each $p(z)$, so these point estimates can be easily derived from the stored $p(z)$.
-
-There are several common point estimators of photo-$z$ posteriors employed by different codes, e.~g.
-the mode, mean, median of the $p(z)$ distribution.
-In addition, many of the machine learning based estimators can be set up to return a single redshift solution.
-For example, SkyNet can be configured to run as a regressor that returns a single float rather than a classifier that returns a 200-bin $p(z)$ estimate.
-The single value returned by a machine learning based code may not correspond to a particular measure such as the mode or mean, and so to avoid interpretation of results that might be introduced by variations in choice of specific point-estimate implementation per code, we discard the code-specific point estimates.
-We instead calculate point estimates more uniformly across the codes directly from the $p(z)$ using two measures, $z_{PEAK}$ and $z_{WEIGHT}$.
-$z_{PEAK}$  is simply the maximum value attained for each galaxy p(z), the mode of the probability distribution.
-$z_{WEIGHT}$ is defined similarly to how it is defined in \citet{Dahlen:13}, as the weighted mean of the redshift over the {\it main peak} of $p(z)$ containing the $z_{PEAK}$ value.
-The main peak is defined by subtracting 0.05$\,\times\,z_{PEAK}$ from $p(z)$ and identifying the roots to isolate the peak containing $z_{PEAK}$, $z_{WEIGHT}$ is defined as the weighted mean redshift within this peak.
-We restrict to a single peak in order to avoid confusion from bimodal and multimodal $p(z)$ such as those shown in bottom panels of Figure~\ref{fig:pz_examples}.
-For example, for a bimodal probability distribution a weighted mean calculated over both peaks would fall between the peaks, at a redshift where the probability is minimal.
-Restricting the weighting to a single peak ensures that the point estimate will fall in the region of maximum redshift probability.
-%\red{Have Dritan check zweight description to make sure this is what was done.}
-% The most common point estimators of photo-$z$ PDFs are the mode, median, and mean of the distribution.
-Following \cite{Tanaka:17}, we report the values of the following metrics using whichever of these point estimators performs best for each method, noting which point reduction is used in each case.
-
- \subsection{Point Estimate Metrics}
- \label{sec:point_metrics}
-We calculate the commonly used point estimate metrics of the overall photo-$z$ scatter ($\sigma_{z}$, the standard deviation of the photo-$z$ residuals), bias, and ``catastrophic outlier rate''.
-Specifically, we calculate the metrics as follows:
-we define $e_{z}$ as
-
-\begin{equation}
-e_{z} = \frac{z_{P} -z_{S}}{1+z_{S}} 
-\end{equation}
-where $z_{P}$ is the point estimate and $z_{S}$ is the true redshift.
-In practice, because the standard deviation calculation is quite sensitive to the outliers, we define the photo-$z$ scatter, $\sigma$ in terms of the Interquartile Range (IQR), the difference between the 75th and 25th percentiles of the $e_{z}$ distribution.
-In order to match the usual meaning of a 1$\sigma$ interval, we scale the IQR and define $\sigma_{IQR} = IQR/1.349$, as there is a factor of 1.349 difference between the IQR and the standard deviation of a Normal distribution.
-%$\sigma_{z}$ is, in practice, defined in terms the Interquartile Range (IQR) of $e_{z}$ values as $IQR/1.349$, where $IQR$ is the difference between the 75th and 25th percentiles of the $e_{z}$ distribution.
-The factor of 1.349 relates the IQR of a Normal distribution to the standard 68\% range of a one $\sigma$ uncertainty.
-%\scc{usually the statistical indicator defined as $\sigma$ is the standard deviation I know that is not a rule but if someone skips the test and go directly to the table (and we all know that a lot of people skips the definition of statistical indicator) could be confusing, could we report it as $\frac{IQR(e_z)}{1.349}$ or $\sigma_{IQR}$ that should be clear without any risk of confusion?}\red{Using IQR is done because IQR is less sensitive to outliers than calculating a sigma.
-I've rewritten the text explaining this above.--SJS}
-While many other studies define the bias based on the {\it mean} offset between true and estimated redshift, in this study we define the bias as the median value of $e_{z}$ for the sample.
-We use median as it is, once again, less sensitive to outliers than the mean.
-The catastrophic outlier fraction is defined as the fraction of galaxies with $e_{z}$ greater than the {\it larger} of $3\sigma_{IQR}$ or 0.06, i.e.
-3$\sigma$ outliers with a floor of $\sigma_{IQR}$=0.02.
-%\scc{again, is clearly not a rule but a huge number of paper refers to the mean as bias, could we use the word median, which is clear without any doubt rather than bias?}
-%\red{I also have code to calculate $\sigma_{MAD}$, should we include this as well?  It's almost always within a few thousandths of $\sigma_{IQR}$, so I left it out for now}.
-%\scc{we may just say that the two indicators reports almost the same value therefore there is no need for both of them?}
-For reference, the goals stated in Section 3.8 of the LSST Science Book \citep{Abell:09} for photo-$z$ performance in these metrics, assuming perfect training knowledge (as we are testing in this paper) are:
-\begin{itemize}
-\item RMS scatter$ < 0.02(1+z)$ 
-%\scc{actually the SB reports in page 75 that the goal of 0.02 is for the RMS of $\frac{\sigma_{z}}{(1+z)}$ while we are using IQR}
-\item bias $<$ 0.003 %\scc{in page 518 of the SB is clearly reported that the mean is expected to be less than 0.003 in our case we have defined as bias the median and not the mean }
-\item catastrophic outlier rate $<$ 10\% %\scc{clearly the number of outliers since are defined above $3\sigma$ strictly depends on the sigma definition, in the science book it seems to refer to the RMS of the $e_z$ distribution even ef it could be the standard deviation, but again I don't think it refers to IQR, therefore we could not consider exactly this numbers as reference.}
-%\red{I am going off of conversations with Zeljko and how we {\it actually} computed the metrics in the Science Book (I have the scripts).
-You are correct that several of these are slightly different than actually stated in the Science Book, or where the Science Book does not use very precise language as to what was done to compute the RMS or define the bias.
-We can also cut out those specifics, or maybe say Ivezic private communication, maybe?
-Also $\sigma_{z}$ defined in terms of IQR is exactly equivalent to the standard deviation when scaled by 1.349 for a Gaussian distribution, so the numbers are appropriate to cite in this context.
-IQR is just a more robust way of calculating sigma for a distribution with outliers.--SJS }
-%\scc{I am very sorry I am not saying that IQR is not fine, I want to keep the IQR my concerns are related to the notation and the \textit{reader understanding}, if I wrote $\sigma$ somewhere in a table the reader will understand for sure that is the traditional standard deviation (a lot of people skips the indicators definition for the \textit{obvious} ones), I am suggesting to replace the notation $\sigma$ with the notation $\sigma_{IQR}$ (or something like this) to avoid any confusion, and also to change from $bias$ to $median$ for the same reason.
-I agree that say Ivezic private communication could be better.}
-\end{itemize}
-These definitions are similar, but not exactly the same, as the $\sigma_{IQR}$ and median bias calculated here, but are similar enough for qualitative comparisons to the LSST goals.
-
-%Scatter: SRD $\sigma<0.02(1+z)$
-%Catastrophic Outler Rate: SRD $3\sigma$ ``catastrophic outlier'' rate below 10 per cent
-%Bias: SRD bias$<0.003$
-
-Fig.~\ref{fig:pz_pointestimates}  shows the point estimates for both $z_{PEAK}$ and $z_{WEIGHT}$.
-Point density is shown with mixed contours to emphasize that most of the galaxies do fall close to the $z_{phot}=z_{spec}$ line, while blue points show differing characteristics of the outlier populations.
-The red dashed lines indicated the cutoff for catastrophic outliers, defined as: $max(0.06,3\sigma_{IQR})$.
-As with the full $p(z)$ results, a variety of behaviours are evident in the different codes.
-Table~\ref{tab:pointestimates} lists the scatter, bias, and catastrophic outlier fractions for the codes.
-The performance of the codes for point metrics is highly correlated with performance on $p(z)$ based tests, which is to be expected, given that the point-estimates were derived from the $p(z)$.
-Some discretization is evident in $z_{PEAK}$, particularly for \textsc{SkyNet}, due to the finite grid spacing of the reported $p(z)$.
-These discreteness effects are mitigated by the weighting of $z_{WEIGHT}$, resulting in a smoother distribution of redshift estimates.
-Several features perpendicular to the main $z_{phot}=z_{spec}$ line are evident.
-These features are due to the 4000 angstrom break passing through the gaps between adjacent LSST filters.
-These features are most prominent in template-based codes, but appear to some degree in all codes tested.
-
-In even the best performing codes, there are visible occupied regions away from the $z_{phot}=z_{spec}$ line, corresponding to degenerate redshift solutions for certain LSST magnitudes and colors.
-While use of the full information available via $p(z)$ mitigates their impact, a full understanding of the outlier population is critical for LSST science, particularly in tomographic applications %\red{I forget what exactly I was trying to say here, if someone else wants to take a crack at this it might be helpful --SJS}
-%\scc{may be worth to say that the usage of further bands could mitigate this effect by removing the degeneration  usually induced from emission lines moving across different filters?}
-
-Finally, we note that all eleven codes are at or near the goals for point-estimates as outlined in the LSST Science Requirements Document\footnote{available at: \url{http://ls.st/srd}} and \citet{Graham:17}.
-This is to be expected, given that the requirements were designed such that a point estimate photo-z would meet these requirements for perfect training data to a depth of $i<25$.
-But, it is still an encouraging sign, given an updated mock galaxy simulation and the expanded set of photo-$z$ codes tested.
-%\red{maybe say consistent with expectations, and a reference to Melissa's paper in here?  What else to say about nearly meeting point goals?}
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.9\textwidth]{figures/ZPEAK_szpz_threecolumn_12codes_navy.jpg}
-\includegraphics[width=0.9\textwidth]{figures/ZWEIGHT_szpz_threecolumn_12codes_navy.jpg}
-\caption{Point estimate photo-z's derived from the posteriors.
-Top panel shows $z_{PEAK}$, while bottom panel shows $z_{WEIGHT}$.
-Point estimate density is represented with fixed density contours, while outliers at lower density are represented by blue points.
-While use of point-estimate photo-$z$'s is not recommended, they do make for useful comparative and visual diagnostics.
-In the lower-right panel of each plot, the \trainz\ estimator results in identical photo-$z$ estimates at the mode and mean of the training set $N'(z)$ distribution for all galaxies.} \label{fig:pz_pointestimates}
-\end{figure*}
-
-\begin{table*}
-\begin{center}
-\caption{Point estimate statistics}\label{tab:pointestimates}
-\begin{tabular}{lcccccc}
-\hline
-\hline
-                  &            & $Z_{PEAK}$  &          &  & $Z_{WEIGHT}$          &\\
-\hline
-Photo-z Code       & $\frac{\sigma_{IQR}}{(1+z)}$ & median  & \multicolumn{1}{|p{0.75cm}|}{\centering outlier \\fraction} & $\frac{\sigma_{IQR}}{(1+z)}$ & median & \multicolumn{1}{|p{0.75cm}|}{\centering outlier \\fraction}\\
-\hline
-\textsc{ANNz2}     & 0.0270  &  0.00063  & 0.044      & 0.0244  &  0.000307  & 0.047  \\ 
-\textsc{BPZ}       & 0.0215  & -0.00175  & 0.035      & 0.0215  & -0.002005  & 0.032 \\ 
-\textsc{Delight}   & 0.0212  & -0.00185  & 0.038      & 0.0216  & -0.002158  & 0.038 \\ 
-\textsc{EAZY}      & 0.0225  & -0.00218  & 0.034      & 0.0226  & -0.003765  & 0.029 \\ 
-\textsc{FlexZBoost}& 0.0154  & -0.00027  & 0.020      & 0.0148  & -0.000211  & 0.017 \\ 
-\textsc{GPz}       & 0.0202  & -0.00091  & 0.036      & 0.0201  & -0.000950  & 0.037 \\ 
-\textsc{LePhare}   & 0.0236  & -0.00161  & 0.058      & 0.0239  & -0.002007  & 0.056 \\ 
-\textsc{METAPhoR}  & 0.0264  &  0.00000  & 0.037      & 0.0262  &  0.001333  & 0.048 \\ 
-\textsc{CMNN}        & 0.0184  & -0.00132  & 0.035      & 0.0170  & -0.001049  & 0.034 \\
-\textsc{Skynet}    & 0.0219  & -0.00167  & 0.036      & 0.0218  &  0.000174  & 0.037 \\
-\textsc{TPZ}       & 0.0161  &  0.00309  & 0.033      & 0.0166  &  0.003048  & 0.031 \\ 
-\hline
-\trainz	   & 0.1808  &  -0.2086  & 0.000	  & 0.2335  & 0.022135  & 0.000\\
-\end{tabular}
-\end{center} 
-\end{table*}
-
-\bibliographystyle{mn2e}
-\bibliography{references}
-\end{document}

### 2018-08-19 paper/tex/results.tex
should now compile!

--- 2018-08-12 paper/main.tex
@@
-%\documentclass[a4paper]{article}
-\documentclass[usenatbib]{mn2e}
-%% Language and font encodings
-\usepackage[english]{babel}
-\usepackage[utf8x]{inputenc}
-\usepackage[T1]{fontenc}
-\usepackage{lineno}
-\linenumbers
-\modulolinenumbers[2]
-
-\usepackage{natbib}
-%% Sets page size and margins
-\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
-
-%% Useful packages
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{graphicx}
-\usepackage{subcaption}
-\usepackage[colorinlistoftodos]{todonotes}
-\usepackage[colorlinks=true, allcolors=blue]{hyperref}
-\usepackage{multirow}
-
-\newcommand       \apj  {ApJ}
-\newcommand       \apjl  {ApJL}
-\newcommand       \apjs  {ApJS}
-\newcommand       \aj  {AJ}
-\newcommand       \aap {A\&A}
-\newcommand       \mnras {MNRAS}
-\newcommand       \nat {Nature}
-\newcommand       \araa {ARA\&A}
-\newcommand       \physrep {Phys.~Reports}
-\newcommand       \prd {Phys. Rev. D}
-
-\newcommand{\trainz}{\textsc{trainZ}}
-
-\newcommand{\textul}{\underline}
-\newcommand*\mathinhead[2]{\texorpdfstring{$\boldsymbol{#1}$}{#2}}
-
-\newcommand{\red}[1]{\textcolor{red}{#1}}
-% Claim a color for your comments
-\newcommand{\aim}[1]{\textcolor{green}{#1}}%Alex Malz comments in green
-\newcommand{\blue}[1]{\textcolor{blue}{#1}}%unknown person comments in blue
-\definecolor{scc}{rgb}{0.0, 0.26, 0.15}
-\newcommand{\scc}[1]{\textcolor{scc}{#1}}%Stefano Cavuoti Comments in dark green
-\definecolor{newpink}{rgb}{0.858, 0.188, 0.478}
-\newcommand{\erfan}[1]{\textcolor{newpink}{#1}} % Erfan Nourbakhsh comments in pink
-\newcommand{\jan}[1]{\textcolor{orange}{#1}}%Alex Malz comments in green
-
-\def\X{{\mathbf{X}}}
-\def\x{{\mathbf{x}}}
-\def\E{{\mathbb E}}
-
-\title{An assessment of photometric redshift PDF performance in the context of LSST}
-% Alternate title proposal: Assessing the characterization of photometric redshift uncertainties in the regime of LSST
-
-\author{LSST-DESC Photometric Redshift Working Group}
-
-\begin{document}
-\maketitle
-
-\begin{abstract}
-Photometric redshift (photo-$z$) probability distribution functions (PDFs) are a key data product of nearly all upcoming galaxy imaging surveys.
-However, the photo-$z$ PDFs resulting from different techniques are not in general consistent with one another, and an optimal method for obtaining an accurate PDF remains unclear.  
-We present the results of an initial study of the Large Synoptic Survey Telescope Dark Energy Science Collaboration (LSST-DESC), the first in a planned series of papers testing multiple photo-$z$ codes on simulations of upcoming LSST galaxy photometry catalogues.
-This initial test evaluates photo-$z$ algorithms in the presence of representative training data and in the absence of several common sources of systematic errors that affect the procedures by which photo-$z$ PDFs are derived.
-The photo-$z$ PDFs are evaluated using multiple metrics including the Kolmogorov-Smirnoff statistic, Cramer-von Mises statistic, Anderson-Darling statistic, Kullback-Leibler divergence, $N(z)$ moments, quantile-quantile plots and probability integral transform.
-We observe several trends, including an overall over/under-prediction in the broadness of the PDFs for several of the codes.
-A careful accounting of all systematics discovered will be necessary for the codes employed in upcoming analyses in order to achieve unbiased cosmological measurements.
-\end{abstract}
-\begin{keywords}
-galaxies: distances and redshifts -- galaxies: statistics -- methods: statistical
-\end{keywords}
-%\noindent comments color-code:\\
-%\red{SJS},
-%\aim{AIM},
-%\scc{SC},
-%\textcolor{brown}{MB},\jan{JAN}
-%\blue{unknown}\\
-%\red{note: not all red text is mine, at some point various people were using red --SJS.}
-
-% Questions left unanswered:
-% 1. Which journal will this paper be published? 
-% Answer: AJ / ApJ / MNRAS (no charges, but not sure if it will accept a paper based on a DC1 effort and the associated simulations)
-% 2. What is the meaning of $p(z)$ in different codes? (Ofer)
-% Answer: ?
-%
-
-\section{Introduction}
-\label{sec:intro}
-
-%(Sam Schmidt, Ofer Lahav, Jeff Newman, Alex Malz, John Soo, Tony Tyson)
-
-%\jan{Is the intention to keep these parenthetical author sets?
-I find it a bit offputting in the text.
-It might be better to put this for sections in the acknowledgements section, and for individual codes/statistics specify who ran/developed it with a sentence.}\red{I was just about to comment these out before finalizing the internal review draft of the paper.
-They are in place to remind me to contact a few people who still have not filled out their authors.csv info.}
-
-Large-scale photometric galaxy surveys are entering a new era with currently or soon-to-be running Stage III and Stage IV dark energy experiments like the Dark Energy Survey \citep[DES,][]{Abbott:05}, the Kilo-Degree Survey \citep[KiDS,][]{de_Jong:13}, Large Synoptic Survey Telescope \citep[LSST,][]{Abell:09}, Euclid \citep{Laureijs:11}, and Wide-Field Infrared Survey Telescope \citep[WFIRST,][]{Green:12}.
-The move to imaging based surveys, rather than spectroscopic based, for cosmological measurements makes proper understanding of photometric redshifts (``photo-$z$'s'') of paramount importance, as cosmological distance measures for statistical samples are directly dependent on photo-$z$ measurements.
-
-The unprecedented sample size of LSST galaxies, expected to number several billion for the main cosmological sample, necessitates stringent constraints on photo-$z$ accuracy if systematic errors are not to dominate the statistical errors.
-The LSST Science Requirements Document (SRD)\footnote{available at \url{https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17}} lists the photometric redshift goals for a magnitude limited sample with $i<25$ as: root-mean-square error with a goal of $\sigma_z<0.02(1+z)$; $3\sigma$ ``catastrophic outlier'' rate below $10\%$; bias below $0.003$\footnote{
-Note that at the time the SRD was written, these goals were stated in terms of a photo-$z$ point estimate for each galaxy, as was standard in many previous studies, while in this paper we emphasize the importance of using a full photo-$z$ PDF.}.
-
-The tremendous size of LSST's galaxy catalog will be enabled by its exceptional depth, pushing to fainter magnitudes and deeper imaging and including galaxies of lower luminosity and higher redshift than ever before.
-In addition to the contribution of low signal-to-noise photometry to the systematic error of photo-$z$s, these populations introduce major physical degeneracies, for example the Lyman break/Balmer break degeneracy, that were not present in the populations covered in previous large area surveys like the Sloan Digital Sky Survey \citep[SDSS,][]{York:00} and the Two Micron All Sky Survey \citep[2MASS,][]{Skrutskie:06}.
-In order to meet LSST's demanding error budget, it will be necessary to fully characterize those degeneracies wherein multiple redshift solutions have comparable likelihood.
-
-There is often a desire to have a single valued ``point-estimate'' redshift for an individual galaxy.
-However, the complex, non-linear (and often non-unique) nature of the mapping between broad band fluxes and redshift means that a single value is unable to capture the full redshift information encoded in a galaxy's magnitudes.
-For example, a common point-estimate for a template-based method is taking the highest likelihood solution as the point photo-$z$.
-A single valued redshift ignores degenerate redshift solutions of lower probability, potentially biasing photometric redshift estimates both for individual galaxies and ensemble distributions.
-Storing more information is necessary, most often photo-$z$ codes output the redshift probability density function (PDF), also often referred to as $p(z)$, describing the relative likelihood as a function of redshift.
-%\red{should we move some of AIM's p(z) vs p(z|d,I) to here?}
-Early template methods such as \citet{Fernandezsoto:99} converted relative $\chi^2$ values of template spectra to likelihoods to estimate $p(z)$.
-Soon after, codes such as \citet{Benitez:00} added a Bayesian prior and output a posterior probability distribution.
-While many early machine learning based algorithms focused on a point-estimate, \citet{Firth:03} used a neural net with 1000 realizations scattered within the photometric errors to estimate a $p(z)$.
-As more groups began to employ photometric redshifts in their cosmological analyses, realization that point-estimate photo-$z$'s were inadequate for precision cosmology measurements \citep{Mandelbaum:2008}.
-From around this point onward, most photo-$z$ algorithms have attempted to implement some estimate of the overall redshift probability in their outputs, and some surveys began supplying a full $p(z)$ rather than a simple redshift point-estimate and error \citep[e.~g.~][]{de_Jong:17}.
-
-%\aim{We should spend a whole paragraph establishing that photo-$z$ PDFs are a thing, i.e.
-review the motivation for PDFs over point estimates, basic definition, early methods papers, surveys that have published them, etc.}
-%\scc{I agree with Alex, we should introduce here what a pdf is and which survey have already published some pdfs (KiDS dr3 for example) or will produce them (LSST, Euclid etc.)}
-%\red{See above paragraph, does this adequately address this?--SJS}\scc{fine for me}
-
-There are numerous techniques for deriving photo-$z$ PDFs from photometry, yet no one method has yet been established as clearly superior.
-Quantitative comparisons of photo-$z$ methods have been made before.  
-The Photo-$z$ Accuracy And Testing \citep[PHAT,][]{Hildebrandt:10} effort focused on point estimates derived from many photometric bands. 
-DES compared several codes for point estimates \citep{Sanchez:14} and a summary statistic of photo-$z$ interim posteriors for tomographically binned galaxy subsamples \citep{Bonnett:16}. 
-This paper is distinguished by its inclusion of metrics of photo-$z$ interim posteriors themselves and consideration of both classic and state-of-the-art photo-$z$ algorithms, comparing the performance of several of the most widely employed codes as well as some that have been developed only recently on the basis of metrics appropriate for a probabilistic data product.
-The results presented in this work are 
-%part of the Data Challenge 1 (DC1) 
- a major focus of the Photometric Redshift working group of the LSST Dark Energy Science Collaboration (LSST-DESC).
-This work is laid out in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}} as one of the critical activities to be completed in preparation for dark energy science analysis on the first year LSST data.
-This is the first of multiple papers by the working group, which will grow in sophistication.
-In this initial paper we focus on evaluating the performance of photometric redshift codes and their ability to produce accurate PDFs in the presence of representative training sets.
-This can be thought of as an initial test under near perfect conditions, before further complexities are added in future papers.
-Comparing the relative performance of the codes enables us to evaluate whether each code is using information in an optimal way, and may reveal enhancements in some codes and deficiencies in others, either in the fundamental algorithm, or in specific implementation.
-
-Certain science cases need redshift information on individual objects, e.~g.~identification of host galaxy redshift for supernova classification, or identifying potential cluster membership.  
-Other science cases need only ensemble redshift information; for instance current weak lensing techniques require the overall redshift distribution $N(z)$ for tomographic redshift samples, but do not need single galaxy estimates.
-In the case of the multiple types of probes of cosmology enabled by the
-LSST cosmology sample of several billion galaxies, the number of redshift bins and their photo-$z$ requirements vary with the specific probe;  2-point angular correlations benefit from many bins, while weak lensing probes do not (due to the wide lensing kernel).
-Large photometric surveys such as LSST must develop algorithms that meet the needs of all such science cases.  
-In order to meet these ambitious goals for photo-$z$ accuracy, every aspect of photo-$z$ estimation will have to be optimized: the algorithms employed, both template and machine-learning based (both in design and implementation); the spectroscopic data used as a training set for machine learning algorithms or to estimate template sets and train Bayesian priors; and probabilistic catalog compression schemes that balance information retention against limited storage resources.
-
-Before moving forward, we must address how the best methods may be unique to the performance metrics and science cases considered and what distinguishes photo-$z$ PDFs of different methods from one another.
-Though photo-$z$ PDFs are often written simply as $p(z)$, the PDF itself must be an interim posterior distribution $p(z | d, I)$, the probability of redshift conditioned on photometric data $d$ that has actually been observed and the prior information $I$ that guides how a redshift is extracted from the photometry.
-If we run multiple photo-$z$ codes on a single dataset, the photo-$z$ interim posteriors will not be identical because each code is based on assumptions in the form of an interim prior --- these assumptions form the premise for photo-$z$ estimation as a whole and are the only way to introduce differences in estimates of what would otherwise be a shared photo-$z$ posterior $p(z | d)$ regardless of the code used to obtain it.
-Though explicit knowledge of the interim prior is necessary to use photo-$z$ interim posteriors self-consistently in physical inference, the interim prior of a particular methodology is often implicit and not necessarily shared among all galaxies in the catalog.
-
-This paper therefore aims (1) to constrain the impact of the interim prior $I$ by separating it into a component $I_{H}$ representing the method itself and a component $I_{D}$ representing physical information, such as a training set or SED template library and (2) to present a procedure for evaluating the performance of photo-$z$ codes in generic tests that may include many more systematics in the interim prior $I$.
-In order to isolate the effects encapsulated by $I_{H}$ of variation between codes from issues with the training set or template library encapsulated by $I_{D}$, we use an identical set of simulated galaxies for every code and construct a template library and training sample that are {\it complete and representative} and shared among all codes; that is, our training sample for machine learning codes is drawn from the same underlying galaxy population as our test set, with no additional selections, and the SED library used for template-based codes is the same as the one used to generate the photometric data.
-We explore a number of performance metrics in this paper, not to make a conclusion regarding the superiority or even relative favorability of each code but to establish a method for comparing photo-$z$ PDFs derived by different methods.
-These test conditions set the stage for addressing in a future paper the crucial issue of incomplete and non-representative prior information.
-
-The outline of the paper is as follows: in \S\,\ref{sec:sims} we present the simulated data set; in \S\,\ref{sec:pzcodes} we describe the current generation codes employed in the paper;  in \S\,\ref{sec:metrics} we discuss the interpretation of photo-$z$ PDFs in terms of metrics of accuracy; in \S\,\ref{sec:results} we show our results and compare the performance of the codes; in \S\,\ref{sec:discussion} we offer our conclusions and discuss future extensions of this work.
-
-\section{The simulation and mock galaxy catalog}
-\label{sec:sims}
-%(Sam Schmidt, Eve Kovacs, Tina Peters)
-
-In order to test the current generation codes, we employ a simulated galaxy catalogue.
-The simulation is completely catalogue-based, with no image construction or mock measurements made.
-We describe these in detail below.
-
-\subsection{Buzzard-v1.0 simulation}
-\label{sec:buzzard}
-The \textsc{Buzzard-highres-v1.0} \red{put in cites to in prep Buzzard papers} catalogue construction started with a dark matter only simulation.
-This N-body simulation contained $2048^3$ particles in a $400$ Mpc h$^{-1}$ box.
-\red{[N]} snapshots (with smoothing and interpolation between snapshots) were saved in order to construct a lightcone.
-Dark matter halos were identified using the \textsc{Rockstar} software package \citep{Behroozi:13}.
-These dark matter halos were populated with galaxies with a stellar mass and absolute $r$-band magnitude in the SDSS system determined using a sub-halo abundance matching model constrained to match both projected two-point galaxy clustering statistics and an observed conditional stellar mass function \citep{Reddick:13}.
-
-To assign an SED to each galaxy, the {\it Adding Density Dependent Spectral Energy Distributions} (\textsc{ADDSEDS}, deRose in prep.)\footnote{\url{https://github.com/vipasu/addseds}} procedure was used.
-This consisted of training an empirical relation between absolute $r$-band magnitude, local galaxy density, and SED using a sample of $\sim 5e^{5}$ galaxies from the magnitude-limited Sloan Digital Sky Survey Data Release 6 Value Added Galaxy Catalog \citep{Blanton:05}\red{[Note: is this the proper reference to SDSS-NYU VAGC?
-File is called combined\_dr6\_cooper.fits, but I don't see which Cooper et al 2006 this is supposed to refer to?]}.
-Each SDSS spectrum is fit with a sum of five SED components using the \textsc{k-correct \red{v?}} software package\footnote{\url{http://kcorrect.org}} \citep{Blanton:07}, thus each galaxy SED is parameterized as five weights for the basis SEDs.
-The distance to the spatial projected fifth-nearest neighbour was used as a proxy for local density in the SDSS training sample.
-For each simulated galaxy, a ``random'' \red{[details]} galaxy with ``similar'' \red{[details]} absolute $r$-band magnitude and local galaxy density was chosen from the training set, and that training galaxy's SED was assigned to the simulated galaxy.
-Given the SED, absolute $r$-band magnitude and redshift, we computed apparent magnitudes in the six LSST filter passbands, $ugrizy$.
-We assigned magnitude errors in the six bands using the simple model described in \citet{Ivezic:08}, assuming full 10-year depth observations had been completed.
-The number of total 30-second visits assumed when generating the photometric errors differs slightly from the fiducial numbers assumed for LSST: we assume 60 visits in u-band, 80 visits in g-band, 180 visits in r-band, 180 visits in i-band, 160 visits in z-band, and 160 visits in y-band.
-%numbers taken from Alex Abate's photErrorModel.py script in PhotoZDC1 repository, which was used with nYrObs=10.
-
-\subsubsection{Selection of training and test sets}
-\label{sec:buzztraining}
-The total catalogue covered $400$ square degrees and contained $238$ million galaxies to an apparent magnitude limit of $r\!=\!29$ and spanning the redshift range $0\!<\!z\!\leq\!8.7$.
-This catalogue contained two orders of magnitude more galaxies than were needed for this study, so only $\sim\!8$ square degrees were used.
-Systematic problems with galaxy colors above $z\!>\!2$ were observed, so the catalogue was trimmed to include only galaxies in the redshift range $0\!<\!z\!\leq\!2.0$.
-A random subset of the the remaining galaxies was chosen, and placed at random into either a ``training'' set ($10$ per cent of the sample), for which the galaxies true redshifts will be supplied, or a ``test'' set (the remaining $90$ per cent of the sample), for which each code will need to predict a redshift PDF for each galaxy.
-The resulting catalogues contain $111\,171$ training galaxies and $1\,000\,883$ test galaxies.
-We restrict our analysis to a sample with $i<25.3$, which give a signal-to-noise $\sim\,30$ for most galaxies, a cut often referred to as the expected ``LSST Gold Sample''.
-This magnitude cut results in a training set with $44\,404$ galaxies and a test set containing $399\,356$ galaxies.
-All subsequent results will evaluate this ``gold sample'' test set.
-
-\subsubsection{Templates}
-\label{sec:buzztemplates}
-As mentioned in Section~\ref{sec:buzzard}, the SEDs in the Buzzard simulation are drawn from an empirical set of SEDs taken from the SDSS DR6 NYU-VAGC, a sample of roughly $\sim5e^{5}$ galaxies with spectra in SDSS.
-To determine a finite set of templates to use with template fitting codes we take the five SED weight coefficients for each of the $\sim 500\,000$ galaxies in the SDSS sample and run a simple K-means clustering algorithm on this five dimensional space.
-The K-means cluster centres span the space of coefficients and properly reflect the underlying density in the coefficient space, thus providing a reasonable approximation for a spanning SED set.
-An ad-hoc number of $K=100$ was chosen and the $100$ K-means centre positions are taken as the weights for the \textsc{k-correct} SED components to construct one hundred template SEDs.
-These $100$ templates were provided,
-%\red{(JS: if I remembered correctly, weren't there 150 templates given?
-Answer: no, initially I supplied a list of 150 that were drawn with a different algorithm, but once we switched to the final k-means set I only supplied 100 (the initial algorithm over-weighted outlier SEDs, which is why we needed the extra 50)--SJS)}
-however not every template code uses this set of one hundred templates: because EAZY was designed and written to use the same five basis templates employed by \textsc{k-correct} when constructing our mock galaxies, EAZY was run using linear combinations of these five templates rather than using the 100 discrete templates.
-
-\subsubsection{Limitations}
-\label{sec:buzzlimitations}
-For our initial investigation of photometric redshift codes, we begin with a data set that is somewhat idealized, and does not contain all of the complicating factors present in real data.
-In several cases, the simplification is done with a purpose, with potentially confounding effects excluded in order to better isolate the differences between current-generation photo-$z$ codes, and their causes.
-We list several of the simulations limitations in this section.
-As the simulation is catalogue-based, no image level effects, such as photometric measurement effects, object blending, contamination from sky background (Zodiacal light, scattered light, etc...), lensing magnification, or Galactic reddening are included.
-No stars are included in the catalogue, nor are the effects of AGN.
-As all SEDs are constructed from only five basis templates, properties of the galaxy population will be restricted to follow linear combinations of the characteristics of the five basis templates, so certain non-linear features, for example the full range of emission line fluxes relative to the continuum, will not be included in the model galaxy population.
-No additional dust reddening intrinsic to the host galaxy is included, the only approximation of dust extinction comes in the form of dust encoded in the five basis SEDs via the training set used to create the basis templates.
-Simple linear combinations of these basis templats will, once again, not explore the full range of realistic dust extinction observed in galaxy populations.
-
-%\subsection{ABC-Galacticus simulation}
-%\label{sec:galacticus}
-%\red{The Argonne-Berkeley-Carnegie (ABC) Galacticus cosmological simulation uses a semi-analytic model to simulate the galaxy properties and is not directly tuned to observations.
-Each simulated SED is replaced with a SED drawn from the distribution of empirical SEDs given by the SED Atlas of \citet{Brown:14} (diffusion maps method in development).}
-
-\section{Methods}
-\label{sec:pzcodes}
-
-%(Sam Schmidt, Rongpu Zhou, Ibrahim Almosallam, Eric Nuss, Johann Cohen Tanugi, John Soo)
-
-Here we outline the photo-$z$ PDF codes tested in this study.
-In total, eleven distinct codes are tested.
-This sample is not comprehensive, but does cover a broad range of current-generation codes.
-Both template-based and machine learning approaches are included and each are described separately in Secs.~\ref{sec:templatecodes} and \ref{sec:trainingcodes} respectively.
-The list of codes are summarized in Table.~\ref{tab:list_of_codes}.
-
-\begin{table*}  %%% DATA TABLE %%%
-\caption{List of photo-$z$ codes featured in this study. ML here means machine learning.} \label{tab:list_of_codes}\resizebox{\textwidth}{!}{
-\begin{tabular}{llll}
-\hline
-\bf Code 			& \bf Type & \bf Paper & \bf Website \\
-\hline
-\textsc{BPz} 		& template 			& \citet{Benitez:00}		& \url{http://www.stsci.edu/~dcoe/BPZ/} \\
-\textsc{EAZY} 		& template 			& \citet{Brammer:08}		& \url{https://github.com/gbrammer/eazy-photoz} \\
-\textsc{LePhare} 	& template 			& \citet{Arnouts:99}		& \url{http://www.cfht.hawaii.edu/~arnouts/lephare.html} \\
-\textsc{ANNz2} 		& ML 	& \citet{Sadeh:16}			& \url{https://github.com/IftachSadeh/ANNZ} \\
-\textsc{Delight} 	& ML/template 	& \citet{Leistedt:17}		& \url{https://github.com/ixkael/Delight} \\
-\multirow{ 2}{*}{\textsc{FlexZBoost}} & \multirow{ 2}{*}{ML} 	& \multirow{ 2}{*}{\citet{Izbicki:17}}		& \url{https://github.com/tpospisi/flexcode};\\ 
-					& 		&							& \url{https://github.com/rizbicki/FlexCoDE}\\
-%\textsc{Frankenz} 	& ML 	& \citet{Speagle:frankenz}	& \url{https://github.com/joshspeagle/frankenz} \\
-\textsc{GPz} 		& ML	& \citet{Almosallam:15b}	& \url{https://github.com/OxfordML/GPz} \\
-\textsc{METAPhoR} 	& ML 	& \citet{Cavuoti:17}		& \url{http://dame.dsf.unina.it}\\
-\textsc{CMNN} 		& ML 	& \citet{Graham:17}			& - \\
-\textsc{SkyNet} 	& ML 	& \citet{Graff:14}			& \url{http://ccpforge.cse.rl.ac.uk/gf/project/skynet/} \\
-\textsc{TPZ} 		& ML 	& \citet{Carrasco_Kind:13}	& \url{https://github.com/mgckind/MLZ} \\
-\hline
-\trainz 	& N/A	& See Section~\ref{sec:method:trainz} & \\
-\end{tabular}}
-\end{table*}
-
-%\red{The questions that must be answered for each code are\dots}
-The questions that must be answered for each code are: what unique features are included in the specific implementation that influence the output $p(z)$.
-What form of validation was performed with the training data, how were photometric uncertainties employed in the analysis, how were negative fluxes treated, what specific prior form was employed (for template based codes), or what specific machine learning architcture was used (for ML codes)?
-%\scc{which parameter (in a general meaning) have been used, which kind of validation has been performed (if any)}
-\subsection{Template-based Approaches}
-\label{sec:templatecodes}
-
-\subsubsection{BPZ}
-\label{sec:BPZ}
-%(Sam Schmidt)
-
-\textsc{BPZ}\footnote{\url{http://www.stsci.edu/~dcoe/BPZ/}} \citep[Bayesian Photometric Redshift,][]{Benitez:00} is a template-based photo-$z$ code that compares the expected colors ($C$) calculated for a set of spectral energy distribution (SED) types/templates ($T$) to the observed colors to calculate the likelihood of observing colors at each redshift for each type, $p(C|z,T)$.
-The code employs an empirically determined Bayesian prior in apparent magnitude ($m_0$) and SED-type.
-Assuming that the SED-types are spanning and exclusive, we can determine the redshift posterior $p(z|C,m_0)$ by marginalizing over all SED-types with a simple sum \citep[Eq.~3 from][]{Benitez:00}:
-
-\begin{equation} \label{eq:redshift_posterior}
-p(z|C,m_0)\,\propto\, \sum_{T}p(z,T|m_0)\,p(C|z,T)
-\end{equation}
-\noindent where the first term on the right-hand side is the Bayesian prior and the second term is the traditional likelihood.
-The prior is assumed to have the form: $p(z,T|m_0)=p(T|m_0)\,p(z|T,m_0)$, i.e.
-it parameterizes the prior as an evolving type fraction with apparent magnitude, combined with a prior on the expected redshift probability distribution as a function of both apparent magnitude and SED-type.
-
-In this paper we use \textsc{BPZ v 1.99.3}. The template set employed here is the set of $100$ discrete SEDs described in Section~\ref{sec:buzztemplates}
-%, and the $129$ Brown SEDs for the Galacticus dataset. 
-To keep the number of free parameters to a manageable level the SEDs in the training set are sorted by the rest-frame $u$-$g$ colour and split into three ``broad'' SED classes, equivalent to the \texttt{E}, \texttt{Sp} and \texttt{Im/SB} types in \citet{Benitez:00}.
-We assume the same functional form for the Bayesian priors as used by \citet{Benitez:00}, and utilize the training-set galaxies with known SED-type, redshift, and apparent magnitude to determine the type fractions and the best fit for the eleven free parameters of the prior.
-%For photo-$z$ point estimates we use the \texttt{Z\_B} output parameter. 
-For galaxies with negative flux in a measured band, the placeholder value is replaced with an estimate one $\sigma$ detection limit in that particular band, i.~e.~a value close to the estimated sky noise threshold.
-The type-marginalized $p(z)$ is generated by setting the parameter \texttt{PROBS\_LITE=TRUE} in the \textsc{BPZ} parameter file.
-
-\subsubsection{EAZY} \label{sec:eazy}
-%(Rongpu Zhou)
-
-\textsc{EAZY}\footnote{\url{https://github.com/gbrammer/eazy-photoz}} \citep[Easy and Accurate Photometric Redshifts from Yale,][]{Brammer:08} is a template-based photo-$z$ code that includes several features that improve on the basic $\chi^2$ fit used in many template codes.
-It can fit the observed photometry with SEDs created from a linear combination of a set of templates at each redshift, and the best-fit SED is found by simultaneously fitting one, two or all of the templates by minimizing $\chi^2$.
-The minimized $\chi^2(z)$ is then combined with an apparent magnitude prior to obtain the posterior redshift probability distribution, although some argue that this is not the mathematically correct way of calculating the posteriors.
-\textsc{EAZY} can also account for the uncertainties in the templates by adding an empirically derived template error in quadrature as a function of redshift to the flux errors.
-
-In this paper we use the all-templates mode, which fits the photometric data with a linear combination of the five basis templates.
-We employed the $5$ basis templates described in Section~\ref{sec:buzzard}, and set the template error to zero since these same templates were used to produce the simulated catalog photometry.
-%\red{The quality parameter \texttt{Qz} is...}
-The likelihoods are calculated on a 200-point redshift grid spanning $0\leq z \leq 2$, and include the application of a type-independent apparent magnitude prior estimated from the training data.
-
-\subsubsection{LePhare}\label{sec:lephare}
-%(C\'ecile Roucelle, Eric Nuss, Johann Cohen-Tanugi)
-
-\textsc{LePhare}\footnote{\url{http://www.cfht.hawaii.edu/~arnouts/lephare.html}}\citep[Photometric Analysis for Redshift Estimate,][]{Arnouts:99,Ilbert:06} is a photo-$z$ reconstruction code based on a $\chi^2$ template-fitting procedure.
-The observed colors are matched with the colours predicted from a set of spectral energy distribution (SED) which can be either synthetic or based on a semi-empirical approach.
-\textsc{LePhare} has been used to produce the COSMOS2015 photo-$z$ catalogue \citep{Laigle:16}.
-
-Each SED is redshifted in steps of $\Delta z = 0.01$ and convolved with the simulated LSST filter transmission curves (accounting for instrument efficiency).
-The opacity of the inter-galactic medium has been set to zero as no additional reddening has been included in the Buzzard simulations.
-The computed photo-$z$ is then the value that minimizes the merit function $\chi^2 (z,T,A)$ from \citet{Arnouts:99}:
-
-\begin{equation}
-\chi^2(z,T,A) = \sum_f^{N_f}\left(\frac{F^f_{\rm{obs}}−A \times F^f_{\rm{pred}}(T,z)}{\sigma^f_{\rm{obs}}}\right)^2 
-\end{equation}
-\noindent where $A$ is a normalization factor, $F^f_{\rm{pred}}(T,z)$ is the flux predicted for a template $T$ at redshift $z$.
-$F^f_{\rm{obs}}$ is the observed flux in a given band $f$ and $\sigma^f_{\rm{obs}}$ the associated observational error.
-The index $f$ refers to the considered band and $N_f$ is the total number of filters.
-
-In this paper we use \textsc{LePhare v 2.2}. 
-The set of templates used for fitting the photo-$z$'s are the 100 discrete Buzzard SED templates as described in section~\ref{sec:buzztemplates}.
-%No correction for the photometric zero-points was applied for this work. 
-The full $p(z)$ corresponds to the likelihoods calculated at each point on our $z$-grid.
-
-\subsection{Training-based Codes}
-\label{sec:trainingcodes}
-
-\subsubsection{ANNz2}
-\label{sec:annz2}
-%(John Soo)
-
-\textsc{ANNz2}\footnote{\url{https://github.com/IftachSadeh/ANNZ}} \citep{Sadeh:16} is a powerful package that has the ability to employ several machine learning algorithms, including artificial neural networks (ANN), boosted decision tree (BDT) and k-nearest neighbour (KNN).
-Using the Toolkit for Multivariate Data Analysis (TMVA) with ROOT\footnote{\url{http://tmva.sourceforge.net/}}, it can run multiple machine learning algorithms for a single training and outputs photo-$z$'s based on a weighted average of their performances.
-
-\textsc{ANNz2} is capable of producing both photo-$z$ point estimates and redshift posterior probability distributions $p(z)$, it could also conduct classifications and supports reweighting between samples.
-The PDFs are produced by propagating the intrinsic uncertainty on the input parameters and the uncertainty in the machine learning method to the expected photo-$z$ solution, averaged over multiple runs weighted based on the performance of each run.
-\textsc{ANNz2} presents its photo-$z$ uncertainty different from many codes by using the KNN method: it estimates the photo-$z$ bias between each object and a fixed number of nearest neighbours in parameter space, it then takes the $68$th percentile width of the distribution of the bias.
-This is based on the implication that objects with similar photometric properties would have similar uncertainties, and therefore the photometric errors of the inputs are not propagated into the code.
-
-In this study, \textsc{ANNz2 v.
-2.0.4} was used.
-The full PDF for each galaxy is also produced with a linear stepsize of $z=0.01$ for $0 \leq z \leq 2$.
-A set of $5$ ANNs with architecture $6:12:12:1$ ($6$ $ugrizy$ inputs, $2$ hidden layers with $12$ nodes each, and $1$ output) with different random seeds are used during each training.
-%Training was conducted on a subset of the given training sample which satisfies $i\le25.3$ as it would have improved the results compared to training on the entire sample.
-Half of the training set is used as a validation set to prevent overtraining.
-All training objects are set to have detected magnitudes, however the non-detections (mag$=-99$) in the testing set are replaced with the mean of that particular band.
-
-\subsubsection{Color-Matched Nearest-Neighbours}
-\label{sec:cmnn}
-%(Melissa Graham)
-
-The nearest-neighbours color-matching photometric redshift estimator (\textsc{CMNN}) is presented in \cite[][herafter G18]{Graham:17}.
-This method uses a training set of galaxies with known redshifts that has equivalent or better photometry as the test set in terms of quality and filter coverage.
-For each galaxy in the test set we identify a color-matched subset of training galaxies, choose one (e.g.
-the nearest-neighbour or a random selection), and use its known redshift as the photo-$z$.
-This color-matched subset is identified by first calculating the Mahalanobis distance $D_M$ in color-space between the test galaxy and all training-set galaxies: the difference between the test and a training set galaxy's color divided by the photometric error, summed over all colors (i.e., $u$-$g$, $g$-$r$, $r$-$i$, $i$-$z$, and $z$-$y$).
-Then, the threshold value for $D_M$ that define a good color match is set by the percent point function (PPF): for example, for $N_{\rm dof}=5$, PPF$=95$ per cent of all training galaxies consistent with the test galaxy will have $D_M < 11.07$ (where $N_{\rm dof}$, the number of degrees of freedom, is the number of colors).
-For a given test galaxy, the $p(z)$ is the normalized distribution of the true catalogue redshifts of this color-matched subset of training galaxies, and the standard deviation of the color-matched subset is used as the photo-$z$ uncertainty.
-
-We have applied the nearest-neighbours color-matching photometric redshift estimator described in G18 to the simulated data.
-Compared to its application in G18, there are some minor differences in the application of this estimator to the Buzzard catalogue.
-First, we do not impose non-detections on galaxies with a magnitude fainter than the expected LSST 10-year limiting magnitude or bright enough to saturate with LSST: {\it all} of the photometry for all the galaxies in the test and training sets are used for this experiment.
-Second, as in G18 we do apply an initial cut in color to the training set before calculating the Mahalanobis distance in order to accelerate processing, and also use a magnitude pseudo-prior to improve photo-$z$ estimates, but for both we have used different cut-off values that are appropriate for the Buzzard galaxies' colors and magnitudes.
-Third, we set different parameters for the identification of the color-matched subset of training galaxies and the selection of a photometric redshift estimate.
-In G18 we used a percent point function (PPF) value of 0.68 to identify the color-matched subset of training galaxies and used the redshift of nearest neighbour in color-space as the photo-$z$ estimate.
-These choices work well when the desire is to obtain accurate photo-$z$ estimates for most test-set galaxies, but does not return a robust $p(z)$ in all cases -- especially for galaxies that are bright and/or have few matches in color-space.
-Since a robust estimate of $p(z)$ is desired for this work we make several changes to our implementation of the \textsc{CMNN} photo-$z$ estimator.
-We continue to use a percent point function of ${\rm PPF} = 0.95$ to generate the subset of color-matched training galaxies, but weight them by the inverse of their Mahalanobis distance.
-This weighting maintains some of the accuracy that was previously achieved by simply using the nearest neighbour in color-space.
-We then use the weights to create the $p(z)$ instead of having the redshift of each color-matched training-set galaxy count equally.
-To obtain a robust estimate of the $p(z)$ for galaxies with a small number of color-matched training set galaxies, when this number is less than $20$ the nearest $20$ neighbours in color-space are used instead, and we convolve the $p(z)$ with a Gaussian with a standard deviation of:
-
-\begin{equation}
-\sigma = \sigma_{\rm train} \sqrt{({\rm PPF}_{20}/0.95)^2 -1}
-\end{equation}
-
-\noindent to appropriately broaden it so that the $p(z)$ for these test galaxies represents the enlarged PPF value associated with it.
-Overall, these three changes will yield poorer accuracy photo-$z$ compared to those presented in G18, but they will all have significantly more robust estimates of the $p(z)$, particularly for the brightest test galaxies.
-This is sufficient for this work because, as described in G18, the goal of the \textsc{CMNN} photo-$z$ estimator was never to provide the ``best'' (or even competitive) estimates in the first place, given its reliance on a deep training set, but rather to provide a means for direct comparisons between LSST photometric quality and photo-$z$ estimates.
-With this work we show how the input parameters should be set in order to return robust $p(z)$ estimates in addition to point value estimates.
-
-\subsubsection{Delight}
-\label{sec:delight}
-%(John Soo)
-
-\textsc{Delight}\footnote{\url{https://github.com/ixkael/Delight}} \citep{Leistedt:17} infers photo-$z$'s by using a data-driven model of latent SEDs and a physical model of photometric fluxes as a function of redshift.
-Generally, machine learning methods rely on representative training data with similar band passes, while template based methods rely on a complete library of templates based on physical models constructed.
-\textsc{Delight} is constructed in attempt to combine the advantages and eliminate the disadvantages of both template-based and machine learning algorithms: it constructs a large collection of latent SED templates (or physical flux-redshift models) from training data, with a template SED library as a guide to the learning of the model.
-The advantage of \textsc{Delight} is that it neither needs representative training data in the same photometric bands, nor does it need detailed galaxy SED models to work.
-
-This conceptually novel approach is done by using Gaussian processes operating in flux-redshift space.
-The posterior distribution on the redshift of a target galaxy is obtained via a pairwise comparison with training galaxies,
-
-\begin{equation}
-p(z|\mathbf{\hat{F}}) \approx \sum_i p(\mathbf{\hat{F}}|z,t_i)\, p(z|t_i)p(t_i),
-\end{equation}
-\noindent where $p(z|t_i)p(t_i)$ captures prior information about the redshift distributions and abundances of the galaxies, with $t_i$ denoting the galaxy template; while $p(\mathbf{\hat{F}}|z,t_i)$ is the posterior of noisy flux $\mathbf{\hat{F}}$ at redshift $z$.
-For each training-target pair, $p(\mathbf{\hat{F}}|z,t_i)$ is evaluated as follows:
-
-\begin{equation} \label{eq:delight_noisy}
-p(\mathbf{\hat{F}}|z,t_i) = \int p(\mathbf{\hat{F}}|\mathbf{F})\, p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)\, d\mathbf{F},
-\end{equation}
-where $p(\mathbf{\hat{F}}|\mathbf{F})$ is the likelihood function, it compares the noisy real flux $\mathbf{\hat{F}}$ with the noiseless flux $\mathbf{F}$ obtained from the linear combination of template models, carefully constructed to account for model uncertainties and different normalization of the same SED; while $p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)$ is the prediction of flux at a different redshift $z$ with respect to the training object with redshift $z_i$ and flux $\mathbf{\hat{F}}_i$.
-Eq.~\ref{eq:delight_noisy} is essentially the probability that the training and the target galaxies having the same SED but at a different redshift.
-The flux prediction $p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)$ of the training galaxy at redshift $z$ is modeled via a Gaussian process,
-
-\begin{equation} \label{eq:delight_gp}
-F_b \sim \mathcal{GP}\left( \mu^F,k^F \right),
-\end{equation}
-\noindent with mean function $\mu^F$ and kernel $k^F$, both imposed to capture expected correlations resulting from the known underlying physics (i.e., fluxes resulting from observing SEDs through filter response, and the SEDs being redshifted).
-The reader should refer to \citet{Leistedt:17} for further details.
-
-In this study, all $100$ ordered Buzzard templates, as described in Section~\ref{sec:buzztemplates}, were used in \textsc{Delight}, and the Gaussian process was trained with a subset of $50\,000$ galaxies.
-Photometric uncertainties from the inputs are propagated into the code, while non-detections for each band are set to the mean of the respective bands.
-Default settings of \textsc{Delight} were use, with the exception that the PDF bins were set to be linear instead of logarithmic, with $200$ equally-spaced bins between $0.0<z<2.0$.
-In this study a flat prior is assumed.
-
-\subsubsection{FlexZBoost}
-\label{sec:flexzboost}
-%(Ann Lee, Rafael Izbicki, Taylor Pospisil, Peter Freeman)
-
-\textsc{FlexZBoost}\footnote{\url{https://github.com/tpospisi/flexcode};\\ \url{https://github.com/rizbicki/FlexCoDE}} \citep{Izbicki:17} is a particular realization of \texttt{FlexCode}, which is a general-purpose methodology for converting any conditional mean point estimator of $z$ to a conditional {\em density} estimator $f(z \vert \x)$, where $\x$ here represents our photometric covariates and errors.\footnote{Instead of $p(z)$, we use the notation $f(z \vert \x)$ to explicitly show the dependence on $\x$.} The key idea is to expand the unknown function $f(z \vert \x)$ in an orthonormal basis $\{\phi_i(z)\}_{i}$:
-
-\begin{equation}
-f(z|\x)=\sum_{i}\beta_{i }(\x)\phi_i(z). \label{eq::series_expansion}
-\end{equation}
-By the orthogonality property, the expansion coefficients are just conditional means
-
-\begin{equation}
-\beta_{i }(\x) =  \mathbb{E}\left[\phi_i(z)|\x\right] \equiv \int f(z|\x)   \phi_i(z) dz. \label{eq:FlexZBoost_coeffs}
-\end{equation}
-These coefficients can easily be estimated from data by regression.
-
-In this paper, we use \textsc{xgboost}~\citep{Chen:16} for the regression part as these techniques scale well for massive data; it should however be noted that \textsc{FlexCode-RF} (also on GitHub), based on Random Forests, generally performs better for smaller data sets.
-As our basis, we choose a standard Fourier basis.
-There are two tuning parameters in our $p(z)$ estimate: (i) the number of terms, $I$, in the series expansion in Eq.~\ref{eq::series_expansion}, and (ii) an exponent $\alpha$ that we use to sharpen the computed density estimates $\widehat{f}(z|\x)$, according to $\widetilde{f}(z|\x) \propto \widehat{f}(z|\x)^\alpha$.
-We split the ``train data'' into a training set (85\%) and a validation set (15\%), and choose both $I$ and $\alpha$ in an automated way by minimizing the weighted $L_2$-loss function \citep[Eq.
-5 in][]{Izbicki:17} on the  validation set.
-
-Although {\tt FlexCode} offers a {\em lossless compression} of the photo-$z$ estimates (in this study, one can reconstruct $\widetilde{f}(z|\x)$ exactly at any resolution from estimates of the first 35 coefficients, Eq.~\ref{eq:FlexZBoost_coeffs}, for a Fourier basis $\{\phi_i(z)\}_{i}$), we discretize our final estimates into 200 bins linearly spaced in $0 < z < 2$ for easy comparison with other algorithms.
-Using a higher resolution may yield better results (with no added cost in storage).
-
-%\subsubsection{Frankenz}
-%\label{sec:frankenz}
-%(seeking volunteers)
-%
-%\red{\textsc{Frankenz}\footnote{\url{https://github.com/joshspeagle/frankenz}} \cite{Speagle:frankenz} is...}
-%
-
-\subsubsection{GPz}
-\label{sec:gpz}
-%(Ibrahim Almosallam)
-
-\textsc{GPz}\footnote{\url{https://github.com/OxfordML/GPz}} \citep{Almosallam:16a,Almosallam:15b} is a sparse Gaussian process based code, a fast and a scalable approximation of full Gaussian Processes \citep{Rasmussen:06}, with the added feature of being able to produce input-dependent variance estimations (heteroscedastic noise).
-The model assumes that the probability of the output $y$, the redshift, given the input $x$, the photometry, is $p(y|x)=\mathcal{N}\left(y|\mu(x),\sigma(x)^{2}\right)$.
-The mean function, $\mu(x)$, and the variance function $\sigma(x)^{2}$ are both linear combinations of basis functions that take the following form:
-
-\begin{align}
-f(x)=\sum_{i=1}^{m}\phi_{i}(x)w_{i},
-\end{align}
-where $\left\{\phi_{i}(x)\right\}_{i=1}^{m}$ and $\left\{w_{i}\right\}_{i=1}^{m}$ are sets of $m$ basis functions and their associated weights respectively.
-Basis function models (BFM), for specific classes of basis functions such as the sigmoid or the squared exponential, have the advantage of being universal approximators, i.e.
-there exist a function of that form that can approximate any function, with mild assumptions, to any desired degree of accuracy.
-The details on how to learn the parameters of the model and the hyper-parameters of the basis functions are described in \citet{Almosallam:15b}.
-
-A unique feature in \textsc{GPz}, is that the variance estimate is composed of two terms each quantifying a different source of uncertainty.
-One term (the model uncertainty) reflects how much of the uncertainty is due to lack of training samples at the location of interest, whereas the second term (the noise uncertainty) reflects how much of the uncertainty is caused from observing many noisy samples at that location.
-Thus, the predictive variance can determine whether we need more representative samples or more precise samples for any particular location in the input space.
-\textsc{GPz} can also emphasize the importance of some samples as weights.
-This weight can be for example $|z_{\rm{spec}}-z_{\rm{phot}}|/(1+z_{\rm{spec}})$ to target the desired objective of minimizing the normalized redshift error or as a function of their probability in the test set relative to the training set in order to pressure the model to better fit samples that are rare in the training set but are expected to be abundant during testing.
-
-The data is prepared for \textsc{GPz} by taking the log of the magnitude errors, decorrelating the data set using PCA and imputing the missing values using a simple linear model that estimates the missing variables given the observed ones.
-The log transformation helps to smooth the long tail distribution of the magnitude errors, which is more stable numerically and makes the optimization process unconstrained.
-The missing values are imputed by computing the mean of the training set $\mu$ and its covariance $\Sigma$, then we use the following equation to estimate the missing values from the observed ones %\[
-
-\begin{equation}
-x_{u} = \mu_{u}+\Sigma_{uo}\Sigma_{oo}^{-1}(x_{o}-\mu_{o}),
-\end{equation}  %\]
-where the subscript $o$ in $x_{o}$ indexes the \emph{observed} part of the input $x$, whereas the subscript $u$ indexes the \emph{unobserved} set (similarly for $\mu$ and $\Sigma$).
-This is the optimal expected value of the unobserved variables given the observed ones if the distribution is jointly Gaussian, note that if the variables are independent, i.e.
-$\Sigma_{uo}=0$, this will reduce to a simple average predictor.
-
-We use the Variable Covariance (VC) option in \textsc{GPz} with 200 basis functions after we note that there is no significant increase in the performance on the validation set (using 80\%-20\% training-validation split) and with no cost-sensitive learning applied.
-We do note that \textsc{GPz} is trained using a data set that includes galaxies fainter than those in the test set, which was truncated at $i<25.3$.
-This mismatch in training vs.~test results in a slightly different distribution in galaxy photometry and redshift distribution between the training and test set, which most likely degraded results compared to a run where the training set more closely matched the test set.
-
-\subsubsection{METAPhoR}
-\label{sec:metaphor}
-%(Stefano Cavuoti, Massimo Brescia, Giuseppe Longo)
-%\textcolor{brown}{general review done on this section}
-\textsc{METAPhoR} \citep[Machine-learning Estimation Tool for Accurate Photometric Redshifts,][]{Cavuoti:17} is a pipeline designed to provide photo-$z$'s point estimates and a reliable PDF for machine learning (ML) based techniques.
-It includes pre- and post-processing phases, hosting a photo-$z$ prediction engine based on the Multi Layer Perceptron with Quasi Newton Algorithm (MLPQNA), already validated on photo-$z$'s in several cases \citep{de_Jong:17,Cavuoti:17b,Cavuoti:15,Brescia:14,Brescia:13,Biviano:13}.
-Due to its plug-in based modular nature, \textsc{METAPhoR} can be easily replaced by any other photo-$z$ prediction kernel, regardless its implementation, by taking the I/O interface compliance as unique constrain.
-
-At a higher level, the pipeline mainly consists of three modules: (i) \textit{data pre-processing}, including a catalogue cross-matching sub-module \citep[based on the tool C3, ][]{Riccio:17}, a sub-module for photometric evaluation and error estimation of the multi-band catalogue used as Knowledge Base (KB), and a sub-module dedicated to the perturbation of the photometric KB, propaedeutic to the PDF estimation; (ii) \textit{photo-$z$ prediction}, which is the training/validation/test phase, producing the photo-$z$'s point estimates, based on a pre-selected ML method; (iii) \textit{PDF estimation}, specifically designed to calculate the PDF of the photo-$z$ estimation errors.
-The last module includes also a post-processing tool, providing some statistics on the produced point estimates and PDFs.
-
-The photometry perturbation law is based on the formula $m_{ij} = m_{ij} + \alpha_{i}F_{ij}*u_{\mu=0,\sigma=1}$, where $\alpha_{i}$ is a user selected multiplicative constant (useful in case of multi-survey photometry), $u_{\mu=0,\sigma=1}$ is a random value from the standard normal distribution and $F_{ij}$ is a bimodal function (a constant function + polynomial fitting of the mean magnitude errors on the binned bands), heuristically tuned in such a way that the constant component is the threshold under which the polynomial function is considered too low to provide a significant noise contribution to the photometry perturbation.
-
-As introduced, the photo-$z$ point estimate prediction engine of \textsc{METAPhoR} is based on the MLPQNA model, whose photo-$z$ regression training error, used by the quasi Newton learning rule, is based on the least square error and Tikhonov $L_{2}$-norm regularization \citep{Hofmann:18}.
-
-As main prerogative, \textsc{METAPhoR} is able to provide a PDF for ML methods by taking into account the photometric errors provided with data, by running $N$ trainings on the same training set, or $M$ trainings on $M$ different random extractions from the KB.
-The different test sets, used to produce the PDF, are thus obtained by introducing a proper perturbation, parametrized from the photometric error distribution in each band, on the photometric data populating the original test set \citep{Brescia:18}.
-
-For the present work since it was required to produce a redshift (and a PDF) for each object of the test set we decided to apply a hierarchical kNN to fill the missing detection, it goes without saying that for such points the reliability of PDFs and point estimation is lower.
-No cross validation has been used.
-
-\subsubsection{SkyNet}
-\label{sec:skynet}
-%(J. Cohen-Tanugi and Hugo Tranin)
-
-\textsc{SkyNet}\footnote{\url{http://ccpforge.cse.rl.ac.uk/gf/project/skynet/}} \citep{Graff:14} is a publicly available neural network software, based on a 2nd order conjugate gradient optimization scheme \citep[see][for further details]{Graff:14}.
-It has been used efficiently for redshift PDF estimates \citep{Sanchez:14,Bonnett:15,Bonnett:16}.
-
-%In the present work, we use \textsc{SkyNet} both as a regressor (for redshift point estimates) and a classifier (for PDF estimates). 
-The neural network is configured as a standard multilayer perceptron with three hidden layers and one input layer with $12$ nodes (the $6$ magnitudes and their errors). 
-%For the regressor, the output layer is a single linear node, while each hidden layer has $10$ nodes with a tanh activation function. 
-The classifier is laid out such that the hidden layers have 20:40:40 nodes each, all rectified linear units, and the output layer has $200$ nodes (corresponding to 200 bins for the PDF) activated with a ``softmax'' function so that they automatically sum to $1$.
-
-To avoid over-fitting, a $30$ per cent fraction of the training set is used as validation, and the training is stopped as soon as the error rate begins to increase in the validation set.
-The weights are randomly initialized based on normal sampling.
-The error function is a standard chi-square function for the regressor, and a cross-entropy function for the classifier.
-Finally, the data are all whitened before processing, with magnitudes pegged to (45,45,40,35,42,42) and their errors pegged to (20,20,10,5,15,15) for $ugrizy$ filters, respectively.
-
-\subsubsection{TPZ}
-\label{sec:tpz}
-%(Erfan Nourbakhsh)
-
-\textsc{TPZ}\footnote{\url{https://github.com/mgckind/MLZ}} \citep[Trees for Photo-$z$,][]{Carrasco_Kind:13,Carrascokind:14} is a parallel machine learning algorithm that generates photometric redshift PDFs using prediction trees and random forest techniques.
-The code recursively splits the input data (i.~e.~the training sample), into two branches, one after another, until a terminal leaf is created that meets a termination criterion (e.~g.~a minimum leaf size or a variance threshold).
-Bootstrap samples from the training data and associated errors are used to build a set of prediction trees.
-In order to minimize correlation between the trees, the data is divided in such a way that the highest information gain among the random subsample of features is obtained at every point.
-The regions in each terminal leaf node corresponds to a specific subsample of the entire data that possesses similar properties.
-
-The training data is examined before running TPZ.
-Since TPZ does not handle non-detections (magnitudes flagged as 99.0), we replace these values with an approximation of the $1\sigma$ detection threshold, i.~e.~a signal to noise ratio of 1 in terms of magnitude uncertainty using the equation $dm = 2.5 ~ \log ( 1 + N/S )$ where $dm \sim 0.7526 ~ mag$ for $N/S=1$.
-That is, for each band, we replace the non-detection with the magnitude corresponding to the error of 0.7526 from the error model forecasted for 10-year LSST data.
-The Out-of-Bag \citep{Breiman:84,Carrasco_Kind:13} cross-validation technique is used within TPZ to evaluate its predictive validity and determine the relative importance of the different input attributes.
-We employed this information to calibrate our algorithm.
-
-In the present work, the LSST magnitudes $u,~g,~r,~i$ and colors $u-g,~g-r,~r-i,~i-z,~z-y$ and their associated errors are used in the process of growing 100 trees with a minimum leaf size of 5 (the $z$ and $y$ magnitudes did not show significant correlation with the redshift in our cross-validation, so we did not use them when constructing our trees).
-We partitioned our redshift space into 100 bins from $z=0.005$ to $z=2.0$ and smoothed each individual PDF with a smoothing scale of twice the bin size.
-
-\subsection{Simple Ensemble Estimator}\label{sec:method:trainz}
-In addition to the main photo-$z$ algorithms described above we also include a very simple method.  For \trainz, as we will we call this simple estimator, we well define $p(z)$ as simply:
-
-\begin{equation}
-p(z) = \frac{1}{N_{train}}\sum_{i=1}^{N_{train}}z_{train}
-\end{equation}
-That is, we simply set the redshift PDF of every galaxy equal to the normalized $N(z)$ of the training sample.
-As the training sample is drawn from the same underlying distribution as the test sample, modulo small deviations due to sample size, the quantiles of the training and test distributions should be identical.
-This is a wildly unrealistic estimator, as it assigns all galaxies, no matter their apparent magnitude, colour, or true redshift, the same redshift PDF, and is thus uninformative at the level of individual object redshifts, but is designed to perform very well for the ensemble of all objects.
-We will discuss this method and cautions relative to metrics in Section~\ref{sec:caution}.
-
-\section{Metrics for quantifying PDF comparisons}
-\label{sec:metrics}
-%(Alex Malz, Rongpu Zhou, Jeff Newman, Ofer Lahav)
-
-%\red{Maybe it would be best to introduce the summary statistics (KS, CvM, AD) up front and then the ensemble functions and stacked n(z).}
-
-The overloaded ``$p(z)$'' is a widespread abuse of notation; we would like the outputs of photo-$z$ PDF codes to be interpretable as probabilities. 
-Obviously photo-$z$ PDFs must not take negative values and must integrate to unity over the range of possible redshifts. 
-Additionally, an estimator derived by method $H$ for the photo-$z$ PDF of galaxy $i$ must be understood as a posterior probability distribution
-
-\begin{equation}
-\hat{p}_{j}(z_{i}) = p(z | d_{i}, I_{D}, I_{H}),
-\end{equation}
-conditioned not only on the photometric data $d_{i}$ for that galaxy but also on parameters encompassing a number of things that will differ depending on the method $H$ used to produce it, namely the assumptions $I_{H}$ necessary for the method to be valid and any inputs $I_{D}$ it takes as prior information, such as a template library or training set.
-Because of this, direct comparison of photo-$z$ PDFs produced by different methods is in some sense impossible; even if they share the same prior information $I_{D}$, by definition they cannot be conditioned on the same assumptions $I_{H}$, otherwise they would not be distinct methods at all.
-
-In this study, we isolate the differences in prior information specific to each method by using a single training set $I_{D}^{ML}$ for all machine learning-based codes and a single template library $I_{D}^{T}$ for all template-based codes, and these sets of prior information are carefully constructed to be representative and complete, we have $I_{D}^{ML}\equiv I_{D}^{T}$ for every method $H$.
-Thus, we are saying
-
-\begin{equation}
-\frac{\hat{p}_{i, H}(z)}{\hat{p}_{i, H'}(z)} \approx \frac{p(z | d_{i}, I_{H})}{p(z | d_{i}, I_{H'})},
-\end{equation}
-meaning that we assume comparisons of $\hat{p}_{i, H}(z)$ isolate the effect of the method used to obtain the estimator, which should make examination of differences caused by specifics of the method implementations easier to isolate.
-
-As mentioned previously, there are cosmology probes that require knowledge of individual galaxy $p(z)$ and others that require only knowledge of the ensemble redshift distribution, $N(z)$.  
-Due to the paucity of principled techniques for using and validating photo-$z$ PDFs, there are few alternatives to the common practice of reducing photo-$z$ PDFs to point estimates. 
-Though this practice should not be encouraged, we also calculate traditional metrics based on the most common point estimators derived from photo-$z$ PDFs.  
-Those seeking to establish a connection to traditional ways of thinking about redshift estimation may consult the Appendix for these results.
-
-There are a number of metrics that can be used to test the accuracy of a photo-$z$ interim posterior as an estimator of a true photo-$z$ posterior if it is known.
-Even for simulated data, the true photo-$z$ PDF is in general not accessible unless the redshifts are in fact drawn from the true photo-$z$ PDFs, a mock catalogue generation procedure that has not yet appeared in the literature.
-Furthermore, only limited applications of photo-$z$ PDFs that could be used as the basis for a metric have been presented in the literature. 
-The most popular application by far is the calculation of the overall redshift distribution $N(z)$, the true value of which is known for the \textsc{Buzzard} simulation and will be denoted as $N'(z)$.
-Though alternatives exist \citep{Malz:chippr}, stacking according to
-
-\begin{equation} \label{eq:stacked}
-\hat{N}^{H}(z) \approx \frac{1}{N_{tot}}\ \sum_{i}^{N_{tot}}\ \hat{p}^{H}_{i}(z)
-\end{equation}
-is the most widely accepted method for estimating the redshift distribution from photo-$z$ PDFs. 
-If we assume that the response of estimators of $N(z)$ is uniform across all approaches $H$, then we may interpret metrics on the accuracy of $\hat{N}(z)$ obtained in this way. 
-We must note, however, that this is a poor assumption in general. 
-Under the setup of this paper, the true redshift distribution $N'(z) = p(z | I_{D})$ (i.e.
-because our training data is representative, the interim prior is the truth).
-In this ideal case, the method that would give the best approximation to $N'(z)$ would be one that neglects all the information contained in the photometry $\{d_{i}\}_{N_{tot}}$ and gives every galaxy the same photo-$z$ PDF $\hat{p}_{i}(z) = N'(z)$ for all $i$.
-This is the exact estimator, \trainz, that we have described in Section~\ref{sec:method:trainz}, and which will serve as a point of reference for the other codes.
-
-The exact implementation of the stacked estimator $\hat{N}^{H}(z)$ will depend on the parametrization of the photo-$z$ PDFs, which may differ across codes and can affect the precision of the estimator \citep{Malz:qp}; even considering a single method under the same parametrization, say a piecewise constant function over bins or a set of samples from the posterior, an estimator using $2N$ bins or samples will trivially be more precise than an estimator using $N$ bins or samples.
-%\red{Note: Let's address how the gridding (endpoints and spacing) affects the metrics, from the email correspondence Re: [LSST-DESC-MEMBERS] Summary of Nov 29 PZ Telecon. (AIM)}
-In order to minimize the effects of such choices, we asked those running all eleven codes to output $p(z)$ parameterized with a generous $\approx 200$ piecewise constant bins spanning $0<z<2$.  
-The piecewise constant format is chosen because of its established presence in the literature, and the choice of 200 bins was motivated by the approximate number of columns expected to be available for storage of $p(z)$ for the final LSST Project tables.\footnote{See, e.~g.~the LSST Data Products Definition Document, available at: \url{https://ls.st/dpdd}}
-All $p(z)$ catalogues are processed using the \textsc{qp} software package \citep{Malz:qp}\footnote{available at: \url{http://github.com/aimalz/qp/}} for manipulating and calculating metrics of 1-dimensional PDFs.
-We will discuss the choice of $p(z)$ parameterization further in Section~\ref{sec:results}.
-
-%\red{Divide metrics by application to ensemble of PDFs vs. estimators of the redshift distribution}
-
-\subsection{Metrics of an ensemble of photo-\mathinhead{$z$}{z} interim posteriors}
-\label{sec:qualmet}
-
-\subsubsection{Probability integral transform (PIT)}
-\label{sec:pit}
-%(Alex Malz, Peter Freeman, Sam Schmidt)
-
-The probability integral transform (PIT) \citep{Polsterer:16} is defined for each individual galaxy as:%essentially the derivative of the QQ plot, a histogram of
-
-\begin{align}
-\label{eq:pit}
-\mathrm{PIT} &= \int_{-\infty}^{z_{\rm true}}p(z)\, dz \,.
-\end{align}
-The distribution of PIT values quantifies the behavior of the {\it ensemble} of photo-$z$ PDFs, enabling us to evaluate whether the $p(z)$ is, on average, accurate:  The PIT value is the Cumulative Distribution Function (CDF) of the $p(z)$ evaluated at the true redshift.
-A catalogue of photo-$z$ PDFs that are accurate should have a flat PIT histogram (i.e., the individual PIT values as samples from each CDF should match a Uniform(0,1) distribution if the CDFs are accurate).
-Specific deviations from flatness indicate inaccuracy: overly broad photo-$z$ PDFs would manifest as underrepresentation of the lowest and highest PIT values, whereas overly narrow photo-$z$ PDFs would manifest as over-representation of the lowest and highest PIT values.
-High frequency at only $\mathrm{PIT}\approx0$ and $\mathrm{PIT}\approx1$ indicates the presence of catastrophic outliers with highly inaccurate photo-$z$ PDFs where the true redshift is outside of the support of $p(z)$.
-\citet{Tanaka:17} use the histogram of PIT values as a diagnostic indicator of overall code performance, while \citet{Freeman:17} independently define the PIT and demonstrate how its individual values may be used both to perform hypothesis testing (via, e.g., the KS, CvM, and AD tests; see below) and to construct quantile-quantile plots.
-%\red{We might want to divide this by redshift to get trends, as the overall PIT could be impacted by true n(z).  
-%For example, a PIT with nontrivial shape could be due to that behavior being uniform as a function of redshift or significant inaccuracy only in a small range of redshifts where $n(z)$ is large and a flat PIT elsewhere.}
-
-\subsubsection{Quantile-quantile (QQ) plot}
-\label{sec:qq}
-%(Rongpu Zhou, Jeffrey Newman, Alex Malz)
-
-The quantile-quantile (QQ) plot is a graphical method for comparing two distributions, where the quantiles of one distribution are plotted against the quantiles of the other distribution (A quantile being defined by partitioning a distribution into consecutive intervals containing equal amounts of probability, or equal numbers of objects in each interval).
-%\blue{Include an equation to ensure quantile definition is crystal clear before building upon it for PIT.}\red{Is maybe the above text clear enough?
-Text may be more clear than any equation I can come up with--SJS.}
-In this paper we show the quantiles of the PIT values compared to the quantiles of the Uniform distribution that we expect the PIT values to match if $p(z)$ is an accurate probability distribution for all objects.
-The QQ plot provides an easy way to qualitatively assess the differences in various properties such as the moments of an estimating distribution relative to a true distribution.  
-%location, scale and skewness.  
-In this paper, QQ plots are used for two purposes: (1) for comparing $N(z)$ from photo-$z$ PDFs (estimated using Eq.~\ref{eq:stacked}) with the true $N(z)$, i.e.
-comparing the estimated distribution of redshifts with the true redshift distribution, and (2) for assessing the overall consistency of an ensemble of photo-$z$ PDFs with their true redshifts on a population level, where the distribution of the PIT values (see previous section) is compared to a uniform distribution between $0$ and $1$.
-The QQ plot contains very similar information to that shown in the PIT histogram plot, we include both forms, as visually they each convey the information in a somewhat distinct manner.
-%\red{is this enough reason to include both?--SJS}
-%\aim{Maybe we should explain why we expect the distribution of CDFs to be uniform.}\red{Note: this is just describing the PIT.
-I'm going to swap the order of PIT to come before this, and then say we are evaluating the distribution of PIT values.
-I will also add why we expect this to be uniform--SJS}
-
-\subsubsection{Conditional density estimation loss}
-\label{sec:CDE_loss}
-%(Alex Malz, Ann Lee)
-
-With the conditional density estimation loss (CDE loss) we can compare how well different methods estimate individual PDFs for photometric covariates $\x$ rather than looking only at the ensemble distribution.
-As in Section~\ref{sec:flexzboost}, we use the notation $f(z \vert \x)$ instead of $p(z)$ to explicitly show the dependence on $\x$.
-
-%\aim{Perhaps this should be rewritten in the same notation used elsewhere in the paper.} \red{I agree, does someone who speaks both stats and astro want to take a shot at this?}
-%\textcolor{cyan}{Ann: We are looking at individual PDFs so the dependence on $\x$ needs to be explicit; this is probably the most accurate notation as it shows both the conditioning clearly as well as whether $\x$ is fixed or random.}
-
-The CDE loss is defined as:
-\begin{equation} \label{eq:cde-loss}
-L(f, \widehat{f}) = \int \int (f(z \mid \x) - \widehat{f}(z \mid \x))^{2} dz dP(\x)
-\end{equation}
-This loss is the CDE equivalent of the RMSE in regression. To estimate this loss we rewrite the loss as
-\begin{equation} \label{eq:estimated-cde-loss}
-\mathbb{E}_\X \left[\int \widehat{f}(z \mid \X)^{2} dz\right] - 2 \mathbb{E}_{\X,Z}\left[\widehat{f}(Z \mid \X)\right] + K_{f},
-\end{equation}
-where the first expectation is with respect to the marginal distribution of the covariates $\X$, the second expectation is with respect to the joint distribution of $\X$ and $Z$, and $K_{f}$ is a constant depending only upon the true conditional densities $f(z\mid \x)$.
-For each method we can estimate these expectations as empirical expectations on the test or validation data \citep[Eq.~7 in][]{Izbicki:17b} without knowledge of the true densities.
-
-%\subsubsection{Continuous Ranked Probability Score (CRPS)}
-%\label{sec:crps}
-%\red{The CRPS is used to assess the accuracy of two probabilistic models respectively.}
-
-\subsection{Metrics over estimated probability distributions}
-\label{sec:quantmet}
-
-In tandem with the QQ and PIT metrics introduced above, we additionally compute the following metrics comparing the empirical CDF of a distribution to the true or expected distribution.
-These metrics give a more quantitative measure of the departure from ideal than the more visual PIT histogram and QQ plot.
-We compute metrics comparing the CDF of PIT values to a the CDF of a Uniform distribution, and also compute the CDF of the true redshift distribution $N'(z)$ compared the $\hat{N}(z)$ distribution derived from summing the $p(z)$ as described in Eq.~\ref{eq:stacked}.
-
-\subsubsection{Root-mean-square error (RMSE)}
-\label{sec:rmse}
-%(Alex Malz)
-
-We employ the familiar root-mean-square error:
-
-\begin{equation}
-\mathrm{RMSE} = \sqrt{\int_{-\infty}^{\infty}\left(\hat{f}(z)-f'(z)\right)^{2}\,dz},
-\end{equation}
-%may be used for two purposes in this analysis. 
-%We calculate it between the true redshift distribution $N'(z)$ and each estimator $\hat{N}(z)$ (derived by Eq.~\ref{eq:stacked}) as well as between the ideal QQ and PIT curves from each method studied and the ideal curves.
-Though this metric does not account for the fact that the redshift distribution function is, in fact, a probability distribution, it can still be interpreted as a measure of the integrated difference between the estimated distribution and the true distribution, and it can be used to quantify the otherwise qualitative metrics.
-
-\subsubsection{Kolmogorov-Smirnov (KS) and related statistics}
-\label{sec:ks}
-%(John Soo, Rongpu Zhou)
-
-The \textit{Kolmogorov-Smirnov statistic} $N_{\rm KS}$ is the maximum difference between $F_{\rm phot}(z)$ and $F_{\rm spec}(z)$, the CDFs of the photo-$z$ and spectroscopic redshift respectively:
-
-\begin{equation}
-N_{\rm KS} = \max_{z}\left( \left| F_{\rm phot}(z) - F_{\rm spec}(z) \right| \right).
-\end{equation}
-%The KS test quantifies whether the two redshift distributions are drawn from a similar distribution, independent of binning. 
-The KS test quantifies the similarity between two distributions, independent of binning. 
-A lower $N_{\rm KS}$ value corresponds to more similar distributions.
-%We calculate two KS test statistics, one to test the distribution of $p(z)$ values, and one to test the stacked estimator $\hat{n}(z)$ of the redshift distribution.  
-%The $p(z)$ test compares the CDF of the PIT values described in the previous section to that of a uniform distribution.  
-%The $n(z)$ test compares the CDF of the distribution formed by stacking the $p(z)$ for every object in our test set against the CDF of the corresponding true redshifts.
-%In this case a lower $N_{\rm KS}$ value means that the photo-$z$ distribution fits the spectroscopic redshift distribution well.
-$F_{\rm phot}(z)$ is calculated by stacking the PDFs of every single object in the sample, producing a smooth cumulative $n(z)$ to be compared with $F_{\rm spec}(z)$.
-
-We also consider two variants of the KS statistic: the Cramer-von Mises (CvM) and Anderson-Darling (AD) statistics.
-The CvM statistic is similar to the KS statistic as it is also computed from the distance between the measured CDF and the ideal CDF, but instead of the maximum distance, the CvM statistic calculates the average of the distance squared:
-
-\begin{equation}
-\label{eq:cvm}
-\omega^2 = \int_{-\infty}^{+\infty}\big(F_{\mathrm{meas.}}(x) - F_{\mathrm{ideal}}(x)\big)^2\mathrm{d}F_{\mathrm{ideal}}
-\end{equation}
-% In this paper we use the CvM statistic to evaluate the performance of both $n(z)$ and $p(z)$.
-The AD statistic is a weighted version of the CvM statistic, making it more sensitive to the tails of the distribution:
-
-\begin{equation} \label{eq:ad}
-A^2 = n\int_{-\infty}^{+\infty} \frac{\big(F_{\mathrm{meas.}}(x) - F_{\mathrm{ideal}}(x)\big)^2} {F_{\mathrm{ideal}}(x) (1-F_{\mathrm{ideal}}(x))}\mathrm{d}F_{\mathrm{ideal}}
-\end{equation}
-where $n$ is the sample size. 
-%The KS statistic and the CvM and AD statistics related to it are all used to evaluate the performance of both $\hat{n}(z)$ and the QQ and PIT of the CDFs of the ensemble of photo-$z$ interim posteriors.
-
-%\subsubsection{Kullback-Leibler Divergence (KLD)}
-%\label{sec:kl}
-%(Alex Malz)
-%
-%The KLD is a measure of the loss in information when using a probability distribution that is an estimate of a known true probability distribution, so we use 
-%\begin{equation}
-%\mathrm{KLD} = \int_{-\infty}^{\infty}n'(z)\log\left[\frac{n'(z)}{\hat{n}(z)}\right]dz
-%\end{equation}
-%as a metric on the stacked redshift distribution estimator of Eq.~\ref{eq:stacked}.
-Noting the assumptions made in using Eq.~\ref{eq:stacked}, we say that a photo-$z$ PDFs method whose stacked redshift distribution estimator $\hat{n}(z)$ more closely resembles the true redshift distribution $n'(z)$ is desired, i.e.
-we seek to minimize the KLD.
-
-\subsubsection{Moments}
-\label{sec:moments}
-%(Alex Malz)
-For the $\hat{N}(z)$ distributions we additionaly calculate the first three moments of the estimated redshift distribution for each code and compare them to the moments of the true redshift distribution $N'(z)$.
-The $m$th moment of a distribution is defined as
-
-\begin{equation}
-\langle z^{m}\rangle = \int_{-\infty}^{\infty}z^{m}N(z)dz.
-\end{equation}
-Here, we use the moments of the stacked estimator of the redshift distribution function as the basis for a metric.
-The closer the moments of $\hat{N}(z)$ for a photo-$z$ PDF method are to the moments of the true redshift distribution function $N'(z)$, the better the photo-$z$ PDF method.
-
 \section{Results}
@@
 %\centering
-\begin{subfigure}[b]{0.49\textwidth}
-\includegraphics[width=\textwidth]{figures/pz_12codes_261931_crop.jpg}
-\end{subfigure}
-\begin{subfigure}[b]{0.49\textwidth}
-\includegraphics[width=\textwidth]{figures/pz_12codes_471167_crop.jpg}
-\end{subfigure}
-\begin{subfigure}[b]{0.49\textwidth}
-\includegraphics[width=\textwidth]
-{figures/pz_12codes_713178_crop.jpg}
-\end{subfigure}
-\begin{subfigure}[b]{0.49\textwidth}
-\includegraphics[width=\textwidth]{figures/pz_12codes_982747_crop.jpg}
-\end{subfigure}
+\includegraphics[width=0.49\textwidth]{fig/pz_12codes_261931_crop.jpg}\includegraphics[width=0.5\textwidth]{fig/pz_12codes_471167_crop.jpg}\\
+\includegraphics[width=\textwidth]{fig/pz_12codes_713178_crop.jpg}\includegraphics[width=\textwidth]{fig/pz_12codes_982747_crop.jpg}
 \caption{Four illustrative examples of individual p(z) distributions produced by the codes.
@@
 %\centering
-%\includegraphics[width=\textwidth]{figures/QQplot_11codes.jpg}
+%\includegraphics[width=\textwidth]{fig/QQplot_11codes.jpg}
 %\caption{The QQ plots produced for every photo-$z$ code.
@@
 \centering
-\includegraphics[width=\textwidth]{figures/PITANDQQplot_12codes_crop.jpg}
+\includegraphics[width=\textwidth]{fig/PITANDQQplot_12codes_crop.jpg}
 \caption{Summary plots for all eleven photo-$z$ codes illustrating performance for the interim posterior statistics.
@@
 %\centering
-%\includegraphics[width=\textwidth]{figures/PIT_histogram_plot_11codes.jpg}
+%\includegraphics[width=\textwidth]{fig/PIT_histogram_plot_11codes.jpg}
 %\caption{The PIT histogram plots produced for every photo-$z$ code.
@@
 \centering
-\includegraphics[width=\textwidth]{figures/KSvsCvMvsAD_PIT_withnull_jpg.jpg}
+\includegraphics[width=\textwidth]{fig/KSvsCvMvsAD_PIT_withnull_jpg.jpg}
 \caption{A visual representation of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the PIT distributions.
@@
 \centering
-\includegraphics[width=\textwidth]{figures/NZsumplot_12codes.jpg}
+\includegraphics[width=\textwidth]{fig/NZsumplot_12codes.jpg}
 \caption{The stacked $p(z)$ produced by each photo-$z$ code ($\hat{N}(z)$, red) compared to the spectroscopic redshift distribution ($N'(z)$, blue).
@@
 \centering
-\includegraphics[width=\textwidth]{figures/KSvsCvMvsAD_NZ_withnull_jpg.jpg}
+\includegraphics[width=\textwidth]{fig/KSvsCvMvsAD_NZ_withnull_jpg.jpg}
 \caption{A visual representation of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the $\hat{N}(z)$ distributions.
@@
 \end{table}
-
-\section{Summary and Discussion}
-\label{sec:discussion}
-%(Sam Schmidt, Ofer Lahav, Jeff Newman, Alex Malz, Eve Kovacs, Tony Tyson, Tina Peters)
-In this paper we presented results evaluating the photometric redshift PDF computation for eleven photo-$z$ codes.
-As discussed in Section~\ref{sec:metrics} the $p(z)$ should accurately reflect the relative likelihood as a function of redshift for each galaxy.
-All codes were provided a set of representative training data and tested on an idealized set of model galaxies with high signal-to-noise and photometry with no confounding effects due to blending, instrumental effects, the night sky, etc...
-included.
-The goal was not to determine a ``best'' photo-$z$ code: in many ways, this was a baseline test of a ``best case scenario'' to predict the expected photo-$z$ performance if a stage IV dark energy survey was to obtain complete training samples and perfectly calibrated their multi-band photometry.
-Given these idealized conditions, any deficiencies observed in a photo-$z$ code's performance should be a cause for concern, and may be evidence in a problem with either/both of the specific code implementation or the underlying algorithm.
-In order to meet the stringent LSST requirements on photo-$z$ performance, identifying and correcting such problems is an important first step before tackling more realistic data in future challenges.
-Most of the codes tested performed well, however, several did not meet the stringent goals that have been laid out for LSST photometric redshift performance.
-This is a cause for concern, given the idealized conditions, and the individual code responses will be studied in detail moving forward.
-One obvious trend in several of the codes tested was an overall over or underprediction of the widths of $p(z)$, as evidenced by the QQ plots and PIT histograms shown in Fig.~\ref{fig:pitqq}.
-A more careful tuning of bandwidth or smoothing during the validation process appears to be necessary for many of the machine learning based codes in order to improve the accuracy of $p(z)$.
-For narrow peaked $p(z)$ the parameterization of the PDF as evaluated on a fixed redshift grid could also have contributed to some overestimates of $p(z)$ width simply due to the finite resolution.
-After evaluating resuts such as those presented in \citet[]{Malz:qp}, in future analyses we plan to switch from a fixed grid to quantile-based storage of $p(z)$ in order to more efficiently and accurately store redshift PDF results.
-
-Another important factor to keep in mind when examining the results presented in this paper is the fact that they are at some level dependent on the metrics that we aim to optimize: in this case code participants were asked to submit their optimal measures of an accurate $p(z)$, so participants used the training/validation data to optimize their codes accordingly.
-Had we, instead, asked for an optimal $\hat{N(z)}$ the resulting metrics would be different for most, if not all, of the codes, as they would optimize toward a different goal.
-Specific metric choice can affect which codes are among the ``best'' codes.
-As stated earlier, there are cosmological science cases that require either individual galaxy photo-$z$ measures, or ensemble $\hat{N}(z)$ measures.
-We must be aware of that the optimal method for one is not necessarily optimal for the other, and in fact several photo-$z$ algorithms may be necessary in the final cosmological analysis in order to satisfy the requirements of all science use cases.
-The example of the simple \trainz\ estimator described in Section~\ref{sec:caution} shows a simple model with a $p(z)$ that is unrealistic for individual objects can still score very well on many of our metrics.
-It is important to look at {\it all} metrics, and keep in mind what information each metric conveys.
-%\red{mention 'null' and how 'dumb' algorithms can score well on metrics, need to keep all metrics in mind}
-%\aim{[reiterate the meaning of $p(z)$ and the goal -- emphasize finding ``best'' code in terms of impact of assumptions underlying the method \textit{or} establishing methodology for the realistic case of biased prior information and other science goals]}
-%\red{[the challenges faced in this project]}
-%\red{mention z$<$2, missing some important degeneracies, new sim will go to higher z}
-%\red{no quality cuts included, could identify outliers, but could also induce biases}
-%\red{[how optimization defined has impact, here we optimized p(z), not n(z), could be science-case specific if stringent requirements are needed, though that may be a computational/storage challenge if too many use cases needed.]}
-%\red{some discussion of output parameterization, limitations of 200 point fixed grid, in future switch to something like quantiles (cite qp paper)}
-We re-emphasize that the dataset tested was quite idealized, and discuss enhancements that will be added in future simulations to test photo-$z$ codes on increasingly realistic conditions in the following section.
-
-\subsection{Future work}
-\label{sec:futurework}
-The work presented in this paper is only the first step in characterizing current photo-$z$ codes and moving toward an improved photometric redshift estimator.
-This initial paper explored code performance in idealized conditions with perfect catalog-based photometry and representative training data.
-As mentioned in Section~\ref{sec:stackedmetrics} for the stacked $N(z)$ metrics we examined only the entire galaxy population with no selections in either photo-$z$ ``quality" or redshift.
-The cosmological analyses for weak lensing and large scale structure based measures plan to break galaxy samples into tomographic redshift bins, using photo-$z$ $p(z)$ to infer the redshift distribution for each bin.
-The specific selection used to determine these bins, both algorithmically and the specific bin boundaries, could induce biases due to indirect selections inherent in the photo-$z$ or other bin selection parameters.
-The effects of tomographic bin selection will be explored in a dedicated future paper.
-\red{[are there any references for this?
-I remember Gary Bernstein talking about this at a photo-z workshop in Japan, but I don't know that it was published.
-I believe Michael Troxel has discussed this as well.]}
-We also plan to propagate the uncertainties measured in a set of fiducial tomographic redshift bins in order to estimate impact on cosmological parameter estimation.
-%\item \red{cosmological parameters in DC2}
-%\item \red{mention z$<$2, missing some important degeneracies, new sim will go to higher z}
-%\red{no quality cuts included, could identify outliers, but could also induce biases}
-
-In future papers we will add more and more complexity to our simulated data in order to test photo-$z$ algorithms in increasingly realistic conditions.
-The most pressing concern is the impact of incomplete spectroscopic training samples.
-As discussed extensively in \citet{Newman:2015} a representative set of spectroscopically confirmed galaxies spanning the full range of both redshift and apparent magnitude is necessary as a training set to characterize the mapping from broad-band fluxes to photometric redshifts.
-However, due to a combination of factors due to both the galaxy SEDs and limitations of spectrographic instruments, redshift samples are known to be systematically incomplete, where certain galaxy types and redshift intervals fail to yield a redshift even at the longest integration times on current and near-future instruments.
-The more representative the training data, the better the performance of photo-$z$ algorithms will be.
-Current and upcoming surveys are putting in significant effort into obtaining these training samples \citep[e.~g.\,][]{Masters:2017}, however we still expect significant incompleteness for LSST-like samples, particularly at faint magnitudes.
-One major focus of an upcoming {\it LSST Dark Energy Science Collaboration Photo-z Working Group} data challenge is to produce a realistically incomplete training set of spectroscopic galaxies, modeling the performance of spectrographs, emission-line properties, and expected signal-to-noise to determine which galaxies will fail to yield a secure redshift.
-In addition to outright redshift failures we will model the inclusion of a small number of falsely identified secure redshifts where misidentified emission lines or noise spikes cause an incorrect redshift solution to be marked as a high quality identification.
-Even sub-per cent level contamination by false redshifts can impact photo-$z$ solutions at levels comparable to the stringent  requirements of some LSST science cases.
-We expect different systematics to occur in different photo-$z$ codes in response to training on incomplete data, particularly some of the machine learning methods.
-The response of the codes will inform future directions of code development.
-
-This initial paper explored a data set that was constructed at the catalog level, with no inclusion of the complications that come from measuring photometry from images.
-Future data challenges will move to catalogs constructed from mock images, including effects that will have great impact on photo-$z$ measurements.
-Object blending will be a major area of investigation, as the mixing of flux from multiple objects and the resultant change in measured colours is predicted to affect a large fraction of LSST galaxies \citep{Dawson:2016}, and will be one of the major contributing systematics for photo-$z$'s.
-Inclusion of differing observing conditions (seeing, clouds, variations in filter curves, Galactic dust, ...), as well as models for instrumental and system effects, sky masks, will all impact object photometry, and will be explored in the upcoming data challenge and their impacts described in upcoming papers.
-All underlying SEDs were parameterized as a weighted combination of five basis SEDs, with no additional accounting for host galaxy dust obscuration beyond what was encoded in the basis templates.
-This, in effect, limited the simulation to a very simple model of internal obscuration.
-Future simulations will include a more complicated and realistic treatment of host galaxy dust.
-
-The underlying simulation used in this work was based on a light-cone constructed to a maximum redshift of $z=2$.
-LSST imaging after 10 years of observations will include a significant number of $z>2$ galaxies in expected cosmology samples, and their inclusion does have potential significant implications for photo-$z$ measures: the high redshift galaxies lie at fainter apparent magnitudes and can have anomalous colours due to evolution of stellar populations and the shift to rest-frame magnitudes probing UV features of the underlying SED.
-More importantly, one of the most common ``catastrophic outlier'' degeneracies observed in deep photometric samples occurs when the Lyman break is mistaken for the Balmer break, leading to multiple redshift solutions at $z\sim0.2-0.3$ and $z\sim2-3$ \citep{Massarotti:2001}.
-This degeneracy, along with other potential degeneracies, are currently not covered by the limited redshift range of this initial paper, which could mean that we are not probing the full range of potential extreme outlier populations and how our photo-$z$ estimators respond to them.
-Extending simulations to include the high-redshift galaxy population will be a priority in future data challenges.
-
-%\item \red{inclusion of incompleteness effects on training sample, ``wrong redshifts'', filter curves}
-%\item \red{inclusion of photometric errors, image based effects, blending, lensing, sky, mask, etc...}
-
-In this study we have not accounted for the presence of Active Galactic Nuclei (AGN) contributions to galaxy fluxes.
-In some cases, AGN will be easily identified from the colors and morphologies, i.e.
-the case of the brightest quasars where the nuclear activity outshines the host galaxy, and numerous studies have utilized color selection to create large samples of quasars \citep[e.g.][]{Richards:06,Maddox:08,Richards:15}.
-In current deep fields, similar in depth to what we expect from LSST, variability information and multi-wavelength data have been critical to not only identify AGN dominated galaxies, but also obtain more accurate photometric redshifts \citep[e.g][]{Salvato:11}.
-
-In addition to AGN dominated galaxies, those with lower levels of nuclear activity present a more insidious problem, where AGN features may not be apparent, but the colors and other host galaxy properties are perturbed relative to galaxies with an inactive nucleus.
-In such cases, the presence of the AGN may induce a bias if the template SEDs or empirical datasets do not include low-level AGN counterparts.
-For LSST, we will need to identify and obtain accurate photometric redshifts of all types of AGN for a range of science goals, whether it is to eliminate such objects from cosmology experiments, or to use them with confidence, all the way through to understanding galaxy evolution and the role that AGN may play in influencing galaxy properties over cosmic time.
-
-A promising route to classifying and obtaining accurate photometric redshifts for the AGN population is by combining machine learning with template-fitting techniques, as has recently been demonstrated by \citet{Duncan:18} for radio-selected AGN.
-This is because AGN are relatively easy to obtain spectroscopic redshifts for over all redshifts due to the strong emission lines that they exhibit, allowing very good training sets for machine learning algorithms to use.
-Whereas for those galaxies where the AGN is sub-dominant the galaxy templates are still adequate for obtaining reasonable photometric redshifts.
-
-In addition to these improvements, the DESC Photo-$z$ group plans to look at all potential methods to combine the results from multiple photo-$z$ codes to improve $p(z)$ accuracy, similar to the work presented in \citet{Dahlen:13,Carrascokind:14,Duncan:18}.
-Taking advantage of multiple algorithms that use observables in slightly different ways has shown promise, however we must be very conscious of whether a potential combination properly treats the covariance between the methods, given that they are estimating quantities based on the same underlying observables.
-Several science cases wish to estimate physical quantities along with redshift, for example galaxy stellar mass and star formation rate.
-Proper joint estimation of redshift and physical quantities requires an in depth understanding of galaxy evolution, and progress on accurate bivariate redshift probability distributions will go hand in had with progress on understanding galaxies themselves.
-Parameterization and storage of a complex 2-dimensional probability surface for potentially billions of galaxies (or even subsets of hundreds of thousand of particular interest) pose a potential challenge.
-These issues will be examined in another future paper.
-
-Finally, while this paper and future papers discussed above focus on photometric redshift codes and estimating accurate $p(z)$ from training data, we plan a separate, but complementary, project to examine calibration of the resultant redshifts via spatial cross-correlations \citep{Newman:2008}, which will be explored in a separate series of future papers.
-The overarching plan describing everything laid out in this section is described in more detail in the LSST DESC Science Roadmap (see Footnote in Section~\ref{sec:intro}).
-These plans will require significant effort, but they are necessary if we are to make optimal use of the LSST data for astrophysical and cosmological analyses.
-%\begin{enumerate}
-%\item \red{the combination of codes}
-
-%\item \red{$p(z,\alpha)$}
-%\item \red{tomographic analysis later on}
-%\red{this paper deals with training, future papers will also explore calibration via cross-correlation methods}
-%\end{enumerate}
-%\red{mention the SRM once again, say that the plans are extensive, but we do have a plan and a rough timeline.}
-
-\section*{acknowledgements}
-\red{[summary of effort breakdown], will need to be in authors.csv for auto-generation by start\_paper}
-
-\red{[personal funding sources]}
-
-\red{[NERSC computation acknowledgement}
-
-The LSST Dark Energy Science Collaboration acknowledges generous ongoing support from the agencies and institutes supporting the collaboration.
-These include the Institut National de Physique Nucl\'eaire et de Physique des Particules in France, the Science \& Technology Facilities Council in the United Kingdom, and the Department of Energy, the National Science Foundation, and the LSST Corporation in the United States.
-This research used resources of the National Energy Research Scientific Computing Center, a DOE Office of Science User Facility supported by the Office of Science of the U.S.
-Department of Energy under Contract No.
-DE-AC02-05CH11231.
-Part of this work was undertaken on STFC DiRAC HPC Facilities, funded by UK BIS National E-infrastructure capital grants, and on the UK particle physics grid, supported by the GridPP Collaboration.
-
-\appendix
-
- \section{Point Estimate Photometric Redshifts}
- \label{sec:pointmetrics}
- While we do not recommend the use of single point estimates of redshift for most science applications, plots of the point estimates can be a useful qualitative diagnostic of photo-z code performance, i.~e.~examining point photo-$z$ vs.~spec-$z$ plots visually can give a quick impression of some common trends in different codes.
-Computing point estimate statistics may also be useful for more direct comparisons with previous photo-z evaluations.
-If a point-estimate is preferred for a specific science case, it is fairly simple to compute the mean, mode, or some other simple estimator from each $p(z)$, so these point estimates can be easily derived from the stored $p(z)$.
-
-There are several common point estimators of photo-$z$ posteriors employed by different codes, e.~g.
-the mode, mean, median of the $p(z)$ distribution.
-In addition, many of the machine learning based estimators can be set up to return a single redshift solution.
-For example, SkyNet can be configured to run as a regressor that returns a single float rather than a classifier that returns a 200-bin $p(z)$ estimate.
-The single value returned by a machine learning based code may not correspond to a particular measure such as the mode or mean, and so to avoid interpretation of results that might be introduced by variations in choice of specific point-estimate implementation per code, we discard the code-specific point estimates.
-We instead calculate point estimates more uniformly across the codes directly from the $p(z)$ using two measures, $z_{PEAK}$ and $z_{WEIGHT}$.
-$z_{PEAK}$  is simply the maximum value attained for each galaxy p(z), the mode of the probability distribution.
-$z_{WEIGHT}$ is defined similarly to how it is defined in \citet{Dahlen:13}, as the weighted mean of the redshift over the {\it main peak} of $p(z)$ containing the $z_{PEAK}$ value.
-The main peak is defined by subtracting 0.05$\,\times\,z_{PEAK}$ from $p(z)$ and identifying the roots to isolate the peak containing $z_{PEAK}$, $z_{WEIGHT}$ is defined as the weighted mean redshift within this peak.
-We restrict to a single peak in order to avoid confusion from bimodal and multimodal $p(z)$ such as those shown in bottom panels of Figure~\ref{fig:pz_examples}.
-For example, for a bimodal probability distribution a weighted mean calculated over both peaks would fall between the peaks, at a redshift where the probability is minimal.
-Restricting the weighting to a single peak ensures that the point estimate will fall in the region of maximum redshift probability.
-%\red{Have Dritan check zweight description to make sure this is what was done.}
-% The most common point estimators of photo-$z$ PDFs are the mode, median, and mean of the distribution.
-Following \cite{Tanaka:17}, we report the values of the following metrics using whichever of these point estimators performs best for each method, noting which point reduction is used in each case.
-
- \subsection{Point Estimate Metrics}
- \label{sec:point_metrics}
-We calculate the commonly used point estimate metrics of the overall photo-$z$ scatter ($\sigma_{z}$, the standard deviation of the photo-$z$ residuals), bias, and ``catastrophic outlier rate''.
-Specifically, we calculate the metrics as follows:
-we define $e_{z}$ as
-
-\begin{equation}
-e_{z} = \frac{z_{P} -z_{S}}{1+z_{S}} 
-\end{equation}
-where $z_{P}$ is the point estimate and $z_{S}$ is the true redshift.
-In practice, because the standard deviation calculation is quite sensitive to the outliers, we define the photo-$z$ scatter, $\sigma$ in terms of the Interquartile Range (IQR), the difference between the 75th and 25th percentiles of the $e_{z}$ distribution.
-In order to match the usual meaning of a 1$\sigma$ interval, we scale the IQR and define $\sigma_{IQR} = IQR/1.349$, as there is a factor of 1.349 difference between the IQR and the standard deviation of a Normal distribution.
-%$\sigma_{z}$ is, in practice, defined in terms the Interquartile Range (IQR) of $e_{z}$ values as $IQR/1.349$, where $IQR$ is the difference between the 75th and 25th percentiles of the $e_{z}$ distribution.
-The factor of 1.349 relates the IQR of a Normal distribution to the standard 68\% range of a one $\sigma$ uncertainty.
-%\scc{usually the statistical indicator defined as $\sigma$ is the standard deviation I know that is not a rule but if someone skips the test and go directly to the table (and we all know that a lot of people skips the definition of statistical indicator) could be confusing, could we report it as $\frac{IQR(e_z)}{1.349}$ or $\sigma_{IQR}$ that should be clear without any risk of confusion?}\red{Using IQR is done because IQR is less sensitive to outliers than calculating a sigma.
-I've rewritten the text explaining this above.--SJS}
-While many other studies define the bias based on the {\it mean} offset between true and estimated redshift, in this study we define the bias as the median value of $e_{z}$ for the sample.
-We use median as it is, once again, less sensitive to outliers than the mean.
-The catastrophic outlier fraction is defined as the fraction of galaxies with $e_{z}$ greater than the {\it larger} of $3\sigma_{IQR}$ or 0.06, i.e.
-3$\sigma$ outliers with a floor of $\sigma_{IQR}$=0.02.
-%\scc{again, is clearly not a rule but a huge number of paper refers to the mean as bias, could we use the word median, which is clear without any doubt rather than bias?}
-%\red{I also have code to calculate $\sigma_{MAD}$, should we include this as well?  It's almost always within a few thousandths of $\sigma_{IQR}$, so I left it out for now}.
-%\scc{we may just say that the two indicators reports almost the same value therefore there is no need for both of them?}
-For reference, the goals stated in Section 3.8 of the LSST Science Book \citep{Abell:09} for photo-$z$ performance in these metrics, assuming perfect training knowledge (as we are testing in this paper) are:
-\begin{itemize}
-\item RMS scatter$ < 0.02(1+z)$ 
-%\scc{actually the SB reports in page 75 that the goal of 0.02 is for the RMS of $\frac{\sigma_{z}}{(1+z)}$ while we are using IQR}
-\item bias $<$ 0.003 %\scc{in page 518 of the SB is clearly reported that the mean is expected to be less than 0.003 in our case we have defined as bias the median and not the mean }
-\item catastrophic outlier rate $<$ 10\% %\scc{clearly the number of outliers since are defined above $3\sigma$ strictly depends on the sigma definition, in the science book it seems to refer to the RMS of the $e_z$ distribution even ef it could be the standard deviation, but again I don't think it refers to IQR, therefore we could not consider exactly this numbers as reference.}
-%\red{I am going off of conversations with Zeljko and how we {\it actually} computed the metrics in the Science Book (I have the scripts).
-You are correct that several of these are slightly different than actually stated in the Science Book, or where the Science Book does not use very precise language as to what was done to compute the RMS or define the bias.
-We can also cut out those specifics, or maybe say Ivezic private communication, maybe?
-Also $\sigma_{z}$ defined in terms of IQR is exactly equivalent to the standard deviation when scaled by 1.349 for a Gaussian distribution, so the numbers are appropriate to cite in this context.
-IQR is just a more robust way of calculating sigma for a distribution with outliers.--SJS }
-%\scc{I am very sorry I am not saying that IQR is not fine, I want to keep the IQR my concerns are related to the notation and the \textit{reader understanding}, if I wrote $\sigma$ somewhere in a table the reader will understand for sure that is the traditional standard deviation (a lot of people skips the indicators definition for the \textit{obvious} ones), I am suggesting to replace the notation $\sigma$ with the notation $\sigma_{IQR}$ (or something like this) to avoid any confusion, and also to change from $bias$ to $median$ for the same reason.
-I agree that say Ivezic private communication could be better.}
-\end{itemize}
-These definitions are similar, but not exactly the same, as the $\sigma_{IQR}$ and median bias calculated here, but are similar enough for qualitative comparisons to the LSST goals.
-
-%Scatter: SRD $\sigma<0.02(1+z)$
-%Catastrophic Outler Rate: SRD $3\sigma$ ``catastrophic outlier'' rate below 10 per cent
-%Bias: SRD bias$<0.003$
-
-Fig.~\ref{fig:pz_pointestimates}  shows the point estimates for both $z_{PEAK}$ and $z_{WEIGHT}$.
-Point density is shown with mixed contours to emphasize that most of the galaxies do fall close to the $z_{phot}=z_{spec}$ line, while blue points show differing characteristics of the outlier populations.
-The red dashed lines indicated the cutoff for catastrophic outliers, defined as: $max(0.06,3\sigma_{IQR})$.
-As with the full $p(z)$ results, a variety of behaviours are evident in the different codes.
-Table~\ref{tab:pointestimates} lists the scatter, bias, and catastrophic outlier fractions for the codes.
-The performance of the codes for point metrics is highly correlated with performance on $p(z)$ based tests, which is to be expected, given that the point-estimates were derived from the $p(z)$.
-Some discretization is evident in $z_{PEAK}$, particularly for \textsc{SkyNet}, due to the finite grid spacing of the reported $p(z)$.
-These discreteness effects are mitigated by the weighting of $z_{WEIGHT}$, resulting in a smoother distribution of redshift estimates.
-Several features perpendicular to the main $z_{phot}=z_{spec}$ line are evident.
-These features are due to the 4000 angstrom break passing through the gaps between adjacent LSST filters.
-These features are most prominent in template-based codes, but appear to some degree in all codes tested.
-
-In even the best performing codes, there are visible occupied regions away from the $z_{phot}=z_{spec}$ line, corresponding to degenerate redshift solutions for certain LSST magnitudes and colors.
-While use of the full information available via $p(z)$ mitigates their impact, a full understanding of the outlier population is critical for LSST science, particularly in tomographic applications %\red{I forget what exactly I was trying to say here, if someone else wants to take a crack at this it might be helpful --SJS}
-%\scc{may be worth to say that the usage of further bands could mitigate this effect by removing the degeneration  usually induced from emission lines moving across different filters?}
-
-Finally, we note that all eleven codes are at or near the goals for point-estimates as outlined in the LSST Science Requirements Document\footnote{available at: \url{http://ls.st/srd}} and \citet{Graham:17}.
-This is to be expected, given that the requirements were designed such that a point estimate photo-z would meet these requirements for perfect training data to a depth of $i<25$.
-But, it is still an encouraging sign, given an updated mock galaxy simulation and the expanded set of photo-$z$ codes tested.
-%\red{maybe say consistent with expectations, and a reference to Melissa's paper in here?  What else to say about nearly meeting point goals?}
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.9\textwidth]{figures/ZPEAK_szpz_threecolumn_12codes_navy.jpg}
-\includegraphics[width=0.9\textwidth]{figures/ZWEIGHT_szpz_threecolumn_12codes_navy.jpg}
-\caption{Point estimate photo-z's derived from the posteriors.
-Top panel shows $z_{PEAK}$, while bottom panel shows $z_{WEIGHT}$.
-Point estimate density is represented with fixed density contours, while outliers at lower density are represented by blue points.
-While use of point-estimate photo-$z$'s is not recommended, they do make for useful comparative and visual diagnostics.
-In the lower-right panel of each plot, the \trainz\ estimator results in identical photo-$z$ estimates at the mode and mean of the training set $N'(z)$ distribution for all galaxies.} \label{fig:pz_pointestimates}
-\end{figure*}
-
-\begin{table*}
-\begin{center}
-\caption{Point estimate statistics}\label{tab:pointestimates}
-\begin{tabular}{lcccccc}
-\hline
-\hline
-                  &            & $Z_{PEAK}$  &          &  & $Z_{WEIGHT}$          &\\
-\hline
-Photo-z Code       & $\frac{\sigma_{IQR}}{(1+z)}$ & median  & \multicolumn{1}{|p{0.75cm}|}{\centering outlier \\fraction} & $\frac{\sigma_{IQR}}{(1+z)}$ & median & \multicolumn{1}{|p{0.75cm}|}{\centering outlier \\fraction}\\
-\hline
-\textsc{ANNz2}     & 0.0270  &  0.00063  & 0.044      & 0.0244  &  0.000307  & 0.047  \\ 
-\textsc{BPZ}       & 0.0215  & -0.00175  & 0.035      & 0.0215  & -0.002005  & 0.032 \\ 
-\textsc{Delight}   & 0.0212  & -0.00185  & 0.038      & 0.0216  & -0.002158  & 0.038 \\ 
-\textsc{EAZY}      & 0.0225  & -0.00218  & 0.034      & 0.0226  & -0.003765  & 0.029 \\ 
-\textsc{FlexZBoost}& 0.0154  & -0.00027  & 0.020      & 0.0148  & -0.000211  & 0.017 \\ 
-\textsc{GPz}       & 0.0202  & -0.00091  & 0.036      & 0.0201  & -0.000950  & 0.037 \\ 
-\textsc{LePhare}   & 0.0236  & -0.00161  & 0.058      & 0.0239  & -0.002007  & 0.056 \\ 
-\textsc{METAPhoR}  & 0.0264  &  0.00000  & 0.037      & 0.0262  &  0.001333  & 0.048 \\ 
-\textsc{CMNN}        & 0.0184  & -0.00132  & 0.035      & 0.0170  & -0.001049  & 0.034 \\
-\textsc{Skynet}    & 0.0219  & -0.00167  & 0.036      & 0.0218  &  0.000174  & 0.037 \\
-\textsc{TPZ}       & 0.0161  &  0.00309  & 0.033      & 0.0166  &  0.003048  & 0.031 \\ 
-\hline
-\trainz	   & 0.1808  &  -0.2086  & 0.000	  & 0.2335  & 0.022135  & 0.000\\
-\end{tabular}
-\end{center} 
-\end{table*}
-
-\bibliographystyle{mn2e}
-\bibliography{references}
-\end{document}

### 2018-08-19 paper/tex/results.tex
fix '\&' in affiliations, and runaway italic in results.tex

@@
 a horizontal line at the mode and mean of the training set distribution respectively.
-The explicit dependence on the \it{individual} posteriors in the calculation of the CDE loss, described in Section~\ref{sec:CDE_loss}, distinguishes this metric from the other $p(z)$ metrics that test the overall ensemble of $p(z)$ distributions.
+The explicit dependence on the {\it individual} posteriors in the calculation of the CDE loss, described in Section~\ref{sec:CDE_loss}, distinguishes this metric from the other $p(z)$ metrics that test the overall ensemble of $p(z)$ distributions.
 With a representative training set, \trainz\ will score well on the ensemble metrics, but fails miserably for metrics tied to individual redshifts.

### 2018-08-20 paper/tex/appendix.tex
fixed figure sizes
resize fig A1 to fit caption

@@
 \centering
-\includegraphics[width=0.9\textwidth]{fig/ZPEAK_szpz_threecolumn_12codes_navy.jpg}
-\includegraphics[width=0.9\textwidth]{fig/ZWEIGHT_szpz_threecolumn_12codes_navy.jpg}
+\includegraphics[width=0.8\textwidth]{fig/ZPEAK_szpz_threecolumn_12codes_navy.jpg}\\
+\includegraphics[width=0.8\textwidth]{fig/ZWEIGHT_szpz_threecolumn_12codes_navy.jpg}
 \caption{Point estimate photo-z's derived from the posteriors.

### 2018-08-20 paper/main.tex

@@
-%\documentclass[a4paper]{article}
-\documentclass[usenatbib]{mn2e}
-%% Language and font encodings
-\usepackage[english]{babel}
-\usepackage[utf8x]{inputenc}
-\usepackage[T1]{fontenc}
-\usepackage{lineno}
-\linenumbers
-\modulolinenumbers[2]
+\RequirePackage[switch, columnwise, running, mathlines, displaymath,
+mathlines]{lineno}
+\RequirePackage{docswitch}
+% \flag is set by the user, through the makefile:
+%    make note
+%    make apj
+% etc.
+\setjournal{\flag}
 
-\usepackage{natbib}
-%% Sets page size and margins
-\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
+\documentclass[\docopts]{\docclass}
 
-%% Useful packages
+\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
+  \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
+  \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
+  \renewenvironment{#1}%
+     {\linenomath\csname old#1\endcsname}%
+     {\csname oldend#1\endcsname\endlinenomath}}%
+\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
+  \patchAmsMathEnvironmentForLineno{#1}%
+  \patchAmsMathEnvironmentForLineno{#1*}}%
+\AtBeginDocument{%
+\patchBothAmsMathEnvironmentsForLineno{equation}%
+\patchBothAmsMathEnvironmentsForLineno{align}%
+\patchBothAmsMathEnvironmentsForLineno{flalign}%
+\patchBothAmsMathEnvironmentsForLineno{alignat}%
+\patchBothAmsMathEnvironmentsForLineno{gather}%
+\patchBothAmsMathEnvironmentsForLineno{multline}%
+}
+
+% You could also define the document class directly
+%\documentclass[]{emulateapj}
+
+% Custom commands from LSST DESC, see texmf/styles/lsstdesc_macros.sty
+\usepackage{lsstdesc_macros}
+% \usepackage{natbib}
+% \usepackage{subcaption}
+\usepackage[colorinlistoftodos]{todonotes}
+% \usepackage[colorlinks=true, allcolors=blue]{hyperref}
+\usepackage{multirow}
+\usepackage{xspace}
+\usepackage{enumitem}
+\usepackage{graphicx}
+\usepackage{wasysym}
 \usepackage{amsmath}
 \usepackage{amsfonts}
-\usepackage{graphicx}
-\usepackage{subcaption}
-\usepackage[colorinlistoftodos]{todonotes}
-\usepackage[colorlinks=true, allcolors=blue]{hyperref}
-\usepackage{multirow}
+\graphicspath{{./}{./figures/}}
+\bibliographystyle{mnras}%mn2e
 
-\newcommand       \apj  {ApJ}
-\newcommand       \apjl  {ApJL}
-\newcommand       \apjs  {ApJS}
-\newcommand       \aj  {AJ}
-\newcommand       \aap {A\&A}
-\newcommand       \mnras {MNRAS}
-\newcommand       \nat {Nature}
-\newcommand       \araa {ARA\&A}
-\newcommand       \physrep {Phys.~Reports}
-\newcommand       \prd {Phys. Rev. D}
+% Add your own macros here:
+
+\newcommand{\lsst}{\textsc{LSST}}
+\newcommand{\lsstdesc}{\lsst\textsc{-DESC}}
+\newcommand{\pz}{photo-$z$}
+\newcommand{\pzpdf}{\pz\ PDF}% can easily change this to specify posterior throughout
 
@@
 \def\x{{\mathbf{x}}}
-\def\E{{\mathbb E}}
+\def\E{{\mathbb{E}}}
+
+% ======================================================================
+
+\begin{document}
+\linenumbers
 
 \title{An assessment of photometric redshift PDF performance in the context of LSST}
-% Alternate title proposal: Assessing the characterization of photometric redshift uncertainties in the regime of LSST
+% \title{Photometric redshift uncertainties for LSST \romannumeral{1}: A unified methodology for assessing photo-$z$ PDF performance}
 
-\author{LSST-DESC Photometric Redshift Working Group}
-
-\begin{document}
-\maketitle
+\maketitlepre
 
 \begin{abstract}
+
   Photometric redshift (photo-$z$) probability distribution functions (PDFs) are a key data product of nearly all upcoming galaxy imaging surveys.
   However, the photo-$z$ PDFs resulting from different techniques are not in general consistent with one another, and an optimal method for obtaining an accurate PDF remains unclear.
-We present the results of an initial study of the Large Synoptic Survey Telescope Dark Energy Science Collaboration (LSST-DESC), the first in a planned series of papers testing multiple photo-$z$ codes on simulations of upcoming LSST galaxy photometry catalogues.
+  We present the results of an initial study of the Large Synoptic Survey Telescope Dark Energy Science Collaboration (\textsc{LSST-DESC}), the first in a planned series of papers testing multiple photo-$z$ codes on simulations of upcoming LSST galaxy photometry catalogues.
   This initial test evaluates photo-$z$ algorithms in the presence of representative training data and in the absence of several common sources of systematic errors that affect the procedures by which photo-$z$ PDFs are derived.
@@
   A careful accounting of all systematics discovered will be necessary for the codes employed in upcoming analyses in order to achieve unbiased cosmological measurements.
+  % Photometric redshift (photo-$z$) probability distribution functions (PDFs) more completely capture the complex nature of photo-$z$ uncertainties and are thus a key data product of nearly all upcoming galaxy imaging surveys, including the Large Synoptic Survey Telescope (LSST).
+  % Photo-$z$ PDFs produced by different algorithms do not in general agree with one another, and no one technique has yet been identified as optimal, leading to a proliferation of photo-$z$ PDF estimation codes.
+  % We present a unified comparison of twelve such codes using simulated LSST photometry, the first in a series of papers on photo-$z$ PDF performance for the LSST Dark Energy Science Collaboration (DESC).
+  % In this initial study, we isolate the effects of the estimation procedure by running each code using perfectly representative training data (for machine learning methods) or the true SED template library (for template-fitting methods) and confirm the mutual inconsistency of the resulting photo-$z$ PDF estimates.
+  % We apply a comprehensive suite of metrics to an the population of photo-$z$ PDFs as well as point estimates derived thereof and an accepted estimator of the redshift distribution $N(z)$.
+  % % The photo-$z$ PDFs are evaluated using multiple metrics including the Kolmogorov-Smirnoff statistic, Cramer-von Mises statistic, Anderson-Darling statistic, Kullback-Leibler divergence, $N(z)$ moments, quantile-quantile plots, and probability integral transform.
+  % We interpret the performance of the codes under each metric in terms of the assumptions underlying the overall approach of the methods and audit the sentivity of the metrics to the failure modes of different techniques, culminating in a discussion predicting the impact of introducing realistic systematics and the direction of attention to the choice of metric when identifying the most appropriate code for a given science use case.
+
 \end{abstract}
-\begin{keywords}
-galaxies: distances and redshifts -- galaxies: statistics -- methods: statistical
-\end{keywords}
-%\noindent comments color-code:\\
-%\red{SJS},
-%\aim{AIM},
-%\scc{SC},
-%\textcolor{brown}{MB},\jan{JAN}
-%\blue{unknown}\\
-%\red{note: not all red text is mine, at some point various people were using red --SJS.}
 
-% Questions left unanswered:
-% 1. Which journal will this paper be published? 
-% Answer: AJ / ApJ / MNRAS (no charges, but not sure if it will accept a paper based on a DC1 effort and the associated simulations)
-% 2. What is the meaning of $p(z)$ in different codes? (Ofer)
-% Answer: ?
+% Keywords are ignored in the LSST DESC Note style:
+\dockeys{galaxies: distances and redshifts -- galaxies: statistics -- methods: statistical}
+
+\maketitlepost
+
+% ----------------------------------------------------------------------
+% {% raw %}
+
+\input{./tex/introduction}
+
+\input{./tex/data}
+
+\input{./tex/methods}
+
+\input{./tex/analysis}
+
+\input{./tex/results}
+
+\input{./tex/discussion}
+
+\input{./tex/conclusion}
+
+% ----------------------------------------------------------------------
+
+\subsection*{Acknowledgments}
+
+%%% Here is where you should add your specific acknowledgments, remembering that some standard thanks will be added via the \code{desc-tex/ack/*.tex} and \code{contributions.tex} files.
+
+%This paper has undergone internal review in the LSST Dark Energy Science Collaboration. % REQUIRED if true
+
+\input{contributions} % Standard papers only: author contribution statements. For examples, see http://blogs.nature.com/nautilus/2007/11/post_12.html
+
+% This work used TBD kindly provided by Not-A-DESC Member and benefitted from comments by Another Non-DESC person.
+
+% Standard papers only: A.B.C. acknowledges support from grant 1234 from ...
+
+The authors would like to thank their LSST-DESC publication review committee.
+
+AIM is advised by David W. Hogg and was supported by National Science Foundation grant AST-1517237.
+
+\input{desc-tex/ack/standard} % also available: key standard_short
+
+% This work used some telescope which is operated/funded by some agency or consortium or foundation ...
+
+% The DESC acknowledges ongoing support from the Institut National de Physique Nucleaire et de Physique des Particules in France; the Science \& Technology Facilities Council in the United Kingdom; and the Department of Energy, the National Science Foundation, and the LSST Corporation in the United States.
 %
+% DESC uses resources of the IN2P3 Computing Center (CC-IN2P3--Lyon/Villeurbanne - France) funded by the Centre National de la Recherche Scientifique; the National Energy Research Scientific Computing Center, a DOE Office of Science User Facility supported by the Office of Science of the U.S.
+Department of Energy under Contract No.
+DE-AC02-05CH11231; STFC DiRAC HPC Facilities, funded by UK BIS National E-infrastructure capital grants; and the UK particle physics grid, supported by the GridPP Collaboration.
+%
+% This work was performed in part under DOE Contract DE-AC02-76SF00515.
 
-\section{Introduction}
-\label{sec:intro}
+% We acknowledge the use of An-External-Tool-like-NED-or-ADS.
 
-%(Sam Schmidt, Ofer Lahav, Jeff Newman, Alex Malz, John Soo, Tony Tyson)
+%{\it Facilities:} \facility{LSST}
 
-%\jan{Is the intention to keep these parenthetical author sets?
-I find it a bit offputting in the text.
-It might be better to put this for sections in the acknowledgements section, and for individual codes/statistics specify who ran/developed it with a sentence.}\red{I was just about to comment these out before finalizing the internal review draft of the paper.
-They are in place to remind me to contact a few people who still have not filled out their authors.csv info.}
+\input{./tex/appendix}
 
-Large-scale photometric galaxy surveys are entering a new era with currently or soon-to-be running Stage III and Stage IV dark energy experiments like the Dark Energy Survey \citep[DES,][]{Abbott:05}, the Kilo-Degree Survey \citep[KiDS,][]{de_Jong:13}, Large Synoptic Survey Telescope \citep[LSST,][]{Abell:09}, Euclid \citep{Laureijs:11}, and Wide-Field Infrared Survey Telescope \citep[WFIRST,][]{Green:12}.
-The move to imaging based surveys, rather than spectroscopic based, for cosmological measurements makes proper understanding of photometric redshifts (``photo-$z$'s'') of paramount importance, as cosmological distance measures for statistical samples are directly dependent on photo-$z$ measurements.
+% Include both collaboration papers and external citations:
+\bibliography{main,lsstdesc}
 
-The unprecedented sample size of LSST galaxies, expected to number several billion for the main cosmological sample, necessitates stringent constraints on photo-$z$ accuracy if systematic errors are not to dominate the statistical errors.
-The LSST Science Requirements Document (SRD)\footnote{available at \url{https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17}} lists the photometric redshift goals for a magnitude limited sample with $i<25$ as: root-mean-square error with a goal of $\sigma_z<0.02(1+z)$; $3\sigma$ ``catastrophic outlier'' rate below $10\%$; bias below $0.003$\footnote{
-Note that at the time the SRD was written, these goals were stated in terms of a photo-$z$ point estimate for each galaxy, as was standard in many previous studies, while in this paper we emphasize the importance of using a full photo-$z$ PDF.}.
+\end{document}
 
-The tremendous size of LSST's galaxy catalog will be enabled by its exceptional depth, pushing to fainter magnitudes and deeper imaging and including galaxies of lower luminosity and higher redshift than ever before.
-In addition to the contribution of low signal-to-noise photometry to the systematic error of photo-$z$s, these populations introduce major physical degeneracies, for example the Lyman break/Balmer break degeneracy, that were not present in the populations covered in previous large area surveys like the Sloan Digital Sky Survey \citep[SDSS,][]{York:00} and the Two Micron All Sky Survey \citep[2MASS,][]{Skrutskie:06}.
-In order to meet LSST's demanding error budget, it will be necessary to fully characterize those degeneracies wherein multiple redshift solutions have comparable likelihood.
-
-There is often a desire to have a single valued ``point-estimate'' redshift for an individual galaxy.
-However, the complex, non-linear (and often non-unique) nature of the mapping between broad band fluxes and redshift means that a single value is unable to capture the full redshift information encoded in a galaxy's magnitudes.
-For example, a common point-estimate for a template-based method is taking the highest likelihood solution as the point photo-$z$.
-A single valued redshift ignores degenerate redshift solutions of lower probability, potentially biasing photometric redshift estimates both for individual galaxies and ensemble distributions.
-Storing more information is necessary, most often photo-$z$ codes output the redshift probability density function (PDF), also often referred to as $p(z)$, describing the relative likelihood as a function of redshift.
-%\red{should we move some of AIM's p(z) vs p(z|d,I) to here?}
-Early template methods such as \citet{Fernandezsoto:99} converted relative $\chi^2$ values of template spectra to likelihoods to estimate $p(z)$.
-Soon after, codes such as \citet{Benitez:00} added a Bayesian prior and output a posterior probability distribution.
-While many early machine learning based algorithms focused on a point-estimate, \citet{Firth:03} used a neural net with 1000 realizations scattered within the photometric errors to estimate a $p(z)$.
-As more groups began to employ photometric redshifts in their cosmological analyses, realization that point-estimate photo-$z$'s were inadequate for precision cosmology measurements \citep{Mandelbaum:2008}.
-From around this point onward, most photo-$z$ algorithms have attempted to implement some estimate of the overall redshift probability in their outputs, and some surveys began supplying a full $p(z)$ rather than a simple redshift point-estimate and error \citep[e.~g.~][]{de_Jong:17}.
-
-%\aim{We should spend a whole paragraph establishing that photo-$z$ PDFs are a thing, i.e.
-review the motivation for PDFs over point estimates, basic definition, early methods papers, surveys that have published them, etc.}
-%\scc{I agree with Alex, we should introduce here what a pdf is and which survey have already published some pdfs (KiDS dr3 for example) or will produce them (LSST, Euclid etc.)}
-%\red{See above paragraph, does this adequately address this?--SJS}\scc{fine for me}
-
-There are numerous techniques for deriving photo-$z$ PDFs from photometry, yet no one method has yet been established as clearly superior.
-Quantitative comparisons of photo-$z$ methods have been made before.  
-The Photo-$z$ Accuracy And Testing \citep[PHAT,][]{Hildebrandt:10} effort focused on point estimates derived from many photometric bands. 
-DES compared several codes for point estimates \citep{Sanchez:14} and a summary statistic of photo-$z$ interim posteriors for tomographically binned galaxy subsamples \citep{Bonnett:16}. 
-This paper is distinguished by its inclusion of metrics of photo-$z$ interim posteriors themselves and consideration of both classic and state-of-the-art photo-$z$ algorithms, comparing the performance of several of the most widely employed codes as well as some that have been developed only recently on the basis of metrics appropriate for a probabilistic data product.
-The results presented in this work are 
-%part of the Data Challenge 1 (DC1) 
- a major focus of the Photometric Redshift working group of the LSST Dark Energy Science Collaboration (LSST-DESC).
-This work is laid out in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}} as one of the critical activities to be completed in preparation for dark energy science analysis on the first year LSST data.
-This is the first of multiple papers by the working group, which will grow in sophistication.
-In this initial paper we focus on evaluating the performance of photometric redshift codes and their ability to produce accurate PDFs in the presence of representative training sets.
-This can be thought of as an initial test under near perfect conditions, before further complexities are added in future papers.
-Comparing the relative performance of the codes enables us to evaluate whether each code is using information in an optimal way, and may reveal enhancements in some codes and deficiencies in others, either in the fundamental algorithm, or in specific implementation.
-
-Certain science cases need redshift information on individual objects, e.~g.~identification of host galaxy redshift for supernova classification, or identifying potential cluster membership.  
-Other science cases need only ensemble redshift information; for instance current weak lensing techniques require the overall redshift distribution $N(z)$ for tomographic redshift samples, but do not need single galaxy estimates.
-In the case of the multiple types of probes of cosmology enabled by the
-LSST cosmology sample of several billion galaxies, the number of redshift bins and their photo-$z$ requirements vary with the specific probe;  2-point angular correlations benefit from many bins, while weak lensing probes do not (due to the wide lensing kernel).
-Large photometric surveys such as LSST must develop algorithms that meet the needs of all such science cases.  
-In order to meet these ambitious goals for photo-$z$ accuracy, every aspect of photo-$z$ estimation will have to be optimized: the algorithms employed, both template and machine-learning based (both in design and implementation); the spectroscopic data used as a training set for machine learning algorithms or to estimate template sets and train Bayesian priors; and probabilistic catalog compression schemes that balance information retention against limited storage resources.
-
-Before moving forward, we must address how the best methods may be unique to the performance metrics and science cases considered and what distinguishes photo-$z$ PDFs of different methods from one another.
-Though photo-$z$ PDFs are often written simply as $p(z)$, the PDF itself must be an interim posterior distribution $p(z | d, I)$, the probability of redshift conditioned on photometric data $d$ that has actually been observed and the prior information $I$ that guides how a redshift is extracted from the photometry.
-If we run multiple photo-$z$ codes on a single dataset, the photo-$z$ interim posteriors will not be identical because each code is based on assumptions in the form of an interim prior --- these assumptions form the premise for photo-$z$ estimation as a whole and are the only way to introduce differences in estimates of what would otherwise be a shared photo-$z$ posterior $p(z | d)$ regardless of the code used to obtain it.
-Though explicit knowledge of the interim prior is necessary to use photo-$z$ interim posteriors self-consistently in physical inference, the interim prior of a particular methodology is often implicit and not necessarily shared among all galaxies in the catalog.
-
-This paper therefore aims (1) to constrain the impact of the interim prior $I$ by separating it into a component $I_{H}$ representing the method itself and a component $I_{D}$ representing physical information, such as a training set or SED template library and (2) to present a procedure for evaluating the performance of photo-$z$ codes in generic tests that may include many more systematics in the interim prior $I$.
-In order to isolate the effects encapsulated by $I_{H}$ of variation between codes from issues with the training set or template library encapsulated by $I_{D}$, we use an identical set of simulated galaxies for every code and construct a template library and training sample that are {\it complete and representative} and shared among all codes; that is, our training sample for machine learning codes is drawn from the same underlying galaxy population as our test set, with no additional selections, and the SED library used for template-based codes is the same as the one used to generate the photometric data.
-We explore a number of performance metrics in this paper, not to make a conclusion regarding the superiority or even relative favorability of each code but to establish a method for comparing photo-$z$ PDFs derived by different methods.
-These test conditions set the stage for addressing in a future paper the crucial issue of incomplete and non-representative prior information.
-
-The outline of the paper is as follows: in \S\,\ref{sec:sims} we present the simulated data set; in \S\,\ref{sec:pzcodes} we describe the current generation codes employed in the paper;  in \S\,\ref{sec:metrics} we discuss the interpretation of photo-$z$ PDFs in terms of metrics of accuracy; in \S\,\ref{sec:results} we show our results and compare the performance of the codes; in \S\,\ref{sec:discussion} we offer our conclusions and discuss future extensions of this work.
-
-\section{The simulation and mock galaxy catalog}
-\label{sec:sims}
-%(Sam Schmidt, Eve Kovacs, Tina Peters)
-
-In order to test the current generation codes, we employ a simulated galaxy catalogue.
-The simulation is completely catalogue-based, with no image construction or mock measurements made.
-We describe these in detail below.
-
-\subsection{Buzzard-v1.0 simulation}
-\label{sec:buzzard}
-The \textsc{Buzzard-highres-v1.0} \red{put in cites to in prep Buzzard papers} catalogue construction started with a dark matter only simulation.
-This N-body simulation contained $2048^3$ particles in a $400$ Mpc h$^{-1}$ box.
-\red{[N]} snapshots (with smoothing and interpolation between snapshots) were saved in order to construct a lightcone.
-Dark matter halos were identified using the \textsc{Rockstar} software package \citep{Behroozi:13}.
-These dark matter halos were populated with galaxies with a stellar mass and absolute $r$-band magnitude in the SDSS system determined using a sub-halo abundance matching model constrained to match both projected two-point galaxy clustering statistics and an observed conditional stellar mass function \citep{Reddick:13}.
-
-To assign an SED to each galaxy, the {\it Adding Density Dependent Spectral Energy Distributions} (\textsc{ADDSEDS}, deRose in prep.)\footnote{\url{https://github.com/vipasu/addseds}} procedure was used.
-This consisted of training an empirical relation between absolute $r$-band magnitude, local galaxy density, and SED using a sample of $\sim 5e^{5}$ galaxies from the magnitude-limited Sloan Digital Sky Survey Data Release 6 Value Added Galaxy Catalog \citep{Blanton:05}\red{[Note: is this the proper reference to SDSS-NYU VAGC?
-File is called combined\_dr6\_cooper.fits, but I don't see which Cooper et al 2006 this is supposed to refer to?]}.
-Each SDSS spectrum is fit with a sum of five SED components using the \textsc{k-correct \red{v?}} software package\footnote{\url{http://kcorrect.org}} \citep{Blanton:07}, thus each galaxy SED is parameterized as five weights for the basis SEDs.
-The distance to the spatial projected fifth-nearest neighbour was used as a proxy for local density in the SDSS training sample.
-For each simulated galaxy, a ``random'' \red{[details]} galaxy with ``similar'' \red{[details]} absolute $r$-band magnitude and local galaxy density was chosen from the training set, and that training galaxy's SED was assigned to the simulated galaxy.
-Given the SED, absolute $r$-band magnitude and redshift, we computed apparent magnitudes in the six LSST filter passbands, $ugrizy$.
-We assigned magnitude errors in the six bands using the simple model described in \citet{Ivezic:08}, assuming full 10-year depth observations had been completed.
-The number of total 30-second visits assumed when generating the photometric errors differs slightly from the fiducial numbers assumed for LSST: we assume 60 visits in u-band, 80 visits in g-band, 180 visits in r-band, 180 visits in i-band, 160 visits in z-band, and 160 visits in y-band.
-%numbers taken from Alex Abate's photErrorModel.py script in PhotoZDC1 repository, which was used with nYrObs=10.
-
-\subsubsection{Selection of training and test sets}
-\label{sec:buzztraining}
-The total catalogue covered $400$ square degrees and contained $238$ million galaxies to an apparent magnitude limit of $r\!=\!29$ and spanning the redshift range $0\!<\!z\!\leq\!8.7$.
-This catalogue contained two orders of magnitude more galaxies than were needed for this study, so only $\sim\!8$ square degrees were used.
-Systematic problems with galaxy colors above $z\!>\!2$ were observed, so the catalogue was trimmed to include only galaxies in the redshift range $0\!<\!z\!\leq\!2.0$.
-A random subset of the the remaining galaxies was chosen, and placed at random into either a ``training'' set ($10$ per cent of the sample), for which the galaxies true redshifts will be supplied, or a ``test'' set (the remaining $90$ per cent of the sample), for which each code will need to predict a redshift PDF for each galaxy.
-The resulting catalogues contain $111\,171$ training galaxies and $1\,000\,883$ test galaxies.
-We restrict our analysis to a sample with $i<25.3$, which give a signal-to-noise $\sim\,30$ for most galaxies, a cut often referred to as the expected ``LSST Gold Sample''.
-This magnitude cut results in a training set with $44\,404$ galaxies and a test set containing $399\,356$ galaxies.
-All subsequent results will evaluate this ``gold sample'' test set.
-
-\subsubsection{Templates}
-\label{sec:buzztemplates}
-As mentioned in Section~\ref{sec:buzzard}, the SEDs in the Buzzard simulation are drawn from an empirical set of SEDs taken from the SDSS DR6 NYU-VAGC, a sample of roughly $\sim5e^{5}$ galaxies with spectra in SDSS.
-To determine a finite set of templates to use with template fitting codes we take the five SED weight coefficients for each of the $\sim 500\,000$ galaxies in the SDSS sample and run a simple K-means clustering algorithm on this five dimensional space.
-The K-means cluster centres span the space of coefficients and properly reflect the underlying density in the coefficient space, thus providing a reasonable approximation for a spanning SED set.
-An ad-hoc number of $K=100$ was chosen and the $100$ K-means centre positions are taken as the weights for the \textsc{k-correct} SED components to construct one hundred template SEDs.
-These $100$ templates were provided,
-%\red{(JS: if I remembered correctly, weren't there 150 templates given?
-Answer: no, initially I supplied a list of 150 that were drawn with a different algorithm, but once we switched to the final k-means set I only supplied 100 (the initial algorithm over-weighted outlier SEDs, which is why we needed the extra 50)--SJS)}
-however not every template code uses this set of one hundred templates: because EAZY was designed and written to use the same five basis templates employed by \textsc{k-correct} when constructing our mock galaxies, EAZY was run using linear combinations of these five templates rather than using the 100 discrete templates.
-
-\subsubsection{Limitations}
-\label{sec:buzzlimitations}
-For our initial investigation of photometric redshift codes, we begin with a data set that is somewhat idealized, and does not contain all of the complicating factors present in real data.
-In several cases, the simplification is done with a purpose, with potentially confounding effects excluded in order to better isolate the differences between current-generation photo-$z$ codes, and their causes.
-We list several of the simulations limitations in this section.
-As the simulation is catalogue-based, no image level effects, such as photometric measurement effects, object blending, contamination from sky background (Zodiacal light, scattered light, etc...), lensing magnification, or Galactic reddening are included.
-No stars are included in the catalogue, nor are the effects of AGN.
-As all SEDs are constructed from only five basis templates, properties of the galaxy population will be restricted to follow linear combinations of the characteristics of the five basis templates, so certain non-linear features, for example the full range of emission line fluxes relative to the continuum, will not be included in the model galaxy population.
-No additional dust reddening intrinsic to the host galaxy is included, the only approximation of dust extinction comes in the form of dust encoded in the five basis SEDs via the training set used to create the basis templates.
-Simple linear combinations of these basis templats will, once again, not explore the full range of realistic dust extinction observed in galaxy populations.
-
-%\subsection{ABC-Galacticus simulation}
-%\label{sec:galacticus}
-%\red{The Argonne-Berkeley-Carnegie (ABC) Galacticus cosmological simulation uses a semi-analytic model to simulate the galaxy properties and is not directly tuned to observations.
-Each simulated SED is replaced with a SED drawn from the distribution of empirical SEDs given by the SED Atlas of \citet{Brown:14} (diffusion maps method in development).}
-
-\section{Methods}
-\label{sec:pzcodes}
-
-%(Sam Schmidt, Rongpu Zhou, Ibrahim Almosallam, Eric Nuss, Johann Cohen Tanugi, John Soo)
-
-Here we outline the photo-$z$ PDF codes tested in this study.
-In total, eleven distinct codes are tested.
-This sample is not comprehensive, but does cover a broad range of current-generation codes.
-Both template-based and machine learning approaches are included and each are described separately in Secs.~\ref{sec:templatecodes} and \ref{sec:trainingcodes} respectively.
-The list of codes are summarized in Table.~\ref{tab:list_of_codes}.
-
-\begin{table*}  %%% DATA TABLE %%%
-\caption{List of photo-$z$ codes featured in this study. ML here means machine learning.} \label{tab:list_of_codes}\resizebox{\textwidth}{!}{
-\begin{tabular}{llll}
-\hline
-\bf Code 			& \bf Type & \bf Paper & \bf Website \\
-\hline
-\textsc{BPz} 		& template 			& \citet{Benitez:00}		& \url{http://www.stsci.edu/~dcoe/BPZ/} \\
-\textsc{EAZY} 		& template 			& \citet{Brammer:08}		& \url{https://github.com/gbrammer/eazy-photoz} \\
-\textsc{LePhare} 	& template 			& \citet{Arnouts:99}		& \url{http://www.cfht.hawaii.edu/~arnouts/lephare.html} \\
-\textsc{ANNz2} 		& ML 	& \citet{Sadeh:16}			& \url{https://github.com/IftachSadeh/ANNZ} \\
-\textsc{Delight} 	& ML/template 	& \citet{Leistedt:17}		& \url{https://github.com/ixkael/Delight} \\
-\multirow{ 2}{*}{\textsc{FlexZBoost}} & \multirow{ 2}{*}{ML} 	& \multirow{ 2}{*}{\citet{Izbicki:17}}		& \url{https://github.com/tpospisi/flexcode};\\ 
-					& 		&							& \url{https://github.com/rizbicki/FlexCoDE}\\
-%\textsc{Frankenz} 	& ML 	& \citet{Speagle:frankenz}	& \url{https://github.com/joshspeagle/frankenz} \\
-\textsc{GPz} 		& ML	& \citet{Almosallam:15b}	& \url{https://github.com/OxfordML/GPz} \\
-\textsc{METAPhoR} 	& ML 	& \citet{Cavuoti:17}		& \url{http://dame.dsf.unina.it}\\
-\textsc{CMNN} 		& ML 	& \citet{Graham:17}			& - \\
-\textsc{SkyNet} 	& ML 	& \citet{Graff:14}			& \url{http://ccpforge.cse.rl.ac.uk/gf/project/skynet/} \\
-\textsc{TPZ} 		& ML 	& \citet{Carrasco_Kind:13}	& \url{https://github.com/mgckind/MLZ} \\
-\hline
-\trainz 	& N/A	& See Section~\ref{sec:method:trainz} & \\
-\end{tabular}}
-\end{table*}
-
-%\red{The questions that must be answered for each code are\dots}
-The questions that must be answered for each code are: what unique features are included in the specific implementation that influence the output $p(z)$.
-What form of validation was performed with the training data, how were photometric uncertainties employed in the analysis, how were negative fluxes treated, what specific prior form was employed (for template based codes), or what specific machine learning architcture was used (for ML codes)?
-%\scc{which parameter (in a general meaning) have been used, which kind of validation has been performed (if any)}
-\subsection{Template-based Approaches}
-\label{sec:templatecodes}
-
-\subsubsection{BPZ}
-\label{sec:BPZ}
-%(Sam Schmidt)
-
-\textsc{BPZ}\footnote{\url{http://www.stsci.edu/~dcoe/BPZ/}} \citep[Bayesian Photometric Redshift,][]{Benitez:00} is a template-based photo-$z$ code that compares the expected colors ($C$) calculated for a set of spectral energy distribution (SED) types/templates ($T$) to the observed colors to calculate the likelihood of observing colors at each redshift for each type, $p(C|z,T)$.
-The code employs an empirically determined Bayesian prior in apparent magnitude ($m_0$) and SED-type.
-Assuming that the SED-types are spanning and exclusive, we can determine the redshift posterior $p(z|C,m_0)$ by marginalizing over all SED-types with a simple sum \citep[Eq.~3 from][]{Benitez:00}:
-
-\begin{equation} \label{eq:redshift_posterior}
-p(z|C,m_0)\,\propto\, \sum_{T}p(z,T|m_0)\,p(C|z,T)
-\end{equation}
-\noindent where the first term on the right-hand side is the Bayesian prior and the second term is the traditional likelihood.
-The prior is assumed to have the form: $p(z,T|m_0)=p(T|m_0)\,p(z|T,m_0)$, i.e.
-it parameterizes the prior as an evolving type fraction with apparent magnitude, combined with a prior on the expected redshift probability distribution as a function of both apparent magnitude and SED-type.
-
-In this paper we use \textsc{BPZ v 1.99.3}. The template set employed here is the set of $100$ discrete SEDs described in Section~\ref{sec:buzztemplates}
-%, and the $129$ Brown SEDs for the Galacticus dataset. 
-To keep the number of free parameters to a manageable level the SEDs in the training set are sorted by the rest-frame $u$-$g$ colour and split into three ``broad'' SED classes, equivalent to the \texttt{E}, \texttt{Sp} and \texttt{Im/SB} types in \citet{Benitez:00}.
-We assume the same functional form for the Bayesian priors as used by \citet{Benitez:00}, and utilize the training-set galaxies with known SED-type, redshift, and apparent magnitude to determine the type fractions and the best fit for the eleven free parameters of the prior.
-%For photo-$z$ point estimates we use the \texttt{Z\_B} output parameter. 
-For galaxies with negative flux in a measured band, the placeholder value is replaced with an estimate one $\sigma$ detection limit in that particular band, i.~e.~a value close to the estimated sky noise threshold.
-The type-marginalized $p(z)$ is generated by setting the parameter \texttt{PROBS\_LITE=TRUE} in the \textsc{BPZ} parameter file.
-
-\subsubsection{EAZY} \label{sec:eazy}
-%(Rongpu Zhou)
-
-\textsc{EAZY}\footnote{\url{https://github.com/gbrammer/eazy-photoz}} \citep[Easy and Accurate Photometric Redshifts from Yale,][]{Brammer:08} is a template-based photo-$z$ code that includes several features that improve on the basic $\chi^2$ fit used in many template codes.
-It can fit the observed photometry with SEDs created from a linear combination of a set of templates at each redshift, and the best-fit SED is found by simultaneously fitting one, two or all of the templates by minimizing $\chi^2$.
-The minimized $\chi^2(z)$ is then combined with an apparent magnitude prior to obtain the posterior redshift probability distribution, although some argue that this is not the mathematically correct way of calculating the posteriors.
-\textsc{EAZY} can also account for the uncertainties in the templates by adding an empirically derived template error in quadrature as a function of redshift to the flux errors.
-
-In this paper we use the all-templates mode, which fits the photometric data with a linear combination of the five basis templates.
-We employed the $5$ basis templates described in Section~\ref{sec:buzzard}, and set the template error to zero since these same templates were used to produce the simulated catalog photometry.
-%\red{The quality parameter \texttt{Qz} is...}
-The likelihoods are calculated on a 200-point redshift grid spanning $0\leq z \leq 2$, and include the application of a type-independent apparent magnitude prior estimated from the training data.
-
-\subsubsection{LePhare}\label{sec:lephare}
-%(C\'ecile Roucelle, Eric Nuss, Johann Cohen-Tanugi)
-
-\textsc{LePhare}\footnote{\url{http://www.cfht.hawaii.edu/~arnouts/lephare.html}}\citep[Photometric Analysis for Redshift Estimate,][]{Arnouts:99,Ilbert:06} is a photo-$z$ reconstruction code based on a $\chi^2$ template-fitting procedure.
-The observed colors are matched with the colours predicted from a set of spectral energy distribution (SED) which can be either synthetic or based on a semi-empirical approach.
-\textsc{LePhare} has been used to produce the COSMOS2015 photo-$z$ catalogue \citep{Laigle:16}.
-
-Each SED is redshifted in steps of $\Delta z = 0.01$ and convolved with the simulated LSST filter transmission curves (accounting for instrument efficiency).
-The opacity of the inter-galactic medium has been set to zero as no additional reddening has been included in the Buzzard simulations.
-The computed photo-$z$ is then the value that minimizes the merit function $\chi^2 (z,T,A)$ from \citet{Arnouts:99}:
-
-\begin{equation}
-\chi^2(z,T,A) = \sum_f^{N_f}\left(\frac{F^f_{\rm{obs}}−A \times F^f_{\rm{pred}}(T,z)}{\sigma^f_{\rm{obs}}}\right)^2 
-\end{equation}
-\noindent where $A$ is a normalization factor, $F^f_{\rm{pred}}(T,z)$ is the flux predicted for a template $T$ at redshift $z$.
-$F^f_{\rm{obs}}$ is the observed flux in a given band $f$ and $\sigma^f_{\rm{obs}}$ the associated observational error.
-The index $f$ refers to the considered band and $N_f$ is the total number of filters.
-
-In this paper we use \textsc{LePhare v 2.2}. 
-The set of templates used for fitting the photo-$z$'s are the 100 discrete Buzzard SED templates as described in section~\ref{sec:buzztemplates}.
-%No correction for the photometric zero-points was applied for this work. 
-The full $p(z)$ corresponds to the likelihoods calculated at each point on our $z$-grid.
-
-\subsection{Training-based Codes}
-\label{sec:trainingcodes}
-
-\subsubsection{ANNz2}
-\label{sec:annz2}
-%(John Soo)
-
-\textsc{ANNz2}\footnote{\url{https://github.com/IftachSadeh/ANNZ}} \citep{Sadeh:16} is a powerful package that has the ability to employ several machine learning algorithms, including artificial neural networks (ANN), boosted decision tree (BDT) and k-nearest neighbour (KNN).
-Using the Toolkit for Multivariate Data Analysis (TMVA) with ROOT\footnote{\url{http://tmva.sourceforge.net/}}, it can run multiple machine learning algorithms for a single training and outputs photo-$z$'s based on a weighted average of their performances.
-
-\textsc{ANNz2} is capable of producing both photo-$z$ point estimates and redshift posterior probability distributions $p(z)$, it could also conduct classifications and supports reweighting between samples.
-The PDFs are produced by propagating the intrinsic uncertainty on the input parameters and the uncertainty in the machine learning method to the expected photo-$z$ solution, averaged over multiple runs weighted based on the performance of each run.
-\textsc{ANNz2} presents its photo-$z$ uncertainty different from many codes by using the KNN method: it estimates the photo-$z$ bias between each object and a fixed number of nearest neighbours in parameter space, it then takes the $68$th percentile width of the distribution of the bias.
-This is based on the implication that objects with similar photometric properties would have similar uncertainties, and therefore the photometric errors of the inputs are not propagated into the code.
-
-In this study, \textsc{ANNz2 v.
-2.0.4} was used.
-The full PDF for each galaxy is also produced with a linear stepsize of $z=0.01$ for $0 \leq z \leq 2$.
-A set of $5$ ANNs with architecture $6:12:12:1$ ($6$ $ugrizy$ inputs, $2$ hidden layers with $12$ nodes each, and $1$ output) with different random seeds are used during each training.
-%Training was conducted on a subset of the given training sample which satisfies $i\le25.3$ as it would have improved the results compared to training on the entire sample.
-Half of the training set is used as a validation set to prevent overtraining.
-All training objects are set to have detected magnitudes, however the non-detections (mag$=-99$) in the testing set are replaced with the mean of that particular band.
-
-\subsubsection{Color-Matched Nearest-Neighbours}
-\label{sec:cmnn}
-%(Melissa Graham)
-
-The nearest-neighbours color-matching photometric redshift estimator (\textsc{CMNN}) is presented in \cite[][herafter G18]{Graham:17}.
-This method uses a training set of galaxies with known redshifts that has equivalent or better photometry as the test set in terms of quality and filter coverage.
-For each galaxy in the test set we identify a color-matched subset of training galaxies, choose one (e.g.
-the nearest-neighbour or a random selection), and use its known redshift as the photo-$z$.
-This color-matched subset is identified by first calculating the Mahalanobis distance $D_M$ in color-space between the test galaxy and all training-set galaxies: the difference between the test and a training set galaxy's color divided by the photometric error, summed over all colors (i.e., $u$-$g$, $g$-$r$, $r$-$i$, $i$-$z$, and $z$-$y$).
-Then, the threshold value for $D_M$ that define a good color match is set by the percent point function (PPF): for example, for $N_{\rm dof}=5$, PPF$=95$ per cent of all training galaxies consistent with the test galaxy will have $D_M < 11.07$ (where $N_{\rm dof}$, the number of degrees of freedom, is the number of colors).
-For a given test galaxy, the $p(z)$ is the normalized distribution of the true catalogue redshifts of this color-matched subset of training galaxies, and the standard deviation of the color-matched subset is used as the photo-$z$ uncertainty.
-
-We have applied the nearest-neighbours color-matching photometric redshift estimator described in G18 to the simulated data.
-Compared to its application in G18, there are some minor differences in the application of this estimator to the Buzzard catalogue.
-First, we do not impose non-detections on galaxies with a magnitude fainter than the expected LSST 10-year limiting magnitude or bright enough to saturate with LSST: {\it all} of the photometry for all the galaxies in the test and training sets are used for this experiment.
-Second, as in G18 we do apply an initial cut in color to the training set before calculating the Mahalanobis distance in order to accelerate processing, and also use a magnitude pseudo-prior to improve photo-$z$ estimates, but for both we have used different cut-off values that are appropriate for the Buzzard galaxies' colors and magnitudes.
-Third, we set different parameters for the identification of the color-matched subset of training galaxies and the selection of a photometric redshift estimate.
-In G18 we used a percent point function (PPF) value of 0.68 to identify the color-matched subset of training galaxies and used the redshift of nearest neighbour in color-space as the photo-$z$ estimate.
-These choices work well when the desire is to obtain accurate photo-$z$ estimates for most test-set galaxies, but does not return a robust $p(z)$ in all cases -- especially for galaxies that are bright and/or have few matches in color-space.
-Since a robust estimate of $p(z)$ is desired for this work we make several changes to our implementation of the \textsc{CMNN} photo-$z$ estimator.
-We continue to use a percent point function of ${\rm PPF} = 0.95$ to generate the subset of color-matched training galaxies, but weight them by the inverse of their Mahalanobis distance.
-This weighting maintains some of the accuracy that was previously achieved by simply using the nearest neighbour in color-space.
-We then use the weights to create the $p(z)$ instead of having the redshift of each color-matched training-set galaxy count equally.
-To obtain a robust estimate of the $p(z)$ for galaxies with a small number of color-matched training set galaxies, when this number is less than $20$ the nearest $20$ neighbours in color-space are used instead, and we convolve the $p(z)$ with a Gaussian with a standard deviation of:
-
-\begin{equation}
-\sigma = \sigma_{\rm train} \sqrt{({\rm PPF}_{20}/0.95)^2 -1}
-\end{equation}
-
-\noindent to appropriately broaden it so that the $p(z)$ for these test galaxies represents the enlarged PPF value associated with it.
-Overall, these three changes will yield poorer accuracy photo-$z$ compared to those presented in G18, but they will all have significantly more robust estimates of the $p(z)$, particularly for the brightest test galaxies.
-This is sufficient for this work because, as described in G18, the goal of the \textsc{CMNN} photo-$z$ estimator was never to provide the ``best'' (or even competitive) estimates in the first place, given its reliance on a deep training set, but rather to provide a means for direct comparisons between LSST photometric quality and photo-$z$ estimates.
-With this work we show how the input parameters should be set in order to return robust $p(z)$ estimates in addition to point value estimates.
-
-\subsubsection{Delight}
-\label{sec:delight}
-%(John Soo)
-
-\textsc{Delight}\footnote{\url{https://github.com/ixkael/Delight}} \citep{Leistedt:17} infers photo-$z$'s by using a data-driven model of latent SEDs and a physical model of photometric fluxes as a function of redshift.
-Generally, machine learning methods rely on representative training data with similar band passes, while template based methods rely on a complete library of templates based on physical models constructed.
-\textsc{Delight} is constructed in attempt to combine the advantages and eliminate the disadvantages of both template-based and machine learning algorithms: it constructs a large collection of latent SED templates (or physical flux-redshift models) from training data, with a template SED library as a guide to the learning of the model.
-The advantage of \textsc{Delight} is that it neither needs representative training data in the same photometric bands, nor does it need detailed galaxy SED models to work.
-
-This conceptually novel approach is done by using Gaussian processes operating in flux-redshift space.
-The posterior distribution on the redshift of a target galaxy is obtained via a pairwise comparison with training galaxies,
-
-\begin{equation}
-p(z|\mathbf{\hat{F}}) \approx \sum_i p(\mathbf{\hat{F}}|z,t_i)\, p(z|t_i)p(t_i),
-\end{equation}
-\noindent where $p(z|t_i)p(t_i)$ captures prior information about the redshift distributions and abundances of the galaxies, with $t_i$ denoting the galaxy template; while $p(\mathbf{\hat{F}}|z,t_i)$ is the posterior of noisy flux $\mathbf{\hat{F}}$ at redshift $z$.
-For each training-target pair, $p(\mathbf{\hat{F}}|z,t_i)$ is evaluated as follows:
-
-\begin{equation} \label{eq:delight_noisy}
-p(\mathbf{\hat{F}}|z,t_i) = \int p(\mathbf{\hat{F}}|\mathbf{F})\, p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)\, d\mathbf{F},
-\end{equation}
-where $p(\mathbf{\hat{F}}|\mathbf{F})$ is the likelihood function, it compares the noisy real flux $\mathbf{\hat{F}}$ with the noiseless flux $\mathbf{F}$ obtained from the linear combination of template models, carefully constructed to account for model uncertainties and different normalization of the same SED; while $p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)$ is the prediction of flux at a different redshift $z$ with respect to the training object with redshift $z_i$ and flux $\mathbf{\hat{F}}_i$.
-Eq.~\ref{eq:delight_noisy} is essentially the probability that the training and the target galaxies having the same SED but at a different redshift.
-The flux prediction $p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)$ of the training galaxy at redshift $z$ is modeled via a Gaussian process,
-
-\begin{equation} \label{eq:delight_gp}
-F_b \sim \mathcal{GP}\left( \mu^F,k^F \right),
-\end{equation}
-\noindent with mean function $\mu^F$ and kernel $k^F$, both imposed to capture expected correlations resulting from the known underlying physics (i.e., fluxes resulting from observing SEDs through filter response, and the SEDs being redshifted).
-The reader should refer to \citet{Leistedt:17} for further details.
-
-In this study, all $100$ ordered Buzzard templates, as described in Section~\ref{sec:buzztemplates}, were used in \textsc{Delight}, and the Gaussian process was trained with a subset of $50\,000$ galaxies.
-Photometric uncertainties from the inputs are propagated into the code, while non-detections for each band are set to the mean of the respective bands.
-Default settings of \textsc{Delight} were use, with the exception that the PDF bins were set to be linear instead of logarithmic, with $200$ equally-spaced bins between $0.0<z<2.0$.
-In this study a flat prior is assumed.
-
-\subsubsection{FlexZBoost}
-\label{sec:flexzboost}
-%(Ann Lee, Rafael Izbicki, Taylor Pospisil, Peter Freeman)
-
-\textsc{FlexZBoost}\footnote{\url{https://github.com/tpospisi/flexcode};\\ \url{https://github.com/rizbicki/FlexCoDE}} \citep{Izbicki:17} is a particular realization of \texttt{FlexCode}, which is a general-purpose methodology for converting any conditional mean point estimator of $z$ to a conditional {\em density} estimator $f(z \vert \x)$, where $\x$ here represents our photometric covariates and errors.\footnote{Instead of $p(z)$, we use the notation $f(z \vert \x)$ to explicitly show the dependence on $\x$.} The key idea is to expand the unknown function $f(z \vert \x)$ in an orthonormal basis $\{\phi_i(z)\}_{i}$:
-
-\begin{equation}
-f(z|\x)=\sum_{i}\beta_{i }(\x)\phi_i(z). \label{eq::series_expansion}
-\end{equation}
-By the orthogonality property, the expansion coefficients are just conditional means
-
-\begin{equation}
-\beta_{i }(\x) =  \mathbb{E}\left[\phi_i(z)|\x\right] \equiv \int f(z|\x)   \phi_i(z) dz. \label{eq:FlexZBoost_coeffs}
-\end{equation}
-These coefficients can easily be estimated from data by regression.
-
-In this paper, we use \textsc{xgboost}~\citep{Chen:16} for the regression part as these techniques scale well for massive data; it should however be noted that \textsc{FlexCode-RF} (also on GitHub), based on Random Forests, generally performs better for smaller data sets.
-As our basis, we choose a standard Fourier basis.
-There are two tuning parameters in our $p(z)$ estimate: (i) the number of terms, $I$, in the series expansion in Eq.~\ref{eq::series_expansion}, and (ii) an exponent $\alpha$ that we use to sharpen the computed density estimates $\widehat{f}(z|\x)$, according to $\widetilde{f}(z|\x) \propto \widehat{f}(z|\x)^\alpha$.
-We split the ``train data'' into a training set (85\%) and a validation set (15\%), and choose both $I$ and $\alpha$ in an automated way by minimizing the weighted $L_2$-loss function \citep[Eq.
-5 in][]{Izbicki:17} on the  validation set.
-
-Although {\tt FlexCode} offers a {\em lossless compression} of the photo-$z$ estimates (in this study, one can reconstruct $\widetilde{f}(z|\x)$ exactly at any resolution from estimates of the first 35 coefficients, Eq.~\ref{eq:FlexZBoost_coeffs}, for a Fourier basis $\{\phi_i(z)\}_{i}$), we discretize our final estimates into 200 bins linearly spaced in $0 < z < 2$ for easy comparison with other algorithms.
-Using a higher resolution may yield better results (with no added cost in storage).
-
-%\subsubsection{Frankenz}
-%\label{sec:frankenz}
-%(seeking volunteers)
-%
-%\red{\textsc{Frankenz}\footnote{\url{https://github.com/joshspeagle/frankenz}} \cite{Speagle:frankenz} is...}
-%
-
-\subsubsection{GPz}
-\label{sec:gpz}
-%(Ibrahim Almosallam)
-
-\textsc{GPz}\footnote{\url{https://github.com/OxfordML/GPz}} \citep{Almosallam:16a,Almosallam:15b} is a sparse Gaussian process based code, a fast and a scalable approximation of full Gaussian Processes \citep{Rasmussen:06}, with the added feature of being able to produce input-dependent variance estimations (heteroscedastic noise).
-The model assumes that the probability of the output $y$, the redshift, given the input $x$, the photometry, is $p(y|x)=\mathcal{N}\left(y|\mu(x),\sigma(x)^{2}\right)$.
-The mean function, $\mu(x)$, and the variance function $\sigma(x)^{2}$ are both linear combinations of basis functions that take the following form:
-
-\begin{align}
-f(x)=\sum_{i=1}^{m}\phi_{i}(x)w_{i},
-\end{align}
-where $\left\{\phi_{i}(x)\right\}_{i=1}^{m}$ and $\left\{w_{i}\right\}_{i=1}^{m}$ are sets of $m$ basis functions and their associated weights respectively.
-Basis function models (BFM), for specific classes of basis functions such as the sigmoid or the squared exponential, have the advantage of being universal approximators, i.e.
-there exist a function of that form that can approximate any function, with mild assumptions, to any desired degree of accuracy.
-The details on how to learn the parameters of the model and the hyper-parameters of the basis functions are described in \citet{Almosallam:15b}.
-
-A unique feature in \textsc{GPz}, is that the variance estimate is composed of two terms each quantifying a different source of uncertainty.
-One term (the model uncertainty) reflects how much of the uncertainty is due to lack of training samples at the location of interest, whereas the second term (the noise uncertainty) reflects how much of the uncertainty is caused from observing many noisy samples at that location.
-Thus, the predictive variance can determine whether we need more representative samples or more precise samples for any particular location in the input space.
-\textsc{GPz} can also emphasize the importance of some samples as weights.
-This weight can be for example $|z_{\rm{spec}}-z_{\rm{phot}}|/(1+z_{\rm{spec}})$ to target the desired objective of minimizing the normalized redshift error or as a function of their probability in the test set relative to the training set in order to pressure the model to better fit samples that are rare in the training set but are expected to be abundant during testing.
-
-The data is prepared for \textsc{GPz} by taking the log of the magnitude errors, decorrelating the data set using PCA and imputing the missing values using a simple linear model that estimates the missing variables given the observed ones.
-The log transformation helps to smooth the long tail distribution of the magnitude errors, which is more stable numerically and makes the optimization process unconstrained.
-The missing values are imputed by computing the mean of the training set $\mu$ and its covariance $\Sigma$, then we use the following equation to estimate the missing values from the observed ones %\[
-
-\begin{equation}
-x_{u} = \mu_{u}+\Sigma_{uo}\Sigma_{oo}^{-1}(x_{o}-\mu_{o}),
-\end{equation}  %\]
-where the subscript $o$ in $x_{o}$ indexes the \emph{observed} part of the input $x$, whereas the subscript $u$ indexes the \emph{unobserved} set (similarly for $\mu$ and $\Sigma$).
-This is the optimal expected value of the unobserved variables given the observed ones if the distribution is jointly Gaussian, note that if the variables are independent, i.e.
-$\Sigma_{uo}=0$, this will reduce to a simple average predictor.
-
-We use the Variable Covariance (VC) option in \textsc{GPz} with 200 basis functions after we note that there is no significant increase in the performance on the validation set (using 80\%-20\% training-validation split) and with no cost-sensitive learning applied.
-We do note that \textsc{GPz} is trained using a data set that includes galaxies fainter than those in the test set, which was truncated at $i<25.3$.
-This mismatch in training vs.~test results in a slightly different distribution in galaxy photometry and redshift distribution between the training and test set, which most likely degraded results compared to a run where the training set more closely matched the test set.
-
-\subsubsection{METAPhoR}
-\label{sec:metaphor}
-%(Stefano Cavuoti, Massimo Brescia, Giuseppe Longo)
-%\textcolor{brown}{general review done on this section}
-\textsc{METAPhoR} \citep[Machine-learning Estimation Tool for Accurate Photometric Redshifts,][]{Cavuoti:17} is a pipeline designed to provide photo-$z$'s point estimates and a reliable PDF for machine learning (ML) based techniques.
-It includes pre- and post-processing phases, hosting a photo-$z$ prediction engine based on the Multi Layer Perceptron with Quasi Newton Algorithm (MLPQNA), already validated on photo-$z$'s in several cases \citep{de_Jong:17,Cavuoti:17b,Cavuoti:15,Brescia:14,Brescia:13,Biviano:13}.
-Due to its plug-in based modular nature, \textsc{METAPhoR} can be easily replaced by any other photo-$z$ prediction kernel, regardless its implementation, by taking the I/O interface compliance as unique constrain.
-
-At a higher level, the pipeline mainly consists of three modules: (i) \textit{data pre-processing}, including a catalogue cross-matching sub-module \citep[based on the tool C3, ][]{Riccio:17}, a sub-module for photometric evaluation and error estimation of the multi-band catalogue used as Knowledge Base (KB), and a sub-module dedicated to the perturbation of the photometric KB, propaedeutic to the PDF estimation; (ii) \textit{photo-$z$ prediction}, which is the training/validation/test phase, producing the photo-$z$'s point estimates, based on a pre-selected ML method; (iii) \textit{PDF estimation}, specifically designed to calculate the PDF of the photo-$z$ estimation errors.
-The last module includes also a post-processing tool, providing some statistics on the produced point estimates and PDFs.
-
-The photometry perturbation law is based on the formula $m_{ij} = m_{ij} + \alpha_{i}F_{ij}*u_{\mu=0,\sigma=1}$, where $\alpha_{i}$ is a user selected multiplicative constant (useful in case of multi-survey photometry), $u_{\mu=0,\sigma=1}$ is a random value from the standard normal distribution and $F_{ij}$ is a bimodal function (a constant function + polynomial fitting of the mean magnitude errors on the binned bands), heuristically tuned in such a way that the constant component is the threshold under which the polynomial function is considered too low to provide a significant noise contribution to the photometry perturbation.
-
-As introduced, the photo-$z$ point estimate prediction engine of \textsc{METAPhoR} is based on the MLPQNA model, whose photo-$z$ regression training error, used by the quasi Newton learning rule, is based on the least square error and Tikhonov $L_{2}$-norm regularization \citep{Hofmann:18}.
-
-As main prerogative, \textsc{METAPhoR} is able to provide a PDF for ML methods by taking into account the photometric errors provided with data, by running $N$ trainings on the same training set, or $M$ trainings on $M$ different random extractions from the KB.
-The different test sets, used to produce the PDF, are thus obtained by introducing a proper perturbation, parametrized from the photometric error distribution in each band, on the photometric data populating the original test set \citep{Brescia:18}.
-
-For the present work since it was required to produce a redshift (and a PDF) for each object of the test set we decided to apply a hierarchical kNN to fill the missing detection, it goes without saying that for such points the reliability of PDFs and point estimation is lower.
-No cross validation has been used.
-
-\subsubsection{SkyNet}
-\label{sec:skynet}
-%(J. Cohen-Tanugi and Hugo Tranin)
-
-\textsc{SkyNet}\footnote{\url{http://ccpforge.cse.rl.ac.uk/gf/project/skynet/}} \citep{Graff:14} is a publicly available neural network software, based on a 2nd order conjugate gradient optimization scheme \citep[see][for further details]{Graff:14}.
-It has been used efficiently for redshift PDF estimates \citep{Sanchez:14,Bonnett:15,Bonnett:16}.
-
-%In the present work, we use \textsc{SkyNet} both as a regressor (for redshift point estimates) and a classifier (for PDF estimates). 
-The neural network is configured as a standard multilayer perceptron with three hidden layers and one input layer with $12$ nodes (the $6$ magnitudes and their errors). 
-%For the regressor, the output layer is a single linear node, while each hidden layer has $10$ nodes with a tanh activation function. 
-The classifier is laid out such that the hidden layers have 20:40:40 nodes each, all rectified linear units, and the output layer has $200$ nodes (corresponding to 200 bins for the PDF) activated with a ``softmax'' function so that they automatically sum to $1$.
-
-To avoid over-fitting, a $30$ per cent fraction of the training set is used as validation, and the training is stopped as soon as the error rate begins to increase in the validation set.
-The weights are randomly initialized based on normal sampling.
-The error function is a standard chi-square function for the regressor, and a cross-entropy function for the classifier.
-Finally, the data are all whitened before processing, with magnitudes pegged to (45,45,40,35,42,42) and their errors pegged to (20,20,10,5,15,15) for $ugrizy$ filters, respectively.
-
-\subsubsection{TPZ}
-\label{sec:tpz}
-%(Erfan Nourbakhsh)
-
-\textsc{TPZ}\footnote{\url{https://github.com/mgckind/MLZ}} \citep[Trees for Photo-$z$,][]{Carrasco_Kind:13,Carrascokind:14} is a parallel machine learning algorithm that generates photometric redshift PDFs using prediction trees and random forest techniques.
-The code recursively splits the input data (i.~e.~the training sample), into two branches, one after another, until a terminal leaf is created that meets a termination criterion (e.~g.~a minimum leaf size or a variance threshold).
-Bootstrap samples from the training data and associated errors are used to build a set of prediction trees.
-In order to minimize correlation between the trees, the data is divided in such a way that the highest information gain among the random subsample of features is obtained at every point.
-The regions in each terminal leaf node corresponds to a specific subsample of the entire data that possesses similar properties.
-
-The training data is examined before running TPZ.
-Since TPZ does not handle non-detections (magnitudes flagged as 99.0), we replace these values with an approximation of the $1\sigma$ detection threshold, i.~e.~a signal to noise ratio of 1 in terms of magnitude uncertainty using the equation $dm = 2.5 ~ \log ( 1 + N/S )$ where $dm \sim 0.7526 ~ mag$ for $N/S=1$.
-That is, for each band, we replace the non-detection with the magnitude corresponding to the error of 0.7526 from the error model forecasted for 10-year LSST data.
-The Out-of-Bag \citep{Breiman:84,Carrasco_Kind:13} cross-validation technique is used within TPZ to evaluate its predictive validity and determine the relative importance of the different input attributes.
-We employed this information to calibrate our algorithm.
-
-In the present work, the LSST magnitudes $u,~g,~r,~i$ and colors $u-g,~g-r,~r-i,~i-z,~z-y$ and their associated errors are used in the process of growing 100 trees with a minimum leaf size of 5 (the $z$ and $y$ magnitudes did not show significant correlation with the redshift in our cross-validation, so we did not use them when constructing our trees).
-We partitioned our redshift space into 100 bins from $z=0.005$ to $z=2.0$ and smoothed each individual PDF with a smoothing scale of twice the bin size.
-
-\subsection{Simple Ensemble Estimator}\label{sec:method:trainz}
-In addition to the main photo-$z$ algorithms described above we also include a very simple method.  For \trainz, as we will we call this simple estimator, we well define $p(z)$ as simply:
-
-\begin{equation}
-p(z) = \frac{1}{N_{train}}\sum_{i=1}^{N_{train}}z_{train}
-\end{equation}
-That is, we simply set the redshift PDF of every galaxy equal to the normalized $N(z)$ of the training sample.
-As the training sample is drawn from the same underlying distribution as the test sample, modulo small deviations due to sample size, the quantiles of the training and test distributions should be identical.
-This is a wildly unrealistic estimator, as it assigns all galaxies, no matter their apparent magnitude, colour, or true redshift, the same redshift PDF, and is thus uninformative at the level of individual object redshifts, but is designed to perform very well for the ensemble of all objects.
-We will discuss this method and cautions relative to metrics in Section~\ref{sec:caution}.
-
-\section{Metrics for quantifying PDF comparisons}
-\label{sec:metrics}
-%(Alex Malz, Rongpu Zhou, Jeff Newman, Ofer Lahav)
-
-%\red{Maybe it would be best to introduce the summary statistics (KS, CvM, AD) up front and then the ensemble functions and stacked n(z).}
-
-The overloaded ``$p(z)$'' is a widespread abuse of notation; we would like the outputs of photo-$z$ PDF codes to be interpretable as probabilities. 
-Obviously photo-$z$ PDFs must not take negative values and must integrate to unity over the range of possible redshifts. 
-Additionally, an estimator derived by method $H$ for the photo-$z$ PDF of galaxy $i$ must be understood as a posterior probability distribution
-
-\begin{equation}
-\hat{p}_{j}(z_{i}) = p(z | d_{i}, I_{D}, I_{H}),
-\end{equation}
-conditioned not only on the photometric data $d_{i}$ for that galaxy but also on parameters encompassing a number of things that will differ depending on the method $H$ used to produce it, namely the assumptions $I_{H}$ necessary for the method to be valid and any inputs $I_{D}$ it takes as prior information, such as a template library or training set.
-Because of this, direct comparison of photo-$z$ PDFs produced by different methods is in some sense impossible; even if they share the same prior information $I_{D}$, by definition they cannot be conditioned on the same assumptions $I_{H}$, otherwise they would not be distinct methods at all.
-
-In this study, we isolate the differences in prior information specific to each method by using a single training set $I_{D}^{ML}$ for all machine learning-based codes and a single template library $I_{D}^{T}$ for all template-based codes, and these sets of prior information are carefully constructed to be representative and complete, we have $I_{D}^{ML}\equiv I_{D}^{T}$ for every method $H$.
-Thus, we are saying
-
-\begin{equation}
-\frac{\hat{p}_{i, H}(z)}{\hat{p}_{i, H'}(z)} \approx \frac{p(z | d_{i}, I_{H})}{p(z | d_{i}, I_{H'})},
-\end{equation}
-meaning that we assume comparisons of $\hat{p}_{i, H}(z)$ isolate the effect of the method used to obtain the estimator, which should make examination of differences caused by specifics of the method implementations easier to isolate.
-
-As mentioned previously, there are cosmology probes that require knowledge of individual galaxy $p(z)$ and others that require only knowledge of the ensemble redshift distribution, $N(z)$.  
-Due to the paucity of principled techniques for using and validating photo-$z$ PDFs, there are few alternatives to the common practice of reducing photo-$z$ PDFs to point estimates. 
-Though this practice should not be encouraged, we also calculate traditional metrics based on the most common point estimators derived from photo-$z$ PDFs.  
-Those seeking to establish a connection to traditional ways of thinking about redshift estimation may consult the Appendix for these results.
-
-There are a number of metrics that can be used to test the accuracy of a photo-$z$ interim posterior as an estimator of a true photo-$z$ posterior if it is known.
-Even for simulated data, the true photo-$z$ PDF is in general not accessible unless the redshifts are in fact drawn from the true photo-$z$ PDFs, a mock catalogue generation procedure that has not yet appeared in the literature.
-Furthermore, only limited applications of photo-$z$ PDFs that could be used as the basis for a metric have been presented in the literature. 
-The most popular application by far is the calculation of the overall redshift distribution $N(z)$, the true value of which is known for the \textsc{Buzzard} simulation and will be denoted as $N'(z)$.
-Though alternatives exist \citep{Malz:chippr}, stacking according to
-
-\begin{equation} \label{eq:stacked}
-\hat{N}^{H}(z) \approx \frac{1}{N_{tot}}\ \sum_{i}^{N_{tot}}\ \hat{p}^{H}_{i}(z)
-\end{equation}
-is the most widely accepted method for estimating the redshift distribution from photo-$z$ PDFs. 
-If we assume that the response of estimators of $N(z)$ is uniform across all approaches $H$, then we may interpret metrics on the accuracy of $\hat{N}(z)$ obtained in this way. 
-We must note, however, that this is a poor assumption in general. 
-Under the setup of this paper, the true redshift distribution $N'(z) = p(z | I_{D})$ (i.e.
-because our training data is representative, the interim prior is the truth).
-In this ideal case, the method that would give the best approximation to $N'(z)$ would be one that neglects all the information contained in the photometry $\{d_{i}\}_{N_{tot}}$ and gives every galaxy the same photo-$z$ PDF $\hat{p}_{i}(z) = N'(z)$ for all $i$.
-This is the exact estimator, \trainz, that we have described in Section~\ref{sec:method:trainz}, and which will serve as a point of reference for the other codes.
-
-The exact implementation of the stacked estimator $\hat{N}^{H}(z)$ will depend on the parametrization of the photo-$z$ PDFs, which may differ across codes and can affect the precision of the estimator \citep{Malz:qp}; even considering a single method under the same parametrization, say a piecewise constant function over bins or a set of samples from the posterior, an estimator using $2N$ bins or samples will trivially be more precise than an estimator using $N$ bins or samples.
-%\red{Note: Let's address how the gridding (endpoints and spacing) affects the metrics, from the email correspondence Re: [LSST-DESC-MEMBERS] Summary of Nov 29 PZ Telecon. (AIM)}
-In order to minimize the effects of such choices, we asked those running all eleven codes to output $p(z)$ parameterized with a generous $\approx 200$ piecewise constant bins spanning $0<z<2$.  
-The piecewise constant format is chosen because of its established presence in the literature, and the choice of 200 bins was motivated by the approximate number of columns expected to be available for storage of $p(z)$ for the final LSST Project tables.\footnote{See, e.~g.~the LSST Data Products Definition Document, available at: \url{https://ls.st/dpdd}}
-All $p(z)$ catalogues are processed using the \textsc{qp} software package \citep{Malz:qp}\footnote{available at: \url{http://github.com/aimalz/qp/}} for manipulating and calculating metrics of 1-dimensional PDFs.
-We will discuss the choice of $p(z)$ parameterization further in Section~\ref{sec:results}.
-
-%\red{Divide metrics by application to ensemble of PDFs vs. estimators of the redshift distribution}
-
-\subsection{Metrics of an ensemble of photo-\mathinhead{$z$}{z} interim posteriors}
-\label{sec:qualmet}
-
-\subsubsection{Probability integral transform (PIT)}
-\label{sec:pit}
-%(Alex Malz, Peter Freeman, Sam Schmidt)
-
-The probability integral transform (PIT) \citep{Polsterer:16} is defined for each individual galaxy as:%essentially the derivative of the QQ plot, a histogram of
-
-\begin{align}
-\label{eq:pit}
-\mathrm{PIT} &= \int_{-\infty}^{z_{\rm true}}p(z)\, dz \,.
-\end{align}
-The distribution of PIT values quantifies the behavior of the {\it ensemble} of photo-$z$ PDFs, enabling us to evaluate whether the $p(z)$ is, on average, accurate:  The PIT value is the Cumulative Distribution Function (CDF) of the $p(z)$ evaluated at the true redshift.
-A catalogue of photo-$z$ PDFs that are accurate should have a flat PIT histogram (i.e., the individual PIT values as samples from each CDF should match a Uniform(0,1) distribution if the CDFs are accurate).
-Specific deviations from flatness indicate inaccuracy: overly broad photo-$z$ PDFs would manifest as underrepresentation of the lowest and highest PIT values, whereas overly narrow photo-$z$ PDFs would manifest as over-representation of the lowest and highest PIT values.
-High frequency at only $\mathrm{PIT}\approx0$ and $\mathrm{PIT}\approx1$ indicates the presence of catastrophic outliers with highly inaccurate photo-$z$ PDFs where the true redshift is outside of the support of $p(z)$.
-\citet{Tanaka:17} use the histogram of PIT values as a diagnostic indicator of overall code performance, while \citet{Freeman:17} independently define the PIT and demonstrate how its individual values may be used both to perform hypothesis testing (via, e.g., the KS, CvM, and AD tests; see below) and to construct quantile-quantile plots.
-%\red{We might want to divide this by redshift to get trends, as the overall PIT could be impacted by true n(z).  
-%For example, a PIT with nontrivial shape could be due to that behavior being uniform as a function of redshift or significant inaccuracy only in a small range of redshifts where $n(z)$ is large and a flat PIT elsewhere.}
-
-\subsubsection{Quantile-quantile (QQ) plot}
-\label{sec:qq}
-%(Rongpu Zhou, Jeffrey Newman, Alex Malz)
-
-The quantile-quantile (QQ) plot is a graphical method for comparing two distributions, where the quantiles of one distribution are plotted against the quantiles of the other distribution (A quantile being defined by partitioning a distribution into consecutive intervals containing equal amounts of probability, or equal numbers of objects in each interval).
-%\blue{Include an equation to ensure quantile definition is crystal clear before building upon it for PIT.}\red{Is maybe the above text clear enough?
-Text may be more clear than any equation I can come up with--SJS.}
-In this paper we show the quantiles of the PIT values compared to the quantiles of the Uniform distribution that we expect the PIT values to match if $p(z)$ is an accurate probability distribution for all objects.
-The QQ plot provides an easy way to qualitatively assess the differences in various properties such as the moments of an estimating distribution relative to a true distribution.  
-%location, scale and skewness.  
-In this paper, QQ plots are used for two purposes: (1) for comparing $N(z)$ from photo-$z$ PDFs (estimated using Eq.~\ref{eq:stacked}) with the true $N(z)$, i.e.
-comparing the estimated distribution of redshifts with the true redshift distribution, and (2) for assessing the overall consistency of an ensemble of photo-$z$ PDFs with their true redshifts on a population level, where the distribution of the PIT values (see previous section) is compared to a uniform distribution between $0$ and $1$.
-The QQ plot contains very similar information to that shown in the PIT histogram plot, we include both forms, as visually they each convey the information in a somewhat distinct manner.
-%\red{is this enough reason to include both?--SJS}
-%\aim{Maybe we should explain why we expect the distribution of CDFs to be uniform.}\red{Note: this is just describing the PIT.
-I'm going to swap the order of PIT to come before this, and then say we are evaluating the distribution of PIT values.
-I will also add why we expect this to be uniform--SJS}
-
-\subsubsection{Conditional density estimation loss}
-\label{sec:CDE_loss}
-%(Alex Malz, Ann Lee)
-
-With the conditional density estimation loss (CDE loss) we can compare how well different methods estimate individual PDFs for photometric covariates $\x$ rather than looking only at the ensemble distribution.
-As in Section~\ref{sec:flexzboost}, we use the notation $f(z \vert \x)$ instead of $p(z)$ to explicitly show the dependence on $\x$.
-
-%\aim{Perhaps this should be rewritten in the same notation used elsewhere in the paper.} \red{I agree, does someone who speaks both stats and astro want to take a shot at this?}
-%\textcolor{cyan}{Ann: We are looking at individual PDFs so the dependence on $\x$ needs to be explicit; this is probably the most accurate notation as it shows both the conditioning clearly as well as whether $\x$ is fixed or random.}
-
-The CDE loss is defined as:
-\begin{equation} \label{eq:cde-loss}
-L(f, \widehat{f}) = \int \int (f(z \mid \x) - \widehat{f}(z \mid \x))^{2} dz dP(\x)
-\end{equation}
-This loss is the CDE equivalent of the RMSE in regression. To estimate this loss we rewrite the loss as
-\begin{equation} \label{eq:estimated-cde-loss}
-\mathbb{E}_\X \left[\int \widehat{f}(z \mid \X)^{2} dz\right] - 2 \mathbb{E}_{\X,Z}\left[\widehat{f}(Z \mid \X)\right] + K_{f},
-\end{equation}
-where the first expectation is with respect to the marginal distribution of the covariates $\X$, the second expectation is with respect to the joint distribution of $\X$ and $Z$, and $K_{f}$ is a constant depending only upon the true conditional densities $f(z\mid \x)$.
-For each method we can estimate these expectations as empirical expectations on the test or validation data \citep[Eq.~7 in][]{Izbicki:17b} without knowledge of the true densities.
-
-%\subsubsection{Continuous Ranked Probability Score (CRPS)}
-%\label{sec:crps}
-%\red{The CRPS is used to assess the accuracy of two probabilistic models respectively.}
-
-\subsection{Metrics over estimated probability distributions}
-\label{sec:quantmet}
-
-In tandem with the QQ and PIT metrics introduced above, we additionally compute the following metrics comparing the empirical CDF of a distribution to the true or expected distribution.
-These metrics give a more quantitative measure of the departure from ideal than the more visual PIT histogram and QQ plot.
-We compute metrics comparing the CDF of PIT values to a the CDF of a Uniform distribution, and also compute the CDF of the true redshift distribution $N'(z)$ compared the $\hat{N}(z)$ distribution derived from summing the $p(z)$ as described in Eq.~\ref{eq:stacked}.
-
-\subsubsection{Root-mean-square error (RMSE)}
-\label{sec:rmse}
-%(Alex Malz)
-
-We employ the familiar root-mean-square error:
-
-\begin{equation}
-\mathrm{RMSE} = \sqrt{\int_{-\infty}^{\infty}\left(\hat{f}(z)-f'(z)\right)^{2}\,dz},
-\end{equation}
-%may be used for two purposes in this analysis. 
-%We calculate it between the true redshift distribution $N'(z)$ and each estimator $\hat{N}(z)$ (derived by Eq.~\ref{eq:stacked}) as well as between the ideal QQ and PIT curves from each method studied and the ideal curves.
-Though this metric does not account for the fact that the redshift distribution function is, in fact, a probability distribution, it can still be interpreted as a measure of the integrated difference between the estimated distribution and the true distribution, and it can be used to quantify the otherwise qualitative metrics.
-
-\subsubsection{Kolmogorov-Smirnov (KS) and related statistics}
-\label{sec:ks}
-%(John Soo, Rongpu Zhou)
-
-The \textit{Kolmogorov-Smirnov statistic} $N_{\rm KS}$ is the maximum difference between $F_{\rm phot}(z)$ and $F_{\rm spec}(z)$, the CDFs of the photo-$z$ and spectroscopic redshift respectively:
-
-\begin{equation}
-N_{\rm KS} = \max_{z}\left( \left| F_{\rm phot}(z) - F_{\rm spec}(z) \right| \right).
-\end{equation}
-%The KS test quantifies whether the two redshift distributions are drawn from a similar distribution, independent of binning. 
-The KS test quantifies the similarity between two distributions, independent of binning. 
-A lower $N_{\rm KS}$ value corresponds to more similar distributions.
-%We calculate two KS test statistics, one to test the distribution of $p(z)$ values, and one to test the stacked estimator $\hat{n}(z)$ of the redshift distribution.  
-%The $p(z)$ test compares the CDF of the PIT values described in the previous section to that of a uniform distribution.  
-%The $n(z)$ test compares the CDF of the distribution formed by stacking the $p(z)$ for every object in our test set against the CDF of the corresponding true redshifts.
-%In this case a lower $N_{\rm KS}$ value means that the photo-$z$ distribution fits the spectroscopic redshift distribution well.
-$F_{\rm phot}(z)$ is calculated by stacking the PDFs of every single object in the sample, producing a smooth cumulative $n(z)$ to be compared with $F_{\rm spec}(z)$.
-
-We also consider two variants of the KS statistic: the Cramer-von Mises (CvM) and Anderson-Darling (AD) statistics.
-The CvM statistic is similar to the KS statistic as it is also computed from the distance between the measured CDF and the ideal CDF, but instead of the maximum distance, the CvM statistic calculates the average of the distance squared:
-
-\begin{equation}
-\label{eq:cvm}
-\omega^2 = \int_{-\infty}^{+\infty}\big(F_{\mathrm{meas.}}(x) - F_{\mathrm{ideal}}(x)\big)^2\mathrm{d}F_{\mathrm{ideal}}
-\end{equation}
-% In this paper we use the CvM statistic to evaluate the performance of both $n(z)$ and $p(z)$.
-The AD statistic is a weighted version of the CvM statistic, making it more sensitive to the tails of the distribution:
-
-\begin{equation} \label{eq:ad}
-A^2 = n\int_{-\infty}^{+\infty} \frac{\big(F_{\mathrm{meas.}}(x) - F_{\mathrm{ideal}}(x)\big)^2} {F_{\mathrm{ideal}}(x) (1-F_{\mathrm{ideal}}(x))}\mathrm{d}F_{\mathrm{ideal}}
-\end{equation}
-where $n$ is the sample size. 
-%The KS statistic and the CvM and AD statistics related to it are all used to evaluate the performance of both $\hat{n}(z)$ and the QQ and PIT of the CDFs of the ensemble of photo-$z$ interim posteriors.
-
-%\subsubsection{Kullback-Leibler Divergence (KLD)}
-%\label{sec:kl}
-%(Alex Malz)
-%
-%The KLD is a measure of the loss in information when using a probability distribution that is an estimate of a known true probability distribution, so we use 
-%\begin{equation}
-%\mathrm{KLD} = \int_{-\infty}^{\infty}n'(z)\log\left[\frac{n'(z)}{\hat{n}(z)}\right]dz
-%\end{equation}
-%as a metric on the stacked redshift distribution estimator of Eq.~\ref{eq:stacked}.
-Noting the assumptions made in using Eq.~\ref{eq:stacked}, we say that a photo-$z$ PDFs method whose stacked redshift distribution estimator $\hat{n}(z)$ more closely resembles the true redshift distribution $n'(z)$ is desired, i.e.
-we seek to minimize the KLD.
-
-\subsubsection{Moments}
-\label{sec:moments}
-%(Alex Malz)
-For the $\hat{N}(z)$ distributions we additionaly calculate the first three moments of the estimated redshift distribution for each code and compare them to the moments of the true redshift distribution $N'(z)$.
-The $m$th moment of a distribution is defined as
-
-\begin{equation}
-\langle z^{m}\rangle = \int_{-\infty}^{\infty}z^{m}N(z)dz.
-\end{equation}
-Here, we use the moments of the stacked estimator of the redshift distribution function as the basis for a metric.
-The closer the moments of $\hat{N}(z)$ for a photo-$z$ PDF method are to the moments of the true redshift distribution function $N'(z)$, the better the photo-$z$ PDF method.
-
-\section{Results}
-\label{sec:results}
-%(Sam Schmidt, Bryce Kalmbach, Johann Cohen Tanugi, Rongpu Zhou)
-
-\subsection{Ensembles of photo-\mathinhead{$z$}{z} interim posteriors}
-
-\begin{figure*}
-%\centering
-\begin{subfigure}[b]{0.49\textwidth}
-\includegraphics[width=\textwidth]{figures/pz_12codes_261931_crop.jpg}
-\end{subfigure}
-\begin{subfigure}[b]{0.49\textwidth}
-\includegraphics[width=\textwidth]{figures/pz_12codes_471167_crop.jpg}
-\end{subfigure}
-\begin{subfigure}[b]{0.49\textwidth}
-\includegraphics[width=\textwidth]
-{figures/pz_12codes_713178_crop.jpg}
-\end{subfigure}
-\begin{subfigure}[b]{0.49\textwidth}
-\includegraphics[width=\textwidth]{figures/pz_12codes_982747_crop.jpg}
-\end{subfigure}
-\caption{Four illustrative examples of individual p(z) distributions produced by the codes.
-The red vertical line represents the true redshift.
-Examples are chosen with common features seen in PDFs: tight unimodal $p(z)$ (upper left), broad unimodal $p(z)$ (upper right), bimodal $p(z)$ (lower right), and complex/multimodal $p(z)$ (lower left).
-Codes show varying amounts of small-scale structure in their reconstruction of the posterior distribution.
-We see varying responses from the codes in the presence of color degeneracies and photometric errors, resulting in narrow and broad unimodal, bimodal, and multi-modal $p(z)$ curves.} \label{fig:pz_examples}
-\end{figure*}
-
-%\begin{figure*}
-%\centering
-%\includegraphics[width=\textwidth]{figures/QQplot_11codes.jpg}
-%\caption{The QQ plots produced for every photo-$z$ code.
-\aim{Consider making this a plot of QQ minus that diagonal for better contrast between methods.}\red{The PIT histogram is essentially this, which is one of the main reasons to include both.
-QQ is a fairly standard form, so plotting in this fashion rather than the difference seems better to me, given that we also ahve the PIT histograms.
---SJS}} \label{fig:qq}
-%\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{figures/PITANDQQplot_12codes_crop.jpg}
-\caption{Summary plots for all eleven photo-$z$ codes illustrating performance for the interim posterior statistics.
-The top panel of each pair shows both the Quantile-Quantile (QQ) plot (red) and the histogram of PIT values (blue).
-The desired behavior is a QQ plot that matches the diagonal dashed line, and a PIT histogram that matches a uniform distribution matching the thin horizontal black line.
-The bottom panel of each pair shows the difference between the QQ quantile and the diagonal, illustrating departure from the desired performance.
-Histograms with an overabundance of PIT values at the centre of the distribution indicate $p(z)$ distributions that are overly broad, while an excess of values at the extrema indicate $p(z)$ distributions that are overly narrow.
-Values of PIT=0 and PIT=1 indicate ``catastrophic failures'' where the true redshift is completely outside the support of $p(z)$.
-Asymmetric features are indicative of systematic bias in the redshift predictions.
-A variety of behaviors are evident, and specific details are discussed in the text.}
-\label{fig:pitqq}
-\end{figure*}
-
-Fig.~\ref{fig:pz_examples} Shows the $p(z)$ produced by each of our eleven photo-$z$ codes for four example galaxies which exemplify some prominent cases that arise when estimating photo-$z$ PDFs: a narrow, unimodal redshift solution, a broader unimodal solution, a bimodal distribution, and a complex, multimodal distribution.
-The red vertical line represents the true redshift of the individual galaxy, and the blue curve represents the redshift probability.  
-Several features are obvious even in these illustrative examples.  
-\textsc{ANNz2, METAPhoR, NN,} and \textsc{SkyNet} all show an excess of small-scale features, which appear to be print-through of the underlying training set galaxies.  
-\textsc{GPZ} (in its current implementation), on the other hand, always produces a single Gaussian, which broadens to cover the multi-modal redshift solutions seen in other codes.
-
-As stated in Section~\ref{sec:metrics}, $p(z)$ is parameterized as $\approx 200$ piecewise constant bins covering $0<z<2$ for all eleven codes, giving a grid size of roughly $\delta z = 0.01$ for each code.
-A piecewise constant grid was a natural choice for some photo-$z$ codes, for instance most template-based codes compute likelihoods on a fixed grid.  
-In contrast, FlexZBoost, for example, can return estimates on any grid without compression errors as it’s a basis expansion method where only the expansion coefficients need to be stored.
-Codes with a native output format other than the shared piecewise constant binning scheme (or one that can be losslessly converted to it) may suffer from loss of information when converting to it, which could artificially favor some codes over others.
-
-Furthermore, the fidelity of photo-$z$ interim posteriors in this format varies with the quality of the photometry.  
-For faint galaxies, this redshift resolution is sufficient to capture the shape of $p(z)$ for the majority of the test sample, where photometric errors on the faint galaxies lead to somewhat broad peaks in the redshift posterior.
-However, as can be seen in e.~g.~the top left panel of Fig.~\ref{fig:pz_examples}, for bright galaxies with narrow $p(z)$ the grid spacing of $\delta z = 0.01$ is not sufficient to resolve the peak.  
-This is consistent with the results described in \citet[]{Malz:qp}, who find that quantiles (and, to a lesser degree, samples) often outperform gridded $p(z)$, particularly for bright objects and in the presence of harsher storage constraints.
-With a full 200 numbers to capture the information of each photo-$z$ PDF, any parametrization will perform adequately, but other storage parametrizations and limits on storage resources may be considered in future work.
-%Switching to a quantile based parameterization may be more costly computationally, for example template-based codes would need to test more grid point in order to resolve the quantiles for bright galaxies.
-However, the computational time for template based codes scales roughly linearly with the number of grid points, so this may be a reasonable thing to implement.
-We will discuss this further in Section~\ref{sec:discussion}.
-% \red{someone review this statement to make sure that I'm saying this correctly!}
-
-Fig.~\ref{fig:pitqq} shows both the quantile-quantile plots (red) and the histogram of PIT values (blue) summarizing the results from each photo-$z$ code.
-The red line shows the measured quantiles, while the black diagonal represents the ideal QQ values if the distribution were perfectly reproduced.
-A second panel below the main panel for each code shows the difference between $Q_{data}$ and $Q_{theory}$, i.~e.~the departure from the diagonal, for clarity.
-Biases and trends in whether the average width of the $p(z)$ values being over/under-predicted are evident.
-An overall bias where the predicted redshift is systematically low manifests as the measured QQ value falling above the diagonal, as is the case for \textsc{BPZ} and \textsc{EAZY}, while a systematic overprediction shows up as the measured QQ value falling below the diagonal, as seen in \textsc{TPZ}.
-In terms of PIT histograms, a systematic underprediction of redshift corresponds to fewer PIT values at $PIT<0.5$ and more at $PIT>0.5$, while a systematic overprediction will show the opposite.
-
-Examination of the PIT histograms and QQ plots shows that there are fairly generic issues with the width of $p(z)$ uncertainties: \textsc{Delight, NN, SkyNet} and \textsc{TPZ} all show a PIT histogram with an dearth of low values and and an excess of high values, signs that, on average, their $p(z)$ are more broad than the true distribution of redshifts.
-\textsc{METAPhoR} shows the opposite trend, indicating the the $p(z)$  are more narrow than the distributions given by the true redshifts.
-In all of these code cases there is a free parameter or bandwidth that can be used to tune uncertainties.
-The sensitivity of multiple codes to this bandwidth choice emphasizes the fact that great care must be taken in setting user-defined parameters in photo-$z$ codes, even in the presence of representative training/validation data.
-for \textsc{FlexZBoost} the ``sharpening'' parameter (described in Section~\ref{sec:flexzboost}) plays a key role in improving the results, resulting in a QQ plot that is very nearly diagonal.
-A similar sharpening procedure could be beneficial for several codes.
-Interestingly, the three purely template-based codes, \textsc{BPZ, EAZY}, and \textsc{LePhare}, show relatively well behaved $p(z)$ statistics (albeit with some bias), which may indicate that the likelihood estimation with representative templates is accurately capturing the uncertainties on individual redshifts.
-
-%Fig.~\ref{fig:pitqq} shows the PIT histogram plots produced by each photo-$z$ code. 
-%These plots show information that is very similar to that shown in the QQ plot in many ways. 
-The ideal PIT histogram would follow the black dashed line, representing a uniform distribution of PIT values, equivalent to the diagonal line in the QQ plot.
-Overly broad $p(z)$ values show up as an excess of PIT values near $0.5$ and a dearth of values at the edges, while overly narrow $p(z)$ will have an excess at the edges and will be missing values at the centre.
-Another feature evident in the PIT histograms is the number of ``catastrophic outlier'' values where the true redshift falls outside of the non-zero support of $p(z)$, corresponding to PIT$=0.0$ or $1.0$ is more apparent than in the QQ plots.
-Following Kodra \& Newman (in prep.) we define $f_{\rm O}$ as the fraction of objects with $PIT<0.0001$ or $PIT>0.9999$.
-Table~\ref{tab:pitoutlier} lists these fractions for each of the codes.
-For a proper Uniform distribution we expect a value of $0.0002$.
-Several codes show a marked excess, with \textsc{ANNz2, FlexZBoost, LePhare, and METAPhoR} with $f_{\rm O} > 0.02$, indicating a sizeable number of catastrophic redshift solutions where the true redshift is not covered by the extent of $p(z)$.
-For \textsc{METAPhoR} this may be partially due to an overall underprediction of the $p(z)$ width, however this is not the case for the other codes.
-\textsc{LePhare} is a particular outlier with nearly 5 per cent of objects outside of $p(z)$ support.
-Further study will be necessary to determine what is causing these misclassifications for \textsc{LePhare}.
-As expected, and by design, \textsc{TrainZ} has the proper fraction of outliers for the $f_{\rm O}$ statistic.
-
-\begin{table}
-\setlength{\tabcolsep}{2pt}
-\centering
-\caption{The fraction of ``catastrophic outlier'' PIT values.
-We expect a value of 0.0002 for a proper Uniform distribution.
-An excess over this small value indicates true redshifts that fall outside the non-zero support of the $p(z)$.
-}\label{tab:pitoutlier}
-\begin{tabular}{lc}
-\hline
-\hline
- ``catastrophic outlier''\\ PIT fraction\\
-\hline
-Photo-z Code & fraction PIT$<10^{-4}$ or $>$0.9999\\
-\hline
-\textsc{ANNz2} & 0.0265\\
-\textsc{BPZ} & 0.0192\\
-\textsc{Delight} & 0.0006\\
-\textsc{EAZY} & 0.0154\\ 
-\textsc{FlexZBoost} & 0.0202\\
-\textsc{GPz} & 0.0068\\ 
-\textsc{LePhare} & 0.0486\\
-\textsc{METAPhoR}& 0.0229\\
-\textsc{CMNN} & 0.0034\\ 
-\textsc{Skynet} & 0.0001\\  
-\textsc{TPZ} & 0.0130\\
-\hline
-\textsc{TrainZ} & 0.0002\\
-\end{tabular}
-\end{table}
-
-%\begin{figure*}
-%\centering
-%\includegraphics[width=\textwidth]{figures/PIT_histogram_plot_11codes.jpg}
-%\caption{The PIT histogram plots produced for every photo-$z$ code.
-Ideal behaviour would match a uniform distribution, indicated by the horizontal dashed line.
-Histograms with an overabundance of PIT values at the centre of the distribution indicate $p(z)$ distributions that are overly broad, while an excess of values at the extrema indicate $p(z)$ distributions that are overly narrow.
-Values of PIT=0 and PIT=1 indicate ``catastrophic failures'' where the true redshift is completely outside the support of $p(z)$.
-Asymmetric features are indicative of systematic bias in the redshift predictions.
-%\aim{Could we combine this with the previous figure by plotting both as differences from ideal on one plot with left and right axis labels?
-The information seems degenerate (which makes sense!)}\red{as you say, the information is degenerat, this form emphasizes the differences a bit more, and shows off the catastrophic outliers with PIT=0 and 1 a bit better in my mind.} } \label{fig:pit}
-%\end{figure*}
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{figures/KSvsCvMvsAD_PIT_withnull_jpg.jpg}
-\caption{A visual representation of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the PIT distributions.
-The statistics are often highly correlated, though the AD statistic truncates the extrema of the distribution and can have disparate values compared to KS and CvM.} \label{fig:pit_stats}
-\end{figure*}
-
-Fig.~\ref{fig:pit_stats} shows comparative metric values for the quantitative Kolmogorov-Smirnoff (KS), Cramer-Von Mises (CvM), and Anderson Darling (AD) test statistics for each of the codes based on comparing the distribution of their PIT values to the expected uniform distribution over the interval [0,1].
-The individual values of the statistic are not as important as the comparative score between the different codes.
-%\red{Can p-values be supplied for each statistic?
-The statistics themselves are difficult to interpret, other than ``lower is better'' (p-value in skgof was broken, having trouble finding 1-sample KS calculation for uniform distribution)}
-The AD test statistic diverges for values that include the extrema, and thus is calculated by excluding the edges of the distribution.  
-We calculate the AD statistic over the range of PIT values $v=[0.01,0.99]$.  \textsc{ANNz2} and \textsc{FlexZBoost} score very well for the PIT metrics.  
-\textsc{METAPhoR} and \textsc{LePhare} score very well in the PIT AD statistic, but both have a large number of catastrophic outliers, resulting in higher KS and CvM scores.
-
-Given the near-perfect training data, examining the individual codes for explanations for departures from the expected behaviour will be instructive in avoiding similar problems in future tests. 
-\textsc{ANNz2} performs quite well in $p(z)$ based metrics.
-In the specific implementation employed in this paper, the final $p(z)$ is a weighted average of five neural-nets.
-During the training process \textsc{ANNz2} compares the percentiles of the redshift training sample against the CDFs of the $p(z)$ sample.
-Distributions that more closely match are given extra weight, and the final weights are designed to produce accurate percentiles.
-Given that our metrics are focused on the percentile distributions, it is unsurprising that \textsc{ANNz2} performs well in the given metrics.
-The discreteness in the individual $p(z)$ estimated by \textsc{ANNz2} can be attributed to the fact that the code was run as a classifier, assigning weights to discrete bins of redshift.
-While multiple bins may receive weight, the bins themselves will still be discretized, and no additional smoothing was performed.
-Overall, \textsc{FlexZBoost} and \textsc{ANNz2} show the best ensemble agreement in their distribution of PIT values.
-
-\subsection{Metrics of the stacked estimator of the redshift distribution}\label{sec:stackedmetrics}
-
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{figures/NZsumplot_12codes.jpg}
-\caption{The stacked $p(z)$ produced by each photo-$z$ code ($\hat{N}(z)$, red) compared to the spectroscopic redshift distribution ($N'(z)$, blue).
-Varying levels of small-scale structure are seen in the codes.
-$N'(z)$ is smoothed using a single bandwidth chosen via Scott's rule for all codes.} \label{fig:nz}
-%\aim{Arguably this could also be clarified by showing the true $n'(z)$ once and then differences from it for each cod's stacked estimator\dots}\scc{what if we plot the difference between true and stacked rather than soverposing the two lines?
-Sam please do not hate me.}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{figures/KSvsCvMvsAD_NZ_withnull_jpg.jpg}
-\caption{A visual representation of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the $\hat{N}(z)$ distributions.
-The statistics are correlated, the codes with the lowest KS statistics tend to have the lowest CvM and AD statistics.
-\textsc{CMNN} performs markedly better than the others in reconstructing the overall $N(z)$ distribution, while \textsc{SkyNet} scores poorly due to an overall bias in its redshift predictions.} \label{fig:nz_stats}
-\end{figure*}
-
-Fig.~\ref{fig:nz} shows the stacked $\hat{N}(z)$ distribution compared to the true redshift distribution $N'(z)$ for all tested codes.
-The red line indicates the summed $p(z)$ for each code, while the blue line shows the true redshift distribution smoothed via kernel density estimation (KDE), with a bandwidth chosen via Scott's rule \citep{Scott:1992}.
-While Scott's rule is used to display $N'(z)$ in the figure, all quantitative statistics are computed via the empirical CDF, and are thus unaffected by bandwidth/smoothing choice.
-Several of the codes show an excess at $z \sim 1.4$, particularly the template-based codes \textsc{BPZ}, \textsc{EAZY}, and \textsc{LePhare}.  
-This is likely due to the $4000$ angstrom break passing through the gap between the $z$ and $y$ filters.  
-This feature is one of the most prominent in individual galaxy p(z), and is readily seen in the point-estimate plots shown in Fig.~\ref{fig:pz_pointestimates} and described in the Appendix.  
-Several of the machine learning based codes appear to be over-trained, adding excess galaxy probability to the redshift peaks and missing probability in the troughs.  
-Given that our training data is drawn from the same galaxy population as the test set, and our data has prominent peaks in $N'(z)$, perhaps it is not unexpected that such overtraining occurs. 
-A more extensive training/validation set might allow for a better choice of smoothing parameters in individual codes that would avoid such overtraining.
-
-%\aim{Move the text on Scott's Rule to Discussion?}
-%Scott's rule \citep{Scott:1992} was used to determine the smoothing bandwidth of the true redshift sample: that is, a single smoothing scale was chosen to represent the true redshift sample for all eleven codes.
-
-As with the $p(z)$ values in Figure~\ref{fig:pitqq}, different levels of substructure are obvious for the different codes.  
-While Scott's rule provides a relatively good general smoothing scale to represent the true $N'(z)$, there are smaller scale fluctuations: while \textsc{FlexZBoost} and \textsc{CMNN} appear somewhat discrepant in Fig.~\ref{fig:nz}, they are actually the two most accurate in terms of their quantitative measurements.
-%\textsc{SkyNet} also shows excess small-scale structure, but also shows significant overestimate of the high redshift galaxy population.  
-Interestingly, while \textsc{ANNz2} shows an abundance of small scale structure in individual $p(z)$ measurements (see Fig.~\ref{fig:pz_examples}), the summed $\hat{N}(z)$ is rather smooth, where the small scale features average out.
-This is not the case for the two other codes that show an abundance of substructure in their individual $p(z)$: both \textsc{CMNN} and \textsc{SkyNet} show small scale features both in $p(z)$ and $\hat{N}(z)$.
-For \textsc{CMNN} the $p(z)$ are a simply a weighted histogram of all spectroscopic training galaxies in nearby colour space with no smoothing applied, so the substructure is not unexpected.
-The PIT histogram and shape of the QQ plot in Figure~\ref{fig:pitqq} show that \textsc{CMNN} is producing $p(z)$ that are overly broad, additional smoothing of the $p(z)$ would exacerbate this problem.
-While the $\hat{N}(z)$ plot shows more small scale features than other codes, these features are actually representative of real structure in the true $N'(z)$, as evidenced by the very good metric scores for \textsc{CMNN}.
-\textsc{SkyNet} $p(z)$ were also not smoothed: while previous implementations of the code such as \citet{Sanchez:14} and \citet{Bonnett:15} (see Appendix C.3) implement a ``sliding bin'' smoothing, no such procedure was used in this study.
-In addition to excess substructure, \textsc{SkyNet} shows an obvious redshift bias, evident both visually in Figure~\ref{fig:nz} and in the first moment of N(z) listed in Table~\ref{tab:moments}, where it is clearly an outlier.
-\textsc{SkyNet} employed a method where a random sample of training galaxies was chosen, but there was no test that the subset was completely representative of the overall redshift distribution.
-Also unlike \citet{Bonnett:15}, no effort was made to add extra weight to more rare low and high redshift galaxies.
-Either of these decisions could be the cause of the bias seen in our results.
-Future runs of \textsc{SkyNet} will explore these implementation choices and their effects.
-
-Figure~\ref{fig:nz_stats} shows the quantitative Kolmogorov-Smirnoff (KS), Cramer-Von Mises (CvM), and Anderson Darling (AD) test statistics for each of the codes for the $\hat{N}(z)$ based measures.
-%\red{Can p-values be supplied for each statistic?
-The statistics themselves are difficult to interpret, other than ``lower is better'' (no, p-values are very difficult to compute for non-uniform distributions)}
-\textsc{FlexZBoost}, \textsc{CMNN}, and \textsc{TPZ} outperform the other codes in the $\hat{N}(z)$ metrics.  
-It is unsurprising that \textsc{CMNN} scores well, as with a near perfectly representative training set means that choosing neighbouring points in color/magnitude space should lead to excellent agreement in the final $\hat{N}(z)$ estimate.
-\textsc{TPZ} performed quite poorly in $p(z)$ statistics, but results in a good fit to the overall $N(z)$.
-This is somewhat surprising, as performance was optimized for accurate $p(z)$, not $\hat{N}(z)$.
-During the validation stage for \textsc{TPZ}, there was a trade off between the width of the $p(z)$ when adjusting a smoothing parameter and overall redshift bias.
-The optimal result in the PIT metrics, as illustrated in the shape of the QQ plot, does contain some level of bias as well as a slight underprediction of mean $p(z)$ width, which translates to poor metric scores.
-This is something that will be looked into for \textsc{TPZ} in the future.
-
-It is also of note that all three template-based codes show an excess in their stacked $p(z)$ at $z\,\sim\,1.3-1.4$.
-This redshift range corresponds to the wavelengths where the $4000$ Angstrom break is passing between the borders of the $z$ and $y$ filters.
-This strong break entering the gap between the two reddest filters can cause problems with redshift estimation of individual galaxies, as can be seen in the point-estimate photo-$z$'s shown in Figure~\ref{fig:pz_pointestimates}.
-This is not unique to this dataset, it is a common occurrence in photo-$z$ estimation.
-The fact that similar excesses appear in Figure~\ref{fig:nz} for \textsc{ANNz2} and \textsc{METAPhoR} shows that the effect is not limited to template-based codes.
-However, the lack of such a feature in the other codes shows that it is possible to eliminate the degeneracies.
-Further study on this issue may provide a solution for codes that suffer from this shortcoming.
-
-%\red{discussion of why NN and FZBoost are doing well, how we can try to improve the other codes} 
-%\textcolor{cyan}{Ann: }
-
-Table~\ref{tab:cdeloss} shows the CDE loss statistic for each photo-$z$ code.
-Once again \textsc{FlexZBoost} and \textsc{CMNN} score very well for the stacked $\hat{N}(z)$ metrics, as do \textsc{GPz} and \textsc{TPZ}.
-The CDE loss measures how well individual PDFs are estimated, and codes with a low CDE loss tend to have good $\hat{N}(z)$ estimates (though the reverse is not necessarily true).
-\textsc{FlexZBoost} is optimized to minimize CDE loss which may explain why the method has good ensemble metrics as well.
-Note from Table~\ref{tab:cdeloss} that both \textsc{FlexZBoost} and \textsc{CMNN} have low CDE losses.
-Empirically, we have found that PIT RMSE is not as closely correlated to CDE loss as it is to the $N(z)$ statistics.
-As CDE loss is a better measure of individual redshift performance, rather than ensemble distribution performance, this statistic is a better indicator of which codes will be most likely to perform well for science cases where single objects are employed.
-
-Table~\ref{tab:rmse} gives the root-mean-square-error (RMSE) statistics for both the PIT and N(z) estimators.
-The PIT value calculates the RMSE between the quantiles shown in the QQ plot in Figure~\ref{fig:pitqq} and the diagonal, while the N(z) calculates the RMSE between the cumulative distribution of the stacked $\hat{N}(z)$ and the true redshift distribution $N'(z)$.
-%\red{more about RMSE?
-Or remove it?
-It's barely described in Section~\ref{sec:metrics} and wasn't even referred to in the text until I added this on June 29th--SJS.}
-
-Table~\ref{tab:moments} lists the first three moments of the stacked $\hat{N}(z)$ distribution, including the moments of the ``truth'' distribution for comparison.
-Several codes are able to reproduce the mean and variance of the distribution to less than a per cent, while several codes do not, which may be a cause for concern, given that mean and variance of the redshift distribution are key properties in cosmological analyses.
-We note that this stated goal of the study as defined for participants was to accurately reproduce $p(z)$, the ``stacking'' of the probability distributions to estimate $\hat{N}(z)$ was not the focus as stated to the participants.
-This explains why some of the best-performing empirical codes in terms of $p(z)$ measures (e.~g.~\textsc{FlexZBoost}) do not do as well at reproducing $\hat{N}(z)$ moments.
-Had we defined a different parameter to optimize, in this case overall accuracy of $\hat{N}(z)$ rather than individual $p(z)$, would result in improved performance in a particular metric.
-That is, optimizing photo-$z$ performance for one metric does not automatically give optimal performance for other metrics.
-As previously stated, there are a variety of scientific use cases for photo-$z$'s in large upcoming surveys, and care must be taken in how the metrics used to optimize catalog photometric redshifts are defined as well as in how they are used.
-In addition, very few scientific use cases will employ the overall $\hat{N}(z)$ with no cuts, as we explore in this paper.
-We discuss more realistic tomographic bin selections that will be explored in a follow-up paper in Section~\ref{sec:futurework}.
-%\red{[mention that we are calculating moments for the entire sample, will look at fiducial tomographic bins in a follow-up paper (unless group and reviewers realy think we should include in this paper.]}
-%\red{FlexZBoost has some of the best $n(z)$ statistics, but the moments are not good.
-Why is this?
-Is it the over/underprediction of the high redshift part of the distribution?
-Discuss this after talking with Rafael.}
-
-%\subsection{Response of Individual Codes}
-%\label{sec:res:pz_indiv_codes}
-%\red{this may be incorporated in to 5.1 and 5.2 rather than live in its own section!}
-
-%\red{HOW CODES deal with negative fluxes, and magnitude uncertainties}
-
-%\red{overall conclusions of the Results section}
-
-\subsection{Interpretation of metrics}\label{sec:caution}
-%Caution on metrics and photo-$z$ codes}\label{sec:caution}
-%(Alex Malz, Sam Schmidt)
-
-Samples from accurate photo-$z$ posteriors should reproduce the space of $p(z, data)$.
-However, it is difficult to test this reconstruction given our data set, as the galaxy distributions arise from mock objects pasted on to an underlying dark matter halo catalogue with properties designed to match empirical relations, rather than being drawn from statistical distributions in redshift.
-In previous sections we have mentioned that optimizing for a specific metric does not guarantee good performance on other metrics, nor is there any guarantee that good performance by our metrics corresponds to \textit{accurate} photo-$z$ posteriors.
-%\aim{Explain that samples from accurate photo-$z$ posteriors should reconstruct the space of $p(z, data)$, a test we can't check with our datasets.}  
-In other words, we can construct photo-$z$ estimators that provide good coverage in many of our tests, but which have very little predictive power.
-
-The \trainz\ estimator, which assigns every galaxy a $p(z)$ equal to $N(z)$ of the training set as described in Section~\ref{sec:method:trainz}, is introduced as a ``null test'' to demonstrate this point via \textit{reductio ad absurdum}.
-\trainz\ outperforms all codes on the PIT-based metrics, and all but one code on the $N(z)$ based statistics.  
-Because our training set is perfectly representative of the test set, $N(z)$ should be identical for both sets down to statistical noise.
-%\textit{Explain the connection between \trainz\ and these metrics.}
-
-The CDE loss and point estimate metrics, however, successfully identify problems with \trainz.  
-As shown in Appendix~\ref{sec:pointmetrics}, \trainz has identical $ZPEAK$ and $ZWEIGHT$ values for every galaxy, and thus the photo-$z$s are constant as a function of spec-$z$s, i.e.
-a horizontal line at the mode and mean of the training set distribution respectively.
-The explicit dependence on the \it{individual} posteriors in the calculation of the CDE loss, described in Section~\ref{sec:CDE_loss}, distinguishes this metric from the other $p(z)$ metrics that test the overall ensemble of $p(z)$ distributions.
-With a representative training set, \trainz\ will score well on the ensemble metrics, but fails miserably for metrics tied to individual redshifts.
-We note that many of the ensemble-based metrics are prominent in the photo-$z$ literature despite their inability to identify problems such as those exemplified by \trainz.
-
-%\aim{Explain why CDE loss is different from other metrics (isn't fooled by \trainz) and note that the others have more of a presence in the literature despite their flaw of failing to identify this severe problem.}
-
-% While looking at one, or even most of our metrics would have given the impression that this estimator was nearly optimal, red flags in CDE loss and point estimates reveal the problems.  
-%[more caution on over-reliance on metrics, and making sure that metrics test all of what you want] \red{this definitely still needs work}
-In summary, context is crucial to interpreting metrics and defending against the likes of \trainz.
-The best photo-$z$ method is the one that most effectively achieves our science goals, not the one that performs best on a metric that does not accurately reflect those goals.
-In the absence of clear goals or the information necessary for a principled metric definition, we must think carefully before choosing a single metric
-
-%\begin{table*}  %%% DATA TABLE %%%
-%\caption{KS, CvM and AD statistics for each photo-$z$ code.} \label{tab:results}
-%PIT statistic
-%
-%\begin{tabular}{lrrrr}
-%\hline
-%\bf Photo-$z$ Code & \bf KS & \bf CvM & \bf AD & \bf AD \\
-%                   &        &         & $v=[0.05,0.95]$ & $v=[0.01,0.99]$        \\
-%\hline
-%\textsc{ANNz2} 		& $0.0200$ &   $52.25$ &   $583.0$ &   $759.2$ \\
-%\textsc{BPZ} 		& $0.0388$ &  $280.79$ &  $1361.4$ &  $1557.5$ \\
-%%\textsc{Delight} 	& $0.1380$ & $3040.59$ & $13880.2$ & $17761.4$ \\
-%\textsc{Delight}    & $0.0876$ & $1075.17$ & $4407.8$  & $6167.5$\\
-%%\textsc{EAZY} 		& $0.0690$ &  $994.19$ &  $3479.5$ &  $4158.8$ \\
-%\textsc{EAZY}       & $0.0723$ & $1105.58$ & $3475.0$  & $4418.6$\\
-%%\textsc{FlexZBoost} & $0.0260$ &   $70.58$ &   $595.0$ &   $533.0$ \\
-%\textsc{FlexZBoost} & $0.0240$ &   $68.83$ &   $567.9$ &   $478.8$\\
-%\textsc{GPz} 		& $0.0449$ &  $258.56$ &   $904.8$ &  $1414.8$ \\
-%\textsc{LePhare} 	& $0.0663$ &  $473.05$ &    $85.9$ &   $383.8$ \\
-%\textsc{METAPhoR} 	& $0.0438$ &  $298.56$ &   $248.6$ &   $715.5$ \\
-%%\textsc{CMNN} 		& $-$ & $-$ & $-$ & $-$ \\
-%\textsc{CMNN}         & $0.0795$  &$1011.11$ &  $3908.2$ &  $6307.5$ \\
-%\textsc{SkyNet} 	& $0.0747$ &  $763.00$ &  $2740.2$ &  $4216.4$ \\
-%\textsc{TPZ} 		& $0.1138$ & $1801.74$ &  $9046.2$ & $10565.7$ \\
-%%\textsc{Frankenz}	& $-$ & $-$ & $-$ & $-$ \\
-%\hline
-%\end{tabular}
-
-%Stacked $n(z)$ statistic
-%
-%\begin{tabular}{lrrrr}
-%\hline
-%\bf Photo-$z$ Code & \bf KS & \bf CvM & \bf AD & \bf AD \\
-%                   &        &         & $z=[0.002,2.0]$ & $z=[0.0,2.0]$        \\
-%\hline
-%\textsc{ANNz2} 		& $0.0237$ &  $25.676$ &   $276.4$ &   $276.0$ \\
-%\textsc{BPZ} 		& $0.0131$ &  $17.587$ &   $169.0$ &   $170.9$ \\
-%%\textsc{Delight} 	& $0.0184$ &  $18.975$ &   $151.5$ &   $151.5$ \\
-%\textsc{Delight}    & $0.0256$ &  $42.341$ &   $253.3$ &   $253.2$ \\
-%%\textsc{EAZY} 		& $0.0434$ & $167.329$ &   $811.3$ &   $817.0$ \\
-%\textsc{EAZY}       & $0.0441$ & $169.977$ &   $827.7$ &   $833.3$ \\
-%%\textsc{FlexZBoost} & $0.0065$ &   $2.764$ &    $24.7$ &    $24.6$ \\
-%\textsc{FlexZBoost} & $0.0128$ &   $7.007$ &  $127.02$ &   $126.9$\\
-%\textsc{GPz} 		& $0.0168$ &  $25.605$ &   $307.1$ &   $307.1$ \\
-%\textsc{LePhare} 	& $0.0254$ &  $58.900$ &   $477.4$ &   $472.0$ \\
-%\textsc{METAPhoR} 	& $0.0350$ &  $60.790$ &   $671.9$ &   $672.3$ \\
-%%\textsc{CMNN} 		& $-$ & $-$ & $-$ & $-$ \\
-%\textsc{CMNN}         & $0.0043$ & $0.538$   &   $5.6$   &      $5.8$ \\
-%\textsc{SkyNet} 	& $0.0483$ & $385.830$ &  $2119.4$ &  $2110.3$ \\
-%\textsc{TPZ} 		& $0.0126$ &   $9.483$ &   $110.7$ &   $110.6$ \\
-%%\textsc{Frankenz}	& $-$ & $-$ & $-$ & $-$ \\
-%\hline
-%\end{tabular}
-%\end{table*}
-
-\begin{table}  %%% DATA TABLE %%%
-\centering
-\caption{CDE loss statistic for each photo-$z$ code.} \label{tab:cdeloss}
-\begin{tabular}{lr}
-\hline
-\bf Photo-$z$ Code & \bf CDE Loss \\
-\hline
-\textsc{ANNz2} 		& $-6.88$ \\
-\textsc{BPZ} 		& $-7.82$ \\
-%\textsc{Delight} 	& $-4.06$ \\
-\textsc{Delight}    & $-8.33$\\
-%\textsc{EAZY} 		& $-7.97$ \\
-\textsc{EAZY}       & $-7.07$ \\
-%\textsc{FlexZBoost} & $-11.51$ \\
-\textsc{FlexZBoost} & $-10.60$\\
-\textsc{GPz} 		& $-9.56$ \\
-\textsc{LePhare} 	& $-1.66$ \\
-\textsc{METAPhoR} 	& $-6.28$ \\
-%\textsc{CMNN} 		& $-$ \\
-\textsc{CMNN}         & $-10.43$ \\
-\textsc{SkyNet} 	& $-7.89$ \\
-\textsc{TPZ} 		& $-9.55$ \\
-\hline
-\trainz		& $-0.83$ \\
-%\textsc{Frankenz}	& $-$  \\
-%\hline
-\end{tabular}
-\end{table}
-
-\begin{table}
-\setlength{\tabcolsep}{2pt}
-\begin{center}
-\caption{Root-Mean-Square-Error (RMSE) statistics for the eleven photo-z codes for both PIT and $\hat{N}(z)$ distributions.}\label{tab:rmse}
-\begin{tabular}{lcc}
-\hline
-\hline
-Root-Mean-Square-Error \\
-(RMSE) statistics \\
-\hline
-Photo-z Code & PIT RMSE & N(z) RMSE\\
-\hline
-\textsc{ANNz2}      & 0.019 & 0.0054\\
-\textsc{BPZ}        & 0.032 & 0.0050\\
-\textsc{Delight}    & 0.111 & 0.0056\\
-\textsc{EAZY}       & 0.054 & 0.0102\\
-\textsc{FlexZBoost} & 0.021 & 0.0022\\
-\textsc{GPz}        & 0.048 & 0.0042\\
-\textsc{LePhare}    & 0.028 & 0.0062\\
-\textsc{METAPhoR}   & 0.064 & 0.0081\\
-\textsc{CMNN}         & 0.108 & 0.0009\\
-\textsc{Skynet}     & 0.054 & 0.0144\\ 
-\textsc{TPZ}        & 0.082 & 0.0031\\
-\hline
-\trainz		& 0.0025 & 0.0013\\
-\end{tabular}
-\end{center} 
-\end{table}
-
-%\begin{table}
-%\begin{center}
-%caption{RMSE Statistics for the 11 Photo-z codes for DC1 $i<25.3$ ``gold'' sample}\label{tab:pitrmse}
-%%\begin{tabular}{|l|c|c|c|c|}
-%\begin{tabular}{lc}
-%\hline
-%\hline
-% PIT RMSE statistics \\
-%\hline
-%Photo-z Code & PIT RMSE\\
-%\hline
-%\texttt{ANNz2}      & 0.019\\
-%\texttt{BPZ}        & 0.032\\
-%\texttt{Delight}    & 0.111\\
-%\texttt{EAZY}       & 0.054\\
-%\texttt{FlexZBoost} & 0.021\\
-%\texttt{GPz}        & 0.048\\
-%texttt{LePhare}    & 0.028\\
-%\texttt{METAPhoR}   & 0.064\\
-%\texttt{NN}         & 0.108 \\
-%\texttt{Skynet}     & 0.054\\ 
-%texttt{TPZ}        & 0.082\\
-%\hline
-%\end{tabular}
-%\end{center} 
-%\end{table}
-%
-%\begin{table}
-%\begin{center}
-%\caption{RMSE Statistics for the 11 Photo-z codes for DC1 $i<25.3$ ``gold'' sample}\label{tab:nzrmse}
-%%\begin{tabular}{|l|c|c|c|c|}
-%\begin{tabular}{lc}
-%\hline
-%\hline
-% N(z) RMSE statistics \\
-%\hline
-%hoto-z Code & N(z) RMSE\\
-%\hline
-%\texttt{ANNz2}      & 0.0054\\
-%\texttt{BPZ}        & 0.0050\\
-%\texttt{Delight}    & 0.0056\\
-%\texttt{EAZY}       & 0.0102\\
-%\texttt{FlexZBoost} & 0.0022\\
-%\texttt{GPz}        & 0.0042\\
-%\texttt{LePhare}    & 0.0062\\
-%\texttt{METAPhoR}   & 0.0081\\
-%\texttt{NN}         & 0.0009\\
-%\texttt{Skynet}     & 0.0144\\ 
-%\texttt{TPZ}        & 0.0031\\
-%\end{tabular}
-%\end{center} 
-%\end{table}
-
-\begin{table}
-\setlength{\tabcolsep}{2pt}
-\caption{Moments of the stacked $\hat{N}(z)$ distribution}\label{tab:moments}
-\begin{tabular}{lccc}
-\hline
-\hline
- \multicolumn{4}{l}{Stacked $n(z)$ Moments} \\
-\hline
-              & 1st Moment & 2nd Moment & 3rd Moment \\
-\textsc{Truth}     & 0.701      &   0.630    & 0.671  \\
-\hline
-Photo-z Code       & 1st Moment & 2nd Moment & 3rd Moment\\
-\hline
-\textsc{ANNz2}     & 0.702      & 0.625      & 0.653    \\ 
-\textsc{BPZ}       & 0.699      & 0.629      & 0.671    \\ 
-\textsc{Delight}   & 0.692      & 0.609      & 0.638    \\ 
-\textsc{EAZY}      & 0.681      & 0.595      & 0.619    \\ 
-\textsc{FlexZBoost}& 0.694      & 0.610      & 0.631    \\ 
-\textsc{GPz}       & 0.692      & 0.605      & 0.619    \\ 
-\textsc{LePhare}   & 0.718      & 0.668      & 0.741    \\ 
-\textsc{METAPhoR}  & 0.705      & 0.628      & 0.657    \\ 
-\textsc{CMNN}        & 0.701      & 0.628      & 0.667    \\
-\textsc{Skynet}    & 0.743      & 0.708      & 0.797    \\
-\textsc{TPZ}       & 0.700      & 0.619      & 0.643    \\ 
-\hline
-\trainz	   & 0.699 		& 0.627 	& 0.666 \\
-\end{tabular}
-\end{table}
-
-\section{Summary and Discussion}
-\label{sec:discussion}
-%(Sam Schmidt, Ofer Lahav, Jeff Newman, Alex Malz, Eve Kovacs, Tony Tyson, Tina Peters)
-In this paper we presented results evaluating the photometric redshift PDF computation for eleven photo-$z$ codes.
-As discussed in Section~\ref{sec:metrics} the $p(z)$ should accurately reflect the relative likelihood as a function of redshift for each galaxy.
-All codes were provided a set of representative training data and tested on an idealized set of model galaxies with high signal-to-noise and photometry with no confounding effects due to blending, instrumental effects, the night sky, etc...
-included.
-The goal was not to determine a ``best'' photo-$z$ code: in many ways, this was a baseline test of a ``best case scenario'' to predict the expected photo-$z$ performance if a stage IV dark energy survey was to obtain complete training samples and perfectly calibrated their multi-band photometry.
-Given these idealized conditions, any deficiencies observed in a photo-$z$ code's performance should be a cause for concern, and may be evidence in a problem with either/both of the specific code implementation or the underlying algorithm.
-In order to meet the stringent LSST requirements on photo-$z$ performance, identifying and correcting such problems is an important first step before tackling more realistic data in future challenges.
-Most of the codes tested performed well, however, several did not meet the stringent goals that have been laid out for LSST photometric redshift performance.
-This is a cause for concern, given the idealized conditions, and the individual code responses will be studied in detail moving forward.
-One obvious trend in several of the codes tested was an overall over or underprediction of the widths of $p(z)$, as evidenced by the QQ plots and PIT histograms shown in Fig.~\ref{fig:pitqq}.
-A more careful tuning of bandwidth or smoothing during the validation process appears to be necessary for many of the machine learning based codes in order to improve the accuracy of $p(z)$.
-For narrow peaked $p(z)$ the parameterization of the PDF as evaluated on a fixed redshift grid could also have contributed to some overestimates of $p(z)$ width simply due to the finite resolution.
-After evaluating resuts such as those presented in \citet[]{Malz:qp}, in future analyses we plan to switch from a fixed grid to quantile-based storage of $p(z)$ in order to more efficiently and accurately store redshift PDF results.
-
-Another important factor to keep in mind when examining the results presented in this paper is the fact that they are at some level dependent on the metrics that we aim to optimize: in this case code participants were asked to submit their optimal measures of an accurate $p(z)$, so participants used the training/validation data to optimize their codes accordingly.
-Had we, instead, asked for an optimal $\hat{N(z)}$ the resulting metrics would be different for most, if not all, of the codes, as they would optimize toward a different goal.
-Specific metric choice can affect which codes are among the ``best'' codes.
-As stated earlier, there are cosmological science cases that require either individual galaxy photo-$z$ measures, or ensemble $\hat{N}(z)$ measures.
-We must be aware of that the optimal method for one is not necessarily optimal for the other, and in fact several photo-$z$ algorithms may be necessary in the final cosmological analysis in order to satisfy the requirements of all science use cases.
-The example of the simple \trainz\ estimator described in Section~\ref{sec:caution} shows a simple model with a $p(z)$ that is unrealistic for individual objects can still score very well on many of our metrics.
-It is important to look at {\it all} metrics, and keep in mind what information each metric conveys.
-%\red{mention 'null' and how 'dumb' algorithms can score well on metrics, need to keep all metrics in mind}
-%\aim{[reiterate the meaning of $p(z)$ and the goal -- emphasize finding ``best'' code in terms of impact of assumptions underlying the method \textit{or} establishing methodology for the realistic case of biased prior information and other science goals]}
-%\red{[the challenges faced in this project]}
-%\red{mention z$<$2, missing some important degeneracies, new sim will go to higher z}
-%\red{no quality cuts included, could identify outliers, but could also induce biases}
-%\red{[how optimization defined has impact, here we optimized p(z), not n(z), could be science-case specific if stringent requirements are needed, though that may be a computational/storage challenge if too many use cases needed.]}
-%\red{some discussion of output parameterization, limitations of 200 point fixed grid, in future switch to something like quantiles (cite qp paper)}
-We re-emphasize that the dataset tested was quite idealized, and discuss enhancements that will be added in future simulations to test photo-$z$ codes on increasingly realistic conditions in the following section.
-
-\subsection{Future work}
-\label{sec:futurework}
-The work presented in this paper is only the first step in characterizing current photo-$z$ codes and moving toward an improved photometric redshift estimator.
-This initial paper explored code performance in idealized conditions with perfect catalog-based photometry and representative training data.
-As mentioned in Section~\ref{sec:stackedmetrics} for the stacked $N(z)$ metrics we examined only the entire galaxy population with no selections in either photo-$z$ ``quality" or redshift.
-The cosmological analyses for weak lensing and large scale structure based measures plan to break galaxy samples into tomographic redshift bins, using photo-$z$ $p(z)$ to infer the redshift distribution for each bin.
-The specific selection used to determine these bins, both algorithmically and the specific bin boundaries, could induce biases due to indirect selections inherent in the photo-$z$ or other bin selection parameters.
-The effects of tomographic bin selection will be explored in a dedicated future paper.
-\red{[are there any references for this?
-I remember Gary Bernstein talking about this at a photo-z workshop in Japan, but I don't know that it was published.
-I believe Michael Troxel has discussed this as well.]}
-We also plan to propagate the uncertainties measured in a set of fiducial tomographic redshift bins in order to estimate impact on cosmological parameter estimation.
-%\item \red{cosmological parameters in DC2}
-%\item \red{mention z$<$2, missing some important degeneracies, new sim will go to higher z}
-%\red{no quality cuts included, could identify outliers, but could also induce biases}
-
-In future papers we will add more and more complexity to our simulated data in order to test photo-$z$ algorithms in increasingly realistic conditions.
-The most pressing concern is the impact of incomplete spectroscopic training samples.
-As discussed extensively in \citet{Newman:2015} a representative set of spectroscopically confirmed galaxies spanning the full range of both redshift and apparent magnitude is necessary as a training set to characterize the mapping from broad-band fluxes to photometric redshifts.
-However, due to a combination of factors due to both the galaxy SEDs and limitations of spectrographic instruments, redshift samples are known to be systematically incomplete, where certain galaxy types and redshift intervals fail to yield a redshift even at the longest integration times on current and near-future instruments.
-The more representative the training data, the better the performance of photo-$z$ algorithms will be.
-Current and upcoming surveys are putting in significant effort into obtaining these training samples \citep[e.~g.\,][]{Masters:2017}, however we still expect significant incompleteness for LSST-like samples, particularly at faint magnitudes.
-One major focus of an upcoming {\it LSST Dark Energy Science Collaboration Photo-z Working Group} data challenge is to produce a realistically incomplete training set of spectroscopic galaxies, modeling the performance of spectrographs, emission-line properties, and expected signal-to-noise to determine which galaxies will fail to yield a secure redshift.
-In addition to outright redshift failures we will model the inclusion of a small number of falsely identified secure redshifts where misidentified emission lines or noise spikes cause an incorrect redshift solution to be marked as a high quality identification.
-Even sub-per cent level contamination by false redshifts can impact photo-$z$ solutions at levels comparable to the stringent  requirements of some LSST science cases.
-We expect different systematics to occur in different photo-$z$ codes in response to training on incomplete data, particularly some of the machine learning methods.
-The response of the codes will inform future directions of code development.
-
-This initial paper explored a data set that was constructed at the catalog level, with no inclusion of the complications that come from measuring photometry from images.
-Future data challenges will move to catalogs constructed from mock images, including effects that will have great impact on photo-$z$ measurements.
-Object blending will be a major area of investigation, as the mixing of flux from multiple objects and the resultant change in measured colours is predicted to affect a large fraction of LSST galaxies \citep{Dawson:2016}, and will be one of the major contributing systematics for photo-$z$'s.
-Inclusion of differing observing conditions (seeing, clouds, variations in filter curves, Galactic dust, ...), as well as models for instrumental and system effects, sky masks, will all impact object photometry, and will be explored in the upcoming data challenge and their impacts described in upcoming papers.
-All underlying SEDs were parameterized as a weighted combination of five basis SEDs, with no additional accounting for host galaxy dust obscuration beyond what was encoded in the basis templates.
-This, in effect, limited the simulation to a very simple model of internal obscuration.
-Future simulations will include a more complicated and realistic treatment of host galaxy dust.
-
-The underlying simulation used in this work was based on a light-cone constructed to a maximum redshift of $z=2$.
-LSST imaging after 10 years of observations will include a significant number of $z>2$ galaxies in expected cosmology samples, and their inclusion does have potential significant implications for photo-$z$ measures: the high redshift galaxies lie at fainter apparent magnitudes and can have anomalous colours due to evolution of stellar populations and the shift to rest-frame magnitudes probing UV features of the underlying SED.
-More importantly, one of the most common ``catastrophic outlier'' degeneracies observed in deep photometric samples occurs when the Lyman break is mistaken for the Balmer break, leading to multiple redshift solutions at $z\sim0.2-0.3$ and $z\sim2-3$ \citep{Massarotti:2001}.
-This degeneracy, along with other potential degeneracies, are currently not covered by the limited redshift range of this initial paper, which could mean that we are not probing the full range of potential extreme outlier populations and how our photo-$z$ estimators respond to them.
-Extending simulations to include the high-redshift galaxy population will be a priority in future data challenges.
-
-%\item \red{inclusion of incompleteness effects on training sample, ``wrong redshifts'', filter curves}
-%\item \red{inclusion of photometric errors, image based effects, blending, lensing, sky, mask, etc...}
-
-In this study we have not accounted for the presence of Active Galactic Nuclei (AGN) contributions to galaxy fluxes.
-In some cases, AGN will be easily identified from the colors and morphologies, i.e.
-the case of the brightest quasars where the nuclear activity outshines the host galaxy, and numerous studies have utilized color selection to create large samples of quasars \citep[e.g.][]{Richards:06,Maddox:08,Richards:15}.
-In current deep fields, similar in depth to what we expect from LSST, variability information and multi-wavelength data have been critical to not only identify AGN dominated galaxies, but also obtain more accurate photometric redshifts \citep[e.g][]{Salvato:11}.
-
-In addition to AGN dominated galaxies, those with lower levels of nuclear activity present a more insidious problem, where AGN features may not be apparent, but the colors and other host galaxy properties are perturbed relative to galaxies with an inactive nucleus.
-In such cases, the presence of the AGN may induce a bias if the template SEDs or empirical datasets do not include low-level AGN counterparts.
-For LSST, we will need to identify and obtain accurate photometric redshifts of all types of AGN for a range of science goals, whether it is to eliminate such objects from cosmology experiments, or to use them with confidence, all the way through to understanding galaxy evolution and the role that AGN may play in influencing galaxy properties over cosmic time.
-
-A promising route to classifying and obtaining accurate photometric redshifts for the AGN population is by combining machine learning with template-fitting techniques, as has recently been demonstrated by \citet{Duncan:18} for radio-selected AGN.
-This is because AGN are relatively easy to obtain spectroscopic redshifts for over all redshifts due to the strong emission lines that they exhibit, allowing very good training sets for machine learning algorithms to use.
-Whereas for those galaxies where the AGN is sub-dominant the galaxy templates are still adequate for obtaining reasonable photometric redshifts.
-
-In addition to these improvements, the DESC Photo-$z$ group plans to look at all potential methods to combine the results from multiple photo-$z$ codes to improve $p(z)$ accuracy, similar to the work presented in \citet{Dahlen:13,Carrascokind:14,Duncan:18}.
-Taking advantage of multiple algorithms that use observables in slightly different ways has shown promise, however we must be very conscious of whether a potential combination properly treats the covariance between the methods, given that they are estimating quantities based on the same underlying observables.
-Several science cases wish to estimate physical quantities along with redshift, for example galaxy stellar mass and star formation rate.
-Proper joint estimation of redshift and physical quantities requires an in depth understanding of galaxy evolution, and progress on accurate bivariate redshift probability distributions will go hand in had with progress on understanding galaxies themselves.
-Parameterization and storage of a complex 2-dimensional probability surface for potentially billions of galaxies (or even subsets of hundreds of thousand of particular interest) pose a potential challenge.
-These issues will be examined in another future paper.
-
-Finally, while this paper and future papers discussed above focus on photometric redshift codes and estimating accurate $p(z)$ from training data, we plan a separate, but complementary, project to examine calibration of the resultant redshifts via spatial cross-correlations \citep{Newman:2008}, which will be explored in a separate series of future papers.
-The overarching plan describing everything laid out in this section is described in more detail in the LSST DESC Science Roadmap (see Footnote in Section~\ref{sec:intro}).
-These plans will require significant effort, but they are necessary if we are to make optimal use of the LSST data for astrophysical and cosmological analyses.
-%\begin{enumerate}
-%\item \red{the combination of codes}
-
-%\item \red{$p(z,\alpha)$}
-%\item \red{tomographic analysis later on}
-%\red{this paper deals with training, future papers will also explore calibration via cross-correlation methods}
-%\end{enumerate}
-%\red{mention the SRM once again, say that the plans are extensive, but we do have a plan and a rough timeline.}
-
-\section*{acknowledgements}
-\red{[summary of effort breakdown], will need to be in authors.csv for auto-generation by start\_paper}
-
-\red{[personal funding sources]}
-
-\red{[NERSC computation acknowledgement}
-
-The LSST Dark Energy Science Collaboration acknowledges generous ongoing support from the agencies and institutes supporting the collaboration.
-These include the Institut National de Physique Nucl\'eaire et de Physique des Particules in France, the Science \& Technology Facilities Council in the United Kingdom, and the Department of Energy, the National Science Foundation, and the LSST Corporation in the United States.
-This research used resources of the National Energy Research Scientific Computing Center, a DOE Office of Science User Facility supported by the Office of Science of the U.S.
-Department of Energy under Contract No.
-DE-AC02-05CH11231.
-Part of this work was undertaken on STFC DiRAC HPC Facilities, funded by UK BIS National E-infrastructure capital grants, and on the UK particle physics grid, supported by the GridPP Collaboration.
-
-\appendix
-
- \section{Point Estimate Photometric Redshifts}
- \label{sec:pointmetrics}
- While we do not recommend the use of single point estimates of redshift for most science applications, plots of the point estimates can be a useful qualitative diagnostic of photo-z code performance, i.~e.~examining point photo-$z$ vs.~spec-$z$ plots visually can give a quick impression of some common trends in different codes.
-Computing point estimate statistics may also be useful for more direct comparisons with previous photo-z evaluations.
-If a point-estimate is preferred for a specific science case, it is fairly simple to compute the mean, mode, or some other simple estimator from each $p(z)$, so these point estimates can be easily derived from the stored $p(z)$.
-
-There are several common point estimators of photo-$z$ posteriors employed by different codes, e.~g.
-the mode, mean, median of the $p(z)$ distribution.
-In addition, many of the machine learning based estimators can be set up to return a single redshift solution.
-For example, SkyNet can be configured to run as a regressor that returns a single float rather than a classifier that returns a 200-bin $p(z)$ estimate.
-The single value returned by a machine learning based code may not correspond to a particular measure such as the mode or mean, and so to avoid interpretation of results that might be introduced by variations in choice of specific point-estimate implementation per code, we discard the code-specific point estimates.
-We instead calculate point estimates more uniformly across the codes directly from the $p(z)$ using two measures, $z_{PEAK}$ and $z_{WEIGHT}$.
-$z_{PEAK}$  is simply the maximum value attained for each galaxy p(z), the mode of the probability distribution.
-$z_{WEIGHT}$ is defined similarly to how it is defined in \citet{Dahlen:13}, as the weighted mean of the redshift over the {\it main peak} of $p(z)$ containing the $z_{PEAK}$ value.
-The main peak is defined by subtracting 0.05$\,\times\,z_{PEAK}$ from $p(z)$ and identifying the roots to isolate the peak containing $z_{PEAK}$, $z_{WEIGHT}$ is defined as the weighted mean redshift within this peak.
-We restrict to a single peak in order to avoid confusion from bimodal and multimodal $p(z)$ such as those shown in bottom panels of Figure~\ref{fig:pz_examples}.
-For example, for a bimodal probability distribution a weighted mean calculated over both peaks would fall between the peaks, at a redshift where the probability is minimal.
-Restricting the weighting to a single peak ensures that the point estimate will fall in the region of maximum redshift probability.
-%\red{Have Dritan check zweight description to make sure this is what was done.}
-% The most common point estimators of photo-$z$ PDFs are the mode, median, and mean of the distribution.
-Following \cite{Tanaka:17}, we report the values of the following metrics using whichever of these point estimators performs best for each method, noting which point reduction is used in each case.
-
- \subsection{Point Estimate Metrics}
- \label{sec:point_metrics}
-We calculate the commonly used point estimate metrics of the overall photo-$z$ scatter ($\sigma_{z}$, the standard deviation of the photo-$z$ residuals), bias, and ``catastrophic outlier rate''.
-Specifically, we calculate the metrics as follows:
-we define $e_{z}$ as
-
-\begin{equation}
-e_{z} = \frac{z_{P} -z_{S}}{1+z_{S}} 
-\end{equation}
-where $z_{P}$ is the point estimate and $z_{S}$ is the true redshift.
-In practice, because the standard deviation calculation is quite sensitive to the outliers, we define the photo-$z$ scatter, $\sigma$ in terms of the Interquartile Range (IQR), the difference between the 75th and 25th percentiles of the $e_{z}$ distribution.
-In order to match the usual meaning of a 1$\sigma$ interval, we scale the IQR and define $\sigma_{IQR} = IQR/1.349$, as there is a factor of 1.349 difference between the IQR and the standard deviation of a Normal distribution.
-%$\sigma_{z}$ is, in practice, defined in terms the Interquartile Range (IQR) of $e_{z}$ values as $IQR/1.349$, where $IQR$ is the difference between the 75th and 25th percentiles of the $e_{z}$ distribution.
-The factor of 1.349 relates the IQR of a Normal distribution to the standard 68\% range of a one $\sigma$ uncertainty.
-%\scc{usually the statistical indicator defined as $\sigma$ is the standard deviation I know that is not a rule but if someone skips the test and go directly to the table (and we all know that a lot of people skips the definition of statistical indicator) could be confusing, could we report it as $\frac{IQR(e_z)}{1.349}$ or $\sigma_{IQR}$ that should be clear without any risk of confusion?}\red{Using IQR is done because IQR is less sensitive to outliers than calculating a sigma.
-I've rewritten the text explaining this above.--SJS}
-While many other studies define the bias based on the {\it mean} offset between true and estimated redshift, in this study we define the bias as the median value of $e_{z}$ for the sample.
-We use median as it is, once again, less sensitive to outliers than the mean.
-The catastrophic outlier fraction is defined as the fraction of galaxies with $e_{z}$ greater than the {\it larger} of $3\sigma_{IQR}$ or 0.06, i.e.
-3$\sigma$ outliers with a floor of $\sigma_{IQR}$=0.02.
-%\scc{again, is clearly not a rule but a huge number of paper refers to the mean as bias, could we use the word median, which is clear without any doubt rather than bias?}
-%\red{I also have code to calculate $\sigma_{MAD}$, should we include this as well?  It's almost always within a few thousandths of $\sigma_{IQR}$, so I left it out for now}.
-%\scc{we may just say that the two indicators reports almost the same value therefore there is no need for both of them?}
-For reference, the goals stated in Section 3.8 of the LSST Science Book \citep{Abell:09} for photo-$z$ performance in these metrics, assuming perfect training knowledge (as we are testing in this paper) are:
-\begin{itemize}
-\item RMS scatter$ < 0.02(1+z)$ 
-%\scc{actually the SB reports in page 75 that the goal of 0.02 is for the RMS of $\frac{\sigma_{z}}{(1+z)}$ while we are using IQR}
-\item bias $<$ 0.003 %\scc{in page 518 of the SB is clearly reported that the mean is expected to be less than 0.003 in our case we have defined as bias the median and not the mean }
-\item catastrophic outlier rate $<$ 10\% %\scc{clearly the number of outliers since are defined above $3\sigma$ strictly depends on the sigma definition, in the science book it seems to refer to the RMS of the $e_z$ distribution even ef it could be the standard deviation, but again I don't think it refers to IQR, therefore we could not consider exactly this numbers as reference.}
-%\red{I am going off of conversations with Zeljko and how we {\it actually} computed the metrics in the Science Book (I have the scripts).
-You are correct that several of these are slightly different than actually stated in the Science Book, or where the Science Book does not use very precise language as to what was done to compute the RMS or define the bias.
-We can also cut out those specifics, or maybe say Ivezic private communication, maybe?
-Also $\sigma_{z}$ defined in terms of IQR is exactly equivalent to the standard deviation when scaled by 1.349 for a Gaussian distribution, so the numbers are appropriate to cite in this context.
-IQR is just a more robust way of calculating sigma for a distribution with outliers.--SJS }
-%\scc{I am very sorry I am not saying that IQR is not fine, I want to keep the IQR my concerns are related to the notation and the \textit{reader understanding}, if I wrote $\sigma$ somewhere in a table the reader will understand for sure that is the traditional standard deviation (a lot of people skips the indicators definition for the \textit{obvious} ones), I am suggesting to replace the notation $\sigma$ with the notation $\sigma_{IQR}$ (or something like this) to avoid any confusion, and also to change from $bias$ to $median$ for the same reason.
-I agree that say Ivezic private communication could be better.}
-\end{itemize}
-These definitions are similar, but not exactly the same, as the $\sigma_{IQR}$ and median bias calculated here, but are similar enough for qualitative comparisons to the LSST goals.
-
-%Scatter: SRD $\sigma<0.02(1+z)$
-%Catastrophic Outler Rate: SRD $3\sigma$ ``catastrophic outlier'' rate below 10 per cent
-%Bias: SRD bias$<0.003$
-
-Fig.~\ref{fig:pz_pointestimates}  shows the point estimates for both $z_{PEAK}$ and $z_{WEIGHT}$.
-Point density is shown with mixed contours to emphasize that most of the galaxies do fall close to the $z_{phot}=z_{spec}$ line, while blue points show differing characteristics of the outlier populations.
-The red dashed lines indicated the cutoff for catastrophic outliers, defined as: $max(0.06,3\sigma_{IQR})$.
-As with the full $p(z)$ results, a variety of behaviours are evident in the different codes.
-Table~\ref{tab:pointestimates} lists the scatter, bias, and catastrophic outlier fractions for the codes.
-The performance of the codes for point metrics is highly correlated with performance on $p(z)$ based tests, which is to be expected, given that the point-estimates were derived from the $p(z)$.
-Some discretization is evident in $z_{PEAK}$, particularly for \textsc{SkyNet}, due to the finite grid spacing of the reported $p(z)$.
-These discreteness effects are mitigated by the weighting of $z_{WEIGHT}$, resulting in a smoother distribution of redshift estimates.
-Several features perpendicular to the main $z_{phot}=z_{spec}$ line are evident.
-These features are due to the 4000 angstrom break passing through the gaps between adjacent LSST filters.
-These features are most prominent in template-based codes, but appear to some degree in all codes tested.
-
-In even the best performing codes, there are visible occupied regions away from the $z_{phot}=z_{spec}$ line, corresponding to degenerate redshift solutions for certain LSST magnitudes and colors.
-While use of the full information available via $p(z)$ mitigates their impact, a full understanding of the outlier population is critical for LSST science, particularly in tomographic applications %\red{I forget what exactly I was trying to say here, if someone else wants to take a crack at this it might be helpful --SJS}
-%\scc{may be worth to say that the usage of further bands could mitigate this effect by removing the degeneration  usually induced from emission lines moving across different filters?}
-
-Finally, we note that all eleven codes are at or near the goals for point-estimates as outlined in the LSST Science Requirements Document\footnote{available at: \url{http://ls.st/srd}} and \citet{Graham:17}.
-This is to be expected, given that the requirements were designed such that a point estimate photo-z would meet these requirements for perfect training data to a depth of $i<25$.
-But, it is still an encouraging sign, given an updated mock galaxy simulation and the expanded set of photo-$z$ codes tested.
-%\red{maybe say consistent with expectations, and a reference to Melissa's paper in here?  What else to say about nearly meeting point goals?}
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.9\textwidth]{figures/ZPEAK_szpz_threecolumn_12codes_navy.jpg}
-\includegraphics[width=0.9\textwidth]{figures/ZWEIGHT_szpz_threecolumn_12codes_navy.jpg}
-\caption{Point estimate photo-z's derived from the posteriors.
-Top panel shows $z_{PEAK}$, while bottom panel shows $z_{WEIGHT}$.
-Point estimate density is represented with fixed density contours, while outliers at lower density are represented by blue points.
-While use of point-estimate photo-$z$'s is not recommended, they do make for useful comparative and visual diagnostics.
-In the lower-right panel of each plot, the \trainz\ estimator results in identical photo-$z$ estimates at the mode and mean of the training set $N'(z)$ distribution for all galaxies.} \label{fig:pz_pointestimates}
-\end{figure*}
-
-\begin{table*}
-\begin{center}
-\caption{Point estimate statistics}\label{tab:pointestimates}
-\begin{tabular}{lcccccc}
-\hline
-\hline
-                  &            & $Z_{PEAK}$  &          &  & $Z_{WEIGHT}$          &\\
-\hline
-Photo-z Code       & $\frac{\sigma_{IQR}}{(1+z)}$ & median  & \multicolumn{1}{|p{0.75cm}|}{\centering outlier \\fraction} & $\frac{\sigma_{IQR}}{(1+z)}$ & median & \multicolumn{1}{|p{0.75cm}|}{\centering outlier \\fraction}\\
-\hline
-\textsc{ANNz2}     & 0.0270  &  0.00063  & 0.044      & 0.0244  &  0.000307  & 0.047  \\ 
-\textsc{BPZ}       & 0.0215  & -0.00175  & 0.035      & 0.0215  & -0.002005  & 0.032 \\ 
-\textsc{Delight}   & 0.0212  & -0.00185  & 0.038      & 0.0216  & -0.002158  & 0.038 \\ 
-\textsc{EAZY}      & 0.0225  & -0.00218  & 0.034      & 0.0226  & -0.003765  & 0.029 \\ 
-\textsc{FlexZBoost}& 0.0154  & -0.00027  & 0.020      & 0.0148  & -0.000211  & 0.017 \\ 
-\textsc{GPz}       & 0.0202  & -0.00091  & 0.036      & 0.0201  & -0.000950  & 0.037 \\ 
-\textsc{LePhare}   & 0.0236  & -0.00161  & 0.058      & 0.0239  & -0.002007  & 0.056 \\ 
-\textsc{METAPhoR}  & 0.0264  &  0.00000  & 0.037      & 0.0262  &  0.001333  & 0.048 \\ 
-\textsc{CMNN}        & 0.0184  & -0.00132  & 0.035      & 0.0170  & -0.001049  & 0.034 \\
-\textsc{Skynet}    & 0.0219  & -0.00167  & 0.036      & 0.0218  &  0.000174  & 0.037 \\
-\textsc{TPZ}       & 0.0161  &  0.00309  & 0.033      & 0.0166  &  0.003048  & 0.031 \\ 
-\hline
-\trainz	   & 0.1808  &  -0.2086  & 0.000	  & 0.2335  & 0.022135  & 0.000\\
-\end{tabular}
-\end{center} 
-\end{table*}
-
-\bibliographystyle{mn2e}
-\bibliography{references}
-\end{document}
+% ======================================================================

### 2018-08-20 paper/tex/analysis.tex

\section{Metrics for quantifying PDF comparisons}
\label{sec:metrics}
%(Alex Malz, Rongpu Zhou, Jeff Newman, Ofer Lahav)

%\red{Maybe it would be best to introduce the summary statistics (KS, CvM, AD) up front and then the ensemble functions and stacked n(z).}

The overloaded ``$p(z)$'' is a widespread abuse of notation; we would like the outputs of photo-$z$ PDF codes to be interpretable as probabilities.
Obviously photo-$z$ PDFs must not take negative values and must integrate to unity over the range of possible redshifts.
Additionally, an estimator derived by method $H$ for the photo-$z$ PDF of galaxy $i$ must be understood as a posterior probability distribution

\begin{equation}
\hat{p}_{j}(z_{i}) = p(z | d_{i}, I_{D}, I_{H}),
\end{equation}
conditioned not only on the photometric data $d_{i}$ for that galaxy but also on parameters encompassing a number of things that will differ depending on the method $H$ used to produce it, namely the assumptions $I_{H}$ necessary for the method to be valid and any inputs $I_{D}$ it takes as prior information, such as a template library or training set.
Because of this, direct comparison of photo-$z$ PDFs produced by different methods is in some sense impossible; even if they share the same prior information $I_{D}$, by definition they cannot be conditioned on the same assumptions $I_{H}$, otherwise they would not be distinct methods at all.

In this study, we isolate the differences in prior information specific to each method by using a single training set $I_{D}^{ML}$ for all machine learning-based codes and a single template library $I_{D}^{T}$ for all template-based codes, and these sets of prior information are carefully constructed to be representative and complete, we have $I_{D}^{ML}\equiv I_{D}^{T}$ for every method $H$.
Thus, we are saying

\begin{equation}
\frac{\hat{p}_{i, H}(z)}{\hat{p}_{i, H'}(z)} \approx \frac{p(z | d_{i}, I_{H})}{p(z | d_{i}, I_{H'})},
\end{equation}
meaning that we assume comparisons of $\hat{p}_{i, H}(z)$ isolate the effect of the method used to obtain the estimator, which should make examination of differences caused by specifics of the method implementations easier to isolate.

As mentioned previously, there are cosmology probes that require knowledge of individual galaxy $p(z)$ and others that require only knowledge of the ensemble redshift distribution, $N(z)$.
Due to the paucity of principled techniques for using and validating photo-$z$ PDFs, there are few alternatives to the common practice of reducing photo-$z$ PDFs to point estimates.
Though this practice should not be encouraged, we also calculate traditional metrics based on the most common point estimators derived from photo-$z$ PDFs.
Those seeking to establish a connection to traditional ways of thinking about redshift estimation may consult the Appendix for these results.

There are a number of metrics that can be used to test the accuracy of a photo-$z$ interim posterior as an estimator of a true photo-$z$ posterior if it is known.
Even for simulated data, the true photo-$z$ PDF is in general not accessible unless the redshifts are in fact drawn from the true photo-$z$ PDFs, a mock catalogue generation procedure that has not yet appeared in the literature.
Furthermore, only limited applications of photo-$z$ PDFs that could be used as the basis for a metric have been presented in the literature.
The most popular application by far is the calculation of the overall redshift distribution $N(z)$, the true value of which is known for the \textsc{Buzzard} simulation and will be denoted as $N'(z)$.
Though alternatives exist \citep{Malz:chippr}, stacking according to

\begin{equation} \label{eq:stacked}
\hat{N}^{H}(z) \approx \frac{1}{N_{tot}}\ \sum_{i}^{N_{tot}}\ \hat{p}^{H}_{i}(z)
\end{equation}
is the most widely accepted method for estimating the redshift distribution from photo-$z$ PDFs.
If we assume that the response of estimators of $N(z)$ is uniform across all approaches $H$, then we may interpret metrics on the accuracy of $\hat{N}(z)$ obtained in this way.
We must note, however, that this is a poor assumption in general.
Under the setup of this paper, the true redshift distribution $N'(z) = p(z | I_{D})$ (i.e. because our training data is representative, the interim prior is the truth).  In this ideal case, the method that would give the best approximation to $N'(z)$ would be one that neglects all the information contained in the photometry $\{d_{i}\}_{N_{tot}}$ and gives every galaxy the same photo-$z$ PDF $\hat{p}_{i}(z) = N'(z)$ for all $i$.  This is the exact estimator, \trainz, that we have described in Section~\ref{sec:method:trainz}, and which will serve as a point of reference for the other codes.

The exact implementation of the stacked estimator $\hat{N}^{H}(z)$ will depend on the parametrization of the photo-$z$ PDFs, which may differ across codes and can affect the precision of the estimator \citep{Malz:qp}; even considering a single method under the same parametrization, say a piecewise constant function over bins or a set of samples from the posterior, an estimator using $2N$ bins or samples will trivially be more precise than an estimator using $N$ bins or samples.
%\red{Note: Let's address how the gridding (endpoints and spacing) affects the metrics, from the email correspondence Re: [LSST-DESC-MEMBERS] Summary of Nov 29 PZ Telecon. (AIM)}
In order to minimize the effects of such choices, we asked those running all eleven codes to output $p(z)$ parameterized with a generous $\approx 200$ piecewise constant bins spanning $0<z<2$.
The piecewise constant format is chosen because of its established presence in the literature, and the choice of 200 bins was motivated by the approximate number of columns expected to be available for storage of $p(z)$ for the final LSST Project tables.\footnote{See, e.~g.~the LSST Data Products Definition Document, available at: \url{https://ls.st/dpdd}}
All $p(z)$ catalogues are processed using the \textsc{qp} software package \citep{Malz:qp}\footnote{available at: \url{http://github.com/aimalz/qp/}} for manipulating and calculating metrics of 1-dimensional PDFs.
We will discuss the choice of $p(z)$ parameterization further in Section~\ref{sec:results}.

%\red{Divide metrics by application to ensemble of PDFs vs. estimators of the redshift distribution}

\subsection{Metrics of an ensemble of photo-\mathinhead{$z$}{z} interim posteriors}
\label{sec:qualmet}

\subsubsection{Probability integral transform (PIT)}
\label{sec:pit}
%(Alex Malz, Peter Freeman, Sam Schmidt)

The probability integral transform (PIT) \citep{Polsterer:16} is defined for each individual galaxy as:%essentially the derivative of the QQ plot, a histogram of

\begin{align}
\label{eq:pit}
\mathrm{PIT} &= \int_{-\infty}^{z_{\rm true}}p(z)\, dz \,.
\end{align}
The distribution of PIT values quantifies the behavior of the {\it ensemble} of photo-$z$ PDFs, enabling us to evaluate whether the $p(z)$ is, on average, accurate:  The PIT value is the Cumulative Distribution Function (CDF) of the $p(z)$ evaluated at the true redshift.  A catalogue of photo-$z$ PDFs that are accurate should have a flat PIT histogram (i.e., the individual PIT values as samples from each CDF should match a Uniform(0,1) distribution if the CDFs are accurate).  Specific deviations from flatness indicate inaccuracy: overly broad photo-$z$ PDFs would manifest as underrepresentation of the lowest and highest PIT values, whereas overly narrow photo-$z$ PDFs would manifest as over-representation of the lowest and highest PIT values.
High frequency at only $\mathrm{PIT}\approx0$ and $\mathrm{PIT}\approx1$ indicates the presence of catastrophic outliers with highly inaccurate photo-$z$ PDFs where the true redshift is outside of the support of $p(z)$.
\citet{Tanaka:17} use the histogram of PIT values as a diagnostic indicator of overall code performance, while \citet{Freeman:17} independently define the PIT and demonstrate how its individual values may be used both to perform hypothesis testing (via, e.g., the KS, CvM, and AD tests; see below) and to construct quantile-quantile plots.
%\red{We might want to divide this by redshift to get trends, as the overall PIT could be impacted by true n(z).
%For example, a PIT with nontrivial shape could be due to that behavior being uniform as a function of redshift or significant inaccuracy only in a small range of redshifts where $n(z)$ is large and a flat PIT elsewhere.}

\subsubsection{Quantile-quantile (QQ) plot}
\label{sec:qq}
%(Rongpu Zhou, Jeffrey Newman, Alex Malz)

The quantile-quantile (QQ) plot is a graphical method for comparing two distributions, where the quantiles of one distribution are plotted against the quantiles of the other distribution (A quantile being defined by partitioning a distribution into consecutive intervals containing equal amounts of probability, or equal numbers of objects in each interval).
%\blue{Include an equation to ensure quantile definition is crystal clear before building upon it for PIT.}\red{Is maybe the above text clear enough?  Text may be more clear than any equation I can come up with--SJS.}
In this paper we show the quantiles of the PIT values compared to the quantiles of the Uniform distribution that we expect the PIT values to match if $p(z)$ is an accurate probability distribution for all objects.
The QQ plot provides an easy way to qualitatively assess the differences in various properties such as the moments of an estimating distribution relative to a true distribution.
%location, scale and skewness.
In this paper, QQ plots are used for two purposes: (1) for comparing $N(z)$ from photo-$z$ PDFs (estimated using Eq.~\ref{eq:stacked}) with the true $N(z)$, i.e. comparing the estimated distribution of redshifts with the true redshift distribution, and (2) for assessing the overall consistency of an ensemble of photo-$z$ PDFs with their true redshifts on a population level, where the distribution of the PIT values (see previous section) is compared to a uniform distribution between $0$ and $1$.  The QQ plot contains very similar information to that shown in the PIT histogram plot, we include both forms, as visually they each convey the information in a somewhat distinct manner. %\red{is this enough reason to include both?--SJS}
%\aim{Maybe we should explain why we expect the distribution of CDFs to be uniform.}\red{Note: this is just describing the PIT.  I'm going to swap the order of PIT to come before this, and then say we are evaluating the distribution of PIT values.  I will also add why we expect this to be uniform--SJS}

\subsubsection{Conditional density estimation loss}
\label{sec:CDE_loss}
%(Alex Malz, Ann Lee)

With the conditional density estimation loss (CDE loss) we can compare how well different methods estimate individual PDFs for photometric covariates $\x$ rather than looking only at the ensemble distribution.  As in Section~\ref{sec:flexzboost}, we use the notation $f(z \vert \x)$ instead of $p(z)$ to explicitly show the dependence on $\x$.

%\aim{Perhaps this should be rewritten in the same notation used elsewhere in the paper.} \red{I agree, does someone who speaks both stats and astro want to take a shot at this?}
%\textcolor{cyan}{Ann: We are looking at individual PDFs so the dependence on $\x$ needs to be explicit; this is probably the most accurate notation as it shows both the conditioning clearly as well as whether $\x$ is fixed or random.}

The CDE loss is defined as:
\begin{equation} \label{eq:cde-loss}
L(f, \widehat{f}) = \int \int (f(z \mid \x) - \widehat{f}(z \mid \x))^{2} dz dP(\x)
\end{equation}
This loss is the CDE equivalent of the RMSE in regression. To estimate this loss we rewrite the loss as
\begin{equation} \label{eq:estimated-cde-loss}
\mathbb{E}_\X \left[\int \widehat{f}(z \mid \X)^{2} dz\right] - 2 \mathbb{E}_{\X,Z}\left[\widehat{f}(Z \mid \X)\right] + K_{f},
\end{equation}
where the first expectation is with respect to the marginal distribution of the covariates $\X$, the second expectation is with respect to the joint distribution of $\X$ and $Z$, and $K_{f}$ is a constant depending only upon the true conditional densities $f(z\mid \x)$.
For each method we can estimate these expectations as empirical expectations on the test or validation data \citep[Eq.~7 in][]{Izbicki:17b} without knowledge of the true densities.

%\subsubsection{Continuous Ranked Probability Score (CRPS)}
%\label{sec:crps}
%\red{The CRPS is used to assess the accuracy of two probabilistic models respectively.}

\subsection{Metrics over estimated probability distributions}
\label{sec:quantmet}

In tandem with the QQ and PIT metrics introduced above, we additionally compute the following metrics comparing the empirical CDF of a distribution to the true or expected distribution.  These metrics give a more quantitative measure of the departure from ideal than the more visual PIT histogram and QQ plot.  We compute metrics comparing the CDF of PIT values to a the CDF of a Uniform distribution, and also compute the CDF of the true redshift distribution $N'(z)$ compared the $\hat{N}(z)$ distribution derived from summing the $p(z)$ as described in Eq.~\ref{eq:stacked}.

\subsubsection{Root-mean-square error (RMSE)}
\label{sec:rmse}
%(Alex Malz)

We employ the familiar root-mean-square error:

\begin{equation}
\mathrm{RMSE} = \sqrt{\int_{-\infty}^{\infty}\left(\hat{f}(z)-f'(z)\right)^{2}\,dz},
\end{equation}
%may be used for two purposes in this analysis.
%We calculate it between the true redshift distribution $N'(z)$ and each estimator $\hat{N}(z)$ (derived by Eq.~\ref{eq:stacked}) as well as between the ideal QQ and PIT curves from each method studied and the ideal curves.
Though this metric does not account for the fact that the redshift distribution function is, in fact, a probability distribution, it can still be interpreted as a measure of the integrated difference between the estimated distribution and the true distribution, and it can be used to quantify the otherwise qualitative metrics.

\subsubsection{Kolmogorov-Smirnov (KS) and related statistics}
\label{sec:ks}
%(John Soo, Rongpu Zhou)

The \textit{Kolmogorov-Smirnov statistic} $N_{\rm KS}$ is the maximum difference between $F_{\rm phot}(z)$ and $F_{\rm spec}(z)$, the CDFs of the photo-$z$ and spectroscopic redshift respectively:

\begin{equation}
N_{\rm KS} = \max_{z}\left( \left| F_{\rm phot}(z) - F_{\rm spec}(z) \right| \right).
\end{equation}
%The KS test quantifies whether the two redshift distributions are drawn from a similar distribution, independent of binning.
The KS test quantifies the similarity between two distributions, independent of binning.
A lower $N_{\rm KS}$ value corresponds to more similar distributions.
%We calculate two KS test statistics, one to test the distribution of $p(z)$ values, and one to test the stacked estimator $\hat{n}(z)$ of the redshift distribution.
%The $p(z)$ test compares the CDF of the PIT values described in the previous section to that of a uniform distribution.
%The $n(z)$ test compares the CDF of the distribution formed by stacking the $p(z)$ for every object in our test set against the CDF of the corresponding true redshifts.
%In this case a lower $N_{\rm KS}$ value means that the photo-$z$ distribution fits the spectroscopic redshift distribution well. $F_{\rm phot}(z)$ is calculated by stacking the PDFs of every single object in the sample, producing a smooth cumulative $n(z)$ to be compared with $F_{\rm spec}(z)$.

We also consider two variants of the KS statistic: the Cramer-von Mises (CvM) and Anderson-Darling (AD) statistics.
The CvM statistic is similar to the KS statistic as it is also computed from the distance between the measured CDF and the ideal CDF, but instead of the maximum distance, the CvM statistic calculates the average of the distance squared:

\begin{equation}
\label{eq:cvm}
\omega^2 = \int_{-\infty}^{+\infty}\big(F_{\mathrm{meas.}}(x) - F_{\mathrm{ideal}}(x)\big)^2\mathrm{d}F_{\mathrm{ideal}}
\end{equation}
% In this paper we use the CvM statistic to evaluate the performance of both $n(z)$ and $p(z)$.
The AD statistic is a weighted version of the CvM statistic, making it more sensitive to the tails of the distribution:

\begin{equation} \label{eq:ad}
A^2 = n\int_{-\infty}^{+\infty} \frac{\big(F_{\mathrm{meas.}}(x) - F_{\mathrm{ideal}}(x)\big)^2} {F_{\mathrm{ideal}}(x) (1-F_{\mathrm{ideal}}(x))}\mathrm{d}F_{\mathrm{ideal}}
\end{equation}
where $n$ is the sample size.
%The KS statistic and the CvM and AD statistics related to it are all used to evaluate the performance of both $\hat{n}(z)$ and the QQ and PIT of the CDFs of the ensemble of photo-$z$ interim posteriors.

%\subsubsection{Kullback-Leibler Divergence (KLD)}
%\label{sec:kl}
%(Alex Malz)
%
%The KLD is a measure of the loss in information when using a probability distribution that is an estimate of a known true probability distribution, so we use
%\begin{equation}
%\mathrm{KLD} = \int_{-\infty}^{\infty}n'(z)\log\left[\frac{n'(z)}{\hat{n}(z)}\right]dz
%\end{equation}
%as a metric on the stacked redshift distribution estimator of Eq.~\ref{eq:stacked}. Noting the assumptions made in using Eq.~\ref{eq:stacked}, we say that a photo-$z$ PDFs method whose stacked redshift distribution estimator $\hat{n}(z)$ more closely resembles the true redshift distribution $n'(z)$ is desired, i.e. we seek to minimize the KLD.

\subsubsection{Moments}
\label{sec:moments}
%(Alex Malz)
For the $\hat{N}(z)$ distributions we additionaly calculate the first three moments of the estimated redshift distribution for each code and compare them to the moments of the true redshift distribution $N'(z)$.
The $m$th moment of a distribution is defined as

\begin{equation}
\langle z^{m}\rangle = \int_{-\infty}^{\infty}z^{m}N(z)dz.
\end{equation}
Here, we use the moments of the stacked estimator of the redshift distribution function as the basis for a metric. The closer the moments of $\hat{N}(z)$ for a photo-$z$ PDF method are to the moments of the true redshift distribution function $N'(z)$, the better the photo-$z$ PDF method.

### 2018-08-20 paper/tex/conclusion.tex

\section{Conclusion}
\label{sec:conclusion}
%(Sam Schmidt, Ofer Lahav, Jeff Newman, Alex Malz, Eve Kovacs, Tony Tyson, Tina Peters)

%\item \red{inclusion of incompleteness effects on training sample, ``wrong redshifts'', filter curves}
%\item \red{inclusion of photometric errors, image based effects, blending, lensing, sky, mask, etc...}

In this study we have not accounted for the presence of Active Galactic Nuclei (AGN) contributions to galaxy fluxes. In some cases, AGN will be easily identified from the colors and morphologies, i.e. the case of the brightest quasars where the nuclear activity outshines the host galaxy, and numerous studies have utilized color selection to create large samples of quasars \citep[e.g.][]{Richards:06,Maddox:08,Richards:15}.  In current deep fields, similar in depth to what we expect from LSST, variability information and multi-wavelength data have been critical to not only identify AGN dominated galaxies, but also obtain more accurate photometric redshifts \citep[e.g][]{Salvato:11}.

In addition to AGN dominated galaxies, those with lower levels of nuclear activity present a more insidious problem, where AGN features may not be apparent, but the colors and other host galaxy properties are perturbed relative to galaxies with an inactive nucleus.  In such cases, the presence of the AGN may induce a bias if the template SEDs or empirical datasets do not include low-level AGN counterparts.
For LSST, we will need to identify and obtain accurate photometric redshifts of all types of AGN for a range of science goals, whether it is to eliminate such objects from cosmology experiments, or to use them with confidence, all the way through to understanding galaxy evolution and the role that AGN may play in influencing galaxy properties over cosmic time.

A promising route to classifying and obtaining accurate photometric redshifts for the AGN population is by combining machine learning with template-fitting techniques, as has recently been demonstrated by \citet{Duncan:18} for radio-selected AGN. This is because AGN are relatively easy to obtain spectroscopic redshifts for over all redshifts due to the strong emission lines that they exhibit, allowing very good training sets for machine learning algorithms to use. Whereas for those galaxies where the AGN is sub-dominant the galaxy templates are still adequate for obtaining reasonable photometric redshifts.

In addition to these improvements, the DESC Photo-$z$ group plans to look at all potential methods to combine the results from multiple photo-$z$ codes to improve $p(z)$ accuracy, similar to the work presented in \citet{Dahlen:13,Carrascokind:14,Duncan:18}.  Taking advantage of multiple algorithms that use observables in slightly different ways has shown promise, however we must be very conscious of whether a potential combination properly treats the covariance between the methods, given that they are estimating quantities based on the same underlying observables.
Several science cases wish to estimate physical quantities along with redshift, for example galaxy stellar mass and star formation rate.  Proper joint estimation of redshift and physical quantities requires an in depth understanding of galaxy evolution, and progress on accurate bivariate redshift probability distributions will go hand in had with progress on understanding galaxies themselves.  Parameterization and storage of a complex 2-dimensional probability surface for potentially billions of galaxies (or even subsets of hundreds of thousand of particular interest) pose a potential challenge.  These issues will be examined in another future paper.

Finally, while this paper and future papers discussed above focus on photometric redshift codes and estimating accurate $p(z)$ from training data, we plan a separate, but complementary, project to examine calibration of the resultant redshifts via spatial cross-correlations \citep{Newman:2008}, which will be explored in a separate series of future papers.  The overarching plan describing everything laid out in this section is described in more detail in the LSST DESC Science Roadmap (see Footnote in Section~\ref{sec:intro}).  These plans will require significant effort, but they are necessary if we are to make optimal use of the LSST data for astrophysical and cosmological analyses.
%\begin{enumerate}
%\item \red{the combination of codes}

%\item \red{$p(z,\alpha)$}
%\item \red{tomographic analysis later on}
%\red{this paper deals with training, future papers will also explore calibration via cross-correlation methods}
%\end{enumerate}
%\red{mention the SRM once again, say that the plans are extensive, but we do have a plan and a rough timeline.}

### 2018-08-20 paper/tex/discussion.tex

\section{Discussion}
\label{sec:discussion}
%(Sam Schmidt, Ofer Lahav, Jeff Newman, Alex Malz, Eve Kovacs, Tony Tyson, Tina Peters)
In this paper we presented results evaluating the photometric redshift PDF computation for eleven photo-$z$ codes.  As discussed in Section~\ref{sec:metrics} the $p(z)$ should accurately reflect the relative likelihood as a function of redshift for each galaxy.  All codes were provided a set of representative training data and tested on an idealized set of model galaxies with high signal-to-noise and photometry with no confounding effects due to blending, instrumental effects, the night sky, etc... included.  The goal was not to determine a ``best'' photo-$z$ code: in many ways, this was a baseline test of a ``best case scenario'' to predict the expected photo-$z$ performance if a stage IV dark energy survey was to obtain complete training samples and perfectly calibrated their multi-band photometry.  Given these idealized conditions, any deficiencies observed in a photo-$z$ code's performance should be a cause for concern, and may be evidence in a problem with either/both of the specific code implementation or the underlying algorithm.  In order to meet the stringent LSST requirements on photo-$z$ performance, identifying and correcting such problems is an important first step before tackling more realistic data in future challenges.  Most of the codes tested performed well, however, several did not meet the stringent goals that have been laid out for LSST photometric redshift performance.  This is a cause for concern, given the idealized conditions, and the individual code responses will be studied in detail moving forward.  One obvious trend in several of the codes tested was an overall over or underprediction of the widths of $p(z)$, as evidenced by the QQ plots and PIT histograms shown in Fig.~\ref{fig:pitqq}.  A more careful tuning of bandwidth or smoothing during the validation process appears to be necessary for many of the machine learning based codes in order to improve the accuracy of $p(z)$.  For narrow peaked $p(z)$ the parameterization of the PDF as evaluated on a fixed redshift grid could also have contributed to some overestimates of $p(z)$ width simply due to the finite resolution.  After evaluating resuts such as those presented in \citet[]{Malz:qp}, in future analyses we plan to switch from a fixed grid to quantile-based storage of $p(z)$ in order to more efficiently and accurately store redshift PDF results.

Another important factor to keep in mind when examining the results presented in this paper is the fact that they are at some level dependent on the metrics that we aim to optimize: in this case code participants were asked to submit their optimal measures of an accurate $p(z)$, so participants used the training/validation data to optimize their codes accordingly.  Had we, instead, asked for an optimal $\hat{N(z)}$ the resulting metrics would be different for most, if not all, of the codes, as they would optimize toward a different goal.  Specific metric choice can affect which codes are among the ``best'' codes.  As stated earlier, there are cosmological science cases that require either individual galaxy photo-$z$ measures, or ensemble $\hat{N}(z)$ measures.  We must be aware of that the optimal method for one is not necessarily optimal for the other, and in fact several photo-$z$ algorithms may be necessary in the final cosmological analysis in order to satisfy the requirements of all science use cases.  The example of the simple \trainz\ estimator described in Section~\ref{sec:caution} shows a simple model with a $p(z)$ that is unrealistic for individual objects can still score very well on many of our metrics.  It is important to look at {\it all} metrics, and keep in mind what information each metric conveys.
%\red{mention 'null' and how 'dumb' algorithms can score well on metrics, need to keep all metrics in mind}
%\aim{[reiterate the meaning of $p(z)$ and the goal -- emphasize finding ``best'' code in terms of impact of assumptions underlying the method \textit{or} establishing methodology for the realistic case of biased prior information and other science goals]}
%\red{[the challenges faced in this project]}
%\red{mention z$<$2, missing some important degeneracies, new sim will go to higher z}
%\red{no quality cuts included, could identify outliers, but could also induce biases}
%\red{[how optimization defined has impact, here we optimized p(z), not n(z), could be science-case specific if stringent requirements are needed, though that may be a computational/storage challenge if too many use cases needed.]}
%\red{some discussion of output parameterization, limitations of 200 point fixed grid, in future switch to something like quantiles (cite qp paper)}
We re-emphasize that the dataset tested was quite idealized, and discuss enhancements that will be added in future simulations to test photo-$z$ codes on increasingly realistic conditions in the following section.

\subsection{Future work}
\label{sec:futurework}
The work presented in this paper is only the first step in characterizing current photo-$z$ codes and moving toward an improved photometric redshift estimator.  This initial paper explored code performance in idealized conditions with perfect catalog-based photometry and representative training data.  As mentioned in Section~\ref{sec:stackedmetrics} for the stacked $N(z)$ metrics we examined only the entire galaxy population with no selections in either photo-$z$ ``quality" or redshift.  The cosmological analyses for weak lensing and large scale structure based measures plan to break galaxy samples into tomographic redshift bins, using photo-$z$ $p(z)$ to infer the redshift distribution for each bin.  The specific selection used to determine these bins, both algorithmically and the specific bin boundaries, could induce biases due to indirect selections inherent in the photo-$z$ or other bin selection parameters.  The effects of tomographic bin selection will be explored in a dedicated future paper. \red{[are there any references for this?  I remember Gary Bernstein talking about this at a photo-z workshop in Japan, but I don't know that it was published.  I believe Michael Troxel has discussed this as well.]}
We also plan to propagate the uncertainties measured in a set of fiducial tomographic redshift bins in order to estimate impact on cosmological parameter estimation.
%\item \red{cosmological parameters in DC2}
%\item \red{mention z$<$2, missing some important degeneracies, new sim will go to higher z}
%\red{no quality cuts included, could identify outliers, but could also induce biases}

In future papers we will add more and more complexity to our simulated data in order to test photo-$z$ algorithms in increasingly realistic conditions.  The most pressing concern is the impact of incomplete spectroscopic training samples.  As discussed extensively in \citet{Newman:2015} a representative set of spectroscopically confirmed galaxies spanning the full range of both redshift and apparent magnitude is necessary as a training set to characterize the mapping from broad-band fluxes to photometric redshifts.  However, due to a combination of factors due to both the galaxy SEDs and limitations of spectrographic instruments, redshift samples are known to be systematically incomplete, where certain galaxy types and redshift intervals fail to yield a redshift even at the longest integration times on current and near-future instruments.  The more representative the training data, the better the performance of photo-$z$ algorithms will be.  Current and upcoming surveys are putting in significant effort into obtaining these training samples \citep[e.~g.\,][]{Masters:2017}, however we still expect significant incompleteness for LSST-like samples, particularly at faint magnitudes.  One major focus of an upcoming {\it LSST Dark Energy Science Collaboration Photo-z Working Group} data challenge is to produce a realistically incomplete training set of spectroscopic galaxies, modeling the performance of spectrographs, emission-line properties, and expected signal-to-noise to determine which galaxies will fail to yield a secure redshift.  In addition to outright redshift failures we will model the inclusion of a small number of falsely identified secure redshifts where misidentified emission lines or noise spikes cause an incorrect redshift solution to be marked as a high quality identification.  Even sub-per cent level contamination by false redshifts can impact photo-$z$ solutions at levels comparable to the stringent  requirements of some LSST science cases.
We expect different systematics to occur in different photo-$z$ codes in response to training on incomplete data, particularly some of the machine learning methods.  The response of the codes will inform future directions of code development.

This initial paper explored a data set that was constructed at the catalog level, with no inclusion of the complications that come from measuring photometry from images.  Future data challenges will move to catalogs constructed from mock images, including effects that will have great impact on photo-$z$ measurements.  Object blending will be a major area of investigation, as the mixing of flux from multiple objects and the resultant change in measured colours is predicted to affect a large fraction of LSST galaxies \citep{Dawson:2016}, and will be one of the major contributing systematics for photo-$z$'s.  Inclusion of differing observing conditions (seeing, clouds, variations in filter curves, Galactic dust, ...), as well as models for instrumental and system effects, sky masks, will all impact object photometry, and will be explored in the upcoming data challenge and their impacts described in upcoming papers.
All underlying SEDs were parameterized as a weighted combination of five basis SEDs, with no additional accounting for host galaxy dust obscuration beyond what was encoded in the basis templates.  This, in effect, limited the simulation to a very simple model of internal obscuration.  Future simulations will include a more complicated and realistic treatment of host galaxy dust.

The underlying simulation used in this work was based on a light-cone constructed to a maximum redshift of $z=2$.  LSST imaging after 10 years of observations will include a significant number of $z>2$ galaxies in expected cosmology samples, and their inclusion does have potential significant implications for photo-$z$ measures: the high redshift galaxies lie at fainter apparent magnitudes and can have anomalous colours due to evolution of stellar populations and the shift to rest-frame magnitudes probing UV features of the underlying SED.  More importantly, one of the most common ``catastrophic outlier'' degeneracies observed in deep photometric samples occurs when the Lyman break is mistaken for the Balmer break, leading to multiple redshift solutions at $z\sim0.2-0.3$ and $z\sim2-3$ \citep{Massarotti:2001}.  This degeneracy, along with other potential degeneracies, are currently not covered by the limited redshift range of this initial paper, which could mean that we are not probing the full range of potential extreme outlier populations and how our photo-$z$ estimators respond to them. Extending simulations to include the high-redshift galaxy population will be a priority in future data challenges.

### 2018-08-20 paper/tex/introduction.tex

\section{Introduction}
\label{sec:intro}

%(Sam Schmidt, Ofer Lahav, Jeff Newman, Alex Malz, John Soo, Tony Tyson)

%\jan{Is the intention to keep these parenthetical author sets?  I find it a bit offputting in the text.  It might be better to put this for sections in the acknowledgements section, and for individual codes/statistics specify who ran/developed it with a sentence.}\red{I was just about to comment these out before finalizing the internal review draft of the paper. They are in place to remind me to contact a few people who still have not filled out their authors.csv info.}

Large-scale photometric galaxy surveys are entering a new era with currently or soon-to-be running Stage III and Stage IV dark energy experiments like the Dark Energy Survey \citep[DES,][]{Abbott:05}, the Kilo-Degree Survey \citep[KiDS,][]{de_Jong:13}, Large Synoptic Survey Telescope \citep[LSST,][]{Abell:09}, Euclid \citep{Laureijs:11}, and Wide-Field Infrared Survey Telescope \citep[WFIRST,][]{Green:12}.
The move to imaging based surveys, rather than spectroscopic based, for cosmological measurements makes proper understanding of photometric redshifts (``photo-$z$'s'') of paramount importance, as cosmological distance measures for statistical samples are directly dependent on photo-$z$ measurements.

The unprecedented sample size of LSST galaxies, expected to number several billion for the main cosmological sample, necessitates stringent constraints on photo-$z$ accuracy if systematic errors are not to dominate the statistical errors.
The LSST Science Requirements Document (SRD)\footnote{available at \url{https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17}} lists the photometric redshift goals for a magnitude limited sample with $i<25$ as: root-mean-square error with a goal of $\sigma_z<0.02(1+z)$; $3\sigma$ ``catastrophic outlier'' rate below $10\%$; bias below $0.003$\footnote{
Note that at the time the SRD was written, these goals were stated in terms of a photo-$z$ point estimate for each galaxy, as was standard in many previous studies, while in this paper we emphasize the importance of using a full photo-$z$ PDF.}.

The tremendous size of LSST's galaxy catalog will be enabled by its exceptional depth, pushing to fainter magnitudes and deeper imaging and including galaxies of lower luminosity and higher redshift than ever before.
In addition to the contribution of low signal-to-noise photometry to the systematic error of photo-$z$s, these populations introduce major physical degeneracies, for example the Lyman break/Balmer break degeneracy, that were not present in the populations covered in previous large area surveys like the Sloan Digital Sky Survey \citep[SDSS,][]{York:00} and the Two Micron All Sky Survey \citep[2MASS,][]{Skrutskie:06}.
In order to meet LSST's demanding error budget, it will be necessary to fully characterize those degeneracies wherein multiple redshift solutions have comparable likelihood.

There is often a desire to have a single valued ``point-estimate'' redshift for an individual galaxy.  However, the complex, non-linear (and often non-unique) nature of the mapping between broad band fluxes and redshift means that a single value is unable to capture the full redshift information encoded in a galaxy's magnitudes.  For example, a common point-estimate for a template-based method is taking the highest likelihood solution as the point photo-$z$.  A single valued redshift ignores degenerate redshift solutions of lower probability, potentially biasing photometric redshift estimates both for individual galaxies and ensemble distributions.  Storing more information is necessary, most often photo-$z$ codes output the redshift probability density function (PDF), also often referred to as $p(z)$, describing the relative likelihood as a function of redshift. %\red{should we move some of AIM's p(z) vs p(z|d,I) to here?}
Early template methods such as \citet{Fernandezsoto:99} converted relative $\chi^2$ values of template spectra to likelihoods to estimate $p(z)$.
Soon after, codes such as \citet{Benitez:00} added a Bayesian prior and output a posterior probability distribution.
While many early machine learning based algorithms focused on a point-estimate, \citet{Firth:03} used a neural net with 1000 realizations scattered within the photometric errors to estimate a $p(z)$.
As more groups began to employ photometric redshifts in their cosmological analyses, realization that point-estimate photo-$z$'s were inadequate for precision cosmology measurements \citep{Mandelbaum:2008}.
From around this point onward, most photo-$z$ algorithms have attempted to implement some estimate of the overall redshift probability in their outputs, and some surveys began supplying a full $p(z)$ rather than a simple redshift point-estimate and error \citep[e.~g.~][]{de_Jong:17}.

%\aim{We should spend a whole paragraph establishing that photo-$z$ PDFs are a thing, i.e. review the motivation for PDFs over point estimates, basic definition, early methods papers, surveys that have published them, etc.}
%\scc{I agree with Alex, we should introduce here what a pdf is and which survey have already published some pdfs (KiDS dr3 for example) or will produce them (LSST, Euclid etc.)}
%\red{See above paragraph, does this adequately address this?--SJS}\scc{fine for me}

There are numerous techniques for deriving photo-$z$ PDFs from photometry, yet no one method has yet been established as clearly superior.
Quantitative comparisons of photo-$z$ methods have been made before.
The Photo-$z$ Accuracy And Testing \citep[PHAT,][]{Hildebrandt:10} effort focused on point estimates derived from many photometric bands.
DES compared several codes for point estimates \citep{Sanchez:14} and a summary statistic of photo-$z$ interim posteriors for tomographically binned galaxy subsamples \citep{Bonnett:16}.
This paper is distinguished by its inclusion of metrics of photo-$z$ interim posteriors themselves and consideration of both classic and state-of-the-art photo-$z$ algorithms, comparing the performance of several of the most widely employed codes as well as some that have been developed only recently on the basis of metrics appropriate for a probabilistic data product.
The results presented in this work are
%part of the Data Challenge 1 (DC1)
a major focus of the Photometric Redshift working group of the LSST Dark Energy Science Collaboration (LSST-DESC).
This work is laid out in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}} as one of the critical activities to be completed in preparation for dark energy science analysis on the first year LSST data.
This is the first of multiple papers by the working group, which will grow in sophistication.
In this initial paper we focus on evaluating the performance of photometric redshift codes and their ability to produce accurate PDFs in the presence of representative training sets.
This can be thought of as an initial test under near perfect conditions, before further complexities are added in future papers.
Comparing the relative performance of the codes enables us to evaluate whether each code is using information in an optimal way, and may reveal enhancements in some codes and deficiencies in others, either in the fundamental algorithm, or in specific implementation.

Certain science cases need redshift information on individual objects, e.~g.~identification of host galaxy redshift for supernova classification, or identifying potential cluster membership.
Other science cases need only ensemble redshift information; for instance current weak lensing techniques require the overall redshift distribution $N(z)$ for tomographic redshift samples, but do not need single galaxy estimates.
In the case of the multiple types of probes of cosmology enabled by the
LSST cosmology sample of several billion galaxies, the number of redshift bins and their photo-$z$ requirements vary with the specific probe;  2-point angular correlations benefit from many bins, while weak lensing probes do not (due to the wide lensing kernel).
Large photometric surveys such as LSST must develop algorithms that meet the needs of all such science cases.
In order to meet these ambitious goals for photo-$z$ accuracy, every aspect of photo-$z$ estimation will have to be optimized: the algorithms employed, both template and machine-learning based (both in design and implementation); the spectroscopic data used as a training set for machine learning algorithms or to estimate template sets and train Bayesian priors; and probabilistic catalog compression schemes that balance information retention against limited storage resources.

Before moving forward, we must address how the best methods may be unique to the performance metrics and science cases considered and what distinguishes photo-$z$ PDFs of different methods from one another.
Though photo-$z$ PDFs are often written simply as $p(z)$, the PDF itself must be an interim posterior distribution $p(z | d, I)$, the probability of redshift conditioned on photometric data $d$ that has actually been observed and the prior information $I$ that guides how a redshift is extracted from the photometry.
If we run multiple photo-$z$ codes on a single dataset, the photo-$z$ interim posteriors will not be identical because each code is based on assumptions in the form of an interim prior --- these assumptions form the premise for photo-$z$ estimation as a whole and are the only way to introduce differences in estimates of what would otherwise be a shared photo-$z$ posterior $p(z | d)$ regardless of the code used to obtain it.
Though explicit knowledge of the interim prior is necessary to use photo-$z$ interim posteriors self-consistently in physical inference, the interim prior of a particular methodology is often implicit and not necessarily shared among all galaxies in the catalog.

This paper therefore aims (1) to constrain the impact of the interim prior $I$ by separating it into a component $I_{H}$ representing the method itself and a component $I_{D}$ representing physical information, such as a training set or SED template library and (2) to present a procedure for evaluating the performance of photo-$z$ codes in generic tests that may include many more systematics in the interim prior $I$.
In order to isolate the effects encapsulated by $I_{H}$ of variation between codes from issues with the training set or template library encapsulated by $I_{D}$, we use an identical set of simulated galaxies for every code and construct a template library and training sample that are {\it complete and representative} and shared among all codes; that is, our training sample for machine learning codes is drawn from the same underlying galaxy population as our test set, with no additional selections, and the SED library used for template-based codes is the same as the one used to generate the photometric data.
We explore a number of performance metrics in this paper, not to make a conclusion regarding the superiority or even relative favorability of each code but to establish a method for comparing photo-$z$ PDFs derived by different methods.
These test conditions set the stage for addressing in a future paper the crucial issue of incomplete and non-representative prior information.

The outline of the paper is as follows: in \S\,\ref{sec:sims} we present the simulated data set; in \S\,\ref{sec:pzcodes} we describe the current generation codes employed in the paper;  in \S\,\ref{sec:metrics} we discuss the interpretation of photo-$z$ PDFs in terms of metrics of accuracy; in \S\,\ref{sec:results} we show our results and compare the performance of the codes; in \S\,\ref{sec:discussion} we offer our conclusions and discuss future extensions of this work.

### 2018-08-20 paper/tex/methods.tex

\section{Methods}
\label{sec:pzcodes}

%(Sam Schmidt, Rongpu Zhou, Ibrahim Almosallam, Eric Nuss, Johann Cohen Tanugi, John Soo)

Here we outline the photo-$z$ PDF codes tested in this study. In total, eleven distinct codes are tested.  This sample is not comprehensive, but does cover a broad range of current-generation codes.  Both template-based and machine learning approaches are included and each are described separately in Secs.~\ref{sec:templatecodes} and \ref{sec:trainingcodes} respectively. The list of codes are summarized in Table.~\ref{tab:list_of_codes}.

\begin{table*}  %%% DATA TABLE %%%
\caption{List of photo-$z$ codes featured in this study. ML here means machine learning.} \label{tab:list_of_codes}\resizebox{\textwidth}{!}{
\begin{tabular}{llll}
\hline
\bf Code 			& \bf Type & \bf Paper & \bf Website \\
\hline
\textsc{BPz} 		& template 			& \citet{Benitez:00}		& \url{http://www.stsci.edu/~dcoe/BPZ/} \\
\textsc{EAZY} 		& template 			& \citet{Brammer:08}		& \url{https://github.com/gbrammer/eazy-photoz} \\
\textsc{LePhare} 	& template 			& \citet{Arnouts:99}		& \url{http://www.cfht.hawaii.edu/~arnouts/lephare.html} \\
\textsc{ANNz2} 		& ML 	& \citet{Sadeh:16}			& \url{https://github.com/IftachSadeh/ANNZ} \\
\textsc{Delight} 	& ML/template 	& \citet{Leistedt:17}		& \url{https://github.com/ixkael/Delight} \\
\multirow{ 2}{*}{\textsc{FlexZBoost}} & \multirow{ 2}{*}{ML} 	& \multirow{ 2}{*}{\citet{Izbicki:17}}		& \url{https://github.com/tpospisi/flexcode};\\
					& 		&							& \url{https://github.com/rizbicki/FlexCoDE}\\
%\textsc{Frankenz} 	& ML 	& \citet{Speagle:frankenz}	& \url{https://github.com/joshspeagle/frankenz} \\
\textsc{GPz} 		& ML	& \citet{Almosallam:15b}	& \url{https://github.com/OxfordML/GPz} \\
\textsc{METAPhoR} 	& ML 	& \citet{Cavuoti:17}		& \url{http://dame.dsf.unina.it}\\
\textsc{CMNN} 		& ML 	& \citet{Graham:17}			& - \\
\textsc{SkyNet} 	& ML 	& \citet{Graff:14}			& \url{http://ccpforge.cse.rl.ac.uk/gf/project/skynet/} \\
\textsc{TPZ} 		& ML 	& \citet{Carrasco_Kind:13}	& \url{https://github.com/mgckind/MLZ} \\
\hline
\trainz 	& N/A	& See Section~\ref{sec:method:trainz} & \\
\end{tabular}}
\end{table*}

%\red{The questions that must be answered for each code are\dots}
The questions that must be answered for each code are: what unique features are included in the specific implementation that influence the output $p(z)$.  What form of validation was performed with the training data, how were photometric uncertainties employed in the analysis, how were negative fluxes treated, what specific prior form was employed (for template based codes), or what specific machine learning architcture was used (for ML codes)?
%\scc{which parameter (in a general meaning) have been used, which kind of validation has been performed (if any)}
\subsection{Template-based Approaches}
\label{sec:templatecodes}

\subsubsection{BPZ}
\label{sec:BPZ}
%(Sam Schmidt)

\textsc{BPZ}\footnote{\url{http://www.stsci.edu/~dcoe/BPZ/}} \citep[Bayesian Photometric Redshift,][]{Benitez:00} is a template-based photo-$z$ code that compares the expected colors ($C$) calculated for a set of spectral energy distribution (SED) types/templates ($T$) to the observed colors to calculate the likelihood of observing colors at each redshift for each type, $p(C|z,T)$. The code employs an empirically determined Bayesian prior in apparent magnitude ($m_0$) and SED-type. Assuming that the SED-types are spanning and exclusive, we can determine the redshift posterior $p(z|C,m_0)$ by marginalizing over all SED-types with a simple sum \citep[Eq.~3 from][]{Benitez:00}:

\begin{equation} \label{eq:redshift_posterior}
p(z|C,m_0)\,\propto\, \sum_{T}p(z,T|m_0)\,p(C|z,T)
\end{equation}
\noindent where the first term on the right-hand side is the Bayesian prior and the second term is the traditional likelihood. The prior is assumed to have the form: $p(z,T|m_0)=p(T|m_0)\,p(z|T,m_0)$, i.e. it parameterizes the prior as an evolving type fraction with apparent magnitude, combined with a prior on the expected redshift probability distribution as a function of both apparent magnitude and SED-type.

In this paper we use \textsc{BPZ v 1.99.3}. The template set employed here is the set of $100$ discrete SEDs described in Section~\ref{sec:buzztemplates}
%, and the $129$ Brown SEDs for the Galacticus dataset.
To keep the number of free parameters to a manageable level the SEDs in the training set are sorted by the rest-frame $u$-$g$ colour and split into three ``broad'' SED classes, equivalent to the \texttt{E}, \texttt{Sp} and \texttt{Im/SB} types in \citet{Benitez:00}. We assume the same functional form for the Bayesian priors as used by \citet{Benitez:00}, and utilize the training-set galaxies with known SED-type, redshift, and apparent magnitude to determine the type fractions and the best fit for the eleven free parameters of the prior.
%For photo-$z$ point estimates we use the \texttt{Z\_B} output parameter.
For galaxies with negative flux in a measured band, the placeholder value is replaced with an estimate one $\sigma$ detection limit in that particular band, i.~e.~a value close to the estimated sky noise threshold.
The type-marginalized $p(z)$ is generated by setting the parameter \texttt{PROBS\_LITE=TRUE} in the \textsc{BPZ} parameter file.

\subsubsection{EAZY} \label{sec:eazy}
%(Rongpu Zhou)

\textsc{EAZY}\footnote{\url{https://github.com/gbrammer/eazy-photoz}} \citep[Easy and Accurate Photometric Redshifts from Yale,][]{Brammer:08} is a template-based photo-$z$ code that includes several features that improve on the basic $\chi^2$ fit used in many template codes. It can fit the observed photometry with SEDs created from a linear combination of a set of templates at each redshift, and the best-fit SED is found by simultaneously fitting one, two or all of the templates by minimizing $\chi^2$. The minimized $\chi^2(z)$ is then combined with an apparent magnitude prior to obtain the posterior redshift probability distribution, although some argue that this is not the mathematically correct way of calculating the posteriors. \textsc{EAZY} can also account for the uncertainties in the templates by adding an empirically derived template error in quadrature as a function of redshift to the flux errors.

In this paper we use the all-templates mode, which fits the photometric data with a linear combination of the five basis templates. We employed the $5$ basis templates described in Section~\ref{sec:buzzard}, and set the template error to zero since these same templates were used to produce the simulated catalog photometry. %\red{The quality parameter \texttt{Qz} is...}
The likelihoods are calculated on a 200-point redshift grid spanning $0\leq z \leq 2$, and include the application of a type-independent apparent magnitude prior estimated from the training data.

\subsubsection{LePhare}\label{sec:lephare}
%(C\'ecile Roucelle, Eric Nuss, Johann Cohen-Tanugi)

\textsc{LePhare}\footnote{\url{http://www.cfht.hawaii.edu/~arnouts/lephare.html}}\citep[Photometric Analysis for Redshift Estimate,][]{Arnouts:99,Ilbert:06} is a photo-$z$ reconstruction code based on a $\chi^2$ template-fitting procedure. The observed colors are matched with the colours predicted from a set of spectral energy distribution (SED) which can be either synthetic or based on a semi-empirical approach. \textsc{LePhare} has been used to produce the COSMOS2015 photo-$z$ catalogue \citep{Laigle:16}.

Each SED is redshifted in steps of $\Delta z = 0.01$ and convolved with the simulated LSST filter transmission curves (accounting for instrument efficiency). The opacity of the inter-galactic medium has been set to zero as no additional reddening has been included in the Buzzard simulations. The computed photo-$z$ is then the value that minimizes the merit function $\chi^2 (z,T,A)$ from \citet{Arnouts:99}:

\begin{equation}
\chi^2(z,T,A) = \sum_f^{N_f}\left(\frac{F^f_{\rm{obs}}−A \times F^f_{\rm{pred}}(T,z)}{\sigma^f_{\rm{obs}}}\right)^2
\end{equation}
\noindent where $A$ is a normalization factor, $F^f_{\rm{pred}}(T,z)$ is the flux predicted for a template $T$ at redshift $z$. $F^f_{\rm{obs}}$ is the observed flux in a given band $f$ and $\sigma^f_{\rm{obs}}$ the associated observational error. The index $f$ refers to the considered band and $N_f$ is the total number of filters.

In this paper we use \textsc{LePhare v 2.2}.
The set of templates used for fitting the photo-$z$'s are the 100 discrete Buzzard SED templates as described in section~\ref{sec:buzztemplates}.
%No correction for the photometric zero-points was applied for this work.
The full $p(z)$ corresponds to the likelihoods calculated at each point on our $z$-grid.

\subsection{Training-based Codes}
\label{sec:trainingcodes}

\subsubsection{ANNz2}
\label{sec:annz2}
%(John Soo)

\textsc{ANNz2}\footnote{\url{https://github.com/IftachSadeh/ANNZ}} \citep{Sadeh:16} is a powerful package that has the ability to employ several machine learning algorithms, including artificial neural networks (ANN), boosted decision tree (BDT) and k-nearest neighbour (KNN). Using the Toolkit for Multivariate Data Analysis (TMVA) with ROOT\footnote{\url{http://tmva.sourceforge.net/}}, it can run multiple machine learning algorithms for a single training and outputs photo-$z$'s based on a weighted average of their performances.

\textsc{ANNz2} is capable of producing both photo-$z$ point estimates and redshift posterior probability distributions $p(z)$, it could also conduct classifications and supports reweighting between samples. The PDFs are produced by propagating the intrinsic uncertainty on the input parameters and the uncertainty in the machine learning method to the expected photo-$z$ solution, averaged over multiple runs weighted based on the performance of each run. \textsc{ANNz2} presents its photo-$z$ uncertainty different from many codes by using the KNN method: it estimates the photo-$z$ bias between each object and a fixed number of nearest neighbours in parameter space, it then takes the $68$th percentile width of the distribution of the bias. This is based on the implication that objects with similar photometric properties would have similar uncertainties, and therefore the photometric errors of the inputs are not propagated into the code.

In this study, \textsc{ANNz2 v. 2.0.4} was used. The full PDF for each galaxy is also produced with a linear stepsize of $z=0.01$ for $0 \leq z \leq 2$. A set of $5$ ANNs with architecture $6:12:12:1$ ($6$ $ugrizy$ inputs, $2$ hidden layers with $12$ nodes each, and $1$ output) with different random seeds are used during each training. %Training was conducted on a subset of the given training sample which satisfies $i\le25.3$ as it would have improved the results compared to training on the entire sample.
Half of the training set is used as a validation set to prevent overtraining. All training objects are set to have detected magnitudes, however the non-detections (mag$=-99$) in the testing set are replaced with the mean of that particular band.

\subsubsection{Color-Matched Nearest-Neighbours}
\label{sec:cmnn}
%(Melissa Graham)

The nearest-neighbours color-matching photometric redshift estimator (\textsc{CMNN}) is presented in \cite[][herafter G18]{Graham:17}. This method uses a training set of galaxies with known redshifts that has equivalent or better photometry as the test set in terms of quality and filter coverage. For each galaxy in the test set we identify a color-matched subset of training galaxies, choose one (e.g. the nearest-neighbour or a random selection), and use its known redshift as the photo-$z$. This color-matched subset is identified by first calculating the Mahalanobis distance $D_M$ in color-space between the test galaxy and all training-set galaxies: the difference between the test and a training set galaxy's color divided by the photometric error, summed over all colors (i.e., $u$-$g$, $g$-$r$, $r$-$i$, $i$-$z$, and $z$-$y$). Then, the threshold value for $D_M$ that define a good color match is set by the percent point function (PPF): for example, for $N_{\rm dof}=5$, PPF$=95$ per cent of all training galaxies consistent with the test galaxy will have $D_M < 11.07$ (where $N_{\rm dof}$, the number of degrees of freedom, is the number of colors). For a given test galaxy, the $p(z)$ is the normalized distribution of the true catalogue redshifts of this color-matched subset of training galaxies, and the standard deviation of the color-matched subset is used as the photo-$z$ uncertainty.

We have applied the nearest-neighbours color-matching photometric redshift estimator described in G18 to the simulated data. Compared to its application in G18, there are some minor differences in the application of this estimator to the Buzzard catalogue. First, we do not impose non-detections on galaxies with a magnitude fainter than the expected LSST 10-year limiting magnitude or bright enough to saturate with LSST: {\it all} of the photometry for all the galaxies in the test and training sets are used for this experiment. Second, as in G18 we do apply an initial cut in color to the training set before calculating the Mahalanobis distance in order to accelerate processing, and also use a magnitude pseudo-prior to improve photo-$z$ estimates, but for both we have used different cut-off values that are appropriate for the Buzzard galaxies' colors and magnitudes. Third, we set different parameters for the identification of the color-matched subset of training galaxies and the selection of a photometric redshift estimate. In G18 we used a percent point function (PPF) value of 0.68 to identify the color-matched subset of training galaxies and used the redshift of nearest neighbour in color-space as the photo-$z$ estimate. These choices work well when the desire is to obtain accurate photo-$z$ estimates for most test-set galaxies, but does not return a robust $p(z)$ in all cases -- especially for galaxies that are bright and/or have few matches in color-space. Since a robust estimate of $p(z)$ is desired for this work we make several changes to our implementation of the \textsc{CMNN} photo-$z$ estimator. We continue to use a percent point function of ${\rm PPF} = 0.95$ to generate the subset of color-matched training galaxies, but weight them by the inverse of their Mahalanobis distance. This weighting maintains some of the accuracy that was previously achieved by simply using the nearest neighbour in color-space. We then use the weights to create the $p(z)$ instead of having the redshift of each color-matched training-set galaxy count equally. To obtain a robust estimate of the $p(z)$ for galaxies with a small number of color-matched training set galaxies, when this number is less than $20$ the nearest $20$ neighbours in color-space are used instead, and we convolve the $p(z)$ with a Gaussian with a standard deviation of:

\begin{equation}
\sigma = \sigma_{\rm train} \sqrt{({\rm PPF}_{20}/0.95)^2 -1}
\end{equation}

\noindent to appropriately broaden it so that the $p(z)$ for these test galaxies represents the enlarged PPF value associated with it. Overall, these three changes will yield poorer accuracy photo-$z$ compared to those presented in G18, but they will all have significantly more robust estimates of the $p(z)$, particularly for the brightest test galaxies. This is sufficient for this work because, as described in G18, the goal of the \textsc{CMNN} photo-$z$ estimator was never to provide the ``best'' (or even competitive) estimates in the first place, given its reliance on a deep training set, but rather to provide a means for direct comparisons between LSST photometric quality and photo-$z$ estimates. With this work we show how the input parameters should be set in order to return robust $p(z)$ estimates in addition to point value estimates.

\subsubsection{Delight}
\label{sec:delight}
%(John Soo)

\textsc{Delight}\footnote{\url{https://github.com/ixkael/Delight}} \citep{Leistedt:17} infers photo-$z$'s by using a data-driven model of latent SEDs and a physical model of photometric fluxes as a function of redshift. Generally, machine learning methods rely on representative training data with similar band passes, while template based methods rely on a complete library of templates based on physical models constructed. \textsc{Delight} is constructed in attempt to combine the advantages and eliminate the disadvantages of both template-based and machine learning algorithms: it constructs a large collection of latent SED templates (or physical flux-redshift models) from training data, with a template SED library as a guide to the learning of the model. The advantage of \textsc{Delight} is that it neither needs representative training data in the same photometric bands, nor does it need detailed galaxy SED models to work.

This conceptually novel approach is done by using Gaussian processes operating in flux-redshift space. The posterior distribution on the redshift of a target galaxy is obtained via a pairwise comparison with training galaxies,

\begin{equation}
p(z|\mathbf{\hat{F}}) \approx \sum_i p(\mathbf{\hat{F}}|z,t_i)\, p(z|t_i)p(t_i),
\end{equation}
\noindent where $p(z|t_i)p(t_i)$ captures prior information about the redshift distributions and abundances of the galaxies, with $t_i$ denoting the galaxy template; while $p(\mathbf{\hat{F}}|z,t_i)$ is the posterior of noisy flux $\mathbf{\hat{F}}$ at redshift $z$. For each training-target pair, $p(\mathbf{\hat{F}}|z,t_i)$ is evaluated as follows:

\begin{equation} \label{eq:delight_noisy}
p(\mathbf{\hat{F}}|z,t_i) = \int p(\mathbf{\hat{F}}|\mathbf{F})\, p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)\, d\mathbf{F},
\end{equation}
where $p(\mathbf{\hat{F}}|\mathbf{F})$ is the likelihood function, it compares the noisy real flux $\mathbf{\hat{F}}$ with the noiseless flux $\mathbf{F}$ obtained from the linear combination of template models, carefully constructed to account for model uncertainties and different normalization of the same SED; while $p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)$ is the prediction of flux at a different redshift $z$ with respect to the training object with redshift $z_i$ and flux $\mathbf{\hat{F}}_i$. Eq.~\ref{eq:delight_noisy} is essentially the probability that the training and the target galaxies having the same SED but at a different redshift. The flux prediction $p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)$ of the training galaxy at redshift $z$ is modeled via a Gaussian process,

\begin{equation} \label{eq:delight_gp}
F_b \sim \mathcal{GP}\left( \mu^F,k^F \right),
\end{equation}
\noindent with mean function $\mu^F$ and kernel $k^F$, both imposed to capture expected correlations resulting from the known underlying physics (i.e., fluxes resulting from observing SEDs through filter response, and the SEDs being redshifted). The reader should refer to \citet{Leistedt:17} for further details.

In this study, all $100$ ordered Buzzard templates, as described in Section~\ref{sec:buzztemplates}, were used in \textsc{Delight}, and the Gaussian process was trained with a subset of $50\,000$ galaxies. Photometric uncertainties from the inputs are propagated into the code, while non-detections for each band are set to the mean of the respective bands. Default settings of \textsc{Delight} were use, with the exception that the PDF bins were set to be linear instead of logarithmic, with $200$ equally-spaced bins between $0.0<z<2.0$. In this study a flat prior is assumed.

\subsubsection{FlexZBoost}
\label{sec:flexzboost}
%(Ann Lee, Rafael Izbicki, Taylor Pospisil, Peter Freeman)

\textsc{FlexZBoost}\footnote{\url{https://github.com/tpospisi/flexcode};\\ \url{https://github.com/rizbicki/FlexCoDE}} \citep{Izbicki:17} is a particular realization of \texttt{FlexCode}, which is a general-purpose methodology for converting any conditional mean point estimator of $z$ to a conditional {\em density} estimator $f(z \vert \x)$, where $\x$ here represents our photometric covariates and errors.\footnote{Instead of $p(z)$, we use the notation $f(z \vert \x)$ to explicitly show the dependence on $\x$.} The key idea is to expand the unknown function $f(z \vert \x)$ in an orthonormal basis $\{\phi_i(z)\}_{i}$:

\begin{equation}
f(z|\x)=\sum_{i}\beta_{i }(\x)\phi_i(z). \label{eq::series_expansion}
\end{equation}
By the orthogonality property, the expansion coefficients are just conditional means

\begin{equation}
\beta_{i }(\x) =  \mathbb{E}\left[\phi_i(z)|\x\right] \equiv \int f(z|\x)   \phi_i(z) dz. \label{eq:FlexZBoost_coeffs}
\end{equation}
These coefficients can easily be estimated from data by regression.

In this paper, we use \textsc{xgboost}~\citep{Chen:16} for the regression part as these techniques scale well for massive data; it should however be noted that \textsc{FlexCode-RF} (also on GitHub), based on Random Forests, generally performs better for smaller data sets. As our basis, we choose a standard Fourier basis. There are two tuning parameters in our $p(z)$ estimate: (i) the number of terms, $I$, in the series expansion in Eq.~\ref{eq::series_expansion}, and (ii) an exponent $\alpha$ that we use to sharpen the computed density estimates $\widehat{f}(z|\x)$, according to $\widetilde{f}(z|\x) \propto \widehat{f}(z|\x)^\alpha$. We split the ``train data'' into a training set (85\%) and a validation set (15\%), and choose both $I$ and $\alpha$ in an automated way by minimizing the weighted $L_2$-loss function \citep[Eq. 5 in][]{Izbicki:17} on the  validation set.

Although {\tt FlexCode} offers a {\em lossless compression} of the photo-$z$ estimates (in this study, one can reconstruct $\widetilde{f}(z|\x)$ exactly at any resolution from estimates of the first 35 coefficients, Eq.~\ref{eq:FlexZBoost_coeffs}, for a Fourier basis $\{\phi_i(z)\}_{i}$), we discretize our final estimates into 200 bins linearly spaced in $0 < z < 2$ for easy comparison with other algorithms. Using a higher resolution may yield better results (with no added cost in storage).

%\subsubsection{Frankenz}
%\label{sec:frankenz}
%(seeking volunteers)
%
%\red{\textsc{Frankenz}\footnote{\url{https://github.com/joshspeagle/frankenz}} \cite{Speagle:frankenz} is...}
%

\subsubsection{GPz}
\label{sec:gpz}
%(Ibrahim Almosallam)

\textsc{GPz}\footnote{\url{https://github.com/OxfordML/GPz}} \citep{Almosallam:16a,Almosallam:15b} is a sparse Gaussian process based code, a fast and a scalable approximation of full Gaussian Processes \citep{Rasmussen:06}, with the added feature of being able to produce input-dependent variance estimations (heteroscedastic noise). The model assumes that the probability of the output $y$, the redshift, given the input $x$, the photometry, is $p(y|x)=\mathcal{N}\left(y|\mu(x),\sigma(x)^{2}\right)$. The mean function, $\mu(x)$, and the variance function $\sigma(x)^{2}$ are both linear combinations of basis functions that take the following form:

\begin{align}
f(x)=\sum_{i=1}^{m}\phi_{i}(x)w_{i},
\end{align}
where $\left\{\phi_{i}(x)\right\}_{i=1}^{m}$ and $\left\{w_{i}\right\}_{i=1}^{m}$ are sets of $m$ basis functions and their associated weights respectively. Basis function models (BFM), for specific classes of basis functions such as the sigmoid or the squared exponential, have the advantage of being universal approximators, i.e. there exist a function of that form that can approximate any function, with mild assumptions, to any desired degree of accuracy. The details on how to learn the parameters of the model and the hyper-parameters of the basis functions are described in \citet{Almosallam:15b}.

A unique feature in \textsc{GPz}, is that the variance estimate is composed of two terms each quantifying a different source of uncertainty. One term (the model uncertainty) reflects how much of the uncertainty is due to lack of training samples at the location of interest, whereas the second term (the noise uncertainty) reflects how much of the uncertainty is caused from observing many noisy samples at that location. Thus, the predictive variance can determine whether we need more representative samples or more precise samples for any particular location in the input space. \textsc{GPz} can also emphasize the importance of some samples as weights. This weight can be for example $|z_{\rm{spec}}-z_{\rm{phot}}|/(1+z_{\rm{spec}})$ to target the desired objective of minimizing the normalized redshift error or as a function of their probability in the test set relative to the training set in order to pressure the model to better fit samples that are rare in the training set but are expected to be abundant during testing.

The data is prepared for \textsc{GPz} by taking the log of the magnitude errors, decorrelating the data set using PCA and imputing the missing values using a simple linear model that estimates the missing variables given the observed ones. The log transformation helps to smooth the long tail distribution of the magnitude errors, which is more stable numerically and makes the optimization process unconstrained. The missing values are imputed by computing the mean of the training set $\mu$ and its covariance $\Sigma$, then we use the following equation to estimate the missing values from the observed ones %\[

\begin{equation}
x_{u} = \mu_{u}+\Sigma_{uo}\Sigma_{oo}^{-1}(x_{o}-\mu_{o}),
\end{equation}  %\]
where the subscript $o$ in $x_{o}$ indexes the \emph{observed} part of the input $x$, whereas the subscript $u$ indexes the \emph{unobserved} set (similarly for $\mu$ and $\Sigma$). This is the optimal expected value of the unobserved variables given the observed ones if the distribution is jointly Gaussian, note that if the variables are independent, i.e. $\Sigma_{uo}=0$, this will reduce to a simple average predictor.

We use the Variable Covariance (VC) option in \textsc{GPz} with 200 basis functions after we note that there is no significant increase in the performance on the validation set (using 80\%-20\% training-validation split) and with no cost-sensitive learning applied.  We do note that \textsc{GPz} is trained using a data set that includes galaxies fainter than those in the test set, which was truncated at $i<25.3$.  This mismatch in training vs.~test results in a slightly different distribution in galaxy photometry and redshift distribution between the training and test set, which most likely degraded results compared to a run where the training set more closely matched the test set.

\subsubsection{METAPhoR}
\label{sec:metaphor}
%(Stefano Cavuoti, Massimo Brescia, Giuseppe Longo)
%\textcolor{brown}{general review done on this section}
\textsc{METAPhoR} \citep[Machine-learning Estimation Tool for Accurate Photometric Redshifts,][]{Cavuoti:17} is a pipeline designed to provide photo-$z$'s point estimates and a reliable PDF for machine learning (ML) based techniques. It includes pre- and post-processing phases, hosting a photo-$z$ prediction engine based on the Multi Layer Perceptron with Quasi Newton Algorithm (MLPQNA), already validated on photo-$z$'s in several cases \citep{de_Jong:17,Cavuoti:17b,Cavuoti:15,Brescia:14,Brescia:13,Biviano:13}. Due to its plug-in based modular nature, \textsc{METAPhoR} can be easily replaced by any other photo-$z$ prediction kernel, regardless its implementation, by taking the I/O interface compliance as unique constrain.

At a higher level, the pipeline mainly consists of three modules: (i) \textit{data pre-processing}, including a catalogue cross-matching sub-module \citep[based on the tool C3, ][]{Riccio:17}, a sub-module for photometric evaluation and error estimation of the multi-band catalogue used as Knowledge Base (KB), and a sub-module dedicated to the perturbation of the photometric KB, propaedeutic to the PDF estimation; (ii) \textit{photo-$z$ prediction}, which is the training/validation/test phase, producing the photo-$z$'s point estimates, based on a pre-selected ML method; (iii) \textit{PDF estimation}, specifically designed to calculate the PDF of the photo-$z$ estimation errors. The last module includes also a post-processing tool, providing some statistics on the produced point estimates and PDFs.

The photometry perturbation law is based on the formula $m_{ij} = m_{ij} + \alpha_{i}F_{ij}*u_{\mu=0,\sigma=1}$, where $\alpha_{i}$ is a user selected multiplicative constant (useful in case of multi-survey photometry), $u_{\mu=0,\sigma=1}$ is a random value from the standard normal distribution and $F_{ij}$ is a bimodal function (a constant function + polynomial fitting of the mean magnitude errors on the binned bands), heuristically tuned in such a way that the constant component is the threshold under which the polynomial function is considered too low to provide a significant noise contribution to the photometry perturbation.

As introduced, the photo-$z$ point estimate prediction engine of \textsc{METAPhoR} is based on the MLPQNA model, whose photo-$z$ regression training error, used by the quasi Newton learning rule, is based on the least square error and Tikhonov $L_{2}$-norm regularization \citep{Hofmann:18}.

As main prerogative, \textsc{METAPhoR} is able to provide a PDF for ML methods by taking into account the photometric errors provided with data, by running $N$ trainings on the same training set, or $M$ trainings on $M$ different random extractions from the KB. The different test sets, used to produce the PDF, are thus obtained by introducing a proper perturbation, parametrized from the photometric error distribution in each band, on the photometric data populating the original test set \citep{Brescia:18}.

For the present work since it was required to produce a redshift (and a PDF) for each object of the test set we decided to apply a hierarchical kNN to fill the missing detection, it goes without saying that for such points the reliability of PDFs and point estimation is lower. No cross validation has been used.

\subsubsection{SkyNet}
\label{sec:skynet}
%(J. Cohen-Tanugi and Hugo Tranin)

\textsc{SkyNet}\footnote{\url{http://ccpforge.cse.rl.ac.uk/gf/project/skynet/}} \citep{Graff:14} is a publicly available neural network software, based on a 2nd order conjugate gradient optimization scheme \citep[see][for further details]{Graff:14}. It has been used efficiently for redshift PDF estimates \citep{Sanchez:14,Bonnett:15,Bonnett:16}.

%In the present work, we use \textsc{SkyNet} both as a regressor (for redshift point estimates) and a classifier (for PDF estimates).
The neural network is configured as a standard multilayer perceptron with three hidden layers and one input layer with $12$ nodes (the $6$ magnitudes and their errors).
%For the regressor, the output layer is a single linear node, while each hidden layer has $10$ nodes with a tanh activation function.
The classifier is laid out such that the hidden layers have 20:40:40 nodes each, all rectified linear units, and the output layer has $200$ nodes (corresponding to 200 bins for the PDF) activated with a ``softmax'' function so that they automatically sum to $1$.

To avoid over-fitting, a $30$ per cent fraction of the training set is used as validation, and the training is stopped as soon as the error rate begins to increase in the validation set. The weights are randomly initialized based on normal sampling. The error function is a standard chi-square function for the regressor, and a cross-entropy function for the classifier. Finally, the data are all whitened before processing, with magnitudes pegged to (45,45,40,35,42,42) and their errors pegged to (20,20,10,5,15,15) for $ugrizy$ filters, respectively.

\subsubsection{TPZ}
\label{sec:tpz}
%(Erfan Nourbakhsh)

\textsc{TPZ}\footnote{\url{https://github.com/mgckind/MLZ}} \citep[Trees for Photo-$z$,][]{Carrasco_Kind:13,Carrascokind:14} is a parallel machine learning algorithm that generates photometric redshift PDFs using prediction trees and random forest techniques.
The code recursively splits the input data (i.~e.~the training sample), into two branches, one after another, until a terminal leaf is created that meets a termination criterion (e.~g.~a minimum leaf size or a variance threshold).
Bootstrap samples from the training data and associated errors are used to build a set of prediction trees.  In order to minimize correlation between the trees, the data is divided in such a way that the highest information gain among the random subsample of features is obtained at every point. The regions in each terminal leaf node corresponds to a specific subsample of the entire data that possesses similar properties.

The training data is examined before running TPZ. Since TPZ does not handle non-detections (magnitudes flagged as 99.0), we replace these values with an approximation of the $1\sigma$ detection threshold, i.~e.~a signal to noise ratio of 1 in terms of magnitude uncertainty using the equation $dm = 2.5 ~ \log ( 1 + N/S )$ where $dm \sim 0.7526 ~ mag$ for $N/S=1$.  That is, for each band, we replace the non-detection with the magnitude corresponding to the error of 0.7526 from the error model forecasted for 10-year LSST data. The Out-of-Bag \citep{Breiman:84,Carrasco_Kind:13} cross-validation technique is used within TPZ to evaluate its predictive validity and determine the relative importance of the different input attributes. We employed this information to calibrate our algorithm.

In the present work, the LSST magnitudes $u,~g,~r,~i$ and colors $u-g,~g-r,~r-i,~i-z,~z-y$ and their associated errors are used in the process of growing 100 trees with a minimum leaf size of 5 (the $z$ and $y$ magnitudes did not show significant correlation with the redshift in our cross-validation, so we did not use them when constructing our trees). We partitioned our redshift space into 100 bins from $z=0.005$ to $z=2.0$ and smoothed each individual PDF with a smoothing scale of twice the bin size.

\subsection{Simple Ensemble Estimator}\label{sec:method:trainz}
In addition to the main photo-$z$ algorithms described above we also include a very simple method.  For \trainz, as we will we call this simple estimator, we well define $p(z)$ as simply:

\begin{equation}
p(z) = \frac{1}{N_{train}}\sum_{i=1}^{N_{train}}z_{train}
\end{equation}
That is, we simply set the redshift PDF of every galaxy equal to the normalized $N(z)$ of the training sample.  As the training sample is drawn from the same underlying distribution as the test sample, modulo small deviations due to sample size, the quantiles of the training and test distributions should be identical.  This is a wildly unrealistic estimator, as it assigns all galaxies, no matter their apparent magnitude, colour, or true redshift, the same redshift PDF, and is thus uninformative at the level of individual object redshifts, but is designed to perform very well for the ensemble of all objects.  We will discuss this method and cautions relative to metrics in Section~\ref{sec:caution}.

### 2018-08-21 paper/tex/methods.tex
started fixing minor typos

@@
 The questions that must be answered for each code are: what unique features are included in the specific implementation that influence the output $p(z)$.
-What form of validation was performed with the training data, how were photometric uncertainties employed in the analysis, how were negative fluxes treated, what specific prior form was employed (for template based codes), or what specific machine learning architcture was used (for ML codes)?
+What form of validation was performed with the training data, how were photometric uncertainties employed in the analysis, how were negative fluxes treated, what specific prior form was employed (for template based codes), or what specific machine learning architecture was used (for ML codes)?
 %\scc{which parameter (in a general meaning) have been used, which kind of validation has been performed (if any)}
@@
 %For photo-$z$ point estimates we use the \texttt{Z\_B} output parameter.
-For galaxies with negative flux in a measured band, the placeholder value is replaced with an estimate one $\sigma$ detection limit in that particular band, i.~e.~a value close to the estimated sky noise threshold.
+For galaxies that are not detected in a measured band, the placeholder value is replaced with an estimate of the one $\sigma$ detection limit in that particular band, i.~e.~a value close to the estimated sky noise threshold.
 The type-marginalized $p(z)$ is generated by setting the parameter \texttt{PROBS\_LITE=TRUE} in the \textsc{BPZ} parameter file.
@@
 Each SED is redshifted in steps of $\Delta z = 0.01$ and convolved with the simulated LSST filter transmission curves (accounting for instrument efficiency).
-The opacity of the inter-galactic medium has been set to zero as no additional reddening has been included in the Buzzard simulations.
 The computed photo-$z$ is then the value that minimizes the merit function $\chi^2 (z,T,A)$ from \citet{Arnouts:99}:
@@
 
-\textsc{ANNz2} is capable of producing both photo-$z$ point estimates and redshift posterior probability distributions $p(z)$, it could also conduct classifications and supports reweighting between samples.
+\textsc{ANNz2} is capable of producing both photo-$z$ point estimates and redshift posterior probability distributions $p(z)$.
+It can also perform classifications, and supports reweighting between samples.
 The PDFs are produced by propagating the intrinsic uncertainty on the input parameters and the uncertainty in the machine learning method to the expected photo-$z$ solution, averaged over multiple runs weighted based on the performance of each run.
@@
 
-The nearest-neighbours color-matching photometric redshift estimator (\textsc{CMNN}) is presented in \cite[][herafter G18]{Graham:17}.
+The nearest-neighbours color-matching photometric redshift estimator (\textsc{CMNN}) is presented in \citet[][herafter G18]{Graham:17}.
 This method uses a training set of galaxies with known redshifts that has equivalent or better photometry as the test set in terms of quality and filter coverage.
@@
 
-In this study, all $100$ ordered Buzzard templates, as described in Section~\ref{sec:buzztemplates}, were used in \textsc{Delight}, and the Gaussian process was trained with a subset of $50\,000$ galaxies.
+In this study, all $100$ ordered Buzzard templates, as described in Section~\ref{sec:buzztemplates}, were used in \textsc{Delight}, and the Gaussian process was trained using the provided training sample.
 Photometric uncertainties from the inputs are propagated into the code, while non-detections for each band are set to the mean of the respective bands.
-Default settings of \textsc{Delight} were use, with the exception that the PDF bins were set to be linear instead of logarithmic, with $200$ equally-spaced bins between $0.0<z<2.0$.
+The default settings of \textsc{Delight} were used, with the exception that the PDF bins were set to be linear instead of logarithmic, with $200$ equally-spaced bins between $0.0<z<2.0$.
 In this study a flat prior is assumed.
@@
 %\textcolor{brown}{general review done on this section}
-\textsc{METAPhoR} \citep[Machine-learning Estimation Tool for Accurate Photometric Redshifts,][]{Cavuoti:17} is a pipeline designed to provide photo-$z$'s point estimates and a reliable PDF for machine learning (ML) based techniques.
+\textsc{METAPhoR} \citep[Machine-learning Estimation Tool for Accurate Photometric Redshifts,][]{Cavuoti:17} is a pipeline designed to provide photo-$z$ point estimates and a reliable PDF for machine learning (ML) based techniques.
 It includes pre- and post-processing phases, hosting a photo-$z$ prediction engine based on the Multi Layer Perceptron with Quasi Newton Algorithm (MLPQNA), already validated on photo-$z$'s in several cases \citep{de_Jong:17,Cavuoti:17b,Cavuoti:15,Brescia:14,Brescia:13,Biviano:13}.
-Due to its plug-in based modular nature, \textsc{METAPhoR} can be easily replaced by any other photo-$z$ prediction kernel, regardless its implementation, by taking the I/O interface compliance as unique constrain.
+Due to its plug-in based modular nature, \textsc{METAPhoR} can be easily replaced by any other photo-$z$ prediction kernel, regardless of specific implementation, while still taking advantage of the overall code interface.
 

### 2018-08-22 paper/tex/methods.tex
address a few more referee points

@@
 In total, eleven distinct codes are tested.
-This sample is not comprehensive, but does cover a broad range of current-generation codes.
+This sample is not comprehensive, codes were chosen based on the expertise available within the group; however, those chosen do cover a broad range of the current-generation codes.
 Both template-based and machine learning approaches are included and each are described separately in Secs.~\ref{sec:templatecodes} and \ref{sec:trainingcodes} respectively.
 The list of codes are summarized in Table.~\ref{tab:list_of_codes}.
+All code runners were asked to output redshift estimates on linear-spaced bins between redshifts 0 and 2.
 
@@
 %\red{The quality parameter \texttt{Qz} is...}
-The likelihoods are calculated on a 200-point redshift grid spanning $0\leq z \leq 2$, and include the application of a type-independent apparent magnitude prior estimated from the training data.
+The likelihoods include the application of a type-independent apparent magnitude prior estimated from the training data.
 
@@
 2.0.4} was used.
-The full PDF for each galaxy is also produced with a linear stepsize of $z=0.01$ for $0 \leq z \leq 2$.
 A set of $5$ ANNs with architecture $6:12:12:1$ ($6$ $ugrizy$ inputs, $2$ hidden layers with $12$ nodes each, and $1$ output) with different random seeds are used during each training.
@@
 Photometric uncertainties from the inputs are propagated into the code, while non-detections for each band are set to the mean of the respective bands.
-The default settings of \textsc{Delight} were used, with the exception that the PDF bins were set to be linear instead of logarithmic, with $200$ equally-spaced bins between $0.0<z<2.0$.
+The default settings of \textsc{Delight} were used, with the exception that the PDF bins were set to be linearly-spaced rather than logarithmic.
 In this study a flat prior is assumed.
@@
 In the present work, the LSST magnitudes $u,~g,~r,~i$ and colors $u-g,~g-r,~r-i,~i-z,~z-y$ and their associated errors are used in the process of growing 100 trees with a minimum leaf size of 5 (the $z$ and $y$ magnitudes did not show significant correlation with the redshift in our cross-validation, so we did not use them when constructing our trees).
-We partitioned our redshift space into 100 bins from $z=0.005$ to $z=2.0$ and smoothed each individual PDF with a smoothing scale of twice the bin size.
-
+We partitioned our redshift space into 200 bins and smoothed each individual PDF with a smoothing scale of twice the bin size.
+%there was a typo here that had 100 bins, I double checked, it is 200 in the final run.
 \subsection{Simple Ensemble Estimator}\label{sec:method:trainz}
@@
 \begin{equation}
-p(z) = \frac{1}{N_{train}}\sum_{i=1}^{N_{train}}z_{train}
+p(z) = \frac{1}{N_{ \mathrm train}}\sum_{\mathrm i=1}^{N_{\mathrm train}}z_{\mathrm train}
 \end{equation}

### 2018-08-29 paper/tex/methods.tex
edits on zgrid and other issues

@@
 The list of codes are summarized in Table.~\ref{tab:list_of_codes}.
-All code runners were asked to output redshift estimates on linear-spaced bins between redshifts 0 and 2.
+All code runners were asked to output redshift estimates on 200 linear-spaced bins between redshifts 0 and 2.
 
@@
 \label{sec:templatecodes}
+We test three publicly available and commonly used template-based codes: \textsc{BPZ}, \textsc{EAZY}, and \textsc{LePhare}.
+All three codes follow the standard procedure for template-based redshift estimation: calculate model fluxes for a set of template spectral energy distributions (SEDs) on a grid of redshift values and calculating a $\chi^2$ merit function using the observed and model fluxes:
+\begin{equation} \label{eq_temp_chi}
+\chi^2(z,T,A) = \sum_i^{N_{\rm filt}}\left(\frac{F^i_{\rm{obs}}−A \times F^i_{\rm{pred}}(T,z)}{\sigma^i_{\rm{obs}}}\right)^2
+\end{equation}
+\noindent  where $A$ is a normalization factor, $F^i_{\rm{pred}}(T,z)$ is the flux predicted for a template $T$ at redshift $z$.
+$F^i_{\rm{obs}}$ is the observed flux in a given band $i$ and $\sigma^i_{\rm{obs}}$ is the observed flux error.
+$N_{\rm tot}$ is the total number of filters, in our case the six $ugrizy$ LSST filters.
+Specific implementation details of each code, e.~g.~prior form and implementation, are described below.
 
@@
 \textsc{BPZ}\footnote{\url{http://www.stsci.edu/~dcoe/BPZ/}} \citep[Bayesian Photometric Redshift,][]{Benitez:00} is a template-based photo-$z$ code that compares the expected colors ($C$) calculated for a set of spectral energy distribution (SED) types/templates ($T$) to the observed colors to calculate the likelihood of observing colors at each redshift for each type, $p(C|z,T)$.
+The likelihoods at each redshift are related to the $\chi^2$ in Equation~\ref{eq_temp_chi} by the simple form: likelihood $\propto \, e^{-\chi^2/2}$.
 The code employs an empirically determined Bayesian prior in apparent magnitude ($m_0$) and SED-type.
@@
 The observed colors are matched with the colours predicted from a set of spectral energy distribution (SED) which can be either synthetic or based on a semi-empirical approach.
-\textsc{LePhare} has been used to produce the COSMOS2015 photo-$z$ catalogue \citep{Laigle:16}.
+%\textsc{LePhare} has been used to produce the COSMOS2015 photo-$z$ catalogue \citep{Laigle:16}.
 
-Each SED is redshifted in steps of $\Delta z = 0.01$ and convolved with the simulated LSST filter transmission curves (accounting for instrument efficiency).
-The computed photo-$z$ is then the value that minimizes the merit function $\chi^2 (z,T,A)$ from \citet{Arnouts:99}:
-
-\begin{equation}
-\chi^2(z,T,A) = \sum_f^{N_f}\left(\frac{F^f_{\rm{obs}}−A \times F^f_{\rm{pred}}(T,z)}{\sigma^f_{\rm{obs}}}\right)^2
-\end{equation}
-\noindent where $A$ is a normalization factor, $F^f_{\rm{pred}}(T,z)$ is the flux predicted for a template $T$ at redshift $z$.
-$F^f_{\rm{obs}}$ is the observed flux in a given band $f$ and $\sigma^f_{\rm{obs}}$ the associated observational error.
-The index $f$ refers to the considered band and $N_f$ is the total number of filters.
+Each SED is convolved with the simulated LSST filter transmission curves (accounting for instrument efficiency).
+The computed photo-$z$ is then the value that minimizes the merit function $\chi^2 (z,T,A)$ from \citet[]{Arnouts:99}, and given in Equation~\ref{eq_temp_chi}.
 
 In this paper we use \textsc{LePhare v 2.2}.
-The set of templates used for fitting the photo-$z$'s are the 100 discrete Buzzard SED templates as described in section~\ref{sec:buzztemplates}.
-%No correction for the photometric zero-points was applied for this work.
-The full $p(z)$ corresponds to the likelihoods calculated at each point on our $z$-grid.
+The set of templates used for fitting the photo-$z$'s are the 100 discrete Buzzard SED templates as described in section~\ref{sec:buzztemplates}, and the full $p(z)$ corresponds to the likelihoods calculated at each point on our $z$-grid.
 
@@
 \label{sec:trainingcodes}
+The training-based codes use a variety of algorithms in order to estimate $p(z)$. 
+Some aspects of data treatment were left to the individual code runners, for example, whether/how to split the available data with known redshifts into separate training and validation sets.
+The treatment of objects not detected in one or more bands was also left to the code runners.
+There are varying conventions among training-based codes, and no one prescription dominates in the photo-$z$ literature.
+This and other details of the code implementations are described below.
 
@@
 
-\textsc{ANNz2}\footnote{\url{https://github.com/IftachSadeh/ANNZ}} \citep{Sadeh:16} is a powerful package that has the ability to employ several machine learning algorithms, including artificial neural networks (ANN), boosted decision tree (BDT) and k-nearest neighbour (KNN).
+\textsc{ANNz2}\footnote{\url{https://github.com/IftachSadeh/ANNZ}} \citep{Sadeh:16} is a software package that has the ability to employ several machine learning algorithms, including artificial neural networks (ANN), boosted decision tree (BDT) and k-nearest neighbour (KNN).
 Using the Toolkit for Multivariate Data Analysis (TMVA) with ROOT\footnote{\url{http://tmva.sourceforge.net/}}, it can run multiple machine learning algorithms for a single training and outputs photo-$z$'s based on a weighted average of their performances.
@@
 Half of the training set is used as a validation set to prevent overtraining.
-All training objects are set to have detected magnitudes, however the non-detections (mag$=-99$) in the testing set are replaced with the mean of that particular band.
+All training objects are set to have detected magnitudes, however the non-detections (mag$=99$) in the testing set are replaced with the mean of that particular band.
 
@@
 
-\textsc{GPz}\footnote{\url{https://github.com/OxfordML/GPz}} \citep{Almosallam:16a,Almosallam:15b} is a sparse Gaussian process based code, a fast and a scalable approximation of full Gaussian Processes \citep{Rasmussen:06}, with the added feature of being able to produce input-dependent variance estimations (heteroscedastic noise).
+\textsc{GPz}\footnote{\url{https://github.com/OxfordML/GPz}} \citep{Almosallam:16a,Almosallam:15b} is a sparse Gaussian process based code, a scalable approximation of full Gaussian Processes \citep{Rasmussen:06}, with the added feature of being able to produce input-dependent variance estimations (heteroscedastic noise).
 The model assumes that the probability of the output $y$, the redshift, given the input $x$, the photometry, is $p(y|x)=\mathcal{N}\left(y|\mu(x),\sigma(x)^{2}\right)$.
@@
 \textsc{METAPhoR} \citep[Machine-learning Estimation Tool for Accurate Photometric Redshifts,][]{Cavuoti:17} is a pipeline designed to provide photo-$z$ point estimates and a reliable PDF for machine learning (ML) based techniques.
-It includes pre- and post-processing phases, hosting a photo-$z$ prediction engine based on the Multi Layer Perceptron with Quasi Newton Algorithm (MLPQNA), already validated on photo-$z$'s in several cases \citep{de_Jong:17,Cavuoti:17b,Cavuoti:15,Brescia:14,Brescia:13,Biviano:13}.
+It includes pre- and post-processing phases, hosting a photo-$z$ prediction engine based on the Multi Layer Perceptron with Quasi Newton Algorithm (MLPQNA).
+%, already validated on photo-$z$'s in several cases \citep{de_Jong:17,Cavuoti:17b,Cavuoti:15,Brescia:14,Brescia:13,Biviano:13}.
 Due to its plug-in based modular nature, \textsc{METAPhoR} can be easily replaced by any other photo-$z$ prediction kernel, regardless of specific implementation, while still taking advantage of the overall code interface.
@@
 \textsc{SkyNet}\footnote{\url{http://ccpforge.cse.rl.ac.uk/gf/project/skynet/}} \citep{Graff:14} is a publicly available neural network software, based on a 2nd order conjugate gradient optimization scheme \citep[see][for further details]{Graff:14}.
-It has been used efficiently for redshift PDF estimates \citep{Sanchez:14,Bonnett:15,Bonnett:16}.
+%It has been used efficiently for redshift PDF estimates \citep{Sanchez:14,Bonnett:15,Bonnett:16}.
 
@@
 %there was a typo here that had 100 bins, I double checked, it is 200 in the final run.
+
 \subsection{Simple Ensemble Estimator}\label{sec:method:trainz}
-In addition to the main photo-$z$ algorithms described above we also include a very simple method.  For \trainz, as we will we call this simple estimator, we well define $p(z)$ as simply:
+In addition to the main photo-$z$ algorithms described above we also include a very simple method as a pathological example.
+For \trainz, as we will we call this simple estimator, we well define $p(z)$ as simply:
 
@@
 This is a wildly unrealistic estimator, as it assigns all galaxies, no matter their apparent magnitude, colour, or true redshift, the same redshift PDF, and is thus uninformative at the level of individual object redshifts, but is designed to perform very well for the ensemble of all objects.
+If the training set was not representative, this estimator would produce biased results, and any attempts to break up the sample into tomographic bins will fail, as every galaxy has an identical $p(z)$.
 We will discuss this method and cautions relative to metrics in Section~\ref{sec:caution}.

### 2018-09-02 paper/tex/results.tex
fixed figure sizes
add scotts rule smoothing to Figure 4

--- 2018-08-19 paper/tex/results.tex
@@
 \includegraphics[width=0.49\textwidth]{fig/pz_12codes_261931_crop.jpg}\includegraphics[width=0.5\textwidth]{fig/pz_12codes_471167_crop.jpg}\\
-\includegraphics[width=\textwidth]{fig/pz_12codes_713178_crop.jpg}\includegraphics[width=\textwidth]{fig/pz_12codes_982747_crop.jpg}
+\includegraphics[width=0.49\textwidth]{fig/pz_12codes_713178_crop.jpg}\includegraphics[width=0.49\textwidth]{fig/pz_12codes_982747_crop.jpg}
 \caption{Four illustrative examples of individual p(z) distributions produced by the codes.
@@
 \textsc{FlexZBoost} & 0.0202\\
-\textsc{GPz} & 0.0068\\
+\textsc{GPz} & 0.0058\\
 \textsc{LePhare} & 0.0486\\
@@
 \centering
-\includegraphics[width=\textwidth]{fig/NZsumplot_12codes.jpg}
+\includegraphics[width=\textwidth]{fig/NZsumplot_12codes_scottsrule.jpg}
 \caption{The stacked $p(z)$ produced by each photo-$z$ code ($\hat{N}(z)$, red) compared to the spectroscopic redshift distribution ($N'(z)$, blue).
 Varying levels of small-scale structure are seen in the codes.
-$N'(z)$ is smoothed using a single bandwidth chosen via Scott's rule for all codes.} \label{fig:nz}
+Both $\hat{N}(z)$ and $N'(z)$ are smoothed using a single bandwidth chosen via Scott's rule for all codes.} \label{fig:nz}
 %\aim{Arguably this could also be clarified by showing the true $n'(z)$ once and then differences from it for each cod's stacked estimator\dots}\scc{what if we plot the difference between true and stacked rather than soverposing the two lines?
@@
 \textsc{FlexZBoost} & $-10.60$\\
-\textsc{GPz} 		& $-9.56$ \\
+\textsc{GPz} 		& $-9.93$ \\
 \textsc{LePhare} 	& $-1.66$ \\
@@
 \textsc{FlexZBoost} & 0.021 & 0.0022\\
-\textsc{GPz}        & 0.048 & 0.0042\\
+\textsc{GPz}        & 0.027 & 0.0042\\
 \textsc{LePhare}    & 0.028 & 0.0062\\
@@
 \textsc{FlexZBoost}& 0.694      & 0.610      & 0.631    \\
-\textsc{GPz}       & 0.692      & 0.605      & 0.619    \\
+\textsc{GPz}       & 0.696      & 0.615      & 0.639    \\
 \textsc{LePhare}   & 0.718      & 0.668      & 0.741    \\

### 2018-09-02 plotting_scripts/plot_nzsums_scottsrule.py

@@ -1,0 +1,88 @@
+import sys,os
+import numpy as np
+import matplotlib
+matplotlib.use('Agg')
+import matplotlib.pyplot as plt
+import seaborn as sns
+import scipy.ndimage.filters as scifilt
+
+def main(argv):
+    """
+    Sept 1, 2018: Add Scott's rule smoothing to all curves
+    Scott's rule bandwidth for scipy.stats.gaussian_kde that qp uses is 
+    given by   
+    n**(-1./(d+4)), so n^(-.2)                         
+    for 399,356 galaxies in interval [0,2] we have a bandwidth of 
+    sigma = 0.07581.  Given the array spacing of the stored NZPLOT_vectors is
+    0.001, this corresponds to a bandwidth of 75.8 pixels (and an extra fudge
+    factor of a 0.4 that is obviously necessary to match TrainZ that I don't 
+    understand)
+    """
+
+    smooth_sigma = 75.81*0.4 #see note in docstring on bandwidth choice
+    basepathpart = "./TESTDC1"
+    codes = ("ANNZ2","BPZ","DELIGHT","EAZY","FLEXZ","GPZ/AUG2018","LEPHARE",
+             "METAPHOR","NN","SKYNET","TPZ","NULL2")
+    labels = ("ANNz2","BPZ","Delight","EAZY","FlexZBoost","GPz","LePhare",
+              "METAPhoR","CMNN","SkyNet","TPZ","TrainZ")
+    labeldict = dict(zip(codes,labels))
+    
+    nzvectorfile = "NZPLOT_vectors.out"
+    
+    numcodes = len(codes)
+    numzs = 2001 #number of zbins in each NZPLOT_vectors.out file
+    szvals = np.zeros([numcodes,numzs])
+    pzvals = np.zeros([numcodes,numzs])
+    z_array = np.zeros(numzs)
+
+    for i,xfile in enumerate(codes):
+        direcpath = "%s%s"%(basepathpart,xfile)
+        fullpath = os.path.join(direcpath,nzvectorfile)
+        #print fullpath
+        data = np.loadtxt(fullpath,skiprows=1)
+        if i == 0:
+            z_array = data[:,0]
+        szvals[i,:] = data[:,1]
+        pzvals[i,:] = data[:,2]
+        smooth_pzvals = scifilt.gaussian_filter1d(pzvals,smooth_sigma,
+                                                  mode='constant')
[…]

### 2018-09-04 paper/tex/results.tex
add scotts rule smoothing to Figure 4
add/modify text describing Fig 4 update

@@
 \includegraphics[width=\textwidth]{fig/PITANDQQplot_12codes_crop.jpg}
-\caption{Summary plots for all eleven photo-$z$ codes illustrating performance for the interim posterior statistics.
+\caption{Summary plots for all twelve photo-$z$ codes illustrating performance for the interim posterior statistics.
 The top panel of each pair shows both the Quantile-Quantile (QQ) plot (red) and the histogram of PIT values (blue).
@@
 
-Fig.~\ref{fig:pz_examples} Shows the $p(z)$ produced by each of our eleven photo-$z$ codes for four example galaxies which exemplify some prominent cases that arise when estimating photo-$z$ PDFs: a narrow, unimodal redshift solution, a broader unimodal solution, a bimodal distribution, and a complex, multimodal distribution.
+Fig.~\ref{fig:pz_examples} Shows the $p(z)$ produced by each of our twelve photo-$z$ codes for four example galaxies which exemplify some prominent cases that arise when estimating photo-$z$ PDFs: a narrow, unimodal redshift solution, a broader unimodal solution, a bimodal distribution, and a complex, multimodal distribution.
 The red vertical line represents the true redshift of the individual galaxy, and the blue curve represents the redshift probability.
@@
 \textsc{ANNz2, METAPhoR, NN,} and \textsc{SkyNet} all show an excess of small-scale features, which appear to be print-through of the underlying training set galaxies.
+For example, in \textsc{CMNN} the $p(z)$ are a simply a weighted histogram of all spectroscopic training galaxies in nearby colour space with no smoothing applied, so the substructure is due to the finite number of neighbours, and is not unexpected.
 \textsc{GPZ} (in its current implementation), on the other hand, always produces a single Gaussian, which broadens to cover the multi-modal redshift solutions seen in other codes.
+%Interestingly, while \textsc{ANNz2} shows an abundance of small scale structure in individual $p(z)$ measurements (see Fig.~\ref{fig:pz_examples}), the summed $\hat{N}(z)$ is rather smooth, where the small scale features average out.
+This is not the case for the two other codes that show an abundance of substructure in their individual $p(z)$: both \textsc{CMNN} and \textsc{SkyNet} show small scale features both in $p(z)$ and $\hat{N}(z)$.
 
-As stated in Section~\ref{sec:metrics}, $p(z)$ is parameterized as $\approx 200$ piecewise constant bins covering $0<z<2$ for all eleven codes, giving a grid size of roughly $\delta z = 0.01$ for each code.
+As stated in Section~\ref{sec:metrics}, $p(z)$ is parameterized as $\approx 200$ piecewise constant bins covering $0<z<2$ for all twelve codes, giving a grid size of roughly $\delta z = 0.01$ for each code.
 A piecewise constant grid was a natural choice for some photo-$z$ codes, for instance most template-based codes compute likelihoods on a fixed grid.
@@
 
-Examination of the PIT histograms and QQ plots shows that there are fairly generic issues with the width of $p(z)$ uncertainties: \textsc{Delight, NN, SkyNet} and \textsc{TPZ} all show a PIT histogram with an dearth of low values and and an excess of high values, signs that, on average, their $p(z)$ are more broad than the true distribution of redshifts.
+Examination of the PIT histograms and QQ plots shows that there are fairly generic issues with the width of $p(z)$ uncertainties: \textsc{Delight, CMNN, SkyNet} and \textsc{TPZ} all show a PIT histogram with an dearth of low values and and an excess of high values, signs that, on average, their $p(z)$ are more broad than the true distribution of redshifts.
 \textsc{METAPhoR} shows the opposite trend, indicating the the $p(z)$  are more narrow than the distributions given by the true redshifts.
@@
 Fig.~\ref{fig:nz} shows the stacked $\hat{N}(z)$ distribution compared to the true redshift distribution $N'(z)$ for all tested codes.
-The red line indicates the summed $p(z)$ for each code, while the blue line shows the true redshift distribution smoothed via kernel density estimation (KDE), with a bandwidth chosen via Scott's rule \citep{Scott:1992}.
+The red line indicates the summed $p(z)$ for each code, while the blue line shows the true redshift distribution.
+All distributions are smoothed via kernel density estimation (KDE) with a common bandwidth chosen via Scott's rule \citep{Scott:1992} in order to minimize differences in small-scale features and make for a more uniform comparison between codes.
 While Scott's rule is used to display $N'(z)$ in the figure, all quantitative statistics are computed via the empirical CDF, and are thus unaffected by bandwidth/smoothing choice.
-Several of the codes show an excess at $z \sim 1.4$, particularly the template-based codes \textsc{BPZ}, \textsc{EAZY}, and \textsc{LePhare}.
-This is likely due to the $4000$ angstrom break passing through the gap between the $z$ and $y$ filters.
-This feature is one of the most prominent in individual galaxy p(z), and is readily seen in the point-estimate plots shown in Fig.~\ref{fig:pz_pointestimates} and described in the Appendix.
-Several of the machine learning based codes appear to be over-trained, adding excess galaxy probability to the redshift peaks and missing probability in the troughs.
-Given that our training data is drawn from the same galaxy population as the test set, and our data has prominent peaks in $N'(z)$, perhaps it is not unexpected that such overtraining occurs.
-A more extensive training/validation set might allow for a better choice of smoothing parameters in individual codes that would avoid such overtraining.
-
+%New text for the new figure
+As expected, \textsc{TrainZ} is in excellent agreement with true redshift distribution: as the training sample is selected from the same underlying distribution as the test set, the redshift distributions are identical, up to Poisson fluctuations due to the finite number of sample galaxies.
+\textsc{CMNN} is also in excellent agreement for similar reasons: with a representative training sample of galaxies spanning the colour-space, the sum of the colour-matched neighbour redshifts should return the true redshift distribution.
+\textsc{FlexZBoost} and \textsc{TPZ} also show very good agreement, with only slight departures, with an over/under-prediction in the high redshift tail of $\hat{N}(z)$ evident around $z\sim1.4$.
+In fact, several of the other codes show an excess at $z \sim 1.4$, particularly the template-based codes \textsc{BPZ}, \textsc{EAZY}, and \textsc{LePhare}.
+This is likely due to the $4000$ angstrom break passing through the gap between the $z$ and $y$ filters, resulting in a drastic change in $z-y$ colour for galaxies in this redshift range.
+The relative dearth of strong features blue-ward of the 4000 angstrom break seems to cause problems for the redshift range that corresponds to the break passing between the two reddest filters bandpasses of a survey.
+The $z\sim1.4$ feature is one of the most prominent sources of larger uncertainty in individual galaxy $p(z)$, and is readily seen in the point-estimate plots shown in Fig.~\ref{fig:pz_pointestimates} and described in the Appendix.
+%end of new text (and slightly modified text)
+Two of the machine learning based codes, \textsc{ANNz2} and \textsc{METAPhoR}, appear to be over-trained, adding excess galaxy probability to the redshift peaks and missing probability in the troughs.
+Given that our training data is drawn from the same galaxy population as the test set, and our data has prominent peaks in $N'(z)$, perhaps it is not unexpected that such overtraining occurs in some codes, though the fact that it does not occur in all training-based codes indicates that it may be due to specifics of the implementations of \textsc{ANNz2} and \textsc{METAPhoR}.
+A more extensive training/validation set might allow for a better choice of smoothing parameters in individual codes that would avoid such overtraining for these particular codes.
 %\aim{Move the text on Scott's Rule to Discussion?}
-%Scott's rule \citep{Scott:1992} was used to determine the smoothing bandwidth of the true redshift sample: that is, a single smoothing scale was chosen to represent the true redshift sample for all eleven codes.
-
-As with the $p(z)$ values in Figure~\ref{fig:pitqq}, different levels of substructure are obvious for the different codes.
-While Scott's rule provides a relatively good general smoothing scale to represent the true $N'(z)$, there are smaller scale fluctuations: while \textsc{FlexZBoost} and \textsc{CMNN} appear somewhat discrepant in Fig.~\ref{fig:nz}, they are actually the two most accurate in terms of their quantitative measurements.
+%Scott's rule \citep{Scott:1992} was used to determine the smoothing bandwidth of the true redshift sample: that is, a single smoothing scale was chosen to represent the true redshift sample for all twelve codes.
+%As with the $p(z)$ values in Figure~\ref{fig:pitqq}, different levels of substructure are obvious for the different codes.
+%While Scott's rule provides a relatively good general smoothing scale to represent the true $N'(z)$, there are smaller scale fluctuations: while \textsc{FlexZBoost} and \textsc{CMNN} appear somewhat discrepant in Fig.~\ref{fig:nz}, they are actually the two most accurate in terms of their quantitative measurements.
 %\textsc{SkyNet} also shows excess small-scale structure, but also shows significant overestimate of the high redshift galaxy population.
-Interestingly, while \textsc{ANNz2} shows an abundance of small scale structure in individual $p(z)$ measurements (see Fig.~\ref{fig:pz_examples}), the summed $\hat{N}(z)$ is rather smooth, where the small scale features average out.
-This is not the case for the two other codes that show an abundance of substructure in their individual $p(z)$: both \textsc{CMNN} and \textsc{SkyNet} show small scale features both in $p(z)$ and $\hat{N}(z)$.
-For \textsc{CMNN} the $p(z)$ are a simply a weighted histogram of all spectroscopic training galaxies in nearby colour space with no smoothing applied, so the substructure is not unexpected.
-The PIT histogram and shape of the QQ plot in Figure~\ref{fig:pitqq} show that \textsc{CMNN} is producing $p(z)$ that are overly broad, additional smoothing of the $p(z)$ would exacerbate this problem.
-While the $\hat{N}(z)$ plot shows more small scale features than other codes, these features are actually representative of real structure in the true $N'(z)$, as evidenced by the very good metric scores for \textsc{CMNN}.
-\textsc{SkyNet} $p(z)$ were also not smoothed: while previous implementations of the code such as \citet{Sanchez:14} and \citet{Bonnett:15} (see Appendix C.3) implement a ``sliding bin'' smoothing, no such procedure was used in this study.
-In addition to excess substructure, \textsc{SkyNet} shows an obvious redshift bias, evident both visually in Figure~\ref{fig:nz} and in the first moment of N(z) listed in Table~\ref{tab:moments}, where it is clearly an outlier.
+\textsc{SkyNet} shows an obvious redshift bias, evident both visually in Figure~\ref{fig:nz} and in the first moment of N(z) listed in Table~\ref{tab:moments}, where it is clearly an outlier.
 \textsc{SkyNet} employed a method where a random sample of training galaxies was chosen, but there was no test that the subset was completely representative of the overall redshift distribution.
-Also unlike \citet{Bonnett:15}, no effort was made to add extra weight to more rare low and high redshift galaxies.
+Unlike the previous implementation of \textsc{SkyNet} in \citet{Bonnett:15}, no effort was made to add extra weight to more rare low and high redshift galaxies.
 Either of these decisions could be the cause of the bias seen in our results.
@@
 \begin{center}
-\caption{Root-Mean-Square-Error (RMSE) statistics for the eleven photo-z codes for both PIT and $\hat{N}(z)$ distributions.}\label{tab:rmse}
+\caption{Root-Mean-Square-Error (RMSE) statistics for the twelve photo-z codes for both PIT and $\hat{N}(z)$ distributions.}\label{tab:rmse}
 \begin{tabular}{lcc}

### 2018-09-04 paper/tex/results.tex
add/modify text describing Fig 4 update
add clarification on overtraining

@@
 \textsc{FlexZBoost} & 0.0202\\
-\textsc{GPz} & 0.0058\\
+\textsc{GPz} & 0.0068\\
 \textsc{LePhare} & 0.0486\\
@@
 %end of new text (and slightly modified text)
-Two of the machine learning based codes, \textsc{ANNz2} and \textsc{METAPhoR}, appear to be over-trained, adding excess galaxy probability to the redshift peaks and missing probability in the troughs.
+Two of the machine learning based codes, \textsc{ANNz2} and \textsc{METAPhoR}, appear to be over-trained, adding excess galaxy probability to the redshift peaks with the largest number of training galaxies, and missing probability in the troughs where training galaxies are of fewer number.
 Given that our training data is drawn from the same galaxy population as the test set, and our data has prominent peaks in $N'(z)$, perhaps it is not unexpected that such overtraining occurs in some codes, though the fact that it does not occur in all training-based codes indicates that it may be due to specifics of the implementations of \textsc{ANNz2} and \textsc{METAPhoR}.
@@
 \textsc{FlexZBoost} & $-10.60$\\
-\textsc{GPz} 		& $-9.93$ \\
+\textsc{GPz} 		& $-9.56$ \\
 \textsc{LePhare} 	& $-1.66$ \\
@@
 \textsc{FlexZBoost} & 0.021 & 0.0022\\
-\textsc{GPz}        & 0.027 & 0.0042\\
+\textsc{GPz}        & 0.048 & 0.0042\\
 \textsc{LePhare}    & 0.028 & 0.0062\\
@@
 \textsc{FlexZBoost}& 0.694      & 0.610      & 0.631    \\
-\textsc{GPz}       & 0.696      & 0.615      & 0.639    \\
+\textsc{GPz}       & 0.692      & 0.605      & 0.619    \\
 \textsc{LePhare}   & 0.718      & 0.668      & 0.741    \\

### 2018-09-09 paper/tex/methods.tex
move note on smoothing from results to skynet section

--- 2018-08-21 paper/tex/methods.tex
@@
 The questions that must be answered for each code are: what unique features are included in the specific implementation that influence the output $p(z)$.
-What form of validation was performed with the training data, how were photometric uncertainties employed in the analysis, how were negative fluxes treated, what specific prior form was employed (for template based codes), or what specific machine learning architecture was used (for ML codes)?
+What form of validation was performed with the training data, how were photometric uncertainties employed in the analysis, how were negative fluxes treated, what specific prior form was employed (for template based codes), or what specific machine learning architcture was used (for ML codes)?
 %\scc{which parameter (in a general meaning) have been used, which kind of validation has been performed (if any)}
@@
 %For photo-$z$ point estimates we use the \texttt{Z\_B} output parameter.
-For galaxies that are not detected in a measured band, the placeholder value is replaced with an estimate of the one $\sigma$ detection limit in that particular band, i.~e.~a value close to the estimated sky noise threshold.
+For galaxies with negative flux in a measured band, the placeholder value is replaced with an estimate one $\sigma$ detection limit in that particular band, i.~e.~a value close to the estimated sky noise threshold.
 The type-marginalized $p(z)$ is generated by setting the parameter \texttt{PROBS\_LITE=TRUE} in the \textsc{BPZ} parameter file.
@@
 Each SED is redshifted in steps of $\Delta z = 0.01$ and convolved with the simulated LSST filter transmission curves (accounting for instrument efficiency).
+The opacity of the inter-galactic medium has been set to zero as no additional reddening has been included in the Buzzard simulations.
 The computed photo-$z$ is then the value that minimizes the merit function $\chi^2 (z,T,A)$ from \citet{Arnouts:99}:
@@
 
-\textsc{ANNz2} is capable of producing both photo-$z$ point estimates and redshift posterior probability distributions $p(z)$.
-It can also perform classifications, and supports reweighting between samples.
+\textsc{ANNz2} is capable of producing both photo-$z$ point estimates and redshift posterior probability distributions $p(z)$, it could also conduct classifications and supports reweighting between samples.
 The PDFs are produced by propagating the intrinsic uncertainty on the input parameters and the uncertainty in the machine learning method to the expected photo-$z$ solution, averaged over multiple runs weighted based on the performance of each run.
@@
 
-The nearest-neighbours color-matching photometric redshift estimator (\textsc{CMNN}) is presented in \citet[][herafter G18]{Graham:17}.
+The nearest-neighbours color-matching photometric redshift estimator (\textsc{CMNN}) is presented in \cite[][herafter G18]{Graham:17}.
 This method uses a training set of galaxies with known redshifts that has equivalent or better photometry as the test set in terms of quality and filter coverage.
@@
 
-In this study, all $100$ ordered Buzzard templates, as described in Section~\ref{sec:buzztemplates}, were used in \textsc{Delight}, and the Gaussian process was trained using the provided training sample.
+In this study, all $100$ ordered Buzzard templates, as described in Section~\ref{sec:buzztemplates}, were used in \textsc{Delight}, and the Gaussian process was trained with a subset of $50\,000$ galaxies.
 Photometric uncertainties from the inputs are propagated into the code, while non-detections for each band are set to the mean of the respective bands.
-The default settings of \textsc{Delight} were used, with the exception that the PDF bins were set to be linear instead of logarithmic, with $200$ equally-spaced bins between $0.0<z<2.0$.
+Default settings of \textsc{Delight} were use, with the exception that the PDF bins were set to be linear instead of logarithmic, with $200$ equally-spaced bins between $0.0<z<2.0$.
 In this study a flat prior is assumed.
@@
 We use the Variable Covariance (VC) option in \textsc{GPz} with 200 basis functions after we note that there is no significant increase in the performance on the validation set (using 80\%-20\% training-validation split) and with no cost-sensitive learning applied.
-We do note that \textsc{GPz} is trained using a data set that includes galaxies fainter than those in the test set, which was truncated at $i<25.3$.
+%We do note that \textsc{GPz} is trained using a data set that includes galaxies fainter than those in the test set, which was truncated at $i<25.3$.
 This mismatch in training vs.~test results in a slightly different distribution in galaxy photometry and redshift distribution between the training and test set, which most likely degraded results compared to a run where the training set more closely matched the test set.
@@
 %\textcolor{brown}{general review done on this section}
-\textsc{METAPhoR} \citep[Machine-learning Estimation Tool for Accurate Photometric Redshifts,][]{Cavuoti:17} is a pipeline designed to provide photo-$z$ point estimates and a reliable PDF for machine learning (ML) based techniques.
+\textsc{METAPhoR} \citep[Machine-learning Estimation Tool for Accurate Photometric Redshifts,][]{Cavuoti:17} is a pipeline designed to provide photo-$z$'s point estimates and a reliable PDF for machine learning (ML) based techniques.
 It includes pre- and post-processing phases, hosting a photo-$z$ prediction engine based on the Multi Layer Perceptron with Quasi Newton Algorithm (MLPQNA), already validated on photo-$z$'s in several cases \citep{de_Jong:17,Cavuoti:17b,Cavuoti:15,Brescia:14,Brescia:13,Biviano:13}.
-Due to its plug-in based modular nature, \textsc{METAPhoR} can be easily replaced by any other photo-$z$ prediction kernel, regardless of specific implementation, while still taking advantage of the overall code interface.
+Due to its plug-in based modular nature, \textsc{METAPhoR} can be easily replaced by any other photo-$z$ prediction kernel, regardless its implementation, by taking the I/O interface compliance as unique constrain.
 

### 2018-09-10 paper/tex/results.tex
add clarification on overtraining
rebase after GPz merge, edit text on z sim 1.4 range

@@
 %New text for the new figure
-As expected, \textsc{TrainZ} is in excellent agreement with true redshift distribution: as the training sample is selected from the same underlying distribution as the test set, the redshift distributions are identical, up to Poisson fluctuations due to the finite number of sample galaxies.
+As expected, \textsc{TrainZ} is in excellent agreement with the true redshift distribution: as the training sample is selected from the same underlying distribution as the test set, the redshift distributions are identical, up to Poisson fluctuations due to the finite number of sample galaxies.
 \textsc{CMNN} is also in excellent agreement for similar reasons: with a representative training sample of galaxies spanning the colour-space, the sum of the colour-matched neighbour redshifts should return the true redshift distribution.
@@
 In fact, several of the other codes show an excess at $z \sim 1.4$, particularly the template-based codes \textsc{BPZ}, \textsc{EAZY}, and \textsc{LePhare}.
-This is likely due to the $4000$ angstrom break passing through the gap between the $z$ and $y$ filters, resulting in a drastic change in $z-y$ colour for galaxies in this redshift range.
-The relative dearth of strong features blue-ward of the 4000 angstrom break seems to cause problems for the redshift range that corresponds to the break passing between the two reddest filters bandpasses of a survey.
-The $z\sim1.4$ feature is one of the most prominent sources of larger uncertainty in individual galaxy $p(z)$, and is readily seen in the point-estimate plots shown in Fig.~\ref{fig:pz_pointestimates} and described in the Appendix.
+This is likely due to the $4000\ {\rm \AA}$ break passing through the gap between the $z$ and $y$ filters, resulting in a drastic change in $z-y$ colour for galaxies in this redshift range.
+With a relative dearth of strong features blue-ward of the $4000\ {\rm \AA}$ break in most galaxy SEDs, the colour change in the two reddest filter bandpasses of a survey has a large influence on the redshift determination.
+The $z\sim1.4$ feature is one of the most prominent sources of larger uncertainty in individual galaxy $p(z)$.
+In our sample individual galaxy $p(z)$'s tend to be broader around $z\sim1.4$ and point estimates are more uncertain in this regime, as is readily seen in the point-estimate plots shown in Fig.~\ref{fig:pz_pointestimates} and described in the Appendix.
+This feature is not unique to this dataset, it is a common occurrence in photo-$z$ estimation.
+The fact that similar excesses appear in Figure~\ref{fig:nz} for \textsc{ANNz2} and \textsc{METAPhoR} shows that the effect is not limited to template-based codes.
+However, the lack of such a feature in the other codes shows that it is possible to eliminate the degeneracies.
+Further study on this issue may provide a solution for codes that suffer from this shortcoming.
+
 %end of new text (and slightly modified text)
@@
 This is something that will be looked into for \textsc{TPZ} in the future.
-
-It is also of note that all three template-based codes show an excess in their stacked $p(z)$ at $z\,\sim\,1.3-1.4$.
-This redshift range corresponds to the wavelengths where the $4000$ Angstrom break is passing between the borders of the $z$ and $y$ filters.
-This strong break entering the gap between the two reddest filters can cause problems with redshift estimation of individual galaxies, as can be seen in the point-estimate photo-$z$'s shown in Figure~\ref{fig:pz_pointestimates}.
-This is not unique to this dataset, it is a common occurrence in photo-$z$ estimation.
-The fact that similar excesses appear in Figure~\ref{fig:nz} for \textsc{ANNz2} and \textsc{METAPhoR} shows that the effect is not limited to template-based codes.
-However, the lack of such a feature in the other codes shows that it is possible to eliminate the degeneracies.
-Further study on this issue may provide a solution for codes that suffer from this shortcoming.
 

### 2018-09-27 paper/tex/introduction.tex
start of introduction edits

@@
 
-Large-scale photometric galaxy surveys are entering a new era with currently or soon-to-be running Stage III and Stage IV dark energy experiments like the Dark Energy Survey \citep[DES,][]{Abbott:05}, the Kilo-Degree Survey \citep[KiDS,][]{de_Jong:13}, Large Synoptic Survey Telescope \citep[LSST,][]{Abell:09}, Euclid \citep{Laureijs:11}, and Wide-Field Infrared Survey Telescope \citep[WFIRST,][]{Green:12}.
+Large-scale photometric galaxy surveys are entering a new era with currently or soon-to-be running Stage III and Stage IV dark energy experiments like the Dark Energy Survey \citep[DES,][]{Abbott:05}, the Kilo-Degree Survey \citep[KiDS,][]{de_Jong:13}, Hyper Suprime-Cam (HSC) Survey \citep[]{Aihara:2018a,Aihara:2018b}, Large Synoptic Survey Telescope \citep[LSST,][]{Abell:09}, Euclid \citep{Laureijs:11}, and Wide-Field Infrared Survey Telescope \citep[WFIRST,][]{Green:12}.
 The move to imaging based surveys, rather than spectroscopic based, for cosmological measurements makes proper understanding of photometric redshifts (``photo-$z$'s'') of paramount importance, as cosmological distance measures for statistical samples are directly dependent on photo-$z$ measurements.
+
+For cosmological measurements, certain science cases require redshift information on individual objects, e.~g.~identification of host galaxy redshift for supernova classification, or identifying potential cluster membership.
+Other science cases seem to need only ensemble redshift information; for instance current weak lensing techniques require only the overall redshift distribution $N(z)$ for tomographic redshift samples.
+However, recent data-driven techniques employing hierarchical Bayesian or Gaussian Process methods have emerged that calibrate redshift distributions using individual $p(z)$ estimates \citep[e.~g.~][]{Sanchez:2018}.
+These methods assume that the $p(z)$ for each galaxy is an accurate PDF, and such methods break down if this assumption is invalid.
+Thus, even methods that seem to need only ensemble $N(z)$ may actually require accurate $p(z)$ in order to meet stringent survey requirements.
+%In the case of the multiple types of probes of cosmology enabled by the
+%LSST cosmology sample of several billion galaxies, the number of redshift bins and their photo-$z$ requirements vary with the specific probe;  2-point angular correlations benefit from many bins, while weak lensing probes do not (due to the wide lensing kernel).
+Large photometric surveys such as LSST must develop algorithms that simultaneously meet the needs of all science cases.
+In order to meet these ambitious goals for photo-$z$ accuracy, every aspect of photo-$z$ estimation will have to be optimized: the algorithms employed, both template and machine-learning based (both in design and implementation); the spectroscopic data used as a training set for machine learning algorithms or to estimate template sets and train Bayesian priors; and probabilistic catalog compression schemes that balance information retention against limited storage resources.
 
@@
 The tremendous size of LSST's galaxy catalog will be enabled by its exceptional depth, pushing to fainter magnitudes and deeper imaging and including galaxies of lower luminosity and higher redshift than ever before.
-In addition to the contribution of low signal-to-noise photometry to the systematic error of photo-$z$s, these populations introduce major physical degeneracies, for example the Lyman break/Balmer break degeneracy, that were not present in the populations covered in previous large area surveys like the Sloan Digital Sky Survey \citep[SDSS,][]{York:00} and the Two Micron All Sky Survey \citep[2MASS,][]{Skrutskie:06}.
+The inclusion of these populations introduce major physical degeneracies, for example the Lyman break/Balmer break degeneracy, that were not present in the populations covered in shallower large area surveys like the Sloan Digital Sky Survey \citep[SDSS,][]{York:00} and the Two Micron All Sky Survey \citep[2MASS,][]{Skrutskie:06}.
 In order to meet LSST's demanding error budget, it will be necessary to fully characterize those degeneracies wherein multiple redshift solutions have comparable likelihood.
@@
 While many early machine learning based algorithms focused on a point-estimate, \citet{Firth:03} used a neural net with 1000 realizations scattered within the photometric errors to estimate a $p(z)$.
-As more groups began to employ photometric redshifts in their cosmological analyses, realization that point-estimate photo-$z$'s were inadequate for precision cosmology measurements \citep{Mandelbaum:2008}.
+As more groups began to employ photometric redshifts in their cosmological analyses, there was a realization that point-estimate photo-$z$'s were inadequate for precision cosmology measurements \citep{Mandelbaum:2008}.
 From around this point onward, most photo-$z$ algorithms have attempted to implement some estimate of the overall redshift probability in their outputs, and some surveys began supplying a full $p(z)$ rather than a simple redshift point-estimate and error \citep[e.~g.~][]{de_Jong:17}.
@@
 The Photo-$z$ Accuracy And Testing \citep[PHAT,][]{Hildebrandt:10} effort focused on point estimates derived from many photometric bands.
+\citet{Rau:2015} introduced a new method for improving redshift PDFs using an ordinal classification algorithm.
 DES compared several codes for point estimates \citep{Sanchez:14} and a summary statistic of photo-$z$ interim posteriors for tomographically binned galaxy subsamples \citep{Bonnett:16}.  
@@
 Comparing the relative performance of the codes enables us to evaluate whether each code is using information in an optimal way, and may reveal enhancements in some codes and deficiencies in others, either in the fundamental algorithm, or in specific implementation.
-
-Certain science cases need redshift information on individual objects, e.~g.~identification of host galaxy redshift for supernova classification, or identifying potential cluster membership.
-Other science cases need only ensemble redshift information; for instance current weak lensing techniques require the overall redshift distribution $N(z)$ for tomographic redshift samples, but do not need single galaxy estimates.
-In the case of the multiple types of probes of cosmology enabled by the
-LSST cosmology sample of several billion galaxies, the number of redshift bins and their photo-$z$ requirements vary with the specific probe;  2-point angular correlations benefit from many bins, while weak lensing probes do not (due to the wide lensing kernel).
-Large photometric surveys such as LSST must develop algorithms that meet the needs of all such science cases.
-In order to meet these ambitious goals for photo-$z$ accuracy, every aspect of photo-$z$ estimation will have to be optimized: the algorithms employed, both template and machine-learning based (both in design and implementation); the spectroscopic data used as a training set for machine learning algorithms or to estimate template sets and train Bayesian priors; and probabilistic catalog compression schemes that balance information retention against limited storage resources.
 

### 2018-09-28 paper/tex/introduction.tex
few more edits, add DESC SRD ref

@@
 Note that at the time the SRD was written, these goals were stated in terms of a photo-$z$ point estimate for each galaxy, as was standard in many previous studies, while in this paper we emphasize the importance of using a full photo-$z$ PDF.}.
-
+The LSST Dark Energy Science Collaboration (LSST-DESC) developed a separate Science Requirements Document \citep{Mandelbaum:2018}, which forecasts the constraining power of five cosmological probes using somewhat conservative assumptions to define requirements on systematic errors for several measurements.
+These include even more stringent requirements on photometric redshift performance than those included in the LSST SRD, though the initial LSST-DESC requirements on tomographic bin populations.
 The tremendous size of LSST's galaxy catalog will be enabled by its exceptional depth, pushing to fainter magnitudes and deeper imaging and including galaxies of lower luminosity and higher redshift than ever before.
 The inclusion of these populations introduce major physical degeneracies, for example the Lyman break/Balmer break degeneracy, that were not present in the populations covered in shallower large area surveys like the Sloan Digital Sky Survey \citep[SDSS,][]{York:00} and the Two Micron All Sky Survey \citep[2MASS,][]{Skrutskie:06}.
-In order to meet LSST's demanding error budget, it will be necessary to fully characterize those degeneracies wherein multiple redshift solutions have comparable likelihood.
+In order to meet the demanding error budgets of LSST and LSST-DESC it will be necessary to fully characterize those degeneracies wherein multiple redshift solutions have comparable likelihood.
 
@@
 %part of the Data Challenge 1 (DC1)
-a major focus of the Photometric Redshift working group of the LSST Dark Energy Science Collaboration (LSST-DESC).
+a major focus of the Photometric Redshift working group of the LSST-DESC.
 This work is laid out in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}} as one of the critical activities to be completed in preparation for dark energy science analysis on the first year LSST data.
 This is the first of multiple papers by the working group, which will grow in sophistication.
-In this initial paper we focus on evaluating the performance of photometric redshift codes and their ability to produce accurate PDFs in the presence of representative training sets.
-This can be thought of as an initial test under near perfect conditions, before further complexities are added in future papers.
+In this initial paper we focus on evaluating the performance of photometric redshift codes and PDF-based performance metrics in the presence of representative training sets.
 Comparing the relative performance of the codes enables us to evaluate whether each code is using information in an optimal way, and may reveal enhancements in some codes and deficiencies in others, either in the fundamental algorithm, or in specific implementation.
+We note that these initial tests are a necessary requirement for photo-$z$ codes that will be used in cosmological analyses; however, meeting these requirements is only the first stage in the process of accounting for incomplete training samples, object blending, and other complications that realistic samples will encounter in estimating photometric redshifts.
+This can be thought of as an initial test under near perfect conditions before further complexities are added in future papers.
 

### 2018-09-28 paper/tex/introduction.tex
more intro edits

@@
 
+The unprecedented sample size of LSST galaxies, expected to number several billion for the main cosmological sample, necessitates stringent constraints on photo-$z$ accuracy if systematic errors are not to dominate the statistical errors.
+The LSST Science Requirements Document (SRD)\footnote{available at \url{https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17}} lists the photometric redshift goals for a magnitude limited sample with $i<25$ as: root-mean-square error with a goal of $\sigma_z<0.02(1+z)$; $3\sigma$ ``catastrophic outlier'' rate below $10\%$; bias below $0.003$\footnote{
+Note that at the time the SRD was written, these goals were stated in terms of a photo-$z$ point estimate for each galaxy, as was standard in many previous studies, while in this paper we emphasize the importance of using a full photo-$z$ PDF.}.
+The LSST Dark Energy Science Collaboration (LSST-DESC) developed a separate Science Requirements Document \citep{Mandelbaum:2018}, which forecasts the constraining power of five cosmological probes using somewhat conservative assumptions to define requirements on systematic errors for several measurements.
+These include even more stringent requirements on photometric redshift performance than those included in the LSST SRD, though the initial LSST-DESC requirements on tomographic bin populations.
+The tremendous size of LSST's galaxy catalogue will be enabled by its exceptional depth, pushing to fainter magnitudes and deeper imaging and including galaxies of lower luminosity and higher redshift than ever before.
+The inclusion of these populations introduce major physical degeneracies, for example the Lyman break/Balmer break degeneracy, that were not present in the populations covered in shallower large area surveys like the Sloan Digital Sky Survey \citep[SDSS,][]{York:00} and the Two Micron All Sky Survey \citep[2MASS,][]{Skrutskie:06}.
+These issues are not unique to LSST, and are present in Stage III Dark Energy surveys; however,
+in order to meet the demanding error budgets of Stage IV projects such as LSST and LSST-DESC it will be necessary to fully characterize those degeneracies wherein multiple redshift solutions have comparable likelihood to per cent level accuracy.
+
+There is often a desire to have a single valued ``point-estimate'' redshift for an individual galaxy.
+However, the complex, non-linear (and often non-unique) nature of the mapping between broad band fluxes and redshift means that a single value is unable to capture the full redshift information encoded in a galaxy's magnitudes.
+For example, a common point-estimate for a template-based method is taking the highest likelihood solution as the point photo-$z$.
+A single valued redshift ignores degenerate redshift solutions of lower probability, potentially biasing photometric redshift estimates both for individual galaxies and ensemble distributions.
+Storing more information is necessary, most often photo-$z$ codes output the redshift probability density function, also often referred to as $p(z)$, describing the relative likelihood as a function of redshift.
+%\red{should we move some of AIM's p(z) vs p(z|d,I) to here?}
+Early template methods such as \citet{Fernandezsoto:99} converted relative $\chi^2$ values of template spectra to likelihoods to estimate $p(z)$.
+Soon after, codes such as \citet{Benitez:00} added a Bayesian prior and output a posterior probability distribution.
+While many early machine learning based algorithms focused on a point-estimate, \citet{Firth:03} used a neural net with 1000 realizations scattered within the photometric errors to estimate a $p(z)$.
+As more groups began to employ photometric redshifts in their cosmological analyses, there was a realization that point-estimate photo-$z$'s were inadequate for precision cosmology measurements \citep{Mandelbaum:2008}.
+From around this point onward, most photo-$z$ algorithms have attempted to implement some estimate of the overall redshift probability in their outputs, and some surveys began supplying a full $p(z)$ rather than a simple redshift point-estimate and error \citep[e.~g.~][]{de_Jong:17}.
+
 For cosmological measurements, certain science cases require redshift information on individual objects, e.~g.~identification of host galaxy redshift for supernova classification, or identifying potential cluster membership.
@@
 Large photometric surveys such as LSST must develop algorithms that simultaneously meet the needs of all science cases.
-In order to meet these ambitious goals for photo-$z$ accuracy, every aspect of photo-$z$ estimation will have to be optimized: the algorithms employed, both template and machine-learning based (both in design and implementation); the spectroscopic data used as a training set for machine learning algorithms or to estimate template sets and train Bayesian priors; and probabilistic catalog compression schemes that balance information retention against limited storage resources.
-
-The unprecedented sample size of LSST galaxies, expected to number several billion for the main cosmological sample, necessitates stringent constraints on photo-$z$ accuracy if systematic errors are not to dominate the statistical errors.
-The LSST Science Requirements Document (SRD)\footnote{available at \url{https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17}} lists the photometric redshift goals for a magnitude limited sample with $i<25$ as: root-mean-square error with a goal of $\sigma_z<0.02(1+z)$; $3\sigma$ ``catastrophic outlier'' rate below $10\%$; bias below $0.003$\footnote{
-Note that at the time the SRD was written, these goals were stated in terms of a photo-$z$ point estimate for each galaxy, as was standard in many previous studies, while in this paper we emphasize the importance of using a full photo-$z$ PDF.}.
-The LSST Dark Energy Science Collaboration (LSST-DESC) developed a separate Science Requirements Document \citep{Mandelbaum:2018}, which forecasts the constraining power of five cosmological probes using somewhat conservative assumptions to define requirements on systematic errors for several measurements.
-These include even more stringent requirements on photometric redshift performance than those included in the LSST SRD, though the initial LSST-DESC requirements on tomographic bin populations.
-The tremendous size of LSST's galaxy catalog will be enabled by its exceptional depth, pushing to fainter magnitudes and deeper imaging and including galaxies of lower luminosity and higher redshift than ever before.
-The inclusion of these populations introduce major physical degeneracies, for example the Lyman break/Balmer break degeneracy, that were not present in the populations covered in shallower large area surveys like the Sloan Digital Sky Survey \citep[SDSS,][]{York:00} and the Two Micron All Sky Survey \citep[2MASS,][]{Skrutskie:06}.
-In order to meet the demanding error budgets of LSST and LSST-DESC it will be necessary to fully characterize those degeneracies wherein multiple redshift solutions have comparable likelihood.
-
-There is often a desire to have a single valued ``point-estimate'' redshift for an individual galaxy.
-However, the complex, non-linear (and often non-unique) nature of the mapping between broad band fluxes and redshift means that a single value is unable to capture the full redshift information encoded in a galaxy's magnitudes.
-For example, a common point-estimate for a template-based method is taking the highest likelihood solution as the point photo-$z$.
-A single valued redshift ignores degenerate redshift solutions of lower probability, potentially biasing photometric redshift estimates both for individual galaxies and ensemble distributions.
-Storing more information is necessary, most often photo-$z$ codes output the redshift probability density function (PDF), also often referred to as $p(z)$, describing the relative likelihood as a function of redshift.
-%\red{should we move some of AIM's p(z) vs p(z|d,I) to here?}
-Early template methods such as \citet{Fernandezsoto:99} converted relative $\chi^2$ values of template spectra to likelihoods to estimate $p(z)$.
-Soon after, codes such as \citet{Benitez:00} added a Bayesian prior and output a posterior probability distribution.
-While many early machine learning based algorithms focused on a point-estimate, \citet{Firth:03} used a neural net with 1000 realizations scattered within the photometric errors to estimate a $p(z)$.
-As more groups began to employ photometric redshifts in their cosmological analyses, there was a realization that point-estimate photo-$z$'s were inadequate for precision cosmology measurements \citep{Mandelbaum:2008}.
-From around this point onward, most photo-$z$ algorithms have attempted to implement some estimate of the overall redshift probability in their outputs, and some surveys began supplying a full $p(z)$ rather than a simple redshift point-estimate and error \citep[e.~g.~][]{de_Jong:17}.
-
-%\aim{We should spend a whole paragraph establishing that photo-$z$ PDFs are a thing, i.e.
-review the motivation for PDFs over point estimates, basic definition, early methods papers, surveys that have published them, etc.}
-%\scc{I agree with Alex, we should introduce here what a pdf is and which survey have already published some pdfs (KiDS dr3 for example) or will produce them (LSST, Euclid etc.)}
-%\red{See above paragraph, does this adequately address this?--SJS}\scc{fine for me}
+In order to meet these ambitious goals for photo-$z$ accuracy, every aspect of photo-$z$ estimation will have to be optimized: the algorithms employed, both template and machine-learning based (both in design and implementation); the spectroscopic data used as a training set for machine learning algorithms or to estimate template sets and train Bayesian priors; and probabilistic catalogue compression schemes that balance information retention against limited storage resources.
 
@@
 DES compared several codes for point estimates \citep{Sanchez:14} and a summary statistic of photo-$z$ interim posteriors for tomographically binned galaxy subsamples \citep{Bonnett:16}.  
-This paper is distinguished by its inclusion of metrics of photo-$z$ interim posteriors themselves and consideration of both classic and state-of-the-art photo-$z$ algorithms, comparing the performance of several of the most widely employed codes as well as some that have been developed only recently on the basis of metrics appropriate for a probabilistic data product.
+This paper is distinguished by its focus on metrics of photo-$z$ interim posteriors themselves and consideration of both classic and state-of-the-art photo-$z$ algorithms, comparing the performance of several of the most widely employed codes as well as some that have been developed only recently on the basis of metrics appropriate for a probabilistic data product.
 The results presented in this work are
@@
 If we run multiple photo-$z$ codes on a single dataset, the photo-$z$ interim posteriors will not be identical because each code is based on assumptions in the form of an interim prior --- these assumptions form the premise for photo-$z$ estimation as a whole and are the only way to introduce differences in estimates of what would otherwise be a shared photo-$z$ posterior $p(z | d)$ regardless of the code used to obtain it.
-Though explicit knowledge of the interim prior is necessary to use photo-$z$ interim posteriors self-consistently in physical inference, the interim prior of a particular methodology is often implicit and not necessarily shared among all galaxies in the catalog.
+Though explicit knowledge of the interim prior is necessary to use photo-$z$ interim posteriors self-consistently in physical inference, the interim prior of a particular methodology is often implicit and not necessarily shared among all galaxies in the catalogue.
 

### 2018-10-10 paper/tex/results.tex
fix merge cruft in results.tex
remove seaborn from Fig 4

@@
 \caption{The stacked $p(z)$ produced by each photo-$z$ code ($\hat{N}(z)$, red) compared to the spectroscopic redshift distribution ($N'(z)$, blue).
-Varying levels of small-scale structure are seen in the codes.
-Both $\hat{N}(z)$ and $N'(z)$ are smoothed using a single bandwidth chosen via Scott's rule for all codes.} \label{fig:nz}
+Varying levels of agreement are seen in the codes.
+Both $\hat{N}(z)$ and $N'(z)$ in all codes are smoothed using a single bandwidth chosen via Scott's rule.} \label{fig:nz}
 %\aim{Arguably this could also be clarified by showing the true $n'(z)$ once and then differences from it for each cod's stacked estimator\dots}\scc{what if we plot the difference between true and stacked rather than soverposing the two lines?

### 2018-10-10 plotting_scripts/plot_nzsums_scottsrule.py

@@ -3,7 +3,7 @@ import numpy as np
 import matplotlib
 matplotlib.use('Agg')
 import matplotlib.pyplot as plt
-import seaborn as sns
+#import seaborn as sns
 import scipy.ndimage.filters as scifilt
 
 def main(argv):
@@ -46,7 +46,7 @@ def main(argv):
         pzvals[i,:] = data[:,2]
         smooth_pzvals = scifilt.gaussian_filter1d(pzvals,smooth_sigma,
                                                   mode='constant')
-        print "read in data for %s"%xfile
+        print ("read in data for %s"%xfile)
         
 
     numrows = 3
@@ -55,19 +55,28 @@ def main(argv):
 
     fig,axes = plt.subplots(numcodes, sharex=True, sharey=True, figsize=(12,9))
     fig.subplots_adjust(hspace=0.025, wspace=0.055)
-    sns.set()
+#    sns.set()
     
     for i in range(numcodes):
         ax = plt.subplot(numrows,numcols,i+1)
-        plt.plot(z_array,szvals[i],c='b',linestyle='-',linewidth=2,
-                 label="zspec KDE sum", alpha=0.85)
-        tmplabel = "%s N(z) sum"%(labels[i])
+        if i == 0: #only include label for first entry
+            plt.plot(z_array,szvals[i],c='b',linestyle='-',linewidth=2,
+                     label="zspec", alpha=0.85)
+        else:
+            plt.plot(z_array,szvals[i],c='b',linestyle='-',linewidth=2,
+                     alpha=0.85)
+        tmplabel = "%s"%(labels[i])
         plt.plot(z_array,smooth_pzvals[i],color='r',linestyle='-',
                  linewidth=2,label=tmplabel, alpha=0.85)
 
         ax.set_xlim([0.0,1.99])
-        ax.set_ylim([0.0,1.99])
-        ax.legend(loc="upper left",fontsize=11)
+        ax.set_ylim([0.0,1.79])
+        ax.yaxis.set_ticks(np.arange(0.,1.55,0.5))
+        if i ==0: #include lines only for first entry
+            ax.legend(loc="upper left",fontsize=11)
+        else: 
+            ax.legend(loc="upper left",fontsize=11, handletextpad=0,
+                      handlelength=0)
         if (i<(numcodes-4)):
             ax.set_xticklabels([])
         else:

### 2018-10-12 paper/main.tex
address more referee issues in intro

@@
   This initial test evaluates photo-$z$ algorithms in the presence of representative training data and in the absence of several common sources of systematic errors that affect the procedures by which photo-$z$ PDFs are derived.
-  The photo-$z$ PDFs are evaluated using multiple metrics including the Kolmogorov-Smirnoff statistic, Cramer-von Mises statistic, Anderson-Darling statistic, Kullback-Leibler divergence, $N(z)$ moments, quantile-quantile plots and probability integral transform.
-  We observe several trends, including an overall over/under-prediction in the broadness of the PDFs for several of the codes.
+%  The photo-$z$ PDFs are evaluated using multiple metrics including the Kolmogorov-Smirnoff statistic, Cramer-von Mises statistic, Anderson-Darling statistic, Kullback-Leibler divergence, $N(z)$ moments, quantile-quantile plots and probability integral transform.
+  The photo-$z$ PDFs are evaluated using multiple metrics and we observe several trends, including an overall over/under-prediction in the broadness of the PDFs for several of the codes.
   A careful accounting of all systematics discovered will be necessary for the codes employed in upcoming analyses in order to achieve unbiased cosmological measurements.

### 2018-10-12 paper/tex/introduction.tex
address more referee issues in intro

@@
 The LSST Dark Energy Science Collaboration (LSST-DESC) developed a separate Science Requirements Document \citep{Mandelbaum:2018}, which forecasts the constraining power of five cosmological probes using somewhat conservative assumptions to define requirements on systematic errors for several measurements.
-These include even more stringent requirements on photometric redshift performance than those included in the LSST SRD, though the initial LSST-DESC requirements on tomographic bin populations.
+These include even more stringent requirements on photometric redshift performance than those included in the LSST SRD, though the initial LSST-DESC requirements are defined for tomographic bin populations.
 The tremendous size of LSST's galaxy catalogue will be enabled by its exceptional depth, pushing to fainter magnitudes and deeper imaging and including galaxies of lower luminosity and higher redshift than ever before.
@@
 For cosmological measurements, certain science cases require redshift information on individual objects, e.~g.~identification of host galaxy redshift for supernova classification, or identifying potential cluster membership.
-Other science cases seem to need only ensemble redshift information; for instance current weak lensing techniques require only the overall redshift distribution $N(z)$ for tomographic redshift samples.
+Other science cases seem to need only ensemble redshift information; for instance many current cosmic shear techniques require only the overall redshift distribution $N(z)$ for tomographic redshift samples.
 However, recent data-driven techniques employing hierarchical Bayesian or Gaussian Process methods have emerged that calibrate redshift distributions using individual $p(z)$ estimates \citep[e.~g.~][]{Sanchez:2018}.
@@
 Comparing the relative performance of the codes enables us to evaluate whether each code is using information in an optimal way, and may reveal enhancements in some codes and deficiencies in others, either in the fundamental algorithm, or in specific implementation.
-We note that these initial tests are a necessary requirement for photo-$z$ codes that will be used in cosmological analyses; however, meeting these requirements is only the first stage in the process of accounting for incomplete training samples, object blending, and other complications that realistic samples will encounter in estimating photometric redshifts.
+We note that these initial tests are a necessary requirement for photo-$z$ codes that will be used in cosmological analyses; however, meeting these requirements is only the first stage in the process of accounting for incomplete training samples, object blending,  the robustness of estimated redshift distributions when combining redshift estimates in tomographic bins, and other complications that realistic samples will encounter in estimating photometric redshifts.
 This can be thought of as an initial test under near perfect conditions before further complexities are added in future papers.

### 2018-10-30 paper/tex/data.tex
finish Section 2 edits

@@
 
-In order to test the current generation codes, we employ a simulated galaxy catalogue.
+In order to test the current generation codes, we employ an existing simulated galaxy catalogue.
 The simulation is completely catalogue-based, with no image construction or mock measurements made.
@@
 \label{sec:buzzard}
-The \textsc{Buzzard-highres-v1.0} \red{put in cites to in prep Buzzard papers} catalogue construction started with a dark matter only simulation.
+The \textsc{Buzzard-highres-v1.0} (De Rose et al., in prep; Wechsler et al., in prep) catalogue construction started with a dark matter only simulation.
 This N-body simulation contained $2048^3$ particles in a $400$ Mpc h$^{-1}$ box.
-\red{[N]} snapshots (with smoothing and interpolation between snapshots) were saved in order to construct a lightcone.
+A set of time snapshots (with smoothing and interpolation between snapshots) were saved in order to construct a lightcone.
 Dark matter halos were identified using the \textsc{Rockstar} software package \citep{Behroozi:13}.
@@
 To assign an SED to each galaxy, the {\it Adding Density Dependent Spectral Energy Distributions} (\textsc{ADDSEDS}, deRose in prep.)\footnote{\url{https://github.com/vipasu/addseds}} procedure was used.
-This consisted of training an empirical relation between absolute $r$-band magnitude, local galaxy density, and SED using a sample of $\sim 5e^{5}$ galaxies from the magnitude-limited Sloan Digital Sky Survey Data Release 6 Value Added Galaxy Catalog \citep{Blanton:05}\red{[Note: is this the proper reference to SDSS-NYU VAGC?
-File is called combined\_dr6\_cooper.fits, but I don't see which Cooper et al 2006 this is supposed to refer to?]}.
-Each SDSS spectrum is fit with a sum of five SED components using the \textsc{k-correct \red{v?}} software package\footnote{\url{http://kcorrect.org}} \citep{Blanton:07}, thus each galaxy SED is parameterized as five weights for the basis SEDs.
+This consisted of training an empirical relation between absolute $r$-band magnitude, local galaxy density, and SED using a sample of $\sim 5\times 10^{5}$ galaxies from the magnitude-limited Sloan Digital Sky Survey Data Release 6 Value Added Galaxy Catalog \citep{Blanton:05}.
+Each SDSS spectrum is fit with a sum of five SED components using the \textsc{k-correct \red{v4\_3?}} software package\footnote{\url{http://kcorrect.org}} \citep{Blanton:07}, thus each galaxy SED is parameterized as five weights for the basis SEDs.
 The distance to the spatial projected fifth-nearest neighbour was used as a proxy for local density in the SDSS training sample.
-For each simulated galaxy, a ``random'' \red{[details]} galaxy with ``similar'' \red{[details]} absolute $r$-band magnitude and local galaxy density was chosen from the training set, and that training galaxy's SED was assigned to the simulated galaxy.
+For each simulated galaxy, a galaxy with similar absolute $r$-band magnitude and local galaxy density was chosen from the training set, and that training galaxy's SED was assigned to the simulated galaxy.
+This process is done in such a way as to preserve the colour-density relation of galaxy environment.
 Given the SED, absolute $r$-band magnitude and redshift, we computed apparent magnitudes in the six LSST filter passbands, $ugrizy$.
@@
 %numbers taken from Alex Abate's photErrorModel.py script in PhotoZDC1 repository, which was used with nYrObs=10.
+In the course of simulating Gaussian photometric errors, we add noise to objects fluxes, and some of these noisy fluxes will become negative in one or more bands.
+We call such negative fluxes ``non-detections'' and signify them with a placholder magnitude of 99.0 in the catalog.
+Thus, further mentions of ``non-detections'' refer to objects that would be ``looked at but not seen'' in multi-band forced photometry, and the photo-z codes will treat them as such.
 
@@
 The total catalogue covered $400$ square degrees and contained $238$ million galaxies to an apparent magnitude limit of $r\!=\!29$ and spanning the redshift range $0\!<\!z\!\leq\!8.7$.
-This catalogue contained two orders of magnitude more galaxies than were needed for this study, so only $\sim\!8$ square degrees were used.
-Systematic problems with galaxy colors above $z\!>\!2$ were observed, so the catalogue was trimmed to include only galaxies in the redshift range $0\!<\!z\!\leq\!2.0$.
+In order for statistical errors not to dominate, we need less than one million galaxies in our sample.
+Several studies claim that only a few tens of thousands of spectra are necessary to calibrate photo-z surveys to Stage IV requirements (e.~g.~\citet{Bernstein:10}, \citet{Masters:2017}).
+Therefore, we aim for a final number of training galaxies between $3\times 10^{4}$ and $5\times 10^{4}$ in our sample.
+%This catalogue contained two orders of magnitude more galaxies than were necessary to determine statistics for this study,
+In order to reduce our sample to a reasonable size, we limit our dataset to a subset of $\sim\!16.8$ square degrees selected from five separate spatial regions of the simulation.
+Systematic problems with galaxy colors above $z\!>\!2$ were observed, so the catalogue was limited to include only galaxies in the redshift range $0\!<\!z\!\leq\!2.0$.
 A random subset of the the remaining galaxies was chosen, and placed at random into either a ``training'' set ($10$ per cent of the sample), for which the galaxies true redshifts will be supplied, or a ``test'' set (the remaining $90$ per cent of the sample), for which each code will need to predict a redshift PDF for each galaxy.
-The resulting catalogues contain $111\,171$ training galaxies and $1\,000\,883$ test galaxies.
-We restrict our analysis to a sample with $i<25.3$, which give a signal-to-noise $\sim\,30$ for most galaxies, a cut often referred to as the expected ``LSST Gold Sample''.
+%The resulting catalogues contain $111\,171$ training galaxies and $1\,000\,883$ test galaxies.
+Finally, we restrict our analysis to a sample with an apparent magnitude limite $i<25.3$, which give a signal-to-noise $\sim\,30$ for most galaxies, a cut often referred to as the expected ``LSST Gold Sample''.
 This magnitude cut results in a training set with $44\,404$ galaxies and a test set containing $399\,356$ galaxies.
@@
 \label{sec:buzztemplates}
-As mentioned in Section~\ref{sec:buzzard}, the SEDs in the Buzzard simulation are drawn from an empirical set of SEDs taken from the SDSS DR6 NYU-VAGC, a sample of roughly $\sim5e^{5}$ galaxies with spectra in SDSS.
-To determine a finite set of templates to use with template fitting codes we take the five SED weight coefficients for each of the $\sim 500\,000$ galaxies in the SDSS sample and run a simple K-means clustering algorithm on this five dimensional space.
-The K-means cluster centres span the space of coefficients and properly reflect the underlying density in the coefficient space, thus providing a reasonable approximation for a spanning SED set.
+As mentioned in Section~\ref{sec:buzzard}, the SEDs in the Buzzard simulation are drawn from an empirical set of SEDs taken from the SDSS DR6 NYU-VAGC, a sample of roughly $\sim5\times 10^{5}$ galaxies with spectra in SDSS.
+To determine a finite set of templates to use with template fitting codes we take the five SED weight coefficients for each of the galaxies in the SDSS sample and run a simple K-means clustering algorithm on this five dimensional space.
+Each dimension was normalized such that it spanned an interval $[0,1]$.
+The K-means clusters partition the five-dimensional space of coefficients into Voronoi cells, spanning the space of coefficients in a way that properly reflects the underlying density in the coefficients.
+Thus, the resultant SEDs constructed using the cell centers as weight coefficients will provide a reasonable spanning SED set.
 An ad-hoc number of $K=100$ was chosen and the $100$ K-means centre positions are taken as the weights for the \textsc{k-correct} SED components to construct one hundred template SEDs.
@@
 Answer: no, initially I supplied a list of 150 that were drawn with a different algorithm, but once we switched to the final k-means set I only supplied 100 (the initial algorithm over-weighted outlier SEDs, which is why we needed the extra 50)--SJS)}
-however not every template code uses this set of one hundred templates: because EAZY was designed and written to use the same five basis templates employed by \textsc{k-correct} when constructing our mock galaxies, EAZY was run using linear combinations of these five templates rather than using the 100 discrete templates.
+and the templates were used by both \textsc{BPZ} and \textsc{LePhare}; however, because EAZY was designed and written to use the same five basis templates employed by \textsc{k-correct} when constructing our mock galaxies, EAZY was run using linear combinations of these five templates rather than using the 100 discrete templates.
+The ability to fit for linear combinations of templates highlights an important implementation difference between similar photo-z codes.
 
@@
 As all SEDs are constructed from only five basis templates, properties of the galaxy population will be restricted to follow linear combinations of the characteristics of the five basis templates, so certain non-linear features, for example the full range of emission line fluxes relative to the continuum, will not be included in the model galaxy population.
+Moreover, the linear combinations of templates are modeled on the $\sim5\times 10^{5}$ SDSS galaxies discussed in Section~\ref{sec:buzzard}, and thus only galaxies that resemble those spectroscopically observed by the SDSS will be included in the sample.
 No additional dust reddening intrinsic to the host galaxy is included, the only approximation of dust extinction comes in the form of dust encoded in the five basis SEDs via the training set used to create the basis templates.
-Simple linear combinations of these basis templats will, once again, not explore the full range of realistic dust extinction observed in galaxy populations.
+Simple linear combinations of these basis templates will, once again, not explore the full range of realistic dust extinction observed in galaxy populations.
+While these idealized conditions limit the realism of our galaxy population, some are also by design.
+We aim to test the photo-z codes at a very basic level, and a simplified model assures that differences in results seen between the codes are due to fundamental differences in their underlying assumptions and implementation details, rather than more nuanced implementation details.
 

### 2018-11-07 paper/tex/methods.tex
address reviewer comments in method section

--- 2018-08-29 paper/tex/methods.tex
@@
 \label{sec:pzcodes}
-
 %(Sam Schmidt, Rongpu Zhou, Ibrahim Almosallam, Eric Nuss, Johann Cohen Tanugi, John Soo)
@@
 \label{sec:trainingcodes}
-The training-based codes use a variety of algorithms in order to estimate $p(z)$. 
+The training-based codes use a variety of algorithms in order to estimate $p(z)$, specifics of each implementation are described in the subsections.
 Some aspects of data treatment were left to the individual code runners, for example, whether/how to split the available data with known redshifts into separate training and validation sets.
-The treatment of objects not detected in one or more bands was also left to the code runners.
-There are varying conventions among training-based codes, and no one prescription dominates in the photo-$z$ literature.
-This and other details of the code implementations are described below.
+Another key difference is the treatment of non-detections in one or more bands.
+Some codes choose to ignore a band, others replace the value with either an estimate for the detection limit, the mean of other values in the training set, or another default value.
+There are varying conventions among training-based codes for treatment of non-detections, and no one prescription dominates in the photo-$z$ literature.
+The specific choices for each code affect the results, and contribute to the implicit prior influencing their output.
+However, we remind the reader that only 2.0 per cent of our sample has non-detections, almost exclusively in the u-band, and thus should not dominate the code performance differences.
 
@@
 \textsc{ANNz2}\footnote{\url{https://github.com/IftachSadeh/ANNZ}} \citep{Sadeh:16} is a software package that has the ability to employ several machine learning algorithms, including artificial neural networks (ANN), boosted decision tree (BDT) and k-nearest neighbour (KNN).
-Using the Toolkit for Multivariate Data Analysis (TMVA) with ROOT\footnote{\url{http://tmva.sourceforge.net/}}, it can run multiple machine learning algorithms for a single training and outputs photo-$z$'s based on a weighted average of their performances.
-
-\textsc{ANNz2} is capable of producing both photo-$z$ point estimates and redshift posterior probability distributions $p(z)$.
+Using the Toolkit for Multivariate Data Analysis (TMVA) with ROOT\footnote{\url{http://tmva.sourceforge.net/}}, it can either run a single machine learning algorithm and return results, or it can run multiple algorithms simultaneously and output photo-$z$'s as a weighted combination of the different algorithms.
+In this study, only ANNs were employed.
+The redshift PDFs are produced by running an ensemble set of ANNs, each with different random seeds used in initialization of input parameters for training.
+Uncertainties for each method are estimated from a KNN-uncertainty estimator \citep{Oyaizu:08}.
+The final PDF can either be the ``best'' of the candidate PDFs, or a weighted average of the PDFs based on their error estimates for each of the ensemble members.
+%\textsc{ANNz2} is capable of producing both photo-$z$ point estimates and redshift posterior probability distributions $p(z)$.
 It can also perform classifications, and supports reweighting between samples.
 The PDFs are produced by propagating the intrinsic uncertainty on the input parameters and the uncertainty in the machine learning method to the expected photo-$z$ solution, averaged over multiple runs weighted based on the performance of each run.
-\textsc{ANNz2} presents its photo-$z$ uncertainty different from many codes by using the KNN method: it estimates the photo-$z$ bias between each object and a fixed number of nearest neighbours in parameter space, it then takes the $68$th percentile width of the distribution of the bias.
-This is based on the implication that objects with similar photometric properties would have similar uncertainties, and therefore the photometric errors of the inputs are not propagated into the code.
 
@@
 
-\subsubsection{Color-Matched Nearest-Neighbours}
+\subsubsection{Colour-Matched Nearest-Neighbours}
 \label{sec:cmnn}
@@
 
-The nearest-neighbours color-matching photometric redshift estimator (\textsc{CMNN}) is presented in \citet[][herafter G18]{Graham:17}.
+The nearest-neighbours colour-matching photometric redshift estimator (\textsc{CMNN}) is presented in \citet[][hereafter G18]{Graham:17}.
 This method uses a training set of galaxies with known redshifts that has equivalent or better photometry as the test set in terms of quality and filter coverage.
-For each galaxy in the test set we identify a color-matched subset of training galaxies, choose one (e.g.
-the nearest-neighbour or a random selection), and use its known redshift as the photo-$z$.
-This color-matched subset is identified by first calculating the Mahalanobis distance $D_M$ in color-space between the test galaxy and all training-set galaxies: the difference between the test and a training set galaxy's color divided by the photometric error, summed over all colors (i.e., $u$-$g$, $g$-$r$, $r$-$i$, $i$-$z$, and $z$-$y$).
-Then, the threshold value for $D_M$ that define a good color match is set by the percent point function (PPF): for example, for $N_{\rm dof}=5$, PPF$=95$ per cent of all training galaxies consistent with the test galaxy will have $D_M < 11.07$ (where $N_{\rm dof}$, the number of degrees of freedom, is the number of colors).
-For a given test galaxy, the $p(z)$ is the normalized distribution of the true catalogue redshifts of this color-matched subset of training galaxies, and the standard deviation of the color-matched subset is used as the photo-$z$ uncertainty.
+For each galaxy in the test set we identify a colour-matched subset of training galaxies.
+This subset is identified by first calculating the Mahalanobis distance $D_M$ in colour-space between the test galaxy and all training-set galaxies:
+\begin{equation} 
+D_{\rm M} = \sum^{N_{\rm colours}}\frac{(c_{\rm train}-c_{\rm test})^2}{(\delta c_{\rm test})^2}
+\end{equation}
+where c colour, $\delta c_{\rm test}$ is the measurement error on the colour, and $N_{\rm colours}$ is the total number of colors (i.e., in our case $u$-$g$, $g$-$r$, $r$-$i$, $i$-$z$, and $z$-$y$).
+Then, we choose a threshold value for $D_M$ that defines the colour match set based on a set value of the percent point function (PPF): for example, for $N_{\rm dof}=5$, choosing a vPPF$=95$ per cent of all training galaxies consistent with the test galaxy will have $D_M < 11.07$ (where $N_{\rm dof}$ is the number of degrees of freedom, in this case the number of colours).
+If a galaxy had a non-detection in a band, that band was dropped and $N_{\rm dof}$ was reduced by one in the colour-matching space.
+For a given test galaxy, the $p(z)$ is the normalized distribution of the true catalogue redshifts of this colour-matched subset of training galaxies.
 
-We have applied the nearest-neighbours color-matching photometric redshift estimator described in G18 to the simulated data.
+We have applied the nearest-neighbours colour-matching photometric redshift estimator described in G18 to the simulated data.
 Compared to its application in G18, there are some minor differences in the application of this estimator to the Buzzard catalogue.
 First, we do not impose non-detections on galaxies with a magnitude fainter than the expected LSST 10-year limiting magnitude or bright enough to saturate with LSST: {\it all} of the photometry for all the galaxies in the test and training sets are used for this experiment.
-Second, as in G18 we do apply an initial cut in color to the training set before calculating the Mahalanobis distance in order to accelerate processing, and also use a magnitude pseudo-prior to improve photo-$z$ estimates, but for both we have used different cut-off values that are appropriate for the Buzzard galaxies' colors and magnitudes.
-Third, we set different parameters for the identification of the color-matched subset of training galaxies and the selection of a photometric redshift estimate.
-In G18 we used a percent point function (PPF) value of 0.68 to identify the color-matched subset of training galaxies and used the redshift of nearest neighbour in color-space as the photo-$z$ estimate.
-These choices work well when the desire is to obtain accurate photo-$z$ estimates for most test-set galaxies, but does not return a robust $p(z)$ in all cases -- especially for galaxies that are bright and/or have few matches in color-space.
-Since a robust estimate of $p(z)$ is desired for this work we make several changes to our implementation of the \textsc{CMNN} photo-$z$ estimator.
-We continue to use a percent point function of ${\rm PPF} = 0.95$ to generate the subset of color-matched training galaxies, but weight them by the inverse of their Mahalanobis distance.
-This weighting maintains some of the accuracy that was previously achieved by simply using the nearest neighbour in color-space.
-We then use the weights to create the $p(z)$ instead of having the redshift of each color-matched training-set galaxy count equally.
-To obtain a robust estimate of the $p(z)$ for galaxies with a small number of color-matched training set galaxies, when this number is less than $20$ the nearest $20$ neighbours in color-space are used instead, and we convolve the $p(z)$ with a Gaussian with a standard deviation of:
+Second, as in G18 we do apply an initial cut in colour to the training set before calculating the Mahalanobis distance in order to accelerate processing, and also use a magnitude pseudo-prior to improve photo-$z$ estimates, but for both we have used different cut-off values that are appropriate for the Buzzard galaxies' colours and magnitudes.
+Third, we set different parameters for the identification of the colour-matched subset of training galaxies and the selection of a photometric redshift estimate.
+In G18 we used a percent point function (PPF) value of 0.68 to identify the colour-matched subset of training galaxies and used the redshift of nearest neighbour in colour-space as the photo-$z$ estimate.
+These choices work well when the desire is to obtain accurate photo-$z$ estimates for most test-set galaxies, but does not return an accurate $p(z)$ in all cases -- especially for galaxies that are bright and/or have few matches in colour-space.
+Since an accurate estimate of $p(z)$ is desired for this work we make several changes to our implementation of the \textsc{CMNN} photo-$z$ estimator.
+We continue to use a percent point function of ${\rm PPF} = 0.95$ to generate the subset of colour-matched training galaxies, but weight them by the inverse of their Mahalanobis distance.
+This weighting maintains some of the accuracy that was previously achieved by simply using the nearest neighbour in colour-space.
+We then use the weights to create the $p(z)$ instead of having the redshift of each colour-matched training-set galaxy count equally.
+To obtain a robust estimate of the $p(z)$ for galaxies with a small number of colour-matched training set galaxies, when this number is less than $20$ the nearest $20$ neighbours in colour-space are used instead, and we convolve the $p(z)$ with a Gaussian with a standard deviation of:
 
@@
 \noindent to appropriately broaden it so that the $p(z)$ for these test galaxies represents the enlarged PPF value associated with it.
-Overall, these three changes will yield poorer accuracy photo-$z$ compared to those presented in G18, but they will all have significantly more robust estimates of the $p(z)$, particularly for the brightest test galaxies.
+Overall, these three changes will yield less precise photo-$z$ estimates compared to those presented in G18, but they will all have significantly more accurate estimates of the $p(z)$, particularly for the brightest test galaxies.
 This is sufficient for this work because, as described in G18, the goal of the \textsc{CMNN} photo-$z$ estimator was never to provide the ``best'' (or even competitive) estimates in the first place, given its reliance on a deep training set, but rather to provide a means for direct comparisons between LSST photometric quality and photo-$z$ estimates.
-With this work we show how the input parameters should be set in order to return robust $p(z)$ estimates in addition to point value estimates.
+With this work we show how the input parameters should be set in order to return accurate $p(z)$ estimates in addition to point value estimates.
 
@@
 \textsc{Delight}\footnote{\url{https://github.com/ixkael/Delight}} \citep{Leistedt:17} infers photo-$z$'s by using a data-driven model of latent SEDs and a physical model of photometric fluxes as a function of redshift.
+Delight models the underlying latent SEDs as a linear combination of a set of pre-defined template SEDs, plus zero mean Gaussian processes with factorized kernels.
 Generally, machine learning methods rely on representative training data with similar band passes, while template based methods rely on a complete library of templates based on physical models constructed.
@@
 The default settings of \textsc{Delight} were used, with the exception that the PDF bins were set to be linearly-spaced rather than logarithmic.
-In this study a flat prior is assumed.
+In this study a flat prior in magnitude/type is assumed.
 
@@
 
-In this paper, we use \textsc{xgboost}~\citep{Chen:16} for the regression part as these techniques scale well for massive data; it should however be noted that \textsc{FlexCode-RF} (also on GitHub), based on Random Forests, generally performs better for smaller data sets.
+In this paper, we use \textsc{xgboost}~\citep{Chen:16} for the regression part; it should however be noted that \textsc{FlexCode-RF} (also on GitHub), based on Random Forests, generally performs better for smaller data sets.
 As our basis, we choose a standard Fourier basis.
 There are two tuning parameters in our $p(z)$ estimate: (i) the number of terms, $I$, in the series expansion in Eq.~\ref{eq::series_expansion}, and (ii) an exponent $\alpha$ that we use to sharpen the computed density estimates $\widehat{f}(z|\x)$, according to $\widetilde{f}(z|\x) \propto \widehat{f}(z|\x)^\alpha$.
-We split the ``train data'' into a training set (85\%) and a validation set (15\%), and choose both $I$ and $\alpha$ in an automated way by minimizing the weighted $L_2$-loss function \citep[Eq.
+We reserve 15\% of the training set data as a validation set, and choose both $I$ and $\alpha$ in an automated way by minimizing the weighted $L_2$-loss function \citep[Eq.
 5 in][]{Izbicki:17} on the  validation set.
+While the native storage format for \textsc{FlexCode} encodes the PDF using the coefficients shown in Equation~\ref{eq:FlexZBoost_coeffs}, to match the output format requested of other codes we discretize our final estimates into 200 bins linearly spaced in $0 < z < 2$.
 
-Although {\tt FlexCode} offers a {\em lossless compression} of the photo-$z$ estimates (in this study, one can reconstruct $\widetilde{f}(z|\x)$ exactly at any resolution from estimates of the first 35 coefficients, Eq.~\ref{eq:FlexZBoost_coeffs}, for a Fourier basis $\{\phi_i(z)\}_{i}$), we discretize our final estimates into 200 bins linearly spaced in $0 < z < 2$ for easy comparison with other algorithms.
+%Although {\tt FlexCode} offers a lossless compression of the photo-$z$ estimates (in this study, one can reconstruct $\widetilde{f}(z|\x)$ exactly at any resolution from estimates of the first 35 coefficients, Eq.~\ref{eq:FlexZBoost_coeffs}, for a Fourier basis $\{\phi_i(z)\}_{i}$), we discretize our final estimates into 200 bins linearly spaced in $0 < z < 2$ for easy comparison with other algorithms.
 Using a higher resolution may yield better results (with no added cost in storage).
@@
 
-The data is prepared for \textsc{GPz} by taking the log of the magnitude errors, decorrelating the data set using PCA and imputing the missing values using a simple linear model that estimates the missing variables given the observed ones.
+The data is prepared for \textsc{GPz} by taking the log of the magnitude errors, decorrelating the data set using PCA and imputing any missing magnitude values using a simple linear model that estimates the missing magnitudes given the observed ones.
 The log transformation helps to smooth the long tail distribution of the magnitude errors, which is more stable numerically and makes the optimization process unconstrained.
@@
 $\Sigma_{uo}=0$, this will reduce to a simple average predictor.
-
 We use the Variable Covariance (VC) option in \textsc{GPz} with 200 basis functions after we note that there is no significant increase in the performance on the validation set (using 80\%-20\% training-validation split) and with no cost-sensitive learning applied.
-We do note that \textsc{GPz} is trained using a data set that includes galaxies fainter than those in the test set, which was truncated at $i<25.3$.
-This mismatch in training vs.~test results in a slightly different distribution in galaxy photometry and redshift distribution between the training and test set, which most likely degraded results compared to a run where the training set more closely matched the test set.
 
@@
 %, already validated on photo-$z$'s in several cases \citep{de_Jong:17,Cavuoti:17b,Cavuoti:15,Brescia:14,Brescia:13,Biviano:13}.
-Due to its plug-in based modular nature, \textsc{METAPhoR} can be easily replaced by any other photo-$z$ prediction kernel, regardless of specific implementation, while still taking advantage of the overall code interface.
 
-At a higher level, the pipeline mainly consists of three modules: (i) \textit{data pre-processing}, including a catalogue cross-matching sub-module \citep[based on the tool C3, ][]{Riccio:17}, a sub-module for photometric evaluation and error estimation of the multi-band catalogue used as Knowledge Base (KB), and a sub-module dedicated to the perturbation of the photometric KB, propaedeutic to the PDF estimation; (ii) \textit{photo-$z$ prediction}, which is the training/validation/test phase, producing the photo-$z$'s point estimates, based on a pre-selected ML method; (iii) \textit{PDF estimation}, specifically designed to calculate the PDF of the photo-$z$ estimation errors.
+\textsc{METAPhoR} includes data modules for pre-processing, photo-$z$ estimation, and PDF estimation, and post-processing.
+The pre-processing includes a model for perturbation of the photometry that is employed in calculating the PDF of the photo-$z$ estimation errors.
+The photometric perturbation is defined as: $m_{ij} = m_{ij} + \alpha_{i}F_{ij}*u_{\mu=0,\sigma=1}$, where $\alpha_{i}$ is a user selected multiplicative constant (useful in case of multi-survey photometry), $u_{\mu=0,\sigma=1}$ is a random value from the standard normal distribution and $F_{ij}$ is a bimodal function (a constant function + polynomial fitting of the mean magnitude errors on the binned bands), heuristically tuned in such a way that the constant component is the threshold under which the polynomial function is considered too low to provide a significant noise contribution to the photometry perturbation.
+
+%At a higher level, the pipeline mainly consists of three modules: (i) \textit{data pre-processing}, including a catalogue cross-matching sub-module \citep[based on the tool C3, ][]{Riccio:17}, a sub-module for photometric evaluation and error estimation of the multi-band catalogue used as Knowledge Base (KB), and a sub-module dedicated to the perturbation of the photometric KB, propaedeutic to the PDF estimation; (ii) \textit{photo-$z$ prediction}, which is the training/validation/test phase, producing the photo-$z$'s point estimates, based on a pre-selected ML method; (iii) \textit{PDF estimation}, specifically designed to calculate the PDF of the photo-$z$ estimation errors.
 The last module includes also a post-processing tool, providing some statistics on the produced point estimates and PDFs.
 
-The photometry perturbation law is based on the formula $m_{ij} = m_{ij} + \alpha_{i}F_{ij}*u_{\mu=0,\sigma=1}$, where $\alpha_{i}$ is a user selected multiplicative constant (useful in case of multi-survey photometry), $u_{\mu=0,\sigma=1}$ is a random value from the standard normal distribution and $F_{ij}$ is a bimodal function (a constant function + polynomial fitting of the mean magnitude errors on the binned bands), heuristically tuned in such a way that the constant component is the threshold under which the polynomial function is considered too low to provide a significant noise contribution to the photometry perturbation.
-
-As introduced, the photo-$z$ point estimate prediction engine of \textsc{METAPhoR} is based on the MLPQNA model, whose photo-$z$ regression training error, used by the quasi Newton learning rule, is based on the least square error and Tikhonov $L_{2}$-norm regularization \citep{Hofmann:18}.
+%As introduced, the photo-$z$ point estimate prediction engine of \textsc{METAPhoR} is based on the MLPQNA model, whose photo-$z$ regression training error, used by the quasi Newton learning rule, is based on the least square error and Tikhonov $L_{2}$-norm regularization \citep{Hofmann:18}.
 
@@
 The different test sets, used to produce the PDF, are thus obtained by introducing a proper perturbation, parametrized from the photometric error distribution in each band, on the photometric data populating the original test set \citep{Brescia:18}.
-
-For the present work since it was required to produce a redshift (and a PDF) for each object of the test set we decided to apply a hierarchical kNN to fill the missing detection, it goes without saying that for such points the reliability of PDFs and point estimation is lower.
+For the present work since it was required to produce a redshift (and a PDF) for each object of the test set we decided to apply a hierarchical kNN to replace the missing detections with values based on their neighbors.
+The reliability of PDFs and point estimation is lower.
 No cross validation has been used.
@@
 The classifier is laid out such that the hidden layers have 20:40:40 nodes each, all rectified linear units, and the output layer has $200$ nodes (corresponding to 200 bins for the PDF) activated with a ``softmax'' function so that they automatically sum to $1$.
+While previous implementations of the code, such as \citet{Sanchez:14} and \citet{Bonnett:15} (see Appendix C.3), implement a ``sliding bin'' smoothing, no such procedure was used in this study.
 
@@
 
-In the present work, the LSST magnitudes $u,~g,~r,~i$ and colors $u-g,~g-r,~r-i,~i-z,~z-y$ and their associated errors are used in the process of growing 100 trees with a minimum leaf size of 5 (the $z$ and $y$ magnitudes did not show significant correlation with the redshift in our cross-validation, so we did not use them when constructing our trees).
+In the present work, the LSST magnitudes $u,~g,~r,~i$ and colours $u-g,~g-r,~r-i,~i-z,~z-y$ and their associated errors are used in the process of growing 100 trees with a minimum leaf size of 5 (the $z$ and $y$ magnitudes did not show significant correlation with the redshift in our cross-validation, so we did not use them when constructing our trees).
 We partitioned our redshift space into 200 bins and smoothed each individual PDF with a smoothing scale of twice the bin size.
@@
 That is, we simply set the redshift PDF of every galaxy equal to the normalized $N(z)$ of the training sample.
-As the training sample is drawn from the same underlying distribution as the test sample, modulo small deviations due to sample size, the quantiles of the training and test distributions should be identical.
+This estimator is essentially a k nearest-neighbour estimator with k equal to the number of galaxies in the training sample.
+As the training sample is drawn from the same underlying distribution as the test sample, modulo small deviations due to sample size, the quantiles of the training and test distributions should be identical, modulo fluctuations due to finite sample size.
 This is a wildly unrealistic estimator, as it assigns all galaxies, no matter their apparent magnitude, colour, or true redshift, the same redshift PDF, and is thus uninformative at the level of individual object redshifts, but is designed to perform very well for the ensemble of all objects.

### 2018-11-08 paper/tex/analysis.tex
start on section 4 edits

@@
 
-As mentioned previously, there are cosmology probes that require knowledge of individual galaxy $p(z)$ and others that require only knowledge of the ensemble redshift distribution, $N(z)$.
-Due to the paucity of principled techniques for using and validating photo-$z$ PDFs, there are few alternatives to the common practice of reducing photo-$z$ PDFs to point estimates.
+As mentioned previously, there are cosmology probes that require knowledge of individual galaxy $p(z)$, for example galaxy intrinsic alignment studies, strong lensing foreground shear prediction, and photometric supernova classification, and others that require only knowledge of the ensemble redshift distribution, $N(z)$.
+Due to the paucity of principled techniques for using and validating individual galaxy photo-$z$ PDFs, there have been few alternatives to the common practice of reducing photo-$z$ PDFs to point estimates when evaluating and comparing photo-$z$ code performance in the literature.
 Though this practice should not be encouraged, we also calculate traditional metrics based on the most common point estimators derived from photo-$z$ PDFs.
@@
 because our training data is representative, the interim prior is the truth).
-In this ideal case, the method that would give the best approximation to $N'(z)$ would be one that neglects all the information contained in the photometry $\{d_{i}\}_{N_{tot}}$ and gives every galaxy the same photo-$z$ PDF $\hat{p}_{i}(z) = N'(z)$ for all $i$.
+In this ideal case, with representative training data, the method that would give the best approximation to $N'(z)$ would be one that neglects all the information contained in the photometry $\{d_{i}\}_{N_{tot}}$ and gives every galaxy the same photo-$z$ PDF $\hat{p}_{i}(z) = N'(z)$ for all $i$.
 This is the exact estimator, \trainz, that we have described in Section~\ref{sec:method:trainz}, and which will serve as a point of reference for the other codes.
@@
 %\red{Note: Let's address how the gridding (endpoints and spacing) affects the metrics, from the email correspondence Re: [LSST-DESC-MEMBERS] Summary of Nov 29 PZ Telecon. (AIM)}
-In order to minimize the effects of such choices, we asked those running all eleven codes to output $p(z)$ parameterized with a generous $\approx 200$ piecewise constant bins spanning $0<z<2$.
+In order to minimize the effects of such choices, we asked those running all eleven codes to output $p(z)$ parameterized with $\approx 200$ piecewise constant bins spanning $0<z<2$.
 The piecewise constant format is chosen because of its established presence in the literature, and the choice of 200 bins was motivated by the approximate number of columns expected to be available for storage of $p(z)$ for the final LSST Project tables.\footnote{See, e.~g.~the LSST Data Products Definition Document, available at: \url{https://ls.st/dpdd}}
@@
 These metrics give a more quantitative measure of the departure from ideal than the more visual PIT histogram and QQ plot.
-We compute metrics comparing the CDF of PIT values to a the CDF of a Uniform distribution, and also compute the CDF of the true redshift distribution $N'(z)$ compared the $\hat{N}(z)$ distribution derived from summing the $p(z)$ as described in Eq.~\ref{eq:stacked}.
+We compute metrics comparing the CDF of PIT values to the CDF of a Uniform distribution, and also compute the CDF of the true redshift distribution $N'(z)$ compared the $\hat{N}(z)$ distribution derived from summing the $p(z)$ as described in Eq.~\ref{eq:stacked}.
 

### 2018-11-13 paper/tex/methods.tex
address EAZY algorithm problem

@@
 In total, eleven distinct codes are tested.
-This sample is not comprehensive, codes were chosen based on the expertise available within the group; however, those chosen do cover a broad range of the current-generation codes.
+This sample is not comprehensive, codes were chosen based on the expertise available within the group; however, those chosen do cover a broad range of the current-generation methods used in the field.
 Both template-based and machine learning approaches are included and each are described separately in Secs.~\ref{sec:templatecodes} and \ref{sec:trainingcodes} respectively.
 The list of codes are summarized in Table.~\ref{tab:list_of_codes}.
-All code runners were asked to output redshift estimates on 200 linear-spaced bins between redshifts 0 and 2.
+All code runners were asked to output redshift posterior estimates on 200 linear-spaced bins between redshifts 0 and 2.
 
@@
 
-\textsc{EAZY}\footnote{\url{https://github.com/gbrammer/eazy-photoz}} \citep[Easy and Accurate Photometric Redshifts from Yale,][]{Brammer:08} is a template-based photo-$z$ code that includes several features that improve on the basic $\chi^2$ fit used in many template codes.
-It can fit the observed photometry with SEDs created from a linear combination of a set of templates at each redshift, and the best-fit SED is found by simultaneously fitting one, two or all of the templates by minimizing $\chi^2$.
-The minimized $\chi^2(z)$ is then combined with an apparent magnitude prior to obtain the posterior redshift probability distribution, although some argue that this is not the mathematically correct way of calculating the posteriors.
+\textsc{EAZY}\footnote{\url{https://github.com/gbrammer/eazy-photoz}} \citep[Easy and Accurate Photometric Redshifts from Yale,][]{Brammer:08} is a template-based photo-$z$ code that includes several features that extend the basic $\chi^2$ fit used in many template codes.
+The code can fit the observed photometry with SEDs created from a linear combination of a set of templates at each redshift, and the best-fit SED is found by simultaneously fitting one, two or all of the templates by minimizing $\chi^2$.
+The minimized $\chi^2(z)$ is then combined with an apparent magnitude prior to obtain the posterior redshift probability distribution.
+On examination of the source code for EAZY, it appears that rather than marginalizing across all templates in the $\chi^2$ calculation, \textsc{EAZY} takes only the minimum value of $\chi^2$ at each redshift.
+This improper marginalization does not lead to the correct posterior distribution, an implementation issue that will need to be addressed in the future.
+%, although some argue that this is not the mathematically correct way of calculating the posteriors.
 \textsc{EAZY} can also account for the uncertainties in the templates by adding an empirically derived template error in quadrature as a function of redshift to the flux errors.

### 2018-11-13 paper/tex/analysis.tex
address all but one comment in 4.1 and 4.2

--- 2018-08-20 paper/tex/analysis.tex
@@
 
-The probability integral transform (PIT) \citep{Polsterer:16} is defined for each individual galaxy as:%essentially the derivative of the QQ plot, a histogram of
+The probability integral transform (PIT) has been employed recently to evaluate fidelity of posterior distributions  \citep[e.~g.~][]{Bordoloi:10,Polsterer:16,Tanaka:17}.
+The PIT value is defined for each individual galaxy as:%essentially the derivative of the QQ plot, a histogram of
 
@@
 \begin{equation} \label{eq:estimated-cde-loss}
-\mathbb{E}_\X \left[\int \widehat{f}(z \mid \X)^{2} dz\right] - 2 \mathbb{E}_{\X,Z}\left[\widehat{f}(Z \mid \X)\right] + K_{f},
+L(f, \widehat{f}) = \mathbb{E}_\X \left[\int \widehat{f}(z \mid \X)^{2} dz\right] - 2 \mathbb{E}_{\X,Z}\left[\widehat{f}(Z \mid \X)\right] + K_{f},
 \end{equation}
@@
 %We calculate it between the true redshift distribution $N'(z)$ and each estimator $\hat{N}(z)$ (derived by Eq.~\ref{eq:stacked}) as well as between the ideal QQ and PIT curves from each method studied and the ideal curves.
-Though this metric does not account for the fact that the redshift distribution function is, in fact, a probability distribution, it can still be interpreted as a measure of the integrated difference between the estimated distribution and the true distribution, and it can be used to quantify the otherwise qualitative metrics.
+Though this metric does not account for the fact that the redshift distribution function is, in fact, a probability distribution, it can still be interpreted as a measure of the integrated difference between the estimated distribution and the true distribution.%, and it can be used to quantify the otherwise qualitative metrics.
 
@@
 
-The \textit{Kolmogorov-Smirnov statistic} $N_{\rm KS}$ is the maximum difference between $F_{\rm phot}(z)$ and $F_{\rm spec}(z)$, the CDFs of the photo-$z$ and spectroscopic redshift respectively:
+The \textit{Kolmogorov-Smirnov statistic} $N_{\rm KS}$ is the maximum difference between $F_{\rm phot}(z)$ and $F_{\rm true}(z)$, the CDFs of the photo-$z$ and true redshift distributions respectively:
 
 \begin{equation}
-N_{\rm KS} = \max_{z}\left( \left| F_{\rm phot}(z) - F_{\rm spec}(z) \right| \right).
+N_{\rm KS} = \max_{z}\left( \left| F_{\rm phot}(z) - F_{\rm true}(z) \right| \right).
 \end{equation}
@@
 % In this paper we use the CvM statistic to evaluate the performance of both $n(z)$ and $p(z)$.
-The AD statistic is a weighted version of the CvM statistic, making it more sensitive to the tails of the distribution:
+While the KS statistic is a good measure near the median of a distribution, the AD statistic, a weighted version of the CvM statistic, is much more sensitive to the tails of the distribution, providing complementary information:
 

### 2018-11-13 paper/tex/results.tex
address comments for section 5.2

@@
 
-As stated in Section~\ref{sec:metrics}, $p(z)$ is parameterized as $\approx 200$ piecewise constant bins covering $0<z<2$ for all twelve codes, giving a grid size of roughly $\delta z = 0.01$ for each code.
+As stated in Section~\ref{sec:metrics}, $p(z)$ is parameterized as $200$ piecewise constant bins covering $0<z<2$ for all twelve codes, giving a grid size of $\delta z = 0.01$ for each code.
 A piecewise constant grid was a natural choice for some photo-$z$ codes, for instance most template-based codes compute likelihoods on a fixed grid.
 In contrast, FlexZBoost, for example, can return estimates on any grid without compression errors as it’s a basis expansion method where only the expansion coefficients need to be stored.
-Codes with a native output format other than the shared piecewise constant binning scheme (or one that can be losslessly converted to it) may suffer from loss of information when converting to it, which could artificially favor some codes over others.
+Codes with a native output format other than the shared piecewise constant binning scheme (or one that can be losslessly converted to it) may suffer from loss of information when converting to it, which could artificially favor some codes over others in a limited number of cases, for example bright galaxies with very narrow $p(z)$ where the true peak falls between grid points.
+We will discuss PDF storage in Section~\ref{sec:discussion}.
 
@@
 The red line shows the measured quantiles, while the black diagonal represents the ideal QQ values if the distribution were perfectly reproduced.
-A second panel below the main panel for each code shows the difference between $Q_{data}$ and $Q_{theory}$, i.~e.~the departure from the diagonal, for clarity.
+A second panel below the main panel for each code shows the difference between $Q_{\rm data}$ and $Q_{\rm theory}$, i.~e.~the departure from the diagonal, for clarity.
 Biases and trends in whether the average width of the $p(z)$ values being over/under-predicted are evident.
@@
 
-Examination of the PIT histograms and QQ plots shows that there are fairly generic issues with the width of $p(z)$ uncertainties: \textsc{Delight, CMNN, SkyNet} and \textsc{TPZ} all show a PIT histogram with an dearth of low values and and an excess of high values, signs that, on average, their $p(z)$ are more broad than the true distribution of redshifts.
+Examination of the PIT histograms and QQ plots shows that there are fairly generic issues with the width of $p(z)$ uncertainties: \textsc{Delight, CMNN, SkyNet} and \textsc{TPZ} all show a PIT histogram with a dearth of low values and and an excess of high values, signs that, on average, their $p(z)$ are more broad than the true distribution of redshifts.
 \textsc{METAPhoR} shows the opposite trend, indicating the the $p(z)$  are more narrow than the distributions given by the true redshifts.
@@
 \caption{A visual representation of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the PIT distributions.
-The statistics are often highly correlated, though the AD statistic truncates the extrema of the distribution and can have disparate values compared to KS and CvM.} \label{fig:pit_stats}
+The statistics are often highly correlated, with similar relative values between metrics for each code; however, the AD statistic excludes some values at the extrema of the distributions, and can have disparate values compared to KS and CvM.} \label{fig:pit_stats}
 \end{figure*}
@@
 Two of the machine learning based codes, \textsc{ANNz2} and \textsc{METAPhoR}, appear to be over-trained, adding excess galaxy probability to the redshift peaks with the largest number of training galaxies, and missing probability in the troughs where training galaxies are of fewer number.
-Given that our training data is drawn from the same galaxy population as the test set, and our data has prominent peaks in $N'(z)$, perhaps it is not unexpected that such overtraining occurs in some codes, though the fact that it does not occur in all training-based codes indicates that it may be due to specifics of the implementations of \textsc{ANNz2} and \textsc{METAPhoR}.
-A more extensive training/validation set might allow for a better choice of smoothing parameters in individual codes that would avoid such overtraining for these particular codes.
+Given that our training data is drawn from the same galaxy population as the test set, and our data has prominent peaks in $N'(z)$, perhaps it is not unexpected that such overtraining occurs in some codes, though the fact that it does not occur in all training-based codes indicates that it may be due to specifics of the implementations and bandwidth choices of \textsc{ANNz2} and \textsc{METAPhoR}.
+This, once again, emphasizes that care much be taken in choice of bandwith parameters in individual codes in order to avoid such overtraining.
 %\aim{Move the text on Scott's Rule to Discussion?}
@@
 As previously stated, there are a variety of scientific use cases for photo-$z$'s in large upcoming surveys, and care must be taken in how the metrics used to optimize catalog photometric redshifts are defined as well as in how they are used.
+This implies that we may need multiple photo-z estimators, tuned to the particular metric, in order to maximize science returns in upcoming surveys.
 In addition, very few scientific use cases will employ the overall $\hat{N}(z)$ with no cuts, as we explore in this paper.

### 2018-11-14 paper/tex/results.tex
address section 5.2 comments, remove seaborn style from fig1 figs

@@
 %\centering
-\includegraphics[width=0.49\textwidth]{fig/pz_12codes_261931_crop.jpg}\includegraphics[width=0.5\textwidth]{fig/pz_12codes_471167_crop.jpg}\\
-\includegraphics[width=0.49\textwidth]{fig/pz_12codes_713178_crop.jpg}\includegraphics[width=0.49\textwidth]{fig/pz_12codes_982747_crop.jpg}
+\includegraphics[width=0.49\textwidth]{fig/pz_12codes_261931_noseaborn_crop.jpg}\includegraphics[width=0.5\textwidth]{fig/pz_12codes_471167_noseaborn_crop.jpg}\\
+\includegraphics[width=0.49\textwidth]{fig/pz_12codes_713178_noseaborn_crop.jpg}\includegraphics[width=0.49\textwidth]{fig/pz_12codes_982747_noseaborn_crop.jpg}
 \caption{Four illustrative examples of individual p(z) distributions produced by the codes.

### 2018-11-15 paper/tex/discussion.tex
address most of section 6 comments

@@
 %(Sam Schmidt, Ofer Lahav, Jeff Newman, Alex Malz, Eve Kovacs, Tony Tyson, Tina Peters)
-In this paper we presented results evaluating the photometric redshift PDF computation for eleven photo-$z$ codes.
+In this paper we presented results evaluating the computation of individual galaxy photometric redshift PDFs for twelve photo-$z$ codes.
 As discussed in Section~\ref{sec:metrics} the $p(z)$ should accurately reflect the relative likelihood as a function of redshift for each galaxy.
-All codes were provided a set of representative training data and tested on an idealized set of model galaxies with high signal-to-noise and photometry with no confounding effects due to blending, instrumental effects, the night sky, etc...
-included.
+All codes were provided a set of representative training data and tested on an idealized set of model galaxies with high signal-to-noise and photometry with no confounding effects due to blending, instrumental effects, the night sky, or other complications included.
 The goal was not to determine a ``best'' photo-$z$ code: in many ways, this was a baseline test of a ``best case scenario'' to predict the expected photo-$z$ performance if a stage IV dark energy survey was to obtain complete training samples and perfectly calibrated their multi-band photometry.
-Given these idealized conditions, any deficiencies observed in a photo-$z$ code's performance should be a cause for concern, and may be evidence in a problem with either/both of the specific code implementation or the underlying algorithm.
-In order to meet the stringent LSST requirements on photo-$z$ performance, identifying and correcting such problems is an important first step before tackling more realistic data in future challenges.
-Most of the codes tested performed well, however, several did not meet the stringent goals that have been laid out for LSST photometric redshift performance.
+Given these idealized conditions, any deficiencies observed in a photo-$z$ code's performance should be a cause for concern, and may be evidence of a problem with either/both of the specific code implementation or the underlying algorithm.
+In order to meet the stringent LSST goals for photo-$z$ performance, identifying and correcting such problems is an important first step before tackling more realistic data in future challenges.
+Many of the codes tested performed well; however, several did not meet the stringent goals that have been laid out for LSST photometric redshift performance for individual galaxies, as laid out in the LSST SRD (See Section~\ref{sec:intro}.
 This is a cause for concern, given the idealized conditions, and the individual code responses will be studied in detail moving forward.
+If methods can not reach the goals on idealized data, then they will almost surely not meet those same goals when the more complex problems that we expect to arise from real LSST data are included.
+The results presented in this paper enable an evaluation of which algorithms are the most promising moving forward, and potentially point to implementation choices or mistakes which could be improved or corrected in others.
+
 One obvious trend in several of the codes tested was an overall over or underprediction of the widths of $p(z)$, as evidenced by the QQ plots and PIT histograms shown in Fig.~\ref{fig:pitqq}.
@@
 For narrow peaked $p(z)$ the parameterization of the PDF as evaluated on a fixed redshift grid could also have contributed to some overestimates of $p(z)$ width simply due to the finite resolution.
-After evaluating resuts such as those presented in \citet[]{Malz:qp}, in future analyses we plan to switch from a fixed grid to quantile-based storage of $p(z)$ in order to more efficiently and accurately store redshift PDF results.
+After evaluating results such as those presented in \citet[]{Malz:qp}, in future analyses we plan to switch from a fixed grid to quantile-based storage of $p(z)$ in order to more efficiently and accurately store redshift PDF results.
 
@@
 The specific selection used to determine these bins, both algorithmically and the specific bin boundaries, could induce biases due to indirect selections inherent in the photo-$z$ or other bin selection parameters.
-The effects of tomographic bin selection will be explored in a dedicated future paper.
-\red{[are there any references for this?
+The effects of tomographic bin selection will be explored in a dedicated future paper, including propagation of redshift uncertainties in a set of fiducial tomographic redshift bins in order to estimate impact on cosmological parameter estimation.
+
+%\red{[are there any references for this?
 I remember Gary Bernstein talking about this at a photo-z workshop in Japan, but I don't know that it was published.
 I believe Michael Troxel has discussed this as well.]}
-We also plan to propagate the uncertainties measured in a set of fiducial tomographic redshift bins in order to estimate impact on cosmological parameter estimation.
 %\item \red{cosmological parameters in DC2}
@@
 
-In future papers we will add more and more complexity to our simulated data in order to test photo-$z$ algorithms in increasingly realistic conditions.
+In future papers a focus of the {\it LSST Dark Energy Science Collaboration Photo-z Working Group} will be to add more and more complexity to our simulated data in order to test photo-$z$ algorithms in increasingly realistic conditions.
 The most pressing concern is the impact of incomplete spectroscopic training samples.
+The SEDs for the galaxy sample in this paper were constructed form linear combinations of five basis SED templates.
+Future simulations will also include more complex SED information, with a more realistic range of physical properties, and the inclusion of AGN effects.
 As discussed extensively in \citet{Newman:2015} a representative set of spectroscopically confirmed galaxies spanning the full range of both redshift and apparent magnitude is necessary as a training set to characterize the mapping from broad-band fluxes to photometric redshifts.
-However, due to a combination of factors due to both the galaxy SEDs and limitations of spectrographic instruments, redshift samples are known to be systematically incomplete, where certain galaxy types and redshift intervals fail to yield a redshift even at the longest integration times on current and near-future instruments.
+%However, due to a combination of factors due to both the galaxy SEDs and limitations of spectrographic instruments, redshift samples are known to be systematically incomplete, where certain galaxy types and redshift intervals fail to yield a redshift even at the longest integration times on current and near-future instruments.
 The more representative the training data, the better the performance of photo-$z$ algorithms will be.
-Current and upcoming surveys are putting in significant effort into obtaining these training samples \citep[e.~g.\,][]{Masters:2017}, however we still expect significant incompleteness for LSST-like samples, particularly at faint magnitudes.
-One major focus of an upcoming {\it LSST Dark Energy Science Collaboration Photo-z Working Group} data challenge is to produce a realistically incomplete training set of spectroscopic galaxies, modeling the performance of spectrographs, emission-line properties, and expected signal-to-noise to determine which galaxies will fail to yield a secure redshift.
+Current and upcoming surveys are putting significant effort into obtaining these training samples \citep[e.~g.\,][]{Masters:2017}, however we still expect significant incompleteness for LSST-like samples, particularly at faint magnitudes.
+We plan to produce a realistically incomplete training set of spectroscopic galaxies, modeling the performance of spectrographs, emission-line properties, and expected signal-to-noise to determine which galaxies will fail to yield a secure redshift.
 In addition to outright redshift failures we will model the inclusion of a small number of falsely identified secure redshifts where misidentified emission lines or noise spikes cause an incorrect redshift solution to be marked as a high quality identification.
@@
 
-This initial paper explored a data set that was constructed at the catalog level, with no inclusion of the complications that come from measuring photometry from images.
-Future data challenges will move to catalogs constructed from mock images, including effects that will have great impact on photo-$z$ measurements.
-Object blending will be a major area of investigation, as the mixing of flux from multiple objects and the resultant change in measured colours is predicted to affect a large fraction of LSST galaxies \citep{Dawson:2016}, and will be one of the major contributing systematics for photo-$z$'s.
-Inclusion of differing observing conditions (seeing, clouds, variations in filter curves, Galactic dust, ...), as well as models for instrumental and system effects, sky masks, will all impact object photometry, and will be explored in the upcoming data challenge and their impacts described in upcoming papers.
-All underlying SEDs were parameterized as a weighted combination of five basis SEDs, with no additional accounting for host galaxy dust obscuration beyond what was encoded in the basis templates.
-This, in effect, limited the simulation to a very simple model of internal obscuration.
-Future simulations will include a more complicated and realistic treatment of host galaxy dust.
-
-The underlying simulation used in this work was based on a light-cone constructed to a maximum redshift of $z=2$.
+The underlying dataset limited this work to a maximum redshift of $z=2$.
 LSST imaging after 10 years of observations will include a significant number of $z>2$ galaxies in expected cosmology samples, and their inclusion does have potential significant implications for photo-$z$ measures: the high redshift galaxies lie at fainter apparent magnitudes and can have anomalous colours due to evolution of stellar populations and the shift to rest-frame magnitudes probing UV features of the underlying SED.
@@
 Extending simulations to include the high-redshift galaxy population will be a priority in future data challenges.
+
+This initial paper explored a data set that was constructed at the catalog level, with no inclusion of the complications that come from measuring photometry from images.
+Future data challenges will move to catalogs constructed from mock images, including effects that will have great impact on photo-$z$ measurements, which will naturally include the complications of object blending, sensor effects, different observing conditions, amongst others.
+Object blending will be a major area of investigation, as the mixing of flux from multiple objects and the resultant change in measured colours is predicted to affect a large fraction of LSST galaxies \citep{Dawson:2016}, and will be one of the major contributing systematics for photo-$z$'s.
+
+Finally, while this paper and future papers discussed above focus on photometric redshift codes and estimating accurate $p(z)$ from training data, we plan a separate, but complementary, project to examine calibration of the resultant redshifts via spatial cross-correlations \citep{Newman:2008}, which will be explored in a separate set of papers.
+The overarching plan describing everything laid out in this section is described in more detail in the LSST DESC Science Roadmap (see Footnote in Section~\ref{sec:intro}).
+These plans will require significant effort, but they are necessary if we are to make optimal use of the LSST data for astrophysical and cosmological analyses.

### 2018-11-16 paper/main.tex
adress appendix comments, add refs for code packages

--- 2018-08-20 paper/main.tex
@@
 
-The authors would like to thank their LSST-DESC publication review committee.
+The authors would like to thank their LSST-DESC publication review committee for comments that improved the paper draft.
 
 AIM is advised by David W. Hogg and was supported by National Science Foundation grant AST-1517237.
+
+In addition to packages cited in the text, analyses performed in this paper used the following software packages: \textsc{Numpy} and \textsc{Scipy} \citep{numpyscipy}, \textsc{Matplotlib} \citep{matplotlib}, \textsc{Seaborn} \citep{seaborn},
 

### 2018-11-16 paper/tex/appendix.tex
finished adding re-run of GPz with proper training
adress appendix comments, add refs for code packages

@@
 \label{sec:pointmetrics}
-While we do not recommend the use of single point estimates of redshift for most science applications, plots of the point estimates can be a useful qualitative diagnostic of photo-z code performance, i.~e.~examining point photo-$z$ vs.~spec-$z$ plots visually can give a quick impression of some common trends in different codes.
-Computing point estimate statistics may also be useful for more direct comparisons with previous photo-z evaluations.
+While we assume that all science analysis will use full PDF information and do not recommend the use of single point estimates of redshift for most science applications, we include a brief evaulation as an Appendix.
+Plots of the point estimates can be a useful qualitative diagnostic of photo-z code performance, i.~e.~examining point photo-$z$ vs.~spec-$z$ plots visually can give a quick impression of some common trends in different codes.
+Computing point estimate statistics may also be useful for more direct historical comparisons from older photo-z evaluations.
 If a point-estimate is preferred for a specific science case, it is fairly simple to compute the mean, mode, or some other simple estimator from each $p(z)$, so these point estimates can be easily derived from the stored $p(z)$.
@@
 
-Finally, we note that all eleven codes are at or near the goals for point-estimates as outlined in the LSST Science Requirements Document\footnote{available at: \url{http://ls.st/srd}} and \citet{Graham:17}.
+Finally, we note that all twelve codes perform at or near the goals for point-estimates as outlined in the LSST Science Requirements Document\footnote{available at: \url{http://ls.st/srd}} and \citet{Graham:17}.
 This is to be expected, given that the requirements were designed such that a point estimate photo-z would meet these requirements for perfect training data to a depth of $i<25$.
@@
 \textsc{FlexZBoost}& 0.0154  & -0.00027  & 0.020      & 0.0148  & -0.000211  & 0.017 \\
-\textsc{GPz}       & 0.0202  & -0.00091  & 0.036      & 0.0201  & -0.000950  & 0.037 \\
+\textsc{GPz}       & 0.0197  & -0.00000  & 0.052      & 0.0195  &  0.000113  & 0.051 \\
 \textsc{LePhare}   & 0.0236  & -0.00161  & 0.058      & 0.0239  & -0.002007  & 0.056 \\

### 2018-11-16 paper/tex/appendix.tex
update appendix figures

@@
 \centering
-\includegraphics[width=0.8\textwidth]{fig/ZPEAK_szpz_threecolumn_12codes_navy.jpg}\\
-\includegraphics[width=0.8\textwidth]{fig/ZWEIGHT_szpz_threecolumn_12codes_navy.jpg}
+\includegraphics[width=0.8\textwidth]{fig/ZPEAK_szpz_threecolumn_12codes_navy_lowalpha.jpg}\\
+\includegraphics[width=0.8\textwidth]{fig/ZWEIGHT_szpz_threecolumn_12codes_navy_lowalpha.jpg}
 \caption{Point estimate photo-z's derived from the posteriors.

### 2018-11-20 paper/main.tex
add more software refs

--- 2018-08-20 paper/main.tex
@@
 
-The authors would like to thank their LSST-DESC publication review committee.
+The authors would like to thank their LSST-DESC publication review committee for comments that improved the paper draft.
 
 AIM is advised by David W. Hogg and was supported by National Science Foundation grant AST-1517237.
+
+In addition to packages cited in the text, analyses performed in this paper used the following software packages: \textsc{Numpy} and \textsc{Scipy} \citep{numpyscipy}, \textsc{Matplotlib} \citep{matplotlib}, \textsc{Seaborn} \citep{seaborn}, \textsc{minFunc} \citep{minfunc}, and {\textsc photUtils} in the LSST simulations package \citep{lsstphotutils}.
 

### 2018-11-20 paper/main.tex
add skynet bonnett ref

--- 2018-11-16 paper/main.tex
@@
 
-In addition to packages cited in the text, analyses performed in this paper used the following software packages: \textsc{Numpy} and \textsc{Scipy} \citep{numpyscipy}, \textsc{Matplotlib} \citep{matplotlib}, \textsc{Seaborn} \citep{seaborn},
+In addition to packages cited in the text, analyses performed in this paper used the following software packages: \textsc{Numpy} and \textsc{Scipy} \citep{numpyscipy}, \textsc{Matplotlib} \citep{matplotlib}, \textsc{Seaborn} \citep{seaborn}, \textsc{minFunc} \citep{minfunc}, \textsc{pySkyNet} \citep{pyskynet}, and {\textsc photUtils} from the LSST simulations package \citep{lsstphotutils}.
 

### 2018-11-21 paper/main.tex
change title (temp) and few edits

--- 2018-11-16 paper/main.tex
@@
 
-\title{An assessment of photometric redshift PDF performance in the context of LSST}
+\title{Implicit assumptions and their impact on photometric redshift PDF performance in the context of LSST}
 % \title{Photometric redshift uncertainties for LSST \romannumeral{1}: A unified methodology for assessing photo-$z$ PDF performance}
@@
 
-The authors would like to thank their LSST-DESC publication review committee for comments that improved the paper draft.
+The authors would like to thank their LSST-DESC publication review committee.
 
 AIM is advised by David W. Hogg and was supported by National Science Foundation grant AST-1517237.
-
-In addition to packages cited in the text, analyses performed in this paper used the following software packages: \textsc{Numpy} and \textsc{Scipy} \citep{numpyscipy}, \textsc{Matplotlib} \citep{matplotlib}, \textsc{Seaborn} \citep{seaborn},
 

### 2018-11-21 paper/tex/introduction.tex
change title (temp) and few edits

@@
 The unprecedented sample size of LSST galaxies, expected to number several billion for the main cosmological sample, necessitates stringent constraints on photo-$z$ accuracy if systematic errors are not to dominate the statistical errors.
-The LSST Science Requirements Document (SRD)\footnote{available at \url{https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17}} lists the photometric redshift goals for a magnitude limited sample with $i<25$ as: root-mean-square error with a goal of $\sigma_z<0.02(1+z)$; $3\sigma$ ``catastrophic outlier'' rate below $10\%$; bias below $0.003$\footnote{
+The LSST Science Requirements Document (SRD)\footnote{available at \url{https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17}} lists the individual galaxy photometric redshift goals for a magnitude limited sample with $i<25$ as: root-mean-square error with a goal of $\sigma_z<0.02(1+z)$; $3\sigma$ ``catastrophic outlier'' rate below $10\%$; bias below $0.003$\footnote{
 Note that at the time the SRD was written, these goals were stated in terms of a photo-$z$ point estimate for each galaxy, as was standard in many previous studies, while in this paper we emphasize the importance of using a full photo-$z$ PDF.}.
 The LSST Dark Energy Science Collaboration (LSST-DESC) developed a separate Science Requirements Document \citep{Mandelbaum:2018}, which forecasts the constraining power of five cosmological probes using somewhat conservative assumptions to define requirements on systematic errors for several measurements.
-These include even more stringent requirements on photometric redshift performance than those included in the LSST SRD, though the initial LSST-DESC requirements are defined for tomographic bin populations.
+These include even more stringent requirements on photometric redshift performance than those included in the LSST SRD, though most of the initial LSST-DESC requirements are defined in terms of tomographic bin populations rather than on individual object redshifts.
 The tremendous size of LSST's galaxy catalogue will be enabled by its exceptional depth, pushing to fainter magnitudes and deeper imaging and including galaxies of lower luminosity and higher redshift than ever before.
@@
 Other science cases seem to need only ensemble redshift information; for instance many current cosmic shear techniques require only the overall redshift distribution $N(z)$ for tomographic redshift samples.
-However, recent data-driven techniques employing hierarchical Bayesian or Gaussian Process methods have emerged that calibrate redshift distributions using individual $p(z)$ estimates \citep[e.~g.~][]{Sanchez:2018}.
+However,  even such cases require individual object redshift estimates for portions of the analysis, for example in determining galaxy intrinsic alignments in weak lensing samples.
+In addition, recent data-driven techniques employing hierarchical Bayesian or Gaussian Process methods have emerged that calibrate redshift distributions using individual $p(z)$ estimates \citep[e.~g.~][]{Sanchez:2018}.
 These methods assume that the $p(z)$ for each galaxy is an accurate PDF, and such methods break down if this assumption is invalid.
@@
 \citet{Rau:2015} introduced a new method for improving redshift PDFs using an ordinal classification algorithm.
-DES compared several codes for point estimates \citep{Sanchez:14} and a summary statistic of photo-$z$ interim posteriors for tomographically binned galaxy subsamples \citep{Bonnett:16}.  
+DES compared several codes for point estimates and a subset with $p(z)$ information \citep{Sanchez:14}.
+A follup up paper examined summary statistics of photo-$z$ interim posteriors for tomographically binned galaxy subsamples \citep{Bonnett:16}.
 This paper is distinguished by its focus on metrics of photo-$z$ interim posteriors themselves and consideration of both classic and state-of-the-art photo-$z$ algorithms, comparing the performance of several of the most widely employed codes as well as some that have been developed only recently on the basis of metrics appropriate for a probabilistic data product.
@@
 This work is laid out in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}} as one of the critical activities to be completed in preparation for dark energy science analysis on the first year LSST data.
-This is the first of multiple papers by the working group, which will grow in sophistication.
+%This is the first of multiple papers by the working group, which will grow in sophistication.
 In this initial paper we focus on evaluating the performance of photometric redshift codes and PDF-based performance metrics in the presence of representative training sets.
 Comparing the relative performance of the codes enables us to evaluate whether each code is using information in an optimal way, and may reveal enhancements in some codes and deficiencies in others, either in the fundamental algorithm, or in specific implementation.
-We note that these initial tests are a necessary requirement for photo-$z$ codes that will be used in cosmological analyses; however, meeting these requirements is only the first stage in the process of accounting for incomplete training samples, object blending,  the robustness of estimated redshift distributions when combining redshift estimates in tomographic bins, and other complications that realistic samples will encounter in estimating photometric redshifts.
-This can be thought of as an initial test under near perfect conditions before further complexities are added in future papers.
+We note that these initial tests are a necessary requirement for photo-$z$ codes that will be used in cosmological analyses; however, meeting these requirements is only the first stage in the process, and can be thought of as an initial test under near perfect conditions before further complexities are added in future analyses.
 
@@
 Though photo-$z$ PDFs are often written simply as $p(z)$, the PDF itself must be an interim posterior distribution $p(z | d, I)$, the probability of redshift conditioned on photometric data $d$ that has actually been observed and the prior information $I$ that guides how a redshift is extracted from the photometry.
-If we run multiple photo-$z$ codes on a single dataset, the photo-$z$ interim posteriors will not be identical because each code is based on assumptions in the form of an interim prior --- these assumptions form the premise for photo-$z$ estimation as a whole and are the only way to introduce differences in estimates of what would otherwise be a shared photo-$z$ posterior $p(z | d)$ regardless of the code used to obtain it.
-Though explicit knowledge of the interim prior is necessary to use photo-$z$ interim posteriors self-consistently in physical inference, the interim prior of a particular methodology is often implicit and not necessarily shared among all galaxies in the catalogue.
+If we run multiple photo-$z$ codes on a single dataset, the photo-$z$  posteriors will not be identical because each code employs assumptions that constitute an interim prior; these assumptions form the premise for photo-$z$ estimation as a whole and are the only way to introduce differences in estimates of what would otherwise be a shared photo-$z$ posterior $p(z | d)$ regardless of the code used to obtain it.
+Though explicit knowledge of the interim prior is necessary to use photo-$z$ posteriors self-consistently in physical inference, the interim prior of a particular methodology is often implicit and not necessarily shared among all galaxies in the catalogue.
 
@@
 In order to isolate the effects encapsulated by $I_{H}$ of variation between codes from issues with the training set or template library encapsulated by $I_{D}$, we use an identical set of simulated galaxies for every code and construct a template library and training sample that are {\it complete and representative} and shared among all codes; that is, our training sample for machine learning codes is drawn from the same underlying galaxy population as our test set, with no additional selections, and the SED library used for template-based codes is the same as the one used to generate the photometric data.
-We explore a number of performance metrics in this paper, not to make a conclusion regarding the superiority or even relative favorability of each code but to establish a method for comparing photo-$z$ PDFs derived by different methods.
+We explore a number of performance metrics in this paper, in the controlled environment of the representative training data in order to establish a method for comparing photo-$z$ PDFs derived by different methods.
+%not to make a conclusion regarding the superiority or even relative favorability of each code but to establish a method for comparing photo-$z$ PDFs derived by different methods.
 These test conditions set the stage for addressing in a future paper the crucial issue of incomplete and non-representative prior information.

### 2018-11-22 paper/tex/data.tex
add sentence with 2percen nondetect
add mention of blinding

@@
 Thus, further mentions of ``non-detections'' refer to objects that would be ``looked at but not seen'' in multi-band forced photometry, and the photo-z codes will treat them as such.
+Only 2.0 per cent of our sample galaxies contain a photometric band with a non-detection, the vast majority of which are in the $u$-band.
 
@@
 All subsequent results will evaluate this ``gold sample'' test set.
+In order to blind results, initially redshifts were not revealed for the ``test'' set, and were only supplied for the training sample galaxies.
+This prevented code runners from tweaking results and over-fitting to the specific test set.
 

### 2018-11-26 paper/tex/analysis.tex
add explanation to eq 17

@@
 \end{equation}
-This loss is the CDE equivalent of the RMSE in regression. To estimate this loss we rewrite the loss as
+This loss is the CDE equivalent of the RMSE in regression. To estimate this loss we rewrite Eq.~\ref{eq:cde-loss} as
 \begin{equation} \label{eq:estimated-cde-loss}
@@
 \end{equation}
-where the first expectation is with respect to the marginal distribution of the covariates $\X$, the second expectation is with respect to the joint distribution of $\X$ and $Z$, and $K_{f}$ is a constant depending only upon the true conditional densities $f(z\mid \x)$.
+where upper-case letters denote random variables and lower-case the observed variables.
+The first expectation is with respect to the marginal distribution of the covariates $\X$, the second expectation is with respect to the joint distribution of $\X$ and $Z$, and $K_{f}$ is a constant depending only upon the true conditional densities $f(z\mid \x)$.
 For each method we can estimate these expectations as empirical expectations on the test or validation data \citep[Eq.~7 in][]{Izbicki:17b} without knowledge of the true densities.

### 2018-11-27 paper/tex/introduction.tex
more edits to the introduction

@@
 A follup up paper examined summary statistics of photo-$z$ interim posteriors for tomographically binned galaxy subsamples \citep{Bonnett:16}.
+
 This paper is distinguished by its focus on metrics of photo-$z$ interim posteriors themselves and consideration of both classic and state-of-the-art photo-$z$ algorithms, comparing the performance of several of the most widely employed codes as well as some that have been developed only recently on the basis of metrics appropriate for a probabilistic data product.
-The results presented in this work are
+The results presented here are
 %part of the Data Challenge 1 (DC1)
@@
 %This is the first of multiple papers by the working group, which will grow in sophistication.
-In this initial paper we focus on evaluating the performance of photometric redshift codes and PDF-based performance metrics in the presence of representative training sets.
+In this initial paper we focus on evaluating the performance of photometric redshift codes and PDF-based performance metrics in the presence of complete and representative training sets.
+Specific implementation choices in each code will influence the resultant posterior distributions, for example choice of prior parameterization in template-based codes, the bandwidth size chosen for machine learning based codes, or even the output format chosen for storing the PDF.
+We have attempted to minimize the impact of many of these factors when comparing codes, for example by using the same template set for all template-based codes, and using a training set that is drawn from the same underlying population as the test sample, to create a controlled environment in which to compare the photo-$z$ PDFs derived from each method.
+We explore a number of performance metrics in this paper that test whether the posterior estimates are actual PDFs.
 Comparing the relative performance of the codes enables us to evaluate whether each code is using information in an optimal way, and may reveal enhancements in some codes and deficiencies in others, either in the fundamental algorithm, or in specific implementation.
-We note that these initial tests are a necessary requirement for photo-$z$ codes that will be used in cosmological analyses; however, meeting these requirements is only the first stage in the process, and can be thought of as an initial test under near perfect conditions before further complexities are added in future analyses.
+Identifying and fixing failure modes within codes may aid us in reaching the stringent photo-$z$ performance goals set out for LSST.
+We note that these initial tests are a necessary requirement for photo-$z$ codes that will be used in cosmological analyses; however, meeting these requirements is only the first stage in the process, and can be thought of as an initial test under near perfect conditions to test for problems before further complexities are added in future analyses.
 
-Before moving forward, we must address how the best methods may be unique to the performance metrics and science cases considered and what distinguishes photo-$z$ PDFs of different methods from one another.
-Though photo-$z$ PDFs are often written simply as $p(z)$, the PDF itself must be an interim posterior distribution $p(z | d, I)$, the probability of redshift conditioned on photometric data $d$ that has actually been observed and the prior information $I$ that guides how a redshift is extracted from the photometry.
-If we run multiple photo-$z$ codes on a single dataset, the photo-$z$  posteriors will not be identical because each code employs assumptions that constitute an interim prior; these assumptions form the premise for photo-$z$ estimation as a whole and are the only way to introduce differences in estimates of what would otherwise be a shared photo-$z$ posterior $p(z | d)$ regardless of the code used to obtain it.
-Though explicit knowledge of the interim prior is necessary to use photo-$z$ posteriors self-consistently in physical inference, the interim prior of a particular methodology is often implicit and not necessarily shared among all galaxies in the catalogue.
+%Before moving forward, we must address how the best methods may be unique to the performance metrics and science cases considered and what distinguishes photo-$z$ PDFs of different methods from one another.
+Though photo-$z$ PDFs are often written simply as $p(z)$, the PDF itself is an interim posterior distribution $p(z | d, I)$, the probability of redshift conditioned on photometric data $d$ that has been observed and the prior information $I$ that guides how a redshift is extracted from the photometry.
+If we run multiple photo-$z$ codes on a single dataset, the photo-$z$ posteriors will not be identical because each code employs assumptions that constitute an interim prior (for example, the specific template set used in a template-based code, or the bandwidth/number of neighbors in a machine learning based code).
+%These assumptions form the premise for photo-$z$ estimation as a whole and are the only way to introduce differences in estimates of what would otherwise be a shared photo-$z$ posterior $p(z | d)$ regardless of the code used to obtain it.
 
-This paper therefore aims (1) to constrain the impact of the interim prior $I$ by separating it into a component $I_{H}$ representing the method itself and a component $I_{D}$ representing physical information, such as a training set or SED template library and (2) to present a procedure for evaluating the performance of photo-$z$ codes in generic tests that may include many more systematics in the interim prior $I$.
-In order to isolate the effects encapsulated by $I_{H}$ of variation between codes from issues with the training set or template library encapsulated by $I_{D}$, we use an identical set of simulated galaxies for every code and construct a template library and training sample that are {\it complete and representative} and shared among all codes; that is, our training sample for machine learning codes is drawn from the same underlying galaxy population as our test set, with no additional selections, and the SED library used for template-based codes is the same as the one used to generate the photometric data.
-We explore a number of performance metrics in this paper, in the controlled environment of the representative training data in order to establish a method for comparing photo-$z$ PDFs derived by different methods.
+%Though explicit knowledge of the interim prior is necessary to use photo-$z$ posteriors self-consistently in physical inference, the interim prior of a particular methodology is often implicit and not necessarily shared among all galaxies in the catalogue.
+
+%This paper aims to constrain the impact of these differing interim priors and isolating the effects due to the differing code methodologies by using a homogeneous
+
+%This paper therefore aims (1) to constrain the impact of the interim prior $I$ by separating it into a component $I_{H}$ representing the method itself and a component $I_{D}$ representing physical information, such as a training set or SED template library and (2) to present a procedure for evaluating the performance of photo-$z$ codes in generic tests that may include many more systematics in the interim prior $I$.
+%In order to isolate the effects encapsulated by $I_{H}$ of variation between codes from issues with the training set or template library encapsulated by $I_{D}$, we use an identical set of simulated galaxies for every code and construct a template library and training sample that are {\it complete and representative} and shared among all codes; that is, our training sample for machine learning codes is drawn from the same underlying galaxy population as our test set, with no additional selections, and the SED library used for template-based codes is the same as the one used to generate the photometric data.
+%We explore a number of performance metrics in this paper, in the controlled environment of the representative training data in order to establish a method for comparing photo-$z$ PDFs derived by different methods.
 %not to make a conclusion regarding the superiority or even relative favorability of each code but to establish a method for comparing photo-$z$ PDFs derived by different methods.
-These test conditions set the stage for addressing in a future paper the crucial issue of incomplete and non-representative prior information.
+%These test conditions set the stage for addressing in a future paper the crucial issue of incomplete and non-representative prior information.
 

### 2018-11-28 paper/tex/discussion.tex
add winnowing comment, few other fixes

@@
 In this paper we presented results evaluating the computation of individual galaxy photometric redshift PDFs for twelve photo-$z$ codes.
-As discussed in Section~\ref{sec:metrics} the $p(z)$ should accurately reflect the relative likelihood as a function of redshift for each galaxy.
+As discussed in Section~\ref{sec:metrics} each $p(z)$ should accurately reflect the relative likelihood as a function of redshift for each galaxy in an informative way; that is, the estimates should provide useful information per individual galaxy, not just the ensemble.
 All codes were provided a set of representative training data and tested on an idealized set of model galaxies with high signal-to-noise and photometry with no confounding effects due to blending, instrumental effects, the night sky, or other complications included.
@@
 Given these idealized conditions, any deficiencies observed in a photo-$z$ code's performance should be a cause for concern, and may be evidence of a problem with either/both of the specific code implementation or the underlying algorithm.
-In order to meet the stringent LSST goals for photo-$z$ performance, identifying and correcting such problems is an important first step before tackling more realistic data in future challenges.
+In order to meet the stringent LSST goals for photo-$z$ performance, identifying and correcting such problems is an important first step before tackling more realistic data in future challenges, and codes that do not perform well may not be worth pursuing in future challenges.
 Many of the codes tested performed well; however, several did not meet the stringent goals that have been laid out for LSST photometric redshift performance for individual galaxies, as laid out in the LSST SRD (See Section~\ref{sec:intro}.

### 2018-11-28 paper/main.tex
remove old conclusions section

--- 2018-11-20 paper/main.tex
@@
 
-\input{./tex/conclusion}
+%\input{./tex/conclusion}
 
@@
 
-The authors would like to thank their LSST-DESC publication review committee for comments that improved the paper draft.
+The authors would like to thank their LSST-DESC publication review committee.
+
+{ \red{personal funding sources}}
 
 AIM is advised by David W. Hogg and was supported by National Science Foundation grant AST-1517237.
-
-In addition to packages cited in the text, analyses performed in this paper used the following software packages: \textsc{Numpy} and \textsc{Scipy} \citep{numpyscipy}, \textsc{Matplotlib} \citep{matplotlib}, \textsc{Seaborn} \citep{seaborn}, \textsc{minFunc} \citep{minfunc}, and {\textsc photUtils} in the LSST simulations package \citep{lsstphotutils}.
 

### 2018-11-28 paper/tex/analysis.tex
addressing comments, propagating macros

--- 2018-11-08 paper/tex/analysis.tex
@@
 
-The overloaded ``$p(z)$'' is a widespread abuse of notation; we would like the outputs of photo-$z$ PDF codes to be interpretable as probabilities.
-Obviously photo-$z$ PDFs must not take negative values and must integrate to unity over the range of possible redshifts.
-Additionally, an estimator derived by method $H$ for the photo-$z$ PDF of galaxy $i$ must be understood as a posterior probability distribution
-
+The overloaded ``$p(z)$'' is a widespread abuse of notation; we would like the outputs of \pzpdf\ codes to be interpretable as probabilities.
+Obviously \pzpdf s must not take negative values and must integrate to unity over the range of possible redshifts.
+Additionally, an estimator derived by method $H$ for the \pzpdf\ of galaxy $i$ must be understood as a posterior probability distribution
 \begin{equation}
@@
 conditioned not only on the photometric data $d_{i}$ for that galaxy but also on parameters encompassing a number of things that will differ depending on the method $H$ used to produce it, namely the assumptions $I_{H}$ necessary for the method to be valid and any inputs $I_{D}$ it takes as prior information, such as a template library or training set.
-Because of this, direct comparison of photo-$z$ PDFs produced by different methods is in some sense impossible; even if they share the same prior information $I_{D}$, by definition they cannot be conditioned on the same assumptions $I_{H}$, otherwise they would not be distinct methods at all.
+Because of this, direct comparison of \pzpdf s produced by different methods is in some sense impossible; even if they share the same prior information $I_{D}$, by definition they cannot be conditioned on the same assumptions $I_{H}$, otherwise they would not be distinct methods at all.
 
@@
 Thus, we are saying
-
 \begin{equation}
@@
 
-As mentioned previously, there are cosmology probes that require knowledge of individual galaxy $p(z)$, for example galaxy intrinsic alignment studies, strong lensing foreground shear prediction, and photometric supernova classification, and others that require only knowledge of the ensemble redshift distribution, $N(z)$.
-Due to the paucity of principled techniques for using and validating individual galaxy photo-$z$ PDFs, there have been few alternatives to the common practice of reducing photo-$z$ PDFs to point estimates when evaluating and comparing photo-$z$ code performance in the literature.
-Though this practice should not be encouraged, we also calculate traditional metrics based on the most common point estimators derived from photo-$z$ PDFs.
+As mentioned previously, there are cosmology probes that require knowledge of individual galaxy's \pzpdf, for example galaxy intrinsic alignment studies, strong lensing foreground shear prediction, and photometric supernova classification, and others that require only knowledge of the ensemble redshift distribution, $N(z)$.
+Due to the paucity of principled techniques for using and validating individual galaxy \pzpdf s, there have been few alternatives to the common practice of reducing \pzpdf s to point estimates when evaluating and comparing \pz\ code performance in the literature.
+Though this practice should not be encouraged, we also calculate traditional metrics based on the most common point estimators derived from \pzpdf s.
 Those seeking to establish a connection to traditional ways of thinking about redshift estimation may consult the Appendix for these results.
 
-There are a number of metrics that can be used to test the accuracy of a photo-$z$ interim posterior as an estimator of a true photo-$z$ posterior if it is known.
-Even for simulated data, the true photo-$z$ PDF is in general not accessible unless the redshifts are in fact drawn from the true photo-$z$ PDFs, a mock catalogue generation procedure that has not yet appeared in the literature.
-Furthermore, only limited applications of photo-$z$ PDFs that could be used as the basis for a metric have been presented in the literature.
-The most popular application by far is the calculation of the overall redshift distribution $N(z)$, the true value of which is known for the \textsc{Buzzard} simulation and will be denoted as $N'(z)$.
-Though alternatives exist \citep{Malz:chippr}, stacking according to
+There are a number of metrics that can be used to test the accuracy of a \pz\ posterior as an estimator of a true \pz\ posterior if it is known.
+Even for simulated data, the true \pzpdf\ is in general not accessible unless the photometry is in fact drawn from the true \pz\ likelihoods, a mock catalogue generation procedure that has not yet appeared in the literature.
+Furthermore, only limited applications of \pzpdf s that could be used as the basis for a metric have been presented in the literature.
 
+The most popular application of \pspdf s by far is the estimation of the overall redshift distribution $N(z)$, the true value of which is known for the \textsc{Buzzard} simulation and will be denoted as $N'(z)$.
+Though alternatives exist \citep{Malz:chippr}, ``stacking'' according to
 \begin{equation} \label{eq:stacked}
-\hat{N}^{H}(z) \approx \frac{1}{N_{tot}}\ \sum_{i}^{N_{tot}}\ \hat{p}^{H}_{i}(z)
+\hat{N}^{H}(z) \equiv \frac{1}{N_{tot}}\ \sum_{i}^{N_{tot}}\ \hat{p}^{H}_{i}(z)
 \end{equation}
-is the most widely accepted method for estimating the redshift distribution from photo-$z$ PDFs.
-If we assume that the response of estimators of $N(z)$ is uniform across all approaches $H$, then we may interpret metrics on the accuracy of $\hat{N}(z)$ obtained in this way.
+is the most widely accepted method for obtaining the stacked estimator $\hat{N}^{H}(z)$ of the redshift distribution from \pzpdf s derived by a method $H$.
+Though we do not endorse the use of the stacked estimator of the redshift distribution, we use it under the assumption that the response of our metrics of $\hat{N}^{H}(z)$ will be analogous to the same metrics applied to a principled estimator of the redshift distribution.
 We must note, however, that this is a poor assumption in general.
-Under the setup of this paper, the true redshift distribution $N'(z) = p(z | I_{D})$ (i.e.
-because our training data is representative, the interim prior is the truth).
-In this ideal case, with representative training data, the method that would give the best approximation to $N'(z)$ would be one that neglects all the information contained in the photometry $\{d_{i}\}_{N_{tot}}$ and gives every galaxy the same photo-$z$ PDF $\hat{p}_{i}(z) = N'(z)$ for all $i$.
-This is the exact estimator, \trainz, that we have described in Section~\ref{sec:method:trainz}, and which will serve as a point of reference for the other codes.
 
-The exact implementation of the stacked estimator $\hat{N}^{H}(z)$ will depend on the parametrization of the photo-$z$ PDFs, which may differ across codes and can affect the precision of the estimator \citep{Malz:qp}; even considering a single method under the same parametrization, say a piecewise constant function over bins or a set of samples from the posterior, an estimator using $2N$ bins or samples will trivially be more precise than an estimator using $N$ bins or samples.
+Returning to the prior of \pzpdf s, the true redshift distribution satisfies the tautology $N'(z) = p(z | I_{D})$, because our training data is representative, $I_{D}$ represents a prior that is also equal to the truth.
+In this ideal case of representative training data, the method that would give the best approximation to $N'(z)$ would be one that neglects all the information contained in the photometry $\{d_{i}\}_{N_{tot}}$ and gives every galaxy the same \pzpdf\ $\hat{p}_{i}(z) = N'(z)$ for all $i$.
+(In fact, including any information from the photometry would only add noise to the optimal result of returning the prior for every galaxy.)
+This is the exact estimator, \trainz, that we have described in Section~\ref{sec:method:trainz}, and which will serve as an experimental control.
+
+The exact implementation of the stacked estimator $\hat{N}^{H}(z)$ should depend on the parametrization of the \pzpdf s, which may differ across codes and can affect the precision of the estimator \citep{Malz:qp}; even considering a single method under the same parametrization, say a piecewise constant function over bins or a set of samples from the posterior, an estimator using $2N$ bins or samples will trivially be more precise than an estimator using $N$ bins or samples.
 %\red{Note: Let's address how the gridding (endpoints and spacing) affects the metrics, from the email correspondence Re: [LSST-DESC-MEMBERS] Summary of Nov 29 PZ Telecon. (AIM)}
-In order to minimize the effects of such choices, we asked those running all eleven codes to output $p(z)$ parameterized with $\approx 200$ piecewise constant bins spanning $0<z<2$.
-The piecewise constant format is chosen because of its established presence in the literature, and the choice of 200 bins was motivated by the approximate number of columns expected to be available for storage of $p(z)$ for the final LSST Project tables.\footnote{See, e.~g.~the LSST Data Products Definition Document, available at: \url{https://ls.st/dpdd}}
-All $p(z)$ catalogues are processed using the \textsc{qp} software package \citep{Malz:qp}\footnote{available at: \url{http://github.com/aimalz/qp/}} for manipulating and calculating metrics of 1-dimensional PDFs.
-We will discuss the choice of $p(z)$ parameterization further in Section~\ref{sec:results}.
+
+In order to minimize the effects of the choice of parameterization, we asked those running all twelve codes to output \pzpdf s parameterized with $\approx 200$ piecewise constant bins spanning $0<z<2$.
+The piecewise constant format is chosen because of its established presence in the literature, and the choice of 200 bins was motivated by the approximate number of columns expected to be available for storage of \pzpdf s for the final \lsst\ Project tables.\footnote{See, e.~g.~the \lsst\ Data Products Definition Document, available at: \url{https://ls.st/dpdd}}
+All the \pzpdf\ catalogs are processed using the \qp\ software package \citep{Malz:qp}\footnote{available at: \url{http://github.com/aimalz/qp/}} for manipulating and calculating metrics of univariate PDFs.
+We will discuss the choice of \pzpdf\ parameterization further in Section~\ref{sec:results}.
 
@@
 
-\subsection{Metrics of an ensemble of photo-\mathinhead{$z$}{z} interim posteriors}
+\subsection{Metrics of an ensemble of photo-\mathinhead{$z$}{z} posteriors}
 \label{sec:qualmet}
+
+Because the \pzpdf s in the \lsst\ catalog will be used for many applications, some of which require accuracy of each individual catalog entry, we consider several metrics that get at the population-level performance of the \pzpdf s as distinct from a summary statistic thereof.
 
@@
 
-The probability integral transform (PIT) \citep{Polsterer:16} is defined for each individual galaxy as:%essentially the derivative of the QQ plot, a histogram of
-
+The probability integral transform (PIT) \citep{Polsterer:16} is defined for each individual galaxy as:
 \begin{align}
@@
 \end{align}
-The distribution of PIT values quantifies the behavior of the {\it ensemble} of photo-$z$ PDFs, enabling us to evaluate whether the $p(z)$ is, on average, accurate:  The PIT value is the Cumulative Distribution Function (CDF) of the $p(z)$ evaluated at the true redshift.
-A catalogue of photo-$z$ PDFs that are accurate should have a flat PIT histogram (i.e., the individual PIT values as samples from each CDF should match a Uniform(0,1) distribution if the CDFs are accurate).
-Specific deviations from flatness indicate inaccuracy: overly broad photo-$z$ PDFs would manifest as underrepresentation of the lowest and highest PIT values, whereas overly narrow photo-$z$ PDFs would manifest as over-representation of the lowest and highest PIT values.
-High frequency at only $\mathrm{PIT}\approx0$ and $\mathrm{PIT}\approx1$ indicates the presence of catastrophic outliers with highly inaccurate photo-$z$ PDFs where the true redshift is outside of the support of $p(z)$.
+The distribution of PIT values quantifies the behavior of the {\it ensemble} of \pzpdf s, enabling us to evaluate whether the population of \pdpdf s is, on average, accurate.
+The PIT value for each galaxy is the Cumulative Distribution Function (CDF) of its \pzpdf\ evaluated at its true redshift.
+A catalogue of \pzpdf s that are accurate should have a flat PIT histogram (i.e., the individual PIT values as samples from each CDF should match a Uniform(0,1) distribution if the CDFs are accurate).
+Specific deviations from flatness indicate inaccuracy: overly broad \pzpdf s would manifest as underrepresentation of the lowest and highest PIT values, whereas overly narrow \pzpdf s would manifest as over-representation of the lowest and highest PIT values.
+High frequency at only $\mathrm{PIT}\approx0$ and $\mathrm{PIT}\approx1$ indicates the presence of catastrophic outliers with highly inaccurate \pzpdf s where the true redshift of a galaxy is outside of the support of its \pzpdf.
 \citet{Tanaka:17} use the histogram of PIT values as a diagnostic indicator of overall code performance, while \citet{Freeman:17} independently define the PIT and demonstrate how its individual values may be used both to perform hypothesis testing (via, e.g., the KS, CvM, and AD tests; see below) and to construct quantile-quantile plots.
@@
 
-The quantile-quantile (QQ) plot is a graphical method for comparing two distributions, where the quantiles of one distribution are plotted against the quantiles of the other distribution (A quantile being defined by partitioning a distribution into consecutive intervals containing equal amounts of probability, or equal numbers of objects in each interval).
+A quantile is defined by partitioning a distribution into consecutive intervals containing equal amounts of probability, or equal numbers of objects in each interval in the case of a distribution of objects.
+The quantile-quantile (QQ) plot serves as a graphical visualization for comparing two distributions, where the quantiles of one distribution are plotted against the quantiles of the other distribution.
 %\blue{Include an equation to ensure quantile definition is crystal clear before building upon it for PIT.}\red{Is maybe the above text clear enough?
 Text may be more clear than any equation I can come up with--SJS.}
-In this paper we show the quantiles of the PIT values compared to the quantiles of the Uniform distribution that we expect the PIT values to match if $p(z)$ is an accurate probability distribution for all objects.
 The QQ plot provides an easy way to qualitatively assess the differences in various properties such as the moments of an estimating distribution relative to a true distribution.
 %location, scale and skewness.
-In this paper, QQ plots are used for two purposes: (1) for comparing $N(z)$ from photo-$z$ PDFs (estimated using Eq.~\ref{eq:stacked}) with the true $N(z)$, i.e.
-comparing the estimated distribution of redshifts with the true redshift distribution, and (2) for assessing the overall consistency of an ensemble of photo-$z$ PDFs with their true redshifts on a population level, where the distribution of the PIT values (see previous section) is compared to a uniform distribution between $0$ and $1$.
-The QQ plot contains very similar information to that shown in the PIT histogram plot, we include both forms, as visually they each convey the information in a somewhat distinct manner.
+
+In this paper, QQ plots are used for two purposes: (1) for comparing $N(z)$ from \pzpdf s (estimated using Eq.~\ref{eq:stacked}) with the true $N(z)$, and (2) for assessing the overall consistency of an ensemble of \pzpdf s with their true redshifts on a population level, where the distribution of the PIT values (see previous section) is compared to a uniform distribution between $0$ and $1$.
+Though the QQ plot contains very similar information to that shown in the PIT histogram plot, due to being the PIT being the derivative of the QQ, we include both forms for completeness and enhanced visual interpretability.
 %\red{is this enough reason to include both?--SJS}
@@
 With the conditional density estimation loss (CDE loss) we can compare how well different methods estimate individual PDFs for photometric covariates $\x$ rather than looking only at the ensemble distribution.
-As in Section~\ref{sec:flexzboost}, we use the notation $f(z \vert \x)$ instead of $p(z)$ to explicitly show the dependence on $\x$.
+As in Section~\ref{sec:flexzboost}, we use the notation $f(z \vert \x)$ instead of $p(z)$ to explicitly show the dependence on the photometry $\x$.
 
@@
 
-The CDE loss is defined as:
+The CDE loss is defined as
 \begin{equation} \label{eq:cde-loss}
-L(f, \widehat{f}) = \int \int (f(z \mid \x) - \widehat{f}(z \mid \x))^{2} dz dP(\x)
+L(f, \widehat{f}) \equiv \int \int (f(z \mid \x) - \widehat{f}(z \mid \x))^{2} dz dP(\x) .
 \end{equation}
-This loss is the CDE equivalent of the RMSE in regression. To estimate this loss we rewrite the loss as
+This loss is the CDE equivalent of the RMSE in regression.
+To estimate this loss we rewrite it as
 \begin{equation} \label{eq:estimated-cde-loss}
@@
 These metrics give a more quantitative measure of the departure from ideal than the more visual PIT histogram and QQ plot.
-We compute metrics comparing the CDF of PIT values to the CDF of a Uniform distribution, and also compute the CDF of the true redshift distribution $N'(z)$ compared the $\hat{N}(z)$ distribution derived from summing the $p(z)$ as described in Eq.~\ref{eq:stacked}.
+We compute metrics comparing the CDF of PIT values to the CDF of a Uniform distribution, and also compute the CDF of the true redshift distribution $N'(z)$ compared the $\hat{N}(z)$ distribution derived from summing the \pzpdf s as described in Eq.~\ref{eq:stacked}.
 
@@
 
-We employ the familiar root-mean-square error:
-
+We employ the familiar root-mean-square error
 \begin{equation}
-\mathrm{RMSE} = \sqrt{\int_{-\infty}^{\infty}\left(\hat{f}(z)-f'(z)\right)^{2}\,dz},
+\mathrm{RMSE} \equiv \sqrt{\int_{-\infty}^{\infty}\left(\hat{f}(z)-f'(z)\right)^{2}\,dz} .
 \end{equation}
@@
 The \textit{Kolmogorov-Smirnov statistic} $N_{\rm KS}$ is the maximum difference between $F_{\rm phot}(z)$ and $F_{\rm spec}(z)$, the CDFs of the photo-$z$ and spectroscopic redshift respectively:
-
 \begin{equation}
-N_{\rm KS} = \max_{z}\left( \left| F_{\rm phot}(z) - F_{\rm spec}(z) \right| \right).
+N_{\rm KS} \equiv \max_{z}\left( \left| F_{\rm phot}(z) - F_{\rm spec}(z) \right| \right).
 \end{equation}
@@
 We also consider two variants of the KS statistic: the Cramer-von Mises (CvM) and Anderson-Darling (AD) statistics.
-The CvM statistic is similar to the KS statistic as it is also computed from the distance between the measured CDF and the ideal CDF, but instead of the maximum distance, the CvM statistic calculates the average of the distance squared:
-
+The CvM statistic is similar to the KS statistic as it is also computed from the distance between the measured CDF and the ideal CDF, but instead of the maximum distance, the CvM statistic
 \begin{equation}
 \label{eq:cvm}
-\omega^2 = \int_{-\infty}^{+\infty}\big(F_{\mathrm{meas.}}(x) - F_{\mathrm{ideal}}(x)\big)^2\mathrm{d}F_{\mathrm{ideal}}
+\omega^2 \equiv \int_{-\infty}^{+\infty}\big(F_{\mathrm{meas.}}(x) - F_{\mathrm{ideal}}(x)\big)^2\mathrm{d}F_{\mathrm{ideal}}
 \end{equation}
+is the average of the distance squared.
 % In this paper we use the CvM statistic to evaluate the performance of both $n(z)$ and $p(z)$.
-The AD statistic is a weighted version of the CvM statistic, making it more sensitive to the tails of the distribution:
-
+The AD statistic
 \begin{equation} \label{eq:ad}
-A^2 = n\int_{-\infty}^{+\infty} \frac{\big(F_{\mathrm{meas.}}(x) - F_{\mathrm{ideal}}(x)\big)^2} {F_{\mathrm{ideal}}(x) (1-F_{\mathrm{ideal}}(x))}\mathrm{d}F_{\mathrm{ideal}}
+A^2 \equiv N_{tot}\int_{-\infty}^{+\infty} \frac{\big(F_{\mathrm{meas.}}(x) - F_{\mathrm{ideal}}(x)\big)^2} {F_{\mathrm{ideal}}(x) (1-F_{\mathrm{ideal}}(x))}\mathrm{d}F_{\mathrm{ideal}}
 \end{equation}
-where $n$ is the sample size.
+is a weighted version of the CvM statistic, making it more sensitive to the tails of the distribution, where $N_{tot}$ is the sample size.
 %The KS statistic and the CvM and AD statistics related to it are all used to evaluate the performance of both $\hat{n}(z)$ and the QQ and PIT of the CDFs of the ensemble of photo-$z$ interim posteriors.
@@
 %(Alex Malz)
-For the $\hat{N}(z)$ distributions we additionaly calculate the first three moments of the estimated redshift distribution for each code and compare them to the moments of the true redshift distribution $N'(z)$.
-The $m$th moment of a distribution is defined as
 
+We additionaly calculate the first three moments of the estimated redshift distribution $\hat{N}^{H}(z)$ for each code and compare them to the moments of the true redshift distribution $N'(z)$.
+The $m^{\mathrm{th}}$ moment of a distribution is defined as
 \begin{equation}
-\langle z^{m}\rangle = \int_{-\infty}^{\infty}z^{m}N(z)dz.
+\langle z^{m}\rangle \equiv \int_{-\infty}^{\infty}z^{m}N(z)dz .
 \end{equation}
 Here, we use the moments of the stacked estimator of the redshift distribution function as the basis for a metric.
-The closer the moments of $\hat{N}(z)$ for a photo-$z$ PDF method are to the moments of the true redshift distribution function $N'(z)$, the better the photo-$z$ PDF method.
+The closer the moments of $\hat{N}(z)$ for a \pzpdf\ method are to the moments of the true redshift distribution function $N'(z)$, the better the \pzpdf\ method.

### 2018-11-29 paper/main.tex
addressing comments, propagating macros

--- 2018-11-21 paper/main.tex
@@
 
-  Photometric redshift (photo-$z$) probability distribution functions (PDFs) are a key data product of nearly all upcoming galaxy imaging surveys.
+In order to maximize scientific returns of current and upcoming galaxy surveys, the photometric redshift (photo-$z$) posterior distributions produced by redshift estimation codes must be accurate probability distribution functions (PDFs).
+However, the posteriors resulting from a number of current techniques are not, in general, consistent with each other, affected by implicit assumptions made by each code, and an optimal method for obtaining an accurate PDF estimate remains unclear.
+We present the results of an initial study of the Large Synoptic Survey Telescope Dark Energy Science Collaboration (\textsc{LSST-DESC}) evaluating twelve photo-$z$ algorithms using complete and representative training data and evaluate multiple metrics to test how accurately the posteriors represent probability distributions.
+We observe several trends, including systematic biases and an overall over/under-prediction in the broadness of the PDFs in many of the codes which may be symptomatic of implementation problems or problems in underlying algorithm design.
+A careful accounting of all systematics discovered will be necessary for the codes employed in upcoming analyses in order to achieve unbiased cosmological measurements.
+
+%  Photometric redshift (photo-$z$) probability distribution functions (PDFs) are a key data product of nearly all upcoming galaxy imaging surveys.
 However, the photo-$z$ PDFs resulting from different techniques are not in general consistent with one another, and an optimal method for obtaining an accurate PDF remains unclear.
@@
 This initial test evaluates photo-$z$ algorithms in the presence of representative training data and in the absence of several common sources of systematic errors that affect the procedures by which photo-$z$ PDFs are derived.
-%  The photo-$z$ PDFs are evaluated using multiple metrics including the Kolmogorov-Smirnoff statistic, Cramer-von Mises statistic, Anderson-Darling statistic, Kullback-Leibler divergence, $N(z)$ moments, quantile-quantile plots and probability integral transform.
-  The photo-$z$ PDFs are evaluated using multiple metrics and we observe several trends, including an overall over/under-prediction in the broadness of the PDFs for several of the codes.
+%%  The photo-$z$ PDFs are evaluated using multiple metrics including the Kolmogorov-Smirnoff statistic, Cramer-von Mises statistic, Anderson-Darling statistic, Kullback-Leibler divergence, $N(z)$ moments, quantile-quantile plots and probability integral transform.
+%  The photo-$z$ PDFs are evaluated using multiple metrics and we observe several trends, including an overall over/under-prediction in the broadness of the PDFs for several of the codes.
 A careful accounting of all systematics discovered will be necessary for the codes employed in upcoming analyses in order to achieve unbiased cosmological measurements.
-  % Photometric redshift (photo-$z$) probability distribution functions (PDFs) more completely capture the complex nature of photo-$z$ uncertainties and are thus a key data product of nearly all upcoming galaxy imaging surveys, including the Large Synoptic Survey Telescope (LSST).
+ %% Photometric redshift (photo-$z$) probability distribution functions (PDFs) more completely capture the complex nature of photo-$z$ uncertainties and are thus a key data product of nearly all upcoming galaxy imaging surveys, including the Large Synoptic Survey Telescope (LSST).
   % Photo-$z$ PDFs produced by different algorithms do not in general agree with one another, and no one technique has yet been identified as optimal, leading to a proliferation of photo-$z$ PDF estimation codes.

### 2018-11-29 paper/tex/methods.tex
address comment on line 688

@@
 \end{equation}
-These coefficients can easily be estimated from data by regression.
+These coefficients can easily be estimated from data by regression since $\mathbb{E}\left[\phi_i(z)|\x\right]$ is the regression of $\phi_i(z)$ (a transformation of $Z$) on $X$.
 
@@
 \end{equation}  %\]
+
 where the subscript $o$ in $x_{o}$ indexes the \emph{observed} part of the input $x$, whereas the subscript $u$ indexes the \emph{unobserved} set (similarly for $\mu$ and $\Sigma$).

### 2018-11-29 paper/tex/analysis.tex

@@
 
-The most popular application of \pspdf s by far is the estimation of the overall redshift distribution $N(z)$, the true value of which is known for the \textsc{Buzzard} simulation and will be denoted as $N'(z)$.
+The most popular application of \pzpdf s by far is the estimation of the overall redshift distribution $N(z)$, the true value of which is known for the \textsc{Buzzard} simulation and will be denoted as $N'(z)$.
 Though alternatives exist \citep{Malz:chippr}, ``stacking'' according to
@@
 \end{align}
-The distribution of PIT values quantifies the behavior of the {\it ensemble} of \pzpdf s, enabling us to evaluate whether the population of \pdpdf s is, on average, accurate.
+The distribution of PIT values quantifies the behavior of the {\it ensemble} of \pzpdf s, enabling us to evaluate whether the population of \pzpdf s is, on average, accurate.
 The PIT value for each galaxy is the Cumulative Distribution Function (CDF) of its \pzpdf\ evaluated at its true redshift.

### 2018-11-29 paper/tex/analysis.tex

@@
 \end{equation}
+
 This loss is the CDE equivalent of the RMSE in regression.
@@
 \begin{equation} \label{eq:estimated-cde-loss}
-\mathbb{E}_\X \left[\int \widehat{f}(z \mid \X)^{2} dz\right] - 2 \mathbb{E}_{\X,Z}\left[\widehat{f}(Z \mid \X)\right] + K_{f},
+L(f, \widehat{f}) = \mathbb{E}_\X \left[\int \widehat{f}(z \mid \X)^{2} dz\right] - 2 \mathbb{E}_{\X,Z}\left[\widehat{f}(Z \mid \X)\right] + K_{f},
 \end{equation}
-where the first expectation is with respect to the marginal distribution of the covariates $\X$, the second expectation is with respect to the joint distribution of $\X$ and $Z$, and $K_{f}$ is a constant depending only upon the true conditional densities $f(z\mid \x)$.
+where upper-case letters denote random variables and lower-case the observed variables.
+The first expectation is with respect to the marginal distribution of the covariates $\X$, the second expectation is with respect to the joint distribution of $\X$ and $Z$, and $K_{f}$ is a constant depending only upon the true conditional densities $f(z\mid \x)$.
 For each method we can estimate these expectations as empirical expectations on the test or validation data \citep[Eq.~7 in][]{Izbicki:17b} without knowledge of the true densities.
@@
 % In this paper we use the CvM statistic to evaluate the performance of both $n(z)$ and $p(z)$.
+
 The AD statistic

### 2018-11-29 paper/tex/results.tex

--- 2018-11-14 paper/tex/results.tex
@@
 
-Samples from accurate photo-$z$ posteriors should reproduce the space of $p(z, data)$.
+Samples from accurate photo-$z$ posteriors should reproduce the space of $p(z,{\rm data})$.
 However, it is difficult to test this reconstruction given our data set, as the galaxy distributions arise from mock objects pasted on to an underlying dark matter halo catalogue with properties designed to match empirical relations, rather than being drawn from statistical distributions in redshift.
-In previous sections we have mentioned that optimizing for a specific metric does not guarantee good performance on other metrics, nor is there any guarantee that good performance by our metrics corresponds to \textit{accurate} photo-$z$ posteriors.
+In previous sections we have mentioned that optimizing for a specific metric does not guarantee good performance on other metrics, nor is there any guarantee that good performance by our metrics corresponds to \textit{accurate} photo-$z$ posteriors, in the sense of predicting redshifts for individual galaxies.
 %\aim{Explain that samples from accurate photo-$z$ posteriors should reconstruct the space of $p(z, data)$, a test we can't check with our datasets.}
@@
 The CDE loss and point estimate metrics, however, successfully identify problems with \trainz.
-As shown in Appendix~\ref{sec:pointmetrics}, \trainz has identical $ZPEAK$ and $ZWEIGHT$ values for every galaxy, and thus the photo-$z$s are constant as a function of spec-$z$s, i.e.
+As shown in Appendix~\ref{sec:pointmetrics}, \trainz ~has identical $ZPEAK$ and $ZWEIGHT$ values for every galaxy, and thus the photo-$z$s are constant as a function of spec-$z$s, i.e.
 a horizontal line at the mode and mean of the training set distribution respectively.
-The explicit dependence on the \it{individual} posteriors in the calculation of the CDE loss, described in Section~\ref{sec:CDE_loss}, distinguishes this metric from the other $p(z)$ metrics that test the overall ensemble of $p(z)$ distributions.
-With a representative training set, \trainz\ will score well on the ensemble metrics, but fails miserably for metrics tied to individual redshifts.
+The explicit dependence on the {\it individual} posteriors in the calculation of the CDE loss, described in Section~\ref{sec:CDE_loss}, distinguishes this metric from the other $p(z)$ metrics that test the overall ensemble of $p(z)$ distributions.
+With a representative training set, \trainz\ will score well on the ensemble metrics, but fails miserably for metrics tied to individual redshifts presented in this paper.
 We note that many of the ensemble-based metrics are prominent in the photo-$z$ literature despite their inability to identify problems such as those exemplified by \trainz.
@@
 The best photo-$z$ method is the one that most effectively achieves our science goals, not the one that performs best on a metric that does not accurately reflect those goals.
-In the absence of clear goals or the information necessary for a principled metric definition, we must think carefully before choosing a single metric
+In the absence of clear goals or the information necessary for a principled metric definition, we must think carefully before choosing a single metric.
 

### 2018-11-29 paper/tex/discussion.tex

@@
-\section{Discussion}
+\section{Discussion and Future Work}
 \label{sec:discussion}
@@
 The SEDs for the galaxy sample in this paper were constructed form linear combinations of five basis SED templates.
-Future simulations will also include more complex SED information, with a more realistic range of physical properties, and the inclusion of AGN effects.
+Future simulations will also include more complex SED information, with a more realistic range of physical properties, and the inclusion of AGN effects, a more insidious problem, where AGN features may not be apparent, but the colors and other host galaxy properties are perturbed relative to galaxies with an inactive nucleus.
+In such cases, the presence of the AGN may induce a bias if the template SEDs or empirical datasets do not include low-level AGN counterparts.
+
 As discussed extensively in \citet{Newman:2015} a representative set of spectroscopically confirmed galaxies spanning the full range of both redshift and apparent magnitude is necessary as a training set to characterize the mapping from broad-band fluxes to photometric redshifts.

### 2018-11-29 paper/main.tex

@@
 \newcommand{\pzpdf}{\pz\ PDF}% can easily change this to specify posterior throughout
+\newcommand{\qp}{\texttt{qp}}
 
@@
 
-The authors would like to thank their LSST-DESC publication review committee.
+The authors would like to thank their LSST-DESC publication review committee for comments that improved the paper draft.
+
+{ \red{personal funding sources}}
 
 AIM is advised by David W. Hogg and was supported by National Science Foundation grant AST-1517237.
+
+In addition to packages cited in the text, analyses performed in this paper used the following software packages: \textsc{Numpy} and \textsc{Scipy} \citep{numpyscipy}, \textsc{Matplotlib} \citep{matplotlib}, \textsc{Seaborn} \citep{seaborn}, \textsc{minFunc} \citep{minfunc}, \textsc{pySkyNet} \citep{pyskynet}, and {\textsc photUtils} from the LSST simulations package \citep{lsstphotutils}.
 

### 2018-12-01 paper/tex/methods.tex
fix huge spacing for two equations

@@
 To obtain a robust estimate of the $p(z)$ for galaxies with a small number of colour-matched training set galaxies, when this number is less than $20$ the nearest $20$ neighbours in colour-space are used instead, and we convolve the $p(z)$ with a Gaussian with a standard deviation of:
-
 \begin{equation}
@@
 \end{equation}
-
 \noindent to appropriately broaden it so that the $p(z)$ for these test galaxies represents the enlarged PPF value associated with it.
@@
 The mean function, $\mu(x)$, and the variance function $\sigma(x)^{2}$ are both linear combinations of basis functions that take the following form:
-
-\begin{align}
+\begin{equation}
 f(x)=\sum_{i=1}^{m}\phi_{i}(x)w_{i},
-\end{align}
+\end{equation}
 where $\left\{\phi_{i}(x)\right\}_{i=1}^{m}$ and $\left\{w_{i}\right\}_{i=1}^{m}$ are sets of $m$ basis functions and their associated weights respectively.

### 2018-12-02 paper/tex/analysis.tex
fix mistaken merge deletion

@@
 
-The probability integral transform (PIT) \citep{Polsterer:16} is defined for each individual galaxy as:
+The probability integral transform (PIT) has been employed recently to evaluate fidelity of posterior distributions \citep[e.~g.~][]{Bordoloi:10,Polsterer:16,Tanaka:17}.
+The PIT value is defined for each individual galaxy as:
 \begin{align}

### 2018-12-03 paper/main.tex
edits based on read-through of paper

@@
 However, the posteriors resulting from a number of current techniques are not, in general, consistent with each other, affected by implicit assumptions made by each code, and an optimal method for obtaining an accurate PDF estimate remains unclear.
-We present the results of an initial study of the Large Synoptic Survey Telescope Dark Energy Science Collaboration (\textsc{LSST-DESC}) evaluating twelve photo-$z$ algorithms using complete and representative training data and evaluate multiple metrics to test how accurately the posteriors represent probability distributions.
+We present the results of an initial study of the Large Synoptic Survey Telescope Dark Energy Science Collaboration (\textsc{LSST-DESC}) testing twelve photo-$z$ algorithms using complete and representative training data and evaluate multiple metrics to test how accurately the posteriors represent probability distributions.
 We observe several trends, including systematic biases and an overall over/under-prediction in the broadness of the PDFs in many of the codes which may be symptomatic of implementation problems or problems in underlying algorithm design.

### 2018-12-03 paper/tex/discussion.tex
edits based on read-through of paper

@@
 All codes were provided a set of representative training data and tested on an idealized set of model galaxies with high signal-to-noise and photometry with no confounding effects due to blending, instrumental effects, the night sky, or other complications included.
-The goal was not to determine a ``best'' photo-$z$ code: in many ways, this was a baseline test of a ``best case scenario'' to predict the expected photo-$z$ performance if a stage IV dark energy survey was to obtain complete training samples and perfectly calibrated their multi-band photometry.
+%The goal was not to determine a ``best'' photo-$z$ code:i
+In many ways, this was a baseline test of a ``best case scenario'' to predict the expected photo-$z$ performance if a stage IV dark energy survey was to obtain complete training samples and perfectly calibrated their multi-band photometry.
 Given these idealized conditions, any deficiencies observed in a photo-$z$ code's performance should be a cause for concern, and may be evidence of a problem with either/both of the specific code implementation or the underlying algorithm.
-In order to meet the stringent LSST goals for photo-$z$ performance, identifying and correcting such problems is an important first step before tackling more realistic data in future challenges, and codes that do not perform well may not be worth pursuing in future challenges.
-Many of the codes tested performed well; however, several did not meet the stringent goals that have been laid out for LSST photometric redshift performance for individual galaxies, as laid out in the LSST SRD (See Section~\ref{sec:intro}.
-This is a cause for concern, given the idealized conditions, and the individual code responses will be studied in detail moving forward.
+Many of the codes tested performed well; however, several did not meet the stringent goals that have been laid out for LSST photometric redshift performance for individual galaxies, as laid out in the LSST SRD (See Section~\ref{sec:intro}).%  This is a cause for concern, and the individual code responses will be studied in detail moving forward.
 If methods can not reach the goals on idealized data, then they will almost surely not meet those same goals when the more complex problems that we expect to arise from real LSST data are included.
 The results presented in this paper enable an evaluation of which algorithms are the most promising moving forward, and potentially point to implementation choices or mistakes which could be improved or corrected in others.
+Codes that do not perform well may not be worth pursuing in future challenges.
 
@@
 We must be aware of that the optimal method for one is not necessarily optimal for the other, and in fact several photo-$z$ algorithms may be necessary in the final cosmological analysis in order to satisfy the requirements of all science use cases.
+
 The example of the simple \trainz\ estimator described in Section~\ref{sec:caution} shows a simple model with a $p(z)$ that is unrealistic for individual objects can still score very well on many of our metrics.

### 2018-12-03 paper/tex/introduction.tex
edits based on read-through of paper

@@
 
-The unprecedented sample size of LSST galaxies, expected to number several billion for the main cosmological sample, necessitates stringent constraints on photo-$z$ accuracy if systematic errors are not to dominate the statistical errors.
+The unprecedented sample size of LSST, expected to number several billion galaxies for the main cosmological sample, necessitates stringent constraints on photo-$z$ accuracy if systematic errors are not to dominate the statistical errors.
 The LSST Science Requirements Document (SRD)\footnote{available at \url{https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17}} lists the individual galaxy photometric redshift goals for a magnitude limited sample with $i<25$ as: root-mean-square error with a goal of $\sigma_z<0.02(1+z)$; $3\sigma$ ``catastrophic outlier'' rate below $10\%$; bias below $0.003$\footnote{
@@
 The LSST Dark Energy Science Collaboration (LSST-DESC) developed a separate Science Requirements Document \citep{Mandelbaum:2018}, which forecasts the constraining power of five cosmological probes using somewhat conservative assumptions to define requirements on systematic errors for several measurements.
-These include even more stringent requirements on photometric redshift performance than those included in the LSST SRD, though most of the initial LSST-DESC requirements are defined in terms of tomographic bin populations rather than on individual object redshifts.
+These include even more stringent requirements on photometric redshift performance than those included in the LSST SRD, though most of the current LSST-DESC requirements are defined in terms of tomographic bin populations rather than on individual object redshifts.
 The tremendous size of LSST's galaxy catalogue will be enabled by its exceptional depth, pushing to fainter magnitudes and deeper imaging and including galaxies of lower luminosity and higher redshift than ever before.
@@
 DES compared several codes for point estimates and a subset with $p(z)$ information \citep{Sanchez:14}.
-A follup up paper examined summary statistics of photo-$z$ interim posteriors for tomographically binned galaxy subsamples \citep{Bonnett:16}.
+A follow up up paper examined summary statistics of photo-$z$ interim posteriors for tomographically binned galaxy subsamples \citep{Bonnett:16}.
 
-This paper is distinguished by its focus on metrics of photo-$z$ interim posteriors themselves and consideration of both classic and state-of-the-art photo-$z$ algorithms, comparing the performance of several of the most widely employed codes as well as some that have been developed only recently on the basis of metrics appropriate for a probabilistic data product.
+This paper is distinguished by its focus on metrics of photo-$z$ interim posteriors themselves and consideration of both classic and state-of-the-art photo-$z$ algorithms, comparing the performance of several of the most widely employed codes, as well as some that have been developed only recently, on the basis of metrics appropriate for a probabilistic data product.
 The results presented here are

### 2018-12-13 paper/main.tex
add funding source

@@
 However, the posteriors resulting from a number of current techniques are not, in general, consistent with each other, affected by implicit assumptions made by each code, and an optimal method for obtaining an accurate PDF estimate remains unclear.
-We present the results of an initial study of the Large Synoptic Survey Telescope Dark Energy Science Collaboration (\textsc{LSST-DESC}) testing twelve photo-$z$ algorithms using complete and representative training data and evaluate multiple metrics to test how accurately the posteriors represent probability distributions.
+We present the results of an initial study by the Large Synoptic Survey Telescope Dark Energy Science Collaboration (\textsc{LSST-DESC}) testing twelve photo-$z$ algorithms using complete and representative training data and evaluate multiple metrics to test how accurately the posteriors represent probability distributions.
 We observe several trends, including systematic biases and an overall over/under-prediction in the broadness of the PDFs in many of the codes which may be symptomatic of implementation problems or problems in underlying algorithm design.
-A careful accounting of all systematics discovered will be necessary for the codes employed in upcoming analyses in order to achieve unbiased cosmological measurements.
+A careful accounting of all photo-$z$ systematics will be necessary for the codes employed in upcoming analyses in order to achieve unbiased cosmological measurements.
 
@@
 { \red{personal funding sources}}
-
-AIM is advised by David W. Hogg and was supported by National Science Foundation grant AST-1517237.
+S.~Schmidt acknowledges support from DOE grant DE-SC0009999 and NSF/AURA grant N56981C.  AIM is advised by David W. Hogg and was supported by National Science Foundation grant AST-1517237.
 

### 2019-02-28 paper/tex/analysis.tex

@@
 %\red{Note: Let's address how the gridding (endpoints and spacing) affects the metrics, from the email correspondence Re: [LSST-DESC-MEMBERS] Summary of Nov 29 PZ Telecon. (AIM)}
-
-In order to minimize the effects of the choice of parameterization, we asked those running all twelve codes to output \pzpdf s parameterized with $\approx 200$ piecewise constant bins spanning $0<z<2$.
+In order to minimize the effects of the choice of parameterization, we asked those running all twelve codes to output \pzpdf s parameterized with $200$ piecewise constant bins spanning $0<z<2$.
 The piecewise constant format is chosen because of its established presence in the literature, and the choice of 200 bins was motivated by the approximate number of columns expected to be available for storage of \pzpdf s for the final \lsst\ Project tables.\footnote{See, e.~g.~the \lsst\ Data Products Definition Document, available at: \url{https://ls.st/dpdd}}
@@
 
-\subsubsection{Root-mean-square error (RMSE)}
-\label{sec:rmse}
-%(Alex Malz)
-
-We employ the familiar root-mean-square error
-\begin{equation}
-\mathrm{RMSE} \equiv \sqrt{\int_{-\infty}^{\infty}\left(\hat{f}(z)-f'(z)\right)^{2}\,dz} .
-\end{equation}
-%may be used for two purposes in this analysis.
-%We calculate it between the true redshift distribution $N'(z)$ and each estimator $\hat{N}(z)$ (derived by Eq.~\ref{eq:stacked}) as well as between the ideal QQ and PIT curves from each method studied and the ideal curves.
-Though this metric does not account for the fact that the redshift distribution function is, in fact, a probability distribution, it can still be interpreted as a measure of the integrated difference between the estimated distribution and the true distribution.%, and it can be used to quantify the otherwise qualitative metrics.
+%\subsubsection{Root-mean-square error (RMSE)}
+%\label{sec:rmse}
+%%(Alex Malz)%
+%
+%We employ the familiar root-mean-square error
+%\begin{equation}
+%\mathrm{RMSE} \equiv \sqrt{\int_{-\infty}^{\infty}\left(\hat{f}(z)-f'(z)\right)^{2}\,dz} .
+%\end{equation}
+%%may be used for two purposes in this analysis.
+%%We calculate it between the true redshift distribution $N'(z)$ and each estimator $\hat{N}(z)$ (derived by Eq.~\ref{eq:stacked}) as well as between the ideal QQ and PIT curves from each method studied and the ideal curves.
+%Though this metric does not account for the fact that the redshift distribution function is, in fact, a probability distribution, it can still be interpreted as a measure of the integrated difference between the estimated distribution and the true distribution.%, and it can be used to quantify the otherwise qualitative metrics.
+%
 
@@
 Here, we use the moments of the stacked estimator of the redshift distribution function as the basis for a metric.
-The closer the moments of $\hat{N}(z)$ for a \pzpdf\ method are to the moments of the true redshift distribution function $N'(z)$, the better the \pzpdf\ method.
+The closer the moments of $\hat{N}(z)$ for a \pzpdf\ method are to the moments of the true redshift distribution function $N'(z)$, the better the \pzpdf\ method is a estimating the overall redshift distribution ``shape''.

### 2019-02-28 paper/tex/data.tex

@@
 \label{sec:buzztraining}
-The total catalogue covered $400$ square degrees and contained $238$ million galaxies to an apparent magnitude limit of $r\!=\!29$ and spanning the redshift range $0\!<\!z\!\leq\!8.7$.
+The total initial catalogue covered $400$ square degrees and contained $238$ million galaxies to an apparent magnitude limit of $r\!=\!29$ and spanning the redshift range $0\!<\!z\!\leq\!8.7$.
 In order for statistical errors not to dominate, we need less than one million galaxies in our sample.
@@
 %The resulting catalogues contain $111\,171$ training galaxies and $1\,000\,883$ test galaxies.
-Finally, we restrict our analysis to a sample with an apparent magnitude limite $i<25.3$, which give a signal-to-noise $\sim\,30$ for most galaxies, a cut often referred to as the expected ``LSST Gold Sample''.
+Finally, we restrict our analysis to a sample with an apparent i-band magnitude limit $i<25.3$, which give a signal-to-noise $\sim\,30$ for most galaxies, a cut often referred to as the expected ``LSST Gold Sample''.
 This magnitude cut results in a training set with $44\,404$ galaxies and a test set containing $399\,356$ galaxies.
@@
 In order to blind results, initially redshifts were not revealed for the ``test'' set, and were only supplied for the training sample galaxies.
-This prevented code runners from tweaking results and over-fitting to the specific test set.
+This prevented code runners from tweaking results and fitting to the specific test set.
 
@@
 For our initial investigation of photometric redshift codes, we begin with a data set that is somewhat idealized, and does not contain all of the complicating factors present in real data.
-In several cases, the simplification is done with a purpose, with potentially confounding effects excluded in order to better isolate the differences between current-generation photo-$z$ codes, and their causes.
-We list several of the simulations limitations in this section.
+In several cases, the simplification is done with a purpose, with potentially confounding effects excluded in order to better isolate the differences between current-generation photo-$z$ codes and their causes.
+We list several of the simulation limitations in this section.
 As the simulation is catalogue-based, no image level effects, such as photometric measurement effects, object blending, contamination from sky background (Zodiacal light, scattered light, etc...), lensing magnification, or Galactic reddening are included.
@@
 While these idealized conditions limit the realism of our galaxy population, some are also by design.
-We aim to test the photo-z codes at a very basic level, and a simplified model assures that differences in results seen between the codes are due to fundamental differences in their underlying assumptions and implementation details, rather than more nuanced implementation details.
+We aim to test the photo-z codes at a very basic level, and a simplified model assures that differences in results seen between the codes are due to fundamental differences in their underlying assumptions and implementation details, rather than more nuanced properties.
 

### 2019-02-28 paper/tex/methods.tex

@@
 Here we outline the photo-$z$ PDF codes tested in this study.
-In total, eleven distinct codes are tested.
+In total, twelve distinct codes are tested.
 This sample is not comprehensive, codes were chosen based on the expertise available within the group; however, those chosen do cover a broad range of the current-generation methods used in the field.
@@
 where c colour, $\delta c_{\rm test}$ is the measurement error on the colour, and $N_{\rm colours}$ is the total number of colors (i.e., in our case $u$-$g$, $g$-$r$, $r$-$i$, $i$-$z$, and $z$-$y$).
-Then, we choose a threshold value for $D_M$ that defines the colour match set based on a set value of the percent point function (PPF): for example, for $N_{\rm dof}=5$, choosing a vPPF$=95$ per cent of all training galaxies consistent with the test galaxy will have $D_M < 11.07$ (where $N_{\rm dof}$ is the number of degrees of freedom, in this case the number of colours).
+Then, we choose a threshold value for $D_M$ that defines the colour match set based on a set value of the percent point function (PPF): for example, for $N_{\rm dof}=5$, choosing a vPPF$=0.95$, $95$ per cent of all training galaxies consistent with the test galaxy will have $D_M < 11.07$ (where $N_{\rm dof}$ is the number of degrees of freedom, in this case the number of colours).
 If a galaxy had a non-detection in a band, that band was dropped and $N_{\rm dof}$ was reduced by one in the colour-matching space.
@@
 Third, we set different parameters for the identification of the colour-matched subset of training galaxies and the selection of a photometric redshift estimate.
-In G18 we used a percent point function (PPF) value of 0.68 to identify the colour-matched subset of training galaxies and used the redshift of nearest neighbour in colour-space as the photo-$z$ estimate.
+In G18 we used a PPF value of 0.68 to identify the colour-matched subset of training galaxies and used the redshift of nearest neighbour in colour-space as the photo-$z$ estimate.
 These choices work well when the desire is to obtain accurate photo-$z$ estimates for most test-set galaxies, but does not return an accurate $p(z)$ in all cases -- especially for galaxies that are bright and/or have few matches in colour-space.
 Since an accurate estimate of $p(z)$ is desired for this work we make several changes to our implementation of the \textsc{CMNN} photo-$z$ estimator.
-We continue to use a percent point function of ${\rm PPF} = 0.95$ to generate the subset of colour-matched training galaxies, but weight them by the inverse of their Mahalanobis distance.
+We continue to use a ${\rm PPF} = 0.95$ to generate the subset of colour-matched training galaxies, but weight them by the inverse of their Mahalanobis distance.
 This weighting maintains some of the accuracy that was previously achieved by simply using the nearest neighbour in colour-space.
@@
 \textsc{GPz}\footnote{\url{https://github.com/OxfordML/GPz}} \citep{Almosallam:16a,Almosallam:15b} is a sparse Gaussian process based code, a scalable approximation of full Gaussian Processes \citep{Rasmussen:06}, with the added feature of being able to produce input-dependent variance estimations (heteroscedastic noise).
-The model assumes that the probability of the output $y$, the redshift, given the input $x$, the photometry, is $p(y|x)=\mathcal{N}\left(y|\mu(x),\sigma(x)^{2}\right)$.
+The model assumes that the probability of the output $y$, the redshift, given the input $x$, the photometry, is $p(z|x)=\mathcal{N}\left(z|\mu(x),\sigma(x)^{2}\right)$.
 The mean function, $\mu(x)$, and the variance function $\sigma(x)^{2}$ are both linear combinations of basis functions that take the following form:
@@
 The pre-processing includes a model for perturbation of the photometry that is employed in calculating the PDF of the photo-$z$ estimation errors.
-The photometric perturbation is defined as: $m_{ij} = m_{ij} + \alpha_{i}F_{ij}*u_{\mu=0,\sigma=1}$, where $\alpha_{i}$ is a user selected multiplicative constant (useful in case of multi-survey photometry), $u_{\mu=0,\sigma=1}$ is a random value from the standard normal distribution and $F_{ij}$ is a bimodal function (a constant function + polynomial fitting of the mean magnitude errors on the binned bands), heuristically tuned in such a way that the constant component is the threshold under which the polynomial function is considered too low to provide a significant noise contribution to the photometry perturbation.
+The photometric perturbation is defined as:
+\begin{equation}
+ m_{ij} = m_{ij} + \alpha_{i}F_{ij}u_{\mu=0,\sigma=1}
+\end{equation}
+\noindent  where $\alpha_{i}$ is a user selected multiplicative constant (useful in case of multi-survey photometry), $u_{\mu=0,\sigma=1}$ is a random value from the standard normal distribution and $F_{ij}$ is a bimodal function (a constant function + polynomial fitting of the mean magnitude errors on the binned bands), heuristically tuned in such a way that the constant component is the threshold under which the polynomial function is considered too low to provide a significant noise contribution to the photometry perturbation.
 

### 2019-02-28 paper/tex/results.tex

@@
 Several features are obvious even in these illustrative examples.
-\textsc{ANNz2, METAPhoR, NN,} and \textsc{SkyNet} all show an excess of small-scale features, which appear to be print-through of the underlying training set galaxies.
+\textsc{ANNz2, METAPhoR, CMNN,} and \textsc{SkyNet} all show an excess of small-scale features, which appear to be print-through of the underlying training set galaxies.
 For example, in \textsc{CMNN} the $p(z)$ are a simply a weighted histogram of all spectroscopic training galaxies in nearby colour space with no smoothing applied, so the substructure is due to the finite number of neighbours, and is not unexpected.
@@
 
-As stated in Section~\ref{sec:metrics}, $p(z)$ is parameterized as $200$ piecewise constant bins covering $0<z<2$ for all twelve codes, giving a grid size of $\delta z = 0.01$ for each code.
-A piecewise constant grid was a natural choice for some photo-$z$ codes, for instance most template-based codes compute likelihoods on a fixed grid.
-In contrast, FlexZBoost, for example, can return estimates on any grid without compression errors as it’s a basis expansion method where only the expansion coefficients need to be stored.
-Codes with a native output format other than the shared piecewise constant binning scheme (or one that can be losslessly converted to it) may suffer from loss of information when converting to it, which could artificially favor some codes over others in a limited number of cases, for example bright galaxies with very narrow $p(z)$ where the true peak falls between grid points.
+As stated in Section~\ref{sec:metrics}, $p(z)$ is parameterized as $200$ piecewise constant bins covering $0<z<2$ for all twelve codes, giving a grid size of $\delta z = 0.01$ for each PDF.
+A piecewise constant grid was a natural choice for some photo-$z$ methods, for instance most template-based codes compute likelihoods on a fixed grid.
+%In contrast, FlexZBoost, for example, can return estimates on any grid without compression errors as it’s a basis expansion method where only the expansion coefficients need to be stored.
+%Codes with a native output format other than the shared piecewise constant binning scheme (or one that can be losslessly converted to it) may suffer from loss of information when converting to it, which could artificially favor some codes over others in a limited number of cases, for example bright galaxies with very narrow $p(z)$ where the true peak falls between grid points.
 We will discuss PDF storage in Section~\ref{sec:discussion}.
-
-Furthermore, the fidelity of photo-$z$ interim posteriors in this format varies with the quality of the photometry.
+%Furthermore, the fidelity of photo-$z$ interim posteriors in this format varies with the quality of the photometry.
 For faint galaxies, this redshift resolution is sufficient to capture the shape of $p(z)$ for the majority of the test sample, where photometric errors on the faint galaxies lead to somewhat broad peaks in the redshift posterior.
-However, as can be seen in e.~g.~the top left panel of Fig.~\ref{fig:pz_examples}, for bright galaxies with narrow $p(z)$ the grid spacing of $\delta z = 0.01$ is not sufficient to resolve the peak.
+However, as can be seen in e.~g.~the top left panel of Fig.~\ref{fig:pz_examples}, for bright galaxies with narrow $p(z)$ the grid spacing of $\delta z = 0.01$ is not always sufficient to resolve the peak.
 This is consistent with the results described in \citet[]{Malz:qp}, who find that quantiles (and, to a lesser degree, samples) often outperform gridded $p(z)$, particularly for bright objects and in the presence of harsher storage constraints.
@@
 Note from Table~\ref{tab:cdeloss} that both \textsc{FlexZBoost} and \textsc{CMNN} have low CDE losses.
-Empirically, we have found that PIT RMSE is not as closely correlated to CDE loss as it is to the $N(z)$ statistics.
+%Empirically, we have found that PIT RMSE is not as closely correlated to CDE loss as it is to the $N(z)$ statistics.
 As CDE loss is a better measure of individual redshift performance, rather than ensemble distribution performance, this statistic is a better indicator of which codes will be most likely to perform well for science cases where single objects are employed.
 
-Table~\ref{tab:rmse} gives the root-mean-square-error (RMSE) statistics for both the PIT and N(z) estimators.
+%Table~\ref{tab:rmse} gives the root-mean-square-error (RMSE) statistics for both the PIT and N(z) estimators.
 The PIT value calculates the RMSE between the quantiles shown in the QQ plot in Figure~\ref{fig:pitqq} and the diagonal, while the N(z) calculates the RMSE between the cumulative distribution of the stacked $\hat{N}(z)$ and the true redshift distribution $N'(z)$.
@@
 As previously stated, there are a variety of scientific use cases for photo-$z$'s in large upcoming surveys, and care must be taken in how the metrics used to optimize catalog photometric redshifts are defined as well as in how they are used.
-This implies that we may need multiple photo-z estimators, tuned to the particular metric, in order to maximize science returns in upcoming surveys.
+This implies that we may need multiple photo-z estimators, tuned to the particular metric, in order to maximize returns over all science cases in upcoming surveys.
 In addition, very few scientific use cases will employ the overall $\hat{N}(z)$ with no cuts, as we explore in this paper.
@@
 %[more caution on over-reliance on metrics, and making sure that metrics test all of what you want] \red{this definitely still needs work}
-In summary, context is crucial to interpreting metrics and defending against the likes of \trainz.
+In summary, context is crucial to interpreting metrics and defending against deceptively well performing methods such as \trainz.
 The best photo-$z$ method is the one that most effectively achieves our science goals, not the one that performs best on a metric that does not accurately reflect those goals.
@@
 
-\begin{table}
-\setlength{\tabcolsep}{2pt}
-\begin{center}
-\caption{Root-Mean-Square-Error (RMSE) statistics for the twelve photo-z codes for both PIT and $\hat{N}(z)$ distributions.}\label{tab:rmse}
-\begin{tabular}{lcc}
-\hline
-\hline
-Root-Mean-Square-Error \\
-(RMSE) statistics \\
-\hline
-Photo-z Code & PIT RMSE & N(z) RMSE\\
-\hline
-\textsc{ANNz2}      & 0.019 & 0.0054\\
-\textsc{BPZ}        & 0.032 & 0.0050\\
-\textsc{Delight}    & 0.111 & 0.0056\\
-\textsc{EAZY}       & 0.054 & 0.0102\\
-\textsc{FlexZBoost} & 0.021 & 0.0022\\
-\textsc{GPz}        & 0.027 & 0.0042\\
-\textsc{LePhare}    & 0.028 & 0.0062\\
-\textsc{METAPhoR}   & 0.064 & 0.0081\\
-\textsc{CMNN}         & 0.108 & 0.0009\\
-\textsc{Skynet}     & 0.054 & 0.0144\\
-\textsc{TPZ}        & 0.082 & 0.0031\\
-\hline
-\trainz		& 0.0025 & 0.0013\\
-\end{tabular}
-\end{center}
-\end{table}
+%\begin{table}
+%\setlength{\tabcolsep}{2pt}
+%\begin{center}
+%\caption{Root-Mean-Square-Error (RMSE) statistics for the twelve photo-z codes for both PIT and $\hat{N}(z)$ distributions.}\label{tab:rmse}
+%\begin{tabular}{lcc}
+%\hline
+%\hline
+%Root-Mean-Square-Error \\
+%(RMSE) statistics \\
+%\hline
+%Photo-z Code & PIT RMSE & N(z) RMSE\\
+%\hline
+%\textsc{ANNz2}      & 0.019 & 0.0054\\
+%\textsc{BPZ}        & 0.032 & 0.0050\\
+%\textsc{Delight}    & 0.111 & 0.0056\\
+%\textsc{EAZY}       & 0.054 & 0.0102\\
+%\textsc{FlexZBoost} & 0.021 & 0.0022\\
+%\textsc{GPz}        & 0.027 & 0.0042\\
+%\textsc{LePhare}    & 0.028 & 0.0062\\
+%\textsc{METAPhoR}   & 0.064 & 0.0081\\
+%\textsc{CMNN}         & 0.108 & 0.0009\\
+%\textsc{Skynet}     & 0.054 & 0.0144\\
+%\textsc{TPZ}        & 0.082 & 0.0031\\
+%\hline
+%\trainz		& 0.0025 & 0.0013\\
+%\end{tabular}
+%\end{center}
+%\end{table}
 

### 2019-03-08 paper/main.tex
update paper title to shorter version
abstract and title

@@
 
-\title{Implicit assumptions and their impact on photometric redshift PDF performance in the context of LSST}
-% \title{Photometric redshift uncertainties for LSST \romannumeral{1}: A unified methodology for assessing photo-$z$ PDF performance}
+\title{Evaluating photometric redshift PDF approaches for \textsc{LSST}}
 
@@
 
-In order to maximize scientific returns of current and upcoming galaxy surveys, the photometric redshift (photo-$z$) posterior distributions produced by redshift estimation codes must be accurate probability distribution functions (PDFs).
-However, the posteriors resulting from a number of current techniques are not, in general, consistent with each other, affected by implicit assumptions made by each code, and an optimal method for obtaining an accurate PDF estimate remains unclear.
-We present the results of an initial study by the Large Synoptic Survey Telescope Dark Energy Science Collaboration (\textsc{LSST-DESC}) testing twelve photo-$z$ algorithms using complete and representative training data and evaluate multiple metrics to test how accurately the posteriors represent probability distributions.
-We observe several trends, including systematic biases and an overall over/under-prediction in the broadness of the PDFs in many of the codes which may be symptomatic of implementation problems or problems in underlying algorithm design.
-A careful accounting of all photo-$z$ systematics will be necessary for the codes employed in upcoming analyses in order to achieve unbiased cosmological measurements.
-
-%  Photometric redshift (photo-$z$) probability distribution functions (PDFs) are a key data product of nearly all upcoming galaxy imaging surveys.
-However, the photo-$z$ PDFs resulting from different techniques are not in general consistent with one another, and an optimal method for obtaining an accurate PDF remains unclear.
-We present the results of an initial study of the Large Synoptic Survey Telescope Dark Energy Science Collaboration (\textsc{LSST-DESC}), the first in a planned series of papers testing multiple photo-$z$ codes on simulations of upcoming LSST galaxy photometry catalogues.
-This initial test evaluates photo-$z$ algorithms in the presence of representative training data and in the absence of several common sources of systematic errors that affect the procedures by which photo-$z$ PDFs are derived.
-%%  The photo-$z$ PDFs are evaluated using multiple metrics including the Kolmogorov-Smirnoff statistic, Cramer-von Mises statistic, Anderson-Darling statistic, Kullback-Leibler divergence, $N(z)$ moments, quantile-quantile plots and probability integral transform.
-%  The photo-$z$ PDFs are evaluated using multiple metrics and we observe several trends, including an overall over/under-prediction in the broadness of the PDFs for several of the codes.
-A careful accounting of all systematics discovered will be necessary for the codes employed in upcoming analyses in order to achieve unbiased cosmological measurements.
- %% Photometric redshift (photo-$z$) probability distribution functions (PDFs) more completely capture the complex nature of photo-$z$ uncertainties and are thus a key data product of nearly all upcoming galaxy imaging surveys, including the Large Synoptic Survey Telescope (LSST).
-  % Photo-$z$ PDFs produced by different algorithms do not in general agree with one another, and no one technique has yet been identified as optimal, leading to a proliferation of photo-$z$ PDF estimation codes.
-  % We present a unified comparison of twelve such codes using simulated LSST photometry, the first in a series of papers on photo-$z$ PDF performance for the LSST Dark Energy Science Collaboration (DESC).
-  % In this initial study, we isolate the effects of the estimation procedure by running each code using perfectly representative training data (for machine learning methods) or the true SED template library (for template-fitting methods) and confirm the mutual inconsistency of the resulting photo-$z$ PDF estimates.
-  % We apply a comprehensive suite of metrics to an the population of photo-$z$ PDFs as well as point estimates derived thereof and an accepted estimator of the redshift distribution $N(z)$.
-  % % The photo-$z$ PDFs are evaluated using multiple metrics including the Kolmogorov-Smirnoff statistic, Cramer-von Mises statistic, Anderson-Darling statistic, Kullback-Leibler divergence, $N(z)$ moments, quantile-quantile plots, and probability integral transform.
-  % We interpret the performance of the codes under each metric in terms of the assumptions underlying the overall approach of the methods and audit the sentivity of the metrics to the failure modes of different techniques, culminating in a discussion predicting the impact of introducing realistic systematics and the direction of attention to the choice of metric when identifying the most appropriate code for a given science use case.
+Many scientific investigations of photometric galaxy surveys require redshift estimates, whose uncertainty properties are best encapsulated by photometric redshift (photo-$z$) posterior probability distribution functions (PDFs).
+A plethora of photo-$z$ PDF estimation methodologies abound, producing discrepant results with no consensus on a preferred approach.
+We present the results of a comprehensive experiment comparing twelve photo-$z$ algorithms on mock data produced for the Large Synoptic Survey Telescope (\textsc{LSST}) Dark Energy Science Collaboration (\textsc{DESC}).
+By supplying perfect prior information, in the form of the complete template library and a representative training set as inputs to each code, we demonstrate the impact of the assumptions underlying each technique on the output photo-$z$ PDFs.
+In the absence of a notion of true, unbiased photo-$z$ PDFs, we evaluate and interpret multiple metrics of the ensemble properties of the derived photo-$z$ PDFs as well as traditional reductions to photo-$z$ point estimates.
+We report systematic biases and overall over/under-breadth of the photo-$z$ PDFs of many popular codes, which may indicate avenues for improvement in the algorithms or implementations.
+Furthermore, we raise attention to the limitations of established metrics for assessing photo-$z$ PDF accuracy and identify the conditional density estimate (CDE) loss as a promising metric of photo-$z$ PDF performance in the case where true redshifts are available but true photo-$z$ PDFs are not.
 

### 2019-03-08 paper/main.tex
clarification in abstract

@@
 We report systematic biases and overall over/under-breadth of the photo-$z$ PDFs of many popular codes, which may indicate avenues for improvement in the algorithms or implementations.
-Furthermore, we raise attention to the limitations of established metrics for assessing photo-$z$ PDF accuracy and identify the conditional density estimate (CDE) loss as a promising metric of photo-$z$ PDF performance in the case where true redshifts are available but true photo-$z$ PDFs are not.
+Furthermore, we raise attention to the limitations of established metrics for assessing photo-$z$ PDF accuracy; though we identify the conditional density estimate (CDE) loss as a promising metric of photo-$z$ PDF performance in the case where true redshifts are available but true photo-$z$ PDFs are not, we emphasize the need for science-specific performance metrics.
 

### 2019-03-08 paper/tex/introduction.tex
tightening up introduction

--- 2018-11-21 paper/tex/introduction.tex
@@
 
-%(Sam Schmidt, Ofer Lahav, Jeff Newman, Alex Malz, John Soo, Tony Tyson)
+The current and next generations of large-scale galaxy surveys, including the Dark Energy Survey \citep[\textsc{DES},][]{Abbott:05}, the Kilo-Degree Survey \citep[\textsc{KiDS},][]{de_Jong:13}, Hyper Suprime-Cam Survey \citep[\textsc{HSC},][]{Aihara:2018a,Aihara:2018b}, Large Synoptic Survey Telescope \citep[\lsst,][]{Abell:09}, Euclid \citep{Laureijs:11}, and Wide-Field Infrared Survey Telescope \citep[\textsc{WFIRST},][]{Green:12}, present a paradigm shift from spectroscopic to photometric galaxy catalogs of substantially larger size at a cost of lacking complete spectroscopically confirmed redshifts.
 
-%\jan{Is the intention to keep these parenthetical author sets?
-I find it a bit offputting in the text.
-It might be better to put this for sections in the acknowledgements section, and for individual codes/statistics specify who ran/developed it with a sentence.}\red{I was just about to comment these out before finalizing the internal review draft of the paper.
-They are in place to remind me to contact a few people who still have not filled out their authors.csv info.}
+Effective astrophysical inference using the catalogs resulting from these ongoing and upcoming missions, however, necessitates accurate and precise photometric redshift (\pz) estimation methodologies.
+As an example, in order for \pz\ systematics to not dominate the statistical noise floor of \lsst's main cosmological sample of $\sim 10^{7}$ galaxies, the \lsst\ Science Requirements Document (SRD)\footnote{available at \url{https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17}} specifies that individual galaxy \pz s must have root-mean-square error $\sigma_z < 0.02 (1+z)$, $3 \sigma$ ``catastrophic outlier'' rate below $10\%$, and bias below $0.003$.
+Specific science cases may have their own requirements on \pz\ performance that exceed those of the survey as a whole.
+In that vein, the \lsst\ Dark Energy Science Collaboration (\lsstdesc) developed a separate SRD \citep{Mandelbaum:2018} that conservatively forecasts the constraining power of five cosmological probes, leading to even more stringent requirements on \pz\ performance, including those defined in terms of tomographically binned subsamples populations rather than individual galaxies.
 
-Large-scale photometric galaxy surveys are entering a new era with currently or soon-to-be running Stage III and Stage IV dark energy experiments like the Dark Energy Survey \citep[DES,][]{Abbott:05}, the Kilo-Degree Survey \citep[KiDS,][]{de_Jong:13}, Hyper Suprime-Cam (HSC) Survey \citep[]{Aihara:2018a,Aihara:2018b}, Large Synoptic Survey Telescope \citep[LSST,][]{Abell:09}, Euclid \citep{Laureijs:11}, and Wide-Field Infrared Survey Telescope \citep[WFIRST,][]{Green:12}.
-The move to imaging based surveys, rather than spectroscopic based, for cosmological measurements makes proper understanding of photometric redshifts (``photo-$z$'s'') of paramount importance, as cosmological distance measures for statistical samples are directly dependent on photo-$z$ measurements.
+Though the standard has long been for each galaxy in a photometric catalog to have a \pz\ point estimate and Gaussian error bar, the nontrivial mapping between broad band fluxes and redshift renders this simplistic summary inadequate to quantify the uncertainty landscape by neglecting degenerate redshift solutions.
+Far from a hypothetical situation, this degeneracy is a real consequence of the same deep imaging that enables larger galaxy catalog sizes.
+The lower luminosity and higher redshift populations captured by deeper imaging introduce major physical systematics to \pz s, among them the Lyman break/Balmer break degeneracy, that did not affect shallower large area surveys like the Sloan Digital Sky Survey \citep[\textsc{SDSS},][]{York:00} and Two Micron All Sky Survey \citep[\textsc{2MASS},][]{Skrutskie:06}.
 
-The unprecedented sample size of LSST galaxies, expected to number several billion for the main cosmological sample, necessitates stringent constraints on photo-$z$ accuracy if systematic errors are not to dominate the statistical errors.
-The LSST Science Requirements Document (SRD)\footnote{available at \url{https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17}} lists the individual galaxy photometric redshift goals for a magnitude limited sample with $i<25$ as: root-mean-square error with a goal of $\sigma_z<0.02(1+z)$; $3\sigma$ ``catastrophic outlier'' rate below $10\%$; bias below $0.003$\footnote{
-Note that at the time the SRD was written, these goals were stated in terms of a photo-$z$ point estimate for each galaxy, as was standard in many previous studies, while in this paper we emphasize the importance of using a full photo-$z$ PDF.}.
-The LSST Dark Energy Science Collaboration (LSST-DESC) developed a separate Science Requirements Document \citep{Mandelbaum:2018}, which forecasts the constraining power of five cosmological probes using somewhat conservative assumptions to define requirements on systematic errors for several measurements.
-These include even more stringent requirements on photometric redshift performance than those included in the LSST SRD, though most of the initial LSST-DESC requirements are defined in terms of tomographic bin populations rather than on individual object redshifts.
-The tremendous size of LSST's galaxy catalogue will be enabled by its exceptional depth, pushing to fainter magnitudes and deeper imaging and including galaxies of lower luminosity and higher redshift than ever before.
-The inclusion of these populations introduce major physical degeneracies, for example the Lyman break/Balmer break degeneracy, that were not present in the populations covered in shallower large area surveys like the Sloan Digital Sky Survey \citep[SDSS,][]{York:00} and the Two Micron All Sky Survey \citep[2MASS,][]{Skrutskie:06}.
-These issues are not unique to LSST, and are present in Stage III Dark Energy surveys; however,
-in order to meet the demanding error budgets of Stage IV projects such as LSST and LSST-DESC it will be necessary to fully characterize those degeneracies wherein multiple redshift solutions have comparable likelihood to per cent level accuracy.
+To fully characterize such physical degeneracies, photometric galaxy catalog data releases from \citep{Mandelbaum:2008} to \citep{de_Jong:17}, provided a more informative \pz\ data product, the \pz\ probability density function (PDF), that describes the relative probability as a function of a galaxy's redshift, conditioned on the observed photometry.
+Early template-based methods such as \citet{Fernandezsoto:99} approximated the likelihood of photometry conditioned on redshift with the relative $\chi^{2}$ values of template spectra.
+Not long after, Bayesian adaptations of template-based approaches such as \citet{Benitez:00} combined the estimated likelihoods with a prior to yield a posterior PDF of redshift conditioned on photometry.
+While the first machine learning based algorithms focused on a point-estimate, \citet{Firth:03} estimated a \pzpdf\ using a neural net with 1000 realizations scattered within the photometric errors.
 
-There is often a desire to have a single valued ``point-estimate'' redshift for an individual galaxy.
-However, the complex, non-linear (and often non-unique) nature of the mapping between broad band fluxes and redshift means that a single value is unable to capture the full redshift information encoded in a galaxy's magnitudes.
-For example, a common point-estimate for a template-based method is taking the highest likelihood solution as the point photo-$z$.
-A single valued redshift ignores degenerate redshift solutions of lower probability, potentially biasing photometric redshift estimates both for individual galaxies and ensemble distributions.
-Storing more information is necessary, most often photo-$z$ codes output the redshift probability density function, also often referred to as $p(z)$, describing the relative likelihood as a function of redshift.
-%\red{should we move some of AIM's p(z) vs p(z|d,I) to here?}
-Early template methods such as \citet{Fernandezsoto:99} converted relative $\chi^2$ values of template spectra to likelihoods to estimate $p(z)$.
-Soon after, codes such as \citet{Benitez:00} added a Bayesian prior and output a posterior probability distribution.
-While many early machine learning based algorithms focused on a point-estimate, \citet{Firth:03} used a neural net with 1000 realizations scattered within the photometric errors to estimate a $p(z)$.
-As more groups began to employ photometric redshifts in their cosmological analyses, there was a realization that point-estimate photo-$z$'s were inadequate for precision cosmology measurements \citep{Mandelbaum:2008}.
-From around this point onward, most photo-$z$ algorithms have attempted to implement some estimate of the overall redshift probability in their outputs, and some surveys began supplying a full $p(z)$ rather than a simple redshift point-estimate and error \citep[e.~g.~][]{de_Jong:17}.
+% MOVE TO WHERE \nz\ IS INTRODUCED AS A METRIC
+% For cosmological measurements, certain science cases require redshift information on individual objects, e.~g.~identification of host galaxy redshift for supernova classification, or identifying potential cluster membership.
+% Other science cases seem to need only ensemble redshift information; for instance many current cosmic shear techniques require only the overall redshift distribution $N(z)$ for tomographic redshift samples.
+% However,  even such cases require individual object redshift estimates for portions of the analysis, for example in determining galaxy intrinsic alignments in weak lensing samples.
+% In addition, recent data-driven techniques employing hierarchical Bayesian or Gaussian Process methods have emerged that calibrate redshift distributions using individual $p(z)$ estimates \citep[e.~g.~][]{Sanchez:2018}.
+% Techniques for using \pzpdf s have lagged behind the development of codes to produce them, however, all existing approaches assume that the \pzpdf\ for each galaxy is an accurate PDF, with failures when the assumption is violated.
+% Thus, even methods that seem to need only ensemble $N(z)$ may actually require accurate $p(z)$ in order to meet stringent survey requirements.
+% Large photometric surveys such as LSST must develop algorithms that simultaneously meet the needs of all science cases.
+% In order to meet these ambitious goals for photo-$z$ accuracy, every aspect of photo-$z$ estimation will have to be optimized: the algorithms employed, both template and machine-learning based (both in design and implementation); the spectroscopic data used as a training set for machine learning algorithms or to estimate template sets and train Bayesian priors; and probabilistic catalogue compression schemes that balance information retention against limited storage resources.
 
-For cosmological measurements, certain science cases require redshift information on individual objects, e.~g.~identification of host galaxy redshift for supernova classification, or identifying potential cluster membership.
-Other science cases seem to need only ensemble redshift information; for instance many current cosmic shear techniques require only the overall redshift distribution $N(z)$ for tomographic redshift samples.
-However,  even such cases require individual object redshift estimates for portions of the analysis, for example in determining galaxy intrinsic alignments in weak lensing samples.
-In addition, recent data-driven techniques employing hierarchical Bayesian or Gaussian Process methods have emerged that calibrate redshift distributions using individual $p(z)$ estimates \citep[e.~g.~][]{Sanchez:2018}.
-These methods assume that the $p(z)$ for each galaxy is an accurate PDF, and such methods break down if this assumption is invalid.
-Thus, even methods that seem to need only ensemble $N(z)$ may actually require accurate $p(z)$ in order to meet stringent survey requirements.
-%In the case of the multiple types of probes of cosmology enabled by the
-%LSST cosmology sample of several billion galaxies, the number of redshift bins and their photo-$z$ requirements vary with the specific probe;  2-point angular correlations benefit from many bins, while weak lensing probes do not (due to the wide lensing kernel).
-Large photometric surveys such as LSST must develop algorithms that simultaneously meet the needs of all science cases.
-In order to meet these ambitious goals for photo-$z$ accuracy, every aspect of photo-$z$ estimation will have to be optimized: the algorithms employed, both template and machine-learning based (both in design and implementation); the spectroscopic data used as a training set for machine learning algorithms or to estimate template sets and train Bayesian priors; and probabilistic catalogue compression schemes that balance information retention against limited storage resources.
+There are numerous techniques for deriving \pzpdf s, yet no one method has yet been established as clearly superior.
+Quantitative comparisons of photo-$z$ methods have been made before.
+The Photo-$z$ Accuracy And Testing \citep[\textsc{PHAT},][]{Hildebrandt:10} effort focused on \pz\ point estimates derived from many photometric bands.
+\citet{Rau:2015} introduced a new method for improving \pzpdf s using an ordinal classification algorithm.
+\textsc{DES} compared several codes for \pz\ point estimates and a subset with \pzpdf\ information \citep{Sanchez:14} and examined summary statistics of \pzpdf s for tomographically binned galaxy subsamples \citep{Bonnett:16}.
 
-There are numerous techniques for deriving photo-$z$ PDFs from photometry, yet no one method has yet been established as clearly superior.
-Quantitative comparisons of photo-$z$ methods have been made before.
-The Photo-$z$ Accuracy And Testing \citep[PHAT,][]{Hildebrandt:10} effort focused on point estimates derived from many photometric bands.
-\citet{Rau:2015} introduced a new method for improving redshift PDFs using an ordinal classification algorithm.
-DES compared several codes for point estimates and a subset with $p(z)$ information \citep{Sanchez:14}.
-A follup up paper examined summary statistics of photo-$z$ interim posteriors for tomographically binned galaxy subsamples \citep{Bonnett:16}.
-This paper is distinguished by its focus on metrics of photo-$z$ interim posteriors themselves and consideration of both classic and state-of-the-art photo-$z$ algorithms, comparing the performance of several of the most widely employed codes as well as some that have been developed only recently on the basis of metrics appropriate for a probabilistic data product.
-The results presented in this work are
-%part of the Data Challenge 1 (DC1)
-a major focus of the Photometric Redshift working group of the LSST-DESC.
-This work is laid out in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}} as one of the critical activities to be completed in preparation for dark energy science analysis on the first year LSST data.
-%This is the first of multiple papers by the working group, which will grow in sophistication.
-In this initial paper we focus on evaluating the performance of photometric redshift codes and PDF-based performance metrics in the presence of representative training sets.
-Comparing the relative performance of the codes enables us to evaluate whether each code is using information in an optimal way, and may reveal enhancements in some codes and deficiencies in others, either in the fundamental algorithm, or in specific implementation.
-We note that these initial tests are a necessary requirement for photo-$z$ codes that will be used in cosmological analyses; however, meeting these requirements is only the first stage in the process, and can be thought of as an initial test under near perfect conditions before further complexities are added in future analyses.
-
-Before moving forward, we must address how the best methods may be unique to the performance metrics and science cases considered and what distinguishes photo-$z$ PDFs of different methods from one another.
-Though photo-$z$ PDFs are often written simply as $p(z)$, the PDF itself must be an interim posterior distribution $p(z | d, I)$, the probability of redshift conditioned on photometric data $d$ that has actually been observed and the prior information $I$ that guides how a redshift is extracted from the photometry.
-If we run multiple photo-$z$ codes on a single dataset, the photo-$z$  posteriors will not be identical because each code employs assumptions that constitute an interim prior; these assumptions form the premise for photo-$z$ estimation as a whole and are the only way to introduce differences in estimates of what would otherwise be a shared photo-$z$ posterior $p(z | d)$ regardless of the code used to obtain it.
-Though explicit knowledge of the interim prior is necessary to use photo-$z$ posteriors self-consistently in physical inference, the interim prior of a particular methodology is often implicit and not necessarily shared among all galaxies in the catalogue.
-
-This paper therefore aims (1) to constrain the impact of the interim prior $I$ by separating it into a component $I_{H}$ representing the method itself and a component $I_{D}$ representing physical information, such as a training set or SED template library and (2) to present a procedure for evaluating the performance of photo-$z$ codes in generic tests that may include many more systematics in the interim prior $I$.
-In order to isolate the effects encapsulated by $I_{H}$ of variation between codes from issues with the training set or template library encapsulated by $I_{D}$, we use an identical set of simulated galaxies for every code and construct a template library and training sample that are {\it complete and representative} and shared among all codes; that is, our training sample for machine learning codes is drawn from the same underlying galaxy population as our test set, with no additional selections, and the SED library used for template-based codes is the same as the one used to generate the photometric data.
-We explore a number of performance metrics in this paper, in the controlled environment of the representative training data in order to establish a method for comparing photo-$z$ PDFs derived by different methods.
-%not to make a conclusion regarding the superiority or even relative favorability of each code but to establish a method for comparing photo-$z$ PDFs derived by different methods.
-These test conditions set the stage for addressing in a future paper the crucial issue of incomplete and non-representative prior information.
+This paper is distinguished by its focus on the evaluation criteria for \pzpdf s and interpretation thereof, as part of the mission of the , as a key project of the Photometric Redshift working group of the \lsstdesc\ laid out in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}}.
+In this initial study, we focus on evaluating the performance of \pzpdf\ codes and PDF-specific performance metrics in a controlled experiment with complete and representative prior information (template libraries and training sets).
+% Specific implementation choices in each code will influence the resultant posterior distributions, for example choice of prior parameterization in template-based codes, the bandwidth size chosen for machine learning based codes, or even the output format chosen for storing the PDF.
+% We have attempted to minimize the impact of many of these factors when comparing codes, for example by using the same template set for all template-based codes, and using a training set that is drawn from the same underlying population as the test sample, to create a controlled environment in which to compare the photo-$z$ PDFs derived from each method.
+% We explore a number of performance metrics in this paper that test whether the posterior estimates are actual PDFs.
+% Comparing the relative performance of the codes enables us to evaluate whether each code is using information in an optimal way, and may reveal enhancements in some codes and deficiencies in others, either in the fundamental algorithm, or in specific implementation.
+% Identifying and fixing failure modes within codes may aid us in reaching the stringent photo-$z$ performance goals set out for LSST.
+% We note that these initial tests are a necessary requirement for photo-$z$ codes that will be used in cosmological analyses; however, meeting these requirements is only the first stage in the process, and can be thought of as an initial test under near perfect conditions to test for problems before further complexities are added in future analyses.
 

### 2019-03-09 paper/tex/introduction.tex
typo

@@
 There are numerous techniques for deriving \pzpdf s, yet no one method has yet been established as clearly superior.
-Quantitative comparisons of photo-$z$ methods have been made before.
-The Photo-$z$ Accuracy And Testing \citep[\textsc{PHAT},][]{Hildebrandt:10} effort focused on \pz\ point estimates derived from many photometric bands.
+Quantitative comparisons of \pz\ methods have been made before.
+The \Pz\ Accuracy And Testing \citep[\textsc{PHAT},][]{Hildebrandt:10} effort focused on \pz\ point estimates derived from many photometric bands.
 \citet{Rau:2015} introduced a new method for improving \pzpdf s using an ordinal classification algorithm.

### 2019-03-09 paper/tex/introduction.tex
typo, adding back some cut content in more appropriate location

@@
 Effective astrophysical inference using the catalogs resulting from these ongoing and upcoming missions, however, necessitates accurate and precise photometric redshift (\pz) estimation methodologies.
-As an example, in order for \pz\ systematics to not dominate the statistical noise floor of \lsst's main cosmological sample of $\sim 10^{7}$ galaxies, the \lsst\ Science Requirements Document (SRD)\footnote{available at \url{https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17}} specifies that individual galaxy \pz s must have root-mean-square error $\sigma_z < 0.02 (1+z)$, $3 \sigma$ ``catastrophic outlier'' rate below $10\%$, and bias below $0.003$.
+As an example, in order for \pz\ systematics to not dominate the statistical noise floor of \lsst's main cosmological sample of $\sim 10^{7}$ galaxies, the \lsst\ Science Requirements Document (SRD)\footnote{available at \url{https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17}} specifies that individual galaxy \pz s must have root-mean-square error $\sigma_z < 0.02 (1+z)$, $3 \sigma$ catastrophic outlier rate below $10\%$, and bias below $0.003$.
 Specific science cases may have their own requirements on \pz\ performance that exceed those of the survey as a whole.
 In that vein, the \lsst\ Dark Energy Science Collaboration (\lsstdesc) developed a separate SRD \citep{Mandelbaum:2018} that conservatively forecasts the constraining power of five cosmological probes, leading to even more stringent requirements on \pz\ performance, including those defined in terms of tomographically binned subsamples populations rather than individual galaxies.
-
-Though the standard has long been for each galaxy in a photometric catalog to have a \pz\ point estimate and Gaussian error bar, the nontrivial mapping between broad band fluxes and redshift renders this simplistic summary inadequate to quantify the uncertainty landscape by neglecting degenerate redshift solutions.
-Far from a hypothetical situation, this degeneracy is a real consequence of the same deep imaging that enables larger galaxy catalog sizes.
-The lower luminosity and higher redshift populations captured by deeper imaging introduce major physical systematics to \pz s, among them the Lyman break/Balmer break degeneracy, that did not affect shallower large area surveys like the Sloan Digital Sky Survey \citep[\textsc{SDSS},][]{York:00} and Two Micron All Sky Survey \citep[\textsc{2MASS},][]{Skrutskie:06}.
-
-To fully characterize such physical degeneracies, photometric galaxy catalog data releases from \citep{Mandelbaum:2008} to \citep{de_Jong:17}, provided a more informative \pz\ data product, the \pz\ probability density function (PDF), that describes the relative probability as a function of a galaxy's redshift, conditioned on the observed photometry.
-Early template-based methods such as \citet{Fernandezsoto:99} approximated the likelihood of photometry conditioned on redshift with the relative $\chi^{2}$ values of template spectra.
-Not long after, Bayesian adaptations of template-based approaches such as \citet{Benitez:00} combined the estimated likelihoods with a prior to yield a posterior PDF of redshift conditioned on photometry.
-While the first machine learning based algorithms focused on a point-estimate, \citet{Firth:03} estimated a \pzpdf\ using a neural net with 1000 realizations scattered within the photometric errors.
 
@@
 
+Though the standard has long been for each galaxy in a photometric catalog to have a \pz\ point estimate and Gaussian error bar, the nontrivial mapping between broad band fluxes and redshift renders this simplistic summary inadequate to quantify the uncertainty landscape by neglecting degenerate redshift solutions.
+Far from a hypothetical situation, this degeneracy is a real consequence of the same deep imaging that enables larger galaxy catalog sizes.
+The lower luminosity and higher redshift populations captured by deeper imaging introduce major physical systematics to \pz s, among them the Lyman break/Balmer break degeneracy, that did not affect shallower large area surveys like the Sloan Digital Sky Survey \citep[\textsc{SDSS},][]{York:00} and Two Micron All Sky Survey \citep[\textsc{2MASS},][]{Skrutskie:06}.
+
+To fully characterize such physical degeneracies, photometric galaxy catalog data releases from \citep{Mandelbaum:2008} to \citep{de_Jong:17}, provided a more informative \pz\ data product, the \pz\ probability density function (PDF), that describes the relative probability as a function of a galaxy's redshift, conditioned on the observed photometry.
+Early template-based methods such as \citet{Fernandezsoto:99} approximated the likelihood of photometry conditioned on redshift with the relative $\chi^{2}$ values of template spectra.
+Not long after, Bayesian adaptations of template-based approaches such as \citet{Benitez:00} combined the estimated likelihoods with a prior to yield a posterior PDF of redshift conditioned on photometry.
+While the first data-driven \pz\ algorithms yielded a point estimate, \citet{Firth:03} estimated a \pzpdf\ using a neural net with realizations scattered within the photometric errors.
+
 There are numerous techniques for deriving \pzpdf s, yet no one method has yet been established as clearly superior.
@@
 
-This paper is distinguished by its focus on the evaluation criteria for \pzpdf s and interpretation thereof, as part of the mission of the , as a key project of the Photometric Redshift working group of the \lsstdesc\ laid out in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}}.
-In this initial study, we focus on evaluating the performance of \pzpdf\ codes and PDF-specific performance metrics in a controlled experiment with complete and representative prior information (template libraries and training sets).
-% Specific implementation choices in each code will influence the resultant posterior distributions, for example choice of prior parameterization in template-based codes, the bandwidth size chosen for machine learning based codes, or even the output format chosen for storing the PDF.
-% We have attempted to minimize the impact of many of these factors when comparing codes, for example by using the same template set for all template-based codes, and using a training set that is drawn from the same underlying population as the test sample, to create a controlled environment in which to compare the photo-$z$ PDFs derived from each method.
-% We explore a number of performance metrics in this paper that test whether the posterior estimates are actual PDFs.
-% Comparing the relative performance of the codes enables us to evaluate whether each code is using information in an optimal way, and may reveal enhancements in some codes and deficiencies in others, either in the fundamental algorithm, or in specific implementation.
-% Identifying and fixing failure modes within codes may aid us in reaching the stringent photo-$z$ performance goals set out for LSST.
-% We note that these initial tests are a necessary requirement for photo-$z$ codes that will be used in cosmological analyses; however, meeting these requirements is only the first stage in the process, and can be thought of as an initial test under near perfect conditions to test for problems before further complexities are added in future analyses.
+This paper is distinguished by its focus on the evaluation criteria for \pzpdf s and interpretation thereof, as a key project of the Photometric Redshift working group of the \lsstdesc\ that aims to perform a comprehensive sensitivity analysis of \pzpdf\ techniques in order to select those that will become part of the \lsst\ pipelines, as laid out in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}}.
+In this initial study, we focus on evaluating the performance of \pzpdf\ codes using PDF-specific performance metrics in a controlled experiment with complete and representative prior information (template libraries and training sets) to set a baseline for subsequent investigations.
+This approach probes how each code considered exploits the information content of the data versus prior information from template libraries and training sets.
 

### 2019-03-09 paper/tex/data.tex
reorganization

@@
-\section{The simulation and mock galaxy catalog}
+\section{Data}
 \label{sec:sims}
-%(Sam Schmidt, Eve Kovacs, Tina Peters)
 
-In order to test the current generation codes, we employ an existing simulated galaxy catalogue.
-The simulation is completely catalogue-based, with no image construction or mock measurements made.
-We describe these in detail below.
+In order to test the current generation of \pzpdf\ codes, we employ an existing simulated galaxy catalog, described in detail in Section~\ref{sec:buzzard}.
+The experimental conditions shared among all codes are motivated by the \lsst\ SRD requirements and implemented for machine learning and template-based \pzpdf\ codes according to the procedures of Sections~\ref{sec:buzztraining} and \ref{sec:buzztemplates} respectively.
+Finally, in Section~\ref{sec:limitations}, we elaborate on the scope of the experimental conditions considered as a result of the sophistication of the mock data.
 
@@
 \label{sec:buzzard}
-The \textsc{Buzzard-highres-v1.0} (De Rose et al., in prep; Wechsler et al., in prep) catalogue construction started with a dark matter only simulation.
+
+The \textsc{Buzzard-highres-v1.0} (De Rose et al., in prep; Wechsler et al., in prep) catalog construction started with a dark matter only simulation.
 This N-body simulation contained $2048^3$ particles in a $400$ Mpc h$^{-1}$ box.
@@
 Dark matter halos were identified using the \textsc{Rockstar} software package \citep{Behroozi:13}.
-These dark matter halos were populated with galaxies with a stellar mass and absolute $r$-band magnitude in the SDSS system determined using a sub-halo abundance matching model constrained to match both projected two-point galaxy clustering statistics and an observed conditional stellar mass function \citep{Reddick:13}.
+These dark matter halos were populated with galaxies with a stellar mass and absolute $r$-band magnitude in the \textsc{SDSS} system determined using a sub-halo abundance matching model constrained to match both projected two-point galaxy clustering statistics and an observed conditional stellar mass function \citep{Reddick:13}.
 
@@
 
-\subsubsection{Selection of training and test sets}
+\subsection{Experimental Conditions}
+\label{sec:controlled}
+
+% NOW MOTIVATE THE SHARED PRIOR INFORMATION
+
+\subsubsection{Training set selection}
 \label{sec:buzztraining}
+
 The total initial catalogue covered $400$ square degrees and contained $238$ million galaxies to an apparent magnitude limit of $r\!=\!29$ and spanning the redshift range $0\!<\!z\!\leq\!8.7$.
@@
 Finally, we restrict our analysis to a sample with an apparent i-band magnitude limit $i<25.3$, which give a signal-to-noise $\sim\,30$ for most galaxies, a cut often referred to as the expected ``LSST Gold Sample''.
-This magnitude cut results in a training set with $44\,404$ galaxies and a test set containing $399\,356$ galaxies.
-All subsequent results will evaluate this ``gold sample'' test set.
+This magnitude cut results in a training set with $44\,404$ galaxies and a test set containing $399\,356$ galaxies.  All subsequent results will evaluate this ``gold sample'' test set.
 In order to blind results, initially redshifts were not revealed for the ``test'' set, and were only supplied for the training sample galaxies.
@@
 
-\subsubsection{Templates}
+\subsubsection{Template library construction}
 \label{sec:buzztemplates}
+
 As mentioned in Section~\ref{sec:buzzard}, the SEDs in the Buzzard simulation are drawn from an empirical set of SEDs taken from the SDSS DR6 NYU-VAGC, a sample of roughly $\sim5\times 10^{5}$ galaxies with spectra in SDSS.
@@
 
-\subsubsection{Limitations}
+\subsection{Limitations}
 \label{sec:buzzlimitations}
+
 For our initial investigation of photometric redshift codes, we begin with a data set that is somewhat idealized, and does not contain all of the complicating factors present in real data.
@@
 We aim to test the photo-z codes at a very basic level, and a simplified model assures that differences in results seen between the codes are due to fundamental differences in their underlying assumptions and implementation details, rather than more nuanced properties.
-
-%\subsection{ABC-Galacticus simulation}
-%\label{sec:galacticus}
-%\red{The Argonne-Berkeley-Carnegie (ABC) Galacticus cosmological simulation uses a semi-analytic model to simulate the galaxy properties and is not directly tuned to observations.
-Each simulated SED is replaced with a SED drawn from the distribution of empirical SEDs given by the SED Atlas of \citet{Brown:14} (diffusion maps method in development).}

### 2019-03-09 paper/tex/data.tex
enforcing British spelling

@@
 
-In order to test the current generation of \pzpdf\ codes, we employ an existing simulated galaxy catalog, described in detail in Section~\ref{sec:buzzard}.
+In order to test the current generation of \pzpdf\ codes, we employ an existing simulated galaxy catalogue, described in detail in Section~\ref{sec:buzzard}.
 The experimental conditions shared among all codes are motivated by the \lsst\ SRD requirements and implemented for machine learning and template-based \pzpdf\ codes according to the procedures of Sections~\ref{sec:buzztraining} and \ref{sec:buzztemplates} respectively.
@@
 
-The \textsc{Buzzard-highres-v1.0} (De Rose et al., in prep; Wechsler et al., in prep) catalog construction started with a dark matter only simulation.
-This N-body simulation contained $2048^3$ particles in a $400$ Mpc h$^{-1}$ box.
-A set of time snapshots (with smoothing and interpolation between snapshots) were saved in order to construct a lightcone.
-Dark matter halos were identified using the \textsc{Rockstar} software package \citep{Behroozi:13}.
-These dark matter halos were populated with galaxies with a stellar mass and absolute $r$-band magnitude in the \textsc{SDSS} system determined using a sub-halo abundance matching model constrained to match both projected two-point galaxy clustering statistics and an observed conditional stellar mass function \citep{Reddick:13}.
+Our mock catalogue is derived from the \textsc{Buzzard-highres-v1.0}  of De Rose et al., in prep, Wechsler et al., in prep) catalogue.
+% UPDATE THIS CITATION NOW THAT THE PAPER IS OUT
+\textsc{Buzzard} is built on a dark matter-only N-body simulation of $2048^{3}$ particles in a $400$ Mpc h$^{-1}$ box.
+The lightcone was constructed from smoothing and interpolation between a set of time snapshots.
+Dark matter halos were identified using the \textsc{Rockstar} software package \citep{Behroozi:13} and then populated with galaxies with a stellar mass and absolute $r$-band magnitude in the \sdss\ system determined using a sub-halo abundance matching model constrained to match both projected two-point galaxy clustering statistics and an observed conditional stellar mass function \citep{Reddick:13}.
 
-To assign an SED to each galaxy, the {\it Adding Density Dependent Spectral Energy Distributions} (\textsc{ADDSEDS}, deRose in prep.)\footnote{\url{https://github.com/vipasu/addseds}} procedure was used.
-This consisted of training an empirical relation between absolute $r$-band magnitude, local galaxy density, and SED using a sample of $\sim 5\times 10^{5}$ galaxies from the magnitude-limited Sloan Digital Sky Survey Data Release 6 Value Added Galaxy Catalog \citep{Blanton:05}.
-Each SDSS spectrum is fit with a sum of five SED components using the \textsc{k-correct \red{v4\_3?}} software package\footnote{\url{http://kcorrect.org}} \citep{Blanton:07}, thus each galaxy SED is parameterized as five weights for the basis SEDs.
-The distance to the spatial projected fifth-nearest neighbour was used as a proxy for local density in the SDSS training sample.
+To assign an SED to each galaxy, the {\it Adding Density Dependent Spectral Energy Distributions} (\texttt{ADDSEDS}, deRose in prep.)\footnote{\url{https://github.com/vipasu/addseds}} procedure was used.
+\texttt{ADDSEDS} uses a sample of $\sim 5\times 10^{5}$ galaxies from the magnitude-limited \sdss\ Data Release 6 Value Added Galaxy Catalogue \citep{Blanton:05} to train an empirical relation between absolute $r$-band magnitude, local galaxy density, and SED.
+Each \sdss\ spectrum is parameterized by five weights corresponding to a weighted sum of five basis SED components using the \texttt{k-correct \red{v4\_3?}} software package\footnote{\url{http://kcorrect.org}} \citep{Blanton:07}.
+
+Correlations between SED and galaxy environment were included so as to preserve the colour-density relation of galaxy environment.
+The distance to the spatially projected fifth-nearest neighbour was used as a proxy for local density in the \sdss\ training sample.
 For each simulated galaxy, a galaxy with similar absolute $r$-band magnitude and local galaxy density was chosen from the training set, and that training galaxy's SED was assigned to the simulated galaxy.
-This process is done in such a way as to preserve the colour-density relation of galaxy environment.
-Given the SED, absolute $r$-band magnitude and redshift, we computed apparent magnitudes in the six LSST filter passbands, $ugrizy$.
-We assigned magnitude errors in the six bands using the simple model described in \citet{Ivezic:08}, assuming full 10-year depth observations had been completed.
+
+Given the SED, absolute $r$-band magnitude, and redshift, we computed apparent magnitudes in the six \lsst\ filter passbands, $ugrizy$.
+We assigned magnitude errors in the six bands using the simple model of \citet{Ivezic:08}, assuming achievement of the full 10-year depth.
 The number of total 30-second visits assumed when generating the photometric errors differs slightly from the fiducial numbers assumed for LSST: we assume 60 visits in u-band, 80 visits in g-band, 180 visits in r-band, 180 visits in i-band, 160 visits in z-band, and 160 visits in y-band.
@@
 In the course of simulating Gaussian photometric errors, we add noise to objects fluxes, and some of these noisy fluxes will become negative in one or more bands.
-We call such negative fluxes ``non-detections'' and signify them with a placholder magnitude of 99.0 in the catalog.
+We call such negative fluxes ``non-detections'' and signify them with a placholder magnitude of 99.0 in the catalogue.
 Thus, further mentions of ``non-detections'' refer to objects that would be ``looked at but not seen'' in multi-band forced photometry, and the photo-z codes will treat them as such.

### 2019-03-09 paper/tex/introduction.tex
enforcing British spelling

@@
 
-The current and next generations of large-scale galaxy surveys, including the Dark Energy Survey \citep[\textsc{DES},][]{Abbott:05}, the Kilo-Degree Survey \citep[\textsc{KiDS},][]{de_Jong:13}, Hyper Suprime-Cam Survey \citep[\textsc{HSC},][]{Aihara:2018a,Aihara:2018b}, Large Synoptic Survey Telescope \citep[\lsst,][]{Abell:09}, Euclid \citep{Laureijs:11}, and Wide-Field Infrared Survey Telescope \citep[\textsc{WFIRST},][]{Green:12}, present a paradigm shift from spectroscopic to photometric galaxy catalogs of substantially larger size at a cost of lacking complete spectroscopically confirmed redshifts.
+The current and next generations of large-scale galaxy surveys, including the Dark Energy Survey \citep[\textsc{DES},][]{Abbott:05}, the Kilo-Degree Survey \citep[\textsc{KiDS},][]{de_Jong:13}, Hyper Suprime-Cam Survey \citep[\textsc{HSC},][]{Aihara:2018a,Aihara:2018b}, Large Synoptic Survey Telescope \citep[\lsst,][]{Abell:09}, Euclid \citep{Laureijs:11}, and Wide-Field Infrared Survey Telescope \citep[\textsc{WFIRST},][]{Green:12}, present a paradigm shift from spectroscopic to photometric galaxy catalogues of substantially larger size at a cost of lacking complete spectroscopically confirmed redshifts.
 
-Effective astrophysical inference using the catalogs resulting from these ongoing and upcoming missions, however, necessitates accurate and precise photometric redshift (\pz) estimation methodologies.
+Effective astrophysical inference using the catalogues resulting from these ongoing and upcoming missions, however, necessitates accurate and precise photometric redshift (\pz) estimation methodologies.
 As an example, in order for \pz\ systematics to not dominate the statistical noise floor of \lsst's main cosmological sample of $\sim 10^{7}$ galaxies, the \lsst\ Science Requirements Document (SRD)\footnote{available at \url{https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17}} specifies that individual galaxy \pz s must have root-mean-square error $\sigma_z < 0.02 (1+z)$, $3 \sigma$ catastrophic outlier rate below $10\%$, and bias below $0.003$.
@@
 
-Though the standard has long been for each galaxy in a photometric catalog to have a \pz\ point estimate and Gaussian error bar, the nontrivial mapping between broad band fluxes and redshift renders this simplistic summary inadequate to quantify the uncertainty landscape by neglecting degenerate redshift solutions.
-Far from a hypothetical situation, this degeneracy is a real consequence of the same deep imaging that enables larger galaxy catalog sizes.
+Though the standard has long been for each galaxy in a photometric catalogue to have a \pz\ point estimate and Gaussian error bar, the nontrivial mapping between broad band fluxes and redshift renders this simplistic summary inadequate to quantify the uncertainty landscape by neglecting degenerate redshift solutions.
+Far from a hypothetical situation, this degeneracy is a real consequence of the same deep imaging that enables larger galaxy catalogue sizes.
 The lower luminosity and higher redshift populations captured by deeper imaging introduce major physical systematics to \pz s, among them the Lyman break/Balmer break degeneracy, that did not affect shallower large area surveys like the Sloan Digital Sky Survey \citep[\textsc{SDSS},][]{York:00} and Two Micron All Sky Survey \citep[\textsc{2MASS},][]{Skrutskie:06}.
 
-To fully characterize such physical degeneracies, photometric galaxy catalog data releases from \citep{Mandelbaum:2008} to \citep{de_Jong:17}, provided a more informative \pz\ data product, the \pz\ probability density function (PDF), that describes the relative probability as a function of a galaxy's redshift, conditioned on the observed photometry.
+To fully characterize such physical degeneracies, photometric galaxy catalogue data releases from \citep{Mandelbaum:2008} to \citep{de_Jong:17}, provided a more informative \pz\ data product, the \pz\ probability density function (PDF), that describes the relative probability as a function of a galaxy's redshift, conditioned on the observed photometry.
 Early template-based methods such as \citet{Fernandezsoto:99} approximated the likelihood of photometry conditioned on redshift with the relative $\chi^{2}$ values of template spectra.

### 2019-03-09 paper/tex/data.tex
data generation cleaned up

@@
 Given the SED, absolute $r$-band magnitude, and redshift, we computed apparent magnitudes in the six \lsst\ filter passbands, $ugrizy$.
-We assigned magnitude errors in the six bands using the simple model of \citet{Ivezic:08}, assuming achievement of the full 10-year depth.
-The number of total 30-second visits assumed when generating the photometric errors differs slightly from the fiducial numbers assumed for LSST: we assume 60 visits in u-band, 80 visits in g-band, 180 visits in r-band, 180 visits in i-band, 160 visits in z-band, and 160 visits in y-band.
+We assigned magnitude errors in the six bands using the simple model of \citet{Ivezic:08}, assuming achievement of the full 10-year depth, with a modification of fiducial \lsst\ total numbers of 30-second visits for photometric error generation: we assume 60 visits in $u$-band, 80 visits in $g$-band, 180 visits in $r$-band, 180 visits in $i$-band, 160 visits in $z$-band, and 160 visits in $y$-band.
 %numbers taken from Alex Abate's photErrorModel.py script in PhotoZDC1 repository, which was used with nYrObs=10.
-In the course of simulating Gaussian photometric errors, we add noise to objects fluxes, and some of these noisy fluxes will become negative in one or more bands.
-We call such negative fluxes ``non-detections'' and signify them with a placholder magnitude of 99.0 in the catalogue.
-Thus, further mentions of ``non-detections'' refer to objects that would be ``looked at but not seen'' in multi-band forced photometry, and the photo-z codes will treat them as such.
-Only 2.0 per cent of our sample galaxies contain a photometric band with a non-detection, the vast majority of which are in the $u$-band.
+
+As a consequence of adding Gaussian-distributed photometric errors, 2.0\% of our galaxies exhibit a negative flux in one or more bands, the vast majority of which are in the $u$-band.
+We deem such negative fluxes \textit{non-detections} and assign a placholder magnitude of 99.0 in the catalogue to indicate to the \pzpdf\ codes that such galaxies would be ``looked at but not seen'' in multi-band forced photometry.
 

### 2019-03-09 paper/tex/data.tex
finished template section

@@
 
-Our mock catalogue is derived from the \textsc{Buzzard-highres-v1.0}  of De Rose et al., in prep, Wechsler et al., in prep) catalogue.
+Our mock catalogue is derived from the \textsc{Buzzard}-highres-v1.0  of De Rose et al., in prep, Wechsler et al., in prep) catalogue.
 % UPDATE THIS CITATION NOW THAT THE PAPER IS OUT
@@
 The lightcone was constructed from smoothing and interpolation between a set of time snapshots.
-Dark matter halos were identified using the \textsc{Rockstar} software package \citep{Behroozi:13} and then populated with galaxies with a stellar mass and absolute $r$-band magnitude in the \sdss\ system determined using a sub-halo abundance matching model constrained to match both projected two-point galaxy clustering statistics and an observed conditional stellar mass function \citep{Reddick:13}.
+Dark matter halos were identified using the \texttt{Rockstar} software package \citep{Behroozi:13} and then populated with galaxies with a stellar mass and absolute $r$-band magnitude in the \sdss\ system determined using a sub-halo abundance matching model constrained to match both projected two-point galaxy clustering statistics and an observed conditional stellar mass function \citep{Reddick:13}.
 
-To assign an SED to each galaxy, the {\it Adding Density Dependent Spectral Energy Distributions} (\texttt{ADDSEDS}, deRose in prep.)\footnote{\url{https://github.com/vipasu/addseds}} procedure was used.
+To assign an SED to each galaxy, the Adding Density Dependent Spectral Energy Distributions (\texttt{ADDSEDS}, deRose in prep.)\footnote{\url{https://github.com/vipasu/addseds}} procedure was used.
 \texttt{ADDSEDS} uses a sample of $\sim 5\times 10^{5}$ galaxies from the magnitude-limited \sdss\ Data Release 6 Value Added Galaxy Catalogue \citep{Blanton:05} to train an empirical relation between absolute $r$-band magnitude, local galaxy density, and SED.
@@
 
+The full \textsc{Buzzard} dataset thus covers $400$ square degrees and contains $238$ million galaxies of redshift $0 < z \leq 8.7$ down to $r = 29$.
+Systematic inconsistencies with galaxy colors at $z > 2$ were observed, so the catalogue was limited to $0 < z \leq 2.0$.
+To obtain a catalogue matching the \lsst\ Gold Sample, we imposed an cut of $i < 25.3$, which gives a signal-to-noise ratio $\sim 30$ for most galaxies.
+In order for statistical errors to be subdominant to the systematic errors we aim to probe, we further reduced the sample size to $<10^{7}$ galaxies by isolating $\sim 16.8$ square degrees selected from five separate spatial regions of the simulation.
+
 \subsection{Experimental Conditions}
@@
 % NOW MOTIVATE THE SHARED PRIOR INFORMATION
+For the purpose of performing a controlled experiment that compares \pzpdf\ codes on equal footing as a baseline for a future sensitivity analysis, we take care to provide each with maximally optimistic prior information.
+Redshift estimation approaches built upon physical modeling and machine learning alike have a notion of prior information considered beyond the photometry of the data for which redshift is to be constrained: that information is derived from a template library for a model-based code and a training set for a data-driven code.
+In this initial study, we seek to set a baseline for a later comparison of the performance of \pzpdf\ codes under incomplete and non-representative prior information that will propagate differently in the space of data-driven and model-based algorithms.
+However, for the baseline case of perfect prior information, physical modeling and machine learning codes can indeed be put on truly equal footing.
+We outline the equivalent ways of providing all codes perfect prior information below.
 
-\subsubsection{Training set selection}
+\subsubsection{Training and test set division}
 \label{sec:buzztraining}
 
-The total initial catalogue covered $400$ square degrees and contained $238$ million galaxies to an apparent magnitude limit of $r\!=\!29$ and spanning the redshift range $0\!<\!z\!\leq\!8.7$.
-In order for statistical errors not to dominate, we need less than one million galaxies in our sample.
-Several studies claim that only a few tens of thousands of spectra are necessary to calibrate photo-z surveys to Stage IV requirements (e.~g.~\citet{Bernstein:10}, \citet{Masters:2017}).
-Therefore, we aim for a final number of training galaxies between $3\times 10^{4}$ and $5\times 10^{4}$ in our sample.
-%This catalogue contained two orders of magnitude more galaxies than were necessary to determine statistics for this study,
-In order to reduce our sample to a reasonable size, we limit our dataset to a subset of $\sim\!16.8$ square degrees selected from five separate spatial regions of the simulation.
-Systematic problems with galaxy colors above $z\!>\!2$ were observed, so the catalogue was limited to include only galaxies in the redshift range $0\!<\!z\!\leq\!2.0$.
-A random subset of the the remaining galaxies was chosen, and placed at random into either a ``training'' set ($10$ per cent of the sample), for which the galaxies true redshifts will be supplied, or a ``test'' set (the remaining $90$ per cent of the sample), for which each code will need to predict a redshift PDF for each galaxy.
+Following the findings of \citet{Bernstein:10}, \citet{Masters:2017} that only $~10^{4}$ spectra are necessary to calibrate \pz s to Stage IV requirements, we aimed to set aside a randomly selected training set of $3-5\times 10^{4}$ galaxies, $\sim 10\%$ of the full sample.
+After all cuts described above, we designated a training set of $44\,404$ galaxies for which both photometry and redshifts would be provided to data-driven codes and a blinded test set of $399\,356$ galaxies for which photometry alone would be provided to all codes and \pzpdf s would be requested.
 %The resulting catalogues contain $111\,171$ training galaxies and $1\,000\,883$ test galaxies.
-Finally, we restrict our analysis to a sample with an apparent i-band magnitude limit $i<25.3$, which give a signal-to-noise $\sim\,30$ for most galaxies, a cut often referred to as the expected ``LSST Gold Sample''.
-This magnitude cut results in a training set with $44\,404$ galaxies and a test set containing $399\,356$ galaxies.  All subsequent results will evaluate this ``gold sample'' test set.
-In order to blind results, initially redshifts were not revealed for the ``test'' set, and were only supplied for the training sample galaxies.
-This prevented code runners from tweaking results and fitting to the specific test set.
 
@@
 
-As mentioned in Section~\ref{sec:buzzard}, the SEDs in the Buzzard simulation are drawn from an empirical set of SEDs taken from the SDSS DR6 NYU-VAGC, a sample of roughly $\sim5\times 10^{5}$ galaxies with spectra in SDSS.
-To determine a finite set of templates to use with template fitting codes we take the five SED weight coefficients for each of the galaxies in the SDSS sample and run a simple K-means clustering algorithm on this five dimensional space.
-Each dimension was normalized such that it spanned an interval $[0,1]$.
-The K-means clusters partition the five-dimensional space of coefficients into Voronoi cells, spanning the space of coefficients in a way that properly reflects the underlying density in the coefficients.
-Thus, the resultant SEDs constructed using the cell centers as weight coefficients will provide a reasonable spanning SED set.
-An ad-hoc number of $K=100$ was chosen and the $100$ K-means centre positions are taken as the weights for the \textsc{k-correct} SED components to construct one hundred template SEDs.
-These $100$ templates were provided,
-%\red{(JS: if I remembered correctly, weren't there 150 templates given?
-Answer: no, initially I supplied a list of 150 that were drawn with a different algorithm, but once we switched to the final k-means set I only supplied 100 (the initial algorithm over-weighted outlier SEDs, which is why we needed the extra 50)--SJS)}
-and the templates were used by both \textsc{BPZ} and \textsc{LePhare}; however, because EAZY was designed and written to use the same five basis templates employed by \textsc{k-correct} when constructing our mock galaxies, EAZY was run using linear combinations of these five templates rather than using the 100 discrete templates.
-The ability to fit for linear combinations of templates highlights an important implementation difference between similar photo-z codes.
+We aimed to provide template-fitting codes with complete yet manageable library of templates spanning the space of SEDs of the \textsc{Buzzard} galaxies.
+We constructed $K=100$ representative templates from the $\sim 5 \times 10^{5}$ SEDs of the \sdss\ DR6 NYU-VAGC by using the five-dimensional vectors of SED weight coefficients described above.
+After regularizing the SED weight coefficients $\in [0, 1]$, we ran a simple K-means clustering algorithm on the five-dimensional space of regularized SED weight coefficients of the \sdss\ galaxy sample.
+The resulting clusters were used to define Voronoi cells in the space of weight coefficients, with centre positions corresponding to weights for the \texttt{k-correct} SED components, yielding 100 templates to be provided to all template-based codes.
+We did not, however, exclude from consideration template-based codes that made modifications in their use of these templates due to architecture limitations (as opposed to knowledge of the experimental conditions that could artificially boost the code's apparent performance), with deviations noted in Section~\ref{sec:pzcodes}.
 

### 2019-03-09 paper/tex/data.tex
polishing chapter

@@
 The experimental conditions shared among all codes are motivated by the \lsst\ SRD requirements and implemented for machine learning and template-based \pzpdf\ codes according to the procedures of Sections~\ref{sec:buzztraining} and \ref{sec:buzztemplates} respectively.
-Finally, in Section~\ref{sec:limitations}, we elaborate on the scope of the experimental conditions considered as a result of the sophistication of the mock data.
 
-\subsection{Buzzard-v1.0 simulation}
+\subsection{The \textsc{Buzzard-v1.0} simulation}
 \label{sec:buzzard}
@@
 For each simulated galaxy, a galaxy with similar absolute $r$-band magnitude and local galaxy density was chosen from the training set, and that training galaxy's SED was assigned to the simulated galaxy.
+In Section~\ref{sec:buzzlimitations}, we critique the realism of this mock data.
+
+\subsubsection{Caveats}
+\label{sec:buzzlimitations}
+
+By necessity, \textsc{Buzzard} does not contain all of the complicating factors present in real data, and here we discuss the most pertinent ways that this limitation affects our experiment.
+% In several cases, the simplification is done with a purpose, with potentially confounding effects excluded in order to better isolate the differences between current-generation photo-$z$ codes and their causes.
+\textsc{Buzzard} includes only galaxies, not stars of AGN.
+The catalogue-based construction excludes image-level effects, such as deblending errors, photometric measurement issues, contamination from sky background (Zodiacal light, scattered light, etc.), lensing magnification, and Galactic reddening.
+
+The SEDs are five-component linear combinations of $\sim 5 \times 10^{5}$ \sdss\ galaxies, so the sample contains only galaxies that resemble linear combinations of those for which \sdss\ obtained spectra.
+The linear combination SEDs also restrict the properties of the galaxy population to linear combinations of the properties corresponding to five basis templates, precluding the modeling of non-linear features such as the full range of emission line fluxes relative to the continuum.
+The only form of intrinsic dust reddening comes from what is already present in the five basis SEDs via the training set used to create the basis templates, and linear combinations thereof do not span the full range of realistic dust extinction observed in galaxy populations.
+
+While these idealized conditions limit the realism of our mock data, they are irrelevant to the controlled experimental conditions of this study, if anything assuring that differentiation in the performance of the \pzpdf\ codes is due to the inferential techniques rather than nuances in the data.
+
+\subsection{\lsst-like mock observations}
+\label{sec:observations}
 
@@
 
-The full \textsc{Buzzard} dataset thus covers $400$ square degrees and contains $238$ million galaxies of redshift $0 < z \leq 8.7$ down to $r = 29$.
+The full dataset thus covers $400$ square degrees and contains $238$ million galaxies of redshift $0 < z \leq 8.7$ down to $r = 29$.
 Systematic inconsistencies with galaxy colors at $z > 2$ were observed, so the catalogue was limited to $0 < z \leq 2.0$.
@@
 In order for statistical errors to be subdominant to the systematic errors we aim to probe, we further reduced the sample size to $<10^{7}$ galaxies by isolating $\sim 16.8$ square degrees selected from five separate spatial regions of the simulation.
+We refer to this final set of galaxies as DC1, for the first \lsstdesc\ Data Challenge.
 
-\subsection{Experimental Conditions}
+\subsection{Shared prior information}
 \label{sec:controlled}
@@
 
-We aimed to provide template-fitting codes with complete yet manageable library of templates spanning the space of SEDs of the \textsc{Buzzard} galaxies.
+We aimed to provide template-fitting codes with complete yet manageable library of templates spanning the space of SEDs of the DC1 galaxies.
 We constructed $K=100$ representative templates from the $\sim 5 \times 10^{5}$ SEDs of the \sdss\ DR6 NYU-VAGC by using the five-dimensional vectors of SED weight coefficients described above.
@@
 We did not, however, exclude from consideration template-based codes that made modifications in their use of these templates due to architecture limitations (as opposed to knowledge of the experimental conditions that could artificially boost the code's apparent performance), with deviations noted in Section~\ref{sec:pzcodes}.
-
-\subsection{Limitations}
-\label{sec:buzzlimitations}
-
-For our initial investigation of photometric redshift codes, we begin with a data set that is somewhat idealized, and does not contain all of the complicating factors present in real data.
-In several cases, the simplification is done with a purpose, with potentially confounding effects excluded in order to better isolate the differences between current-generation photo-$z$ codes and their causes.
-We list several of the simulation limitations in this section.
-As the simulation is catalogue-based, no image level effects, such as photometric measurement effects, object blending, contamination from sky background (Zodiacal light, scattered light, etc...), lensing magnification, or Galactic reddening are included.
-No stars are included in the catalogue, nor are the effects of AGN.
-As all SEDs are constructed from only five basis templates, properties of the galaxy population will be restricted to follow linear combinations of the characteristics of the five basis templates, so certain non-linear features, for example the full range of emission line fluxes relative to the continuum, will not be included in the model galaxy population.
-Moreover, the linear combinations of templates are modeled on the $\sim5\times 10^{5}$ SDSS galaxies discussed in Section~\ref{sec:buzzard}, and thus only galaxies that resemble those spectroscopically observed by the SDSS will be included in the sample.
-No additional dust reddening intrinsic to the host galaxy is included, the only approximation of dust extinction comes in the form of dust encoded in the five basis SEDs via the training set used to create the basis templates.
-Simple linear combinations of these basis templates will, once again, not explore the full range of realistic dust extinction observed in galaxy populations.
-While these idealized conditions limit the realism of our galaxy population, some are also by design.
-We aim to test the photo-z codes at a very basic level, and a simplified model assures that differences in results seen between the codes are due to fundamental differences in their underlying assumptions and implementation details, rather than more nuanced properties.

### 2019-03-10 paper/tex/methods.tex
working through BPZ section

@@
 \label{sec:pzcodes}
-%(Sam Schmidt, Rongpu Zhou, Ibrahim Almosallam, Eric Nuss, Johann Cohen Tanugi, John Soo)
 
-Here we outline the photo-$z$ PDF codes tested in this study.
-In total, twelve distinct codes are tested.
-This sample is not comprehensive, codes were chosen based on the expertise available within the group; however, those chosen do cover a broad range of the current-generation methods used in the field.
-Both template-based and machine learning approaches are included and each are described separately in Secs.~\ref{sec:templatecodes} and \ref{sec:trainingcodes} respectively.
-The list of codes are summarized in Table.~\ref{tab:list_of_codes}.
-All code runners were asked to output redshift posterior estimates on 200 linear-spaced bins between redshifts 0 and 2.
+Here we summarize the twelve \pzpdf\ codes compared in this study, summarized in Table~\ref{tab:list_of_codes}, which include both established and emerging approaches in template fitting and machine learning.
+Though not exhaustive, this sample represents codes for which there was sufficient expertise within the \lsstdesc\ Photometric Redshifts Working Group; the authors welcome interest from those outside \lsstdesc\ to have their codes assessed in future investigations that build upon this one.
 
 \begin{table*}  %%% DATA TABLE %%%
-\caption{List of photo-$z$ codes featured in this study. ML here means machine learning.} \label{tab:list_of_codes}\resizebox{\textwidth}{!}{
-\begin{tabular}{llll}
+\caption{List of \pzpdf\ codes featured in this study} \label{tab:list_of_codes}\resizebox{\textwidth}{!}{
+\begin{tabular}{lll}
 \hline
-\bf Code 			& \bf Type & \bf Paper & \bf Website \\
+\bf Published code & \bf Type & \bf Public source code \\
 \hline
-\textsc{BPz} 		& template 			& \citet{Benitez:00}		& \url{http://www.stsci.edu/~dcoe/BPZ/} \\
-\textsc{EAZY} 		& template 			& \citet{Brammer:08}		& \url{https://github.com/gbrammer/eazy-photoz} \\
-\textsc{LePhare} 	& template 			& \citet{Arnouts:99}		& \url{http://www.cfht.hawaii.edu/~arnouts/lephare.html} \\
-\textsc{ANNz2} 		& ML 	& \citet{Sadeh:16}			& \url{https://github.com/IftachSadeh/ANNZ} \\
-\textsc{Delight} 	& ML/template 	& \citet{Leistedt:17}		& \url{https://github.com/ixkael/Delight} \\
-\multirow{ 2}{*}{\textsc{FlexZBoost}} & \multirow{ 2}{*}{ML} 	& \multirow{ 2}{*}{\citet{Izbicki:17}}		& \url{https://github.com/tpospisi/flexcode};\\
-					& 		&							& \url{https://github.com/rizbicki/FlexCoDE}\\
+\bpz~\citep{Benitez:00} 		   & template fitting	& \url{http://www.stsci.edu/~dcoe/BPZ/} \\
+\eazy~\citep{Brammer:08}		   & template fitting & \url{https://github.com/gbrammer/eazy-photoz} \\
+\lephare~\citep{Arnouts:99}	   & template fitting	& \url{http://www.cfht.hawaii.edu/~arnouts/lephare.html} \\
+\annz~\citep{Sadeh:16}		     & machine learning	& \url{https://github.com/IftachSadeh/ANNZ} \\
+\delight~\citep{Leistedt:17}   & machine learning & \url{https://github.com/ixkael/Delight} \\
+\flexzboost~\citep{Izbicki:17} & machine learning & \url{https://github.com/tpospisi/flexcode}; \url{https://github.com/rizbicki/FlexCoDE}\\
 %\textsc{Frankenz} 	& ML 	& \citet{Speagle:frankenz}	& \url{https://github.com/joshspeagle/frankenz} \\
-\textsc{GPz} 		& ML	& \citet{Almosallam:15b}	& \url{https://github.com/OxfordML/GPz} \\
-\textsc{METAPhoR} 	& ML 	& \citet{Cavuoti:17}		& \url{http://dame.dsf.unina.it}\\
-\textsc{CMNN} 		& ML 	& \citet{Graham:17}			& - \\
-\textsc{SkyNet} 	& ML 	& \citet{Graff:14}			& \url{http://ccpforge.cse.rl.ac.uk/gf/project/skynet/} \\
-\textsc{TPZ} 		& ML 	& \citet{Carrasco_Kind:13}	& \url{https://github.com/mgckind/MLZ} \\
+\gpz~\citep{Almosallam:15b}	   & machine learning	& \url{https://github.com/OxfordML/GPz} \\
+\metaphor~\citep{Cavuoti:17}   & machine learning	& \url{http://dame.dsf.unina.it}\\
+\cmnn~\citep{Graham:17}        & machine learning & N/A \\
+\skynet~\citep{Graff:14}       & machine learning & \url{http://ccpforge.cse.rl.ac.uk/gf/project/skynet/} \\
+\tpz~\citep{Carrasco_Kind:13}	 & machine learning	& \url{https://github.com/mgckind/MLZ} \\
 \hline
-\trainz 	& N/A	& See Section~\ref{sec:method:trainz} & \\
+\trainz 	                             & machine learning	& See Section~\ref{sec:method:trainz} \\
 \end{tabular}}
@@
 
-%\red{The questions that must be answered for each code are\dots}
-The questions that must be answered for each code are: what unique features are included in the specific implementation that influence the output $p(z)$.
-What form of validation was performed with the training data, how were photometric uncertainties employed in the analysis, how were negative fluxes treated, what specific prior form was employed (for template based codes), or what specific machine learning architecture was used (for ML codes)?
-%\scc{which parameter (in a general meaning) have been used, which kind of validation has been performed (if any)}
+We describe the algorithms and implementations of the model-based and data-driven codes in Sections~\ref{sec:templatecodes} and \ref{sec:trainingcodes} respectively.
+For each approach, we note how it accounts for photometric uncertainties, how it interprets negative fluxes, what its native output format of \pzpdf s is, and what, if any, preprocessing, including validation, of the prior information it performs.
+
 \subsection{Template-based Approaches}
 \label{sec:templatecodes}
-We test three publicly available and commonly used template-based codes: \textsc{BPZ}, \textsc{EAZY}, and \textsc{LePhare}.
-All three codes follow the standard procedure for template-based redshift estimation: calculate model fluxes for a set of template spectral energy distributions (SEDs) on a grid of redshift values and calculating a $\chi^2$ merit function using the observed and model fluxes:
-\begin{equation} \label{eq_temp_chi}
-\chi^2(z,T,A) = \sum_i^{N_{\rm filt}}\left(\frac{F^i_{\rm{obs}}−A \times F^i_{\rm{pred}}(T,z)}{\sigma^i_{\rm{obs}}}\right)^2
+
+We test three publicly available and commonly used template-based codes, \bpz, \eazy, and \lephare, that share the standard physically motivated approach of calculating model fluxes for a set of template SEDs on a grid of redshift values and evaluating a merit function using the observed and model fluxes.
+These three also share the $\chi^{2}$ merit function
+\begin{equation}
+	\label{eq_temp_chi}
+	\chi^{2}(z, T, A) \equiv \sum_{b}^{N_{\rm{filt}}} \left(\frac{F^{\rm{obs}}_{b} - A F^{\rm{mod}}_{b}(T, z)}{\sigma^{\rm{obs}}_{b}}\right)^{2},
 \end{equation}
-\noindent  where $A$ is a normalization factor, $F^i_{\rm{pred}}(T,z)$ is the flux predicted for a template $T$ at redshift $z$.
-$F^i_{\rm{obs}}$ is the observed flux in a given band $i$ and $\sigma^i_{\rm{obs}}$ is the observed flux error.
-$N_{\rm tot}$ is the total number of filters, in our case the six $ugrizy$ LSST filters.
-Specific implementation details of each code, e.~g.~prior form and implementation, are described below.
+of normalization factor $A$, template $T$, and redshift $z$, which is a sum of observed flux error $\sigma_{b}^{\rm{obs}}$-weighted squared differences between the observed flux $F^{\rm{obs}}_{b}$ and the normalized predicted flux $F^{\rm{mod}}_{b}(T, z)$ in $N_{\rm{filt}}$ photometric filters $b$, which in our case is the \lsst\ $ugrizy$ filters.
 
@@
 
-\textsc{BPZ}\footnote{\url{http://www.stsci.edu/~dcoe/BPZ/}} \citep[Bayesian Photometric Redshift,][]{Benitez:00} is a template-based photo-$z$ code that compares the expected colors ($C$) calculated for a set of spectral energy distribution (SED) types/templates ($T$) to the observed colors to calculate the likelihood of observing colors at each redshift for each type, $p(C|z,T)$.
-The likelihoods at each redshift are related to the $\chi^2$ in Equation~\ref{eq_temp_chi} by the simple form: likelihood $\propto \, e^{-\chi^2/2}$.
-The code employs an empirically determined Bayesian prior in apparent magnitude ($m_0$) and SED-type.
-Assuming that the SED-types are spanning and exclusive, we can determine the redshift posterior $p(z|C,m_0)$ by marginalizing over all SED-types with a simple sum \citep[Eq.~3 from][]{Benitez:00}:
+\bpz\footnote{\url{http://www.stsci.edu/~dcoe/BPZ/}} \citep[Bayesian Photometric Redshift,][]{Benitez:00} is a template-based \pz\ code that determines the likelihood $p(C \mid z, T)$ of a galaxy's observed colours $C$ for a set of SED templates $T$ at redshifts $z$.
+The \bpz\ likelihood is related to Equation~\ref{eq_temp_chi} by $p(C \mid z, T) \propto \exp[- \chi^{2} / 2]$.
+Given a Bayesian prior $p(z, T \mid m_{0})$ over apparent magnitude $m_0$ and assuming that the SED templates are spanning and exclusive, \bpz\ constructs the redshift posterior $p(z \mid C, m_0)$ by marginalizing over all SED templates as in \citep[Eq.~3 from][]{Benitez:00}, corresponding to setting the parameter \texttt{PROBS\_LITE=TRUE} in the \bpz\ parameter file.
+The \bpz\ prior is the product of an SED template proportion that varies with apparent magnitude $p(T \mid m_{0})$ and a prior $p(z \mid T, m_{0})$ over the expected redshift as a function of apparent magnitude and SED template.
 
-\begin{equation} \label{eq:redshift_posterior}
-p(z|C,m_0)\,\propto\, \sum_{T}p(z,T|m_0)\,p(C|z,T)
-\end{equation}
-\noindent where the first term on the right-hand side is the Bayesian prior and the second term is the traditional likelihood.
-The prior is assumed to have the form: $p(z,T|m_0)=p(T|m_0)\,p(z|T,m_0)$, i.e.
-it parameterizes the prior as an evolving type fraction with apparent magnitude, combined with a prior on the expected redshift probability distribution as a function of both apparent magnitude and SED-type.
+Here we test \bpz-v 1.99.3 with the DC1 template set of Section~\ref{sec:buzztemplates}.
+To keep the number of free parameters manageable, the DC1 template set is pre-sorted by the rest-frame $u$-$g$ colour and split into three broad classes of SED template, equivalent to the E, Sp and Im/SB types in .
+The Bayesian prior term $p(T \mid m_{0})$ was derived directly from the DC1 training set, and the other term $p(z \mid T, m_{0})$ was chosen to be the best fit for the eleven free parameters of the functional form of \citet{Benitez:00}.
+%For photo-$z$ point estimates we use the \texttt{Z\_B} output parameter.
+Prior to running the code, the non-detection placeholder magnitude was replaced with an estimate of the one-$\sigma$ detection limit for the undetected band as a proxy for a value close to the estimated sky noise threshold.
 
-In this paper we use \textsc{BPZ v 1.99.3}. The template set employed here is the set of $100$ discrete SEDs described in Section~\ref{sec:buzztemplates}
-%, and the $129$ Brown SEDs for the Galacticus dataset.
-To keep the number of free parameters to a manageable level the SEDs in the training set are sorted by the rest-frame $u$-$g$ colour and split into three ``broad'' SED classes, equivalent to the \texttt{E}, \texttt{Sp} and \texttt{Im/SB} types in \citet{Benitez:00}.
-We assume the same functional form for the Bayesian priors as used by \citet{Benitez:00}, and utilize the training-set galaxies with known SED-type, redshift, and apparent magnitude to determine the type fractions and the best fit for the eleven free parameters of the prior.
-%For photo-$z$ point estimates we use the \texttt{Z\_B} output parameter.
-For galaxies that are not detected in a measured band, the placeholder value is replaced with an estimate of the one $\sigma$ detection limit in that particular band, i.~e.~a value close to the estimated sky noise threshold.
-The type-marginalized $p(z)$ is generated by setting the parameter \texttt{PROBS\_LITE=TRUE} in the \textsc{BPZ} parameter file.
-
-\subsubsection{EAZY} \label{sec:eazy}
+\subsubsection{EAZY}
+\label{sec:eazy}
 %(Rongpu Zhou)
@@
 In this paper we use the all-templates mode, which fits the photometric data with a linear combination of the five basis templates.
+The EAZY architecture cannot accept a template set other than the same five basis templates employed by \texttt{k-correct} when constructing our mock galaxies.
+EAZY was thus run using linear combinations of these five templates rather than using the 100 discrete templates shared by the other template-fitting codes.
 We employed the $5$ basis templates described in Section~\ref{sec:buzzard}, and set the template error to zero since these same templates were used to produce the simulated catalog photometry.
 %\red{The quality parameter \texttt{Qz} is...}
+The ability to fit for linear combinations of templates highlights an important implementation difference between similar photo-z codes.
 The likelihoods include the application of a type-independent apparent magnitude prior estimated from the training data.
@@
 \label{sec:trainingcodes}
+
 The training-based codes use a variety of algorithms in order to estimate $p(z)$, specifics of each implementation are described in the subsections.
@@
 
-In this study, \textsc{ANNz2 v.
-2.0.4} was used.
+In this study, \textsc{ANNz2 v. 2.0.4} was used.
 A set of $5$ ANNs with architecture $6:12:12:1$ ($6$ $ugrizy$ inputs, $2$ hidden layers with $12$ nodes each, and $1$ output) with different random seeds are used during each training.
@@
 For each training-target pair, $p(\mathbf{\hat{F}}|z,t_i)$ is evaluated as follows:
-
 \begin{equation} \label{eq:delight_noisy}
@@
 
-\textsc{FlexZBoost}\footnote{\url{https://github.com/tpospisi/flexcode};\\ \url{https://github.com/rizbicki/FlexCoDE}} \citep{Izbicki:17} is a particular realization of \texttt{FlexCode}, which is a general-purpose methodology for converting any conditional mean point estimator of $z$ to a conditional {\em density} estimator $f(z \vert \x)$, where $\x$ here represents our photometric covariates and errors.\footnote{Instead of $p(z)$, we use the notation $f(z \vert \x)$ to explicitly show the dependence on $\x$.} The key idea is to expand the unknown function $f(z \vert \x)$ in an orthonormal basis $\{\phi_i(z)\}_{i}$:
-
+\textsc{FlexZBoost}\footnote{\url{https://github.com/tpospisi/flexcode};\\ \url{https://github.com/rizbicki/FlexCoDE}} \citep{Izbicki:17} is a particular realization of \texttt{FlexCode}, which is a general-purpose methodology for converting any conditional mean point estimator of $z$ to a conditional {\em density} estimator $f(z \vert \x)$, where $\x$ here represents our photometric covariates and errors.\footnote{Instead of $p(z)$, we use the notation $f(z \vert \x)$ to explicitly show the dependence on $\x$.}
+The key idea is to expand the unknown function $f(z \vert \x)$ in an orthonormal basis $\{\phi_i(z)\}_{i}$:
 \begin{equation}
@@
 By the orthogonality property, the expansion coefficients are just conditional means
-
 \begin{equation}
@@
 The missing values are imputed by computing the mean of the training set $\mu$ and its covariance $\Sigma$, then we use the following equation to estimate the missing values from the observed ones %\[
-
 \begin{equation}
@@
 \end{equation}  %\]
-
 where the subscript $o$ in $x_{o}$ indexes the \emph{observed} part of the input $x$, whereas the subscript $u$ indexes the \emph{unobserved} set (similarly for $\mu$ and $\Sigma$).
@@
 \end{equation}
-\noindent  where $\alpha_{i}$ is a user selected multiplicative constant (useful in case of multi-survey photometry), $u_{\mu=0,\sigma=1}$ is a random value from the standard normal distribution and $F_{ij}$ is a bimodal function (a constant function + polynomial fitting of the mean magnitude errors on the binned bands), heuristically tuned in such a way that the constant component is the threshold under which the polynomial function is considered too low to provide a significant noise contribution to the photometry perturbation.
+where $\alpha_{i}$ is a user selected multiplicative constant (useful in case of multi-survey photometry), $u_{\mu=0,\sigma=1}$ is a random value from the standard normal distribution and $F_{ij}$ is a bimodal function (a constant function + polynomial fitting of the mean magnitude errors on the binned bands), heuristically tuned in such a way that the constant component is the threshold under which the polynomial function is considered too low to provide a significant noise contribution to the photometry perturbation.
 
@@
 For the present work since it was required to produce a redshift (and a PDF) for each object of the test set we decided to apply a hierarchical kNN to replace the missing detections with values based on their neighbors.
-The reliability of PDFs and point estimation is lower.
-No cross validation has been used.
+The reliability of PDFs and point estimation is lower. No cross validation has been used.
 
@@
 \subsection{Simple Ensemble Estimator}\label{sec:method:trainz}
+
 In addition to the main photo-$z$ algorithms described above we also include a very simple method as a pathological example.
 For \trainz, as we will we call this simple estimator, we well define $p(z)$ as simply:
-
 \begin{equation}

### 2019-03-10 paper/tex/introduction.tex
new macros

@@
 
-The current and next generations of large-scale galaxy surveys, including the Dark Energy Survey \citep[\textsc{DES},][]{Abbott:05}, the Kilo-Degree Survey \citep[\textsc{KiDS},][]{de_Jong:13}, Hyper Suprime-Cam Survey \citep[\textsc{HSC},][]{Aihara:2018a,Aihara:2018b}, Large Synoptic Survey Telescope \citep[\lsst,][]{Abell:09}, Euclid \citep{Laureijs:11}, and Wide-Field Infrared Survey Telescope \citep[\textsc{WFIRST},][]{Green:12}, present a paradigm shift from spectroscopic to photometric galaxy catalogues of substantially larger size at a cost of lacking complete spectroscopically confirmed redshifts.
+The current and next generations of large-scale galaxy surveys, including the Dark Energy Survey \citep[\proj{DES},][]{Abbott:05}, the Kilo-Degree Survey \citep[\proj{KiDS},][]{de_Jong:13}, Hyper Suprime-Cam Survey \citep[\proj{HSC},][]{Aihara:2018a,Aihara:2018b}, Large Synoptic Survey Telescope \citep[\lsst,][]{Abell:09}, Euclid \citep{Laureijs:11}, and Wide-Field Infrared Survey Telescope \citep[\proj{WFIRST},][]{Green:12}, present a paradigm shift from spectroscopic to photometric galaxy catalogues of substantially larger size at a cost of lacking complete spectroscopically confirmed redshifts ($z$).
 
@@
 
-To fully characterize such physical degeneracies, photometric galaxy catalogue data releases from \citep{Mandelbaum:2008} to \citep{de_Jong:17}, provided a more informative \pz\ data product, the \pz\ probability density function (PDF), that describes the relative probability as a function of a galaxy's redshift, conditioned on the observed photometry.
+To fully characterize such physical degeneracies, photometric galaxy catalogue data releases from \citep{Mandelbaum:2008} to \citep{de_Jong:17}, provided a more informative \pz\ data product, the \pz\ probability density function (PDF), that describes the redshift probability $p(z)$ as a function of a galaxy's redshift, conditioned on the observed photometry.
 Early template-based methods such as \citet{Fernandezsoto:99} approximated the likelihood of photometry conditioned on redshift with the relative $\chi^{2}$ values of template spectra.

### 2019-03-10 paper/tex/data.tex
consistency with methods

@@
 
-To assign an SED to each galaxy, the Adding Density Dependent Spectral Energy Distributions (\texttt{ADDSEDS}, deRose in prep.)\footnote{\url{https://github.com/vipasu/addseds}} procedure was used.
+To assign a spectrum to each galaxy, the Adding Density Dependent Spectral Energy Distributions (SEDs) procedure (\texttt{ADDSEDS}, deRose in prep.)\footnote{\url{https://github.com/vipasu/addseds}} was used.
 \texttt{ADDSEDS} uses a sample of $\sim 5\times 10^{5}$ galaxies from the magnitude-limited \sdss\ Data Release 6 Value Added Galaxy Catalogue \citep{Blanton:05} to train an empirical relation between absolute $r$-band magnitude, local galaxy density, and SED.
@@
 Following the findings of \citet{Bernstein:10}, \citet{Masters:2017} that only $~10^{4}$ spectra are necessary to calibrate \pz s to Stage IV requirements, we aimed to set aside a randomly selected training set of $3-5\times 10^{4}$ galaxies, $\sim 10\%$ of the full sample.
-After all cuts described above, we designated a training set of $44\,404$ galaxies for which both photometry and redshifts would be provided to data-driven codes and a blinded test set of $399\,356$ galaxies for which photometry alone would be provided to all codes and \pzpdf s would be requested.
+After all cuts described above, we designated the \textit{DC1 training set} of $44\,404$ galaxies for which observed photometry, true SEDs, and true redshifts would be provided to all codes codes and the blinded \textit{DC1 test set} of $399\,356$ galaxies for which photometry alone would be provided to all codes and \pzpdf s would be requested.
 %The resulting catalogues contain $111\,171$ training galaxies and $1\,000\,883$ test galaxies.
@@
 After regularizing the SED weight coefficients $\in [0, 1]$, we ran a simple K-means clustering algorithm on the five-dimensional space of regularized SED weight coefficients of the \sdss\ galaxy sample.
-The resulting clusters were used to define Voronoi cells in the space of weight coefficients, with centre positions corresponding to weights for the \texttt{k-correct} SED components, yielding 100 templates to be provided to all template-based codes.
+The resulting clusters were used to define Voronoi cells in the space of weight coefficients, with centre positions corresponding to weights for the \texttt{k-correct} SED components, yielding the 100 \textit{DC1 template set} to be provided to all template-based codes.
 We did not, however, exclude from consideration template-based codes that made modifications in their use of these templates due to architecture limitations (as opposed to knowledge of the experimental conditions that could artificially boost the code's apparent performance), with deviations noted in Section~\ref{sec:pzcodes}.

### 2019-03-10 paper/main.tex
clearer title
new macros

@@
 % Add your own macros here:
+\newcommand{\textul}{\underline}
 
-\newcommand{\lsst}{\textsc{LSST}}
-\newcommand{\lsstdesc}{\lsst\textsc{-DESC}}
+% projects
+\newcommand{\proj}[1]{\textsc{#1}}
+\newcommand{\lsst}{\proj{LSST}}
+\newcommand{\lsstdesc}{\lsst-\proj{DESC}}
+\newcommand{\sdss}{\proj{SDSS}}
+\newcommand{\buzz}{\proj{Buzzard}}
+
+% can easily change these to specify posterior throughout, etc
 \newcommand{\pz}{photo-$z$}
-\newcommand{\pzpdf}{\pz\ PDF}% can easily change this to specify posterior throughout
-\newcommand{\qp}{\texttt{qp}}
+\newcommand{\pzpdf}{\pz\ PDF}
+\newcommand{\Pz}{Photo-$z$}
+\newcommand{\Pzpdf}{\Pz\ PDF}
+\newcommand{\nz}{$N(z)$}
 
-\newcommand{\trainz}{\textsc{trainZ}}
+% codes
+\newcommand{\pzcode}[1]{\texttt{#1}}
+\newcommand{\qp}{\pzcode{qp}}
+\newcommand{\annz}{\pzcode{ANNz2}}
+\newcommand{\bpz}{\pzcode{BPz}}
+\newcommand{\cmnn}{\pzcode{CMNN}}
+\newcommand{\delight}{\pzcode{Delight}}
+\newcommand{\eazy}{\pzcode{EAZY}}
+\newcommand{\flexzboost}{\pzcode{FlexZBoost}}
+\newcommand{\gpz}{\pzcode{GPz}}
+\newcommand{\lephare}{\pzcode{LePhare}}
+\newcommand{\metaphor}{\pzcode{METAPhoR}}
+\newcommand{\skynet}{\pzcode{SkyNet}}
+\newcommand{\tpz}{\pzcode{TPZ}}
+\newcommand{\trainz}{\pzcode{trainZ}}
 
-\newcommand{\textul}{\underline}
 \newcommand*\mathinhead[2]{\texorpdfstring{$\boldsymbol{#1}$}{#2}}
@@
 \newcommand{\erfan}[1]{\textcolor{newpink}{#1}} % Erfan Nourbakhsh comments in pink
-\newcommand{\jan}[1]{\textcolor{orange}{#1}}%Alex Malz comments in green
+\newcommand{\jan}[1]{\textcolor{orange}{#1}}%Jeff Newman comments in Orange
 
@@
 
-\title{Evaluating photometric redshift PDF approaches for \textsc{LSST}}
+\title{Evaluating approaches to deriving photometric redshift posterior probability distribution functions for \textsc{LSST}}
 
@@
 { \red{personal funding sources}}
-S.~Schmidt acknowledges support from DOE grant DE-SC0009999 and NSF/AURA grant N56981C.  AIM is advised by David W. Hogg and was supported by National Science Foundation grant AST-1517237.
+S.~Schmidt acknowledges support from DOE grant DE-SC0009999 and NSF/AURA grant N56981C.
+AIM is advised by David W. Hogg and was supported by National Science Foundation grant AST-1517237.
 
-In addition to packages cited in the text, analyses performed in this paper used the following software packages: \textsc{Numpy} and \textsc{Scipy} \citep{numpyscipy}, \textsc{Matplotlib} \citep{matplotlib}, \textsc{Seaborn} \citep{seaborn}, \textsc{minFunc} \citep{minfunc}, \textsc{pySkyNet} \citep{pyskynet}, and {\textsc photUtils} from the LSST simulations package \citep{lsstphotutils}.
+In addition to packages cited in the text, analyses performed in this paper used the following software packages: \textsc{Numpy} and \textsc{Scipy} \citep{numpyscipy}, \textsc{Matplotlib} \citep{matplotlib}, \textsc{Seaborn} \citep{seaborn}, \textsc{minFunc} \citep{minfunc}, \textsc{pySkyNet} \citep{pyskynet}, and {\textsc photUtils} from the \lsst\ simulations package \citep{lsstphotutils}.
 

### 2019-03-10 paper/tex/methods.tex
editing EAZY

@@
 We describe the algorithms and implementations of the model-based and data-driven codes in Sections~\ref{sec:templatecodes} and \ref{sec:trainingcodes} respectively.
-For each approach, we note how it accounts for photometric uncertainties, how it interprets negative fluxes, what its native output format of \pzpdf s is, and what, if any, preprocessing, including validation, of the prior information it performs.
+% For each approach, we note how it accounts for photometric uncertainties, how it interprets negative fluxes, what its native output format of \pzpdf s is, and what, if any, preprocessing, including validation, of the prior information it performs.
 
@@
 We test three publicly available and commonly used template-based codes, \bpz, \eazy, and \lephare, that share the standard physically motivated approach of calculating model fluxes for a set of template SEDs on a grid of redshift values and evaluating a merit function using the observed and model fluxes.
-These three also share the $\chi^{2}$ merit function
+These three also share the \chisq\ merit function
 \begin{equation}
@@
 
-\textsc{EAZY}\footnote{\url{https://github.com/gbrammer/eazy-photoz}} \citep[Easy and Accurate Photometric Redshifts from Yale,][]{Brammer:08} is a template-based photo-$z$ code that includes several features that extend the basic $\chi^2$ fit used in many template codes.
-The code can fit the observed photometry with SEDs created from a linear combination of a set of templates at each redshift, and the best-fit SED is found by simultaneously fitting one, two or all of the templates by minimizing $\chi^2$.
-The minimized $\chi^2(z)$ is then combined with an apparent magnitude prior to obtain the posterior redshift probability distribution.
-On examination of the source code for EAZY, it appears that rather than marginalizing across all templates in the $\chi^2$ calculation, \textsc{EAZY} takes only the minimum value of $\chi^2$ at each redshift.
-This improper marginalization does not lead to the correct posterior distribution, an implementation issue that will need to be addressed in the future.
-%, although some argue that this is not the mathematically correct way of calculating the posteriors.
-\textsc{EAZY} can also account for the uncertainties in the templates by adding an empirically derived template error in quadrature as a function of redshift to the flux errors.
+\eazy\footnote{\url{https://github.com/gbrammer/eazy-photoz}} \citep[Easy and Accurate Photometric Redshifts from Yale,][]{Brammer:08} extends the basic \chisq\ fit procedure that defines template-fitting approaches.
+The algorithm models the observed photometry with a linear combination of template SEDs at each redshift.
+The best-fit SED is found by simultaneously fitting one, two, or all of the templates via \chisq\ minimization, which is distinct from marginalizing across all templates.
+The minimized \chisq\ likelihood at each redshift is then combined with an apparent magnitude prior to obtain the redshift posterior PDF.
+We note that the utilization of the best-fit SED rather than a proper marginalization does not lead to the correct posterior distribution, an implementation issue that has now been identified and will be addressed by the developers in the future.
+\eazy\ can account for uncertainty in the template set by adding in quadrature to the flux errors an empirically derived template error as a function of redshift.
 
-In this paper we use the all-templates mode, which fits the photometric data with a linear combination of the five basis templates.
-The EAZY architecture cannot accept a template set other than the same five basis templates employed by \texttt{k-correct} when constructing our mock galaxies.
-EAZY was thus run using linear combinations of these five templates rather than using the 100 discrete templates shared by the other template-fitting codes.
-We employed the $5$ basis templates described in Section~\ref{sec:buzzard}, and set the template error to zero since these same templates were used to produce the simulated catalog photometry.
-%\red{The quality parameter \texttt{Qz} is...}
-The ability to fit for linear combinations of templates highlights an important implementation difference between similar photo-z codes.
-The likelihoods include the application of a type-independent apparent magnitude prior estimated from the training data.
+The SED-independent apparent magnitude prior was derived empirically from the DC1 training set.
+The \eazy\ architecture cannot accept a template set other than the same five basis templates employed by \texttt{k-correct} when constructing the DC1 catalogue.
+However, \eazy\ does feature a flexible \texttt{all-templates} mode, which fits the photometric data with a linear combination of the five basis templates.
+We set the template error to zero since the same templates were in fact used to produce the DC1 photometry.
 
-\subsubsection{LePhare}\label{sec:lephare}
+\subsubsection{LePhare}
+\label{sec:lephare}
 %(C\'ecile Roucelle, Eric Nuss, Johann Cohen-Tanugi)

### 2019-03-10 paper/tex/introduction.tex
macros

@@
 To fully characterize such physical degeneracies, photometric galaxy catalogue data releases from \citep{Mandelbaum:2008} to \citep{de_Jong:17}, provided a more informative \pz\ data product, the \pz\ probability density function (PDF), that describes the redshift probability $p(z)$ as a function of a galaxy's redshift, conditioned on the observed photometry.
-Early template-based methods such as \citet{Fernandezsoto:99} approximated the likelihood of photometry conditioned on redshift with the relative $\chi^{2}$ values of template spectra.
+Early template-based methods such as \citet{Fernandezsoto:99} approximated the likelihood of photometry conditioned on redshift with the relative \chisq\ values of template spectra.
 Not long after, Bayesian adaptations of template-based approaches such as \citet{Benitez:00} combined the estimated likelihoods with a prior to yield a posterior PDF of redshift conditioned on photometry.

### 2019-03-10 paper/tex/methods.tex
moved LePhare because it has no info beyond the intro

@@
 
-We test three publicly available and commonly used template-based codes, \bpz, \eazy, and \lephare, that share the standard physically motivated approach of calculating model fluxes for a set of template SEDs on a grid of redshift values and evaluating a merit function using the observed and model fluxes.
+We test three publicly available and commonly used template-based codes, \lephare, \bpz, and \eazy, that share the standard physically motivated approach of calculating model fluxes for a set of template SEDs on a grid of redshift values and evaluating a merit function using the observed and model fluxes.
 These three also share the \chisq\ merit function
@@
 
+\subsubsection{LePhare}
+\label{sec:lephare}
+%(C\'ecile Roucelle, Eric Nuss, Johann Cohen-Tanugi)
+
+\lephare \footnote{\url{http://www.cfht.hawaii.edu/~arnouts/lephare.html}}\citep[Photometric Analysis for Redshift Estimate,][]{Arnouts:99,Ilbert:06} matches observed colors with those predicted from a template set, which can be semi-empirical or entirely synthetic, directly according to Equation~\ref{eq_temp_chi} without modification.
+The reported \pzpdf\ is an arbitrary normalization of the likelihood evaluated on the output redshift grid.
+%\textsc{LePhare} has been used to produce the COSMOS2015 photo-$z$ catalogue \citep{Laigle:16}.
+Here we use \lephare-v 2.2 with the DC1 template set of Section~\ref{sec:buzztemplates}.
+
 \subsubsection{BPZ}
@@
 
-\bpz\footnote{\url{http://www.stsci.edu/~dcoe/BPZ/}} \citep[Bayesian Photometric Redshift,][]{Benitez:00} is a template-based \pz\ code that determines the likelihood $p(C \mid z, T)$ of a galaxy's observed colours $C$ for a set of SED templates $T$ at redshifts $z$.
+\bpz \footnote{\url{http://www.stsci.edu/~dcoe/BPZ/}} \citep[Bayesian Photometric Redshift,][]{Benitez:00} determines the likelihood $p(C \mid z, T)$ of a galaxy's observed colours $C$ for a set of SED templates $T$ at redshifts $z$.
 The \bpz\ likelihood is related to Equation~\ref{eq_temp_chi} by $p(C \mid z, T) \propto \exp[- \chi^{2} / 2]$.
@@
 
-\eazy\footnote{\url{https://github.com/gbrammer/eazy-photoz}} \citep[Easy and Accurate Photometric Redshifts from Yale,][]{Brammer:08} extends the basic \chisq\ fit procedure that defines template-fitting approaches.
+\eazy \footnote{\url{https://github.com/gbrammer/eazy-photoz}} \citep[Easy and Accurate Photometric Redshifts from Yale,][]{Brammer:08} extends the basic \chisq\ fit procedure that defines template-fitting approaches.
 The algorithm models the observed photometry with a linear combination of template SEDs at each redshift.
@@
 We set the template error to zero since the same templates were in fact used to produce the DC1 photometry.
-
-\subsubsection{LePhare}
-\label{sec:lephare}
-%(C\'ecile Roucelle, Eric Nuss, Johann Cohen-Tanugi)
-
-\textsc{LePhare}\footnote{\url{http://www.cfht.hawaii.edu/~arnouts/lephare.html}}\citep[Photometric Analysis for Redshift Estimate,][]{Arnouts:99,Ilbert:06} is a photo-$z$ reconstruction code based on a $\chi^2$ template-fitting procedure.
-The observed colors are matched with the colours predicted from a set of spectral energy distribution (SED) which can be either synthetic or based on a semi-empirical approach.
-%\textsc{LePhare} has been used to produce the COSMOS2015 photo-$z$ catalogue \citep{Laigle:16}.
-
-Each SED is convolved with the simulated LSST filter transmission curves (accounting for instrument efficiency).
-The computed photo-$z$ is then the value that minimizes the merit function $\chi^2 (z,T,A)$ from \citet[]{Arnouts:99}, and given in Equation~\ref{eq_temp_chi}.
-
-In this paper we use \textsc{LePhare v 2.2}.
-The set of templates used for fitting the photo-$z$'s are the 100 discrete Buzzard SED templates as described in section~\ref{sec:buzztemplates}, and the full $p(z)$ corresponds to the likelihoods calculated at each point on our $z$-grid.
 

### 2019-03-10 paper/tex/methods.tex
intro to ML redundancy

@@
 
-We test three publicly available and commonly used template-based codes, \lephare, \bpz, and \eazy, that share the standard physically motivated approach of calculating model fluxes for a set of template SEDs on a grid of redshift values and evaluating a merit function using the observed and model fluxes.
+We test three publicly available and commonly used template-based codes that share the standard physically motivated approach of calculating model fluxes for a set of template SEDs on a grid of redshift values and evaluating a merit function using the observed and model fluxes.
 These three also share the \chisq\ merit function
@@
 
-\subsection{Training-based Codes}
+\subsection{Training-based Approaches}
 \label{sec:trainingcodes}
 
-The training-based codes use a variety of algorithms in order to estimate $p(z)$, specifics of each implementation are described in the subsections.
-Some aspects of data treatment were left to the individual code runners, for example, whether/how to split the available data with known redshifts into separate training and validation sets.
-Another key difference is the treatment of non-detections in one or more bands.
-Some codes choose to ignore a band, others replace the value with either an estimate for the detection limit, the mean of other values in the training set, or another default value.
-There are varying conventions among training-based codes for treatment of non-detections, and no one prescription dominates in the photo-$z$ literature.
-The specific choices for each code affect the results, and contribute to the implicit prior influencing their output.
-However, we remind the reader that only 2.0 per cent of our sample has non-detections, almost exclusively in the u-band, and thus should not dominate the code performance differences.
+We compared nine data-driven \pz\ estimation approaches, eight of which are described in this section and one of which is discussed in Section~\ref{sec:trainz}.
+Because the algorithms differ more from one another and the techniques are relative newcomers to the astronomical literature, we provide somewhat more detail about the implementations below.
+%
+% Some aspects of data treatment were left to the individual code runners, for example, whether and how to split the DC1 training set for validation.
+% The codes considered treated
+% Another key difference is the treatment of non-detections in one or more bands.
+% Some codes choose to ignore a band, others replace the value with either an estimate for the detection limit, the mean of other values in the training set, or another default value.
+% There are varying conventions among training-based codes for treatment of non-detections, and no one prescription dominates in the photo-$z$ literature.
+% The specific choices for each code affect the results, and contribute to the implicit prior influencing their output.
+% However, we remind the reader that only 2.0 per cent of our sample has non-detections, almost exclusively in the u-band, and thus should not dominate the code performance differences.
 
@@
 
-\textsc{ANNz2}\footnote{\url{https://github.com/IftachSadeh/ANNZ}} \citep{Sadeh:16} is a software package that has the ability to employ several machine learning algorithms, including artificial neural networks (ANN), boosted decision tree (BDT) and k-nearest neighbour (KNN).
+\annz \footnote{\url{https://github.com/IftachSadeh/ANNZ}} \citep{Sadeh:16} employs several machine learning algorithms, including artificial neural networks (ANN), boosted decision tree (BDT) and k-nearest neighbour (KNN).
 Using the Toolkit for Multivariate Data Analysis (TMVA) with ROOT\footnote{\url{http://tmva.sourceforge.net/}}, it can either run a single machine learning algorithm and return results, or it can run multiple algorithms simultaneously and output photo-$z$'s as a weighted combination of the different algorithms.
@@
 
-\subsection{Simple Ensemble Estimator}\label{sec:method:trainz}
+\subsection{Simple Ensemble Estimator}
+\label{sec:method:trainz}
 

### 2019-03-10 paper/tex/methods.tex
ANNz2 editing

@@
 
-\annz \footnote{\url{https://github.com/IftachSadeh/ANNZ}} \citep{Sadeh:16} employs several machine learning algorithms, including artificial neural networks (ANN), boosted decision tree (BDT) and k-nearest neighbour (KNN).
-Using the Toolkit for Multivariate Data Analysis (TMVA) with ROOT\footnote{\url{http://tmva.sourceforge.net/}}, it can either run a single machine learning algorithm and return results, or it can run multiple algorithms simultaneously and output photo-$z$'s as a weighted combination of the different algorithms.
-In this study, only ANNs were employed.
-The redshift PDFs are produced by running an ensemble set of ANNs, each with different random seeds used in initialization of input parameters for training.
-Uncertainties for each method are estimated from a KNN-uncertainty estimator \citep{Oyaizu:08}.
-The final PDF can either be the ``best'' of the candidate PDFs, or a weighted average of the PDFs based on their error estimates for each of the ensemble members.
+\annz \footnote{\url{https://github.com/IftachSadeh/ANNZ}} \citep{Sadeh:16} employs several machine learning algorithms, including artificial neural networks (ANN), boosted decision tree (BDT), and k-nearest neighbour (KNN) regression.
+In addition to accounting for errors on the input photometry, \annz\ uses the KNN-uncertainty estimate of \citet{Oyaizu:08} to quantify uncertainty in the choice of method over multiple runs.
+Using the Toolkit for Multivariate Data Analysis (TMVA) with ROOT\footnote{\url{http://tmva.sourceforge.net/}}, it can return the results of running a single machine learning algorithm, a ``best'' choice of the results from simultaneously running multiple algorithms, or a combination of the results of multiple algorithms weighted by their method uncertainties averaged over multiple runs.
 %\textsc{ANNz2} is capable of producing both photo-$z$ point estimates and redshift posterior probability distributions $p(z)$.
-It can also perform classifications, and supports reweighting between samples.
-The PDFs are produced by propagating the intrinsic uncertainty on the input parameters and the uncertainty in the machine learning method to the expected photo-$z$ solution, averaged over multiple runs weighted based on the performance of each run.
+% It can also perform classifications, and supports reweighting between samples.
+% \annz\ propagates the intrinsic uncertainty on the input parameters and the uncertainty in the machine learning method to the expected photo-$z$ solution, averaged over multiple runs weighted based on the performance of each run.
 
-In this study, \textsc{ANNz2 v. 2.0.4} was used.
-A set of $5$ ANNs with architecture $6:12:12:1$ ($6$ $ugrizy$ inputs, $2$ hidden layers with $12$ nodes each, and $1$ output) with different random seeds are used during each training.
+In this study, we used \annz-v.2.0.4 to output only the result of the ANN algorithm.
+\Pzpdf s were produced by running an ensemble of 5 ANNs with architecture $6:12:12:1$ (6 $ugrizy$ inputs, 2 hidden layers with 12 nodes each, and 1 output), each trained with different random seeds for the initialization of input parameters.
 %Training was conducted on a subset of the given training sample which satisfies $i\le25.3$ as it would have improved the results compared to training on the entire sample.
-Half of the training set is used as a validation set to prevent overtraining.
-All training objects are set to have detected magnitudes, however the non-detections (mag$=99$) in the testing set are replaced with the mean of that particular band.
+To prevent overtraining, half of the training set was set aside for validation.
+Undetected galaxies were excluded from the training set, and per-band non-detections in the test set were replaced with the mean magnitude in that band within the entire test set.
 

### 2019-03-10 paper/tex/methods.tex
more concise CMNN discussion

@@
 
-The nearest-neighbours colour-matching photometric redshift estimator (\textsc{CMNN}) is presented in \citet[][hereafter G18]{Graham:17}.
-This method uses a training set of galaxies with known redshifts that has equivalent or better photometry as the test set in terms of quality and filter coverage.
-For each galaxy in the test set we identify a colour-matched subset of training galaxies.
-This subset is identified by first calculating the Mahalanobis distance $D_M$ in colour-space between the test galaxy and all training-set galaxies:
-\begin{equation}
-D_{\rm M} = \sum^{N_{\rm colours}}\frac{(c_{\rm train}-c_{\rm test})^2}{(\delta c_{\rm test})^2}
-\end{equation}
-where c colour, $\delta c_{\rm test}$ is the measurement error on the colour, and $N_{\rm colours}$ is the total number of colors (i.e., in our case $u$-$g$, $g$-$r$, $r$-$i$, $i$-$z$, and $z$-$y$).
-Then, we choose a threshold value for $D_M$ that defines the colour match set based on a set value of the percent point function (PPF): for example, for $N_{\rm dof}=5$, choosing a vPPF$=0.95$, $95$ per cent of all training galaxies consistent with the test galaxy will have $D_M < 11.07$ (where $N_{\rm dof}$ is the number of degrees of freedom, in this case the number of colours).
-If a galaxy had a non-detection in a band, that band was dropped and $N_{\rm dof}$ was reduced by one in the colour-matching space.
-For a given test galaxy, the $p(z)$ is the normalized distribution of the true catalogue redshifts of this colour-matched subset of training galaxies.
+The nearest-neighbours colour-matching photometric redshift estimator \citep[\cmnn][]{Graham:17} uses a training set of galaxies with known redshifts that has equivalent or better photometry than the test set in terms of quality and filter coverage.
+For each galaxy in the test set, \cmnn\ identifies a colour-matched subset of training galaxies using a threshold in the Mahalanobis distance $D_M$ in the space of available colours for the test set galaxy
+% between the test galaxy and all training set galaxies:
+% \begin{equation}
+% D_{\rm M} = \sum^{N_{\rm colours}}\frac{(c_{\rm train}-c_{\rm test})^2}{(\delta c_{\rm test})^2}
+% \end{equation}
+% where c colour, $\delta c_{\rm test}$ is the measurement error on the colour, and $N_{\rm colours}$ is the total number of colors (i.e., in our case $u$-$g$, $g$-$r$, $r$-$i$, $i$-$z$, and $z$-$y$).
+that defines the colour match set based on a value of the percent point function (PPF).
+As an example, for $N_{\rm{filt}}=5$ with PPF$=0.95$, $95\%$ of all training galaxies consistent with the test galaxy will have $D_M < 11.07$.
+Undetected bands are dropped, thereby reducing the effective $N_{\rm{filt}}$ for that galaxy.
+The \pzpdf\ of a given test set galaxy is the normalized distribution of redshifts of its colour-matched subset of training set galaxies.
 
-We have applied the nearest-neighbours colour-matching photometric redshift estimator described in G18 to the simulated data.
-Compared to its application in G18, there are some minor differences in the application of this estimator to the Buzzard catalogue.
-First, we do not impose non-detections on galaxies with a magnitude fainter than the expected LSST 10-year limiting magnitude or bright enough to saturate with LSST: {\it all} of the photometry for all the galaxies in the test and training sets are used for this experiment.
-Second, as in G18 we do apply an initial cut in colour to the training set before calculating the Mahalanobis distance in order to accelerate processing, and also use a magnitude pseudo-prior to improve photo-$z$ estimates, but for both we have used different cut-off values that are appropriate for the Buzzard galaxies' colours and magnitudes.
-Third, we set different parameters for the identification of the colour-matched subset of training galaxies and the selection of a photometric redshift estimate.
-In G18 we used a PPF value of 0.68 to identify the colour-matched subset of training galaxies and used the redshift of nearest neighbour in colour-space as the photo-$z$ estimate.
-These choices work well when the desire is to obtain accurate photo-$z$ estimates for most test-set galaxies, but does not return an accurate $p(z)$ in all cases -- especially for galaxies that are bright and/or have few matches in colour-space.
-Since an accurate estimate of $p(z)$ is desired for this work we make several changes to our implementation of the \textsc{CMNN} photo-$z$ estimator.
-We continue to use a ${\rm PPF} = 0.95$ to generate the subset of colour-matched training galaxies, but weight them by the inverse of their Mahalanobis distance.
-This weighting maintains some of the accuracy that was previously achieved by simply using the nearest neighbour in colour-space.
-We then use the weights to create the $p(z)$ instead of having the redshift of each colour-matched training-set galaxy count equally.
-To obtain a robust estimate of the $p(z)$ for galaxies with a small number of colour-matched training set galaxies, when this number is less than $20$ the nearest $20$ neighbours in colour-space are used instead, and we convolve the $p(z)$ with a Gaussian with a standard deviation of:
-\begin{equation}
-\sigma = \sigma_{\rm train} \sqrt{({\rm PPF}_{20}/0.95)^2 -1}
-\end{equation}
-\noindent to appropriately broaden it so that the $p(z)$ for these test galaxies represents the enlarged PPF value associated with it.
-Overall, these three changes will yield less precise photo-$z$ estimates compared to those presented in G18, but they will all have significantly more accurate estimates of the $p(z)$, particularly for the brightest test galaxies.
-This is sufficient for this work because, as described in G18, the goal of the \textsc{CMNN} photo-$z$ estimator was never to provide the ``best'' (or even competitive) estimates in the first place, given its reliance on a deep training set, but rather to provide a means for direct comparisons between LSST photometric quality and photo-$z$ estimates.
-With this work we show how the input parameters should be set in order to return accurate $p(z)$ estimates in addition to point value estimates.
+Here, we make two modifications to the implementation of \citet{Graham:17} to comply with the controlled experimental conditions.
+First, we do not impose non-detections on galaxies fainter than the expected \lsst\ 10-year limiting magnitude or bright enough to saturate with \lsst's CCDs, instead using all of the photometry for the DC1 test and training sets.
+Second, we apply the initial colour cut to the training set before calculating the Mahalanobis distance in order to accelerate processing and use a magnitude pseudo-prior as in \citet{Graham:17}, but for both we use cut-off values corresponding to the DC1 training set galaxies' colours and magnitudes.
+
+We make an additional adaptation to enable the \cmnn\ algorithm to yield accurate \pzpdf s for all galaxies, as the original \citet{Graham:17} algorithm is optimized for \pz\ point estimates and is susceptible to less accurate \pzpdf s for bright galaxies or those with few matches in colour-space.
+We use PPF$=0.95$ rather than PPF$=0.68$ to generate the subset of colour-matched training galaxies, whose redshifts are weighted by their inverse Mahalanobis distances of the when composing the \pzpdf\ rather than weighting all colour-matched training galaxies equally.
+Additionally, when the number of colour-matched training set galaxies is less than 20, the nearest 20 neighbours in color-space are used instead, and the output \pzpdf\ is convolved with a Gaussian kernel of variance $\sigma_{\rm train}^{2}({\rm PPF}_{20}/0.95)^2 -1$ to account foe the corresponding growth of the effective PPF to include 20 neighbors.
 

### 2019-03-10 paper/tex/methods.tex
revising earlier methods

@@
 
-We test three publicly available and commonly used template-based codes that share the standard physically motivated approach of calculating model fluxes for a set of template SEDs on a grid of redshift values and evaluating a merit function using the observed and model fluxes.
-These three also share the \chisq\ merit function
-\begin{equation}
-	\label{eq_temp_chi}
-	\chi^{2}(z, T, A) \equiv \sum_{b}^{N_{\rm{filt}}} \left(\frac{F^{\rm{obs}}_{b} - A F^{\rm{mod}}_{b}(T, z)}{\sigma^{\rm{obs}}_{b}}\right)^{2},
-\end{equation}
-of normalization factor $A$, template $T$, and redshift $z$, which is a sum of observed flux error $\sigma_{b}^{\rm{obs}}$-weighted squared differences between the observed flux $F^{\rm{obs}}_{b}$ and the normalized predicted flux $F^{\rm{mod}}_{b}(T, z)$ in $N_{\rm{filt}}$ photometric filters $b$, which in our case is the \lsst\ $ugrizy$ filters.
+We test three publicly available and commonly used template-based codes that share the standard physically motivated approach of calculating model fluxes for a set of template SEDs on a grid of redshift values and evaluating a \chisq\ merit function using the observed and model fluxes.
 
@@
 
-\lephare \footnote{\url{http://www.cfht.hawaii.edu/~arnouts/lephare.html}}\citep[Photometric Analysis for Redshift Estimate,][]{Arnouts:99,Ilbert:06} matches observed colors with those predicted from a template set, which can be semi-empirical or entirely synthetic, directly according to Equation~\ref{eq_temp_chi} without modification.
+\lephare \footnote{\url{http://www.cfht.hawaii.edu/~arnouts/lephare.html}}\citep[Photometric Analysis for Redshift Estimate,][]{Arnouts:99,Ilbert:06} matches observed colors with those predicted from a template set, which can be semi-empirical or entirely synthetic, directly according to the likelihood $\chi^{2}(z, T, A) \equiv \sum_{b}^{N_{\rm{filt}}} \left((F^{\rm{obs}}_{b} - A F^{\rm{mod}}_{b}(T, z)) / \sigma^{\rm{obs}}_{b}\right)^{2}$ of normalization factor $A$, template $T$, and redshift $z$.
+In words, the likelihood is a sum of observed flux error $\sigma_{b}^{\rm{obs}}$-weighted squared differences between the observed flux $F^{\rm{obs}}_{b}$ and the normalized predicted flux $F^{\rm{mod}}_{b}(T, z)$ in $N_{\rm{filt}}$ photometric filters $b$, which is the \lsst\ $ugrizy$ filters in this case.
 The reported \pzpdf\ is an arbitrary normalization of the likelihood evaluated on the output redshift grid.
 %\textsc{LePhare} has been used to produce the COSMOS2015 photo-$z$ catalogue \citep{Laigle:16}.
+
 Here we use \lephare-v 2.2 with the DC1 template set of Section~\ref{sec:buzztemplates}.
@@
 Because the algorithms differ more from one another and the techniques are relative newcomers to the astronomical literature, we provide somewhat more detail about the implementations below.
-%
+
 % Some aspects of data treatment were left to the individual code runners, for example, whether and how to split the DC1 training set for validation.
@@
 
-\annz \footnote{\url{https://github.com/IftachSadeh/ANNZ}} \citep{Sadeh:16} employs several machine learning algorithms, including artificial neural networks (ANN), boosted decision tree (BDT), and k-nearest neighbour (KNN) regression.
+\annz \footnote{\url{https://github.com/IftachSadeh/ANNZ}} \citep{Sadeh:16} employs several machine learning algorithms, including artificial neural networks (ANN), boosted decision tree, and k-nearest neighbour (KNN) regression.
 In addition to accounting for errors on the input photometry, \annz\ uses the KNN-uncertainty estimate of \citet{Oyaizu:08} to quantify uncertainty in the choice of method over multiple runs.
-Using the Toolkit for Multivariate Data Analysis (TMVA) with ROOT\footnote{\url{http://tmva.sourceforge.net/}}, it can return the results of running a single machine learning algorithm, a ``best'' choice of the results from simultaneously running multiple algorithms, or a combination of the results of multiple algorithms weighted by their method uncertainties averaged over multiple runs.
+Using the Toolkit for Multivariate Data Analysis with ROOT\footnote{\url{http://tmva.sourceforge.net/}}, it can return the results of running a single machine learning algorithm, a ``best'' choice of the results from simultaneously running multiple algorithms, or a combination of the results of multiple algorithms weighted by their method uncertainties averaged over multiple runs.
 %\textsc{ANNz2} is capable of producing both photo-$z$ point estimates and redshift posterior probability distributions $p(z)$.
@@
 In this study, we used \annz-v.2.0.4 to output only the result of the ANN algorithm.
-\Pzpdf s were produced by running an ensemble of 5 ANNs with architecture $6:12:12:1$ (6 $ugrizy$ inputs, 2 hidden layers with 12 nodes each, and 1 output), each trained with different random seeds for the initialization of input parameters.
-%Training was conducted on a subset of the given training sample which satisfies $i\le25.3$ as it would have improved the results compared to training on the entire sample.
-To prevent overtraining, half of the training set was set aside for validation.
+\Pzpdf s were produced by running an ensemble of 5 ANNs with 6 $ugrizy$ inputs, 2 hidden layers with 12 nodes each, and 1 output of redshift).
+Each of the five ANNs was trained with different random seeds for the initialization of input parameters.
+Additionally, all ANNs were trained on only a $i \leq 25.3$ subsample of the DC1 training set, and half of the training set was reserved for validation to prevent overfitting.
 Undetected galaxies were excluded from the training set, and per-band non-detections in the test set were replaced with the mean magnitude in that band within the entire test set.
@@
 
-The nearest-neighbours colour-matching photometric redshift estimator \citep[\cmnn][]{Graham:17} uses a training set of galaxies with known redshifts that has equivalent or better photometry than the test set in terms of quality and filter coverage.
+The nearest-neighbours colour-matching photometric redshift estimator \citep[\cmnn,][]{Graham:17} uses a training set of galaxies with known redshifts that has equivalent or better photometry than the test set in terms of quality and filter coverage.
 For each galaxy in the test set, \cmnn\ identifies a colour-matched subset of training galaxies using a threshold in the Mahalanobis distance $D_M$ in the space of available colours for the test set galaxy
@@
 Here, we make two modifications to the implementation of \citet{Graham:17} to comply with the controlled experimental conditions.
-First, we do not impose non-detections on galaxies fainter than the expected \lsst\ 10-year limiting magnitude or bright enough to saturate with \lsst's CCDs, instead using all of the photometry for the DC1 test and training sets.
+First, we do not impose nondetections on galaxies fainter than the expected \lsst\ 10-year limiting magnitude or bright enough to saturate with \lsst's CCDs, instead using all of the photometry for the DC1 test and training sets.
 Second, we apply the initial colour cut to the training set before calculating the Mahalanobis distance in order to accelerate processing and use a magnitude pseudo-prior as in \citet{Graham:17}, but for both we use cut-off values corresponding to the DC1 training set galaxies' colours and magnitudes.

### 2019-03-10 paper/tex/methods.tex
working through flexzboost and trying to use consistent notation

@@
 \hline
+\lephare~\citep{Arnouts:99}	   & template fitting	& \url{http://www.cfht.hawaii.edu/~arnouts/lephare.html} \\
 \bpz~\citep{Benitez:00} 		   & template fitting	& \url{http://www.stsci.edu/~dcoe/BPZ/} \\
 \eazy~\citep{Brammer:08}		   & template fitting & \url{https://github.com/gbrammer/eazy-photoz} \\
-\lephare~\citep{Arnouts:99}	   & template fitting	& \url{http://www.cfht.hawaii.edu/~arnouts/lephare.html} \\
 \annz~\citep{Sadeh:16}		     & machine learning	& \url{https://github.com/IftachSadeh/ANNZ} \\
-\delight~\citep{Leistedt:17}   & machine learning & \url{https://github.com/ixkael/Delight} \\
 \flexzboost~\citep{Izbicki:17} & machine learning & \url{https://github.com/tpospisi/flexcode}; \url{https://github.com/rizbicki/FlexCoDE}\\
@@
 \tpz~\citep{Carrasco_Kind:13}	 & machine learning	& \url{https://github.com/mgckind/MLZ} \\
+\delight~\citep{Leistedt:17}   & hybrid           & \url{https://github.com/ixkael/Delight} \\
 \hline
@@
 
-\bpz \footnote{\url{http://www.stsci.edu/~dcoe/BPZ/}} \citep[Bayesian Photometric Redshift,][]{Benitez:00} determines the likelihood $p(C \mid z, T)$ of a galaxy's observed colours $C$ for a set of SED templates $T$ at redshifts $z$.
-The \bpz\ likelihood is related to Equation~\ref{eq_temp_chi} by $p(C \mid z, T) \propto \exp[- \chi^{2} / 2]$.
-Given a Bayesian prior $p(z, T \mid m_{0})$ over apparent magnitude $m_0$ and assuming that the SED templates are spanning and exclusive, \bpz\ constructs the redshift posterior $p(z \mid C, m_0)$ by marginalizing over all SED templates as in \citep[Eq.~3 from][]{Benitez:00}, corresponding to setting the parameter \texttt{PROBS\_LITE=TRUE} in the \bpz\ parameter file.
-The \bpz\ prior is the product of an SED template proportion that varies with apparent magnitude $p(T \mid m_{0})$ and a prior $p(z \mid T, m_{0})$ over the expected redshift as a function of apparent magnitude and SED template.
+\bpz \footnote{\url{http://www.stsci.edu/~dcoe/BPZ/}} \citep[Bayesian Photometric Redshift,][]{Benitez:00} determines the likelihood $p(C \vert z, T)$ of a galaxy's observed colours $C$ for a set of SED templates $T$ at redshifts $z$.
+The \bpz\ likelihood is related to the \chisq\ likelihood by $p(C \vert z, T) \propto Equation~\ref{eq_temp_chi} \exp[- \chi^{2} / 2]$.
+Given a Bayesian prior $p(z, T \vert m_{0})$ over apparent magnitude $m_0$ and assuming that the SED templates are spanning and exclusive, \bpz\ constructs the redshift posterior $p(z \vert C, m_0)$ by marginalizing over all SED templates as in \citep[Eq.~3 from][]{Benitez:00}, corresponding to setting the parameter \texttt{PROBS\_LITE=TRUE} in the \bpz\ parameter file.
+The \bpz\ prior is the product of an SED template proportion that varies with apparent magnitude $p(T \vert m_{0})$ and a prior $p(z \vert T, m_{0})$ over the expected redshift as a function of apparent magnitude and SED template.
 
@@
 To keep the number of free parameters manageable, the DC1 template set is pre-sorted by the rest-frame $u$-$g$ colour and split into three broad classes of SED template, equivalent to the E, Sp and Im/SB types in .
-The Bayesian prior term $p(T \mid m_{0})$ was derived directly from the DC1 training set, and the other term $p(z \mid T, m_{0})$ was chosen to be the best fit for the eleven free parameters of the functional form of \citet{Benitez:00}.
+The Bayesian prior term $p(T \vert m_{0})$ was derived directly from the DC1 training set, and the other term $p(z \vert T, m_{0})$ was chosen to be the best fit for the eleven free parameters of the functional form of \citet{Benitez:00}.
 %For photo-$z$ point estimates we use the \texttt{Z\_B} output parameter.
@@
 The nearest-neighbours colour-matching photometric redshift estimator \citep[\cmnn,][]{Graham:17} uses a training set of galaxies with known redshifts that has equivalent or better photometry than the test set in terms of quality and filter coverage.
-For each galaxy in the test set, \cmnn\ identifies a colour-matched subset of training galaxies using a threshold in the Mahalanobis distance $D_M$ in the space of available colours for the test set galaxy
-% between the test galaxy and all training set galaxies:
-% \begin{equation}
-% D_{\rm M} = \sum^{N_{\rm colours}}\frac{(c_{\rm train}-c_{\rm test})^2}{(\delta c_{\rm test})^2}
-% \end{equation}
-% where c colour, $\delta c_{\rm test}$ is the measurement error on the colour, and $N_{\rm colours}$ is the total number of colors (i.e., in our case $u$-$g$, $g$-$r$, $r$-$i$, $i$-$z$, and $z$-$y$).
-that defines the colour match set based on a value of the percent point function (PPF).
+For each galaxy in the test set, \cmnn\ identifies a colour-matched subset of training galaxies using a threshold in the Mahalanobis distance $D_M = \sum_{j}^{N_{\rm colours}} (c^{\rm train}_{j} - c^{\rm test}_{j})^{2} / \delta c_{\rm test}^2$ in the space of available colours $c$, with colour measurement errors $\delta c_{\rm test}$ and $N_{\rm colours} = 5$ colors $j$ defined by the $ugrizy$ filters, which defines the set of colour-matched neighbours based on a value of the percent point function (PPF).
 As an example, for $N_{\rm{filt}}=5$ with PPF$=0.95$, $95\%$ of all training galaxies consistent with the test galaxy will have $D_M < 11.07$.
@@
 
-\subsubsection{Delight}
-\label{sec:delight}
-%(John Soo)
-
-\textsc{Delight}\footnote{\url{https://github.com/ixkael/Delight}} \citep{Leistedt:17} infers photo-$z$'s by using a data-driven model of latent SEDs and a physical model of photometric fluxes as a function of redshift.
-Delight models the underlying latent SEDs as a linear combination of a set of pre-defined template SEDs, plus zero mean Gaussian processes with factorized kernels.
-Generally, machine learning methods rely on representative training data with similar band passes, while template based methods rely on a complete library of templates based on physical models constructed.
-\textsc{Delight} is constructed in attempt to combine the advantages and eliminate the disadvantages of both template-based and machine learning algorithms: it constructs a large collection of latent SED templates (or physical flux-redshift models) from training data, with a template SED library as a guide to the learning of the model.
-The advantage of \textsc{Delight} is that it neither needs representative training data in the same photometric bands, nor does it need detailed galaxy SED models to work.
-
-This conceptually novel approach is done by using Gaussian processes operating in flux-redshift space.
-The posterior distribution on the redshift of a target galaxy is obtained via a pairwise comparison with training galaxies,
-
-\begin{equation}
-p(z|\mathbf{\hat{F}}) \approx \sum_i p(\mathbf{\hat{F}}|z,t_i)\, p(z|t_i)p(t_i),
-\end{equation}
-\noindent where $p(z|t_i)p(t_i)$ captures prior information about the redshift distributions and abundances of the galaxies, with $t_i$ denoting the galaxy template; while $p(\mathbf{\hat{F}}|z,t_i)$ is the posterior of noisy flux $\mathbf{\hat{F}}$ at redshift $z$.
-For each training-target pair, $p(\mathbf{\hat{F}}|z,t_i)$ is evaluated as follows:
-\begin{equation} \label{eq:delight_noisy}
-p(\mathbf{\hat{F}}|z,t_i) = \int p(\mathbf{\hat{F}}|\mathbf{F})\, p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)\, d\mathbf{F},
-\end{equation}
-where $p(\mathbf{\hat{F}}|\mathbf{F})$ is the likelihood function, it compares the noisy real flux $\mathbf{\hat{F}}$ with the noiseless flux $\mathbf{F}$ obtained from the linear combination of template models, carefully constructed to account for model uncertainties and different normalization of the same SED; while $p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)$ is the prediction of flux at a different redshift $z$ with respect to the training object with redshift $z_i$ and flux $\mathbf{\hat{F}}_i$.
-Eq.~\ref{eq:delight_noisy} is essentially the probability that the training and the target galaxies having the same SED but at a different redshift.
-The flux prediction $p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)$ of the training galaxy at redshift $z$ is modeled via a Gaussian process,
-
-\begin{equation} \label{eq:delight_gp}
-F_b \sim \mathcal{GP}\left( \mu^F,k^F \right),
-\end{equation}
-\noindent with mean function $\mu^F$ and kernel $k^F$, both imposed to capture expected correlations resulting from the known underlying physics (i.e., fluxes resulting from observing SEDs through filter response, and the SEDs being redshifted).
-The reader should refer to \citet{Leistedt:17} for further details.
-
-In this study, all $100$ ordered Buzzard templates, as described in Section~\ref{sec:buzztemplates}, were used in \textsc{Delight}, and the Gaussian process was trained using the provided training sample.
-Photometric uncertainties from the inputs are propagated into the code, while non-detections for each band are set to the mean of the respective bands.
-The default settings of \textsc{Delight} were used, with the exception that the PDF bins were set to be linearly-spaced rather than logarithmic.
-In this study a flat prior in magnitude/type is assumed.
-
 \subsubsection{FlexZBoost}
@@
 
-\textsc{FlexZBoost}\footnote{\url{https://github.com/tpospisi/flexcode};\\ \url{https://github.com/rizbicki/FlexCoDE}} \citep{Izbicki:17} is a particular realization of \texttt{FlexCode}, which is a general-purpose methodology for converting any conditional mean point estimator of $z$ to a conditional {\em density} estimator $f(z \vert \x)$, where $\x$ here represents our photometric covariates and errors.\footnote{Instead of $p(z)$, we use the notation $f(z \vert \x)$ to explicitly show the dependence on $\x$.}
-The key idea is to expand the unknown function $f(z \vert \x)$ in an orthonormal basis $\{\phi_i(z)\}_{i}$:
-\begin{equation}
-f(z|\x)=\sum_{i}\beta_{i }(\x)\phi_i(z). \label{eq::series_expansion}
-\end{equation}
-By the orthogonality property, the expansion coefficients are just conditional means
-\begin{equation}
-\beta_{i }(\x) =  \mathbb{E}\left[\phi_i(z)|\x\right] \equiv \int f(z|\x)   \phi_i(z) dz. \label{eq:FlexZBoost_coeffs}
-\end{equation}
-These coefficients can easily be estimated from data by regression since $\mathbb{E}\left[\phi_i(z)|\x\right]$ is the regression of $\phi_i(z)$ (a transformation of $Z$) on $X$.
+\flexzboost \footnote{\url{https://github.com/tpospisi/flexcode};  \url{https://github.com/rizbicki/FlexCoDE} \label{flexzboost_github}} \citep{Izbicki:17} is built on \texttt{FlexCode}, a general-purpose methodology for converting any conditional mean point estimator of $z$ to a conditional density estimator $p(z \vert \x) \equiv f(z \vert \x)$, where $\x$ here represents our photometric covariates and errors.
+\flexzboost\ expands the unknown function $f(z \vert \x) = \sum_{i}\beta_{i}(\x)\phi_{i}(z)$ using an orthonormal basis $\{\phi_{i}(z)\}_{i}$.
+By the orthogonality property, the expansion coefficients $\beta_{i}(\x) =  \mathbb{E}\left[\phi_i(z)|\x\right] \equiv \int f(z \vert \x) \phi_{i}(z) dz$ are thus conditional means.
+The expectation value $\mathbb{E}\left[\phi_i(z) \vert \x\right]$ of the expansion coefficients conditioned on the data is equivalent to the regression of the space of possible redshifts on the space of possible photometry.
+Thus the expansion coefficients $\beta_{i}(\x)$ can be estimated from the data via regression to yield the conditional density estimate $\widehat{f}(z \vert \x)$.
 
-In this paper, we use \textsc{xgboost}~\citep{Chen:16} for the regression part; it should however be noted that \textsc{FlexCode-RF} (also on GitHub), based on Random Forests, generally performs better for smaller data sets.
-As our basis, we choose a standard Fourier basis.
-There are two tuning parameters in our $p(z)$ estimate: (i) the number of terms, $I$, in the series expansion in Eq.~\ref{eq::series_expansion}, and (ii) an exponent $\alpha$ that we use to sharpen the computed density estimates $\widehat{f}(z|\x)$, according to $\widetilde{f}(z|\x) \propto \widehat{f}(z|\x)^\alpha$.
-We reserve 15\% of the training set data as a validation set, and choose both $I$ and $\alpha$ in an automated way by minimizing the weighted $L_2$-loss function \citep[Eq.
-5 in][]{Izbicki:17} on the  validation set.
-While the native storage format for \textsc{FlexCode} encodes the PDF using the coefficients shown in Equation~\ref{eq:FlexZBoost_coeffs}, to match the output format requested of other codes we discretize our final estimates into 200 bins linearly spaced in $0 < z < 2$.
-
-%Although {\tt FlexCode} offers a lossless compression of the photo-$z$ estimates (in this study, one can reconstruct $\widetilde{f}(z|\x)$ exactly at any resolution from estimates of the first 35 coefficients, Eq.~\ref{eq:FlexZBoost_coeffs}, for a Fourier basis $\{\phi_i(z)\}_{i}$), we discretize our final estimates into 200 bins linearly spaced in $0 < z < 2$ for easy comparison with other algorithms.
-Using a higher resolution may yield better results (with no added cost in storage).
+In this paper, we used \texttt{xgboost} \citep{Chen:16} for the regression; it should however be noted that \texttt{FlexCode-RF}\footref{flexzboost_github}, based on Random Forests, generally performs better for smaller datasets.
+As our basis $\phi_{i}(z)$, we choose a standard Fourier basis.
+The two tuning parameters in our \pzpdf\ estimate are the number $I$ of terms in the series expansion and an exponent $\alpha$ that we use to sharpen the computed density estimates $\widetilde{f}(z \vert \x) \propto \widehat{f}(z \vert \x)^{\alpha}$.
+Both $I$ and $\alpha$ were chosen in an automated way by minimizing the weighted $L_2$-loss function \citep[Eq.
+5 in][]{Izbicki:17} on a validation set comprised of a randomly selected 15\% of the DC1 training set.
+While \texttt{FlexCode}'s lossless native encoding stores each \pzpdf\ using the basis coefficients $\beta_{i}(\x)$, we discretized the final estimates into 200 linearly-spaced redshift bins $0 < z < 2$ to match the consistent output format of the experimental conditions.
 
@@
 
-\textsc{GPz}\footnote{\url{https://github.com/OxfordML/GPz}} \citep{Almosallam:16a,Almosallam:15b} is a sparse Gaussian process based code, a scalable approximation of full Gaussian Processes \citep{Rasmussen:06}, with the added feature of being able to produce input-dependent variance estimations (heteroscedastic noise).
-The model assumes that the probability of the output $y$, the redshift, given the input $x$, the photometry, is $p(z|x)=\mathcal{N}\left(z|\mu(x),\sigma(x)^{2}\right)$.
-The mean function, $\mu(x)$, and the variance function $\sigma(x)^{2}$ are both linear combinations of basis functions that take the following form:
-\begin{equation}
-f(x)=\sum_{i=1}^{m}\phi_{i}(x)w_{i},
-\end{equation}
-where $\left\{\phi_{i}(x)\right\}_{i=1}^{m}$ and $\left\{w_{i}\right\}_{i=1}^{m}$ are sets of $m$ basis functions and their associated weights respectively.
+\gpz \footnote{\url{https://github.com/OxfordML/GPz}} \citep{Almosallam:16a,Almosallam:15b} is a sparse Gaussian process based code, a scalable approximation of full Gaussian Processes \citep{Rasmussen:06}, that produces input-dependent variance estimates corresponding to heteroscedastic noise.
+The model assumes a Gaussian posterior probability $p(z \vert \x) = \mathcal{N}\left(z \vert \mu(\x), \sigma(\x)^{2}\right)$ of the output redshift $z$ given the input photometry $\x$.
+The mean $\mu(\x)$ and the variance $\sigma(\x)^{2}$ are modeled as functions $f(\x) = \sum_{i=1}^{m}w_{i}\phi_{i}(\x)$ linear combinations of basis functions $\left\{\phi_{i}(x)\right\}_{i=1}^{m}$ with coefficients $\left\{w_{i}\right\}_{i=1}^{m}$ are sets of $m$ basis functions and their associated weights respectively.
 Basis function models (BFM), for specific classes of basis functions such as the sigmoid or the squared exponential, have the advantage of being universal approximators, i.e.
@@
 
-\subsection{Simple Ensemble Estimator}
+\subsubsection{Delight}
+\label{sec:delight}
+%(John Soo)
+
+\delight \footnote{\url{https://github.com/ixkael/Delight}} \citep{Leistedt:17} infers \pz s with a data-driven model of latent SEDs and a physical model of photometric fluxes as a function of redshift.
+Delight models the underlying latent SEDs as a linear combination of a set of pre-defined template SEDs, plus zero mean Gaussian processes with factorized kernels.
+Generally, machine learning methods rely on representative training data with similar band passes, while template based methods rely on a complete library of templates based on physical models constructed.
+\textsc{Delight} is constructed in attempt to combine the advantages and eliminate the disadvantages of both template-based and machine learning algorithms: it constructs a large collection of latent SED templates (or physical flux-redshift models) from training data, with a template SED library as a guide to the learning of the model.
+The advantage of \textsc{Delight} is that it neither needs representative training data in the same photometric bands, nor does it need detailed galaxy SED models to work.
+
+This conceptually novel approach is done by using Gaussian processes operating in flux-redshift space.
+The posterior distribution on the redshift of a target galaxy is obtained via a pairwise comparison with training galaxies,
+
+\begin{equation}
+p(z|\mathbf{\hat{F}}) \approx \sum_i p(\mathbf{\hat{F}}|z,t_i)\, p(z|t_i)p(t_i),
+\end{equation}
+\noindent where $p(z|t_i)p(t_i)$ captures prior information about the redshift distributions and abundances of the galaxies, with $t_i$ denoting the galaxy template; while $p(\mathbf{\hat{F}}|z,t_i)$ is the posterior of noisy flux $\mathbf{\hat{F}}$ at redshift $z$.
+For each training-target pair, $p(\mathbf{\hat{F}}|z,t_i)$ is evaluated as follows:
+\begin{equation} \label{eq:delight_noisy}
+p(\mathbf{\hat{F}}|z,t_i) = \int p(\mathbf{\hat{F}}|\mathbf{F})\, p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)\, d\mathbf{F},
+\end{equation}
+where $p(\mathbf{\hat{F}}|\mathbf{F})$ is the likelihood function, it compares the noisy real flux $\mathbf{\hat{F}}$ with the noiseless flux $\mathbf{F}$ obtained from the linear combination of template models, carefully constructed to account for model uncertainties and different normalization of the same SED; while $p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)$ is the prediction of flux at a different redshift $z$ with respect to the training object with redshift $z_i$ and flux $\mathbf{\hat{F}}_i$.
+Eq.~\ref{eq:delight_noisy} is essentially the probability that the training and the target galaxies having the same SED but at a different redshift.
+The flux prediction $p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)$ of the training galaxy at redshift $z$ is modeled via a Gaussian process,
+
+\begin{equation} \label{eq:delight_gp}
+F_b \sim \mathcal{GP}\left( \mu^F,k^F \right),
+\end{equation}
+\noindent with mean function $\mu^F$ and kernel $k^F$, both imposed to capture expected correlations resulting from the known underlying physics (i.e., fluxes resulting from observing SEDs through filter response, and the SEDs being redshifted).
+The reader should refer to \citet{Leistedt:17} for further details.
+
+In this study, all $100$ ordered Buzzard templates, as described in Section~\ref{sec:buzztemplates}, were used in \textsc{Delight}, and the Gaussian process was trained using the provided training sample.
+Photometric uncertainties from the inputs are propagated into the code, while non-detections for each band are set to the mean of the respective bands.
+The default settings of \textsc{Delight} were used, with the exception that the PDF bins were set to be linearly-spaced rather than logarithmic.
+In this study a flat prior in magnitude/type is assumed.
+
+\subsection{trainZ: a straw-person \pz\ estimator}
 \label{sec:method:trainz}

### 2019-03-10 paper/tex/methods.tex
worked through GPZ

@@
 
-We describe the algorithms and implementations of the model-based and data-driven codes in Sections~\ref{sec:templatecodes} and \ref{sec:trainingcodes} respectively.
+We describe the algorithms and implementations of the model-based and data-driven codes in Sections~\ref{sec:templatecodes} and \ref{sec:trainingcodes} respectively, with a straw-person approach included in Section~\ref{sec:trainz}.
 % For each approach, we note how it accounts for photometric uncertainties, how it interprets negative fluxes, what its native output format of \pzpdf s is, and what, if any, preprocessing, including validation, of the prior information it performs.
+% In the following sections, we use a consistent notation defined in Table~\ref{tab:variables}.
+%
+% \begin{table}
+% 	\label{tab:variables}
+% 	\caption{Definitions of variables used in outlining \pzpdf\ codes}
+% 	\begin{tabular}{ll}
+% 		\hline
+% 		\bf Symbol & \bf Definition \\
+% 		\hline
+% 		\multirow{ 2}{*}{$b = 1,\dots,B$	& \multirow{ 2}{*}{number $B$ of photometric filters $b$;} \\
+% 														& \lsst's $ugrizy$ filters used here correspond to $B=6$ \\
+% 		\multirow{ 2}{*}{$m_{b} = 1,\dots,B$	& \multirow{ 2}{*}{number $B$ of photometric filters $b$;} \\
+% 																										& \lsst's $ugrizy$ filters used here correspond to $B=6$ \\
+% 		$z$											& redshift \\
+% 		$T$											& galaxy SED \\
+%
+% 	\end{tabular}
+% \end{table}
 
@@
 \bpz \footnote{\url{http://www.stsci.edu/~dcoe/BPZ/}} \citep[Bayesian Photometric Redshift,][]{Benitez:00} determines the likelihood $p(C \vert z, T)$ of a galaxy's observed colours $C$ for a set of SED templates $T$ at redshifts $z$.
-The \bpz\ likelihood is related to the \chisq\ likelihood by $p(C \vert z, T) \propto Equation~\ref{eq_temp_chi} \exp[- \chi^{2} / 2]$.
+The \bpz\ likelihood is related to the \chisq\ likelihood by $p(C \vert z, T) \propto \exp[- \chi^{2} / 2]$.
 Given a Bayesian prior $p(z, T \vert m_{0})$ over apparent magnitude $m_0$ and assuming that the SED templates are spanning and exclusive, \bpz\ constructs the redshift posterior $p(z \vert C, m_0)$ by marginalizing over all SED templates as in \citep[Eq.~3 from][]{Benitez:00}, corresponding to setting the parameter \texttt{PROBS\_LITE=TRUE} in the \bpz\ parameter file.
@@
 Here we test \bpz-v 1.99.3 with the DC1 template set of Section~\ref{sec:buzztemplates}.
-To keep the number of free parameters manageable, the DC1 template set is pre-sorted by the rest-frame $u$-$g$ colour and split into three broad classes of SED template, equivalent to the E, Sp and Im/SB types in .
+To keep the number of free parameters manageable, the DC1 template set is pre-sorted by the rest-frame $u-g$ colour and split into three broad classes of SED template, equivalent to the E, Sp and Im/SB types in .
 The Bayesian prior term $p(T \vert m_{0})$ was derived directly from the DC1 training set, and the other term $p(z \vert T, m_{0})$ was chosen to be the best fit for the eleven free parameters of the functional form of \citet{Benitez:00}.
@@
 The model assumes a Gaussian posterior probability $p(z \vert \x) = \mathcal{N}\left(z \vert \mu(\x), \sigma(\x)^{2}\right)$ of the output redshift $z$ given the input photometry $\x$.
-The mean $\mu(\x)$ and the variance $\sigma(\x)^{2}$ are modeled as functions $f(\x) = \sum_{i=1}^{m}w_{i}\phi_{i}(\x)$ linear combinations of basis functions $\left\{\phi_{i}(x)\right\}_{i=1}^{m}$ with coefficients $\left\{w_{i}\right\}_{i=1}^{m}$ are sets of $m$ basis functions and their associated weights respectively.
-Basis function models (BFM), for specific classes of basis functions such as the sigmoid or the squared exponential, have the advantage of being universal approximators, i.e.
+The mean $\mu(\x)$ and the variance $\sigma(\x)^{2}$ are modeled as functions $f(\x) = \sum_{i=1}^{m}w_{i}\phi_{i}(\x)$ linear combinations of $m$ basis functions $\left\{\phi_{i}(\x)\right\}_{i=1}^{m}$ with associated weights $\left\{w_{i}\right\}_{i=1}^{m}$.
+% Basis function models, for specific classes of basis functions such as the sigmoid or the squared exponential, have the advantage of being universal approximators, i.e.
 there exist a function of that form that can approximate any function, with mild assumptions, to any desired degree of accuracy.
 The details on how to learn the parameters of the model and the hyper-parameters of the basis functions are described in \citet{Almosallam:15b}.
+\gpz's variance estimate is composed of a model uncertainty term corresponding to sparsity of the training set photometry and a noise uncertainty term encompassing noisy photometric observations, enabling quantification of any need for more representative or more precise training samples.
+\gpz\ may also weight training set samples by importance according to $|z_{\rm{spec}} - z_{\rm{phot}}| / (1+z_{\rm{spec}})$ to minimize the normalized \pz\ point estimate error, however, this function may be adapted to \pzpdf s, pressuring the model to dedicate more resources to test set galaxies that are not well-represented in the training set.
 
-A unique feature in \textsc{GPz}, is that the variance estimate is composed of two terms each quantifying a different source of uncertainty.
-One term (the model uncertainty) reflects how much of the uncertainty is due to lack of training samples at the location of interest, whereas the second term (the noise uncertainty) reflects how much of the uncertainty is caused from observing many noisy samples at that location.
-Thus, the predictive variance can determine whether we need more representative samples or more precise samples for any particular location in the input space.
-\textsc{GPz} can also emphasize the importance of some samples as weights.
-This weight can be for example $|z_{\rm{spec}}-z_{\rm{phot}}|/(1+z_{\rm{spec}})$ to target the desired objective of minimizing the normalized redshift error or as a function of their probability in the test set relative to the training set in order to pressure the model to better fit samples that are rare in the training set but are expected to be abundant during testing.
-
-The data is prepared for \textsc{GPz} by taking the log of the magnitude errors, decorrelating the data set using PCA and imputing any missing magnitude values using a simple linear model that estimates the missing magnitudes given the observed ones.
-The log transformation helps to smooth the long tail distribution of the magnitude errors, which is more stable numerically and makes the optimization process unconstrained.
-The missing values are imputed by computing the mean of the training set $\mu$ and its covariance $\Sigma$, then we use the following equation to estimate the missing values from the observed ones %\[
-\begin{equation}
-x_{u} = \mu_{u}+\Sigma_{uo}\Sigma_{oo}^{-1}(x_{o}-\mu_{o}),
-\end{equation}  %\]
-where the subscript $o$ in $x_{o}$ indexes the \emph{observed} part of the input $x$, whereas the subscript $u$ indexes the \emph{unobserved} set (similarly for $\mu$ and $\Sigma$).
-This is the optimal expected value of the unobserved variables given the observed ones if the distribution is jointly Gaussian, note that if the variables are independent, i.e.
-$\Sigma_{uo}=0$, this will reduce to a simple average predictor.
-We use the Variable Covariance (VC) option in \textsc{GPz} with 200 basis functions after we note that there is no significant increase in the performance on the validation set (using 80\%-20\% training-validation split) and with no cost-sensitive learning applied.
+To smooth the long tail in the distribution of magnitude errors, we use the log of the magnitude errors, improving numerical stability and eliminating the need for constraints on the optimization process.
+% We use principal component analysis (PCA) to decorrelate the data
+% WHAT IS BEING DECORRELATED HERE?
+Unobserved magnitudes $x_{\rm u} = \mu_{\rm u} + \Sigma_{\rm uo}\Sigma_{\rm oo}^{-1}(x_{\rm o} - \mu_{\rm o})$ were imputed from observed magnitudes $x_{\rm o}$ and the training set mean $\mu$ and covariance $\Sigma$ using a linear model.
+This is the optimal expected value of the unobserved variables given the observed ones under the assumption that the distribution is jointly Gaussian; note that this reduces to a simple average if the covariates are independent with $\Sigma_{\rm uo} = 0$.
+We reserved for validation 20\% of the training set and used the Variable Covariance option in \textsc{GPz} with 200 basis functions, neglecting to apply cost-sensitive learning options.
 

### 2019-03-10 paper/tex/methods.tex
edited METAPhoR

@@
 %(Stefano Cavuoti, Massimo Brescia, Giuseppe Longo)
-%\textcolor{brown}{general review done on this section}
-\textsc{METAPhoR} \citep[Machine-learning Estimation Tool for Accurate Photometric Redshifts,][]{Cavuoti:17} is a pipeline designed to provide photo-$z$ point estimates and a reliable PDF for machine learning (ML) based techniques.
-It includes pre- and post-processing phases, hosting a photo-$z$ prediction engine based on the Multi Layer Perceptron with Quasi Newton Algorithm (MLPQNA).
+
+\textsc{METAPhoR} \citep[Machine-learning Estimation Tool for Accurate Photometric Redshifts,][]{Cavuoti:17} is based on the Multi Layer Perceptron with Quasi Newton Algorithm (MLPQNA) with the least square error model and Tikhonov $L_{2}$-norm regularization \citep{Hofmann:18}.
 %, already validated on photo-$z$'s in several cases \citep{de_Jong:17,Cavuoti:17b,Cavuoti:15,Brescia:14,Brescia:13,Biviano:13}.
+\Pzpdf s are generated by running $N$ trainings on the same training set, or $M$ trainings on $M$ different random samplings of the training set.
+Upon regression of the test set, the photometry $m_{ij}$ of each test set galaxy $j$ in filter $i$ is perturbed according to $m_{ij}' = m_{ij} + \alpha_{i} F_{ij} \epsilon$ in terms of the standard normal random variable $\epsilon \sim \mathcal{N}(0, 1)$, a multiplicative constant $\alpha_{i}$ permitting accommodation of multi-survey photometry, and a bimodal function $F_{ij}$ composed of a polynomial fit of the mean magnitude errors on the binned bands plus a constant term representing the threshold below which the polynomial's noise contribution is negligible \citep{Brescia:18}.
+% At a higher level, the pipeline mainly consists of three modules: (i) \textit{data pre-processing}, including a catalogue cross-matching sub-module \citep[based on the tool C3, ][]{Riccio:17}, a sub-module for photometric evaluation and error estimation of the multi-band catalogue used as Knowledge Base (KB), and a sub-module dedicated to the perturbation of the photometric KB, propaedeutic to the PDF estimation; (ii) \textit{photo-$z$ prediction}, which is the training/validation/test phase, producing the photo-$z$'s point estimates, based on a pre-selected ML method; (iii) \textit{PDF estimation}, specifically designed to calculate the PDF of the photo-$z$ estimation errors.
+% The last module includes also a post-processing tool, providing some statistics on the produced point estimates and PDFs.
 
-\textsc{METAPhoR} includes data modules for pre-processing, photo-$z$ estimation, and PDF estimation, and post-processing.
-The pre-processing includes a model for perturbation of the photometry that is employed in calculating the PDF of the photo-$z$ estimation errors.
-The photometric perturbation is defined as:
-\begin{equation}
- m_{ij} = m_{ij} + \alpha_{i}F_{ij}u_{\mu=0,\sigma=1}
-\end{equation}
-where $\alpha_{i}$ is a user selected multiplicative constant (useful in case of multi-survey photometry), $u_{\mu=0,\sigma=1}$ is a random value from the standard normal distribution and $F_{ij}$ is a bimodal function (a constant function + polynomial fitting of the mean magnitude errors on the binned bands), heuristically tuned in such a way that the constant component is the threshold under which the polynomial function is considered too low to provide a significant noise contribution to the photometry perturbation.
-
-%At a higher level, the pipeline mainly consists of three modules: (i) \textit{data pre-processing}, including a catalogue cross-matching sub-module \citep[based on the tool C3, ][]{Riccio:17}, a sub-module for photometric evaluation and error estimation of the multi-band catalogue used as Knowledge Base (KB), and a sub-module dedicated to the perturbation of the photometric KB, propaedeutic to the PDF estimation; (ii) \textit{photo-$z$ prediction}, which is the training/validation/test phase, producing the photo-$z$'s point estimates, based on a pre-selected ML method; (iii) \textit{PDF estimation}, specifically designed to calculate the PDF of the photo-$z$ estimation errors.
-The last module includes also a post-processing tool, providing some statistics on the produced point estimates and PDFs.
-
-%As introduced, the photo-$z$ point estimate prediction engine of \textsc{METAPhoR} is based on the MLPQNA model, whose photo-$z$ regression training error, used by the quasi Newton learning rule, is based on the least square error and Tikhonov $L_{2}$-norm regularization \citep{Hofmann:18}.
-
-As main prerogative, \textsc{METAPhoR} is able to provide a PDF for ML methods by taking into account the photometric errors provided with data, by running $N$ trainings on the same training set, or $M$ trainings on $M$ different random extractions from the KB.
-The different test sets, used to produce the PDF, are thus obtained by introducing a proper perturbation, parametrized from the photometric error distribution in each band, on the photometric data populating the original test set \citep{Brescia:18}.
-For the present work since it was required to produce a redshift (and a PDF) for each object of the test set we decided to apply a hierarchical kNN to replace the missing detections with values based on their neighbors.
-The reliability of PDFs and point estimation is lower. No cross validation has been used.
+In this work, we used a hierarchical KNN to replace nondetections with values based on their neighbors.
+The usual cross-validation step was also omitted for this study.
 

### 2019-03-10 paper/tex/methods.tex
edited SkyNet

@@
 In this study, we used \annz-v.2.0.4 to output only the result of the ANN algorithm.
-\Pzpdf s were produced by running an ensemble of 5 ANNs with 6 $ugrizy$ inputs, 2 hidden layers with 12 nodes each, and 1 output of redshift).
+\Pzpdf s were produced by running an ensemble of 5 ANNs with a $6:12:12:1$ architecture corresponding to the 6 $ugrizy$ inputs, 2 hidden layers with 12 nodes each, and 1 output of redshift.
 Each of the five ANNs was trained with different random seeds for the initialization of input parameters.
@@
 
-\textsc{SkyNet}\footnote{\url{http://ccpforge.cse.rl.ac.uk/gf/project/skynet/}} \citep{Graff:14} is a publicly available neural network software, based on a 2nd order conjugate gradient optimization scheme \citep[see][for further details]{Graff:14}.
+\skynet \footnote{\url{http://ccpforge.cse.rl.ac.uk/gf/project/skynet/}} \citep{Graff:14} employs a neural network based on a second order conjugate gradient optimization scheme \citep[see][for further details]{Graff:14}.
 %It has been used efficiently for redshift PDF estimates \citep{Sanchez:14,Bonnett:15,Bonnett:16}.
+The neural network is configured as a standard multilayer perceptron with three hidden layers and one input layer with 12 nodes corresponding to the 6 photometric magnitudes and their measurement errors.
+We use \skynet\ as a regressor for \pz\ point estimation and as a classifier for \pzpdf\ estimation.
 
-%In the present work, we use \textsc{SkyNet} both as a regressor (for redshift point estimates) and a classifier (for PDF estimates).
-The neural network is configured as a standard multilayer perceptron with three hidden layers and one input layer with $12$ nodes (the $6$ magnitudes and their errors).
-%For the regressor, the output layer is a single linear node, while each hidden layer has $10$ nodes with a tanh activation function.
-The classifier is laid out such that the hidden layers have 20:40:40 nodes each, all rectified linear units, and the output layer has $200$ nodes (corresponding to 200 bins for the PDF) activated with a ``softmax'' function so that they automatically sum to $1$.
-While previous implementations of the code, such as \citet{Sanchez:14} and \citet{Bonnett:15} (see Appendix C.3), implement a ``sliding bin'' smoothing, no such procedure was used in this study.
+The regressor used a standard \chisq\ error function with a single linear node as the output layer and 10 nodes with a $\tanh$ activation function for each hidden layer.
+The classifier used a cross-entropy error function with a 20:40:40 node (all rectified linear units) architecture for each hidden layer and an output layer of 200 nodes corresponding to 200 bins for the PDF, with a softmax activation function to enforce the normalization condition that the probabilities sum to unity.
+While previous implementations of the code \citep[see Appendix C.3 of~][]{Sanchez:14,Bonnett:15} implement a sliding bin smoothing, no such procedure was used in this study.
 
-To avoid over-fitting, a $30$ per cent fraction of the training set is used as validation, and the training is stopped as soon as the error rate begins to increase in the validation set.
-The weights are randomly initialized based on normal sampling.
-The error function is a standard chi-square function for the regressor, and a cross-entropy function for the classifier.
-Finally, the data are all whitened before processing, with magnitudes pegged to (45,45,40,35,42,42) and their errors pegged to (20,20,10,5,15,15) for $ugrizy$ filters, respectively.
+We pre-whitened the data by pegging the magnitudes to (45,45,40,35,42,42) and errors to (20,20,10,5,15,15) for $ugrizy$ filters, respectively.
+% NO CLUDE WHAT THIS MEANS
+To avoid over-fitting, $30\%$ of the training set was reserved for validation, and training was halted as soon as the error rate began to increase on the validation set.
+The weights were randomly initialized based on normal sampling.
+% WHAT WEIGHTS?
 

### 2019-03-10 paper/tex/methods.tex
URLs

@@
 
-\lephare \footnote{\url{http://www.cfht.hawaii.edu/~arnouts/lephare.html}}\citep[Photometric Analysis for Redshift Estimate,][]{Arnouts:99,Ilbert:06} matches observed colors with those predicted from a template set, which can be semi-empirical or entirely synthetic, directly according to the likelihood $\chi^{2}(z, T, A) \equiv \sum_{b}^{N_{\rm{filt}}} \left((F^{\rm{obs}}_{b} - A F^{\rm{mod}}_{b}(T, z)) / \sigma^{\rm{obs}}_{b}\right)^{2}$ of normalization factor $A$, template $T$, and redshift $z$.
+Photometric Analysis for Redshift Estimate \citep[\lephare\footnote{\url{http://www.cfht.hawaii.edu/~arnouts/lephare.html}},][]{Arnouts:99,Ilbert:06} matches observed colors with those predicted from a template set, which can be semi-empirical or entirely synthetic, directly according to the likelihood $\chi^{2}(z, T, A) \equiv \sum_{b}^{N_{\rm{filt}}} \left((F^{\rm{obs}}_{b} - A F^{\rm{mod}}_{b}(T, z)) / \sigma^{\rm{obs}}_{b}\right)^{2}$ of normalization factor $A$, template $T$, and redshift $z$.
 In words, the likelihood is a sum of observed flux error $\sigma_{b}^{\rm{obs}}$-weighted squared differences between the observed flux $F^{\rm{obs}}_{b}$ and the normalized predicted flux $F^{\rm{mod}}_{b}(T, z)$ in $N_{\rm{filt}}$ photometric filters $b$, which is the \lsst\ $ugrizy$ filters in this case.
@@
 
-\bpz \footnote{\url{http://www.stsci.edu/~dcoe/BPZ/}} \citep[Bayesian Photometric Redshift,][]{Benitez:00} determines the likelihood $p(C \vert z, T)$ of a galaxy's observed colours $C$ for a set of SED templates $T$ at redshifts $z$.
+Bayesian Photometric Redshift \citep[\bpz\footnote{\url{http://www.stsci.edu/~dcoe/BPZ/}},][]{Benitez:00} determines the likelihood $p(C \vert z, T)$ of a galaxy's observed colours $C$ for a set of SED templates $T$ at redshifts $z$.
 The \bpz\ likelihood is related to the \chisq\ likelihood by $p(C \vert z, T) \propto \exp[- \chi^{2} / 2]$.
@@
 
-\eazy \footnote{\url{https://github.com/gbrammer/eazy-photoz}} \citep[Easy and Accurate Photometric Redshifts from Yale,][]{Brammer:08} extends the basic \chisq\ fit procedure that defines template-fitting approaches.
+Easy and Accurate Photometric Redshifts from Yale \citep[\eazy\footnote{\url{https://github.com/gbrammer/eazy-photoz}},][]{Brammer:08} extends the basic \chisq\ fit procedure that defines template-fitting approaches.
 The algorithm models the observed photometry with a linear combination of template SEDs at each redshift.
@@
 
-\textsc{METAPhoR} \citep[Machine-learning Estimation Tool for Accurate Photometric Redshifts,][]{Cavuoti:17} is based on the Multi Layer Perceptron with Quasi Newton Algorithm (MLPQNA) with the least square error model and Tikhonov $L_{2}$-norm regularization \citep{Hofmann:18}.
+Machine-learning Estimation Tool for Accurate Photometric Redshifts \citep[\metaphor\footnote{\url{http://dame.dsf.unina.it}},][]{Cavuoti:17} is based on the Multi Layer Perceptron with Quasi Newton Algorithm (MLPQNA) with the least square error model and Tikhonov $L_{2}$-norm regularization \citep{Hofmann:18}.
 %, already validated on photo-$z$'s in several cases \citep{de_Jong:17,Cavuoti:17b,Cavuoti:15,Brescia:14,Brescia:13,Biviano:13}.
@@
 
-\textsc{TPZ}\footnote{\url{https://github.com/mgckind/MLZ}} \citep[Trees for Photo-$z$,][]{Carrasco_Kind:13,Carrascokind:14} is a parallel machine learning algorithm that generates photometric redshift PDFs using prediction trees and random forest techniques.
+Trees for \Pz \citep[\tpz\footnote{\url{https://github.com/mgckind/MLZ}},][]{Carrasco_Kind:13,Carrascokind:14} is a parallel machine learning algorithm that generates photometric redshift PDFs using prediction trees and random forest techniques.
 The code recursively splits the input data (i.~e.~the training sample), into two branches, one after another, until a terminal leaf is created that meets a termination criterion (e.~g.~a minimum leaf size or a variance threshold).

### 2019-03-10 paper/tex/methods.tex
edited TPZ

@@
 
-Trees for \Pz \citep[\tpz\footnote{\url{https://github.com/mgckind/MLZ}},][]{Carrasco_Kind:13,Carrascokind:14} is a parallel machine learning algorithm that generates photometric redshift PDFs using prediction trees and random forest techniques.
-The code recursively splits the input data (i.~e.~the training sample), into two branches, one after another, until a terminal leaf is created that meets a termination criterion (e.~g.~a minimum leaf size or a variance threshold).
-Bootstrap samples from the training data and associated errors are used to build a set of prediction trees.
-In order to minimize correlation between the trees, the data is divided in such a way that the highest information gain among the random subsample of features is obtained at every point.
-The regions in each terminal leaf node corresponds to a specific subsample of the entire data that possesses similar properties.
+Trees for \Pz \citep[\tpz\footnote{\url{https://github.com/mgckind/MLZ}},][]{Carrasco_Kind:13,Carrascokind:14} uses prediction trees and random forest techniques to estimate \pzpdf s.
+\tpz\ recursively splits the training set into branch pairs based on maximizing information gain among a random subsample of features, to minimize correlation between the trees, terminating only when a newly created leaf meets a criterion, such as a leaf size minimum or a variance threshold.
+The regions in each terminal leaf node correspond to a subsample of the training set with similar properties.
+Bootstrap samples from the training set photometry and errors are used to build a set of prediction trees.
 
-The training data is examined before running TPZ.
-Since TPZ does not handle non-detections (magnitudes flagged as 99.0), we replace these values with an approximation of the $1\sigma$ detection threshold, i.~e.~a signal to noise ratio of 1 in terms of magnitude uncertainty using the equation $dm = 2.5 ~ \log ( 1 + N/S )$ where $dm \sim 0.7526 ~ mag$ for $N/S=1$.
-That is, for each band, we replace the non-detection with the magnitude corresponding to the error of 0.7526 from the error model forecasted for 10-year LSST data.
-The Out-of-Bag \citep{Breiman:84,Carrasco_Kind:13} cross-validation technique is used within TPZ to evaluate its predictive validity and determine the relative importance of the different input attributes.
-We employed this information to calibrate our algorithm.
-
-In the present work, the LSST magnitudes $u,~g,~r,~i$ and colours $u-g,~g-r,~r-i,~i-z,~z-y$ and their associated errors are used in the process of growing 100 trees with a minimum leaf size of 5 (the $z$ and $y$ magnitudes did not show significant correlation with the redshift in our cross-validation, so we did not use them when constructing our trees).
+To run \tpz, we replaced nondetections with an approximation of the $1\sigma$ detection threshold based on the error forecast of the 10-year \lsst\ data, i.~e.
+$dm = 2.5 \log (1 + N/S)$ where $dm \sim 0.7526$ mag for $N/S = 1$.
+We calibrated \tpz\ with the Out-of-Bag cross-validation technique \citep{Breiman:84,Carrasco_Kind:13} to evaluate its predictive validity and determine the relative importance of the different input attributes.
+We grew 100 trees to a minimum leaf size of 5 using the $ugri$ magnitudes, all $u-g, g-r, r-i, i-z, z-y$ colours, and the associated errors, as the $z$ and $y$ magnitudes did not show significant correlation with the redshift in our cross-validation.
 We partitioned our redshift space into 200 bins and smoothed each individual PDF with a smoothing scale of twice the bin size.
-%there was a typo here that had 100 bins, I double checked, it is 200 in the final run.
 

### 2019-03-11 paper/tex/methods.tex
partially through Delight

@@
 
-\delight \footnote{\url{https://github.com/ixkael/Delight}} \citep{Leistedt:17} infers \pz s with a data-driven model of latent SEDs and a physical model of photometric fluxes as a function of redshift.
-Delight models the underlying latent SEDs as a linear combination of a set of pre-defined template SEDs, plus zero mean Gaussian processes with factorized kernels.
-Generally, machine learning methods rely on representative training data with similar band passes, while template based methods rely on a complete library of templates based on physical models constructed.
-\textsc{Delight} is constructed in attempt to combine the advantages and eliminate the disadvantages of both template-based and machine learning algorithms: it constructs a large collection of latent SED templates (or physical flux-redshift models) from training data, with a template SED library as a guide to the learning of the model.
-The advantage of \textsc{Delight} is that it neither needs representative training data in the same photometric bands, nor does it need detailed galaxy SED models to work.
+\delight \footnote{\url{https://github.com/ixkael/Delight}} \citep{Leistedt:17} is a hybrid technique that infers \pz s with a data-driven model of latent SEDs and a physical model of photometric fluxes as a function of redshift.
+Generally, machine learning methods rely on representative training data with shared band passes, while template based methods rely on a complete library of templates based on physical models constructed.
+\delight\ takes the best aspects of both approaches by constructing a large collection of latent SED templates (or physical flux-redshift models) from training data, with a template SED library as a guide to the learning of the model, thereby circumventing the machine learning prerequisite of representative training data in the same photometric bands and the template fitting requirement of detailed galaxy SED models.
+This conceptually novel approach is achieved with Gaussian processes operating in flux-redshift space.
 
-This conceptually novel approach is done by using Gaussian processes operating in flux-redshift space.
-The posterior distribution on the redshift of a target galaxy is obtained via a pairwise comparison with training galaxies,
+Each test set galaxy $i$ has a posterior $p(z \vert \mathbf{\hat{F}}) \approx \sum_i p(\mathbf{\hat{F}} \vert z, T_i) p(z \vert T_i) p(T_i)$ of redshift $z$ conditioned on noisy flux $\mathbf{\hat{F}}$, where $p(z \vert T_i) p(T_i)$ captures prior information about the redshift distributions and abundances of the galaxy templates $T_i$.
+As in traditional template fitting, each likelihood $p(\mathbf{\hat{F}} \vert \mathbf{F})$ relates the noisy flux $\mathbf{\hat{F}}$ with the noiseless flux $\mathbf{F}$ predicted by the model of a linear combination of templates, carefully constructed to account for model uncertainties and different normalization of the same SED.
+The machine learning approach appears in the inclusion of a pairwise comparison term $p(\mathbf{F} \vert z, z_j, \mathbf{\hat{F}}_j)$ for the prediction of flux at a redshift $z$ with respect to training set galaxy $j$ with redshift $z_j$ and flux $\mathbf{\hat{F}}_j$.
+Thus the \pz\ posterior $p(\mathbf{\hat{F}} \vert z, T_i) = \int p(\mathbf{\hat{F}} \vert \mathbf{F}) p(\mathbf{F} \vert z, z_j, \mathbf{\hat{F}}_j) d\mathbf{F}$ is essentially the probability that the training and the target galaxies have the same SED at a different redshift.
+The flux prediction $p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)$ of the training galaxy at redshift $z$ is modeled via a Gaussian process,
 
-\begin{equation}
-p(z|\mathbf{\hat{F}}) \approx \sum_i p(\mathbf{\hat{F}}|z,t_i)\, p(z|t_i)p(t_i),
-\end{equation}
-\noindent where $p(z|t_i)p(t_i)$ captures prior information about the redshift distributions and abundances of the galaxies, with $t_i$ denoting the galaxy template; while $p(\mathbf{\hat{F}}|z,t_i)$ is the posterior of noisy flux $\mathbf{\hat{F}}$ at redshift $z$.
-For each training-target pair, $p(\mathbf{\hat{F}}|z,t_i)$ is evaluated as follows:
-\begin{equation} \label{eq:delight_noisy}
-p(\mathbf{\hat{F}}|z,t_i) = \int p(\mathbf{\hat{F}}|\mathbf{F})\, p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)\, d\mathbf{F},
-\end{equation}
-where $p(\mathbf{\hat{F}}|\mathbf{F})$ is the likelihood function, it compares the noisy real flux $\mathbf{\hat{F}}$ with the noiseless flux $\mathbf{F}$ obtained from the linear combination of template models, carefully constructed to account for model uncertainties and different normalization of the same SED; while $p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)$ is the prediction of flux at a different redshift $z$ with respect to the training object with redshift $z_i$ and flux $\mathbf{\hat{F}}_i$.
-Eq.~\ref{eq:delight_noisy} is essentially the probability that the training and the target galaxies having the same SED but at a different redshift.
-The flux prediction $p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)$ of the training galaxy at redshift $z$ is modeled via a Gaussian process,
+The underlying latent SEDs are modeled as a linear combination of a set of pre-defined template SEDs, plus zero mean Gaussian processes with factorized kernels.
 

### 2019-03-11 paper/tex/methods.tex
partially through Delight

@@
 Generally, machine learning methods rely on representative training data with shared band passes, while template based methods rely on a complete library of templates based on physical models constructed.
-\delight\ takes the best aspects of both approaches by constructing a large collection of latent SED templates (or physical flux-redshift models) from training data, with a template SED library as a guide to the learning of the model, thereby circumventing the machine learning prerequisite of representative training data in the same photometric bands and the template fitting requirement of detailed galaxy SED models.
-This conceptually novel approach is achieved with Gaussian processes operating in flux-redshift space.
+\delight\ aims to take the best aspects of both approaches by constructing a large collection of latent SED templates (or physical flux-redshift models) from training data, with a template SED library as a guide to the learning of the model, thereby circumventing the machine learning prerequisite of representative training data in the same photometric bands and the template fitting requirement of detailed galaxy SED models.
 
@@
 As in traditional template fitting, each likelihood $p(\mathbf{\hat{F}} \vert \mathbf{F})$ relates the noisy flux $\mathbf{\hat{F}}$ with the noiseless flux $\mathbf{F}$ predicted by the model of a linear combination of templates, carefully constructed to account for model uncertainties and different normalization of the same SED.
+\delight\ models SEDs as a linar combination of template SEDs plus Gaussian processes $F_b \sim \mathcal{GP}\left( \mu^F, k^F \right)$ with zero mean function $\mu^{F}$ and a physically motivated kernel $k^{F}$ that induces realistic correlations in flux-redshift space.
+
 The machine learning approach appears in the inclusion of a pairwise comparison term $p(\mathbf{F} \vert z, z_j, \mathbf{\hat{F}}_j)$ for the prediction of flux at a redshift $z$ with respect to training set galaxy $j$ with redshift $z_j$ and flux $\mathbf{\hat{F}}_j$.
-Thus the \pz\ posterior $p(\mathbf{\hat{F}} \vert z, T_i) = \int p(\mathbf{\hat{F}} \vert \mathbf{F}) p(\mathbf{F} \vert z, z_j, \mathbf{\hat{F}}_j) d\mathbf{F}$ is essentially the probability that the training and the target galaxies have the same SED at a different redshift.
-The flux prediction $p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)$ of the training galaxy at redshift $z$ is modeled via a Gaussian process,
+Thus the \pz\ posterior $p(\mathbf{\hat{F}} \vert z, T_i) = \int p(\mathbf{\hat{F}} \vert \mathbf{F}) p(\mathbf{F} \vert z, z_j, \mathbf{\hat{F}}_j) d\mathbf{F}$ may be interpreted as the probability that the training and the target galaxies have the same SED at a different redshift.
+The flux prediction $p(\mathbf{F} \vert z, z_j, \mathbf{\hat{F}}_j)$ of the training galaxy at redshift $z$ is modeled via a Gaussian process,
 
-The underlying latent SEDs are modeled as a linear combination of a set of pre-defined template SEDs, plus zero mean Gaussian processes with factorized kernels.
-
-\begin{equation} \label{eq:delight_gp}
-F_b \sim \mathcal{GP}\left( \mu^F,k^F \right),
-\end{equation}
-\noindent with mean function $\mu^F$ and kernel $k^F$, both imposed to capture expected correlations resulting from the known underlying physics (i.e., fluxes resulting from observing SEDs through filter response, and the SEDs being redshifted).
 The reader should refer to \citet{Leistedt:17} for further details.

### 2019-03-14 paper/tex/methods.tex
finished delight

@@
 \delight \footnote{\url{https://github.com/ixkael/Delight}} \citep{Leistedt:17} is a hybrid technique that infers \pz s with a data-driven model of latent SEDs and a physical model of photometric fluxes as a function of redshift.
-Generally, machine learning methods rely on representative training data with shared band passes, while template based methods rely on a complete library of templates based on physical models constructed.
+Generally, machine learning methods rely on representative training data with shared photometric filters, while template based methods rely on a complete library of templates based on physical models constructed.
 \delight\ aims to take the best aspects of both approaches by constructing a large collection of latent SED templates (or physical flux-redshift models) from training data, with a template SED library as a guide to the learning of the model, thereby circumventing the machine learning prerequisite of representative training data in the same photometric bands and the template fitting requirement of detailed galaxy SED models.
+It models noisy observed flux $\mathbf{\hat{F}} = \mathbf{F} + F_{b}$ as a sum of a noiseless flux plus a Gaussian processes $F_b \sim \mathcal{GP}\left(\mu^F, k^F \right)$ with zero mean function $\mu^{F}$ and a physically motivated kernel $k^{F}$ that induces realistic correlations in flux-redshift space.
 
-Each test set galaxy $i$ has a posterior $p(z \vert \mathbf{\hat{F}}) \approx \sum_i p(\mathbf{\hat{F}} \vert z, T_i) p(z \vert T_i) p(T_i)$ of redshift $z$ conditioned on noisy flux $\mathbf{\hat{F}}$, where $p(z \vert T_i) p(T_i)$ captures prior information about the redshift distributions and abundances of the galaxy templates $T_i$.
-As in traditional template fitting, each likelihood $p(\mathbf{\hat{F}} \vert \mathbf{F})$ relates the noisy flux $\mathbf{\hat{F}}$ with the noiseless flux $\mathbf{F}$ predicted by the model of a linear combination of templates, carefully constructed to account for model uncertainties and different normalization of the same SED.
-\delight\ models SEDs as a linar combination of template SEDs plus Gaussian processes $F_b \sim \mathcal{GP}\left( \mu^F, k^F \right)$ with zero mean function $\mu^{F}$ and a physically motivated kernel $k^{F}$ that induces realistic correlations in flux-redshift space.
+From a template-fitting perspective, each test set galaxy has a posterior $p(z \vert \mathbf{\hat{F}}) \approx \sum_i p(\mathbf{\hat{F}} \vert z, T_i) p(z \vert T_i) p(T_i)$ of redshift $z$ conditioned on noisy flux $\mathbf{\hat{F}}$, where $p(z \vert T_i) p(T_i)$ captures prior information about the redshift distributions and abundances of the galaxy templates $T_i$.
+As in traditional template fitting, each likelihood $p(\mathbf{\hat{F}} \vert \mathbf{F})$ relates the noisy flux $\mathbf{\hat{F}}$ with the noiseless flux $\mathbf{F}$ predicted by the model of a linear combination of templates, carefully constructed to account for model uncertainties and different normalization of the same SED, plus the Gaussian process term.
 
-The machine learning approach appears in the inclusion of a pairwise comparison term $p(\mathbf{F} \vert z, z_j, \mathbf{\hat{F}}_j)$ for the prediction of flux at a redshift $z$ with respect to training set galaxy $j$ with redshift $z_j$ and flux $\mathbf{\hat{F}}_j$.
-Thus the \pz\ posterior $p(\mathbf{\hat{F}} \vert z, T_i) = \int p(\mathbf{\hat{F}} \vert \mathbf{F}) p(\mathbf{F} \vert z, z_j, \mathbf{\hat{F}}_j) d\mathbf{F}$ may be interpreted as the probability that the training and the target galaxies have the same SED at a different redshift.
-The flux prediction $p(\mathbf{F} \vert z, z_j, \mathbf{\hat{F}}_j)$ of the training galaxy at redshift $z$ is modeled via a Gaussian process,
+The machine learning approach appears in the inclusion of a pairwise comparison term $p(\mathbf{F} \vert z, z_j, \mathbf{\hat{F}}_j)$ for the prediction of model flux $\mathbf{F}$ at a model redshift $z$ with respect to training set galaxy $j$ with redshift $z_j$ and observed flux $\mathbf{\hat{F}}_j$.
+Thus the \pz\ posterior $p(\mathbf{\hat{F}} \vert z, T_i) = \int p(\mathbf{\hat{F}} \vert \mathbf{F}) p(\mathbf{F} \vert z, z_j, \mathbf{\hat{F}}_j) d\mathbf{F}$ may be interpreted as the probability that the training and the target galaxies have the same SED at different redshifts.
+The flux prediction $p(\mathbf{F} \vert z, z_j, \mathbf{\hat{F}}_j)$ of the training galaxy at redshift $z$ is modeled via the Gaussian process described above; more detail is provided in \citet{Leistedt:17}.
 
-The reader should refer to \citet{Leistedt:17} for further details.
-
-In this study, all $100$ ordered Buzzard templates, as described in Section~\ref{sec:buzztemplates}, were used in \textsc{Delight}, and the Gaussian process was trained using the provided training sample.
+In this study, the default settings of \delight\ were used, with the exception that the PDF bins were set to be linearly-spaced rather than logarithmic.
+The Gaussian process was trained using the full DC1 training set.
+We used the full DC1 template set with a flat prior in magnitude and SED type.
 Photometric uncertainties from the inputs are propagated into the code, while non-detections for each band are set to the mean of the respective bands.
-The default settings of \textsc{Delight} were used, with the exception that the PDF bins were set to be linearly-spaced rather than logarithmic.
-In this study a flat prior in magnitude/type is assumed.
 

### 2019-03-14 paper/tex/methods.tex
edited trainZ

@@
 
-\subsection{trainZ: a straw-person \pz\ estimator}
+\subsection{trainZ: a pathological \pz\ estimator}
 \label{sec:method:trainz}
 
-In addition to the main photo-$z$ algorithms described above we also include a very simple method as a pathological example.
-For \trainz, as we will we call this simple estimator, we well define $p(z)$ as simply:
-\begin{equation}
-p(z) = \frac{1}{N_{ \mathrm train}}\sum_{\mathrm i=1}^{N_{\mathrm train}}z_{\mathrm train}
-\end{equation}
-That is, we simply set the redshift PDF of every galaxy equal to the normalized $N(z)$ of the training sample.
-This estimator is essentially a k nearest-neighbour estimator with k equal to the number of galaxies in the training sample.
-As the training sample is drawn from the same underlying distribution as the test sample, modulo small deviations due to sample size, the quantiles of the training and test distributions should be identical, modulo fluctuations due to finite sample size.
-This is a wildly unrealistic estimator, as it assigns all galaxies, no matter their apparent magnitude, colour, or true redshift, the same redshift PDF, and is thus uninformative at the level of individual object redshifts, but is designed to perform very well for the ensemble of all objects.
-If the training set was not representative, this estimator would produce biased results, and any attempts to break up the sample into tomographic bins will fail, as every galaxy has an identical $p(z)$.
-We will discuss this method and cautions relative to metrics in Section~\ref{sec:caution}.
+We also include as an experimental control a pathological \pzpdf\ estimation method, dubbed \trainz, which assigns each test set galaxy a \pzpdf\ equal to the normalized redshift distribution $N(z)$ of the training set, according to $p(z \vert \{z_{j}\}) \equiv \frac{1}{N_{ \mathrm train}}\sum_{\mathrm i=1}^{N_{\mathrm train}} \begin{cases} 1 & \text{if\ } z_{k}\leq z_{i} < z_{k+1}\\ 0 & \text{otherwise} \end{cases}$.
+Unlike the other methods, the \trainz\ estimator is \textit{independent of the photometric data} of the test set galaxies, effectively performing a KNN procedure with $k=N_{\rm train}$.
+
+As the training sample is drawn from the same underlying distribution as the test sample, modulo the Poisson error due to small sample size, the quantiles of the training and test set distributions should be identical.
+If the training set were not representative, this estimator would produce biased results.
+We will evaluate its performance under the metrics of Section~\ref{sec:metrics} discuss it as an illustrative control case in Section~\ref{sec:caution} to highlight the limitations of metrics of \pzpdf s.
+% any attempts to break up the sample into tomographic bins will fail, as every galaxy has an identical $p(z)$.

### 2019-03-14 paper/tex/methods.tex
finalizing methods section edits

@@
 
-\flexzboost \footnote{\url{https://github.com/tpospisi/flexcode};  \url{https://github.com/rizbicki/FlexCoDE} \label{flexzboost_github}} \citep{Izbicki:17} is built on \texttt{FlexCode}, a general-purpose methodology for converting any conditional mean point estimator of $z$ to a conditional density estimator $p(z \vert \x) \equiv f(z \vert \x)$, where $\x$ here represents our photometric covariates and errors.
+\flexzboost \footnote{\url{https://github.com/tpospisi/flexcode};\\ \url{https://github.com/rizbicki/FlexCoDE} \label{flexzboost_github}} \citep{Izbicki:17} is built on \texttt{FlexCode}, a general-purpose methodology for converting any conditional mean point estimator of $z$ to a conditional density estimator $p(z \vert \x) \equiv f(z \vert \x)$, where $\x$ here represents our photometric covariates and errors.
 \flexzboost\ expands the unknown function $f(z \vert \x) = \sum_{i}\beta_{i}(\x)\phi_{i}(z)$ using an orthonormal basis $\{\phi_{i}(z)\}_{i}$.
@@
 
-We also include as an experimental control a pathological \pzpdf\ estimation method, dubbed \trainz, which assigns each test set galaxy a \pzpdf\ equal to the normalized redshift distribution $N(z)$ of the training set, according to $p(z \vert \{z_{j}\}) \equiv \frac{1}{N_{ \mathrm train}}\sum_{\mathrm i=1}^{N_{\mathrm train}} \begin{cases} 1 & \text{if\ } z_{k}\leq z_{i} < z_{k+1}\\ 0 & \text{otherwise} \end{cases}$.
+We also consider a pathological \pzpdf\ estimation method, dubbed \trainz, which assigns each test set galaxy a \pzpdf\ equal to the normalized redshift distribution $N(z)$ of the training set, according to
+\begin{equation}
+	p(z \vert \{z_{j}\}) \equiv \frac{1}{N_{ \mathrm train}}\sum_{\mathrm i=1}^{N_{\mathrm train}} \begin{cases} 1 & \text{if\ } z_{k}\leq z_{i} < z_{k+1}\\ 0 & \text{otherwise} \end{cases}.
+\end{equation}
 Unlike the other methods, the \trainz\ estimator is \textit{independent of the photometric data} of the test set galaxies, effectively performing a KNN procedure with $k=N_{\rm train}$.
 
-As the training sample is drawn from the same underlying distribution as the test sample, modulo the Poisson error due to small sample size, the quantiles of the training and test set distributions should be identical.
-If the training set were not representative, this estimator would produce biased results.
-We will evaluate its performance under the metrics of Section~\ref{sec:metrics} discuss it as an illustrative control case in Section~\ref{sec:caution} to highlight the limitations of metrics of \pzpdf s.
+Though \trainz\ is strongly vulnerable to a nonrepresenative training set, it should optimize performance metrics probing the ensemble properties of the galaxy sample, modulo Poisson error due to small sample size, as the training set and test set are drawn from the same underlying population.
+We will demonstrate its performance under the metrics of Section~\ref{sec:metrics} and discuss it as an illustrative experimental control case in Section~\ref{sec:caution} to highlight the limitations of our evaluation criteria for \pzpdf s.
 % any attempts to break up the sample into tomographic bins will fail, as every galaxy has an identical $p(z)$.

### 2019-03-14 paper/tex/methods.tex
finalizing methods section edits

@@
 \end{equation}
-Unlike the other methods, the \trainz\ estimator is \textit{independent of the photometric data} of the test set galaxies, effectively performing a KNN procedure with $k=N_{\rm train}$.
+Unlike the other methods, the \trainz\ estimator is \textit{independent of the photometric data}, effectively performing a KNN procedure with $k=N_{\rm train}$.
 

### 2019-03-14 paper/tex/introduction.tex
macros

@@
 
-To fully characterize such physical degeneracies, photometric galaxy catalogue data releases from \citep{Mandelbaum:2008} to \citep{de_Jong:17}, provided a more informative \pz\ data product, the \pz\ probability density function (PDF), that describes the redshift probability $p(z)$ as a function of a galaxy's redshift, conditioned on the observed photometry.
+To fully characterize such physical degeneracies, photometric galaxy catalogue data releases from \citep{Mandelbaum:2008} to \citep{de_Jong:17}, provided a more informative \pz\ data product, the \pz\ probability density function (PDF), that describes the redshift probability, commonly denoted as $p(z)$, as a function of a galaxy's redshift, conditioned on the observed photometry.
 Early template-based methods such as \citet{Fernandezsoto:99} approximated the likelihood of photometry conditioned on redshift with the relative \chisq\ values of template spectra.
@@
 
-This paper is distinguished by its focus on the evaluation criteria for \pzpdf s and interpretation thereof, as a key project of the Photometric Redshift working group of the \lsstdesc\ that aims to perform a comprehensive sensitivity analysis of \pzpdf\ techniques in order to select those that will become part of the \lsst\ pipelines, as laid out in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}}.
+This paper is distinguished by its focus on the evaluation criteria for \pzpdf s and interpretation thereof, as a key project of the Photometric Redshifts working group of the \lsstdesc\ that aims to perform a comprehensive sensitivity analysis of \pzpdf\ techniques in order to ultimately select those that will become part of the \lsst\ pipelines, as laid out in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}}.
 In this initial study, we focus on evaluating the performance of \pzpdf\ codes using PDF-specific performance metrics in a controlled experiment with complete and representative prior information (template libraries and training sets) to set a baseline for subsequent investigations.

### 2019-03-14 paper/tex/analysis.tex
removing cosmology emphasis from preamble

@@
-\section{Metrics for quantifying PDF comparisons}
+\section{Analysis}
 \label{sec:metrics}
@@
 
-The overloaded ``$p(z)$'' is a widespread abuse of notation; we would like the outputs of \pzpdf\ codes to be interpretable as probabilities.
+% All code runners were asked to output redshift posterior estimates on 200 linear-spaced bins between redshifts 0 and 2.
+
+The goal of this study is to evaluate the degree to which \pzpdf s of each method can be trusted for a generic probabilistic data analysis.
 Obviously \pzpdf s must not take negative values and must integrate to unity over the range of possible redshifts.
-Additionally, an estimator derived by method $H$ for the \pzpdf\ of galaxy $i$ must be understood as a posterior probability distribution
+The overloaded ``$p(z)$'' is a widespread abuse of notation that obfuscates this goal, so we will dedicate some attention to breaking it apart.
+Galaxies have redshifts $z$ and photometric data $d$ drawn from a joint probability space $p(z, d)$ in nature that is not known to us.
+If it were known, however, each galaxy $i$ would have a \textit{true posterior \pzpdf}\ $p(z \vert d_{i})$ as well as a true likelihood $p(d \vert z_{i})$.
+There are a number of metrics that can be used to test the accuracy of a \pz\ posterior as an estimator of a true \pz\ posterior if it is known.
+Even for simulated data, the true \pzpdf\ is in general not accessible unless the photometry is in fact drawn from the true \pz\ likelihoods, a mock catalogue generation procedure that has not yet appeared in the literature.
+Furthermore, only limited applications of \pzpdf s that could be used as the basis for a metric have been presented in the literature.
+
+A \pzpdf\ estimator derived by method $H$ must be understood as a posterior probability distribution
 \begin{equation}
-\hat{p}_{j}(z_{i}) = p(z | d_{i}, I_{D}, I_{H}),
+\hat{p}^{H}_{i}(z) \equiv p(z \vert d_{i}, I_{D}, I_{H}),
 \end{equation}
@@
 
-In this study, we isolate the differences in prior information specific to each method by using a single training set $I_{D}^{ML}$ for all machine learning-based codes and a single template library $I_{D}^{T}$ for all template-based codes, and these sets of prior information are carefully constructed to be representative and complete, we have $I_{D}^{ML}\equiv I_{D}^{T}$ for every method $H$.
+In this study, we isolate the effect of differences in prior information $I_{H}$ specific to each method by using a single training set $I_{D}^{ML}$ for all machine learning-based codes and a single template library $I_{D}^{T}$ for all template-based codes.
+These sets of prior information are carefully constructed to be representative and complete, so we have $I_{D} \equiv I_{D}^{ML} \equiv I_{D}^{T}$ for every method $H$.
 Thus, we are saying
 \begin{equation}
-\frac{\hat{p}_{i, H}(z)}{\hat{p}_{i, H'}(z)} \approx \frac{p(z | d_{i}, I_{H})}{p(z | d_{i}, I_{H'})},
+\frac{\hat{p}_{i}^{H}(z)}{\hat{p}_{i}^{H'}(z)} \approx \frac{p(z \vert d_{i}, I_{H})}{p(z \vert d_{i}, I_{H'})},
 \end{equation}
-meaning that we assume comparisons of $\hat{p}_{i, H}(z)$ isolate the effect of the method used to obtain the estimator, which should make examination of differences caused by specifics of the method implementations easier to isolate.
+meaning that we assume comparisons of $\hat{p}_{i}^{H}(z)$ isolate the effect of the method used to obtain the estimator, which should enable interpetation of the differences between estimated PDFs in terms of the specifics of the method implementations.
 
-As mentioned previously, there are cosmology probes that require knowledge of individual galaxy's \pzpdf, for example galaxy intrinsic alignment studies, strong lensing foreground shear prediction, and photometric supernova classification, and others that require only knowledge of the ensemble redshift distribution, $N(z)$.
-Due to the paucity of principled techniques for using and validating individual galaxy \pzpdf s, there have been few alternatives to the common practice of reducing \pzpdf s to point estimates when evaluating and comparing \pz\ code performance in the literature.
-Though this practice should not be encouraged, we also calculate traditional metrics based on the most common point estimators derived from \pzpdf s.
-Those seeking to establish a connection to traditional ways of thinking about redshift estimation may consult the Appendix for these results.
+Some scientific applications of \pzpdf s necessitate knowledge of each individual galaxy's \pzpdf, for example galaxy intrinsic alignment studies, strong lensing foreground shear prediction, and photometric supernova classification, while others require only knowledge of the ensemble redshift distribution, $N(z)$.
+Due to the paucity of principled techniques for using and validating individual galaxy \pzpdf s in the literature, there have been few alternatives to the common practice of reducing \pzpdf s to point estimates in prior comparisons of \pz\ code performance.
+Though the outmoded practice should not be encouraged, we also calculate traditional metrics based on the most common point estimators derived from \pzpdf s in Appendix~\ref{sec:pointmetrics}.
 
-There are a number of metrics that can be used to test the accuracy of a \pz\ posterior as an estimator of a true \pz\ posterior if it is known.
-Even for simulated data, the true \pzpdf\ is in general not accessible unless the photometry is in fact drawn from the true \pz\ likelihoods, a mock catalogue generation procedure that has not yet appeared in the literature.
-Furthermore, only limited applications of \pzpdf s that could be used as the basis for a metric have been presented in the literature.
+The most popular application of \pzpdf s is the estimation of the overall redshift distribution $N(z)$, the true value of which is known for the DC1 data set and will be denoted as $\tilde{N}(z)$.
+In terms of the prior information provided to each method, the true redshift distribution satisfies the tautology $\tilde{N}(z) = p(z \vert I_{D})$ due to our experimental set-up; because the DC1 training and template sets are representative and complete, $I_{D}$ represents a prior that is also equal to the truth.
+In this ideal case of complete and representative prior information, the method that would give the best approximation to $\hat{N}(z)$ would be one that neglects all the information contained in the photometry $\{d_{i}\}_{N_{tot}}$ and gives every galaxy the same \pzpdf\ $\hat{p}_{i}(z) = \tilde{N}(z)$ for all $i$; the inclusion of any information from the photometry would only introduce noise to the optimal result of returning the prior.
+This is the exact estimator, \trainz, that we have described in Section~\ref{sec:method:trainz}, and which will serve as an experimental control.
 
-The most popular application of \pzpdf s by far is the estimation of the overall redshift distribution $N(z)$, the true value of which is known for the \textsc{Buzzard} simulation and will be denoted as $N'(z)$.
 Though alternatives exist \citep{Malz:chippr}, ``stacking'' according to
-\begin{equation} \label{eq:stacked}
+\begin{equation}
+  \label{eq:stacked}
   \hat{N}^{H}(z) \equiv \frac{1}{N_{tot}}\ \sum_{i}^{N_{tot}}\ \hat{p}^{H}_{i}(z)
 \end{equation}
-is the most widely accepted method for obtaining the stacked estimator $\hat{N}^{H}(z)$ of the redshift distribution from \pzpdf s derived by a method $H$.
-Though we do not endorse the use of the stacked estimator of the redshift distribution, we use it under the assumption that the response of our metrics of $\hat{N}^{H}(z)$ will be analogous to the same metrics applied to a principled estimator of the redshift distribution.
-We must note, however, that this is a poor assumption in general.
-
-Returning to the prior of \pzpdf s, the true redshift distribution satisfies the tautology $N'(z) = p(z | I_{D})$, because our training data is representative, $I_{D}$ represents a prior that is also equal to the truth.
-In this ideal case of representative training data, the method that would give the best approximation to $N'(z)$ would be one that neglects all the information contained in the photometry $\{d_{i}\}_{N_{tot}}$ and gives every galaxy the same \pzpdf\ $\hat{p}_{i}(z) = N'(z)$ for all $i$.
-(In fact, including any information from the photometry would only add noise to the optimal result of returning the prior for every galaxy.)
-This is the exact estimator, \trainz, that we have described in Section~\ref{sec:method:trainz}, and which will serve as an experimental control.
-
-The exact implementation of the stacked estimator $\hat{N}^{H}(z)$ should depend on the parametrization of the \pzpdf s, which may differ across codes and can affect the precision of the estimator \citep{Malz:qp}; even considering a single method under the same parametrization, say a piecewise constant function over bins or a set of samples from the posterior, an estimator using $2N$ bins or samples will trivially be more precise than an estimator using $N$ bins or samples.
+is the most widely accepted method for obtaining $\hat{N}^{H}(z)$ as an estimator of the redshift distribution from \pzpdf s derived by a method $H$.
+Though we do not endorse the use of the stacked estimator of the redshift distribution, we use it under the untested assumption that the response of our metrics of $\hat{N}^{H}(z)$ will be analogous to the same metrics applied to a principled estimator of the redshift distribution.
+The exact implementation of the stacked estimator $\hat{N}^{H}(z)$ depends on the parametrization of the \pzpdf s, which may differ across codes and can affect the precision of the estimator \citep{Malz:qp}.
+Even considering a single method under the same parametrization, such as the $200$-bin $0<z<2$ piecewise constant function used here, the exact bin definitions will affect the result.
 %\red{Note: Let's address how the gridding (endpoints and spacing) affects the metrics, from the email correspondence Re: [LSST-DESC-MEMBERS] Summary of Nov 29 PZ Telecon. (AIM)}
-In order to minimize the effects of the choice of parameterization, we asked those running all twelve codes to output \pzpdf s parameterized with $200$ piecewise constant bins spanning $0<z<2$.
 The piecewise constant format is chosen because of its established presence in the literature, and the choice of 200 bins was motivated by the approximate number of columns expected to be available for storage of \pzpdf s for the final \lsst\ Project tables.\footnote{See, e.~g.~the \lsst\ Data Products Definition Document, available at: \url{https://ls.st/dpdd}}
@@
 We will discuss the choice of \pzpdf\ parameterization further in Section~\ref{sec:results}.
-
-%\red{Divide metrics by application to ensemble of PDFs vs. estimators of the redshift distribution}
 

### 2019-03-14 paper/tex/analysis.tex
preamble and PIT

@@
 The overloaded ``$p(z)$'' is a widespread abuse of notation that obfuscates this goal, so we will dedicate some attention to breaking it apart.
-Galaxies have redshifts $z$ and photometric data $d$ drawn from a joint probability space $p(z, d)$ in nature that is not known to us.
-If it were known, however, each galaxy $i$ would have a \textit{true posterior \pzpdf}\ $p(z \vert d_{i})$ as well as a true likelihood $p(d \vert z_{i})$.
-There are a number of metrics that can be used to test the accuracy of a \pz\ posterior as an estimator of a true \pz\ posterior if it is known.
-Even for simulated data, the true \pzpdf\ is in general not accessible unless the photometry is in fact drawn from the true \pz\ likelihoods, a mock catalogue generation procedure that has not yet appeared in the literature.
-Furthermore, only limited applications of \pzpdf s that could be used as the basis for a metric have been presented in the literature.
+Galaxies have redshifts $z$ and photometric data $d$ drawn from a joint probability space $p(z, d)$ in nature.
+As a result, each observed galaxy $i$ has a \textit{true posterior \pzpdf}\ $p(z \vert d_{i})$ as well as a true likelihood $p(d \vert z_{i})$.
+There are a number of metrics that can be used to test the accuracy of a \pz\ posterior as an estimator of a true \pz\ posterior if the true \pzpdf\ is known.
+However, the true \pzpdf\ is in general not accessible unless the photometry is in fact drawn from the true \pz\ likelihoods, a mock catalogue generation procedure that has not yet appeared in the literature and was not performed for this work.
 
+Before describing the metrics appropriate to the DC1 data set, we outline the philosophy behind our choices.
 A \pzpdf\ estimator derived by method $H$ must be understood as a posterior probability distribution
@@
 
-In this study, we isolate the effect of differences in prior information $I_{H}$ specific to each method by using a single training set $I_{D}^{ML}$ for all machine learning-based codes and a single template library $I_{D}^{T}$ for all template-based codes.
-These sets of prior information are carefully constructed to be representative and complete, so we have $I_{D} \equiv I_{D}^{ML} \equiv I_{D}^{T}$ for every method $H$.
+In this study, we isolate the effect of differences in prior information $I_{H}$ specific to each method by using a single training set $I_{D}^{\rm ML}$ for all machine learning-based codes and a single template library $I_{D}^{\rm T}$ for all template-based codes.
+These sets of prior information are carefully constructed to be representative and complete, so we have $I_{D} \equiv I_{D}^{\rm ML} \equiv I_{D}^{\rm T}$ for every method $H$.
 Thus, we are saying
@@
 
-Some scientific applications of \pzpdf s necessitate knowledge of each individual galaxy's \pzpdf, for example galaxy intrinsic alignment studies, strong lensing foreground shear prediction, and photometric supernova classification, while others require only knowledge of the ensemble redshift distribution, $N(z)$.
-Due to the paucity of principled techniques for using and validating individual galaxy \pzpdf s in the literature, there have been few alternatives to the common practice of reducing \pzpdf s to point estimates in prior comparisons of \pz\ code performance.
-Though the outmoded practice should not be encouraged, we also calculate traditional metrics based on the most common point estimators derived from \pzpdf s in Appendix~\ref{sec:pointmetrics}.
-
-The most popular application of \pzpdf s is the estimation of the overall redshift distribution $N(z)$, the true value of which is known for the DC1 data set and will be denoted as $\tilde{N}(z)$.
-In terms of the prior information provided to each method, the true redshift distribution satisfies the tautology $\tilde{N}(z) = p(z \vert I_{D})$ due to our experimental set-up; because the DC1 training and template sets are representative and complete, $I_{D}$ represents a prior that is also equal to the truth.
-In this ideal case of complete and representative prior information, the method that would give the best approximation to $\hat{N}(z)$ would be one that neglects all the information contained in the photometry $\{d_{i}\}_{N_{tot}}$ and gives every galaxy the same \pzpdf\ $\hat{p}_{i}(z) = \tilde{N}(z)$ for all $i$; the inclusion of any information from the photometry would only introduce noise to the optimal result of returning the prior.
-This is the exact estimator, \trainz, that we have described in Section~\ref{sec:method:trainz}, and which will serve as an experimental control.
-
-Though alternatives exist \citep{Malz:chippr}, ``stacking'' according to
-\begin{equation}
-  \label{eq:stacked}
-  \hat{N}^{H}(z) \equiv \frac{1}{N_{tot}}\ \sum_{i}^{N_{tot}}\ \hat{p}^{H}_{i}(z)
-\end{equation}
-is the most widely accepted method for obtaining $\hat{N}^{H}(z)$ as an estimator of the redshift distribution from \pzpdf s derived by a method $H$.
-Though we do not endorse the use of the stacked estimator of the redshift distribution, we use it under the untested assumption that the response of our metrics of $\hat{N}^{H}(z)$ will be analogous to the same metrics applied to a principled estimator of the redshift distribution.
-The exact implementation of the stacked estimator $\hat{N}^{H}(z)$ depends on the parametrization of the \pzpdf s, which may differ across codes and can affect the precision of the estimator \citep{Malz:qp}.
-Even considering a single method under the same parametrization, such as the $200$-bin $0<z<2$ piecewise constant function used here, the exact bin definitions will affect the result.
+The exact implementation of the metrics theoretically depends on the parametrization of the \pzpdf s, which may differ across codes and can affect the precision of the estimator \citep{Malz:qp}.
+Even considering a single method under the same parametrization, such as the 200-bin $0 < z < 2$ piecewise constant function used here, the exact bin definitions will affect the result.
 %\red{Note: Let's address how the gridding (endpoints and spacing) affects the metrics, from the email correspondence Re: [LSST-DESC-MEMBERS] Summary of Nov 29 PZ Telecon. (AIM)}
 The piecewise constant format is chosen because of its established presence in the literature, and the choice of 200 bins was motivated by the approximate number of columns expected to be available for storage of \pzpdf s for the final \lsst\ Project tables.\footnote{See, e.~g.~the \lsst\ Data Products Definition Document, available at: \url{https://ls.st/dpdd}}
-All the \pzpdf\ catalogs are processed using the \qp\ software package \citep{Malz:qp}\footnote{available at: \url{http://github.com/aimalz/qp/}} for manipulating and calculating metrics of univariate PDFs.
 We will discuss the choice of \pzpdf\ parameterization further in Section~\ref{sec:results}.
 
-\subsection{Metrics of an ensemble of photo-\mathinhead{$z$}{z} posteriors}
+This analysis is conducted using the \qp\footnote{\url{http://github.com/aimalz/qp/}}\ software package \citep{Malz:qp} for manipulating and calculating metrics of univariate PDFs.
+We present the metrics of \pzpdf s that address our goals in the sections below.
+Though the outmoded practice should not be encouraged, those seeking a connection to previous comparison studies will find metrics of common redshift point estimators derived from \pzpdf s in Appendix~\ref{sec:pointmetrics}.
+
+\subsection{Metrics of an ensemble of photo-\mathinhead{$z$}{} posteriors}
 \label{sec:qualmet}
 
-Because the \pzpdf s in the \lsst\ catalog will be used for many applications, some of which require accuracy of each individual catalog entry, we consider several metrics that get at the population-level performance of the \pzpdf s as distinct from a summary statistic thereof.
+Because \lsst's \pzpdf s will be used for many scientific applications, some of which require accuracy of each individual catalog entry, we consider several metrics that probe the population-level performance of the \pzpdf s.
 
@@
 
-The probability integral transform (PIT) has been employed recently to evaluate fidelity of posterior distributions \citep[e.~g.~][]{Bordoloi:10,Polsterer:16,Tanaka:17}.
-The PIT value is defined for each individual galaxy as:
+The probability integral transform (PIT)
 \begin{align}
 \label{eq:pit}
-\mathrm{PIT} &= \int_{-\infty}^{z_{\rm true}}p(z)\, dz \,.
+\mathrm{PIT} &\equiv \int_{-\infty}^{z_{\rm true}} p(z) dz
 \end{align}
-The distribution of PIT values quantifies the behavior of the {\it ensemble} of \pzpdf s, enabling us to evaluate whether the population of \pzpdf s is, on average, accurate.
-The PIT value for each galaxy is the Cumulative Distribution Function (CDF) of its \pzpdf\ evaluated at its true redshift.
-A catalogue of \pzpdf s that are accurate should have a flat PIT histogram (i.e., the individual PIT values as samples from each CDF should match a Uniform(0,1) distribution if the CDFs are accurate).
-Specific deviations from flatness indicate inaccuracy: overly broad \pzpdf s would manifest as underrepresentation of the lowest and highest PIT values, whereas overly narrow \pzpdf s would manifest as over-representation of the lowest and highest PIT values.
-High frequency at only $\mathrm{PIT}\approx0$ and $\mathrm{PIT}\approx1$ indicates the presence of catastrophic outliers with highly inaccurate \pzpdf s where the true redshift of a galaxy is outside of the support of its \pzpdf.
+is the Cumulative Distribution Function (CDF) of a \pzpdf\ evaluated at its true redshift, and the distribution of PIT values probes the average accuracy of the \pzpdf s of an ensemble of galaxies.
+A catalogue of accurate \pzpdf s should have a PIT distribution that is uniform $U(0,1)$, and deviations from flatness are interpretable: overly broad \pzpdf s induce underrepresentation of the lowest and highest PIT values, whereas overly narrow \pzpdf s induce overrepresentation of the lowest and highest PIT values.
+Catastrophic outliers with a true redshift outside the support of its \pzpdf\ have $\mathrm{PIT} \approx 0$ or $\mathrm{PIT} \approx 1$.
+
+The PIT distribution has been used to quantify the performance of \pzpdf\ methods in the past \citep[e.~g.~][]{Bordoloi:10,Polsterer:16,Tanaka:17}.
 \citet{Tanaka:17} use the histogram of PIT values as a diagnostic indicator of overall code performance, while \citet{Freeman:17} independently define the PIT and demonstrate how its individual values may be used both to perform hypothesis testing (via, e.g., the KS, CvM, and AD tests; see below) and to construct quantile-quantile plots.
@@
 
-\subsubsection{Moments}
+\subsubsection{Metrics of a science-motivated quantity}
 \label{sec:moments}
 %(Alex Malz)
+
+The most popular application of \pzpdf s is the estimation of the overall redshift distribution $N(z)$, the true value of which is known for the DC1 data set and will be denoted as $\tilde{N}(z)$.
+In terms of the prior information provided to each method, the true redshift distribution satisfies the tautology $\tilde{N}(z) = p(z \vert I_{D})$ due to our experimental set-up; because the DC1 training and template sets are representative and complete, $I_{D}$ represents a prior that is also equal to the truth.
+In this ideal case of complete and representative prior information, the method that would give the best approximation to $\hat{N}(z)$ would be one that neglects all the information contained in the photometry $\{d_{i}\}_{N_{tot}}$ and gives every galaxy the same \pzpdf\ $\hat{p}_{i}(z) = \tilde{N}(z)$ for all $i$; the inclusion of any information from the photometry would only introduce noise to the optimal result of returning the prior.
+This is the exact estimator, \trainz, that we have described in Section~\ref{sec:method:trainz}, and which will serve as an experimental control.
+
+Though alternatives exist \citep{Malz:chippr}, ``stacking'' according to
+\begin{equation}
+  \label{eq:stacked}
+  \hat{N}^{H}(z) \equiv \frac{1}{N_{tot}}\ \sum_{i}^{N_{tot}}\ \hat{p}^{H}_{i}(z)
+\end{equation}
+is the most widely accepted method for obtaining $\hat{N}^{H}(z)$ as an estimator of the redshift distribution from \pzpdf s derived by a method $H$.
+Though we do not endorse the use of the stacked estimator of the redshift distribution, we use it under the untested assumption that the response of our metrics of $\hat{N}^{H}(z)$ will be analogous to the same metrics applied to a principled estimator of the redshift distribution.
 

### 2019-03-14 paper/tex/analysis.tex
combining QQ and PIT

@@
 \label{sec:qualmet}
+%(Rongpu Zhou, Jeffrey Newman, Alex Malz)
+%(Alex Malz, Peter Freeman, Sam Schmidt)
 
 Because \lsst's \pzpdf s will be used for many scientific applications, some of which require accuracy of each individual catalog entry, we consider several metrics that probe the population-level performance of the \pzpdf s.
+Because we have the true redshifts but not true \pzpdf s for comparison, we remind the reader of the Cumulative Distribution Function (CDF)
+\begin{equation}
+  \label{eq:cdf}
+  \rm{CDF}(q) \equiv \int_{-\infty}^{q} p(z) dz,
+\end{equation}
+of a \pzpdf, which is used as the basis for several of our metrics.
 
-\subsubsection{Probability integral transform (PIT)}
-\label{sec:pit}
-%(Alex Malz, Peter Freeman, Sam Schmidt)
+A quantile of a distribution is the value $q$ at which the CDF of the distribution is equal to $Q$; percentiles and quartiles are familiar examples of linearly spaced sets of 100 and 4 quantiles, respectively.
+The quantile-quantile (QQ) plot serves as a graphical visualization for comparing two distributions, where the quantiles of one distribution are plotted against the quantiles of the other distribution,
+%\blue{Include an equation to ensure quantile definition is crystal clear before building upon it for PIT.}\red{Is maybe the above text clear enough?
+Text may be more clear than any equation I can come up with--SJS.}
+providing an easy way to qualitatively assess the consistency between an estimating distribution and a true distribution.
+%location, scale and skewness.
+The closer the QQ plot is to diagonal, the closer the match between the distributions.
+% In this paper, QQ plots are used for two purposes: (1) for comparing $N(z)$ from \pzpdf s (estimated using Eq.~\ref{eq:stacked}) with the true $N(z)$, and (2) for assessing the overall consistency of an ensemble of \pzpdf s with their true redshifts on a population level, where the distribution of the PIT values (see previous section) is compared to a uniform distribution between $0$ and $1$.
+%\red{is this enough reason to include both?--SJS}
+%\aim{Maybe we should explain why we expect the distribution of CDFs to be uniform.}\red{Note: this is just describing the PIT.
+I'm going to swap the order of PIT to come before this, and then say we are evaluating the distribution of PIT values.
+I will also add why we expect this to be uniform--SJS}
 
@@
 \end{align}
-is the Cumulative Distribution Function (CDF) of a \pzpdf\ evaluated at its true redshift, and the distribution of PIT values probes the average accuracy of the \pzpdf s of an ensemble of galaxies.
+is the CDF of a \pzpdf\ evaluated at its true redshift, and the distribution of PIT values probes the average accuracy of the \pzpdf s of an ensemble of galaxies.
+The distribution of PIT values is effectively the derivative of the QQ plot.
 A catalogue of accurate \pzpdf s should have a PIT distribution that is uniform $U(0,1)$, and deviations from flatness are interpretable: overly broad \pzpdf s induce underrepresentation of the lowest and highest PIT values, whereas overly narrow \pzpdf s induce overrepresentation of the lowest and highest PIT values.
@@
 %For example, a PIT with nontrivial shape could be due to that behavior being uniform as a function of redshift or significant inaccuracy only in a small range of redshifts where $n(z)$ is large and a flat PIT elsewhere.}
+Though the QQ plot contains very similar information to that shown in the PIT histogram plot, due to being the PIT being the derivative of the QQ, we include both forms for completeness and enhanced visual interpretability.
 
-\subsubsection{Quantile-quantile (QQ) plot}
-\label{sec:qq}
-%(Rongpu Zhou, Jeffrey Newman, Alex Malz)
-
-A quantile is defined by partitioning a distribution into consecutive intervals containing equal amounts of probability, or equal numbers of objects in each interval in the case of a distribution of objects.
-The quantile-quantile (QQ) plot serves as a graphical visualization for comparing two distributions, where the quantiles of one distribution are plotted against the quantiles of the other distribution.
-%\blue{Include an equation to ensure quantile definition is crystal clear before building upon it for PIT.}\red{Is maybe the above text clear enough?
-Text may be more clear than any equation I can come up with--SJS.}
-The QQ plot provides an easy way to qualitatively assess the differences in various properties such as the moments of an estimating distribution relative to a true distribution.
-%location, scale and skewness.
-
-In this paper, QQ plots are used for two purposes: (1) for comparing $N(z)$ from \pzpdf s (estimated using Eq.~\ref{eq:stacked}) with the true $N(z)$, and (2) for assessing the overall consistency of an ensemble of \pzpdf s with their true redshifts on a population level, where the distribution of the PIT values (see previous section) is compared to a uniform distribution between $0$ and $1$.
-Though the QQ plot contains very similar information to that shown in the PIT histogram plot, due to being the PIT being the derivative of the QQ, we include both forms for completeness and enhanced visual interpretability.
-%\red{is this enough reason to include both?--SJS}
-%\aim{Maybe we should explain why we expect the distribution of CDFs to be uniform.}\red{Note: this is just describing the PIT.
-I'm going to swap the order of PIT to come before this, and then say we are evaluating the distribution of PIT values.
-I will also add why we expect this to be uniform--SJS}
-
-\subsubsection{Conditional density estimation loss}
+\subsection{Conditional density estimation loss}
 \label{sec:CDE_loss}
@@
 
+We use the QQ plot to visualize the relationship between $\hat{N}^{H}(z)$ and $\tilde{N}(z)$.
 We additionaly calculate the first three moments of the estimated redshift distribution $\hat{N}^{H}(z)$ for each code and compare them to the moments of the true redshift distribution $N'(z)$.

### 2019-03-15 paper/tex/analysis.tex
finished statistics

@@
 
+\subsection{Metric of individual photo-\mathinhead{$z$}{} posteriors}
+\label{sec:CDE_loss}
+%(Alex Malz, Ann Lee)
+
+To the knowledge of the authors, there is only one metric that can be used to evaluate the performance of individual \pzpdf\ estimators in the absence of true \pz\ posteriors.
+Using the notation introduced in Section~\ref{sec:flexzboost}, the conditional density estimation (CDE) loss is defined as
+\begin{equation}
+  \label{eq:cde-loss}
+  L(f, \widehat{f}) \equiv \int \int (f(z \vert \x) - \widehat{f}(z \vert \x))^{2} dz dP(\x)
+\end{equation}
+in terms of the photometry $\x$, an analogue to the familiar root-mean-square-error used in conventional regression.
+We estimate the CDE loss via
+\begin{equation}
+  \label{eq:estimated-cde-loss}
+  \hat{L}(f, \widehat{f}) = \mathbb{E}_\X \left[\int \widehat{f}(z \mid \X)^{2} dz\right] - 2 \mathbb{E}_{\X, Z}\left[\widehat{f}(Z \mid \X)\right] + K_{f},
+\end{equation}
+where the first term is the expectation value of the \pz\ posterior with respect to the marginal distribution of the photometric covariates $\X$, the second term is the expectation value with respect to the joint distribution of $\X$ and the space $Z$ of all possible redshifts, and the third term $K_{f}$ is a constant depending only upon the true conditional densities $f(z \vert \x)$.
+We may estimate these expectations empirically on the test or validation data \citep[Eq.~7 in][]{Izbicki:17b} without knowledge of the true densities.
+
 \subsection{Metrics of an ensemble of photo-\mathinhead{$z$}{} posteriors}
@@
 %(Alex Malz, Peter Freeman, Sam Schmidt)
+%(John Soo, Rongpu Zhou)
 
@@
   \label{eq:cdf}
-  \rm{CDF}(q) \equiv \int_{-\infty}^{q} p(z) dz,
+  \rm{CDF}[f, q] \equiv \int_{-\infty}^{q} f(z) dz,
 \end{equation}
-of a \pzpdf, which is used as the basis for several of our metrics.
+of a generic univariate PDF $f(z)$, which is used as the basis for several of our metrics.
 
@@
 \label{eq:pit}
-\mathrm{PIT} &\equiv \int_{-\infty}^{z_{\rm true}} p(z) dz
+\mathrm{PIT} &\equiv \rm{CDF}[\hat{p}, z_{\rm true}] %\int_{-\infty}^{z_{\rm true}} p(z) dz
 \end{align}
@@
 %For example, a PIT with nontrivial shape could be due to that behavior being uniform as a function of redshift or significant inaccuracy only in a small range of redshifts where $n(z)$ is large and a flat PIT elsewhere.}
-Though the QQ plot contains very similar information to that shown in the PIT histogram plot, due to being the PIT being the derivative of the QQ, we include both forms for completeness and enhanced visual interpretability.
-
-\subsection{Conditional density estimation loss}
-\label{sec:CDE_loss}
-%(Alex Malz, Ann Lee)
-
-With the conditional density estimation loss (CDE loss) we can compare how well different methods estimate individual PDFs for photometric covariates $\x$ rather than looking only at the ensemble distribution.
-As in Section~\ref{sec:flexzboost}, we use the notation $f(z \vert \x)$ instead of $p(z)$ to explicitly show the dependence on the photometry $\x$.
-
-%\aim{Perhaps this should be rewritten in the same notation used elsewhere in the paper.} \red{I agree, does someone who speaks both stats and astro want to take a shot at this?}
-%\textcolor{cyan}{Ann: We are looking at individual PDFs so the dependence on $\x$ needs to be explicit; this is probably the most accurate notation as it shows both the conditioning clearly as well as whether $\x$ is fixed or random.}
-
-The CDE loss is defined as
-\begin{equation} \label{eq:cde-loss}
-L(f, \widehat{f}) \equiv \int \int (f(z \mid \x) - \widehat{f}(z \mid \x))^{2} dz dP(\x) .
-\end{equation}
-
-This loss is the CDE equivalent of the RMSE in regression.
-To estimate this loss we rewrite it as
-\begin{equation} \label{eq:estimated-cde-loss}
-L(f, \widehat{f}) = \mathbb{E}_\X \left[\int \widehat{f}(z \mid \X)^{2} dz\right] - 2 \mathbb{E}_{\X,Z}\left[\widehat{f}(Z \mid \X)\right] + K_{f},
-\end{equation}
-where upper-case letters denote random variables and lower-case the observed variables.
-The first expectation is with respect to the marginal distribution of the covariates $\X$, the second expectation is with respect to the joint distribution of $\X$ and $Z$, and $K_{f}$ is a constant depending only upon the true conditional densities $f(z\mid \x)$.
-For each method we can estimate these expectations as empirical expectations on the test or validation data \citep[Eq.~7 in][]{Izbicki:17b} without knowledge of the true densities.
 
@@
 
-\subsection{Metrics over estimated probability distributions}
-\label{sec:quantmet}
+% \subsection{Metrics over estimated probability distributions}
+% \label{sec:quantmet}
 
-In tandem with the QQ and PIT metrics introduced above, we additionally compute the following metrics comparing the empirical CDF of a distribution to the true or expected distribution.
-These metrics give a more quantitative measure of the departure from ideal than the more visual PIT histogram and QQ plot.
-We compute metrics comparing the CDF of PIT values to the CDF of a Uniform distribution, and also compute the CDF of the true redshift distribution $N'(z)$ compared the $\hat{N}(z)$ distribution derived from summing the \pzpdf s as described in Eq.~\ref{eq:stacked}.
+% We compute metrics comparing the CDF of PIT values to the CDF of a Uniform distribution, and also compute the CDF of the true redshift distribution $N'(z)$ compared the $\hat{N}(z)$ distribution derived from summing the \pzpdf s as described in Eq.~\ref{eq:stacked}.
 
@@
 %%(Alex Malz)%
-%
 %We employ the familiar root-mean-square error
@@
 %Though this metric does not account for the fact that the redshift distribution function is, in fact, a probability distribution, it can still be interpreted as a measure of the integrated difference between the estimated distribution and the true distribution.%, and it can be used to quantify the otherwise qualitative metrics.
-%
 
-\subsubsection{Kolmogorov-Smirnov (KS) and related statistics}
-\label{sec:ks}
-%(John Soo, Rongpu Zhou)
+% \subsubsection{Kolmogorov-Smirnov (KS) and related statistics}
+% \label{sec:ks}
 
-The \textit{Kolmogorov-Smirnov statistic} $N_{\rm KS}$ is the maximum difference between $F_{\rm phot}(z)$ and $F_{\rm spec}(z)$, the CDFs of the photo-$z$ and spectroscopic redshift respectively:
+We evaluate a number of quantitative metrics derived from the visually interpretable QQ plot and PIT histogram, built on the Kolmogorov-Smirnov (KS) statistic
 \begin{equation}
-N_{\rm KS} \equiv \max_{z}\left( \left| F_{\rm phot}(z) - F_{\rm spec}(z) \right| \right).
+  \label{eq:ks}
+  \rm{KS} \equiv \max_{z} \left( \left| \rm{CDF}[\hat{f}, z] - \rm{CDF}[\tilde{f}, z] \right| \right),
 \end{equation}
+interpretable as the maximum difference between the CDFs of an approximating univariate distribution $\hat{f}(z)$ and a reference distribution $\tilde{f}(z)$.
+% $CDF_{\rm phot}(z)$ and $CDF_{\rm spec}(z)$, the CDFs of the photo-$z$ and spectroscopic redshift respectively:
 %The KS test quantifies whether the two redshift distributions are drawn from a similar distribution, independent of binning.
-The KS test quantifies the similarity between two distributions, independent of binning.
-A lower $N_{\rm KS}$ value corresponds to more similar distributions.
+% The KS test quantifies the similarity between two distributions, independent of binning.
+% A lower $N_{\rm KS}$ value corresponds to more similar distributions.
 %We calculate two KS test statistics, one to test the distribution of $p(z)$ values, and one to test the stacked estimator $\hat{n}(z)$ of the redshift distribution.
@@
 $F_{\rm phot}(z)$ is calculated by stacking the PDFs of every single object in the sample, producing a smooth cumulative $n(z)$ to be compared with $F_{\rm spec}(z)$.
-
-We also consider two variants of the KS statistic: the Cramer-von Mises (CvM) and Anderson-Darling (AD) statistics.
-The CvM statistic is similar to the KS statistic as it is also computed from the distance between the measured CDF and the ideal CDF, but instead of the maximum distance, the CvM statistic
+We also consider two variants of the KS statistic.
+A cousin of the KS statistic, the Cramer-von Mises (CvM) statistic
 \begin{equation}
 \label{eq:cvm}
-\omega^2 \equiv \int_{-\infty}^{+\infty}\big(F_{\mathrm{meas.}}(x) - F_{\mathrm{ideal}}(x)\big)^2\mathrm{d}F_{\mathrm{ideal}}
+  \rm{CvM}^{2} \equiv \int_{-\infty}^{+\infty} \big(\rm{CDF}[\hat{f}, z] - \rm{CDF}[\tilde{f}, z]\big)^2 \mathrm{d}\rm{CDF}[\tilde{f}, z]
 \end{equation}
-is the average of the distance squared.
+is the mean-squared difference between the CDFs of an approximate and true PDF.
+The Anderson-Darling (AD) statistic
+\begin{equation} \label{eq:ad}
+  \rm{AD}^2 \equiv N_{tot}\int_{-\infty}^{+\infty} \frac{\big(\rm{CDF}[\hat{f}, z] - \rm{CDF}[\tilde{f}, z]\big)^2} {\rm{CDF}[\tilde{f}, z] (1 -\rm{CDF}[\tilde{f}, z]} \mathrm{d}\rm{CDF}[\tilde{f}, z]
+\end{equation}
+is a weighted mean-squared difference featuring enhanced sensitivity to discrepancies in the tails of the distribution.
+% The CvM statistic is similar to the KS statistic as it is also computed from the distance between the measured CDF and the ideal CDF, but instead of the maximum distance, the CvM statistic
+% is the average of the distance squared.
 % In this paper we use the CvM statistic to evaluate the performance of both $n(z)$ and $p(z)$.
-
-The AD statistic
-\begin{equation} \label{eq:ad}
-A^2 \equiv N_{tot}\int_{-\infty}^{+\infty} \frac{\big(F_{\mathrm{meas.}}(x) - F_{\mathrm{ideal}}(x)\big)^2} {F_{\mathrm{ideal}}(x) (1-F_{\mathrm{ideal}}(x))}\mathrm{d}F_{\mathrm{ideal}}
-\end{equation}
-is a weighted version of the CvM statistic, making it more sensitive to the tails of the distribution, where $N_{tot}$ is the sample size.
+% The AD statistic
+% is a weighted version of the CvM statistic, making it more sensitive to the tails of the distribution, where $N_{tot}$ is the sample size.
 %The KS statistic and the CvM and AD statistics related to it are all used to evaluate the performance of both $\hat{n}(z)$ and the QQ and PIT of the CDFs of the ensemble of photo-$z$ interim posteriors.
@@
 
-The most popular application of \pzpdf s is the estimation of the overall redshift distribution $N(z)$, the true value of which is known for the DC1 data set and will be denoted as $\tilde{N}(z)$.
+Perhaps the most popular application of \pzpdf s is the estimation of the overall redshift distribution $N(z)$, a quantity that enters some cosmological calculations and the true value of which is known for the DC1 data set and will be denoted as $\tilde{N}(z)$.
 In terms of the prior information provided to each method, the true redshift distribution satisfies the tautology $\tilde{N}(z) = p(z \vert I_{D})$ due to our experimental set-up; because the DC1 training and template sets are representative and complete, $I_{D}$ represents a prior that is also equal to the truth.
@@
 
-We use the QQ plot to visualize the relationship between $\hat{N}^{H}(z)$ and $\tilde{N}(z)$.
-We additionaly calculate the first three moments of the estimated redshift distribution $\hat{N}^{H}(z)$ for each code and compare them to the moments of the true redshift distribution $N'(z)$.
-The $m^{\mathrm{th}}$ moment of a distribution is defined as
+As $N(z)$ is itself a univariate PDF, we apply the metrics of the previous sections to it as well.
+We additionaly calculate the first three moments
 \begin{equation}
-\langle z^{m}\rangle \equiv \int_{-\infty}^{\infty}z^{m}N(z)dz .
+  \label{eq:moment}
+\langle z^{m}\rangle \equiv \int_{-\infty}^{\infty} z^{m} N(z) dz
 \end{equation}
-Here, we use the moments of the stacked estimator of the redshift distribution function as the basis for a metric.
-The closer the moments of $\hat{N}(z)$ for a \pzpdf\ method are to the moments of the true redshift distribution function $N'(z)$, the better the \pzpdf\ method is a estimating the overall redshift distribution ``shape''.
+of the estimated redshift distribution $\hat{N}^{H}(z)$ for each code and compare them to the moments of the true redshift distribution $\tilde{N}(z)$.
+Under the assumption that the stacked estimator is unbiased, a superior method minimizes the difference between the true and estimated moments.

### 2019-03-15 paper/tex/discussion.tex
consistent messaging around figure 1

@@
 %(Sam Schmidt, Ofer Lahav, Jeff Newman, Alex Malz, Eve Kovacs, Tony Tyson, Tina Peters)
-In this paper we presented results evaluating the computation of individual galaxy photometric redshift PDFs for twelve photo-$z$ codes.
+In this paper we present results evaluating the computation of individual galaxy photometric redshift PDFs for twelve photo-$z$ codes.
 As discussed in Section~\ref{sec:metrics} each $p(z)$ should accurately reflect the relative likelihood as a function of redshift for each galaxy in an informative way; that is, the estimates should provide useful information per individual galaxy, not just the ensemble.
@@
 \label{sec:futurework}
+
 The work presented in this paper is only the first step in characterizing current photo-$z$ codes and moving toward an improved photometric redshift estimator.

### 2019-03-15 paper/tex/results.tex
consistent messaging around figure 1

@@
 \label{sec:results}
-%(Sam Schmidt, Bryce Kalmbach, Johann Cohen Tanugi, Rongpu Zhou)
+%(Sam Schmidt, Bryce Kalmbach, Johann Cohen Tanugi, Rongpu Zhou, Alex Malz)
 
-\subsection{Ensembles of photo-\mathinhead{$z$}{z} interim posteriors}
+We begin with a demonstrative visual inspection of the \pzpdf s produced by each code for individual galaxies.
+Fig.~\ref{fig:pz_examples} shows the \pzpdf s for four galaxies chosen as examples of \pzpdf\ archetypes: a narrow unimodal redshift solution, a broad unimodal solution, a bimodal distribution, and a multimodal distribution.
 
@@
 %\centering
-\includegraphics[width=0.49\textwidth]{fig/pz_12codes_261931_noseaborn_crop.jpg}\includegraphics[width=0.5\textwidth]{fig/pz_12codes_471167_noseaborn_crop.jpg}\\
+\includegraphics[width=0.49\textwidth]{fig/pz_12codes_261931_noseaborn_crop.jpg}\includegraphics[width=0.49\textwidth]{fig/pz_12codes_471167_noseaborn_crop.jpg}\\
 \includegraphics[width=0.49\textwidth]{fig/pz_12codes_713178_noseaborn_crop.jpg}\includegraphics[width=0.49\textwidth]{fig/pz_12codes_982747_noseaborn_crop.jpg}
-\caption{Four illustrative examples of individual p(z) distributions produced by the codes.
-The red vertical line represents the true redshift.
-Examples are chosen with common features seen in PDFs: tight unimodal $p(z)$ (upper left), broad unimodal $p(z)$ (upper right), bimodal $p(z)$ (lower right), and complex/multimodal $p(z)$ (lower left).
-Codes show varying amounts of small-scale structure in their reconstruction of the posterior distribution.
-We see varying responses from the codes in the presence of color degeneracies and photometric errors, resulting in narrow and broad unimodal, bimodal, and multi-modal $p(z)$ curves.} \label{fig:pz_examples}
+\caption{The individual \pzpdf s (blue) distributions produced by the twelve codes (small panels) on four exemplary galaxies' photometry (large panels) with different true redshifts (red).
+The \pzpdf s of all codes share some features for the example galaxies due to physical color degeneracies and photometric errors: tight unimodal $p(z)$ (upper left), broad unimodal $p(z)$ (upper right), bimodal $p(z)$ (lower right), and complex/multimodal $p(z)$ (lower left).
+The diverse algorithms and implementations induce differences in small-scale structure and sensitivity to physical systematics.}
+\label{fig:pz_examples}
 \end{figure*}
 
-%\begin{figure*}
-%\centering
-%\includegraphics[width=\textwidth]{fig/QQplot_11codes.jpg}
-%\caption{The QQ plots produced for every photo-$z$ code.
-\aim{Consider making this a plot of QQ minus that diagonal for better contrast between methods.}\red{The PIT histogram is essentially this, which is one of the main reasons to include both.
-QQ is a fairly standard form, so plotting in this fashion rather than the difference seems better to me, given that we also ahve the PIT histograms.
---SJS}} \label{fig:qq}
-%\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{fig/PITANDQQplot_12codes_crop.jpg}
-\caption{Summary plots for all twelve photo-$z$ codes illustrating performance for the interim posterior statistics.
-The top panel of each pair shows both the Quantile-Quantile (QQ) plot (red) and the histogram of PIT values (blue).
-The desired behavior is a QQ plot that matches the diagonal dashed line, and a PIT histogram that matches a uniform distribution matching the thin horizontal black line.
-The bottom panel of each pair shows the difference between the QQ quantile and the diagonal, illustrating departure from the desired performance.
-Histograms with an overabundance of PIT values at the centre of the distribution indicate $p(z)$ distributions that are overly broad, while an excess of values at the extrema indicate $p(z)$ distributions that are overly narrow.
-Values of PIT=0 and PIT=1 indicate ``catastrophic failures'' where the true redshift is completely outside the support of $p(z)$.
-Asymmetric features are indicative of systematic bias in the redshift predictions.
-A variety of behaviors are evident, and specific details are discussed in the text.}
-\label{fig:pitqq}
-\end{figure*}
-
-Fig.~\ref{fig:pz_examples} Shows the $p(z)$ produced by each of our twelve photo-$z$ codes for four example galaxies which exemplify some prominent cases that arise when estimating photo-$z$ PDFs: a narrow, unimodal redshift solution, a broader unimodal solution, a bimodal distribution, and a complex, multimodal distribution.
-The red vertical line represents the true redshift of the individual galaxy, and the blue curve represents the redshift probability.
-Several features are obvious even in these illustrative examples.
-\textsc{ANNz2, METAPhoR, CMNN,} and \textsc{SkyNet} all show an excess of small-scale features, which appear to be print-through of the underlying training set galaxies.
-For example, in \textsc{CMNN} the $p(z)$ are a simply a weighted histogram of all spectroscopic training galaxies in nearby colour space with no smoothing applied, so the substructure is due to the finite number of neighbours, and is not unexpected.
-\textsc{GPZ} (in its current implementation), on the other hand, always produces a single Gaussian, which broadens to cover the multi-modal redshift solutions seen in other codes.
+The most striking differences between codes are due to small-scale features induced by the interaction between the shared piecewise constant parameterization of $200$ bins $0 < z < 2$ of Section~\ref{sec:metrics} and the smoothing conditions or lack thereof in each algorithm.
+The $\rm{d}z = 0.01$ redshift resolution is sufficient to capture the broad peaks of faint galaxies' \pzpdf s with large photometric errors but is too broad to resolve the narrow peaks for bright galaxies' \pzpdf s with small photometric errors.
+This observation is consistent with the findings of \citet[]{Malz:qp} that the piecewise constant parameterization underperforms in the presence of small-scale structures.
+However, the shared small-scale features of \annz, \metaphor, \cmnn, and \skynet\ are a smoking gun for overfitting of the training set, a result of various weighted sums of the limited number of training set galaxies with colors similar to those of the test set galaxy in question.
+The settings used on \gpz\ in this work forced broadening of the single Gaussian to cover the multimodal redshift solutions of the other codes.
 %Interestingly, while \textsc{ANNz2} shows an abundance of small scale structure in individual $p(z)$ measurements (see Fig.~\ref{fig:pz_examples}), the summed $\hat{N}(z)$ is rather smooth, where the small scale features average out.
 This is not the case for the two other codes that show an abundance of substructure in their individual $p(z)$: both \textsc{CMNN} and \textsc{SkyNet} show small scale features both in $p(z)$ and $\hat{N}(z)$.
-
-As stated in Section~\ref{sec:metrics}, $p(z)$ is parameterized as $200$ piecewise constant bins covering $0<z<2$ for all twelve codes, giving a grid size of $\delta z = 0.01$ for each PDF.
-A piecewise constant grid was a natural choice for some photo-$z$ methods, for instance most template-based codes compute likelihoods on a fixed grid.
 %In contrast, FlexZBoost, for example, can return estimates on any grid without compression errors as it’s a basis expansion method where only the expansion coefficients need to be stored.
@@
 %Furthermore, the fidelity of photo-$z$ interim posteriors in this format varies with the quality of the photometry.
-For faint galaxies, this redshift resolution is sufficient to capture the shape of $p(z)$ for the majority of the test sample, where photometric errors on the faint galaxies lead to somewhat broad peaks in the redshift posterior.
-However, as can be seen in e.~g.~the top left panel of Fig.~\ref{fig:pz_examples}, for bright galaxies with narrow $p(z)$ the grid spacing of $\delta z = 0.01$ is not always sufficient to resolve the peak.
-This is consistent with the results described in \citet[]{Malz:qp}, who find that quantiles (and, to a lesser degree, samples) often outperform gridded $p(z)$, particularly for bright objects and in the presence of harsher storage constraints.
-With a full 200 numbers to capture the information of each photo-$z$ PDF, any parametrization will perform adequately, but other storage parametrizations and limits on storage resources may be considered in future work.
 %Switching to a quantile based parameterization may be more costly computationally, for example template-based codes would need to test more grid point in order to resolve the quantiles for bright galaxies.
 However, the computational time for template based codes scales roughly linearly with the number of grid points, so this may be a reasonable thing to implement.
-We will discuss this further in Section~\ref{sec:discussion}.
+% We will discuss this further in Section~\ref{sec:discussion}.
 % \red{someone review this statement to make sure that I'm saying this correctly!}
+
+\subsection{Ensembles of \pzpdf s}
+
+\begin{figure*}
+\centering
+\includegraphics[width=0.74\textwidth]{fig/PITANDQQplot_12codes_crop.jpg}
+\caption{The QQ plot (red) and PIT histogram (blue) of the \pzpdf\ codes (panels) along with the ideal QQ (black dashed diagonal) and ideal PIT (gray horizontal) curves, as well as a difference plot for the QQ difference from the ideal diagonal (lower inset).
+The twelve codes exhibit varying degrees of four deviations from perfection: an overabundance of PIT values at the centre of the distribution indicate a catalogue of overly broad \pzpdf s, an excess of PIT values at the extrema indicates a catalogue of overly narrow \pzpdf s, catastrophic outliers manifest as overabundances at PIT values of $0$ and $1$, and asymmetry indicates systematic bias, a form of model misspecification.}
+\label{fig:pitqq}
+\end{figure*}
 

### 2019-03-15 paper/tex/results.tex
spacing

@@
 
-Fig.~\ref{fig:pitqq} shows both the quantile-quantile plots (red) and the histogram of PIT values (blue) summarizing the results from each photo-$z$ code.
-The red line shows the measured quantiles, while the black diagonal represents the ideal QQ values if the distribution were perfectly reproduced.
-A second panel below the main panel for each code shows the difference between $Q_{\rm data}$ and $Q_{\rm theory}$, i.~e.~the departure from the diagonal, for clarity.
+Fig.~\ref{fig:pitqq} shows both the QQ plots and the histogram of PIT values, QQ plot, and a QQ difference plot relative to the ideal diagonal.
 Biases and trends in whether the average width of the $p(z)$ values being over/under-predicted are evident.
@@
 We expect a value of 0.0002 for a proper Uniform distribution.
-An excess over this small value indicates true redshifts that fall outside the non-zero support of the $p(z)$.
-}\label{tab:pitoutlier}
+An excess over this small value indicates true redshifts that fall outside the non-zero support of the $p(z)$.}
+\label{tab:pitoutlier}
 \begin{tabular}{lc}
@@
 Given the near-perfect training data, examining the individual codes for explanations for departures from the expected behaviour will be instructive in avoiding similar problems in future tests.
-\textsc{ANNz2} performs quite well in $p(z)$ based metrics.
-In the specific implementation employed in this paper, the final $p(z)$ is a weighted average of five neural-nets.
+\textsc{ANNz2} performs quite well in $p(z)$ based metrics.  In the specific implementation employed in this paper, the final $p(z)$ is a weighted average of five neural-nets.
 During the training process \textsc{ANNz2} compares the percentiles of the redshift training sample against the CDFs of the $p(z)$ sample.
@@
 Varying levels of agreement are seen in the codes.
-Both $\hat{N}(z)$ and $N'(z)$ in all codes are smoothed using a single bandwidth chosen via Scott's rule.} \label{fig:nz}
+Both $\hat{N}(z)$ and $N'(z)$ in all codes are smoothed using a single bandwidth chosen via Scott's rule.}
+\label{fig:nz}
 %\aim{Arguably this could also be clarified by showing the true $n'(z)$ once and then differences from it for each cod's stacked estimator\dots}\scc{what if we plot the difference between true and stacked rather than soverposing the two lines?
@@
 The statistics are correlated, the codes with the lowest KS statistics tend to have the lowest CvM and AD statistics.
-\textsc{CMNN} performs markedly better than the others in reconstructing the overall $N(z)$ distribution, while \textsc{SkyNet} scores poorly due to an overall bias in its redshift predictions.} \label{fig:nz_stats}
+\textsc{CMNN} performs markedly better than the others in reconstructing the overall $N(z)$ distribution, while \textsc{SkyNet} scores poorly due to an overall bias in its redshift predictions.}
+\label{fig:nz_stats}
 \end{figure*}

### 2019-03-15 paper/tex/analysis.tex
consistent messaging around figure 1
fixing a bug I introduced

@@
 
-\subsection{Metric of individual photo-\mathinhead{$z$}{} posteriors}
+\subsection{Metric of individual \pzpdf s}
 \label{sec:CDE_loss}
@@
 
-\subsection{Metrics of an ensemble of photo-\mathinhead{$z$}{} posteriors}
+\subsection{Metrics of an ensemble of \pzpdf s}
 \label{sec:qualmet}
@@
 In this ideal case of complete and representative prior information, the method that would give the best approximation to $\hat{N}(z)$ would be one that neglects all the information contained in the photometry $\{d_{i}\}_{N_{tot}}$ and gives every galaxy the same \pzpdf\ $\hat{p}_{i}(z) = \tilde{N}(z)$ for all $i$; the inclusion of any information from the photometry would only introduce noise to the optimal result of returning the prior.
-This is the exact estimator, \trainz, that we have described in Section~\ref{sec:method:trainz}, and which will serve as an experimental control.
+This is the exact estimator, \trainz, that we have described in Section~\ref{sec:trainz}, and which will serve as an experimental control.
 

### 2019-03-15 paper/tex/discussion.tex
fixing a bug I introduced

@@
 This initial paper explored code performance in idealized conditions with perfect catalog-based photometry and representative training data.
-As mentioned in Section~\ref{sec:stackedmetrics} for the stacked $N(z)$ metrics we examined only the entire galaxy population with no selections in either photo-$z$ ``quality" or redshift.
+As mentioned in Section~\ref{sec:metrics} for the stacked $N(z)$ metrics we examined only the entire galaxy population with no selections in either photo-$z$ ``quality" or redshift.
 The cosmological analyses for weak lensing and large scale structure based measures plan to break galaxy samples into tomographic redshift bins, using photo-$z$ $p(z)$ to infer the redshift distribution for each bin.

### 2019-03-15 paper/tex/results.tex
fixing a bug I introduced

@@
 \caption{The QQ plot (red) and PIT histogram (blue) of the \pzpdf\ codes (panels) along with the ideal QQ (black dashed diagonal) and ideal PIT (gray horizontal) curves, as well as a difference plot for the QQ difference from the ideal diagonal (lower inset).
-The twelve codes exhibit varying degrees of four deviations from perfection: an overabundance of PIT values at the centre of the distribution indicate a catalogue of overly broad \pzpdf s, an excess of PIT values at the extrema indicates a catalogue of overly narrow \pzpdf s, catastrophic outliers manifest as overabundances at PIT values of $0$ and $1$, and asymmetry indicates systematic bias, a form of model misspecification.}
+The twelve codes exhibit varying degrees of four deviations from perfection: an overabundance of PIT values at the centre of the distribution indicate a catalogue of overly broad \pzpdf s, an excess of PIT values at the extrema indicates a catalogue of overly narrow \pzpdf s, catastrophic outliers manifest as overabundances at PIT values of 0 and 1, and asymmetry indicates systematic bias, a form of model misspecification.}
 \label{fig:pitqq}
@@
 
-Examination of the PIT histograms and QQ plots shows that there are fairly generic issues with the width of $p(z)$ uncertainties: \textsc{Delight, CMNN, SkyNet} and \textsc{TPZ} all show a PIT histogram with a dearth of low values and and an excess of high values, signs that, on average, their $p(z)$ are more broad than the true distribution of redshifts.
+Examination of the PIT histograms and QQ plots smethod:hows that there are fairly generic issues with the width of $p(z)$ uncertainties: \textsc{Delight, CMNN, SkyNet} and \textsc{TPZ} all show a PIT histogram with a dearth of low values and and an excess of high values, signs that, on average, their $p(z)$ are more broad than the true distribution of redshifts.
 \textsc{METAPhoR} shows the opposite trend, indicating the the $p(z)$  are more narrow than the distributions given by the true redshifts.
@@
 
-\subsection{Metrics of the stacked estimator of the redshift distribution}\label{sec:stackedmetrics}
+\subsection{Metrics of the stacked estimator of the redshift distribution}
+\label{sec:stackedmetrics}
 
@@
 %\red{discussion of why NN and FZBoost are doing well, how we can try to improve the other codes}
-%\textcolor{cyan}{Ann: }
 
@@
 
-\subsection{Interpretation of metrics}\label{sec:caution}
+\subsection{Interpretation of metrics}
+\label{sec:caution}
 %Caution on metrics and photo-$z$ codes}\label{sec:caution}
@@
 
-The \trainz\ estimator, which assigns every galaxy a $p(z)$ equal to $N(z)$ of the training set as described in Section~\ref{sec:method:trainz}, is introduced as a ``null test'' to demonstrate this point via \textit{reductio ad absurdum}.
+The \trainz\ estimator, which assigns every galaxy a $p(z)$ equal to $N(z)$ of the training set as described in Section~\ref{sec:trainz}, is introduced as a ``null test'' to demonstrate this point via \textit{reductio ad absurdum}.
 \trainz\ outperforms all codes on the PIT-based metrics, and all but one code on the $N(z)$ based statistics.

### 2019-03-15 paper/tex/results.tex
text around Figure 2

@@
 
-Fig.~\ref{fig:pitqq} shows both the QQ plots and the histogram of PIT values, QQ plot, and a QQ difference plot relative to the ideal diagonal.
-Biases and trends in whether the average width of the $p(z)$ values being over/under-predicted are evident.
-An overall bias where the predicted redshift is systematically low manifests as the measured QQ value falling above the diagonal, as is the case for \textsc{BPZ} and \textsc{EAZY}, while a systematic overprediction shows up as the measured QQ value falling below the diagonal, as seen in \textsc{TPZ}.
-In terms of PIT histograms, a systematic underprediction of redshift corresponds to fewer PIT values at $PIT<0.5$ and more at $PIT>0.5$, while a systematic overprediction will show the opposite.
+Fig.~\ref{fig:pitqq} shows a histogram of PIT values, QQ plot, and QQ difference plot relative to the ideal diagonal, showcasing the biases and trends in the average accuracy of the \pzpdf s for each code.
+The high QQ values (more high than low PIT values) of \bpz, \cmnn, \delight, \eazy, and \gpz\ indicate \pzpdf s biased low, and the low QQ values (more low than high PIT values) of \skynet\ and \tpz\ indicate \pzpdf s biased high.
 
-Examination of the PIT histograms and QQ plots smethod:hows that there are fairly generic issues with the width of $p(z)$ uncertainties: \textsc{Delight, CMNN, SkyNet} and \textsc{TPZ} all show a PIT histogram with a dearth of low values and and an excess of high values, signs that, on average, their $p(z)$ are more broad than the true distribution of redshifts.
-\textsc{METAPhoR} shows the opposite trend, indicating the the $p(z)$  are more narrow than the distributions given by the true redshifts.
-In all of these code cases there is a free parameter or bandwidth that can be used to tune uncertainties.
-The sensitivity of multiple codes to this bandwidth choice emphasizes the fact that great care must be taken in setting user-defined parameters in photo-$z$ codes, even in the presence of representative training/validation data.
-for \textsc{FlexZBoost} the ``sharpening'' parameter (described in Section~\ref{sec:flexzboost}) plays a key role in improving the results, resulting in a QQ plot that is very nearly diagonal.
-A similar sharpening procedure could be beneficial for several codes.
-Interestingly, the three purely template-based codes, \textsc{BPZ, EAZY}, and \textsc{LePhare}, show relatively well behaved $p(z)$ statistics (albeit with some bias), which may indicate that the likelihood estimation with representative templates is accurately capturing the uncertainties on individual redshifts.
+The PIT histograms of \delight, \cmnn, \skynet, and \tpz\ feature an underrepresentation of extreme values, indicative of overly broad \pzpdf s, while the overrepresentation of extreme values for for \metaphor indicate overly narrow \pzpdf s.
+These five codes in particular have a free parameter for bandwidth, which may be responsible for this vulnerability, in spite of the opportunity for fine-tuning with perfect prior information.
+\flexzboost's ``sharpening'' parameter (described in Section~\ref{sec:flexzboost}) played a key role in diagonalizing the QQ plot, indicating a common avenue for improvement in the approaches that share this type of parameter.
+On the other hand, the three purely template-based codes, \bpz, \eazy, and \lephare, do not exhibit much systematic broadening or narrowing, which may indicate that complete template coverage effectively defends from these effects.
 
-%Fig.~\ref{fig:pitqq} shows the PIT histogram plots produced by each photo-$z$ code.
-%These plots show information that is very similar to that shown in the QQ plot in many ways.
-The ideal PIT histogram would follow the black dashed line, representing a uniform distribution of PIT values, equivalent to the diagonal line in the QQ plot.
-Overly broad $p(z)$ values show up as an excess of PIT values near $0.5$ and a dearth of values at the edges, while overly narrow $p(z)$ will have an excess at the edges and will be missing values at the centre.
-Another feature evident in the PIT histograms is the number of ``catastrophic outlier'' values where the true redshift falls outside of the non-zero support of $p(z)$, corresponding to PIT$=0.0$ or $1.0$ is more apparent than in the QQ plots.
-Following Kodra \& Newman (in prep.) we define $f_{\rm O}$ as the fraction of objects with $PIT<0.0001$ or $PIT>0.9999$.
-Table~\ref{tab:pitoutlier} lists these fractions for each of the codes.
+\begin{table}
+\setlength{\tabcolsep}{2pt}
+\centering
+\caption{The catastrophic outlier rate as defined by extreme PIT values.
+We expect a value of 0.0002 for a proper Uniform distribution.
+An excess over this small value indicates true redshifts that fall outside the non-zero support of the $p(z)$.}
+\label{tab:pitoutlier}
+\begin{tabular}{lc}
+\hline
+Photo-z Code & fraction PIT$<10^{-4}$ or $>$0.9999\\
+\hline
+\annz       & 0.0265\\
+\bpz        & 0.0192\\
+\delight    & 0.0006\\
+\eazy       & 0.0154\\
+\flexzboost & 0.0202\\
+\gpz        & 0.0058\\
+\lephare    & 0.0486\\
+\metaphor   & 0.0229\\
+\cmnn       & 0.0034\\
+\skynet     & 0.0001\\
+\tpz        & 0.0130\\
+\hline
+\trainz     & 0.0002\\
+\end{tabular}
+\end{table}
+
+Following Kodra \& Newman (in prep.) we define the PIT-based catastrophic outlier rate $f_{\rm O}$ as the fraction of galaxies with $\rm{PIT} < 0.0001$ or $\rm{PIT} > 0.9999$.
+Though these spikes in the first and last bin of the PIT histogram were cut out of Figure~\ref{fig:pitqq} for visualization, the catastrophic outlier rates are provided in Table~\ref{tab:pitoutlier}.
+
 For a proper Uniform distribution we expect a value of $0.0002$.
@@
 As expected, and by design, \textsc{TrainZ} has the proper fraction of outliers for the $f_{\rm O}$ statistic.
-
-\begin{table}
-\setlength{\tabcolsep}{2pt}
-\centering
-\caption{The fraction of ``catastrophic outlier'' PIT values.
-We expect a value of 0.0002 for a proper Uniform distribution.
-An excess over this small value indicates true redshifts that fall outside the non-zero support of the $p(z)$.}
-\label{tab:pitoutlier}
-\begin{tabular}{lc}
-\hline
-\hline
- ``catastrophic outlier''\\ PIT fraction\\
-\hline
-Photo-z Code & fraction PIT$<10^{-4}$ or $>$0.9999\\
-\hline
-\textsc{ANNz2} & 0.0265\\
-\textsc{BPZ} & 0.0192\\
-\textsc{Delight} & 0.0006\\
-\textsc{EAZY} & 0.0154\\
-\textsc{FlexZBoost} & 0.0202\\
-\textsc{GPz} & 0.0058\\
-\textsc{LePhare} & 0.0486\\
-\textsc{METAPhoR}& 0.0229\\
-\textsc{CMNN} & 0.0034\\
-\textsc{Skynet} & 0.0001\\
-\textsc{TPZ} & 0.0130\\
-\hline
-\textsc{TrainZ} & 0.0002\\
-\end{tabular}
-\end{table}
 

### 2019-03-15 paper/tex/results.tex
text around Table 2

@@
 
-\subsection{Ensembles of \pzpdf s}
+\subsection{\Pzpdf\ ensembles}
+\label{sec:pitqq}
+
+Fig.~\ref{fig:pitqq} shows a histogram of PIT values, QQ plot, and QQ difference plot relative to the ideal diagonal, showcasing the biases and trends in the average accuracy of the \pzpdf s for each code.
+The high QQ values (more high than low PIT values) of \bpz, \cmnn, \delight, \eazy, and \gpz\ indicate \pzpdf s biased low, and the low QQ values (more low than high PIT values) of \skynet\ and \tpz\ indicate \pzpdf s biased high.
 
@@
 \end{figure*}
-
-Fig.~\ref{fig:pitqq} shows a histogram of PIT values, QQ plot, and QQ difference plot relative to the ideal diagonal, showcasing the biases and trends in the average accuracy of the \pzpdf s for each code.
-The high QQ values (more high than low PIT values) of \bpz, \cmnn, \delight, \eazy, and \gpz\ indicate \pzpdf s biased low, and the low QQ values (more low than high PIT values) of \skynet\ and \tpz\ indicate \pzpdf s biased high.
 
@@
 
-Following Kodra \& Newman (in prep.) we define the PIT-based catastrophic outlier rate $f_{\rm O}$ as the fraction of galaxies with $\rm{PIT} < 0.0001$ or $\rm{PIT} > 0.9999$.
-Though these spikes in the first and last bin of the PIT histogram were cut out of Figure~\ref{fig:pitqq} for visualization, the catastrophic outlier rates are provided in Table~\ref{tab:pitoutlier}.
+Though the spikes in the first and last bin of the PIT histogram were cut off in Figure~\ref{fig:pitqq} for visualization, the catastrophic outlier rates are provided in Table~\ref{tab:pitoutlier}.
+As expected, \trainz\ achieves precisely the 0.0002 value expected of an ideal PIT distribution.
+\annz, \flexzboost, \lephare, and \metaphor\ have notably high catastrophic outlier rates $> 0.02$, exceeding 100 times the ideal PIT rate, meriting further investigation.
 
-For a proper Uniform distribution we expect a value of $0.0002$.
-Several codes show a marked excess, with \textsc{ANNz2, FlexZBoost, LePhare, and METAPhoR} with $f_{\rm O} > 0.02$, indicating a sizeable number of catastrophic redshift solutions where the true redshift is not covered by the extent of $p(z)$.
-For \textsc{METAPhoR} this may be partially due to an overall underprediction of the $p(z)$ width, however this is not the case for the other codes.
-\textsc{LePhare} is a particular outlier with nearly 5 per cent of objects outside of $p(z)$ support.
-Further study will be necessary to determine what is causing these misclassifications for \textsc{LePhare}.
-As expected, and by design, \textsc{TrainZ} has the proper fraction of outliers for the $f_{\rm O}$ statistic.
-
-%\begin{figure*}
-%\centering
-%\includegraphics[width=\textwidth]{fig/PIT_histogram_plot_11codes.jpg}
-%\caption{The PIT histogram plots produced for every photo-$z$ code.
-Ideal behaviour would match a uniform distribution, indicated by the horizontal dashed line.
-Histograms with an overabundance of PIT values at the centre of the distribution indicate $p(z)$ distributions that are overly broad, while an excess of values at the extrema indicate $p(z)$ distributions that are overly narrow.
-Values of PIT=0 and PIT=1 indicate ``catastrophic failures'' where the true redshift is completely outside the support of $p(z)$.
-Asymmetric features are indicative of systematic bias in the redshift predictions.
-%\aim{Could we combine this with the previous figure by plotting both as differences from ideal on one plot with left and right axis labels?
-The information seems degenerate (which makes sense!)}\red{as you say, the information is degenerat, this form emphasizes the differences a bit more, and shows off the catastrophic outliers with PIT=0 and 1 a bit better in my mind.} } \label{fig:pit}
-%\end{figure*}
 \begin{figure*}

### 2019-03-15 paper/tex/analysis.tex
moving content from results
moving content

@@
 %For example, a PIT with nontrivial shape could be due to that behavior being uniform as a function of redshift or significant inaccuracy only in a small range of redshifts where $n(z)$ is large and a flat PIT elsewhere.}
+Following Kodra \& Newman (in prep.) we define the PIT-based catastrophic outlier rate as the fraction of galaxies with $\rm{PIT} < 0.0001$ or $\rm{PIT} > 0.9999$, which should total 0.0002 for an ideal uniform distribution.
 
@@
 %The KS statistic and the CvM and AD statistics related to it are all used to evaluate the performance of both $\hat{n}(z)$ and the QQ and PIT of the CDFs of the ensemble of photo-$z$ interim posteriors.
+In anticipation of a substantial fraction of galaxies having PIT of 0 or 1, a consequence of catastrophic outliers, we evaluate the AD statistic with modified bounds of integration $(0.01, 0.99)$ to exclude those extremes.
 

### 2019-03-15 paper/tex/results.tex
moving content

@@
 \centering
-\includegraphics[width=\textwidth]{fig/KSvsCvMvsAD_PIT_withnull_jpg.jpg}
-\caption{A visual representation of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the PIT distributions.
-The statistics are often highly correlated, with similar relative values between metrics for each code; however, the AD statistic excludes some values at the extrema of the distributions, and can have disparate values compared to KS and CvM.} \label{fig:pit_stats}
+\includegraphics[width=0.74\textwidth]{fig/KSvsCvMvsAD_PIT_withnull_jpg.jpg}
+\caption{A visualization of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the PIT distributions.
+There is generally good agreement between these statistics, with differences corresponding to the codes with outstanding catastrophic outlier rates, a reflection in the differences in how each statistic weights the tails of the distribution.}
+\label{fig:pit_stats}
 \end{figure*}
 
-Fig.~\ref{fig:pit_stats} shows comparative metric values for the quantitative Kolmogorov-Smirnoff (KS), Cramer-Von Mises (CvM), and Anderson Darling (AD) test statistics for each of the codes based on comparing the distribution of their PIT values to the expected uniform distribution over the interval [0,1].
-The individual values of the statistic are not as important as the comparative score between the different codes.
+Fig.~\ref{fig:pit_stats} displays the values of the KS, CvM, and AD test statistics between the PIT distribution and a uniform distribution $U(0, 1)$, highlighting the relative rather than absolute numbers.
 %\red{Can p-values be supplied for each statistic?
 The statistics themselves are difficult to interpret, other than ``lower is better'' (p-value in skgof was broken, having trouble finding 1-sample KS calculation for uniform distribution)}
-The AD test statistic diverges for values that include the extrema, and thus is calculated by excluding the edges of the distribution.
-We calculate the AD statistic over the range of PIT values $v=[0.01,0.99]$.  \textsc{ANNz2} and \textsc{FlexZBoost} score very well for the PIT metrics.
+\textsc{ANNz2} and \textsc{FlexZBoost} score very well for the PIT metrics.
 \textsc{METAPhoR} and \textsc{LePhare} score very well in the PIT AD statistic, but both have a large number of catastrophic outliers, resulting in higher KS and CvM scores.

### 2019-03-15 paper/tex/analysis.tex
moving content

@@
 %The KS statistic and the CvM and AD statistics related to it are all used to evaluate the performance of both $\hat{n}(z)$ and the QQ and PIT of the CDFs of the ensemble of photo-$z$ interim posteriors.
-In anticipation of a substantial fraction of galaxies having PIT of 0 or 1, a consequence of catastrophic outliers, we evaluate the AD statistic with modified bounds of integration $(0.01, 0.99)$ to exclude those extremes.
+In anticipation of a substantial fraction of galaxies having PIT of 0 or 1, a consequence of catastrophic outliers, we evaluate the AD statistic with modified bounds of integration $(0.01, 0.99)$ to exclude those extremes in the name of numerical stability.
 

### 2019-03-15 paper/tex/results.tex
removing redundant text

@@
 This observation is consistent with the findings of \citet[]{Malz:qp} that the piecewise constant parameterization underperforms in the presence of small-scale structures.
-However, the shared small-scale features of \annz, \metaphor, \cmnn, and \skynet\ are a smoking gun for overfitting of the training set, a result of various weighted sums of the limited number of training set galaxies with colors similar to those of the test set galaxy in question.
+However, the shared small-scale features of \annz, \metaphor, \cmnn, and \skynet\ are a result of various weighted sums of the limited number of training set galaxies with colors similar to those of the test set galaxy in question, with behavior closer to classification than regression in the case of \annz.
 The settings used on \gpz\ in this work forced broadening of the single Gaussian to cover the multimodal redshift solutions of the other codes.
@@
 The statistics themselves are difficult to interpret, other than ``lower is better'' (p-value in skgof was broken, having trouble finding 1-sample KS calculation for uniform distribution)}
-\textsc{ANNz2} and \textsc{FlexZBoost} score very well for the PIT metrics.
-\textsc{METAPhoR} and \textsc{LePhare} score very well in the PIT AD statistic, but both have a large number of catastrophic outliers, resulting in higher KS and CvM scores.
-
-Given the near-perfect training data, examining the individual codes for explanations for departures from the expected behaviour will be instructive in avoiding similar problems in future tests.
-\textsc{ANNz2} performs quite well in $p(z)$ based metrics.  In the specific implementation employed in this paper, the final $p(z)$ is a weighted average of five neural-nets.
-During the training process \textsc{ANNz2} compares the percentiles of the redshift training sample against the CDFs of the $p(z)$ sample.
-Distributions that more closely match are given extra weight, and the final weights are designed to produce accurate percentiles.
-Given that our metrics are focused on the percentile distributions, it is unsurprising that \textsc{ANNz2} performs well in the given metrics.
-The discreteness in the individual $p(z)$ estimated by \textsc{ANNz2} can be attributed to the fact that the code was run as a classifier, assigning weights to discrete bins of redshift.
-While multiple bins may receive weight, the bins themselves will still be discretized, and no additional smoothing was performed.
-Overall, \textsc{FlexZBoost} and \textsc{ANNz2} show the best ensemble agreement in their distribution of PIT values.
+\metaphor\ and \lephare\ perform well under the AD but poorly under the KS and CvM due to their high catastrophic outlier rates.
+\annz\ and \flexzboost\  are the top scorers under these metrics of the PIT distribution.
+\annz's strong performance can be attributed to an aspect of the training process in which training set galaxies with a PIT that more closely matches the percentiles of the DC1 training set's redshift distribution are upweighted; in effect, these quantile-based metrics were part of the algorithm itself.
 

### 2019-03-15 paper/tex/results.tex
removing redundant text

@@
 \annz\ and \flexzboost\  are the top scorers under these metrics of the PIT distribution.
-\annz's strong performance can be attributed to an aspect of the training process in which training set galaxies with a PIT that more closely matches the percentiles of the DC1 training set's redshift distribution are upweighted; in effect, these quantile-based metrics were part of the algorithm itself.
+\annz's strong performance can be attributed to an aspect of the training process in which training set galaxies with a PIT that more closely matches the percentiles of the DC1 training set's redshift distribution are upweighted; in effect, these quantile-based metrics were part of the algorithm itself that may or may not serve it well under more realistic experimental conditions.
 

### 2019-03-15 paper/tex/results.tex
moving Scotts Rule and resizing excessively large figures

@@
 \centering
-\includegraphics[width=\textwidth]{fig/NZsumplot_12codes_scottsrule.jpg}
-\caption{The stacked $p(z)$ produced by each photo-$z$ code ($\hat{N}(z)$, red) compared to the spectroscopic redshift distribution ($N'(z)$, blue).
-Varying levels of agreement are seen in the codes.
-Both $\hat{N}(z)$ and $N'(z)$ in all codes are smoothed using a single bandwidth chosen via Scott's rule.}
+\includegraphics[width=0.74\textwidth]{fig/NZsumplot_12codes_scottsrule.jpg}
+\caption{The smoothed stacked estimator $\hat{N}(z)$ of the redshift distribution (red) produced by each code (panels) compared to the true redshift distribution $\tilde{N}(z)$ (blue).
+Varying levels of agreement are seen among the codes, with the smallest deviations for \cmnn, \flexzboost, \tpz, and \trainz.}
 \label{fig:nz}
@@
 
-\begin{figure*}
+\begin{figure}
 \centering
-\includegraphics[width=\textwidth]{fig/KSvsCvMvsAD_NZ_withnull_jpg.jpg}
-\caption{A visual representation of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the $\hat{N}(z)$ distributions.
-The statistics are correlated, the codes with the lowest KS statistics tend to have the lowest CvM and AD statistics.
-\textsc{CMNN} performs markedly better than the others in reconstructing the overall $N(z)$ distribution, while \textsc{SkyNet} scores poorly due to an overall bias in its redshift predictions.}
+\includegraphics[width=0.49\textwidth]{fig/KSvsCvMvsAD_NZ_withnull_jpg.jpg}
+\caption{A visualization of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the $\hat{N}(z)$ distributions.
+We make the observation that these related statistics do not disagree significantly with one another.
+\cmnn\ performs comparably well to \trainz, the control case, while \textsc{SkyNet} scores poorly due to an overall bias in its redshift predictions.}
 \label{fig:nz_stats}
-\end{figure*}
+\end{figure}
 

### 2019-03-15 paper/tex/results.tex
moving all the things around

@@
 
+Fig.~\ref{fig:nz} shows the stacked estimator $\hat{N}(z)$ of the redshift distribution for each code compared to the true redshift distribution $\tilde{N}(z)$, where the stacked estimator has been smoothed for each code in the plot using a kernel density estimate (KDE) with a bandwidth chosen by Scott's Rule \citep{Scott:1992} in order to minimize visual differences in small-scale features; the quantiative statistics, however, are calculated using the empirical CDF which is not smoothed.
+
 \begin{figure*}
@@
 
-\begin{figure}
+Many of the codes, including all the model-fitting approaches and \annz, \gpz, \metaphor, and \skynet\ from the data-driven camp, overestimate the redshift density at $z \sim 1.4$.
+This behavior is a consequence of the $4000\ {\rm \AA}$ break passing through the gap between the $z$ and $y$ filters, which induces a genuine discontinuity in the $z - y$ colour as a function of redshift that can sway the \pzpdf\ estimates in the absence of bluer spectral features.
+
+\annz, \gpz, and \metaphor\ show signs of overtraining, estimating enhanced peaks and diminished troughs relative to the training set, an obstacle that may be overcome with adjustment of the implementation.
+
+\skynet\ exhibits an obvious redshift bias, evident both visually in Figure~\ref{fig:nz} and in the first moment of N(z) listed in Table~\ref{tab:moments}, where it is clearly an outlier.
+\textsc{SkyNet} employed a method where a random sample of training galaxies was chosen, but there was no test that the subset was completely representative of the overall redshift distribution.
+Unlike the previous implementation of \textsc{SkyNet} in \citet{Bonnett:15}, no effort was made to add extra weight to more rare low and high redshift galaxies.
+Either of these decisions could be the cause of the bias seen in our results.
+Future runs of \textsc{SkyNet} will explore these implementation choices and their effects.
+
+As expected, \trainz\ perfectly recovers the true redshift distribution: as the training sample is selected from the same underlying distribution as the test set, the redshift distributions are identical, up to Poisson fluctuations due to the finite number of sample galaxies.
+\cmnn\ is also in excellent agreement for similar reasons: with a representative training sample of galaxies spanning the colour-space, the sum of the colour-matched neighbour redshifts should return the true redshift distribution.
+\flexzboost\ and \tpz\ also perform superb recovery of the true redshift distribution, with only a slight deviation at $z \sim 1.4$.
+Our metrics, however, cannot discern whether these four approaches, as well as \delight, are spared the $z \sim 1.4$ degeneracy in $\hat{N}(z)$ because they have more effectively used information in the data or if the impact is simply washed out by the effective average over the test set galaxy sample.
+See Appendix~\ref{sec:pointmetrics} for further discussion of the $z \sim 1.4$ issue.
+
+\begin{figure*}
 \centering
-\includegraphics[width=0.49\textwidth]{fig/KSvsCvMvsAD_NZ_withnull_jpg.jpg}
+\includegraphics[width=0.74\textwidth]{fig/KSvsCvMvsAD_NZ_withnull_jpg.jpg}
 \caption{A visualization of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the $\hat{N}(z)$ distributions.
@@
 \label{fig:nz_stats}
-\end{figure}
-
-Fig.~\ref{fig:nz} shows the stacked $\hat{N}(z)$ distribution compared to the true redshift distribution $N'(z)$ for all tested codes.
-The red line indicates the summed $p(z)$ for each code, while the blue line shows the true redshift distribution.
-All distributions are smoothed via kernel density estimation (KDE) with a common bandwidth chosen via Scott's rule \citep{Scott:1992} in order to minimize differences in small-scale features and make for a more uniform comparison between codes.
-While Scott's rule is used to display $N'(z)$ in the figure, all quantitative statistics are computed via the empirical CDF, and are thus unaffected by bandwidth/smoothing choice.
-%New text for the new figure
-As expected, \textsc{TrainZ} is in excellent agreement with the true redshift distribution: as the training sample is selected from the same underlying distribution as the test set, the redshift distributions are identical, up to Poisson fluctuations due to the finite number of sample galaxies.
-\textsc{CMNN} is also in excellent agreement for similar reasons: with a representative training sample of galaxies spanning the colour-space, the sum of the colour-matched neighbour redshifts should return the true redshift distribution.
-\textsc{FlexZBoost} and \textsc{TPZ} also show very good agreement, with only slight departures, with an over/under-prediction in the high redshift tail of $\hat{N}(z)$ evident around $z\sim1.4$.
-In fact, several of the other codes show an excess at $z \sim 1.4$, particularly the template-based codes \textsc{BPZ}, \textsc{EAZY}, and \textsc{LePhare}.
-This is likely due to the $4000\ {\rm \AA}$ break passing through the gap between the $z$ and $y$ filters, resulting in a drastic change in $z-y$ colour for galaxies in this redshift range.
-With a relative dearth of strong features blue-ward of the $4000\ {\rm \AA}$ break in most galaxy SEDs, the colour change in the two reddest filter bandpasses of a survey has a large influence on the redshift determination.
-The $z\sim1.4$ feature is one of the most prominent sources of larger uncertainty in individual galaxy $p(z)$.
-In our sample individual galaxy $p(z)$'s tend to be broader around $z\sim1.4$ and point estimates are more uncertain in this regime, as is readily seen in the point-estimate plots shown in Fig.~\ref{fig:pz_pointestimates} and described in the Appendix.
-This feature is not unique to this dataset, it is a common occurrence in photo-$z$ estimation.
-The fact that similar excesses appear in Figure~\ref{fig:nz} for \textsc{ANNz2} and \textsc{METAPhoR} shows that the effect is not limited to template-based codes.
-However, the lack of such a feature in the other codes shows that it is possible to eliminate the degeneracies.
-Further study on this issue may provide a solution for codes that suffer from this shortcoming.
-
-%end of new text (and slightly modified text)
-Two of the machine learning based codes, \textsc{ANNz2} and \textsc{METAPhoR}, appear to be over-trained, adding excess galaxy probability to the redshift peaks with the largest number of training galaxies, and missing probability in the troughs where training galaxies are of fewer number.
-Given that our training data is drawn from the same galaxy population as the test set, and our data has prominent peaks in $N'(z)$, perhaps it is not unexpected that such overtraining occurs in some codes, though the fact that it does not occur in all training-based codes indicates that it may be due to specifics of the implementations and bandwidth choices of \textsc{ANNz2} and \textsc{METAPhoR}.
-This, once again, emphasizes that care much be taken in choice of bandwith parameters in individual codes in order to avoid such overtraining.
-%\aim{Move the text on Scott's Rule to Discussion?}
-%Scott's rule \citep{Scott:1992} was used to determine the smoothing bandwidth of the true redshift sample: that is, a single smoothing scale was chosen to represent the true redshift sample for all twelve codes.
-%As with the $p(z)$ values in Figure~\ref{fig:pitqq}, different levels of substructure are obvious for the different codes.
-%While Scott's rule provides a relatively good general smoothing scale to represent the true $N'(z)$, there are smaller scale fluctuations: while \textsc{FlexZBoost} and \textsc{CMNN} appear somewhat discrepant in Fig.~\ref{fig:nz}, they are actually the two most accurate in terms of their quantitative measurements.
-%\textsc{SkyNet} also shows excess small-scale structure, but also shows significant overestimate of the high redshift galaxy population.
-\textsc{SkyNet} shows an obvious redshift bias, evident both visually in Figure~\ref{fig:nz} and in the first moment of N(z) listed in Table~\ref{tab:moments}, where it is clearly an outlier.
-\textsc{SkyNet} employed a method where a random sample of training galaxies was chosen, but there was no test that the subset was completely representative of the overall redshift distribution.
-Unlike the previous implementation of \textsc{SkyNet} in \citet{Bonnett:15}, no effort was made to add extra weight to more rare low and high redshift galaxies.
-Either of these decisions could be the cause of the bias seen in our results.
-Future runs of \textsc{SkyNet} will explore these implementation choices and their effects.
+\end{figure*}
 

### 2019-03-15 paper/tex/results.tex
reordering section

@@
 We begin with a demonstrative visual inspection of the \pzpdf s produced by each code for individual galaxies.
-Fig.~\ref{fig:pz_examples} shows the \pzpdf s for four galaxies chosen as examples of \pzpdf\ archetypes: a narrow unimodal redshift solution, a broad unimodal solution, a bimodal distribution, and a multimodal distribution.
+Figure~\ref{fig:pz_examples} shows the \pzpdf s for four galaxies chosen as examples of \pzpdf\ archetypes: a narrow unimodal redshift solution, a broad unimodal solution, a bimodal distribution, and a multimodal distribution.
 
@@
 
-\subsection{\Pzpdf\ ensembles}
+\subsection{Performance on \pzpdf\ ensembles}
 \label{sec:pitqq}
 
-Fig.~\ref{fig:pitqq} shows a histogram of PIT values, QQ plot, and QQ difference plot relative to the ideal diagonal, showcasing the biases and trends in the average accuracy of the \pzpdf s for each code.
+Figure~\ref{fig:pitqq} shows a histogram of PIT values, QQ plot, and QQ difference plot relative to the ideal diagonal, showcasing the biases and trends in the average accuracy of the \pzpdf s for each code.
 The high QQ values (more high than low PIT values) of \bpz, \cmnn, \delight, \eazy, and \gpz\ indicate \pzpdf s biased low, and the low QQ values (more low than high PIT values) of \skynet\ and \tpz\ indicate \pzpdf s biased high.
@@
 
-Fig.~\ref{fig:pit_stats} displays the values of the KS, CvM, and AD test statistics between the PIT distribution and a uniform distribution $U(0, 1)$, highlighting the relative rather than absolute numbers.
+Figure~\ref{fig:pit_stats} displays the values of the KS, CvM, and AD test statistics between the PIT distribution and a uniform distribution $U(0, 1)$, highlighting the relative rather than absolute numbers.
 %\red{Can p-values be supplied for each statistic?
@@
 
-\subsection{Metrics of the stacked estimator of the redshift distribution}
+\subsection{Performance on the stacked estimator of the redshift distribution}
 \label{sec:stackedmetrics}
 
-Fig.~\ref{fig:nz} shows the stacked estimator $\hat{N}(z)$ of the redshift distribution for each code compared to the true redshift distribution $\tilde{N}(z)$, where the stacked estimator has been smoothed for each code in the plot using a kernel density estimate (KDE) with a bandwidth chosen by Scott's Rule \citep{Scott:1992} in order to minimize visual differences in small-scale features; the quantiative statistics, however, are calculated using the empirical CDF which is not smoothed.
+Figure~\ref{fig:nz} shows the stacked estimator $\hat{N}(z)$ of the redshift distribution for each code compared to the true redshift distribution $\tilde{N}(z)$, where the stacked estimator has been smoothed for each code in the plot using a kernel density estimate (KDE) with a bandwidth chosen by Scott's Rule \citep{Scott:1992} in order to minimize visual differences in small-scale features; the quantiative statistics, however, are calculated using the empirical CDF which is not smoothed.
 
@@
 
-\skynet\ exhibits an obvious redshift bias, evident both visually in Figure~\ref{fig:nz} and in the first moment of N(z) listed in Table~\ref{tab:moments}, where it is clearly an outlier.
-\textsc{SkyNet} employed a method where a random sample of training galaxies was chosen, but there was no test that the subset was completely representative of the overall redshift distribution.
-Unlike the previous implementation of \textsc{SkyNet} in \citet{Bonnett:15}, no effort was made to add extra weight to more rare low and high redshift galaxies.
-Either of these decisions could be the cause of the bias seen in our results.
-Future runs of \textsc{SkyNet} will explore these implementation choices and their effects.
-
 As expected, \trainz\ perfectly recovers the true redshift distribution: as the training sample is selected from the same underlying distribution as the test set, the redshift distributions are identical, up to Poisson fluctuations due to the finite number of sample galaxies.
@@
 \flexzboost\ and \tpz\ also perform superb recovery of the true redshift distribution, with only a slight deviation at $z \sim 1.4$.
-Our metrics, however, cannot discern whether these four approaches, as well as \delight, are spared the $z \sim 1.4$ degeneracy in $\hat{N}(z)$ because they have more effectively used information in the data or if the impact is simply washed out by the effective average over the test set galaxy sample.
+Our metrics, however, cannot discern whether these four approaches, as well as \delight, are spared the $z \sim 1.4$ degeneracy in $\hat{N}(z)$ because they have more effectively used information in the data or if the impact is simply washed out by the stacked estimator's effective average over the test set galaxy sample.
 See Appendix~\ref{sec:pointmetrics} for further discussion of the $z \sim 1.4$ issue.
+
+Figure~\ref{fig:nz_stats} shows the quantitative Kolmogorov-Smirnoff (KS), Cramer-Von Mises (CvM), and Anderson Darling (AD) test statistics for each of the codes for the $\hat{N}(z)$ based measures.
+%\red{Can p-values be supplied for each statistic?
+The statistics themselves are difficult to interpret, other than ``lower is better'' (no, p-values are very difficult to compute for non-uniform distributions)}
+The stacked estimators of the redshift distribution for \cmnn\ and \trainz\ best estimate $\tilde{N}(z)$ under these metrics, and the only codes that do especially poorly are \eazy, \lephare, \metaphor, and \skynet.
+It is unsurprising that \cmnn\ scores well, as with a near perfectly representative training set means that choosing neighbouring points in color/magnitude space should lead to excellent agreement in the final $\hat{N}(z)$ estimate.
 
@@
 \caption{A visualization of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the $\hat{N}(z)$ distributions.
-We make the observation that these related statistics do not disagree significantly with one another.
-\cmnn\ performs comparably well to \trainz, the control case, while \textsc{SkyNet} scores poorly due to an overall bias in its redshift predictions.}
+We make the reassuring observation that these related statistics do not disagree significantly with one another.
+\cmnn\ performs comparably well to \trainz, the control case, while \skynet\ scores poorly due to an overall bias in its redshift predictions.}
 \label{fig:nz_stats}
@@
 
-Figure~\ref{fig:nz_stats} shows the quantitative Kolmogorov-Smirnoff (KS), Cramer-Von Mises (CvM), and Anderson Darling (AD) test statistics for each of the codes for the $\hat{N}(z)$ based measures.
-%\red{Can p-values be supplied for each statistic?
-The statistics themselves are difficult to interpret, other than ``lower is better'' (no, p-values are very difficult to compute for non-uniform distributions)}
-\textsc{FlexZBoost}, \textsc{CMNN}, and \textsc{TPZ} outperform the other codes in the $\hat{N}(z)$ metrics.
-It is unsurprising that \textsc{CMNN} scores well, as with a near perfectly representative training set means that choosing neighbouring points in color/magnitude space should lead to excellent agreement in the final $\hat{N}(z)$ estimate.
-\textsc{TPZ} performed quite poorly in $p(z)$ statistics, but results in a good fit to the overall $N(z)$.
-This is somewhat surprising, as performance was optimized for accurate $p(z)$, not $\hat{N}(z)$.
-During the validation stage for \textsc{TPZ}, there was a trade off between the width of the $p(z)$ when adjusting a smoothing parameter and overall redshift bias.
-The optimal result in the PIT metrics, as illustrated in the shape of the QQ plot, does contain some level of bias as well as a slight underprediction of mean $p(z)$ width, which translates to poor metric scores.
-This is something that will be looked into for \textsc{TPZ} in the future.
+It is, however, surprising that \tpz\ does well on $\hat{N}(z)$ given its poor performance on the ensemble \pzpdf, especially knowing that \tpz\ was optimized for \pzpdf\ ensemble metrics rather than the stacked estimator of the redshift distribution.
+A possible explanation is the choice of smoothing parameter chosen during validation, which affects \pzpdf\ widths as well as overall redshift bias and could be modified to improve performance under the \pzpdf\ metrics.
 
-%\red{discussion of why NN and FZBoost are doing well, how we can try to improve the other codes}
-
-Table~\ref{tab:cdeloss} shows the CDE loss statistic for each photo-$z$ code.
-Once again \textsc{FlexZBoost} and \textsc{CMNN} score very well for the stacked $\hat{N}(z)$ metrics, as do \textsc{GPz} and \textsc{TPZ}.
-The CDE loss measures how well individual PDFs are estimated, and codes with a low CDE loss tend to have good $\hat{N}(z)$ estimates (though the reverse is not necessarily true).
-\textsc{FlexZBoost} is optimized to minimize CDE loss which may explain why the method has good ensemble metrics as well.
-Note from Table~\ref{tab:cdeloss} that both \textsc{FlexZBoost} and \textsc{CMNN} have low CDE losses.
-%Empirically, we have found that PIT RMSE is not as closely correlated to CDE loss as it is to the $N(z)$ statistics.
-As CDE loss is a better measure of individual redshift performance, rather than ensemble distribution performance, this statistic is a better indicator of which codes will be most likely to perform well for science cases where single objects are employed.
-
-%Table~\ref{tab:rmse} gives the root-mean-square-error (RMSE) statistics for both the PIT and N(z) estimators.
-The PIT value calculates the RMSE between the quantiles shown in the QQ plot in Figure~\ref{fig:pitqq} and the diagonal, while the N(z) calculates the RMSE between the cumulative distribution of the stacked $\hat{N}(z)$ and the true redshift distribution $N'(z)$.
-%\red{more about RMSE?
-Or remove it?
-It's barely described in Section~\ref{sec:metrics} and wasn't even referred to in the text until I added this on June 29th--SJS.}
+\begin{table}
+\setlength{\tabcolsep}{2pt}
+\caption{Moments of the stacked $\hat{N}(z)$ distribution}\label{tab:moments}
+\begin{tabular}{lccc}
+\hline
+\hline
+ \multicolumn{4}{l}{Stacked $n(z)$ Moments} \\
+\hline
+              & 1st Moment & 2nd Moment & 3rd Moment \\
+\textsc{Truth}     & 0.701      &   0.630    & 0.671  \\
+\hline
+Photo-z Code       & 1st Moment & 2nd Moment & 3rd Moment\\
+\hline
+\textsc{ANNz2}     & 0.702      & 0.625      & 0.653    \\
+\textsc{BPZ}       & 0.699      & 0.629      & 0.671    \\
+\textsc{Delight}   & 0.692      & 0.609      & 0.638    \\
+\textsc{EAZY}      & 0.681      & 0.595      & 0.619    \\
+\textsc{FlexZBoost}& 0.694      & 0.610      & 0.631    \\
+\textsc{GPz}       & 0.696      & 0.615      & 0.639    \\
+\textsc{LePhare}   & 0.718      & 0.668      & 0.741    \\
+\textsc{METAPhoR}  & 0.705      & 0.628      & 0.657    \\
+\textsc{CMNN}        & 0.701      & 0.628      & 0.667    \\
+\textsc{Skynet}    & 0.743      & 0.708      & 0.797    \\
+\textsc{TPZ}       & 0.700      & 0.619      & 0.643    \\
+\hline
+\trainz	   & 0.699 		& 0.627 	& 0.666 \\
+\end{tabular}
+\end{table}
 
@@
 Discuss this after talking with Rafael.}
+
+\skynet\ exhibits an obvious redshift bias, evident both visually in Figure~\ref{fig:nz} and in the first moment of N(z) listed in Table~\ref{tab:moments}, where it is clearly an outlier.
+\textsc{SkyNet} employed a method where a random sample of training galaxies was chosen, but there was no test that the subset was completely representative of the overall redshift distribution.
+Unlike the previous implementation of \textsc{SkyNet} in \citet{Bonnett:15}, no effort was made to add extra weight to more rare low and high redshift galaxies.
+Either of these decisions could be the cause of the bias seen in our results.
+Future runs of \textsc{SkyNet} will explore these implementation choices and their effects.
+
+\subsection{Performance on individual \pzpdf s}
+
+\begin{table}  %%% DATA TABLE %%%
+\centering
+\caption{CDE loss statistic for each photo-$z$ code.} \label{tab:cdeloss}
+\begin{tabular}{lr}
+\hline
+\bf Photo-$z$ Code & \bf CDE Loss \\
+\hline
+\textsc{ANNz2} 		& $-6.88$ \\
+\textsc{BPZ} 		& $-7.82$ \\
+%\textsc{Delight} 	& $-4.06$ \\
+\textsc{Delight}    & $-8.33$\\
+%\textsc{EAZY} 		& $-7.97$ \\
+\textsc{EAZY}       & $-7.07$ \\
+%\textsc{FlexZBoost} & $-11.51$ \\
+\textsc{FlexZBoost} & $-10.60$\\
+\textsc{GPz} 		& $-9.93$ \\
+\textsc{LePhare} 	& $-1.66$ \\
+\textsc{METAPhoR} 	& $-6.28$ \\
+%\textsc{CMNN} 		& $-$ \\
+\textsc{CMNN}         & $-10.43$ \\
+\textsc{SkyNet} 	& $-7.89$ \\
+\textsc{TPZ} 		& $-9.55$ \\
+\hline
+\trainz		& $-0.83$ \\
+%\textsc{Frankenz}	& $-$  \\
+%\hline
+\end{tabular}
+\end{table}
+
+Table~\ref{tab:cdeloss} contains the CDE loss statistic for each code.
+Once again \textsc{FlexZBoost} and \textsc{CMNN} score very well for the stacked $\hat{N}(z)$ metrics, as do \textsc{GPz} and \textsc{TPZ}.
+The CDE loss measures how well individual PDFs are estimated, and codes with a low CDE loss tend to have good $\hat{N}(z)$ estimates (though the reverse is not necessarily true).
+\textsc{FlexZBoost} is optimized to minimize CDE loss which may explain why the method has good ensemble metrics as well.
+Note from Table~\ref{tab:cdeloss} that both \textsc{FlexZBoost} and \textsc{CMNN} have low CDE losses.
+%Empirically, we have found that PIT RMSE is not as closely correlated to CDE loss as it is to the $N(z)$ statistics.
+As CDE loss is a better measure of individual redshift performance, rather than ensemble distribution performance, this statistic is a better indicator of which codes will be most likely to perform well for science cases where single objects are employed.
 
@@
 
-\begin{table}  %%% DATA TABLE %%%
-\centering
-\caption{CDE loss statistic for each photo-$z$ code.} \label{tab:cdeloss}
-\begin{tabular}{lr}
-\hline
-\bf Photo-$z$ Code & \bf CDE Loss \\
-\hline
-\textsc{ANNz2} 		& $-6.88$ \\
-\textsc{BPZ} 		& $-7.82$ \\
-%\textsc{Delight} 	& $-4.06$ \\
-\textsc{Delight}    & $-8.33$\\
-%\textsc{EAZY} 		& $-7.97$ \\
-\textsc{EAZY}       & $-7.07$ \\
-%\textsc{FlexZBoost} & $-11.51$ \\
-\textsc{FlexZBoost} & $-10.60$\\
-\textsc{GPz} 		& $-9.93$ \\
-\textsc{LePhare} 	& $-1.66$ \\
-\textsc{METAPhoR} 	& $-6.28$ \\
-%\textsc{CMNN} 		& $-$ \\
-\textsc{CMNN}         & $-10.43$ \\
-\textsc{SkyNet} 	& $-7.89$ \\
-\textsc{TPZ} 		& $-9.55$ \\
-\hline
-\trainz		& $-0.83$ \\
-%\textsc{Frankenz}	& $-$  \\
-%\hline
-\end{tabular}
-\end{table}
-
 %\begin{table}
@@
 %\end{table}
-
-\begin{table}
-\setlength{\tabcolsep}{2pt}
-\caption{Moments of the stacked $\hat{N}(z)$ distribution}\label{tab:moments}
-\begin{tabular}{lccc}
-\hline
-\hline
- \multicolumn{4}{l}{Stacked $n(z)$ Moments} \\
-\hline
-              & 1st Moment & 2nd Moment & 3rd Moment \\
-\textsc{Truth}     & 0.701      &   0.630    & 0.671  \\
-\hline
-Photo-z Code       & 1st Moment & 2nd Moment & 3rd Moment\\
-\hline
-\textsc{ANNz2}     & 0.702      & 0.625      & 0.653    \\
-\textsc{BPZ}       & 0.699      & 0.629      & 0.671    \\
-\textsc{Delight}   & 0.692      & 0.609      & 0.638    \\
-\textsc{EAZY}      & 0.681      & 0.595      & 0.619    \\
-\textsc{FlexZBoost}& 0.694      & 0.610      & 0.631    \\
-\textsc{GPz}       & 0.696      & 0.615      & 0.639    \\
-\textsc{LePhare}   & 0.718      & 0.668      & 0.741    \\
-\textsc{METAPhoR}  & 0.705      & 0.628      & 0.657    \\
-\textsc{CMNN}        & 0.701      & 0.628      & 0.667    \\
-\textsc{Skynet}    & 0.743      & 0.708      & 0.797    \\
-\textsc{TPZ}       & 0.700      & 0.619      & 0.643    \\
-\hline
-\trainz	   & 0.699 		& 0.627 	& 0.666 \\
-\end{tabular}
-\end{table}

### 2019-03-15 paper/tex/results.tex
moving content

@@
 
-It is, however, surprising that \tpz\ does well on $\hat{N}(z)$ given its poor performance on the ensemble \pzpdf, especially knowing that \tpz\ was optimized for \pzpdf\ ensemble metrics rather than the stacked estimator of the redshift distribution.
+It is, however, surprising that \tpz\ does well on $\hat{N}(z)$ given its poor performance on the ensemble \pzpdf s, especially knowing that \tpz\ was optimized for \pzpdf\ ensemble metrics rather than the stacked estimator of the redshift distribution.
 A possible explanation is the choice of smoothing parameter chosen during validation, which affects \pzpdf\ widths as well as overall redshift bias and could be modified to improve performance under the \pzpdf\ metrics.
+
+Table~\ref{tab:moments} lists the first three moments of the stacked $\hat{N}(z)$ distribution relative to the empirical estimate of the truth distribution.
+Accuracy of the moments varies widely between codes, raising concerns about the propagation to cosmological analyses.
+%\red{[mention that we are calculating moments for the entire sample, will look at fiducial tomographic bins in a follow-up paper (unless group and reviewers realy think we should include in this paper.]}
+%\red{FlexZBoost has some of the best $n(z)$ statistics, but the moments are not good.
+Why is this?
+Is it the over/underprediction of the high redshift part of the distribution?
+Discuss this after talking with Rafael.}
 
@@
 \setlength{\tabcolsep}{2pt}
-\caption{Moments of the stacked $\hat{N}(z)$ distribution}\label{tab:moments}
+\caption{Moments of the stacked $\hat{n}(z)$ distribution}\label{tab:moments}
 \begin{tabular}{lccc}
@@
               & 1st Moment & 2nd Moment & 3rd Moment \\
-\textsc{Truth}     & 0.701      &   0.630    & 0.671  \\
+Empirical estimate of truth & 0.701 & 0.630 & 0.671  \\
 \hline
@@
 \hline
-\textsc{ANNz2}     & 0.702      & 0.625      & 0.653    \\
-\textsc{BPZ}       & 0.699      & 0.629      & 0.671    \\
-\textsc{Delight}   & 0.692      & 0.609      & 0.638    \\
-\textsc{EAZY}      & 0.681      & 0.595      & 0.619    \\
-\textsc{FlexZBoost}& 0.694      & 0.610      & 0.631    \\
-\textsc{GPz}       & 0.696      & 0.615      & 0.639    \\
-\textsc{LePhare}   & 0.718      & 0.668      & 0.741    \\
-\textsc{METAPhoR}  & 0.705      & 0.628      & 0.657    \\
-\textsc{CMNN}        & 0.701      & 0.628      & 0.667    \\
-\textsc{Skynet}    & 0.743      & 0.708      & 0.797    \\
-\textsc{TPZ}       & 0.700      & 0.619      & 0.643    \\
-\hline
+\annz       & 0.702      & 0.625      & 0.653    \\
+\bpz        & 0.699      & 0.629      & 0.671    \\
+\delight    & 0.692      & 0.609      & 0.638    \\
+\eazy       & 0.681      & 0.595      & 0.619    \\
+\flexzboost & 0.694      & 0.610      & 0.631    \\
+\gpz        & 0.696      & 0.615      & 0.639    \\
+\lephare    & 0.718      & 0.668      & 0.741    \\
+\metaphor   & 0.705      & 0.628      & 0.657    \\
+\cmnn       & 0.701      & 0.628      & 0.667    \\
+\skynet     & 0.743      & 0.708      & 0.797    \\
+\tpz        & 0.700      & 0.619      & 0.643    \\
 \trainz	    & 0.699 		 & 0.627 	    & 0.666 \\
@@
 \end{table}
-
-Table~\ref{tab:moments} lists the first three moments of the stacked $\hat{N}(z)$ distribution, including the moments of the ``truth'' distribution for comparison.
-Several codes are able to reproduce the mean and variance of the distribution to less than a per cent, while several codes do not, which may be a cause for concern, given that mean and variance of the redshift distribution are key properties in cosmological analyses.
-We note that this stated goal of the study as defined for participants was to accurately reproduce $p(z)$, the ``stacking'' of the probability distributions to estimate $\hat{N}(z)$ was not the focus as stated to the participants.
-This explains why some of the best-performing empirical codes in terms of $p(z)$ measures (e.~g.~\textsc{FlexZBoost}) do not do as well at reproducing $\hat{N}(z)$ moments.
-Had we defined a different parameter to optimize, in this case overall accuracy of $\hat{N}(z)$ rather than individual $p(z)$, would result in improved performance in a particular metric.
-That is, optimizing photo-$z$ performance for one metric does not automatically give optimal performance for other metrics.
-As previously stated, there are a variety of scientific use cases for photo-$z$'s in large upcoming surveys, and care must be taken in how the metrics used to optimize catalog photometric redshifts are defined as well as in how they are used.
-This implies that we may need multiple photo-z estimators, tuned to the particular metric, in order to maximize returns over all science cases in upcoming surveys.
-In addition, very few scientific use cases will employ the overall $\hat{N}(z)$ with no cuts, as we explore in this paper.
-We discuss more realistic tomographic bin selections that will be explored in a follow-up paper in Section~\ref{sec:futurework}.
-%\red{[mention that we are calculating moments for the entire sample, will look at fiducial tomographic bins in a follow-up paper (unless group and reviewers realy think we should include in this paper.]}
-%\red{FlexZBoost has some of the best $n(z)$ statistics, but the moments are not good.
-Why is this?
-Is it the over/underprediction of the high redshift part of the distribution?
-Discuss this after talking with Rafael.}
 
@@
 %(Alex Malz, Sam Schmidt)
+
+We remind the reader that contributed codes were given a goal of obtaining accurate \pzpdf s, not an accurate stacked estimator of the redshift distribution, so we do not expect the same codes to necessarily perform well for both classes of metrics.
+As previously stated, there are a variety of scientific use cases for \pz s, and metric-specific relative performance implies that we may need multiple \pzpdf\ approaches tuned to each metric in order to maximize returns over all science cases in large upcoming surveys.
 

### 2019-03-15 paper/tex/results.tex
edited CDE loss section for messaging

@@
 
-Table~\ref{tab:moments} lists the first three moments of the stacked $\hat{N}(z)$ distribution relative to the empirical estimate of the truth distribution.
+The first three moments of the stacked $\hat{N}(z)$ distribution relative to the empirical estimate of the truth distribution are given in Table~\ref{tab:moments}.
 Accuracy of the moments varies widely between codes, raising concerns about the propagation to cosmological analyses.
@@
 \hline
-              & 1st Moment & 2nd Moment & 3rd Moment \\
-Empirical estimate of truth & 0.701 & 0.630 & 0.671  \\
-\hline
-Photo-z Code       & 1st Moment & 2nd Moment & 3rd Moment\\
+              & mean & variance & skewness \\
+Empirical ``truth'' & 0.701 & 0.630 & 0.671  \\
 \hline
@@
 
-\skynet\ exhibits an obvious redshift bias, evident both visually in Figure~\ref{fig:nz} and in the first moment of N(z) listed in Table~\ref{tab:moments}, where it is clearly an outlier.
-\textsc{SkyNet} employed a method where a random sample of training galaxies was chosen, but there was no test that the subset was completely representative of the overall redshift distribution.
-Unlike the previous implementation of \textsc{SkyNet} in \citet{Bonnett:15}, no effort was made to add extra weight to more rare low and high redshift galaxies.
-Either of these decisions could be the cause of the bias seen in our results.
-Future runs of \textsc{SkyNet} will explore these implementation choices and their effects.
+\skynet\ exhibits redshift bias in Figure~\ref{fig:nz} and is a clear outlier in the first moment of $\hat{N}(z)$ in Table~\ref{tab:moments}.
+The \skynet\ algorithm employs a random subsampling of the training set without testing that the subset is representative of the full population, and the implementation used here does not upweight rarer low- and high-redshift galaxies, as in \citet{Bonnett:15}, suggesting a possible cause that may be addressed in future work.
 
 \subsection{Performance on individual \pzpdf s}
+\label{sec:cdelossresults}
+
+The values of the CDE loss statistic of individual \pzpdf\ accuracy are provided in Table~\ref{tab:cdeloss}.
+It is worth noting that strong performance on the CDE loss should imply strong performance on the other metrics, though the inverse is not necessarily true.
+Thus the CDE loss is the most effective metric for generic science cases.
 
@@
 \hline
-\textsc{ANNz2} 		& $-6.88$ \\
-\textsc{BPZ} 		& $-7.82$ \\
+\annz 		& $-6.88$ \\
+\bpz 		& $-7.82$ \\
 %\textsc{Delight} 	& $-4.06$ \\
-\textsc{Delight}    & $-8.33$\\
+\delight    & $-8.33$\\
 %\textsc{EAZY} 		& $-7.97$ \\
-\textsc{EAZY}       & $-7.07$ \\
+\eazy       & $-7.07$ \\
 %\textsc{FlexZBoost} & $-11.51$ \\
-\textsc{FlexZBoost} & $-10.60$\\
-\textsc{GPz} 		& $-9.93$ \\
-\textsc{LePhare} 	& $-1.66$ \\
-\textsc{METAPhoR} 	& $-6.28$ \\
+\flexzboost & $-10.60$\\
+\gpz		& $-9.93$ \\
+\lephare 	& $-1.66$ \\
+\metaphor 	& $-6.28$ \\
 %\textsc{CMNN} 		& $-$ \\
-\textsc{CMNN}         & $-10.43$ \\
-\textsc{SkyNet} 	& $-7.89$ \\
-\textsc{TPZ} 		& $-9.55$ \\
+\cmnn         & $-10.43$ \\
+\skynet 	& $-7.89$ \\
+\tpz 		& $-9.55$ \\
 \hline
@@
 
-Table~\ref{tab:cdeloss} contains the CDE loss statistic for each code.
-Once again \textsc{FlexZBoost} and \textsc{CMNN} score very well for the stacked $\hat{N}(z)$ metrics, as do \textsc{GPz} and \textsc{TPZ}.
-The CDE loss measures how well individual PDFs are estimated, and codes with a low CDE loss tend to have good $\hat{N}(z)$ estimates (though the reverse is not necessarily true).
-\textsc{FlexZBoost} is optimized to minimize CDE loss which may explain why the method has good ensemble metrics as well.
-Note from Table~\ref{tab:cdeloss} that both \textsc{FlexZBoost} and \textsc{CMNN} have low CDE losses.
+This metric is the only one that can appropriately penalize \trainz\ and indicates strong performance for \cmnn\ and \flexzboost, the latter of which is optimized for this metric.
 %Empirically, we have found that PIT RMSE is not as closely correlated to CDE loss as it is to the $N(z)$ statistics.
-As CDE loss is a better measure of individual redshift performance, rather than ensemble distribution performance, this statistic is a better indicator of which codes will be most likely to perform well for science cases where single objects are employed.
 

### 2019-03-15 paper/tex/analysis.tex
moving Scotts Rule and resizing excessively large figures
moving content

@@
 
+The \buzz\ simulation process precludes testing the degree to which samples from our \pz\ posteriors reconstruct the space of $p(z, {\rm data})$.
 To the knowledge of the authors, there is only one metric that can be used to evaluate the performance of individual \pzpdf\ estimators in the absence of true \pz\ posteriors.
@@
 Though we do not endorse the use of the stacked estimator of the redshift distribution, we use it under the untested assumption that the response of our metrics of $\hat{N}^{H}(z)$ will be analogous to the same metrics applied to a principled estimator of the redshift distribution.
+In addition, very few scientific use cases will employ the overall $\hat{N}(z)$ with no cuts, as we explore in this paper.
+We discuss in Section~\ref{sec:futurework} variants of the stacked estimator, such as realistic tomographic binning schemes, that will be explored in future work.
 

### 2019-03-15 paper/tex/results.tex
moving content

@@
 We remind the reader that contributed codes were given a goal of obtaining accurate \pzpdf s, not an accurate stacked estimator of the redshift distribution, so we do not expect the same codes to necessarily perform well for both classes of metrics.
+In previous sections we have mentioned that optimizing for a specific metric does not guarantee good performance on other metrics, nor are our metrics able to assess the fidelity of individual \pzpdf s relative to true posteriors.
 As previously stated, there are a variety of scientific use cases for \pz s, and metric-specific relative performance implies that we may need multiple \pzpdf\ approaches tuned to each metric in order to maximize returns over all science cases in large upcoming surveys.
 
-Samples from accurate photo-$z$ posteriors should reproduce the space of $p(z,{\rm data})$.
-However, it is difficult to test this reconstruction given our data set, as the galaxy distributions arise from mock objects pasted on to an underlying dark matter halo catalogue with properties designed to match empirical relations, rather than being drawn from statistical distributions in redshift.
-In previous sections we have mentioned that optimizing for a specific metric does not guarantee good performance on other metrics, nor is there any guarantee that good performance by our metrics corresponds to \textit{accurate} photo-$z$ posteriors, in the sense of predicting redshifts for individual galaxies.
 %\aim{Explain that samples from accurate photo-$z$ posteriors should reconstruct the space of $p(z, data)$, a test we can't check with our datasets.}

### 2019-03-15 paper/tex/results.tex
finished editing results section

@@
 \hline
-Photo-z Code & fraction PIT$<10^{-4}$ or $>$0.9999\\
+\hline
+\Pz\ Code & fraction PIT$<10^{-4}$ or $>$0.9999\\
 \hline
@@
 \setlength{\tabcolsep}{2pt}
-\caption{Moments of the stacked $\hat{n}(z)$ distribution}\label{tab:moments}
+\caption{Moments of the stacked estimator $\hat{N}(z)$ of the redshift distribution.
+Most of the codes considered recover the moments of $\tilde{N}(z)$}
+\label{tab:moments}
 \begin{tabular}{lccc}
@@
 \hline
- \multicolumn{4}{l}{Stacked $n(z)$ Moments} \\
+ \multicolumn{4}{l}{Moments of $\hat{N}(z)$} \\
 \hline
-              & mean & variance & skewness \\
+Estimator  & mean       & variance    & skewness \\
 Empirical ``truth'' & 0.701 & 0.630 & 0.671  \\
@@
 \tpz        & 0.700      & 0.619      & 0.643    \\
+\hline
 \trainz	    & 0.699 		 & 0.627 	    & 0.666 \\
@@
 \centering
-\caption{CDE loss statistic for each photo-$z$ code.} \label{tab:cdeloss}
+\caption{CDE loss statistic of the individual \pzpdf s for each code.
+A lower value of the CDE loss indicates more accurate individual \pzpdf s, with \cmnn\ and \flexzboost\ performing best under this metric.}
+\label{tab:cdeloss}
 \begin{tabular}{lr}
 \hline
-\bf Photo-$z$ Code & \bf CDE Loss \\
+Photo-$z$ Code & CDE Loss \\
 \hline
@@
 We remind the reader that contributed codes were given a goal of obtaining accurate \pzpdf s, not an accurate stacked estimator of the redshift distribution, so we do not expect the same codes to necessarily perform well for both classes of metrics.
-In previous sections we have mentioned that optimizing for a specific metric does not guarantee good performance on other metrics, nor are our metrics able to assess the fidelity of individual \pzpdf s relative to true posteriors.
-As previously stated, there are a variety of scientific use cases for \pz s, and metric-specific relative performance implies that we may need multiple \pzpdf\ approaches tuned to each metric in order to maximize returns over all science cases in large upcoming surveys.
+Furthermore, our metrics are not necessarily able to assess the fidelity of individual \pzpdf s relative to true posteriors.
+Metric-specific performance implies that we may need multiple \pzpdf\ approaches tuned to each metric in order to maximize returns over all science cases in large upcoming surveys.
 
-%\aim{Explain that samples from accurate photo-$z$ posteriors should reconstruct the space of $p(z, data)$, a test we can't check with our datasets.}
-In other words, we can construct photo-$z$ estimators that provide good coverage in many of our tests, but which have very little predictive power.
-
-The \trainz\ estimator, which assigns every galaxy a $p(z)$ equal to $N(z)$ of the training set as described in Section~\ref{sec:trainz}, is introduced as a ``null test'' to demonstrate this point via \textit{reductio ad absurdum}.
-\trainz\ outperforms all codes on the PIT-based metrics, and all but one code on the $N(z)$ based statistics.
+The \trainz\ estimator, which assigns every galaxy a \pzpdf\ equal to $N(z)$ of the training set as described in Section~\ref{sec:trainz}, is introduced as a control case to demonstrate this point via \textit{reductio ad absurdum}.
 Because our training set is perfectly representative of the test set, $N(z)$ should be identical for both sets down to statistical noise.
+\textit{We make the alarming observation that \trainz\ outperforms all codes on the PIT-based metrics, and all but one code on the $N(z)$ based statistics.}
 %\textit{Explain the connection between \trainz\ and these metrics.}
 
-The CDE loss and point estimate metrics, however, successfully identify problems with \trainz.
-As shown in Appendix~\ref{sec:pointmetrics}, \trainz ~has identical $ZPEAK$ and $ZWEIGHT$ values for every galaxy, and thus the photo-$z$s are constant as a function of spec-$z$s, i.e.
+The CDE loss and point estimate metrics appropriately penalize \trainz's naivete.
+As shown in Appendix~\ref{sec:pointmetrics}, \trainz ~has identical $ZPEAK$ and $ZWEIGHT$ values for every galaxy, and thus the \pz\ point estimats are constant as a function of true redshift, i.e.
 a horizontal line at the mode and mean of the training set distribution respectively.
-The explicit dependence on the {\it individual} posteriors in the calculation of the CDE loss, described in Section~\ref{sec:CDE_loss}, distinguishes this metric from the other $p(z)$ metrics that test the overall ensemble of $p(z)$ distributions.
-With a representative training set, \trainz\ will score well on the ensemble metrics, but fails miserably for metrics tied to individual redshifts presented in this paper.
-We note that many of the ensemble-based metrics are prominent in the photo-$z$ literature despite their inability to identify problems such as those exemplified by \trainz.
-
-%\aim{Explain why CDE loss is different from other metrics (isn't fooled by \trainz) and note that the others have more of a presence in the literature despite their flaw of failing to identify this severe problem.}
+The explicit dependence on the individual posteriors in the calculation of the CDE loss, described in Section~\ref{sec:cdelossresults}, distinguishes this metric from those of the \pzpdf\ ensemble and stacked estimator of the redshift distribution, despite their prevalence in the \pz\ literature.
 
@@
 %[more caution on over-reliance on metrics, and making sure that metrics test all of what you want] \red{this definitely still needs work}
-In summary, context is crucial to interpreting metrics and defending against deceptively well performing methods such as \trainz.
-The best photo-$z$ method is the one that most effectively achieves our science goals, not the one that performs best on a metric that does not accurately reflect those goals.
+In summary, context is crucial to defending against deceptively strong performers such as \trainz; the best \pzpdf\ method is the one that most effectively achieves our science goals, not the one that performs best on a metric that does not reflect those goals.
 In the absence of clear goals or the information necessary for a principled metric definition, we must think carefully before choosing a single metric.
-
-%\begin{table*}  %%% DATA TABLE %%%
-%\caption{KS, CvM and AD statistics for each photo-$z$ code.} \label{tab:results}
-%PIT statistic
-%
-%\begin{tabular}{lrrrr}
-%\hline
-%\bf Photo-$z$ Code & \bf KS & \bf CvM & \bf AD & \bf AD \\
-%                   &        &         & $v=[0.05,0.95]$ & $v=[0.01,0.99]$        \\
-%\hline
-%\textsc{ANNz2} 		& $0.0200$ &   $52.25$ &   $583.0$ &   $759.2$ \\
-%\textsc{BPZ} 		& $0.0388$ &  $280.79$ &  $1361.4$ &  $1557.5$ \\
-%%\textsc{Delight} 	& $0.1380$ & $3040.59$ & $13880.2$ & $17761.4$ \\
-%\textsc{Delight}    & $0.0876$ & $1075.17$ & $4407.8$  & $6167.5$\\
-%%\textsc{EAZY} 		& $0.0690$ &  $994.19$ &  $3479.5$ &  $4158.8$ \\
-%\textsc{EAZY}       & $0.0723$ & $1105.58$ & $3475.0$  & $4418.6$\\
-%%\textsc{FlexZBoost} & $0.0260$ &   $70.58$ &   $595.0$ &   $533.0$ \\
-%\textsc{FlexZBoost} & $0.0240$ &   $68.83$ &   $567.9$ &   $478.8$\\
-%\textsc{GPz} 		& $0.0449$ &  $258.56$ &   $904.8$ &  $1414.8$ \\
-%\textsc{LePhare} 	& $0.0663$ &  $473.05$ &    $85.9$ &   $383.8$ \\
-%\textsc{METAPhoR} 	& $0.0438$ &  $298.56$ &   $248.6$ &   $715.5$ \\
-%%\textsc{CMNN} 		& $-$ & $-$ & $-$ & $-$ \\
-%\textsc{CMNN}         & $0.0795$  &$1011.11$ &  $3908.2$ &  $6307.5$ \\
-%\textsc{SkyNet} 	& $0.0747$ &  $763.00$ &  $2740.2$ &  $4216.4$ \\
-%\textsc{TPZ} 		& $0.1138$ & $1801.74$ &  $9046.2$ & $10565.7$ \\
-%%\textsc{Frankenz}	& $-$ & $-$ & $-$ & $-$ \\
-%\hline
-%\end{tabular}
-
-%Stacked $n(z)$ statistic
-%
-%\begin{tabular}{lrrrr}
-%\hline
-%\bf Photo-$z$ Code & \bf KS & \bf CvM & \bf AD & \bf AD \\
-%                   &        &         & $z=[0.002,2.0]$ & $z=[0.0,2.0]$        \\
-%\hline
-%\textsc{ANNz2} 		& $0.0237$ &  $25.676$ &   $276.4$ &   $276.0$ \\
-%\textsc{BPZ} 		& $0.0131$ &  $17.587$ &   $169.0$ &   $170.9$ \\
-%%\textsc{Delight} 	& $0.0184$ &  $18.975$ &   $151.5$ &   $151.5$ \\
-%\textsc{Delight}    & $0.0256$ &  $42.341$ &   $253.3$ &   $253.2$ \\
-%%\textsc{EAZY} 		& $0.0434$ & $167.329$ &   $811.3$ &   $817.0$ \\
-%\textsc{EAZY}       & $0.0441$ & $169.977$ &   $827.7$ &   $833.3$ \\
-%%\textsc{FlexZBoost} & $0.0065$ &   $2.764$ &    $24.7$ &    $24.6$ \\
-%\textsc{FlexZBoost} & $0.0128$ &   $7.007$ &  $127.02$ &   $126.9$\\
-%\textsc{GPz} 		& $0.0168$ &  $25.605$ &   $307.1$ &   $307.1$ \\
-%\textsc{LePhare} 	& $0.0254$ &  $58.900$ &   $477.4$ &   $472.0$ \\
-%\textsc{METAPhoR} 	& $0.0350$ &  $60.790$ &   $671.9$ &   $672.3$ \\
-%%\textsc{CMNN} 		& $-$ & $-$ & $-$ & $-$ \\
-%\textsc{CMNN}         & $0.0043$ & $0.538$   &   $5.6$   &      $5.8$ \\
-%\textsc{SkyNet} 	& $0.0483$ & $385.830$ &  $2119.4$ &  $2110.3$ \\
-%\textsc{TPZ} 		& $0.0126$ &   $9.483$ &   $110.7$ &   $110.6$ \\
-%%\textsc{Frankenz}	& $-$ & $-$ & $-$ & $-$ \\
-%\hline
-%\end{tabular}
-%\end{table*}
-
-%\begin{table}
-%\setlength{\tabcolsep}{2pt}
-%\begin{center}
-%\caption{Root-Mean-Square-Error (RMSE) statistics for the twelve photo-z codes for both PIT and $\hat{N}(z)$ distributions.}\label{tab:rmse}
-%\begin{tabular}{lcc}
-%\hline
-%\hline
-%Root-Mean-Square-Error \\
-%(RMSE) statistics \\
-%\hline
-%Photo-z Code & PIT RMSE & N(z) RMSE\\
-%\hline
-%\textsc{ANNz2}      & 0.019 & 0.0054\\
-%\textsc{BPZ}        & 0.032 & 0.0050\\
-%\textsc{Delight}    & 0.111 & 0.0056\\
-%\textsc{EAZY}       & 0.054 & 0.0102\\
-%\textsc{FlexZBoost} & 0.021 & 0.0022\\
-%\textsc{GPz}        & 0.027 & 0.0042\\
-%\textsc{LePhare}    & 0.028 & 0.0062\\
-%\textsc{METAPhoR}   & 0.064 & 0.0081\\
-%\textsc{CMNN}         & 0.108 & 0.0009\\
-%\textsc{Skynet}     & 0.054 & 0.0144\\
-%\textsc{TPZ}        & 0.082 & 0.0031\\
-%\hline
-%\trainz		& 0.0025 & 0.0013\\
-%\end{tabular}
-%\end{center}
-%\end{table}
-
-%\begin{table}
-%\begin{center}
-%caption{RMSE Statistics for the 11 Photo-z codes for DC1 $i<25.3$ ``gold'' sample}\label{tab:pitrmse}
-%%\begin{tabular}{|l|c|c|c|c|}
-%\begin{tabular}{lc}
-%\hline
-%\hline
-% PIT RMSE statistics \\
-%\hline
-%Photo-z Code & PIT RMSE\\
-%\hline
-%\texttt{ANNz2}      & 0.019\\
-%\texttt{BPZ}        & 0.032\\
-%\texttt{Delight}    & 0.111\\
-%\texttt{EAZY}       & 0.054\\
-%\texttt{FlexZBoost} & 0.021\\
-%\texttt{GPz}        & 0.048\\
-%texttt{LePhare}    & 0.028\\
-%\texttt{METAPhoR}   & 0.064\\
-%\texttt{NN}         & 0.108 \\
-%\texttt{Skynet}     & 0.054\\
-%texttt{TPZ}        & 0.082\\
-%\hline
-%\end{tabular}
-%\end{center}
-%\end{table}
-%
-%\begin{table}
-%\begin{center}
-%\caption{RMSE Statistics for the 11 Photo-z codes for DC1 $i<25.3$ ``gold'' sample}\label{tab:nzrmse}
-%%\begin{tabular}{|l|c|c|c|c|}
-%\begin{tabular}{lc}
-%\hline
-%\hline
-% N(z) RMSE statistics \\
-%\hline
-%hoto-z Code & N(z) RMSE\\
-%\hline
-%\texttt{ANNz2}      & 0.0054\\
-%\texttt{BPZ}        & 0.0050\\
-%\texttt{Delight}    & 0.0056\\
-%\texttt{EAZY}       & 0.0102\\
-%\texttt{FlexZBoost} & 0.0022\\
-%\texttt{GPz}        & 0.0042\\
-%\texttt{LePhare}    & 0.0062\\
-%\texttt{METAPhoR}   & 0.0081\\
-%\texttt{NN}         & 0.0009\\
-%\texttt{Skynet}     & 0.0144\\
-%\texttt{TPZ}        & 0.0031\\
-%\end{tabular}
-%\end{center}
-%\end{table}

### 2019-03-15 paper/tex/analysis.tex
major rearrangement of content

--- 2019-03-15 paper/tex/analysis.tex
@@
 we seek to minimize the KLD.
-
-\subsubsection{Metrics of a science-motivated quantity}
-\label{sec:moments}
-%(Alex Malz)
-
-Perhaps the most popular application of \pzpdf s is the estimation of the overall redshift distribution $N(z)$, a quantity that enters some cosmological calculations and the true value of which is known for the DC1 data set and will be denoted as $\tilde{N}(z)$.
-In terms of the prior information provided to each method, the true redshift distribution satisfies the tautology $\tilde{N}(z) = p(z \vert I_{D})$ due to our experimental set-up; because the DC1 training and template sets are representative and complete, $I_{D}$ represents a prior that is also equal to the truth.
-In this ideal case of complete and representative prior information, the method that would give the best approximation to $\hat{N}(z)$ would be one that neglects all the information contained in the photometry $\{d_{i}\}_{N_{tot}}$ and gives every galaxy the same \pzpdf\ $\hat{p}_{i}(z) = \tilde{N}(z)$ for all $i$; the inclusion of any information from the photometry would only introduce noise to the optimal result of returning the prior.
-This is the exact estimator, \trainz, that we have described in Section~\ref{sec:trainz}, and which will serve as an experimental control.
-
-Though alternatives exist \citep{Malz:chippr}, ``stacking'' according to
-\begin{equation}
-  \label{eq:stacked}
-  \hat{N}^{H}(z) \equiv \frac{1}{N_{tot}}\ \sum_{i}^{N_{tot}}\ \hat{p}^{H}_{i}(z)
-\end{equation}
-is the most widely accepted method for obtaining $\hat{N}^{H}(z)$ as an estimator of the redshift distribution from \pzpdf s derived by a method $H$.
-Though we do not endorse the use of the stacked estimator of the redshift distribution, we use it under the untested assumption that the response of our metrics of $\hat{N}^{H}(z)$ will be analogous to the same metrics applied to a principled estimator of the redshift distribution.
-
-As $N(z)$ is itself a univariate PDF, we apply the metrics of the previous sections to it as well.
-We additionaly calculate the first three moments
-\begin{equation}
-  \label{eq:moment}
-  \langle z^{m}\rangle \equiv \int_{-\infty}^{\infty} z^{m} N(z) dz
-\end{equation}
-of the estimated redshift distribution $\hat{N}^{H}(z)$ for each code and compare them to the moments of the true redshift distribution $\tilde{N}(z)$.
-Under the assumption that the stacked estimator is unbiased, a superior method minimizes the difference between the true and estimated moments.

### 2019-03-15 paper/tex/appendix.tex
major rearrangement of content

--- 2018-11-16 paper/tex/appendix.tex
@@
 
-\section{Point Estimate Photometric Redshifts}
+\section{Evaluation of the redshift distribution}
+\label{sec:moments}
+%(Alex Malz)
+
+Perhaps the most popular application of \pzpdf s is the estimation of the overall redshift distribution $N(z)$, a quantity that enters some cosmological calculations and the true value of which is known for the DC1 data set and will be denoted as $\tilde{N}(z)$.
+In terms of the prior information provided to each method, the true redshift distribution satisfies the tautology $\tilde{N}(z) = p(z \vert I_{D})$ due to our experimental set-up; because the DC1 training and template sets are representative and complete, $I_{D}$ represents a prior that is also equal to the truth.
+In this ideal case of complete and representative prior information, the method that would give the best approximation to $\hat{N}(z)$ would be one that neglects all the information contained in the photometry $\{d_{i}\}_{N_{tot}}$ and gives every galaxy the same \pzpdf\ $\hat{p}_{i}(z) = \tilde{N}(z)$ for all $i$; the inclusion of any information from the photometry would only introduce noise to the optimal result of returning the prior.
+This is the exact estimator, \trainz, that we have described in Section~\ref{sec:trainz}, and which will serve as an experimental control.
+
+\subsection{Metrics of the stacked estimator of the redshift distribution}
+\label{sec:stackedmetrics}
+
+Though alternatives exist \citep{Malz:chippr}, ``stacking'' according to
+\begin{equation}
+  \label{eq:stacked}
+  \hat{N}^{H}(z) \equiv \frac{1}{N_{tot}}\ \sum_{i}^{N_{tot}}\ \hat{p}^{H}_{i}(z)
+\end{equation}
+is the most widely accepted method for obtaining $\hat{N}^{H}(z)$ as an estimator of the redshift distribution from \pzpdf s derived by a method $H$.
+Though we do not endorse the use of the stacked estimator of the redshift distribution, we use it under the untested assumption that the response of our metrics of $\hat{N}^{H}(z)$ will be analogous to the same metrics applied to a principled estimator of the redshift distribution.
+In addition, very few scientific use cases will employ the overall $\hat{N}(z)$ with no cuts, as we explore in this paper.
+We discuss in Section~\ref{sec:futurework} variants of the stacked estimator, such as realistic tomographic binning schemes, that will be explored in future work.
+
+As $N(z)$ is itself a univariate PDF, we apply the metrics of the previous sections to it as well.
+We additionaly calculate the first three moments
+\begin{equation}
+  \label{eq:moment}
+  \langle z^{m}\rangle \equiv \int_{-\infty}^{\infty} z^{m} N(z) dz
+\end{equation}
+of the estimated redshift distribution $\hat{N}^{H}(z)$ for each code and compare them to the moments of the true redshift distribution $\tilde{N}(z)$.
+Under the assumption that the stacked estimator is unbiased, a superior method minimizes the difference between the true and estimated moments.
+
+\subsection{Performance on the stacked estimator of the redshift distribution}
+\label{sec:stackedmetrics_results}
+
+Figure~\ref{fig:nz} shows the stacked estimator $\hat{N}(z)$ of the redshift distribution for each code compared to the true redshift distribution $\tilde{N}(z)$, where the stacked estimator has been smoothed for each code in the plot using a kernel density estimate (KDE) with a bandwidth chosen by Scott's Rule \citep{Scott:1992} in order to minimize visual differences in small-scale features; the quantiative statistics, however, are calculated using the empirical CDF which is not smoothed.
+
+\begin{figure*}
+\centering
+\includegraphics[width=0.74\textwidth]{fig/NZsumplot_12codes_scottsrule.jpg}
+\caption{The smoothed stacked estimator $\hat{N}(z)$ of the redshift distribution (red) produced by each code (panels) compared to the true redshift distribution $\tilde{N}(z)$ (blue).
+Varying levels of agreement are seen among the codes, with the smallest deviations for \cmnn, \flexzboost, \tpz, and \trainz.}
+\label{fig:nz}
+%\aim{Arguably this could also be clarified by showing the true $n'(z)$ once and then differences from it for each cod's stacked estimator\dots}\scc{what if we plot the difference between true and stacked rather than soverposing the two lines?
+Sam please do not hate me.}
+\end{figure*}
+
+Many of the codes, including all the model-fitting approaches and \annz, \gpz, \metaphor, and \skynet\ from the data-driven camp, overestimate the redshift density at $z \sim 1.4$.
+This behavior is a consequence of the $4000\ {\rm \AA}$ break passing through the gap between the $z$ and $y$ filters, which induces a genuine discontinuity in the $z - y$ colour as a function of redshift that can sway the \pzpdf\ estimates in the absence of bluer spectral features.
+
+\annz, \gpz, and \metaphor\ show signs of overtraining, estimating enhanced peaks and diminished troughs relative to the training set, an obstacle that may be overcome with adjustment of the implementation.
+
+As expected, \trainz\ perfectly recovers the true redshift distribution: as the training sample is selected from the same underlying distribution as the test set, the redshift distributions are identical, up to Poisson fluctuations due to the finite number of sample galaxies.
+\cmnn\ is also in excellent agreement for similar reasons: with a representative training sample of galaxies spanning the colour-space, the sum of the colour-matched neighbour redshifts should return the true redshift distribution.
+\flexzboost\ and \tpz\ also perform superb recovery of the true redshift distribution, with only a slight deviation at $z \sim 1.4$.
+Our metrics, however, cannot discern whether these four approaches, as well as \delight, are spared the $z \sim 1.4$ degeneracy in $\hat{N}(z)$ because they have more effectively used information in the data or if the impact is simply washed out by the stacked estimator's effective average over the test set galaxy sample.
+See Appendix~\ref{sec:pointmetrics} for further discussion of the $z \sim 1.4$ issue.
+
+Figure~\ref{fig:nz_stats} shows the quantitative Kolmogorov-Smirnoff (KS), Cramer-Von Mises (CvM), and Anderson Darling (AD) test statistics for each of the codes for the $\hat{N}(z)$ based measures.
+%\red{Can p-values be supplied for each statistic?
+The statistics themselves are difficult to interpret, other than ``lower is better'' (no, p-values are very difficult to compute for non-uniform distributions)}
+The stacked estimators of the redshift distribution for \cmnn\ and \trainz\ best estimate $\tilde{N}(z)$ under these metrics, and the only codes that do especially poorly are \eazy, \lephare, \metaphor, and \skynet.
+It is unsurprising that \cmnn\ scores well, as with a near perfectly representative training set means that choosing neighbouring points in color/magnitude space should lead to excellent agreement in the final $\hat{N}(z)$ estimate.
+
+\begin{figure*}
+\centering
+\includegraphics[width=0.74\textwidth]{fig/KSvsCvMvsAD_NZ_withnull_jpg.jpg}
+\caption{A visualization of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the $\hat{N}(z)$ distributions.
+We make the reassuring observation that these related statistics do not disagree significantly with one another.
+\cmnn\ performs comparably well to \trainz, the control case, while \skynet\ scores poorly due to an overall bias in its redshift predictions.}
+\label{fig:nz_stats}
+\end{figure*}
+
+It is, however, surprising that \tpz\ does well on $\hat{N}(z)$ given its poor performance on the ensemble \pzpdf s, especially knowing that \tpz\ was optimized for \pzpdf\ ensemble metrics rather than the stacked estimator of the redshift distribution.
+A possible explanation is the choice of smoothing parameter chosen during validation, which affects \pzpdf\ widths as well as overall redshift bias and could be modified to improve performance under the \pzpdf\ metrics.
+
+The first three moments of the stacked $\hat{N}(z)$ distribution relative to the empirical estimate of the truth distribution are given in Table~\ref{tab:moments}.
+Accuracy of the moments varies widely between codes, raising concerns about the propagation to cosmological analyses.
+%\red{[mention that we are calculating moments for the entire sample, will look at fiducial tomographic bins in a follow-up paper (unless group and reviewers realy think we should include in this paper.]}
+%\red{FlexZBoost has some of the best $n(z)$ statistics, but the moments are not good.
+Why is this?
+Is it the over/underprediction of the high redshift part of the distribution?
+Discuss this after talking with Rafael.}
+
+\begin{table}
+\setlength{\tabcolsep}{2pt}
+\caption{Moments of the stacked estimator $\hat{N}(z)$ of the redshift distribution.
+Most of the codes considered recover the moments of $\tilde{N}(z)$}
+\label{tab:moments}
+\begin{tabular}{lccc}
+\hline
+\hline
+ \multicolumn{4}{l}{Moments of $\hat{N}(z)$} \\
+\hline
+Estimator  & mean       & variance    & skewness \\
+Empirical ``truth'' & 0.701 & 0.630 & 0.671  \\
+\hline
+\annz       & 0.702      & 0.625      & 0.653    \\
+\bpz        & 0.699      & 0.629      & 0.671    \\
+\delight    & 0.692      & 0.609      & 0.638    \\
+\eazy       & 0.681      & 0.595      & 0.619    \\
+\flexzboost & 0.694      & 0.610      & 0.631    \\
+\gpz        & 0.696      & 0.615      & 0.639    \\
+\lephare    & 0.718      & 0.668      & 0.741    \\
+\metaphor   & 0.705      & 0.628      & 0.657    \\
+\cmnn       & 0.701      & 0.628      & 0.667    \\
+\skynet     & 0.743      & 0.708      & 0.797    \\
+\tpz        & 0.700      & 0.619      & 0.643    \\
+\hline
+\trainz	    & 0.699 		 & 0.627 	    & 0.666 \\
+\end{tabular}
+\end{table}
+
+\skynet\ exhibits redshift bias in Figure~\ref{fig:nz} and is a clear outlier in the first moment of $\hat{N}(z)$ in Table~\ref{tab:moments}.
+The \skynet\ algorithm employs a random subsampling of the training set without testing that the subset is representative of the full population, and the implementation used here does not upweight rarer low- and high-redshift galaxies, as in \citet{Bonnett:15}, suggesting a possible cause that may be addressed in future work.
+
+\section{\Pz\ point estimation and metrics}
 \label{sec:pointmetrics}
+
 While we assume that all science analysis will use full PDF information and do not recommend the use of single point estimates of redshift for most science applications, we include a brief evaulation as an Appendix.
@@
 
-There are several common point estimators of photo-$z$ posteriors employed by different codes, e.~g.
-the mode, mean, median of the $p(z)$ distribution.
+\subsection{Reduction of \pzpdf s to point estimates}
+\label{sec:pointest}
+
+There are several common point estimators of photo-$z$ posteriors employed by different codes, e.~g. the mode, mean, median of the $p(z)$ distribution.
 In addition, many of the machine learning based estimators can be set up to return a single redshift solution.
-For example, SkyNet can be configured to run as a regressor that returns a single float rather than a classifier that returns a 200-bin $p(z)$ estimate.
+For example, \skynet\ can be configured to run as a regressor that returns a single float rather than a classifier that returns a 200-bin $p(z)$ estimate.
 The single value returned by a machine learning based code may not correspond to a particular measure such as the mode or mean, and so to avoid interpretation of results that might be introduced by variations in choice of specific point-estimate implementation per code, we discard the code-specific point estimates.
@@
 Following \cite{Tanaka:17}, we report the values of the following metrics using whichever of these point estimators performs best for each method, noting which point reduction is used in each case.
+%
+\subsection{Metrics of \pz\ point estimates}
+\label{sec:point_metrics}
 
-\subsection{Point Estimate Metrics}
-\label{sec:point_metrics}
 We calculate the commonly used point estimate metrics of the overall photo-$z$ scatter ($\sigma_{z}$, the standard deviation of the photo-$z$ residuals), bias, and ``catastrophic outlier rate''.
@@
 we define $e_{z}$ as
-
 \begin{equation}
@@
 
-Fig.~\ref{fig:pz_pointestimates}  shows the point estimates for both $z_{PEAK}$ and $z_{WEIGHT}$.
+\subsection{Comparison of \pz\ point estimate metrics}
+\label{sec:pointmetrics_results}
+
+\begin{figure*}
+\centering
+\includegraphics[width=0.49\textwidth]{fig/ZPEAK_szpz_threecolumn_12codes_navy_lowalpha.jpg}
+\includegraphics[width=0.49\textwidth]{fig/ZWEIGHT_szpz_threecolumn_12codes_navy_lowalpha.jpg}
+\caption{\Pz\ point estimates reduced from the \pzpdf s, via the mode ($z_{PEAK}$, left panel) and mean ($z_{WEIGHT}$, right panel).
+Point estimate density is represented with fixed density contours, while outliers at lower density are represented by blue points.
+While use of point-estimate photo-$z$'s is not recommended, they do make for useful comparative and visual diagnostics.
+In the lower-right panel of each plot, the \trainz\ estimator results in identical photo-$z$ estimates at the mode and mean of the training set $N'(z)$ distribution for all galaxies.}
+\label{fig:pz_pointestimates}
+\end{figure*}
+
+Figure~\ref{fig:pz_pointestimates} shows the point estimates for both $z_{PEAK}$ and $z_{WEIGHT}$.
 Point density is shown with mixed contours to emphasize that most of the galaxies do fall close to the $z_{phot}=z_{spec}$ line, while blue points show differing characteristics of the outlier populations.
 The red dashed lines indicated the cutoff for catastrophic outliers, defined as: $max(0.06,3\sigma_{IQR})$.
-As with the full $p(z)$ results, a variety of behaviours are evident in the different codes.
-Table~\ref{tab:pointestimates} lists the scatter, bias, and catastrophic outlier fractions for the codes.
+As with the full $p(z)$ results, a variety of behaviours are evident in the different codes.  Table~\ref{tab:pointestimates} lists the scatter, bias, and catastrophic outlier fractions for the codes.
 The performance of the codes for point metrics is highly correlated with performance on $p(z)$ based tests, which is to be expected, given that the point-estimates were derived from the $p(z)$.
@@
 
-Finally, we note that all twelve codes perform at or near the goals for point-estimates as outlined in the LSST Science Requirements Document\footnote{available at: \url{http://ls.st/srd}} and \citet{Graham:17}.
-This is to be expected, given that the requirements were designed such that a point estimate photo-z would meet these requirements for perfect training data to a depth of $i<25$.
-But, it is still an encouraging sign, given an updated mock galaxy simulation and the expanded set of photo-$z$ codes tested.
-%\red{maybe say consistent with expectations, and a reference to Melissa's paper in here?  What else to say about nearly meeting point goals?}
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.8\textwidth]{fig/ZPEAK_szpz_threecolumn_12codes_navy.jpg}\\
-\includegraphics[width=0.8\textwidth]{fig/ZWEIGHT_szpz_threecolumn_12codes_navy.jpg}
-\caption{Point estimate photo-z's derived from the posteriors.
-Top panel shows $z_{PEAK}$, while bottom panel shows $z_{WEIGHT}$.
-Point estimate density is represented with fixed density contours, while outliers at lower density are represented by blue points.
-While use of point-estimate photo-$z$'s is not recommended, they do make for useful comparative and visual diagnostics.
-In the lower-right panel of each plot, the \trainz\ estimator results in identical photo-$z$ estimates at the mode and mean of the training set $N'(z)$ distribution for all galaxies.} \label{fig:pz_pointestimates}
-\end{figure*}
-
 \begin{table*}
 \begin{center}
-\caption{Point estimate statistics}\label{tab:pointestimates}
+\caption{\Pz\ point estimate statistics}\label{tab:pointestimates}
 \begin{tabular}{lcccccc}
@@
 \hline
-Photo-z Code       & $\frac{\sigma_{IQR}}{(1+z)}$ & median  & \multicolumn{1}{|p{0.75cm}|}{\centering outlier \\fraction} & $\frac{\sigma_{IQR}}{(1+z)}$ & median & \multicolumn{1}{|p{0.75cm}|}{\centering outlier \\fraction}\\
+\Pzpdf\ Code       & $\frac{\sigma_{IQR}}{(1+z)}$ & median  & \multicolumn{1}{|p{0.75cm}|}{\centering outlier \\fraction} & $\frac{\sigma_{IQR}}{(1+z)}$ & median & \multicolumn{1}{|p{0.75cm}|}{\centering outlier \\fraction}\\
 \hline
-\textsc{ANNz2}     & 0.0270  &  0.00063  & 0.044      & 0.0244  &  0.000307  & 0.047  \\
-\textsc{BPZ}       & 0.0215  & -0.00175  & 0.035      & 0.0215  & -0.002005  & 0.032 \\
-\textsc{Delight}   & 0.0212  & -0.00185  & 0.038      & 0.0216  & -0.002158  & 0.038 \\
-\textsc{EAZY}      & 0.0225  & -0.00218  & 0.034      & 0.0226  & -0.003765  & 0.029 \\
-\textsc{FlexZBoost}& 0.0154  & -0.00027  & 0.020      & 0.0148  & -0.000211  & 0.017 \\
-\textsc{GPz}       & 0.0197  & -0.00000  & 0.052      & 0.0195  &  0.000113  & 0.051 \\
-\textsc{LePhare}   & 0.0236  & -0.00161  & 0.058      & 0.0239  & -0.002007  & 0.056 \\
-\textsc{METAPhoR}  & 0.0264  &  0.00000  & 0.037      & 0.0262  &  0.001333  & 0.048 \\
-\textsc{CMNN}        & 0.0184  & -0.00132  & 0.035      & 0.0170  & -0.001049  & 0.034 \\
-\textsc{Skynet}    & 0.0219  & -0.00167  & 0.036      & 0.0218  &  0.000174  & 0.037 \\
-\textsc{TPZ}       & 0.0161  &  0.00309  & 0.033      & 0.0166  &  0.003048  & 0.031 \\
+\annz     & 0.0270  &  0.00063  & 0.044      & 0.0244  &  0.000307  & 0.047  \\
+\bpz       & 0.0215  & -0.00175  & 0.035      & 0.0215  & -0.002005  & 0.032 \\
+\delight   & 0.0212  & -0.00185  & 0.038      & 0.0216  & -0.002158  & 0.038 \\
+\eazy      & 0.0225  & -0.00218  & 0.034      & 0.0226  & -0.003765  & 0.029 \\
+\flexzboost& 0.0154  & -0.00027  & 0.020      & 0.0148  & -0.000211  & 0.017 \\
+\gpz       & 0.0197  & -0.00000  & 0.052      & 0.0195  &  0.000113  & 0.051 \\
+\lephare   & 0.0236  & -0.00161  & 0.058      & 0.0239  & -0.002007  & 0.056 \\
+\metaphor  & 0.0264  &  0.00000  & 0.037      & 0.0262  &  0.001333  & 0.048 \\
+\cmnn        & 0.0184  & -0.00132  & 0.035      & 0.0170  & -0.001049  & 0.034 \\
+\skynet    & 0.0219  & -0.00167  & 0.036      & 0.0218  &  0.000174  & 0.037 \\
+\tpz       & 0.0161  &  0.00309  & 0.033      & 0.0166  &  0.003048  & 0.031 \\
 \hline
@@
 \end{table*}
+
+Finally, we note that all twelve codes perform at or near the goals for point-estimates as outlined in the LSST Science Requirements Document\footnote{available at: \url{http://ls.st/srd}} and \citet{Graham:17}.
+This is to be expected, given that the requirements were designed such that a point estimate photo-z would meet these requirements for perfect training data to a depth of $i<25$.
+But, it is still an encouraging sign, given an updated mock galaxy simulation and the expanded set of photo-$z$ codes tested.
+%\red{maybe say consistent with expectations, and a reference to Melissa's paper in here?  What else to say about nearly meeting point goals?}

### 2019-03-15 paper/tex/conclusion.tex
major rearrangement of content

@@
 
-%\item \red{inclusion of incompleteness effects on training sample, ``wrong redshifts'', filter curves}
-%\item \red{inclusion of photometric errors, image based effects, blending, lensing, sky, mask, etc...}
+This paper compares the performance of twelve \pzpdf\ approaches under idealized experimental conditions of representative and complete prior information to set a baseline for an upcoming sensitivity analysis by first isolating the impact of the \pzpdf\ estimation technique from realistic physical systematics of the data.
+Though the mock data set of this experiment did not include true \pz\ posteriors for comparison, \textbf{we interpret deviations from perfect results given perfect prior information as the imprint of the implicit assumptions underlying the estimation approach}.
 
-In this study we have not accounted for the presence of Active Galactic Nuclei (AGN) contributions to galaxy fluxes.
-In some cases, AGN will be easily identified from the colors and morphologies, i.e.
-the case of the brightest quasars where the nuclear activity outshines the host galaxy, and numerous studies have utilized color selection to create large samples of quasars \citep[e.g.][]{Richards:06,Maddox:08,Richards:15}.
-In current deep fields, similar in depth to what we expect from LSST, variability information and multi-wavelength data have been critical to not only identify AGN dominated galaxies, but also obtain more accurate photometric redshifts \citep[e.g][]{Salvato:11}.
-
-In addition to AGN dominated galaxies, those with lower levels of nuclear activity present a more insidious problem, where AGN features may not be apparent, but the colors and other host galaxy properties are perturbed relative to galaxies with an inactive nucleus.
-In such cases, the presence of the AGN may induce a bias if the template SEDs or empirical datasets do not include low-level AGN counterparts.
-For LSST, we will need to identify and obtain accurate photometric redshifts of all types of AGN for a range of science goals, whether it is to eliminate such objects from cosmology experiments, or to use them with confidence, all the way through to understanding galaxy evolution and the role that AGN may play in influencing galaxy properties over cosmic time.
-
-A promising route to classifying and obtaining accurate photometric redshifts for the AGN population is by combining machine learning with template-fitting techniques, as has recently been demonstrated by \citet{Duncan:18} for radio-selected AGN.
-This is because AGN are relatively easy to obtain spectroscopic redshifts for over all redshifts due to the strong emission lines that they exhibit, allowing very good training sets for machine learning algorithms to use.
-Whereas for those galaxies where the AGN is sub-dominant the galaxy templates are still adequate for obtaining reasonable photometric redshifts.
-
-In addition to these improvements, the DESC Photo-$z$ group plans to look at all potential methods to combine the results from multiple photo-$z$ codes to improve $p(z)$ accuracy, similar to the work presented in \citet{Dahlen:13,Carrascokind:14,Duncan:18}.
-Taking advantage of multiple algorithms that use observables in slightly different ways has shown promise, however we must be very conscious of whether a potential combination properly treats the covariance between the methods, given that they are estimating quantities based on the same underlying observables.
-Several science cases wish to estimate physical quantities along with redshift, for example galaxy stellar mass and star formation rate.
-Proper joint estimation of redshift and physical quantities requires an in depth understanding of galaxy evolution, and progress on accurate bivariate redshift probability distributions will go hand in had with progress on understanding galaxies themselves.
-Parameterization and storage of a complex 2-dimensional probability surface for potentially billions of galaxies (or even subsets of hundreds of thousand of particular interest) pose a potential challenge.
-These issues will be examined in another future paper.
-
-Finally, while this paper and future papers discussed above focus on photometric redshift codes and estimating accurate $p(z)$ from training data, we plan a separate, but complementary, project to examine calibration of the resultant redshifts via spatial cross-correlations \citep{Newman:2008}, which will be explored in a separate series of future papers.
-The overarching plan describing everything laid out in this section is described in more detail in the LSST DESC Science Roadmap (see Footnote in Section~\ref{sec:intro}).
-These plans will require significant effort, but they are necessary if we are to make optimal use of the LSST data for astrophysical and cosmological analyses.
-%\begin{enumerate}
-%\item \red{the combination of codes}
-
-%\item \red{$p(z,\alpha)$}
-%\item \red{tomographic analysis later on}
-%\red{this paper deals with training, future papers will also explore calibration via cross-correlation methods}
-%\end{enumerate}
-%\red{mention the SRM once again, say that the plans are extensive, but we do have a plan and a rough timeline.}
+We evaluated the codes under science-agnostic metrics both established and emerging to stress-test the ensemble properties of \pzpdf\ catalogues derived by each method, as well as metrics of a prevalent summary statistic of \pzpdf\ catalogues used in cosmological analyses.
+We observe that no one code dominates in all metrics, and that the standard metrics of \pzpdf s and the stacked estimator of the redshift distribution can be gamed by a trivial approach to \pzpdf\ estimation.
+We emphasize to the \pz\ community that \textbf{metrics used to vet \pzpdf\ methods must be tailored to science cases with self-consistent analysis approaches that respect the inherently probabilistic nature of \pzpdf\ catalogs}.

### 2019-03-15 paper/tex/discussion.tex
moving all the things around
major rearrangement of content

@@
-\section{Discussion and Future Work}
+\section{Discussion and future work}
 \label{sec:discussion}
 %(Sam Schmidt, Ofer Lahav, Jeff Newman, Alex Malz, Eve Kovacs, Tony Tyson, Tina Peters)
-In this paper we present results evaluating the computation of individual galaxy photometric redshift PDFs for twelve photo-$z$ codes.
-As discussed in Section~\ref{sec:metrics} each $p(z)$ should accurately reflect the relative likelihood as a function of redshift for each galaxy in an informative way; that is, the estimates should provide useful information per individual galaxy, not just the ensemble.
-All codes were provided a set of representative training data and tested on an idealized set of model galaxies with high signal-to-noise and photometry with no confounding effects due to blending, instrumental effects, the night sky, or other complications included.
-%The goal was not to determine a ``best'' photo-$z$ code:i
-In many ways, this was a baseline test of a ``best case scenario'' to predict the expected photo-$z$ performance if a stage IV dark energy survey was to obtain complete training samples and perfectly calibrated their multi-band photometry.
-Given these idealized conditions, any deficiencies observed in a photo-$z$ code's performance should be a cause for concern, and may be evidence of a problem with either/both of the specific code implementation or the underlying algorithm.
-Many of the codes tested performed well; however, several did not meet the stringent goals that have been laid out for LSST photometric redshift performance for individual galaxies, as laid out in the LSST SRD (See Section~\ref{sec:intro}).%  This is a cause for concern, and the individual code responses will be studied in detail moving forward.
-If methods can not reach the goals on idealized data, then they will almost surely not meet those same goals when the more complex problems that we expect to arise from real LSST data are included.
-The results presented in this paper enable an evaluation of which algorithms are the most promising moving forward, and potentially point to implementation choices or mistakes which could be improved or corrected in others.
-Codes that do not perform well may not be worth pursuing in future challenges.
 
-One obvious trend in several of the codes tested was an overall over or underprediction of the widths of $p(z)$, as evidenced by the QQ plots and PIT histograms shown in Fig.~\ref{fig:pitqq}.
+%The goal was not to determine a ``best'' photo-$z$ code.
+
+% If methods can not reach the goals on idealized data, then they will almost surely not meet those same goals when the more complex problems that we expect to arise from real LSST data are included.
+% The results presented in this paper enable an evaluation of which algorithms are the most promising moving forward, and potentially point to avenues for improvement.
+
+In contrast with other \pzpdf\ comparison papers that have aimed to identify the ``best'' code for a given survey, we have focused on the somewhat more philosophical questions of how to assess \pzpdf\ methods and how to interpret differences between codes in terms of \pzpdf\ performance.
+
+\subsection{Interpretation of metrics}
+\label{sec:caution}
+%Caution on metrics and photo-$z$ codes}\label{sec:caution}
+%(Alex Malz, Sam Schmidt)
+
+We remind the reader that contributed codes were given a goal of obtaining accurate \pzpdf s, not an accurate stacked estimator of the redshift distribution, so we do not expect the same codes to necessarily perform well for both classes of metrics.
+Furthermore, our metrics are not necessarily able to assess the fidelity of individual \pzpdf s relative to true posteriors.
+Metric-specific performance implies that we may need multiple \pzpdf\ approaches tuned to each metric in order to maximize returns over all science cases in large upcoming surveys.
+
+The \trainz\ estimator, which assigns every galaxy a \pzpdf\ equal to $N(z)$ of the training set as described in Section~\ref{sec:trainz}, is introduced as a control case to demonstrate this point via \textit{reductio ad absurdum}.
+Because our training set is perfectly representative of the test set, $N(z)$ should be identical for both sets down to statistical noise.
+\textit{We make the alarming observation that \trainz\ outperforms all codes on the PIT-based metrics, and all but one code on the $N(z)$ based statistics.}
+%\textit{Explain the connection between \trainz\ and these metrics.}
+
+The CDE loss and point estimate metrics appropriately penalize \trainz's naivete.
+As shown in Appendix~\ref{sec:pointmetrics}, \trainz ~has identical $ZPEAK$ and $ZWEIGHT$ values for every galaxy, and thus the \pz\ point estimats are constant as a function of true redshift, i.e.
+a horizontal line at the mode and mean of the training set distribution respectively.
+The explicit dependence on the individual posteriors in the calculation of the CDE loss, described in Section~\ref{sec:cdelossresults}, distinguishes this metric from those of the \pzpdf\ ensemble and stacked estimator of the redshift distribution, despite their prevalence in the \pz\ literature.
+
+% While looking at one, or even most of our metrics would have given the impression that this estimator was nearly optimal, red flags in CDE loss and point estimates reveal the problems.
+%[more caution on over-reliance on metrics, and making sure that metrics test all of what you want] \red{this definitely still needs work}
+In summary, context is crucial to defending against deceptively strong performers such as \trainz; the best \pzpdf\ method is the one that most effectively achieves our science goals, not the one that performs best on a metric that does not reflect those goals.
+In the absence of clear goals or the information necessary for a principled metric definition, we must think carefully before choosing a single metric.
+
+Trends in the metrics shared among codes indicate the
+
+One obvious trend in several of the codes tested was an overall over or underprediction of the widths of \pzpdf s, as evidenced by the QQ plots and PIT histograms shown in Figure~\ref{fig:pitqq}.
 A more careful tuning of bandwidth or smoothing during the validation process appears to be necessary for many of the machine learning based codes in order to improve the accuracy of $p(z)$.
@@
 %\red{some discussion of output parameterization, limitations of 200 point fixed grid, in future switch to something like quantiles (cite qp paper)}
-We re-emphasize that the dataset tested was quite idealized, and discuss enhancements that will be added in future simulations to test photo-$z$ codes on increasingly realistic conditions in the following section.
 
-\subsection{Future work}
-\label{sec:futurework}
+\subsubsection{Nontrivial experimental design}
+\label{sec:futureexperiments}
+
+%\begin{enumerate}
+%\item \red{the combination of codes}
+
+%\item \red{$p(z,\alpha)$}
+%\item \red{tomographic analysis later on}
+%\red{this paper deals with training, future papers will also explore calibration via cross-correlation methods}
+%\end{enumerate}
+%\red{mention the SRM once again, say that the plans are extensive, but we do have a plan and a rough timeline.}
 
@@
 
-In future papers a focus of the {\it LSST Dark Energy Science Collaboration Photo-z Working Group} will be to add more and more complexity to our simulated data in order to test photo-$z$ algorithms in increasingly realistic conditions.
+We re-emphasize that the dataset tested was quite idealized, and discuss enhancements that will be added in future simulations to test photo-$z$ codes on increasingly realistic conditions in the following section.
+
+\subsubsection{More realistic mock data}
+\label{sec:futuredata}
+
+%\item \red{inclusion of incompleteness effects on training sample, ``wrong redshifts'', filter curves}
+%\item \red{inclusion of photometric errors, image based effects, blending, lensing, sky, mask, etc...}
+
+In future papers a focus of the \lsstdesc\ \Pz\ Working Group will be to add more and more complexity to our simulated data in order to test photo-$z$ algorithms in increasingly realistic conditions.
 The most pressing concern is the impact of incomplete spectroscopic training samples.
@@
 In addition to outright redshift failures we will model the inclusion of a small number of falsely identified secure redshifts where misidentified emission lines or noise spikes cause an incorrect redshift solution to be marked as a high quality identification.
-Even sub-per cent level contamination by false redshifts can impact photo-$z$ solutions at levels comparable to the stringent  requirements of some LSST science cases.
-We expect different systematics to occur in different photo-$z$ codes in response to training on incomplete data, particularly some of the machine learning methods.
+Even sub-per cent level contamination by false redshifts can impact \pz\ solutions at levels comparable to the stringent  requirements of some LSST science cases.
+We expect different systematics to occur in different \pzpdf\ codes in response to training on incomplete data, particularly some of the machine learning methods.
 The response of the codes will inform future directions of code development.
@@
 More importantly, one of the most common ``catastrophic outlier'' degeneracies observed in deep photometric samples occurs when the Lyman break is mistaken for the Balmer break, leading to multiple redshift solutions at $z\sim0.2-0.3$ and $z\sim2-3$ \citep{Massarotti:2001}.
-This degeneracy, along with other potential degeneracies, are currently not covered by the limited redshift range of this initial paper, which could mean that we are not probing the full range of potential extreme outlier populations and how our photo-$z$ estimators respond to them.
+This degeneracy, along with other potential degeneracies, are currently not covered by the limited redshift range of this initial paper, which could mean that we are not probing the full range of potential extreme outlier populations and how our \pz\ estimators respond to them.
 Extending simulations to include the high-redshift galaxy population will be a priority in future data challenges.
@@
 This initial paper explored a data set that was constructed at the catalog level, with no inclusion of the complications that come from measuring photometry from images.
-Future data challenges will move to catalogs constructed from mock images, including effects that will have great impact on photo-$z$ measurements, which will naturally include the complications of object blending, sensor effects, different observing conditions, amongst others.
+Future data challenges will move to catalogs constructed from mock images, including effects that will have great impact on \pz\ estimates, which will naturally include the complications of object blending, sensor effects, different observing conditions, amongst others.
 Object blending will be a major area of investigation, as the mixing of flux from multiple objects and the resultant change in measured colours is predicted to affect a large fraction of LSST galaxies \citep{Dawson:2016}, and will be one of the major contributing systematics for photo-$z$'s.
 
-Finally, while this paper and future papers discussed above focus on photometric redshift codes and estimating accurate $p(z)$ from training data, we plan a separate, but complementary, project to examine calibration of the resultant redshifts via spatial cross-correlations \citep{Newman:2008}, which will be explored in a separate set of papers.
-The overarching plan describing everything laid out in this section is described in more detail in the LSST DESC Science Roadmap (see Footnote in Section~\ref{sec:intro}).
+In this study we have not accounted for the presence of Active Galactic Nuclei (AGN) contributions to galaxy fluxes.
+In some cases, AGN will be easily identified from the colors and morphologies, i.e.
+the case of the brightest quasars where the nuclear activity outshines the host galaxy, and numerous studies have utilized color selection to create large samples of quasars \citep[e.g.][]{Richards:06,Maddox:08,Richards:15}.
+In current deep fields, similar in depth to what we expect from \lsst, variability information and multi-wavelength data have been critical to not only identify AGN dominated galaxies, but also obtain more accurate photometric redshifts \citep[e.g][]{Salvato:11}.
+
+In addition to AGN dominated galaxies, those with lower levels of nuclear activity present a more insidious problem, where AGN features may not be apparent, but the colours and other host galaxy properties are perturbed relative to galaxies with an inactive nucleus.
+In such cases, the presence of the AGN may induce a bias if the template SEDs or empirical datasets do not include low-level AGN counterparts.
+For LSST, we will need to identify and obtain accurate photometric redshifts of all types of AGN for a range of science goals, whether it is to eliminate such objects from cosmology experiments, or to use them with confidence, all the way through to understanding galaxy evolution and the role that AGN may play in influencing galaxy properties over cosmic time.
+
+A promising route to classifying and obtaining accurate photometric redshifts for the AGN population is by combining machine learning with template-fitting techniques, as has recently been demonstrated by \citet{Duncan:18} for radio-selected AGN.
+This is because AGN are relatively easy to obtain spectroscopic redshifts for over all redshifts due to the strong emission lines that they exhibit, allowing very good training sets for machine learning algorithms to use.
+Whereas for those galaxies where the AGN is sub-dominant the galaxy templates are still adequate for obtaining reasonable photometric redshifts.
+
+In addition to these improvements, the \lsstdesc \Pz\ Working Group plans to look at all potential methods to combine the results from multiple \pzpdf\ codes to improve accuracy, similar to the work presented in \citet{Dahlen:13,Carrascokind:14,Duncan:18}.
+Taking advantage of multiple algorithms that use observables in slightly different ways has shown promise, however we must be very conscious of whether a potential combination properly treats the covariance between the methods, given that they are estimating quantities based on the same underlying observables.
+Several science cases wish to estimate physical quantities along with redshift, for example galaxy stellar mass and star formation rate.
+Proper joint estimation of redshift and physical quantities requires an in depth understanding of galaxy evolution, and progress on accurate bivariate redshift probability distributions will go hand in had with progress on understanding galaxies themselves.
+Parameterization and storage of a complex 2-dimensional probability surface for potentially billions of galaxies (or even subsets of hundreds of thousand of particular interest) pose a potential challenge.
+These issues will be examined in another future paper.
+
+Finally, while this paper and future papers discussed above focus on photometric redshift codes and estimating accurate \pzpdf s from training data, we plan a separate, but complementary, project to examine calibration of the resultant redshifts via spatial cross-correlations \citep{Newman:2008}, which will be explored in a separate series of future papers.
+The overarching plan describing everything laid out in this section is described in more detail in the \lsstdesc\ Science Roadmap (see Footnote in Section~\ref{sec:intro}).
 These plans will require significant effort, but they are necessary if we are to make optimal use of the LSST data for astrophysical and cosmological analyses.

### 2019-03-15 paper/tex/results.tex
major rearrangement of content

@@
 
-\subsection{Performance on the stacked estimator of the redshift distribution}
-\label{sec:stackedmetrics}
-
-Figure~\ref{fig:nz} shows the stacked estimator $\hat{N}(z)$ of the redshift distribution for each code compared to the true redshift distribution $\tilde{N}(z)$, where the stacked estimator has been smoothed for each code in the plot using a kernel density estimate (KDE) with a bandwidth chosen by Scott's Rule \citep{Scott:1992} in order to minimize visual differences in small-scale features; the quantiative statistics, however, are calculated using the empirical CDF which is not smoothed.
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.74\textwidth]{fig/NZsumplot_12codes_scottsrule.jpg}
-\caption{The smoothed stacked estimator $\hat{N}(z)$ of the redshift distribution (red) produced by each code (panels) compared to the true redshift distribution $\tilde{N}(z)$ (blue).
-Varying levels of agreement are seen among the codes, with the smallest deviations for \cmnn, \flexzboost, \tpz, and \trainz.}
-\label{fig:nz}
-%\aim{Arguably this could also be clarified by showing the true $n'(z)$ once and then differences from it for each cod's stacked estimator\dots}\scc{what if we plot the difference between true and stacked rather than soverposing the two lines?
-Sam please do not hate me.}
-\end{figure*}
-
-Many of the codes, including all the model-fitting approaches and \annz, \gpz, \metaphor, and \skynet\ from the data-driven camp, overestimate the redshift density at $z \sim 1.4$.
-This behavior is a consequence of the $4000\ {\rm \AA}$ break passing through the gap between the $z$ and $y$ filters, which induces a genuine discontinuity in the $z - y$ colour as a function of redshift that can sway the \pzpdf\ estimates in the absence of bluer spectral features.
-
-\annz, \gpz, and \metaphor\ show signs of overtraining, estimating enhanced peaks and diminished troughs relative to the training set, an obstacle that may be overcome with adjustment of the implementation.
-
-As expected, \trainz\ perfectly recovers the true redshift distribution: as the training sample is selected from the same underlying distribution as the test set, the redshift distributions are identical, up to Poisson fluctuations due to the finite number of sample galaxies.
-\cmnn\ is also in excellent agreement for similar reasons: with a representative training sample of galaxies spanning the colour-space, the sum of the colour-matched neighbour redshifts should return the true redshift distribution.
-\flexzboost\ and \tpz\ also perform superb recovery of the true redshift distribution, with only a slight deviation at $z \sim 1.4$.
-Our metrics, however, cannot discern whether these four approaches, as well as \delight, are spared the $z \sim 1.4$ degeneracy in $\hat{N}(z)$ because they have more effectively used information in the data or if the impact is simply washed out by the stacked estimator's effective average over the test set galaxy sample.
-See Appendix~\ref{sec:pointmetrics} for further discussion of the $z \sim 1.4$ issue.
-
-Figure~\ref{fig:nz_stats} shows the quantitative Kolmogorov-Smirnoff (KS), Cramer-Von Mises (CvM), and Anderson Darling (AD) test statistics for each of the codes for the $\hat{N}(z)$ based measures.
-%\red{Can p-values be supplied for each statistic?
-The statistics themselves are difficult to interpret, other than ``lower is better'' (no, p-values are very difficult to compute for non-uniform distributions)}
-The stacked estimators of the redshift distribution for \cmnn\ and \trainz\ best estimate $\tilde{N}(z)$ under these metrics, and the only codes that do especially poorly are \eazy, \lephare, \metaphor, and \skynet.
-It is unsurprising that \cmnn\ scores well, as with a near perfectly representative training set means that choosing neighbouring points in color/magnitude space should lead to excellent agreement in the final $\hat{N}(z)$ estimate.
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.74\textwidth]{fig/KSvsCvMvsAD_NZ_withnull_jpg.jpg}
-\caption{A visualization of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the $\hat{N}(z)$ distributions.
-We make the reassuring observation that these related statistics do not disagree significantly with one another.
-\cmnn\ performs comparably well to \trainz, the control case, while \skynet\ scores poorly due to an overall bias in its redshift predictions.}
-\label{fig:nz_stats}
-\end{figure*}
-
-It is, however, surprising that \tpz\ does well on $\hat{N}(z)$ given its poor performance on the ensemble \pzpdf s, especially knowing that \tpz\ was optimized for \pzpdf\ ensemble metrics rather than the stacked estimator of the redshift distribution.
-A possible explanation is the choice of smoothing parameter chosen during validation, which affects \pzpdf\ widths as well as overall redshift bias and could be modified to improve performance under the \pzpdf\ metrics.
-
-The first three moments of the stacked $\hat{N}(z)$ distribution relative to the empirical estimate of the truth distribution are given in Table~\ref{tab:moments}.
-Accuracy of the moments varies widely between codes, raising concerns about the propagation to cosmological analyses.
-%\red{[mention that we are calculating moments for the entire sample, will look at fiducial tomographic bins in a follow-up paper (unless group and reviewers realy think we should include in this paper.]}
-%\red{FlexZBoost has some of the best $n(z)$ statistics, but the moments are not good.
-Why is this?
-Is it the over/underprediction of the high redshift part of the distribution?
-Discuss this after talking with Rafael.}
-
-\begin{table}
-\setlength{\tabcolsep}{2pt}
-\caption{Moments of the stacked estimator $\hat{N}(z)$ of the redshift distribution.
-Most of the codes considered recover the moments of $\tilde{N}(z)$}
-\label{tab:moments}
-\begin{tabular}{lccc}
-\hline
-\hline
- \multicolumn{4}{l}{Moments of $\hat{N}(z)$} \\
-\hline
-Estimator  & mean       & variance    & skewness \\
-Empirical ``truth'' & 0.701 & 0.630 & 0.671  \\
-\hline
-\annz       & 0.702      & 0.625      & 0.653    \\
-\bpz        & 0.699      & 0.629      & 0.671    \\
-\delight    & 0.692      & 0.609      & 0.638    \\
-\eazy       & 0.681      & 0.595      & 0.619    \\
-\flexzboost & 0.694      & 0.610      & 0.631    \\
-\gpz        & 0.696      & 0.615      & 0.639    \\
-\lephare    & 0.718      & 0.668      & 0.741    \\
-\metaphor   & 0.705      & 0.628      & 0.657    \\
-\cmnn       & 0.701      & 0.628      & 0.667    \\
-\skynet     & 0.743      & 0.708      & 0.797    \\
-\tpz        & 0.700      & 0.619      & 0.643    \\
-\hline
-\trainz	    & 0.699 		 & 0.627 	    & 0.666 \\
-\end{tabular}
-\end{table}
-
-\skynet\ exhibits redshift bias in Figure~\ref{fig:nz} and is a clear outlier in the first moment of $\hat{N}(z)$ in Table~\ref{tab:moments}.
-The \skynet\ algorithm employs a random subsampling of the training set without testing that the subset is representative of the full population, and the implementation used here does not upweight rarer low- and high-redshift galaxies, as in \citet{Bonnett:15}, suggesting a possible cause that may be addressed in future work.
-
 \subsection{Performance on individual \pzpdf s}
@@
 %\red{overall conclusions of the Results section}
-
-\subsection{Interpretation of metrics}
-\label{sec:caution}
-%Caution on metrics and photo-$z$ codes}\label{sec:caution}
-%(Alex Malz, Sam Schmidt)
-
-We remind the reader that contributed codes were given a goal of obtaining accurate \pzpdf s, not an accurate stacked estimator of the redshift distribution, so we do not expect the same codes to necessarily perform well for both classes of metrics.
-Furthermore, our metrics are not necessarily able to assess the fidelity of individual \pzpdf s relative to true posteriors.
-Metric-specific performance implies that we may need multiple \pzpdf\ approaches tuned to each metric in order to maximize returns over all science cases in large upcoming surveys.
-
-The \trainz\ estimator, which assigns every galaxy a \pzpdf\ equal to $N(z)$ of the training set as described in Section~\ref{sec:trainz}, is introduced as a control case to demonstrate this point via \textit{reductio ad absurdum}.
-Because our training set is perfectly representative of the test set, $N(z)$ should be identical for both sets down to statistical noise.
-\textit{We make the alarming observation that \trainz\ outperforms all codes on the PIT-based metrics, and all but one code on the $N(z)$ based statistics.}
-%\textit{Explain the connection between \trainz\ and these metrics.}
-
-The CDE loss and point estimate metrics appropriately penalize \trainz's naivete.
-As shown in Appendix~\ref{sec:pointmetrics}, \trainz ~has identical $ZPEAK$ and $ZWEIGHT$ values for every galaxy, and thus the \pz\ point estimats are constant as a function of true redshift, i.e.
-a horizontal line at the mode and mean of the training set distribution respectively.
-The explicit dependence on the individual posteriors in the calculation of the CDE loss, described in Section~\ref{sec:cdelossresults}, distinguishes this metric from those of the \pzpdf\ ensemble and stacked estimator of the redshift distribution, despite their prevalence in the \pz\ literature.
-
-% While looking at one, or even most of our metrics would have given the impression that this estimator was nearly optimal, red flags in CDE loss and point estimates reveal the problems.
-%[more caution on over-reliance on metrics, and making sure that metrics test all of what you want] \red{this definitely still needs work}
-In summary, context is crucial to defending against deceptively strong performers such as \trainz; the best \pzpdf\ method is the one that most effectively achieves our science goals, not the one that performs best on a metric that does not reflect those goals.
-In the absence of clear goals or the information necessary for a principled metric definition, we must think carefully before choosing a single metric.

### 2019-03-15 paper/tex/analysis.tex
cutting cruft

@@
 
-The goal of this study is to evaluate the degree to which \pzpdf s of each method can be trusted for a generic probabilistic data analysis.
-Obviously \pzpdf s must not take negative values and must integrate to unity over the range of possible redshifts.
+The goal of this study is to evaluate the degree to which \pzpdf s of each method can be trusted for a generic analysis.
 The overloaded ``$p(z)$'' is a widespread abuse of notation that obfuscates this goal, so we will dedicate some attention to breaking it apart.
@@
 The piecewise constant format is chosen because of its established presence in the literature, and the choice of 200 bins was motivated by the approximate number of columns expected to be available for storage of \pzpdf s for the final \lsst\ Project tables.\footnote{See, e.~g.~the \lsst\ Data Products Definition Document, available at: \url{https://ls.st/dpdd}}
-We will discuss the choice of \pzpdf\ parameterization further in Section~\ref{sec:results}.
+We will discuss the choice of \pzpdf\ parameterization further in Section~\ref{sec:discussion}.
 
@@
 We present the metrics of \pzpdf s that address our goals in the sections below.
-Though the outmoded practice should not be encouraged, those seeking a connection to previous comparison studies will find metrics of common redshift point estimators derived from \pzpdf s in Appendix~\ref{sec:pointmetrics}.
-
-\subsection{Metric of individual \pzpdf s}
-\label{sec:CDE_loss}
-%(Alex Malz, Ann Lee)
-
-The \buzz\ simulation process precludes testing the degree to which samples from our \pz\ posteriors reconstruct the space of $p(z, {\rm data})$.
-To the knowledge of the authors, there is only one metric that can be used to evaluate the performance of individual \pzpdf\ estimators in the absence of true \pz\ posteriors.
-Using the notation introduced in Section~\ref{sec:flexzboost}, the conditional density estimation (CDE) loss is defined as
-\begin{equation}
-  \label{eq:cde-loss}
-  L(f, \widehat{f}) \equiv \int \int (f(z \vert \x) - \widehat{f}(z \vert \x))^{2} dz dP(\x)
-\end{equation}
-in terms of the photometry $\x$, an analogue to the familiar root-mean-square-error used in conventional regression.
-We estimate the CDE loss via
-\begin{equation}
-  \label{eq:estimated-cde-loss}
-  \hat{L}(f, \widehat{f}) = \mathbb{E}_\X \left[\int \widehat{f}(z \mid \X)^{2} dz\right] - 2 \mathbb{E}_{\X, Z}\left[\widehat{f}(Z \mid \X)\right] + K_{f},
-\end{equation}
-where the first term is the expectation value of the \pz\ posterior with respect to the marginal distribution of the photometric covariates $\X$, the second term is the expectation value with respect to the joint distribution of $\X$ and the space $Z$ of all possible redshifts, and the third term $K_{f}$ is a constant depending only upon the true conditional densities $f(z \vert \x)$.
-We may estimate these expectations empirically on the test or validation data \citep[Eq.~7 in][]{Izbicki:17b} without knowledge of the true densities.
+Section~\ref{sec:qualmet} outlines aggregate metrics of a catalogue of \pzpdf s, and Section~\ref{sec:CDE_loss} presents a metric of individual \pzpdf s in the absence of true \pzpdf s.
+Though the outmoded practices should not be encouraged, those seeking a connection to previous comparison studies will find metrics of redshift point estimate reductions of \pzpdf s in Appendix~\ref{sec:pointmetrics} and metrics of a science-specific summary statistics heuristically derived from \pzpdf s in Appendix~\ref{sec:moments}.
 
@@
 A quantile of a distribution is the value $q$ at which the CDF of the distribution is equal to $Q$; percentiles and quartiles are familiar examples of linearly spaced sets of 100 and 4 quantiles, respectively.
-The quantile-quantile (QQ) plot serves as a graphical visualization for comparing two distributions, where the quantiles of one distribution are plotted against the quantiles of the other distribution,
-%\blue{Include an equation to ensure quantile definition is crystal clear before building upon it for PIT.}\red{Is maybe the above text clear enough?
-Text may be more clear than any equation I can come up with--SJS.}
-providing an easy way to qualitatively assess the consistency between an estimating distribution and a true distribution.
-%location, scale and skewness.
+The quantile-quantile (QQ) plot serves as a graphical visualization for comparing two distributions, where the quantiles of one distribution are plotted against the quantiles of the other distribution, providing an easy way to qualitatively assess the consistency between an estimating distribution and a true distribution.
 The closer the QQ plot is to diagonal, the closer the match between the distributions.
@@
 \label{eq:pit}
-\mathrm{PIT} &\equiv \rm{CDF}[\hat{p}, z_{\rm true}] %\int_{-\infty}^{z_{\rm true}} p(z) dz
+\mathrm{PIT} &\equiv \rm{CDF}[\hat{p}, z_{\rm true}]
 \end{align}
@@
 
-% \subsection{Metrics over estimated probability distributions}
-% \label{sec:quantmet}
-
-% We compute metrics comparing the CDF of PIT values to the CDF of a Uniform distribution, and also compute the CDF of the true redshift distribution $N'(z)$ compared the $\hat{N}(z)$ distribution derived from summing the \pzpdf s as described in Eq.~\ref{eq:stacked}.
-
 %\subsubsection{Root-mean-square error (RMSE)}
@@
 %Though this metric does not account for the fact that the redshift distribution function is, in fact, a probability distribution, it can still be interpreted as a measure of the integrated difference between the estimated distribution and the true distribution.%, and it can be used to quantify the otherwise qualitative metrics.
-
-% \subsubsection{Kolmogorov-Smirnov (KS) and related statistics}
-% \label{sec:ks}
 
@@
 is a weighted mean-squared difference featuring enhanced sensitivity to discrepancies in the tails of the distribution.
-% The CvM statistic is similar to the KS statistic as it is also computed from the distance between the measured CDF and the ideal CDF, but instead of the maximum distance, the CvM statistic
-% is the average of the distance squared.
-% In this paper we use the CvM statistic to evaluate the performance of both $n(z)$ and $p(z)$.
-% The AD statistic
-% is a weighted version of the CvM statistic, making it more sensitive to the tails of the distribution, where $N_{tot}$ is the sample size.
 %The KS statistic and the CvM and AD statistics related to it are all used to evaluate the performance of both $\hat{n}(z)$ and the QQ and PIT of the CDFs of the ensemble of photo-$z$ interim posteriors.
@@
 %(Alex Malz)
-%
 %The KLD is a measure of the loss in information when using a probability distribution that is an estimate of a known true probability distribution, so we use
@@
 we seek to minimize the KLD.
+
+\subsection{Conditional Density Estimate (CDE) Loss: a metric of individual \pzpdf s}
+\label{sec:CDE_loss}
+%(Alex Malz, Ann Lee)
+
+The \buzz\ simulation process precludes testing the degree to which samples from our \pz\ posteriors reconstruct the space of $p(z, {\rm data})$.
+To the knowledge of the authors, there is only one metric that can be used to evaluate the performance of individual \pzpdf\ estimators in the absence of true \pz\ posteriors.
+Using the notation introduced in Section~\ref{sec:flexzboost}, the conditional density estimation (CDE) loss is defined as
+\begin{equation}
+  \label{eq:cde-loss}
+  L(f, \widehat{f}) \equiv \int \int (f(z \vert \x) - \widehat{f}(z \vert \x))^{2} dz dP(\x)
+\end{equation}
+in terms of the photometry $\x$, an analogue to the familiar root-mean-square-error used in conventional regression.
+We estimate the CDE loss via
+\begin{equation}
+  \label{eq:estimated-cde-loss}
+  \hat{L}(f, \widehat{f}) = \mathbb{E}_\X \left[\int \widehat{f}(z \mid \X)^{2} dz\right] - 2 \mathbb{E}_{\X, Z}\left[\widehat{f}(Z \mid \X)\right] + K_{f},
+\end{equation}
+where the first term is the expectation value of the \pz\ posterior with respect to the marginal distribution of the photometric covariates $\X$, the second term is the expectation value with respect to the joint distribution of $\X$ and the space $Z$ of all possible redshifts, and the third term $K_{f}$ is a constant depending only upon the true conditional densities $f(z \vert \x)$.
+We may estimate these expectations empirically on the test or validation data \citep[Eq.~7 in][]{Izbicki:17b} without knowledge of the true densities.

### 2019-03-15 paper/tex/conclusion.tex
minot clarification

@@
 
-This paper compares the performance of twelve \pzpdf\ approaches under idealized experimental conditions of representative and complete prior information to set a baseline for an upcoming sensitivity analysis by first isolating the impact of the \pzpdf\ estimation technique from realistic physical systematics of the data.
+This paper compares the performance of twelve \pzpdf\ approaches under controlled experimental conditions of representative and complete prior information to set a baseline for an upcoming sensitivity analysis by first isolating the impact of the \pzpdf\ estimation technique from realistic physical systematics of the data.
 Though the mock data set of this experiment did not include true \pz\ posteriors for comparison, \textbf{we interpret deviations from perfect results given perfect prior information as the imprint of the implicit assumptions underlying the estimation approach}.
 
-We evaluated the codes under science-agnostic metrics both established and emerging to stress-test the ensemble properties of \pzpdf\ catalogues derived by each method, as well as metrics of a prevalent summary statistic of \pzpdf\ catalogues used in cosmological analyses.
+We evaluate the codes under science-agnostic metrics both established and emerging to stress-test the ensemble properties of \pzpdf\ catalogues derived by each method, as well as metrics of a prevalent summary statistic of \pzpdf\ catalogues used in cosmological analyses.
 We observe that no one code dominates in all metrics, and that the standard metrics of \pzpdf s and the stacked estimator of the redshift distribution can be gamed by a trivial approach to \pzpdf\ estimation.

### 2019-03-15 paper/tex/analysis.tex
introduce implicit prior language

@@
 %(Alex Malz, Rongpu Zhou, Jeff Newman, Ofer Lahav)
-
-%\red{Maybe it would be best to introduce the summary statistics (KS, CvM, AD) up front and then the ensemble functions and stacked n(z).}
-
-% All code runners were asked to output redshift posterior estimates on 200 linear-spaced bins between redshifts 0 and 2.
 
@@
 \begin{equation}
+  \label{eq:pzpdf}
 \hat{p}^{H}_{i}(z) \equiv p(z \vert d_{i}, I_{D}, I_{H}),
 \end{equation}
-conditioned not only on the photometric data $d_{i}$ for that galaxy but also on parameters encompassing a number of things that will differ depending on the method $H$ used to produce it, namely the assumptions $I_{H}$ necessary for the method to be valid and any inputs $I_{D}$ it takes as prior information, such as a template library or training set.
-Because of this, direct comparison of \pzpdf s produced by different methods is in some sense impossible; even if they share the same prior information $I_{D}$, by definition they cannot be conditioned on the same assumptions $I_{H}$, otherwise they would not be distinct methods at all.
+conditioned not only on the photometric data $d_{i}$ for that galaxy but also on parameters encompassing a number of things that will differ depending on the method $H$ used to produce it, namely the often implicit assumptions $I_{H}$ necessary for the method to be valid and any inputs $I_{D}$ it takes as prior information, such as a template library or training set.
+Because of this, direct comparison of \pzpdf s produced by different methods is in some sense impossible; even if they share the same external prior information $I_{D}$, by definition they cannot be conditioned on the same assumptions $I_{H}$, otherwise they would not be distinct methods at all.
+We call $I_{H}$ the \textit{implicit prior} specific to the method, though some aspects of its nature may be discerned.
 
@@
 These sets of prior information are carefully constructed to be representative and complete, so we have $I_{D} \equiv I_{D}^{\rm ML} \equiv I_{D}^{\rm T}$ for every method $H$.
-Thus, we are saying
-\begin{equation}
-\frac{\hat{p}_{i}^{H}(z)}{\hat{p}_{i}^{H'}(z)} \approx \frac{p(z \vert d_{i}, I_{H})}{p(z \vert d_{i}, I_{H'})},
-\end{equation}
-meaning that we assume comparisons of $\hat{p}_{i}^{H}(z)$ isolate the effect of the method used to obtain the estimator, which should enable interpetation of the differences between estimated PDFs in terms of the specifics of the method implementations.
+Under this assumption, a ratio of posteriors of codes is in effect a ratio of the implicit posteriors $p(z \vert d_{i}, I_{H'})$ since the external prior information $I_{D}$ is present in the numerator and denominator.
+Thus comparisons of $\hat{p}_{i}^{H}(z)$ isolate the effect of the method used to obtain the estimator, which should enable interpretation of the differences between estimated PDFs in terms of the specifics of the method implementations.
 

### 2019-03-15 paper/tex/discussion.tex
checking references and moving content
moving content

@@
 %(Sam Schmidt, Ofer Lahav, Jeff Newman, Alex Malz, Eve Kovacs, Tony Tyson, Tina Peters)
-
-%The goal was not to determine a ``best'' photo-$z$ code.
 
@@
 In contrast with other \pzpdf\ comparison papers that have aimed to identify the ``best'' code for a given survey, we have focused on the somewhat more philosophical questions of how to assess \pzpdf\ methods and how to interpret differences between codes in terms of \pzpdf\ performance.
+In Section~\ref{sec:caution}, we reframe the strong performance of our pathological \pzpdf\ technique, \trainz, as a cautionary tale about the importance of choosing appropriate comparison metrics.
+In Section~\ref{sec:newcodes}, we interpret the performance of the genuine \pzpdf\ codes in terms of concrete improvements that should be pursued by the code developers.
+In Section~\ref{sec:futureexperiments}, we outline the experiments we intend to build upon this study
+In Section~\ref{sec:futuredata}, we discuss the enhancements of the mock data set that will be necessary to enable the future experiments.
 
@@
 \label{sec:caution}
-%Caution on metrics and photo-$z$ codes}\label{sec:caution}
 %(Alex Malz, Sam Schmidt)
@@
 We remind the reader that contributed codes were given a goal of obtaining accurate \pzpdf s, not an accurate stacked estimator of the redshift distribution, so we do not expect the same codes to necessarily perform well for both classes of metrics.
+Indeed, the codes were optimized for their interpretation of our request for ``accurate \pzpdf s,'' and we expect that the implementations would have been adjusted had we requested optimization of the traditional metrics of Appendices~\ref{sec:moments, sec:pointmetrics}.
+
 Furthermore, our metrics are not necessarily able to assess the fidelity of individual \pzpdf s relative to true posteriors.
@@
 
-The \trainz\ estimator, which assigns every galaxy a \pzpdf\ equal to $N(z)$ of the training set as described in Section~\ref{sec:trainz}, is introduced as a control case to demonstrate this point via \textit{reductio ad absurdum}.
+The \trainz\ estimator, which assigns every galaxy a \pzpdf\ equal to $N(z)$ of the training set as described in Section~\ref{sec:trainz}, is introduced as an experimental control or null test to demonstrate this point via \textit{reductio ad absurdum}.
 Because our training set is perfectly representative of the test set, $N(z)$ should be identical for both sets down to statistical noise.
@@
 
-Trends in the metrics shared among codes indicate the
+\subsection{Evolution of \pzpdf\ codes}
+\label{sec:newcodes}
 
-One obvious trend in several of the codes tested was an overall over or underprediction of the widths of \pzpdf s, as evidenced by the QQ plots and PIT histograms shown in Figure~\ref{fig:pitqq}.
-A more careful tuning of bandwidth or smoothing during the validation process appears to be necessary for many of the machine learning based codes in order to improve the accuracy of $p(z)$.
-For narrow peaked $p(z)$ the parameterization of the PDF as evaluated on a fixed redshift grid could also have contributed to some overestimates of $p(z)$ width simply due to the finite resolution.
-After evaluating results such as those presented in \citet[]{Malz:qp}, in future analyses we plan to switch from a fixed grid to quantile-based storage of $p(z)$ in order to more efficiently and accurately store redshift PDF results.
+Trends in the metrics shared among codes indicate avenues for improvement of each approach.
 
-Another important factor to keep in mind when examining the results presented in this paper is the fact that they are at some level dependent on the metrics that we aim to optimize: in this case code participants were asked to submit their optimal measures of an accurate $p(z)$, so participants used the training/validation data to optimize their codes accordingly.
-Had we, instead, asked for an optimal $\hat{N(z)}$ the resulting metrics would be different for most, if not all, of the codes, as they would optimize toward a different goal.
-Specific metric choice can affect which codes are among the ``best'' codes.
-As stated earlier, there are cosmological science cases that require either individual galaxy photo-$z$ measures, or ensemble $\hat{N}(z)$ measures.
-We must be aware of that the optimal method for one is not necessarily optimal for the other, and in fact several photo-$z$ algorithms may be necessary in the final cosmological analysis in order to satisfy the requirements of all science use cases.
+Section~\ref{sec:pitqq} identifies over- and under-estimation of the widths of \pzpdf s by certain codes.
+We suggest that the smoothing parameter present in \bpz, \cmnn, \delight, \eazy, \gpz, \skynet, and \tpz\ be chosen more carefully to improve their performance in the QQ plots and PIT histograms of
 
@@
 
-\subsubsection{Nontrivial experimental design}
+\subsection{Extended experimental design}
 \label{sec:futureexperiments}
@@
 The work presented in this paper is only the first step in characterizing current photo-$z$ codes and moving toward an improved photometric redshift estimator.
+
+As stated earlier, there are cosmological science cases that require either individual galaxy photo-$z$ measures, or ensemble $\hat{N}(z)$ measures.
+We must be aware of that the optimal method for one is not necessarily optimal for the other, and in fact several photo-$z$ algorithms may be necessary in the final cosmological analysis in order to satisfy the requirements of all science use cases.
+
 This initial paper explored code performance in idealized conditions with perfect catalog-based photometry and representative training data.
@@
 
-\subsubsection{More realistic mock data}
+Sequels to this study will also address some shortcomings of our experimental procedure.
+For narrow peaked $p(z)$ the parameterization of the PDF as evaluated on a fixed redshift grid could also have contributed to some overestimates of $p(z)$ width simply due to the finite resolution.
+After evaluating results such as those presented in \citet[]{Malz:qp}, in future analyses we plan to switch from a fixed grid to quantile-based storage of $p(z)$ in order to more efficiently and accurately store redshift PDF results.
+
+\subsection{Realistic mock data}
 \label{sec:futuredata}

### 2019-03-15 paper/tex/results.tex
moving content

@@
 We begin with a demonstrative visual inspection of the \pzpdf s produced by each code for individual galaxies.
-Figure~\ref{fig:pz_examples} shows the \pzpdf s for four galaxies chosen as examples of \pzpdf\ archetypes: a narrow unimodal redshift solution, a broad unimodal solution, a bimodal distribution, and a multimodal distribution.
+Figure~\ref{fig:pz_examples} shows the \pzpdf s for four galaxies chosen as examples of \pzpdf\ archetypes: a narrow unimodal PDF, a broad unimodal PDF, a bimodal PDF, and a multimodal PDF.
+We reiterate that under our idealized experimental conditions, differences between codes are the isolated signature of the implicit prior due to the method by which the \pzpdf s were derived.
 
@@
 This observation is consistent with the findings of \citet[]{Malz:qp} that the piecewise constant parameterization underperforms in the presence of small-scale structures.
+
 However, the shared small-scale features of \annz, \metaphor, \cmnn, and \skynet\ are a result of various weighted sums of the limited number of training set galaxies with colors similar to those of the test set galaxy in question, with behavior closer to classification than regression in the case of \annz.

### 2019-03-15 paper/tex/discussion.tex
consolidating

@@
 In Section~\ref{sec:caution}, we reframe the strong performance of our pathological \pzpdf\ technique, \trainz, as a cautionary tale about the importance of choosing appropriate comparison metrics.
-In Section~\ref{sec:newcodes}, we interpret the performance of the genuine \pzpdf\ codes in terms of concrete improvements that should be pursued by the code developers.
 In Section~\ref{sec:futureexperiments}, we outline the experiments we intend to build upon this study
@@
 
-The \trainz\ estimator, which assigns every galaxy a \pzpdf\ equal to $N(z)$ of the training set as described in Section~\ref{sec:trainz}, is introduced as an experimental control or null test to demonstrate this point via \textit{reductio ad absurdum}.
+The \trainz\ estimator of Section~\ref{sec:trainz}, which assigns every galaxy a \pzpdf\ equal to $N(z)$ of the training set, is introduced as an experimental control or null test to demonstrate this point via \textit{reductio ad absurdum}.
 Because our training set is perfectly representative of the test set, $N(z)$ should be identical for both sets down to statistical noise.
@@
 %[more caution on over-reliance on metrics, and making sure that metrics test all of what you want] \red{this definitely still needs work}
-In summary, context is crucial to defending against deceptively strong performers such as \trainz; the best \pzpdf\ method is the one that most effectively achieves our science goals, not the one that performs best on a metric that does not reflect those goals.
-In the absence of clear goals or the information necessary for a principled metric definition, we must think carefully before choosing a single metric.
-
-\subsection{Evolution of \pzpdf\ codes}
-\label{sec:newcodes}
-
-Trends in the metrics shared among codes indicate avenues for improvement of each approach.
-
-Section~\ref{sec:pitqq} identifies over- and under-estimation of the widths of \pzpdf s by certain codes.
-We suggest that the smoothing parameter present in \bpz, \cmnn, \delight, \eazy, \gpz, \skynet, and \tpz\ be chosen more carefully to improve their performance in the QQ plots and PIT histograms of
-
-The example of the simple \trainz\ estimator described in Section~\ref{sec:caution} shows a simple model with a $p(z)$ that is unrealistic for individual objects can still score very well on many of our metrics.
-It is important to look at {\it all} metrics, and keep in mind what information each metric conveys.
-%\red{mention 'null' and how 'dumb' algorithms can score well on metrics, need to keep all metrics in mind}
+In summary, context is crucial to defending against deceptively strong performers such as \trainz; \textbf{the best \pzpdf\ method is the one that most effectively achieves our science goals}, not the one that performs best on a metric that does not reflect those goals.
+In the absence of a single scientific motivation or the information necessary for a principled metric definition, we must consider many metrics and be critical of the information transmitted by each.
+% As stated earlier, science applications may be sensitive to individual \pzpdf s, \pzpdf\ ensembles, \pz\ point estimates, or science-specific statistics like $\hat{N}(z)$.
+% We must be aware of that the optimal method for one is not necessarily optimal for the other, and in fact several photo-$z$ algorithms may be necessary in the final cosmological analysis in order to satisfy the requirements of all science use cases.
+% The example of the simple \trainz\ estimator described in Section~\ref{sec:caution} shows a simple model with a $p(z)$ that is unrealistic for individual objects can still score very well on many of our metrics.
+% It is important to look at {\it all} metrics, and keep in mind what information each metric conveys.
 %\aim{[reiterate the meaning of $p(z)$ and the goal -- emphasize finding ``best'' code in terms of impact of assumptions underlying the method \textit{or} establishing methodology for the realistic case of biased prior information and other science goals]}
@@
 
-\subsection{Extended experimental design}
+\subsection{Extensions to the experimental design}
 \label{sec:futureexperiments}
@@
 %\item \red{the combination of codes}
-
 %\item \red{$p(z,\alpha)$}
@@
 
-The work presented in this paper is only the first step in characterizing current photo-$z$ codes and moving toward an improved photometric redshift estimator.
-
-As stated earlier, there are cosmological science cases that require either individual galaxy photo-$z$ measures, or ensemble $\hat{N}(z)$ measures.
-We must be aware of that the optimal method for one is not necessarily optimal for the other, and in fact several photo-$z$ algorithms may be necessary in the final cosmological analysis in order to satisfy the requirements of all science use cases.
+The work presented in this paper is only a first step in assessing \pzpdf\ approaches and moving toward an improved photometric redshift estimator.
+Here we discuss the next steps for subsequent investigations.
 
 This initial paper explored code performance in idealized conditions with perfect catalog-based photometry and representative training data.
-As mentioned in Section~\ref{sec:metrics} for the stacked $N(z)$ metrics we examined only the entire galaxy population with no selections in either photo-$z$ ``quality" or redshift.
-The cosmological analyses for weak lensing and large scale structure based measures plan to break galaxy samples into tomographic redshift bins, using photo-$z$ $p(z)$ to infer the redshift distribution for each bin.
-The specific selection used to determine these bins, both algorithmically and the specific bin boundaries, could induce biases due to indirect selections inherent in the photo-$z$ or other bin selection parameters.
-The effects of tomographic bin selection will be explored in a dedicated future paper, including propagation of redshift uncertainties in a set of fiducial tomographic redshift bins in order to estimate impact on cosmological parameter estimation.
+A top priority for a follow-up study is to test realistic forms of incomplete and non-representative template libraries and training sets as well as the impact of other forms of external priors that must be ingested by the codes.
+
+Appendix~\ref{sec:moments} only addresses the stacked estimator of the redshift distribution of the entire galaxy catalogue rather than subsets in bins, tomographic or otherwise.
+The effects of tomographic binning scheme will be explored in a dedicated future paper, including propagation of redshift uncertainties in a set of fiducial tomographic redshift bins in order to estimate impact on cosmological parameter estimation.
 
@@
 
-We re-emphasize that the dataset tested was quite idealized, and discuss enhancements that will be added in future simulations to test photo-$z$ codes on increasingly realistic conditions in the following section.
-
 Sequels to this study will also address some shortcomings of our experimental procedure.
-For narrow peaked $p(z)$ the parameterization of the PDF as evaluated on a fixed redshift grid could also have contributed to some overestimates of $p(z)$ width simply due to the finite resolution.
-After evaluating results such as those presented in \citet[]{Malz:qp}, in future analyses we plan to switch from a fixed grid to quantile-based storage of $p(z)$ in order to more efficiently and accurately store redshift PDF results.
+The fixed redshift grid shared between the codes may have unfairly penalized codes with a different native parameterization, as precision is lost when converting between formats.
+Performance on sharply peaked \pzpdf s may have been suppressed across all codes due to the insufficient resolution of the redshift grid.
+In light of the results of \citet[]{Malz:qp}, in future analyses we plan to switch from a fixed grid to the quantile parameterization or to permit each code to use its native storage format provided they all store the same number of parameters.
 
@@
 \label{sec:futuredata}
+
+We re-emphasize that the dataset tested was quite idealized, and discuss enhancements that will be added in future simulations to test photo-$z$ codes on increasingly realistic conditions in the following section.
 

### 2019-03-15 paper/tex/results.tex
consolidating

@@
 
-Figure~\ref{fig:pitqq} shows a histogram of PIT values, QQ plot, and QQ difference plot relative to the ideal diagonal, showcasing the biases and trends in the average accuracy of the \pzpdf s for each code.
+The histogram of PIT values, QQ plot, and QQ difference plot relative to the ideal diagonal are provided in Figure~\ref{fig:pitqq}, showcasing the biases and trends in the average accuracy of the \pzpdf s for each code.
 The high QQ values (more high than low PIT values) of \bpz, \cmnn, \delight, \eazy, and \gpz\ indicate \pzpdf s biased low, and the low QQ values (more low than high PIT values) of \skynet\ and \tpz\ indicate \pzpdf s biased high.

### 2019-03-15 paper/tex/discussion.tex
consolidation

@@
 This initial paper explored code performance in idealized conditions with perfect catalog-based photometry and representative training data.
-A top priority for a follow-up study is to test realistic forms of incomplete and non-representative template libraries and training sets as well as the impact of other forms of external priors that must be ingested by the codes.
+A top priority for a follow-up study is to test realistic forms of incomplete, erroneous, and non-representative template libraries and training sets as well as the impact of other forms of external priors that must be ingested by the codes, major concerns in \citet{Newman:2015, Masters:2017}.
+We plan to perform a full sensitivity analysis on a realistically incomplete training set of spectroscopic galaxies, modeling the performance of spectrographs, emission-line properties, and expected signal-to-noise to determine which potential training set galaxies are most likely to be excluded.
+In addition to outright redshift failures we will model the inclusion of a small number of high-confidence yet false redshifts due to emission line misidentification or noise spikes.
 
@@
 
-We re-emphasize that the dataset tested was quite idealized, and discuss enhancements that will be added in future simulations to test photo-$z$ codes on increasingly realistic conditions in the following section.
-
-%\item \red{inclusion of incompleteness effects on training sample, ``wrong redshifts'', filter curves}
+The goals of future investigations will require a realistically complex data set of higher sophistication than the DC1 data set used here.
 %\item \red{inclusion of photometric errors, image based effects, blending, lensing, sky, mask, etc...}
-
-In future papers a focus of the \lsstdesc\ \Pz\ Working Group will be to add more and more complexity to our simulated data in order to test photo-$z$ algorithms in increasingly realistic conditions.
-The most pressing concern is the impact of incomplete spectroscopic training samples.
-The SEDs for the galaxy sample in this paper were constructed form linear combinations of five basis SED templates.
-Future simulations will also include more complex SED information, with a more realistic range of physical properties, and the inclusion of AGN effects, a more insidious problem, where AGN features may not be apparent, but the colors and other host galaxy properties are perturbed relative to galaxies with an inactive nucleus.
-In such cases, the presence of the AGN may induce a bias if the template SEDs or empirical datasets do not include low-level AGN counterparts.
-
-As discussed extensively in \citet{Newman:2015} a representative set of spectroscopically confirmed galaxies spanning the full range of both redshift and apparent magnitude is necessary as a training set to characterize the mapping from broad-band fluxes to photometric redshifts.
-%However, due to a combination of factors due to both the galaxy SEDs and limitations of spectrographic instruments, redshift samples are known to be systematically incomplete, where certain galaxy types and redshift intervals fail to yield a redshift even at the longest integration times on current and near-future instruments.
-The more representative the training data, the better the performance of photo-$z$ algorithms will be.
-Current and upcoming surveys are putting significant effort into obtaining these training samples \citep[e.~g.\,][]{Masters:2017}, however we still expect significant incompleteness for LSST-like samples, particularly at faint magnitudes.
-We plan to produce a realistically incomplete training set of spectroscopic galaxies, modeling the performance of spectrographs, emission-line properties, and expected signal-to-noise to determine which galaxies will fail to yield a secure redshift.
-In addition to outright redshift failures we will model the inclusion of a small number of falsely identified secure redshifts where misidentified emission lines or noise spikes cause an incorrect redshift solution to be marked as a high quality identification.
-Even sub-per cent level contamination by false redshifts can impact \pz\ solutions at levels comparable to the stringent  requirements of some LSST science cases.
-We expect different systematics to occur in different \pzpdf\ codes in response to training on incomplete data, particularly some of the machine learning methods.
-The response of the codes will inform future directions of code development.
 
@@
 Object blending will be a major area of investigation, as the mixing of flux from multiple objects and the resultant change in measured colours is predicted to affect a large fraction of LSST galaxies \citep{Dawson:2016}, and will be one of the major contributing systematics for photo-$z$'s.
+
+The DC1 galaxy SEDs were linear combinations of just five basis SED templates, but a next generation of data for \pzpdf\ investigations must include a broader range of physical properties.
+A subsequent study must also have a data set that includes active galactic nuclei (AGN) features in the SEDs, which perturb colours and other host galaxy properties.
 

### 2019-03-15 paper/tex/discussion.tex
finished editing Discussion

@@
 Performance on sharply peaked \pzpdf s may have been suppressed across all codes due to the insufficient resolution of the redshift grid.
-In light of the results of \citet[]{Malz:qp}, in future analyses we plan to switch from a fixed grid to the quantile parameterization or to permit each code to use its native storage format provided they all store the same number of parameters.
+In light of the results of \citet[]{Malz:qp}, in future analyses we plan to switch from a fixed grid to the quantile parameterization or to permit each code to use its native storage format under a shared number of parameters.
 
@@
 
-The goals of future investigations will require a realistically complex data set of higher sophistication than the DC1 data set used here.
+To make optimal use of the \lsst\ data for cosmological and other astrophysical analyses of the Science Roadmap, future investigations that build upon this one will require a more sophisticated set of galaxy photometry and redshifts.
 %\item \red{inclusion of photometric errors, image based effects, blending, lensing, sky, mask, etc...}
-
-The underlying dataset limited this work to a maximum redshift of $z=2$.
-LSST imaging after 10 years of observations will include a significant number of $z>2$ galaxies in expected cosmology samples, and their inclusion does have potential significant implications for photo-$z$ measures: the high redshift galaxies lie at fainter apparent magnitudes and can have anomalous colours due to evolution of stellar populations and the shift to rest-frame magnitudes probing UV features of the underlying SED.
-More importantly, one of the most common ``catastrophic outlier'' degeneracies observed in deep photometric samples occurs when the Lyman break is mistaken for the Balmer break, leading to multiple redshift solutions at $z\sim0.2-0.3$ and $z\sim2-3$ \citep{Massarotti:2001}.
-This degeneracy, along with other potential degeneracies, are currently not covered by the limited redshift range of this initial paper, which could mean that we are not probing the full range of potential extreme outlier populations and how our \pz\ estimators respond to them.
-Extending simulations to include the high-redshift galaxy population will be a priority in future data challenges.
-
 This initial paper explored a data set that was constructed at the catalog level, with no inclusion of the complications that come from measuring photometry from images.
-Future data challenges will move to catalogs constructed from mock images, including effects that will have great impact on \pz\ estimates, which will naturally include the complications of object blending, sensor effects, different observing conditions, amongst others.
-Object blending will be a major area of investigation, as the mixing of flux from multiple objects and the resultant change in measured colours is predicted to affect a large fraction of LSST galaxies \citep{Dawson:2016}, and will be one of the major contributing systematics for photo-$z$'s.
+Future data challenges will move to catalogs constructed from mock images, including the complications of deblending, sensor inefficiencies, and heterogeneous observing conditions, all anticipated to affect the measured colours of \lsst's galaxy sample \citep{Dawson:2016}.
 
 The DC1 galaxy SEDs were linear combinations of just five basis SED templates, but a next generation of data for \pzpdf\ investigations must include a broader range of physical properties.
-A subsequent study must also have a data set that includes active galactic nuclei (AGN) features in the SEDs, which perturb colours and other host galaxy properties.
+Though we only considered $z < 2$ here, \lsst\ 10-year data will contain $z > 2$ galaxies, plagued by fainter apparent magnitudes and anomalous colours due to stellar evolution.
+A subsequent study must also have a data set that includes low-level active galactic nuclei (AGN) features in the SEDs, which perturb colours and other host galaxy properties.
+An observational degeneracy between the Lyman break of a $z \sim 2-3$ galaxy from the Balmer break of a $z \sim 0.2-0.3$ galaxy is a known source of catastrophic outliers \citep{Massarotti:2001} that was not effectively included in this study.
+To gauge the sensitivity of \pzpdf\ estimators to catastrophic outliers, our data set must include realistic high-redshift galaxy populations.
 
-In this study we have not accounted for the presence of Active Galactic Nuclei (AGN) contributions to galaxy fluxes.
-In some cases, AGN will be easily identified from the colors and morphologies, i.e.
+% In this study we have not accounted for the presence of Active Galactic Nuclei (AGN) contributions to galaxy fluxes.
+% In some cases, AGN will be easily identified from the colors and morphologies, i.e.
 the case of the brightest quasars where the nuclear activity outshines the host galaxy, and numerous studies have utilized color selection to create large samples of quasars \citep[e.g.][]{Richards:06,Maddox:08,Richards:15}.
-In current deep fields, similar in depth to what we expect from \lsst, variability information and multi-wavelength data have been critical to not only identify AGN dominated galaxies, but also obtain more accurate photometric redshifts \citep[e.g][]{Salvato:11}.
+% In current deep fields, similar in depth to what we expect from \lsst, variability information and multi-wavelength data have been critical to not only identify AGN dominated galaxies, but also obtain more accurate photometric redshifts \citep[e.g][]{Salvato:11}.
+%
+% In addition to AGN dominated galaxies, those with lower levels of nuclear activity present a more insidious problem, where AGN features may not be apparent, but the colours and other host galaxy properties are perturbed relative to galaxies with an inactive nucleus.
+% In such cases, the presence of the AGN may induce a bias if the template SEDs or empirical datasets do not include low-level AGN counterparts.
+% For LSST, we will need to identify and obtain accurate photometric redshifts of all types of AGN for a range of science goals, whether it is to eliminate such objects from cosmology experiments, or to use them with confidence, all the way through to understanding galaxy evolution and the role that AGN may play in influencing galaxy properties over cosmic time.
 
-In addition to AGN dominated galaxies, those with lower levels of nuclear activity present a more insidious problem, where AGN features may not be apparent, but the colours and other host galaxy properties are perturbed relative to galaxies with an inactive nucleus.
-In such cases, the presence of the AGN may induce a bias if the template SEDs or empirical datasets do not include low-level AGN counterparts.
-For LSST, we will need to identify and obtain accurate photometric redshifts of all types of AGN for a range of science goals, whether it is to eliminate such objects from cosmology experiments, or to use them with confidence, all the way through to understanding galaxy evolution and the role that AGN may play in influencing galaxy properties over cosmic time.
+% A promising route to classifying and obtaining accurate photometric redshifts for the AGN population is by combining machine learning with template-fitting techniques, as has recently been demonstrated by \citet{Duncan:18} for radio-selected AGN.
+% This is because AGN are relatively easy to obtain spectroscopic redshifts for over all redshifts due to the strong emission lines that they exhibit, allowing very good training sets for machine learning algorithms to use.
+% Whereas for those galaxies where the AGN is sub-dominant the galaxy templates are still adequate for obtaining reasonable photometric redshifts.
 
-A promising route to classifying and obtaining accurate photometric redshifts for the AGN population is by combining machine learning with template-fitting techniques, as has recently been demonstrated by \citet{Duncan:18} for radio-selected AGN.
-This is because AGN are relatively easy to obtain spectroscopic redshifts for over all redshifts due to the strong emission lines that they exhibit, allowing very good training sets for machine learning algorithms to use.
-Whereas for those galaxies where the AGN is sub-dominant the galaxy templates are still adequate for obtaining reasonable photometric redshifts.
+% In addition to these improvements, the \lsstdesc \Pz\ Working Group plans to look at all potential methods to combine the results from multiple \pzpdf\ codes to improve accuracy, similar to the work presented in \citet{Dahlen:13,Carrascokind:14,Duncan:18}.
+% Taking advantage of multiple algorithms that use observables in slightly different ways has shown promise, however we must be very conscious of whether a potential combination properly treats the covariance between the methods, given that they are estimating quantities based on the same underlying observables.
+% Several science cases wish to estimate physical quantities along with redshift, for example galaxy stellar mass and star formation rate.
+% Proper joint estimation of redshift and physical quantities requires an in depth understanding of galaxy evolution, and progress on accurate bivariate redshift probability distributions will go hand in had with progress on understanding galaxies themselves.
+% Parameterization and storage of a complex 2-dimensional probability surface for potentially billions of galaxies (or even subsets of hundreds of thousand of particular interest) pose a potential challenge.
+% These issues will be examined in another future paper.
 
-In addition to these improvements, the \lsstdesc \Pz\ Working Group plans to look at all potential methods to combine the results from multiple \pzpdf\ codes to improve accuracy, similar to the work presented in \citet{Dahlen:13,Carrascokind:14,Duncan:18}.
-Taking advantage of multiple algorithms that use observables in slightly different ways has shown promise, however we must be very conscious of whether a potential combination properly treats the covariance between the methods, given that they are estimating quantities based on the same underlying observables.
-Several science cases wish to estimate physical quantities along with redshift, for example galaxy stellar mass and star formation rate.
-Proper joint estimation of redshift and physical quantities requires an in depth understanding of galaxy evolution, and progress on accurate bivariate redshift probability distributions will go hand in had with progress on understanding galaxies themselves.
-Parameterization and storage of a complex 2-dimensional probability surface for potentially billions of galaxies (or even subsets of hundreds of thousand of particular interest) pose a potential challenge.
-These issues will be examined in another future paper.
-
-Finally, while this paper and future papers discussed above focus on photometric redshift codes and estimating accurate \pzpdf s from training data, we plan a separate, but complementary, project to examine calibration of the resultant redshifts via spatial cross-correlations \citep{Newman:2008}, which will be explored in a separate series of future papers.
+% Finally, while this paper and future papers discussed above focus on photometric redshift codes and estimating accurate \pzpdf s from training data, we plan a separate, but complementary, project to examine calibration of the resultant redshifts via spatial cross-correlations \citep{Newman:2008}, which will be explored in a separate series of future papers.
 The overarching plan describing everything laid out in this section is described in more detail in the \lsstdesc\ Science Roadmap (see Footnote in Section~\ref{sec:intro}).
-These plans will require significant effort, but they are necessary if we are to make optimal use of the LSST data for astrophysical and cosmological analyses.

### 2019-03-15 paper/tex/conclusion.tex
rewording

@@
 
-This paper compares the performance of twelve \pzpdf\ approaches under controlled experimental conditions of representative and complete prior information to set a baseline for an upcoming sensitivity analysis by first isolating the impact of the \pzpdf\ estimation technique from realistic physical systematics of the data.
-Though the mock data set of this experiment did not include true \pz\ posteriors for comparison, \textbf{we interpret deviations from perfect results given perfect prior information as the imprint of the implicit assumptions underlying the estimation approach}.
+This paper compares twelve \pzpdf\ codes under controlled experimental conditions of representative and complete prior information to set a baseline for an upcoming sensitivity analysis.
+This work isolates the impact on metrics of \pzpdf\ accuracy due to the estimation technique as opposed to the complications of realistic physical systematics of the photometry.
+Though the mock data set of this investigation did not include true \pz\ posteriors for comparison, \textbf{we interpret deviations from perfect results given perfect prior information as the imprint of the implicit assumptions underlying the estimation approach}.
 
-We evaluate the codes under science-agnostic metrics both established and emerging to stress-test the ensemble properties of \pzpdf\ catalogues derived by each method, as well as metrics of a prevalent summary statistic of \pzpdf\ catalogues used in cosmological analyses.
-We observe that no one code dominates in all metrics, and that the standard metrics of \pzpdf s and the stacked estimator of the redshift distribution can be gamed by a trivial approach to \pzpdf\ estimation.
-We emphasize to the \pz\ community that \textbf{metrics used to vet \pzpdf\ methods must be tailored to science cases with self-consistent analysis approaches that respect the inherently probabilistic nature of \pzpdf\ catalogs}.
+We evaluate the twelve codes under science-agnostic metrics both established and emerging to stress-test the ensemble properties of \pzpdf\ catalogues derived by each method.
+In appendices, we also present metrics of point estimates and a prevalent summary statistic of \pzpdf\ catalogues used in cosmological analyses to enable the reader to relate this work to studies of similar scope.
+We observe that no one code dominates in all metrics, and that the standard metrics of \pzpdf s and the stacked estimator of the redshift distribution can be gamed by a vacuously wrong procedure that asserts the prior over the data.
+We emphasize to the \pz\ community that \textbf{metrics used to vet \pzpdf\ methods must be scrutinized to ensure they correspond to the quantities that matter to our science}.

### 2019-03-15 paper/tex/appendix.tex
streamlining point estimate metrics

@@
 Though we do not endorse the use of the stacked estimator of the redshift distribution, we use it under the untested assumption that the response of our metrics of $\hat{N}^{H}(z)$ will be analogous to the same metrics applied to a principled estimator of the redshift distribution.
-In addition, very few scientific use cases will employ the overall $\hat{N}(z)$ with no cuts, as we explore in this paper.
-We discuss in Section~\ref{sec:futurework} variants of the stacked estimator, such as realistic tomographic binning schemes, that will be explored in future work.
 
@@
 
-While we assume that all science analysis will use full PDF information and do not recommend the use of single point estimates of redshift for most science applications, we include a brief evaulation as an Appendix.
-Plots of the point estimates can be a useful qualitative diagnostic of photo-z code performance, i.~e.~examining point photo-$z$ vs.~spec-$z$ plots visually can give a quick impression of some common trends in different codes.
-Computing point estimate statistics may also be useful for more direct historical comparisons from older photo-z evaluations.
-If a point-estimate is preferred for a specific science case, it is fairly simple to compute the mean, mode, or some other simple estimator from each $p(z)$, so these point estimates can be easily derived from the stored $p(z)$.
+While this work assumes that science applications value the information of the full \pzpdf, we present conventional metrics of \pz\ point estimates as a quick and dirty diagnostic tool and to facilitate direct comparisons to historical studies.
 
@@
 
-There are several common point estimators of photo-$z$ posteriors employed by different codes, e.~g. the mode, mean, median of the $p(z)$ distribution.
-In addition, many of the machine learning based estimators can be set up to return a single redshift solution.
-For example, \skynet\ can be configured to run as a regressor that returns a single float rather than a classifier that returns a 200-bin $p(z)$ estimate.
-The single value returned by a machine learning based code may not correspond to a particular measure such as the mode or mean, and so to avoid interpretation of results that might be introduced by variations in choice of specific point-estimate implementation per code, we discard the code-specific point estimates.
-We instead calculate point estimates more uniformly across the codes directly from the $p(z)$ using two measures, $z_{PEAK}$ and $z_{WEIGHT}$.
+Though we acknowledge that many of the codes can also return a native \pz\ point estimate, we consider two common \pz\ point estimators, the mode $z_{PEAK}$ and mean $z_{WEIGHT}$, derived from \pzpdf s.
 $z_{PEAK}$  is simply the maximum value attained for each galaxy p(z), the mode of the probability distribution.

### 2019-03-15 paper/tex/appendix.tex
streamlining definition of point estimate

@@
 
-Though we acknowledge that many of the codes can also return a native \pz\ point estimate, we consider two common \pz\ point estimators, the mode $z_{PEAK}$ and mean $z_{WEIGHT}$, derived from \pzpdf s.
-$z_{PEAK}$  is simply the maximum value attained for each galaxy p(z), the mode of the probability distribution.
-$z_{WEIGHT}$ is defined similarly to how it is defined in \citet{Dahlen:13}, as the weighted mean of the redshift over the {\it main peak} of $p(z)$ containing the $z_{PEAK}$ value.
-The main peak is defined by subtracting 0.05$\,\times\,z_{PEAK}$ from $p(z)$ and identifying the roots to isolate the peak containing $z_{PEAK}$, $z_{WEIGHT}$ is defined as the weighted mean redshift within this peak.
-We restrict to a single peak in order to avoid confusion from bimodal and multimodal $p(z)$ such as those shown in bottom panels of Figure~\ref{fig:pz_examples}.
-For example, for a bimodal probability distribution a weighted mean calculated over both peaks would fall between the peaks, at a redshift where the probability is minimal.
-Restricting the weighting to a single peak ensures that the point estimate will fall in the region of maximum redshift probability.
-%\red{Have Dritan check zweight description to make sure this is what was done.}
-% The most common point estimators of photo-$z$ PDFs are the mode, median, and mean of the distribution.
-Following \cite{Tanaka:17}, we report the values of the following metrics using whichever of these point estimators performs best for each method, noting which point reduction is used in each case.
-%
+Though we acknowledge that many of the codes can also return a native \pz\ point estimate, we put all codes on equal footing by considering two generic \pz\ point estimators, the mode $z_{PEAK}$ and main-peak-mean $z_{WEIGHT}$ \citep{Dahlen:13}, a weighted mean within the bounds of the main peak, as identified by the roots of $p(z) - 0.05 \times z_{PEAK}$.
+Though $z_{WEIGHT}$ neglects information in a secondary peak of a, say, bimodal distribution, it avoids the pitfall of reducing the \pzpdf\ to a redshift between peaks where there is low probability.
+
 \subsection{Metrics of \pz\ point estimates}

### 2019-03-15 paper/main.tex
macros
worked through GPZ
shorter title
removing unused macro
adding back conclusion
code macros

@@
 \usepackage{amsfonts}
+\usepackage{footmisc}
 \graphicspath{{./}{./figures/}}
@@
 \newcommand{\Pzpdf}{\Pz\ PDF}
-\newcommand{\nz}{$N(z)$}
+\newcommand{\chisq}{$\chi^{2}$}
 
@@
 
-\title{Evaluating approaches to deriving photometric redshift posterior probability distribution functions for \textsc{LSST}}
+\title{Evaluation of probabilistic photometric redshift estimation approaches for \textsc{LSST}}
 
@@
 
-%\input{./tex/conclusion}
+\input{./tex/conclusion}
 
@@
 
-In addition to packages cited in the text, analyses performed in this paper used the following software packages: \textsc{Numpy} and \textsc{Scipy} \citep{numpyscipy}, \textsc{Matplotlib} \citep{matplotlib}, \textsc{Seaborn} \citep{seaborn}, \textsc{minFunc} \citep{minfunc}, \textsc{pySkyNet} \citep{pyskynet}, and {\textsc photUtils} from the \lsst\ simulations package \citep{lsstphotutils}.
+In addition to packages cited in the text, analyses performed in this paper used the following software packages: \code{Numpy} and \code{Scipy} \citep{numpyscipy}, \code{Matplotlib} \citep{matplotlib}, \code{Seaborn} \citep{seaborn}, \code{minFunc} \citep{minfunc}, \code{pySkyNet} \citep{pyskynet}, and \code{photUtils} from the \lsst\ simulations package \citep{lsstphotutils}.
 
@@
 
-% This work used some telescope which is operated/funded by some agency or consortium or foundation ...
-
-% The DESC acknowledges ongoing support from the Institut National de Physique Nucleaire et de Physique des Particules in France; the Science \& Technology Facilities Council in the United Kingdom; and the Department of Energy, the National Science Foundation, and the LSST Corporation in the United States.
-%
-% DESC uses resources of the IN2P3 Computing Center (CC-IN2P3--Lyon/Villeurbanne - France) funded by the Centre National de la Recherche Scientifique; the National Energy Research Scientific Computing Center, a DOE Office of Science User Facility supported by the Office of Science of the U.S.
-Department of Energy under Contract No.
-DE-AC02-05CH11231; STFC DiRAC HPC Facilities, funded by UK BIS National E-infrastructure capital grants; and the UK particle physics grid, supported by the GridPP Collaboration.
-%
-% This work was performed in part under DOE Contract DE-AC02-76SF00515.
-
-% We acknowledge the use of An-External-Tool-like-NED-or-ADS.
-
-%{\it Facilities:} \facility{LSST}
-
 \input{./tex/appendix}
 
-% Include both collaboration papers and external citations:
 \bibliography{main,lsstdesc}

### 2019-03-16 paper/tex/appendix.tex
finished second appendix editing

@@
 
-We calculate the commonly used point estimate metrics of the overall photo-$z$ scatter ($\sigma_{z}$, the standard deviation of the photo-$z$ residuals), bias, and ``catastrophic outlier rate''.
-Specifically, we calculate the metrics as follows:
-we define $e_{z}$ as
-\begin{equation}
-e_{z} = \frac{z_{P} -z_{S}}{1+z_{S}}
-\end{equation}
-where $z_{P}$ is the point estimate and $z_{S}$ is the true redshift.
-In practice, because the standard deviation calculation is quite sensitive to the outliers, we define the photo-$z$ scatter, $\sigma$ in terms of the Interquartile Range (IQR), the difference between the 75th and 25th percentiles of the $e_{z}$ distribution.
-In order to match the usual meaning of a 1$\sigma$ interval, we scale the IQR and define $\sigma_{IQR} = IQR/1.349$, as there is a factor of 1.349 difference between the IQR and the standard deviation of a Normal distribution.
-%$\sigma_{z}$ is, in practice, defined in terms the Interquartile Range (IQR) of $e_{z}$ values as $IQR/1.349$, where $IQR$ is the difference between the 75th and 25th percentiles of the $e_{z}$ distribution.
-The factor of 1.349 relates the IQR of a Normal distribution to the standard 68\% range of a one $\sigma$ uncertainty.
-%\scc{usually the statistical indicator defined as $\sigma$ is the standard deviation I know that is not a rule but if someone skips the test and go directly to the table (and we all know that a lot of people skips the definition of statistical indicator) could be confusing, could we report it as $\frac{IQR(e_z)}{1.349}$ or $\sigma_{IQR}$ that should be clear without any risk of confusion?}\red{Using IQR is done because IQR is less sensitive to outliers than calculating a sigma.
-I've rewritten the text explaining this above.--SJS}
-While many other studies define the bias based on the {\it mean} offset between true and estimated redshift, in this study we define the bias as the median value of $e_{z}$ for the sample.
-We use median as it is, once again, less sensitive to outliers than the mean.
-The catastrophic outlier fraction is defined as the fraction of galaxies with $e_{z}$ greater than the {\it larger} of $3\sigma_{IQR}$ or 0.06, i.e.
-3$\sigma$ outliers with a floor of $\sigma_{IQR}$=0.02.
-%\scc{again, is clearly not a rule but a huge number of paper refers to the mean as bias, could we use the word median, which is clear without any doubt rather than bias?}
-%\red{I also have code to calculate $\sigma_{MAD}$, should we include this as well?  It's almost always within a few thousandths of $\sigma_{IQR}$, so I left it out for now}.
-%\scc{we may just say that the two indicators reports almost the same value therefore there is no need for both of them?}
-For reference, the goals stated in Section 3.8 of the LSST Science Book \citep{Abell:09} for photo-$z$ performance in these metrics, assuming perfect training knowledge (as we are testing in this paper) are:
+We calculate the commonly used point estimate metrics of the overall intrinsic scatter, bias, and catastrophic outlier rate, defined in terms of the standard error $e_{z} \equiv (z_{PEAK} - z_{\rm true}) / (1 + z_{\rm true})$.
+Because the standard deviation of the \pz\ residuals is sensitive to outliers, we define the scatter in terms of the Interquartile Range (IQR), the difference between the 75th and 25th percentiles of the distribution of $e_{z}$, imposing the scaling $\sigma_{\rm IQR} = \rm{IQR} / 1.349$ to ensure that the area within $\sigma_{\rm IQR}$ is the same as that within one standard deviation from a standard Normal distribution.
+We also resist the effect of catastrophic outliers by defining the bias $b_{z}$ as the median rather than mean value of $e_{z}$.
+The catastrophic outlier rate $f_{\rm out}$ is defined as the fraction of galaxies with $e_{z}$ greater than $\max(3 \sigma_{\rm IQR}, 0.06)$.
+
+For reference, Section 3.8 of the \lsst\ Science Book \citep{Abell:09} uses the standard definitions of these parameters in requiring
 \begin{itemize}
-\item RMS scatter$ < 0.02(1+z)$
+\item RMS scatter $\sigma < 0.02 (1 + z_{\rm true})$
 %\scc{actually the SB reports in page 75 that the goal of 0.02 is for the RMS of $\frac{\sigma_{z}}{(1+z)}$ while we are using IQR}
-\item bias $<$ 0.003 %\scc{in page 518 of the SB is clearly reported that the mean is expected to be less than 0.003 in our case we have defined as bias the median and not the mean }
-\item catastrophic outlier rate $<$ 10\% %\scc{clearly the number of outliers since are defined above $3\sigma$ strictly depends on the sigma definition, in the science book it seems to refer to the RMS of the $e_z$ distribution even ef it could be the standard deviation, but again I don't think it refers to IQR, therefore we could not consider exactly this numbers as reference.}
+\item bias $b_{z} < 0.003$ %\scc{in page 518 of the SB is clearly reported that the mean is expected to be less than 0.003 in our case we have defined as bias the median and not the mean }
+\item catastrophic outlier rate $f_{\rm out} < 10\%$ %\scc{clearly the number of outliers since are defined above $3\sigma$ strictly depends on the sigma definition, in the science book it seems to refer to the RMS of the $e_z$ distribution even ef it could be the standard deviation, but again I don't think it refers to IQR, therefore we could not consider exactly this numbers as reference.}
 %\red{I am going off of conversations with Zeljko and how we {\it actually} computed the metrics in the Science Book (I have the scripts).
@@
 I agree that say Ivezic private communication could be better.}
+.
 \end{itemize}
-These definitions are similar, but not exactly the same, as the $\sigma_{IQR}$ and median bias calculated here, but are similar enough for qualitative comparisons to the LSST goals.
-
-%Scatter: SRD $\sigma<0.02(1+z)$
-%Catastrophic Outler Rate: SRD $3\sigma$ ``catastrophic outlier'' rate below 10 per cent
-%Bias: SRD bias$<0.003$
 
@@
 \label{sec:pointmetrics_results}
+
+Figure~\ref{fig:pz_pointestimates} shows both point estimates forall codes  both $z_{PEAK}$ and $z_{WEIGHT}$.
+Point density is shown with mixed contours to emphasize that most of the galaxies do fall close to the $z_{phot} = z_{spec}$ line, while points trace the details of the catastrophic outlier populations.
 
@@
 \includegraphics[width=0.49\textwidth]{fig/ZWEIGHT_szpz_threecolumn_12codes_navy_lowalpha.jpg}
-\caption{\Pz\ point estimates reduced from the \pzpdf s, via the mode ($z_{PEAK}$, left panel) and mean ($z_{WEIGHT}$, right panel).
-Point estimate density is represented with fixed density contours, while outliers at lower density are represented by blue points.
-While use of point-estimate photo-$z$'s is not recommended, they do make for useful comparative and visual diagnostics.
-In the lower-right panel of each plot, the \trainz\ estimator results in identical photo-$z$ estimates at the mode and mean of the training set $N'(z)$ distribution for all galaxies.}
+\caption{The density of \pz\ point estimates (contours) reduced from the \pzpdf s with outliers (blue) beyond the outlier cutoff (red dashed lines), via the mode ($z_{PEAK}$, left panel) and main-peak-mean ($z_{WEIGHT}$, right panel).
+The \trainz\ estimator (lower right sub-panels) has a shared $z_{PEAK}$ and $z_{WEIGHT}$ for the entire test set galaxy sample.}
 \label{fig:pz_pointestimates}
@@
 
-Figure~\ref{fig:pz_pointestimates} shows the point estimates for both $z_{PEAK}$ and $z_{WEIGHT}$.
-Point density is shown with mixed contours to emphasize that most of the galaxies do fall close to the $z_{phot}=z_{spec}$ line, while blue points show differing characteristics of the outlier populations.
-The red dashed lines indicated the cutoff for catastrophic outliers, defined as: $max(0.06,3\sigma_{IQR})$.
-As with the full $p(z)$ results, a variety of behaviours are evident in the different codes.  Table~\ref{tab:pointestimates} lists the scatter, bias, and catastrophic outlier fractions for the codes.
-The performance of the codes for point metrics is highly correlated with performance on $p(z)$ based tests, which is to be expected, given that the point-estimates were derived from the $p(z)$.
-Some discretization is evident in $z_{PEAK}$, particularly for \textsc{SkyNet}, due to the finite grid spacing of the reported $p(z)$.
-These discreteness effects are mitigated by the weighting of $z_{WEIGHT}$, resulting in a smoother distribution of redshift estimates.
-Several features perpendicular to the main $z_{phot}=z_{spec}$ line are evident.
-These features are due to the 4000 angstrom break passing through the gaps between adjacent LSST filters.
-These features are most prominent in template-based codes, but appear to some degree in all codes tested.
-
-In even the best performing codes, there are visible occupied regions away from the $z_{phot}=z_{spec}$ line, corresponding to degenerate redshift solutions for certain LSST magnitudes and colors.
-While use of the full information available via $p(z)$ mitigates their impact, a full understanding of the outlier population is critical for LSST science, particularly in tomographic applications %\red{I forget what exactly I was trying to say here, if someone else wants to take a crack at this it might be helpful --SJS}
+The finite grid spacing of the \pzpdf s induces some discretization in $z_{PEAK}$.%, particularly for \skynet.
+The features perpendicular to the $z_{phot} = z_{spec}$ line are due to the $4000 \rm{\AA}$ break passing through the gaps between adjacent filters.% and logically are more prominent for template-based codes
+Even the strongest codes feature populations far from the $z_{phot} = z_{spec}$ line representing a degeneracy in the space of colours and redshifts.
+% While use of the full information available via $p(z)$ mitigates their impact, a full understanding of the outlier population is critical for LSST science, particularly in tomographic applications.
+%\red{I forget what exactly I was trying to say here, if someone else wants to take a crack at this it might be helpful --SJS}
 %\scc{may be worth to say that the usage of further bands could mitigate this effect by removing the degeneration  usually induced from emission lines moving across different filters?}
@@
 
-Finally, we note that all twelve codes perform at or near the goals for point-estimates as outlined in the LSST Science Requirements Document\footnote{available at: \url{http://ls.st/srd}} and \citet{Graham:17}.
-This is to be expected, given that the requirements were designed such that a point estimate photo-z would meet these requirements for perfect training data to a depth of $i<25$.
-But, it is still an encouraging sign, given an updated mock galaxy simulation and the expanded set of photo-$z$ codes tested.
+The intrinsic scatter, bias, and catastrophic outlier rate are given in Table~\ref{tab:pointestimates}.
+Perhaps unsurprisingly, performance under these metrics largely tracks that of the metrics of Section~\ref{sec:metrics} of the \pzpdf s from which the point estimates were derived.
+All twelve codes perform at or near the goals of the \lsst\ Science Requirements Document\footnote{available at: \url{http://ls.st/srd}} and \citet{Graham:17}, which is encouraging if not unexpected for $i < 25$.
+% But, it is still an encouraging sign, given an updated mock galaxy simulation and the expanded set of photo-$z$ codes tested.
 %\red{maybe say consistent with expectations, and a reference to Melissa's paper in here?  What else to say about nearly meeting point goals?}

### 2019-03-30 paper/tex/introduction.tex
fix characterization of Mandelbaum08 and add CFHT Erben ref

@@
 
-Though the standard has long been for each galaxy in a photometric catalogue to have a \pz\ point estimate and Gaussian error bar, the nontrivial mapping between broad band fluxes and redshift renders this simplistic summary inadequate to quantify the uncertainty landscape by neglecting degenerate redshift solutions.
+Though the standard has long been for each galaxy in a photometric catalogue to have a \pz\ point estimate and Gaussian error bar; however, as use in precision cosmology measurements became more prevalent there was a realization that point-estimate photo-$z$'s were inadequte \citep{mandelbaum:2008}.
+The nontrivial mapping between broad band fluxes and redshift renders a simplistic point estimate inadequate to quantify the uncertainty landscape by neglecting degenerate redshift solutions.
 Far from a hypothetical situation, this degeneracy is a real consequence of the same deep imaging that enables larger galaxy catalogue sizes.
@@
 
-To fully characterize such physical degeneracies, photometric galaxy catalogue data releases from \citep{Mandelbaum:2008} to \citep{de_Jong:17}, provided a more informative \pz\ data product, the \pz\ probability density function (PDF), that describes the redshift probability, commonly denoted as $p(z)$, as a function of a galaxy's redshift, conditioned on the observed photometry.
+To fully characterize such physical degeneracies, photometric galaxy catalogue data releases, (e.~g.~\citet{Erben:2013}, \citet{de_Jong:17}), provided a more informative \pz\ data product, the \pz\ probability density function (PDF), that describes the redshift probability, commonly denoted as $p(z)$, as a function of a galaxy's redshift, conditioned on the observed photometry.
 Early template-based methods such as \citet{Fernandezsoto:99} approximated the likelihood of photometry conditioned on redshift with the relative \chisq\ values of template spectra.

### 2019-03-30 paper/tex/analysis.tex
fix i.e. and e.g. format

@@
 The PIT distribution has been used to quantify the performance of \pzpdf\ methods in the past \citep[e.~g.~][]{Bordoloi:10,Polsterer:16,Tanaka:17}.
-\citet{Tanaka:17} use the histogram of PIT values as a diagnostic indicator of overall code performance, while \citet{Freeman:17} independently define the PIT and demonstrate how its individual values may be used both to perform hypothesis testing (via, e.g., the KS, CvM, and AD tests; see below) and to construct quantile-quantile plots.
+\citet{Tanaka:17} use the histogram of PIT values as a diagnostic indicator of overall code performance, while \citet{Freeman:17} independently define the PIT and demonstrate how its individual values may be used both to perform hypothesis testing (via, e.~g.~ the KS, CvM, and AD tests; see below) and to construct quantile-quantile plots.
 %\red{We might want to divide this by redshift to get trends, as the overall PIT could be impacted by true n(z).

### 2019-03-30 paper/tex/discussion.tex
fix i.e. and e.g. format

@@
 The CDE loss and point estimate metrics appropriately penalize \trainz's naivete.
-As shown in Appendix~\ref{sec:pointmetrics}, \trainz ~has identical $ZPEAK$ and $ZWEIGHT$ values for every galaxy, and thus the \pz\ point estimats are constant as a function of true redshift, i.e.
-a horizontal line at the mode and mean of the training set distribution respectively.
+As shown in Appendix~\ref{sec:pointmetrics}, \trainz ~has identical $ZPEAK$ and $ZWEIGHT$ values for every galaxy, and thus the \pz\ point estimats are constant as a function of true redshift, i.~e.~a horizontal line at the mode and mean of the training set distribution respectively.
 The explicit dependence on the individual posteriors in the calculation of the CDE loss, described in Section~\ref{sec:cdelossresults}, distinguishes this metric from those of the \pzpdf\ ensemble and stacked estimator of the redshift distribution, despite their prevalence in the \pz\ literature.

### 2019-03-30 paper/tex/methods.tex
fixing formatting
fixed LaTeX-breaking typo
fix i.e. and e.g. format

@@
 \hline
-\trainz 	                             & machine learning	& See Section~\ref{sec:method:trainz} \\
+\trainz 	                             & machine learning	& See Section~\ref{sec:trainz} \\
 \end{tabular}}
@@
 To smooth the long tail in the distribution of magnitude errors, we use the log of the magnitude errors, improving numerical stability and eliminating the need for constraints on the optimization process.
-% We use principal component analysis (PCA) to decorrelate the data
+% We use principaphoto-\mathinhead{$z$}{z} interim posteriorsl component analysis (PCA) to decorrelate the data
 % WHAT IS BEING DECORRELATED HERE?
@@
 
-To run \tpz, we replaced nondetections with an approximation of the $1\sigma$ detection threshold based on the error forecast of the 10-year \lsst\ data, i.~e.
-$dm = 2.5 \log (1 + N/S)$ where $dm \sim 0.7526$ mag for $N/S = 1$.
+To run \tpz, we replaced nondetections with an approximation of the $1\sigma$ detection threshold based on the error forecast of the 10-year \lsst\ data, i.~e.~$dm = 2.5 \log (1 + N/S)$ where $dm \sim 0.7526$ mag for $N/S = 1$.
 We calibrated \tpz\ with the Out-of-Bag cross-validation technique \citep{Breiman:84,Carrasco_Kind:13} to evaluate its predictive validity and determine the relative importance of the different input attributes.
@@
 \subsection{trainZ: a pathological \pz\ estimator}
-\label{sec:method:trainz}
+\label{sec:trainz}
 

### 2019-03-30 paper/tex/methods.tex
address more Troxel comments, add interp=2 to bpz

@@
 
-We test three publicly available and commonly used template-based codes that share the standard physically motivated approach of calculating model fluxes for a set of template SEDs on a grid of redshift values and evaluating a \chisq\ merit function using the observed and model fluxes.
+We test three publicly available and commonly used template-based codes that share the standard physically motivated approach of calculating model fluxes for a set of template SEDs on a grid of redshift values and evaluating a \chisq\ merit function using the observed and model fluxes:
+\begin{equation} \label{eq_temp_chi}
+\chi^2(z,T,A) = \sum_i^{N_{\rm filt}}\left(\frac{F^i_{\rm{obs}} - A \, F^i_{\rm{pred}}(T,z)}{\sigma^i_{\rm{obs}}}\right)^2
+\end{equation}
+\noindent  where $A$ is a normalization factor, $F^i_{\rm{pred}}(T,z)$ is the flux predicted for a template $T$ at redshift $z$.
+$F^i_{\rm{obs}}$ is the observed flux in a given band $i$ and $\sigma^i_{\rm{obs}}$ is the observed flux error.
+$N_{\rm filt}$ is the total number of filters, in our case the six $ugrizy$ LSST filters.
+Specific implementation details of each code, e.~g.~prior form and implementation, are described below.
 
@@
 
-Photometric Analysis for Redshift Estimate \citep[\lephare\footnote{\url{http://www.cfht.hawaii.edu/~arnouts/lephare.html}},][]{Arnouts:99,Ilbert:06} matches observed colors with those predicted from a template set, which can be semi-empirical or entirely synthetic, directly according to the likelihood $\chi^{2}(z, T, A) \equiv \sum_{b}^{N_{\rm{filt}}} \left((F^{\rm{obs}}_{b} - A F^{\rm{mod}}_{b}(T, z)) / \sigma^{\rm{obs}}_{b}\right)^{2}$ of normalization factor $A$, template $T$, and redshift $z$.
+Photometric Analysis for Redshift Estimate \citep[\lephare\footnote{\url{http://www.cfht.hawaii.edu/~arnouts/lephare.html}},][]{Arnouts:99,Ilbert:06} matches observed colors with those predicted from a template set, which can be semi-empirical or entirely synthetic, directly according to the $\chi^2$ form given in Equation~\ref{eq_temp_chi}.
 In words, the likelihood is a sum of observed flux error $\sigma_{b}^{\rm{obs}}$-weighted squared differences between the observed flux $F^{\rm{obs}}_{b}$ and the normalized predicted flux $F^{\rm{mod}}_{b}(T, z)$ in $N_{\rm{filt}}$ photometric filters $b$, which is the \lsst\ $ugrizy$ filters in this case.
@@
 The \bpz\ likelihood is related to the \chisq\ likelihood by $p(C \vert z, T) \propto \exp[- \chi^{2} / 2]$.
-Given a Bayesian prior $p(z, T \vert m_{0})$ over apparent magnitude $m_0$ and assuming that the SED templates are spanning and exclusive, \bpz\ constructs the redshift posterior $p(z \vert C, m_0)$ by marginalizing over all SED templates as in \citep[Eq.~3 from][]{Benitez:00}, corresponding to setting the parameter \texttt{PROBS\_LITE=TRUE} in the \bpz\ parameter file.
+Given a Bayesian prior $p(z, T \vert m_{0})$ over apparent magnitude $m_0$ and type $T$, and assuming that the SED templates are spanning and exclusive, \bpz\ constructs the redshift posterior $p(z \vert C, m_0)$ by marginalizing over all SED templates as in \citep[Eq.~3 from][]{Benitez:00}, corresponding to setting the parameter \texttt{PROBS\_LITE=TRUE} in the \bpz\ parameter file.
 The \bpz\ prior is the product of an SED template proportion that varies with apparent magnitude $p(T \vert m_{0})$ and a prior $p(z \vert T, m_{0})$ over the expected redshift as a function of apparent magnitude and SED template.
@@
 The Bayesian prior term $p(T \vert m_{0})$ was derived directly from the DC1 training set, and the other term $p(z \vert T, m_{0})$ was chosen to be the best fit for the eleven free parameters of the functional form of \citet{Benitez:00}.
+We use template interpolation, creating two linearly interpolated templates between each basis SED (sorted by rest-frame $u-g$ colour) by setting the parameter \texttt{INTERP=2}.
 %For photo-$z$ point estimates we use the \texttt{Z\_B} output parameter.

### 2019-04-10 paper/main.tex
address more Troxel comments, add interp=2 to bpz
added additional funding sourc

--- 2019-03-08 paper/main.tex
@@
 
-\title{Evaluating photometric redshift PDF approaches for \textsc{LSST}}
+\title{Implicit assumptions and their impact on photometric redshift PDF performance for LSST}
+% \title{Photometric redshift uncertainties for LSST \romannumeral{1}: A unified methodology for assessing photo-$z$ PDF performance}
 
@@
 
-Many scientific investigations of photometric galaxy surveys require redshift estimates, whose uncertainty properties are best encapsulated by photometric redshift (photo-$z$) posterior probability distribution functions (PDFs).
-A plethora of photo-$z$ PDF estimation methodologies abound, producing discrepant results with no consensus on a preferred approach.
-We present the results of a comprehensive experiment comparing twelve photo-$z$ algorithms on mock data produced for the Large Synoptic Survey Telescope (\textsc{LSST}) Dark Energy Science Collaboration (\textsc{DESC}).
-By supplying perfect prior information, in the form of the complete template library and a representative training set as inputs to each code, we demonstrate the impact of the assumptions underlying each technique on the output photo-$z$ PDFs.
-In the absence of a notion of true, unbiased photo-$z$ PDFs, we evaluate and interpret multiple metrics of the ensemble properties of the derived photo-$z$ PDFs as well as traditional reductions to photo-$z$ point estimates.
-We report systematic biases and overall over/under-breadth of the photo-$z$ PDFs of many popular codes, which may indicate avenues for improvement in the algorithms or implementations.
-Furthermore, we raise attention to the limitations of established metrics for assessing photo-$z$ PDF accuracy; though we identify the conditional density estimate (CDE) loss as a promising metric of photo-$z$ PDF performance in the case where true redshifts are available but true photo-$z$ PDFs are not, we emphasize the need for science-specific performance metrics.
+In order to maximize scientific returns of current and upcoming galaxy surveys, the photometric redshift (photo-$z$) posterior distributions produced by redshift estimation codes must be accurate probability distribution functions (PDFs).
+However, the posteriors resulting from a number of current techniques are not, in general, consistent with each other, affected by implicit assumptions made by each code, and an optimal method for obtaining an accurate PDF estimate remains unclear.
+We present the results of an initial study by the Large Synoptic Survey Telescope Dark Energy Science Collaboration (\textsc{LSST-DESC}) testing twelve photo-$z$ algorithms using complete and representative training data and evaluate multiple metrics to test how accurately the posteriors represent probability distributions.
+We observe several trends, including systematic biases and an overall over/under-prediction in the broadness of the PDFs in many of the codes which may be symptomatic of implementation problems or problems in underlying algorithm design.
+A careful accounting of all photo-$z$ systematics will be necessary for the codes employed in upcoming analyses in order to achieve unbiased cosmological measurements.
+
+%  Photometric redshift (photo-$z$) probability distribution functions (PDFs) are a key data product of nearly all upcoming galaxy imaging surveys.
+However, the photo-$z$ PDFs resulting from different techniques are not in general consistent with one another, and an optimal method for obtaining an accurate PDF remains unclear.
+We present the results of an initial study of the Large Synoptic Survey Telescope Dark Energy Science Collaboration (\textsc{LSST-DESC}), the first in a planned series of papers testing multiple photo-$z$ codes on simulations of upcoming LSST galaxy photometry catalogues.
+This initial test evaluates photo-$z$ algorithms in the presence of representative training data and in the absence of several common sources of systematic errors that affect the procedures by which photo-$z$ PDFs are derived.
+%%  The photo-$z$ PDFs are evaluated using multiple metrics including the Kolmogorov-Smirnoff statistic, Cramer-von Mises statistic, Anderson-Darling statistic, Kullback-Leibler divergence, $N(z)$ moments, quantile-quantile plots and probability integral transform.
+%  The photo-$z$ PDFs are evaluated using multiple metrics and we observe several trends, including an overall over/under-prediction in the broadness of the PDFs for several of the codes.
+A careful accounting of all systematics discovered will be necessary for the codes employed in upcoming analyses in order to achieve unbiased cosmological measurements.
+ %% Photometric redshift (photo-$z$) probability distribution functions (PDFs) more completely capture the complex nature of photo-$z$ uncertainties and are thus a key data product of nearly all upcoming galaxy imaging surveys, including the Large Synoptic Survey Telescope (LSST).
+  % Photo-$z$ PDFs produced by different algorithms do not in general agree with one another, and no one technique has yet been identified as optimal, leading to a proliferation of photo-$z$ PDF estimation codes.
+  % We present a unified comparison of twelve such codes using simulated LSST photometry, the first in a series of papers on photo-$z$ PDF performance for the LSST Dark Energy Science Collaboration (DESC).
+  % In this initial study, we isolate the effects of the estimation procedure by running each code using perfectly representative training data (for machine learning methods) or the true SED template library (for template-fitting methods) and confirm the mutual inconsistency of the resulting photo-$z$ PDF estimates.
+  % We apply a comprehensive suite of metrics to an the population of photo-$z$ PDFs as well as point estimates derived thereof and an accepted estimator of the redshift distribution $N(z)$.
+  % % The photo-$z$ PDFs are evaluated using multiple metrics including the Kolmogorov-Smirnoff statistic, Cramer-von Mises statistic, Anderson-Darling statistic, Kullback-Leibler divergence, $N(z)$ moments, quantile-quantile plots, and probability integral transform.
+  % We interpret the performance of the codes under each metric in terms of the assumptions underlying the overall approach of the methods and audit the sentivity of the metrics to the failure modes of different techniques, culminating in a discussion predicting the impact of introducing realistic systematics and the direction of attention to the choice of metric when identifying the most appropriate code for a given science use case.
 
@@
 { \red{personal funding sources}}
-S.~Schmidt acknowledges support from DOE grant DE-SC0009999 and NSF/AURA grant N56981C.  AIM is advised by David W. Hogg and was supported by National Science Foundation grant AST-1517237.
+S.~Schmidt acknowledges support from DOE grant DE-SC0009999 and NSF/AURA grant N56981C.
+AIM is advised by David W. Hogg and was supported by National Science Foundation grant AST-1517237.
+AIM was also supported by the U.S. Department of Energy, Office of 
+Science, Office of Workforce Development for Teachers and Scientists, Office of 
+Science Graduate Student Research (SCGSR) program, administered by the Oak 
+Ridge Institute for Science and Education for the DOE under contract number 
+DE‐SC0014664.
 

### 2019-05-27 paper/tex/discussion.tex
add bootstrap error lines to fig 3 and a little discussion

@@
 We remind the reader that contributed codes were given a goal of obtaining accurate \pzpdf s, not an accurate stacked estimator of the redshift distribution, so we do not expect the same codes to necessarily perform well for both classes of metrics.
-Indeed, the codes were optimized for their interpretation of our request for ``accurate \pzpdf s,'' and we expect that the implementations would have been adjusted had we requested optimization of the traditional metrics of Appendices~\ref{sec:moments, sec:pointmetrics}.
+Indeed, the codes were optimized for their interpretation of our request for ``accurate \pzpdf s,'' and we expect that the implementations would have been adjusted had we requested optimization of the traditional metrics of Appendices~\ref{sec:moments} and~\ref{sec:pointmetrics}.
 

### 2019-05-27 paper/tex/results.tex
add bootstrap error lines to fig 3 and a little discussion

@@
 The settings used on \gpz\ in this work forced broadening of the single Gaussian to cover the multimodal redshift solutions of the other codes.
-%Interestingly, while \textsc{ANNz2} shows an abundance of small scale structure in individual $p(z)$ measurements (see Fig.~\ref{fig:pz_examples}), the summed $\hat{N}(z)$ is rather smooth, where the small scale features average out.
-This is not the case for the two other codes that show an abundance of substructure in their individual $p(z)$: both \textsc{CMNN} and \textsc{SkyNet} show small scale features both in $p(z)$ and $\hat{N}(z)$.
-%In contrast, FlexZBoost, for example, can return estimates on any grid without compression errors as it’s a basis expansion method where only the expansion coefficients need to be stored.
-%Codes with a native output format other than the shared piecewise constant binning scheme (or one that can be losslessly converted to it) may suffer from loss of information when converting to it, which could artificially favor some codes over others in a limited number of cases, for example bright galaxies with very narrow $p(z)$ where the true peak falls between grid points.
-We will discuss PDF storage in Section~\ref{sec:discussion}.
-%Furthermore, the fidelity of photo-$z$ interim posteriors in this format varies with the quality of the photometry.
-%Switching to a quantile based parameterization may be more costly computationally, for example template-based codes would need to test more grid point in order to resolve the quantiles for bright galaxies.
-However, the computational time for template based codes scales roughly linearly with the number of grid points, so this may be a reasonable thing to implement.
-% We will discuss this further in Section~\ref{sec:discussion}.
-% \red{someone review this statement to make sure that I'm saying this correctly!}
 
@@
 The high QQ values (more high than low PIT values) of \bpz, \cmnn, \delight, \eazy, and \gpz\ indicate \pzpdf s biased low, and the low QQ values (more low than high PIT values) of \skynet\ and \tpz\ indicate \pzpdf s biased high.
+The gray shaded band marks the $2\sigma$ variance in PIT values found using the \trainz\ algorithm with a bootstrap resampling of the training set and a sample size of 30,000 galaxies.
+This represents a very conservative estimate of the representative training sample size, and thus an approximate minimal error significance compared to ideal performance.
+Deviations in the PIT histograms outside of this range show that significant biases are present for some codes.
 
@@
 \centering
-\includegraphics[width=0.74\textwidth]{fig/PITANDQQplot_12codes_crop.jpg}
+\includegraphics[width=0.74\textwidth]{fig/PITANDQQplot_12codes_withsigmaband_crop.jpg}
 \caption{The QQ plot (red) and PIT histogram (blue) of the \pzpdf\ codes (panels) along with the ideal QQ (black dashed diagonal) and ideal PIT (gray horizontal) curves, as well as a difference plot for the QQ difference from the ideal diagonal (lower inset).
-The twelve codes exhibit varying degrees of four deviations from perfection: an overabundance of PIT values at the centre of the distribution indicate a catalogue of overly broad \pzpdf s, an excess of PIT values at the extrema indicates a catalogue of overly narrow \pzpdf s, catastrophic outliers manifest as overabundances at PIT values of 0 and 1, and asymmetry indicates systematic bias, a form of model misspecification.}
+The gray shaded region indicates the $2 \sigma$ range from a bootstrap resampling of the training set with a size of 30,000 galaxies using \trainz.
+The twelve codes exhibit varying degrees of four deviations from perfection: an overabundance of PIT values at the centre of the distribution indicate a catalogue of overly broad \pzpdf s, an excess of PIT values at the extrema indicates a catalogue of overly narrow \pzpdf s, catastrophic outliers manifest as overabundances at PIT values of 0 and 1, and asymmetry indicates systematic bias, a form of model misspecification.
+Values in excess of the $2\sigma$ shaded region show that for some codes these errors will be significant given expected training sample sizes.}
 \label{fig:pitqq}
@@
 
-The PIT histograms of \delight, \cmnn, \skynet, and \tpz\ feature an underrepresentation of extreme values, indicative of overly broad \pzpdf s, while the overrepresentation of extreme values for for \metaphor indicate overly narrow \pzpdf s.
+The PIT histograms of \delight, \cmnn, \skynet, and \tpz\ feature an underrepresentation of extreme values, indicative of overly broad \pzpdf s, while the overrepresentation of extreme values for \metaphor\ indicate overly narrow \pzpdf s.
 These five codes in particular have a free parameter for bandwidth, which may be responsible for this vulnerability, in spite of the opportunity for fine-tuning with perfect prior information.
@@
 
-Though the spikes in the first and last bin of the PIT histogram were cut off in Figure~\ref{fig:pitqq} for visualization, the catastrophic outlier rates are provided in Table~\ref{tab:pitoutlier}.
+Close inspection of the extremes at PIT values of 0 and 1 reveal spikes in the first and last bin of the PIT histogram for some codes in Figure~\ref{fig:pitqq}, corresponding to ``catastrophic outliers'' where the true redshift lies outside of the support of the $p(z)$.
+The catastrophic outlier rates are provided in Table~\ref{tab:pitoutlier}.
 As expected, \trainz\ achieves precisely the 0.0002 value expected of an ideal PIT distribution.
@@
 \centering
-\includegraphics[width=0.74\textwidth]{fig/KSvsCvMvsAD_PIT_withnull_jpg.jpg}
+\includegraphics[width=0.74\textwidth]{fig/KSvsCvMvsAD_PIT_withbootstrap.jpg}
 \caption{A visualization of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the PIT distributions.
-There is generally good agreement between these statistics, with differences corresponding to the codes with outstanding catastrophic outlier rates, a reflection in the differences in how each statistic weights the tails of the distribution.}
+There is generally good agreement between these statistics, with differences corresponding to the codes with outstanding catastrophic outlier rates, a reflection in the differences in how each statistic weights the tails of the distribution.
+Horizontal lines indicate the level of uncertainty found by bootstrapping a training set sample of 30,000 galaxies using \trainz, none of the codes reach this conservative ideal floor in expected uncertainty.}
 \label{fig:pit_stats}
@@
 
-Figure~\ref{fig:pit_stats} displays the values of the KS, CvM, and AD test statistics between the PIT distribution and a uniform distribution $U(0, 1)$, highlighting the relative rather than absolute numbers.
-%\red{Can p-values be supplied for each statistic?
-The statistics themselves are difficult to interpret, other than ``lower is better'' (p-value in skgof was broken, having trouble finding 1-sample KS calculation for uniform distribution)}
+Figure~\ref{fig:pit_stats} displays the values of the KS, CvM, and AD test statistics calculated by comparing the PIT distribution and a uniform distribution $U(0, 1)$.
+This plot highlights the relative rather than absolute numbers.
 \metaphor\ and \lephare\ perform well under the AD but poorly under the KS and CvM due to their high catastrophic outlier rates.
@@
 \annz's strong performance can be attributed to an aspect of the training process in which training set galaxies with a PIT that more closely matches the percentiles of the DC1 training set's redshift distribution are upweighted; in effect, these quantile-based metrics were part of the algorithm itself that may or may not serve it well under more realistic experimental conditions.
+Similar to what was done for the PIT histograms in Figure~\ref{fig:pitqq} we create bootstrap training samples of 30,000 galaxies for use with \trainz\ in order to estimate a conservative statistical floor that we would expect in real data.
+No code reaches this idealized floor, indicating that all codes suffer some degradation from the ideal when employing their implicit priors, though \annz, \flexzboost, and \gpz\ are within a factor of two.
 
@@
 This metric is the only one that can appropriately penalize \trainz\ and indicates strong performance for \cmnn\ and \flexzboost, the latter of which is optimized for this metric.
-%Empirically, we have found that PIT RMSE is not as closely correlated to CDE loss as it is to the $N(z)$ statistics.
-
-%\subsection{Response of Individual Codes}
-%\label{sec:res:pz_indiv_codes}
-%\red{this may be incorporated in to 5.1 and 5.2 rather than live in its own section!}
-
-%\red{HOW CODES deal with negative fluxes, and magnitude uncertainties}
-
-%\red{overall conclusions of the Results section}

### 2019-05-28 paper/tex/appendix.tex
add lines to NZ KSCVMAD fig, add Rongpu to author list and move ext collab to end

@@
 Figure~\ref{fig:nz_stats} shows the quantitative Kolmogorov-Smirnoff (KS), Cramer-Von Mises (CvM), and Anderson Darling (AD) test statistics for each of the codes for the $\hat{N}(z)$ based measures.
-%\red{Can p-values be supplied for each statistic?
-The statistics themselves are difficult to interpret, other than ``lower is better'' (no, p-values are very difficult to compute for non-uniform distributions)}
-The stacked estimators of the redshift distribution for \cmnn\ and \trainz\ best estimate $\tilde{N}(z)$ under these metrics, and the only codes that do especially poorly are \eazy, \lephare, \metaphor, and \skynet.
+The horizontal lines show the the result of a bootstrap resampling of the training set using 30,000 samples for \trainz.
+As mentioned in Section~\ref{sec:pitqq} this represents a conservative idealized limit on expected performance for a modest-sized representative training set of galaxies.
+The AD bootstrap statistic is elevated due to this statistic's sensitivity to distribution tails.
+The stacked estimators of the redshift distribution for \cmnn\ and \trainz\ best estimate $\tilde{N}(z)$ under these metrics, and \bpz, \gpz, and \tpz\ are within a factor of two of the conservative limit for all statistics.
+The only codes that do especially poorly are \eazy, \lephare, \metaphor, and \skynet.
 It is unsurprising that \cmnn\ scores well, as with a near perfectly representative training set means that choosing neighbouring points in color/magnitude space should lead to excellent agreement in the final $\hat{N}(z)$ estimate.
@@
 \centering
-\includegraphics[width=0.74\textwidth]{fig/KSvsCvMvsAD_NZ_withnull_jpg.jpg}
+\includegraphics[width=0.74\textwidth]{fig/KSvsCvMvsAD_NZ_withbootstrap.jpg}
 \caption{A visualization of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the $\hat{N}(z)$ distributions.
+Horizontal lines indicate the statistic values (including uncertainty) achieved using \trainz\ via bootstrap resampling a training set containing 30,000 redshifts.
 We make the reassuring observation that these related statistics do not disagree significantly with one another.
-\cmnn\ performs comparably well to \trainz, the control case, while \skynet\ scores poorly due to an overall bias in its redshift predictions.}
+\cmnn\ actually outperforms the control case, \trainz, and several codes are within a factor of two of this conservative idealized limit.
+\skynet\ scores poorly due to an overall bias in its redshift predictions.}
 \label{fig:nz_stats}

### 2019-08-22 paper/tex/analysis.tex
address most remaining comments

@@
 
-\subsection{Metrics of an ensemble of \pzpdf s}
+\subsection{Metrics of \pzpdf ensembles}
 \label{sec:qualmet}
@@
 A quantile of a distribution is the value $q$ at which the CDF of the distribution is equal to $Q$; percentiles and quartiles are familiar examples of linearly spaced sets of 100 and 4 quantiles, respectively.
-The quantile-quantile (QQ) plot serves as a graphical visualization for comparing two distributions, where the quantiles of one distribution are plotted against the quantiles of the other distribution, providing an easy way to qualitatively assess the consistency between an estimating distribution and a true distribution.
+The quantile-quantile (QQ) plot serves as a graphical visualization for comparing two distributions, where the quantiles of one distribution are plotted against the quantiles of the other distribution, providing an easy way to qualitatively assess the consistency between an estimated distribution and a true distribution.
 The closer the QQ plot is to diagonal, the closer the match between the distributions.

### 2019-08-22 paper/tex/appendix.tex
address most remaining comments

@@
 is the most widely accepted method for obtaining $\hat{N}^{H}(z)$ as an estimator of the redshift distribution from \pzpdf s derived by a method $H$.
-Though we do not endorse the use of the stacked estimator of the redshift distribution, we use it under the untested assumption that the response of our metrics of $\hat{N}^{H}(z)$ will be analogous to the same metrics applied to a principled estimator of the redshift distribution.
+Though the use of the stacked estimator of the redshift distribution is not formally correct, we use it under the untested assumption that the response of our metrics of $\hat{N}^{H}(z)$ will be analogous to the same metrics applied to a principled estimator of the redshift distribution.
 
@@
 
-\annz, \gpz, and \metaphor\ show signs of overtraining, estimating enhanced peaks and diminished troughs relative to the training set, an obstacle that may be overcome with adjustment of the implementation.
+\annz, \gpz, and \metaphor\ show possible signs of overtraining: showing exaggerated peaks and troughs relative to the training set.
+Further investigation on overtraining is needed, if present this is an obstacle that may be overcome with adjustment of the implementation.
 
@@
 The only codes that do especially poorly are \eazy, \lephare, \metaphor, and \skynet.
-It is unsurprising that \cmnn\ scores well, as with a near perfectly representative training set means that choosing neighbouring points in color/magnitude space should lead to excellent agreement in the final $\hat{N}(z)$ estimate.
+It is unsurprising that \cmnn\ scores well, as with a nearly complete and representative training set choosing neighbouring points in color/magnitude space to construct an estimator should lead to excellent agreement in the final $\hat{N}(z)$.
 

### 2019-08-22 paper/tex/data.tex
moving information from LePhare
address most remaining comments

@@
 
-Our mock catalogue is derived from the \textsc{Buzzard}-highres-v1.0  of De Rose et al., in prep, Wechsler et al., in prep) catalogue.
-% UPDATE THIS CITATION NOW THAT THE PAPER IS OUT
+Our mock catalogue is derived from the \textsc{Buzzard}-highres-v1.0 catalogue  \citep[][Wechsler et al., in prep]{DeRose:19}.
 \textsc{Buzzard} is built on a dark matter-only N-body simulation of $2048^{3}$ particles in a $400$ Mpc h$^{-1}$ box.
@@
 \texttt{ADDSEDS} uses a sample of $\sim 5\times 10^{5}$ galaxies from the magnitude-limited \sdss\ Data Release 6 Value Added Galaxy Catalogue \citep{Blanton:05} to train an empirical relation between absolute $r$-band magnitude, local galaxy density, and SED.
-Each \sdss\ spectrum is parameterized by five weights corresponding to a weighted sum of five basis SED components using the \texttt{k-correct \red{v4\_3?}} software package\footnote{\url{http://kcorrect.org}} \citep{Blanton:07}.
+Each \sdss\ spectrum is parameterized by five weights corresponding to a weighted sum of five basis SED components using the \texttt{k-correct} software package\footnote{\url{http://kcorrect.org}} \citep{Blanton:07}.
 
@@
 %The resulting catalogues contain $111\,171$ training galaxies and $1\,000\,883$ test galaxies.
+The \lsst\ photometric filter transmission curves were also considered public information that could be used by any code.
 

### 2019-08-22 paper/tex/methods.tex
address most remaining comments

@@
 Here we summarize the twelve \pzpdf\ codes compared in this study, summarized in Table~\ref{tab:list_of_codes}, which include both established and emerging approaches in template fitting and machine learning.
-Though not exhaustive, this sample represents codes for which there was sufficient expertise within the \lsstdesc\ Photometric Redshifts Working Group; the authors welcome interest from those outside \lsstdesc\ to have their codes assessed in future investigations that build upon this one.
+Though not exhaustive, this sample represents codes for which there was sufficient expertise within the \lsstdesc\ Photometric Redshifts Working Group.
+Some aspects of data treatment were left to the individual code runners, for example, whether/how to split the available data with known redshifts into separate training and validation sets.
+Another key difference is the treatment of non-detections in one or more bands: some codes choose to ignore a band, others replace the value with either an estimate for the detection limit, the mean of other values in the training set, or another default value.
+There are varying conventions among machine learning based codes for treatment of non-detections, and no one prescription dominates in the photo-z literature.
+The specific choices for each code affect the results, and contribute to the implicit prior influencing their output.
+However, we remind the reader that only 2.0 per cent of our sample has non-detections, almost exclusively in the u-band, and thus should not dominate the code performance differences.
+The authors welcome interest from those outside \lsstdesc\ to have their codes assessed in future investigations that build upon this one.
 
@@
 
-\subsection{Training-based Approaches}
+\subsection{Machine Learning-based Approaches}
 \label{sec:trainingcodes}
@@
 Here, we make two modifications to the implementation of \citet{Graham:17} to comply with the controlled experimental conditions.
-First, we do not impose nondetections on galaxies fainter than the expected \lsst\ 10-year limiting magnitude or bright enough to saturate with \lsst's CCDs, instead using all of the photometry for the DC1 test and training sets.
+First, we do not impose non-detections on galaxies fainter than the expected \lsst\ 10-year limiting magnitude or bright enough to saturate with \lsst's CCDs, instead using all of the photometry for the DC1 test and training sets.
 Second, we apply the initial colour cut to the training set before calculating the Mahalanobis distance in order to accelerate processing and use a magnitude pseudo-prior as in \citet{Graham:17}, but for both we use cut-off values corresponding to the DC1 training set galaxies' colours and magnitudes.
@@
 This is the optimal expected value of the unobserved variables given the observed ones under the assumption that the distribution is jointly Gaussian; note that this reduces to a simple average if the covariates are independent with $\Sigma_{\rm uo} = 0$.
-We reserved for validation 20\% of the training set and used the Variable Covariance option in \textsc{GPz} with 200 basis functions, neglecting to apply cost-sensitive learning options.
+We reserved for validation 20\% of the training set and used the Variable Covariance option in \textsc{GPz} with 200 basis functions (see \citet{Almosallam:15b} for details), neglecting to apply cost-sensitive learning options.
 
@@
 
-In this work, we used a hierarchical KNN to replace nondetections with values based on their neighbors.
-The usual cross-validation step was also omitted for this study.
+In this work, we used a hierarchical KNN to replace non-detections with values based on their neighbors.
+The usual cross-validation of redshift estimates and PDFs was also omitted for this study.
 
@@
 
-To run \tpz, we replaced nondetections with an approximation of the $1\sigma$ detection threshold based on the error forecast of the 10-year \lsst\ data, i.~e.~$dm = 2.5 \log (1 + N/S)$ where $dm \sim 0.7526$ mag for $N/S = 1$.
+To run \tpz, we replaced non-detections with an approximation of the $1\sigma$ detection threshold based on the error forecast of the 10-year \lsst\ data, i.~e.~$dm = 2.5 \log (1 + N/S)$ where $dm \sim 0.7526$ magnitudes for $N/S = 1$.
 We calibrated \tpz\ with the Out-of-Bag cross-validation technique \citep{Breiman:84,Carrasco_Kind:13} to evaluate its predictive validity and determine the relative importance of the different input attributes.

### 2019-08-22 paper/tex/methods.tex
fix eqn and awkward phrase

@@
 The \bpz\ likelihood is related to the \chisq\ likelihood by $p(C \vert z, T) \propto \exp[- \chi^{2} / 2]$.
-Given a Bayesian prior $p(z, T \vert m_{0})$ over apparent magnitude $m_0$ and type $T$, and assuming that the SED templates are spanning and exclusive, \bpz\ constructs the redshift posterior $p(z \vert C, m_0)$ by marginalizing over all SED templates as in \citep[Eq.~3 from][]{Benitez:00}, corresponding to setting the parameter \texttt{PROBS\_LITE=TRUE} in the \bpz\ parameter file.
+Given a Bayesian prior $p(z, T \vert m_{0})$ over apparent magnitude $m_0$ and type $T$, and assuming that the SED templates are spanning and exclusive, \bpz\ constructs the redshift posterior $p(z \vert C, m_0)$ by marginalizing over all SED templates with the form $p(z|C,m_0)\,\propto\, \sum_{T}p(C|z,T)\,p(z,T|m_0)$
+ \citep[Eq.~3 from][]{Benitez:00}, corresponding to setting the parameter \texttt{PROBS\_LITE=TRUE} in the \bpz\ parameter file.
 The \bpz\ prior is the product of an SED template proportion that varies with apparent magnitude $p(T \vert m_{0})$ and a prior $p(z \vert T, m_{0})$ over the expected redshift as a function of apparent magnitude and SED template.

### 2019-08-22 paper/tex/introduction.tex
address more Troxel comments, add interp=2 to bpz
fix very long sentence in intro

@@
 
-There are numerous techniques for deriving \pzpdf s, yet no one method has yet been established as clearly superior.
+There are numerous techniques for deriving \pzpdf s, yet no one method has been established as clearly superior.
 Quantitative comparisons of \pz\ methods have been made before.
@@
 
-This paper is distinguished by its focus on the evaluation criteria for \pzpdf s and interpretation thereof, as a key project of the Photometric Redshifts working group of the \lsstdesc\ that aims to perform a comprehensive sensitivity analysis of \pzpdf\ techniques in order to ultimately select those that will become part of the \lsst\ pipelines, as laid out in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}}.
+This paper is distinguished by its focus on the evaluation criteria for \pzpdf s and interpretation thereof.
+This is a key project of the Photometric Redshifts working group of the \lsstdesc\ that aims to perform a comprehensive sensitivity analysis of \pzpdf\ techniques in order to ultimately select those that will become part of the \lsst\ pipelines.
+These plans are laid out in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}}.
 In this initial study, we focus on evaluating the performance of \pzpdf\ codes using PDF-specific performance metrics in a controlled experiment with complete and representative prior information (template libraries and training sets) to set a baseline for subsequent investigations.

### 2019-08-23 paper/tex/analysis.tex
add some discussion on individual PDF

@@
 There are a number of metrics that can be used to test the accuracy of a \pz\ posterior as an estimator of a true \pz\ posterior if the true \pzpdf\ is known.
-However, the true \pzpdf\ is in general not accessible unless the photometry is in fact drawn from the true \pz\ likelihoods, a mock catalogue generation procedure that has not yet appeared in the literature and was not performed for this work.
+However, the true \pzpdf\ is in general not accessible unless the photometry is in fact drawn from the true \pz\ likelihoods, which is not how most existing datasets have been constructed.
+Instead, data is often taken from actual observations, or generated from mock catalogs derived from N-body simulations of the Universe (as is the case in this study).
+In these cases there is no ``true PDF'' for an individual object, and most measures of PDF fidelity will need to rely on ensemble properties of the similar objects, or the entire sample.
+We mention a potential project addressing this issue in \S\,\ref{sec:futureexperiments}.
+%a mock catalogue generation procedure that has not yet appeared in the literature and was not performed for this work.
 

### 2019-08-23 paper/tex/discussion.tex
add some discussion on individual PDF

@@
 In addition to outright redshift failures we will model the inclusion of a small number of high-confidence yet false redshifts due to emission line misidentification or noise spikes.
+Section~\ref{sec:metrics} discussed the difficulty in evaluating PDF accuracy for individual objects.
+We are planning a project to generate ``true PDF'' distributions using generative adversarial networks\citep[(GANs)][]{Goodfellow:14} trained on the simulation data.
+This dataset will enable a test of PDF accuracy for individual galaxies rather than resorting to ensembles, as required for many tests in the current work.
 

### 2019-08-23 paper/tex/introduction.tex
add some discussion on individual PDF

@@
 There are numerous techniques for deriving \pzpdf s, yet no one method has been established as clearly superior.
-Quantitative comparisons of \pz\ methods have been made before.
-The \Pz\ Accuracy And Testing \citep[\textsc{PHAT},][]{Hildebrandt:10} effort focused on \pz\ point estimates derived from many photometric bands.
+In fact, in the absence of simulated data drawn from known redshift distributions, the very concept of a ``true PDF'' for an individual galaxy is somewhat ill-defined, and we must instead rely on measures of ensemble behaviour to characterize PDF quality (see \S\,\ref{sec:metrics} for further discussion).
+This caveat aside, quantitative comparisons of \pz\ methods have been made before; the \Pz\ Accuracy And Testing \citep[\textsc{PHAT},][]{Hildebrandt:10} effort focused on \pz\ point estimates derived from many photometric bands.
 \citet{Rau:2015} introduced a new method for improving \pzpdf s using an ordinal classification algorithm.

### 2019-08-26 paper/tex/analysis.tex
few more words on individual PDFs

@@
 However, the true \pzpdf\ is in general not accessible unless the photometry is in fact drawn from the true \pz\ likelihoods, which is not how most existing datasets have been constructed.
-Instead, data is often taken from actual observations, or generated from mock catalogs derived from N-body simulations of the Universe (as is the case in this study).
-In these cases there is no ``true PDF'' for an individual object, and most measures of PDF fidelity will need to rely on ensemble properties of the similar objects, or the entire sample.
-We mention a potential project addressing this issue in \S\,\ref{sec:futureexperiments}.
+Instead, data is often taken from empirical observations, or generated from mock catalogs derived from N-body simulations of the Universe (as is the case in this study).
+In these cases there is no ``true PDF'' for an individual object given its observed properties, as for real data the true redshift distribution is unkown, and for simulations the photometric data is usually generated by empirical or semi-analytic relations tied to the underlying dark matter distribution and local environment.
+As a consequence, most measures of PDF fidelity will need to rely on ensemble properties of the similar objects, or the entire sample.
+We do discuss a metric that can measure performance of individual PDF quality, the conditional density estimate loss, in \S\,\ref{sec:CDE_loss}.
+We also mention a potential future project addressing this issue in \S\,\ref{sec:futureexperiments}.
 %a mock catalogue generation procedure that has not yet appeared in the literature and was not performed for this work.
@@
 
-\subsection{Metrics of \pzpdf ensembles}
+\subsection{Metrics of \pzpdf \ ensembles}
 \label{sec:qualmet}

### 2019-08-26 paper/tex/discussion.tex
few more words on individual PDFs

@@
 
-Furthermore, our metrics are not necessarily able to assess the fidelity of individual \pzpdf s relative to true posteriors.
-Metric-specific performance implies that we may need multiple \pzpdf\ approaches tuned to each metric in order to maximize returns over all science cases in large upcoming surveys.
+Furthermore, our metrics are not necessarily able to assess the fidelity of individual \pzpdf s relative to true posteriors: in the absence of a ``true PDF'' from which redshifts are drawn, it is difficult to construct metrics to measure performance for individual galaxies rather than ensembles, with the exception of the CDE Loss metric of section~\ref{sec:CDE_loss}, where knowledge of the true densities is not required in computing the Loss function for an individual galaxy.
+A lack of appropriate metrics more sophisticated than the CDE Loss remains an open issue for science cases requiring accurate individual galaxy PDFs.
+The metric-specific performance demonstrated in this paper implies that we may need multiple \pzpdf\ approaches tuned to each metric in order to maximize returns over all science cases in large upcoming surveys.
 

### 2019-08-27 paper/tex/methods.tex
add annz2 detail

@@
 In addition to accounting for errors on the input photometry, \annz\ uses the KNN-uncertainty estimate of \citet{Oyaizu:08} to quantify uncertainty in the choice of method over multiple runs.
-Using the Toolkit for Multivariate Data Analysis with ROOT\footnote{\url{http://tmva.sourceforge.net/}}, it can return the results of running a single machine learning algorithm, a ``best'' choice of the results from simultaneously running multiple algorithms, or a combination of the results of multiple algorithms weighted by their method uncertainties averaged over multiple runs.
+Using the Toolkit for Multivariate Data Analysis with ROOT\footnote{\url{http://tmva.sourceforge.net/}}, it can return the results of running a single machine learning algorithm, a ``best'' choice of the results from simultaneously running multiple algorithms (based on evaluation the cumulative distribution functions of validation set objects), or a combination of the results of multiple algorithms weighted by their method uncertainties averaged over multiple runs.
 %\textsc{ANNz2} is capable of producing both photo-$z$ point estimates and redshift posterior probability distributions $p(z)$.

### 2019-09-04 paper/tex/analysis.tex
fixed minor grammar errors, added affiliation

--- 2019-08-22 paper/tex/analysis.tex
@@
 The goal of this study is to evaluate the degree to which \pzpdf s of each method can be trusted for a generic analysis.
-The overloaded ``$p(z)$'' is a widespread abuse of notation that obfuscates this goal, so we will dedicate some attention to breaking it apart.
+The overloaded ``$p(z)$'' is a widespread abuse of notation that obfuscates this goal, so we dedicate attention to dismantling it here.
 Galaxies have redshifts $z$ and photometric data $d$ drawn from a joint probability space $p(z, d)$ in nature.
@@
 There are a number of metrics that can be used to test the accuracy of a \pz\ posterior as an estimator of a true \pz\ posterior if the true \pzpdf\ is known.
-However, the true \pzpdf\ is in general not accessible unless the photometry is in fact drawn from the true \pz\ likelihoods, a mock catalogue generation procedure that has not yet appeared in the literature and was not performed for this work.
+However, the true \pzpdf\ is in general not accessible unless the photometry is in fact drawn from a ground truth for the joint probability density of redshift and photometry $p(z, d)$.
+In contrast, existing mock catalogs produce redshift-photometry pairs $(z, d)$ by a deterministic algorithm that does not correspond to a joint probability density from which one can take samples.
+In these cases there is no ``true PDF'' for an individual object, and most measures of PDF fidelity will necessarily be restricted to probing the quality of the ensemble of \pzpdf s.
+(See \S,\ref{sec:futureexperiments} for a discussion of how one might circumvent this limitation.)
 
@@
 A quantile of a distribution is the value $q$ at which the CDF of the distribution is equal to $Q$; percentiles and quartiles are familiar examples of linearly spaced sets of 100 and 4 quantiles, respectively.
-The quantile-quantile (QQ) plot serves as a graphical visualization for comparing two distributions, where the quantiles of one distribution are plotted against the quantiles of the other distribution, providing an easy way to qualitatively assess the consistency between an estimated distribution and a true distribution.
+The quantile-quantile (QQ) plot serves as a graphical visualization for comparing two distributions, where the quantiles of one distribution are plotted against the quantiles of the other distribution, providing an intuitive way to qualitatively assess the consistency between an estimated distribution and a true distribution.
 The closer the QQ plot is to diagonal, the closer the match between the distributions.

### 2019-09-04 paper/tex/appendix.tex
fixed minor grammar errors, added affiliation

@@
 
-\annz, \gpz, and \metaphor\ show possible signs of overtraining: showing exaggerated peaks and troughs relative to the training set.
+\annz, \gpz, and \metaphor\ feature exaggerated peaks and troughs relative to the training set, a potential sign of overtraining.
 Further investigation on overtraining is needed, if present this is an obstacle that may be overcome with adjustment of the implementation.
@@
 Figure~\ref{fig:nz_stats} shows the quantitative Kolmogorov-Smirnoff (KS), Cramer-Von Mises (CvM), and Anderson Darling (AD) test statistics for each of the codes for the $\hat{N}(z)$ based measures.
-The horizontal lines show the the result of a bootstrap resampling of the training set using 30,000 samples for \trainz.
-As mentioned in Section~\ref{sec:pitqq} this represents a conservative idealized limit on expected performance for a modest-sized representative training set of galaxies.
-The AD bootstrap statistic is elevated due to this statistic's sensitivity to distribution tails.
-The stacked estimators of the redshift distribution for \cmnn\ and \trainz\ best estimate $\tilde{N}(z)$ under these metrics, and \bpz, \gpz, and \tpz\ are within a factor of two of the conservative limit for all statistics.
-The only codes that do especially poorly are \eazy, \lephare, \metaphor, and \skynet.
+The horizontal lines show the the result of a bootstrap resampling of the training set using 30,000 samples for \trainz, representing a conservative idealized limit on expected performance for a modest-sized representative training set of galaxies, as mentioned in Section~\ref{sec:pitqq}.
+The AD bootstrap statistic is elevated due to its sensitivity to the tails of distributions.
+The stacked estimators of the redshift distribution for \cmnn\ and \trainz\ best estimate $\tilde{N}(z)$ under these metrics, whereas \eazy, \lephare, \metaphor, and \skynet\ underperform; \bpz, \gpz, and \tpz\ are within a factor of two of the conservative limit for all statistics.
 It is unsurprising that \cmnn\ scores well, as with a nearly complete and representative training set choosing neighbouring points in color/magnitude space to construct an estimator should lead to excellent agreement in the final $\hat{N}(z)$.
@@
 We make the reassuring observation that these related statistics do not disagree significantly with one another.
-\cmnn\ actually outperforms the control case, \trainz, and several codes are within a factor of two of this conservative idealized limit.
+\cmnn\ outperforms the control case, \trainz, and several codes are within a factor of two of this conservative idealized limit.
 \skynet\ scores poorly due to an overall bias in its redshift predictions.}

### 2019-09-04 paper/tex/discussion.tex
fixed minor grammar errors, added affiliation

--- 2019-05-27 paper/tex/discussion.tex
@@
 
-Furthermore, our metrics are not necessarily able to assess the fidelity of individual \pzpdf s relative to true posteriors.
-Metric-specific performance implies that we may need multiple \pzpdf\ approaches tuned to each metric in order to maximize returns over all science cases in large upcoming surveys.
+Furthermore, our metrics are not necessarily able to assess the fidelity of individual \pzpdf s relative to true posteriors: in the absence of a ``true PDF'' from which redshifts are drawn, it is difficult to construct metrics to measure performance for individual galaxies rather than ensembles.
+(The CDE Loss metric of section~\ref{sec:CDE_loss} is an exception to this rule.)
+A lack of appropriate metrics more sophisticated than the CDE Loss remains an open issue for science cases requiring accurate individual galaxy PDFs.
+The metric-specific performance demonstrated in this paper implies that we may need multiple \pzpdf\ approaches tuned to each metric in order to maximize returns over all science cases in large upcoming surveys.
 
@@
 A top priority for a follow-up study is to test realistic forms of incomplete, erroneous, and non-representative template libraries and training sets as well as the impact of other forms of external priors that must be ingested by the codes, major concerns in \citet{Newman:2015, Masters:2017}.
+Outright redshift failures due to emission line misidentification or noise spikes may be modeled by the inclusion of a small number of high-confidence yet false redshifts.
 We plan to perform a full sensitivity analysis on a realistically incomplete training set of spectroscopic galaxies, modeling the performance of spectrographs, emission-line properties, and expected signal-to-noise to determine which potential training set galaxies are most likely to be excluded.
-In addition to outright redshift failures we will model the inclusion of a small number of high-confidence yet false redshifts due to emission line misidentification or noise spikes.
 
@@
 In light of the results of \citet[]{Malz:qp}, in future analyses we plan to switch from a fixed grid to the quantile parameterization or to permit each code to use its native storage format under a shared number of parameters.
+
+Section~\ref{sec:metrics} discussed the difficulty in evaluating PDF accuracy for individual objects.
+In a follow-up study, we will
+ % use generative adversarial networks\citep[(GANs)][]{Goodfellow:14} trained on the simulation data to
+generate ``true PDF'' distributions, yielding a dataset that enables a test of PDF accuracy for individual galaxies rather than solely ensembles.
 

### 2019-09-04 paper/tex/introduction.tex
fixed minor grammar errors, added affiliation

@@
 
-Though the standard has long been for each galaxy in a photometric catalogue to have a \pz\ point estimate and Gaussian error bar; however, as use in precision cosmology measurements became more prevalent there was a realization that point-estimate photo-$z$'s were inadequte \citep{mandelbaum:2008}.
-The nontrivial mapping between broad band fluxes and redshift renders a simplistic point estimate inadequate to quantify the uncertainty landscape by neglecting degenerate redshift solutions.
+Though the standard has long been for each galaxy in a photometric catalogue to have a \pz\ point estimate and Gaussian error bar, even early applications of \pz s in precision cosmology indicate the inadequacy of point estimates \citep{mandelbaum:2008} to encapsulate the degeneracies resulting from the nontrivial mapping between broad band fluxes and redshift.
 Far from a hypothetical situation, this degeneracy is a real consequence of the same deep imaging that enables larger galaxy catalogue sizes.
@@
 
-To fully characterize such physical degeneracies, photometric galaxy catalogue data releases, (e.~g.~\citet{Erben:2013}, \citet{de_Jong:17}), provided a more informative \pz\ data product, the \pz\ probability density function (PDF), that describes the redshift probability, commonly denoted as $p(z)$, as a function of a galaxy's redshift, conditioned on the observed photometry.
+To fully characterize such physical degeneracies, later photometric galaxy catalogue data releases, (e.~g.~\citet{Erben:2013}, \citet{de_Jong:17}), provide a more informative \pz\ data product, the \pz\ probability density function (PDF), that describes the redshift probability, commonly denoted as $p(z)$, as a function of a galaxy's redshift, conditioned on the observed photometry.
 Early template-based methods such as \citet{Fernandezsoto:99} approximated the likelihood of photometry conditioned on redshift with the relative \chisq\ values of template spectra.
@@
 There are numerous techniques for deriving \pzpdf s, yet no one method has been established as clearly superior.
-In fact, in the absence of simulated data drawn from known redshift distributions, the very concept of a ``true PDF'' for an individual galaxy is somewhat ill-defined, and we must instead rely on measures of ensemble behaviour to characterize PDF quality (see \S\,\ref{sec:metrics} for further discussion).
+In fact, in the absence of simulated data drawn from known redshift distributions, the very concept of a ``true PDF'' for an individual galaxy is unavailable, and we must instead rely on measures of ensemble behaviour to characterize PDF quality (see \S\,\ref{sec:metrics} for further discussion).
 This caveat aside, quantitative comparisons of \pz\ methods have been made before; the \Pz\ Accuracy And Testing \citep[\textsc{PHAT},][]{Hildebrandt:10} effort focused on \pz\ point estimates derived from many photometric bands.
@@
 
-This paper is distinguished by its focus on the evaluation criteria for \pzpdf s and interpretation thereof.
-This is a key project of the Photometric Redshifts working group of the \lsstdesc\ that aims to perform a comprehensive sensitivity analysis of \pzpdf\ techniques in order to ultimately select those that will become part of the \lsst\ pipelines.
-These plans are laid out in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}}.
+This paper is distinguished from other comparisons of \pz\ methods by its focus on the evaluation criteria for \pzpdf s and interpretation thereof.
+We aim to perform a comprehensive sensitivity analysis of \pzpdf\ techniques in order to ultimately select those that will become part of the \lsst\ pipelines, as part of a key project of the Photometric Redshifts working group of the \lsstdesc, described in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}}.
 In this initial study, we focus on evaluating the performance of \pzpdf\ codes using PDF-specific performance metrics in a controlled experiment with complete and representative prior information (template libraries and training sets) to set a baseline for subsequent investigations.

### 2019-09-04 paper/tex/methods.tex
fixed minor grammar errors, added affiliation

--- 2019-08-22 paper/tex/methods.tex
@@
 
-Here we summarize the twelve \pzpdf\ codes compared in this study, summarized in Table~\ref{tab:list_of_codes}, which include both established and emerging approaches in template fitting and machine learning.
+Here we summarize the twelve \pzpdf\ codes compared in this study, also in Table~\ref{tab:list_of_codes}, which include both established and emerging approaches in template fitting and machine learning.
 Though not exhaustive, this sample represents codes for which there was sufficient expertise within the \lsstdesc\ Photometric Redshifts Working Group.
 Some aspects of data treatment were left to the individual code runners, for example, whether/how to split the available data with known redshifts into separate training and validation sets.
-Another key difference is the treatment of non-detections in one or more bands: some codes choose to ignore a band, others replace the value with either an estimate for the detection limit, the mean of other values in the training set, or another default value.
-There are varying conventions among machine learning based codes for treatment of non-detections, and no one prescription dominates in the photo-z literature.
-The specific choices for each code affect the results, and contribute to the implicit prior influencing their output.
-However, we remind the reader that only 2.0 per cent of our sample has non-detections, almost exclusively in the u-band, and thus should not dominate the code performance differences.
+
+Another key difference is the treatment of non-detections in one or more bands: some codes ignore incomplete bands, while others replace the value with either an estimate for the detection limit, the mean of other values in the training set, or another default value.
+There are varying conventions among machine learning based codes for treatment of non-detections, and no one prescription dominates in the \pz\ literature.
+The specific choices for each code affect the results and contribute to the implicit prior influencing their output.
+However, we remind the reader that only 2.0 per cent of our sample has non-detections, almost exclusively in the $u$-band, and thus should not dominate the code performance differences.
 The authors welcome interest from those outside \lsstdesc\ to have their codes assessed in future investigations that build upon this one.
@@
 \end{equation}
-\noindent  where $A$ is a normalization factor, $F^i_{\rm{pred}}(T,z)$ is the flux predicted for a template $T$ at redshift $z$.
-$F^i_{\rm{obs}}$ is the observed flux in a given band $i$ and $\sigma^i_{\rm{obs}}$ is the observed flux error.
-$N_{\rm filt}$ is the total number of filters, in our case the six $ugrizy$ LSST filters.
+\noindent where $A$ is a normalization factor, $F^i_{\rm{pred}}(T,z)$ is the flux predicted for a template $T$ at redshift $z$, $F^i_{\rm{obs}}$ is the observed flux in a given band $i$, $\sigma^i_{\rm{obs}}$ is the observed flux error, and $N_{\rm filt}$ is the total number of filters, in our case the six $ugrizy$ LSST filters.
+In words, the likelihood is a sum of observed flux error $\sigma_{b}^{\rm{obs}}$-weighted squared differences between the observed flux $F^{\rm{obs}}_{b}$ and the normalized predicted flux $F^{\rm{mod}}_{b}(T, z)$ in $N_{\rm{filt}}$ photometric filters $b$, which is the \lsst\ $ugrizy$ filters in this case.
 Specific implementation details of each code, e.~g.~prior form and implementation, are described below.
@@
 
-Photometric Analysis for Redshift Estimate \citep[\lephare\footnote{\url{http://www.cfht.hawaii.edu/~arnouts/lephare.html}},][]{Arnouts:99,Ilbert:06} matches observed colors with those predicted from a template set, which can be semi-empirical or entirely synthetic, directly according to the $\chi^2$ form given in Equation~\ref{eq_temp_chi}.
-In words, the likelihood is a sum of observed flux error $\sigma_{b}^{\rm{obs}}$-weighted squared differences between the observed flux $F^{\rm{obs}}_{b}$ and the normalized predicted flux $F^{\rm{mod}}_{b}(T, z)$ in $N_{\rm{filt}}$ photometric filters $b$, which is the \lsst\ $ugrizy$ filters in this case.
+Photometric Analysis for Redshift Estimate \citep[\lephare\footnote{\url{http://www.cfht.hawaii.edu/~arnouts/lephare.html}},][]{Arnouts:99,Ilbert:06} uses the $\chi^2$ of Equation~\ref{eq_temp_chi} to match observed colors with those predicted from a template set, which can be semi-empirical or entirely synthetic, directly according to the
 The reported \pzpdf\ is an arbitrary normalization of the likelihood evaluated on the output redshift grid.
@@
 The \bpz\ likelihood is related to the \chisq\ likelihood by $p(C \vert z, T) \propto \exp[- \chi^{2} / 2]$.
-Given a Bayesian prior $p(z, T \vert m_{0})$ over apparent magnitude $m_0$ and type $T$, and assuming that the SED templates are spanning and exclusive, \bpz\ constructs the redshift posterior $p(z \vert C, m_0)$ by marginalizing over all SED templates as in \citep[Eq.~3 from][]{Benitez:00}, corresponding to setting the parameter \texttt{PROBS\_LITE=TRUE} in the \bpz\ parameter file.
+Given a Bayesian prior $p(z, T \vert m_{0})$ over apparent magnitude $m_0$ and type $T$, and assuming that the SED templates are spanning and exclusive, \bpz\ constructs the redshift posterior $p(z \vert C, m_0)$ by marginalizing over all SED templates with the form $p(z|C,m_0)\,\propto\, \sum_{T}p(C|z,T)\,p(z,T|m_0)$ \citep[Eq.~3 from][]{Benitez:00}, corresponding to setting the parameter \texttt{PROBS\_LITE=TRUE} in the \bpz\ parameter file.
 The \bpz\ prior is the product of an SED template proportion that varies with apparent magnitude $p(T \vert m_{0})$ and a prior $p(z \vert T, m_{0})$ over the expected redshift as a function of apparent magnitude and SED template.
@@
 Here we test \bpz-v 1.99.3 with the DC1 template set of Section~\ref{sec:buzztemplates}.
-To keep the number of free parameters manageable, the DC1 template set is pre-sorted by the rest-frame $u-g$ colour and split into three broad classes of SED template, equivalent to the E, Sp and Im/SB types in .
+To keep the number of free parameters manageable, the DC1 template set is pre-sorted by the rest-frame $u-g$ colour and split into three broad classes of SED template, equivalent to the E, Sp and Im/SB types.
 The Bayesian prior term $p(T \vert m_{0})$ was derived directly from the DC1 training set, and the other term $p(z \vert T, m_{0})$ was chosen to be the best fit for the eleven free parameters of the functional form of \citet{Benitez:00}.
@@
 In addition to accounting for errors on the input photometry, \annz\ uses the KNN-uncertainty estimate of \citet{Oyaizu:08} to quantify uncertainty in the choice of method over multiple runs.
-Using the Toolkit for Multivariate Data Analysis with ROOT\footnote{\url{http://tmva.sourceforge.net/}}, it can return the results of running a single machine learning algorithm, a ``best'' choice of the results from simultaneously running multiple algorithms, or a combination of the results of multiple algorithms weighted by their method uncertainties averaged over multiple runs.
+Using the Toolkit for Multivariate Data Analysis with ROOT\footnote{\url{http://tmva.sourceforge.net/}}, \annz\ can return the results of running a single machine learning algorithm, a ``best'' choice of the results from simultaneously running multiple algorithms (based on evaluation the cumulative distribution functions of validation set objects), or a combination of the results of multiple algorithms weighted by their method uncertainties averaged over multiple runs.
 %\textsc{ANNz2} is capable of producing both photo-$z$ point estimates and redshift posterior probability distributions $p(z)$.
@@
 Here, we make two modifications to the implementation of \citet{Graham:17} to comply with the controlled experimental conditions.
-First, we do not impose non-detections on galaxies fainter than the expected \lsst\ 10-year limiting magnitude or bright enough to saturate with \lsst's CCDs, instead using all of the photometry for the DC1 test and training sets.
+First, we do not impose non-detections on galaxies fainter than the expected \lsst\ 10-year limiting magnitude nor galaxies bright enough to saturate with \lsst's CCDs, instead using all of the photometry for the DC1 test and training sets.
 Second, we apply the initial colour cut to the training set before calculating the Mahalanobis distance in order to accelerate processing and use a magnitude pseudo-prior as in \citet{Graham:17}, but for both we use cut-off values corresponding to the DC1 training set galaxies' colours and magnitudes.
@@
 This is the optimal expected value of the unobserved variables given the observed ones under the assumption that the distribution is jointly Gaussian; note that this reduces to a simple average if the covariates are independent with $\Sigma_{\rm uo} = 0$.
-We reserved for validation 20\% of the training set and used the Variable Covariance option in \textsc{GPz} with 200 basis functions (see \citet{Almosallam:15b} for details), neglecting to apply cost-sensitive learning options.
+We reserved for validation 20\% of the training set and used the Variable Covariance option in \gpz\ with 200 basis functions (see \citet{Almosallam:15b} for details), neglecting to apply cost-sensitive learning options.
 

### 2019-09-04 paper/tex/results.tex
fixed minor grammar errors, added affiliation

@@
 The histogram of PIT values, QQ plot, and QQ difference plot relative to the ideal diagonal are provided in Figure~\ref{fig:pitqq}, showcasing the biases and trends in the average accuracy of the \pzpdf s for each code.
-The high QQ values (more high than low PIT values) of \bpz, \cmnn, \delight, \eazy, and \gpz\ indicate \pzpdf s biased low, and the low QQ values (more low than high PIT values) of \skynet\ and \tpz\ indicate \pzpdf s biased high.
-The gray shaded band marks the $2\sigma$ variance in PIT values found using the \trainz\ algorithm with a bootstrap resampling of the training set and a sample size of 30,000 galaxies.
-This represents a very conservative estimate of the representative training sample size, and thus an approximate minimal error significance compared to ideal performance.
+The high QQ values (i.~e.~ more high than low PIT values) of \bpz, \cmnn, \delight, \eazy, and \gpz\ indicate \pzpdf s biased low, and the low QQ values (more low than high PIT values) of \skynet\ and \tpz\ indicate \pzpdf s biased high.
+The gray shaded band marks the $2\sigma$ variance in PIT values found using the \trainz\ algorithm with a bootstrap resampling of the training set and a sample size of 30,000 galaxies, representing a very conservative estimate of the representative training sample size, and thus an approximate minimal error significance compared to ideal performance.
 Deviations in the PIT histograms outside of this range show that significant biases are present for some codes.
@@
 
-Close inspection of the extremes at PIT values of 0 and 1 reveal spikes in the first and last bin of the PIT histogram for some codes in Figure~\ref{fig:pitqq}, corresponding to ``catastrophic outliers'' where the true redshift lies outside of the support of the $p(z)$.
+Close inspection of the extremes at PIT values of 0 and 1 reveal spikes in the first and last bin of the PIT histogram for some codes in Figure~\ref{fig:pitqq}, corresponding to catastrophic outliers where the true redshift lies outside of the support of the $p(z)$.
 The catastrophic outlier rates are provided in Table~\ref{tab:pitoutlier}.
@@
 There is generally good agreement between these statistics, with differences corresponding to the codes with outstanding catastrophic outlier rates, a reflection in the differences in how each statistic weights the tails of the distribution.
-Horizontal lines indicate the level of uncertainty found by bootstrapping a training set sample of 30,000 galaxies using \trainz, none of the codes reach this conservative ideal floor in expected uncertainty.}
+Horizontal lines indicate the level of uncertainty found by bootstrapping a training set sample of 30,000 galaxies using \trainz; none of the codes reach this conservative ideal floor in expected uncertainty.}
 \label{fig:pit_stats}
@@
 
-Figure~\ref{fig:pit_stats} displays the values of the KS, CvM, and AD test statistics calculated by comparing the PIT distribution and a uniform distribution $U(0, 1)$.
-This plot highlights the relative rather than absolute numbers.
+Figure~\ref{fig:pit_stats} highlights the relative values of the KS, CvM, and AD test statistics calculated by comparing the PIT distribution and a uniform distribution $U(0, 1)$.
 \metaphor\ and \lephare\ perform well under the AD but poorly under the KS and CvM due to their high catastrophic outlier rates.
@@
 \annz's strong performance can be attributed to an aspect of the training process in which training set galaxies with a PIT that more closely matches the percentiles of the DC1 training set's redshift distribution are upweighted; in effect, these quantile-based metrics were part of the algorithm itself that may or may not serve it well under more realistic experimental conditions.
-Similar to what was done for the PIT histograms in Figure~\ref{fig:pitqq} we create bootstrap training samples of 30,000 galaxies for use with \trainz\ in order to estimate a conservative statistical floor that we would expect in real data.
+Similar to what was done for the PIT histograms in Figure~\ref{fig:pitqq}, we create bootstrap training samples of 30,000 galaxies for use with \trainz\ in order to estimate a conservative statistical floor that we would expect in real data.
 No code reaches this idealized floor, indicating that all codes suffer some degradation from the ideal when employing their implicit priors, though \annz, \flexzboost, and \gpz\ are within a factor of two.

### 2019-09-04 paper/main.tex
add bootstrap error lines to fig 3 and a little discussion
addressing final comments from PR

--- 2019-03-15 paper/main.tex
@@
 \newcommand{\annz}{\pzcode{ANNz2}}
-\newcommand{\bpz}{\pzcode{BPz}}
+\newcommand{\bpz}{\pzcode{BPZ}}
 \newcommand{\cmnn}{\pzcode{CMNN}}
@@
 % ======================================================================
-
+%\hypersetup{draft}
 \begin{document}
@@
 A plethora of photo-$z$ PDF estimation methodologies abound, producing discrepant results with no consensus on a preferred approach.
-We present the results of a comprehensive experiment comparing twelve photo-$z$ algorithms on mock data produced for the Large Synoptic Survey Telescope (\textsc{LSST}) Dark Energy Science Collaboration (\textsc{DESC}).
+We present the results of a comprehensive experiment comparing twelve photo-$z$ algorithms applied to mock data produced for the Large Synoptic Survey Telescope (\textsc{LSST}) Dark Energy Science Collaboration (\textsc{DESC}).
 By supplying perfect prior information, in the form of the complete template library and a representative training set as inputs to each code, we demonstrate the impact of the assumptions underlying each technique on the output photo-$z$ PDFs.
@@
 S.~Schmidt acknowledges support from DOE grant DE-SC0009999 and NSF/AURA grant N56981C.
-AIM is advised by David W. Hogg and was supported by National Science Foundation grant AST-1517237.
+AIM acknowledges support from the Max Planck Society and the Alexander von Humboldt Foundation in the framework of the Max Planck-Humboldt Research Award endowed by the Federal Ministry of Education and Research.
+During the completion of this work, AIM was advised by David W. Hogg and was supported by National Science Foundation grant AST-1517237.
 

### 2019-09-04 paper/main.tex
addressing comments from @sschmidt23

@@
 \usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{amsfonts}
@@
 
-In addition to packages cited in the text, analyses performed in this paper used the following software packages: \code{Numpy} and \code{Scipy} \citep{numpyscipy}, \code{Matplotlib} \citep{matplotlib}, \code{Seaborn} \citep{seaborn}, \code{minFunc} \citep{minfunc}, \code{pySkyNet} \citep{pyskynet}, and \code{photUtils} from the \lsst\ simulations package \citep{lsstphotutils}.
+In addition to packages cited in the text, analyses performed in this paper used the following software packages: \code{Numpy} and \code{Scipy} \citep{numpyscipy}, \code{Matplotlib} \citep{matplotlib}, \code{Seaborn} \citep{seaborn}, \code{minFunc} \citep{minfunc}, \code{qp} \citep{Malz:qp, Malz:2018}, \code{pySkyNet} \citep{pyskynet}, and \code{photUtils} from the \lsst\ simulations package \citep{lsstphotutils}.
 

### 2019-09-04 paper/tex/analysis.tex
addressing comments from @sschmidt23

@@
 
-In this study, we isolate the effect of differences in prior information $I_{H}$ specific to each method by using a single training set $I_{D}^{\rm ML}$ for all machine learning-based codes and a single template library $I_{D}^{\rm T}$ for all template-based codes.
-These sets of prior information are carefully constructed to be representative and complete, so we have $I_{D} \equiv I_{D}^{\rm ML} \equiv I_{D}^{\rm T}$ for every method $H$.
+In this study, we isolate the effect of differences in prior information $I_{H}$ specific to each method by using a single training set $I_{D}^{\mathrm{ML}}$ for all machine learning-based codes and a single template library $I_{D}^{\mathrm{T}}$ for all template-based codes.
+These sets of prior information are carefully constructed to be representative and complete, so we have $I_{D} \equiv I_{D}^{\mathrm{ML}} \equiv I_{D}^{\mathrm{T}}$ for every method $H$.
 Under this assumption, a ratio of posteriors of codes is in effect a ratio of the implicit posteriors $p(z \vert d_{i}, I_{H'})$ since the external prior information $I_{D}$ is present in the numerator and denominator.
@@
 
-The exact implementation of the metrics theoretically depends on the parametrization of the \pzpdf s, which may differ across codes and can affect the precision of the estimator \citep{Malz:qp}.
+The exact implementation of the metrics theoretically depends on the parametrization of the \pzpdf s, which may differ across codes and can affect the precision of the estimator \citep{Malz:2018}.
 Even considering a single method under the same parametrization, such as the 200-bin $0 < z < 2$ piecewise constant function used here, the exact bin definitions will affect the result.
@@
   \label{eq:cdf}
-  \rm{CDF}[f, q] \equiv \int_{-\infty}^{q} f(z) dz,
+  \mathrm{CDF}[f, q] \equiv \int_{-\infty}^{q} f(z) dz,
 \end{equation}
@@
 \label{eq:pit}
-\mathrm{PIT} &\equiv \rm{CDF}[\hat{p}, z_{\rm true}]
+\mathrm{PIT} &\equiv \mathrm{CDF}[\hat{p}, z_{\mathrm true}]
 \end{align}
@@
 %For example, a PIT with nontrivial shape could be due to that behavior being uniform as a function of redshift or significant inaccuracy only in a small range of redshifts where $n(z)$ is large and a flat PIT elsewhere.}
-Following Kodra \& Newman (in prep.) we define the PIT-based catastrophic outlier rate as the fraction of galaxies with $\rm{PIT} < 0.0001$ or $\rm{PIT} > 0.9999$, which should total 0.0002 for an ideal uniform distribution.
+Following Kodra \& Newman (in prep.) we define the PIT-based catastrophic outlier rate as the fraction of galaxies with $\mathrm{PIT} < 0.0001$ or $\mathrm{PIT} > 0.9999$, which should total 0.0002 for an ideal uniform distribution.
 
@@
   \label{eq:ks}
-  \rm{KS} \equiv \max_{z} \left( \left| \rm{CDF}[\hat{f}, z] - \rm{CDF}[\tilde{f}, z] \right| \right),
+  \mathrm{KS} \equiv \max_{z} \left( \left| \mathrm{CDF}[\hat{f}, z] - \mathrm{CDF}[\tilde{f}, z] \right| \right),
 \end{equation}
-interpretable as the maximum difference between the CDFs of an approximating univariate distribution $\hat{f}(z)$ and a reference distribution $\tilde{f}(z)$.
+interpretable as the maximum difference between the CDFs of an approximating univariate distribution $\hat{f}(z)$ and a reference distribution $\tilde{f}(z)$, in this case $U(0,1)$.
 % $CDF_{\rm phot}(z)$ and $CDF_{\rm spec}(z)$, the CDFs of the photo-$z$ and spectroscopic redshift respectively:
@@
 \label{eq:cvm}
-  \rm{CvM}^{2} \equiv \int_{-\infty}^{+\infty} \big(\rm{CDF}[\hat{f}, z] - \rm{CDF}[\tilde{f}, z]\big)^2 \mathrm{d}\rm{CDF}[\tilde{f}, z]
+  \mathrm{CvM}^{2} \equiv \int_{-\infty}^{+\infty} \big(\mathrm{CDF}[\hat{f}, z] - \mathrm{CDF}[\tilde{f}, z]\big)^2 \mathrm{d}\mathrm{CDF}[\tilde{f}, z]
 \end{equation}
@@
 \begin{equation} \label{eq:ad}
-  \rm{AD}^2 \equiv N_{tot}\int_{-\infty}^{+\infty} \frac{\big(\rm{CDF}[\hat{f}, z] - \rm{CDF}[\tilde{f}, z]\big)^2} {\rm{CDF}[\tilde{f}, z] (1 -\rm{CDF}[\tilde{f}, z]} \mathrm{d}\rm{CDF}[\tilde{f}, z]
+  \mathrm{AD}^2 \equiv N_{tot}\int_{-\infty}^{+\infty} \frac{\big(\mathrm{CDF}[\hat{f}, z] - \mathrm{CDF}[\tilde{f}, z]\big)^2} {\mathrm{CDF}[\tilde{f}, z] (1 -\mathrm{CDF}[\tilde{f}, z]} \mathrm{d}\mathrm{CDF}[\tilde{f}, z]
 \end{equation}
@@
 
-The \buzz\ simulation process precludes testing the degree to which samples from our \pz\ posteriors reconstruct the space of $p(z, {\rm data})$.
+The \buzz\ simulation process precludes testing the degree to which samples from our \pz\ posteriors reconstruct the space of $p(z, \mathrm{data})$.
 To the knowledge of the authors, there is only one metric that can be used to evaluate the performance of individual \pzpdf\ estimators in the absence of true \pz\ posteriors.
@@
   \label{eq:cde-loss}
-  L(f, \widehat{f}) \equiv \int \int (f(z \vert \x) - \widehat{f}(z \vert \x))^{2} dz dP(\x)
+  L(f, \widehat{f}) \equiv \int \int (f(z \vert \x) - \widehat{f}(z \vert \x))^{2} \mathrm{d}z \mathrm{d}P(\x)
 \end{equation}

### 2019-09-04 paper/tex/appendix.tex
addressing comments from @sschmidt23

@@
 Many of the codes, including all the model-fitting approaches and \annz, \gpz, \metaphor, and \skynet\ from the data-driven camp, overestimate the redshift density at $z \sim 1.4$.
-This behavior is a consequence of the $4000\ {\rm \AA}$ break passing through the gap between the $z$ and $y$ filters, which induces a genuine discontinuity in the $z - y$ colour as a function of redshift that can sway the \pzpdf\ estimates in the absence of bluer spectral features.
+This behavior is a consequence of the $4000$ \AA break passing through the gap between the $z$ and $y$ filters, which induces a genuine discontinuity in the $z - y$ colour as a function of redshift that can sway the \pzpdf\ estimates in the absence of bluer spectral features.
 
@@
 
-While this work assumes that science applications value the information of the full \pzpdf, we present conventional metrics of \pz\ point estimates as a quick and dirty diagnostic tool and to facilitate direct comparisons to historical studies.
+While this work assumes that science applications value the information of the full \pzpdf, we present conventional metrics of \pz\ point estimates as a quick and dirty visual diagnostic tool and to facilitate direct comparisons to historical studies.
 
@@
 Though we acknowledge that many of the codes can also return a native \pz\ point estimate, we put all codes on equal footing by considering two generic \pz\ point estimators, the mode $z_{PEAK}$ and main-peak-mean $z_{WEIGHT}$ \citep{Dahlen:13}, a weighted mean within the bounds of the main peak, as identified by the roots of $p(z) - 0.05 \times z_{PEAK}$.
-Though $z_{WEIGHT}$ neglects information in a secondary peak of a, say, bimodal distribution, it avoids the pitfall of reducing the \pzpdf\ to a redshift between peaks where there is low probability.
+Though $z_{WEIGHT}$ neglects information in a secondary peak of e.~g.~ a bimodal distribution, it avoids the pitfall of reducing the \pzpdf\ to a redshift between peaks where there is low probability.
 
@@
 
-We calculate the commonly used point estimate metrics of the overall intrinsic scatter, bias, and catastrophic outlier rate, defined in terms of the standard error $e_{z} \equiv (z_{PEAK} - z_{\rm true}) / (1 + z_{\rm true})$.
-Because the standard deviation of the \pz\ residuals is sensitive to outliers, we define the scatter in terms of the Interquartile Range (IQR), the difference between the 75th and 25th percentiles of the distribution of $e_{z}$, imposing the scaling $\sigma_{\rm IQR} = \rm{IQR} / 1.349$ to ensure that the area within $\sigma_{\rm IQR}$ is the same as that within one standard deviation from a standard Normal distribution.
+We calculate the commonly used point estimate metrics of the overall intrinsic scatter, bias, and catastrophic outlier rate, defined in terms of the standard error $e_{z} \equiv (z_{PEAK} - z_{\mathrm{true}}) / (1 + z_{\mathrm{true}})$.
+Because the standard deviation of the \pz\ residuals is sensitive to outliers, we define the scatter in terms of the Interquartile Range (IQR), the difference between the 75th and 25th percentiles of the distribution of $e_{z}$, imposing the scaling $\sigma_{\mathrm{IQR}} = \mathrm{IQR} / 1.349$ to ensure that the area within $\sigma_{\mathrm{IQR}}$ is the same as that within one standard deviation from a standard Normal distribution.
 We also resist the effect of catastrophic outliers by defining the bias $b_{z}$ as the median rather than mean value of $e_{z}$.
-The catastrophic outlier rate $f_{\rm out}$ is defined as the fraction of galaxies with $e_{z}$ greater than $\max(3 \sigma_{\rm IQR}, 0.06)$.
+The catastrophic outlier rate $f_{\mathrm{out}}$ is defined as the fraction of galaxies with $e_{z}$ greater than $\max(3 \sigma_{\mathrm{IQR}}, 0.06)$.
 
@@
 \begin{itemize}
-\item RMS scatter $\sigma < 0.02 (1 + z_{\rm true})$
+\item RMS scatter $\sigma < 0.02 (1 + z_{\mathrm{true}})$
 %\scc{actually the SB reports in page 75 that the goal of 0.02 is for the RMS of $\frac{\sigma_{z}}{(1+z)}$ while we are using IQR}
 \item bias $b_{z} < 0.003$ %\scc{in page 518 of the SB is clearly reported that the mean is expected to be less than 0.003 in our case we have defined as bias the median and not the mean }
-\item catastrophic outlier rate $f_{\rm out} < 10\%$ %\scc{clearly the number of outliers since are defined above $3\sigma$ strictly depends on the sigma definition, in the science book it seems to refer to the RMS of the $e_z$ distribution even ef it could be the standard deviation, but again I don't think it refers to IQR, therefore we could not consider exactly this numbers as reference.}
+\item catastrophic outlier rate $f_{\mathrm{out}} < 10\%$ %\scc{clearly the number of outliers since are defined above $3\sigma$ strictly depends on the sigma definition, in the science book it seems to refer to the RMS of the $e_z$ distribution even ef it could be the standard deviation, but again I don't think it refers to IQR, therefore we could not consider exactly this numbers as reference.}
 %\red{I am going off of conversations with Zeljko and how we {\it actually} computed the metrics in the Science Book (I have the scripts).
@@
 
-Figure~\ref{fig:pz_pointestimates} shows both point estimates forall codes  both $z_{PEAK}$ and $z_{WEIGHT}$.
+Figure~\ref{fig:pz_pointestimates} shows both point estimates for all codes  both $z_{PEAK}$ and $z_{WEIGHT}$.
 Point density is shown with mixed contours to emphasize that most of the galaxies do fall close to the $z_{phot} = z_{spec}$ line, while points trace the details of the catastrophic outlier populations.
@@
 
-The finite grid spacing of the \pzpdf s induces some discretization in $z_{PEAK}$.%, particularly for \skynet.
-The features perpendicular to the $z_{phot} = z_{spec}$ line are due to the $4000 \rm{\AA}$ break passing through the gaps between adjacent filters.% and logically are more prominent for template-based codes
+The finite grid spacing of the \pzpdf s induces some discretization in $z_{PEAK}$.
+%, particularly for \skynet.
+The features perpendicular to the $z_{phot} = z_{spec}$ line are due to the $4000$ \AA break passing through the gaps between adjacent filters.
+% and logically are more prominent for template-based codes
 Even the strongest codes feature populations far from the $z_{phot} = z_{spec}$ line representing a degeneracy in the space of colours and redshifts.
@@
 Perhaps unsurprisingly, performance under these metrics largely tracks that of the metrics of Section~\ref{sec:metrics} of the \pzpdf s from which the point estimates were derived.
-All twelve codes perform at or near the goals of the \lsst\ Science Requirements Document\footnote{available at: \url{http://ls.st/srd}} and \citet{Graham:17}, which is encouraging if not unexpected for $i < 25$.
+All twelve codes perform at or near the goals of the \lsst\ Science Requirements Document\footnote{available at: \url{http://ls.st/srd}} and \citet{Graham:17}, which is encouraging if not unexpected for $i < 25.3$.
 % But, it is still an encouraging sign, given an updated mock galaxy simulation and the expanded set of photo-$z$ codes tested.

### 2019-09-04 paper/tex/data.tex
addressing comments from @sschmidt23

@@
 For each simulated galaxy, a galaxy with similar absolute $r$-band magnitude and local galaxy density was chosen from the training set, and that training galaxy's SED was assigned to the simulated galaxy.
-In Section~\ref{sec:buzzlimitations}, we critique the realism of this mock data.
+In Section~\ref{sec:buzzlimitations}, we discuss the limited realism of this mock data.
 
@@
 
-By necessity, \textsc{Buzzard} does not contain all of the complicating factors present in real data, and here we discuss the most pertinent ways that this limitation affects our experiment.
+By necessity, \textsc{Buzzard} does not contain all of the complicating factors present in real data, and here we discuss the most pertinent ways that these limitations affect our experiment.
 % In several cases, the simplification is done with a purpose, with potentially confounding effects excluded in order to better isolate the differences between current-generation photo-$z$ codes and their causes.
-\textsc{Buzzard} includes only galaxies, not stars of AGN.
+\textsc{Buzzard} includes only galaxies, not stars nor AGN.
 The catalogue-based construction excludes image-level effects, such as deblending errors, photometric measurement issues, contamination from sky background (Zodiacal light, scattered light, etc.), lensing magnification, and Galactic reddening.
 
-The SEDs are five-component linear combinations of $\sim 5 \times 10^{5}$ \sdss\ galaxies, so the sample contains only galaxies that resemble linear combinations of those for which \sdss\ obtained spectra.
+The \textsc{Buzzard} SEDs are drawn from a set of $\sim 5 \times 10^{5}$ SEDs, which themselves are derived from a five-component linear combination fit to $\sim 5 \times 10^{5}$ \sdss\ galaxies; thus the sample contains only galaxies that resemble linear combinations of those for which \sdss\ obtained spectra, and there are necessarily duplicates.
 The linear combination SEDs also restrict the properties of the galaxy population to linear combinations of the properties corresponding to five basis templates, precluding the modeling of non-linear features such as the full range of emission line fluxes relative to the continuum.
@@
 Systematic inconsistencies with galaxy colors at $z > 2$ were observed, so the catalogue was limited to $0 < z \leq 2.0$.
-To obtain a catalogue matching the \lsst\ Gold Sample, we imposed an cut of $i < 25.3$, which gives a signal-to-noise ratio $\sim 30$ for most galaxies.
+To obtain a catalogue matching the \lsst\ Gold Sample, we imposed an cut of $i < 25.3$, which gives a signal-to-noise ratio $\gtrsim 30$ for most galaxies.
 In order for statistical errors to be subdominant to the systematic errors we aim to probe, we further reduced the sample size to $<10^{7}$ galaxies by isolating $\sim 16.8$ square degrees selected from five separate spatial regions of the simulation.
@@
 %The resulting catalogues contain $111\,171$ training galaxies and $1\,000\,883$ test galaxies.
-The \lsst\ photometric filter transmission curves were also considered public information that could be used by any code.
+The exact form of \lsst\ photometric filter transmission curves were also considered public information that could be used by any code.
 
@@
 After regularizing the SED weight coefficients $\in [0, 1]$, we ran a simple K-means clustering algorithm on the five-dimensional space of regularized SED weight coefficients of the \sdss\ galaxy sample.
-The resulting clusters were used to define Voronoi cells in the space of weight coefficients, with centre positions corresponding to weights for the \texttt{k-correct} SED components, yielding the 100 \textit{DC1 template set} to be provided to all template-based codes.
+The resulting clusters were used to define Voronoi cells in the space of weight coefficients, with centre positions corresponding to weights for the \texttt{k-correct} SED components, yielding the 100 SEDs that comprise the \textit{DC1 template set} provided to all template-based codes.
 We did not, however, exclude from consideration template-based codes that made modifications in their use of these templates due to architecture limitations (as opposed to knowledge of the experimental conditions that could artificially boost the code's apparent performance), with deviations noted in Section~\ref{sec:pzcodes}.

### 2019-09-04 paper/tex/discussion.tex
addressing comments from @sschmidt23

@@
 Because our training set is perfectly representative of the test set, $N(z)$ should be identical for both sets down to statistical noise.
-\textit{We make the alarming observation that \trainz\ outperforms all codes on the PIT-based metrics, and all but one code on the $N(z)$ based statistics.}
+\textbf{We make the alarming observation that \trainz, the experimental control, outperforms all codes on the CDF-based metrics, and all but one code on the $N(z)$ based statistics.}
+The PIT and other CDF-based metrics upon which modern \pzpdf\ comparisons are built can be gamed by a trivial estimator that yields only an affirmation of prior knowledge uninformed by the data.
+In other words, such ensemble metrics are not appropriate for the task of selecting \pzpdf\ codes for analysis pipelines.
 %\textit{Explain the connection between \trainz\ and these metrics.}
@@
 %[more caution on over-reliance on metrics, and making sure that metrics test all of what you want] \red{this definitely still needs work}
-In summary, context is crucial to defending against deceptively strong performers such as \trainz; \textbf{the best \pzpdf\ method is the one that most effectively achieves our science goals}, not the one that performs best on a metric that does not reflect those goals.
+In summary, context is crucial to defend against deceptively strong performers such as \trainz; \textbf{the best \pzpdf\ method is the one that most effectively achieves our science goals}, not the one that performs best on a metric that does not reflect those goals.
 In the absence of a single scientific motivation or the information necessary for a principled metric definition, we must consider many metrics and be critical of the information transmitted by each.
@@
 
-This initial paper explored code performance in idealized conditions with perfect catalog-based photometry and representative training data.
+This initial paper explores \pzpdf\ code performance in idealized conditions with perfect catalog-based photometry and representative training data, but the resilience of each code to such realistic imperfections in prior information has not yet been evaluated.
 A top priority for a follow-up study is to test realistic forms of incomplete, erroneous, and non-representative template libraries and training sets as well as the impact of other forms of external priors that must be ingested by the codes, major concerns in \citet{Newman:2015, Masters:2017}.
@@
 The fixed redshift grid shared between the codes may have unfairly penalized codes with a different native parameterization, as precision is lost when converting between formats.
-Performance on sharply peaked \pzpdf s may have been suppressed across all codes due to the insufficient resolution of the redshift grid.
-In light of the results of \citet[]{Malz:qp}, in future analyses we plan to switch from a fixed grid to the quantile parameterization or to permit each code to use its native storage format under a shared number of parameters.
+Performance on the (admittedly small) population of sharply peaked \pzpdf s may have been suppressed across all codes due to the insufficient resolution of the redshift grid.
+In light of the results of \citet[]{Malz:2018}, in future analyses we plan to switch from a fixed grid to the quantile parameterization or to permit each code to use its native storage format under a shared number of parameters.
 
@@
 
-To make optimal use of the \lsst\ data for cosmological and other astrophysical analyses of the Science Roadmap, future investigations that build upon this one will require a more sophisticated set of galaxy photometry and redshifts.
+To make optimal use of the \lsst\ data for cosmological and other astrophysical analyses of the \lsstdesc\ Science Roadmap, future investigations that build upon this one will require a more sophisticated set of galaxy photometry and redshifts.
 %\item \red{inclusion of photometric errors, image based effects, blending, lensing, sky, mask, etc...}

### 2019-09-04 paper/tex/introduction.tex
addressing comments from @sschmidt23

@@
 
-The current and next generations of large-scale galaxy surveys, including the Dark Energy Survey \citep[\proj{DES},][]{Abbott:05}, the Kilo-Degree Survey \citep[\proj{KiDS},][]{de_Jong:13}, Hyper Suprime-Cam Survey \citep[\proj{HSC},][]{Aihara:2018a,Aihara:2018b}, Large Synoptic Survey Telescope \citep[\lsst,][]{Abell:09}, Euclid \citep{Laureijs:11}, and Wide-Field Infrared Survey Telescope \citep[\proj{WFIRST},][]{Green:12}, present a paradigm shift from spectroscopic to photometric galaxy catalogues of substantially larger size at a cost of lacking complete spectroscopically confirmed redshifts ($z$).
+The current and next generations of large-scale galaxy surveys, including the Dark Energy Survey \citep[\proj{DES},][]{Abbott:05}, the Kilo-Degree Survey \citep[\proj{KiDS},][]{de_Jong:13}, Hyper Suprime-Cam Survey \citep[\proj{HSC},][]{Aihara:2018a,Aihara:2018b}, Large Synoptic Survey Telescope \citep[\lsst,][]{Abell:09}, Euclid \citep{Laureijs:11}, and Wide-Field Infrared Survey Telescope \citep[\proj{WFIRST},][]{Green:12}, represent a paradigm shift to reliance on photometric, rather than solely spectroscopic, galaxy catalogues of substantially larger size at a cost of lacking complete spectroscopically confirmed redshifts ($z$).
 
@@
 
-To fully characterize such physical degeneracies, later photometric galaxy catalogue data releases, (e.~g.~\citet{Erben:2013}, \citet{de_Jong:17}), provide a more informative \pz\ data product, the \pz\ probability density function (PDF), that describes the redshift probability, commonly denoted as $p(z)$, as a function of a galaxy's redshift, conditioned on the observed photometry.
+To fully characterize such physical degeneracies, later photometric galaxy catalogue data releases, (e.~g.~\citet{Sheldon:2012, Erben:2013, de_Jong:17}), provide a more informative \pz\ data product, the \pz\ probability density function (PDF), that describes the redshift probability, commonly denoted as $p(z)$, as a function of a galaxy's redshift, conditioned on the observed photometry.
 Early template-based methods such as \citet{Fernandezsoto:99} approximated the likelihood of photometry conditioned on redshift with the relative \chisq\ values of template spectra.

### 2019-09-04 paper/tex/methods.tex
addressing comments from @sschmidt23

@@
 \begin{equation} \label{eq_temp_chi}
-\chi^2(z,T,A) = \sum_i^{N_{\rm filt}}\left(\frac{F^i_{\rm{obs}} - A \, F^i_{\rm{pred}}(T,z)}{\sigma^i_{\rm{obs}}}\right)^2
+\chi^{2}(z,T,A) = \sum_{i}^{N_{\mathrm{filt}}}\left(\frac{F^{i}_{\mathrm{obs}} - A \, F^{i}_{\mathrm{pred}}(T,z)}{\sigma^{i}_{\mathrm{obs}}}\right)^2
 \end{equation}
-\noindent where $A$ is a normalization factor, $F^i_{\rm{pred}}(T,z)$ is the flux predicted for a template $T$ at redshift $z$, $F^i_{\rm{obs}}$ is the observed flux in a given band $i$, $\sigma^i_{\rm{obs}}$ is the observed flux error, and $N_{\rm filt}$ is the total number of filters, in our case the six $ugrizy$ LSST filters.
-In words, the likelihood is a sum of observed flux error $\sigma_{b}^{\rm{obs}}$-weighted squared differences between the observed flux $F^{\rm{obs}}_{b}$ and the normalized predicted flux $F^{\rm{mod}}_{b}(T, z)$ in $N_{\rm{filt}}$ photometric filters $b$, which is the \lsst\ $ugrizy$ filters in this case.
+\noindent where $A$ is a normalization factor, $F^i_{\mathrm{pred}}(T,z)$ is the flux predicted for a template $T$ at redshift $z$, $F^i_{\mathrm{obs}}$ is the observed flux in a given band $i$, $\sigma^i_{\mathrm{obs}}$ is the observed flux error, and $N_{\mathrm filt}$ is the total number of filters, in our case the six $ugrizy$ LSST filters.
+In words, the likelihood is a sum of observed flux error $\sigma_{b}^{\mathrm{obs}}$-weighted squared differences between the observed flux $F^{\mathrm{obs}}_{b}$ and the normalized predicted flux $F^{\mathrm{mod}}_{b}(T, z)$ in $N_{\mathrm{filt}}$ photometric filters $b$, which is the \lsst\ $ugrizy$ filters in this case.
 Specific implementation details of each code, e.~g.~prior form and implementation, are described below.
+
+\subsubsection{BPZ}
+\label{sec:BPZ}
+%(Sam Schmidt)
+
+Bayesian Photometric Redshift \citep[\bpz\footnote{\url{http://www.stsci.edu/~dcoe/BPZ/}},][]{Benitez:00} determines the likelihood $p(C \vert z, T)$ of a galaxy's observed colours $C$ for a set of SED templates $T$ at redshifts $z$.
+The \bpz\ likelihood is related to the \chisq\ likelihood by $p(C \vert z, T) \propto \exp[- \chi^{2} / 2]$.
+Given a Bayesian prior $p(z, T \vert m_{0})$ over apparent magnitude $m_0$ and type $T$, and assuming that the SED templates are spanning and exclusive, \bpz\ constructs the redshift posterior $p(z \vert C, m_0)$ by marginalizing over all SED templates with the form $p(z|C,m_0)\,\propto\, \sum_{T}p(C|z,T)\,p(z,T|m_0)$ \citep[Eq.~3 from][]{Benitez:00}, corresponding to setting the parameter \texttt{PROBS\_LITE=TRUE} in the \bpz\ parameter file.
+The \bpz\ prior is the product of an SED template proportion that varies with apparent magnitude $p(T \vert m_{0})$ and a prior $p(z \vert T, m_{0})$ over the expected redshift as a function of apparent magnitude and SED template.
+We anticipate \bpz\ to outperform other template-based approaches due to the prior that both comprehensively accounts for SED type and is calibrated to the training set.
+
+Here we test \bpz-v 1.99.3 \citep{Benitez:00} with the DC1 template set of Section~\ref{sec:buzztemplates}.
+To keep the number of free parameters manageable, the DC1 template set is pre-sorted by the rest-frame $u-g$ colour and split into three broad classes of SED template, equivalent to the E, Sp and Im/SB types.
+The Bayesian prior term $p(T \vert m_{0})$ was derived directly from the DC1 training set, and the other term $p(z \vert T, m_{0})$ was chosen to be the best fit for the eleven free parameters from the functional form of \citet{Benitez:00}.
+We use template interpolation, creating two linearly interpolated templates between each basis SED (sorted by rest-frame $u-g$ colour) by setting the parameter \texttt{INTERP=2}.
+%For photo-$z$ point estimates we use the \texttt{Z\_B} output parameter.
+Prior to running the code, the non-detection placeholder magnitude was replaced with an estimate of the one-$\sigma$ detection limit for the undetected band as a proxy for a value close to the estimated sky noise threshold.
+
+\subsubsection{EAZY}
+\label{sec:eazy}
+%(Rongpu Zhou)
+
+Easy and Accurate Photometric Redshifts from Yale \citep[\eazy\footnote{\url{https://github.com/gbrammer/eazy-photoz}},][]{Brammer:08} extends the basic \chisq\ fit procedure that defines template-fitting approaches.
+The algorithm models the observed photometry with a linear combination of template SEDs at each redshift.
+The best-fit SED at each redshift is found by simultaneously fitting one, two, or all of the templates via \chisq\ minimization, which is distinct from marginalizing across all templates.
+The minimized \chisq\ likelihood at each redshift is then combined with an apparent magnitude prior to obtain the redshift posterior PDF.
+We note that the utilization of the best-fit SED conditioned on redshift rather than a proper marginalization does not lead to the correct posterior distribution, an implementation issue that has now been identified and will be addressed by the developers in the future.
+
+In contrast with \bpz, \eazy's apparent magnitude prior is independent of SED, though it was derived empirically from the DC1 training set.
+The \eazy\ architecture cannot accept a template set other than the same five basis templates employed by \texttt{k-correct} when constructing the DC1 catalogue, but, for consistency with the experimental scope of perfect prior information, \eazy's flexible \texttt{all-templates} mode was used to fit the photometric data with a linear combination of the five basis templates.
+Though \eazy\ can account for uncertainty in the template set by adding in quadrature to the flux errors an empirically derived template error as a function of redshift, we set the template error to zero since the same templates were in fact used to produce the DC1 photometry.
 
@@
 Here we use \lephare-v 2.2 with the DC1 template set of Section~\ref{sec:buzztemplates}.
-
-\subsubsection{BPZ}
-\label{sec:BPZ}
-%(Sam Schmidt)
-
-Bayesian Photometric Redshift \citep[\bpz\footnote{\url{http://www.stsci.edu/~dcoe/BPZ/}},][]{Benitez:00} determines the likelihood $p(C \vert z, T)$ of a galaxy's observed colours $C$ for a set of SED templates $T$ at redshifts $z$.
-The \bpz\ likelihood is related to the \chisq\ likelihood by $p(C \vert z, T) \propto \exp[- \chi^{2} / 2]$.
-Given a Bayesian prior $p(z, T \vert m_{0})$ over apparent magnitude $m_0$ and type $T$, and assuming that the SED templates are spanning and exclusive, \bpz\ constructs the redshift posterior $p(z \vert C, m_0)$ by marginalizing over all SED templates with the form $p(z|C,m_0)\,\propto\, \sum_{T}p(C|z,T)\,p(z,T|m_0)$ \citep[Eq.~3 from][]{Benitez:00}, corresponding to setting the parameter \texttt{PROBS\_LITE=TRUE} in the \bpz\ parameter file.
-The \bpz\ prior is the product of an SED template proportion that varies with apparent magnitude $p(T \vert m_{0})$ and a prior $p(z \vert T, m_{0})$ over the expected redshift as a function of apparent magnitude and SED template.
-
-Here we test \bpz-v 1.99.3 with the DC1 template set of Section~\ref{sec:buzztemplates}.
-To keep the number of free parameters manageable, the DC1 template set is pre-sorted by the rest-frame $u-g$ colour and split into three broad classes of SED template, equivalent to the E, Sp and Im/SB types.
-The Bayesian prior term $p(T \vert m_{0})$ was derived directly from the DC1 training set, and the other term $p(z \vert T, m_{0})$ was chosen to be the best fit for the eleven free parameters of the functional form of \citet{Benitez:00}.
-We use template interpolation, creating two linearly interpolated templates between each basis SED (sorted by rest-frame $u-g$ colour) by setting the parameter \texttt{INTERP=2}.
-%For photo-$z$ point estimates we use the \texttt{Z\_B} output parameter.
-Prior to running the code, the non-detection placeholder magnitude was replaced with an estimate of the one-$\sigma$ detection limit for the undetected band as a proxy for a value close to the estimated sky noise threshold.
-
-\subsubsection{EAZY}
-\label{sec:eazy}
-%(Rongpu Zhou)
-
-Easy and Accurate Photometric Redshifts from Yale \citep[\eazy\footnote{\url{https://github.com/gbrammer/eazy-photoz}},][]{Brammer:08} extends the basic \chisq\ fit procedure that defines template-fitting approaches.
-The algorithm models the observed photometry with a linear combination of template SEDs at each redshift.
-The best-fit SED is found by simultaneously fitting one, two, or all of the templates via \chisq\ minimization, which is distinct from marginalizing across all templates.
-The minimized \chisq\ likelihood at each redshift is then combined with an apparent magnitude prior to obtain the redshift posterior PDF.
-We note that the utilization of the best-fit SED rather than a proper marginalization does not lead to the correct posterior distribution, an implementation issue that has now been identified and will be addressed by the developers in the future.
-\eazy\ can account for uncertainty in the template set by adding in quadrature to the flux errors an empirically derived template error as a function of redshift.
-
-The SED-independent apparent magnitude prior was derived empirically from the DC1 training set.
-The \eazy\ architecture cannot accept a template set other than the same five basis templates employed by \texttt{k-correct} when constructing the DC1 catalogue.
-However, \eazy\ does feature a flexible \texttt{all-templates} mode, which fits the photometric data with a linear combination of the five basis templates.
-We set the template error to zero since the same templates were in fact used to produce the DC1 photometry.
+Unlike both \bpz\ and \eazy, \lephare\ uses generic, SED-independent priors that are not tuned to the DC1 data set.
 
@@
 
-\annz \footnote{\url{https://github.com/IftachSadeh/ANNZ}} \citep{Sadeh:16} employs several machine learning algorithms, including artificial neural networks (ANN), boosted decision tree, and k-nearest neighbour (KNN) regression.
+\annz \footnote{\url{https://github.com/IftachSadeh/ANNZ}} \citep{Sadeh:16} supports several machine learning algorithms, including artificial neural networks (ANN), boosted decision tree, and k-nearest neighbour (KNN) regression.
 In addition to accounting for errors on the input photometry, \annz\ uses the KNN-uncertainty estimate of \citet{Oyaizu:08} to quantify uncertainty in the choice of method over multiple runs.
@@
 \Pzpdf s were produced by running an ensemble of 5 ANNs with a $6:12:12:1$ architecture corresponding to the 6 $ugrizy$ inputs, 2 hidden layers with 12 nodes each, and 1 output of redshift.
-Each of the five ANNs was trained with different random seeds for the initialization of input parameters.
-Additionally, all ANNs were trained on only a $i \leq 25.3$ subsample of the DC1 training set, and half of the training set was reserved for validation to prevent overfitting.
+Each of the five ANNs was trained with different random seeds for the initialization of input parameters, reserving half of the training set for validation to prevent overfitting.
 Undetected galaxies were excluded from the training set, and per-band non-detections in the test set were replaced with the mean magnitude in that band within the entire test set.
@@
 
-The nearest-neighbours colour-matching photometric redshift estimator \citep[\cmnn,][]{Graham:17} uses a training set of galaxies with known redshifts that has equivalent or better photometry than the test set in terms of quality and filter coverage.
-For each galaxy in the test set, \cmnn\ identifies a colour-matched subset of training galaxies using a threshold in the Mahalanobis distance $D_M = \sum_{j}^{N_{\rm colours}} (c^{\rm train}_{j} - c^{\rm test}_{j})^{2} / \delta c_{\rm test}^2$ in the space of available colours $c$, with colour measurement errors $\delta c_{\rm test}$ and $N_{\rm colours} = 5$ colors $j$ defined by the $ugrizy$ filters, which defines the set of colour-matched neighbours based on a value of the percent point function (PPF).
-As an example, for $N_{\rm{filt}}=5$ with PPF$=0.95$, $95\%$ of all training galaxies consistent with the test galaxy will have $D_M < 11.07$.
-Undetected bands are dropped, thereby reducing the effective $N_{\rm{filt}}$ for that galaxy.
+The colour-matched nearest-neighbours photometric redshift estimator \citep[\cmnn,][]{Graham:17} uses a training set of galaxies with known redshifts that has equivalent or better photometry than the test set in terms of quality and filter coverage.
+For each galaxy in the test set, \cmnn\ identifies a colour-matched subset of training galaxies using a threshold in the Mahalanobis distance $D_M = \sum_{j}^{N_{\mathrm{colours}}} (c^{\mathrm{train}}_{j} - c^{\mathrm{test}}_{j})^{2} / \delta c_{\mathrm{test}}^2$ in the space of available colours $c$, with colour measurement errors $\delta c_{\mathrm{test}}$ and $N_{\mathrm{colours}} = 5$ colors $j$ defined by the $ugrizy$ filters, which defines the set of colour-matched neighbours based on a value of the percent point function (PPF).
+As an example, for $N_{\mathrm{filt}}=5$ with PPF$=0.95$, $95\%$ of all training galaxies consistent with the test galaxy will have $D_M < 11.07$.
+Undetected bands are dropped, thereby reducing the effective $N_{\mathrm{filt}}$ for that galaxy.
 The \pzpdf\ of a given test set galaxy is the normalized distribution of redshifts of its colour-matched subset of training set galaxies.
@@
 We use PPF$=0.95$ rather than PPF$=0.68$ to generate the subset of colour-matched training galaxies, whose redshifts are weighted by their inverse Mahalanobis distances of the when composing the \pzpdf\ rather than weighting all colour-matched training galaxies equally.
-Additionally, when the number of colour-matched training set galaxies is less than 20, the nearest 20 neighbours in color-space are used instead, and the output \pzpdf\ is convolved with a Gaussian kernel of variance $\sigma_{\rm train}^{2}({\rm PPF}_{20}/0.95)^2 -1$ to account foe the corresponding growth of the effective PPF to include 20 neighbors.
+Additionally, when the number of colour-matched training set galaxies is less than 20, the nearest 20 neighbours in color-space are used instead, and the output \pzpdf\ is convolved with a Gaussian kernel of variance $\sigma_{\mathrm{ train}}^{2}(\mathrm{PPF}_{20}/0.95)^2 -1$ to account for the corresponding growth of the effective PPF to include 20 neighbors.
+
+\subsubsection{Delight}
+\label{sec:delight}
+%(John Soo)
+
+\delight \footnote{\url{https://github.com/ixkael/Delight}} \citep{Leistedt:17} is a hybrid technique that infers \pz s with a data-driven model of latent SEDs and a physical model of photometric fluxes as a function of redshift.
+Generally, machine learning methods rely on representative training data with shared photometric filters, while template based methods rely on a complete library of templates based on physical models constructed.
+\delight\ aims to take the best aspects of both approaches by constructing a large collection of latent SED templates (or physical flux-redshift models) from training data, with a template SED library as a guide to the learning of the model, thereby circumventing the machine learning prerequisite of representative training data in the same photometric bands and the template fitting requirement of detailed galaxy SED models.
+It models noisy observed flux $\mathbf{\hat{F}} = \mathbf{F} + F_{b}$ as a sum of a noiseless flux plus a Gaussian processes $F_b \sim \mathcal{GP}\left(\mu^F, k^F \right)$ with zero mean function $\mu^{F}$ and a physically motivated kernel $k^{F}$ that induces realistic correlations in flux-redshift space.
+
+From a template-fitting perspective, each test set galaxy has a posterior $p(z \vert \mathbf{\hat{F}}) \approx \sum_i p(\mathbf{\hat{F}} \vert z, T_i) p(z \vert T_i) p(T_i)$ of redshift $z$ conditioned on noisy flux $\mathbf{\hat{F}}$, where $p(z \vert T_i) p(T_i)$ captures prior information about the redshift distributions and abundances of the galaxy templates $T_i$.
+As in traditional template fitting, each likelihood $p(\mathbf{\hat{F}} \vert \mathbf{F})$ relates the noisy flux $\mathbf{\hat{F}}$ with the noiseless flux $\mathbf{F}$ predicted by the model of a linear combination of templates, carefully constructed to account for model uncertainties and different normalization of the same SED, plus the Gaussian process term.
+
+The machine learning approach appears in the inclusion of a pairwise comparison term $p(\mathbf{F} \vert z, z_j, \mathbf{\hat{F}}_j)$ for the prediction of model flux $\mathbf{F}$ at a model redshift $z$ with respect to training set galaxy $j$ with redshift $z_j$ and observed flux $\mathbf{\hat{F}}_j$.
+Thus the \pz\ posterior $p(\mathbf{\hat{F}} \vert z, T_i) = \int p(\mathbf{\hat{F}} \vert \mathbf{F}) p(\mathbf{F} \vert z, z_j, \mathbf{\hat{F}}_j) d\mathbf{F}$ may be interpreted as the probability that the training and the target galaxies have the same SED at different redshifts.
+The flux prediction $p(\mathbf{F} \vert z, z_j, \mathbf{\hat{F}}_j)$ of the training galaxy at redshift $z$ is modeled via the Gaussian process described above; more detail is provided in \citet{Leistedt:17}.
+
+In this study, the default settings of \delight\ were used, with the exception that the PDF bins were set to be linearly-spaced rather than logarithmic.
+The Gaussian process was trained using the full DC1 training set.
+We used the full DC1 template set with a flat prior in magnitude and SED type.
+Photometric uncertainties from the inputs are propagated into the code, while non-detections for each band are set to the mean of the respective bands.
 
@@
 \gpz's variance estimate is composed of a model uncertainty term corresponding to sparsity of the training set photometry and a noise uncertainty term encompassing noisy photometric observations, enabling quantification of any need for more representative or more precise training samples.
-\gpz\ may also weight training set samples by importance according to $|z_{\rm{spec}} - z_{\rm{phot}}| / (1+z_{\rm{spec}})$ to minimize the normalized \pz\ point estimate error, however, this function may be adapted to \pzpdf s, pressuring the model to dedicate more resources to test set galaxies that are not well-represented in the training set.
+\gpz\ may also weight training set samples by importance according to $|z_{\mathrm{spec}} - z_{\mathrm{phot}}| / (1+z_{\mathrm{spec}})$ to minimize the normalized \pz\ point estimate error, however, this function may be adapted to \pzpdf s, pressuring the model to dedicate more resources to test set galaxies that are not well-represented in the training set.
 
@@
 % WHAT IS BEING DECORRELATED HERE?
-Unobserved magnitudes $x_{\rm u} = \mu_{\rm u} + \Sigma_{\rm uo}\Sigma_{\rm oo}^{-1}(x_{\rm o} - \mu_{\rm o})$ were imputed from observed magnitudes $x_{\rm o}$ and the training set mean $\mu$ and covariance $\Sigma$ using a linear model.
-This is the optimal expected value of the unobserved variables given the observed ones under the assumption that the distribution is jointly Gaussian; note that this reduces to a simple average if the covariates are independent with $\Sigma_{\rm uo} = 0$.
+Unobserved magnitudes $x_{\mathrm{u}} = \mu_{\mathrm{u}} + \Sigma_{\mathrm{uo}}\Sigma_{\mathrm{oo}}^{-1}(x_{\mathrm{o}} - \mu_{\mathrm{o}})$ were imputed from observed magnitudes $x_{\mathrm{o}}$ and the training set mean $\mu$ and covariance $\Sigma$ using a linear model.
+This is the optimal expected value of the unobserved variables given the observed ones under the assumption that the distribution is jointly Gaussian; note that this reduces to a simple average if the covariates are independent with $\Sigma_{\mathrm{uo}} = 0$.
 We reserved for validation 20\% of the training set and used the Variable Covariance option in \gpz\ with 200 basis functions (see \citet{Almosallam:15b} for details), neglecting to apply cost-sensitive learning options.
@@
 The neural network is configured as a standard multilayer perceptron with three hidden layers and one input layer with 12 nodes corresponding to the 6 photometric magnitudes and their measurement errors.
-We use \skynet\ as a regressor for \pz\ point estimation and as a classifier for \pzpdf\ estimation.
+% We use \skynet\ as a regressor for \pz\ point estimation and as a classifier for \pzpdf\ estimation.
 
-The regressor used a standard \chisq\ error function with a single linear node as the output layer and 10 nodes with a $\tanh$ activation function for each hidden layer.
-The classifier used a cross-entropy error function with a 20:40:40 node (all rectified linear units) architecture for each hidden layer and an output layer of 200 nodes corresponding to 200 bins for the PDF, with a softmax activation function to enforce the normalization condition that the probabilities sum to unity.
+% The regressor used a standard \chisq\ error function with a single linear node as the output layer and 10 nodes with a $\tanh$ activation function for each hidden layer.
+\skynet's classifier mode uses a cross-entropy error function with a 20:40:40 node (all rectified linear units) architecture for each hidden layer and an output layer of 200 nodes corresponding to 200 bins for the PDF, with a softmax activation function to enforce the normalization condition that the probabilities sum to unity.
 While previous implementations of the code \citep[see Appendix C.3 of~][]{Sanchez:14,Bonnett:15} implement a sliding bin smoothing, no such procedure was used in this study.
@@
 
-\subsubsection{Delight}
-\label{sec:delight}
-%(John Soo)
-
-\delight \footnote{\url{https://github.com/ixkael/Delight}} \citep{Leistedt:17} is a hybrid technique that infers \pz s with a data-driven model of latent SEDs and a physical model of photometric fluxes as a function of redshift.
-Generally, machine learning methods rely on representative training data with shared photometric filters, while template based methods rely on a complete library of templates based on physical models constructed.
-\delight\ aims to take the best aspects of both approaches by constructing a large collection of latent SED templates (or physical flux-redshift models) from training data, with a template SED library as a guide to the learning of the model, thereby circumventing the machine learning prerequisite of representative training data in the same photometric bands and the template fitting requirement of detailed galaxy SED models.
-It models noisy observed flux $\mathbf{\hat{F}} = \mathbf{F} + F_{b}$ as a sum of a noiseless flux plus a Gaussian processes $F_b \sim \mathcal{GP}\left(\mu^F, k^F \right)$ with zero mean function $\mu^{F}$ and a physically motivated kernel $k^{F}$ that induces realistic correlations in flux-redshift space.
-
-From a template-fitting perspective, each test set galaxy has a posterior $p(z \vert \mathbf{\hat{F}}) \approx \sum_i p(\mathbf{\hat{F}} \vert z, T_i) p(z \vert T_i) p(T_i)$ of redshift $z$ conditioned on noisy flux $\mathbf{\hat{F}}$, where $p(z \vert T_i) p(T_i)$ captures prior information about the redshift distributions and abundances of the galaxy templates $T_i$.
-As in traditional template fitting, each likelihood $p(\mathbf{\hat{F}} \vert \mathbf{F})$ relates the noisy flux $\mathbf{\hat{F}}$ with the noiseless flux $\mathbf{F}$ predicted by the model of a linear combination of templates, carefully constructed to account for model uncertainties and different normalization of the same SED, plus the Gaussian process term.
-
-The machine learning approach appears in the inclusion of a pairwise comparison term $p(\mathbf{F} \vert z, z_j, \mathbf{\hat{F}}_j)$ for the prediction of model flux $\mathbf{F}$ at a model redshift $z$ with respect to training set galaxy $j$ with redshift $z_j$ and observed flux $\mathbf{\hat{F}}_j$.
-Thus the \pz\ posterior $p(\mathbf{\hat{F}} \vert z, T_i) = \int p(\mathbf{\hat{F}} \vert \mathbf{F}) p(\mathbf{F} \vert z, z_j, \mathbf{\hat{F}}_j) d\mathbf{F}$ may be interpreted as the probability that the training and the target galaxies have the same SED at different redshifts.
-The flux prediction $p(\mathbf{F} \vert z, z_j, \mathbf{\hat{F}}_j)$ of the training galaxy at redshift $z$ is modeled via the Gaussian process described above; more detail is provided in \citet{Leistedt:17}.
-
-In this study, the default settings of \delight\ were used, with the exception that the PDF bins were set to be linearly-spaced rather than logarithmic.
-The Gaussian process was trained using the full DC1 training set.
-We used the full DC1 template set with a flat prior in magnitude and SED type.
-Photometric uncertainties from the inputs are propagated into the code, while non-detections for each band are set to the mean of the respective bands.
-
-\subsection{trainZ: a pathological \pz\ estimator}
+\subsection{trainZ: a pathological \pzpdf\ estimator}
 \label{sec:trainz}
@@
 \end{equation}
-Unlike the other methods, the \trainz\ estimator is \textit{independent of the photometric data}, effectively performing a KNN procedure with $k=N_{\rm train}$.
+Unlike the other methods, the \trainz\ estimator is \textit{independent of the photometric data}, effectively performing a KNN procedure with $k=N_{\mathrm{train}}$.
 

### 2019-09-04 paper/tex/old-results.tex
addressing comments from @sschmidt23

\section{Results}
\label{sec:results}
%(Sam Schmidt, Bryce Kalmbach, Johann Cohen Tanugi, Rongpu Zhou)

\subsection{Ensembles of photo-\mathinhead{$z$}{z} interim posteriors}

\begin{figure*}
%\centering
\includegraphics[width=0.49\textwidth]{fig/pz_12codes_261931_crop.jpg}\includegraphics[width=0.5\textwidth]{fig/pz_12codes_471167_crop.jpg}\\
\includegraphics[width=0.49\textwidth]{fig/pz_12codes_713178_crop.jpg}\includegraphics[width=0.49\textwidth]{fig/pz_12codes_982747_crop.jpg}
\caption{Four illustrative examples of individual p(z) distributions produced by the codes.  The red vertical line represents the true redshift.  Examples are chosen with common features seen in PDFs: tight unimodal $p(z)$ (upper left), broad unimodal $p(z)$ (upper right), bimodal $p(z)$ (lower right), and complex/multimodal $p(z)$ (lower left).  Codes show varying amounts of small-scale structure in their reconstruction of the posterior distribution.  We see varying responses from the codes in the presence of color degeneracies and photometric errors, resulting in narrow and broad unimodal, bimodal, and multi-modal $p(z)$ curves.} \label{fig:pz_examples}
\end{figure*}

%\begin{figure*}
%\centering
%\includegraphics[width=\textwidth]{fig/QQplot_11codes.jpg}
%\caption{The QQ plots produced for every photo-$z$ code.  \aim{Consider making this a plot of QQ minus that diagonal for better contrast between methods.}\red{The PIT histogram is essentially this, which is one of the main reasons to include both.  QQ is a fairly standard form, so plotting in this fashion rather than the difference seems better to me, given that we also ahve the PIT histograms. --SJS}} \label{fig:qq}
%\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{fig/PITANDQQplot_12codes_crop.jpg}
\caption{Summary plots for all twelve photo-$z$ codes illustrating performance for the interim posterior statistics. The top panel of each pair shows both the Quantile-Quantile (QQ) plot (red) and the histogram of PIT values (blue).  The desired behavior is a QQ plot that matches the diagonal dashed line, and a PIT histogram that matches a uniform distribution matching the thin horizontal black line.  The bottom panel of each pair shows the difference between the QQ quantile and the diagonal, illustrating departure from the desired performance.  Histograms with an overabundance of PIT values at the centre of the distribution indicate $p(z)$ distributions that are overly broad, while an excess of values at the extrema indicate $p(z)$ distributions that are overly narrow.  Values of PIT=0 and PIT=1 indicate ``catastrophic failures'' where the true redshift is completely outside the support of $p(z)$.  Asymmetric features are indicative of systematic bias in the redshift predictions.  A variety of behaviors are evident, and specific details are discussed in the text.}
\label{fig:pitqq}
\end{figure*}

Fig.~\ref{fig:pz_examples} Shows the $p(z)$ produced by each of our twelve photo-$z$ codes for four example galaxies which exemplify some prominent cases that arise when estimating photo-$z$ PDFs: a narrow, unimodal redshift solution, a broader unimodal solution, a bimodal distribution, and a complex, multimodal distribution.
The red vertical line represents the true redshift of the individual galaxy, and the blue curve represents the redshift probability.
Several features are obvious even in these illustrative examples.
\textsc{ANNz2, METAPhoR, NN,} and \textsc{SkyNet} all show an excess of small-scale features, which appear to be print-through of the underlying training set galaxies.  For example, in \textsc{CMNN} the $p(z)$ are a simply a weighted histogram of all spectroscopic training galaxies in nearby colour space with no smoothing applied, so the substructure is due to the finite number of neighbours, and is not unexpected.
\textsc{GPZ} (in its current implementation), on the other hand, always produces a single Gaussian, which broadens to cover the multi-modal redshift solutions seen in other codes.
%Interestingly, while \textsc{ANNz2} shows an abundance of small scale structure in individual $p(z)$ measurements (see Fig.~\ref{fig:pz_examples}), the summed $\hat{N}(z)$ is rather smooth, where the small scale features average out.  This is not the case for the two other codes that show an abundance of substructure in their individual $p(z)$: both \textsc{CMNN} and \textsc{SkyNet} show small scale features both in $p(z)$ and $\hat{N}(z)$.

As stated in Section~\ref{sec:metrics}, $p(z)$ is parameterized as $\approx 200$ piecewise constant bins covering $0<z<2$ for all twelve codes, giving a grid size of roughly $\delta z = 0.01$ for each code.
A piecewise constant grid was a natural choice for some photo-$z$ codes, for instance most template-based codes compute likelihoods on a fixed grid.
In contrast, FlexZBoost, for example, can return estimates on any grid without compression errors as it’s a basis expansion method where only the expansion coefficients need to be stored.
Codes with a native output format other than the shared piecewise constant binning scheme (or one that can be losslessly converted to it) may suffer from loss of information when converting to it, which could artificially favor some codes over others.

Furthermore, the fidelity of photo-$z$ interim posteriors in this format varies with the quality of the photometry.
For faint galaxies, this redshift resolution is sufficient to capture the shape of $p(z)$ for the majority of the test sample, where photometric errors on the faint galaxies lead to somewhat broad peaks in the redshift posterior.
However, as can be seen in e.~g.~the top left panel of Fig.~\ref{fig:pz_examples}, for bright galaxies with narrow $p(z)$ the grid spacing of $\delta z = 0.01$ is not sufficient to resolve the peak.
This is consistent with the results described in \citet[]{Malz:qp}, who find that quantiles (and, to a lesser degree, samples) often outperform gridded $p(z)$, particularly for bright objects and in the presence of harsher storage constraints.
With a full 200 numbers to capture the information of each photo-$z$ PDF, any parametrization will perform adequately, but other storage parametrizations and limits on storage resources may be considered in future work.
%Switching to a quantile based parameterization may be more costly computationally, for example template-based codes would need to test more grid point in order to resolve the quantiles for bright galaxies.  However, the computational time for template based codes scales roughly linearly with the number of grid points, so this may be a reasonable thing to implement.
We will discuss this further in Section~\ref{sec:discussion}.
% \red{someone review this statement to make sure that I'm saying this correctly!}

Fig.~\ref{fig:pitqq} shows both the quantile-quantile plots (red) and the histogram of PIT values (blue) summarizing the results from each photo-$z$ code.  The red line shows the measured quantiles, while the black diagonal represents the ideal QQ values if the distribution were perfectly reproduced.  A second panel below the main panel for each code shows the difference between $Q_{data}$ and $Q_{theory}$, i.~e.~the departure from the diagonal, for clarity.
Biases and trends in whether the average width of the $p(z)$ values being over/under-predicted are evident.  An overall bias where the predicted redshift is systematically low manifests as the measured QQ value falling above the diagonal, as is the case for \textsc{BPZ} and \textsc{EAZY}, while a systematic overprediction shows up as the measured QQ value falling below the diagonal, as seen in \textsc{TPZ}.  In terms of PIT histograms, a systematic underprediction of redshift corresponds to fewer PIT values at $PIT<0.5$ and more at $PIT>0.5$, while a systematic overprediction will show the opposite.

Examination of the PIT histograms and QQ plots shows that there are fairly generic issues with the width of $p(z)$ uncertainties: \textsc{Delight, CMNN, SkyNet} and \textsc{TPZ} all show a PIT histogram with an dearth of low values and and an excess of high values, signs that, on average, their $p(z)$ are more broad than the true distribution of redshifts.  \textsc{METAPhoR} shows the opposite trend, indicating the the $p(z)$  are more narrow than the distributions given by the true redshifts.  In all of these code cases there is a free parameter or bandwidth that can be used to tune uncertainties.  The sensitivity of multiple codes to this bandwidth choice emphasizes the fact that great care must be taken in setting user-defined parameters in photo-$z$ codes, even in the presence of representative training/validation data.  for \textsc{FlexZBoost} the ``sharpening'' parameter (described in Section~\ref{sec:flexzboost}) plays a key role in improving the results, resulting in a QQ plot that is very nearly diagonal.  A similar sharpening procedure could be beneficial for several codes.
Interestingly, the three purely template-based codes, \textsc{BPZ, EAZY}, and \textsc{LePhare}, show relatively well behaved $p(z)$ statistics (albeit with some bias), which may indicate that the likelihood estimation with representative templates is accurately capturing the uncertainties on individual redshifts.

%Fig.~\ref{fig:pitqq} shows the PIT histogram plots produced by each photo-$z$ code.
%These plots show information that is very similar to that shown in the QQ plot in many ways.
The ideal PIT histogram would follow the black dashed line, representing a uniform distribution of PIT values, equivalent to the diagonal line in the QQ plot.
Overly broad $p(z)$ values show up as an excess of PIT values near $0.5$ and a dearth of values at the edges, while overly narrow $p(z)$ will have an excess at the edges and will be missing values at the centre.
Another feature evident in the PIT histograms is the number of ``catastrophic outlier'' values where the true redshift falls outside of the non-zero support of $p(z)$, corresponding to PIT$=0.0$ or $1.0$ is more apparent than in the QQ plots.  Following Kodra \& Newman (in prep.) we define $f_{\rm O}$ as the fraction of objects with $PIT<0.0001$ or $PIT>0.9999$.  Table~\ref{tab:pitoutlier} lists these fractions for each of the codes. For a proper Uniform distribution we expect a value of $0.0002$.  Several codes show a marked excess, with \textsc{ANNz2, FlexZBoost, LePhare, and METAPhoR} with $f_{\rm O} > 0.02$, indicating a sizeable number of catastrophic redshift solutions where the true redshift is not covered by the extent of $p(z)$.  For \textsc{METAPhoR} this may be partially due to an overall underprediction of the $p(z)$ width, however this is not the case for the other codes.  \textsc{LePhare} is a particular outlier with nearly 5 per cent of objects outside of $p(z)$ support.  Further study will be necessary to determine what is causing these misclassifications for \textsc{LePhare}.  As expected, and by design, \textsc{TrainZ} has the proper fraction of outliers for the $f_{\rm O}$ statistic.

\begin{table}
\setlength{\tabcolsep}{2pt}
\centering
\caption{The fraction of ``catastrophic outlier'' PIT values.  We expect a value of 0.0002 for a proper Uniform distribution.  An excess over this small value indicates true redshifts that fall outside the non-zero support of the $p(z)$.  }\label{tab:pitoutlier}
\begin{tabular}{lc}
\hline
\hline
 ``catastrophic outlier''\\ PIT fraction\\
\hline
Photo-z Code & fraction PIT$<10^{-4}$ or $>$0.9999\\
\hline
\textsc{ANNz2} & 0.0265\\
\textsc{BPZ} & 0.0192\\
\textsc{Delight} & 0.0006\\
\textsc{EAZY} & 0.0154\\
\textsc{FlexZBoost} & 0.0202\\
\textsc{GPz} & 0.0058\\
\textsc{LePhare} & 0.0486\\
\textsc{METAPhoR}& 0.0229\\
\textsc{CMNN} & 0.0034\\
\textsc{Skynet} & 0.0001\\
\textsc{TPZ} & 0.0130\\
\hline
\textsc{TrainZ} & 0.0002\\
\end{tabular}
\end{table}

%\begin{figure*}
%\centering
%\includegraphics[width=\textwidth]{fig/PIT_histogram_plot_11codes.jpg}
%\caption{The PIT histogram plots produced for every photo-$z$ code.  Ideal behaviour would match a uniform distribution, indicated by the horizontal dashed line.  Histograms with an overabundance of PIT values at the centre of the distribution indicate $p(z)$ distributions that are overly broad, while an excess of values at the extrema indicate $p(z)$ distributions that are overly narrow.  Values of PIT=0 and PIT=1 indicate ``catastrophic failures'' where the true redshift is completely outside the support of $p(z)$.  Asymmetric features are indicative of systematic bias in the redshift predictions.
%\aim{Could we combine this with the previous figure by plotting both as differences from ideal on one plot with left and right axis labels?  The information seems degenerate (which makes sense!)}\red{as you say, the information is degenerat, this form emphasizes the differences a bit more, and shows off the catastrophic outliers with PIT=0 and 1 a bit better in my mind.} } \label{fig:pit}
%\end{figure*}
\begin{figure*}
\centering
\includegraphics[width=\textwidth]{fig/KSvsCvMvsAD_PIT_withnull_jpg.jpg}
\caption{A visual representation of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the PIT distributions.  The statistics are often highly correlated, though the AD statistic truncates the extrema of the distribution and can have disparate values compared to KS and CvM.} \label{fig:pit_stats}
\end{figure*}

Fig.~\ref{fig:pit_stats} shows comparative metric values for the quantitative Kolmogorov-Smirnoff (KS), Cramer-Von Mises (CvM), and Anderson Darling (AD) test statistics for each of the codes based on comparing the distribution of their PIT values to the expected uniform distribution over the interval [0,1].  The individual values of the statistic are not as important as the comparative score between the different codes.
%\red{Can p-values be supplied for each statistic? The statistics themselves are difficult to interpret, other than ``lower is better'' (p-value in skgof was broken, having trouble finding 1-sample KS calculation for uniform distribution)}
The AD test statistic diverges for values that include the extrema, and thus is calculated by excluding the edges of the distribution.
We calculate the AD statistic over the range of PIT values $v=[0.01,0.99]$.  \textsc{ANNz2} and \textsc{FlexZBoost} score very well for the PIT metrics.
\textsc{METAPhoR} and \textsc{LePhare} score very well in the PIT AD statistic, but both have a large number of catastrophic outliers, resulting in higher KS and CvM scores.

Given the near-perfect training data, examining the individual codes for explanations for departures from the expected behaviour will be instructive in avoiding similar problems in future tests.
\textsc{ANNz2} performs quite well in $p(z)$ based metrics.  In the specific implementation employed in this paper, the final $p(z)$ is a weighted average of five neural-nets.  During the training process \textsc{ANNz2} compares the percentiles of the redshift training sample against the CDFs of the $p(z)$ sample.  Distributions that more closely match are given extra weight, and the final weights are designed to produce accurate percentiles.  Given that our metrics are focused on the percentile distributions, it is unsurprising that \textsc{ANNz2} performs well in the given metrics.  The discreteness in the individual $p(z)$ estimated by \textsc{ANNz2} can be attributed to the fact that the code was run as a classifier, assigning weights to discrete bins of redshift.  While multiple bins may receive weight, the bins themselves will still be discretized, and no additional smoothing was performed.
Overall, \textsc{FlexZBoost} and \textsc{ANNz2} show the best ensemble agreement in their distribution of PIT values.

\subsection{Metrics of the stacked estimator of the redshift distribution}\label{sec:stackedmetrics}

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{fig/NZsumplot_12codes_scottsrule.jpg}
\caption{The stacked $p(z)$ produced by each photo-$z$ code ($\hat{N}(z)$, red) compared to the spectroscopic redshift distribution ($N'(z)$, blue).  Varying levels of agreement are seen in the codes.  Both $\hat{N}(z)$ and $N'(z)$ in all codes are smoothed using a single bandwidth chosen via Scott's rule.} \label{fig:nz}
%\aim{Arguably this could also be clarified by showing the true $n'(z)$ once and then differences from it for each cod's stacked estimator\dots}\scc{what if we plot the difference between true and stacked rather than soverposing the two lines? Sam please do not hate me.}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{fig/KSvsCvMvsAD_NZ_withnull_jpg.jpg}
\caption{A visual representation of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the $\hat{N}(z)$ distributions. The statistics are correlated, the codes with the lowest KS statistics tend to have the lowest CvM and AD statistics.  \textsc{CMNN} performs markedly better than the others in reconstructing the overall $N(z)$ distribution, while \textsc{SkyNet} scores poorly due to an overall bias in its redshift predictions.} \label{fig:nz_stats}
\end{figure*}

Fig.~\ref{fig:nz} shows the stacked $\hat{N}(z)$ distribution compared to the true redshift distribution $N'(z)$ for all tested codes.  The red line indicates the summed $p(z)$ for each code, while the blue line shows the true redshift distribution. All distributions are smoothed via kernel density estimation (KDE) with a common bandwidth chosen via Scott's rule \citep{Scott:1992} in order to minimize differences in small-scale features and make for a more uniform comparison between codes.  While Scott's rule is used to display $N'(z)$ in the figure, all quantitative statistics are computed via the empirical CDF, and are thus unaffected by bandwidth/smoothing choice.
%New text for the new figure
As expected, \textsc{TrainZ} is in excellent agreement with the true redshift distribution: as the training sample is selected from the same underlying distribution as the test set, the redshift distributions are identical, up to Poisson fluctuations due to the finite number of sample galaxies.  \textsc{CMNN} is also in excellent agreement for similar reasons: with a representative training sample of galaxies spanning the colour-space, the sum of the colour-matched neighbour redshifts should return the true redshift distribution. \textsc{FlexZBoost} and \textsc{TPZ} also show very good agreement, with only slight departures, with an over/under-prediction in the high redshift tail of $\hat{N}(z)$ evident around $z\sim1.4$.  In fact, several of the other codes show an excess at $z \sim 1.4$, particularly the template-based codes \textsc{BPZ}, \textsc{EAZY}, and \textsc{LePhare}.  This is likely due to the $4000\ {\rm \AA}$ break passing through the gap between the $z$ and $y$ filters, resulting in a drastic change in $z-y$ colour for galaxies in this redshift range.  With a relative dearth of strong features blue-ward of the $4000\ {\rm \AA}$ break in most galaxy SEDs, the colour change in the two reddest filter bandpasses of a survey has a large influence on the redshift determination.  The $z\sim1.4$ feature is one of the most prominent sources of larger uncertainty in individual galaxy $p(z)$.
In our sample individual galaxy $p(z)$'s tend to be broader around $z\sim1.4$ and point estimates are more uncertain in this regime, as is readily seen in the point-estimate plots shown in Fig.~\ref{fig:pz_pointestimates} and described in the Appendix.  This feature is not unique to this dataset, it is a common occurrence in photo-$z$ estimation.  The fact that similar excesses appear in Figure~\ref{fig:nz} for \textsc{ANNz2} and \textsc{METAPhoR} shows that the effect is not limited to template-based codes.  However, the lack of such a feature in the other codes shows that it is possible to eliminate the degeneracies.  Further study on this issue may provide a solution for codes that suffer from this shortcoming.

%end of new text (and slightly modified text)
Two of the machine learning based codes, \textsc{ANNz2} and \textsc{METAPhoR}, appear to be over-trained, adding excess galaxy probability to the redshift peaks with the largest number of training galaxies, and missing probability in the troughs where training galaxies are of fewer number.
Given that our training data is drawn from the same galaxy population as the test set, and our data has prominent peaks in $N'(z)$, perhaps it is not unexpected that such overtraining occurs in some codes, though the fact that it does not occur in all training-based codes indicates that it may be due to specifics of the implementations of \textsc{ANNz2} and \textsc{METAPhoR}.
A more extensive training/validation set might allow for a better choice of smoothing parameters in individual codes that would avoid such overtraining for these particular codes.
%\aim{Move the text on Scott's Rule to Discussion?}
%Scott's rule \citep{Scott:1992} was used to determine the smoothing bandwidth of the true redshift sample: that is, a single smoothing scale was chosen to represent the true redshift sample for all twelve codes.
%As with the $p(z)$ values in Figure~\ref{fig:pitqq}, different levels of substructure are obvious for the different codes.
%While Scott's rule provides a relatively good general smoothing scale to represent the true $N'(z)$, there are smaller scale fluctuations: while \textsc{FlexZBoost} and \textsc{CMNN} appear somewhat discrepant in Fig.~\ref{fig:nz}, they are actually the two most accurate in terms of their quantitative measurements.
%\textsc{SkyNet} also shows excess small-scale structure, but also shows significant overestimate of the high redshift galaxy population.
\textsc{SkyNet} shows an obvious redshift bias, evident both visually in Figure~\ref{fig:nz} and in the first moment of N(z) listed in Table~\ref{tab:moments}, where it is clearly an outlier.  \textsc{SkyNet} employed a method where a random sample of training galaxies was chosen, but there was no test that the subset was completely representative of the overall redshift distribution.  Unlike the previous implementation of \textsc{SkyNet} in \citet{Bonnett:15}, no effort was made to add extra weight to more rare low and high redshift galaxies.  Either of these decisions could be the cause of the bias seen in our results.  Future runs of \textsc{SkyNet} will explore these implementation choices and their effects.

Figure~\ref{fig:nz_stats} shows the quantitative Kolmogorov-Smirnoff (KS), Cramer-Von Mises (CvM), and Anderson Darling (AD) test statistics for each of the codes for the $\hat{N}(z)$ based measures.
%\red{Can p-values be supplied for each statistic? The statistics themselves are difficult to interpret, other than ``lower is better'' (no, p-values are very difficult to compute for non-uniform distributions)}
\textsc{FlexZBoost}, \textsc{CMNN}, and \textsc{TPZ} outperform the other codes in the $\hat{N}(z)$ metrics.
It is unsurprising that \textsc{CMNN} scores well, as with a near perfectly representative training set means that choosing neighbouring points in color/magnitude space should lead to excellent agreement in the final $\hat{N}(z)$ estimate.  \textsc{TPZ} performed quite poorly in $p(z)$ statistics, but results in a good fit to the overall $N(z)$.  This is somewhat surprising, as performance was optimized for accurate $p(z)$, not $\hat{N}(z)$.  During the validation stage for \textsc{TPZ}, there was a trade off between the width of the $p(z)$ when adjusting a smoothing parameter and overall redshift bias.  The optimal result in the PIT metrics, as illustrated in the shape of the QQ plot, does contain some level of bias as well as a slight underprediction of mean $p(z)$ width, which translates to poor metric scores.  This is something that will be looked into for \textsc{TPZ} in the future.

%\red{discussion of why NN and FZBoost are doing well, how we can try to improve the other codes}
%\textcolor{cyan}{Ann: }

Table~\ref{tab:cdeloss} shows the CDE loss statistic for each photo-$z$ code.  Once again \textsc{FlexZBoost} and \textsc{CMNN} score very well for the stacked $\hat{N}(z)$ metrics, as do \textsc{GPz} and \textsc{TPZ}.  The CDE loss measures how well individual PDFs are estimated, and codes with a low CDE loss tend to have good $\hat{N}(z)$ estimates (though the reverse is not necessarily true). \textsc{FlexZBoost} is optimized to minimize CDE loss which may explain why the method has good ensemble metrics as well. Note from Table~\ref{tab:cdeloss} that both \textsc{FlexZBoost} and \textsc{CMNN} have low CDE losses.  Empirically, we have found that PIT RMSE is not as closely correlated to CDE loss as it is to the $N(z)$ statistics.  As CDE loss is a better measure of individual redshift performance, rather than ensemble distribution performance, this statistic is a better indicator of which codes will be most likely to perform well for science cases where single objects are employed.

Table~\ref{tab:rmse} gives the root-mean-square-error (RMSE) statistics for both the PIT and N(z) estimators.  The PIT value calculates the RMSE between the quantiles shown in the QQ plot in Figure~\ref{fig:pitqq} and the diagonal, while the N(z) calculates the RMSE between the cumulative distribution of the stacked $\hat{N}(z)$ and the true redshift distribution $N'(z)$.  %\red{more about RMSE?  Or remove it? It's barely described in Section~\ref{sec:metrics} and wasn't even referred to in the text until I added this on June 29th--SJS.}

Table~\ref{tab:moments} lists the first three moments of the stacked $\hat{N}(z)$ distribution, including the moments of the ``truth'' distribution for comparison.  Several codes are able to reproduce the mean and variance of the distribution to less than a per cent, while several codes do not, which may be a cause for concern, given that mean and variance of the redshift distribution are key properties in cosmological analyses.  We note that this stated goal of the study as defined for participants was to accurately reproduce $p(z)$, the ``stacking'' of the probability distributions to estimate $\hat{N}(z)$ was not the focus as stated to the participants.  This explains why some of the best-performing empirical codes in terms of $p(z)$ measures (e.~g.~\textsc{FlexZBoost}) do not do as well at reproducing $\hat{N}(z)$ moments.  Had we defined a different parameter to optimize, in this case overall accuracy of $\hat{N}(z)$ rather than individual $p(z)$, would result in improved performance in a particular metric.  That is, optimizing photo-$z$ performance for one metric does not automatically give optimal performance for other metrics.  As previously stated, there are a variety of scientific use cases for photo-$z$'s in large upcoming surveys, and care must be taken in how the metrics used to optimize catalog photometric redshifts are defined as well as in how they are used.  In addition, very few scientific use cases will employ the overall $\hat{N}(z)$ with no cuts, as we explore in this paper.  We discuss more realistic tomographic bin selections that will be explored in a follow-up paper in Section~\ref{sec:futurework}.
%\red{[mention that we are calculating moments for the entire sample, will look at fiducial tomographic bins in a follow-up paper (unless group and reviewers realy think we should include in this paper.]}
%\red{FlexZBoost has some of the best $n(z)$ statistics, but the moments are not good.  Why is this?  Is it the over/underprediction of the high redshift part of the distribution?  Discuss this after talking with Rafael.}

%\subsection{Response of Individual Codes}
%\label{sec:res:pz_indiv_codes}
%\red{this may be incorporated in to 5.1 and 5.2 rather than live in its own section!}

%\red{HOW CODES deal with negative fluxes, and magnitude uncertainties}

%\red{overall conclusions of the Results section}

\subsection{Interpretation of metrics}\label{sec:caution}
%Caution on metrics and photo-$z$ codes}\label{sec:caution}
%(Alex Malz, Sam Schmidt)

Samples from accurate photo-$z$ posteriors should reproduce the space of $p(z, data)$. 
However, it is difficult to test this reconstruction given our data set, as the galaxy distributions arise from mock objects pasted on to an underlying dark matter halo catalogue with properties designed to match empirical relations, rather than being drawn from statistical distributions in redshift.  
In previous sections we have mentioned that optimizing for a specific metric does not guarantee good performance on other metrics, nor is there any guarantee that good performance by our metrics corresponds to \textit{accurate} photo-$z$ posteriors.
%\aim{Explain that samples from accurate photo-$z$ posteriors should reconstruct the space of $p(z, data)$, a test we can't check with our datasets.}
In other words, we can construct photo-$z$ estimators that provide good coverage in many of our tests, but which have very little predictive power.

The \trainz\ estimator, which assigns every galaxy a $p(z)$ equal to $N(z)$ of the training set as described in Section~\ref{sec:method:trainz}, is introduced as a ``null test'' to demonstrate this point via \textit{reductio ad absurdum}.
\trainz\ outperforms all codes on the PIT-based metrics, and all but one code on the $N(z)$ based statistics.
Because our training set is perfectly representative of the test set, $N(z)$ should be identical for both sets down to statistical noise.
%\textit{Explain the connection between \trainz\ and these metrics.}

The CDE loss and point estimate metrics, however, successfully identify problems with \trainz.
As shown in Appendix~\ref{sec:pointmetrics}, \trainz\ has identical $ZPEAK$ and $ZWEIGHT$ values for every galaxy, and thus the photo-$z$s are constant as a function of spec-$z$s, i.e. a horizontal line at the mode and mean of the training set distribution respectively.  
The explicit dependence on the \textit{individual} posteriors in the calculation of the CDE loss, described in Section~\ref{sec:CDE_loss}, distinguishes this metric from the other $p(z)$ metrics that test the overall ensemble of $p(z)$ distributions.  
With a representative training set, \trainz\ will score well on the ensemble metrics, but fails miserably for metrics tied to individual redshifts.  
We note that many of the ensemble-based metrics are prominent in the photo-$z$ literature despite their inability to identify problems such as those exemplified by \trainz.

%\aim{Explain why CDE loss is different from other metrics (isn't fooled by \trainz) and note that the others have more of a presence in the literature despite their flaw of failing to identify this severe problem.}

% While looking at one, or even most of our metrics would have given the impression that this estimator was nearly optimal, red flags in CDE loss and point estimates reveal the problems.
%[more caution on over-reliance on metrics, and making sure that metrics test all of what you want] \red{this definitely still needs work}
In summary, context is crucial to interpreting metrics and defending against the likes of \trainz.
The best photo-$z$ method is the one that most effectively achieves our science goals, not the one that performs best on a metric that does not accurately reflect those goals.
In the absence of clear goals or the information necessary for a principled metric definition, we must think carefully before choosing a single metric

%\begin{table*}  %%% DATA TABLE %%%
%\caption{KS, CvM and AD statistics for each photo-$z$ code.} \label{tab:results}
%PIT statistic
%
%\begin{tabular}{lrrrr}
%\hline
%\bf Photo-$z$ Code & \bf KS & \bf CvM & \bf AD & \bf AD \\
%                   &        &         & $v=[0.05,0.95]$ & $v=[0.01,0.99]$        \\
%\hline
%\textsc{ANNz2} 		& $0.0200$ &   $52.25$ &   $583.0$ &   $759.2$ \\
%\textsc{BPZ} 		& $0.0388$ &  $280.79$ &  $1361.4$ &  $1557.5$ \\
%%\textsc{Delight} 	& $0.1380$ & $3040.59$ & $13880.2$ & $17761.4$ \\
%\textsc{Delight}    & $0.0876$ & $1075.17$ & $4407.8$  & $6167.5$\\
%%\textsc{EAZY} 		& $0.0690$ &  $994.19$ &  $3479.5$ &  $4158.8$ \\
%\textsc{EAZY}       & $0.0723$ & $1105.58$ & $3475.0$  & $4418.6$\\
%%\textsc{FlexZBoost} & $0.0260$ &   $70.58$ &   $595.0$ &   $533.0$ \\
%\textsc{FlexZBoost} & $0.0240$ &   $68.83$ &   $567.9$ &   $478.8$\\
%\textsc{GPz} 		& $0.0449$ &  $258.56$ &   $904.8$ &  $1414.8$ \\
%\textsc{LePhare} 	& $0.0663$ &  $473.05$ &    $85.9$ &   $383.8$ \\
%\textsc{METAPhoR} 	& $0.0438$ &  $298.56$ &   $248.6$ &   $715.5$ \\
%%\textsc{CMNN} 		& $-$ & $-$ & $-$ & $-$ \\
%\textsc{CMNN}         & $0.0795$  &$1011.11$ &  $3908.2$ &  $6307.5$ \\
%\textsc{SkyNet} 	& $0.0747$ &  $763.00$ &  $2740.2$ &  $4216.4$ \\
%\textsc{TPZ} 		& $0.1138$ & $1801.74$ &  $9046.2$ & $10565.7$ \\
%%\textsc{Frankenz}	& $-$ & $-$ & $-$ & $-$ \\
%\hline
%\end{tabular}

%Stacked $n(z)$ statistic
%
%\begin{tabular}{lrrrr}
%\hline
%\bf Photo-$z$ Code & \bf KS & \bf CvM & \bf AD & \bf AD \\
%                   &        &         & $z=[0.002,2.0]$ & $z=[0.0,2.0]$        \\
%\hline
%\textsc{ANNz2} 		& $0.0237$ &  $25.676$ &   $276.4$ &   $276.0$ \\
%\textsc{BPZ} 		& $0.0131$ &  $17.587$ &   $169.0$ &   $170.9$ \\
%%\textsc{Delight} 	& $0.0184$ &  $18.975$ &   $151.5$ &   $151.5$ \\
%\textsc{Delight}    & $0.0256$ &  $42.341$ &   $253.3$ &   $253.2$ \\
%%\textsc{EAZY} 		& $0.0434$ & $167.329$ &   $811.3$ &   $817.0$ \\
%\textsc{EAZY}       & $0.0441$ & $169.977$ &   $827.7$ &   $833.3$ \\
%%\textsc{FlexZBoost} & $0.0065$ &   $2.764$ &    $24.7$ &    $24.6$ \\
%\textsc{FlexZBoost} & $0.0128$ &   $7.007$ &  $127.02$ &   $126.9$\\
%\textsc{GPz} 		& $0.0168$ &  $25.605$ &   $307.1$ &   $307.1$ \\
%\textsc{LePhare} 	& $0.0254$ &  $58.900$ &   $477.4$ &   $472.0$ \\
%\textsc{METAPhoR} 	& $0.0350$ &  $60.790$ &   $671.9$ &   $672.3$ \\
%%\textsc{CMNN} 		& $-$ & $-$ & $-$ & $-$ \\
%\textsc{CMNN}         & $0.0043$ & $0.538$   &   $5.6$   &      $5.8$ \\
%\textsc{SkyNet} 	& $0.0483$ & $385.830$ &  $2119.4$ &  $2110.3$ \\
%\textsc{TPZ} 		& $0.0126$ &   $9.483$ &   $110.7$ &   $110.6$ \\
%%\textsc{Frankenz}	& $-$ & $-$ & $-$ & $-$ \\
%\hline
%\end{tabular}
%\end{table*}

\begin{table}  %%% DATA TABLE %%%
\centering
\caption{CDE loss statistic for each photo-$z$ code.} \label{tab:cdeloss}
\begin{tabular}{lr}
\hline
\bf Photo-$z$ Code & \bf CDE Loss \\
\hline
\textsc{ANNz2} 		& $-6.88$ \\
\textsc{BPZ} 		& $-7.82$ \\
%\textsc{Delight} 	& $-4.06$ \\
\textsc{Delight}    & $-8.33$\\
%\textsc{EAZY} 		& $-7.97$ \\
\textsc{EAZY}       & $-7.07$ \\
%\textsc{FlexZBoost} & $-11.51$ \\
\textsc{FlexZBoost} & $-10.60$\\
\textsc{GPz} 		& $-9.93$ \\
\textsc{LePhare} 	& $-1.66$ \\
\textsc{METAPhoR} 	& $-6.28$ \\
%\textsc{CMNN} 		& $-$ \\
\textsc{CMNN}         & $-10.43$ \\
\textsc{SkyNet} 	& $-7.89$ \\
\textsc{TPZ} 		& $-9.55$ \\
\hline
\trainz		& $-0.83$ \\
%\textsc{Frankenz}	& $-$  \\
%\hline
\end{tabular}
\end{table}

\begin{table}
\setlength{\tabcolsep}{2pt}
\begin{center}
\caption{Root-Mean-Square-Error (RMSE) statistics for the twelve photo-z codes for both PIT and $\hat{N}(z)$ distributions.}\label{tab:rmse}
\begin{tabular}{lcc}
\hline
\hline
Root-Mean-Square-Error \\
(RMSE) statistics \\
\hline
Photo-z Code & PIT RMSE & N(z) RMSE\\
\hline
\textsc{ANNz2}      & 0.019 & 0.0054\\
\textsc{BPZ}        & 0.032 & 0.0050\\
\textsc{Delight}    & 0.111 & 0.0056\\
\textsc{EAZY}       & 0.054 & 0.0102\\
\textsc{FlexZBoost} & 0.021 & 0.0022\\
\textsc{GPz}        & 0.027 & 0.0042\\
\textsc{LePhare}    & 0.028 & 0.0062\\
\textsc{METAPhoR}   & 0.064 & 0.0081\\
\textsc{CMNN}         & 0.108 & 0.0009\\
\textsc{Skynet}     & 0.054 & 0.0144\\
\textsc{TPZ}        & 0.082 & 0.0031\\
\hline
\trainz		& 0.0025 & 0.0013\\
\end{tabular}
\end{center}
\end{table}

%\begin{table}
%\begin{center}
%caption{RMSE Statistics for the 11 Photo-z codes for DC1 $i<25.3$ ``gold'' sample}\label{tab:pitrmse}
%%\begin{tabular}{|l|c|c|c|c|}
%\begin{tabular}{lc}
%\hline
%\hline
% PIT RMSE statistics \\
%\hline
%Photo-z Code & PIT RMSE\\
%\hline
%\texttt{ANNz2}      & 0.019\\
%\texttt{BPZ}        & 0.032\\
%\texttt{Delight}    & 0.111\\
%\texttt{EAZY}       & 0.054\\
%\texttt{FlexZBoost} & 0.021\\
%\texttt{GPz}        & 0.048\\
%texttt{LePhare}    & 0.028\\
%\texttt{METAPhoR}   & 0.064\\
%\texttt{NN}         & 0.108 \\
%\texttt{Skynet}     & 0.054\\
%texttt{TPZ}        & 0.082\\
%\hline
%\end{tabular}
%\end{center}
%\end{table}
%
%\begin{table}
%\begin{center}
%\caption{RMSE Statistics for the 11 Photo-z codes for DC1 $i<25.3$ ``gold'' sample}\label{tab:nzrmse}
%%\begin{tabular}{|l|c|c|c|c|}
%\begin{tabular}{lc}
%\hline
%\hline
% N(z) RMSE statistics \\
%\hline
%hoto-z Code & N(z) RMSE\\
%\hline
%\texttt{ANNz2}      & 0.0054\\
%\texttt{BPZ}        & 0.0050\\
%\texttt{Delight}    & 0.0056\\
%\texttt{EAZY}       & 0.0102\\
%\texttt{FlexZBoost} & 0.0022\\
%\texttt{GPz}        & 0.0042\\
%\texttt{LePhare}    & 0.0062\\
%\texttt{METAPhoR}   & 0.0081\\
%\texttt{NN}         & 0.0009\\
%\texttt{Skynet}     & 0.0144\\
%\texttt{TPZ}        & 0.0031\\
%\end{tabular}
%\end{center}
%\end{table}

\begin{table}
\setlength{\tabcolsep}{2pt}
\caption{Moments of the stacked $\hat{N}(z)$ distribution}\label{tab:moments}
\begin{tabular}{lccc}
\hline
\hline
 \multicolumn{4}{l}{Stacked $n(z)$ Moments} \\
\hline
              & 1st Moment & 2nd Moment & 3rd Moment \\
\textsc{Truth}     & 0.701      &   0.630    & 0.671  \\
\hline
Photo-z Code       & 1st Moment & 2nd Moment & 3rd Moment\\
\hline
\textsc{ANNz2}     & 0.702      & 0.625      & 0.653    \\
\textsc{BPZ}       & 0.699      & 0.629      & 0.671    \\
\textsc{Delight}   & 0.692      & 0.609      & 0.638    \\
\textsc{EAZY}      & 0.681      & 0.595      & 0.619    \\
\textsc{FlexZBoost}& 0.694      & 0.610      & 0.631    \\
\textsc{GPz}       & 0.696      & 0.615      & 0.639    \\
\textsc{LePhare}   & 0.718      & 0.668      & 0.741    \\
\textsc{METAPhoR}  & 0.705      & 0.628      & 0.657    \\
\textsc{CMNN}        & 0.701      & 0.628      & 0.667    \\
\textsc{Skynet}    & 0.743      & 0.708      & 0.797    \\
\textsc{TPZ}       & 0.700      & 0.619      & 0.643    \\
\hline
\trainz	   & 0.699 		& 0.627 	& 0.666 \\
\end{tabular}
\end{table}

### 2019-09-04 paper/tex/results.tex
addressing comments from @sschmidt23

@@
 
-The most striking differences between codes are due to small-scale features induced by the interaction between the shared piecewise constant parameterization of $200$ bins $0 < z < 2$ of Section~\ref{sec:metrics} and the smoothing conditions or lack thereof in each algorithm.
-The $\rm{d}z = 0.01$ redshift resolution is sufficient to capture the broad peaks of faint galaxies' \pzpdf s with large photometric errors but is too broad to resolve the narrow peaks for bright galaxies' \pzpdf s with small photometric errors.
-This observation is consistent with the findings of \citet[]{Malz:qp} that the piecewise constant parameterization underperforms in the presence of small-scale structures.
+The most striking differences between codes are the small-scale features induced by the interaction between the shared piecewise constant parameterization of $200$ bins $0 < z < 2$ of Section~\ref{sec:metrics} and the smoothing conditions or lack thereof in each algorithm.
+The $\mathrm{d}z = 0.01$ redshift resolution is sufficient to capture the broad peaks of faint galaxies' \pzpdf s with large photometric errors but is too broad to resolve the narrow peaks for bright galaxies' \pzpdf s with small photometric errors.
+This observation is consistent with the findings of \citet[]{Malz:2018} that the piecewise constant parameterization underperforms in the presence of small-scale structures.
 
@@
 The values of the CDE loss statistic of individual \pzpdf\ accuracy are provided in Table~\ref{tab:cdeloss}.
-It is worth noting that strong performance on the CDE loss should imply strong performance on the other metrics, though the inverse is not necessarily true.
+It is worth noting that strong performance on the CDE loss, corresponding to lower values of the metric, should imply strong performance on the other metrics, though the inverse is not necessarily true.
 Thus the CDE loss is the most effective metric for generic science cases.
@@
 
-This metric is the only one that can appropriately penalize \trainz\ and indicates strong performance for \cmnn\ and \flexzboost, the latter of which is optimized for this metric.
+Of the metrics we were able to consider in this experiment, \textbf{the CDE Loss is the only metric that can appropriately penalize the pathological\trainz}.
+Additionally, it favors \cmnn\ and \flexzboost, the latter of which is optimized for this metric.

### 2019-09-04 paper/tex/analysis.tex
incorporated feedback from @morriscb

@@
 The exact implementation of the metrics theoretically depends on the parametrization of the \pzpdf s, which may differ across codes and can affect the precision of the estimator \citep{Malz:2018}.
-Even considering a single method under the same parametrization, such as the 200-bin $0 < z < 2$ piecewise constant function used here, the exact bin definitions will affect the result.
+Even considering a single method under the same parametrization, such as the 200-bin $0 < z < 2$ piecewise constant function used here, the exact bin definitions must affect the result.
 %\red{Note: Let's address how the gridding (endpoints and spacing) affects the metrics, from the email correspondence Re: [LSST-DESC-MEMBERS] Summary of Nov 29 PZ Telecon. (AIM)}
@@
 Because \lsst's \pzpdf s will be used for many scientific applications, some of which require accuracy of each individual catalog entry, we consider several metrics that probe the population-level performance of the \pzpdf s.
-Because we have the true redshifts but not true \pzpdf s for comparison, we remind the reader of the Cumulative Distribution Function (CDF)
+As we have the true redshifts but not true \pzpdf s for comparison, we remind the reader of the Cumulative Distribution Function (CDF)
 \begin{equation}
@@
 of a generic univariate PDF $f(z)$, which is used as the basis for several of our metrics.
+We describe metrics based on the CDF in Section~\ref{sec:qqpit} and metrics of summary statistics thereof in Section~\ref{sec:summqqpit}.
+
+\subsubsection{CDF-based metrics}
+\label{sec:qqpit}
 
@@
 %Though this metric does not account for the fact that the redshift distribution function is, in fact, a probability distribution, it can still be interpreted as a measure of the integrated difference between the estimated distribution and the true distribution.%, and it can be used to quantify the otherwise qualitative metrics.
+
+\subsubsection{Summary statistics of CDF-based metrics}
+\label{sec:summqqpit}
 
@@
 To the knowledge of the authors, there is only one metric that can be used to evaluate the performance of individual \pzpdf\ estimators in the absence of true \pz\ posteriors.
-Using the notation introduced in Section~\ref{sec:flexzboost}, the conditional density estimation (CDE) loss is defined as
+The conditional density estimation (CDE) loss is an analogue to the familiar root-mean-square-error used in conventional regression, defined as
 \begin{equation}
   \label{eq:cde-loss}
-  L(f, \widehat{f}) \equiv \int \int (f(z \vert \x) - \widehat{f}(z \vert \x))^{2} \mathrm{d}z \mathrm{d}P(\x)
+  L(f, \widehat{f}) \equiv \int \int (f(z \vert \x) - \widehat{f}(z \vert \x))^{2} \mathrm{d}z \mathrm{d}P(\x) ,
 \end{equation}
-in terms of the photometry $\x$, an analogue to the familiar root-mean-square-error used in conventional regression.
+where $f(z \vert \x)$ is the true \pzpdf\ that we do not have and $\widehat{f}(z \vert \x)$ is an estimate thereof, in terms of the photometry $\x$.
+(See Section~\ref{sec:flexzboost} for a review of the notation.)
 We estimate the CDE loss via

### 2019-09-04 paper/tex/discussion.tex
incorporated feedback from @morriscb

@@
 \textbf{We make the alarming observation that \trainz, the experimental control, outperforms all codes on the CDF-based metrics, and all but one code on the $N(z)$ based statistics.}
-The PIT and other CDF-based metrics upon which modern \pzpdf\ comparisons are built can be gamed by a trivial estimator that yields only an affirmation of prior knowledge uninformed by the data.
+The PIT and other CDF-based metrics upon which modern \pzpdf\ comparisons are built \citep{Bordoloi:10,Polsterer:16,Tanaka:17} can be gamed by a trivial estimator that yields only an affirmation of prior knowledge uninformed by the data.
 In other words, such ensemble metrics are not appropriate for the task of selecting \pzpdf\ codes for analysis pipelines.

### 2019-09-04 paper/tex/introduction.tex
incorporated feedback from @morriscb

@@
 The current and next generations of large-scale galaxy surveys, including the Dark Energy Survey \citep[\proj{DES},][]{Abbott:05}, the Kilo-Degree Survey \citep[\proj{KiDS},][]{de_Jong:13}, Hyper Suprime-Cam Survey \citep[\proj{HSC},][]{Aihara:2018a,Aihara:2018b}, Large Synoptic Survey Telescope \citep[\lsst,][]{Abell:09}, Euclid \citep{Laureijs:11}, and Wide-Field Infrared Survey Telescope \citep[\proj{WFIRST},][]{Green:12}, represent a paradigm shift to reliance on photometric, rather than solely spectroscopic, galaxy catalogues of substantially larger size at a cost of lacking complete spectroscopically confirmed redshifts ($z$).
+Effective astrophysical inference using the catalogues resulting from these ongoing and upcoming missions, however, necessitates accurate and precise photometric redshift (\pz) estimation methodologies.
 
-Effective astrophysical inference using the catalogues resulting from these ongoing and upcoming missions, however, necessitates accurate and precise photometric redshift (\pz) estimation methodologies.
 As an example, in order for \pz\ systematics to not dominate the statistical noise floor of \lsst's main cosmological sample of $\sim 10^{7}$ galaxies, the \lsst\ Science Requirements Document (SRD)\footnote{available at \url{https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17}} specifies that individual galaxy \pz s must have root-mean-square error $\sigma_z < 0.02 (1+z)$, $3 \sigma$ catastrophic outlier rate below $10\%$, and bias below $0.003$.
@@
 Though the standard has long been for each galaxy in a photometric catalogue to have a \pz\ point estimate and Gaussian error bar, even early applications of \pz s in precision cosmology indicate the inadequacy of point estimates \citep{mandelbaum:2008} to encapsulate the degeneracies resulting from the nontrivial mapping between broad band fluxes and redshift.
-Far from a hypothetical situation, this degeneracy is a real consequence of the same deep imaging that enables larger galaxy catalogue sizes.
+Far from a hypothetical situation, such degeneracies are real consequences of the same deep imaging that enables larger galaxy catalogue sizes.
 The lower luminosity and higher redshift populations captured by deeper imaging introduce major physical systematics to \pz s, among them the Lyman break/Balmer break degeneracy, that did not affect shallower large area surveys like the Sloan Digital Sky Survey \citep[\textsc{SDSS},][]{York:00} and Two Micron All Sky Survey \citep[\textsc{2MASS},][]{Skrutskie:06}.
@@
 There are numerous techniques for deriving \pzpdf s, yet no one method has been established as clearly superior.
-In fact, in the absence of simulated data drawn from known redshift distributions, the very concept of a ``true PDF'' for an individual galaxy is unavailable, and we must instead rely on measures of ensemble behaviour to characterize PDF quality (see \S\,\ref{sec:metrics} for further discussion).
-This caveat aside, quantitative comparisons of \pz\ methods have been made before; the \Pz\ Accuracy And Testing \citep[\textsc{PHAT},][]{Hildebrandt:10} effort focused on \pz\ point estimates derived from many photometric bands.
+Consistent experimental conditions enable the quantification if not isolation of their differences, which can be interpreted as a sort of \textit{implicit prior} imparted by the method itself.
+Comprehensive comparisons of \pz\ methods have been made before; the \Pz\ Accuracy And Testing \citep[\textsc{PHAT},][]{Hildebrandt:10} effort focused on \pz\ point estimates derived from many photometric bands.
 \citet{Rau:2015} introduced a new method for improving \pzpdf s using an ordinal classification algorithm.
@@
 This paper is distinguished from other comparisons of \pz\ methods by its focus on the evaluation criteria for \pzpdf s and interpretation thereof.
+In the absence of simulated data drawn from known redshift distributions, the very concept of a ``true PDF'' for an individual galaxy is unavailable, and we must instead rely on measures of ensemble behaviour to characterize PDF quality (see \S\,\ref{sec:metrics} for further discussion).
 We aim to perform a comprehensive sensitivity analysis of \pzpdf\ techniques in order to ultimately select those that will become part of the \lsst\ pipelines, as part of a key project of the Photometric Redshifts working group of the \lsstdesc, described in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}}.
-In this initial study, we focus on evaluating the performance of \pzpdf\ codes using PDF-specific performance metrics in a controlled experiment with complete and representative prior information (template libraries and training sets) to set a baseline for subsequent investigations.
+In this initial study, we focus on evaluating the performance of \pzpdf\ codes using PDF-specific performance metrics in a formally controlled experiment with complete and representative prior information (template libraries and training sets) to set a baseline for subsequent investigations.
 This approach probes how each code considered exploits the information content of the data versus prior information from template libraries and training sets.

### 2019-09-04 paper/tex/appendix.tex
taking out CHIPPR reference?

@@
 
-Though alternatives exist \citep{Malz:chippr}, ``stacking'' according to
+``Stacking'' according to
 \begin{equation}
@@
 is the most widely accepted method for obtaining $\hat{N}^{H}(z)$ as an estimator of the redshift distribution from \pzpdf s derived by a method $H$.
-Though the use of the stacked estimator of the redshift distribution is not formally correct, we use it under the untested assumption that the response of our metrics of $\hat{N}^{H}(z)$ will be analogous to the same metrics applied to a principled estimator of the redshift distribution.
+Though the use of the stacked estimator of the redshift distribution is not formally correct \citep{Malz:chippr}, we use it under the untested assumption that the response of our metrics of $\hat{N}^{H}(z)$ will be analogous to the same metrics applied to a principled estimator of the redshift distribution.
 

### 2019-09-19 paper/tex/conclusion.tex

@@
 In appendices, we also present metrics of point estimates and a prevalent summary statistic of \pzpdf\ catalogues used in cosmological analyses to enable the reader to relate this work to studies of similar scope.
-We observe that no one code dominates in all metrics, and that the standard metrics of \pzpdf s and the stacked estimator of the redshift distribution can be gamed by a vacuously wrong procedure that asserts the prior over the data.
+We observe that no one code dominates in all metrics, and that the standard metrics of \pzpdf s and the stacked estimator of the redshift distribution can be gamed by a very simplistic procedure that asserts the prior over the data.
 We emphasize to the \pz\ community that \textbf{metrics used to vet \pzpdf\ methods must be scrutinized to ensure they correspond to the quantities that matter to our science}.

### 2019-09-19 paper/tex/data.tex

@@
 For each simulated galaxy, a galaxy with similar absolute $r$-band magnitude and local galaxy density was chosen from the training set, and that training galaxy's SED was assigned to the simulated galaxy.
-In Section~\ref{sec:buzzlimitations}, we discuss the limited realism of this mock data.
 
@@
 
-Given the SED, absolute $r$-band magnitude, and redshift, we computed apparent magnitudes in the six \lsst\ filter passbands, $ugrizy$.
+Given the SED, absolute $r$-band magnitude, and true redshift of each simulated galaxy, we computed apparent magnitudes in the six \lsst\ filter passbands, $ugrizy$.
 We assigned magnitude errors in the six bands using the simple model of \citet{Ivezic:08}, assuming achievement of the full 10-year depth, with a modification of fiducial \lsst\ total numbers of 30-second visits for photometric error generation: we assume 60 visits in $u$-band, 80 visits in $g$-band, 180 visits in $r$-band, 180 visits in $i$-band, 160 visits in $z$-band, and 160 visits in $y$-band.
@@
 Following the findings of \citet{Bernstein:10}, \citet{Masters:2017} that only $~10^{4}$ spectra are necessary to calibrate \pz s to Stage IV requirements, we aimed to set aside a randomly selected training set of $3-5\times 10^{4}$ galaxies, $\sim 10\%$ of the full sample.
-After all cuts described above, we designated the \textit{DC1 training set} of $44\,404$ galaxies for which observed photometry, true SEDs, and true redshifts would be provided to all codes codes and the blinded \textit{DC1 test set} of $399\,356$ galaxies for which photometry alone would be provided to all codes and \pzpdf s would be requested.
+After all cuts described above, we designated the \textit{DC1 training set} of $44\,404$ galaxies for which observed photometry, true SEDs, and true redshifts would be provided to all codes and the blinded \textit{DC1 test set} of $399\,356$ galaxies for which photometry alone would be provided to all codes and \pzpdf s would be requested.
 %The resulting catalogues contain $111\,171$ training galaxies and $1\,000\,883$ test galaxies.

### 2019-09-19 paper/tex/methods.tex

@@
 \noindent where $A$ is a normalization factor, $F^i_{\mathrm{pred}}(T,z)$ is the flux predicted for a template $T$ at redshift $z$, $F^i_{\mathrm{obs}}$ is the observed flux in a given band $i$, $\sigma^i_{\mathrm{obs}}$ is the observed flux error, and $N_{\mathrm filt}$ is the total number of filters, in our case the six $ugrizy$ LSST filters.
-In words, the likelihood is a sum of observed flux error $\sigma_{b}^{\mathrm{obs}}$-weighted squared differences between the observed flux $F^{\mathrm{obs}}_{b}$ and the normalized predicted flux $F^{\mathrm{mod}}_{b}(T, z)$ in $N_{\mathrm{filt}}$ photometric filters $b$, which is the \lsst\ $ugrizy$ filters in this case.
+The likelihood is a sum of observed flux error $\sigma_{b}^{\mathrm{obs}}$-weighted squared differences between the observed flux $F^{\mathrm{obs}}_{b}$ and the normalized predicted flux $F^{\mathrm{mod}}_{b}(T, z)$ in $N_{\mathrm{filt}}$ photometric filters $b$, which is the \lsst\ $ugrizy$ filters in this case.
 Specific implementation details of each code, e.~g.~prior form and implementation, are described below.
@@
 
-Photometric Analysis for Redshift Estimate \citep[\lephare\footnote{\url{http://www.cfht.hawaii.edu/~arnouts/lephare.html}},][]{Arnouts:99,Ilbert:06} uses the $\chi^2$ of Equation~\ref{eq_temp_chi} to match observed colors with those predicted from a template set, which can be semi-empirical or entirely synthetic, directly according to the
+Photometric Analysis for Redshift Estimate \citep[\lephare\footnote{\url{http://www.cfht.hawaii.edu/~arnouts/lephare.html}},][]{Arnouts:99,Ilbert:06} uses the $\chi^2$ of Equation~\ref{eq_temp_chi} to match observed colors with those predicted from a template set.
+The template set can be semi-empirical or entirely synthetic.
 The reported \pzpdf\ is an arbitrary normalization of the likelihood evaluated on the output redshift grid.
@@
 
-In this paper, we used \texttt{xgboost} \citep{Chen:16} for the regression; it should however be noted that \texttt{FlexCode-RF}\footref{flexzboost_github}, based on Random Forests, generally performs better for smaller datasets.
+In this paper, we used \texttt{xgboost} \citep{Chen:16} for the regression; it should, however, be noted that \texttt{FlexCode-RF}\footref{flexzboost_github}, based on Random Forests, generally performs better on smaller datasets.
 As our basis $\phi_{i}(z)$, we choose a standard Fourier basis.

### 2019-09-23 paper/tex/introduction.tex
fix citep on line 49

@@
 
-To fully characterize such physical degeneracies, later photometric galaxy catalogue data releases, (e.~g.~\citet{Sheldon:2012, Erben:2013, de_Jong:17}), provide a more informative \pz\ data product, the \pz\ probability density function (PDF), that describes the redshift probability, commonly denoted as $p(z)$, as a function of a galaxy's redshift, conditioned on the observed photometry.
+To fully characterize such physical degeneracies, later photometric galaxy catalogue data releases, \citep[e.~g.~][]{Sheldon:2012, Erben:2013, de_Jong:17}, provide a more informative \pz\ data product, the \pz\ probability density function (PDF), that describes the redshift probability, commonly denoted as $p(z)$, as a function of a galaxy's redshift, conditioned on the observed photometry.
 Early template-based methods such as \citet{Fernandezsoto:99} approximated the likelihood of photometry conditioned on redshift with the relative \chisq\ values of template spectra.

### 2019-09-27 paper/tex/appendix.tex
add new moments figure and some text to support it

@@
 
-The first three moments of the stacked $\hat{N}(z)$ distribution relative to the empirical estimate of the truth distribution are given in Table~\ref{tab:moments}.
+\begin{figure*}
+\centering
+\includegraphics[width=0.75\textwidth]{fig/moments_figure.jpg}
+\caption{Residuals of the first three moments of the stacked $\hat{N}(z)$ distribution.
+Red and blue horizontal lines indicate the Year 1 DESC SRD requirements on accuracy of the mean and variance respectively.
+Only a small number of codes are able to meet these specifications even with perfect training data.
+}
+\label{fig:moments}
+\end{figure*}
+
+We calculated the first three moments of the stacked $\hat{N}(z)$ distribution of all galaxies and compared it to the moments of the true redshift distribution.
+Figure~\ref{fig:moments} shows the residuals of the moments for all codes.
 Accuracy of the moments varies widely between codes, raising concerns about the propagation to cosmological analyses.
-%\red{[mention that we are calculating moments for the entire sample, will look at fiducial tomographic bins in a follow-up paper (unless group and reviewers realy think we should include in this paper.]}
-%\red{FlexZBoost has some of the best $n(z)$ statistics, but the moments are not good.
-Why is this?
-Is it the over/underprediction of the high redshift part of the distribution?
-Discuss this after talking with Rafael.}
+The DESC SRD \citep{Mandelbaum:2018} lists stringent requirements on how well the mean and variance of tomographic redshift bins must be known for each of the main DESC science cases.
+We indicate the Year 1 requirements assuming our true mean redshift of $z=0.701$ as dashed lines.
+In this study with representative training data, \annz, \bpz, \cmnn, \tpz, and our pathological \trainz\ estimator meet the Y1 requirement on the mean redshift.
+Only \annz, \bpz, \cmnn, and \trainz\ meet both requirements.
+One should be concerned that many codes fail to meet this ambitious limit under perfect prior information because all codes are anticipated to do no better under realistically imperfect prior information, and indicates that additional calibration to remove these systamatic offsets \citep[e.~g.~][]{Newman:08} will likely be necessary in order to meet these stringent goals.
 
-\begin{table}
-\setlength{\tabcolsep}{2pt}
-\caption{Moments of the stacked estimator $\hat{N}(z)$ of the redshift distribution.
-Most of the codes considered recover the moments of $\tilde{N}(z)$}
-\label{tab:moments}
-\begin{tabular}{lccc}
-\hline
-\hline
- \multicolumn{4}{l}{Moments of $\hat{N}(z)$} \\
-\hline
-Estimator  & mean       & variance    & skewness \\
-Empirical ``truth'' & 0.701 & 0.630 & 0.671  \\
-\hline
-\annz       & 0.702      & 0.625      & 0.653    \\
-\bpz        & 0.699      & 0.629      & 0.671    \\
-\delight    & 0.692      & 0.609      & 0.638    \\
-\eazy       & 0.681      & 0.595      & 0.619    \\
-\flexzboost & 0.694      & 0.610      & 0.631    \\
-\gpz        & 0.696      & 0.615      & 0.639    \\
-\lephare    & 0.718      & 0.668      & 0.741    \\
-\metaphor   & 0.705      & 0.628      & 0.657    \\
-\cmnn       & 0.701      & 0.628      & 0.667    \\
-\skynet     & 0.743      & 0.708      & 0.797    \\
-\tpz        & 0.700      & 0.619      & 0.643    \\
-\hline
-\trainz	    & 0.699 		 & 0.627 	    & 0.666 \\
-\end{tabular}
-\end{table}
+%\begin{table}
+%\setlength{\tabcolsep}{2pt}
+%\caption{Moments of the stacked estimator $\hat{N}(z)$ of the redshift distribution.
+%Most of the codes considered recover the moments of $\tilde{N}(z)$}
+%\label{tab:moments}
+%\begin{tabular}{lccc}
+%\hline
+%\hline
+% \multicolumn{4}{l}{Moments of $\hat{N}(z)$} \\
+%\hline
+%Estimator  & mean       & variance    & skewness \\
+%Empirical ``truth'' & 0.701 & 0.630 & 0.671  \\
+%\hline
+%\annz       & 0.702      & 0.625      & 0.653    \\
+%\bpz        & 0.699      & 0.629      & 0.671    \\
+%\delight    & 0.692      & 0.609      & 0.638    \\
+%\eazy       & 0.681      & 0.595      & 0.619    \\
+%\flexzboost & 0.694      & 0.610      & 0.631    \\
+%\gpz        & 0.696      & 0.615      & 0.639    \\
+%\lephare    & 0.718      & 0.668      & 0.741    \\
+%\metaphor   & 0.705      & 0.628      & 0.657    \\
+%\cmnn       & 0.701      & 0.628      & 0.667    \\
+%\skynet     & 0.743      & 0.708      & 0.797    \\
+%\tpz        & 0.700      & 0.619      & 0.643    \\
+%\hline
+%\trainz	    & 0.699 		 & 0.627 	    & 0.666 \\
+%\end{tabular}
+%\end{table}
 

### 2019-09-27 paper/tex/appendix.tex
typo in Newman reference

@@
 Only \annz, \bpz, \cmnn, and \trainz\ meet both requirements.
-One should be concerned that many codes fail to meet this ambitious limit under perfect prior information because all codes are anticipated to do no better under realistically imperfect prior information, and indicates that additional calibration to remove these systamatic offsets \citep[e.~g.~][]{Newman:08} will likely be necessary in order to meet these stringent goals.
+One should be concerned that many codes fail to meet this ambitious limit under perfect prior information because all codes are anticipated to do no better under realistically imperfect prior information, and indicates that additional calibration to remove these systamatic offsets \citep[e.~g.~][]{Newman:2008} will likely be necessary in order to meet these stringent goals.
 

### 2019-10-01 paper/tex/analysis.tex
added paragraph on implicit prior

@@
 \end{equation}
-conditioned not only on the photometric data $d_{i}$ for that galaxy but also on parameters encompassing a number of things that will differ depending on the method $H$ used to produce it, namely the often implicit assumptions $I_{H}$ necessary for the method to be valid and any inputs $I_{D}$ it takes as prior information, such as a template library or training set.
-Because of this, direct comparison of \pzpdf s produced by different methods is in some sense impossible; even if they share the same external prior information $I_{D}$, by definition they cannot be conditioned on the same assumptions $I_{H}$, otherwise they would not be distinct methods at all.
-We call $I_{H}$ the \textit{implicit prior} specific to the method, though some aspects of its nature may be discerned.
+conditioned not only on the photometric data $d_{i}$ for that galaxy but also on parameters encompassing prior information $I_{D}$ shared, in our experiment, among all \pzpdf\ codes and $I_{H}$ that will differ depending on the method $H$ used to produce it.
+To be concrete, $I_{D}$ takes the form of a training set for the machine learning codes and a template library for the model fitting codes.
 
+The interpretation of the information $I_{H}$ is more subtle.
+This investigation is built upon the knowledge that two codes taking the same approach, among choices of model fitting or machine learning, are nonetheless expected to yield different results even if they take the same external prior information $I_{D}$.
+$I_{H}$ represents the projection of the code's architecture onto the estimated posteriors over redshift, specific to each code, and even the tunable parameters or random seeds of a specific run of a code with a random component.
+We thus refer to $I_{H}$ as the \textit{implicit prior}, in contrast with the training set or template library provided to a given code explicitly by the researcher.
+
+The presence of the implicit prior in some sense makes a direct comparison of \pzpdf s produced by different methods impossible; even if they share the same external prior information $I_{D}$, by definition they cannot be conditioned on the same assumptions $I_{H}$, otherwise they would not be distinct methods at all.
 In this study, we isolate the effect of differences in prior information $I_{H}$ specific to each method by using a single training set $I_{D}^{\mathrm{ML}}$ for all machine learning-based codes and a single template library $I_{D}^{\mathrm{T}}$ for all template-based codes.

### 2019-10-10 paper/tex/appendix.tex
fix typos

--- 2019-09-27 paper/tex/appendix.tex
@@
 Only \annz, \bpz, \cmnn, and \trainz\ meet both requirements.
-One should be concerned that many codes fail to meet this ambitious limit under perfect prior information because all codes are anticipated to do no better under realistically imperfect prior information, and indicates that additional calibration to remove these systamatic offsets \citep[e.~g.~][]{Newman:08} will likely be necessary in order to meet these stringent goals.
+One should be concerned that many codes fail to meet this ambitious limit under perfect prior information because all codes are anticipated to do no better under realistically imperfect prior information, and indicates that additional calibration to remove these systematic offsets \citep[e.~g.~][]{Newman:2008} will likely be necessary in order to meet these stringent goals.
 
@@
 
-\skynet\ exhibits redshift bias in Figure~\ref{fig:nz} and is a clear outlier in the first moment of $\hat{N}(z)$ in Table~\ref{tab:moments}.
+\skynet\ exhibits redshift bias in Figure~\ref{fig:nz} and is a clear outlier in the first moment of $\hat{N}(z)$ in Figure~\ref{fig:moments}.
 The \skynet\ algorithm employs a random subsampling of the training set without testing that the subset is representative of the full population, and the implementation used here does not upweight rarer low- and high-redshift galaxies, as in \citet{Bonnett:15}, suggesting a possible cause that may be addressed in future work.

### 2019-10-10 paper/main.tex
incorporated feedback from @morriscb
auncomment line on internal review and add names

@@
 
-%This paper has undergone internal review in the LSST Dark Energy Science Collaboration. % REQUIRED if true
+This paper has undergone internal review in the LSST Dark Energy Science Collaboration, the internal reviewers were: Daniel Gruen, Markus Rau, and Michael Troxel. % REQUIRED if true
 
@@
 The authors would like to thank their LSST-DESC publication review committee for comments that improved the paper draft.
+% hack event acknowledgments are important
+%We thank Stony Brook University for hosting the Summer 2017 LSST-DESC Hack Day at which this work was partially completed.
 
 { \red{personal funding sources}}
-S.~Schmidt acknowledges support from DOE grant DE-SC0009999 and NSF/AURA grant N56981C.
+SJS acknowledges support from DOE grant DE-SC0009999 and NSF/AURA grant N56981C.
 AIM acknowledges support from the Max Planck Society and the Alexander von Humboldt Foundation in the framework of the Max Planck-Humboldt Research Award endowed by the Federal Ministry of Education and Research.
 During the completion of this work, AIM was advised by David W. Hogg and was supported by National Science Foundation grant AST-1517237.
+%JYHS is advised by Ofer Lahav and acknowledges support from the MyBrainSc Scholarship, provided by the Ministry of Education, Malaysia.
+JS would also like to thank Antonella Palmese and Boris Leistedt for guidance on the use of the algorithms ANNz2 and Delight respectively.
 

### 2019-10-10 paper/tex/analysis.tex
add sentence on implicit prior

@@
 $I_{H}$ represents the projection of the code's architecture onto the estimated posteriors over redshift, specific to each code, and even the tunable parameters or random seeds of a specific run of a code with a random component.
-We thus refer to $I_{H}$ as the \textit{implicit prior}, in contrast with the training set or template library provided to a given code explicitly by the researcher.
+We refer to $I_{H}$ as the \textit{implicit prior}, in contrast with the training set or template library provided to a given code explicitly by the researcher.
+In simple terms, the implicit prior is the collection of the many different assumptions, coding choices, algorithm selections, and other implementation details that are specific to each code, the ensemble of which results in differing estimates of redshift when combined with the data and prior information in common to all codes.
 

### 2019-10-14 paper/tex/analysis.tex
clarified description of CDF metrics

@@
 \end{equation}
-interpretable as the maximum difference between the CDFs of an approximating univariate distribution $\hat{f}(z)$ and a reference distribution $\tilde{f}(z)$, in this case $U(0,1)$.
+interpretable as the maximum difference between the CDFs of the empirical distribution of PIT values for the test sample $\hat{f}(z)$ and a reference distribution $\tilde{f}(z)$, in this case $U(0,1)$, for the ideal distribution of PIT values.
 % $CDF_{\rm phot}(z)$ and $CDF_{\rm spec}(z)$, the CDFs of the photo-$z$ and spectroscopic redshift respectively:

### 2019-10-14 paper/tex/analysis.tex
paragraph breaks

@@
 The overloaded ``$p(z)$'' is a widespread abuse of notation that obfuscates this goal, so we dedicate attention to dismantling it here.
+
 Galaxies have redshifts $z$ and photometric data $d$ drawn from a joint probability space $p(z, d)$ in nature.
@@
 There are a number of metrics that can be used to test the accuracy of a \pz\ posterior as an estimator of a true \pz\ posterior if the true \pzpdf\ is known.
-However, the true \pzpdf\ is in general not accessible unless the photometry is in fact drawn from a ground truth for the joint probability density of redshift and photometry $p(z, d)$.
-In contrast, existing mock catalogs produce redshift-photometry pairs $(z, d)$ by a deterministic algorithm that does not correspond to a joint probability density from which one can take samples.
+However, the true \pzpdf\ of observed data is not accessible, and existing mock catalogs produce redshift-photometry pairs $(z, d)$ by a deterministic algorithm that does not correspond to a joint probability density from which one can take samples.
 In these cases there is no ``true PDF'' for an individual object, and most measures of PDF fidelity will necessarily be restricted to probing the quality of the ensemble of \pzpdf s.

### 2019-10-31 paper/tex/analysis.tex
remove phrase on true likelihood

@@
 
-Galaxies have redshifts $z$ and photometric data $d$ drawn from a joint probability space $p(z, d)$ in nature.
-As a result, each observed galaxy $i$ has a \textit{true posterior \pzpdf}\ $p(z \vert d_{i})$ as well as a true likelihood $p(d \vert z_{i})$.
+Galaxies have redshifts $z$ and photometric data $d$ drawn from a joint probability space $p(z, d)$ in nature, and each observed galaxy $i$ has a \textit{true posterior \pzpdf}\ $p(z \vert d_{i})$.
 There are a number of metrics that can be used to test the accuracy of a \pz\ posterior as an estimator of a true \pz\ posterior if the true \pzpdf\ is known.

### 2019-11-12 paper/tex/results.tex
change line format on Figures 3 and A2 and modify legends, add plot script

@@
 \centering
-\includegraphics[width=0.74\textwidth]{fig/KSvsCvMvsAD_PIT_withbootstrap.jpg}
+\includegraphics[width=0.74\textwidth]{fig/KSvsCvMvsAD_PIT_withnull_jpg.jpg}
 \caption{A visualization of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the PIT distributions.
@@
 
-Of the metrics we were able to consider in this experiment, \textbf{the CDE Loss is the only metric that can appropriately penalize the pathological\trainz}.
+Of the metrics we were able to consider in this experiment, \textbf{the CDE Loss is the only metric that can appropriately penalize the pathological \trainz}.
 Additionally, it favors \cmnn\ and \flexzboost, the latter of which is optimized for this metric.

### 2019-11-12 paper/tex/data.tex
fix minor typos

@@
 Systematic inconsistencies with galaxy colors at $z > 2$ were observed, so the catalogue was limited to $0 < z \leq 2.0$.
-To obtain a catalogue matching the \lsst\ Gold Sample, we imposed an cut of $i < 25.3$, which gives a signal-to-noise ratio $\gtrsim 30$ for most galaxies.
+To obtain a catalogue matching the \lsst\ Gold Sample, we imposed a cut of $i < 25.3$, which gives a signal-to-noise ratio $\gtrsim 30$ for most galaxies.
 In order for statistical errors to be subdominant to the systematic errors we aim to probe, we further reduced the sample size to $<10^{7}$ galaxies by isolating $\sim 16.8$ square degrees selected from five separate spatial regions of the simulation.

### 2019-11-12 paper/tex/introduction.tex
fix minor typos

@@
 
-As an example, in order for \pz\ systematics to not dominate the statistical noise floor of \lsst's main cosmological sample of $\sim 10^{7}$ galaxies, the \lsst\ Science Requirements Document (SRD)\footnote{available at \url{https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17}} specifies that individual galaxy \pz s must have root-mean-square error $\sigma_z < 0.02 (1+z)$, $3 \sigma$ catastrophic outlier rate below $10\%$, and bias below $0.003$.
+As an example, in order for \pz\ systematics to not dominate the statistical noise floor of \lsst's main cosmological sample of several $10^{9}$ galaxies, the \lsst\ Science Requirements Document (SRD)\footnote{available at \url{https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17}} specifies that individual galaxy \pz s must have root-mean-square error $\sigma_z < 0.02 (1+z)$, $3 \sigma$ catastrophic outlier rate below $10\%$, and bias below $0.003$.
 Specific science cases may have their own requirements on \pz\ performance that exceed those of the survey as a whole.
-In that vein, the \lsst\ Dark Energy Science Collaboration (\lsstdesc) developed a separate SRD \citep{Mandelbaum:2018} that conservatively forecasts the constraining power of five cosmological probes, leading to even more stringent requirements on \pz\ performance, including those defined in terms of tomographically binned subsamples populations rather than individual galaxies.
+In that vein, the \lsst\ Dark Energy Science Collaboration (\lsstdesc) developed a separate SRD \citep{Mandelbaum:2018} that conservatively forecasts the constraining power of five cosmological probes, leading to even more stringent requirements on \pz\ performance, including those defined in terms of tomographically binned subsample populations rather than individual galaxies.
 

### 2019-11-13 paper/tex/introduction.tex

@@
 Specific science cases may have their own requirements on \pz\ performance that exceed those of the survey as a whole.
-In that vein, the \lsst\ Dark Energy Science Collaboration (\lsstdesc) developed a separate SRD \citep{Mandelbaum:2018} that conservatively forecasts the constraining power of five cosmological probes, leading to even more stringent requirements on \pz\ performance, including those defined in terms of tomographically binned subsample populations rather than individual galaxies.
+In that vein, the \lsst\ Dark Energy Science Collaboration (\lsstdesc) developed a separate SRD \citep{Mandelbaum:2018} that conservatively forecasts the constraining power of five cosmological probes, leading to even more stringent requirements on \pz\ performance, including those defined in terms of tomographically binned subsamples populations rather than individual galaxies.
 

### 2019-11-13 paper/tex/discussion.tex
elaborated on true PDFs

@@
 
-Section~\ref{sec:metrics} discussed the difficulty in evaluating PDF accuracy for individual objects.
+Section~\ref{sec:metrics} discussed the difficulty in evaluating PDF accuracy for individual objects with known $(z, d)$ information but without a known $p(z, d)$.
 In a follow-up study, we will
  % use generative adversarial networks\citep[(GANs)][]{Goodfellow:14} trained on the simulation data to
-generate ``true PDF'' distributions, yielding a dataset that enables a test of PDF accuracy for individual galaxies rather than solely ensembles.
+generate mock data probabilistically, yielding true PDFs in addition to true redshifts and photometric data.
+This future data set will enable tests of PDF accuracy for individual galaxies rather than solely ensembles.
 

### 2019-11-13 paper/tex/analysis.tex

@@
 Section~\ref{sec:qualmet} outlines aggregate metrics of a catalogue of \pzpdf s, and Section~\ref{sec:CDE_loss} presents a metric of individual \pzpdf s in the absence of true \pzpdf s.
-Though the outmoded practices should not be encouraged, those seeking a connection to previous comparison studies will find metrics of redshift point estimate reductions of \pzpdf s in Appendix~\ref{sec:pointmetrics} and metrics of a science-specific summary statistics heuristically derived from \pzpdf s in Appendix~\ref{sec:moments}.
+Those seeking a connection to previous comparison studies will find metrics of redshift point estimate reductions of \pzpdf s in Appendix~\ref{sec:pointmetrics} and metrics of a science-specific summary statistic heuristically derived from \pzpdf s in Appendix~\ref{sec:moments}.
 

### 2019-11-13 paper/tex/appendix.tex

@@
 Many of the codes, including all the model-fitting approaches and \annz, \gpz, \metaphor, and \skynet\ from the data-driven camp, overestimate the redshift density at $z \sim 1.4$.
-This behavior is a consequence of the $4000$ \AA break passing through the gap between the $z$ and $y$ filters, which induces a genuine discontinuity in the $z - y$ colour as a function of redshift that can sway the \pzpdf\ estimates in the absence of bluer spectral features.
+This behavior is a consequence of the $4000$ \AA\ break passing through the gap between the $z$ and $y$ filters, which induces a genuine discontinuity in the $z - y$ colour as a function of redshift that can sway the \pzpdf\ estimates in the absence of bluer spectral features.
 
@@
 \centering
-\includegraphics[width=0.74\textwidth]{fig/KSvsCvMvsAD_NZ_withbootstrap.jpg}
+\includegraphics[width=0.74\textwidth]{fig/KSvsCvMvsAD_NZ_withnull_jpg.jpg}
 \caption{A visualization of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the $\hat{N}(z)$ distributions.
@@
 %, particularly for \skynet.
-The features perpendicular to the $z_{phot} = z_{spec}$ line are due to the $4000$ \AA break passing through the gaps between adjacent filters.
+The features perpendicular to the $z_{phot} = z_{spec}$ line are due to the $4000$ \AA\ break passing through the gaps between adjacent filters.
 % and logically are more prominent for template-based codes

### 2019-11-13 paper/tex/data.tex
add ref to training sample size

@@
 
-Following the findings of \citet{Bernstein:10}, \citet{Masters:2017} that only $~10^{4}$ spectra are necessary to calibrate \pz s to Stage IV requirements, we aimed to set aside a randomly selected training set of $3-5\times 10^{4}$ galaxies, $\sim 10\%$ of the full sample.
+Following the findings of \citet{Bernstein:10}, \citet{Newman:2015}, and \citet{Masters:2017} that only $~10^{4}$ spectra are necessary to calibrate \pz s to Stage IV requirements, we aimed to set aside a randomly selected training set of $3-5\times 10^{4}$ galaxies, $\sim 10\%$ of the full sample.
 After all cuts described above, we designated the \textit{DC1 training set} of $44\,404$ galaxies for which observed photometry, true SEDs, and true redshifts would be provided to all codes and the blinded \textit{DC1 test set} of $399\,356$ galaxies for which photometry alone would be provided to all codes and \pzpdf s would be requested.

### 2019-11-14 paper/tex/appendix.tex
update moments plot to Y10 values

@@
 \centering
-\includegraphics[width=0.75\textwidth]{fig/moments_figure.jpg}
+\includegraphics[width=0.75\textwidth]{fig/momentsplot_10yr.jpg}
 \caption{Residuals of the first three moments of the stacked $\hat{N}(z)$ distribution.
-Red and blue horizontal lines indicate the Year 1 DESC SRD requirements on accuracy of the mean and variance respectively.
+Red and blue horizontal lines indicate the Year 10 DESC SRD requirements on accuracy of the mean and variance respectively.
 Only a small number of codes are able to meet these specifications even with perfect training data.
@@
 The DESC SRD \citep{Mandelbaum:2018} lists stringent requirements on how well the mean and variance of tomographic redshift bins must be known for each of the main DESC science cases.
-We indicate the Year 1 requirements assuming our true mean redshift of $z=0.701$ as dashed lines.
-In this study with representative training data, \annz, \bpz, \cmnn, \tpz, and our pathological \trainz\ estimator meet the Y1 requirement on the mean redshift.
+We indicate the Year 10 (Y10) requirements assuming our true mean redshift of $z=0.701$ as dashed lines.
+In this study with representative training data, \annz, \bpz, \cmnn, \tpz, and our pathological \trainz\ estimator meet the Y10 requirement on the mean redshift.
 Only \annz, \bpz, \cmnn, and \trainz\ meet both requirements.

### 2019-11-15 paper/tex/data.tex
fix missing tilde

--- 2019-11-12 paper/tex/data.tex
@@
 
-Following the findings of \citet{Bernstein:10}, \citet{Masters:2017} that only $~10^{4}$ spectra are necessary to calibrate \pz s to Stage IV requirements, we aimed to set aside a randomly selected training set of $3-5\times 10^{4}$ galaxies, $\sim 10\%$ of the full sample.
+Following the findings of \citet{Bernstein:10}, \citet{Newman:2015}, and \citet{Masters:2017} that only $\sim\!10^{4}$ spectra are necessary to calibrate \pz s to Stage IV requirements, we aimed to set aside a randomly selected training set of $3-5\times 10^{4}$ galaxies, $\sim 10\%$ of the full sample.
 After all cuts described above, we designated the \textit{DC1 training set} of $44\,404$ galaxies for which observed photometry, true SEDs, and true redshifts would be provided to all codes and the blinded \textit{DC1 test set} of $399\,356$ galaxies for which photometry alone would be provided to all codes and \pzpdf s would be requested.

### 2019-11-16 paper/tex/appendix.tex
remove BPZ from list meeting reqs

@@
 We indicate the Year 10 (Y10) requirements assuming our true mean redshift of $z=0.701$ as dashed lines.
-In this study with representative training data, \annz, \bpz, \cmnn, \tpz, and our pathological \trainz\ estimator meet the Y10 requirement on the mean redshift.
-Only \annz, \bpz, \cmnn, and \trainz\ meet both requirements.
+In this study with representative training data, \annz, \cmnn, \tpz, and our pathological \trainz\ estimator meet the Y10 requirement on the mean redshift.
+Only \annz, \cmnn, and \trainz\ meet both requirements.
 One should be concerned that many codes fail to meet this ambitious limit under perfect prior information because all codes are anticipated to do no better under realistically imperfect prior information, and indicates that additional calibration to remove these systematic offsets \citep[e.~g.~][]{Newman:2008} will likely be necessary in order to meet these stringent goals.

### 2019-11-18 paper/tex/appendix.tex
clarify chippr reference and move to footnote

@@
 \end{equation}
-is the most widely accepted method for obtaining $\hat{N}^{H}(z)$ as an estimator of the redshift distribution from \pzpdf s derived by a method $H$.
-Though the use of the stacked estimator of the redshift distribution is not formally correct \citep{Malz:chippr}, we use it under the untested assumption that the response of our metrics of $\hat{N}^{H}(z)$ will be analogous to the same metrics applied to a principled estimator of the redshift distribution.
+ is the most widely used method for obtaining $\hat{N}^{H}(z)$ as an estimator of the redshift distribution from \pzpdf s derived by a method $H$.
+While the stacked estimator of the redshift distribution violates the mathematical definition of statistical independence and is thus not formally correct\footnote{\citet{Malz:chippr} shows how the stacking procedure can lead to bias in the estimate of $N(z)$ and presents a principled alternative to this commonly employed method.}, we use it as a basis for comparison of \pzpdf\ methods under the untested assumption that the response of our metrics of $\hat{N}^{H}(z)$ will be analogous to the same metrics applied to a principled estimator of the redshift distribution.
+
+%is the most widely accepted method for obtaining $\hat{N}^{H}(z)$ as an estimator of the redshift distribution from \pzpdf s derived by a method $H$.
+This procedure is not formally correct, basic statistics tells us that joint probabilities are not formed by adding probabilities.
+\citet{Malz:chippr} shows how the stacking procedure can lead to bias in redshift estimates, and develop a principled alternative to this commonly employed method.
+However, while  the use of the stacked estimator of the redshift distribution is not formally correct, we use it under the untested assumption that the response of our metrics of $\hat{N}^{H}(z)$ will be analogous to the same metrics applied to a principled estimator of the redshift distribution.
 

### 2019-11-21 paper/tex/analysis.tex
fix most of typos pointed out by Seth

@@
 There are a number of metrics that can be used to test the accuracy of a \pz\ posterior as an estimator of a true \pz\ posterior if the true \pzpdf\ is known.
-However, the true \pzpdf\ of observed data is not accessible, and existing mock catalogs produce redshift-photometry pairs $(z, d)$ by a deterministic algorithm that does not correspond to a joint probability density from which one can take samples.
+However, the true \pzpdf\ of observed data is not accessible, and existing mock catalogues produce redshift-photometry pairs $(z, d)$ by a deterministic algorithm that does not correspond to a joint probability density from which one can take samples.
 In these cases there is no ``true PDF'' for an individual object, and most measures of PDF fidelity will necessarily be restricted to probing the quality of the ensemble of \pzpdf s.
@@
 
-Because \lsst's \pzpdf s will be used for many scientific applications, some of which require accuracy of each individual catalog entry, we consider several metrics that probe the population-level performance of the \pzpdf s.
+Because \lsst's \pzpdf s will be used for many scientific applications, some of which require each individual catalogue entry to be accurate, we consider several metrics that probe the population-level performance of the \pzpdf s.
 As we have the true redshifts but not true \pzpdf s for comparison, we remind the reader of the Cumulative Distribution Function (CDF)
@@
 The PIT distribution has been used to quantify the performance of \pzpdf\ methods in the past \citep[e.~g.~][]{Bordoloi:10,Polsterer:16,Tanaka:17}.
-\citet{Tanaka:17} use the histogram of PIT values as a diagnostic indicator of overall code performance, while \citet{Freeman:17} independently define the PIT and demonstrate how its individual values may be used both to perform hypothesis testing (via, e.~g.~ the KS, CvM, and AD tests; see below) and to construct quantile-quantile plots.
+\citet{Tanaka:17} use the histogram of PIT values as a diagnostic indicator of overall code performance, while \citet{Freeman:17} independently define the PIT and demonstrate how its individual values may be used both to perform hypothesis testing (via, e.~g.~ the KS, CvM, and AD tests; see below) and to construct QQ plots.
 %\red{We might want to divide this by redshift to get trends, as the overall PIT could be impacted by true n(z).

### 2019-11-21 paper/tex/appendix.tex
fix most of typos pointed out by Seth

@@
 The stacked estimators of the redshift distribution for \cmnn\ and \trainz\ best estimate $\tilde{N}(z)$ under these metrics, whereas \eazy, \lephare, \metaphor, and \skynet\ underperform; \bpz, \gpz, and \tpz\ are within a factor of two of the conservative limit for all statistics.
-It is unsurprising that \cmnn\ scores well, as with a nearly complete and representative training set choosing neighbouring points in color/magnitude space to construct an estimator should lead to excellent agreement in the final $\hat{N}(z)$.
+It is unsurprising that \cmnn\ scores well, as with a nearly complete and representative training set choosing neighbouring points in colour/magnitude space to construct an estimator should lead to excellent agreement in the final $\hat{N}(z)$.
 
@@
 \item bias $b_{z} < 0.003$ %\scc{in page 518 of the SB is clearly reported that the mean is expected to be less than 0.003 in our case we have defined as bias the median and not the mean }
-\item catastrophic outlier rate $f_{\mathrm{out}} < 10\%$ %\scc{clearly the number of outliers since are defined above $3\sigma$ strictly depends on the sigma definition, in the science book it seems to refer to the RMS of the $e_z$ distribution even ef it could be the standard deviation, but again I don't think it refers to IQR, therefore we could not consider exactly this numbers as reference.}
+\item catastrophic outlier rate $f_{\mathrm{out}} < 10$ per cent %\scc{clearly the number of outliers since are defined above $3\sigma$ strictly depends on the sigma definition, in the science book it seems to refer to the RMS of the $e_z$ distribution even ef it could be the standard deviation, but again I don't think it refers to IQR, therefore we could not consider exactly this numbers as reference.}
 %\red{I am going off of conversations with Zeljko and how we {\it actually} computed the metrics in the Science Book (I have the scripts).

### 2019-11-21 paper/tex/data.tex
fix most of typos pointed out by Seth

--- 2019-11-12 paper/tex/data.tex
@@
 The lightcone was constructed from smoothing and interpolation between a set of time snapshots.
-Dark matter halos were identified using the \texttt{Rockstar} software package \citep{Behroozi:13} and then populated with galaxies with a stellar mass and absolute $r$-band magnitude in the \sdss\ system determined using a sub-halo abundance matching model constrained to match both projected two-point galaxy clustering statistics and an observed conditional stellar mass function \citep{Reddick:13}.
+Dark matter halos were identified using the \texttt{Rockstar} software package \citep{Behroozi:13} and then populated with galaxies with stellar masses and absolute $r$-band magnitudes in the \sdss\ system determined using a sub-halo abundance matching model constrained to match both projected two-point galaxy clustering statistics and an observed conditional stellar mass function \citep{Reddick:13}.
 
 To assign a spectrum to each galaxy, the Adding Density Dependent Spectral Energy Distributions (SEDs) procedure (\texttt{ADDSEDS}, deRose in prep.)\footnote{\url{https://github.com/vipasu/addseds}} was used.
-\texttt{ADDSEDS} uses a sample of $\sim 5\times 10^{5}$ galaxies from the magnitude-limited \sdss\ Data Release 6 Value Added Galaxy Catalogue \citep{Blanton:05} to train an empirical relation between absolute $r$-band magnitude, local galaxy density, and SED.
+\texttt{ADDSEDS} uses a sample of $\sim 5\times 10^{5}$ galaxies from the magnitude-limited \sdss\ Data Release 6 Value Added Galaxy Catalogue \citep[NYU-VAGC,][]{Blanton:05} to train an empirical relation between absolute $r$-band magnitude, local galaxy density, and SED.
 Each \sdss\ spectrum is parameterized by five weights corresponding to a weighted sum of five basis SED components using the \texttt{k-correct} software package\footnote{\url{http://kcorrect.org}} \citep{Blanton:07}.
 
-Correlations between SED and galaxy environment were included so as to preserve the colour-density relation of galaxy environment.
+Correlations between SED and galaxy environment were included so as to preserve the colour-density relation of galaxy environments.
 The distance to the spatially projected fifth-nearest neighbour was used as a proxy for local density in the \sdss\ training sample.
@@
 
-As a consequence of adding Gaussian-distributed photometric errors, 2.0\% of our galaxies exhibit a negative flux in one or more bands, the vast majority of which are in the $u$-band.
+As a consequence of adding Gaussian-distributed photometric errors, 2.0 per cent of our galaxies exhibit a negative flux in one or more bands, the vast majority of which are in the $u$-band.
 We deem such negative fluxes \textit{non-detections} and assign a placholder magnitude of 99.0 in the catalogue to indicate to the \pzpdf\ codes that such galaxies would be ``looked at but not seen'' in multi-band forced photometry.
@@
 % NOW MOTIVATE THE SHARED PRIOR INFORMATION
-For the purpose of performing a controlled experiment that compares \pzpdf\ codes on equal footing as a baseline for a future sensitivity analysis, we take care to provide each with maximally optimistic prior information.
+For the purpose of performing a controlled experiment that compares \pzpdf\ codes on equal footing as a baseline for a future sensitivity analysis, we take care to provide each with optimal prior information.
 Redshift estimation approaches built upon physical modeling and machine learning alike have a notion of prior information considered beyond the photometry of the data for which redshift is to be constrained: that information is derived from a template library for a model-based code and a training set for a data-driven code.
@@
 
-Following the findings of \citet{Bernstein:10}, \citet{Masters:2017} that only $~10^{4}$ spectra are necessary to calibrate \pz s to Stage IV requirements, we aimed to set aside a randomly selected training set of $3-5\times 10^{4}$ galaxies, $\sim 10\%$ of the full sample.
+Following the findings of \citet{Bernstein:10}, \citet{Newman:2015}, and \citet{Masters:2017} that only $\sim\!10^{4}$ spectra are necessary to calibrate \pz s to Stage IV requirements, we aimed to set aside a randomly selected training set of $3-5\times 10^{4}$ galaxies, $\sim 10$ per cent of the full sample.
 After all cuts described above, we designated the \textit{DC1 training set} of $44\,404$ galaxies for which observed photometry, true SEDs, and true redshifts would be provided to all codes and the blinded \textit{DC1 test set} of $399\,356$ galaxies for which photometry alone would be provided to all codes and \pzpdf s would be requested.

### 2019-11-21 paper/tex/discussion.tex
fix most of typos pointed out by Seth

@@
 %(Sam Schmidt, Ofer Lahav, Jeff Newman, Alex Malz, Eve Kovacs, Tony Tyson, Tina Peters)
-
-% If methods can not reach the goals on idealized data, then they will almost surely not meet those same goals when the more complex problems that we expect to arise from real LSST data are included.
-% The results presented in this paper enable an evaluation of which algorithms are the most promising moving forward, and potentially point to avenues for improvement.
-
 In contrast with other \pzpdf\ comparison papers that have aimed to identify the ``best'' code for a given survey, we have focused on the somewhat more philosophical questions of how to assess \pzpdf\ methods and how to interpret differences between codes in terms of \pzpdf\ performance.
@@
 
-This initial paper explores \pzpdf\ code performance in idealized conditions with perfect catalog-based photometry and representative training data, but the resilience of each code to such realistic imperfections in prior information has not yet been evaluated.
+This initial paper explores \pzpdf\ code performance in idealized conditions with perfect catalogue-based photometry and representative training data, but the resilience of each code to such realistic imperfections in prior information has not yet been evaluated.
 A top priority for a follow-up study is to test realistic forms of incomplete, erroneous, and non-representative template libraries and training sets as well as the impact of other forms of external priors that must be ingested by the codes, major concerns in \citet{Newman:2015, Masters:2017}.
@@
 %\item \red{inclusion of photometric errors, image based effects, blending, lensing, sky, mask, etc...}
-This initial paper explored a data set that was constructed at the catalog level, with no inclusion of the complications that come from measuring photometry from images.
-Future data challenges will move to catalogs constructed from mock images, including the complications of deblending, sensor inefficiencies, and heterogeneous observing conditions, all anticipated to affect the measured colours of \lsst's galaxy sample \citep{Dawson:2016}.
+This initial paper explored a data set that was constructed at the catalogue level, with no inclusion of the complications that come from measuring photometry from images.
+Future data challenges will move to catalogues constructed from mock images, including the complications of deblending, sensor inefficiencies, and heterogeneous observing conditions, all anticipated to affect the measured colours of \lsst's galaxy sample \citep{Dawson:2016}.
 

### 2019-11-21 paper/tex/introduction.tex
fix most of typos pointed out by Seth

--- 2019-11-12 paper/tex/introduction.tex
@@
 
-As an example, in order for \pz\ systematics to not dominate the statistical noise floor of \lsst's main cosmological sample of several $10^{9}$ galaxies, the \lsst\ Science Requirements Document (SRD)\footnote{available at \url{https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17}} specifies that individual galaxy \pz s must have root-mean-square error $\sigma_z < 0.02 (1+z)$, $3 \sigma$ catastrophic outlier rate below $10\%$, and bias below $0.003$.
+As an example, in order for \pz\ systematics to not dominate the statistical noise floor of \lsst's main cosmological sample of several $10^{9}$ galaxies, the \lsst\ Science Requirements Document (SRD)\footnote{available at \url{https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17}} specifies that individual galaxy \pz s must have root-mean-square error $\sigma_z < 0.02 (1+z)$, $3 \sigma$ catastrophic outlier rate below $10$ per cent, and bias below $0.003$.
 Specific science cases may have their own requirements on \pz\ performance that exceed those of the survey as a whole.
@@
 
-To fully characterize such physical degeneracies, later photometric galaxy catalogue data releases, \citep[e.~g.~][]{Sheldon:2012, Erben:2013, de_Jong:17}, provide a more informative \pz\ data product, the \pz\ probability density function (PDF), that describes the redshift probability, commonly denoted as $p(z)$, as a function of a galaxy's redshift, conditioned on the observed photometry.
+To fully characterize such physical degeneracies, subsequent photometric galaxy catalogue data releases, \citep[e.~g.~][]{Sheldon:2012, Erben:2013, de_Jong:17}, provide a more informative \pz\ data product, the \pz\ probability density function (PDF), that describes the redshift probability, commonly denoted as $p(z)$, as a function of a galaxy's redshift, conditioned on the observed photometry.
 Early template-based methods such as \citet{Fernandezsoto:99} approximated the likelihood of photometry conditioned on redshift with the relative \chisq\ values of template spectra.
@@
 In the absence of simulated data drawn from known redshift distributions, the very concept of a ``true PDF'' for an individual galaxy is unavailable, and we must instead rely on measures of ensemble behaviour to characterize PDF quality (see \S\,\ref{sec:metrics} for further discussion).
-We aim to perform a comprehensive sensitivity analysis of \pzpdf\ techniques in order to ultimately select those that will become part of the \lsst\ pipelines, as part of a key project of the Photometric Redshifts working group of the \lsstdesc, described in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}}.
+We aim to perform a comprehensive sensitivity analysis of \pzpdf\ techniques in order to ultimately select those that will become part of the \lsstdesc\ pipelines, described in the Science Roadmap (SRM)\footnote{Available at: \url{https://lsstdesc.org/assets/pdf/docs/DESC_SRM_latest.pdf}}.
 In this initial study, we focus on evaluating the performance of \pzpdf\ codes using PDF-specific performance metrics in a formally controlled experiment with complete and representative prior information (template libraries and training sets) to set a baseline for subsequent investigations.

### 2019-11-21 paper/tex/methods.tex
commenting out confusing line about implicit prior
fix most of typos pointed out by Seth

@@
 
-Here we summarize the twelve \pzpdf\ codes compared in this study, also in Table~\ref{tab:list_of_codes}, which include both established and emerging approaches in template fitting and machine learning.
+Here we summarize the twelve \pzpdf\ codes compared in this study, listed in Table~\ref{tab:list_of_codes}, which include both established and emerging approaches in template fitting and machine learning.
 Though not exhaustive, this sample represents codes for which there was sufficient expertise within the \lsstdesc\ Photometric Redshifts Working Group.
@@
 Another key difference is the treatment of non-detections in one or more bands: some codes ignore incomplete bands, while others replace the value with either an estimate for the detection limit, the mean of other values in the training set, or another default value.
-There are varying conventions among machine learning based codes for treatment of non-detections, and no one prescription dominates in the \pz\ literature.
-The specific choices for each code affect the results and contribute to the implicit prior influencing their output.
+There are varying conventions among machine learning-based codes for treatment of non-detections, and no one prescription dominates in the \pz\ literature.
 However, we remind the reader that only 2.0 per cent of our sample has non-detections, almost exclusively in the $u$-band, and thus should not dominate the code performance differences.
+%The specific choices for each code affect the results and contribute to the implicit prior influencing their output.
+
 The authors welcome interest from those outside \lsstdesc\ to have their codes assessed in future investigations that build upon this one.
@@
 \noindent where $A$ is a normalization factor, $F^i_{\mathrm{pred}}(T,z)$ is the flux predicted for a template $T$ at redshift $z$, $F^i_{\mathrm{obs}}$ is the observed flux in a given band $i$, $\sigma^i_{\mathrm{obs}}$ is the observed flux error, and $N_{\mathrm filt}$ is the total number of filters, in our case the six $ugrizy$ LSST filters.
-The likelihood is a sum of observed flux error $\sigma_{b}^{\mathrm{obs}}$-weighted squared differences between the observed flux $F^{\mathrm{obs}}_{b}$ and the normalized predicted flux $F^{\mathrm{mod}}_{b}(T, z)$ in $N_{\mathrm{filt}}$ photometric filters $b$, which is the \lsst\ $ugrizy$ filters in this case.
+The likelihood is a sum of observed flux error $\sigma_{b}^{\mathrm{obs}}$-weighted squared differences between the observed flux $F^{\mathrm{obs}}_{b}$ and the normalized predicted flux $F^{\mathrm{mod}}_{b}(T, z)$ in $N_{\mathrm{filt}}$ photometric filters $b$, which are the \lsst\ $ugrizy$ filters in this case.
 Specific implementation details of each code, e.~g.~prior form and implementation, are described below.
@@
 
-Bayesian Photometric Redshift \citep[\bpz\footnote{\url{http://www.stsci.edu/~dcoe/BPZ/}},][]{Benitez:00} determines the likelihood $p(C \vert z, T)$ of a galaxy's observed colours $C$ for a set of SED templates $T$ at redshifts $z$.
+Bayesian Photometric Redshift \citep[\bpz,][]{Benitez:00} determines the likelihood $p(C \vert z, T)$ of a galaxy's observed colours $C$ for a set of SED templates $T$ at redshifts $z$.
 The \bpz\ likelihood is related to the \chisq\ likelihood by $p(C \vert z, T) \propto \exp[- \chi^{2} / 2]$.
@@
 %For photo-$z$ point estimates we use the \texttt{Z\_B} output parameter.
-Prior to running the code, the non-detection placeholder magnitude was replaced with an estimate of the one-$\sigma$ detection limit for the undetected band as a proxy for a value close to the estimated sky noise threshold.
+Prior to running the code, the non-detection placeholder magnitude was replaced with an estimate of the 1-$\sigma$ detection limit for the undetected band as a proxy for a value close to the estimated sky noise threshold.
 
@@
 
-Easy and Accurate Photometric Redshifts from Yale \citep[\eazy\footnote{\url{https://github.com/gbrammer/eazy-photoz}},][]{Brammer:08} extends the basic \chisq\ fit procedure that defines template-fitting approaches.
+Easy and Accurate Photometric Redshifts from Yale \citep[\eazy,][]{Brammer:08} extends the basic \chisq\ fit procedure that defines template-fitting approaches.
 The algorithm models the observed photometry with a linear combination of template SEDs at each redshift.
@@
 
-Photometric Analysis for Redshift Estimate \citep[\lephare\footnote{\url{http://www.cfht.hawaii.edu/~arnouts/lephare.html}},][]{Arnouts:99,Ilbert:06} uses the $\chi^2$ of Equation~\ref{eq_temp_chi} to match observed colors with those predicted from a template set.
+Photometric Analysis for Redshift Estimate \citep[\lephare,][]{Arnouts:99,Ilbert:06} uses the $\chi^2$ of Equation~\ref{eq_temp_chi} to match observed colours with those predicted from a template set.
 The template set can be semi-empirical or entirely synthetic.
@@
 
-\annz \footnote{\url{https://github.com/IftachSadeh/ANNZ}} \citep{Sadeh:16} supports several machine learning algorithms, including artificial neural networks (ANN), boosted decision tree, and k-nearest neighbour (KNN) regression.
+\annz \citep{Sadeh:16} supports several machine learning algorithms, including artificial neural networks (ANN), boosted decision tree, and k-nearest neighbour (KNN) regression.
 In addition to accounting for errors on the input photometry, \annz\ uses the KNN-uncertainty estimate of \citet{Oyaizu:08} to quantify uncertainty in the choice of method over multiple runs.
@@
 The colour-matched nearest-neighbours photometric redshift estimator \citep[\cmnn,][]{Graham:17} uses a training set of galaxies with known redshifts that has equivalent or better photometry than the test set in terms of quality and filter coverage.
-For each galaxy in the test set, \cmnn\ identifies a colour-matched subset of training galaxies using a threshold in the Mahalanobis distance $D_M = \sum_{j}^{N_{\mathrm{colours}}} (c^{\mathrm{train}}_{j} - c^{\mathrm{test}}_{j})^{2} / \delta c_{\mathrm{test}}^2$ in the space of available colours $c$, with colour measurement errors $\delta c_{\mathrm{test}}$ and $N_{\mathrm{colours}} = 5$ colors $j$ defined by the $ugrizy$ filters, which defines the set of colour-matched neighbours based on a value of the percent point function (PPF).
-As an example, for $N_{\mathrm{filt}}=5$ with PPF$=0.95$, $95\%$ of all training galaxies consistent with the test galaxy will have $D_M < 11.07$.
+For each galaxy in the test set, \cmnn\ identifies a colour-matched subset of training galaxies using a threshold in the Mahalanobis distance $D_M = \sum_{j}^{N_{\mathrm{colours}}} (c^{\mathrm{train}}_{j} - c^{\mathrm{test}}_{j})^{2} / \delta c_{\mathrm{test}}^2$ in the space of available colours $c$, with colour measurement errors $\delta c_{\mathrm{test}}$ and $N_{\mathrm{colours}} = 5$ colours $j$ defined by the $ugrizy$ filters, which defines the set of colour-matched neighbours based on a value of the percent point function (PPF).
+As an example, for $N_{\mathrm{filt}}=5$ with PPF$=0.95$, $95$ per cent of all training galaxies consistent with the test galaxy will have $D_M < 11.07$.
 Undetected bands are dropped, thereby reducing the effective $N_{\mathrm{filt}}$ for that galaxy.
@@
 We make an additional adaptation to enable the \cmnn\ algorithm to yield accurate \pzpdf s for all galaxies, as the original \citet{Graham:17} algorithm is optimized for \pz\ point estimates and is susceptible to less accurate \pzpdf s for bright galaxies or those with few matches in colour-space.
-We use PPF$=0.95$ rather than PPF$=0.68$ to generate the subset of colour-matched training galaxies, whose redshifts are weighted by their inverse Mahalanobis distances of the when composing the \pzpdf\ rather than weighting all colour-matched training galaxies equally.
-Additionally, when the number of colour-matched training set galaxies is less than 20, the nearest 20 neighbours in color-space are used instead, and the output \pzpdf\ is convolved with a Gaussian kernel of variance $\sigma_{\mathrm{ train}}^{2}(\mathrm{PPF}_{20}/0.95)^2 -1$ to account for the corresponding growth of the effective PPF to include 20 neighbors.
+We use PPF$=0.95$ rather than PPF$=0.68$ to generate the subset of colour-matched training galaxies, whose redshifts are weighted by their inverse Mahalanobis distances when composing the \pzpdf\ rather than weighting all colour-matched training galaxies equally.
+Additionally, when the number of colour-matched training set galaxies is less than 20, the nearest 20 neighbours in colour-space are used instead, and the output \pzpdf\ is convolved with a Gaussian kernel of variance $\sigma_{\mathrm{ train}}^{2}(\mathrm{PPF}_{20}/0.95)^2 -1$ to account for the corresponding growth of the effective PPF to include 20 neighbors.
 
@@
 
-\delight \footnote{\url{https://github.com/ixkael/Delight}} \citep{Leistedt:17} is a hybrid technique that infers \pz s with a data-driven model of latent SEDs and a physical model of photometric fluxes as a function of redshift.
-Generally, machine learning methods rely on representative training data with shared photometric filters, while template based methods rely on a complete library of templates based on physical models constructed.
+\delight \citep{Leistedt:17} is a hybrid technique that infers \pz s with a data-driven model of latent SEDs and a physical model of photometric fluxes as a function of redshift.
+Generally, machine learning methods rely on representative training data with shared photometric filters, while template-based methods rely on a complete library of templates based on physical models constructed.
 \delight\ aims to take the best aspects of both approaches by constructing a large collection of latent SED templates (or physical flux-redshift models) from training data, with a template SED library as a guide to the learning of the model, thereby circumventing the machine learning prerequisite of representative training data in the same photometric bands and the template fitting requirement of detailed galaxy SED models.
@@
 
-In this study, the default settings of \delight\ were used, with the exception that the PDF bins were set to be linearly-spaced rather than logarithmic.
+In this study, the default settings of \delight\ were used, with the exception that the PDF bins were set to be linearly spaced rather than logarithmically.
 The Gaussian process was trained using the full DC1 training set.
@@
 
-\flexzboost \footnote{\url{https://github.com/tpospisi/flexcode};\\ \url{https://github.com/rizbicki/FlexCoDE} \label{flexzboost_github}} \citep{Izbicki:17} is built on \texttt{FlexCode}, a general-purpose methodology for converting any conditional mean point estimator of $z$ to a conditional density estimator $p(z \vert \x) \equiv f(z \vert \x)$, where $\x$ here represents our photometric covariates and errors.
+\flexzboost \citep{Izbicki:17} is built on \texttt{FlexCode}, a general-purpose methodology for converting any conditional mean point estimator of $z$ to a conditional density estimator $p(z \vert \x) \equiv f(z \vert \x)$, where $\x$ here represents our photometric covariates and errors.
 \flexzboost\ expands the unknown function $f(z \vert \x) = \sum_{i}\beta_{i}(\x)\phi_{i}(z)$ using an orthonormal basis $\{\phi_{i}(z)\}_{i}$.
@@
 
-In this paper, we used \texttt{xgboost} \citep{Chen:16} for the regression; it should, however, be noted that \texttt{FlexCode-RF}\footref{flexzboost_github}, based on Random Forests, generally performs better on smaller datasets.
+In this paper, we used \texttt{xgboost} \citep{Chen:16} for the regression; it should, however, be noted that \texttt{FlexCode-RF}\footnote{\url{https://github.com/tpospisi/flexcode};\\ \url{https://github.com/rizbicki/FlexCoDE} \label{flexzboost_github}}, based on Random Forests, generally performs better on smaller datasets.
 As our basis $\phi_{i}(z)$, we choose a standard Fourier basis.
@@
 Both $I$ and $\alpha$ were chosen in an automated way by minimizing the weighted $L_2$-loss function \citep[Eq.
-5 in][]{Izbicki:17} on a validation set comprised of a randomly selected 15\% of the DC1 training set.
-While \texttt{FlexCode}'s lossless native encoding stores each \pzpdf\ using the basis coefficients $\beta_{i}(\x)$, we discretized the final estimates into 200 linearly-spaced redshift bins $0 < z < 2$ to match the consistent output format of the experimental conditions.
+5 in][]{Izbicki:17} on a validation set comprised of a randomly selected 15 per cent of the DC1 training set.
+While \texttt{FlexCode}'s lossless native encoding stores each \pzpdf\ using the basis coefficients $\beta_{i}(\x)$, we discretized the final estimates into 200 linearly spaced redshift bins $0 < z < 2$ to match the consistent output format of the experimental conditions.
 
@@
 
-\gpz \footnote{\url{https://github.com/OxfordML/GPz}} \citep{Almosallam:16a,Almosallam:15b} is a sparse Gaussian process based code, a scalable approximation of full Gaussian Processes \citep{Rasmussen:06}, that produces input-dependent variance estimates corresponding to heteroscedastic noise.
+\gpz \citep{Almosallam:16a,Almosallam:15b} is a sparse Gaussian process-based code, a scalable approximation of full Gaussian Processes \citep{Rasmussen:06}, that produces input-dependent variance estimates corresponding to heteroscedastic noise.
 The model assumes a Gaussian posterior probability $p(z \vert \x) = \mathcal{N}\left(z \vert \mu(\x), \sigma(\x)^{2}\right)$ of the output redshift $z$ given the input photometry $\x$.
-The mean $\mu(\x)$ and the variance $\sigma(\x)^{2}$ are modeled as functions $f(\x) = \sum_{i=1}^{m}w_{i}\phi_{i}(\x)$ linear combinations of $m$ basis functions $\left\{\phi_{i}(\x)\right\}_{i=1}^{m}$ with associated weights $\left\{w_{i}\right\}_{i=1}^{m}$.
+The mean $\mu(\x)$ and the variance $\sigma(\x)^{2}$ are modeled as functions $f(\x) = \sum_{i=1}^{m}w_{i}\phi_{i}(\x)$ that are linear combinations of $m$ basis functions $\left\{\phi_{i}(\x)\right\}_{i=1}^{m}$ with associated weights $\left\{w_{i}\right\}_{i=1}^{m}$.
 % Basis function models, for specific classes of basis functions such as the sigmoid or the squared exponential, have the advantage of being universal approximators, i.e.
@@
 \gpz's variance estimate is composed of a model uncertainty term corresponding to sparsity of the training set photometry and a noise uncertainty term encompassing noisy photometric observations, enabling quantification of any need for more representative or more precise training samples.
-\gpz\ may also weight training set samples by importance according to $|z_{\mathrm{spec}} - z_{\mathrm{phot}}| / (1+z_{\mathrm{spec}})$ to minimize the normalized \pz\ point estimate error, however, this function may be adapted to \pzpdf s, pressuring the model to dedicate more resources to test set galaxies that are not well-represented in the training set.
+\gpz\ may also weight training set samples by importance according to $|z_{\mathrm{spec}} - z_{\mathrm{phot}}| / (1+z_{\mathrm{spec}})$ to minimize the normalized \pz\ point estimate error.
+However, this function may be adapted to \pzpdf s, pressuring the model to dedicate more resources to test set galaxies that are not well-represented in the training set.
 
-To smooth the long tail in the distribution of magnitude errors, we use the log of the magnitude errors, improving numerical stability and eliminating the need for constraints on the optimization process.
+To smooth the long tail in the distribution of magnitude errors, we use the logarithm of the magnitude errors, improving numerical stability and eliminating the need for constraints on the optimization process.
 % We use principaphoto-\mathinhead{$z$}{z} interim posteriorsl component analysis (PCA) to decorrelate the data
@@
 This is the optimal expected value of the unobserved variables given the observed ones under the assumption that the distribution is jointly Gaussian; note that this reduces to a simple average if the covariates are independent with $\Sigma_{\mathrm{uo}} = 0$.
-We reserved for validation 20\% of the training set and used the Variable Covariance option in \gpz\ with 200 basis functions (see \citet{Almosallam:15b} for details), neglecting to apply cost-sensitive learning options.
+We reserved for validation 20 per cent of the training set and used the Variable Covariance option in \gpz\ with 200 basis functions (see \citet{Almosallam:15b} for details), and did not apply cost-sensitive learning options.
 
@@
 
-Machine-learning Estimation Tool for Accurate Photometric Redshifts \citep[\metaphor\footnote{\url{http://dame.dsf.unina.it}},][]{Cavuoti:17} is based on the Multi Layer Perceptron with Quasi Newton Algorithm (MLPQNA) with the least square error model and Tikhonov $L_{2}$-norm regularization \citep{Hofmann:18}.
+Machine-learning Estimation Tool for Accurate Photometric Redshifts \citep[\metaphor,][]{Cavuoti:17} is based on the Multi Layer Perceptron with Quasi Newton Algorithm (MLPQNA) with the least square error model and Tikhonov $L_{2}$-norm regularization \citep{Hofmann:18}.
 %, already validated on photo-$z$'s in several cases \citep{de_Jong:17,Cavuoti:17b,Cavuoti:15,Brescia:14,Brescia:13,Biviano:13}.
@@
 
-\skynet \footnote{\url{http://ccpforge.cse.rl.ac.uk/gf/project/skynet/}} \citep{Graff:14} employs a neural network based on a second order conjugate gradient optimization scheme \citep[see][for further details]{Graff:14}.
+\skynet \citep{Graff:14} employs a neural network based on a second-order conjugate gradient optimization scheme \citep[see][for further details]{Graff:14}.
 %It has been used efficiently for redshift PDF estimates \citep{Sanchez:14,Bonnett:15,Bonnett:16}.
@@
 % NO CLUDE WHAT THIS MEANS
-To avoid over-fitting, $30\%$ of the training set was reserved for validation, and training was halted as soon as the error rate began to increase on the validation set.
+To avoid over-fitting, $30$ per cent of the training set was reserved for validation, and training was halted as soon as the error rate began to increase on the validation set.
 The weights were randomly initialized based on normal sampling.
@@
 
-Trees for \Pz \citep[\tpz\footnote{\url{https://github.com/mgckind/MLZ}},][]{Carrasco_Kind:13,Carrascokind:14} uses prediction trees and random forest techniques to estimate \pzpdf s.
+Trees for \Pz\ \citep[\tpz,][]{Carrasco_Kind:13,Carrascokind:14} uses prediction trees and random forest techniques to estimate \pzpdf s.
 \tpz\ recursively splits the training set into branch pairs based on maximizing information gain among a random subsample of features, to minimize correlation between the trees, terminating only when a newly created leaf meets a criterion, such as a leaf size minimum or a variance threshold.
@@
 
-To run \tpz, we replaced non-detections with an approximation of the $1\sigma$ detection threshold based on the error forecast of the 10-year \lsst\ data, i.~e.~$dm = 2.5 \log (1 + N/S)$ where $dm \sim 0.7526$ magnitudes for $N/S = 1$.
+To run \tpz, we replaced non-detections with an approximation of the $1\sigma$ detection threshold based on the signal-to-noise-based error forecast of the 10-year \lsst\ data, i.~e.~$dm = 2.5 \log (1 + N/S)$ where $dm \sim 0.7526$ magnitudes for $N/S = 1$ (where $N$ and $S$ are the noise and signal).
 We calibrated \tpz\ with the Out-of-Bag cross-validation technique \citep{Breiman:84,Carrasco_Kind:13} to evaluate its predictive validity and determine the relative importance of the different input attributes.

### 2019-11-21 paper/tex/results.tex
fix most of typos pointed out by Seth

@@
 \caption{The individual \pzpdf s (blue) distributions produced by the twelve codes (small panels) on four exemplary galaxies' photometry (large panels) with different true redshifts (red).
-The \pzpdf s of all codes share some features for the example galaxies due to physical color degeneracies and photometric errors: tight unimodal $p(z)$ (upper left), broad unimodal $p(z)$ (upper right), bimodal $p(z)$ (lower right), and complex/multimodal $p(z)$ (lower left).
+The \pzpdf s of all codes share some features for the example galaxies due to physical colour degeneracies and photometric errors: tight unimodal $p(z)$ (upper left), broad unimodal $p(z)$ (upper right), bimodal $p(z)$ (lower right), and complex/multimodal $p(z)$ (lower left).
 The diverse algorithms and implementations induce differences in small-scale structure and sensitivity to physical systematics.}
@@
 
-However, the shared small-scale features of \annz, \metaphor, \cmnn, and \skynet\ are a result of various weighted sums of the limited number of training set galaxies with colors similar to those of the test set galaxy in question, with behavior closer to classification than regression in the case of \annz.
+However, the shared small-scale features of \annz, \metaphor, \cmnn, and \skynet\ are a result of various weighted sums of the limited number of training set galaxies with colours similar to those of the test set galaxy in question, with behavior closer to classification than regression in the case of \annz.
 The settings used on \gpz\ in this work forced broadening of the single Gaussian to cover the multimodal redshift solutions of the other codes.

### 2019-11-22 paper/main.tex
added acknowledgment

@@
 
-This paper has undergone internal review in the LSST Dark Energy Science Collaboration, the internal reviewers were: Daniel Gruen, Markus Rau, and Michael Troxel. % REQUIRED if true
+This paper has undergone internal review in the LSST Dark Energy Science Collaboration.%, the internal reviewers were: Daniel Gruen, Markus Rau, and Michael Troxel. % REQUIRED if true
 
@@
 
-The authors would like to thank their LSST-DESC publication review committee for comments that improved the paper draft.
+The authors express gratitude to Alex Abate, without whom this paper would not have gotten started.
+% The authors would like to thank their LSST-DESC publication review committee for comments that improved the paper draft.
 % hack event acknowledgments are important

### 2019-11-22 paper/main.tex
naming the reviewers

@@
 
-This paper has undergone internal review in the LSST Dark Energy Science Collaboration.%, the internal reviewers were: Daniel Gruen, Markus Rau, and Michael Troxel. % REQUIRED if true
+This paper has undergone internal review in the LSST Dark Energy Science Collaboration.
+The authors acknowledge feedback from the internal reviewers: Daniel Gruen, Markus Rau, and Michael Troxel. % REQUIRED if true
 
@@
 
-The authors express gratitude to Alex Abate, without whom this paper would not have gotten started.
+The authors express immense gratitude to Alex Abate, without whom this paper would not have gotten started.
 % The authors would like to thank their LSST-DESC publication review committee for comments that improved the paper draft.

### 2019-11-22 paper/tex/discussion.tex
address more Seth typos

@@
 In Section~\ref{sec:caution}, we reframe the strong performance of our pathological \pzpdf\ technique, \trainz, as a cautionary tale about the importance of choosing appropriate comparison metrics.
-In Section~\ref{sec:futureexperiments}, we outline the experiments we intend to build upon this study
+In Section~\ref{sec:futureexperiments}, we outline the experiments we intend to build upon this study.
 In Section~\ref{sec:futuredata}, we discuss the enhancements of the mock data set that will be necessary to enable the future experiments.
@@
 
-We remind the reader that contributed codes were given a goal of obtaining accurate \pzpdf s, not an accurate stacked estimator of the redshift distribution, so we do not expect the same codes to necessarily perform well for both classes of metrics.
+We remind the reader that codes utilized in this study were given a goal of obtaining accurate \pzpdf s, not an accurate stacked estimator of the redshift distribution, so we do not expect the same codes to necessarily perform well for both classes of metrics.
 Indeed, the codes were optimized for their interpretation of our request for ``accurate \pzpdf s,'' and we expect that the implementations would have been adjusted had we requested optimization of the traditional metrics of Appendices~\ref{sec:moments} and~\ref{sec:pointmetrics}.
@@
 The CDE loss and point estimate metrics appropriately penalize \trainz's naivete.
-As shown in Appendix~\ref{sec:pointmetrics}, \trainz ~has identical $ZPEAK$ and $ZWEIGHT$ values for every galaxy, and thus the \pz\ point estimats are constant as a function of true redshift, i.~e.~a horizontal line at the mode and mean of the training set distribution respectively.
+As shown in Appendix~\ref{sec:pointmetrics}, \trainz ~has identical $ZPEAK$ and $ZWEIGHT$ values for every galaxy, and thus the \pz\ point estimates are constant as a function of true redshift, i.e.~a horizontal line at the mode and mean of the training set distribution respectively.
 The explicit dependence on the individual posteriors in the calculation of the CDE loss, described in Section~\ref{sec:cdelossresults}, distinguishes this metric from those of the \pzpdf\ ensemble and stacked estimator of the redshift distribution, despite their prevalence in the \pz\ literature.
@@
 %[more caution on over-reliance on metrics, and making sure that metrics test all of what you want] \red{this definitely still needs work}
-In summary, context is crucial to defend against deceptively strong performers such as \trainz; \textbf{the best \pzpdf\ method is the one that most effectively achieves our science goals}, not the one that performs best on a metric that does not reflect those goals.
+In summary, context is crucial to defend against deceptively strong performers such as \trainz; \textbf{the best \pzpdf\ method is the one that most effectively achieves the science goals of a particular study}, not the one that performs best on a metric that does not reflect those goals.
 In the absence of a single scientific motivation or the information necessary for a principled metric definition, we must consider many metrics and be critical of the information transmitted by each.
@@
 
-This initial paper explores \pzpdf\ code performance in idealized conditions with perfect catalogue-based photometry and representative training data, but the resilience of each code to such realistic imperfections in prior information has not yet been evaluated.
+This initial paper explores \pzpdf\ code performance in idealized conditions with perfect catalogue-based photometry and representative training data, but the resilience of each code to realistic imperfections in prior information has not yet been evaluated.
 A top priority for a follow-up study is to test realistic forms of incomplete, erroneous, and non-representative template libraries and training sets as well as the impact of other forms of external priors that must be ingested by the codes, major concerns in \citet{Newman:2015, Masters:2017}.
@@
 Appendix~\ref{sec:moments} only addresses the stacked estimator of the redshift distribution of the entire galaxy catalogue rather than subsets in bins, tomographic or otherwise.
-The effects of tomographic binning scheme will be explored in a dedicated future paper, including propagation of redshift uncertainties in a set of fiducial tomographic redshift bins in order to estimate impact on cosmological parameter estimation.
+The effects of tomographic binning schemes will be explored in a dedicated future paper, including propagation of redshift uncertainties in a set of fiducial tomographic redshift bins in order to estimate impact on cosmological parameter estimation.
 
@@
 %\item \red{inclusion of photometric errors, image based effects, blending, lensing, sky, mask, etc...}
-This initial paper explored a data set that was constructed at the catalogue level, with no inclusion of the complications that come from measuring photometry from images.
+This initial paper explored a data set that was constructed at the catalogue level, with no inclusion of the complications that arise from photometric measurements of imaging data.
 Future data challenges will move to catalogues constructed from mock images, including the complications of deblending, sensor inefficiencies, and heterogeneous observing conditions, all anticipated to affect the measured colours of \lsst's galaxy sample \citep{Dawson:2016}.
 
-The DC1 galaxy SEDs were linear combinations of just five basis SED templates, but a next generation of data for \pzpdf\ investigations must include a broader range of physical properties.
+The DC1 galaxy SEDs were linear combinations of just five basis SED templates, and the next generation of data for \pzpdf\ investigations must include a broader range of physical properties.
 Though we only considered $z < 2$ here, \lsst\ 10-year data will contain $z > 2$ galaxies, plagued by fainter apparent magnitudes and anomalous colours due to stellar evolution.

### 2019-11-22 paper/tex/results.tex
address more Seth typos

@@
 \caption{The individual \pzpdf s (blue) distributions produced by the twelve codes (small panels) on four exemplary galaxies' photometry (large panels) with different true redshifts (red).
+All \pzpdf s have been scaled to the same peak value.
 The \pzpdf s of all codes share some features for the example galaxies due to physical colour degeneracies and photometric errors: tight unimodal $p(z)$ (upper left), broad unimodal $p(z)$ (upper right), bimodal $p(z)$ (lower right), and complex/multimodal $p(z)$ (lower left).
@@
 
-The most striking differences between codes are the small-scale features induced by the interaction between the shared piecewise constant parameterization of $200$ bins $0 < z < 2$ of Section~\ref{sec:metrics} and the smoothing conditions or lack thereof in each algorithm.
+The most striking differences between codes are the small-scale features induced by the interaction between the shared piecewise constant parameterization of $200$ bins for $0 < z < 2$ of Section~\ref{sec:metrics} and the smoothing conditions or lack thereof in each algorithm.
 The $\mathrm{d}z = 0.01$ redshift resolution is sufficient to capture the broad peaks of faint galaxies' \pzpdf s with large photometric errors but is too broad to resolve the narrow peaks for bright galaxies' \pzpdf s with small photometric errors.
-This observation is consistent with the findings of \citet[]{Malz:2018} that the piecewise constant parameterization underperforms in the presence of small-scale structures.
+This observation is consistent with the findings of \citet[]{Malz:2018} that the piecewise constant form underperforms other parameterizations in the presence of small-scale structures.
 
@@
 
-The PIT histograms of \delight, \cmnn, \skynet, and \tpz\ feature an underrepresentation of extreme values, indicative of overly broad \pzpdf s, while the overrepresentation of extreme values for \metaphor\ indicate overly narrow \pzpdf s.
+The PIT histograms of \delight, \cmnn, \skynet, and \tpz\ feature an underrepresentation of extreme values, indicative of overly broad \pzpdf s, while the overrepresentation of extreme values for \metaphor\ indicates overly narrow \pzpdf s.
 These five codes in particular have a free parameter for bandwidth, which may be responsible for this vulnerability, in spite of the opportunity for fine-tuning with perfect prior information.
@@
 As expected, \trainz\ achieves precisely the 0.0002 value expected of an ideal PIT distribution.
-\annz, \flexzboost, \lephare, and \metaphor\ have notably high catastrophic outlier rates $> 0.02$, exceeding 100 times the ideal PIT rate, meriting further investigation.
+\annz, \flexzboost, \lephare, and \metaphor\ have notably high catastrophic outlier rates $> 0.02$, exceeding 100 times the ideal PIT rate, meriting further investigation elsewhere.
 
@@
 \caption{A visualization of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the PIT distributions.
-There is generally good agreement between these statistics, with differences corresponding to the codes with outstanding catastrophic outlier rates, a reflection in the differences in how each statistic weights the tails of the distribution.
+There is generally good agreement between these statistics, with differences corresponding to the codes with outstanding catastrophic outlier rates, a reflection of the differences in how each statistic weights the tails of the distribution.
 Horizontal lines indicate the level of uncertainty found by bootstrapping a training set sample of 30,000 galaxies using \trainz; none of the codes reach this conservative ideal floor in expected uncertainty.}
@@
 \annz\ and \flexzboost\  are the top scorers under these metrics of the PIT distribution.
-\annz's strong performance can be attributed to an aspect of the training process in which training set galaxies with a PIT that more closely matches the percentiles of the DC1 training set's redshift distribution are upweighted; in effect, these quantile-based metrics were part of the algorithm itself that may or may not serve it well under more realistic experimental conditions.
+\annz's strong performance can be attributed to an aspect of the training process in which training set galaxies with PIT values that more closely match the percentiles of the DC1 training set's redshift distribution are upweighted; in effect, these quantile-based metrics were part of the algorithm itself that may or may not serve it well under more realistic experimental conditions.
 Similar to what was done for the PIT histograms in Figure~\ref{fig:pitqq}, we create bootstrap training samples of 30,000 galaxies for use with \trainz\ in order to estimate a conservative statistical floor that we would expect in real data.

### 2019-11-22 paper/tex/discussion.tex
more responses to Seth comments

@@
 The PIT and other CDF-based metrics upon which modern \pzpdf\ comparisons are built \citep{Bordoloi:10,Polsterer:16,Tanaka:17} can be gamed by a trivial estimator that yields only an affirmation of prior knowledge uninformed by the data.
-In other words, such ensemble metrics are not appropriate for the task of selecting \pzpdf\ codes for analysis pipelines.
+In other words, such ensemble metrics are insufficient for the task of selecting \pzpdf\ codes for analysis pipelines.
 %\textit{Explain the connection between \trainz\ and these metrics.}
@@
 Outright redshift failures due to emission line misidentification or noise spikes may be modeled by the inclusion of a small number of high-confidence yet false redshifts.
-We plan to perform a full sensitivity analysis on a realistically incomplete training set of spectroscopic galaxies, modeling the performance of spectrographs, emission-line properties, and expected signal-to-noise to determine which potential training set galaxies are most likely to be excluded.
+We plan to perform a full sensitivity analysis on a realistically incomplete training set of spectroscopic galaxies, modeling the performance of spectrographs, emission-line properties, and expected signal-to-noise to determine which potential training set galaxies are most likely not to yield a secure redshift.
 

### 2019-11-22 paper/tex/results.tex
more responses to Seth comments

@@
 The high QQ values (i.~e.~ more high than low PIT values) of \bpz, \cmnn, \delight, \eazy, and \gpz\ indicate \pzpdf s biased low, and the low QQ values (more low than high PIT values) of \skynet\ and \tpz\ indicate \pzpdf s biased high.
-The gray shaded band marks the $2\sigma$ variance in PIT values found using the \trainz\ algorithm with a bootstrap resampling of the training set and a sample size of 30,000 galaxies, representing a very conservative estimate of the representative training sample size, and thus an approximate minimal error significance compared to ideal performance.
-Deviations in the PIT histograms outside of this range show that significant biases are present for some codes.
+The gray shaded band marks the $2\sigma$ variance in PIT values found using the \trainz\ algorithm with a bootstrap resampling of the training set and a sample size of 30,000 galaxies, representing a very conservative estimate of the representative training sample size estimated as being required for direct \pz\ calibration \citep{Newman:2015}, and thus an approximate minimal error significance compared to ideal performance.
+The existence of deviations in the PIT histograms outside of this gray shaded uncertainty range show that significant biases are present for some codes.
 

### 2019-11-22 paper/tex/appendix.tex
revise to address mroe Seth comments

@@
  is the most widely used method for obtaining $\hat{N}^{H}(z)$ as an estimator of the redshift distribution from \pzpdf s derived by a method $H$.
-While the stacked estimator of the redshift distribution violates the mathematical definition of statistical independence and is thus not formally correct\footnote{\citet{Malz:chippr} shows how the stacking procedure can lead to bias in the estimate of $N(z)$ and presents a principled alternative to this commonly employed method.}, we use it as a basis for comparison of \pzpdf\ methods under the untested assumption that the response of our metrics of $\hat{N}^{H}(z)$ will be analogous to the same metrics applied to a principled estimator of the redshift distribution.
+While the stacked estimator of the redshift distribution violates the mathematical definition of statistical independence and is thus not formally correct\footnote{Malz \& Hogg (in prep) shows how the stacking procedure can lead to bias in the estimate of $N(z)$ and presents a principled alternative to this commonly employed method.
+See \url{https://github.com/aimalz/chippr} for details.}, we use it as a basis for comparison of \pzpdf\ methods under the untested assumption that the response of our metrics of $\hat{N}^{H}(z)$ will be analogous to the same metrics applied to a principled estimator of the redshift distribution.
 

### 2019-11-22 paper/tex/discussion.tex
revise to address mroe Seth comments

@@
 
-The work presented in this paper is only a first step in assessing \pzpdf\ approaches and moving toward an improved photometric redshift estimator.
+The work presented in this paper is only a first step in assessing \pzpdf\ approaches and moving toward a photometric redshift estimator that will be employed for LSST analyses.
 Here we discuss the next steps for subsequent investigations.

### 2019-11-22 paper/tex/methods.tex
address more Seth typos
revise to address mroe Seth comments

@@
 \hline
-\lephare~\citep{Arnouts:99}	   & template fitting	& \url{http://www.cfht.hawaii.edu/~arnouts/lephare.html} \\
 \bpz~\citep{Benitez:00} 		   & template fitting	& \url{http://www.stsci.edu/~dcoe/BPZ/} \\
 \eazy~\citep{Brammer:08}		   & template fitting & \url{https://github.com/gbrammer/eazy-photoz} \\
+\lephare~\citep{Arnouts:99}	   & template fitting	& \url{http://www.cfht.hawaii.edu/~arnouts/lephare.html} \\
 \annz~\citep{Sadeh:16}		     & machine learning	& \url{https://github.com/IftachSadeh/ANNZ} \\
+\cmnn~\citep{Graham:17}        & machine learning & N/A \\
+\delight~\citep{Leistedt:17}   & hybrid           & \url{https://github.com/ixkael/Delight} \\
 \flexzboost~\citep{Izbicki:17} & machine learning & \url{https://github.com/tpospisi/flexcode}; \url{https://github.com/rizbicki/FlexCoDE}\\
-%\textsc{Frankenz} 	& ML 	& \citet{Speagle:frankenz}	& \url{https://github.com/joshspeagle/frankenz} \\
 \gpz~\citep{Almosallam:15b}	   & machine learning	& \url{https://github.com/OxfordML/GPz} \\
 \metaphor~\citep{Cavuoti:17}   & machine learning	& \url{http://dame.dsf.unina.it}\\
-\cmnn~\citep{Graham:17}        & machine learning & N/A \\
 \skynet~\citep{Graff:14}       & machine learning & \url{http://ccpforge.cse.rl.ac.uk/gf/project/skynet/} \\
 \tpz~\citep{Carrasco_Kind:13}	 & machine learning	& \url{https://github.com/mgckind/MLZ} \\
-\delight~\citep{Leistedt:17}   & hybrid           & \url{https://github.com/ixkael/Delight} \\
 \hline
@@
 \gpz\ may also weight training set samples by importance according to $|z_{\mathrm{spec}} - z_{\mathrm{phot}}| / (1+z_{\mathrm{spec}})$ to minimize the normalized \pz\ point estimate error.
-However, this function may be adapted to \pzpdf s, pressuring the model to dedicate more resources to test set galaxies that are not well-represented in the training set.
+However, this function may be adapted to \pzpdf s, adding weight to test set galaxies that are not well-represented in the training set.
 

### 2019-11-22 paper/tex/discussion.tex
remove SRM sentence
address final batch of Seth comments

@@
 The work presented in this paper is only a first step in assessing \pzpdf\ approaches and moving toward a photometric redshift estimator that will be employed for LSST analyses.
-Here we discuss the next steps for subsequent investigations.
+Extensions of the experimental design will require further rounds of analyses, and the authors welcome interest from those outside \lsstdesc\ to have their codes assessed in these future investigations.
+%Here we discuss the next steps for subsequent investigations.
 
@@
 % Finally, while this paper and future papers discussed above focus on photometric redshift codes and estimating accurate \pzpdf s from training data, we plan a separate, but complementary, project to examine calibration of the resultant redshifts via spatial cross-correlations \citep{Newman:2008}, which will be explored in a separate series of future papers.
-The overarching plan describing everything laid out in this section is described in more detail in the \lsstdesc\ Science Roadmap (see Footnote in Section~\ref{sec:intro}).
+%The overarching plan describing everything laid out in this section is described in more detail in the \lsstdesc\ Science Roadmap (see Footnote in Section~\ref{sec:intro}).

### 2019-11-22 paper/tex/results.tex
address final batch of Seth comments

@@
 Figure~\ref{fig:pz_examples} shows the \pzpdf s for four galaxies chosen as examples of \pzpdf\ archetypes: a narrow unimodal PDF, a broad unimodal PDF, a bimodal PDF, and a multimodal PDF.
-We reiterate that under our idealized experimental conditions, differences between codes are the isolated signature of the implicit prior due to the method by which the \pzpdf s were derived.
+We reiterate that under our idealized experimental conditions, differences between the resulting \pzpdf s are the isolated signature of the implicit prior due to the method by which the \pzpdf s were derived.
 

### 2019-11-30 paper/tex/analysis.tex
fix missing parenthesis in eq 8

@@
 \begin{equation} \label{eq:ad}
-  \mathrm{AD}^2 \equiv N_{tot}\int_{-\infty}^{+\infty} \frac{\big(\mathrm{CDF}[\hat{f}, z] - \mathrm{CDF}[\tilde{f}, z]\big)^2} {\mathrm{CDF}[\tilde{f}, z] (1 -\mathrm{CDF}[\tilde{f}, z]} \mathrm{d}\mathrm{CDF}[\tilde{f}, z]
+  \mathrm{AD}^2 \equiv N_{tot}\int_{-\infty}^{+\infty} \frac{\big(\mathrm{CDF}[\hat{f}, z] - \mathrm{CDF}[\tilde{f}, z]\big)^2} {\mathrm{CDF}[\tilde{f}, z] (1 -\mathrm{CDF}[\tilde{f}, z])} \mathrm{d}\mathrm{CDF}[\tilde{f}, z]
 \end{equation}

### 2019-12-01 paper/tex/methods.tex
address final batch of Seth comments
remove label for trainZ in table
change neighbor spelling to British

@@
 
-The authors welcome interest from those outside \lsstdesc\ to have their codes assessed in future investigations that build upon this one.
+%The authors welcome interest from those outside \lsstdesc\ to have their codes assessed in future investigations that build upon this one.
 
@@
 \hline
-\trainz 	                             & machine learning	& See Section~\ref{sec:trainz} \\
+\trainz 	                             & --	& See Section~\ref{sec:trainz} \\
 \end{tabular}}
@@
 We use PPF$=0.95$ rather than PPF$=0.68$ to generate the subset of colour-matched training galaxies, whose redshifts are weighted by their inverse Mahalanobis distances when composing the \pzpdf\ rather than weighting all colour-matched training galaxies equally.
-Additionally, when the number of colour-matched training set galaxies is less than 20, the nearest 20 neighbours in colour-space are used instead, and the output \pzpdf\ is convolved with a Gaussian kernel of variance $\sigma_{\mathrm{ train}}^{2}(\mathrm{PPF}_{20}/0.95)^2 -1$ to account for the corresponding growth of the effective PPF to include 20 neighbors.
+Additionally, when the number of colour-matched training set galaxies is less than 20, the nearest 20 neighbours in colour-space are used instead, and the output \pzpdf\ is convolved with a Gaussian kernel of variance $\sigma_{\mathrm{ train}}^{2}(\mathrm{PPF}_{20}/0.95)^2 -1$ to account for the corresponding growth of the effective PPF to include 20 neighbours.
 
@@
 
-In this work, we used a hierarchical KNN to replace non-detections with values based on their neighbors.
+In this work, we used a hierarchical KNN to replace non-detections with values based on their neighbours.
 The usual cross-validation of redshift estimates and PDFs was also omitted for this study.

### 2019-12-05 paper/main.tex
add affiliation and acknowledgement for John Soo
partial update on author affil and acknowl

@@
 During the completion of this work, AIM was advised by David W. Hogg and was supported by National Science Foundation grant AST-1517237.
-%JYHS is advised by Ofer Lahav and acknowledges support from the MyBrainSc Scholarship, provided by the Ministry of Education, Malaysia.
-JS would also like to thank Antonella Palmese and Boris Leistedt for guidance on the use of the algorithms ANNz2 and Delight respectively.
+JYHS acknowledges financial support from the MyBrainSc Scholarship (Ministry of Education, Malaysia), and the supervision of Ofer Lahav and Benjamin Joachimi.
+JYHS would also like to thank Antonella Palmese and Boris Leistedt for guidance on the use of the algorithms ANNz2 and Delight respectively.
+MB acknowledges support from the Agreement ASI/INAF 2018-23-HH.0 - Phase D.
+AJC and JBK acknowledges support from DOE grant DE-SC-0011635.
+AJC, JBK, MLG, CBM acknowledge support from the DIRAC Institute in the Department of Astronomy at the University of Washington.
+The DIRAC Institute is supported through generous gifts from the Charles and Lisa Simonyi Fund for Arts and Sciences, and the Washington Research Foundation.
+MJJ acknowledges support from Oxford Hintze Centre for Astrophysical Surveys which is funded through generous support from the Hintze Family Charitable Foundation.
 

### 2019-12-05 paper/main.tex
add Morrison acknowl

@@
 MJJ acknowledges support from Oxford Hintze Centre for Astrophysical Surveys which is funded through generous support from the Hintze Family Charitable Foundation.
+CBM is supported in part by the National Science Foundation through Cooperative Agreement 1258333 managed by the Association of Universities for Research in Astronomy(AURA), and the Department of Energy under Contract No.
+DE-AC02-76SF00515 with the SLAC National Accelerator Laboratory.
+Additional LSST funding comes from private donations, grants to universities, and in-kind support from LSSTC Institutional Members.
 

### 2019-12-05 paper/main.tex
more affil and funding updates

@@
 { \red{personal funding sources}}
-SJS acknowledges support from DOE grant DE-SC0009999 and NSF/AURA grant N56981C.
+SJS and EN acknowledge support from DOE grant DE-SC0009999.  SJS acknowledges support from NSF/AURA grant N56981C.
 AIM acknowledges support from the Max Planck Society and the Alexander von Humboldt Foundation in the framework of the Max Planck-Humboldt Research Award endowed by the Federal Ministry of Education and Research.
@@
 Additional LSST funding comes from private donations, grants to universities, and in-kind support from LSSTC Institutional Members.
+RI acknowledges support from FAPESP grant 2019/11321-9 and CNPq grant 306943/2017-4.
 

### 2019-12-09 paper/main.tex
add Newman acknowl
more affiliation and funding additions

@@
 JYHS would also like to thank Antonella Palmese and Boris Leistedt for guidance on the use of the algorithms ANNz2 and Delight respectively.
+I.A. acknowledges the support of King Abdulaziz City for Science and Technology.
 MB acknowledges support from the Agreement ASI/INAF 2018-23-HH.0 - Phase D.
@@
 The DIRAC Institute is supported through generous gifts from the Charles and Lisa Simonyi Fund for Arts and Sciences, and the Washington Research Foundation.
+PEF acknowledges support from NSF grant 1521786.
 MJJ acknowledges support from Oxford Hintze Centre for Astrophysical Surveys which is funded through generous support from the Hintze Family Charitable Foundation.
+GL acknowledges partial funding from the EU funded ITN Marie Curie Network SUNDIAL.
 CBM is supported in part by the National Science Foundation through Cooperative Agreement 1258333 managed by the Association of Universities for Research in Astronomy(AURA), and the Department of Energy under Contract No.
@@
 Additional LSST funding comes from private donations, grants to universities, and in-kind support from LSSTC Institutional Members.
+JAN and RZ were supported by the U.S. Department of Energy, Office of Science, Office of High Energy Physics under award number DE-SC0007914.
 RI acknowledges support from FAPESP grant 2019/11321-9 and CNPq grant 306943/2017-4.

### 2019-12-09 paper/tex/methods.tex
add url for CMNN code
added text emphasizing that trainZ is absurd

@@
 \annz~\citep{Sadeh:16}		     & machine learning	& \url{https://github.com/IftachSadeh/ANNZ} \\
-\cmnn~\citep{Graham:17}        & machine learning & N/A \\
+\cmnn~\citep{Graham:17}        & machine learning & \url{https://github.com/dirac-institute/CMNN_Photoz_Estimator} \\
 \delight~\citep{Leistedt:17}   & hybrid           & \url{https://github.com/ixkael/Delight} \\
@@
 \end{equation}
-Unlike the other methods, the \trainz\ estimator is \textit{independent of the photometric data}, effectively performing a KNN procedure with $k=N_{\mathrm{train}}$.
+Unlike the other methods, the \trainz\ estimator is \textit{independent of the photometric data}, effectively performing a KNN procedure with $k=N_{\mathrm{train}}$, a limiting case of a \pzpdf\ estimator dominated by the shared prior information of the training set.
+In this way, \trainz\ serves as an experimental control that is not a competitive \pzpdf\ method that would be used by any real survey.
 

### 2019-12-09 paper/tex/discussion.tex
added footnote and references for trainZ cousins

@@
 Because our training set is perfectly representative of the test set, $N(z)$ should be identical for both sets down to statistical noise.
-\textbf{We make the alarming observation that \trainz, the experimental control, outperforms all codes on the CDF-based metrics, and all but one code on the $N(z)$ based statistics.}
+\textbf{We make the alarming observation that \trainz, the absurd experimental control, outperforms all codes on the CDF-based metrics, and all but one code on the $N(z)$ based statistics.}
 The PIT and other CDF-based metrics upon which modern \pzpdf\ comparisons are built \citep{Bordoloi:10,Polsterer:16,Tanaka:17} can be gamed by a trivial estimator that yields only an affirmation of prior knowledge uninformed by the data.
-In other words, such ensemble metrics are insufficient for the task of selecting \pzpdf\ codes for analysis pipelines.
+In other words, such ensemble metrics are insufficient for the task of selecting \pzpdf\ codes for analysis pipelines\footnote{That being said, we note that close relatives of \trainz\ have been employed by weak lensing surveys of the past and present \citep{Lima:08, Hildebrandt:17, Hoyle:18} to estimate $N(z)$ by assigning each test set galaxy the redshift distribution of a \emph{subset} of the training set, where the subset is defined as similar to the test set galaxy in the space of whose photometric data.
+The specific science goal naturally guides the choice of metric to focus on $N(z)$ rather than individual \pzpdf s, and on the basis of that metric such improvements upon \trainz\ are guaranteed to be more robust to training set imbalance.}.
 %\textit{Explain the connection between \trainz\ and these metrics.}
@@
 %[more caution on over-reliance on metrics, and making sure that metrics test all of what you want] \red{this definitely still needs work}
-In summary, context is crucial to defend against deceptively strong performers such as \trainz; \textbf{the best \pzpdf\ method is the one that most effectively achieves the science goals of a particular study}, not the one that performs best on a metric that does not reflect those goals.
+In summary, context is crucial to defend against deceptively strong performers such as \trainz;
+\textbf{the best \pzpdf\ method is the one that most effectively achieves the science goals of a particular study}, not the one that performs best on a metric that does not reflect those goals.
 In the absence of a single scientific motivation or the information necessary for a principled metric definition, we must consider many metrics and be critical of the information transmitted by each.

### 2019-12-09 paper/tex/discussion.tex
fixed typo

@@
 The PIT and other CDF-based metrics upon which modern \pzpdf\ comparisons are built \citep{Bordoloi:10,Polsterer:16,Tanaka:17} can be gamed by a trivial estimator that yields only an affirmation of prior knowledge uninformed by the data.
-In other words, such ensemble metrics are insufficient for the task of selecting \pzpdf\ codes for analysis pipelines\footnote{That being said, we note that close relatives of \trainz\ have been employed by weak lensing surveys of the past and present \citep{Lima:08, Hildebrandt:17, Hoyle:18} to estimate $N(z)$ by assigning each test set galaxy the redshift distribution of a \emph{subset} of the training set, where the subset is defined as similar to the test set galaxy in the space of whose photometric data.
+In other words, such ensemble metrics are insufficient for the task of selecting \pzpdf\ codes for analysis pipelines\footnote{That being said, we note that close relatives of \trainz\ have been employed by weak lensing surveys of the past and present \citep{Lima:08, Hildebrandt:17, Hoyle:18} to estimate $N(z)$ by assigning each test set galaxy the redshift distribution of a \emph{subset} of the training set, where the subset is defined as similar to the test set galaxy in the space of photometric data.
 The specific science goal naturally guides the choice of metric to focus on $N(z)$ rather than individual \pzpdf s, and on the basis of that metric such improvements upon \trainz\ are guaranteed to be more robust to training set imbalance.}.

### 2019-12-11 paper/tex/data.tex
fix references, add reference to Chinchilla in data section

@@
 
-Our mock catalogue is derived from the \textsc{Buzzard}-highres-v1.0 catalogue  \citep[][Wechsler et al., in prep]{DeRose:19}.
-\textsc{Buzzard} is built on a dark matter-only N-body simulation of $2048^{3}$ particles in a $400$ Mpc h$^{-1}$ box.
+Our mock catalogue is derived from the \textsc{Buzzard}-highres-v1.0 catalogue (DeRose et al., in prep).
+\textsc{Buzzard} is built on the \texttt{Chinchilla-400} \citep{Mao:15} dark matter-only N-body simulation consisting of $2048^{3}$ particles in a $400$ Mpc h$^{-1}$ box.
 The lightcone was constructed from smoothing and interpolation between a set of time snapshots.
@@
 
-To assign a spectrum to each galaxy, the Adding Density Dependent Spectral Energy Distributions (SEDs) procedure (\texttt{ADDSEDS}, deRose in prep.)\footnote{\url{https://github.com/vipasu/addseds}} was used.
+To assign a spectrum to each galaxy, the Adding Density Dependent Spectral Energy Distributions (SEDs) procedure \citep[\texttt{ADDSEDS,}][Wechsler et al., in prep,]{DeRose:19} was used.
 \texttt{ADDSEDS} uses a sample of $\sim 5\times 10^{5}$ galaxies from the magnitude-limited \sdss\ Data Release 6 Value Added Galaxy Catalogue \citep[NYU-VAGC,][]{Blanton:05} to train an empirical relation between absolute $r$-band magnitude, local galaxy density, and SED.

### 2019-12-11 paper/tex/methods.tex
add Dalmasso:2019 reference

--- 2019-11-22 paper/tex/methods.tex
@@
 
-Here we summarize the twelve \pzpdf\ codes compared in this study, listed in Table~\ref{tab:list_of_codes}, which include both established and emerging approaches in template fitting and machine learning.
+Here we summarize the twelve \pzpdf\ codes compared in this study, also in Table~\ref{tab:list_of_codes}, which include both established and emerging approaches in template fitting and machine learning.
 Though not exhaustive, this sample represents codes for which there was sufficient expertise within the \lsstdesc\ Photometric Redshifts Working Group.
@@
 Another key difference is the treatment of non-detections in one or more bands: some codes ignore incomplete bands, while others replace the value with either an estimate for the detection limit, the mean of other values in the training set, or another default value.
-There are varying conventions among machine learning-based codes for treatment of non-detections, and no one prescription dominates in the \pz\ literature.
+There are varying conventions among machine learning based codes for treatment of non-detections, and no one prescription dominates in the \pz\ literature.
 However, we remind the reader that only 2.0 per cent of our sample has non-detections, almost exclusively in the $u$-band, and thus should not dominate the code performance differences.
@@
 \hline
+\lephare~\citep{Arnouts:99}	   & template fitting	& \url{http://www.cfht.hawaii.edu/~arnouts/lephare.html} \\
 \bpz~\citep{Benitez:00} 		   & template fitting	& \url{http://www.stsci.edu/~dcoe/BPZ/} \\
 \eazy~\citep{Brammer:08}		   & template fitting & \url{https://github.com/gbrammer/eazy-photoz} \\
-\lephare~\citep{Arnouts:99}	   & template fitting	& \url{http://www.cfht.hawaii.edu/~arnouts/lephare.html} \\
 \annz~\citep{Sadeh:16}		     & machine learning	& \url{https://github.com/IftachSadeh/ANNZ} \\
-\cmnn~\citep{Graham:17}        & machine learning & N/A \\
-\delight~\citep{Leistedt:17}   & hybrid           & \url{https://github.com/ixkael/Delight} \\
 \flexzboost~\citep{Izbicki:17} & machine learning & \url{https://github.com/tpospisi/flexcode}; \url{https://github.com/rizbicki/FlexCoDE}\\
+%\textsc{Frankenz} 	& ML 	& \citet{Speagle:frankenz}	& \url{https://github.com/joshspeagle/frankenz} \\
 \gpz~\citep{Almosallam:15b}	   & machine learning	& \url{https://github.com/OxfordML/GPz} \\
 \metaphor~\citep{Cavuoti:17}   & machine learning	& \url{http://dame.dsf.unina.it}\\
+\cmnn~\citep{Graham:17}        & machine learning & N/A \\
 \skynet~\citep{Graff:14}       & machine learning & \url{http://ccpforge.cse.rl.ac.uk/gf/project/skynet/} \\
 \tpz~\citep{Carrasco_Kind:13}	 & machine learning	& \url{https://github.com/mgckind/MLZ} \\
+\delight~\citep{Leistedt:17}   & hybrid           & \url{https://github.com/ixkael/Delight} \\
 \hline
@@
 \noindent where $A$ is a normalization factor, $F^i_{\mathrm{pred}}(T,z)$ is the flux predicted for a template $T$ at redshift $z$, $F^i_{\mathrm{obs}}$ is the observed flux in a given band $i$, $\sigma^i_{\mathrm{obs}}$ is the observed flux error, and $N_{\mathrm filt}$ is the total number of filters, in our case the six $ugrizy$ LSST filters.
-The likelihood is a sum of observed flux error $\sigma_{b}^{\mathrm{obs}}$-weighted squared differences between the observed flux $F^{\mathrm{obs}}_{b}$ and the normalized predicted flux $F^{\mathrm{mod}}_{b}(T, z)$ in $N_{\mathrm{filt}}$ photometric filters $b$, which are the \lsst\ $ugrizy$ filters in this case.
+The likelihood is a sum of observed flux error $\sigma_{b}^{\mathrm{obs}}$-weighted squared differences between the observed flux $F^{\mathrm{obs}}_{b}$ and the normalized predicted flux $F^{\mathrm{mod}}_{b}(T, z)$ in $N_{\mathrm{filt}}$ photometric filters $b$, which is the \lsst\ $ugrizy$ filters in this case.
 Specific implementation details of each code, e.~g.~prior form and implementation, are described below.
@@
 
-Bayesian Photometric Redshift \citep[\bpz,][]{Benitez:00} determines the likelihood $p(C \vert z, T)$ of a galaxy's observed colours $C$ for a set of SED templates $T$ at redshifts $z$.
+Bayesian Photometric Redshift \citep[\bpz\footnote{\url{http://www.stsci.edu/~dcoe/BPZ/}},][]{Benitez:00} determines the likelihood $p(C \vert z, T)$ of a galaxy's observed colours $C$ for a set of SED templates $T$ at redshifts $z$.
 The \bpz\ likelihood is related to the \chisq\ likelihood by $p(C \vert z, T) \propto \exp[- \chi^{2} / 2]$.
@@
 %For photo-$z$ point estimates we use the \texttt{Z\_B} output parameter.
-Prior to running the code, the non-detection placeholder magnitude was replaced with an estimate of the 1-$\sigma$ detection limit for the undetected band as a proxy for a value close to the estimated sky noise threshold.
+Prior to running the code, the non-detection placeholder magnitude was replaced with an estimate of the one-$\sigma$ detection limit for the undetected band as a proxy for a value close to the estimated sky noise threshold.
 
@@
 
-Easy and Accurate Photometric Redshifts from Yale \citep[\eazy,][]{Brammer:08} extends the basic \chisq\ fit procedure that defines template-fitting approaches.
+Easy and Accurate Photometric Redshifts from Yale \citep[\eazy\footnote{\url{https://github.com/gbrammer/eazy-photoz}},][]{Brammer:08} extends the basic \chisq\ fit procedure that defines template-fitting approaches.
 The algorithm models the observed photometry with a linear combination of template SEDs at each redshift.
@@
 
-Photometric Analysis for Redshift Estimate \citep[\lephare,][]{Arnouts:99,Ilbert:06} uses the $\chi^2$ of Equation~\ref{eq_temp_chi} to match observed colours with those predicted from a template set.
+Photometric Analysis for Redshift Estimate \citep[\lephare\footnote{\url{http://www.cfht.hawaii.edu/~arnouts/lephare.html}},][]{Arnouts:99,Ilbert:06} uses the $\chi^2$ of Equation~\ref{eq_temp_chi} to match observed colors with those predicted from a template set.
 The template set can be semi-empirical or entirely synthetic.
@@
 
-\annz \citep{Sadeh:16} supports several machine learning algorithms, including artificial neural networks (ANN), boosted decision tree, and k-nearest neighbour (KNN) regression.
+\annz \footnote{\url{https://github.com/IftachSadeh/ANNZ}} \citep{Sadeh:16} supports several machine learning algorithms, including artificial neural networks (ANN), boosted decision tree, and k-nearest neighbour (KNN) regression.
 In addition to accounting for errors on the input photometry, \annz\ uses the KNN-uncertainty estimate of \citet{Oyaizu:08} to quantify uncertainty in the choice of method over multiple runs.
@@
 The colour-matched nearest-neighbours photometric redshift estimator \citep[\cmnn,][]{Graham:17} uses a training set of galaxies with known redshifts that has equivalent or better photometry than the test set in terms of quality and filter coverage.
-For each galaxy in the test set, \cmnn\ identifies a colour-matched subset of training galaxies using a threshold in the Mahalanobis distance $D_M = \sum_{j}^{N_{\mathrm{colours}}} (c^{\mathrm{train}}_{j} - c^{\mathrm{test}}_{j})^{2} / \delta c_{\mathrm{test}}^2$ in the space of available colours $c$, with colour measurement errors $\delta c_{\mathrm{test}}$ and $N_{\mathrm{colours}} = 5$ colours $j$ defined by the $ugrizy$ filters, which defines the set of colour-matched neighbours based on a value of the percent point function (PPF).
-As an example, for $N_{\mathrm{filt}}=5$ with PPF$=0.95$, $95$ per cent of all training galaxies consistent with the test galaxy will have $D_M < 11.07$.
+For each galaxy in the test set, \cmnn\ identifies a colour-matched subset of training galaxies using a threshold in the Mahalanobis distance $D_M = \sum_{j}^{N_{\mathrm{colours}}} (c^{\mathrm{train}}_{j} - c^{\mathrm{test}}_{j})^{2} / \delta c_{\mathrm{test}}^2$ in the space of available colours $c$, with colour measurement errors $\delta c_{\mathrm{test}}$ and $N_{\mathrm{colours}} = 5$ colors $j$ defined by the $ugrizy$ filters, which defines the set of colour-matched neighbours based on a value of the percent point function (PPF).
+As an example, for $N_{\mathrm{filt}}=5$ with PPF$=0.95$, $95\%$ of all training galaxies consistent with the test galaxy will have $D_M < 11.07$.
 Undetected bands are dropped, thereby reducing the effective $N_{\mathrm{filt}}$ for that galaxy.
@@
 We make an additional adaptation to enable the \cmnn\ algorithm to yield accurate \pzpdf s for all galaxies, as the original \citet{Graham:17} algorithm is optimized for \pz\ point estimates and is susceptible to less accurate \pzpdf s for bright galaxies or those with few matches in colour-space.
-We use PPF$=0.95$ rather than PPF$=0.68$ to generate the subset of colour-matched training galaxies, whose redshifts are weighted by their inverse Mahalanobis distances when composing the \pzpdf\ rather than weighting all colour-matched training galaxies equally.
-Additionally, when the number of colour-matched training set galaxies is less than 20, the nearest 20 neighbours in colour-space are used instead, and the output \pzpdf\ is convolved with a Gaussian kernel of variance $\sigma_{\mathrm{ train}}^{2}(\mathrm{PPF}_{20}/0.95)^2 -1$ to account for the corresponding growth of the effective PPF to include 20 neighbors.
+We use PPF$=0.95$ rather than PPF$=0.68$ to generate the subset of colour-matched training galaxies, whose redshifts are weighted by their inverse Mahalanobis distances of the when composing the \pzpdf\ rather than weighting all colour-matched training galaxies equally.
+Additionally, when the number of colour-matched training set galaxies is less than 20, the nearest 20 neighbours in color-space are used instead, and the output \pzpdf\ is convolved with a Gaussian kernel of variance $\sigma_{\mathrm{ train}}^{2}(\mathrm{PPF}_{20}/0.95)^2 -1$ to account for the corresponding growth of the effective PPF to include 20 neighbors.
 
@@
 
-\delight \citep{Leistedt:17} is a hybrid technique that infers \pz s with a data-driven model of latent SEDs and a physical model of photometric fluxes as a function of redshift.
-Generally, machine learning methods rely on representative training data with shared photometric filters, while template-based methods rely on a complete library of templates based on physical models constructed.
+\delight \footnote{\url{https://github.com/ixkael/Delight}} \citep{Leistedt:17} is a hybrid technique that infers \pz s with a data-driven model of latent SEDs and a physical model of photometric fluxes as a function of redshift.
+Generally, machine learning methods rely on representative training data with shared photometric filters, while template based methods rely on a complete library of templates based on physical models constructed.
 \delight\ aims to take the best aspects of both approaches by constructing a large collection of latent SED templates (or physical flux-redshift models) from training data, with a template SED library as a guide to the learning of the model, thereby circumventing the machine learning prerequisite of representative training data in the same photometric bands and the template fitting requirement of detailed galaxy SED models.
@@
 
-In this study, the default settings of \delight\ were used, with the exception that the PDF bins were set to be linearly spaced rather than logarithmically.
+In this study, the default settings of \delight\ were used, with the exception that the PDF bins were set to be linearly-spaced rather than logarithmic.
 The Gaussian process was trained using the full DC1 training set.
@@
 
-\flexzboost \citep{Izbicki:17} is built on \texttt{FlexCode}, a general-purpose methodology for converting any conditional mean point estimator of $z$ to a conditional density estimator $p(z \vert \x) \equiv f(z \vert \x)$, where $\x$ here represents our photometric covariates and errors.
+\flexzboost \footnote{\url{https://github.com/tpospisi/flexcode};\\ \url{https://github.com/rizbicki/FlexCoDE} \label{flexzboost_github}} \citep{Izbicki:17,Dalmasso:2019} is built on \texttt{FlexCode}, a general-purpose methodology for converting any conditional mean point estimator of $z$ to a conditional density estimator $p(z \vert \x) \equiv f(z \vert \x)$, where $\x$ here represents our photometric covariates and errors.
 \flexzboost\ expands the unknown function $f(z \vert \x) = \sum_{i}\beta_{i}(\x)\phi_{i}(z)$ using an orthonormal basis $\{\phi_{i}(z)\}_{i}$.
@@
 
-In this paper, we used \texttt{xgboost} \citep{Chen:16} for the regression; it should, however, be noted that \texttt{FlexCode-RF}\footnote{\url{https://github.com/tpospisi/flexcode};\\ \url{https://github.com/rizbicki/FlexCoDE} \label{flexzboost_github}}, based on Random Forests, generally performs better on smaller datasets.
+In this paper, we used \texttt{xgboost} \citep{Chen:16} for the regression; it should, however, be noted that \texttt{FlexCode-RF}\footref{flexzboost_github}, based on Random Forests, generally performs better on smaller datasets.
 As our basis $\phi_{i}(z)$, we choose a standard Fourier basis.
@@
 Both $I$ and $\alpha$ were chosen in an automated way by minimizing the weighted $L_2$-loss function \citep[Eq.
-5 in][]{Izbicki:17} on a validation set comprised of a randomly selected 15 per cent of the DC1 training set.
-While \texttt{FlexCode}'s lossless native encoding stores each \pzpdf\ using the basis coefficients $\beta_{i}(\x)$, we discretized the final estimates into 200 linearly spaced redshift bins $0 < z < 2$ to match the consistent output format of the experimental conditions.
+5 in][]{Izbicki:17} on a validation set comprised of a randomly selected 15\% of the DC1 training set.
+While \texttt{FlexCode}'s lossless native encoding stores each \pzpdf\ using the basis coefficients $\beta_{i}(\x)$, we discretized the final estimates into 200 linearly-spaced redshift bins $0 < z < 2$ to match the consistent output format of the experimental conditions.
 
@@
 
-\gpz \citep{Almosallam:16a,Almosallam:15b} is a sparse Gaussian process-based code, a scalable approximation of full Gaussian Processes \citep{Rasmussen:06}, that produces input-dependent variance estimates corresponding to heteroscedastic noise.
+\gpz \footnote{\url{https://github.com/OxfordML/GPz}} \citep{Almosallam:16a,Almosallam:15b} is a sparse Gaussian process based code, a scalable approximation of full Gaussian Processes \citep{Rasmussen:06}, that produces input-dependent variance estimates corresponding to heteroscedastic noise.
 The model assumes a Gaussian posterior probability $p(z \vert \x) = \mathcal{N}\left(z \vert \mu(\x), \sigma(\x)^{2}\right)$ of the output redshift $z$ given the input photometry $\x$.
-The mean $\mu(\x)$ and the variance $\sigma(\x)^{2}$ are modeled as functions $f(\x) = \sum_{i=1}^{m}w_{i}\phi_{i}(\x)$ that are linear combinations of $m$ basis functions $\left\{\phi_{i}(\x)\right\}_{i=1}^{m}$ with associated weights $\left\{w_{i}\right\}_{i=1}^{m}$.
+The mean $\mu(\x)$ and the variance $\sigma(\x)^{2}$ are modeled as functions $f(\x) = \sum_{i=1}^{m}w_{i}\phi_{i}(\x)$ linear combinations of $m$ basis functions $\left\{\phi_{i}(\x)\right\}_{i=1}^{m}$ with associated weights $\left\{w_{i}\right\}_{i=1}^{m}$.
 % Basis function models, for specific classes of basis functions such as the sigmoid or the squared exponential, have the advantage of being universal approximators, i.e.
@@
 \gpz's variance estimate is composed of a model uncertainty term corresponding to sparsity of the training set photometry and a noise uncertainty term encompassing noisy photometric observations, enabling quantification of any need for more representative or more precise training samples.
-\gpz\ may also weight training set samples by importance according to $|z_{\mathrm{spec}} - z_{\mathrm{phot}}| / (1+z_{\mathrm{spec}})$ to minimize the normalized \pz\ point estimate error.
-However, this function may be adapted to \pzpdf s, adding weight to test set galaxies that are not well-represented in the training set.
+\gpz\ may also weight training set samples by importance according to $|z_{\mathrm{spec}} - z_{\mathrm{phot}}| / (1+z_{\mathrm{spec}})$ to minimize the normalized \pz\ point estimate error, however, this function may be adapted to \pzpdf s, pressuring the model to dedicate more resources to test set galaxies that are not well-represented in the training set.
 
-To smooth the long tail in the distribution of magnitude errors, we use the logarithm of the magnitude errors, improving numerical stability and eliminating the need for constraints on the optimization process.
+To smooth the long tail in the distribution of magnitude errors, we use the log of the magnitude errors, improving numerical stability and eliminating the need for constraints on the optimization process.
 % We use principaphoto-\mathinhead{$z$}{z} interim posteriorsl component analysis (PCA) to decorrelate the data
@@
 This is the optimal expected value of the unobserved variables given the observed ones under the assumption that the distribution is jointly Gaussian; note that this reduces to a simple average if the covariates are independent with $\Sigma_{\mathrm{uo}} = 0$.
-We reserved for validation 20 per cent of the training set and used the Variable Covariance option in \gpz\ with 200 basis functions (see \citet{Almosallam:15b} for details), and did not apply cost-sensitive learning options.
+We reserved for validation 20\% of the training set and used the Variable Covariance option in \gpz\ with 200 basis functions (see \citet{Almosallam:15b} for details), neglecting to apply cost-sensitive learning options.
 
@@
 
-Machine-learning Estimation Tool for Accurate Photometric Redshifts \citep[\metaphor,][]{Cavuoti:17} is based on the Multi Layer Perceptron with Quasi Newton Algorithm (MLPQNA) with the least square error model and Tikhonov $L_{2}$-norm regularization \citep{Hofmann:18}.
+Machine-learning Estimation Tool for Accurate Photometric Redshifts \citep[\metaphor\footnote{\url{http://dame.dsf.unina.it}},][]{Cavuoti:17} is based on the Multi Layer Perceptron with Quasi Newton Algorithm (MLPQNA) with the least square error model and Tikhonov $L_{2}$-norm regularization \citep{Hofmann:18}.
 %, already validated on photo-$z$'s in several cases \citep{de_Jong:17,Cavuoti:17b,Cavuoti:15,Brescia:14,Brescia:13,Biviano:13}.
@@
 
-\skynet \citep{Graff:14} employs a neural network based on a second-order conjugate gradient optimization scheme \citep[see][for further details]{Graff:14}.
+\skynet \footnote{\url{http://ccpforge.cse.rl.ac.uk/gf/project/skynet/}} \citep{Graff:14} employs a neural network based on a second order conjugate gradient optimization scheme \citep[see][for further details]{Graff:14}.
 %It has been used efficiently for redshift PDF estimates \citep{Sanchez:14,Bonnett:15,Bonnett:16}.
@@
 % NO CLUDE WHAT THIS MEANS
-To avoid over-fitting, $30$ per cent of the training set was reserved for validation, and training was halted as soon as the error rate began to increase on the validation set.
+To avoid over-fitting, $30\%$ of the training set was reserved for validation, and training was halted as soon as the error rate began to increase on the validation set.
 The weights were randomly initialized based on normal sampling.
@@
 
-Trees for \Pz\ \citep[\tpz,][]{Carrasco_Kind:13,Carrascokind:14} uses prediction trees and random forest techniques to estimate \pzpdf s.
+Trees for \Pz \citep[\tpz\footnote{\url{https://github.com/mgckind/MLZ}},][]{Carrasco_Kind:13,Carrascokind:14} uses prediction trees and random forest techniques to estimate \pzpdf s.
 \tpz\ recursively splits the training set into branch pairs based on maximizing information gain among a random subsample of features, to minimize correlation between the trees, terminating only when a newly created leaf meets a criterion, such as a leaf size minimum or a variance threshold.
@@
 
-To run \tpz, we replaced non-detections with an approximation of the $1\sigma$ detection threshold based on the signal-to-noise-based error forecast of the 10-year \lsst\ data, i.~e.~$dm = 2.5 \log (1 + N/S)$ where $dm \sim 0.7526$ magnitudes for $N/S = 1$ (where $N$ and $S$ are the noise and signal).
+To run \tpz, we replaced non-detections with an approximation of the $1\sigma$ detection threshold based on the error forecast of the 10-year \lsst\ data, i.~e.~$dm = 2.5 \log (1 + N/S)$ where $dm \sim 0.7526$ magnitudes for $N/S = 1$.
 We calibrated \tpz\ with the Out-of-Bag cross-validation technique \citep{Breiman:84,Carrasco_Kind:13} to evaluate its predictive validity and determine the relative importance of the different input attributes.

### 2019-12-11 paper/tex/methods.tex
add fzboost reference

--- 2019-12-09 paper/tex/methods.tex
@@
 
-\flexzboost \citep{Izbicki:17} is built on \texttt{FlexCode}, a general-purpose methodology for converting any conditional mean point estimator of $z$ to a conditional density estimator $p(z \vert \x) \equiv f(z \vert \x)$, where $\x$ here represents our photometric covariates and errors.
+\flexzboost \citep{Izbicki:17,Dalmasso:2019} is built on \texttt{FlexCode}, a general-purpose methodology for converting any conditional mean point estimator of $z$ to a conditional density estimator $p(z \vert \x) \equiv f(z \vert \x)$, where $\x$ here represents our photometric covariates and errors.
 \flexzboost\ expands the unknown function $f(z \vert \x) = \sum_{i}\beta_{i}(\x)\phi_{i}(z)$ using an orthonormal basis $\{\phi_{i}(z)\}_{i}$.

### 2019-12-11 paper/tex/discussion.tex
add footnote about limited scope, change Roadmap to SRM

@@
 
-To make optimal use of the \lsst\ data for cosmological and other astrophysical analyses of the \lsstdesc\ Science Roadmap, future investigations that build upon this one will require a more sophisticated set of galaxy photometry and redshifts.
+To make optimal use of the \lsst\ data for cosmological and other astrophysical analyses of the \lsstdesc\ SRM, future investigations that build upon this one will require a more sophisticated set of galaxy photometry and redshifts.
 %\item \red{inclusion of photometric errors, image based effects, blending, lensing, sky, mask, etc...}

### 2019-12-11 paper/tex/introduction.tex
add footnote about limited scope, change Roadmap to SRM

@@
 Specific science cases may have their own requirements on \pz\ performance that exceed those of the survey as a whole.
-In that vein, the \lsst\ Dark Energy Science Collaboration (\lsstdesc) developed a separate SRD \citep{Mandelbaum:2018} that conservatively forecasts the constraining power of five cosmological probes, leading to even more stringent requirements on \pz\ performance, including those defined in terms of tomographically binned subsample populations rather than individual galaxies.
+In that vein, the \lsst\ Dark Energy Science Collaboration (\lsstdesc) developed a separate SRD \citep{Mandelbaum:2018} that conservatively forecasts the constraining power of five cosmological probes, leading to even more stringent requirements on \pz\ performance, including those defined in terms of tomographically binned subsample populations\footnote{While tomographic samples will play a prominent role in some science cases, estimation of tomographic distributions is a distinct problem with distinct solutions.
+In this paper we focus solely on individual galaxy redshift estimates.} rather than individual galaxies.
 

### 2019-12-12 paper/tex/introduction.tex
clarifying sensitivity analysis

@@
 In the absence of simulated data drawn from known redshift distributions, the very concept of a ``true PDF'' for an individual galaxy is unavailable, and we must instead rely on measures of ensemble behaviour to characterize PDF quality (see \S\,\ref{sec:metrics} for further discussion).
-We aim to perform a comprehensive sensitivity analysis of \pzpdf\ techniques in order to ultimately select those that will become part of the \lsstdesc\ pipelines, described in the Science Roadmap (SRM)\footnote{Available at: \url{https://lsstdesc.org/assets/pdf/docs/DESC_SRM_latest.pdf}}.
+We aim to perform a comprehensive sensitivity analysis of the dependence of \pzpdf s on the code used to produce them in order to ultimately select those that will become part of the \lsstdesc\ pipelines, described in the Science Roadmap (SRM)\footnote{Available at: \url{https://lsstdesc.org/assets/pdf/docs/DESC_SRM_latest.pdf}}.
 In this initial study, we focus on evaluating the performance of \pzpdf\ codes using PDF-specific performance metrics in a formally controlled experiment with complete and representative prior information (template libraries and training sets) to set a baseline for subsequent investigations.

### 2019-12-13 paper/tex/data.tex
fix incorrect Masters reference

@@
 
-Following the findings of \citet{Bernstein:10}, \citet{Newman:2015}, and \citet{Masters:2017} that only $\sim\!10^{4}$ spectra are necessary to calibrate \pz s to Stage IV requirements, we aimed to set aside a randomly selected training set of $3-5\times 10^{4}$ galaxies, $\sim 10$ per cent of the full sample.
+Following the findings of \citet{Bernstein:10}, \citet{Newman:2015}, and \citet{Masters:2015} that only $\sim\!10^{4}$ spectra are necessary to calibrate \pz s to Stage IV requirements, we aimed to set aside a randomly selected training set of $3-5\times 10^{4}$ galaxies, $\sim 10$ per cent of the full sample.
 After all cuts described above, we designated the \textit{DC1 training set} of $44\,404$ galaxies for which observed photometry, true SEDs, and true redshifts would be provided to all codes and the blinded \textit{DC1 test set} of $399\,356$ galaxies for which photometry alone would be provided to all codes and \pzpdf s would be requested.

### 2019-12-16 paper/tex/appendix.tex
remove superfluous comments from tex

@@
 \label{sec:moments}
-%(Alex Malz)
 
@@
 See \url{https://github.com/aimalz/chippr} for details.}, we use it as a basis for comparison of \pzpdf\ methods under the untested assumption that the response of our metrics of $\hat{N}^{H}(z)$ will be analogous to the same metrics applied to a principled estimator of the redshift distribution.
-
-%is the most widely accepted method for obtaining $\hat{N}^{H}(z)$ as an estimator of the redshift distribution from \pzpdf s derived by a method $H$.
-This procedure is not formally correct, basic statistics tells us that joint probabilities are not formed by adding probabilities.
-\citet{Malz:chippr} shows how the stacking procedure can lead to bias in redshift estimates, and develop a principled alternative to this commonly employed method.
-However, while  the use of the stacked estimator of the redshift distribution is not formally correct, we use it under the untested assumption that the response of our metrics of $\hat{N}^{H}(z)$ will be analogous to the same metrics applied to a principled estimator of the redshift distribution.
 
@@
 \label{fig:nz}
-%\aim{Arguably this could also be clarified by showing the true $n'(z)$ once and then differences from it for each cod's stacked estimator\dots}\scc{what if we plot the difference between true and stacked rather than soverposing the two lines?
-Sam please do not hate me.}
 \end{figure*}
@@
 
-%\begin{table}
-%\setlength{\tabcolsep}{2pt}
-%\caption{Moments of the stacked estimator $\hat{N}(z)$ of the redshift distribution.
-%Most of the codes considered recover the moments of $\tilde{N}(z)$}
-%\label{tab:moments}
-%\begin{tabular}{lccc}
-%\hline
-%\hline
-% \multicolumn{4}{l}{Moments of $\hat{N}(z)$} \\
-%\hline
-%Estimator  & mean       & variance    & skewness \\
-%Empirical ``truth'' & 0.701 & 0.630 & 0.671  \\
-%\hline
-%\annz       & 0.702      & 0.625      & 0.653    \\
-%\bpz        & 0.699      & 0.629      & 0.671    \\
-%\delight    & 0.692      & 0.609      & 0.638    \\
-%\eazy       & 0.681      & 0.595      & 0.619    \\
-%\flexzboost & 0.694      & 0.610      & 0.631    \\
-%\gpz        & 0.696      & 0.615      & 0.639    \\
-%\lephare    & 0.718      & 0.668      & 0.741    \\
-%\metaphor   & 0.705      & 0.628      & 0.657    \\
-%\cmnn       & 0.701      & 0.628      & 0.667    \\
-%\skynet     & 0.743      & 0.708      & 0.797    \\
-%\tpz        & 0.700      & 0.619      & 0.643    \\
-%\hline
-%\trainz	    & 0.699 		 & 0.627 	    & 0.666 \\
-%\end{tabular}
-%\end{table}
-
 \skynet\ exhibits redshift bias in Figure~\ref{fig:nz} and is a clear outlier in the first moment of $\hat{N}(z)$ in Figure~\ref{fig:moments}.
@@
 \item RMS scatter $\sigma < 0.02 (1 + z_{\mathrm{true}})$
-%\scc{actually the SB reports in page 75 that the goal of 0.02 is for the RMS of $\frac{\sigma_{z}}{(1+z)}$ while we are using IQR}
-\item bias $b_{z} < 0.003$ %\scc{in page 518 of the SB is clearly reported that the mean is expected to be less than 0.003 in our case we have defined as bias the median and not the mean }
-\item catastrophic outlier rate $f_{\mathrm{out}} < 10$ per cent %\scc{clearly the number of outliers since are defined above $3\sigma$ strictly depends on the sigma definition, in the science book it seems to refer to the RMS of the $e_z$ distribution even ef it could be the standard deviation, but again I don't think it refers to IQR, therefore we could not consider exactly this numbers as reference.}
-%\red{I am going off of conversations with Zeljko and how we {\it actually} computed the metrics in the Science Book (I have the scripts).
-You are correct that several of these are slightly different than actually stated in the Science Book, or where the Science Book does not use very precise language as to what was done to compute the RMS or define the bias.
-We can also cut out those specifics, or maybe say Ivezic private communication, maybe?
-Also $\sigma_{z}$ defined in terms of IQR is exactly equivalent to the standard deviation when scaled by 1.349 for a Gaussian distribution, so the numbers are appropriate to cite in this context.
-IQR is just a more robust way of calculating sigma for a distribution with outliers.--SJS }
-%\scc{I am very sorry I am not saying that IQR is not fine, I want to keep the IQR my concerns are related to the notation and the \textit{reader understanding}, if I wrote $\sigma$ somewhere in a table the reader will understand for sure that is the traditional standard deviation (a lot of people skips the indicators definition for the \textit{obvious} ones), I am suggesting to replace the notation $\sigma$ with the notation $\sigma_{IQR}$ (or something like this) to avoid any confusion, and also to change from $bias$ to $median$ for the same reason.
-I agree that say Ivezic private communication could be better.}
-.
+\item bias $b_{z} < 0.003$ 
+\item catastrophic outlier rate $f_{\mathrm{out}} < 10$ per cent 
 \end{itemize}
@@
 The finite grid spacing of the \pzpdf s induces some discretization in $z_{PEAK}$.
-%, particularly for \skynet.
 The features perpendicular to the $z_{phot} = z_{spec}$ line are due to the $4000$ \AA\ break passing through the gaps between adjacent filters.
-% and logically are more prominent for template-based codes
 Even the strongest codes feature populations far from the $z_{phot} = z_{spec}$ line representing a degeneracy in the space of colours and redshifts.
-% While use of the full information available via $p(z)$ mitigates their impact, a full understanding of the outlier population is critical for LSST science, particularly in tomographic applications.
-%\red{I forget what exactly I was trying to say here, if someone else wants to take a crack at this it might be helpful --SJS}
-%\scc{may be worth to say that the usage of further bands could mitigate this effect by removing the degeneration  usually induced from emission lines moving across different filters?}
 
@@
 All twelve codes perform at or near the goals of the \lsst\ Science Requirements Document\footnote{available at: \url{http://ls.st/srd}} and \citet{Graham:17}, which is encouraging if not unexpected for $i < 25.3$.
-% But, it is still an encouraging sign, given an updated mock galaxy simulation and the expanded set of photo-$z$ codes tested.
-%\red{maybe say consistent with expectations, and a reference to Melissa's paper in here?  What else to say about nearly meeting point goals?}

### 2019-12-17 paper/tex/appendix.tex
switch to pdf figs

@@
 \centering
-\includegraphics[width=0.74\textwidth]{fig/NZsumplot_12codes_scottsrule.jpg}
+\includegraphics[width=0.74\textwidth]{fig/NZsumplot_12codes_scottsrule.pdf}
 \caption{The smoothed stacked estimator $\hat{N}(z)$ of the redshift distribution (red) produced by each code (panels) compared to the true redshift distribution $\tilde{N}(z)$ (blue).
@@
 \centering
-\includegraphics[width=0.74\textwidth]{fig/KSvsCvMvsAD_NZ_withnull_jpg.jpg}
+\includegraphics[width=0.74\textwidth]{fig/KSvsCvMvsAD_NZ_withnull_jpg.pdf}
 \caption{A visualization of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the $\hat{N}(z)$ distributions.
@@
 \centering
-\includegraphics[width=0.75\textwidth]{fig/momentsplot_10yr.jpg}
+\includegraphics[width=0.75\textwidth]{fig/momentsplot_10yr.pdf}
 \caption{Residuals of the first three moments of the stacked $\hat{N}(z)$ distribution.
@@
 \centering
-\includegraphics[width=0.49\textwidth]{fig/ZPEAK_szpz_threecolumn_12codes_navy_lowalpha.jpg}
-\includegraphics[width=0.49\textwidth]{fig/ZWEIGHT_szpz_threecolumn_12codes_navy_lowalpha.jpg}
+\includegraphics[width=0.49\textwidth]{fig/ZPEAK_szpz_threecolumn_12codes_navy_lowalpha.pdf}
+\includegraphics[width=0.49\textwidth]{fig/ZWEIGHT_szpz_threecolumn_12codes_navy_lowalpha.pdf}
 \caption{The density of \pz\ point estimates (contours) reduced from the \pzpdf s with outliers (blue) beyond the outlier cutoff (red dashed lines), via the mode ($z_{PEAK}$, left panel) and main-peak-mean ($z_{WEIGHT}$, right panel).

### 2019-12-17 paper/tex/results.tex
switch to pdf figs

@@
 \label{sec:results}
-%(Sam Schmidt, Bryce Kalmbach, Johann Cohen Tanugi, Rongpu Zhou, Alex Malz)
 
@@
 %\centering
-\includegraphics[width=0.49\textwidth]{fig/pz_12codes_261931_noseaborn_crop.jpg}\includegraphics[width=0.49\textwidth]{fig/pz_12codes_471167_noseaborn_crop.jpg}\\
-\includegraphics[width=0.49\textwidth]{fig/pz_12codes_713178_noseaborn_crop.jpg}\includegraphics[width=0.49\textwidth]{fig/pz_12codes_982747_noseaborn_crop.jpg}
+\includegraphics[width=0.49\textwidth]{fig/pz_12codes_261931_noseaborn_crop.pdf}\includegraphics[width=0.49\textwidth]{fig/pz_12codes_471167_noseaborn_crop.pdf}\\
+\includegraphics[width=0.49\textwidth]{fig/pz_12codes_713178_noseaborn_crop.pdf}\includegraphics[width=0.49\textwidth]{fig/pz_12codes_982747_noseaborn_crop.pdf}
 \caption{The individual \pzpdf s (blue) distributions produced by the twelve codes (small panels) on four exemplary galaxies' photometry (large panels) with different true redshifts (red).
@@
 \centering
-\includegraphics[width=0.74\textwidth]{fig/PITANDQQplot_12codes_withsigmaband_crop.jpg}
+\includegraphics[width=0.74\textwidth]{fig/PITANDQQplot_12codes_withsigmaband_crop.pdf}
 \caption{The QQ plot (red) and PIT histogram (blue) of the \pzpdf\ codes (panels) along with the ideal QQ (black dashed diagonal) and ideal PIT (gray horizontal) curves, as well as a difference plot for the QQ difference from the ideal diagonal (lower inset).
@@
 \centering
-\includegraphics[width=0.74\textwidth]{fig/KSvsCvMvsAD_PIT_withnull_jpg.jpg}
+\includegraphics[width=0.74\textwidth]{fig/KSvsCvMvsAD_PIT_withnull_jpg.pdf}
 \caption{A visualization of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the PIT distributions.
@@
 \bpz 		    & $-7.82$ \\
-%\textsc{Delight} 	& $-4.06$ \\
 \delight    & $-8.33$\\
-%\textsc{EAZY} 		& $-7.97$ \\
 \eazy       & $-7.07$ \\
-%\textsc{FlexZBoost} & $-11.51$ \\
 \flexzboost & $-10.60$\\
@@
 \metaphor 	& $-6.28$ \\
-%\textsc{CMNN} 		& $-$ \\
 \cmnn       & $-10.43$ \\
@@
 \trainz		  & $-0.83$ \\
-%\textsc{Frankenz}	& $-$  \\
-%\hline
 \end{tabular}

### 2019-12-17 paper/main.tex
fix references, add reference to Chinchilla in data section
remove red personal funding reminder text
add Risa and move affiliations to end of paper, modify LSST DESC in auth list

@@
-\RequirePackage[switch, columnwise, running, mathlines, displaymath,
-mathlines]{lineno}
+\RequirePackage[switch, columnwise, running, mathlines, displaymath
+]{lineno}
 \RequirePackage{docswitch}
@@
 \begin{document}
-\linenumbers
+%\linenumbers
 
-\title{Evaluation of probabilistic photometric redshift estimation approaches for \textsc{LSST}}
+\title[Photo-z estimation approaches for LSST]{Evaluation of probabilistic photometric redshift estimation approaches for \textsc{LSST}}
 
@@
 
-Many scientific investigations of photometric galaxy surveys require redshift estimates, whose uncertainty properties are best encapsulated by photometric redshift (photo-$z$) posterior probability distribution functions (PDFs).
+Many scientific investigations of photometric galaxy surveys require redshift estimates, whose uncertainty properties are best encapsulated by photometric redshift (photo-$z$) posterior probability density functions (PDFs).
 A plethora of photo-$z$ PDF estimation methodologies abound, producing discrepant results with no consensus on a preferred approach.
@@
 
-{ \red{personal funding sources}}
 SJS and EN acknowledge support from DOE grant DE-SC0009999.  SJS acknowledges support from NSF/AURA grant N56981C.
@@
 MJJ acknowledges support from Oxford Hintze Centre for Astrophysical Surveys which is funded through generous support from the Hintze Family Charitable Foundation.
+ABL and TP acknowledge support from NSF DMS grant 1520786.
 GL acknowledges partial funding from the EU funded ITN Marie Curie Network SUNDIAL.
@@
 
+\section*{Affiliations}
+$^{1}$ Department of Physics, University of California, One Shields Ave., Davis, CA, 95616, USA\\
+$^{2}$ German Centre of Cosmological Lensing, Ruhr-Universitaet Bochum, Universitaetsstra{\ss}e 150, 44801 Bochum, Germany\\
+$^{3}$ Center for Cosmology and Particle Physics, New York University, 726 Broadway, New York, 10003, USA\\
+$^{4}$ Department of Physics, New York University, 726 Broadway, New York, 10003, USA\\
+$^{5}$ School of Physics, University of Science Malaysia, 11800 USM, Pulau Pinang, Malaysia\\
+$^{6}$ Department of Physics and Astronomy, University College London, Gower Street, London WC1E 6BT, UK\\
+$^{7}$ King Abdulaziz City for Science and Technology, Riyadh 11442, Saudi Arabia\\
+$^{8}$ Information Engineering, Parks Road, Oxford, OX1 3PJ, UK\\
+$^{9}$ INAF-Astronomical Observatory of Capodimonte, Salita Moiariello 16, 80131 Napoli, Italy\\
+$^{10}$ Department of Physics E. Pancini, University Federico II, via Cinthia 6, I-80126, Napoli, Italy\\
+$^{11}$ Laboratoire Univers et Particules de Montpellier, Universit\'e de Montpellier, CNRS, Montpellier, France\\
+$^{12}$ DIRAC Institute and Department of Astronomy, University of Washington, Box 351580, U.W., Seattle WA 98195, USA\\
+$^{13}$ Santa Cruz Institute for Particle Physics, Santa Cruz, CA 95064, USA\\
+$^{14}$ Berkeley Center for Cosmological Physics, Department of Physics, University of California, Berkeley CA 94720\\
+$^{15}$ Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, CA 94720, USA\\
+$^{16}$ Kavli Institute for Particle Astrophysics and Cosmology and Department of Physics, Stanford University, Stanford, CA 94305, USA\\
+$^{17}$ Department of Particle Physics and Astrophysics, SLAC National Accelerator Laboratory, Stanford, CA 94305, USA\\
+$^{18}$ Department of Statistics \& Data Science, Carnegie Mellon University, 5000 Forbes Avenue, Pittsburgh, PA 15213, USA\\
+$^{19}$ Dunlap Institute for Astronomy \& Astrophysics, University of Toronto, 50 St. George Street, Toronto, ON M5S 3H4, Canada\\
+$^{20}$ Department of Physics and Astronomy, Rutgers, The State University of New Jersey, 136 Frelinghuysen Road, Piscataway, NJ 08854-8019 USA\\
+$^{21}$ Astrophysics, Department of Physics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford, OX1 3RH, UK\\
+$^{22}$ Department of Physics and Astronomy, University of the Western Cape, Bellville 7535, South Africa\\
+$^{23}$ Argonne National Laboratory, Lemont, IL 60439, USA\\
+$^{24}$ Department of Physics and Astronomy and the Pittsburgh Particle Physics, Astrophysics and Cosmology Center (PITT PACC), University of Pittsburgh, Pittsburgh, PA 15260, USA\\
+$^{25}$ Department of Physics, Stanford University, 382 Via Pueblo Mall, Stanford, CA 94305, USA\\
+$^{26}$ SLAC National Accelerator Laboratory, Menlo Park, CA, 94025, USA\\
+$^{27}$ Department of Statistics, Federal University of Sao Carlos, Sao Carlos, Brazil\\
+$^{28}$ External collaborator\\
+
 \end{document}

### 2019-12-17 paper/authors.tex
modify deRose and Wechsler contributions

% Author list file generated with: mkauthlist 1.2.3 
% mkauthlist -j mnras -f -c LSST Dark Energy Science Collaboration --cntrb contributions.tex authors.csv authors.tex

\author[LSST Dark Energy Science Collaboration]{
\parbox{\textwidth}{
\Large
S.J.~Schmidt$^{1}$,
A.I.~Malz$^{2,3,4}$,
J.Y.H.~Soo$^{5,6}$,
I.A.~Almosallam$^{7,8}$,
M.~Brescia$^{9}$,
S.~Cavuoti$^{9,10}$,
J.~Cohen-Tanugi$^{11}$,
A.J.~Connolly$^{12}$,
J.~DeRose$^{13,14,15,16,17}$,
P.E.~Freeman$^{18}$,
M.L.~Graham$^{12}$,
K.G.~Iyer$^{19,20}$,
M.J.~Jarvis$^{21,22}$,
J.B.~Kalmbach$^{12}$,
E.~Kovacs$^{23}$,
A.B.~Lee$^{18}$,
G.~Longo$^{10}$,
C.B.~Morrison$^{12}$,
J.A.~Newman$^{24}$,
E.~Nourbakhsh$^{1}$,
E.~Nuss$^{11}$,
T.~Pospisil$^{18}$,
H.~Tranin$^{11}$,
R.H.~Wechsler$^{25,16,26}$,
R.~Zhou$^{15,24}$,
R.~Izbicki$^{27,28}$
\begin{center} (LSST Dark Energy Science Collaboration) \end{center}
}
\vspace{0.4cm}
\\
\parbox{\textwidth}{
%\scriptsize
$^{1}$ Department of Physics, University of California, One Shields Ave., Davis, CA, 95616, USA\\
$^{2}$ German Centre of Cosmological Lensing, Ruhr-Universitaet Bochum, Universitaetsstra{\ss}e 150, 44801 Bochum, Germany\\
$^{3}$ Center for Cosmology and Particle Physics, New York University, 726 Broadway, New York, 10003, USA\\
$^{4}$ Department of Physics, New York University, 726 Broadway, New York, 10003, USA\\
$^{5}$ School of Physics, University of Science Malaysia, 11800 USM, Pulau Pinang, Malaysia\\
$^{6}$ Department of Physics and Astronomy, University College London, Gower Street, London WC1E 6BT, UK\\
$^{7}$ King Abdulaziz City for Science and Technology, Riyadh 11442, Saudi Arabia\\
$^{8}$ Information Engineering, Parks Road, Oxford, OX1 3PJ, UK\\
$^{9}$ INAF-Astronomical Observatory of Capodimonte, Salita Moiariello 16, 80131 Napoli, Italy\\
$^{10}$ Department of Physics E. Pancini, University Federico II, via Cinthia 6, I-80126, Napoli, Italy\\
$^{11}$ Laboratoire Univers et Particules de Montpellier, Universit\'e de Montpellier, CNRS, Montpellier, France\\
$^{12}$ DIRAC Institute and Department of Astronomy, University of Washington, Box 351580, U.W., Seattle WA 98195, USA\\
$^{13}$ Santa Cruz Institute for Particle Physics, Santa Cruz, CA 95064, USA\\
$^{14}$ Berkeley Center for Cosmological Physics, Department of Physics, University of California, Berkeley CA 94720\\
$^{15}$ Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, CA 94720, USA\\
$^{16}$ Kavli Institute for Particle Astrophysics and Cosmology and Department of Physics, Stanford University, Stanford, CA 94305, USA\\
$^{17}$ Department of Particle Physics and Astrophysics, SLAC National Accelerator Laboratory, Stanford, CA 94305, USA\\
$^{18}$ Department of Statistics \& Data Science, Carnegie Mellon University, 5000 Forbes Avenue, Pittsburgh, PA 15213, USA\\
$^{19}$ Dunlap Institute for Astronomy \& Astrophysics, University of Toronto, 50 St. George Street, Toronto, ON M5S 3H4, Canada\\
$^{20}$ Department of Physics and Astronomy, Rutgers, The State University of New Jersey, 136 Frelinghuysen Road, Piscataway, NJ 08854-8019 USA\\
$^{21}$ Astrophysics, Department of Physics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford, OX1 3RH, UK\\
$^{22}$ Department of Physics and Astronomy, University of the Western Cape, Bellville 7535, South Africa\\
$^{23}$ Argonne National Laboratory, Lemont, IL 60439, USA\\
$^{24}$ Department of Physics and Astronomy and the Pittsburgh Particle Physics, Astrophysics and Cosmology Center (PITT PACC), University of Pittsburgh, Pittsburgh, PA 15260, USA\\
$^{25}$ Department of Physics, Stanford University, 382 Via Pueblo Mall, Stanford, CA 94305, USA\\
$^{26}$ SLAC National Accelerator Laboratory, Menlo Park, CA, 94025, USA\\
$^{27}$ Department of Statistics, Federal University of Sao Carlos, Sao Carlos, Brazil\\
$^{28}$ External collaborator\\
}
}

### 2019-12-17 paper/authors.tex
re-fix affiliations and modify Mahalanobis formulat in CMNN

@@
 R.~Zhou$^{15,24}$,
-R.~Izbicki$^{27,28}$
-\begin{center} (LSST Dark Energy Science Collaboration) \end{center}
+R.~Izbicki$^{27,28}$,
+and The LSST Dark Energy Science Collaboration
 }
@@
 \parbox{\textwidth}{
-%\scriptsize
-$^{1}$ Department of Physics, University of California, One Shields Ave., Davis, CA, 95616, USA\\
-$^{2}$ German Centre of Cosmological Lensing, Ruhr-Universitaet Bochum, Universitaetsstra{\ss}e 150, 44801 Bochum, Germany\\
-$^{3}$ Center for Cosmology and Particle Physics, New York University, 726 Broadway, New York, 10003, USA\\
-$^{4}$ Department of Physics, New York University, 726 Broadway, New York, 10003, USA\\
-$^{5}$ School of Physics, University of Science Malaysia, 11800 USM, Pulau Pinang, Malaysia\\
-$^{6}$ Department of Physics and Astronomy, University College London, Gower Street, London WC1E 6BT, UK\\
-$^{7}$ King Abdulaziz City for Science and Technology, Riyadh 11442, Saudi Arabia\\
-$^{8}$ Information Engineering, Parks Road, Oxford, OX1 3PJ, UK\\
-$^{9}$ INAF-Astronomical Observatory of Capodimonte, Salita Moiariello 16, 80131 Napoli, Italy\\
-$^{10}$ Department of Physics E. Pancini, University Federico II, via Cinthia 6, I-80126, Napoli, Italy\\
-$^{11}$ Laboratoire Univers et Particules de Montpellier, Universit\'e de Montpellier, CNRS, Montpellier, France\\
-$^{12}$ DIRAC Institute and Department of Astronomy, University of Washington, Box 351580, U.W., Seattle WA 98195, USA\\
-$^{13}$ Santa Cruz Institute for Particle Physics, Santa Cruz, CA 95064, USA\\
-$^{14}$ Berkeley Center for Cosmological Physics, Department of Physics, University of California, Berkeley CA 94720\\
-$^{15}$ Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, CA 94720, USA\\
-$^{16}$ Kavli Institute for Particle Astrophysics and Cosmology and Department of Physics, Stanford University, Stanford, CA 94305, USA\\
-$^{17}$ Department of Particle Physics and Astrophysics, SLAC National Accelerator Laboratory, Stanford, CA 94305, USA\\
-$^{18}$ Department of Statistics \& Data Science, Carnegie Mellon University, 5000 Forbes Avenue, Pittsburgh, PA 15213, USA\\
-$^{19}$ Dunlap Institute for Astronomy \& Astrophysics, University of Toronto, 50 St. George Street, Toronto, ON M5S 3H4, Canada\\
-$^{20}$ Department of Physics and Astronomy, Rutgers, The State University of New Jersey, 136 Frelinghuysen Road, Piscataway, NJ 08854-8019 USA\\
-$^{21}$ Astrophysics, Department of Physics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford, OX1 3RH, UK\\
-$^{22}$ Department of Physics and Astronomy, University of the Western Cape, Bellville 7535, South Africa\\
-$^{23}$ Argonne National Laboratory, Lemont, IL 60439, USA\\
-$^{24}$ Department of Physics and Astronomy and the Pittsburgh Particle Physics, Astrophysics and Cosmology Center (PITT PACC), University of Pittsburgh, Pittsburgh, PA 15260, USA\\
-$^{25}$ Department of Physics, Stanford University, 382 Via Pueblo Mall, Stanford, CA 94305, USA\\
-$^{26}$ SLAC National Accelerator Laboratory, Menlo Park, CA, 94025, USA\\
-$^{27}$ Department of Statistics, Federal University of Sao Carlos, Sao Carlos, Brazil\\
-$^{28}$ External collaborator\\
+(Affiliations are listed at the end of the paper)
+\vspace{0.8cm}
 }

### 2019-12-17 paper/tex/methods.tex
remove superfluous comments from tex
re-fix affiliations and modify Mahalanobis formulat in CMNN

@@
 However, we remind the reader that only 2.0 per cent of our sample has non-detections, almost exclusively in the $u$-band, and thus should not dominate the code performance differences.
-%The specific choices for each code affect the results and contribute to the implicit prior influencing their output.
-
-%The authors welcome interest from those outside \lsstdesc\ to have their codes assessed in future investigations that build upon this one.
 
@@
 We describe the algorithms and implementations of the model-based and data-driven codes in Sections~\ref{sec:templatecodes} and \ref{sec:trainingcodes} respectively, with a straw-person approach included in Section~\ref{sec:trainz}.
-% For each approach, we note how it accounts for photometric uncertainties, how it interprets negative fluxes, what its native output format of \pzpdf s is, and what, if any, preprocessing, including validation, of the prior information it performs.
-% In the following sections, we use a consistent notation defined in Table~\ref{tab:variables}.
-%
-% \begin{table}
-% 	\label{tab:variables}
-% 	\caption{Definitions of variables used in outlining \pzpdf\ codes}
-% 	\begin{tabular}{ll}
-% 		\hline
-% 		\bf Symbol & \bf Definition \\
-% 		\hline
-% 		\multirow{ 2}{*}{$b = 1,\dots,B$	& \multirow{ 2}{*}{number $B$ of photometric filters $b$;} \\
-% 														& \lsst's $ugrizy$ filters used here correspond to $B=6$ \\
-% 		\multirow{ 2}{*}{$m_{b} = 1,\dots,B$	& \multirow{ 2}{*}{number $B$ of photometric filters $b$;} \\
-% 																										& \lsst's $ugrizy$ filters used here correspond to $B=6$ \\
-% 		$z$											& redshift \\
-% 		$T$											& galaxy SED \\
-%
-% 	\end{tabular}
-% \end{table}
 
@@
 We use template interpolation, creating two linearly interpolated templates between each basis SED (sorted by rest-frame $u-g$ colour) by setting the parameter \texttt{INTERP=2}.
-%For photo-$z$ point estimates we use the \texttt{Z\_B} output parameter.
 Prior to running the code, the non-detection placeholder magnitude was replaced with an estimate of the 1-$\sigma$ detection limit for the undetected band as a proxy for a value close to the estimated sky noise threshold.
@@
 \label{sec:eazy}
-%(Rongpu Zhou)
 
@@
 \label{sec:lephare}
-%(C\'ecile Roucelle, Eric Nuss, Johann Cohen-Tanugi)
 
@@
 The reported \pzpdf\ is an arbitrary normalization of the likelihood evaluated on the output redshift grid.
-%\textsc{LePhare} has been used to produce the COSMOS2015 photo-$z$ catalogue \citep{Laigle:16}.
 
@@
 
-% Some aspects of data treatment were left to the individual code runners, for example, whether and how to split the DC1 training set for validation.
-% The codes considered treated
-% Another key difference is the treatment of non-detections in one or more bands.
-% Some codes choose to ignore a band, others replace the value with either an estimate for the detection limit, the mean of other values in the training set, or another default value.
-% There are varying conventions among training-based codes for treatment of non-detections, and no one prescription dominates in the photo-$z$ literature.
-% The specific choices for each code affect the results, and contribute to the implicit prior influencing their output.
-% However, we remind the reader that only 2.0 per cent of our sample has non-detections, almost exclusively in the u-band, and thus should not dominate the code performance differences.
-
 \subsubsection{ANNz2}
 \label{sec:annz2}
-%(John Soo)
 
@@
 Using the Toolkit for Multivariate Data Analysis with ROOT\footnote{\url{http://tmva.sourceforge.net/}}, \annz\ can return the results of running a single machine learning algorithm, a ``best'' choice of the results from simultaneously running multiple algorithms (based on evaluation the cumulative distribution functions of validation set objects), or a combination of the results of multiple algorithms weighted by their method uncertainties averaged over multiple runs.
-%\textsc{ANNz2} is capable of producing both photo-$z$ point estimates and redshift posterior probability distributions $p(z)$.
-% It can also perform classifications, and supports reweighting between samples.
-% \annz\ propagates the intrinsic uncertainty on the input parameters and the uncertainty in the machine learning method to the expected photo-$z$ solution, averaged over multiple runs weighted based on the performance of each run.
 
@@
 \label{sec:cmnn}
-%(Melissa Graham)
 
 The colour-matched nearest-neighbours photometric redshift estimator \citep[\cmnn,][]{Graham:17} uses a training set of galaxies with known redshifts that has equivalent or better photometry than the test set in terms of quality and filter coverage.
-For each galaxy in the test set, \cmnn\ identifies a colour-matched subset of training galaxies using a threshold in the Mahalanobis distance $D_M = \sum_{j}^{N_{\mathrm{colours}}} (c^{\mathrm{train}}_{j} - c^{\mathrm{test}}_{j})^{2} / \delta c_{\mathrm{test}}^2$ in the space of available colours $c$, with colour measurement errors $\delta c_{\mathrm{test}}$ and $N_{\mathrm{colours}} = 5$ colours $j$ defined by the $ugrizy$ filters, which defines the set of colour-matched neighbours based on a value of the percent point function (PPF).
+For each galaxy in the test set, \cmnn\ identifies a colour-matched subset of training galaxies using a threshold in the Mahalanobis distance $D_M = \sum_{j}^{N_{\mathrm{colours}}} (c_{\mathrm{j,train}} - c_{\mathrm{j,test}})^{2} / \delta c_{\mathrm{j,test}}^2$ in the space of available colours $c$, with colour measurement errors $\delta c_{\mathrm{test}}$ and $N_{\mathrm{colours}} = 5$ colours $j$ defined by the $ugrizy$ filters, which defines the set of colour-matched neighbours based on a value of the percent point function (PPF).
 As an example, for $N_{\mathrm{filt}}=5$ with PPF$=0.95$, $95$ per cent of all training galaxies consistent with the test galaxy will have $D_M < 11.07$.
@@
 \label{sec:flexzboost}
-%(Ann Lee, Rafael Izbicki, Taylor Pospisil, Peter Freeman)
 
@@
 
-%\subsubsection{Frankenz}
-%\label{sec:frankenz}
-%(seeking volunteers)
-%
-%\red{\textsc{Frankenz}\footnote{\url{https://github.com/joshspeagle/frankenz}} \cite{Speagle:frankenz} is...}
-%
-
 \subsubsection{GPz}
 \label{sec:gpz}
-%(Ibrahim Almosallam)
 
@@
 The mean $\mu(\x)$ and the variance $\sigma(\x)^{2}$ are modeled as functions $f(\x) = \sum_{i=1}^{m}w_{i}\phi_{i}(\x)$ that are linear combinations of $m$ basis functions $\left\{\phi_{i}(\x)\right\}_{i=1}^{m}$ with associated weights $\left\{w_{i}\right\}_{i=1}^{m}$.
-% Basis function models, for specific classes of basis functions such as the sigmoid or the squared exponential, have the advantage of being universal approximators, i.e.
-there exist a function of that form that can approximate any function, with mild assumptions, to any desired degree of accuracy.
 The details on how to learn the parameters of the model and the hyper-parameters of the basis functions are described in \citet{Almosallam:15b}.
@@
 To smooth the long tail in the distribution of magnitude errors, we use the logarithm of the magnitude errors, improving numerical stability and eliminating the need for constraints on the optimization process.
-% We use principaphoto-\mathinhead{$z$}{z} interim posteriorsl component analysis (PCA) to decorrelate the data
-% WHAT IS BEING DECORRELATED HERE?
 Unobserved magnitudes $x_{\mathrm{u}} = \mu_{\mathrm{u}} + \Sigma_{\mathrm{uo}}\Sigma_{\mathrm{oo}}^{-1}(x_{\mathrm{o}} - \mu_{\mathrm{o}})$ were imputed from observed magnitudes $x_{\mathrm{o}}$ and the training set mean $\mu$ and covariance $\Sigma$ using a linear model.
@@
 \label{sec:metaphor}
-%(Stefano Cavuoti, Massimo Brescia, Giuseppe Longo)
 
 Machine-learning Estimation Tool for Accurate Photometric Redshifts \citep[\metaphor,][]{Cavuoti:17} is based on the Multi Layer Perceptron with Quasi Newton Algorithm (MLPQNA) with the least square error model and Tikhonov $L_{2}$-norm regularization \citep{Hofmann:18}.
-%, already validated on photo-$z$'s in several cases \citep{de_Jong:17,Cavuoti:17b,Cavuoti:15,Brescia:14,Brescia:13,Biviano:13}.
 \Pzpdf s are generated by running $N$ trainings on the same training set, or $M$ trainings on $M$ different random samplings of the training set.
 Upon regression of the test set, the photometry $m_{ij}$ of each test set galaxy $j$ in filter $i$ is perturbed according to $m_{ij}' = m_{ij} + \alpha_{i} F_{ij} \epsilon$ in terms of the standard normal random variable $\epsilon \sim \mathcal{N}(0, 1)$, a multiplicative constant $\alpha_{i}$ permitting accommodation of multi-survey photometry, and a bimodal function $F_{ij}$ composed of a polynomial fit of the mean magnitude errors on the binned bands plus a constant term representing the threshold below which the polynomial's noise contribution is negligible \citep{Brescia:18}.
-% At a higher level, the pipeline mainly consists of three modules: (i) \textit{data pre-processing}, including a catalogue cross-matching sub-module \citep[based on the tool C3, ][]{Riccio:17}, a sub-module for photometric evaluation and error estimation of the multi-band catalogue used as Knowledge Base (KB), and a sub-module dedicated to the perturbation of the photometric KB, propaedeutic to the PDF estimation; (ii) \textit{photo-$z$ prediction}, which is the training/validation/test phase, producing the photo-$z$'s point estimates, based on a pre-selected ML method; (iii) \textit{PDF estimation}, specifically designed to calculate the PDF of the photo-$z$ estimation errors.
-% The last module includes also a post-processing tool, providing some statistics on the produced point estimates and PDFs.
 
@@
 \label{sec:skynet}
-%(J. Cohen-Tanugi and Hugo Tranin)
 
 \skynet \citep{Graff:14} employs a neural network based on a second-order conjugate gradient optimization scheme \citep[see][for further details]{Graff:14}. 
-%It has been used efficiently for redshift PDF estimates \citep{Sanchez:14,Bonnett:15,Bonnett:16}.
 The neural network is configured as a standard multilayer perceptron with three hidden layers and one input layer with 12 nodes corresponding to the 6 photometric magnitudes and their measurement errors.
-% We use \skynet\ as a regressor for \pz\ point estimation and as a classifier for \pzpdf\ estimation.
 
-% The regressor used a standard \chisq\ error function with a single linear node as the output layer and 10 nodes with a $\tanh$ activation function for each hidden layer.
 \skynet's classifier mode uses a cross-entropy error function with a 20:40:40 node (all rectified linear units) architecture for each hidden layer and an output layer of 200 nodes corresponding to 200 bins for the PDF, with a softmax activation function to enforce the normalization condition that the probabilities sum to unity.
@@
 We pre-whitened the data by pegging the magnitudes to (45,45,40,35,42,42) and errors to (20,20,10,5,15,15) for $ugrizy$ filters, respectively.
-% NO CLUDE WHAT THIS MEANS
 To avoid over-fitting, $30$ per cent of the training set was reserved for validation, and training was halted as soon as the error rate began to increase on the validation set.
 The weights were randomly initialized based on normal sampling.
-% WHAT WEIGHTS?
 
@@
 \label{sec:tpz}
-%(Erfan Nourbakhsh)
 
@@
 We will demonstrate its performance under the metrics of Section~\ref{sec:metrics} and discuss it as an illustrative experimental control case in Section~\ref{sec:caution} to highlight the limitations of our evaluation criteria for \pzpdf s.
-% any attempts to break up the sample into tomographic bins will fail, as every galaxy has an identical $p(z)$.

### 2019-12-17 paper/authors.tex
update contact author emails

@@
 \Large
-S.J.~Schmidt$^{1}$,
-A.I.~Malz$^{2,3,4}$,
+S.J.~Schmidt$^{1}$\thanks{Email: sschmidt@physics.ucdavis.edu},
+A.I.~Malz$^{2,3,4}$\thanks{Email: aimalz@astro.ruhr-uni-bochum.de},
 J.Y.H.~Soo$^{5,6}$,

### 2019-12-17 paper/contributions.tex

Author contributions are listed below. \\
S.J.~Schmidt: Co-led the project. (conceptualization, data curation, formal analysis, investigation, methodology, project administration, resources, software, supervision, visualization, writing -- original draft, writing -- review \& editing). \\
A.I.~Malz: Co-led the project, contributed to choice of metrics, implementation in code, and writing. (conceptualization, methodology, project administration, resources, software, visualization, writing -- original draft, writing -- review \& editing). \\
J.Y.H.~Soo: Ran ANNz2 and Delight, updated abstract, edited sections 1 through 6, added tables in Methods and Results, updated references.bib and added references throughout the paper. \\
I.A.~Almosallam: vetted the early versions of the data set and ran many photo-z codes on it, applied GPz to the final version and wrote the GPz subsection. \\
M.~Brescia: Main-ideator of MLPQNA and co-ideator of METAPHOR; modification of METAPHOR pipeline to fit the LSST data structure and requirements. \\
S.~Cavuoti: Co-ideator of METAPHOR, contributed to choice and test of metrics, ran METAPHOR, minor text editing. \\
J.~Cohen-Tanugi: contributed to running code, analysis discussion, and editing, reviewing the paper. \\
A.J.~Connolly: Developed the colour-matched nearest-neighbours photo-z code; participated in discussions of the analysis. \\
J.~DeRose: One of the primary developers of the Buzzard-highres simulated galaxy catalogue employed in the analysis. \\
P.E.~Freeman: Contributed to choice of CDE metrics and to implementation of FlexZBoost. \\
M.L.~Graham: Ran the colour-matched nearest-neighbours photo-z code on the Buzzard catalogue and wrote the relevant piece of Section 2; participated in discussions of the analysis. \\
K.G.~Iyer: assisted in writing metric functions used to evaluate codes. \\
M.J.~Jarvis: Contributed text on AGN to Discussion section and portions of GPz work. \\
J.B.~Kalmbach: Worked on preparing the figures for the paper. \\
E.~Kovacs: Ran simulations, discussed data format and properties for SEDs, dust, and ELG corrrections. \\
A.B.~Lee: Co-developed FlexZBoost and the CDE loss statistic, wrote text on the work, and supervised the development of FlexZBoost software packages. \\
G.~Longo: Scientific advise, test and validation of the modified METAPHOR pipeline, text of the METAPHOR section. \\
C.B.~Morrison: Managerial support; Discussions with authors regarding metrics and style; Some coding contribution to metric computation. \\
J.A.~Newman: Contributions to overall strategy, design of metrics, and supervision of work done by Rongpu Zhou. \\
E.~Nourbakhsh: Ran and optimized TPZ code and wrote a subsection of Section 2 for TPZ. \\
E.~Nuss: contributed to running code, analysis discussion, and editing,reviewing the paper. \\
T.~Pospisil: Co-developed FlexZBoost software and CDE loss calculation code. \\
H.~Tranin: contributed to providing SkyNet results and writing the relevant section. \\
R.H.~Wechsler: Project lead for Buzzard-highres simulated galaxy catalogue employed in analysis. \\
R.~Zhou: Optimized and ran EAZY and contributed to the draft. \\
R.~Izbicki: Co-developed FlexZBoost and the CDE loss statistic, and wrote software for FlexZBoost \\

### 2019-12-17 paper/tex/discussion.tex

@@
 \label{sec:discussion}
-%(Sam Schmidt, Ofer Lahav, Jeff Newman, Alex Malz, Eve Kovacs, Tony Tyson, Tina Peters)
 In contrast with other \pzpdf\ comparison papers that have aimed to identify the ``best'' code for a given survey, we have focused on the somewhat more philosophical questions of how to assess \pzpdf\ methods and how to interpret differences between codes in terms of \pzpdf\ performance.
@@
 \label{sec:caution}
-%(Alex Malz, Sam Schmidt)
 
@@
 The specific science goal naturally guides the choice of metric to focus on $N(z)$ rather than individual \pzpdf s, and on the basis of that metric such improvements upon \trainz\ are guaranteed to be more robust to training set imbalance.}.
-%\textit{Explain the connection between \trainz\ and these metrics.}
 
@@
 
-% While looking at one, or even most of our metrics would have given the impression that this estimator was nearly optimal, red flags in CDE loss and point estimates reveal the problems.
-%[more caution on over-reliance on metrics, and making sure that metrics test all of what you want] \red{this definitely still needs work}
 In summary, context is crucial to defend against deceptively strong performers such as \trainz;
@@
 In the absence of a single scientific motivation or the information necessary for a principled metric definition, we must consider many metrics and be critical of the information transmitted by each.
-% As stated earlier, science applications may be sensitive to individual \pzpdf s, \pzpdf\ ensembles, \pz\ point estimates, or science-specific statistics like $\hat{N}(z)$.
-% We must be aware of that the optimal method for one is not necessarily optimal for the other, and in fact several photo-$z$ algorithms may be necessary in the final cosmological analysis in order to satisfy the requirements of all science use cases.
-% The example of the simple \trainz\ estimator described in Section~\ref{sec:caution} shows a simple model with a $p(z)$ that is unrealistic for individual objects can still score very well on many of our metrics.
-% It is important to look at {\it all} metrics, and keep in mind what information each metric conveys.
-%\aim{[reiterate the meaning of $p(z)$ and the goal -- emphasize finding ``best'' code in terms of impact of assumptions underlying the method \textit{or} establishing methodology for the realistic case of biased prior information and other science goals]}
-%\red{[the challenges faced in this project]}
-%\red{mention z$<$2, missing some important degeneracies, new sim will go to higher z}
-%\red{no quality cuts included, could identify outliers, but could also induce biases}
-%\red{[how optimization defined has impact, here we optimized p(z), not n(z), could be science-case specific if stringent requirements are needed, though that may be a computational/storage challenge if too many use cases needed.]}
-%\red{some discussion of output parameterization, limitations of 200 point fixed grid, in future switch to something like quantiles (cite qp paper)}
 
@@
 
-%\begin{enumerate}
-%\item \red{the combination of codes}
-%\item \red{$p(z,\alpha)$}
-%\item \red{tomographic analysis later on}
-%\red{this paper deals with training, future papers will also explore calibration via cross-correlation methods}
-%\end{enumerate}
-%\red{mention the SRM once again, say that the plans are extensive, but we do have a plan and a rough timeline.}
-
 The work presented in this paper is only a first step in assessing \pzpdf\ approaches and moving toward a photometric redshift estimator that will be employed for LSST analyses.
 Extensions of the experimental design will require further rounds of analyses, and the authors welcome interest from those outside \lsstdesc\ to have their codes assessed in these future investigations.
-%Here we discuss the next steps for subsequent investigations.
 
@@
 
-%\red{[are there any references for this?
-I remember Gary Bernstein talking about this at a photo-z workshop in Japan, but I don't know that it was published.
-I believe Michael Troxel has discussed this as well.]}
-%\item \red{cosmological parameters in DC2}
-%\item \red{mention z$<$2, missing some important degeneracies, new sim will go to higher z}
-%\red{no quality cuts included, could identify outliers, but could also induce biases}
-
 Sequels to this study will also address some shortcomings of our experimental procedure.
@@
 In a follow-up study, we will
- % use generative adversarial networks\citep[(GANs)][]{Goodfellow:14} trained on the simulation data to
 generate mock data probabilistically, yielding true PDFs in addition to true redshifts and photometric data.
@@
 To make optimal use of the \lsst\ data for cosmological and other astrophysical analyses of the \lsstdesc\ SRM, future investigations that build upon this one will require a more sophisticated set of galaxy photometry and redshifts.
-%\item \red{inclusion of photometric errors, image based effects, blending, lensing, sky, mask, etc...}
 This initial paper explored a data set that was constructed at the catalogue level, with no inclusion of the complications that arise from photometric measurements of imaging data.
@@
 To gauge the sensitivity of \pzpdf\ estimators to catastrophic outliers, our data set must include realistic high-redshift galaxy populations.
-
-% In this study we have not accounted for the presence of Active Galactic Nuclei (AGN) contributions to galaxy fluxes.
-% In some cases, AGN will be easily identified from the colors and morphologies, i.e.
-the case of the brightest quasars where the nuclear activity outshines the host galaxy, and numerous studies have utilized color selection to create large samples of quasars \citep[e.g.][]{Richards:06,Maddox:08,Richards:15}.
-% In current deep fields, similar in depth to what we expect from \lsst, variability information and multi-wavelength data have been critical to not only identify AGN dominated galaxies, but also obtain more accurate photometric redshifts \citep[e.g][]{Salvato:11}.
-%
-% In addition to AGN dominated galaxies, those with lower levels of nuclear activity present a more insidious problem, where AGN features may not be apparent, but the colours and other host galaxy properties are perturbed relative to galaxies with an inactive nucleus.
-% In such cases, the presence of the AGN may induce a bias if the template SEDs or empirical datasets do not include low-level AGN counterparts.
-% For LSST, we will need to identify and obtain accurate photometric redshifts of all types of AGN for a range of science goals, whether it is to eliminate such objects from cosmology experiments, or to use them with confidence, all the way through to understanding galaxy evolution and the role that AGN may play in influencing galaxy properties over cosmic time.
-
-% A promising route to classifying and obtaining accurate photometric redshifts for the AGN population is by combining machine learning with template-fitting techniques, as has recently been demonstrated by \citet{Duncan:18} for radio-selected AGN.
-% This is because AGN are relatively easy to obtain spectroscopic redshifts for over all redshifts due to the strong emission lines that they exhibit, allowing very good training sets for machine learning algorithms to use.
-% Whereas for those galaxies where the AGN is sub-dominant the galaxy templates are still adequate for obtaining reasonable photometric redshifts.
-
-% In addition to these improvements, the \lsstdesc \Pz\ Working Group plans to look at all potential methods to combine the results from multiple \pzpdf\ codes to improve accuracy, similar to the work presented in \citet{Dahlen:13,Carrascokind:14,Duncan:18}.
-% Taking advantage of multiple algorithms that use observables in slightly different ways has shown promise, however we must be very conscious of whether a potential combination properly treats the covariance between the methods, given that they are estimating quantities based on the same underlying observables.
-% Several science cases wish to estimate physical quantities along with redshift, for example galaxy stellar mass and star formation rate.
-% Proper joint estimation of redshift and physical quantities requires an in depth understanding of galaxy evolution, and progress on accurate bivariate redshift probability distributions will go hand in had with progress on understanding galaxies themselves.
-% Parameterization and storage of a complex 2-dimensional probability surface for potentially billions of galaxies (or even subsets of hundreds of thousand of particular interest) pose a potential challenge.
-% These issues will be examined in another future paper.
-
-% Finally, while this paper and future papers discussed above focus on photometric redshift codes and estimating accurate \pzpdf s from training data, we plan a separate, but complementary, project to examine calibration of the resultant redshifts via spatial cross-correlations \citep{Newman:2008}, which will be explored in a separate series of future papers.
-%The overarching plan describing everything laid out in this section is described in more detail in the \lsstdesc\ Science Roadmap (see Footnote in Section~\ref{sec:intro}).

### 2020-02-28 paper/main.tex
fix LSST naming and typo in appendix

@@
 \usepackage{footmisc}
+\usepackage{ulem}
 \graphicspath{{./}{./figures/}}
@@
 
+\newcommand{\boldblue}[1]{{\bf \textcolor{blue}{#1}}}
+
 \def\X{{\mathbf{X}}}
@@
 
-\title[Photo-z estimation approaches for LSST]{Evaluation of probabilistic photometric redshift estimation approaches for \textsc{LSST}}
+\title[Photo-z estimation approaches for LSST]{Evaluation of probabilistic photometric redshift estimation approaches for \boldblue{ \textsc{The Rubin Observatory Legacy Survey of Space and Time} (LSST)}}
 
@@
 A plethora of photo-$z$ PDF estimation methodologies abound, producing discrepant results with no consensus on a preferred approach.
-We present the results of a comprehensive experiment comparing twelve photo-$z$ algorithms applied to mock data produced for the Large Synoptic Survey Telescope (\textsc{LSST}) Dark Energy Science Collaboration (\textsc{DESC}).
+We present the results of a comprehensive experiment comparing twelve photo-$z$ algorithms applied to mock data produced for \sout{Large Synoptic Survey Telescope} \boldblue{The Rubin Observatory Legacy Survey of Space and Time} (\textsc{LSST}) Dark Energy Science Collaboration (\textsc{DESC}).
 By supplying perfect prior information, in the form of the complete template library and a representative training set as inputs to each code, we demonstrate the impact of the assumptions underlying each technique on the output photo-$z$ PDFs.
@@
 % hack event acknowledgments are important
-%We thank Stony Brook University for hosting the Summer 2017 LSST-DESC Hack Day at which this work was partially completed.
+We thank Stony Brook University for hosting the Summer 2017 LSST-DESC Hack Day at which this work was partially completed.
 

### 2020-02-28 paper/tex/introduction.tex
fix LSST naming and typo in appendix

@@
 
-The current and next generations of large-scale galaxy surveys, including the Dark Energy Survey \citep[\proj{DES},][]{Abbott:05}, the Kilo-Degree Survey \citep[\proj{KiDS},][]{de_Jong:13}, Hyper Suprime-Cam Survey \citep[\proj{HSC},][]{Aihara:2018a,Aihara:2018b}, Large Synoptic Survey Telescope \citep[\lsst,][]{Abell:09}, Euclid \citep{Laureijs:11}, and Wide-Field Infrared Survey Telescope \citep[\proj{WFIRST},][]{Green:12}, represent a paradigm shift to reliance on photometric, rather than solely spectroscopic, galaxy catalogues of substantially larger size at a cost of lacking complete spectroscopically confirmed redshifts ($z$).
+The current and next generations of large-scale galaxy surveys, including the Dark Energy Survey \citep[\proj{DES},][]{Abbott:05}, the Kilo-Degree Survey \citep[\proj{KiDS},][]{de_Jong:13}, Hyper Suprime-Cam Survey \citep[\proj{HSC},][]{Aihara:2018a,Aihara:2018b}, \sout{Large Synoptic Survey Telescope} \boldblue{The Rubin Observatory Legacy Survey of Space and Time} \citep[\lsst,][]{Abell:09}, Euclid \citep{Laureijs:11}, and Wide-Field Infrared Survey Telescope \citep[\proj{WFIRST},][]{Green:12}, represent a paradigm shift to reliance on photometric, rather than solely spectroscopic, galaxy catalogues of substantially larger size at a cost of lacking complete spectroscopically confirmed redshifts ($z$).
 Effective astrophysical inference using the catalogues resulting from these ongoing and upcoming missions, however, necessitates accurate and precise photometric redshift (\pz) estimation methodologies.
@@
 In this paper we focus solely on individual galaxy redshift estimates.} rather than individual galaxies.
-
-% MOVE TO WHERE \nz\ IS INTRODUCED AS A METRIC
-% For cosmological measurements, certain science cases require redshift information on individual objects, e.~g.~identification of host galaxy redshift for supernova classification, or identifying potential cluster membership.
-% Other science cases seem to need only ensemble redshift information; for instance many current cosmic shear techniques require only the overall redshift distribution $N(z)$ for tomographic redshift samples.
-% However,  even such cases require individual object redshift estimates for portions of the analysis, for example in determining galaxy intrinsic alignments in weak lensing samples.
-% In addition, recent data-driven techniques employing hierarchical Bayesian or Gaussian Process methods have emerged that calibrate redshift distributions using individual $p(z)$ estimates \citep[e.~g.~][]{Sanchez:2018}.
-% Techniques for using \pzpdf s have lagged behind the development of codes to produce them, however, all existing approaches assume that the \pzpdf\ for each galaxy is an accurate PDF, with failures when the assumption is violated.
-% Thus, even methods that seem to need only ensemble $N(z)$ may actually require accurate $p(z)$ in order to meet stringent survey requirements.
-% Large photometric surveys such as LSST must develop algorithms that simultaneously meet the needs of all science cases.
-% In order to meet these ambitious goals for photo-$z$ accuracy, every aspect of photo-$z$ estimation will have to be optimized: the algorithms employed, both template and machine-learning based (both in design and implementation); the spectroscopic data used as a training set for machine learning algorithms or to estimate template sets and train Bayesian priors; and probabilistic catalogue compression schemes that balance information retention against limited storage resources.
 

### 2020-03-04 paper/tex/data.tex
address some referee comments for sec 2 and 3

@@
 By necessity, \textsc{Buzzard} does not contain all of the complicating factors present in real data, and here we discuss the most pertinent ways that these limitations affect our experiment.
-% In several cases, the simplification is done with a purpose, with potentially confounding effects excluded in order to better isolate the differences between current-generation photo-$z$ codes and their causes.
 \textsc{Buzzard} includes only galaxies, not stars nor AGN.
@@
 We assigned magnitude errors in the six bands using the simple model of \citet{Ivezic:08}, assuming achievement of the full 10-year depth, with a modification of fiducial \lsst\ total numbers of 30-second visits for photometric error generation: we assume 60 visits in $u$-band, 80 visits in $g$-band, 180 visits in $r$-band, 180 visits in $i$-band, 160 visits in $z$-band, and 160 visits in $y$-band.
-%numbers taken from Alex Abate's photErrorModel.py script in PhotoZDC1 repository, which was used with nYrObs=10.
 
@@
 The full dataset thus covers $400$ square degrees and contains $238$ million galaxies of redshift $0 < z \leq 8.7$ down to $r = 29$.
-Systematic inconsistencies with galaxy colors at $z > 2$ were observed, so the catalogue was limited to $0 < z \leq 2.0$.
+\boldblue{However, a problem with the k-correction calculation led to incorrect colors at $z>2$, which necessitated limiting the redshift range to $0 < z \leq 2.0$.}
+\sout{Systematic inconsistencies with galaxy colors at $z > 2$ were observed, so the catalogue was limited to $0 < z \leq 2.0$.}
 To obtain a catalogue matching the \lsst\ Gold Sample, we imposed a cut of $i < 25.3$, which gives a signal-to-noise ratio $\gtrsim 30$ for most galaxies.
@@
 
-% NOW MOTIVATE THE SHARED PRIOR INFORMATION
 For the purpose of performing a controlled experiment that compares \pzpdf\ codes on equal footing as a baseline for a future sensitivity analysis, we take care to provide each with optimal prior information.
@@
 
-Following the findings of \citet{Bernstein:10}, \citet{Newman:2015}, and \citet{Masters:2015} that only $\sim\!10^{4}$ spectra are necessary to calibrate \pz s to Stage IV requirements, we aimed to set aside a randomly selected training set of $3-5\times 10^{4}$ galaxies, $\sim 10$ per cent of the full sample.
+Following the findings of \citet{Bernstein:10}, \citet{Newman:2015}, and \citet{Masters:2015} that\boldblue{, in the idealized case of representative training data,} only $\sim\!10^{4}$ spectra are necessary to calibrate \pz s to Stage IV requirements, we aimed to set aside a randomly selected training set of $3-5\times 10^{4}$ galaxies, $\sim 10$ per cent of the full sample.
 After all cuts described above, we designated the \textit{DC1 training set} of $44\,404$ galaxies for which observed photometry, true SEDs, and true redshifts would be provided to all codes and the blinded \textit{DC1 test set} of $399\,356$ galaxies for which photometry alone would be provided to all codes and \pzpdf s would be requested.
-%The resulting catalogues contain $111\,171$ training galaxies and $1\,000\,883$ test galaxies.
 The exact form of \lsst\ photometric filter transmission curves were also considered public information that could be used by any code.

### 2020-03-04 paper/tex/methods.tex
address some referee comments for sec 2 and 3

@@
 
-Another key difference is the treatment of non-detections in one or more bands: some codes ignore incomplete bands, while others replace the value with either an estimate for the detection limit, the mean of other values in the training set, or another default value.
+\sout{Another key difference is the treatment of non-detections in one or more bands: some codes ignore incomplete bands, while others replace the value with either an estimate for the detection limit, the mean of other values in the training set, or another default value.
 There are varying conventions among machine learning-based codes for treatment of non-detections, and no one prescription dominates in the \pz\ literature.
-However, we remind the reader that only 2.0 per cent of our sample has non-detections, almost exclusively in the $u$-band, and thus should not dominate the code performance differences.
+However, we remind the reader that only 2.0 per cent of our sample has non-detections, almost exclusively in the $u$-band, and thus should not dominate the code performance differences.}
 
@@
 \noindent where $A$ is a normalization factor, $F^i_{\mathrm{pred}}(T,z)$ is the flux predicted for a template $T$ at redshift $z$, $F^i_{\mathrm{obs}}$ is the observed flux in a given band $i$, $\sigma^i_{\mathrm{obs}}$ is the observed flux error, and $N_{\mathrm filt}$ is the total number of filters, in our case the six $ugrizy$ LSST filters.
-The likelihood is a sum of observed flux error $\sigma_{b}^{\mathrm{obs}}$-weighted squared differences between the observed flux $F^{\mathrm{obs}}_{b}$ and the normalized predicted flux $F^{\mathrm{mod}}_{b}(T, z)$ in $N_{\mathrm{filt}}$ photometric filters $b$, which are the \lsst\ $ugrizy$ filters in this case.
+\boldblue{All three template-based codes replace per-band non-detections with an estimate of the 1$\sigma$ magnitude limit as estimated from the photometric error model.}
+\sout{The likelihood is a sum of observed flux error $\sigma_{b}^{\mathrm{obs}}$-weighted squared differences between the observed flux $F^{\mathrm{obs}}_{b}$ and the normalized predicted flux $F^{\mathrm{mod}}_{b}(T, z)$ in $N_{\mathrm{filt}}$ photometric filters $b$, which are the \lsst\ $ugrizy$ filters in this case.}
 Specific implementation details of each code, e.~g.~prior form and implementation, are described below.
@@
 Unlike both \bpz\ and \eazy, \lephare\ uses generic, SED-independent priors that are not tuned to the DC1 data set.
+\boldblue{The choice to use the generic prior form could lead to a degradation in code performance, given the slight mismatch between the galaxy distributions in the studied dataset and that parameterized in the \lephare\ prior.}
 
@@
 Because the algorithms differ more from one another and the techniques are relative newcomers to the astronomical literature, we provide somewhat more detail about the implementations below.
+\boldblue{A key difference between the codes is the treatment of non-detections in one or more bands: some codes ignore per-band non-detections, while others replace the value with either an estimate for the detection limit, the mean of other values in the training set, or another default value.
+There are varying conventions among machine learning-based codes for treatment of non-detections, and no one prescription dominates in the \pz\ literature.
+However, we remind the reader that only 2.0 per cent of our sample has non-detections, almost exclusively in the $u$-band, and thus should not dominate the code performance differences.
+}
 
@@
 
-\annz \citep{Sadeh:16} supports several machine learning algorithms, including artificial neural networks (ANN), boosted decision tree, and k-nearest neighbour (KNN) regression.
+\annz\ \citep{Sadeh:16} supports several machine learning algorithms, including artificial neural networks (ANN), boosted decision tree, and k-nearest neighbour (KNN) regression.
 In addition to accounting for errors on the input photometry, \annz\ uses the KNN-uncertainty estimate of \citet{Oyaizu:08} to quantify uncertainty in the choice of method over multiple runs.
@@
 Each of the five ANNs was trained with different random seeds for the initialization of input parameters, reserving half of the training set for validation to prevent overfitting.
-Undetected galaxies were excluded from the training set, and per-band non-detections in the test set were replaced with the mean magnitude in that band within the entire test set.
+\boldblue{Galaxies not detected in the $u$-band} \sout{Undetected galaxies} were excluded from the training set, and per-band non-detections in the test set were replaced with the mean magnitude \boldblue{of the test set} in that band.
+\sout{within the entire test set.}
 
@@
 
-\delight \citep{Leistedt:17} is a hybrid technique that infers \pz s with a data-driven model of latent SEDs and a physical model of photometric fluxes as a function of redshift.
+\delight\ \citep{Leistedt:17} is a hybrid technique that infers \pz s with a data-driven model of latent SEDs and a physical model of photometric fluxes as a function of redshift.
 Generally, machine learning methods rely on representative training data with shared photometric filters, while template-based methods rely on a complete library of templates based on physical models constructed.
@@
 
-\flexzboost \citep{Izbicki:17,Dalmasso:2019} is built on \texttt{FlexCode}, a general-purpose methodology for converting any conditional mean point estimator of $z$ to a conditional density estimator $p(z \vert \x) \equiv f(z \vert \x)$, where $\x$ here represents our photometric covariates and errors.
+\flexzboost\ \citep{Izbicki:17,Dalmasso:2019} is built on \texttt{FlexCode}, a general-purpose methodology for converting any conditional mean point estimator of $z$ to a conditional density estimator $p(z \vert \x) \equiv f(z \vert \x)$, where $\x$ here represents our photometric covariates and errors.
 \flexzboost\ expands the unknown function $f(z \vert \x) = \sum_{i}\beta_{i}(\x)\phi_{i}(z)$ using an orthonormal basis $\{\phi_{i}(z)\}_{i}$.
@@
 
-\gpz \citep{Almosallam:16a,Almosallam:15b} is a sparse Gaussian process-based code, a scalable approximation of full Gaussian Processes \citep{Rasmussen:06}, that produces input-dependent variance estimates corresponding to heteroscedastic noise.
+\gpz\ \citep{Almosallam:16a,Almosallam:15b} is a sparse Gaussian process-based code, a scalable approximation of full Gaussian Processes \citep{Rasmussen:06}, that produces input-dependent variance estimates corresponding to heteroscedastic noise.
 The model assumes a Gaussian posterior probability $p(z \vert \x) = \mathcal{N}\left(z \vert \mu(\x), \sigma(\x)^{2}\right)$ of the output redshift $z$ given the input photometry $\x$.
@@
 
-\skynet \citep{Graff:14} employs a neural network based on a second-order conjugate gradient optimization scheme \citep[see][for further details]{Graff:14}. 
+\skynet\ \citep{Graff:14} employs a neural network based on a second-order conjugate gradient optimization scheme \citep[see][for further details]{Graff:14}. 
 The neural network is configured as a standard multilayer perceptron with three hidden layers and one input layer with 12 nodes corresponding to the 6 photometric magnitudes and their measurement errors.

### 2020-03-06 paper/tex/discussion.tex
add paragraph to intro

@@
 The PIT and other CDF-based metrics upon which modern \pzpdf\ comparisons are built \citep{Bordoloi:10,Polsterer:16,Tanaka:17} can be gamed by a trivial estimator that yields only an affirmation of prior knowledge uninformed by the data.
+\boldblue{The extreme case of \trainz\ also exemplifies the fact that characteristics of the training data are imprinted on the ensemble distributions via the prior, and any imperfections and biases in the training sample will be propagated to the measured \pzpdf s.}
 In other words, such ensemble metrics are insufficient for the task of selecting \pzpdf\ codes for analysis pipelines\footnote{That being said, we note that close relatives of \trainz\ have been employed by weak lensing surveys of the past and present \citep{Lima:08, Hildebrandt:17, Hoyle:18} to estimate $N(z)$ by assigning each test set galaxy the redshift distribution of a \emph{subset} of the training set, where the subset is defined as similar to the test set galaxy in the space of photometric data.
@@
 In summary, context is crucial to defend against deceptively strong performers such as \trainz;
-\textbf{the best \pzpdf\ method is the one that most effectively achieves the science goals of a particular study}, not the one that performs best on a metric that does not reflect those goals.
+the best \pzpdf\ method is the one that most effectively achieves the science goals of a particular study, not the one that performs best on a metric that does not reflect those goals.
 In the absence of a single scientific motivation or the information necessary for a principled metric definition, we must consider many metrics and be critical of the information transmitted by each.

### 2020-03-06 paper/tex/introduction.tex
add paragraph to intro

@@
 
+\boldblue{In currently available spectroscopic training sets systematic incompleteness, particularly at faint magnitudes, and incorrectly assigned redshifts are a harrowing problem.
+Certain subpopulations can potentially fail to yield secure redshifts, given limited wavelength coverage.
+Alternately, redshifts may be  incorrectly assigned for certain populations due to misidentification of spectral features.
+These missing or misidentified populations would yield training sets that are not representative of the true underlying galaxy population, which can bias redshift estimates well beyond the levels required for upcoming analyses.
+Characterizing and mitigating these problems will be a major undertaking which we will cover in future papers, as discussed briefly in \S,\ref{sec:futureexperiments}.
+Before taking on this very complex issue, however, we must first investigate a more fundamental one: how do current methods perform in the presence of representative data?}
+
 This paper is distinguished from other comparisons of \pz\ methods by its focus on the evaluation criteria for \pzpdf s and interpretation thereof.

### 2020-03-08 paper/tex/appendix.tex
fix LSST naming and typo in appendix
add sentences on potential bias causes

--- 2019-12-16 paper/tex/appendix.tex
@@
 
-Figure~\ref{fig:pz_pointestimates} shows both point estimates for all codes  both $z_{PEAK}$ and $z_{WEIGHT}$.
+Figure~\ref{fig:pz_pointestimates} shows \sout{both} point estimates \sout{both} $z_{PEAK}$ and $z_{WEIGHT}$ \boldblue{versus true redshift for all codes}.
 Point density is shown with mixed contours to emphasize that most of the galaxies do fall close to the $z_{phot} = z_{spec}$ line, while points trace the details of the catastrophic outlier populations.
@@
 The features perpendicular to the $z_{phot} = z_{spec}$ line are due to the $4000$ \AA\ break passing through the gaps between adjacent filters.
-Even the strongest codes feature populations far from the $z_{phot} = z_{spec}$ line representing a degeneracy in the space of colours and redshifts.
+\boldblue{In the presence of genuine ambiguities in the colour space to redshift mapping, the compression of information down to a single number that is the point estimate redshift fails to adequately capture the more complex nature of the multimodal redshift matching: using the mode selects the highest peak while ignoring posterior probability in lower valued redshift peaks, while measuring the mean of a multimodal PDF often results in a value lying at a redshift between the peaks, removed from either/any of the most likely redshift solutions.
+Thus,} \sout{E}even the strongest codes feature populations far from the $z_{phot} = z_{spec}$ line representing a \boldblue{genuine} degeneracy in the space of colours and redshifts.
 

### 2020-03-08 paper/tex/results.tex
add sentences on potential bias causes

--- 2019-11-22 paper/tex/results.tex
@@
 The high QQ values (i.~e.~ more high than low PIT values) of \bpz, \cmnn, \delight, \eazy, and \gpz\ indicate \pzpdf s biased low, and the low QQ values (more low than high PIT values) of \skynet\ and \tpz\ indicate \pzpdf s biased high.
+\boldblue{All three template-based codes use a parameterized functional form for the prior, and deviations from this form in the test set could account for some of the biases seen.
+In addition, the \eazy\ and \lephare\ codes both seem to contain an error in how the marginalize over likelihoods, which likely contributes to their observed biases.
+For machine learning codes, \delight\ employed a uniform prior in both apparent magnitude and SED type, which could have biased the resultant PDFs, and \gpz\ parameterized all PDFs as a single Gaussian, which may not have been a flexible enough parameterization to capture the full complexity of the posteriors.
+In other cases, an obvious source for bias was not discovered, though we note in the particular instance of \tpz\,the results showed a tradeoff in bias versus smoothing width during validation, and a smoothing value that balanced the two effects was chosen.}
+
 The gray shaded band marks the $2\sigma$ variance in PIT values found using the \trainz\ algorithm with a bootstrap resampling of the training set and a sample size of 30,000 galaxies, representing a very conservative estimate of the representative training sample size estimated as being required for direct \pz\ calibration \citep{Newman:2015}, and thus an approximate minimal error significance compared to ideal performance.
@@
 These five codes in particular have a free parameter for bandwidth, which may be responsible for this vulnerability, in spite of the opportunity for fine-tuning with perfect prior information.
+\boldblue{In many cases this bandwidth parameter can be thought of as tuning the size of the neighbourhood of influence in the parameter space of the training data.
+Choosing too small of a bandwidth can lead to overfitting and overconfidence, leading to narrow PDFs, while choosing too large of a neighbourhood can oversmooth and lead to overly broad PDFs.}
 \flexzboost's ``sharpening'' parameter (described in Section~\ref{sec:flexzboost}) played a key role in diagonalizing the QQ plot, indicating a common avenue for improvement in the approaches that share this type of parameter.

### 2020-03-09 paper/tex/introduction.tex
add Wolf 2009 reference and corresponding sentence

@@
 While the first data-driven \pz\ algorithms yielded a point estimate, \citet{Firth:03} estimated a \pzpdf\ using a neural net with realizations scattered within the photometric errors.
+\boldblue{Several years later \citet{Wolf:09} used a data driven $\chi^{2}$ approach to estimate redshift PDFs, and specifically noted an intrinsic floor in redshift uncertainties even under ideal conditions can arise due to the intrinsic spread of redshifts within the training set even in well localized regions of parameter space.}
 

### 2020-03-10 paper/tex/results.tex
bigger labels for fig1

@@
 %\centering
-\includegraphics[width=0.49\textwidth]{fig/pz_12codes_261931_noseaborn_crop.jpg}\includegraphics[width=0.49\textwidth]{fig/pz_12codes_471167_noseaborn_crop.jpg}\\
-\includegraphics[width=0.49\textwidth]{fig/pz_12codes_713178_noseaborn_crop.jpg}\includegraphics[width=0.49\textwidth]{fig/pz_12codes_982747_noseaborn_crop.jpg}
+\includegraphics[width=0.49\textwidth]{fig/pz_12codes_261931_biglabels_crop.jpg}\includegraphics[width=0.49\textwidth]{fig/pz_12codes_471167_biglabels_crop.jpg}\\
+\includegraphics[width=0.49\textwidth]{fig/pz_12codes_713178_biglabels_crop.jpg}\includegraphics[width=0.49\textwidth]{fig/pz_12codes_982747_biglabels_crop.jpg}
 \caption{The individual \pzpdf s (blue) distributions produced by the twelve codes (small panels) on four exemplary galaxies' photometry (large panels) with different true redshifts (red).

### 2020-03-10 plotting_scripts/plot_nzsums_scottsrule_biglabels.py
embiggen Nzsum labels, add ANNz2 sentences

@@ -1,0 +1,97 @@
+import sys,os
+import numpy as np
+import matplotlib
+matplotlib.use('Agg')
+import matplotlib.pyplot as plt
+#import seaborn as sns
+import scipy.ndimage.filters as scifilt
+
+def main(argv):
+    """
+    Sept 1, 2018: Add Scott's rule smoothing to all curves
+    Scott's rule bandwidth for scipy.stats.gaussian_kde that qp uses is 
+    given by   
+    n**(-1./(d+4)), so n^(-.2)                         
+    for 399,356 galaxies in interval [0,2] we have a bandwidth of 
+    sigma = 0.07581.  Given the array spacing of the stored NZPLOT_vectors is
+    0.001, this corresponds to a bandwidth of 75.8 pixels (and an extra fudge
+    factor of a 0.4 that is obviously necessary to match TrainZ that I don't 
+    understand)
+    """
+
+    smooth_sigma = 75.81*0.4 #see note in docstring on bandwidth choice
+    basepathpart = "./TESTDC1"
+    codes = ("ANNZ2","BPZ","DELIGHT","EAZY","FLEXZ","GPZ/AUG2018","LEPHARE",
+             "METAPHOR","NN","SKYNET","TPZ","NULL2")
+    labels = ("ANNz2","BPZ","Delight","EAZY","FlexZBoost","GPz","LePhare",
+              "METAPhoR","CMNN","SkyNet","TPZ","TrainZ")
+    labeldict = dict(zip(codes,labels))
+    
+    nzvectorfile = "NZPLOT_vectors.out"
+    
+    numcodes = len(codes)
+    numzs = 2001 #number of zbins in each NZPLOT_vectors.out file
+    szvals = np.zeros([numcodes,numzs])
+    pzvals = np.zeros([numcodes,numzs])
+    z_array = np.zeros(numzs)
+
+    for i,xfile in enumerate(codes):
+        direcpath = "%s%s"%(basepathpart,xfile)
+        fullpath = os.path.join(direcpath,nzvectorfile)
+        #print fullpath
+        data = np.loadtxt(fullpath,skiprows=1)
+        if i == 0:
+            z_array = data[:,0]
+        szvals[i,:] = data[:,1]
+        pzvals[i,:] = data[:,2]
+        smooth_pzvals = scifilt.gaussian_filter1d(pzvals,smooth_sigma,
+                                                  mode='constant')
+        print ("read in data for %s"%xfile)
+        
+
+    numrows = 3
[…]

### 2020-03-11 plotting_scripts/pitqq_biglabel.py
embiggen labels on Fig 2

@@ -1,0 +1,145 @@
+#file to read in all the point pz data
+#Oct 12, 2017
+import sys,os
+import numpy as np
+import matplotlib
+matplotlib.use('Agg')
+import matplotlib.pyplot as plt
+import seaborn as sns
+import matplotlib.gridspec as gridspec
+import matplotlib.ticker as ticker
+
+def main(argv):
+    basepathpart = "./TESTDC1"
+    codes = ("ANNZ2","BPZ","DELIGHT","EAZY/NEWSCATMAG","FLEXZ/MAR5RESULTS","GPZ/AUG2018","LEPHARE","METAPHOR","NN/DEC8RESULTS","SKYNET","TPZ","NULL")
+    labels = ("ANNz2","BPZ","Delight","EAZY","FlexZBoost","GPz","LePhare","METAPhoR","CMNN","SkyNet","TPZ","TrainZ")
+    labeldict = dict(zip(codes,labels))
+    
+    qqvectorfile = "TESTQQvectors.out"
+
+    PITfile = "TESTPITVALS.out"
+    
+    gridindeces = np.array([0,1,2,3,8,9,10,11,16,17,18,19])
+    
+    numcodes = len(codes)
+    numquants = 1001 #number of zbins in each NZPLOT_vectors.out file
+    numgals = 399356 #number of galaxies in each file
+    xlevel = float(numgals/101.) #flat line definition for PIT hist
+    trainz_1sig = 501.2# 2sigma 250.6#1sigma
+    xlevel_low = xlevel-trainz_1sig
+    xlevel_high = xlevel+trainz_1sig
+    
+    pitvals = np.zeros([numcodes,numgals])
+    qtvals = np.zeros([numcodes,numquants])
+    qdvals = np.zeros([numcodes,numquants])
+    diffvals = np.zeros([numcodes,numquants])
+
+    for i,xfile in enumerate(codes):
+        direcpath = "%s%s"%(basepathpart,xfile)
+        fullpath = os.path.join(direcpath,qqvectorfile)
+        #print fullpath
+        data = np.loadtxt(fullpath,skiprows=1)
+        qtvals[i,:] = data[:,0]
+        qdvals[i,:] = data[:,1]
+        diffvals[i,:] = np.subtract(data[:,1],data[:,0])
+        print ("read in QQ data for %s"%xfile)
+        pitpath = os.path.join(direcpath,PITfile)
+        pitdata = np.loadtxt(pitpath,skiprows=1)
[…]

### 2020-03-12 paper/tex/appendix.tex
embiggen Nzsum labels, add ANNz2 sentences
add z=1.4 text and embiggen labels on point figures

@@
 \centering
-\includegraphics[width=0.74\textwidth]{fig/NZsumplot_12codes_scottsrule.jpg}
+\includegraphics[width=0.74\textwidth]{fig/NZsumplot_12codes_scottsrule_biglabels.jpg}
 \caption{The smoothed stacked estimator $\hat{N}(z)$ of the redshift distribution (red) produced by each code (panels) compared to the true redshift distribution $\tilde{N}(z)$ (blue).
@@
 Many of the codes, including all the model-fitting approaches and \annz, \gpz, \metaphor, and \skynet\ from the data-driven camp, overestimate the redshift density at $z \sim 1.4$.
-This behavior is a consequence of the $4000$ \AA\ break passing through the gap between the $z$ and $y$ filters, which induces a genuine discontinuity in the $z - y$ colour as a function of redshift that can sway the \pzpdf\ estimates in the absence of bluer spectral features.
+This behavior is a consequence of the $4000$ \AA\ break passing through the gap between the $z$ and $y$ filters, which induces a \sout{genuine discontinuity}\boldblue{dramatic rise and fall} in the $z - y$ colour as a function of redshift\boldblue{.
+With a dearth of strong spectral features blue-ward of the 4000 \AA\ break in most galaxy SEDs, the degeneracies on either side of the peak in $z - y$ colour tend to broaden the redshift PDFs near $z \sim 1.4$, which can lead to the ``bump'' seen in the stacked $\hat{N}(z)$ estimate.} \sout{that can sway the \pzpdf\ estimates in the absence of bluer spectral features.}
 
@@
 \flexzboost\ and \tpz\ also perform superb recovery of the true redshift distribution, with only a slight deviation at $z \sim 1.4$.
-Our metrics, however, cannot discern whether these four approaches, as well as \delight, are spared the $z \sim 1.4$ degeneracy in $\hat{N}(z)$ because they have more effectively used information in the data or if the impact is simply washed out by the stacked estimator's effective average over the test set galaxy sample.
-See Appendix~\ref{sec:pointmetrics} for further discussion of the $z \sim 1.4$ issue.
+\sout{Our metrics, however, cannot discern whether these four approaches, as well as \delight, are spared the $z \sim 1.4$ degeneracy in $\hat{N}(z)$ because they have more effectively used information in the data or if the impact is simply washed out by the stacked estimator's effective average over the test set galaxy sample.
+See Appendix~\ref{sec:pointmetrics} for further discussion of the $z \sim 1.4$ issue.}
 
@@
 \centering
-\includegraphics[width=0.49\textwidth]{fig/ZPEAK_szpz_threecolumn_12codes_navy_lowalpha.jpg}
-\includegraphics[width=0.49\textwidth]{fig/ZWEIGHT_szpz_threecolumn_12codes_navy_lowalpha.jpg}
+\includegraphics[width=0.49\textwidth]{fig/ZPEAK_szpz_threecolumn_12codes_biglabels_crop.jpg}
+\includegraphics[width=0.49\textwidth]{fig/ZWEIGHT_szpz_threecolumn_12codes_biglabels_crop.jpg}
 \caption{The density of \pz\ point estimates (contours) reduced from the \pzpdf s with outliers (blue) beyond the outlier cutoff (red dashed lines), via the mode ($z_{PEAK}$, left panel) and main-peak-mean ($z_{WEIGHT}$, right panel).
@@
 The finite grid spacing of the \pzpdf s induces some discretization in $z_{PEAK}$.
-The features perpendicular to the $z_{phot} = z_{spec}$ line are due to the $4000$ \AA\ break passing through the gaps between adjacent filters.
-\boldblue{In the presence of genuine ambiguities in the colour space to redshift mapping, the compression of information down to a single number that is the point estimate redshift fails to adequately capture the more complex nature of the multimodal redshift matching: using the mode selects the highest peak while ignoring posterior probability in lower valued redshift peaks, while measuring the mean of a multimodal PDF often results in a value lying at a redshift between the peaks, removed from either/any of the most likely redshift solutions.
+The features perpendicular to the $z_{phot} = z_{spec}$ line are due to the $4000$ \AA\ break passing through the gaps between adjacent filters\boldblue{,  most notably at $z \sim 0.4$ between the g and r filters, and at $z \sim 1.4$ between the z and y filters.
+All codes show some level of concentrated catastrophic outliers, often due to degenerate redshift solutions for a single area of color space.
+In the presence of genuine ambiguities in the colour space to redshift mapping, the compression of information down to a single number that is the point estimate redshift fails to adequately capture the more complex nature of the multimodal redshift matching: using the mode selects the highest peak while ignoring posterior probability in lower valued redshift peaks, while measuring the mean of a multimodal PDF often results in a value lying at a redshift between the peaks, removed from either/any of the most likely redshift solutions.
 Thus,} \sout{E}even the strongest codes feature populations far from the $z_{phot} = z_{spec}$ line representing a \boldblue{genuine} degeneracy in the space of colours and redshifts.

### 2020-03-13 paper/tex/appendix.tex
change bootstraps to 44k and address point 17

@@
 Figure~\ref{fig:nz_stats} shows the quantitative Kolmogorov-Smirnoff (KS), Cramer-Von Mises (CvM), and Anderson Darling (AD) test statistics for each of the codes for the $\hat{N}(z)$ based measures.
-The horizontal lines show the the result of a bootstrap resampling of the training set using 30,000 samples for \trainz, representing a conservative idealized limit on expected performance for a modest-sized representative training set of galaxies, as mentioned in Section~\ref{sec:pitqq}.
+The horizontal lines show the the result of a bootstrap resampling of the training set using 30,000 samples for \trainz, representing a conservative \sout{idealized limit on expected performance for}\boldblue{error estimate assuming} a modest-sized representative training set of galaxies, as mentioned in Section~\ref{sec:pitqq}.
 The AD bootstrap statistic is elevated due to its sensitivity to the tails of distributions.

### 2020-03-13 paper/tex/results.tex
embiggen labels on Fig 2
change bootstraps to 44k and address point 17

@@
 
-The gray shaded band marks the $2\sigma$ variance in PIT values found using the \trainz\ algorithm with a bootstrap resampling of the training set and a sample size of 30,000 galaxies, representing a very conservative estimate of the representative training sample size estimated as being required for direct \pz\ calibration \citep{Newman:2015}, and thus an approximate minimal error significance compared to ideal performance.
+The gray shaded band marks the $2\sigma$ variance in PIT values found using the \trainz\ algorithm with a bootstrap resampling of the training set and a sample size of \sout{30}\boldblue{44},000 galaxies, representing a \sout{very} conservative estimate of the \boldblue{expected errors due to the training sample size} \sout{representative training sample size estimated as being required for direct \pz\ calibration  }, and thus an approximate minimal error significance compared to ideal performance.
 The existence of deviations in the PIT histograms outside of this gray shaded uncertainty range show that significant biases are present for some codes.
@@
 \centering
-\includegraphics[width=0.74\textwidth]{fig/PITANDQQplot_12codes_withsigmaband_crop.jpg}
+\includegraphics[width=0.74\textwidth]{fig/PITANDQQplot_12codes_withsigmaband_biglabels_crop.jpg}
 \caption{The QQ plot (red) and PIT histogram (blue) of the \pzpdf\ codes (panels) along with the ideal QQ (black dashed diagonal) and ideal PIT (gray horizontal) curves, as well as a difference plot for the QQ difference from the ideal diagonal (lower inset).
-The gray shaded region indicates the $2 \sigma$ range from a bootstrap resampling of the training set with a size of 30,000 galaxies using \trainz.
+The gray shaded region indicates the $2 \sigma$ range from a bootstrap resampling of the training set with a size of \sout{30}\boldblue{44},000 galaxies using \trainz.
 The twelve codes exhibit varying degrees of four deviations from perfection: an overabundance of PIT values at the centre of the distribution indicate a catalogue of overly broad \pzpdf s, an excess of PIT values at the extrema indicates a catalogue of overly narrow \pzpdf s, catastrophic outliers manifest as overabundances at PIT values of 0 and 1, and asymmetry indicates systematic bias, a form of model misspecification.
@@
 \centering
-\includegraphics[width=0.74\textwidth]{fig/KSvsCvMvsAD_PIT_withnull_jpg.jpg}
+\includegraphics[width=0.74\textwidth]{fig/KSvsCvMvsAD_PIT_withnull_44k.jpg}
 \caption{A visualization of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the PIT distributions.
 There is generally good agreement between these statistics, with differences corresponding to the codes with outstanding catastrophic outlier rates, a reflection of the differences in how each statistic weights the tails of the distribution.
-Horizontal lines indicate the level of uncertainty found by bootstrapping a training set sample of 30,000 galaxies using \trainz; none of the codes reach this conservative ideal floor in expected uncertainty.}
+Horizontal lines indicate the level of uncertainty found by bootstrapping a training set sample of \sout{30}\boldblue{44},000 galaxies using \trainz; none of the codes reach this conservative ideal floor in expected uncertainty.}
 \label{fig:pit_stats}
@@
 \annz's strong performance can be attributed to an aspect of the training process in which training set galaxies with PIT values that more closely match the percentiles of the DC1 training set's redshift distribution are upweighted; in effect, these quantile-based metrics were part of the algorithm itself that may or may not serve it well under more realistic experimental conditions.
-Similar to what was done for the PIT histograms in Figure~\ref{fig:pitqq}, we create bootstrap training samples of 30,000 galaxies for use with \trainz\ in order to estimate a conservative statistical floor that we would expect in real data.
-No code reaches this idealized floor, indicating that all codes suffer some degradation from the ideal when employing their implicit priors, though \annz, \flexzboost, and \gpz\ are within a factor of two.
+Similar to what was done for the PIT histograms in Figure~\ref{fig:pitqq}, we create bootstrap training samples of \sout{30}\boldblue{44},000 galaxies for use with \trainz\ in order to estimate a conservative statistical floor that we would expect in real data.
+No code reaches this idealized floor, indicating that all codes suffer some degradation from the ideal when employing their implicit priors, though \annz, \flexzboost, and \gpz\ are within \boldblue{roughly} a factor of two.
 

### 2020-03-14 paper/tex/appendix.tex
add final bootstrap fig update

--- 2020-03-12 paper/tex/appendix.tex
@@
 Figure~\ref{fig:nz_stats} shows the quantitative Kolmogorov-Smirnoff (KS), Cramer-Von Mises (CvM), and Anderson Darling (AD) test statistics for each of the codes for the $\hat{N}(z)$ based measures.
-The horizontal lines show the the result of a bootstrap resampling of the training set using 30,000 samples for \trainz, representing a conservative idealized limit on expected performance for a modest-sized representative training set of galaxies, as mentioned in Section~\ref{sec:pitqq}.
+The horizontal lines show the the result of a bootstrap resampling of the training set using \sout{30}\boldblue{44},000 samples for \trainz, representing a conservative \sout{idealized limit on expected performance for a} \boldblue{error estimate assuming our} modest-sized representative training set of galaxies, as mentioned in Section~\ref{sec:pitqq}.
 The AD bootstrap statistic is elevated due to its sensitivity to the tails of distributions.
@@
 \centering
-\includegraphics[width=0.74\textwidth]{fig/KSvsCvMvsAD_NZ_withnull_jpg.jpg}
+\includegraphics[width=0.74\textwidth]{fig/KSvsCvMvsAD_NZ_withnull_44k.jpg}
 \caption{A visualization of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the $\hat{N}(z)$ distributions.
-Horizontal lines indicate the statistic values (including uncertainty) achieved using \trainz\ via bootstrap resampling a training set containing 30,000 redshifts.
+Horizontal lines indicate the statistic values (including uncertainty) achieved using \trainz\ via bootstrap resampling a training set containing \sout{30}\boldblue{44},000 redshifts.
 We make the reassuring observation that these related statistics do not disagree significantly with one another.

### 2020-03-14 plotting_scripts/pitqq_biglabel.py
other plotting script updates

@@ -25,9 +25,9 @@ def main(argv):
     numquants = 1001 #number of zbins in each NZPLOT_vectors.out file
     numgals = 399356 #number of galaxies in each file
     xlevel = float(numgals/101.) #flat line definition for PIT hist
-    trainz_1sig = 501.2# 2sigma 250.6#1sigma
-    xlevel_low = xlevel-trainz_1sig
-    xlevel_high = xlevel+trainz_1sig
+    trainz_2sig = 446.5# 2sigma 223.27#1sigma
+    xlevel_low = xlevel-trainz_2sig
+    xlevel_high = xlevel+trainz_2sig
     
     pitvals = np.zeros([numcodes,numgals])
     qtvals = np.zeros([numcodes,numquants])

### 2020-03-14 plotting_scripts/plot_nzsums_scottsrule_biglabels.py

@@ -22,7 +22,7 @@ def main(argv):
     smooth_sigma = 75.81*0.4 #see note in docstring on bandwidth choice
     basepathpart = "./TESTDC1"
     codes = ("ANNZ2","BPZ","DELIGHT","EAZY","FLEXZ","GPZ/AUG2018","LEPHARE",
-             "METAPHOR","NN","SKYNET","TPZ","NULL2")
+             "METAPHOR","NN","SKYNET","TPZ","NULL")
     labels = ("ANNz2","BPZ","Delight","EAZY","FlexZBoost","GPz","LePhare",
               "METAPhoR","CMNN","SkyNet","TPZ","TrainZ")
     labeldict = dict(zip(codes,labels))
@@ -46,7 +46,7 @@ def main(argv):
         pzvals[i,:] = data[:,2]
         smooth_pzvals = scifilt.gaussian_filter1d(pzvals,smooth_sigma,
                                                   mode='constant')
-        print ("read in data for %s"%xfile)
+        print "read in data for %s"%xfile
         
 
     numrows = 3
@@ -59,7 +59,7 @@ def main(argv):
     
     for i in range(numcodes):
         ax = plt.subplot(numrows,numcols,i+1)
-        if i == 0: #only include label for first entry
+        if i == 0:
             plt.plot(z_array,szvals[i],c='b',linestyle='-',linewidth=2,
                      label="zspec", alpha=0.85)
         else:
@@ -70,12 +70,12 @@ def main(argv):
                  linewidth=2,label=tmplabel, alpha=0.85)
 
         ax.set_xlim([0.0,1.99])
-        ax.set_ylim([0.0,1.79])
+        ax.set_ylim([0.0,1.99])
         ax.yaxis.set_ticks(np.arange(0.,1.55,0.5))
-        if i ==0: #include lines only for first entry
-            ax.legend(loc="upper right",fontsize=14)
-        else: 
-            ax.legend(loc="upper right",fontsize=14, handletextpad=0,
+        if i == 0:
+            ax.legend(loc="upper left",fontsize=14)
+        else:
+            ax.legend(loc="upper left",fontsize=14,handletextpad=0.0,
                       handlelength=0)
         if (i<(numcodes-4)):
             ax.set_xticklabels([])
@@ -90,7 +90,7 @@ def main(argv):
         plt.yticks(fontsize=12)
 
    
[…]

### 2020-04-04 paper/authors.tex
fixing affiliation

--- 2019-12-17 paper/authors.tex
@@
 $^{1}$ Department of Physics, University of California, One Shields Ave., Davis, CA, 95616, USA\\
-$^{2}$ German Centre of Cosmological Lensing, Ruhr-Universitaet Bochum, Universitaetsstra{\ss}e 150, 44801 Bochum, Germany\\
+$^{2}$ Ruhr-University Bochum, Astronomical Institute, German Centre for Cosmological Lensing, Universitaetsstr. 150, 44801 Bochum, Germany\\
 $^{3}$ Center for Cosmology and Particle Physics, New York University, 726 Broadway, New York, 10003, USA\\

### 2020-05-15 paper/tex/analysis.tex
remove superfluous comments from tex
address cde loss concerns

@@
 \label{sec:metrics}
-%(Alex Malz, Rongpu Zhou, Jeff Newman, Ofer Lahav)
 
@@
 There are a number of metrics that can be used to test the accuracy of a \pz\ posterior as an estimator of a true \pz\ posterior if the true \pzpdf\ is known.
-However, the true \pzpdf\ of observed data is not accessible, and existing mock catalogues produce redshift-photometry pairs $(z, d)$ by a deterministic algorithm that does not correspond to a joint probability density from which one can take samples.
+However, the true \pzpdf\ of the observed data is not accessible, as existing mock catalogues produce redshift-photometry pairs $(z, d)$ by a deterministic algorithm that does not correspond to a joint probability density from which one can take samples.
 In these cases there is no ``true PDF'' for an individual object, and most measures of PDF fidelity will necessarily be restricted to probing the quality of the ensemble of \pzpdf s.
+\boldblue{While a discrete approximation to the true $p(z,d)$ is possible by sampling the local neighbourhood of parameter space in large datasets with methods like a nearest neighbor or conditional density estimate, we do not make an explicit computation of such an approximate distribution.
+Rather, we note that the CMNN and FlexZBoost algorithms are specific implementations of such algorithms which can be used as examples of such approximations in the absence of knowledge of the true $p(z,d)$.}
 (See \S,\ref{sec:futureexperiments} for a discussion of how one might circumvent this limitation.)
@@
 Even considering a single method under the same parametrization, such as the 200-bin $0 < z < 2$ piecewise constant function used here, the exact bin definitions must affect the result.
-%\red{Note: Let's address how the gridding (endpoints and spacing) affects the metrics, from the email correspondence Re: [LSST-DESC-MEMBERS] Summary of Nov 29 PZ Telecon. (AIM)}
 The piecewise constant format is chosen because of its established presence in the literature, and the choice of 200 bins was motivated by the approximate number of columns expected to be available for storage of \pzpdf s for the final \lsst\ Project tables.\footnote{See, e.~g.~the \lsst\ Data Products Definition Document, available at: \url{https://ls.st/dpdd}}
@@
 \label{sec:qualmet}
-%(Rongpu Zhou, Jeffrey Newman, Alex Malz)
-%(Alex Malz, Peter Freeman, Sam Schmidt)
-%(John Soo, Rongpu Zhou)
 
@@
 The closer the QQ plot is to diagonal, the closer the match between the distributions.
-% In this paper, QQ plots are used for two purposes: (1) for comparing $N(z)$ from \pzpdf s (estimated using Eq.~\ref{eq:stacked}) with the true $N(z)$, and (2) for assessing the overall consistency of an ensemble of \pzpdf s with their true redshifts on a population level, where the distribution of the PIT values (see previous section) is compared to a uniform distribution between $0$ and $1$.
-%\red{is this enough reason to include both?--SJS}
-%\aim{Maybe we should explain why we expect the distribution of CDFs to be uniform.}\red{Note: this is just describing the PIT.
-I'm going to swap the order of PIT to come before this, and then say we are evaluating the distribution of PIT values.
-I will also add why we expect this to be uniform--SJS}
 
@@
 \citet{Tanaka:17} use the histogram of PIT values as a diagnostic indicator of overall code performance, while \citet{Freeman:17} independently define the PIT and demonstrate how its individual values may be used both to perform hypothesis testing (via, e.~g.~ the KS, CvM, and AD tests; see below) and to construct QQ plots.
-%\red{We might want to divide this by redshift to get trends, as the overall PIT could be impacted by true n(z).
-%For example, a PIT with nontrivial shape could be due to that behavior being uniform as a function of redshift or significant inaccuracy only in a small range of redshifts where $n(z)$ is large and a flat PIT elsewhere.}
 Following Kodra \& Newman (in prep.) we define the PIT-based catastrophic outlier rate as the fraction of galaxies with $\mathrm{PIT} < 0.0001$ or $\mathrm{PIT} > 0.9999$, which should total 0.0002 for an ideal uniform distribution.
-
-%\subsubsection{Continuous Ranked Probability Score (CRPS)}
-%\label{sec:crps}
-%\red{The CRPS is used to assess the accuracy of two probabilistic models respectively.}
-
-%\subsubsection{Root-mean-square error (RMSE)}
-%\label{sec:rmse}
-%%(Alex Malz)%
-%We employ the familiar root-mean-square error
-%\begin{equation}
-%\mathrm{RMSE} \equiv \sqrt{\int_{-\infty}^{\infty}\left(\hat{f}(z)-f'(z)\right)^{2}\,dz} .
-%\end{equation}
-%%may be used for two purposes in this analysis.
-%%We calculate it between the true redshift distribution $N'(z)$ and each estimator $\hat{N}(z)$ (derived by Eq.~\ref{eq:stacked}) as well as between the ideal QQ and PIT curves from each method studied and the ideal curves.
-%Though this metric does not account for the fact that the redshift distribution function is, in fact, a probability distribution, it can still be interpreted as a measure of the integrated difference between the estimated distribution and the true distribution.%, and it can be used to quantify the otherwise qualitative metrics.
 
@@
 interpretable as the maximum difference between the CDFs of the empirical distribution of PIT values for the test sample $\hat{f}(z)$ and a reference distribution $\tilde{f}(z)$, in this case $U(0,1)$, for the ideal distribution of PIT values.
-% $CDF_{\rm phot}(z)$ and $CDF_{\rm spec}(z)$, the CDFs of the photo-$z$ and spectroscopic redshift respectively:
-%The KS test quantifies whether the two redshift distributions are drawn from a similar distribution, independent of binning.
-% The KS test quantifies the similarity between two distributions, independent of binning.
-% A lower $N_{\rm KS}$ value corresponds to more similar distributions.
-%We calculate two KS test statistics, one to test the distribution of $p(z)$ values, and one to test the stacked estimator $\hat{n}(z)$ of the redshift distribution.
-%The $p(z)$ test compares the CDF of the PIT values described in the previous section to that of a uniform distribution.
-%The $n(z)$ test compares the CDF of the distribution formed by stacking the $p(z)$ for every object in our test set against the CDF of the corresponding true redshifts.
-%In this case a lower $N_{\rm KS}$ value means that the photo-$z$ distribution fits the spectroscopic redshift distribution well.
-$F_{\rm phot}(z)$ is calculated by stacking the PDFs of every single object in the sample, producing a smooth cumulative $n(z)$ to be compared with $F_{\rm spec}(z)$.
 We also consider two variants of the KS statistic.
@@
 is a weighted mean-squared difference featuring enhanced sensitivity to discrepancies in the tails of the distribution.
-%The KS statistic and the CvM and AD statistics related to it are all used to evaluate the performance of both $\hat{n}(z)$ and the QQ and PIT of the CDFs of the ensemble of photo-$z$ interim posteriors.
 In anticipation of a substantial fraction of galaxies having PIT of 0 or 1, a consequence of catastrophic outliers, we evaluate the AD statistic with modified bounds of integration $(0.01, 0.99)$ to exclude those extremes in the name of numerical stability.
-
-%\subsubsection{Kullback-Leibler Divergence (KLD)}
-%\label{sec:kl}
-%(Alex Malz)
-%The KLD is a measure of the loss in information when using a probability distribution that is an estimate of a known true probability distribution, so we use
-%\begin{equation}
-%\mathrm{KLD} = \int_{-\infty}^{\infty}n'(z)\log\left[\frac{n'(z)}{\hat{n}(z)}\right]dz
-%\end{equation}
-%as a metric on the stacked redshift distribution estimator of Eq.~\ref{eq:stacked}.
-Noting the assumptions made in using Eq.~\ref{eq:stacked}, we say that a photo-$z$ PDFs method whose stacked redshift distribution estimator $\hat{n}(z)$ more closely resembles the true redshift distribution $n'(z)$ is desired, i.e.
-we seek to minimize the KLD.
 
@@
 \label{sec:CDE_loss}
-%(Alex Malz, Ann Lee)
 
@@
 where the first term is the expectation value of the \pz\ posterior with respect to the marginal distribution of the photometric covariates $\X$, the second term is the expectation value with respect to the joint distribution of $\X$ and the space $Z$ of all possible redshifts, and the third term $K_{f}$ is a constant depending only upon the true conditional densities $f(z \vert \x)$.
-We may estimate these expectations empirically on the test or validation data \citep[Eq.~7 in][]{Izbicki:17b} without knowledge of the true densities.
+\sout{We may estimate these expectations empirically on the test or validation data} \citep[Eq.~7 in][]{Izbicki:17b} \sout{without knowledge of the true densities.}
+\boldblue{One of the most powerful features of the CDE loss is the ability to estimate the loss function up to a constant even in the absence of knowledge of the true underlying distribution (see Eq.~7 and accompanying discussion in \citet{Izbicki:17} for more details).
+This feature enables a quantitative comparison of our estimation methods with our current dataset where we lack access to the true $f(z \vert \x)$.}

### 2020-05-15 paper/tex/analysis.tex
move analysis sentences to footnote

@@
 However, the true \pzpdf\ of the observed data is not accessible, as existing mock catalogues produce redshift-photometry pairs $(z, d)$ by a deterministic algorithm that does not correspond to a joint probability density from which one can take samples.
-In these cases there is no ``true PDF'' for an individual object, and most measures of PDF fidelity will necessarily be restricted to probing the quality of the ensemble of \pzpdf s.
-\boldblue{While a discrete approximation to the true $p(z,d)$ is possible by sampling the local neighbourhood of parameter space in large datasets with methods like a nearest neighbor or conditional density estimate, we do not make an explicit computation of such an approximate distribution.
-Rather, we note that the CMNN and FlexZBoost algorithms are specific implementations of such algorithms which can be used as examples of such approximations in the absence of knowledge of the true $p(z,d)$.}
+In these cases there is no ``true PDF'' for an individual object
+\footnote{\boldblue{While a discrete approximation to the true $p(z,d)$ is possible by sampling the local neighbourhood of parameter space in large datasets with methods like a nearest neighbor or conditional density estimate, we do not make an explicit computation of such an approximate distribution.
+Rather, we note that the CMNN and FlexZBoost algorithms are specific implementations of such algorithms which can be used as examples of such approximations in the absence of knowledge of the true $p(z,d)$.}}, and most measures of PDF fidelity will necessarily be restricted to probing the quality of the ensemble of \pzpdf s.
 (See \S,\ref{sec:futureexperiments} for a discussion of how one might circumvent this limitation.)

### 2020-06-12 paper/tex/introduction.tex
fix minor typo

@@
 These missing or misidentified populations would yield training sets that are not representative of the true underlying galaxy population, which can bias redshift estimates well beyond the levels required for upcoming analyses.
-Characterizing and mitigating these problems will be a major undertaking which we will cover in future papers, as discussed briefly in \S,\ref{sec:futureexperiments}.
+Characterizing and mitigating these problems will be a major undertaking which we will cover in future papers, as discussed briefly in \S \ref{sec:futureexperiments}.
 Before taking on this very complex issue, however, we must first investigate a more fundamental one: how do current methods perform in the presence of representative data?}

### 2020-08-18 paper/main.tex
add couple acknowledgements
add data availability statement

@@
 %%% Here is where you should add your specific acknowledgments, remembering that some standard thanks will be added via the \code{desc-tex/ack/*.tex} and \code{contributions.tex} files.
-
+The authors would like to thank the journal referee, who made many insightful suggestions that improved the draft.
 This paper has undergone internal review in the LSST Dark Energy Science Collaboration.
@@
 The authors express immense gratitude to Alex Abate, without whom this paper would not have gotten started.
+We would also like to thank Ofer Lahav and Will Hartley for contributions during the writing of the draft.
 % The authors would like to thank their LSST-DESC publication review committee for comments that improved the paper draft.
@@
 \input{./tex/appendix}
+
+\section*{Data Availability}
+The data underlying this article were provided by the makers of the Buzzard Flock simulation group by permission.
+Data will be shared on request with the corresponding authors upon permission of the Buzzard Flock simulation group.
+All analysis and plotting scripts used in the writing of this paper are available on GitHub in the \texttt{metric\_scripts} and \texttt{plotting\_scripts} directories of the following repository: \url{https://github.com/LSSTDESC/PZDC1paper}.
 

### 2020-08-21 paper/main.tex
add sentence on Buzzard data request

@@
 The data underlying this article were provided by the makers of the Buzzard Flock simulation group by permission.
-Data will be shared on request with the corresponding authors upon permission of the Buzzard Flock simulation group.
+Data will be shared on request to the corresponding authors upon permission of the Buzzard Flock simulation group.
+For updated and larger datasets, requests can be made directly to the authors of the Buzzard Flock simulations \citep{DeRose:19}.
 All analysis and plotting scripts used in the writing of this paper are available on GitHub in the \texttt{metric\_scripts} and \texttt{plotting\_scripts} directories of the following repository: \url{https://github.com/LSSTDESC/PZDC1paper}.

### 2020-09-02 paper/authors.tex

--- 2019-12-17 paper/authors.tex
@@
 R.~Zhou$^{15,24}$,
-R.~Izbicki$^{27,28}$,
-and The LSST Dark Energy Science Collaboration
+R.~Izbicki$^{27,28}$
+\begin{center} (LSST Dark Energy Science Collaboration) \end{center}
 }

### 2020-09-02 paper/tex/methods.tex
add Wolf 2009 reference and corresponding sentence

@@
 
-In this study, we used \annz-v.2.0.4 to output only the result of the ANN algorithm.
+In this study, we use \annz-v.2.0.4.
+\boldblue{Several combinations of the ANN, BDT, and kNN machine learning algorithms were tested during the training/validation stage.
+For this particular dataset, the optimal and most stable results came from a setup consisting solely of ANNs.
+We present results for this case and do not include results from a mixture of multiple algorithms.
+}
+\sout{to output only the result of the ANN algorithm.}
 \Pzpdf s were produced by running an ensemble of 5 ANNs with a $6:12:12:1$ architecture corresponding to the 6 $ugrizy$ inputs, 2 hidden layers with 12 nodes each, and 1 output of redshift.
@@
 As an example, for $N_{\mathrm{filt}}=5$ with PPF$=0.95$, $95$ per cent of all training galaxies consistent with the test galaxy will have $D_M < 11.07$.
-Undetected bands are dropped, thereby reducing the effective $N_{\mathrm{filt}}$ for that galaxy.
+\sout{Undetected bands} \boldblue{Bands in which the galaxy is not detected} are dropped, thereby reducing the effective $N_{\mathrm{filt}}$ for that galaxy.
 The \pzpdf\ of a given test set galaxy is the normalized distribution of redshifts of its colour-matched subset of training set galaxies.

### 2020-09-02 paper/tex/results.tex

@@
 The values of the CDE loss statistic of individual \pzpdf\ accuracy are provided in Table~\ref{tab:cdeloss}.
+\boldblue{Even in the absence of knowledge of the true underlying distribution the CDE loss can be determined up to a constant, thus the relative values provide a quantitative measure of code performance.}
 It is worth noting that strong performance on the CDE loss, corresponding to lower values of the metric, should imply strong performance on the other metrics, though the inverse is not necessarily true.

### 2020-09-03 paper/tex/analysis.tex
redefine macros, fix a few spacing issues

@@
 \end{equation}
-where the first term is the expectation value of the \pz\ posterior with respect to the marginal distribution of the photometric covariates $\X$, the second term is the expectation value with respect to the joint distribution of $\X$ and the space $Z$ of all possible redshifts, and the third term $K_{f}$ is a constant depending only upon the true conditional densities $f(z \vert \x)$.
-\sout{We may estimate these expectations empirically on the test or validation data} \citep[Eq.~7 in][]{Izbicki:17b} \sout{without knowledge of the true densities.}
+where the first term is the expectation value of the \pz\ posterior with respect to the marginal distribution of the photometric covariates $\X$, the second term is the expectation value with respect to the joint distribution of $\X$ and the space $Z$ of all possible redshifts, and the third term $K_{f}$ is a constant depending only upon the true conditional densities $f(z \vert \x)$\xsout{We may estimate these expectations empirically on the test or validation data} \citep[Eq.~7 in][]{Izbicki:17b}.
+\xsout{without knowledge of the true densities.}
 \boldblue{One of the most powerful features of the CDE loss is the ability to estimate the loss function up to a constant even in the absence of knowledge of the true underlying distribution (see Eq.~7 and accompanying discussion in \citet{Izbicki:17} for more details).

### 2020-09-03 paper/tex/data.tex
redefine macros, fix a few spacing issues

@@
 \boldblue{However, a problem with the k-correction calculation led to incorrect colors at $z>2$, which necessitated limiting the redshift range to $0 < z \leq 2.0$.}
-\sout{Systematic inconsistencies with galaxy colors at $z > 2$ were observed, so the catalogue was limited to $0 < z \leq 2.0$.}
+\xsout{Systematic inconsistencies with galaxy colors at $z > 2$ were observed, so the catalogue was limited to $0 < z \leq 2.0$.}
 To obtain a catalogue matching the \lsst\ Gold Sample, we imposed a cut of $i < 25.3$, which gives a signal-to-noise ratio $\gtrsim 30$ for most galaxies.

### 2020-09-03 paper/tex/methods.tex
redefine macros, fix a few spacing issues

@@
 
-\sout{Another key difference is the treatment of non-detections in one or more bands: some codes ignore incomplete bands, while others replace the value with either an estimate for the detection limit, the mean of other values in the training set, or another default value.
+\xsout{Another key difference is the treatment of non-detections in one or more bands: some codes ignore incomplete bands, while others replace the value with either an estimate for the detection limit, the mean of other values in the training set, or another default value.
 There are varying conventions among machine learning-based codes for treatment of non-detections, and no one prescription dominates in the \pz\ literature.
@@
 \boldblue{All three template-based codes replace per-band non-detections with an estimate of the 1$\sigma$ magnitude limit as estimated from the photometric error model.}
-\sout{The likelihood is a sum of observed flux error $\sigma_{b}^{\mathrm{obs}}$-weighted squared differences between the observed flux $F^{\mathrm{obs}}_{b}$ and the normalized predicted flux $F^{\mathrm{mod}}_{b}(T, z)$ in $N_{\mathrm{filt}}$ photometric filters $b$, which are the \lsst\ $ugrizy$ filters in this case.}
+\xsout{The likelihood is a sum of observed flux error $\sigma_{b}^{\mathrm{obs}}$-weighted squared differences between the observed flux $F^{\mathrm{obs}}_{b}$ and the normalized predicted flux $F^{\mathrm{mod}}_{b}(T, z)$ in $N_{\mathrm{filt}}$ photometric filters $b$, which are the \lsst\ $ugrizy$ filters in this case.}
 Specific implementation details of each code, e.~g.~prior form and implementation, are described below.
@@
 }
-\sout{to output only the result of the ANN algorithm.}
+\xsout{to output only the result of the ANN algorithm.}
 \Pzpdf s were produced by running an ensemble of 5 ANNs with a $6:12:12:1$ architecture corresponding to the 6 $ugrizy$ inputs, 2 hidden layers with 12 nodes each, and 1 output of redshift.
 Each of the five ANNs was trained with different random seeds for the initialization of input parameters, reserving half of the training set for validation to prevent overfitting.
-\boldblue{Galaxies not detected in the $u$-band} \sout{Undetected galaxies} were excluded from the training set, and per-band non-detections in the test set were replaced with the mean magnitude \boldblue{of the test set} in that band.
-\sout{within the entire test set.}
+\boldblue{Galaxies not detected in the $u$-band} \xsout{Undetected galaxies}were excluded from the training set, and per-band non-detections in the test set were replaced with the mean magnitude \boldblue{of the test set} in that band.
+\xsout{within the entire test set.}
 
@@
 As an example, for $N_{\mathrm{filt}}=5$ with PPF$=0.95$, $95$ per cent of all training galaxies consistent with the test galaxy will have $D_M < 11.07$.
-\sout{Undetected bands} \boldblue{Bands in which the galaxy is not detected} are dropped, thereby reducing the effective $N_{\mathrm{filt}}$ for that galaxy.
+\xsout{Undetected bands}\boldblue{Bands in which the galaxy is not detected} are dropped, thereby reducing the effective $N_{\mathrm{filt}}$ for that galaxy.
 The \pzpdf\ of a given test set galaxy is the normalized distribution of redshifts of its colour-matched subset of training set galaxies.

### 2020-09-23 paper/tex/appendix.tex

@@
 Many of the codes, including all the model-fitting approaches and \annz, \gpz, \metaphor, and \skynet\ from the data-driven camp, overestimate the redshift density at $z \sim 1.4$.
-This behavior is a consequence of the $4000$ \AA\ break passing through the gap between the $z$ and $y$ filters, which induces a \sout{genuine discontinuity}\boldblue{dramatic rise and fall} in the $z - y$ colour as a function of redshift\boldblue{.
-With a dearth of strong spectral features blue-ward of the 4000 \AA\ break in most galaxy SEDs, the degeneracies on either side of the peak in $z - y$ colour tend to broaden the redshift PDFs near $z \sim 1.4$, which can lead to the ``bump'' seen in the stacked $\hat{N}(z)$ estimate.} \sout{that can sway the \pzpdf\ estimates in the absence of bluer spectral features.}
+This behavior is a consequence of the $4000$ \AA\ break passing through the gap between the $z$ and $y$ filters, which induces a\xsout{genuine discontinuity}\boldblue{dramatic rise and fall} in the $z - y$ colour as a function of redshift\boldblue{.
+With a dearth of strong spectral features blue-ward of the 4000 \AA\ break in most galaxy SEDs, the degeneracies on either side of the peak in $z - y$ colour tend to broaden the redshift PDFs near $z \sim 1.4$, which can lead to the ``bump'' seen in the stacked $\hat{N}(z)$ estimate.}\xsout{that can sway the \pzpdf\ estimates in the absence of bluer spectral features.}
 
@@
 \flexzboost\ and \tpz\ also perform superb recovery of the true redshift distribution, with only a slight deviation at $z \sim 1.4$.
-\sout{Our metrics, however, cannot discern whether these four approaches, as well as \delight, are spared the $z \sim 1.4$ degeneracy in $\hat{N}(z)$ because they have more effectively used information in the data or if the impact is simply washed out by the stacked estimator's effective average over the test set galaxy sample.
+\xsout{Our metrics, however, cannot discern whether these four approaches, as well as \delight, are spared the $z \sim 1.4$ degeneracy in $\hat{N}(z)$ because they have more effectively used information in the data or if the impact is simply washed out by the stacked estimator's effective average over the test set galaxy sample.
 See Appendix~\ref{sec:pointmetrics} for further discussion of the $z \sim 1.4$ issue.}
@@
 Figure~\ref{fig:nz_stats} shows the quantitative Kolmogorov-Smirnoff (KS), Cramer-Von Mises (CvM), and Anderson Darling (AD) test statistics for each of the codes for the $\hat{N}(z)$ based measures.
-The horizontal lines show the the result of a bootstrap resampling of the training set using \sout{30}\boldblue{44},000 samples for \trainz, representing a conservative \sout{idealized limit on expected performance for a} \boldblue{error estimate assuming our} modest-sized representative training set of galaxies, as mentioned in Section~\ref{sec:pitqq}.
+The horizontal lines show the the result of a bootstrap resampling of the training set using\xsout{30}\boldblue{44},000 samples for \trainz, representing a conservative\xsout{idealized limit on expected performance for a}\boldblue{error estimate assuming our} modest-sized representative training set of galaxies, as mentioned in Section~\ref{sec:pitqq}.
 The AD bootstrap statistic is elevated due to its sensitivity to the tails of distributions.
@@
 \caption{A visualization of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the $\hat{N}(z)$ distributions.
-Horizontal lines indicate the statistic values (including uncertainty) achieved using \trainz\ via bootstrap resampling a training set containing \sout{30}\boldblue{44},000 redshifts.
+Horizontal lines indicate the statistic values (including uncertainty) achieved using \trainz\ via bootstrap resampling a training set containing\xsout{30}\boldblue{44},000 redshifts.
 We make the reassuring observation that these related statistics do not disagree significantly with one another.
@@
 
-Figure~\ref{fig:pz_pointestimates} shows \sout{both} point estimates \sout{both} $z_{PEAK}$ and $z_{WEIGHT}$ \boldblue{versus true redshift for all codes}.
+Figure~\ref{fig:pz_pointestimates} shows\xsout{both} point estimates\xsout{both} $z_{PEAK}$ and $z_{WEIGHT}$ \boldblue{versus true redshift for all codes}.
 Point density is shown with mixed contours to emphasize that most of the galaxies do fall close to the $z_{phot} = z_{spec}$ line, while points trace the details of the catastrophic outlier populations.
@@
 In the presence of genuine ambiguities in the colour space to redshift mapping, the compression of information down to a single number that is the point estimate redshift fails to adequately capture the more complex nature of the multimodal redshift matching: using the mode selects the highest peak while ignoring posterior probability in lower valued redshift peaks, while measuring the mean of a multimodal PDF often results in a value lying at a redshift between the peaks, removed from either/any of the most likely redshift solutions.
-Thus,} \sout{E}even the strongest codes feature populations far from the $z_{phot} = z_{spec}$ line representing a \boldblue{genuine} degeneracy in the space of colours and redshifts.
+Thus,}\xsout{E}even the strongest codes feature populations far from the $z_{phot} = z_{spec}$ line representing a \boldblue{genuine} degeneracy in the space of colours and redshifts.
 

### 2020-09-23 paper/tex/results.tex

@@
 
-The gray shaded band marks the $2\sigma$ variance in PIT values found using the \trainz\ algorithm with a bootstrap resampling of the training set and a sample size of \sout{30}\boldblue{44},000 galaxies, representing a \sout{very} conservative estimate of the \boldblue{expected errors due to the training sample size} \sout{representative training sample size estimated as being required for direct \pz\ calibration  }, and thus an approximate minimal error significance compared to ideal performance.
+The gray shaded band marks the $2\sigma$ variance in PIT values found using the \trainz\ algorithm with a bootstrap resampling of the training set and a sample size of \xsout{30}\boldblue{44},000 galaxies, representing a \xsout{very}conservative estimate of the \boldblue{expected errors due to the training sample size}\xsout{representative training sample size estimated as being required for direct \pz\ calibration  }, and thus an approximate minimal error significance compared to ideal performance.
 The existence of deviations in the PIT histograms outside of this gray shaded uncertainty range show that significant biases are present for some codes.
@@
 \caption{The QQ plot (red) and PIT histogram (blue) of the \pzpdf\ codes (panels) along with the ideal QQ (black dashed diagonal) and ideal PIT (gray horizontal) curves, as well as a difference plot for the QQ difference from the ideal diagonal (lower inset).
-The gray shaded region indicates the $2 \sigma$ range from a bootstrap resampling of the training set with a size of \sout{30}\boldblue{44},000 galaxies using \trainz.
+The gray shaded region indicates the $2 \sigma$ range from a bootstrap resampling of the training set with a size of \xsout{30}\boldblue{44},000 galaxies using \trainz.
 The twelve codes exhibit varying degrees of four deviations from perfection: an overabundance of PIT values at the centre of the distribution indicate a catalogue of overly broad \pzpdf s, an excess of PIT values at the extrema indicates a catalogue of overly narrow \pzpdf s, catastrophic outliers manifest as overabundances at PIT values of 0 and 1, and asymmetry indicates systematic bias, a form of model misspecification.
@@
 There is generally good agreement between these statistics, with differences corresponding to the codes with outstanding catastrophic outlier rates, a reflection of the differences in how each statistic weights the tails of the distribution.
-Horizontal lines indicate the level of uncertainty found by bootstrapping a training set sample of \sout{30}\boldblue{44},000 galaxies using \trainz; none of the codes reach this conservative ideal floor in expected uncertainty.}
+Horizontal lines indicate the level of uncertainty found by bootstrapping a training set sample of \xsout{30}\boldblue{44},000 galaxies using \trainz; none of the codes reach this conservative ideal floor in expected uncertainty.}
 \label{fig:pit_stats}
@@
 \annz's strong performance can be attributed to an aspect of the training process in which training set galaxies with PIT values that more closely match the percentiles of the DC1 training set's redshift distribution are upweighted; in effect, these quantile-based metrics were part of the algorithm itself that may or may not serve it well under more realistic experimental conditions.
-Similar to what was done for the PIT histograms in Figure~\ref{fig:pitqq}, we create bootstrap training samples of \sout{30}\boldblue{44},000 galaxies for use with \trainz\ in order to estimate a conservative statistical floor that we would expect in real data.
+Similar to what was done for the PIT histograms in Figure~\ref{fig:pitqq}, we create bootstrap training samples of \xsout{30}\boldblue{44},000 galaxies for use with \trainz\ in order to estimate a conservative statistical floor that we would expect in real data.
 No code reaches this idealized floor, indicating that all codes suffer some degradation from the ideal when employing their implicit priors, though \annz, \flexzboost, and \gpz\ are within \boldblue{roughly} a factor of two.

### 2020-09-23 paper/main.tex

@@
 
-\newcommand{\boldblue}[1]{{\bf \textcolor{blue}{#1}}}
+\newcommand{\boldblue}[1]{{ \textcolor{black}{#1}}}
+\newcommand{\xsout}[1]{}
 
@@
 A plethora of photo-$z$ PDF estimation methodologies abound, producing discrepant results with no consensus on a preferred approach.
-We present the results of a comprehensive experiment comparing twelve photo-$z$ algorithms applied to mock data produced for \sout{Large Synoptic Survey Telescope} \boldblue{The Rubin Observatory Legacy Survey of Space and Time} (\textsc{LSST}) Dark Energy Science Collaboration (\textsc{DESC}).
+We present the results of a comprehensive experiment comparing twelve photo-$z$ algorithms applied to mock data produced for\xsout{Large Synoptic Survey Telescope} \boldblue{The Rubin Observatory Legacy Survey of Space and Time} (\textsc{LSST}) Dark Energy Science Collaboration (\textsc{DESC}).
 By supplying perfect prior information, in the form of the complete template library and a representative training set as inputs to each code, we demonstrate the impact of the assumptions underlying each technique on the output photo-$z$ PDFs.

### 2020-09-23 paper/tex/introduction.tex

@@
 
-The current and next generations of large-scale galaxy surveys, including the Dark Energy Survey \citep[\proj{DES},][]{Abbott:05}, the Kilo-Degree Survey \citep[\proj{KiDS},][]{de_Jong:13}, Hyper Suprime-Cam Survey \citep[\proj{HSC},][]{Aihara:2018a,Aihara:2018b}, \sout{Large Synoptic Survey Telescope} \boldblue{The Rubin Observatory Legacy Survey of Space and Time} \citep[\lsst,][]{Abell:09}, Euclid \citep{Laureijs:11}, and Wide-Field Infrared Survey Telescope \citep[\proj{WFIRST},][]{Green:12}, represent a paradigm shift to reliance on photometric, rather than solely spectroscopic, galaxy catalogues of substantially larger size at a cost of lacking complete spectroscopically confirmed redshifts ($z$).
+The current and next generations of large-scale galaxy surveys, including the Dark Energy Survey \citep[\proj{DES},][]{Abbott:05}, the Kilo-Degree Survey \citep[\proj{KiDS},][]{de_Jong:13}, Hyper Suprime-Cam Survey \citep[\proj{HSC},][]{Aihara:2018a,Aihara:2018b}, \xsout{Large Synoptic Survey Telescope}\boldblue{The Rubin Observatory Legacy Survey of Space and Time} \citep[\lsst,][]{Abell:09}, Euclid \citep{Laureijs:11}, and Wide-Field Infrared Survey Telescope \citep[\proj{WFIRST},][]{Green:12}, represent a paradigm shift to reliance on photometric, rather than solely spectroscopic, galaxy catalogues of substantially larger size at a cost of lacking complete spectroscopically confirmed redshifts ($z$).
 Effective astrophysical inference using the catalogues resulting from these ongoing and upcoming missions, however, necessitates accurate and precise photometric redshift (\pz) estimation methodologies.
